### 2022-10-21 Optimizationnotes.tex
Add files via upload

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=20mm}
\usepackage{graphicx}
\usepackage{subcaption}

\title{notes on how light scattering set up is evaluated and optimised}

\begin{document}

\maketitle

\section*{Abstract}
Obtaining Static Light Scattering profiles from optically trapped entities poses considerable engineering challenge due to space constraints in a typical Optical Trap setup. We propose here a three-angle light scattering technique that can not only help characterize the structure but also be useful in understanding the dynamics of microscopic objects that are trapped optically. In this work, we demonstrate the use of such a measurement scheme in determining the optimal angles for light scattering measurements and probing the orientational dynamics of a microsphere dimer.

\section*{Introduction}
Optical trapping has been a vital tool in the study of single molecule properties, the ability to deliver a stable and precise force vector has proved invaluable for studying biological groups in particular \cite{2}. Recent years there has been a growing interest in the study of colloid aggregation and growth; where the optical binding effect resulted in rapid aggregation of nanoparticles around the trapping focus. Raman scattering techniques can be used to elucidate particle characteristics and have been applied to optical trapping methods to determine compositional information of the trapped entity \cite{4,7}. However, the particle's dynamics in the trap can not be elucidated based purely on the spectral data. For simple near spherical particles the motion can be largely attributed to Brownian motion which can be readily measured using a quadrant photo detector. It is not unusual however, that two microspheres to become attached to one another; in the presence of a focused Gaussian beam the microsphere dimer will reorient itself to align along the propagation axis of the beam. If the dimer is symmetrical then this reorientation motion is instantaneous, in the case of an asymmetrical dimer the alignment rapidly fluctuates about the propagation axis before eventually achieving an equilibrated orientation. Developing a method to model a dimer's motion within an optical trap based on information gathered in situ will prove useful for understanding the dynamics of more complexly shaped particle's that cannot be directly imaged.

This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{1}. They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis. In the case where the trapped object scatters anisotropically - such as for an asymmetrical dimer - it is necessary to utilise multiple detection fibres to build a better picture of the trapped object's orientation based on the produced light scattering. However, for inverted microscope optical traps there is a significant engineering challenge of placing enough detection fibres at the correct angles within a close enough distance to the trapping focus to collect a clear signal. This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres. Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be

\section*{Theory}

Consider a dimer with unequal sphere diameters $a_1, a_2$ in an optical trap positioned in some orientation $\hat{s}$. Due to the Brownian motion from the surrounding fluid the dimer's orientation changes with time. Assuming we cannot accurately determine the dimer's orientation from imaging alone we instead rely on it's light scattering. We collect the near field intensity, via optical fibres,from a probe laser at 3 predetermined angles $\theta_1, \theta_2, \theta_3$. To determine the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.

\begin{table}[tb]
\caption{\label{tab:n}}
\begin{tabular}{rrrr}
\hline\hline
orientation & $\hat{n}_x$ &  $\hat{n}_y$ &  $\hat{n}_z$ \\
\hline
$1$ & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$2$ & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$3$ & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$4$ & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$5$ & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$6$ & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$7$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$8$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$9$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$10$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
$11$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
$12$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
$13$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$14$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$15$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$16$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$17$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$18$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$19$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$20$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$21$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$22$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
$23$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
$24$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
\hline\hline
\end{tabular}
\end{table}

We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{6} to determine the light scattering intensity of our dimer in each reference orientation. We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k. Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern produced belongs to the reference orientation $\hat{n}$

\begin{align}
	p(y(\hat{s})\parallel\hat{n}) = \Pi^3_{k=1} 
	(2\pi\sigma_k^2)^{-1/2} e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
\end{align}

We can view this result as a conditional probability, if the orientation is determined ($\hat{n}$) then what is the probability that expected scattering signal matches our dimer's scattering signal. We instead want to know the inverse conditional, that with a given signal y the dimer was in orientation $\hat{n}$. We can calculate this using Bayes' theory:

\begin{align}
	p(\hat{n}\parallel y(\hat{s})) = \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
\end{align}

Where $p(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and scattering signals respectively. The former can be estimated by some Boltzmann distribution based on the radial distance between neighbouring reference orientations. This essentially is biasing our estimate by tying it to its proximity in the orientation space. The latter priori can be taken as the integral of  $p(y(\hat{s})\parallel \hat{n})$ and $p(\hat{n})$ across all possible reference orientations. In order to evaluate if our model is accurate we need first test it using an idealised simulation of the target dimer.

\subsection*{Testing the model}
We use the Brownian OT package by Jerome Fung \cite{5} to simulate how our target dimer moves while subject to an optical trap. At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern. We then use the above method to estimate the dimer orientation based on its scattering pattern. The priori estimate of the particle's orientation can be defined as a Boltzmann's distribution based on the radial distance between our previous estimate and each reference orientation. 
\begin{align}
	r_j(t, \hat{n}) = \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
	p(\hat{n}) = \frac{e^{r_j(t,\hat{n})}}
	{\Sigma_{i=1}^{24}e^{r_i(t, \hat{n})}}
\end{align}

Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know if our model can correctly predict the closest reference orientation. We take the dot product between our estimated $\hat{n}$ and the correct $\hat{n}$ and calculate the angle between the two. Performing a summation over the full simulation data we should see that the model has a total

$$ {\LARGE K_l(B \parallel D) = - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})}$$

Where $B_n$ is the Bayesian estimate and $D_n$ is the dot product result; the closer our estimate is to the dot product result the more accurate the model, and the divergence will approach 0. We can therefore evaluate the overall accuracy of our estimate based on the divergence - which is determined primarily by the detection angles.

\subparagraph*{Measurement of divergence parameter space}
The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.

Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{3}. Basic fitting methods (i.e. sum of least squares) are intended for finding the parameter choice with the best overall fit to the data. While this works well for functions with few fitting parameters and where the parameters are already constrained to predefined values; the computation time required for more functions where the parameters are less constrained and/or have no direct connection to the final returned result.

In cases like this its best to implement Monte Carlo techniques to intelligently sample across the entire parameter space. Ultra-nest implements a variation of this called nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation.

\section*{Results}
The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap. The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.

\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width =\textwidth]{traj.png}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth]}
		\includegraphics[width=\textwidth]{pos.png}
	\end{subfigure}
\end{figure}

\includegraphics{Divergance.png}[h]

\bibliography{bib}
\bibliographystyle{ieeetr}
\end{document}

### 2022-10-24 Optimizationnotes.tex
Add files via upload

@@
 \usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage{babel}
 
-\title{notes on how light scattering set up is evaluated and optimised}
+\title{Investigating the applications and limits of single particle light scattering using limited detection schemes}
+
+\author{Maciver, Daniel}
 
@@
 To determine the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
-
-\begin{table}[tb]
-\caption{\label{tab:n}}
+\begin{figure}
+	\begin{subfigure}{0.45\textwidth}
 			\begin{tabular}{rrrr}
@@
 			\end{tabular}
-\end{table}
+	\end{subfigure}
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=1.35\textwidth]{junk_00000.png}
+	\end{subfigure}
+\caption{Reference orientations plotted in the physical space. Green arrow points in the direction of the dimers orientation.}
+\end{figure}
 
@@
 \end{align}
+
+Where $\sigma_k$ is the average error in our signal collection; we assumed that in a best case scenario for our set up there will be roughly a $17\%$ error in the signal we collected.
+This was implemented by adjusting the values of $y(\hat{s}_k)$.
+\begin{eqnarray*}
+	\sigma_k = 0.17\bar{y}(\hat{n}_k) \\
+	\Rightarrow y(\hat{s}_k) = y(\hat{s})_k \pm n\sigma_k \\ 
+	n \in \mathbb{R}, n\in[0,1]
+\end{eqnarray*}
 
@@
 
-$$ {\LARGE K_l(B \parallel D) = - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})}$$
+$$  K_l(B \parallel D) = - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})$$
 
@@
 
-\begin{figure}
-	\begin{subfigure}{0.5\textwidth}
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width =\textwidth]{traj.png}
 	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth]}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\textwidth]{pos.png}
 	\end{subfigure}
+	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
 \end{figure}
 
-\includegraphics{Divergance.png}[h]
+As can be seen from Figure 2, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
+This rotation is caused by the dimer realigning itself so that the gradient forces are minimised.
+Its interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
+In Vigilante's work with dimers \cite{5} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
+
+The data from Figure 2a was used to test the efficacy of our predictive model.
+By recording how much our estimate diverges from the true orientation we get a rough score of how well our model can predict the dimer's orientation.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth]{Divergance.png}
+	\caption{Divergence between our model estimate and best estimate calculated via taking the dot product}
+\end{figure}
+
+Large spikes in the divergence plot indicate that the model is deviating far from the desired prediction.
+From Figure 3 we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off.
+This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error.
+Reduction of the signal error results in closer and closer
 
 \bibliography{bib}
-\bibliographystyle{ieeetr}
+\bibliographystyle{ieeetran}
 \end{document}

### 2022-10-27 Optimizationnotes.tex
Add files via upload

@@
 This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres.
-Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be
+Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be extracted from the trapped particle.
+
+The experiemental set up that this is based on is a inverted microscope optical trap, a suspension of microparticles are loaded onto a plastic cover slide.
+Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm HeNe laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
+The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microcopes camera.
+The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lense of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by []'s paper.
 
@@
 \section*{Results}
+\subparagraph*{Asymmetric dimer dynamics}
 The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap.
@@
 As can be seen from Figure 2, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
-This rotation is caused by the dimer realigning itself so that the gradient forces are minimised.
-Its interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
+Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
+It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
+Further simulations of differently sized dimers showed similar results, but only when $a_1 \geq 2a_2$.
+Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position.
 In Vigilante's work with dimers \cite{5} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
@@
 
+\subparagraph*{Predictive efficacy of model}
 The data from Figure 2a was used to test the efficacy of our predictive model.
@@
 
-Large spikes in the divergence plot indicate that the model is deviating far from the desired prediction.
+Spikes in the divergence plot indicate that the model is deviating  from the desired prediction.
 From Figure 3 we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off.
 This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error.
-Reduction of the signal error results in closer and closer
+Reduction of the signal error results in a higher accuracy in the models predictive efficacy to an extent; in truth the most significant factor in the prediction is our priori estimate of the dimer orientation $p(\hat{n})$.
+Since our choice of reference orientations have mirror copies in the physical space the conditional probabilities $p(y_k\parallel\hat{n})$ can be very close to each other.
+The addition of $p(\hat{n})$ means results in a significant improvement in the prediction by restraining our result to orientations close together in the physical space.
+
+\subsection*{Further improvements}
+While our model provides a satisfactory prediction of the dimer's motion; ideally this model can be exteneded to any particle characteristic that has a clear effect on the scattering from the trapped entity.
+For example using the mstm package we can determine how differently sized particles in similar orientations will produce different scattering patterns.
+
+\section*{Conclusion}
+We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. The
 

### 2022-11-01 Investigation into low light scattering.tex
Add files via upload

--- 2022-10-27 Optimizationnotes.tex
@@
 	\section*{Abstract}
-Obtaining Static Light Scattering profiles from optically trapped entities poses considerable engineering challenge due to space constraints in a typical Optical Trap setup.
-We propose here a three-angle light scattering technique that can not only help characterize the structure but also be useful in understanding the dynamics of microscopic objects that are trapped optically.
-In this work, we demonstrate the use of such a measurement scheme in determining the optimal angles for light scattering measurements and probing the orientational dynamics of a microsphere dimer.
+	Combing light scattering techniques with optical trapping set ups poses a significant engineering challenge.
+We propose a optical tweezer set up that will allow for light scattering measurements of a trapped micro particle.
+The set up is constrained by the limited space available to place optical fibres within range of the trapped entity.
+This paper discusses how light scattering data from a trapped dimer can be used to predict the orientation of the dimer within the trap.
+This method can be extended for determining characteristics that influence the light scattering pattern.
+We also discuss the how to determine best possible arrangement of optical fibres that will provide the greatest accuracy in our prediction.
+We hope to take this work further with real experimental data from our own light scattering measurements.
 
 	\section*{Introduction}
-Optical trapping has been a vital tool in the study of single molecule properties, the ability to deliver a stable and precise force vector has proved invaluable for studying biological groups in particular \cite{2}.
-Recent years there has been a growing interest in the study of colloid aggregation and growth; where the optical binding effect resulted in rapid aggregation of nanoparticles around the trapping focus.
-Raman scattering techniques can be used to elucidate particle characteristics and have been applied to optical trapping methods to determine compositional information of the trapped entity \cite{4,7}.
-However, the particle's dynamics in the trap can not be elucidated based purely on the spectral data.
+	Optical trapping has been a vital tool in the study of single molecule properties, the ability to deliver a stable and precise force vector has proved invaluable for studying biological groups in particular \cite{Biophysics_trapping}.
+Recent years there has been a growing interest in the study of colloid aggregation and growth; where the optical binding effect resulted in aggregation of particles around the trapping focus \cite{Colloids_and_tweezers}.
+Raman scattering techniques can be used to determine particle characteristics in a bulk or individual instance and have been applied to optical trapping methods to determine compositional information of the trapped entity \cite{Raman_Trapping,Raman_trapping_Biophysics}.
+However, the particle's dynamics in the trap can not be elucidated based purely on spectroscopy.
 For simple near spherical particles the motion can be largely attributed to Brownian motion which can be readily measured using a quadrant photo detector.
@@
 
-This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{1}.
+	This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{Dynamic_trap_scattering}.
 They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis.
@@
 
-The experiemental set up that this is based on is a inverted microscope optical trap, a suspension of microparticles are loaded onto a plastic cover slide.
-Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm HeNe laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
-The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microcopes camera.
-The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lense of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by []'s paper.
+	The experimental set up that this is based on is a inverted microscope optical trap, a suspension of micro particles are loaded onto a plastic cover slide.
+Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm He-Ne laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
+The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microscopes camera.
+The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lens of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by \cite{Long_distance_trap}'s paper.
 
@@
 To determine the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
-\begin{figure}
-	\begin{subfigure}{0.45\textwidth}
-			\begin{tabular}{rrrr}
-				\hline\hline
-				orientation & $\hat{n}_x$ &  $\hat{n}_y$ &  $\hat{n}_z$ \\
-				\hline
-				$1$ & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
-				$2$ & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
-				$3$ & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-				$4$ & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
-				$5$ & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
-				$6$ & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
-				$7$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
-				$8$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
-				$9$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
-				$10$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
-				$11$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
-				$12$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
-				$13$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
-				$14$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
-				$15$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-				$16$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
-				$17$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
-				$18$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
-				$19$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
-				$20$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
-				$21$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
-				$22$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
-				$23$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
-				$24$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
-				\hline\hline
-			\end{tabular}
-	\end{subfigure}
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[width=1.35\textwidth]{junk_00000.png}
-	\end{subfigure}
-\caption{Reference orientations plotted in the physical space. Green arrow points in the direction of the dimers orientation.}
+
+	\begin{figure}[t]
+		\centering
+		\includegraphics{junk_00078.png}
+		\caption{Reference orientations plotted in the physical space.
+Green arrow points in the direction of the dimers orientation.
+The red point is our estimation, and the green point is the best possible estimate.}
 	\end{figure}
 
-We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{6} to determine the light scattering intensity of our dimer in each reference orientation.
+	We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
 We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k.
@@
 	\begin{align}
-	p(y(\hat{s})\parallel\hat{n}) = \Pi^3_{k=1} 
+		p(y(\hat{s})\parallel\hat{n})&= \Pi^3_{k=1} 
 		(2\pi\sigma_k^2)^{-1/2} e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
@@
 	\begin{align}
-	p(\hat{n}\parallel y(\hat{s})) = \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
+		p(\hat{n}\parallel y(\hat{s}))&= \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
 	\end{align}
@@
 	\subsection*{Testing the model}
-We use the Brownian OT package by Jerome Fung \cite{5} to simulate how our target dimer moves while subject to an optical trap.
+	We use the Brownian OT package by Jerome Fung \cite{Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
 At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern.
@@
 	\begin{align}
-	r_j(t, \hat{n}) = \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
-	p(\hat{n}) = \frac{e^{r_j(t,\hat{n})}}
+		r_j(t, \hat{n})&= \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
+		p(\hat{n})&= \frac{e^{r_j(t,\hat{n})}}
 		{\Sigma_{i=1}^{24}e^{r_i(t, \hat{n})}}
@@
 
-$$  K_l(B \parallel D) = - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})$$
+	\begin{align}
+		K_l(B \parallel D) &= - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})
+	\end{align}
 
@@
 
-\subparagraph*{Measurement of divergence parameter space}
+	\subparagraph*{Optimization of angle choice}
 	The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
 
-Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{3}.
+	Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}.
 Basic fitting methods (i.e.
 sum of least squares) are intended for finding the parameter choice with the best overall fit to the data.
-While this works well for functions with few fitting parameters and where the parameters are already constrained to predefined values; the computation time required for more functions where the parameters are less constrained and/or have no direct connection to the final returned result.
-
+While this works well for functions with few fitting parameters and where the parameters are already constrained to predefined values; the computation time required for more functions where the parameters are less constrained or where there is no analytical solution connecting the parameters to the final output.
 In cases like this its best to implement Monte Carlo techniques to intelligently sample across the entire parameter space.
 Ultra-nest implements a variation of this called nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation.
+
+	We used the Ultra-nest software to create a set of live points of different choices of detection angles, with each point being evaluated by the divergence ($K_l(B\parallel D)$) produced.
+After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
+This results in a convergence on parameter choices that result in an average divergence result of almost 0.
+A common issue with complex parameter spaces, such as ours, is the issue of degeneracies (where two or more parameters have identical parameter spaces), to combat this we implemented a random step sampler that selects a new live point some 'n' steps away in the parameter space.
 
@@
 
-As can be seen from Figure 2, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
+	As can be seen from \figurename{ 2}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
 Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
@@
 Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position.
-In Vigilante's work with dimers \cite{5} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+	In Vigilante's work with dimers \cite{Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
 Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
@@
 	Spikes in the divergence plot indicate that the model is deviating from the desired prediction.
-From Figure 3 we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off.
+From \figurename{ 3} we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off.
 This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error.
@@
 Since our choice of reference orientations have mirror copies in the physical space the conditional probabilities $p(y_k\parallel\hat{n})$ can be very close to each other.
-The addition of $p(\hat{n})$ means results in a significant improvement in the prediction by restraining our result to orientations close together in the physical space.
+The addition of $p(\hat{n})$ results in a significant improvement in the prediction by restraining our prediction to orientations close by in the physical space.
 
-\subsection*{Further improvements}
-While our model provides a satisfactory prediction of the dimer's motion; ideally this model can be exteneded to any particle characteristic that has a clear effect on the scattering from the trapped entity.
-For example using the mstm package we can determine how differently sized particles in similar orientations will produce different scattering patterns.
+	\subsection*{Ultra-nest results}
+	We first allowed ultra nest to traverse the parameter space while each angle was constrained to a predefined area.
+We assumed that the each fibre would be placed within a roughly $30^{\circ}$ range so that we get a sampling of the entire light scattering pattern, we didn't search further than $100^{\circ}$ as the signals would be to low to detect for a micron scale scatterer.
+The sampler ran until the divergence result was in the order of $10^1$.
+
+	\begin{figure} [t]
+		\centering
+		\includegraphics[width=0.4\textwidth]{Cornerangleconstrained-1.png}
+		\caption{Corner plot from ultra-nest, darker regions indicate a better divergence result.
+$\theta_1$ was constrained between $15^{\circ}$ and $25^{\circ}$; $\theta_2$ was constrained between $30^{\circ}$ and $60^{\circ}$; $\theta_3$ was constrained between $65^{\circ}$ and $100^{\circ}$.}
+	\end{figure}
+	As can be seen from \figurename{ 4} the parameter space for our angle choice is essentially flat, with no explicit maxima to be found, the sudden increase in our divergence around $\theta_1 = 18^{\circ}$ is likely due to the fact that if we consider the scattering curves for our reference orientations there exist local minima on at roughly $\theta_1 = 13^{\circ}$ and $25^{\circ}$.
+Therefore between these two minima the model struggles to differentiate between reference orientations given that the light intensity is similar for different orientations.
+
+	We then decided to remove the limitations on our choice of angles, the only constraining factor being that the fibre's could not be placed within $10^{\circ}$ of each other.
+
+	\begin{figure} [h]
+		\centering
+		\includegraphics[width=0.4\textwidth]{corneranglesfree-1.png}
+		\caption{Corner plot from ultra-nest, darker regions indicate a better divergence result. Choice of angles is not limited to a predefined region.}
+	\end{figure}
+
+	Now we see that allowing the angles to be anywhere in the parameter space results in a more focused result for both $\theta_2$ and $\theta_3$ but still a flat parameter space for $\theta_1$.
+The result suggests that the model is more successful when we allow for backscattering to be considered in our model.
+While this works for a perfect system where noise is a known quantity it is unlikely if the signals are clear enough under laboratory conditions.
 
 	\section*{Conclusion}
-We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. The
+	We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
+The model can successfully predict the dimer's motion using its prior estimation to restrain the estimate in the physical space.
+
+	Analysis of the parameter space for our choice of detection angles shows a strong preference for angles far from the $0^{\circ}$ suggesting a more accurate estimation can be achieved by considering back scattering signals.
+We hope to validate our model by collecting light scattering signals from a trapped dimer and testing it against the video and QPD signals generated from our result.
 
 	\bibliography{bib} 
-\bibliographystyle{ieeetran}
+	\bibliographystyle{ieeetr}
+
 \end{document}

### 2022-11-02 Investigation into low light scattering.tex
Add files via upload

@@
 The latter priori can be taken as the integral of $p(y(\hat{s})\parallel \hat{n})$ and $p(\hat{n})$ across all possible reference orientations.
-In order to evaluate if our model is accurate we need first test it using an idealised simulation of the target dimer.
+In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
 
@@
 
-	Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know if our model can correctly predict the closest reference orientation.
-We take the dot product between our estimated $\hat{n}$ and the correct $\hat{n}$ and calculate the angle between the two.
-Performing a summation over the full simulation data we should see that the model has a total
+	Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know how often our model can determine the closest reference orientation $\hat{n}$ to the true orientation $\hat{s}$, and how precise our prediction is.
+For a perfect model our probability distribution should be peaked at the correct result and zero elsewhere.
+	\begin{equation}
+		p_{ideal}=
+		\begin{cases}
+			1 & \text{when $\hat{n}$ = $\hat{n}_{best}$}\\
+			0 & \text{anywhere else}
+		\end{cases}
+	\end{equation}
+
+	In reality our model will be some probability distribution that has some probability of any orientation being correct.
+To evaluate our model's accuracy and confidence we calculate the Kullback-Leibler divergence of the two probability distributions.
 
 	\begin{align}
-		K_l(B \parallel D) &= - \sum_{n=1}^{no. steps} B_n \ln(\frac{D_n}{B_n})
+		K_{l, \#}(p_{guess} \parallel p_{ideal}) &= 
+		p(\hat{n}\parallel y_k(\hat{s})) \ln \left[
+		\frac{p_{ideal}}{p(\hat{n}\parallel y_k(\hat{s}))}
+		\right]
 	\end{align}
 
-	Where $B_n$ is the Bayesian estimate and $D_n$ is the dot product result; the closer our estimate is to the dot product result the more accurate the model, and the divergence will approach 0.
-We can therefore evaluate the overall accuracy of our estimate based on the divergence - which is determined primarily by the detection angles.
+	Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of both the accuracy and the spread of our probability distribution.
+We sum the divergence of each measurement across the entire simulation to get a score for the entire simulation.
+
+	\begin{align}
+		K_{l,total} = \sum\limits_{\# =1}^{no. of steps} K_{l,\#}
+	\end{align}
+
+	Since our divergence result is a product of our choice of angles we can see analyse how the choice of angles can be altered to improve our estimate.
 
@@
 
-	We used the Ultra-nest software to create a set of live points of different choices of detection angles, with each point being evaluated by the divergence ($K_l(B\parallel D)$) produced.
+	We used the Ultra-nest software to create a set of live points of different choices of detection angles, with each point being evaluated by the divergence ($K_{l, total}$) produced.
 After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.

### 2022-11-03 Investigation into low light scattering.tex
Add files via upload

@@
 
-	Where $\sigma_k$ is the average error in our signal collection; we assumed that in a best case scenario for our set up there will be roughly a $17\%$ error in the signal we collected.
+	Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
 This was implemented by adjusting the values of $y(\hat{s}_k)$.
@@
 	Where $p(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and scattering signals respectively.
-The former can be estimated by some Boltzmann distribution based on the radial distance between neighbouring reference orientations.
-This essentially is biasing our estimate by tying it to its proximity in the orientation space.
-The latter priori can be taken as the integral of $p(y(\hat{s})\parallel \hat{n})$ and $p(\hat{n})$ across all possible reference orientations.
+The former can be estimated by some Boltzmann distribution based on the spacing between reference orientations.
+The latter priori $p(y)$ is the probability of our dimer producing the signal $y$ and is given as.
+
+	\begin{align}
+		p(y) = \int p(y(\hat{s})\parallel \hat{n}) p(\hat{n}) d\hat{n}
+	\end{align}
 	In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
@@
 We then use the above method to estimate the dimer orientation based on its scattering pattern.
-The priori estimate of the particle's orientation can be defined as a Boltzmann's distribution based on the radial distance between our previous estimate and each reference orientation.
+The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the radial distance between our previous estimate and each reference orientation.
 	\begin{align}
@@
 		K_{l, \#}(p_{guess} \parallel p_{ideal}) &= 
-		p(\hat{n}\parallel y_k(\hat{s})) \ln \left[
+		p_{ideal} \ln \left[
 		\frac{p_{ideal}}{p(\hat{n}\parallel y_k(\hat{s}))}
@@
 
-	Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of both the accuracy and the spread of our probability distribution.
-We sum the divergence of each measurement across the entire simulation to get a score for the entire simulation.
+	Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of both how confident our model is in it's estimation.
+We sum the divergence of each measurement across the entire simulation to get an idea of how well the model performs across the entire simulation.
 
@@
 	The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
+Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}.
+Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation.
 
-	Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}.
-Basic fitting methods (i.e.
-sum of least squares) are intended for finding the parameter choice with the best overall fit to the data.
-While this works well for functions with few fitting parameters and where the parameters are already constrained to predefined values; the computation time required for more functions where the parameters are less constrained or where there is no analytical solution connecting the parameters to the final output.
-In cases like this its best to implement Monte Carlo techniques to intelligently sample across the entire parameter space.
-Ultra-nest implements a variation of this called nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation.
-
-	We used the Ultra-nest software to create a set of live points of different choices of detection angles, with each point being evaluated by the divergence ($K_{l, total}$) produced.
+	We used the Ultra-nest software to create a set of live points of different choices of $\theta_1, \theta_2, \ \& \ \theta_3$, with each point being evaluated by the divergence ($K_{l, total}$) produced.
 After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
@@
 Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
+Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
 
 	\subparagraph*{Predictive efficacy of model}
-	The data from Figure 2a was used to test the efficacy of our predictive model.
-By recording how much our estimate diverges from the true orientation we get a rough score of how well our model can predict the dimer's orientation.
+	The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
+To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
+For a perfect estimation these two plots should overlap.
 
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.5\textwidth]{Divergance.png}
-		\caption{Divergence between our model estimate and best estimate calculated via taking the dot product}
+	\begin{figure}
+
 	\end{figure}
 
-	Spikes in the divergence plot indicate that the model is deviating from the desired prediction.
-From \figurename{ 3} we can see our model has short bursts of being highly accurate in its prediction before suddenly shooting off.
-This can be in part be due to the fact that our $y(\hat{s}_k)$ values are not set but fluctuate due to the signal error.
-Reduction of the signal error results in a higher accuracy in the models predictive efficacy to an extent; in truth the most significant factor in the prediction is our priori estimate of the dimer orientation $p(\hat{n})$.
-Since our choice of reference orientations have mirror copies in the physical space the conditional probabilities $p(y_k\parallel\hat{n})$ can be very close to each other.
-The addition of $p(\hat{n})$ results in a significant improvement in the prediction by restraining our prediction to orientations close by in the physical space.
+	We first tested how varying the error value would effect our model's efficacy; as can be seen from \figurename{ 3} our model actually benefits from some degree of signal noise.
+The introduction of noise allows the prediction to correct itself if it strays to much from the correct result, the trade off comes in a reduction in our model's precision and the occurence of random spikes away from the result.
+To combat these spikes we time averaged our estimation over $10 \ \mu s$ increments to reduce the effects of noise.
+This significantly reduced the sharp deviations from the correct result while still allowing the model to correct itself if it ever deviates to far from the correct orientation.
+Our model is therefore capable of taking a noisy signal and converting it into a close approximation of the
 

### 2022-11-04 Investigation into low light scattering.tex
Add files via upload

@@
 We sum the divergence of each measurement across the entire simulation to get an idea of how well the model performs across the entire simulation.
+To measure how much our result had improved we compare it to a worse case scenario, where every $\hat{n}$ is given uniform probability, and evaluate the improvement factor $F(K_l)$
 
@@
 		K_{l,total} = \sum\limits_{\# =1}^{no. of steps} K_{l,\#}
+		F(K_l) = \frac{K_{l,\ worst}}{K_{l,total}}
 	\end{align}
 
+	To measure how much our result had improved we compare it to a worse case scenario, where every $\hat{n}$ is given uniform probability.
 Since our divergence result is a product of our choice of angles we can see analyse how the choice of angles can be altered to improve our estimate.
@@
 
-	\begin{figure}
-
+	We first tested how varying the error value would effect our model's efficacy; calculating the improvement factor $F(K_l)$ with each test.
+	As can be seen from \figurename{ 3} our model actually performs better with noisy signals, to the extent that a signal error of $10\%$ resulted in a worse performance than if we had just guessed each orientation at random.
+	\begin{figure}[b]
+		\begin{subfigure}{0.33\textwidth}
+			\includegraphics[width=\textwidth]{Error_10_percent.png}
+			\subcaption{$F(K_l) = 0.9422$}
+		\end{subfigure}
+		\begin{subfigure}{0.33\textwidth}
+			\includegraphics[width=\textwidth]{Error_20_percent.png}
+			\subcaption{$F(K_l) = 1.366$}
+		\end{subfigure}
+		\begin{subfigure}{0.33\textwidth}
+			\includegraphics[width=\textwidth]{Error_30_percent.png}
+			\subcaption{$F(K_l)=1.440$}
+		\end{subfigure}
+		\caption{Model predictions with signal error at a) $10 \%$, b)  $20\%$, and c) $30\%$}
 	\end{figure}
-
-	We first tested how varying the error value would effect our model's efficacy; as can be seen from \figurename{ 3} our model actually benefits from some degree of signal noise.
-The introduction of noise allows the prediction to correct itself if it strays to much from the correct result, the trade off comes in a reduction in our model's precision and the occurence of random spikes away from the result.
-To combat these spikes we time averaged our estimation over $10 \ \mu s$ increments to reduce the effects of noise.
-This significantly reduced the sharp deviations from the correct result while still allowing the model to correct itself if it ever deviates to far from the correct orientation.
-Our model is therefore capable of taking a noisy signal and converting it into a close approximation of the
 

### 2022-11-07 Manuscript.tex
Add files via upload

--- 2022-11-04 Investigation into low light scattering.tex
@@
-\documentclass[11pt]{article}
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
 
-\usepackage{amsmath,amssymb}
-\usepackage{geometry}
-\geometry{margin=20mm}
-\usepackage{graphicx}
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
+%%
+%% 
+%%
+%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
+%%
+%%
+\documentclass[review,3p]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+\usepackage{amsthm}
 \usepackage{subcaption}
@@
 \usepackage{babel}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
+
+\journal{Journal of Quantitative Spectroscopy $\&$ Radiative Transfer}
+
+\begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%% \fntext[label3]{}
 
@@
 
-\author{Maciver, Daniel}
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
 
-\begin{document}
+\author[aff1]{Dan Maciver}
+\affiliation[aff1]{organization={Department of Chemical Engineering,
+			University of Strathclyde},
+            addressline={16 Richmond Street}, 
+            city={Glasgow},
+            postcode={G1 1XQ}, 
+            country={Scotland}}
+\author[aff1]{Praveen Parthasarathi}
+\author[aff1]{Mark Haw}
+\author[aff1]{Leo Lue}
+\author[aff1]{Jan Sefcik}
 
-	\maketitle
-
-	\section*{Abstract}
+\begin{abstract}
 Combing light scattering techniques with optical trapping set ups poses a significant engineering challenge.
@@
 
-	\section*{Introduction}
-	Optical trapping has been a vital tool in the study of single molecule properties, the ability to deliver a stable and precise force vector has proved invaluable for studying biological groups in particular \cite{Biophysics_trapping}.
-Recent years there has been a growing interest in the study of colloid aggregation and growth; where the optical binding effect resulted in aggregation of particles around the trapping focus \cite{Colloids_and_tweezers}.
-Raman scattering techniques can be used to determine particle characteristics in a bulk or individual instance and have been applied to optical trapping methods to determine compositional information of the trapped entity \cite{Raman_Trapping,Raman_trapping_Biophysics}.
-However, the particle's dynamics in the trap can not be elucidated based purely on spectroscopy.
-For simple near spherical particles the motion can be largely attributed to Brownian motion which can be readily measured using a quadrant photo detector.
-It is not unusual however, that two microspheres to become attached to one another; in the presence of a focused Gaussian beam the microsphere dimer will reorient itself to align along the propagation axis of the beam.
-If the dimer is symmetrical then this reorientation motion is instantaneous, in the case of an asymmetrical dimer the alignment rapidly fluctuates about the propagation axis before eventually achieving an equilibrated orientation.
-Developing a method to model a dimer's motion within an optical trap based on information gathered in situ will prove useful for understanding the dynamics of more complexly shaped particle's that cannot be directly imaged.
+\end{abstract}
 
-	This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{Dynamic_trap_scattering}.
-They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis.
-In the case where the trapped object scatters anisotropically - such as for an asymmetrical dimer - it is necessary to utilise multiple detection fibres to build a better picture of the trapped object's orientation based on the produced light scattering.
-However, for inverted microscope optical traps there is a significant engineering challenge of placing enough detection fibres at the correct angles within a close enough distance to the trapping focus to collect a clear signal.
-This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres.
-Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be extracted from the trapped particle.
+%%Graphical abstract
+\begin{graphicalabstract}
+	\centering
+	\includegraphics[width=0.8\textwidth, height = 0.6\textheight]{fig1a.png}
+\end{graphicalabstract}
 
-	The experimental set up that this is based on is a inverted microscope optical trap, a suspension of micro particles are loaded onto a plastic cover slide.
-Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm He-Ne laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
-The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microscopes camera.
-The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lens of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by \cite{Long_distance_trap}'s paper.
+%%Research highlights
+\begin{highlights}
+\item Asymmetric dimers of undergo a full inversion in the presence of an potential well
+\item Orientation can be classified by applying Bayesian Statistics to light intensity readings
+\item Bayesian model shows increased accuracy when processing noisy signal measurements. 
+\item Multiple estimations can be  used to cancels out artefacts of noise. 
+\end{highlights}
 
-	\section*{Theory}
+\begin{keyword}
+	Optical Trapping \sep Light Scattering \sep Measurements \sep Statistics
+
+\end{keyword}
+
+\end{frontmatter}
+
+%%\linenumbers
+
+%% main text
+\section{Introduction}
+\label{1}
+
+Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics~\cite{Wang 2021} to those that test the fundamental assumptions of Quantum Mechanics ~\cite{Duan 2013} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few pico-newton.
+Recent applications of the Optical Tweezers in metrological measurements ~\cite{Scott Waitukaitis} and colloidal aggregation ~\cite{Optical Binding} demand the tweezer to also be able to characterise  the trapped entity.
+To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering ~\cite{P K Gupta} and dynamical characterisation has been achieved mainly by following the centre-of-mass Brownian motion of the trapped entity using a Quadrant Photo Detector~\cite{Alexander Rohrbach}.
+While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision~\cite{Alexander Rohrbach} and  rotation~\cite{recent paper} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in ~\cite{New Zealand paper} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
+
+Possibility of integrating light scattering with optical trapping was demonstrated by Bar-Ziv et.al.,~\cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's Brownian motion was studied.
+While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
+In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in ~\cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.
+
+\section{Theory}
+\label{2}
 
@@
 	\centering
-		\includegraphics{junk_00078.png}
+	\includegraphics[width = 0.5\textwidth]{junk_00078.png}
 	\caption{Reference orientations plotted in the physical space.
@@
 
-	\begin{align}
+	\begin{eqnarray}
 		p(y(\hat{s})\parallel\hat{n})&= \Pi^3_{k=1} 
 		(2\pi\sigma_k^2)^{-1/2} e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
-	\end{align}
+	\end{eqnarray}
 
@@
 This was implemented by adjusting the values of $y(\hat{s}_k)$.
-	\begin{eqnarray*}
+\begin{eqnarray}
 	\sigma_k = 0.17\bar{y}(\hat{n}_k) \\
@@
 	n \in \mathbb{R}, n\in[0,1]
-	\end{eqnarray*}
+\end{eqnarray}
 
@@
 
-	\begin{align}
+\begin{eqnarray}
 	p(\hat{n}\parallel y(\hat{s}))&= \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
-	\end{align}
+\end{eqnarray}
 
@@
 
-	\begin{align}
+\begin{eqnarray}
 	p(y) = \int p(y(\hat{s})\parallel \hat{n}) p(\hat{n}) d\hat{n}
-	\end{align}
+\end{eqnarray}
 In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
 
-	\subsection*{Testing the model}
+\subsection{Testing the model}
+\label{2.1}
 We use the Brownian OT package by Jerome Fung \cite{Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
@@
 We then use the above method to estimate the dimer orientation based on its scattering pattern.
-The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the radial distance between our previous estimate and each reference orientation.
-	\begin{align}
-		r_j(t, \hat{n})&= \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
-		p(\hat{n})&= \frac{e^{r_j(t,\hat{n})}}
+The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
+\begin{eqnarray}
+	r_j(t, \hat{n})= \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
+	p(\hat{n})= \frac{e^{(r_j(t,\hat{n})-1)}}
 	{\Sigma_{i=1}^{24}e^{r_i(t, \hat{n})}}
-	\end{align}
+\end{eqnarray}
 
 Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know how often our model can determine the closest reference orientation $\hat{n}$ to the true orientation $\hat{s}$, and how precise our prediction is.
-For a perfect model our probability distribution should be peaked at the correct result and zero elsewhere.
-	\begin{equation}
-		p_{ideal}=
-		\begin{cases}
-			1 & \text{when $\hat{n}$ = $\hat{n}_{best}$}\\
-			0 & \text{anywhere else}
-		\end{cases}
-	\end{equation}
-
+For a perfect model our probability distribution should equal 1 at the correct result and zero elsewhere.
 In reality our model will be some probability distribution that has some probability of any orientation being correct.
@@
 
-	\begin{align}
+\begin{eqnarray}
 	K_{l, \#}(p_{guess} \parallel p_{ideal}) &= 
@@
 	\right]
-	\end{align}
+\end{eqnarray}
 
@@
 
-	\begin{align}
-		K_{l,total} = \sum\limits_{\# =1}^{no. of steps} K_{l,\#}
+\begin{eqnarray}
+	K_{l,total} = \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
 	F(K_l) = \frac{K_{l,\ worst}}{K_{l,total}}
-	\end{align}
+\end{eqnarray}
 
@@
 
-	\subparagraph*{Optimization of angle choice}
+\subparagraph{Optimization of angle choice}
+\label{2.2}
 The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
@@
 
-	\section*{Results}
-	\subparagraph*{Asymmetric dimer dynamics}
+\section{Results}
+\label{3}
+\subsection{Asymmetric dimer dynamics}
+\label{3.1}
 The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap.
@@
 
-	\subparagraph*{Predictive efficacy of model}
+\subsection{Predictive efficacy of model}
+\label{3.2}
 The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
@@
 As can be seen from \figurename{ 3} our model actually performs better with noisy signals, to the extent that a signal error of $10\%$ resulted in a worse performance than if we had just guessed each orientation at random.
-	\begin{figure}[b]
+\begin{figure}
 	\begin{subfigure}{0.33\textwidth}
@@
 
-	\subsection*{Ultra-nest results}
+The issue with having a high signal error is that our model's prediction can be very far off from the correct answer and still be confident with it's results; fortunately our priori for the dimer's orientation can help to correct our estimate by limiting the probability in the physical space.
+We found that the inclusion of our priori estimate compared to a uniform priori improved our results by around $20 \%$.
+We can alter the influence of our priori by adding an additional exponential
+
+\subsection{Ultra-nest results}
+\label{3.3}
 We first allowed ultra nest to traverse the parameter space while each angle was constrained to a predefined area.
@@
 	\centering
-		\includegraphics[width=0.4\textwidth]{corneranglesfree-1.png}
+	\includegraphics[width=0.4\textwidth]{corneranglesfreed-1.png}
 	\caption{Corner plot from ultra-nest, darker regions indicate a better divergence result. Choice of angles is not limited to a predefined region.}
@@
 
-	\section*{Conclusion}
+\section{Conclusion}
+\label{4}
 We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
-The model can successfully predict the dimer's motion using its prior estimation to restrain the estimate in the physical space.
+Compared to a random selection of the particle's orientation we find that our model has a $150\%$ increase in it's accuracy.
+Even more interestingly, our model performs better under noisy signal conditions by allowing our estimate to self-correct itself, in the future we hope to remove the artefacts of the signal noise by combing multiple estimates using different detection angles.
+This can be achieved via the inclusion of additional signal detections, we refrained from collecting more signal point for two key reasons; firstly, we wanted to show the extent of the mathematical model in the case of limited detection schemes, and secondly, we wanted to limit our model to what is possible in a real world experimental set up.
+Once we have a clearer understanding of the maximum limit of detection fibres that can be placed we can update our model to account for this and improve the accuracy further.
 
-	Analysis of the parameter space for our choice of detection angles shows a strong preference for angles far from the $0^{\circ}$ suggesting a more accurate estimation can be achieved by considering back scattering signals.
-We hope to validate our model by collecting light scattering signals from a trapped dimer and testing it against the video and QPD signals generated from our result.
+Our simulations of asymmetric dimers found that for large sphere ratios the dimer will prefer to be orientated against the beam propagation and will even move further from the trapping focus to invert itself before coming back to the laser focus.
+It is interesting that is only the case for dimers where the size ratio is greater than 1:2, understanding how the optical torque varies with size ratio can provide unique insight into the dynamics of non-spherical trapping targets.
+
+Overall the model developed can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
+The MSTM package is flexible in  calculating the light scattering for a variety of micro particles, so long as we can calculate the light scattering we can apply our model to help characterise the trapped model.
+We hope to extend this work for characterising continuos processes such as nucleation within an optical trap.
+
+The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.
 
@@
 \bibliographystyle{ieeetr}
-
 \end{document}
+\endinput

### 2022-11-07 Manuscript.tex
Add files via upload

@@
 %%
-\documentclass[review,3p]{elsarticle}
+\documentclass[final,3p, twocolumn]{elsarticle}
 
@@
 %% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-\usepackage{amsthm}
+\usepackage{amssymb, amsthm}
+\usepackage{amsmath}
+%% The amsthm package provides extended theorem environment
 \usepackage{subcaption}
@@
 
-\author[aff1]{Dan Maciver}
+\author[aff1]{Dan Maciver\corref{cor1}} %
+\ead{Daniel.Maciver.2016@uni.strath.ac.uk}
+
+\author[aff1]{Praveen Parthasarathi}
+
+\author[aff1]{Mark Haw}
+
+\author[aff1]{Leo Lue}
+
+\author[aff1]{Jan Sefcik}
+
+\cortext[cor1]{Corresponding author}
 \affiliation[aff1]{organization={Department of Chemical Engineering,
@@
             country={Scotland}}
-\author[aff1]{Praveen Parthasarathi}
-\author[aff1]{Mark Haw}
-\author[aff1]{Leo Lue}
-\author[aff1]{Jan Sefcik}
 
@@
 \begin{keyword}
-	Optical Trapping \sep Light Scattering \sep Measurements \sep Statistics
+	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics
 
@@
 Assuming we cannot accurately determine the dimer's orientation from imaging alone we instead rely on it's light scattering.
-We collect the near field intensity, via optical fibres,from a probe laser at 3 predetermined angles $\theta_1, \theta_2, \theta_3$.
-To determine the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
+We collect the near field intensity, via optical fibres,from a probe laser at 3 angles $\theta_1, \theta_2, \theta_3$.
+To classify the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
 
@@
 
-	\begin{eqnarray}
-		p(y(\hat{s})\parallel\hat{n})&= \Pi^3_{k=1} 
-		(2\pi\sigma_k^2)^{-1/2} e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
-	\end{eqnarray}
+\begin{align}
+	p(y(\hat{s})\parallel\hat{n}) =& \Pi^3_{k=1}
+	(2\pi\sigma_k^2)^{-1/2} \nonumber \\     &e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
+\end{align}
 
-Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
+Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of signal error would effect our predictive model.
 This was implemented by adjusting the values of $y(\hat{s}_k)$.
-\begin{eqnarray}
-	\sigma_k = 0.17\bar{y}(\hat{n}_k) \\
-	\Rightarrow y(\hat{s}_k) = y(\hat{s})_k \pm n\sigma_k \\ 
-	n \in \mathbb{R}, n\in[0,1]
-\end{eqnarray}
+
+\begin{align}
+	\centering
+	\sigma_k =& 0.17\bar{y}(\hat{n}_k) \\
+	\Rightarrow y(\hat{s}_k) =& y(\hat{s})_k \pm n\sigma_k \\ 
+	n \in \mathbb{R}, \ \ & n\in[0,1] \nonumber
+\end{align}
 
@@
 
-\begin{eqnarray}
+\begin{align}
 	p(\hat{n}\parallel y(\hat{s}))&= \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
-\end{eqnarray}
+\end{align}
 
@@
 
-\begin{eqnarray}
+\begin{align}
 	p(y) = \int p(y(\hat{s})\parallel \hat{n}) p(\hat{n}) d\hat{n}
-\end{eqnarray}
-In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
+\end{align}
+
+The inclusion of $p(y)$ normalises our probability distribution. In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
 
@@
 We then use the above method to estimate the dimer orientation based on its scattering pattern.
-The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
-\begin{eqnarray}
-	r_j(t, \hat{n})= \hat{n}_{t} \cdot \hat{n}_{t-\Delta t} \\
-	p(\hat{n})= \frac{e^{(r_j(t,\hat{n})-1)}}
-	{\Sigma_{i=1}^{24}e^{r_i(t, \hat{n})}}
-\end{eqnarray}
+The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
 
+\begin{align}
+	r_j(t, \hat{n})= \frac{1}{2}((\hat{n}_{t} \cdot \hat{n}_{t-\Delta t})-1) \\
+	p(\hat{n})= \frac{e^{\beta (r_j(t,\hat{n})-1)}}
+	{\Sigma_{i=1}^{24}e^{\beta(r_i(t, \hat{n}))}}
+\end{align}
+
+Where $\beta$ is a weighting factor that controls the significance of our priori estimate.
 Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know how often our model can determine the closest reference orientation $\hat{n}$ to the true orientation $\hat{s}$, and how precise our prediction is.
@@
 
-\begin{eqnarray}
+\begin{align}
 	K_{l, \#}(p_{guess} \parallel p_{ideal}) &= 
-	p_{ideal} \ln \left[
-	\frac{p_{ideal}}{p(\hat{n}\parallel y_k(\hat{s}))}
+	p_{ideal} \nonumber \\ 
+	&\ln \left[\frac{p_{ideal}}{p(\hat{n}\parallel y_k(\hat{s}))}
 	\right]
-\end{eqnarray}
+\end{align}
 
@@
 
-\begin{eqnarray}
+\begin{align}
 	K_{l,total} = \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
 	F(K_l) = \frac{K_{l,\ worst}}{K_{l,total}}
-\end{eqnarray}
+\end{align}
 
@@
 
-\subparagraph{Optimization of angle choice}
+\subparagraph{Optimization of parameters}
 \label{2.2}
@@
 Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}.
-Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation.
+Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
 
-We used the Ultra-nest software to create a set of live points of different choices of $\theta_1, \theta_2, \ \& \ \theta_3$, with each point being evaluated by the divergence ($K_{l, total}$) produced.
+We used the Ultra-nest software to create a set of live points of different choices of $\theta_1, \ \theta_2, \ \& \ \theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$) produced.
 After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
@@
 	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
 		\includegraphics[width =\textwidth]{traj.png}
@@
 	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{pos.png}
@@
 
-We first tested how varying the error value would effect our model's efficacy; calculating the improvement factor $F(K_l)$ with each test.
-As can be seen from \figurename{ 3} our model actually performs better with noisy signals, to the extent that a signal error of $10\%$ resulted in a worse performance than if we had just guessed each orientation at random.
-\begin{figure}
+We first tested how varying the error value would effect our model's efficacy; calculating the improvement factor $F(K_l)$ with each test, the measurements were made with detection angles of $15^{\circ}, \ 45^{\circ}, \ \& \ 90^{\circ}$.
+As can be seen from \figurename{ 3} our model actually performs better with noisy signals, to the extent that there was a 0.2 difference between our estimation with $10 \%$ signal error and $30\%$.
+The reasoning for this being that a noisy signal means that any time our model deviates from the correct response it will quickly self correct due to the larger variance.
+The issue with having a high signal error is that our model's prediction can be very far off from the correct answer and still be confident with it's results; fortunately our priori for the dimer's orientation can help to correct our estimate by limiting the probability in the physical space.
+We found that the inclusion of our priori estimate compared to a uniform priori improved our results by around $20 \%$.
+
+\begin{figure}[t]
+	\centering
 	\begin{subfigure}{0.33\textwidth}
@@
 	\end{subfigure}
-	\caption{Model predictions with signal error at a) $10 \%$, b)  $20\%$, and c) $30\%$}
+	\caption{Model predictions with signal error at a) $10 \%$, b)  $20\%$, and c) $30\%$. Dotted line represents minimum distance between reference orientations}
 \end{figure}
 
-The issue with having a high signal error is that our model's prediction can be very far off from the correct answer and still be confident with it's results; fortunately our priori for the dimer's orientation can help to correct our estimate by limiting the probability in the physical space.
-We found that the inclusion of our priori estimate compared to a uniform priori improved our results by around $20 \%$.
-We can alter the influence of our priori by adding an additional exponential
+We estimate that at minimum our experimental setup should have a $17\%$ signal error, assuming near perfect alignment of fibres and minimal interference from other microspheres in the solution.
+Therefore the fact that our model provides a better estimation of the dimer's motion with worse noise is promising for the implementation of our model into a experimental setup.
+To assess the impact of signal error, angle choice, and weighting factor we utilised ultranest to step through the entire parameter space.
 
@@
 \label{3.3}
-We first allowed ultra nest to traverse the parameter space while each angle was constrained to a predefined area.
-We assumed that the each fibre would be placed within a roughly $30^{\circ}$ range so that we get a sampling of the entire light scattering pattern, we didn't search further than $100^{\circ}$ as the signals would be to low to detect for a micron scale scatterer.
-The sampler ran until the divergence result was in the order of $10^1$.
+We first wanted to measure analyse the impact of our choice of angles in our predictive model.
 
@@
 \end{figure}
+
 As can be seen from \figurename{ 4} the parameter space for our angle choice is essentially flat, with no explicit maxima to be found, the sudden increase in our divergence around $\theta_1 = 18^{\circ}$ is likely due to the fact that if we consider the scattering curves for our reference orientations there exist local minima on at roughly $\theta_1 = 13^{\circ}$ and $25^{\circ}$.

### 2022-11-08 Manuscript.tex
Add files via upload

@@
 
-Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics~\cite{Wang 2021} to those that test the fundamental assumptions of Quantum Mechanics ~\cite{Duan 2013} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few pico-newton.
-Recent applications of the Optical Tweezers in metrological measurements ~\cite{Scott Waitukaitis} and colloidal aggregation ~\cite{Optical Binding} demand the tweezer to also be able to characterise  the trapped entity.
-To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering ~\cite{P K Gupta} and dynamical characterisation has been achieved mainly by following the centre-of-mass Brownian motion of the trapped entity using a Quadrant Photo Detector~\cite{Alexander Rohrbach}.
-While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision~\cite{Alexander Rohrbach} and  rotation~\cite{recent paper} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in ~\cite{New Zealand paper} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
+\indent Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few piconewton.
+Recent applications of the Optical Tweezers in metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical} demand the tweezer to also be able to characterise  the trapped entity.
+To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman} and dynamical characterisation has been achieved mainly by following the centre-of-mass brownian motion of the trapped entity using a Quadrant Photo Detector \cite{friedrich2012tuning}.
+While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision \cite{friedrich2012tuning} and  rotation \cite{yifat2021facile} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
 
-Possibility of integrating light scattering with optical trapping was demonstrated by Bar-Ziv et.al.,~\cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's Brownian motion was studied.
+Possibility of integrating light scattering with optical trapping was demonstrated by Saffran and 
+co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's brownian motion was studied.
 While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
-In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in ~\cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.
+In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.
 
@@
 
-We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
+We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{Mishchenko1996MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
 We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k.
@@
 \label{2.1}
-We use the Brownian OT package by Jerome Fung \cite{Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
+We use the Brownian OT package by Jerome Fung \cite{Vigilante2020Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
 At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern.
@@
 The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
-Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Ultranest}.
+Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Buchner2016Ultranest}.
 Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
@@
 Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position.
-In Vigilante's work with dimers \cite{Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
 Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.

### 2022-11-09 Manuscript.tex
Update Manuscript.tex

@@
 \begin{abstract}
-Combing light scattering techniques with optical trapping set ups poses a significant engineering challenge.
-We propose a optical tweezer set up that will allow for light scattering measurements of a trapped micro particle.
-The set up is constrained by the limited space available to place optical fibres within range of the trapped entity.
-This paper discusses how light scattering data from a trapped dimer can be used to predict the orientation of the dimer within the trap.
-This method can be extended for determining characteristics that influence the light scattering pattern.
-We also discuss the how to determine best possible arrangement of optical fibres that will provide the greatest accuracy in our prediction.
-We hope to take this work further with real experimental data from our own light scattering measurements.
+While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem.
+Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup.
+We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entitites.
+To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped.
+We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
 

### 2022-11-09 Manuscript.tex
Add files via upload

@@
 %%
-\documentclass[final,3p, twocolumn]{elsarticle}
+\documentclass[preprint,3p]{elsarticle}
 
@@
 \usepackage{babel}
+\usepackage{wrapfig}
+\usepackage{geometry}
+\usepackage{fancyref}
 %% The lineno packages adds line numbers. Start line numbering with
@@
 
-\title{Investigating the applications and limits of single particle light scattering using limited detection schemes}
+\title{Investigating the Applications and Limits of Single asymmetric Particle Light Scattering using Limited Detection Schemes}
 
@@
 %%             country={}}
-%%
+%%Saffran
 %% \affiliation[label2]{organization={},
@@
 			University of Strathclyde},
-            addressline={16 Richmond Street}, 
+            addressline={75 Montrose Street}, 
             city={Glasgow},
-            postcode={G1 1XQ}, 
+            postcode={G1 1XL}, 
             country={Scotland}}
@@
 Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup.
-We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entitites.
+We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities.
 To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped.
 We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
-
 \end{abstract}
-
-%%Graphical abstract
-\begin{graphicalabstract}
-	\centering
-	\includegraphics[width=0.8\textwidth, height = 0.6\textheight]{fig1a.png}
-\end{graphicalabstract}
-
 %%Research highlights
@@
 \begin{keyword}
-	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics
-
+	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics \sep Theoretical 
 \end{keyword}
@@
 
-%%\linenumbers
+\linenumbers
 
@@
 \section{Introduction}
-\label{1}
+\label{sec:Intro}
 
@@
 Recent applications of the Optical Tweezers in metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical} demand the tweezer to also be able to characterise  the trapped entity.
-To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman} and dynamical characterisation has been achieved mainly by following the centre-of-mass brownian motion of the trapped entity using a Quadrant Photo Detector \cite{friedrich2012tuning}.
+To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman} and dynamical characterisation has been achieved mainly by following the centre-of-mass Brownian motion of the trapped entity using a Quadrant Photo Detector \cite{friedrich2012tuning}.
+\begin{wrapfigure}[12]{r}{0.3\textwidth}
+	\centering
+	\includegraphics[width=0.3\textwidth]{fig1a}
+	\caption{Experimental set up}
+	\label{fig:1}
+	\vspace{-100pt}
+\end{wrapfigure}
 While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision \cite{friedrich2012tuning} and  rotation \cite{yifat2021facile} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
-
-Possibility of integrating light scattering with optical trapping was demonstrated by Saffran and 
-co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's brownian motion was studied.
+Possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's Brownian motion was studied.
 While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
-In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure 1.
+\newpage
+In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:1}.
 
 \section{Theory}
-\label{2}
+\label{sec:Theory}
 
-Consider a dimer with unequal sphere diameters $a_1, a_2$ in an optical trap positioned in some orientation $\hat{s}$.
+Consider a dimer with unequal sphere diameters $a_1, a_2$ in an optical trap in some orientation $\hat{s}$.
 Due to the Brownian motion from the surrounding fluid the dimer's orientation changes with time.
 Assuming we cannot accurately determine the dimer's orientation from imaging alone we instead rely on it's light scattering.
-We collect the near field intensity, via optical fibres,from a probe laser at 3 angles $\theta_1, \theta_2, \theta_3$.
+Rather than directly calculate the orientation vector we instead discretise the orientation space into 30 reference orientations.
+See \ref{Appendix 1} for the exact numerical values of reference orientations.
+The following records how we apply a Bayesian model to determine the estimate the closest reference orientation based on an instantaneous light scattering measurement.
+
+We use a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{Mishchenko1996MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
+The raw signal intensities for at 3 different angles ($\theta_1, \ \theta_2, \ \& \ \theta_3$) are calculated and then scaled down so that the signal response $y_k$ is normalised for each angle $k$:
+
+\begin{align}
+	y_k(\hat{n}) = 
+	 \frac{\left[I(\hat{n}, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
+	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
+\end{align}
+
+See \ref{Appendix 1} for a breakdown of the raw intensities and scaled signal responses.
 To classify the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width = 0.5\textwidth]{junk_00078.png}
-	\caption{Reference orientations plotted in the physical space.
-Green arrow points in the direction of the dimers orientation.
-The red point is our estimation, and the green point is the best possible estimate.}
-\end{figure}
-
-We use the a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{Mishchenko1996MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
 We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k.
-Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern produced belongs to the reference orientation $\hat{n}$
+Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern belongs to the reference orientation $\hat{n}$
 
@@
 	p(y(\hat{s})\parallel\hat{n}) =& \Pi^3_{k=1}
-	(2\pi\sigma_k^2)^{-1/2} \nonumber \\     &e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma_k^2}
+	(2\pi\sigma^2)^{-1/2} \nonumber \\     &e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma^2}
 \end{align}
 
-Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of signal error would effect our predictive model.
-This was implemented by adjusting the values of $y(\hat{s}_k)$.
-
-\begin{align}
-	\centering
-	\sigma_k =& 0.17\bar{y}(\hat{n}_k) \\
-	\Rightarrow y(\hat{s}_k) =& y(\hat{s})_k \pm n\sigma_k \\ 
-	n \in \mathbb{R}, \ \ & n\in[0,1] \nonumber
-\end{align}
-
+Where $\sigma$ is the minimum distance between our reference orientations signals ($y_1, \ y_2, \ \& \ y_3$), .
 We can view this result as a conditional probability, if the orientation is determined ($\hat{n}$) then what is the probability that expected scattering signal matches our dimer's scattering signal.
@@
 
-\subsection{Testing the model}
-\label{2.1}
+\subsection{Brownian Simulation}
+\label{sec:2.1}
 We use the Brownian OT package by Jerome Fung \cite{Vigilante2020Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
-At each time step we determine the dimer's orientation and use MSTM to calculate its scattering pattern.
-We then use the above method to estimate the dimer orientation based on its scattering pattern.
-The priori estimate $p(\hat{n})$ can be defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
+The package is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) to simulate the motion of arbitrary shaped particles.
+
+For any trapped particle their are two major sources of motion, Brownian motion and optical forces.
+The Brownian displacement ($\Delta q^{\textbf{B}}$) is approximated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
+The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}.
+
+\begin{align}
+	\langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle &= 2D_{ij} \Delta t 
+\end{align}
+
+For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
+The T-matirx describes the scattering process by relating the scattering wave to the incident wave in an elegant equation.
+
+\begin{align}
+	\begin{pmatrix}
+		p_{mn} \\
+		q_{mn}
+	\end{pmatrix}
+	= \textbf{T}
+	\begin{pmatrix}
+		a_{mn} \\
+		b_{mn}
+	\end{pmatrix}
+\end{align}
+
+MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
+The benefit of this approach is that MSTM can accurately compute the T-matrix for the dimer once and then \textit{ott} can calculate the optical force after each time step.
+The dimer displacement is the sum of both the optical and Brownian contributions.
+Brownian OT returns the displacement as a trajectory file; with the [x,y,z] position being relative to the trap focus, and the orientation being returned as a quaternion.
+
+For our model we simulate the motion of silica dimer trapped in a 5mW Gaussian beam; the dimer is comprised of two spheres with radii $1\\mu m, \ \& 0.5 \ \mu m$ respectively.
+
+\subsection{Testing our Model}
+\label{sec:2.2}
+
+Since MSTM has computed the T-matrix for our dimer, we can reuse it to calculate.
+We then apply \eqref{3} - covered in Section~\ref{sec:2.1} - to estimate the dimer orientation based on its scattering pattern.
+The priori estimate $p(\hat{n})$ is defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
 
@@
 	p(\hat{n})= \frac{e^{\beta (r_j(t,\hat{n})-1)}}
-	{\Sigma_{i=1}^{24}e^{\beta(r_i(t, \hat{n}))}}
+	{\Sigma_{i=1}^{30}e^{\beta(r_i(t, \hat{n})-1)}}
 \end{align}
@@
 
-Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of both how confident our model is in it's estimation.
+Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of how confident our model is in it's estimation.
 We sum the divergence of each measurement across the entire simulation to get an idea of how well the model performs across the entire simulation.
@@
 \begin{align}
-	K_{l,total} = \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
-	F(K_l) = \frac{K_{l,\ worst}}{K_{l,total}}
+	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
+	K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
+	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
 \end{align}
 
-To measure how much our result had improved we compare it to a worse case scenario, where every $\hat{n}$ is given uniform probability.
-Since our divergence result is a product of our choice of angles we can see analyse how the choice of angles can be altered to improve our estimate.
+The worst case scenario can be akin to randomly choosing a reference orientation at each time step.
+Because our model is dependent on several parameters at once we need to a more sophisticated method for finding the optimum choice of parameters.
 
 \subparagraph{Optimization of parameters}
-\label{2.2}
-The efficacy of our estimation is a function of our detection angles, testing every possible configuration manually is time consuming, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
+\label{2.3}
+The efficacy of our estimation is a function of our detection angles, signal error, and weighting of $p(\hat{n})$.
+Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
 Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Buchner2016Ultranest}.
-Ultra-nest implements a variation of Monte Carlo integration,nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
+Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}, \ 0 \leq \epsilon \leq 1, \ 0 \leq \beta \leq 5$).
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\centering
@@
 		\subcaption{}
-		\includegraphics[width =\textwidth]{traj.png}
+		\includegraphics[width =\textwidth]{./Images/traj.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{pos.png}
+		\includegraphics[width=\textwidth]{./Images/pos.png}
 	\end{subfigure}
@@
 
+\newpage
 \subsection{Predictive efficacy of model}
-\label{3.2}
+\label{sec:3.2}
 The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
@@
 For a perfect estimation these two plots should overlap.
+Overall their our three main factors that effect the performance of our prediction; the signal error $\sigma$, our choice of angles $\theta_1, \theta_2, \ \& \theta_3$, and our weighting factor $\beta$.
 
-We first tested how varying the error value would effect our model's efficacy; calculating the improvement factor $F(K_l)$ with each test, the measurements were made with detection angles of $15^{\circ}, \ 45^{\circ}, \ \& \ 90^{\circ}$.
-As can be seen from \figurename{ 3} our model actually performs better with noisy signals, to the extent that there was a 0.2 difference between our estimation with $10 \%$ signal error and $30\%$.
-The reasoning for this being that a noisy signal means that any time our model deviates from the correct response it will quickly self correct due to the larger variance.
-The issue with having a high signal error is that our model's prediction can be very far off from the correct answer and still be confident with it's results; fortunately our priori for the dimer's orientation can help to correct our estimate by limiting the probability in the physical space.
-We found that the inclusion of our priori estimate compared to a uniform priori improved our results by around $20 \%$.
+We tested how changing our weighting factor for a set choice of angles would influence the performance of our model.
+We first set our choice of angles to $15^{\circ}, \ 45^{\circ},\ \& \ 90^{\circ}$ and varied $\beta$ between 0.5-1.5.
 
-\begin{figure}[t]
-	\centering
+\begin{figure}[b]
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{Error_10_percent.png}
-		\subcaption{$F(K_l) = 0.9422$}
+		\includegraphics[width=\textwidth]{./Images/154590_Beta=1.png}
 	\end{subfigure}
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{Error_20_percent.png}
-		\subcaption{$F(K_l) = 1.366$}
+		\includegraphics[width=\textwidth]{./Images/154590_Beta=05.png}
 	\end{subfigure}
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{Error_30_percent.png}
-		\subcaption{$F(K_l)=1.440$}
+		\includegraphics[width=\textwidth]{./Images/154590_Beta=15.png}
 	\end{subfigure}
-	\caption{Model predictions with signal error at a) $10 \%$, b)  $20\%$, and c) $30\%$. Dotted line represents minimum distance between reference orientations}
 \end{figure}
 
-We estimate that at minimum our experimental setup should have a $17\%$ signal error, assuming near perfect alignment of fibres and minimal interference from other microspheres in the solution.
-Therefore the fact that our model provides a better estimation of the dimer's motion with worse noise is promising for the implementation of our model into a experimental setup.
-To assess the impact of signal error, angle choice, and weighting factor we utilised ultranest to step through the entire parameter space.
+\newpage
+\subsection{Ultranest Results}
+\label{3.3}
+We wanted to see if our choice of angles plays a significant role in our model's efficacy.
+We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
+We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
 
-\subsection{Ultra-nest results}
-\label{3.3}
-We first wanted to measure analyse the impact of our choice of angles in our predictive model.
-
-\begin{figure} [t]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=0.4\textwidth]{Cornerangleconstrained-1.png}
-	\caption{Corner plot from ultra-nest, darker regions indicate a better divergence result.
-$\theta_1$ was constrained between $15^{\circ}$ and $25^{\circ}$; $\theta_2$ was constrained between $30^{\circ}$ and $60^{\circ}$; $\theta_3$ was constrained between $65^{\circ}$ and $100^{\circ}$.}
+	\includegraphics[width=0.5\textwidth]{corneranglesfreed-1.png}
+	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
 \end{figure}
 
-As can be seen from \figurename{ 4} the parameter space for our angle choice is essentially flat, with no explicit maxima to be found, the sudden increase in our divergence around $\theta_1 = 18^{\circ}$ is likely due to the fact that if we consider the scattering curves for our reference orientations there exist local minima on at roughly $\theta_1 = 13^{\circ}$ and $25^{\circ}$.
-Therefore between these two minima the model struggles to differentiate between reference orientations given that the light intensity is similar for different orientations.
+\figurename{ 4} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
+Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle.
+While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we can't achieve any greater accuracy than dividing the orientation space into 30 reference orientations.
+We hope to develop our model further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.
 
-We then decided to remove the limitations on our choice of angles, the only constraining factor being that the fibre's could not be placed within $10^{\circ}$ of each other.
-
-\begin{figure} [h]
-	\centering
-	\includegraphics[width=0.4\textwidth]{corneranglesfreed-1.png}
-	\caption{Corner plot from ultra-nest, darker regions indicate a better divergence result. Choice of angles is not limited to a predefined region.}
-\end{figure}
-
-Now we see that allowing the angles to be anywhere in the parameter space results in a more focused result for both $\theta_2$ and $\theta_3$ but still a flat parameter space for $\theta_1$.
-The result suggests that the model is more successful when we allow for backscattering to be considered in our model.
-While this works for a perfect system where noise is a known quantity it is unlikely if the signals are clear enough under laboratory conditions.
+As covered in \ref{sec:3.2} our choice for beta can help to improve our estimation somewhat by limiting our estimation in the physical space.
 
@@
 \bibliographystyle{ieeetr}
+\newpage
+
+\section*{Appendix}
+\label{Appendix 1}
+\begin{table}[h]
+	\begin{center}
+		\caption{\label{tab:n} Reference Orientations vector components$^*$}
+		\begin{tabular}{|c|c|c|c|}
+			\hline\hline
+			Reference Orientation & $\hat{n}_x$ &  $\hat{n}_y$ &  $\hat{n}_z$ \\
+			\hline
+			$1$ & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
+			$2$ & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
+			$3$ & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
+			$4$ & $1.0000$ &  $0.00000$ & $0.00000$ \\
+			$5$ & $0.00000$ &  $1.0000$ & $0.00000$ \\
+			$6$ & $0.00000$ &  $0.00000$ & $1.0000$ \\
+			$7$ & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
+			$8$ & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
+			$9$ & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
+			$10$ & $0.00000$ &  $0.00000$ & $-1.0000$ \\
+			$11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
+			$12$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
+			$13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
+			$14$ & $0.00000$ &  $-1.0000$ & $0.00000$ \\
+			$15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
+			$16$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
+			$17$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
+			$18$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
+			$19$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
+			$20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
+			$21$ & $-1.0000$ &  $0.00000$ & $0.00000$ \\
+			$22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
+			$23$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
+			$24$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
+			$25$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
+			$26$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
+			$27$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
+			$28$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
+			$28$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
+			$30$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
+			\hline\hline
+			\multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[h]
+	\begin{center}
+		\caption{\label{tab:n} Raw intensities $I_k$$^*$ and scaled intensities $y_k$}
+		\begin{tabular}{|c|c|c|c|c|c|c|}
+			\hline\hline
+			Reference Orientation & $I_1$ &  $I_2$ &  $I_3$ & $y_1$ & $y_2$ & $y_3$ \\
+			\hline
+			$1$ & $13.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
+			$2$ & $13.237$ & $0.0272$ & $0.0102$ & $0.2429$ & $-1.401$ & $-0.7826$ \\
+			$3$ & $15.840$ & $0.1072$ & $0.0234$ & $1.337$ &$-0.6142$&	$1.162$ \\
+			$4$ & $12.657$ & $0.2246$ & $0.0244$ & $-0.0012$ &	$0.5399$ & $0.9695$ \\
+			$5$ & $15.351$ & $0.0527$ & $0.0221$ & $1.132$ &	$-1.150$ & $1.303$ \\
+			$6$ & $12.046$ & $0.0916$ & $0.0142$ & $-0.2584$ &	$-0.7676$ &	$-0.1886$ \\
+			$7$ & $11.376$ & $0.0036$ & $0.0095$ & $-0.5403$ &$-1.633$ & $-0.8792$ \\
+			$8$ & $13.130$ & $0.2346$ & $0.0095$ & $0.1977$ & $0.6379$ & $-0.8794$ \\
+			$9$ & $15.975$ & $0.2893$ & $0.0273$ & $1.38724$ & $1.175$ & $1.726$ \\
+			$10$ & $10.073$ & $0.2933$ & $0.0221$ & $-1.088$	& $1.215$ & $0.9641$ \\
+			$11$ & $11.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
+			$12$ & $13.238$ & $0.0272$ & $0.0102$ & $0.2429$ &	$-1.401$ & $-0.7826$ \\
+			$13$ & $15.840$ & $0.1072$ & $0.0234$ & $1.337$ &	$-0.6142$ & $1.162$ \\
+			$14$ & $15.351$ & $0.0527$ & $0.0244$ & $1.132$ & $-1.150$ & $1.303$ \\
+			$15$ & $13.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.633$ & $-0.8792$ \\
+			$16$ & $13.130$ & $0.2346$ & $0.0095$ & $0.1977$ & $0.6379$ & $-0.8794$ \\
+			$17$ & $15.957$ & $0.2893$ & $0.0273$ & $1.387$	& $1.175$	& $1.726$ \\
+			$18$ & $14.199$ & $0.0935$ & $0.0076$ & $0.6472$ & $-0.7496$ & $-1.158$ \\
+			$19$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.743$ & $1.411$ & $-1.157$ \\
+			$20$ & $12.961$ & $0.2420$ & $0.0188$ & $0.1267$ & $0.7113$ & $0.4765$ \\
+			$21$ & $8.2160$ & $0.3217$ & $0.0142$ & $-1.869$ & $1.495$ & $-0.1890$ \\
+			$22$ & $14.212$ & $0.2537$ & $0.0102$ & $0.6529$ & $0.8265$ & $-0.7835$ \\
+			$23$ & $8.9280$ & $0.1398$ & $0.0123$ & $-1.570$ &	$-0.2934$ & $-0.4840$ \\
+			$24$ & $13.328$ & $0.1964$ & $0.0234$ & $0.2807$ & $0.2624$ & $1.162$ \\
+			$25$ & $14.199$ & $0.0935$ & $0.0076$ & $0.6472$ & $-0.7496$ & $-1.1580$ \\
+			$26$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.743$	& $1.411$ & $-1.157$ \\
+			$27$ & $12.961$ & $0.2420$ & $0.0188$ & $0.1267$	& $0.7113$ & $0.4765$ \\
+			$28$ & $14.212$ & $0.2537$ & $0.0102$ & $0.6529$ & $0.8265$ & $-0.7835$ \\
+			$29$ & $8.9280$ & $0.1398$ & $0.0122$ & $-1.570$ &	$-0.2934$ & $-0.4840$ \\
+			$30$ & $13.328$ & $0.1964$ & $0.0234$ & $0.2807$ & $0.2624$ & $1.162$ \\
+			\hline\hline
+			\multicolumn{7}{l}{\small *$I_k$ values are calculated using MSTM package.}
+		\end{tabular}
+	\end{center}
+\end{table}
+
 \end{document}

### 2022-11-10 Manuscript.tex
Add files via upload

@@
 Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
+The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
+To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
+For a perfect estimation these two plots should overlap, while  we cannot achieve an exact estimation there are several parameters that will effect our model's efficacy.
 
@@
 \label{sec:3.2}
-The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
-To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
-For a perfect estimation these two plots should overlap.
-Overall their our three main factors that effect the performance of our prediction; the signal error $\sigma$, our choice of angles $\theta_1, \theta_2, \ \& \theta_3$, and our weighting factor $\beta$.
-
-We tested how changing our weighting factor for a set choice of angles would influence the performance of our model.
-We first set our choice of angles to $15^{\circ}, \ 45^{\circ},\ \& \ 90^{\circ}$ and varied $\beta$ between 0.5-1.5.
+When we choose our three detction angles we are also inherently setting the lower limit for our signal error based on the minimum spacing between our $y_k$ values.
+If then signal error is smaller than this minimum distance then every refernce orientation will have the same probability $p(\hat{n} \parallel y_k)$ resulting in our model performing worse than just assigning reference orientations at random.
+We can reduce this minimum distance by creating more reference orientations thus increasing the accuracy of our predictive model.
+However if our signal error is to large then we are limited in how many refernece orienations we can use; futhermore, if the signal error is to large then our prediction can be.
 
@@
 
-As covered in \ref{sec:3.2} our choice for beta can help to improve our estimation somewhat by limiting our estimation in the physical space.
+As covered in \ref{sec:3.2} our choice for beta can help to improve our estimation by restricting it in the physical space. By tuning our choices of $\beta$ in accordance with
 
@@
 We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
-Compared to a random selection of the particle's orientation we find that our model has a $150\%$ increase in it's accuracy.
-Even more interestingly, our model performs better under noisy signal conditions by allowing our estimate to self-correct itself, in the future we hope to remove the artefacts of the signal noise by combing multiple estimates using different detection angles.
-This can be achieved via the inclusion of additional signal detections, we refrained from collecting more signal point for two key reasons; firstly, we wanted to show the extent of the mathematical model in the case of limited detection schemes, and secondly, we wanted to limit our model to what is possible in a real world experimental set up.
-Once we have a clearer understanding of the maximum limit of detection fibres that can be placed we can update our model to account for this and improve the accuracy further.
 

### 2022-11-11 Manuscript.tex
Add files via upload

@@
-%% 
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
+\documentclass[final, 3p]{elsarticle}
 
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-%%
-%% 
-%%
-%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
-%%
-%%
-\documentclass[preprint,3p]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
 \usepackage{amssymb, amsthm}
 \usepackage{amsmath}
-%% The amsthm package provides extended theorem environment
 \usepackage{subcaption}
@@
 \usepackage{fancyref}
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
 \usepackage{lineno}
+\usepackage{ragged2e}
+\usepackage{xcolor}
 
@@
 
-%% Title, authors and addresses
+\title{Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes}
 
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%% \fntext[label3]{}
-
-\title{Investigating the Applications and Limits of Single asymmetric Particle Light Scattering using Limited Detection Schemes}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%Saffran
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-
-\author[aff1]{Dan Maciver\corref{cor1}} %
+\author[aff1]{Dan Maciver\corref{cor1}} 
 \ead{Daniel.Maciver.2016@uni.strath.ac.uk}
@@
 
-\author[aff1]{Mark Haw}
-
 \author[aff1]{Leo Lue}
@@
 \author[aff1]{Jan Sefcik}
+
+\author[aff1]{Mark Haw}
 
@@
 \begin{abstract}
-While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem.
-Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup.
-We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities.
-To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped.
-We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
+\justifying
+While optical trapping is a well understood method for force transduction and detection, \textcolor{blue}{ further} characterisation of trapped entities \textcolor{blue}{with complex structure and dynamics} \textcolor{blue}{ would provide valuable information for example on aggregation processes, microscale biological processes and so on.
+Such characterisation using light scattering from the trapped object} poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem.
+\textcolor{blue} {Experimentally,} combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  \textcolor{blue} {while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised.
+To address these issues} we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, \textcolor{blue}{combined with}  a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem.
+\textcolor{blue}{As an example problem} we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal.
+We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
 \end{abstract}
-%%Research highlights
+
 \begin{highlights}
-\item Asymmetric dimers of undergo a full inversion in the presence of an potential well
-\item Orientation can be classified by applying Bayesian Statistics to light intensity readings
-\item Bayesian model shows increased accuracy when processing noisy signal measurements. 
-\item Multiple estimations can be  used to cancels out artefacts of noise. 
+\item Asymmetric dimers undergo a full inversion in the presence of an potential well. This only occurs if the size ratio is greater than 1:2.  
+\item Orientation can be classified by applying Bayesian Statistics to light intensity readings. 
+\item Choice of detection angles sets a lower limit on possible signal error. 
+\item Future applications in continuos processes where dimer shape is changing with time.   
 \end{highlights}
@@
 
-\linenumbers
+%\linenumbers
 
-%% main text
 \section{Introduction}
@@
 
-\indent Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few piconewton.
-Recent applications of the Optical Tweezers in metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical} demand the tweezer to also be able to characterise  the trapped entity.
-To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman} and dynamical characterisation has been achieved mainly by following the centre-of-mass Brownian motion of the trapped entity using a Quadrant Photo Detector \cite{friedrich2012tuning}.
-\begin{wrapfigure}[12]{r}{0.3\textwidth}
+Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons.
+The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
+Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.
+
+The possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and its Brownian motion was studied.
+While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
+Furthermore, this method cannot be extended for trapped entities that would scatter light non isotopically if their motion is large compared to the dimensions of the trap focus.
+In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.
+
+To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity with more complex dynamics and scattering than a single sphere, i.e.
+a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity.
+There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
+Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation form the light scattering signals given a choice of scattering measurement angles.
+
+\textcolor{blue} {not very happy with the above but I am trying to set out clearly in advance the narrative of what we are doing, starting with the general problem, moving to the specific example we are studying, and introducing how we then go about it.
+we need to emphasise how oreitnation of a dimer is our choice of example problem, where a dimer is a paradigm of a 'complex object' and its orientation is a paradigm of 'data' that in some way describes it}
+
+\begin{figure}
 	\centering
-	\includegraphics[width=0.3\textwidth]{fig1a}
-	\caption{Experimental set up}
-	\label{fig:1}
-	\vspace{-100pt}
-\end{wrapfigure}
-While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision \cite{friedrich2012tuning} and  rotation \cite{yifat2021facile} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
-Possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's Brownian motion was studied.
-While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
-\newpage
-In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:1}.
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
+		\subcaption{}
+	\end{subfigure}
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
+		\subcaption{}
+	\end{subfigure}
+	\caption{Experimental set up, a) Side view, b) Top view }
+	\label{fig:setup}
+\end{figure}
 
-\section{Theory}
-\label{sec:Theory}
-
-Consider a dimer with unequal sphere diameters $a_1, a_2$ in an optical trap in some orientation $\hat{s}$.
-Due to the Brownian motion from the surrounding fluid the dimer's orientation changes with time.
-Assuming we cannot accurately determine the dimer's orientation from imaging alone we instead rely on it's light scattering.
-Rather than directly calculate the orientation vector we instead discretise the orientation space into 30 reference orientations.
-See \ref{Appendix 1} for the exact numerical values of reference orientations.
-The following records how we apply a Bayesian model to determine the estimate the closest reference orientation based on an instantaneous light scattering measurement.
-
-We use a Fortran package specialised for calculating non spherical t-matrices (MSTM) \cite{Mishchenko1996MSTM} to determine the light scattering intensity of our dimer in each reference orientation.
-The raw signal intensities for at 3 different angles ($\theta_1, \ \theta_2, \ \& \ \theta_3$) are calculated and then scaled down so that the signal response $y_k$ is normalised for each angle $k$:
-
-\begin{align}
-	y_k(\hat{n}) = 
-	 \frac{\left[I(\hat{n}, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
-	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
-\end{align}
-
-See \ref{Appendix 1} for a breakdown of the raw intensities and scaled signal responses.
-To classify the particle's orientation we compare the produced light scattering signal to a collection of reference orientations that are evenly spaced around the orientation space.
-We assume that the for each detection fibre the intensities can be distributed on a Gaussian curve centred at $y_k$, the normalised intensity at detection fibre k.
-Based on the light scattering from the dimer we can assign a probability that the produced light scattering pattern belongs to the reference orientation $\hat{n}$
-
-\begin{align}
-	p(y(\hat{s})\parallel\hat{n}) =& \Pi^3_{k=1}
-	(2\pi\sigma^2)^{-1/2} \nonumber \\     &e^{-(y(\hat{n})_k-y(\hat{s})_k)^2/2\sigma^2}
-\end{align}
-
-Where $\sigma$ is the minimum distance between our reference orientations signals ($y_1, \ y_2, \ \& \ y_3$), .
-We can view this result as a conditional probability, if the orientation is determined ($\hat{n}$) then what is the probability that expected scattering signal matches our dimer's scattering signal.
-We instead want to know the inverse conditional, that with a given signal y the dimer was in orientation $\hat{n}$.
-We can calculate this using Bayes' theory:
-
-\begin{align}
-	p(\hat{n}\parallel y(\hat{s}))&= \frac{p(y(\hat{s})\parallel\hat{n})p(\hat{n})}{p(y)}
-\end{align}
-
-Where $p(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and scattering signals respectively.
-The former can be estimated by some Boltzmann distribution based on the spacing between reference orientations.
-The latter priori $p(y)$ is the probability of our dimer producing the signal $y$ and is given as.
-
-\begin{align}
-	p(y) = \int p(y(\hat{s})\parallel \hat{n}) p(\hat{n}) d\hat{n}
-\end{align}
-
-The inclusion of $p(y)$ normalises our probability distribution. In order to test the accuracy of our model we first test it using an idealised simulation of the target dimer.
-
+\section{Methodology}
+\label{sec:Method}
 \subsection{Brownian Simulation}
 \label{sec:2.1}
-We use the Brownian OT package by Jerome Fung \cite{Vigilante2020Brownian_OT} to simulate how our target dimer moves while subject to an optical trap.
-The package is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) to simulate the motion of arbitrary shaped particles.
+We use the Brownian OT package developed by Jerome Fung \textit{et al} \cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap.
+Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters.
+We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m, \ \& \ 0.5 \ \mu m$ respectively.
+We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
 
-For any trapped particle their are two major sources of motion, Brownian motion and optical forces.
-The Brownian displacement ($\Delta q^{\textbf{B}}$) is approximated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
+With any trapped particle there are two main sources of motion, Brownian and optical forces.
+The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
 The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}.
-
-\begin{align}
-	\langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle &= 2D_{ij} \Delta t 
-\end{align}
-
 For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
@@
 MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
-The benefit of this approach is that MSTM can accurately compute the T-matrix for the dimer once and then \textit{ott} can calculate the optical force after each time step.
-The dimer displacement is the sum of both the optical and Brownian contributions.
-Brownian OT returns the displacement as a trajectory file; with the [x,y,z] position being relative to the trap focus, and the orientation being returned as a quaternion.
+The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once and then \textit{ott} can calculate the optical forces and toques.
+The total displacement is the sum of the optical and Brownian contributions.
 
-For our model we simulate the motion of silica dimer trapped in a 5mW Gaussian beam; the dimer is comprised of two spheres with radii $1\\mu m, \ \& 0.5 \ \mu m$ respectively.
+\begin{align}
+	\langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle = 2D_{ij} \Delta t \\
+	\Delta q^{opt} = \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F} \\
+	\Delta q^{tot} = \Delta q^{\textbf{B}} + \Delta q^{opt}
+\end{align}
+
+Where \textbf{F} is the a 6 component vector that describes the optical force and torque components.
+Brownian OT returns the dimers motion as a \textit{csv} file; with the $[x,y,z]$ position being relative to the trap focus, and the orientation being returned as a quaternion.
+
+\subsection{Theory}
+\label{sec:2.2}
+
+If we consider our simulated dimer from before; as it moves within the trap its instantaneous orientation can be represented as a three component unit vector, denoted as $\hat{s}$, one that points from the centre of the larger sphere, to the centre of the smaller sphere.
+If we shine a plane wave laser, perpendicular to the trapping laser, at the dimer a scattering pattern will be produced that is dependent upon the dimer's instantaneous orientation, denoted as $I(\hat{s},\ \theta)$.
+This scattering pattern was computed using MSTM for each time step of our Brownian simulation.
+From the scattering pattern, three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the intensity at each angle k was recorded as $I(\hat{s},\ \theta_k)$.
+
+Instead of trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}.
+Using MSTM we compute the scattering patterns produced by our dimer in that orientation, denoted as $I(\hat{n}_i, \ \theta)$.
+We use the same choice of angles to record $I(\hat{n}_i, \ \theta_k)$.
+These raw intensities can vary by several orders of magnitude across the entire range of possible angles, to correct this each intensity is scaled according to \eqref{eq:scale}:
+
+\begin{align}
+	\label{eq:scale}
+	y_k(\hat{n}_i,\ \theta_k) = 
+	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}_i, \ \theta_k) \rangle \right]} 
+	{\langle I^2(\hat{n}_i,\ \theta_k) \rangle -\langle I(\hat{n}_i, \ \theta_k)\rangle^2}
+\end{align}
+
+Where the denominator is simply the standard deviation of $I(\hat{n}_i, \theta_k)$; our dimer's instantaneous intensity is also scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$.
+See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals.
+In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty associated with it.
+We compare our instantaneous signal to our reference signal by assuming that our values of $y(\hat{n}_i,\ \theta_k)$ are normally disturbed at each angle k.
+Combining all three detection angles provides us with the conditional probability that our dimer in orientation $\hat{n}_i$ will produce a signal $y(\hat{s})$:
+
+\begin{align}
+	\label{eq:conditional}
+	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
+	(2\pi\sigma&_k^2)^{-1/2}  \nonumber \\
+	&e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
+\end{align}
+
+Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
+The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.
+
+\begin{eqnarray*}
+	\sigma_k = \epsilon\bar{y}(\hat{n}_k) \\
+	\Rightarrow y(\hat{s}_k) = y(\hat{s})_k \pm n\sigma_k \\ 
+	n \in \mathbb{R}, n\in[0,1]
+\end{eqnarray*}
+
+Where $\epsilon$ is the percent error expected from our detection fibres.
+Equation~\eqref{eq:conditional} is assigning a probability to each reference orientation based on how similar it's signal is to the signal our dimer produced.
+If however, our reference signals are similar to one another it can be difficult to parse the correct reference orientation.
+Bayes Theorem was utilised to update our conditional probability by accounting for the priori distribution of reference orientations and instantaneous signals.
+
+\begin{align}
+	\label{eq:bayes}
+	p_i(\hat{n}\parallel y(\hat{s}))&= \frac{p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})}{p(y)}
+\end{align}
+
+Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and instantaneous signals respectively.
+The former can be estimated by some Boltzmann distribution based on the physical spacing between reference orientations, covered in Section~\ref{sec:2.3}.
+The latter priori $p(y)$ is the probability of our dimer producing the signal $y(\hat{s})$ and is given as.
+
+\begin{align}
+	p(y) = \sum\limits_{i=1}^{30} p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})
+ \end{align}
+
+The implementation of \eqref{eq:bayes} improves our estimate by now accounting for how our reference orientations are distributed in physical space.
+In order to evaluate the accuracy of our model we tested it using our Brownian simulation of the dimer, see Section~\ref{sec:2.1}.
 
 \subsection{Testing our Model}
-\label{sec:2.2}
-
-Since MSTM has computed the T-matrix for our dimer, we can reuse it to calculate.
-We then apply \eqref{3} - covered in Section~\ref{sec:2.1} - to estimate the dimer orientation based on its scattering pattern.
-The priori estimate $p(\hat{n})$ is defined as a Boltzmann's distribution based on the normalised dot product of our previous estimation $\hat{n}_{t-\Delta t}$ and each reference orientation $\hat{n}_t$:
+\label{sec:2.3}
+Using our simulation from Section~\ref{sec:2.1} we calculated the instantaneous signal from our dimer at each time step using MSTM.
+We applied \eqref{eq:bayes} - covered in Section~\ref{sec:2.2} - and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
+The priori of reference orientations $p_i(\hat{n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{n}$
 
 \begin{align}
-	r_j(t, \hat{n})= \frac{1}{2}((\hat{n}_{t} \cdot \hat{n}_{t-\Delta t})-1) \\
-	p(\hat{n})= \frac{e^{\beta (r_j(t,\hat{n})-1)}}
-	{\Sigma_{i=1}^{30}e^{\beta(r_i(t, \hat{n})-1)}}
+	r_i(t, \hat{n}_i)= \frac{1}{2}((\hat{n}_{i,\ t} \cdot \hat{n}_{t-\Delta t})-1) \\
+	p_i(\hat{n})= \frac{e^{\beta (r_i(t,\hat{n}_i)-1)}}
+	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{n}_i)-1)}}
 \end{align}
@@
 Where $\beta$ is a weighting factor that controls the significance of our priori estimate.
-Since we are testing with a simulation we already know the exact orientation of the dimer, we want to know how often our model can determine the closest reference orientation $\hat{n}$ to the true orientation $\hat{s}$, and how precise our prediction is.
-For a perfect model our probability distribution should equal 1 at the correct result and zero elsewhere.
-In reality our model will be some probability distribution that has some probability of any orientation being correct.
-To evaluate our model's accuracy and confidence we calculate the Kullback-Leibler divergence of the two probability distributions.
+The simulation data from Section~\ref{sec:2.1} was used to check if the estimation was correct by comparing the estimate provided by \eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as $\hat{n}_{best}$.
+To measure the error in the model's estimate the distribution's of $p(\hat{n}\parallel y(\hat{s}))$ and an ideal result, $p_ideal$, were compared.
+An ideal distribution is one that is 0 for every $\hat{n}$ apart from the correct result.
 
 \begin{align}
-	K_{l, \#}(p_{guess} \parallel p_{ideal}) &= 
-	p_{ideal} \nonumber \\ 
-	&\ln \left[\frac{p_{ideal}}{p(\hat{n}\parallel y_k(\hat{s}))}
+	p_{ideal} = 
+	\begin{cases}
+		1 & \text{when $\hat{n}_i$ = $\hat{n}_{best}$}\\
+		0 & \text{anywhere else}
+	\end{cases}
+\end{align}
+
+The distribution from \eqref{eq:bayes} will assign some non-zero probability to every reference orientation.
+The model's error was measured by calculating the Kullback-Leibler ($K_l$) divergence between the two distributions
+
+\begin{align}
+	K_{l, \#}(p_i(\hat{n}\parallel y_k(\hat{s})) &\parallel p_{ideal}) =\nonumber \\
+	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{n}_{best}\parallel y(\hat{s}))}
 	\right]
@@
 
-Where the divergence goes to zero if $p_{ideal}=0$, for a single estimation this gives us an idea of how confident our model is in it's estimation.
-We sum the divergence of each measurement across the entire simulation to get an idea of how well the model performs across the entire simulation.
-To measure how much our result had improved we compare it to a worse case scenario, where every $\hat{n}$ is given uniform probability, and evaluate the improvement factor $F(K_l)$
+Where a more negative value of $K_l$ indicates a poorer estimation of the dimer's instantaneous orientation.
+By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
+To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
 
@@
 
-The worst case scenario can be akin to randomly choosing a reference orientation at each time step.
-Because our model is dependent on several parameters at once we need to a more sophisticated method for finding the optimum choice of parameters.
+The worst case scenario is akin to randomly choosing a reference orientation at each time step.
+The greater the value of $F(K_l)$ the lower the error between our estimated orientation and the dimer's instantaneous orientation.
+Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.
 
-\subparagraph{Optimization of parameters}
-\label{2.3}
-The efficacy of our estimation is a function of our detection angles, signal error, and weighting of $p(\hat{n})$.
-Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra nest to apply a Monte Carlo simulation to our model.
-Ultra-nest is a sophisticated python package for fitting models with complex parameter interactions \cite{Buchner2016Ultranest}.
-Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where the likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}, \ 0 \leq \epsilon \leq 1, \ 0 \leq \beta \leq 5$).
+\subsection{Optimization of parameters}
+\label{sec:2.4}
+The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly.
+However, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo simulation to our model.
+Ultra-nest is a sophisticated python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}.
+Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
 
-We used the Ultra-nest software to create a set of live points of different choices of $\theta_1, \ \theta_2, \ \& \ \theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$) produced.
+The Ultra-nest software was used to create a set of live points of different choices of $\theta_1, \ \theta_2, \ \& \ \theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$).
 After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
-This results in a convergence on parameter choices that result in an average divergence result of almost 0.
-A common issue with complex parameter spaces, such as ours, is the issue of degeneracies (where two or more parameters have identical parameter spaces), to combat this we implemented a random step sampler that selects a new live point some 'n' steps away in the parameter space.
+If there is a correlation between the choice of angles and $F(K_l)$ then the live points should congregate around a single (or multiple) choice of angles that will maximise $F(K_l)$ and therefore minimise the error between our estimate and instantaneous orientation.
 
-\section{Results}
-\label{3}
+\section{Discussion}
+\label{sec:Discussion}
 \subsection{Asymmetric dimer dynamics}
-\label{3.1}
+\label{sec:3.1}
 The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap.
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 	\centering
 	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width =\textwidth]{./Images/traj.png}
 		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/traj.png}
 	\end{subfigure}
 	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth]{./Images/pos.png}
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/pos.png}
 	\end{subfigure}
 	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
 \end{figure}
 
-As can be seen from \figurename{ 2}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
+As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
 Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
@@
 To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
-For a perfect estimation these two plots should overlap, while  we cannot achieve an exact estimation there are several parameters that will effect our model's efficacy.
+For a perfect estimation these two plots should overlap, while we cannot achieve an exact estimation we can optimise it by adjusting the expected signal error and weighting factor.
 
-\newpage
-\subsection{Predictive efficacy of model}
+\subsection{Impact of signal error on model predictions}
 \label{sec:3.2}
-When we choose our three detction angles we are also inherently setting the lower limit for our signal error based on the minimum spacing between our $y_k$ values.
-If then signal error is smaller than this minimum distance then every refernce orientation will have the same probability $p(\hat{n} \parallel y_k)$ resulting in our model performing worse than just assigning reference orientations at random.
-We can reduce this minimum distance by creating more reference orientations thus increasing the accuracy of our predictive model.
-However if our signal error is to large then we are limited in how many refernece orienations we can use; futhermore, if the signal error is to large then our prediction can be.
+Signal error is an important parameter to control when it comes to understanding how well our model can predict the dimers motion.
+To this end we ran our model while varying the value of $\epsilon$.
+\figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1.
+The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation
 
-\begin{figure}[b]
-	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/154590_Beta=1.png}
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_5.png}
 	\end{subfigure}
-	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/154590_Beta=05.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_15.png}
 	\end{subfigure}
-	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/154590_Beta=15.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_25.png}
 	\end{subfigure}
+	\caption{Model prediction for signal error of a) $5\%$, $15\%$, and $25\%$}
+	\label{fig:epsilon}
 \end{figure}
 
-\newpage
-\subsection{Ultranest Results}
-\label{3.3}
+As can be seen from Figure~\ref{fig:epsilon} there is seemingly a sweet spot for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an alright estimation but with sudden shifts away from the ideal result.
+The reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.
+As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
+This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
+In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.
+
+The latter issue of a high signal error can be addressed somewhat by adjusting our value of $\beta$ to flatten our probability distribution.
+We repeated our model estimation at $25\%$ at different values of $\beta$ to see the upper limit we could improve our result to.
+
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\textwidth]{./Images/beta_5.png}
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\textwidth]{./Images/beta_6.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
+	\label{fig:beta}
+\end{figure}
+
+As shown by \ref{fig:beta} eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
+The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error.
+In the future we want to also investigate the possibility of extending our priori estimation so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.
+
+\subsection{Ultra-nest Results}
+\label{sec:3.3}
 We wanted to see if our choice of angles plays a significant role in our model's efficacy.
-We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
+We employed ultra-nest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
 We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{corneranglesfreed-1.png}
+	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
 	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
+	\label{fig:corner}
 \end{figure}
 
-\figurename{ 4} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
+Figure~\ref{fig:corner} is the resulting corner plot from our ultra-nest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
 Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle.
@@
 
-As covered in \ref{sec:3.2} our choice for beta can help to improve our estimation by restricting it in the physical space. By tuning our choices of $\beta$ in accordance with
+As covered in Section~\ref{sec:3.2} our choice for beta can help to improve our estimation by restricting it in the physical space.
+In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
 
@@
 We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
+In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
+The MSTM package is flexible in calculating the light scattering for a variety of micro particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
+This has potential to be applied to particles that are changing while in the optical trap.
+We hope to apply this work in the future to  characterise processes such as nucleation within an optical trap.\\
 
-Our simulations of asymmetric dimers found that for large sphere ratios the dimer will prefer to be orientated against the beam propagation and will even move further from the trapping focus to invert itself before coming back to the laser focus.
-It is interesting that is only the case for dimers where the size ratio is greater than 1:2, understanding how the optical torque varies with size ratio can provide unique insight into the dynamics of non-spherical trapping targets.
+\noindent \textbf{Funding.} \\
 
-Overall the model developed can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
-The MSTM package is flexible in  calculating the light scattering for a variety of micro particles, so long as we can calculate the light scattering we can apply our model to help characterise the trapped model.
-We hope to extend this work for characterising continuos processes such as nucleation within an optical trap.
+\noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\
 
-The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.
+\noindent \textbf{Disclosures.} The authors declare no conflict of interest. \\
 
@@
 \bibliographystyle{ieeetr}
+
 \newpage
-
+\appendix
+\onecolumn
 \section*{Appendix}
-\label{Appendix 1}
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\label{sec:App1}
 \begin{table}[h]
+	\label{tab:A1}
 	\begin{center}
-		\caption{\label{tab:n} Reference Orientations vector components$^*$}
+		\caption{Reference Orientations vector components$^*$}
 		\begin{tabular}{|c|c|c|c|}
 			\hline\hline
-			Reference Orientation & $\hat{n}_x$ &  $\hat{n}_y$ &  $\hat{n}_z$ \\
+			Reference Orientation i & $\hat{n}_{i,\ x}$ &  $\hat{n}_{i,\ y}$ &  $\hat{n}_{i,\ z}$ \\
 			\hline
@@
 
+\newpage
 \begin{table}[h]
+	\label{tab:A2}
 	\begin{center}
-		\caption{\label{tab:n} Raw intensities $I_k$$^*$ and scaled intensities $y_k$}
+		\caption{Raw intensities $I_k$$^*$ and scaled intensities $y_k$}
 		\begin{tabular}{|c|c|c|c|c|c|c|}
 			\hline\hline
-			Reference Orientation & $I_1$ &  $I_2$ &  $I_3$ & $y_1$ & $y_2$ & $y_3$ \\
+			Reference Orientation i  & $I(\hat{n}_i,\ 15^{\circ})$ &  $I(\hat{n}_i, \ 45^{\circ})$ &  $I(\hat{n}_i,\ 90^{\circ})$ & $y(\hat{n}_i, \ 15^{\circ})$ & $y(\hat{n}_i, \ 45^{\circ})$ & $y(\hat{n}_i, \ 90^{\circ})$ \\
 			\hline

### 2022-11-14 Manuscript.tex
Add files via upload

@@
 
-\title{Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes}
+\title{Investigations on Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles}
 
@@
             city={Glasgow},
-            postcode={G1 1XL}, 
+            postcode={G1 1XJ}, 
             country={Scotland}}
@@
 \begin{abstract}
+	\label{sec:Abstract}
 \justifying
-While optical trapping is a well understood method for force transduction and detection, \textcolor{blue}{ further} characterisation of trapped entities \textcolor{blue}{with complex structure and dynamics} \textcolor{blue}{ would provide valuable information for example on aggregation processes, microscale biological processes and so on.
-Such characterisation using light scattering from the trapped object} poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem.
-\textcolor{blue} {Experimentally,} combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  \textcolor{blue} {while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised.
-To address these issues} we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, \textcolor{blue}{combined with}  a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem.
-\textcolor{blue}{As an example problem} we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal.
+While optical trapping is a well understood method for force transduction and detection, further characterisation of trapped entities with complex structure and dynamics would provide valuable information for example on aggregation processes, micro-scale biological processes and so on.
+Such characterisation using light scattering from the trapped object poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem.
+Experimentally, combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised.
+To address these issues we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, combined with a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem.
+As an example problem we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal.
 We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
@@
 \item Choice of detection angles sets a lower limit on possible signal error. 
-\item Future applications in continuos processes where dimer shape is changing with time.   
+\item Method can be reworked to determine other characteristics of trapped entities. 
 \end{highlights}
@@
 \begin{keyword}
-	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics \sep Theoretical 
+	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
 \end{keyword}
@@
 
-Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons.
+Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico newtons.
 The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
@@
 a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity.
-There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
-Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation form the light scattering signals given a choice of scattering measurement angles.
-
-\textcolor{blue} {not very happy with the above but I am trying to set out clearly in advance the narrative of what we are doing, starting with the general problem, moving to the specific example we are studying, and introducing how we then go about it.
-we need to emphasise how oreitnation of a dimer is our choice of example problem, where a dimer is a paradigm of a 'complex object' and its orientation is a paradigm of 'data' that in some way describes it}
+There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers.
+Here, we simulate the Brownian motion of an asymmetric dimer, calculate its light scattering pattern, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
 
@@
 We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
+	\caption{Example dimer in orientation $\hat{s}$}
+	\label{fig:dimer}
+\end{figure}
 
@@
 \label{sec:3.2}
-Signal error is an important parameter to control when it comes to understanding how well our model can predict the dimers motion.
-To this end we ran our model while varying the value of $\epsilon$.
-\figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1.
-The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation
+A key assumption of our Bayesian model is that we assume our detected light signal will inherently have some degree of noise.
+This is the central premise that allows us to form equation~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
+To this end we ran our model while varying the value of $\epsilon$; \figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1.
+The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know the instantaneous orientation $\hat{s}$ is between two reference orientation.
 
@@
 
-As can be seen from Figure~\ref{fig:epsilon} there is seemingly a sweet spot for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an alright estimation but with sudden shifts away from the ideal result.
-The reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.
+As can be seen from Figure~\ref{fig:epsilon} there is an optimal point for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an good estimation but with periodic spikes that indicate a very poor classification of our dimer's orientation.
+The reason why our low error estimation performed so poorly is that the minimum difference between our reference signals $y_k(\hat{n})$ is larger than our signal error.
 As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
-This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
-In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.
+For our particular choice of angles the minimum difference was around 0.1734 which works out to around a $15\%$ signal error, lowering the signal error any more than this results in the model's performance dropping substantially.
+This minimum difference is predetermined by our choice of angles and sets a lower limit of how accurate our signals can be.
+In the future we hope to look at increasing the number of reference orientations to see if this could allow us to achieve a higher degree of accuracy in our estimation of the instantaneous orientation.
 
@@
 	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
-	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
+	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement. Contour ranges from $F(K_l) = 1.7 \ \text{to} \ F(K_l)  =  0.1$ }
 	\label{fig:corner}
@@
 In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
-The MSTM package is flexible in calculating the light scattering for a variety of micro particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
-This has potential to be applied to particles that are changing while in the optical trap.
-We hope to apply this work in the future to  characterise processes such as nucleation within an optical trap.\\
+The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
+This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
+The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust
 

### 2022-11-14 Manuscript.tex
Add files via upload

--- 2022-11-11 Manuscript.tex
@@
 \usepackage{lineno}
-\usepackage{ragged2e}
-\usepackage{xcolor}
+\usepackage{color}
+\usepackage{nomecl}
 
@@
 \justifying
-While optical trapping is a well understood method for force transduction and detection, \textcolor{blue}{ further} characterisation of trapped entities \textcolor{blue}{with complex structure and dynamics} \textcolor{blue}{ would provide valuable information for example on aggregation processes, microscale biological processes and so on.
-Such characterisation using light scattering from the trapped object} poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem.
-\textcolor{blue} {Experimentally,} combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  \textcolor{blue} {while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised.
-To address these issues} we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, \textcolor{blue}{combined with}  a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem.
-\textcolor{blue}{As an example problem} we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal.
-We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
+While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem.
+Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup.
+We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities.
+To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped.
+We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
 \end{abstract}
@@
 
+\nomenclature{$\hat{s}$}{Dimer's instantaneous orientation}
+\nomenclature{$\hat{n}_i$}{Reference Orientation i}
+
 %\linenumbers
@@
 Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons.
-The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
+\textit{Further characterisation of the trapped entity would provide useful information, as explored in for example} metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
 Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.
@@
 While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
-Furthermore, this method cannot be extended for trapped entities that would scatter light non isotopically if their motion is large compared to the dimensions of the trap focus.
 In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.
 
-To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity with more complex dynamics and scattering than a single sphere, i.e.
+To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity more complex than a single sphere, i.e.
 a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity.
 There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
-Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation form the light scattering signals given a choice of scattering measurement angles.
-
-\textcolor{blue} {not very happy with the above but I am trying to set out clearly in advance the narrative of what we are doing, starting with the general problem, moving to the specific example we are studying, and introducing how we then go about it.
-we need to emphasise how oreitnation of a dimer is our choice of example problem, where a dimer is a paradigm of a 'complex object' and its orientation is a paradigm of 'data' that in some way describes it}
+Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
 
@@
 Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters.
-We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m, \ \& \ 0.5 \ \mu m$ respectively.
+We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m$ and $ 0.5  \mu m$ respectively.
 We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
+	\caption{Example dimer in orientation $\hat{s}$}
+	\label{fig:dimer}
+\end{figure}
 
@@
 For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
-The T-matirx describes the scattering process by relating the scattering wave to the incident wave in an elegant equation.
+The T-matrix describes the scattering process by relating the scattering wave to the incident wave:
 
@@
 MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
-The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once and then \textit{ott} can calculate the optical forces and toques.
-The total displacement is the sum of the optical and Brownian contributions.
+The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques.
+The total displacement is the sum of the optical and Brownian contributions:
 
@@
 
-Where \textbf{F} is the a 6 component vector that describes the optical force and torque components.
-Brownian OT returns the dimers motion as a \textit{csv} file; with the $[x,y,z]$ position being relative to the trap focus, and the orientation being returned as a quaternion.
+where \textbf{F} is the a 6 component vector that describes the optical force and torque components.
+Brownian OT returns the dimer's motion as a \textit{csv} file containing for each timestep the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.
 
-\subsection{Theory}
+\subsection{Orientation estimation from scattering measurements}
 \label{sec:2.2}
 
-If we consider our simulated dimer from before; as it moves within the trap its instantaneous orientation can be represented as a three component unit vector, denoted as $\hat{s}$, one that points from the centre of the larger sphere, to the centre of the smaller sphere.
-If we shine a plane wave laser, perpendicular to the trapping laser, at the dimer a scattering pattern will be produced that is dependent upon the dimer's instantaneous orientation, denoted as $I(\hat{s},\ \theta)$.
+We represent the instantaneous orientation of the moving dimer in the trap as a three component unit vector,  $\hat{s}$, pointing from the centre of the larger sphere to the centre of the smaller sphere.
+A plane wave 'probe' laser beam directed perpendicular to the trapping laser, incident on the dimer, generates a scattering pattern dependent on the dimer's instantaneous orientation,  $I(\hat{s},\ \theta)$.
 This scattering pattern was computed using MSTM for each time step of our Brownian simulation.
-From the scattering pattern, three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the intensity at each angle k was recorded as $I(\hat{s},\ \theta_k)$.
+To represent a simulated 'measurement', three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the calculated  intensity at each angle k recorded as $I(\hat{s},\ \theta_k)$.
 
-Instead of trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}.
-Using MSTM we compute the scattering patterns produced by our dimer in that orientation, denoted as $I(\hat{n}_i, \ \theta)$.
-We use the same choice of angles to record $I(\hat{n}_i, \ \theta_k)$.
-These raw intensities can vary by several orders of magnitude across the entire range of possible angles, to correct this each intensity is scaled according to \eqref{eq:scale}:
+Our goal is to estimate the dimer orientation $\hat{s}$ over time, based on the scattering measurements  $I(\hat{s},\ \theta_k)$: the question then is what set of angles ($\theta_1,\ \theta_2,\ \& \theta_3$) give us the most useful information on $\hat{s}$.
+To initially simplify the problem, rather than trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}: our question becomes, how can we optimise the experiment to identify the reference orientation $\hat{n}_i$ closest to the real orientation $\hat{s}$.
+(Ultimately our method could be extended to an arbitrarily large number of reference orientations but for clarity and practicability we use a relatively small number here; additionally, the choice of a small number of discrete reference or 'candidate' orientations also accounts for the fact that any experiment will necessarily have a limited precision.)   Using MSTM we compute the scattering that would be produced by the dimer if it were in each of the reference orientations, denoted as $I(\hat{n}_i, \ \theta)$ and obtain 'measurements'  $I(\hat{n}_i, \ \theta_k)$ at our chosen set of  angles $\theta_k$.
+These raw intensities are normalised according to \eqref{eq:scale}:
 
@@
 	y_k(\hat{n}_i,\ \theta_k) = 
-	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}_i, \ \theta_k) \rangle \right]} 
-	{\langle I^2(\hat{n}_i,\ \theta_k) \rangle -\langle I(\hat{n}_i, \ \theta_k)\rangle^2}
+	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
+	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
 \end{align}
 
-Where the denominator is simply the standard deviation of $I(\hat{n}_i, \theta_k)$; our dimer's instantaneous intensity is also scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$.
+Where the denominator is simply the standard deviation for our values of $I(\hat{n}, \theta_k)$.
+The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\hat{s}$  is similarly scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$.
 See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals.
-In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty associated with it.
-We compare our instantaneous signal to our reference signal by assuming that our values of $y(\hat{n}_i,\ \theta_k)$ are normally disturbed at each angle k.
-Combining all three detection angles provides us with the conditional probability that our dimer in orientation $\hat{n}_i$ will produce a signal $y(\hat{s})$:
+In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
+We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\hat{s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{n}_i$, it would produce a signal $y(\hat{s})$:
 
@@
 	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
-	(2\pi\sigma&_k^2)^{-1/2}  \nonumber \\
-	&e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
+	(2\pi\sigma_k^2)^{-1/2} 
+	e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
 \end{align}
 
-Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
+Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\hat{s})$.
+We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
 The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.
@@
 Where $\epsilon$ is the percent error expected from our detection fibres.
-Equation~\eqref{eq:conditional} is assigning a probability to each reference orientation based on how similar it's signal is to the signal our dimer produced.
-If however, our reference signals are similar to one another it can be difficult to parse the correct reference orientation.
-Bayes Theorem was utilised to update our conditional probability by accounting for the priori distribution of reference orientations and instantaneous signals.
+Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{n}_i)$ is to the actual signal $y_{k}(\hat{s})$ generated by the dimer in its real orientation $\hat{s}$.
+However, if our values of $y_k(\hat{n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations.
+Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement.
+We apply this by accounting for our distribution of reference orientations and instantaneous signals.
 
@@
 
-Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and instantaneous signals respectively.
-The former can be estimated by some Boltzmann distribution based on the physical spacing between reference orientations, covered in Section~\ref{sec:2.3}.
-The latter priori $p(y)$ is the probability of our dimer producing the signal $y(\hat{s})$ and is given as.
+Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively.
+The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}.
+The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\hat{s})$ that falls with our distribution of reference signals $y_k(\hat{n})$.
+This is given by taking the discrete integral over collection of reference orientations.
 
@@
 
-The implementation of \eqref{eq:bayes} improves our estimate by now accounting for how our reference orientations are distributed in physical space.
+The implementation of \eqref{eq:bayes} improves our estimate by suppressing reference orientations that are far away from our dimer's orientation vector in physical space.
 In order to evaluate the accuracy of our model we tested it using our Brownian simulation of the dimer, see Section~\ref{sec:2.1}.
@@
 
-\section{Discussion}
+\section{Results $\&$ Discussion}
 \label{sec:Discussion}
@@
 \begin{figure}[h]
+	\label{fig:motion}
 	\centering
 	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
 		\includegraphics[width =\textwidth]{./Images/traj.png}
-		\subcaption{}
 	\end{subfigure}
 	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/pos.png}
-		\subcaption{}
 	\end{subfigure}
 	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
-	\label{fig:motion}
 \end{figure}
@@
 \label{sec:3.2}
-Signal error is an important parameter to control when it comes to understanding how well our model can predict the dimers motion.
+A key assumption of equation~\ref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise.
+This is the central premise that allows us to form equation~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
 To this end we ran our model while varying the value of $\epsilon$.
@@
 \begin{figure}[h]
+
 	\centering
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.49\textwidth}
 		\includegraphics[width=\textwidth]{./Images/epsilon_5.png}
 	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.49\textwidth}
 		\includegraphics[width=\textwidth]{./Images/epsilon_15.png}
 	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.5\textwidth}
 		\includegraphics[width=\textwidth]{./Images/epsilon_25.png}
 	\end{subfigure}
-	\caption{Model prediction for signal error of a) $5\%$, $15\%$, and $25\%$}
+	\caption{Model prediction for signal error of a) $5\%$, b) $15\%$, and c) $25\%$. Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}}
 	\label{fig:epsilon}
@@
 
-\begin{figure}[h]
+\begin{figure}
+	\label{fig:beta}
 	\centering
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.49\textwidth}
 		\includegraphics[width=\textwidth]{./Images/beta_5.png}
 	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.49\textwidth}
 		\includegraphics[width=\textwidth]{./Images/beta_6.png}
@@
 	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
-	\label{fig:beta}
 \end{figure}
@@
 The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error.
-In the future we want to also investigate the possibility of extending our priori estimation so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.
+In the future we want to also investigate the possibility of extending our prior $p(\hat{n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.
 
-\subsection{Ultra-nest Results}
+\subsection{Parameter Optimisation Results}
 \label{sec:3.3}
 We wanted to see if our choice of angles plays a significant role in our model's efficacy.
-We employed ultra-nest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
+We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
 We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
@@
 \begin{figure}[h]
+	\label{fig:corner}
 	\centering
-	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
+	\includegraphics[width=0.45\textwidth]{./Images/corneranglesfreed-1.png}
 	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
-	\label{fig:corner}
 \end{figure}
 
-Figure~\ref{fig:corner} is the resulting corner plot from our ultra-nest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
+Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
 Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle.
@@
 \section{Conclusion}
-\label{4}
+\label{sec:Conclusion}
 We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
 In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
-The MSTM package is flexible in calculating the light scattering for a variety of micro particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
-This has potential to be applied to particles that are changing while in the optical trap.
-We hope to apply this work in the future to  characterise processes such as nucleation within an optical trap.\\
-
-\noindent \textbf{Funding.} \\
+The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
+This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
+The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.\\
 

### 2022-11-14 Manuscript.tex
starting to clean up manuscript

@@
 \usepackage{color}
-\usepackage{nomecl}
+\usepackage{nomencl}
 
-\journal{Journal of Quantitative Spectroscopy $\&$ Radiative Transfer}
+\journal{Journal of Quantitative Spectroscopy \& Radiative Transfer}
 
@@
 \begin{abstract}
-\justifying
-While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem.
-Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup.
-We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities.
-To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped.
-We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
+  % \justifying
+  %
+  While optical trapping is a well understood method for force
+  transduction and detection,characterisation of trapped entities
+  poses a two-fold challenge --- one experimental concerning the
+  arrangement of light detectors and the other, theoretical involving
+  solving of the inverse light scattering problem. Combining static
+  light scattering techniques with optical trapping poses significant
+  engineering challenges due to the space constraints in a
+  conventional optical trapping setup.  We propose here a plausible
+  scenario of detecting scattered light from an optically trapped
+  asymmetric microstructure using a novel, multi-angle, optical-fibre
+  based detection scheme and demonstrate how a Bayesian inference
+  based analysis of the data simulated to mimic light scattering
+  detection signals in such scenarios maybe used for solving the
+  inverse light scattering problem and help characterise trapped
+  entities.  To this end, we discuss the application of our method to
+  infer the instantaneous orientations of an asymmetric microsphere
+  dimer being trapped. We argue that this method can be extended for
+  determining any characteristics of the trapped microstructure that
+  influence the light scattering pattern.
 \end{abstract}
@@
 
-\nomenclature{$\hat{s}$}{Dimer's instantaneous orientation}
-\nomenclature{$\hat{n}_i$}{Reference Orientation i}
+\nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
+\nomenclature{$\hat{\bf n}_i$}{Reference Orientation i}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
 \label{sec:Method}
+
 \subsection{Brownian Simulation}
 \label{sec:2.1}
-We use the Brownian OT package developed by Jerome Fung \textit{et al} \cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap.
-Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters.
-We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m$ and $ 0.5  \mu m$ respectively.
-We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
+
+We use the Brownian OT package developed by Fung~\textit{et~al}.
+\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric
+dimer within an optical trap.  Brownian OT combines MSTM
+\cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox''
+(\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+shaped sphere clusters.  We simulated the motion of silica dimer
+trapped in a $5$\,mW Gaussian beam; the dimer was comprised of two
+spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$,
+respectively.  We ran our simulation over the span of 10 seconds, with
+Brownian OT calculating the dimer's orientation and position every
+$1 \ ms$.
 
@@
 	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
-	\caption{Example dimer in orientation $\hat{s}$}
+	\caption{Example dimer in orientation $\bar{\bf s}$}
 	\label{fig:dimer}
@@
 
-With any trapped particle there are two main sources of motion, Brownian and optical forces.
-The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
-The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}.
-For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
-The T-matrix describes the scattering process by relating the scattering wave to the incident wave:
-
+With any trapped particle there are two main sources of motion,
+Brownian and optical forces.  The dimer's Brownian motion
+($\Delta q^{\textbf{B}}$) is simulated by generating a set of
+normally-distributed numbers whose covariance is given by the dimer's
+diffusion tensor $\textbf{D}$.  The diffusion tensor is computed based
+on Farsund's work \cite{Farsund1996}. For computing the optical force
+Brownian OT uses the T-Matrix method to calculate the scattering
+coefficients of the outgoing scattered wave.  The T-matrix describes
+the scattering process by relating the scattering wave to the incident
+wave:
 \begin{align}
@@
 \end{align}
+MSTM computes the T-matrix of the dimer, and then passes the T-matrix
+to \textit{ott} to compute the optical forces. The benefit of this
+approach is that MSTM can accurately compute the T-matrix for any
+arbitrary collection of spheres once, and then \textit{ott} can
+calculate the optical forces and torques. The total displacement is
+the sum of the optical and Brownian contributions:
+\begin{align}
+  \langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle
+  &= 2D_{ij} \Delta t
+  \\
+  \Delta q^{opt} &= \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F}
+  \\
+\Delta q^{tot} &= \Delta q^{\textbf{B}} + \Delta q^{opt}
+\end{align}
+where \textbf{F} is the a 6-component vector that describes the
+optical force and torque components.  Brownian OT returns the dimer's
+motion as a \textit{csv} file containing for each timestep the
+$[x,y,z]$ position of the dimer's centre of mass relative to the trap
+focus, and the orientation represented as a quaternion.
 
-MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
-The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques.
-The total displacement is the sum of the optical and Brownian contributions:
-
-\begin{align}
-	\langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle = 2D_{ij} \Delta t \\
-	\Delta q^{opt} = \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F} \\
-	\Delta q^{tot} = \Delta q^{\textbf{B}} + \Delta q^{opt}
-\end{align}
-
-where \textbf{F} is the a 6 component vector that describes the optical force and torque components.
-Brownian OT returns the dimer's motion as a \textit{csv} file containing for each timestep the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Orientation estimation from scattering measurements}
@@
 
-We represent the instantaneous orientation of the moving dimer in the trap as a three component unit vector,  $\hat{s}$, pointing from the centre of the larger sphere to the centre of the smaller sphere.
-A plane wave 'probe' laser beam directed perpendicular to the trapping laser, incident on the dimer, generates a scattering pattern dependent on the dimer's instantaneous orientation,  $I(\hat{s},\ \theta)$.
-This scattering pattern was computed using MSTM for each time step of our Brownian simulation.
-To represent a simulated 'measurement', three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the calculated  intensity at each angle k recorded as $I(\hat{s},\ \theta_k)$.
+We represent the instantaneous orientation of the moving dimer in the
+trap as a three component unit vector, $\hat{\bf s}$, pointing from
+the centre of the larger sphere to the centre of the smaller sphere.
+A plane wave ``probe'' laser beam directed perpendicular to the
+trapping laser, incident on the dimer, generates a scattering pattern
+dependent on the dimer's instantaneous orientation,
+$I(\bar{\bf s},\ \theta)$. This scattering pattern was computed using MSTM
+for each time step of our Brownian simulation.  To represent a
+simulated 'measurement', three angles ($\theta_1$, $\theta_2$,
+$\theta_3$) were chosen and the calculated intensity at each angle k
+recorded as $I(\hat{\bf s}, \theta_k)$.
 
-Our goal is to estimate the dimer orientation $\hat{s}$ over time, based on the scattering measurements  $I(\hat{s},\ \theta_k)$: the question then is what set of angles ($\theta_1,\ \theta_2,\ \& \theta_3$) give us the most useful information on $\hat{s}$.
-To initially simplify the problem, rather than trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}: our question becomes, how can we optimise the experiment to identify the reference orientation $\hat{n}_i$ closest to the real orientation $\hat{s}$.
-(Ultimately our method could be extended to an arbitrarily large number of reference orientations but for clarity and practicability we use a relatively small number here; additionally, the choice of a small number of discrete reference or 'candidate' orientations also accounts for the fact that any experiment will necessarily have a limited precision.)   Using MSTM we compute the scattering that would be produced by the dimer if it were in each of the reference orientations, denoted as $I(\hat{n}_i, \ \theta)$ and obtain 'measurements'  $I(\hat{n}_i, \ \theta_k)$ at our chosen set of  angles $\theta_k$.
-These raw intensities are normalised according to \eqref{eq:scale}:
+Our goal is to estimate the dimer orientation $\bar{\bf s}$ over time,
+based on the scattering measurements $I(\bar{\bf s},\ \theta_k)$: the
+question then is what set of angles ($\theta_1$, $\theta_2$,
+$\theta_3$) give us the most useful information on $\hat{\bf s}$.  To
+initially simplify the problem, rather than trying to get an exact
+estimation of $\hat{\bf s}$ we discretise the orientation space into
+$30$ possible reference orientations $\hat{\bf n}_i$ that divide up a
+unit sphere evenly \cite{Rey2006}: our question becomes, how can we
+optimise the experiment to identify the reference orientation
+$\hat{\bf n}_i$ closest to the real orientation $\hat{\bf s}$.
+(Ultimately our method could be extended to an arbitrarily large
+number of reference orientations but for clarity and practicability we
+use a relatively small number here; additionally, the choice of a
+small number of discrete reference or 'candidate' orientations also
+accounts for the fact that any experiment will necessarily have a
+limited precision.)  Using MSTM we compute the scattering that would
+be produced by the dimer if it were in each of the reference
+orientations, denoted as $I(\hat{\bf n}_i, \ \theta)$ and obtain
+'measurements' $I(\hat{\bf n}_i, \ \theta_k)$ at our chosen set of angles
+$\theta_k$. These raw intensities are normalised according to
+\eqref{eq:scale}:
 
@@
 	\label{eq:scale}
-	y_k(\hat{n}_i,\ \theta_k) = 
-	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
-	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
+	y_k(\hat{\bf n}_i,\ \theta_k) = 
+	 \frac{\left[I(\hat{\bf n}_i, \ \theta_k)- \langle I(\hat{\bf n}, \ \theta_k) \rangle \right]} 
+	{\langle I^2(\hat{\bf n},\ \theta_k) \rangle -\langle I(\hat{\bf n}, \ \theta_k)\rangle^2}
 \end{align}
 
-Where the denominator is simply the standard deviation for our values of $I(\hat{n}, \theta_k)$.
-The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\hat{s}$  is similarly scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$.
+Where the denominator is simply the standard deviation for our values of $I(\hat{\bf n}, \theta_k)$.
+The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\bar{\bf s}$  is similarly scaled using \eqref{eq:scale} to give $y(\bar{\bf s},\ \theta_k)$.
 See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals.
-In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
-We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\hat{s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{n}_i$, it would produce a signal $y(\hat{s})$:
+In practice $y(\bar{\bf s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
+We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{\bf n}_i$, it would produce a signal $y(\bar{\bf s})$:
 
@@
 	\label{eq:conditional}
-	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
+	p_i(y(\bar{\bf s})\parallel\hat{\bf n}_i) = \Pi^3_{k=1}
 	(2\pi\sigma_k^2)^{-1/2} 
-	e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
+	e^{-(y_{k}(\hat{\bf n}_i)-y_{k}(\bar{\bf s}))^2/2\sigma_k^2}
 \end{align}
 
-Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\hat{s})$.
+Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\bar{\bf s})$.
 We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
-The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.
+The signal error was simulated in our model by stochastically adjusting our values of $y_k(\bar{\bf s})$.
 
 \begin{eqnarray*}
-	\sigma_k = \epsilon\bar{y}(\hat{n}_k) \\
-	\Rightarrow y(\hat{s}_k) = y(\hat{s})_k \pm n\sigma_k \\ 
+	\sigma_k = \epsilon\bar{y}(\hat{\bf n}_k) \\
+	\Rightarrow y(\bar{\bf s}_k) = y(\bar{\bf s})_k \pm n\sigma_k \\ 
 	n \in \mathbb{R}, n\in[0,1]
@@
 Where $\epsilon$ is the percent error expected from our detection fibres.
-Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{n}_i)$ is to the actual signal $y_{k}(\hat{s})$ generated by the dimer in its real orientation $\hat{s}$.
-However, if our values of $y_k(\hat{n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations.
+Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{\bf n}_i)$ is to the actual signal $y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation $\bar{\bf s}$.
+However, if our values of $y_k(\hat{\bf n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations.
 Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement.
@@
 	\label{eq:bayes}
-	p_i(\hat{n}\parallel y(\hat{s}))&= \frac{p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})}{p(y)}
+	p_i(\hat{\bf n}\parallel y(\bar{\bf s}))&= \frac{p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})}{p(y)}
 \end{align}
 
-Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively.
+Where $p_i(\hat{\bf n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively.
 The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}.
-The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\hat{s})$ that falls with our distribution of reference signals $y_k(\hat{n})$.
+The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\bar{\bf s})$ that falls with our distribution of reference signals $y_k(\hat{\bf n})$.
 This is given by taking the discrete integral over collection of reference orientations.
@@
 \begin{align}
-	p(y) = \sum\limits_{i=1}^{30} p_i(y(\hat{s})\parallel\hat{n})p_i(\hat{n})
+	p(y) = \sum\limits_{i=1}^{30} p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})
  \end{align}
@@
 We applied \eqref{eq:bayes} - covered in Section~\ref{sec:2.2} - and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
-The priori of reference orientations $p_i(\hat{n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{n}$
+The priori of reference orientations $p_i(\hat{\bf n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{\bf n}$
 
 \begin{align}
-	r_i(t, \hat{n}_i)= \frac{1}{2}((\hat{n}_{i,\ t} \cdot \hat{n}_{t-\Delta t})-1) \\
-	p_i(\hat{n})= \frac{e^{\beta (r_i(t,\hat{n}_i)-1)}}
-	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{n}_i)-1)}}
+	r_i(t, \hat{\bf n}_i)= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1) \\
+	p_i(\hat{\bf n})= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
+	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
 \end{align}
-
-Where $\beta$ is a weighting factor that controls the significance of our priori estimate.
-The simulation data from Section~\ref{sec:2.1} was used to check if the estimation was correct by comparing the estimate provided by \eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as $\hat{n}_{best}$.
-To measure the error in the model's estimate the distribution's of $p(\hat{n}\parallel y(\hat{s}))$ and an ideal result, $p_ideal$, were compared.
-An ideal distribution is one that is 0 for every $\hat{n}$ apart from the correct result.
-
+where $\beta$ is a weighting factor that controls the significance of
+our priori estimate. The simulation data from Section~\ref{sec:2.1}
+was used to check if the estimation was correct by comparing the
+estimate provided by \eqref{eq:bayes} to the reference orientation
+closest to the instantaneous orientation, denoted as
+$\hat{\bf n}_{best}$.  To measure the error in the model's estimate the
+distribution's of $p(\hat{\bf n}\parallel y(\bar{\bf s}))$ and an ideal
+result, $p_ideal$, were compared. An ideal distribution is one that is
+0 for every $\hat{\bf n}$ apart from the correct result.
 \begin{align}
@@
 	\begin{cases}
-		1 & \text{when $\hat{n}_i$ = $\hat{n}_{best}$}\\
+		1 & \text{when $\hat{\bf n}_i$ = $\hat{\bf n}_{best}$}\\
 		0 & \text{anywhere else}
@@
 \end{align}
-
 The distribution from \eqref{eq:bayes} will assign some non-zero probability to every reference orientation.
@@
 \begin{align}
-	K_{l, \#}(p_i(\hat{n}\parallel y_k(\hat{s})) &\parallel p_{ideal}) =\nonumber \\
-	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{n}_{best}\parallel y(\hat{s}))}
+	K_{l, \#}(p_i(\hat{\bf n}\parallel y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber \\
+	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}\parallel y(\bar{\bf s}))}
 	\right]
 \end{align}
-
-Where a more negative value of $K_l$ indicates a poorer estimation of the dimer's instantaneous orientation.
-By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
-To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
-
+where a more negative value of $K_l$ indicates a poorer estimation of
+the dimer's instantaneous orientation. By summing the divergence of
+each measurement across the entire simulation we get an evaluation of
+how well the model performed in estimating the dimer's orientation. To
+compare between different iterations of our model the we compare our
+result of $K_l$ to a worst case scenario and evaluate how much it
+improves upon this, denoted as $F(K_l)$:
 \begin{align}
@@
 
-The worst case scenario is akin to randomly choosing a reference orientation at each time step.
-The greater the value of $F(K_l)$ the lower the error between our estimated orientation and the dimer's instantaneous orientation.
-Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.
+The worst case scenario is akin to randomly choosing a reference
+orientation at each time step. The greater the value of $F(K_l)$ the
+lower the error between our estimated orientation and the dimer's
+instantaneous orientation. Because our model is dependent on several
+parameters we need to a sophisticated method for understanding how
+these parameters correlate with $F(K_l)$.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Optimization of parameters}
@@
 
-As can be seen from Figure~\ref{fig:epsilon} there is seemingly a sweet spot for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an alright estimation but with sudden shifts away from the ideal result.
-The reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.
-As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
-This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
-In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.
+As can be seen from Figure~\ref{fig:epsilon}, there is seemingly a
+sweet spot for the signal error; with a $5$\% error giving an
+inaccurate estimation, and with a $25$\% error giving an alright
+estimation but with sudden shifts away from the ideal result. The
+reason why our low error estimation performed so poorly is that the
+minimum difference between our values of $y_k$ is larger than our
+signal error.  As such the model could not distinguish between
+different reference orientations and was essentially just picking
+options at random.  This minimum difference determined by our choice of
+angles and sets a lower limit of how accurate our signals can be.  In
+the future we hope to look at increasing the number of reference
+orientations to see what the true lower bound is.
 
-The latter issue of a high signal error can be addressed somewhat by adjusting our value of $\beta$ to flatten our probability distribution.
-We repeated our model estimation at $25\%$ at different values of $\beta$ to see the upper limit we could improve our result to.
+The latter issue of a high signal error can be addressed somewhat by
+adjusting our value of $\beta$ to flatten our probability
+distribution. We repeated our model estimation at $25\%$ at different
+values of $\beta$ to see the upper limit we could improve our result
+to.
 
@@
 
-As shown by \ref{fig:beta} eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
-The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error.
-In the future we want to also investigate the possibility of extending our prior $p(\hat{n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.
+As shown in Figure~\ref{fig:beta}, eventually $\beta$ begins to over
+correct our estimate to the extent that we are a full radian away from
+the correct result.  The exact value of $\beta$ that will optimise our
+performance varies for each choice of angle, therefore we can use it
+as variable for finely tuning our results to suit our choice of angles
+and our signal error.  In future, we want to also investigate the
+possibility of extending our prior $p(\hat{\bf n})$ so we account for not
+just the change in orientation after one time step but after 2 or 3
+time steps to investigate if we can investigate the auto correlation
+of our dimer's motion.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
 \label{sec:Conclusion}
+
 We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{bib} 
@@
 \onecolumn
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Appendix}
@@
 			\hline\hline
-			Reference Orientation i & $\hat{n}_{i,\ x}$ &  $\hat{n}_{i,\ y}$ &  $\hat{n}_{i,\ z}$ \\
+			Reference Orientation i & $\hat{\bf n}_{i,\ x}$ &  $\hat{\bf n}_{i,\ y}$ &  $\hat{\bf n}_{i,\ z}$ \\
 			\hline
@@
 			\hline\hline
-			Reference Orientation i  & $I(\hat{n}_i,\ 15^{\circ})$ &  $I(\hat{n}_i, \ 45^{\circ})$ &  $I(\hat{n}_i,\ 90^{\circ})$ & $y(\hat{n}_i, \ 15^{\circ})$ & $y(\hat{n}_i, \ 45^{\circ})$ & $y(\hat{n}_i, \ 90^{\circ})$ \\
+			Reference Orientation i  & $I(\hat{\bf n}_i,\ 15^{\circ})$ &  $I(\hat{\bf n}_i, \ 45^{\circ})$ &  $I(\hat{\bf n}_i,\ 90^{\circ})$ & $y(\hat{\bf n}_i, \ 15^{\circ})$ & $y(\hat{\bf n}_i, \ 45^{\circ})$ & $y(\hat{\bf n}_i, \ 90^{\circ})$ \\
 			\hline
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2022-11-14 Manuscript.tex
more work on cleaning up text

@@
 \end{align}
-
-Where the denominator is simply the standard deviation for our values of $I(\hat{\bf n}, \theta_k)$.
-The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\bar{\bf s}$  is similarly scaled using \eqref{eq:scale} to give $y(\bar{\bf s},\ \theta_k)$.
-See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals.
-In practice $y(\bar{\bf s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
-We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{\bf n}_i$, it would produce a signal $y(\bar{\bf s})$:
-
+where the denominator is simply the standard deviation for our values
+of $I(\hat{\bf n}, \theta_k)$.  The dimer's \emph{actual}
+instantaneous scattered intensity based on the real instantaneous
+orientation $\bar{\bf s}$ is similarly scaled using \eqref{eq:scale}
+to give $y(\bar{\bf s},\ \theta_k)$.  See Tables A.1 and \ref{tab:A2}
+for numerical values of our reference orientation, raw intensities,
+and scaled signals.  In practice, $y(\bar{\bf s}\, \theta_k)$ will
+have an inherent amount of uncertainty in our measurement due to noise
+from our detector.  We compare our instantaneous signal to our
+reference signal by assuming that the measurement uncertainty for each
+detection fibre is Gaussian and is centred about $y_k(\bar{\bf s})$
+Combining all three detection angles provides us with the conditional
+probability that, should the dimer be in orientation $\hat{\bf n}_i$,
+it would produce a signal $y(\bar{\bf s})$:
 \begin{align}
 	\label{eq:conditional}
-	p_i(y(\bar{\bf s})\parallel\hat{\bf n}_i) = \Pi^3_{k=1}
+	p_i(y(\bar{\bf s})|\hat{\bf n}_i) = \Pi^3_{k=1}
 	(2\pi\sigma_k^2)^{-1/2} 
@@
 \end{align}
-
-Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\bar{\bf s})$.
-We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model.
-The signal error was simulated in our model by stochastically adjusting our values of $y_k(\bar{\bf s})$.
-
+where $\sigma_k$ is the average error in the detected signal from our
+dimer $y_k(\bar{\bf s})$.  We left this as a tuneable variable to
+measure how the influence of experimental error would effect our
+predictive model.  The signal error was simulated in our model by
+stochastically adjusting our values of $y_k(\bar{\bf s})$.
 \begin{eqnarray*}
-	\sigma_k = \epsilon\bar{y}(\hat{\bf n}_k) \\
-	\Rightarrow y(\bar{\bf s}_k) = y(\bar{\bf s})_k \pm n\sigma_k \\ 
+  \sigma_k
+  &= \epsilon\bar{y}(\hat{\bf n}_k)
+  \\
+  \Rightarrow y(\bar{\bf s}_k) &= y(\bar{\bf s})_k \pm n\sigma_k
+  \\ 
 n \in \mathbb{R}, n\in[0,1]
 \end{eqnarray*} 
-
-Where $\epsilon$ is the percent error expected from our detection fibres.
-Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{\bf n}_i)$ is to the actual signal $y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation $\bar{\bf s}$.
-However, if our values of $y_k(\hat{\bf n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations.
-Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement.
-We apply this by accounting for our distribution of reference orientations and instantaneous signals.
-
+where $\epsilon$ is the percent error expected from our detection
+fibres.  Equation~\eqref{eq:conditional} assigns a probability to each
+reference orientation based on how similar the 'reference orientation
+signal' $y_{k}(\hat{\bf n}_i)$ is to the actual signal
+$y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation
+$\bar{\bf s}$.  However, if our values of $y_k(\hat{\bf n}_i)$ are
+similar to one another the probability distribution from
+Eq.~\eqref{eq:conditional} will therefore assign similar
+probability for multiple reference orientations. Bayesian inference
+allows one to update a probability distribution by accounting for
+information known prior to making a measurement.  We apply this by
+accounting for our distribution of reference orientations and
+instantaneous signals.
 \begin{align}
   \label{eq:bayes}
-	p_i(\hat{\bf n}\parallel y(\bar{\bf s}))&= \frac{p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})}{p(y)}
+  p_i(\hat{\bf n}| y(\bar{\bf s}))
+  &=
+    \frac{p_i(y(\bar{\bf s})|\hat{\bf n})p_i(\hat{\bf n})}{p(y)}
 \end{align}
+where $p_i(\hat{\bf n})$ and $p(y)$ are our estimations of the prior
+distributions of particle orientations and instantaneous signals
+respectively.  The former can be estimated by some Boltzmann
+distribution based on how far apart reference orientations are in
+physical space, covered in Section~\ref{sec:2.3}.  The latter prior
+$p(y)$ is the probability of our dimer producing a signal
+$y_k(\bar{\bf s})$ that falls with our distribution of reference
+signals $y_k(\hat{\bf n})$.  This is given by taking the discrete
+integral over collection of reference orientations.
+\begin{align}
+  p(y)
+  =
+  \sum_{i=1}^{30} p_i(y(\bar{\bf s})|\hat{\bf n})p_i(\hat{\bf n})
+\end{align}
+The implementation of Eq.~\eqref{eq:bayes} improves our estimate by
+suppressing reference orientations that are far away from our dimer's
+orientation vector in physical space.  In order to evaluate the
+accuracy of our method, we tested it using a Brownian dynamics
+simulation of the dimer (see Section~\ref{sec:2.1}).
 
-Where $p_i(\hat{\bf n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively.
-The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}.
-The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\bar{\bf s})$ that falls with our distribution of reference signals $y_k(\hat{\bf n})$.
-This is given by taking the discrete integral over collection of reference orientations.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing the Model}
+\label{sec:2.3}
 
+Using our simulation from Section~\ref{sec:2.1} we calculated the
+instantaneous signal from our dimer at each time step using MSTM.  We
+applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:2.2} ---
+and the reference orientation with the highest probability is taken as
+our estimation of the dimer's instantaneous orientation.  The prior of
+the reference orientations $p_i(\hat{\bf n})$ is defined at each time
+step by accounting for the physical distance between our previous
+estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference
+orientation $\hat{\bf n}$
 \begin{align}
-	p(y) = \sum\limits_{i=1}^{30} p_i(y(\bar{\bf s})\parallel\hat{\bf n})p_i(\hat{\bf n})
- \end{align}
-
-The implementation of \eqref{eq:bayes} improves our estimate by suppressing reference orientations that are far away from our dimer's orientation vector in physical space.
-In order to evaluate the accuracy of our model we tested it using our Brownian simulation of the dimer, see Section~\ref{sec:2.1}.
-
-\subsection{Testing our Model}
-\label{sec:2.3}
-Using our simulation from Section~\ref{sec:2.1} we calculated the instantaneous signal from our dimer at each time step using MSTM.
-We applied \eqref{eq:bayes} - covered in Section~\ref{sec:2.2} - and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
-The priori of reference orientations $p_i(\hat{\bf n})$ is defined at each time step by accounting for the physical distance between our previous estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference orientation $\hat{\bf n}$
-
-\begin{align}
-	r_i(t, \hat{\bf n}_i)= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1) \\
-	p_i(\hat{\bf n})= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
-	{\Sigma_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
+  r_i(t, \hat{\bf n}_i)
+  &= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1)
+  \\
+  p_i(\hat{\bf n})
+  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
+{\sum_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
 \end{align}
@@
 was used to check if the estimation was correct by comparing the
-estimate provided by \eqref{eq:bayes} to the reference orientation
+estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation
 closest to the instantaneous orientation, denoted as
-$\hat{\bf n}_{best}$.  To measure the error in the model's estimate the
-distribution's of $p(\hat{\bf n}\parallel y(\bar{\bf s}))$ and an ideal
-result, $p_ideal$, were compared. An ideal distribution is one that is
-0 for every $\hat{\bf n}$ apart from the correct result.
+$\hat{\bf n}_{best}$.  To measure the error in the model's estimate
+the distribution's of $p(\hat{\bf n}| y(\bar{\bf s}))$ and an
+ideal result, $p_ideal$, were compared.  An ideal distribution is one
+that is 0 for every $\hat{\bf n}$ apart from the correct result.
 \begin{align}
@@
 \end{align}
-The distribution from \eqref{eq:bayes} will assign some non-zero probability to every reference orientation.
-The model's error was measured by calculating the Kullback-Leibler ($K_l$) divergence between the two distributions
-
+The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
+probability to every reference orientation.  The model's error was
+measured by calculating the Kullback-Leibler divergence $K_l$ between
+the two distributions
 \begin{align}
-	K_{l, \#}(p_i(\hat{\bf n}\parallel y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber \\
-	&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}\parallel y(\bar{\bf s}))}
+  K_{l, \#}(p_i(\hat{\bf n}| y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber
+  \\
+&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y(\bar{\bf s}))}
 \right]
@@
 each measurement across the entire simulation we get an evaluation of
-how well the model performed in estimating the dimer's orientation. To
-compare between different iterations of our model the we compare our
-result of $K_l$ to a worst case scenario and evaluate how much it
+how well the model performed in estimating the dimer's orientation.
+To compare between different iterations of our model the we compare
+our result of $K_l$ to a worst case scenario and evaluate how much it
 improves upon this, denoted as $F(K_l)$:
@@
 \label{sec:2.4}
-The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly.
-However, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo simulation to our model.
-Ultra-nest is a sophisticated python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}.
-Ultra-nest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
 
-The Ultra-nest software was used to create a set of live points of different choices of $\theta_1, \ \theta_2, \ \& \ \theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$).
-After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
-If there is a correlation between the choice of angles and $F(K_l)$ then the live points should congregate around a single (or multiple) choice of angles that will maximise $F(K_l)$ and therefore minimise the error between our estimate and instantaneous orientation.
+The signal error $\epsilon$ and the weighting factor $\beta$ are known
+quantities whose influence can be measured directly; however, our
+choice of angles and how they correlate with $F(K_l)$ is not easily
+understood Evaluating every possible configuration manually is
+inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo
+simulation to our model. Ultra-nest is a sophisticated python package
+for fitting models with complex parameter correlations
+\cite{Buchner2016Ultranest}.  UltraNest implements a variation of
+Monte Carlo integration, nested-sampling, where a likelihood contour
+is used to update a group of live points chosen from the prior
+estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
 
-\section{Results $\&$ Discussion}
+The UltraNest software was used to create a set of live points of
+different choices of $\theta_1$, $\theta_2$, and $\theta_3$, with each
+point being evaluated by its improvement factor ($F(K_l)$).  After
+evaluating every live point, the worst results are removed and new
+live points are created based on the information provided by the
+previous live points.  If there is a correlation between the choice of
+angles and $F(K_l)$ then the live points should congregate around a
+single (or multiple) choice of angles that will maximise $F(K_l)$ and
+therefore minimise the error between our estimate and instantaneous
+orientation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results and  Discussion}
 \label{sec:Discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Asymmetric dimer dynamics}
 \label{sec:3.1}
-The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap.
-The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
-The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.
+
+The Brownian OT software was used to simulate the motion of a trapped
+dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first $10$\,seconds of
+entering the optical trap.  The initial orientation was assumed as
+strictly vertical (in line with the beam propagation direction). The
+dimer's position orientation was recorded every $10\,{\rm \mu s}$ for
+using as a test dataset for our model.
 
@@
 
-As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
-Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
-It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
-Further simulations of differently sized dimers showed similar results, but only when $a_1 \geq 2a_2$.
-Dimers with more symmetrical size ratios immediately aligned into a fixed vertical position.
-In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT} simulations of trapped symmetrical dimers was investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
-Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
-Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
-The data from \figurename{ 2a} was used to test the efficacy of our predictive model.
-To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
-For a perfect estimation these two plots should overlap, while we cannot achieve an exact estimation we can optimise it by adjusting the expected signal error and weighting factor.
+As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a
+full $180^{\circ}$ rotation upon entering the trap. Typically
+horizontal alignment of a dimer is unstable and will result in the
+particle rotating to align along its vertical axis. It is interesting
+to note that the dimer is furthest from the trap centre as it goes
+into a horizontal orientation before drawing closer again as it
+inverts completely. Further simulations of differently sized dimers
+showed similar results, but only when $a_1 \geq 2a_2$. Dimers with
+more symmetrical size ratios immediately aligned into a fixed vertical
+position.
+%
+In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT}
+simulations of trapped symmetrical dimers was investigated; their
+findings showed that the optical torque on the dimer goes to zero
+while aligned vertically and is at its maximum in a horizontal
+alignment.  Therefore, the inversion of an asymmetric dimer suggests
+that if the size difference is significant the optical torque goes is
+minimal for a dimer in both horizontal and vertical orientations. Once
+we have our experimental set up complete we can confirm this by
+trapping an asymmetric dimer and monitoring its orientation. The data
+from \figurename{ 2a} was used to test the efficacy of our predictive
+model. To visualise how well our model tracks the dimer's orientation
+we plot the radial distance between our model's prediction and the
+true orientation, and between the best possible orientation and the
+true orientation. For a perfect estimation these two plots should
+overlap, while we cannot achieve an exact estimation we can optimise
+it by adjusting the expected signal error and weighting factor.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Impact of signal error on model predictions}
 \label{sec:3.2}
-A key assumption of equation~\ref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise.
-This is the central premise that allows us to form equation~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
+
+A key assumption of Eq.~\eqref{eq:conditional} is that we assume our
+detected light signal will inherently have some degree of noise. This
+is the central premise that allows us to form
+Eq.~\eqref{eq:conditional}, understanding the limits of acceptable
+signal error is necessary for implementing this to experimental work.
 To this end we ran our model while varying the value of $\epsilon$.
-\figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1.
-The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation
+\figurename{ 3} shows the performance of the method with a selection
+of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
+to $1$.  The dotted line indicates the minimum radian distance
+($0.137$ radians) between two neighbouring reference orientations, if
+we are under this line then we know our model is just off from
+selecting the correct orientation
 
@@
 
-We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
-In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
-The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
-This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
-The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.\\
+We have developed a method for interpreting the dynamics of a trapped
+particle based purely on the light scattering pattern. In theory the
+model can be applied to any characteristic that impacts the light
+scattering pattern produced by the trapped entity. The MSTM package is
+flexible in calculating the light scattering for a variety of
+micro-particles; so long as there is a distinguishable difference in
+the light scattering pattern we can apply this method to characterise
+the particle to some degree of accuracy. This has potential to be
+applied to characterise particle size and shape that are changing
+while in the optical trap. The authors would like to acknowledge and
+thank the support for this research from the funding provided by the
+Leverhulme Trust.\\
 
@@
 \begin{table}[h]
-	\label{tab:A1}
 \begin{center}
-		\caption{Reference Orientations vector components$^*$}
+\caption{\label{tab:A1}
+%
+Reference Orientations vector components$^*$}
 \begin{tabular}{|c|c|c|c|}
 \hline\hline
-			Reference Orientation i & $\hat{\bf n}_{i,\ x}$ &  $\hat{\bf n}_{i,\ y}$ &  $\hat{\bf n}_{i,\ z}$ \\
+Reference Orientation & $\hat{\bf n}_{i,\ x}$ &  $\hat{\bf n}_{i,\ y}$ &  $\hat{\bf n}_{i,\ z}$ \\
 \hline
@@
 \newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}[h]
-	\label{tab:A2}
   \begin{center}
-		\caption{Raw intensities $I_k$$^*$ and scaled intensities $y_k$}
+    \caption{\label{tab:A2}
+      % 
+      Raw intensities $I_k^*$ and scaled intensities $y_k$}
     \begin{tabular}{|c|c|c|c|c|c|c|}
       \hline\hline
-			Reference Orientation i  & $I(\hat{\bf n}_i,\ 15^{\circ})$ &  $I(\hat{\bf n}_i, \ 45^{\circ})$ &  $I(\hat{\bf n}_i,\ 90^{\circ})$ & $y(\hat{\bf n}_i, \ 15^{\circ})$ & $y(\hat{\bf n}_i, \ 45^{\circ})$ & $y(\hat{\bf n}_i, \ 90^{\circ})$ \\
+      Reference Orientation  & $I(\hat{\bf n}_i,\ 15^{\circ})$ &  $I(\hat{\bf n}_i, \ 45^{\circ})$ &  $I(\hat{\bf n}_i,\ 90^{\circ})$ & $y(\hat{\bf n}_i, \ 15^{\circ})$ & $y(\hat{\bf n}_i, \ 45^{\circ})$ & $y(\hat{\bf n}_i, \ 90^{\circ})$ \\
 \hline

### 2022-11-14 Manuscript.tex
fixing notation

@@
 
-The possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and its Brownian motion was studied.
-While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
-In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.
+The possibility of integrating light scattering with optical trapping
+was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998}
+where a single-mode optical fibre was aligned to detect the scattered
+light from a trapped bead and its Brownian motion was studied.  While
+this allowed for studying the dynamics, structural information about
+the trapped bead was precluded as the measurement was obtained only at
+a single angle.  In this work, we propose a scheme that expands on the
+technique in \cite{Bar-Ziv_1998} to detect scattered light
+simultaneously at 3 angles as shown in Figure~\ref{fig:setup},
+combined with a novel Bayesian Inference based analysis technique to
+enable interpretation of the resulting multi-angle data as well as
+optimisation of the setup to provide maximal information from the
+signal.
 
-To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity more complex than a single sphere, i.e.
-a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity.
-There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
-Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
+To demonstrate the analysis, we study a simple, illustrative example
+of a trapped entity more complex than a single sphere (i.e.\ a trapped
+asymmetric dimer).  We explore how to optimise the experimental
+measurement to obtain the best estimate of the dimer's instantaneous
+orientation, as a further paradigmatic example of interpretation of
+the scattering data to obtain dynamical information on the trapped
+entity.  There is a paucity of literature on measuring the orientation
+of trapped non-spherical particles till the recent attempt in
+Ref.~\citenum{raudsepp2022estimating} where imaging was employed to
+study the orientation of trapped dimers and a resolution of () was
+achieved.  Here, we simulate the Brownian motion of a dimer and
+calculate the resulting scattered light, and show how Bayesian
+inference can be used to optimise extraction of the true dimer
+orientation from the light scattering signals given a choice of
+scattering measurement angles.
 
@@
 \end{subfigure}
-	\caption{Experimental set up, a) Side view, b) Top view }
-	\label{fig:setup}
+\caption{\label{fig:setup}
+  %
+  Experimental set up, a) Side view, b) Top view
+%
+}
 \end{figure}
@@
 We represent the instantaneous orientation of the moving dimer in the
-trap as a three component unit vector, $\hat{\bf s}$, pointing from
-the centre of the larger sphere to the centre of the smaller sphere.
-A plane wave ``probe'' laser beam directed perpendicular to the
-trapping laser, incident on the dimer, generates a scattering pattern
-dependent on the dimer's instantaneous orientation,
-$I(\bar{\bf s},\ \theta)$. This scattering pattern was computed using MSTM
-for each time step of our Brownian simulation.  To represent a
-simulated 'measurement', three angles ($\theta_1$, $\theta_2$,
-$\theta_3$) were chosen and the calculated intensity at each angle k
-recorded as $I(\hat{\bf s}, \theta_k)$.
+trap as a unit vector $\hat{\bf s}$, pointing from the centre of the
+larger sphere to the centre of the smaller sphere.  A plane wave
+``probe'' laser beam directed perpendicular to the trapping laser,
+incident on the dimer, generates a scattering pattern dependent on the
+dimer's instantaneous orientation, $I(\bar{\bf s},\theta_k)$.  This
+scattering pattern was computed using MSTM for each time step of the
+Brownian dynamics simulation.  To represent a simulated measurement,
+three angles ($\theta_1$, $\theta_2$, $\theta_3$) were chosen, and the
+calculated intensity at each angle $\theta_k$ is recorded as
+$I(\hat{\bf s}, \theta_k)$.
 
@@
 initially simplify the problem, rather than trying to get an exact
-estimation of $\hat{\bf s}$ we discretise the orientation space into
-$30$ possible reference orientations $\hat{\bf n}_i$ that divide up a
-unit sphere evenly \cite{Rey2006}: our question becomes, how can we
-optimise the experiment to identify the reference orientation
-$\hat{\bf n}_i$ closest to the real orientation $\hat{\bf s}$.
-(Ultimately our method could be extended to an arbitrarily large
-number of reference orientations but for clarity and practicability we
-use a relatively small number here; additionally, the choice of a
-small number of discrete reference or 'candidate' orientations also
-accounts for the fact that any experiment will necessarily have a
-limited precision.)  Using MSTM we compute the scattering that would
-be produced by the dimer if it were in each of the reference
-orientations, denoted as $I(\hat{\bf n}_i, \ \theta)$ and obtain
-'measurements' $I(\hat{\bf n}_i, \ \theta_k)$ at our chosen set of angles
-$\theta_k$. These raw intensities are normalised according to
-\eqref{eq:scale}:
-
+estimation of $\hat{\bf s}$, we discretise the orientation into
+$n_{\rm ref}=30$ possible reference orientations $\hat{\bf n}_\alpha$
+that are evenly distributed on the unit sphere \cite{Rey2006}.
+%
+We then consider how to optimise an experiment to identify the
+reference orientation $\hat{\bf n}_\alpha$ that is closest to the
+actual orientation $\hat{\bf s}$ of the dimer.
+%
+The method can be extended to an arbitrarily large number of reference
+orientations, but for clarity and practicability we use a relatively
+small number here; additionally, the choice of a small number of
+discrete reference or ``candidate'' orientations also accounts for the
+fact that any experiment will necessarily have a limited precision.
+%
+Using MSTM we compute the scattering that would be produced by the
+dimer in each of the reference orientations, denoted as
+$I(\hat{\bf n}_\alpha, \theta)$ and obtain ``measurements''
+$I(\hat{\bf n}_\alpha, \theta_k)$ at our chosen set of angles
+($\theta_1$, $\theta_2$, $\theta_3$).
+%
+These raw intensities are normalised according to
 \begin{align}
 \label{eq:scale}
-	y_k(\hat{\bf n}_i,\ \theta_k) = 
-	 \frac{\left[I(\hat{\bf n}_i, \ \theta_k)- \langle I(\hat{\bf n}, \ \theta_k) \rangle \right]} 
-	{\langle I^2(\hat{\bf n},\ \theta_k) \rangle -\langle I(\hat{\bf n}, \ \theta_k)\rangle^2}
+  y_k(\hat{\bf n}_\alpha)
+  = 
+  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
+  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
 \end{align}
@@
 of $I(\hat{\bf n}, \theta_k)$.  The dimer's \emph{actual}
-instantaneous scattered intensity based on the real instantaneous
-orientation $\bar{\bf s}$ is similarly scaled using \eqref{eq:scale}
-to give $y(\bar{\bf s},\ \theta_k)$.  See Tables A.1 and \ref{tab:A2}
-for numerical values of our reference orientation, raw intensities,
-and scaled signals.  In practice, $y(\bar{\bf s}\, \theta_k)$ will
-have an inherent amount of uncertainty in our measurement due to noise
-from our detector.  We compare our instantaneous signal to our
-reference signal by assuming that the measurement uncertainty for each
-detection fibre is Gaussian and is centred about $y_k(\bar{\bf s})$
-Combining all three detection angles provides us with the conditional
-probability that, should the dimer be in orientation $\hat{\bf n}_i$,
-it would produce a signal $y(\bar{\bf s})$:
+instantaneous scattered intensity, based on the real instantaneous
+orientation $\bar{\bf s}$, is similarly scaled using
+Eq.~\eqref{eq:scale} to give $y(\bar{\bf s},\theta_k)$.
+%
+The reference orientations, raw intensities, and scaled signals are
+given in Tables~\ref{tab:A1} and \ref{tab:A2}.
+%
+In practice, $y_k(\bar{\bf s})$ will have an inherent amount of
+uncertainty in our measurement due to noise from our detector.  We
+compare our instantaneous signal to our reference signal by assuming
+that the measurement uncertainty for each detection fibre is Gaussian
+and is centred about $y_k(\bar{\bf s})$ Combining all three detection
+angles provides us with the conditional probability that, should the
+dimer be in orientation $\hat{\bf n}_\alpha$, it would produce a
+signal $y(\bar{\bf s})$:
 \begin{align}
 \label{eq:conditional}
-	p_i(y(\bar{\bf s})|\hat{\bf n}_i) = \Pi^3_{k=1}
+  p(y(\bar{\bf s})|\hat{\bf n}_\alpha)
+  = \prod^3_{k=1}
 (2\pi\sigma_k^2)^{-1/2} 
-	e^{-(y_{k}(\hat{\bf n}_i)-y_{k}(\bar{\bf s}))^2/2\sigma_k^2}
+e^{-(y_{k}(\hat{\bf n}_\alpha)-y_{k}(\bar{\bf s}))^2/2\sigma_k^2}
 \end{align}
-where $\sigma_k$ is the average error in the detected signal from our
-dimer $y_k(\bar{\bf s})$.  We left this as a tuneable variable to
-measure how the influence of experimental error would effect our
-predictive model.  The signal error was simulated in our model by
+where $\sigma_k$ is related to the uncertainty in the detected signal
+from our dimer $y_k(\bar{\bf s})$.  We left this as a tuneable
+parameter to characterize the influence of experimental error on our
+predictive method.  The signal error was simulated in our model by
 stochastically adjusting our values of $y_k(\bar{\bf s})$.
@@
 fibres.  Equation~\eqref{eq:conditional} assigns a probability to each
-reference orientation based on how similar the 'reference orientation
-signal' $y_{k}(\hat{\bf n}_i)$ is to the actual signal
+reference orientation based on how similar the ``reference orientation
+signal'' $y_{k}(\hat{\bf n}_\alpha)$ is to the actual signal
 $y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation
-$\bar{\bf s}$.  However, if our values of $y_k(\hat{\bf n}_i)$ are
-similar to one another the probability distribution from
-Eq.~\eqref{eq:conditional} will therefore assign similar
-probability for multiple reference orientations. Bayesian inference
-allows one to update a probability distribution by accounting for
-information known prior to making a measurement.  We apply this by
-accounting for our distribution of reference orientations and
-instantaneous signals.
+$\bar{\bf s}$.  However, if our values of $y_k(\hat{\bf n}_\alpha)$
+are similar to one another, the probability distribution from
+Eq.~\eqref{eq:conditional} will therefore assign similar probability
+for multiple reference orientations. Bayesian inference allows one to
+update a probability distribution by accounting for information known
+prior to making a measurement.  We apply this by accounting for our
+distribution of reference orientations and instantaneous signals.
 \begin{align}
   \label{eq:bayes}
-  p_i(\hat{\bf n}| y(\bar{\bf s}))
+  p(\hat{\bf n}_\alpha| y_1,y_2,y_3)
   &=
-    \frac{p_i(y(\bar{\bf s})|\hat{\bf n})p_i(\hat{\bf n})}{p(y)}
+    \frac{p(y_1,y_2,y_3|\hat{\bf n}_\alpha)
+    p(\hat{\bf n}_\alpha)}{p(y_1, y_2, y_3)}
 \end{align}
-where $p_i(\hat{\bf n})$ and $p(y)$ are our estimations of the prior
-distributions of particle orientations and instantaneous signals
-respectively.  The former can be estimated by some Boltzmann
-distribution based on how far apart reference orientations are in
-physical space, covered in Section~\ref{sec:2.3}.  The latter prior
-$p(y)$ is the probability of our dimer producing a signal
-$y_k(\bar{\bf s})$ that falls with our distribution of reference
-signals $y_k(\hat{\bf n})$.  This is given by taking the discrete
-integral over collection of reference orientations.
+where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior
+estimates of the distributions of particle orientations and
+instantaneous signals, respectively.  The former can be estimated by
+some Boltzmann distribution based on how far apart reference
+orientations are in physical space, covered in Section~\ref{sec:2.3}.
+The latter prior $p(y)$ is the probability of our dimer producing a
+signal $y_k(\bar{\bf s})$ that falls with our distribution of
+reference signals $y_k(\hat{\bf n})$.  This is given by taking the
+discrete integral over collection of reference orientations.
 \begin{align}
-  p(y)
+  p(y_1, y_2, y_3)
   =
-  \sum_{i=1}^{30} p_i(y(\bar{\bf s})|\hat{\bf n})p_i(\hat{\bf n})
+  \sum_{\alpha=1}^{n_{\rm ref}}
+  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
+  p(\hat{\bf n}_\alpha)
 \end{align}

### 2022-11-14 Manuscript.tex
fixing bibtex entry

@@
 \nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
-\nomenclature{$\hat{\bf n}_i$}{Reference Orientation i}
+\nomenclature{$\hat{\bf n}_\alpha$}{Reference Orientation $\alpha$}
 
@@
 our estimation of the dimer's instantaneous orientation.  The prior of
-the reference orientations $p_i(\hat{\bf n})$ is defined at each time
-step by accounting for the physical distance between our previous
-estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference
-orientation $\hat{\bf n}$
+the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each
+time step by accounting for the physical distance between the previous
+estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference
+orientation $\hat{\bf n}_\alpha$
 \begin{align}
-  r_i(t, \hat{\bf n}_i)
-  &= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1)
+  r_i(t, \hat{\bf n}_\alpha)
+  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
   \\
@@
   &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
-{\sum_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
+{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
 \end{align}
@@
 \begin{cases}
-	1 & \text{when $\hat{\bf n}_i$ = $\hat{\bf n}_{best}$}\\
+	1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
 	0 & \text{anywhere else}
@@
 \begin{align}
-  K_{l, \#}(p_i(\hat{\bf n}| y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber
+  K_{l, \#}(p(\hat{\bf n}_\alpha| y_1, y_2, y_3), p_{ideal})
+  =\nonumber
   \\
-&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y(\bar{\bf s}))}
+&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
 \right]
@@
 understood Evaluating every possible configuration manually is
-inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo
-simulation to our model. Ultra-nest is a sophisticated python package
-for fitting models with complex parameter correlations
+inefficient, therefore we utilised UltraNest to apply a Monte Carlo
+simulation to our model.  UltraNest is a Python package for fitting
+models with complex parameter correlations
 \cite{Buchner2016Ultranest}.  UltraNest implements a variation of
@@
 The Brownian OT software was used to simulate the motion of a trapped
-dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first $10$\,seconds of
-entering the optical trap.  The initial orientation was assumed as
-strictly vertical (in line with the beam propagation direction). The
-dimer's position orientation was recorded every $10\,{\rm \mu s}$ for
-using as a test dataset for our model.
+dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
+$10$\,seconds of entering the optical trap.  The initial orientation
+was assumed as strictly vertical (in line with the beam propagation
+direction). The dimer's position orientation was recorded every
+$10\,{\rm \mu s}$ for using as a test dataset for our model.
 
@@
 %
-In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT}
-simulations of trapped symmetrical dimers was investigated; their
-findings showed that the optical torque on the dimer goes to zero
-while aligned vertically and is at its maximum in a horizontal
-alignment.  Therefore, the inversion of an asymmetric dimer suggests
-that if the size difference is significant the optical torque goes is
-minimal for a dimer in both horizontal and vertical orientations. Once
-we have our experimental set up complete we can confirm this by
-trapping an asymmetric dimer and monitoring its orientation. The data
-from \figurename{ 2a} was used to test the efficacy of our predictive
-model. To visualise how well our model tracks the dimer's orientation
-we plot the radial distance between our model's prediction and the
-true orientation, and between the best possible orientation and the
-true orientation. For a perfect estimation these two plots should
-overlap, while we cannot achieve an exact estimation we can optimise
-it by adjusting the expected signal error and weighting factor.
+In the simulations of Vigilante~\emph{et~al}.\
+\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
+investigated; their findings showed that the optical torque on the
+dimer goes to zero while aligned vertically and is at its maximum in a
+horizontal alignment.  Therefore, the inversion of an asymmetric dimer
+suggests that if the size difference is significant the optical torque
+goes is minimal for a dimer in both horizontal and vertical
+orientations. Once we have our experimental set up complete we can
+confirm this by trapping an asymmetric dimer and monitoring its
+orientation.  The data from \figurename{ 2a} was used to test the
+efficacy of our predictive model.  To visualise how well our model
+tracks the dimer's orientation we plot the radial distance between our
+model's prediction and the true orientation, and between the best
+possible orientation and the true orientation.  For a perfect
+estimation these two plots should overlap, while we cannot achieve an
+exact estimation we can optimise it by adjusting the expected signal
+error and weighting factor.
 
@@
 \label{sec:3.3}
-We wanted to see if our choice of angles plays a significant role in our model's efficacy.
-We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
-We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
+
+We wanted to see if our choice of angles plays a significant role in
+our model's efficacy. We employed ultranest to analyse the parameter
+space for our three angles by sampling across the priori of
+$10^{\circ} \ \& \ 180^{\circ}$.   We kept both $\beta$ and $\sigma$
+constant during this so we could as these are parameters that can be
+changed more freely.
 
@@
 
-Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
-Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle.
-While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we can't achieve any greater accuracy than dividing the orientation space into 30 reference orientations.
-We hope to develop our model further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.
+Figure~\ref{fig:corner} is the resulting corner plot from our
+ultranest sampler, darker areas indicate a choice of angles where the
+model performed better where as lighter areas are where the model did
+poorer.  Overall from the corner plot provided, it is clear that there
+is no direct correlation between our method's performance and the
+choice of detection angle. While the choice of angles does not provide
+a direct improvement in our estimation it does determine how finely we
+can estimate the orientation, for choices of angles that are close
+together the minimum spacing between our signals is so small that we
+cannot achieve any greater accuracy than dividing the orientation
+space into $n_{\rm ref}=30$ reference orientations.  We hope to
+develop our method further in the future by being able to determine
+the maximum accuracy given that we know the signal error of our
+measurements.
 
-As covered in Section~\ref{sec:3.2} our choice for beta can help to improve our estimation by restricting it in the physical space.
-In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
+As covered in Section~\ref{sec:3.2} our choice for beta can help to
+improve our estimation by restricting it in the physical space. In
+addition depending on our choice of angles we are also setting the
+lower limit for our signal error $\epsilon$, so it may be preferable
+to choose a selection of angles that allows for a large signal error
+and then adjust the weighting factor to improve upon our estimation.
 

### 2022-11-14 Manuscript.tex
more updates

@@
 orientation $\bar{\bf s}$, is similarly scaled using
-Eq.~\eqref{eq:scale} to give $y(\bar{\bf s},\theta_k)$.
+Eq.~\eqref{eq:scale} to give $y_1(\bar{\bf s})$, $y_2(\bar{\bf s})$,
+and $y_k(\bar{\bf s})$.
 %
@@
 that the measurement uncertainty for each detection fibre is Gaussian
-and is centred about $y_k(\bar{\bf s})$ Combining all three detection
-angles provides us with the conditional probability that, should the
-dimer be in orientation $\hat{\bf n}_\alpha$, it would produce a
-signal $y(\bar{\bf s})$:
+and is centred about $y_k(\bar{\bf s})$.
+%
+Combining all three detection angles provides us with the conditional
+probability that, should the dimer be in orientation
+$\hat{\bf n}_\alpha$, it would produce a signal ($y_1$, $y_2$, $y_3$):
 \begin{align}
 \label{eq:conditional}
-  p(y(\bar{\bf s})|\hat{\bf n}_\alpha)
+  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
   = \prod^3_{k=1}
 (2\pi\sigma_k^2)^{-1/2} 
-e^{-(y_{k}(\hat{\bf n}_\alpha)-y_{k}(\bar{\bf s}))^2/2\sigma_k^2}
+e^{-(y_{k}-y_{k}(\hat{\bf n}_\alpha))^2/2\sigma_k^2}
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Parameter Optimisation Results}

### 2022-11-15 Manuscript_rev1.tex
Add files via upload

--- 2022-11-14 Manuscript.tex
@@
 \subsection{Brownian Simulation}
-\label{sec:2.1}
+\label{sec:brownian}
 
@@
 \subsection{Orientation estimation from scattering measurements}
-\label{sec:2.2}
+\label{sec:Bayes}
 
-We represent the instantaneous orientation of the moving dimer in the
-trap as a unit vector $\hat{\bf s}$, pointing from the centre of the
-larger sphere to the centre of the smaller sphere.  A plane wave
-``probe'' laser beam directed perpendicular to the trapping laser,
-incident on the dimer, generates a scattering pattern dependent on the
-dimer's instantaneous orientation, $I(\bar{\bf s},\theta_k)$.  This
-scattering pattern was computed using MSTM for each time step of the
-Brownian dynamics simulation.  To represent a simulated measurement,
-three angles ($\theta_1$, $\theta_2$, $\theta_3$) were chosen, and the
-calculated intensity at each angle $\theta_k$ is recorded as
-$I(\hat{\bf s}, \theta_k)$.
+We consider our dimer from Section~\ref{sec:brownian} in an optical trap; at any point in time we can draw a unit vector $\hat{s}$ that points from the centre of larger sphere, to the centre of the smaller sphere.
+A plane wave 'probe' laser, directly perpendicular to the trapping laser, is incident on the dimer generating a scattering pattern that is dependent upon the dimer's orientation $I(\hat{\bf s}, \theta)$, computed using MSTM.
+To represent the experimental set up, three angles  ($\theta_1, \ \theta_2, \ \& \theta_3$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$
 
-Our goal is to estimate the dimer orientation $\bar{\bf s}$ over time,
-based on the scattering measurements $I(\bar{\bf s},\ \theta_k)$: the
-question then is what set of angles ($\theta_1$, $\theta_2$,
-$\theta_3$) give us the most useful information on $\hat{\bf s}$.  To
-initially simplify the problem, rather than trying to get an exact
-estimation of $\hat{\bf s}$, we discretise the orientation into
-$n_{\rm ref}=30$ possible reference orientations $\hat{\bf n}_\alpha$
-that are evenly distributed on the unit sphere \cite{Rey2006}.
-%
-We then consider how to optimise an experiment to identify the
-reference orientation $\hat{\bf n}_\alpha$ that is closest to the
-actual orientation $\hat{\bf s}$ of the dimer.
-%
-The method can be extended to an arbitrarily large number of reference
-orientations, but for clarity and practicability we use a relatively
-small number here; additionally, the choice of a small number of
-discrete reference or ``candidate'' orientations also accounts for the
-fact that any experiment will necessarily have a limited precision.
-%
-Using MSTM we compute the scattering that would be produced by the
-dimer in each of the reference orientations, denoted as
-$I(\hat{\bf n}_\alpha, \theta)$ and obtain ``measurements''
-$I(\hat{\bf n}_\alpha, \theta_k)$ at our chosen set of angles
-($\theta_1$, $\theta_2$, $\theta_3$).
+Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \& \ \theta_3$.
+Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretised into $\textit{n}_ref \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
+Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
 %
@@
 where the denominator is simply the standard deviation for our values
-of $I(\hat{\bf n}, \theta_k)$.  The dimer's \emph{actual}
-instantaneous scattered intensity, based on the real instantaneous
-orientation $\bar{\bf s}$, is similarly scaled using
-Eq.~\eqref{eq:scale} to give $y_1(\bar{\bf s})$, $y_2(\bar{\bf s})$,
-and $y_k(\bar{\bf s})$.
-%
-The reference orientations, raw intensities, and scaled signals are
-given in Tables~\ref{tab:A1} and \ref{tab:A2}.
+of $I(\hat{\bf n}, \theta_k)$.
+
+The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
 %
 In practice, $y_k(\bar{\bf s})$ will have an inherent amount of
-uncertainty in our measurement due to noise from our detector.  We
-compare our instantaneous signal to our reference signal by assuming
-that the measurement uncertainty for each detection fibre is Gaussian
-and is centred about $y_k(\bar{\bf s})$.
+uncertainty in our measurement due to noise from our detector.
+We can compare any instantaneous signal $y_1, y_2, y_3$ to our reference signals by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf n})$.
 %
-Combining all three detection angles provides us with the conditional
-probability that, should the dimer be in orientation
+Combining the signals from all three detection angles provides us with the conditional probability that, should the dimer be in orientation
 $\hat{\bf n}_\alpha$, it would produce a signal ($y_1$, $y_2$, $y_3$):
@@
 where $\sigma_k$ is related to the uncertainty in the detected signal
-from our dimer $y_k(\bar{\bf s})$.  We left this as a tuneable
-parameter to characterize the influence of experimental error on our
-predictive method.  The signal error was simulated in our model by
-stochastically adjusting our values of $y_k(\bar{\bf s})$.
+from our dimer $y_k(\bar{\bf s})$.
+
+We left this as a tuneable parameter to characterize the influence of experimental error on our predictive method.
+The signal error was simulated in our model by stochastically adjusting our values of $y_k(\bar{\bf s})$.
 \begin{eqnarray*}
-  \sigma_k
-  &= \epsilon\bar{y}(\hat{\bf n}_k)
+  \sigma_k&=\epsilon\bar{y}(\hat{\bf n}_k)
   \\
-  \Rightarrow y(\bar{\bf s}_k) &= y(\bar{\bf s})_k \pm n\sigma_k
+  \Rightarrow y(\bar{\bf s}_k)&=y(\bar{\bf s})_k \pm n\sigma_k
   \\ 
-n \in \mathbb{R}, n\in[0,1]
+n \in \mathbb{R},&n\in[0,1]
 \end{eqnarray*} 
 where $\epsilon$ is the percent error expected from our detection
-fibres.  Equation~\eqref{eq:conditional} assigns a probability to each
-reference orientation based on how similar the ``reference orientation
-signal'' $y_{k}(\hat{\bf n}_\alpha)$ is to the actual signal
-$y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation
-$\bar{\bf s}$.  However, if our values of $y_k(\hat{\bf n}_\alpha)$
-are similar to one another, the probability distribution from
-Eq.~\eqref{eq:conditional} will therefore assign similar probability
-for multiple reference orientations.  Bayesian inference allows one to
-update a probability distribution by accounting for information known
-prior to making a measurement.  We apply this by accounting for our
-distribution of reference orientations and instantaneous signals.
+fibres.
+
+Equation~\eqref{eq:conditional} gives the probability that a signal
+($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference
+orientation $\hat{\rm n}_\alpha$.
+%
+Using Bayes' theorem, the probability that the dimer is in reference
+orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
+$y_3$) is measured, can be estimated:
+%.
 \begin{align}
@@
 estimates of the distributions of particle orientations and
-instantaneous signals, respectively.  The former can be estimated by
-some Boltzmann distribution based on how far apart reference
-orientations are in physical space, covered in Section~\ref{sec:2.3}.
-The latter prior $p(y)$ is the probability of our dimer producing a
-signal $y_k(\bar{\bf s})$ that falls with our distribution of
-reference signals $y_k(\hat{\bf n})$.  This is given by taking the
-discrete integral over collection of reference orientations.
+instantaneous signals, respectively.
+%
+Without any prior evidence we assume that the orientation priori of our dimer $p(\hat{\bf n}_{\alpha})$ is uniform.
+Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$ covered in Section~\ref{sec:test}.
+The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).
+This is given by taking the discrete integral over collection of reference orientations.
 \begin{align}
@@
 \end{align}
-The implementation of Eq.~\eqref{eq:bayes} improves our estimate by
-suppressing reference orientations that are far away from our dimer's
-orientation vector in physical space.  In order to evaluate the
-accuracy of our method, we tested it using a Brownian dynamics
-simulation of the dimer (see Section~\ref{sec:2.1}).
+From \eqref{eq:bayes} we get a mass probability distribution that denotes the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a signal ($y_1$, $y_2$, $y_3$).
+In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Calculation of error}
+\label{sec:divergence}
+Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate the orientation of a dimer based on its scattering signal.
+To evaluate the estimation we can use the simulation in Section~\ref{sec:brownian}; we denote the dimer's \emph{actual} orientation as $\hat{\bf s}$ and use MSTM to calculate its light scattering $I(\hat{\bf s}, \theta)$.
+At each angle $\theta_k$ we use \eqref{eq:scale} to scale the measurements collected from our dimer and record them as $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$.
+These measurements signals are used to calculate $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$, because we know $\hat{s}$ we can determine which reference orientation is closest, denoted as $\hat{\bf n}_{best}$
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Testing the Model}
-\label{sec:2.3}
+\label{sec:test}
 
-Using our simulation from Section~\ref{sec:2.1} we calculated the
+Using our simulation from Section~\ref{sec:brownian} we calculated the
 instantaneous signal from our dimer at each time step using MSTM.  We
-applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:2.2} ---
+applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} ---
 and the reference orientation with the highest probability is taken as
@@
 where $\beta$ is a weighting factor that controls the significance of
-our priori estimate.  The simulation data from Section~\ref{sec:2.1}
+our priori estimate. The simulation data from Section~\ref{sec:brownian}
 was used to check if the estimation was correct by comparing the
@@
   K_{l, \#}(p(\hat{\bf n}_\alpha| y_1, y_2, y_3), p_{ideal})
-  =\nonumber
-  \\
-&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
+  = p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
 \right]
@@
 \subsection{Optimization of parameters}
-\label{sec:2.4}
+\label{sec:ultranest}
 
@@
 \subsection{Asymmetric dimer dynamics}
-\label{sec:3.1}
+\label{sec:motion}
 
@@
 \begin{figure}[h]
-	\label{fig:motion}
 	\centering
@@
 	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
 \end{figure}
@@
 \subsection{Impact of signal error on model predictions}
-\label{sec:3.2}
+\label{sec:epsilon}
 
@@
 	\centering
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_5.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_at_5.png}
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_15.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_at_15.png}
 	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_25.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\textwidth]{./Images/epsilon_at_25.png}
 	\end{subfigure}
@@
 
-\begin{figure}
-	\label{fig:beta}
+\begin{figure}[h]
 	\centering
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{./Images/beta_5.png}
+	\begin{subfigure}{0.33\textwidth}
+		\includegraphics[width=\textwidth]{./Images/Beta_2.png}
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{./Images/beta_6.png}
+	\begin{subfigure}{0.33\textwidth}
+		\includegraphics[width=\textwidth]{./Images/Beta_3.png}
 	\end{subfigure}
 	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
+	\label{fig:beta}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 \subsection{Parameter Optimisation Results}
-\label{sec:3.3}
+\label{sec:ultranest_res}
 
@@
 \begin{figure}[h]
-	\label{fig:corner}
 	\centering
@@
 	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
+	\label{fig:corner}
 \end{figure}
@@
 
-As covered in Section~\ref{sec:3.2} our choice for beta can help to
+As covered in Section~\ref{sec:epsilon} our choice for beta can help to
 improve our estimation by restricting it in the physical space. In

### 2022-11-17 Manuscript_rev1.tex
Add files via upload

@@
 At each angle $\theta_k$ we use \eqref{eq:scale} to scale the measurements collected from our dimer and record them as $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$.
-These measurements signals are used to calculate $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$, because we know $\hat{s}$ we can determine which reference orientation is closest, denoted as $\hat{\bf n}_{best}$
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Testing the Model}
-\label{sec:test}
+These measurements signals are used to calculate $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$, because we know $\hat{s}$ we can determine which reference orientation is closest, denoted as $\hat{\bf n}_{best}$.
 
-Using our simulation from Section~\ref{sec:brownian} we calculated the
-instantaneous signal from our dimer at each time step using MSTM.  We
-applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} ---
-and the reference orientation with the highest probability is taken as
-our estimation of the dimer's instantaneous orientation.  The prior of
-the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each
-time step by accounting for the physical distance between the previous
-estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference
-orientation $\hat{\bf n}_\alpha$
-\begin{align}
-  r_i(t, \hat{\bf n}_\alpha)
-  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
-  \\
-  p_i(\hat{\bf n})
-  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
-{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
-\end{align}
-where $\beta$ is a weighting factor that controls the significance of
-our priori estimate. The simulation data from Section~\ref{sec:brownian}
-was used to check if the estimation was correct by comparing the
-estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation
-closest to the instantaneous orientation, denoted as
-$\hat{\bf n}_{best}$.  To measure the error in the model's estimate
-the distribution's of $p(\hat{\bf n}| y(\bar{\bf s}))$ and an
-ideal result, $p_ideal$, were compared.  An ideal distribution is one
-that is 0 for every $\hat{\bf n}$ apart from the correct result.
+To measure the error in the model's estimate the distribution from Eq.~\eqref{eq:bayes} and an ideal result, $p_ideal$, were compared.
+An ideal result would be one that is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$.
 \begin{align}
@@
 measured by calculating the Kullback-Leibler divergence $K_l$ between
-the two distributions
+the two distributions.
 \begin{align}
-  K_{l, \#}(p(\hat{\bf n}_\alpha| y_1, y_2, y_3), p_{ideal})
-  = p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
+	K_{l, \#}(p_{ideal}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
+	= p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}| y_1,y_2,y_3))}
 	\right]
+	\label{eq;kullback}
 \end{align}
-where a more negative value of $K_l$ indicates a poorer estimation of
-the dimer's instantaneous orientation.  By summing the divergence of
-each measurement across the entire simulation we get an evaluation of
-how well the model performed in estimating the dimer's orientation.
-To compare between different iterations of our model the we compare
-our result of $K_l$ to a worst case scenario and evaluate how much it
-improves upon this, denoted as $F(K_l)$:
+where a larger value of $K_l$ indicates that our model is less confident in its prediction of the dimer's orientation.
+The divergence provides us a measure of how confident our measurement is.
+To better reflect the model's accuracy we implemented a penalty condition that worsens the divergence if our model estimates the wrong reference orientation:
+\begin{align}
+	p(\hat{\bf n}|y_1,y_2,y_3) = 
+	\begin{cases}
+		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3) & 
+		\text{when $\hat{\bf n}_{estimate}$ = $\hat{\bf n}_{best}$}\\
+		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3)-p(\hat{\bf n}_{best}|y_1,y_2,y_3) & \text{when $\hat{\bf n}_{estimate} \ne \hat{\bf n}_{best}$}
+	\end{cases}
+\end{align}
+
+This allows us to account for situations when our collected signal is equally likely to be associated with 2 reference signals.
+Section~\ref{sec:Bayes} and Section~\ref{sec:divergence} are all for a single measurement of a light scattering signal from our dimer.
+To properly model it's motion we need to consider how the orientation changes with time.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing the Model}
+\label{sec:test}
+
+Using our simulation from Section~\ref{sec:brownian} we calculated the instantaneous signal from our dimer at each time step using MSTM.
+We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
+To visualise our model's estimation we plotted the distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ with time.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is rather poor, as seen in.
+The prior of the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each time step by accounting for the physical distance between the previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$
+\begin{align}
+  r_i(t, \hat{\bf n}_\alpha)
+  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
+  \\
+  p_i(\hat{\bf n})
+  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
+{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
+\end{align}
+where $\beta$ is a weighting factor that controls the significance of our priori estimate.
+The simulation data from Section~\ref{sec:brownian} was used to check if the estimation was correct by comparing the estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as
+$\hat{\bf n}_{best}$.  By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
+To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
 \begin{align}
 K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
-K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
+K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right] \\
 F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
@@
 we are under this line then we know our model is just off from
-selecting the correct orientation
+selecting the correct orientation:
 

### 2022-11-22 Manuscript_rev1.tex
Add files via upload

@@
 
-\title{Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes}
+\title{Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles}
 
@@
 We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
-To visualise our model's estimation we plotted the distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ with time.
-If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is rather poor, as seen in.
-The prior of the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each time step by accounting for the physical distance between the previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$
+To visualise our model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time.
+For comparison, we also plotted the radial distance between the closest reference orientation, denoted $\hat{\bf n}_{best}$, and the dimer's instantaneous orientation.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is rather poor, as seen in \ref{fig:uniform}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth]{./Images/uniform_pn.png}
+	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.}
+	\label{fig:uniform}
+\end{figure}
+
+The reasoning for this lies with the fact that differences in reference signals and reference orientation spacing are not directly correlated with each other.
+To remedy this we need to account for the fact that a trapped dimer's motion is severely limited due to the trap stiffness.
+Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:
+
 \begin{align}
   r_i(t, \hat{\bf n}_\alpha)
-  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
+  &= (\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_\alpha(t-\Delta t))
   \\
@@
 \end{align}
+
 where $\beta$ is a weighting factor that controls the significance of our priori estimate.
-The simulation data from Section~\ref{sec:brownian} was used to check if the estimation was correct by comparing the estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as
-$\hat{\bf n}_{best}$.  By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
-To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
+The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}.
+By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
+To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
 \begin{align}
@@
 
-The worst case scenario is akin to randomly choosing a reference
-orientation at each time step. The greater the value of $F(K_l)$ the
-lower the error between our estimated orientation and the dimer's
-instantaneous orientation. Because our model is dependent on several
-parameters we need to a sophisticated method for understanding how
-these parameters correlate with $F(K_l)$.
+The worst case scenario is akin to randomly choosing a reference orientation at each time step.
+The greater the value of $F(K_l)$ the better our model fits the dimer's motion.
+Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.
 

### 2022-11-22 Manuscript_rev1.tex
Add files via upload

@@
 \end{align}
-The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
-probability to every reference orientation.  The model's error was
-measured by calculating the Kullback-Leibler divergence $K_l$ between
-the two distributions.
+The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero probability to every reference orientation.
+The model's error was measured by calculating the Kullback-Leibler divergence $K_l$ between the two distributions.
 \begin{align}
@@
 For comparison, we also plotted the radial distance between the closest reference orientation, denoted $\hat{\bf n}_{best}$, and the dimer's instantaneous orientation.
-If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is rather poor, as seen in \ref{fig:uniform}.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in \ref{fig:uniform}.
 
@@
 The reasoning for this lies with the fact that differences in reference signals and reference orientation spacing are not directly correlated with each other.
+In some cases, two reference orientation will have identical signal responses.
 To remedy this we need to account for the fact that a trapped dimer's motion is severely limited due to the trap stiffness.
@@
 
-The signal error $\epsilon$ and the weighting factor $\beta$ are known
-quantities whose influence can be measured directly; however, our
-choice of angles and how they correlate with $F(K_l)$ is not easily
-understood Evaluating every possible configuration manually is
-inefficient, therefore we utilised UltraNest to apply a Monte Carlo
-simulation to our model.  UltraNest is a Python package for fitting
-models with complex parameter correlations
-\cite{Buchner2016Ultranest}.  UltraNest implements a variation of
-Monte Carlo integration, nested-sampling, where a likelihood contour
-is used to update a group of live points chosen from the prior
-estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
+The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly; however, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised UltraNest to apply a Monte Carlo simulation to our model.
+UltraNest is a Python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}.
+UltraNest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
 
-The UltraNest software was used to create a set of live points of
-different choices of $\theta_1$, $\theta_2$, and $\theta_3$, with each
-point being evaluated by its improvement factor ($F(K_l)$).  After
-evaluating every live point, the worst results are removed and new
-live points are created based on the information provided by the
-previous live points.  If there is a correlation between the choice of
-angles and $F(K_l)$ then the live points should congregate around a
-single (or multiple) choice of angles that will maximise $F(K_l)$ and
-therefore minimise the error between our estimate and instantaneous
-orientation.
+The UltraNest software was used to create a set of live points of different choices of $\theta_1$, $\theta_2$, and $\theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$).
+After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
+If there is a correlation between the choice of angles and $F(K_l)$ then the live points should congregate around a single (or multiple) choice of angles that will maximise $F(K_l)$ and therefore minimise the error between our estimate and instantaneous orientation.
 
@@
 \label{sec:Discussion}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \hline\hline
-Reference Orientation & $\hat{\bf n}_{i,\ x}$ &  $\hat{\bf n}_{i,\ y}$ &  $\hat{\bf n}_{i,\ z}$ \\
+Reference Orientation & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
 \hline
@@
 $3$  & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-$4$ & $1.0000$ &  $0.00000$ & $0.00000$ \\
-$5$ & $0.00000$ &  $1.0000$ & $0.00000$ \\
-$6$ & $0.00000$ &  $0.00000$ & $1.0000$ \\
+$4$  & $ 1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
+$5$  & $ 0.00000$ &  $ 1.00000$ & $ 0.00000$ \\
+$6$  & $ 0.00000$ &  $ 0.00000$ & $ 1.00000$ \\
 $7$  & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
@@
 $9$  & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
-$10$ & $0.00000$ &  $0.00000$ & $-1.0000$ \\
+$10$ & $ 0.00000$ &  $ 0.00000$ & $-1.00000$ \\
 $11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
@@
 $13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
-$14$ & $0.00000$ &  $-1.0000$ & $0.00000$ \\
+$14$ & $ 0.00000$ &  $ -1.0000$ & $ 0.00000$ \\
 $15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
@@
 $20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-$21$ & $-1.0000$ &  $0.00000$ & $0.00000$ \\
+$21$ & $-1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
 $22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
@@
       \hline\hline
-      Reference Orientation  & $I(\hat{\bf n}_i,\ 15^{\circ})$ &  $I(\hat{\bf n}_i, \ 45^{\circ})$ &  $I(\hat{\bf n}_i,\ 90^{\circ})$ & $y(\hat{\bf n}_i, \ 15^{\circ})$ & $y(\hat{\bf n}_i, \ 45^{\circ})$ & $y(\hat{\bf n}_i, \ 90^{\circ})$ \\
+      Reference Orientation  & $I(\hat{\bf n}_\alpha\ ,\ 15^{\circ})$ &  $I(\hat{\bf n}_\alpha \ , \ 45^{\circ})$ &  $I(\hat{\bf n}_\alpha \ ,\ 90^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 15^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 45^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 90^{\circ})$ \\
 \hline
 $1$  & $13.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
-$2$ & $13.237$ & $0.0272$ & $0.0102$ & $0.2429$ & $-1.401$ & $-0.7826$ \\
-$3$ & $15.840$ & $0.1072$ & $0.0234$ & $1.337$ &$-0.6142$&	$1.162$ \\
-$4$ & $12.657$ & $0.2246$ & $0.0244$ & $-0.0012$ &	$0.5399$ & $0.9695$ \\
-$5$ & $15.351$ & $0.0527$ & $0.0221$ & $1.132$ &	$-1.150$ & $1.303$ \\
+$2$  & $13.237$ & $0.0272$ & $0.0102$ & $ 0.2429$ & $-1.4010$ & $-0.7826$ \\
+$3$  & $15.840$ & $0.1072$ & $0.0234$ & $ 1.3370$ & $-0.6142$ & $ 1.1620$ \\
+$4$  & $12.657$ & $0.2246$ & $0.0244$ & $-0.0012$ &	$ 0.5399$ & $ 0.9695$ \\
+$5$  & $15.351$ & $0.0527$ & $0.0221$ & $ 1.1320$ & $-1.1500$ & $ 1.3030$ \\
 $6$  & $12.046$ & $0.0916$ & $0.0142$ & $-0.2584$ &	$-0.7676$ &	$-0.1886$ \\
-$7$ & $11.376$ & $0.0036$ & $0.0095$ & $-0.5403$ &$-1.633$ & $-0.8792$ \\
-$8$ & $13.130$ & $0.2346$ & $0.0095$ & $0.1977$ & $0.6379$ & $-0.8794$ \\
-$9$ & $15.975$ & $0.2893$ & $0.0273$ & $1.38724$ & $1.175$ & $1.726$ \\
-$10$ & $10.073$ & $0.2933$ & $0.0221$ & $-1.088$	& $1.215$ & $0.9641$ \\
+$7$  & $11.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.6330$ & $-0.8792$ \\
+$8$  & $13.130$ & $0.2346$ & $0.0095$ & $ 0.1977$ & $ 0.6379$ & $-0.8794$ \\
+$9$  & $15.975$ & $0.2893$ & $0.0273$ & $ 1.3872$ & $ 1.1750$ & $ 1.7260$ \\
+$10$ & $10.073$ & $0.2933$ & $0.0221$ & $-1.0880$ & $ 1.2150$ & $ 0.9641$ \\
 $11$ & $11.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
-$12$ & $13.238$ & $0.0272$ & $0.0102$ & $0.2429$ &	$-1.401$ & $-0.7826$ \\
-$13$ & $15.840$ & $0.1072$ & $0.0234$ & $1.337$ &	$-0.6142$ & $1.162$ \\
-$14$ & $15.351$ & $0.0527$ & $0.0244$ & $1.132$ & $-1.150$ & $1.303$ \\
-$15$ & $13.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.633$ & $-0.8792$ \\
-$16$ & $13.130$ & $0.2346$ & $0.0095$ & $0.1977$ & $0.6379$ & $-0.8794$ \\
-$17$ & $15.957$ & $0.2893$ & $0.0273$ & $1.387$	& $1.175$	& $1.726$ \\
-$18$ & $14.199$ & $0.0935$ & $0.0076$ & $0.6472$ & $-0.7496$ & $-1.158$ \\
-$19$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.743$ & $1.411$ & $-1.157$ \\
-$20$ & $12.961$ & $0.2420$ & $0.0188$ & $0.1267$ & $0.7113$ & $0.4765$ \\
-$21$ & $8.2160$ & $0.3217$ & $0.0142$ & $-1.869$ & $1.495$ & $-0.1890$ \\
-$22$ & $14.212$ & $0.2537$ & $0.0102$ & $0.6529$ & $0.8265$ & $-0.7835$ \\
-$23$ & $8.9280$ & $0.1398$ & $0.0123$ & $-1.570$ &	$-0.2934$ & $-0.4840$ \\
-$24$ & $13.328$ & $0.1964$ & $0.0234$ & $0.2807$ & $0.2624$ & $1.162$ \\
-$25$ & $14.199$ & $0.0935$ & $0.0076$ & $0.6472$ & $-0.7496$ & $-1.1580$ \\
-$26$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.743$	& $1.411$ & $-1.157$ \\
-$27$ & $12.961$ & $0.2420$ & $0.0188$ & $0.1267$	& $0.7113$ & $0.4765$ \\
-$28$ & $14.212$ & $0.2537$ & $0.0102$ & $0.6529$ & $0.8265$ & $-0.7835$ \\
-$29$ & $8.9280$ & $0.1398$ & $0.0122$ & $-1.570$ &	$-0.2934$ & $-0.4840$ \\
-$30$ & $13.328$ & $0.1964$ & $0.0234$ & $0.2807$ & $0.2624$ & $1.162$ \\
+$12$ & $13.238$ & $0.0272$ & $0.0102$ & $ 0.2429$ &	$-1.4010$ & $-0.7826$ \\
+$13$ & $15.840$ & $0.1072$ & $0.0234$ & $ 1.3370$ & $-0.6142$ & $ 1.1620$ \\
+$14$ & $15.351$ & $0.0527$ & $0.0244$ & $ 1.1320$ & $-1.1500$ & $ 1.3030$ \\
+$15$ & $13.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.6330$ & $-0.8792$ \\
+$16$ & $13.130$ & $0.2346$ & $0.0095$ & $ 0.1977$ & $ 0.6379$ & $-0.8794$ \\
+$17$ & $15.957$ & $0.2893$ & $0.0273$ & $ 1.3870$ & $ 1.1750$ & $ 1.7260$ \\
+$18$ & $14.199$ & $0.0935$ & $0.0076$ & $ 0.6472$ & $-0.7496$ & $-1.1580$ \\
+$19$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.7430$ & $ 1.4110$ & $-1.1570$ \\
+$20$ & $12.961$ & $0.2420$ & $0.0188$ & $ 0.1267$ & $ 0.7113$ & $ 0.4765$ \\
+$21$ & $8.2160$ & $0.3217$ & $0.0142$ & $-1.8690$ & $ 1.4950$ & $-0.1890$ \\
+$22$ & $14.212$ & $0.2537$ & $0.0102$ & $ 0.6529$ & $ 0.8265$ & $-0.7835$ \\
+$23$ & $8.9280$ & $0.1398$ & $0.0123$ & $-1.5700$ &	$-0.2934$ & $-0.4840$ \\
+$24$ & $13.328$ & $0.1964$ & $0.0234$ & $ 0.2807$ & $ 0.2624$ & $ 1.1620$ \\
+$25$ & $14.199$ & $0.0935$ & $0.0076$ & $ 0.6472$ & $-0.7496$ & $-1.1580$ \\
+$26$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.7430$ & $ 1.4110$ & $-1.1570$ \\
+$27$ & $12.961$ & $0.2420$ & $0.0188$ & $ 0.1267$ & $ 0.7113$ & $ 0.4765$ \\
+$28$ & $14.212$ & $0.2537$ & $0.0102$ & $ 0.6529$ & $ 0.8265$ & $-0.7835$ \\
+$29$ & $8.9280$ & $0.1398$ & $0.0122$ & $-1.5700$ & $-0.2934$ & $-0.4840$ \\
+$30$ & $13.328$ & $0.1964$ & $0.0234$ & $ 0.2807$ & $ 0.2624$ & $ 1.1620$ \\
 \hline\hline

### 2022-11-29 Manuscript_rev1.tex
Add files via upload

@@
 shaped sphere clusters.  We simulated the motion of silica dimer
-trapped in a $5$\,mW Gaussian beam; the dimer was comprised of two
+trapped in a $5$ mW Gaussian beam; the dimer was comprised of two
 spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$,
@@
 For comparison, we also plotted the radial distance between the closest reference orientation, denoted $\hat{\bf n}_{best}$, and the dimer's instantaneous orientation.
+The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another.
 If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in \ref{fig:uniform}.
@@
 of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
-to $1$.  The dotted line indicates the minimum radian distance
-($0.137$ radians) between two neighbouring reference orientations, if
-we are under this line then we know our model is just off from
-selecting the correct orientation:
+to $1$:
 
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, there is seemingly a
+As can be seen from Figure~\ref{fig:epsilon}, there is a
 sweet spot for the signal error; with a $5$\% error giving an
@@
 
-The latter issue of a high signal error can be addressed somewhat by
-adjusting our value of $\beta$ to flatten our probability
-distribution. We repeated our model estimation at $25\%$ at different
-values of $\beta$ to see the upper limit we could improve our result
-to.
+The latter issue of a high signal error can be addressed somewhat byadjusting our value of $\beta$ to flatten our probability distribution.
+We repeated our model estimation at $25\%$ at different values of $\beta$ to see the upper limit we could improve our result to.
 
@@
 
-As shown in Figure~\ref{fig:beta}, eventually $\beta$ begins to over
-correct our estimate to the extent that we are a full radian away from
-the correct result.  The exact value of $\beta$ that will optimise our
-performance varies for each choice of angle, therefore we can use it
-as variable for finely tuning our results to suit our choice of angles
-and our signal error.  In future, we want to also investigate the
-possibility of extending our prior $p(\hat{\bf n})$ so we account for not
-just the change in orientation after one time step but after 2 or 3
-time steps to investigate if we can investigate the auto correlation
-of our dimer's motion.
+As shown in Figure~\ref{fig:beta}, eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
+The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error.
+In future, we want to also investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate the auto correlation function of our dimer's motion.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
 \subsection{Parameter Optimisation Results}
@@
 
-Figure~\ref{fig:corner} is the resulting corner plot from our
-ultranest sampler, darker areas indicate a choice of angles where the
-model performed better where as lighter areas are where the model did
-poorer.  Overall from the corner plot provided, it is clear that there
-is no direct correlation between our method's performance and the
-choice of detection angle. While the choice of angles does not provide
-a direct improvement in our estimation it does determine how finely we
-can estimate the orientation, for choices of angles that are close
-together the minimum spacing between our signals is so small that we
-cannot achieve any greater accuracy than dividing the orientation
-space into $n_{\rm ref}=30$ reference orientations.  We hope to
-develop our method further in the future by being able to determine
-the maximum accuracy given that we know the signal error of our
-measurements.
+Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
+Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the
+choice of detection angle.
+While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we cannot achieve any greater accuracy than dividing the orientation space into $n_{\rm ref}=30$ reference orientations.
+We hope to develop our method further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.
 
-As covered in Section~\ref{sec:epsilon} our choice for beta can help to
-improve our estimation by restricting it in the physical space. In
-addition depending on our choice of angles we are also setting the
-lower limit for our signal error $\epsilon$, so it may be preferable
-to choose a selection of angles that allows for a large signal error
-and then adjust the weighting factor to improve upon our estimation.
+As covered in Section~\ref{sec:epsilon} our choice for beta can help to improve our estimation by restricting it in the physical space.
+In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
+
+This result only indicates that orientation is not a characteristic that is angularly biased in its scattering.
+It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.
 

### 2022-12-05 Manuscript_rev1.tex
Add files via upload

@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
+	\includegraphics[width=0.9\textwidth]{./Images/fig2.png}
 	\caption{Example dimer in orientation $\bar{\bf s}$}
@@
 optical force and torque components.  Brownian OT returns the dimer's
-motion as a \textit{csv} file containing for each timestep the
+motion as a \textit{csv} file containing for each time step the
 $[x,y,z]$ position of the dimer's centre of mass relative to the trap
@@
 Without any prior evidence we assume that the orientation priori of our dimer $p(\hat{\bf n}_{\alpha})$ is uniform.
-Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$ covered in Section~\ref{sec:test}.
+Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.
 The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).
@@
 The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another.
-If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in \ref{fig:uniform}.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in Figure~\ref{fig:uniform}.
 
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{./Images/uniform_pn.png}
+	\includegraphics[width=0.5\textwidth]{./Images/fig3.png}
 	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.}
@@
 In some cases, two reference orientation will have identical signal responses.
-To remedy this we need to account for the fact that a trapped dimer's motion is severely limited due to the trap stiffness.
+To remedy this we need to account for the fact that a trapped dimer's motion i limited due to the trap stiffness.
 Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:
@@
 To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
+
 \begin{align}
@@
 
-The Brownian OT software was used to simulate the motion of a trapped
-dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
-$10$\,seconds of entering the optical trap.  The initial orientation
-was assumed as strictly vertical (in line with the beam propagation
-direction). The dimer's position orientation was recorded every
-$10\,{\rm \mu s}$ for using as a test dataset for our model.
+The Brownian OT software was used to simulate the motion of a trappeddimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
+$10$ seconds of entering the optical trap.
+The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
+The dimer's position orientation was recorded every $10\,{\rm \mu s}$ for using as a test dataset for our model.
 
@@
 		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/traj.png}
+		\includegraphics[width =\textwidth]{./Images/fig4a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/pos.png}
+		\includegraphics[width=\textwidth]{./Images/fig4b.png}
 	\end{subfigure}
@@
 
-As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a
-full $180^{\circ}$ rotation upon entering the trap.  Typically
-horizontal alignment of a dimer is unstable and will result in the
-particle rotating to align along its vertical axis.  It is interesting
-to note that the dimer is furthest from the trap centre as it goes
-into a horizontal orientation before drawing closer again as it
-inverts completely. Further simulations of differently sized dimers
-showed similar results, but only when $a_1 \geq 2a_2$.  Dimers with
-more symmetrical size ratios immediately aligned into a fixed vertical
-position.
+As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
+Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
+It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
+Further simulations of differently sized dimers showed sa similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.
+Dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
 %
-In the simulations of Vigilante~\emph{et~al}.\
-\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
-investigated; their findings showed that the optical torque on the
-dimer goes to zero while aligned vertically and is at its maximum in a
-horizontal alignment.  Therefore, the inversion of an asymmetric dimer
-suggests that if the size difference is significant the optical torque
-goes is minimal for a dimer in both horizontal and vertical
-orientations. Once we have our experimental set up complete we can
-confirm this by trapping an asymmetric dimer and monitoring its
-orientation.  The data from \figurename{ 2a} was used to test the
-efficacy of our predictive model.  To visualise how well our model
-tracks the dimer's orientation we plot the radial distance between our
-model's prediction and the true orientation, and between the best
-possible orientation and the true orientation.  For a perfect
-estimation these two plots should overlap, while we cannot achieve an
-exact estimation we can optimise it by adjusting the expected signal
-error and weighting factor.
+In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
+Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
+The data from Figure~\ref{fig:motion} was used to test the efficacy of our predictive model.
+To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
+For a perfect estimation these two plots should overlap, while we cannot achieve an exact estimation we can optimise it by adjusting the expected signal error and weighting factor.
 
@@
 
-A key assumption of Eq.~\eqref{eq:conditional} is that we assume our
-detected light signal will inherently have some degree of noise. This
-is the central premise that allows us to form
-Eq.~\eqref{eq:conditional}, understanding the limits of acceptable
-signal error is necessary for implementing this to experimental work.
-To this end we ran our model while varying the value of $\epsilon$.
-\figurename{ 3} shows the performance of the method with a selection
-of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
-to $1$:
+A key assumption of Eq.~\eqref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise.
+This is the central premise that allows us to form Eq.~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
+To this end we ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
 
@@
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_at_5.png}
+		\includegraphics[width=\textwidth]{./Images/fig5a.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_at_15.png}
+		\includegraphics[width=\textwidth]{./Images/fig5b.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/epsilon_at_25.png}
+		\includegraphics[width=\textwidth]{./Images/fig5c.png}
 	\end{subfigure}
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, there is a
-sweet spot for the signal error; with a $5$\% error giving an
-inaccurate estimation, and with a $25$\% error giving an alright
-estimation but with sudden shifts away from the ideal result. The
-reason why our low error estimation performed so poorly is that the
-minimum difference between our values of $y_k$ is larger than our
-signal error. As such the model could not distinguish between
-different reference orientations and was essentially just picking
-options at random. This minimum difference determined by our choice of
-angles and sets a lower limit of how accurate our signals can be. In
-the future we hope to look at increasing the number of reference
-orientations to see what the true lower bound is.
+As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away from the ideal result.
+A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.
+As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
+This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
+In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.
 
-The latter issue of a high signal error can be addressed somewhat byadjusting our value of $\beta$ to flatten our probability distribution.
-We repeated our model estimation at $25\%$ at different values of $\beta$ to see the upper limit we could improve our result to.
+The latter issue of a high signal error can be addressed somewhat by adjusting our value of $\beta$ to flatten our probability distribution.
+We repeated our model estimation at $25\%$ with different values of $\beta$ to see the upper limit we could improve our result to.
 
@@
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/Beta_2.png}
+		\includegraphics[width=\textwidth]{./Images/fig6a.png}
 	\end{subfigure}
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/Beta_3.png}
+		\includegraphics[width=\textwidth]{./Images/fig6b.png}
 	\end{subfigure}
@@
 
-We wanted to see if our choice of angles plays a significant role in
-our model's efficacy. We employed ultranest to analyse the parameter
-space for our three angles by sampling across the priori of
-$10^{\circ} \ \& \ 180^{\circ}$.   We kept both $\beta$ and $\sigma$
-constant during this so we could as these are parameters that can be
-changed more freely.
+We wanted to see if our choice of angles plays a significant role in our model's efficacy.
+We employed UltraNest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
+We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
 
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{./Images/corneranglesfreed-1.png}
+	\includegraphics[width=0.45\textwidth]{./Images/fig7.png}
 	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
@@
 
-Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
-Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the
-choice of detection angle.
+Figure~\ref{fig:corner} is the resulting corner plot from our UltraNest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
+Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the choice of detection angle.
+This result only indicates that orientation is not a characteristic that is angularly biased in its scattering.
+It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.
+
 While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we cannot achieve any greater accuracy than dividing the orientation space into $n_{\rm ref}=30$ reference orientations.
@@
 In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
-
-This result only indicates that orientation is not a characteristic that is angularly biased in its scattering.
-It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We have developed a method for interpreting the dynamics of a trapped
-particle based purely on the light scattering pattern. In theory the
-model can be applied to any characteristic that impacts the light
-scattering pattern produced by the trapped entity. The MSTM package is
-flexible in calculating the light scattering for a variety of
-micro-particles; so long as there is a distinguishable difference in
-the light scattering pattern we can apply this method to characterise
-the particle to some degree of accuracy. This has potential to be
-applied to characterise particle size and shape that are changing
-while in the optical trap. The authors would like to acknowledge and
-thank the support for this research from the funding provided by the
-Leverhulme Trust.\\
+We have developed a method for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
+In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
+The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
+This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
+We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.\\
 
@@
 \renewcommand{\thetable}{A\arabic{table}}
-\label{sec:App1}
 \begin{table}[h]
@@
 	\end{center}
+\label{sec:App1}
 \end{table}
@@
 \hline
-$1$  & $13.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
-$2$  & $13.237$ & $0.0272$ & $0.0102$ & $ 0.2429$ & $-1.4010$ & $-0.7826$ \\
-$3$  & $15.840$ & $0.1072$ & $0.0234$ & $ 1.3370$ & $-0.6142$ & $ 1.1620$ \\
-$4$  & $12.657$ & $0.2246$ & $0.0244$ & $-0.0012$ &	$ 0.5399$ & $ 0.9695$ \\
-$5$  & $15.351$ & $0.0527$ & $0.0221$ & $ 1.1320$ & $-1.1500$ & $ 1.3030$ \\
-$6$  & $12.046$ & $0.0916$ & $0.0142$ & $-0.2584$ &	$-0.7676$ &	$-0.1886$ \\
-$7$  & $11.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.6330$ & $-0.8792$ \\
-$8$  & $13.130$ & $0.2346$ & $0.0095$ & $ 0.1977$ & $ 0.6379$ & $-0.8794$ \\
-$9$  & $15.975$ & $0.2893$ & $0.0273$ & $ 1.3872$ & $ 1.1750$ & $ 1.7260$ \\
-$10$ & $10.073$ & $0.2933$ & $0.0221$ & $-1.0880$ & $ 1.2150$ & $ 0.9641$ \\
-$11$ & $11.371$ & $0.1264$ & $0.0122$ & $-0.5423$ & $-0.4253$ & $-0.4838$ \\
-$12$ & $13.238$ & $0.0272$ & $0.0102$ & $ 0.2429$ &	$-1.4010$ & $-0.7826$ \\
-$13$ & $15.840$ & $0.1072$ & $0.0234$ & $ 1.3370$ & $-0.6142$ & $ 1.1620$ \\
-$14$ & $15.351$ & $0.0527$ & $0.0244$ & $ 1.1320$ & $-1.1500$ & $ 1.3030$ \\
-$15$ & $13.376$ & $0.0036$ & $0.0095$ & $-0.5403$ & $-1.6330$ & $-0.8792$ \\
-$16$ & $13.130$ & $0.2346$ & $0.0095$ & $ 0.1977$ & $ 0.6379$ & $-0.8794$ \\
-$17$ & $15.957$ & $0.2893$ & $0.0273$ & $ 1.3870$ & $ 1.1750$ & $ 1.7260$ \\
-$18$ & $14.199$ & $0.0935$ & $0.0076$ & $ 0.6472$ & $-0.7496$ & $-1.1580$ \\
-$19$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.7430$ & $ 1.4110$ & $-1.1570$ \\
-$20$ & $12.961$ & $0.2420$ & $0.0188$ & $ 0.1267$ & $ 0.7113$ & $ 0.4765$ \\
-$21$ & $8.2160$ & $0.3217$ & $0.0142$ & $-1.8690$ & $ 1.4950$ & $-0.1890$ \\
-$22$ & $14.212$ & $0.2537$ & $0.0102$ & $ 0.6529$ & $ 0.8265$ & $-0.7835$ \\
-$23$ & $8.9280$ & $0.1398$ & $0.0123$ & $-1.5700$ &	$-0.2934$ & $-0.4840$ \\
-$24$ & $13.328$ & $0.1964$ & $0.0234$ & $ 0.2807$ & $ 0.2624$ & $ 1.1620$ \\
-$25$ & $14.199$ & $0.0935$ & $0.0076$ & $ 0.6472$ & $-0.7496$ & $-1.1580$ \\
-$26$ & $8.5164$ & $0.3133$ & $0.0076$ & $-1.7430$ & $ 1.4110$ & $-1.1570$ \\
-$27$ & $12.961$ & $0.2420$ & $0.0188$ & $ 0.1267$ & $ 0.7113$ & $ 0.4765$ \\
-$28$ & $14.212$ & $0.2537$ & $0.0102$ & $ 0.6529$ & $ 0.8265$ & $-0.7835$ \\
-$29$ & $8.9280$ & $0.1398$ & $0.0122$ & $-1.5700$ & $-0.2934$ & $-0.4840$ \\
-$30$ & $13.328$ & $0.1964$ & $0.0234$ & $ 0.2807$ & $ 0.2624$ & $ 1.1620$ \\
+$1$  & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
+$2$  & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
+$3$  & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
+$4$  & $4.5187$ & $0.0459$ & $0.0221$ & $-0.9770$ & $ 0.2062$ & $ 0.9695$ \\
+$5$  & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
+$6$  & $7.1799$ & $0.0377$ & $0.0142$ & $ 0.5458$ & $-0.0383$ & $-0.1886$ \\
+$7$  & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
+$8$  & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
+$9$  & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
+$10$ & $4.6351$ & $0.1536$ & $0.0221$ & $-0.9103$ & $ 3.4040$ & $ 0.9641$ \\
+$11$ & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
+$12$ & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
+$13$ & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
+$14$ & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
+$15$ & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
+$16$ & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
+$17$ & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
+$18$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
+$19$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
+$20$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
+$21$ & $4.2902$ & $0.0103$ & $0.0142$ & $-1.1077$ & $-0.8510$ & $-0.1890$ \\
+$22$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 1.8281$ & $-0.7835$ \\
+$23$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
+$24$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
+$25$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
+$26$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
+$27$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
+$28$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 0.1005$ & $ 0.0102$ \\
+$29$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
+$30$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
 \hline\hline
@@
 \end{center}
+\label{sec:App2}
 \end{table}

### 2022-12-06 Manuscript_rev1.tex
Add files via upload

@@
 \usepackage{nomencl}
+\makenomenclature
 
@@
 
-\nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
-\nomenclature{$\hat{\bf n}_\alpha$}{Reference Orientation $\alpha$}
-
-%\linenumbers
+\linenumbers
 
@@
 of trapped non-spherical particles till the recent attempt in
-Ref.~\citenum{raudsepp2022estimating} where imaging was employed to
+Ref.~\cite{raudsepp2022estimating} where imaging was employed to
 study the orientation of trapped dimers and a resolution of () was
@@
 
+\printnomenclature
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We use the Brownian OT package developed by Fung~\textit{et~al}.
-\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric
-dimer within an optical trap.  Brownian OT combines MSTM
-\cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox''
-(\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
-shaped sphere clusters.  We simulated the motion of silica dimer
-trapped in a $5$ mW Gaussian beam; the dimer was comprised of two
-spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$,
-respectively.  We ran our simulation over the span of 10 seconds, with
-Brownian OT calculating the dimer's orientation and position every
-$1 \ ms$.
+\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap.
+Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+shaped sphere clusters. We simulated the motion of a silica dimer
+trapped in a $5$ mW Gaussian beam; the dimer was comprised of two tangent spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$, respectively.
+We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
 
@@
 
-With any trapped particle there are two main sources of motion,
-Brownian and optical forces.  The dimer's Brownian motion
-($\Delta q^{\textbf{B}}$) is simulated by generating a set of
-normally-distributed numbers whose covariance is given by the dimer's
-diffusion tensor $\textbf{D}$.  The diffusion tensor is computed based
-on Farsund's work \cite{Farsund1996}. For computing the optical force
-Brownian OT uses the T-Matrix method to calculate the scattering
-coefficients of the outgoing scattered wave.  The T-matrix describes
-the scattering process by relating the scattering wave to the incident
-wave:
+With any trapped particle there are two main sources of motion, Brownian and optical forces.
+The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
+The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}.
+For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
+The T-matrix describes the scattering process by relating the scattering wave to the incident wave:
 \begin{align}
@@
 \end{align}
-MSTM computes the T-matrix of the dimer, and then passes the T-matrix
-to \textit{ott} to compute the optical forces. The benefit of this
-approach is that MSTM can accurately compute the T-matrix for any
-arbitrary collection of spheres once, and then \textit{ott} can
-calculate the optical forces and torques. The total displacement is
-the sum of the optical and Brownian contributions:
+MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
+The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques.
+The total displacement is the sum of the optical and Brownian contributions:
 \begin{align}
@@
 \end{align}
-where \textbf{F} is the a 6-component vector that describes the
-optical force and torque components.  Brownian OT returns the dimer's
-motion as a \textit{csv} file containing for each time step the
-$[x,y,z]$ position of the dimer's centre of mass relative to the trap
-focus, and the orientation represented as a quaternion.
+where \textbf{F} is the a 6-component vector that describes the optical force and torque components.
+Brownian OT returns the dimer's motion as a \textit{csv} file containing for each time step the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.
 
@@
 Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \& \ \theta_3$.
-Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretised into $\textit{n}_ref \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
+Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretised into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
 Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
@@
 \end{align}
-where the denominator is simply the standard deviation for our values
-of $I(\hat{\bf n}, \theta_k)$.
-
+where the denominator is simply the standard deviation for our values of $I(\hat{\bf n}, \theta_k)$.
 The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
 %
-In practice, $y_k(\bar{\bf s})$ will have an inherent amount of
-uncertainty in our measurement due to noise from our detector.
+In practice, $y_k(\bar{\bf s})$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
 We can compare any instantaneous signal $y_1, y_2, y_3$ to our reference signals by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf n})$.
@@
 \end{eqnarray*} 
-where $\epsilon$ is the percent error expected from our detection
-fibres.
+where $\epsilon$ is the percent error expected from our detection fibres.
 
@@
 orientation $\hat{\rm n}_\alpha$.
-%
-Using Bayes' theorem, the probability that the dimer is in reference
-orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
+Using Bayes' theorem, the probability that the dimer is in reference orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
 $y_3$) is measured, can be estimated:
@@
 %
-Without any prior evidence we assume that the orientation priori of our dimer $p(\hat{\bf n}_{\alpha})$ is uniform.
-Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.
+Without any prior evidence we assume that the orientation priori of our dimer
+$p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer from previous
+measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.
 The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).
@@
 \end{align}
-From \eqref{eq:bayes} we get a mass probability distribution that denotes the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a signal ($y_1$, $y_2$, $y_3$).
-In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
+From \eqref{eq:bayes} we get a mass probability distribution that denotes the
+probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a
+signal ($y_1$, $y_2$, $y_3$). In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
 
@@
 \label{sec:divergence}
-Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate the orientation of a dimer based on its scattering signal.
+Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate
+the orientation of a dimer based on its scattering signal.
 To evaluate the estimation we can use the simulation in Section~\ref{sec:brownian}; we denote the dimer's \emph{actual} orientation as $\hat{\bf s}$ and use MSTM to calculate its light scattering $I(\hat{\bf s}, \theta)$.
@@
 
-To measure the error in the model's estimate the distribution from Eq.~\eqref{eq:bayes} and an ideal result, $p_ideal$, were compared.
-An ideal result would be one that is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$.
+To measure the error in the model's estimate the distribution from
+Eq.~\eqref{eq:bayes} and an ideal result, $p_ideal$, were compared.  An ideal result would be one that is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$.
 \begin{align}
-	p_{ideal} = 
+	p_{best} = 
 	\begin{cases}
@@
 \end{align}
-The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero probability to every reference orientation.
-The model's error was measured by calculating the Kullback-Leibler divergence $K_l$ between the two distributions.
+The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
+probability to every reference orientation. The model's error was measured by calculating the Kullback-Leibler divergence $K_l$ between the two distributions.
 \begin{align}
-	K_{l, \#}(p_{ideal}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
-	= p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}| y_1,y_2,y_3))}
+	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
+	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}| y_1,y_2,y_3))}
 	\right]
@@
 \end{align}
-where a larger value of $K_l$ indicates that our model is less confident in its prediction of the dimer's orientation.
+where a larger value of $K_l$ indicates that our model is less confident in its 
+prediction of the dimer's orientation.
 The divergence provides us a measure of how confident our measurement is.
@@
 	\includegraphics[width=0.5\textwidth]{./Images/fig3.png}
-	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.}
+	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.
+Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see text for further description.}
 	\label{fig:uniform}
@@
 To remedy this we need to account for the fact that a trapped dimer's motion i limited due to the trap stiffness.
-Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:
+Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{est}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:
 
 \begin{align}
-  r_i(t, \hat{\bf n}_\alpha)
-  &= (\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_\alpha(t-\Delta t))
-  \\
-  p_i(\hat{\bf n})
-  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
-{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
+  p(\hat{\bf n}_\alpha)
+  &=\frac{e^{\beta (\hat{\bf n}_\alpha (t) 
+  	\cdot \hat{\bf n}_{est}(t-\Delta t))}}
+  {\sum_{j=1}^{n_{\rm ref}}
+	e^{\beta (\hat{\bf n}_\alpha (t) 
+	\cdot \hat{\bf n}_{est}(t-\Delta t)}}
 \end{align}

### 2022-12-08 Manuscript_rev1.tex
Add files via upload

@@
 
-Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons.
+Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico newtons.
 \textit{Further characterisation of the trapped entity would provide useful information, as explored in for example} metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
@@
 
-The possibility of integrating light scattering with optical trapping
-was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998}
-where a single-mode optical fibre was aligned to detect the scattered
-light from a trapped bead and its Brownian motion was studied.  While
-this allowed for studying the dynamics, structural information about
-the trapped bead was precluded as the measurement was obtained only at
-a single angle.  In this work, we propose a scheme that expands on the
-technique in \cite{Bar-Ziv_1998} to detect scattered light
-simultaneously at 3 angles as shown in Figure~\ref{fig:setup},
-combined with a novel Bayesian Inference based analysis technique to
-enable interpretation of the resulting multi-angle data as well as
-optimisation of the setup to provide maximal information from the
-signal.
+The possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered
+light from a trapped bead and its Brownian motion was studied.
+While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
+In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.
 
-To demonstrate the analysis, we study a simple, illustrative example
-of a trapped entity more complex than a single sphere (i.e.\ a trapped
-asymmetric dimer).  We explore how to optimise the experimental
-measurement to obtain the best estimate of the dimer's instantaneous
-orientation, as a further paradigmatic example of interpretation of
-the scattering data to obtain dynamical information on the trapped
-entity.  There is a paucity of literature on measuring the orientation
-of trapped non-spherical particles till the recent attempt in
-Ref.~\cite{raudsepp2022estimating} where imaging was employed to
-study the orientation of trapped dimers and a resolution of () was
-achieved.  Here, we simulate the Brownian motion of a dimer and
-calculate the resulting scattered light, and show how Bayesian
-inference can be used to optimise extraction of the true dimer
-orientation from the light scattering signals given a choice of
-scattering measurement angles.
+To demonstrate the analysis, we study a simple, illustrative example of a trapped entity more complex than a single sphere (i.e.\ a trapped asymmetric dimer).
+We explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity.
+There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in Ref.~\cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers.
+Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
 
@@
 shaped sphere clusters. We simulated the motion of a silica dimer
-trapped in a $5$ mW Gaussian beam; the dimer was comprised of two tangent spheres with radii $1\,{\rm \mu m}$ and $ 0.5\,{\rm \mu m}$, respectively.
-We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.
+trapped in a highly focused Gaussian beam.
 
@@
 
-With any trapped particle there are two main sources of motion, Brownian and optical forces.
-The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$.
-The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}.
-For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave.
-The T-matrix describes the scattering process by relating the scattering wave to the incident wave:
-\begin{align}
-	\begin{pmatrix}
-		p_{mn} \\
-		q_{mn}
-	\end{pmatrix}
-	= \textbf{T}
-	\begin{pmatrix}
-		a_{mn} \\
-		b_{mn}
-	\end{pmatrix}
-\end{align}
+The dimer's Brownian motion is simulated by calculating its diffusion tensor according to Farsund's \emph{et al} work \cite{Farsund1996}.
+For computing the optical force Brownian OT uses the T-Matrix method to calculate scattering of the dimer.
 MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
 The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques.
-The total displacement is the sum of the optical and Brownian contributions:
-\begin{align}
-  \langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle
-  &= 2D_{ij} \Delta t
-  \\
-  \Delta q^{opt} &= \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F}
-  \\
-\Delta q^{tot} &= \Delta q^{\textbf{B}} + \Delta q^{opt}
-\end{align}
-where \textbf{F} is the a 6-component vector that describes the optical force and torque components.
+The total displacement is the sum of the optical and Brownian contributions.
+
 Brownian OT returns the dimer's motion as a \textit{csv} file containing for each time step the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.
@@
 Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \& \ \theta_3$.
-Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretised into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
+Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretized into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
 Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
 %
-These raw intensities are normalised according to
+These raw intensities are normalized according to
 \begin{align}
@@
 The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
-%
-In practice, $y_k(\bar{\bf s})$ will have an inherent amount of uncertainty in our measurement due to noise from our detector.
-We can compare any instantaneous signal $y_1, y_2, y_3$ to our reference signals by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\bar{\bf n})$.
+The distribution of collected signals bares no specific relationship with its associated reference orientation, in some cases different reference orientations will have identical signals.
+This is an inherent challenge of the inverse light scattering problem, fortunately we can apply statistical techniques to resolve the lack of relationship between orientation and scattering.
+
+In practice, $y_k(\hat{\bf s})$ will have an inherent amount of uncertainty in its measurement.
+We can compare any instantaneous signal $y_1, y_2, y_3$ to our reference signals by assuming that the measurement uncertainty is Gaussian and is centred about $y_k(\hat{\bf n}_\alpha)$.
 %
@@
 \end{align}
-where $\sigma_k$ is related to the uncertainty in the detected signal
-from our dimer $y_k(\bar{\bf s})$.
-
-We left this as a tuneable parameter to characterize the influence of experimental error on our predictive method.
-The signal error was simulated in our model by stochastically adjusting our values of $y_k(\bar{\bf s})$.
-\begin{eqnarray*}
-  \sigma_k&=\epsilon\bar{y}(\hat{\bf n}_k)
-  \\
-  \Rightarrow y(\bar{\bf s}_k)&=y(\bar{\bf s})_k \pm n\sigma_k
-  \\ 
-n \in \mathbb{R},&n\in[0,1]
-\end{eqnarray*} 
-where $\epsilon$ is the percent error expected from our detection fibres.
-
-Equation~\eqref{eq:conditional} gives the probability that a signal
-($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference
-orientation $\hat{\rm n}_\alpha$.
-Using Bayes' theorem, the probability that the dimer is in reference orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
-$y_3$) is measured, can be estimated:
+where $\sigma_k$ is related to the uncertainty in our measurement $y_k(\hat{\bf s})$.
+We left this as a tuneable parameter to determine the influence of experimental error on our predictive method.
+To simulate uncertainty in our measurement we stochastically adjusted our values of $y_k(\hat{\bf s})$.
+Equation~\eqref{eq:conditional} gives the probability that a signal ($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference orientation $\hat{\rm n}_\alpha$.
+Using Bayes' theorem, the probability that the dimer is in reference orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$, $y_3$) is measured, can be estimated:
 %.
@@
 \end{align}
-where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior
-estimates of the distributions of particle orientations and
-instantaneous signals, respectively.
+where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of the distributions of particle orientations and instantaneous signals, respectively.
 %
@@
 	\begin{cases}
-		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3) & 
+		p(\hat{\bf n}_{est}|y_1,y_2,y_3) & 
 		\text{when $\hat{\bf n}_{estimate}$ = $\hat{\bf n}_{best}$}\\
-		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3)-p(\hat{\bf n}_{best}|y_1,y_2,y_3) & \text{when $\hat{\bf n}_{estimate} \ne \hat{\bf n}_{best}$}
+		p(\hat{\bf n}_{est}|y_1,y_2,y_3)-p(\hat{\bf n}_{best}|y_1,y_2,y_3) & \text{when $\hat{\bf n}_{est} \ne \hat{\bf n}_{best}$}
 	\end{cases}
@@
 
-Using our simulation from Section~\ref{sec:brownian} we calculated the instantaneous signal from our dimer at each time step using MSTM.
+Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer trapped within a 5 mW optical trap.
+The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively.
+We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.
+
 We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation.
@@
   	\cdot \hat{\bf n}_{est}(t-\Delta t))}}
-  {\sum_{j=1}^{n_{\rm ref}}
+  {\sum_{\alpha=1}^{n_{\rm ref}}
 	e^{\beta (\hat{\bf n}_\alpha (t) 
@@
 
-where $\beta$ is a weighting factor that controls the significance of our priori estimate.
+where $\beta$ is a weighting factor that describes the dimer's freedom of motion within the trap.
+By summing over each reference orientation $p(\hat{\bf n}_\alpha)$ is normalized.
 The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}.
@@
 
-The Brownian OT software was used to simulate the motion of a trappeddimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
-$10$ seconds of entering the optical trap.
+The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first $10$ seconds of entering the optical trap.
 The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
-The dimer's position orientation was recorded every $10\,{\rm \mu s}$ for using as a test dataset for our model.
+The dimer's position orientation was recorded every 1 ms for using as a test dataset for our model.
 
@@
 
-A key assumption of Eq.~\eqref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise.
-This is the central premise that allows us to form Eq.~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
-To this end we ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
+A key assumption of Eq.~\eqref{eq:conditional} is that we assume our measurement of scattering signals will inherently have some degree of uncertainty.
+This is the central premise that allows us to form Eq.~\eqref{eq:conditional}.
+In an experimental situation the most common source of uncertainty is from the detection bias.
+Therefore, understanding the limits of acceptable uncertainty is necessary for implementing this to experimental work.
+To this end we redefine $\sigma_k$ as the average reference signal $\bar{y}(\hat{\bf n}_\alpha)$ multiplied by the percent error caused by the signal detectors:
+\begin{align}
+	\sigma_k = \epsilon \bar{y}(\hat{\bf n}_\alpha)
+\end{align} 
+We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
 
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away from the ideal result.
+As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away fromthe ideal result.
 A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.

### 2022-12-09 Manuscript_rev2.tex
Add files via upload

--- 2022-12-08 Manuscript_rev1.tex
@@
 \makenomenclature
+\renewcommand{\nomlabel}[1]{\hfil #1\hfil}
 
@@
 \item Asymmetric dimers undergo a full inversion in the presence of an potential well. This only occurs if the size ratio is greater than 1:2.  
-\item Orientation can be classified by applying Bayesian Statistics to light intensity readings. 
-\item Choice of detection angles sets a lower limit on possible signal error. 
-\item Future applications in continuos processes where dimer shape is changing with time.   
+\item Orientation can be classified by applying Bayesian inference to light intensity signals. 
+\item The choice of detection angles has no noticeable effect on the performance of characterising orientation.
+\item Performance of model is determined in part due to uncertainty in the measurement precision. 
+\item Future applications in continuos processes where particle size is changing with time are being considered.  
 \end{highlights}
@@
 \end{frontmatter}
-
-\linenumbers
 
@@
 \begin{subfigure}{0.45\textwidth}
+	\subcaption{}
 	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
-	\subcaption{}
 \end{subfigure}
 \begin{subfigure}{0.45\textwidth}
+	\subcaption{}
 	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
-	\subcaption{}
 \end{subfigure}
@@
 
-\printnomenclature
+\printnomenclature[0.55in]
 
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.9\textwidth]{./Images/fig2.png}
-	\caption{Example dimer in orientation $\bar{\bf s}$}
-	\label{fig:dimer}
-\end{figure}
-
 The dimer's Brownian motion is simulated by calculating its diffusion tensor according to Farsund's \emph{et al} work \cite{Farsund1996}.
@@
 
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig2a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig2b.png}
+	\end{subfigure}
+	\caption{(a)Example dimer in orientation $\bar{\bf s}$, (b) 30 Referenec orientations pointing from [0,0,0] to each point}
+	\label{fig:dimer}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A plane wave 'probe' laser, directly perpendicular to the trapping laser, is incident on the dimer generating a scattering pattern that is dependent upon the dimer's orientation $I(\hat{\bf s}, \theta)$, computed using MSTM.
-To represent the experimental set up, three angles  ($\theta_1, \ \theta_2, \ \& \theta_3$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$
+To represent the experimental set up, three angles  ($\theta_1, \ \theta_2, \ \theta_3$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$
 
 Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \& \ \theta_3$.
-Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretized into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}.
+Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretized into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}, see Figure~\ref{fig:dimer} b.
+The maximum spacing between two neighbouring reference orientations is roughly 0.895 radians.
+
 Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$. 
@@
 \begin{align}
+	\label{eq:best}
 	p_{best} = 
@@
 	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
-	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}| y_1,y_2,y_3))}
+	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
 	\right]
@@
 \begin{align}
-	p(\hat{\bf n}|y_1,y_2,y_3) = 
+	p(\hat{\bf n}_\alpha|y_1,y_2,y_3) = 
 	\begin{cases}
@@
 	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.
-Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see text for further description.}
+Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
 	\label{fig:uniform}
@@
 	\end{subfigure}
-	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
+	\caption{Simulation results of: (a) the dimers orientation vector with time, (b) the dimer's [x,y,z] position with time.}
 	\label{fig:motion}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Impact of signal error on model predictions}
+\subsection{Impact of measurement precision on model predictions}
 \label{sec:epsilon}
@@
 Therefore, understanding the limits of acceptable uncertainty is necessary for implementing this to experimental work.
-To this end we redefine $\sigma_k$ as the average reference signal $\bar{y}(\hat{\bf n}_\alpha)$ multiplied by the percent error caused by the signal detectors:
+To this end we redefined $\sigma_k$ as the average reference signal $\bar{y}(\hat{\bf n}_\alpha)$ multiplied by the percent error caused by the signal detectors:
 \begin{align}
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of a) $5\%$, b) $15\%$, and c) $25\%$. Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}}
+	\caption{Model prediction for signal error of (a) $5\%$, (b) $15\%$, and (c) $25\%$.
+Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
 	\label{fig:epsilon}
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away fromthe ideal result.
-A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error.
+As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away from the ideal result.
+A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k(\hat{\bf n}_\alpha)$ is larger than our signal error.
 As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
 This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
-In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.
+Because there is not a unique relationship between the reference orientations and their respective signals this cannot be mitigated by increasing $\textit{n}_{ref}$.
 
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
+	\caption{Model prediction for signal error of $25\%$ when (a) $\beta = 5$ and (b) $\beta=6$.
+Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
 	\label{fig:beta}
@@
 We employed UltraNest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
-We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.
+We kept both $\beta$ and $\sigma_k$ constant during this so we could as these are parameters that can be changed more freely.
 
@@
 	\includegraphics[width=0.45\textwidth]{./Images/fig7.png}
-	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
+	\caption{Corner plot mapping combinations of $\theta_k$ that generate the best improvement}
 	\label{fig:corner}
@@
 As covered in Section~\ref{sec:epsilon} our choice for beta can help to improve our estimation by restricting it in the physical space.
-In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
+In addition depending on our choice of angles we are also setting the lower limit for our measurement precision, so it may be preferable to choose a selection of angles that allows for a larger measurement uncertainty and then adjust the weighting factor to improve upon our estimation.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 \bibliography{bib} 
@@
 \hline\hline
-Reference Orientation & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
+$\alpha$ & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
 \hline
@@
       \hline\hline
-      Reference Orientation  & $I(\hat{\bf n}_\alpha\ ,\ 15^{\circ})$ &  $I(\hat{\bf n}_\alpha \ , \ 45^{\circ})$ &  $I(\hat{\bf n}_\alpha \ ,\ 90^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 15^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 45^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 90^{\circ})$ \\
+      $\alpha$  & $I(\hat{\bf n}_\alpha\ ,\ 15^{\circ})$ &  $I(\hat{\bf n}_\alpha \ , \ 55^{\circ})$ &  $I(\hat{\bf n}_\alpha \ ,\ 90^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 15^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 55^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 90^{\circ})$ \\
 \hline

### 2022-12-09 Experiment3.py
Add files via upload

@@ -1,0 +1,648 @@
+#!/usr/bin/python3
+
+import heapq
+import math
+import csv
+import numpy as np
+import pickle
+import pylab as plt
+from statistics import mean, stdev
+from scipy.interpolate import splrep, splev, splint
+from scipy.integrate import tplquad
+from itertools import product
+from numpy.linalg import norm
+import random
+import argparse
+import itertools
+import xml.etree.cElementTree as ET
+from xml.dom import minidom
+import multiprocessing
+
+
+from Beam import *
+from MSTM import *
+from Particle import *
+
+neighbor_list = [np.array(vec) for vec in product([-1, 0, 1], repeat=3)]
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
+class Experiment3(object):
+
+
+    def __init__(self, pklfile='fit.pkl', angle_list=None):
+        
+        self.theta_list = None
+        self.n_list = self.set_measurement_angles        
+        self.sigma = 0.1e1
+        self.fit_dict = self.read_pickle(pklfile)
+        self.N = 10000
+        self.mstm = MSTM()
+
+        if (angle_list is None):
+            angle_list = [15.0, 45.0, 90.0]
+        self.set_measurement_angles(angle_list)
+
+        
+################################################################################
+    def read_pickle(self, filename):
+        pklfile = open(filename, 'rb')
+        fit_dict = pickle.load(pklfile)
+        pklfile.close()
+        return fit_dict
+
+
+################################################################################
+    def set_measurement_angles(self, theta_list):
+
+        self.theta_list = theta_list
+        
+        n_list = []
+        Iraw_measurements = [[] for theta in theta_list]
+        Deriv_measurements =[[] for theta in theta_list]
+        sum_prob = 0.0
+        for n, fit in self.fit_dict.items():
+            Iraw = [ float(splev(theta, fit) ) for theta in theta_list]
+            Deriv = [float(splev(theta, fit, der =1)) for theta in theta_list]
+            for index, I in enumerate(Iraw):
+                Iraw_measurements[index].append(I)
+            prob = 1.0
+            sum_prob += prob
[…]

### 2022-12-09 Manuscript_rev2.tex
Add files via upload

@@
 We left this as a tuneable parameter to determine the influence of experimental error on our predictive method.
-To simulate uncertainty in our measurement we stochastically adjusted our values of $y_k(\hat{\bf s})$.
 Equation~\eqref{eq:conditional} gives the probability that a signal ($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference orientation $\hat{\rm n}_\alpha$.
@@
 The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another.
-If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in Figure~\ref{fig:uniform}.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the model is somewhat random in it's estimation of the orientation, as seen in Figure~\ref{fig:uniform}.
 
@@
 	\cdot \hat{\bf n}_{est}(t-\Delta t)}}
+	\label{eq:boltz}
 \end{align}
@@
 By summing over each reference orientation $p(\hat{\bf n}_\alpha)$ is normalized.
+Implementation of Eq~\eqref{eq:boltz} helps to reduce the random elements seen in  Figure~\ref{fig:uniform} to give a better estimation, as shown in Figure~\ref{fig:biased}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth]{./Images/fig4.png}
+	\caption{Model's estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ being defined by Eq~\eqref{eq:boltz}.
+Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
+	\label{fig:biased}
+\end{figure}
+
 The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}.
@@
 Therefore, understanding the limits of acceptable uncertainty is necessary for implementing this to experimental work.
-To this end we redefined $\sigma_k$ as the average reference signal $\bar{y}(\hat{\bf n}_\alpha)$ multiplied by the percent error caused by the signal detectors:
+To this end we implemented implemented a stochastic variable that adjusts our detected signal by some amount of our measurement uncertainty:
 \begin{align}
-	\sigma_k = \epsilon \bar{y}(\hat{\bf n}_\alpha)
+	\eta = \epsilon \sigma_k
+	\rightarrow \ y_k(\hat{\bf{s}}) = y_k(\hat{\bf{s}}) \pm \eta
 \end{align}
-We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
+Where $\eta$ is our percent error in our detection bias.
+We ran our model while varying the value of $\eta$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
 
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $5\%$, (b) $15\%$, and (c) $25\%$.
+	\caption{Model prediction for signal error of (a) $5\%$, (b) $50\%$, and (c) $100\%$.
 Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away from the ideal result.
-A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k(\hat{\bf n}_\alpha)$ is larger than our signal error.
-As such the model could not distinguish between different reference orientations and was essentially just picking options at random.
-This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.
-Because there is not a unique relationship between the reference orientations and their respective signals this cannot be mitigated by increasing $\textit{n}_{ref}$.
-
-The latter issue of a high signal error can be addressed somewhat by adjusting our value of $\beta$ to flatten our probability distribution.
-We repeated our model estimation at $25\%$ with different values of $\beta$ to see the upper limit we could improve our result to.
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of an signal error variable results in a gradual decrease in our model's performance.
+While still providing a decent estimate of the dimer orientation the model eventually is nearly as correct as it is incorrect.
+Fortunately we can amend this by adjusting the weighting of our priori distribution $p(\hat{\bf n})$.
+We repeated our model estimation with different values of $\beta$ to see the upper limit we could improve our result to.
 
@@
 
-As shown in Figure~\ref{fig:beta}, eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
-The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error.
+Eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
+The exact value of $\beta$ that will optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
 In future, we want to also investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate the auto correlation function of our dimer's motion.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
 \bibliography{bib} 

### 2022-12-12 Manuscript_rev2.tex
Add files via upload

@@
 Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
+Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.
 
@@
 \end{align}
-where a larger value of $K_l$ indicates that our model is less confident in its 
-prediction of the dimer's orientation.
-The divergence provides us a measure of how confident our measurement is.
-To better reflect the model's accuracy we implemented a penalty condition that worsens the divergence if our model estimates the wrong reference orientation:
-\begin{align}
-	p(\hat{\bf n}_\alpha|y_1,y_2,y_3) = 
-	\begin{cases}
-		p(\hat{\bf n}_{est}|y_1,y_2,y_3) & 
-		\text{when $\hat{\bf n}_{estimate}$ = $\hat{\bf n}_{best}$}\\
-		p(\hat{\bf n}_{est}|y_1,y_2,y_3)-p(\hat{\bf n}_{best}|y_1,y_2,y_3) & \text{when $\hat{\bf n}_{est} \ne \hat{\bf n}_{best}$}
-	\end{cases}
-\end{align}
-
-This allows us to account for situations when our collected signal is equally likely to be associated with 2 reference signals.
+where a larger value of $K_l$ indicates that our model is less confident in its prediction of the dimer's orientation.
+While the divergence calculations are useful for showing our model's confidence in the prediction it cannot accurately reflect if the model is correct or not with it's estimation.
 Section~\ref{sec:Bayes} and Section~\ref{sec:divergence} are all for a single measurement of a light scattering signal from our dimer.
@@
 The worst case scenario is akin to randomly choosing a reference orientation at each time step.
-The greater the value of $F(K_l)$ the better our model fits the dimer's motion.
+The greater the value of $F(K_l)$ the better our model's confidence is in characterising the dimer's motion.
 Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.
@@
 		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/fig4a.png}
+		\includegraphics[width =\textwidth]{./Images/fig5a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig4b.png}
+		\includegraphics[width=\textwidth]{./Images/fig5b.png}
 	\end{subfigure}
@@
 It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely.
-Further simulations of differently sized dimers showed sa similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.
+Further simulations of differently sized dimers showed a similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.
 Dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
 %
-In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
-Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations.
+In the simulations of Vigilante~\emph{et~al}.\ = \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.
 Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
-The data from Figure~\ref{fig:motion} was used to test the efficacy of our predictive model.
-To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.
-For a perfect estimation these two plots should overlap, while we cannot achieve an exact estimation we can optimise it by adjusting the expected signal error and weighting factor.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This is the central premise that allows us to form Eq.~\eqref{eq:conditional}.
-In an experimental situation the most common source of uncertainty is from the detection bias.
+In an experimental situation the most common source of uncertainty is from the bias in our detection fibres.
 Therefore, understanding the limits of acceptable uncertainty is necessary for implementing this to experimental work.
-To this end we implemented implemented a stochastic variable that adjusts our detected signal by some amount of our measurement uncertainty:
+To this end we simulated the uncertainty in our signal measurements via the following:
 \begin{align}
@@
 \end{align}
-Where $\eta$ is our percent error in our detection bias.
-We ran our model while varying the value of $\eta$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
+Where $\eta$ is the signal noise and $\epsilon$ is our percent error in our signal measurement.
+We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
 
@@
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/fig5a.png}
+		\includegraphics[width=\textwidth]{./Images/fig6a.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/fig5b.png}
+		\includegraphics[width=\textwidth]{./Images/fig6b.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\textwidth]{./Images/fig5c.png}
+		\includegraphics[width=\textwidth]{./Images/fig6c.png}
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $5\%$, (b) $50\%$, and (c) $100\%$.
+	\caption{Model prediction for signal error of (a) $5\%$, (b) $50\%$, and (c) $75\%$.
 Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, the inclusion of an signal error variable results in a gradual decrease in our model's performance.
-While still providing a decent estimate of the dimer orientation the model eventually is nearly as correct as it is incorrect.
-Fortunately we can amend this by adjusting the weighting of our priori distribution $p(\hat{\bf n})$.
-We repeated our model estimation with different values of $\beta$ to see the upper limit we could improve our result to.
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise results in a gradual decrease in our model's performance.
+Eventually after the signal exceeds $X \ \%$ the performance drops of significantly as our estimation begins to fluctuate seemingly at random, similar to the behaviour shown in Figure~\ref{fig:uniform}.
+One common feature of our model's estimations is the large hump from $t = 1 \ s$ to $t = 4 \ s$; looking at the simulation data this correlates with the point where the dimers is moving very quickly as it inverts within the trap.
+One way we can alleviate the randomness is by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
+We repeated the model at $75\%$ signal error but this time increasing $\beta$ in order to demonstrate its effect on our model's performance.
 
@@
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/fig6a.png}
+		\includegraphics[width=\textwidth]{./Images/fig7a.png}
 	\end{subfigure}
 	\begin{subfigure}{0.33\textwidth}
-		\includegraphics[width=\textwidth]{./Images/fig6b.png}
+		\includegraphics[width=\textwidth]{./Images/fig7b.png}
 	\end{subfigure}
@@
 
+The exact value of $\beta$ that will best optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
 Eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
-The exact value of $\beta$ that will optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
+The other downside of increasing $\beta$ is that it also reduces our model's confidence significantly; this is not so much a problem for orientation where we can measure the difference between different reference orientation.
+It may however, pose an issue in scenarios where we want to characterise something more discrete such as scattering shape or size.
+
 In future, we want to also investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate the auto correlation function of our dimer's motion.
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{./Images/fig7.png}
+	\includegraphics[width=0.45\textwidth]{./Images/fig8.png}
 	\caption{Corner plot mapping combinations of $\theta_k$ that generate the best improvement}

### 2022-12-19 Manuscript_rev2.tex
Add files via upload

@@
 There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in Ref.~\cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers.
-Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
+Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, we utilise a neural network to determine the corresponding orientation.
+We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.
 Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.
@@
 A plane wave 'probe' laser, directly perpendicular to the trapping laser, is incident on the dimer generating a scattering pattern that is dependent upon the dimer's orientation $I(\hat{\bf s}, \theta)$, computed using MSTM.
-To represent the experimental set up, three angles  ($\theta_1, \ \theta_2, \ \theta_3$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$
+To represent the experimental set up, four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$.
 
-Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \& \ \theta_3$.
+Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \theta3, \ \& \ \theta_4$.
 Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretized into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}, see Figure~\ref{fig:dimer} b.
@@
 Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
+We noticed that keeping all four angles in the same plane resulted in identical signals, to counter this we projected one of our angles out of the horizontal plane and into the vertical plane resulting in unique signals for each reference orientation.
 %
@@
 The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
-The distribution of collected signals bares no specific relationship with its associated reference orientation, in some cases different reference orientations will have identical signals.
-This is an inherent challenge of the inverse light scattering problem, fortunately we can apply statistical techniques to resolve the lack of relationship between orientation and scattering.
+The distribution of collected signals bares no specific relationship with its associated reference orientation, this is an inherent problem of the inverse light scattering pattern.
+Fortunately, for cases where the uncertainty in our signal measurements is low, we can easily predict the orientation by utilising computational techniques such as neural networks.
 
-In practice, $y_k(\hat{\bf s})$ will have an inherent amount of uncertainty in its measurement.
-We can compare any instantaneous signal $y_1, y_2, y_3$ to our reference signals by assuming that the measurement uncertainty is Gaussian and is centred about $y_k(\hat{\bf n}_\alpha)$.
+We utilised a python machine learning program called scikit-learn to build a neural network for identifying the dimer's orientation from its light scattering signal.
+The network was trained by generating a database of random orientation vectors, calculating the light scattering signals, and then having the network estimate the probability of the signal coming from our dimer in that corresponding reference orientation.
+After making its estimation the network's loss function was evaluated and used to improve the estimation, we trained the network until the improvement in the loss function was less than 0.0001.
 %
-Combining the signals from all three detection angles provides us with the conditional probability that, should the dimer be in orientation
-$\hat{\bf n}_\alpha$, it would produce a signal ($y_1$, $y_2$, $y_3$):
-\begin{align}
-\label{eq:conditional}
-  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
-  = \prod^3_{k=1}
-(2\pi\sigma_k^2)^{-1/2} 
-e^{-(y_{k}-y_{k}(\hat{\bf n}_\alpha))^2/2\sigma_k^2}
-\end{align}
-where $\sigma_k$ is related to the uncertainty in our measurement $y_k(\hat{\bf s})$.
-We left this as a tuneable parameter to determine the influence of experimental error on our predictive method.
-Equation~\eqref{eq:conditional} gives the probability that a signal ($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference orientation $\hat{\rm n}_\alpha$.
-Using Bayes' theorem, the probability that the dimer is in reference orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$, $y_3$) is measured, can be estimated:
-%.
+The estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer.
+We therefore, utilise Bayesian inference to update our neural networks estimation.
+
 \begin{align}
   \label{eq:bayes}
-  p(\hat{\bf n}_\alpha| y_1,y_2,y_3)
+  p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
   &=
-    \frac{p(y_1,y_2,y_3|\hat{\bf n}_\alpha)
-    p(\hat{\bf n}_\alpha)}{p(y_1, y_2, y_3)}
+    \frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
+    p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
 \end{align}
@@
 
-Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer trapped within a 5 mW optical trap.
+Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
+The trapping laser is 1064nm NIR focused through a 1.25 NA objective.
 The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively.
@@
 The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another.
-If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the model is somewhat random in it's estimation of the orientation, as seen in Figure~\ref{fig:uniform}.
+If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the neural network's predictions are fairly accurate with the occasional random jump away from the correct result.
 
@@
 
-The reasoning for this lies with the fact that differences in reference signals and reference orientation spacing are not directly correlated with each other.
-In some cases, two reference orientation will have identical signal responses.
-To remedy this we need to account for the fact that a trapped dimer's motion i limited due to the trap stiffness.
-Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{est}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:
+The reasoning for this is that the distance between reference signals ($\hat{\bf n}_\alpha$) does not correlate with the spacing between reference orientations.
+To improve our estimation we can account for the fact that the motion of the dimer is limited due to the trapping stiffness.
+Therefore, our prior of the reference orientations $p(\hat{\bf n}_\alpha$ was redefined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$
 
@@
   p(\hat{\bf n}_\alpha)
-  &=\frac{e^{\beta (\hat{\bf n}_\alpha (t) 
+  &=\frac{e^{\beta (\hat{\bf n}_\alpha 
   	\cdot \hat{\bf n}_{est}(t-\Delta t))}}
   {\sum_{\alpha=1}^{n_{\rm ref}}
-	e^{\beta (\hat{\bf n}_\alpha (t) 
+	e^{\beta (\hat{\bf n}_\alpha 
 	\cdot \hat{\bf n}_{est}(t-\Delta t)}}
@@
 %
-In the simulations of Vigilante~\emph{et~al}.\ = \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
 Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.
@@
 \begin{align}
-	\eta = \epsilon \sigma_k
+	\eta_k = \epsilon \bar{y_k}(\hat{\bf n}_\alpha)
 	\rightarrow \ y_k(\hat{\bf{s}}) = y_k(\hat{\bf{s}}) \pm \eta
@@
 As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise results in a gradual decrease in our model's performance.
-Eventually after the signal exceeds $X \ \%$ the performance drops of significantly as our estimation begins to fluctuate seemingly at random, similar to the behaviour shown in Figure~\ref{fig:uniform}.
+Eventually after the signal exceeds $X \ \%$ the accuracy drops of significantly as our estimation begins to fluctuate seemingly at random.
+Similar to the behaviour shown in Figure~\ref{fig:uniform}.
 One common feature of our model's estimations is the large hump from $t = 1 \ s$ to $t = 4 \ s$; looking at the simulation data this correlates with the point where the dimers is moving very quickly as it inverts within the trap.
-One way we can alleviate the randomness is by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
+We can alleviate the randomness by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
 We repeated the model at $75\%$ signal error but this time increasing $\beta$ in order to demonstrate its effect on our model's performance.
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of $25\%$ when (a) $\beta = 5$ and (b) $\beta=6$.
+	\caption{Model prediction for signal error of $75\%$ when (a) $\beta = 5$ and (b) $\beta=15$.
 Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
@@
 
+As seen from Figure~\ref{fig:beta} $\beta$ helps to significantly reduce the random fluctuations in our model, eventually $\beta$ begins to over correct our estimate as seen in 7b.
 The exact value of $\beta$ that will best optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
-Eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result.
-The other downside of increasing $\beta$ is that it also reduces our model's confidence significantly; this is not so much a problem for orientation where we can measure the difference between different reference orientation.
+The downside of increasing $\beta$ is that it also reduces our model's confidence significantly; this is not so much a problem for orientation where we can measure the difference between reference orientation.
 It may however, pose an issue in scenarios where we want to characterise something more discrete such as scattering shape or size.
 
-In future, we want to also investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate the auto correlation function of our dimer's motion.
+In future, we want to investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 'n' time steps to investigate the auto correlation function of our dimer's motion.
 

### 2022-12-19 Manuscript_rev2.tex
Add files via upload

@@
 Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
-We noticed that keeping all four angles in the same plane resulted in identical signals, to counter this we projected one of our angles out of the horizontal plane and into the vertical plane resulting in unique signals for each reference orientation.
+We noticed that keeping all four angles in the same plane resulted in identical signals, to counter this we projected $\theta_4$ out of the horizontal plane and into the vertical plane ($\phi = 90$) resulting in unique signals for each reference orientation.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Optimization of parameters}
-\label{sec:ultranest}
-
-The signal error $\epsilon$ and the weighting factor $\beta$ are known quantities whose influence can be measured directly; however, our choice of angles and how they correlate with $F(K_l)$ is not easily understood Evaluating every possible configuration manually is inefficient, therefore we utilised UltraNest to apply a Monte Carlo simulation to our model.
-UltraNest is a Python package for fitting models with complex parameter correlations \cite{Buchner2016Ultranest}.
-UltraNest implements a variation of Monte Carlo integration, nested-sampling, where a likelihood contour is used to update a group of live points chosen from the prior estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).
-
-The UltraNest software was used to create a set of live points of different choices of $\theta_1$, $\theta_2$, and $\theta_3$, with each point being evaluated by its improvement factor ($F(K_l)$).
-After evaluating every live point, the worst results are removed and new live points are created based on the information provided by the previous live points.
-If there is a correlation between the choice of angles and $F(K_l)$ then the live points should congregate around a single (or multiple) choice of angles that will maximise $F(K_l)$ and therefore minimise the error between our estimate and instantaneous orientation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-A key assumption of Eq.~\eqref{eq:conditional} is that we assume our measurement of scattering signals will inherently have some degree of uncertainty.
-This is the central premise that allows us to form Eq.~\eqref{eq:conditional}.
-In an experimental situation the most common source of uncertainty is from the bias in our detection fibres.
-Therefore, understanding the limits of acceptable uncertainty is necessary for implementing this to experimental work.
-To this end we simulated the uncertainty in our signal measurements via the following:
+A key assumption of our neural network is that the scattering signal we detect has no uncertainty associated with it; in reality it is guaranteed that the signals we detect will have an inherent amount of measurement uncertainty.
+To demonstrate the impact of measurement uncertainty on our models performance we introduced a Gaussian noise to our measured signal.
 \begin{align}
-	\eta_k = \epsilon \bar{y_k}(\hat{\bf n}_\alpha)
-	\rightarrow \ y_k(\hat{\bf{s}}) = y_k(\hat{\bf{s}}) \pm \eta
+	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \eta I(\hat{\bf s})
 \end{align}
-Where $\eta$ is the signal noise and $\epsilon$ is our percent error in our signal measurement.
-We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:
+Where $\eta$ is the percentage error associated with our measurement.
+We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and the vertical detector at $75^{\circ}$, with $\beta$ set to $1$:
 
 \begin{figure}[h]
-
 	\centering
 	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/fig6a.png}
@@
 	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/fig6b.png}
@@
 	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/fig6c.png}
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $5\%$, (b) $50\%$, and (c) $75\%$.
-Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
+	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=1.174]$, (b) $15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~  $[F(K_l)=0.457]$.}
 	\label{fig:epsilon}
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise results in a gradual decrease in our model's performance.
-Eventually after the signal exceeds $X \ \%$ the accuracy drops of significantly as our estimation begins to fluctuate seemingly at random.
-Similar to the behaviour shown in Figure~\ref{fig:uniform}.
-One common feature of our model's estimations is the large hump from $t = 1 \ s$ to $t = 4 \ s$; looking at the simulation data this correlates with the point where the dimers is moving very quickly as it inverts within the trap.
-We can alleviate the randomness by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
-We repeated the model at $75\%$ signal error but this time increasing $\beta$ in order to demonstrate its effect on our model's performance.
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise quickly leads to a decrease in our model's performance.
+This is due to an inherent feature of the inverse scattering problem; that being that a two distinct regions of orientation space will become heavily intertwined in the intensity space resulting in poor predictions from our neural network.
+Fortunately, we can alleviate this by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
+We repeated the model at for the above signal errors, but increasing $\beta$ until the model best fit our dimer's motion
 
@@
 	\centering
-	\begin{subfigure}{0.33\textwidth}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/fig7a.png}
 	\end{subfigure}
-	\begin{subfigure}{0.33\textwidth}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
 		\includegraphics[width=\textwidth]{./Images/fig7b.png}
 	\end{subfigure}
-	\caption{Model prediction for signal error of $75\%$ when (a) $\beta = 5$ and (b) $\beta=15$.
-Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig7c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of $1\%$, $15\%$, and $25\%$ with (a) $\beta=4$ ~ $[F(K_l)=1.914]$, (b) $\beta=15$ ~  $[F(K_l)=0.734]$, and (c) $\beta=22$ ~ $[F(K_l)=0.621]$.} 
 	\label{fig:beta}
@@
 
-As seen from Figure~\ref{fig:beta} $\beta$ helps to significantly reduce the random fluctuations in our model, eventually $\beta$ begins to over correct our estimate as seen in 7b.
+As seen from Figure~\ref{fig:beta} $\beta$ helps to significantly reduce the random fluctuations in our model, eventually $\beta$ will begin to over correct our estimation.
 The exact value of $\beta$ that will best optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
-The downside of increasing $\beta$ is that it also reduces our model's confidence significantly; this is not so much a problem for orientation where we can measure the difference between reference orientation.
-It may however, pose an issue in scenarios where we want to characterise something more discrete such as scattering shape or size.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter Optimisation Results}
-\label{sec:ultranest_res}
-
-We wanted to see if our choice of angles plays a significant role in our model's efficacy.
-We employed UltraNest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.
-We kept both $\beta$ and $\sigma_k$ constant during this so we could as these are parameters that can be changed more freely.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.45\textwidth]{./Images/fig8.png}
-	\caption{Corner plot mapping combinations of $\theta_k$ that generate the best improvement}
-	\label{fig:corner}
-\end{figure}
-
-Figure~\ref{fig:corner} is the resulting corner plot from our UltraNest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer.
-Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the choice of detection angle.
-This result only indicates that orientation is not a characteristic that is angularly biased in its scattering.
-It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.
-
-While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we cannot achieve any greater accuracy than dividing the orientation space into $n_{\rm ref}=30$ reference orientations.
-We hope to develop our method further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.
-
-As covered in Section~\ref{sec:epsilon} our choice for beta can help to improve our estimation by restricting it in the physical space.
-In addition depending on our choice of angles we are also setting the lower limit for our measurement precision, so it may be preferable to choose a selection of angles that allows for a larger measurement uncertainty and then adjust the weighting factor to improve upon our estimation.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-12-21 Manuscript_rev2.tex
working on manuscript

@@
 \item Asymmetric dimers undergo a full inversion in the presence of an potential well. This only occurs if the size ratio is greater than 1:2.  
-\item Orientation can be classified by applying Bayesian inference to light intensity signals. 
-\item The choice of detection angles has no noticeable effect on the performance of characterising orientation.
-\item Performance of model is determined in part due to uncertainty in the measurement precision. 
+\item Orientations are classified by a combination of neural networks and Bayesian inference. 
+\item Signal error has a significant effect on the performance of the neural network, this can be countered via biasing of the priori distribution. 
 \item Future applications in continuos processes where particle size is changing with time are being considered.  
@@
 
-Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico newtons.
-\textit{Further characterisation of the trapped entity would provide useful information, as explored in for example} metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
-Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.
+Since the invention of optical tweezers in the late 1980's, they have
+found applications in experiments ranging from single molecule
+biophysics \cite{Bustamante2021Biophysics} to those that test the
+fundamental assumptions of quantum mechanics \cite{yin2013large}
+thanks mainly to the ability of the tweezer to transduce and detect
+forces on the order of a few piconewtons. \textit{Further
+  characterisation of the trapped entity would provide useful
+  information, as explored in for example} metrological measurements
+\cite{arita2020coherent} and colloidal aggregation
+\cite{burns1990optical}. Chemical characterisation of the trapped
+entities has been achieved using spectroscopic techniques such as
+Raman Scattering \cite{gupta2014raman}, while dynamical
+characterisation has been demonstrated using a Quadrant Photo
+Detector, by following the centre-of-mass Brownian motion of the
+trapped entity \cite{friedrich2012tuning} and measuring rotation of
+the centre-of-mass \cite{yifat2021facile}.
 
@@
 
-\subsection{Brownian Simulation}
+\subsection{Brownian Dynamics Simulation}
 \label{sec:brownian}
 
-We use the Brownian OT package developed by Fung~\textit{et~al}.
-\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap.
-Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
-shaped sphere clusters. We simulated the motion of a silica dimer
-trapped in a highly focused Gaussian beam.
-
-The dimer's Brownian motion is simulated by calculating its diffusion tensor according to Farsund's \emph{et al} work \cite{Farsund1996}.
-For computing the optical force Brownian OT uses the T-Matrix method to calculate scattering of the dimer.
-MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces.
-The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques.
-The total displacement is the sum of the optical and Brownian contributions.
-
-Brownian OT returns the dimer's motion as a \textit{csv} file containing for each time step the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.
+Brownian dynamics simulations of the motion of a silica dimer trapped
+in a highly focused Gaussian beam were performed using software
+package Brownian~OT, developed by Fung~\textit{et~al}.\
+\cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric
+dimer within an optical trap.  Brownian~OT combines MSTM
+\cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox''
+(\textit{ott}) \cite{Lenton2020} to determine the optical forces and
+torques on arbitrary clusters of spheres using the T-Matrix method.
+%MSTM
+%computes the T-matrix of the dimer, and then passes the T-matrix to
+%\textit{ott} to compute the optical forces.
+%The benefit of this approach is that MSTM can accurately compute the
+%T-matrix for any arbitrary collection of spheres once, and then
+%\textit{ott} can calculate the optical forces and torques.
+%
+The diffusion tensor of the dimer taken from Ref.~\citenum{Nir_Acrivos_1973}.
+%Ref.~\citenum{Farsund1996}.
+%
+%The total displacement is the sum of the optical and Brownian
+%contributions.
+%
+Brownian OT returns the dimer's motion as a \textit{csv} file
+containing for each time step the $[x,y,z]$ position of the dimer's
+centre of mass relative to the trap focus, and the orientation
+represented as a quaternion.
 
@@
 
-Our goal is to determine the orientation of our trapper dimer based on the scattering measurements $I(\hat{n}, \ \theta_k)$ collected from three detection fibres situated at $\theta_1,\ \theta_2, \ \theta3, \ \& \ \theta_4$.
-Rather than get an exact estimate of the dimer's orientation the possible orientation space is discretized into $\textit{n}_{ref} \ = \ 30$ possible reference orientations $\hat{\bf n}_{\alpha}$ that are evenly distributed on a unit sphere \cite{Rey2006}, see Figure~\ref{fig:dimer} b.
-The maximum spacing between two neighbouring reference orientations is roughly 0.895 radians.
+Our goal is to determine the orientation of our trapper dimer based on
+the scattering measurements $I(\hat{n}, \theta_k)$ collected from
+three detection fibres situated at $\theta_1$, $\theta_2$, $\theta_3$,
+and $\theta_4$.  The dimer's orientation is discretized into
+$n_{ref}=30$ reference orientations $\hat{\bf n}_{\alpha}$ that are
+evenly distributed on a unit sphere \cite{Rey2006} (see
+Figure~\ref{fig:dimer}~b).  The spacing between two neighbouring
+reference orientations is roughly $0.895$~radians.
 
-Using MSTM we compute the raw intensities that would be collected from the dimer in each reference orientation $I(\hat{\bf n}_{\alpha}, \theta_k)$.
-We noticed that keeping all four angles in the same plane resulted in identical signals, to counter this we projected $\theta_4$ out of the horizontal plane and into the vertical plane ($\phi = 90$) resulting in unique signals for each reference orientation.
+Using MSTM, we compute the raw intensities that would be collected
+from the dimer in each reference orientation
+$I(\hat{\bf n}_{\alpha}, \theta_k)$. We noticed that keeping all four
+angles in the same plane resulted in identical signals, to counter
+this we projected $\theta_4$ out of the horizontal plane and into the
+vertical plane ($\phi = 90$) resulting in unique signals for each
+reference orientation.
 %
@@
 \end{align}
-where the denominator is simply the standard deviation for our values of $I(\hat{\bf n}, \theta_k)$.
-The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
-The distribution of collected signals bares no specific relationship with its associated reference orientation, this is an inherent problem of the inverse light scattering pattern.
-Fortunately, for cases where the uncertainty in our signal measurements is low, we can easily predict the orientation by utilising computational techniques such as neural networks.
+where the denominator is simply the standard deviation for our values
+of $I(\hat{\bf n}, \theta_k)$.  The reference orientations, raw
+intensities, and scaled signals are given in Tables~\ref{tab:A1} and
+\ref{tab:A2}.  The distribution of collected signals bares no specific
+relationship with its associated reference orientation, this is an
+inherent problem of the inverse light scattering pattern.
 
-We utilised a python machine learning program called scikit-learn to build a neural network for identifying the dimer's orientation from its light scattering signal.
-The network was trained by generating a database of random orientation vectors, calculating the light scattering signals, and then having the network estimate the probability of the signal coming from our dimer in that corresponding reference orientation.
-After making its estimation the network's loss function was evaluated and used to improve the estimation, we trained the network until the improvement in the loss function was less than 0.0001.
+%Fortunately, for cases where the uncertainty in our signal
+%measurements is low, we can easily predict the orientation by
+%utilising computational techniques such as neural networks.
+
+The inverse mapping from the scaled measured signal to a particular
+reference orientation of the dimer was fit using a multi-layer
+perceptron neural network with four hidden layers, each with 100
+neurons logistic activation function.  This was implemented in Python
+using the scikit-learn \cite{Pedregosa_etal_2011} library.
+%to build a neural network for
+%identifying the dimer's orientation from its light scattering
+%signal.
+The neural network was trained using the stochastic optimizer Adam
+\cite{Kingma_Ba_014} on a data set that was generated by randomly
+choosing $10^4$ orientations and calculating the corresponding scaled
+signals from light scattering.
 %
-The estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer.
-We therefore, utilise Bayesian inference to update our neural networks estimation.
+%
+%and then having the network estimate the probability of the
+%signal coming from our dimer in that corresponding reference
+%orientation.  After making its estimation the network's loss function
+%was evaluated and used to improve the estimation, we trained the
+%network until the improvement in the loss function was less than
+%$10^{-4}$.
 
+The estimate provided by the neural network can be improved further by
+accounting for any prior information we know about the dimer. We
+therefore, utilise Bayesian inference to update our neural networks
+estimation.
 \begin{align}
@@
 As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise quickly leads to a decrease in our model's performance.
-This is due to an inherent feature of the inverse scattering problem; that being that a two distinct regions of orientation space will become heavily intertwined in the intensity space resulting in poor predictions from our neural network.
+This is due to an inherent feature of the inverse scattering problem; that being that a two distinct regions of orientation space will become heavily intertwined in the intensity space despite remaining continuous.
+As the signal error increases our neural network will map the signal to the incorrect reference orientation.
+%
 Fortunately, we can alleviate this by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap.
@@
 We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.\\
+
+Additionally, we found that the due to the inverse light scattering problem that physical space regions can be continuously mapped to a space in the intensity region but with a loss of shape and size.
+In the future, we believe that we can improve our estimation by utilising a greater number of reference orientations and by optimising our choice of detection angles to better discretise our intensity regions.
 

### 2022-12-21 Manuscript_rev2-LL.tex
reinstating old version of Manuscript_rev2.tex

--- 2022-12-21 Manuscript_rev2.tex
@@
 %
-Without any prior evidence we assume that the orientation priori of our dimer
-$p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer from previous
-measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.
-The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).
-This is given by taking the discrete integral over collection of reference orientations.
+Without any prior evidence we assume that the orientation priori of
+our dimer $p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer
+from previous measurement can be used to inform our estimate of
+$p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.  The
+latter prior $p(y)$ is the probability of measuring a signal
+$(y_1, y_2, y_3)$.  This is given by taking the discrete integral over
+collection of reference orientations.
 \begin{align}
@@
 \end{align}
-From \eqref{eq:bayes} we get a mass probability distribution that denotes the
-probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a
-signal ($y_1$, $y_2$, $y_3$). In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
+From Eq.~\eqref{eq:bayes}, we get a mass probability distribution that
+denotes the probability that our dimer is in orientation
+$\hat{\bf n}_{\alpha}$ given a signal $(y_1, y_2, y_3)$.  In order to
+evaluate the accuracy of our method, we tested it using a Brownian
+dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Calculation of error}
-\label{sec:divergence}
+\subsection{Calculation of error
+\label{sec:divergence}}
+
 Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate

### 2022-12-21 Manuscript_rev2-LL.tex
edits to discussion of prior

@@
 
-The estimate provided by the neural network can be improved further by
-accounting for any prior information we know about the dimer. We
-therefore, utilise Bayesian inference to update our neural networks
-estimation.
+The estimate provided by the neural network can be further improved by
+accounting for any prior information known about the dimer orientation
+utilizing Bayesian inference:
 \begin{align}
@@
 \end{align}
-where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of the distributions of particle orientations and instantaneous signals, respectively.
+where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3, y_4)$ are the
+prior estimates of the distributions of particle orientations and
+instantaneous signals, respectively.
 %
-Without any prior evidence we assume that the orientation priori of
-our dimer $p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer
-from previous measurement can be used to inform our estimate of
-$p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.  The
-latter prior $p(y)$ is the probability of measuring a signal
-$(y_1, y_2, y_3)$.  This is given by taking the discrete integral over
-collection of reference orientations.
+Without any prior information, we assume that
+$p(\hat{\bf n}_{\alpha})$ is uniform.  Knowledge of the dimer
+orientation from a previous measurement can be used to inform our
+estimate of $p(\hat{\bf n}_{\alpha})$, covered in
+Section~\ref{sec:test}.  The latter prior $p(y)$ is the probability of
+measuring a signal $(y_1, y_2, y_3, y_4)$.  This is given by taking
+the discrete integral over collection of reference orientations.
 \begin{align}
-  p(y_1, y_2, y_3)
+  p(y_1, y_2, y_3, y_4)
   =
   \sum_{\alpha=1}^{n_{\rm ref}}
-  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
+  p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
   p(\hat{\bf n}_\alpha)
@@
 denotes the probability that our dimer is in orientation
-$\hat{\bf n}_{\alpha}$ given a signal $(y_1, y_2, y_3)$.  In order to
-evaluate the accuracy of our method, we tested it using a Brownian
-dynamics simulation of the dimer (see Section~\ref{sec:brownian}).
+$\hat{\bf n}_{\alpha}$ given a signal $(y_1, y_2, y_3, y_4)$.  In
+order to evaluate the accuracy of our method, we tested it using a
+Brownian dynamics simulation of the dimer (see
+Section~\ref{sec:brownian}).
 

### 2022-12-21 Manuscript_rev2-LL.tex
fixes to acknowledgement and disclosure sections

@@
 
-To measure the error in the model's estimate the distribution from
-Eq.~\eqref{eq:bayes} and an ideal result, $p_ideal$, were compared.  An ideal result would be one that is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$.
+To measure the error in the model's estimate, the distribution from
+Eq.~\eqref{eq:bayes} and an ideal result, $p_{ideal}$, were compared.
+An ideal result would be one that is 0 for every $\hat{\bf n}$ apart
+from $\hat{\bf n}_{best}$.
 \begin{align}
@@
 The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
-probability to every reference orientation. The model's error was measured by calculating the Kullback-Leibler divergence $K_l$ between the two distributions.
+probability to every reference orientation. The model's error was
+quantified by calculating the Kullback-Leibler divergence $K_l$
+between the two distributions:
 \begin{align}
@@
 \end{align}
-where a larger value of $K_l$ indicates that our model is less confident in its prediction of the dimer's orientation.
-While the divergence calculations are useful for showing our model's confidence in the prediction it cannot accurately reflect if the model is correct or not with it's estimation.
-Section~\ref{sec:Bayes} and Section~\ref{sec:divergence} are all for a single measurement of a light scattering signal from our dimer.
-To properly model it's motion we need to consider how the orientation changes with time.
+where a larger value of $K_l$ indicates that our model is less
+confident in its prediction of the dimer's orientation. While the
+divergence calculations are useful for showing our model's confidence
+in the prediction it cannot accurately reflect if the model is correct
+or not with its estimation.  Section~\ref{sec:Bayes} and
+Section~\ref{sec:divergence} are all for a single measurement of a
+light scattering signal from our dimer.  To properly model it's motion
+we need to consider how the orientation changes with time.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
-The trapping laser is 1064nm NIR focused through a 1.25 NA objective.
-The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively.
-We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.
+Using our simulation from Section~\ref{sec:brownian} we simulated the
+motion of a silica dimer (index of refraction $n=1.59$) trapped in
+water ($n = 1.33$) within a $5\,{\rm mW}$ optical trap.  The trapping
+laser is $1064\,{\rm nm}$ NIR focused through a $1.25$~NA objective.
+The dimer is comprised of two tangent spheres with radii
+$1\,\mu{\rm m}$ and $0.5\,\mu{\rm m}$ respectively. We simulated the
+first $10$\,seconds of motion, calculating the orientation and
+position every $1\,{\rm ms}$.
 
@@
 Therefore, our prior of the reference orientations $p(\hat{\bf n}_\alpha$ was redefined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$
-
 \begin{align}
@@
 \end{align}
-
 where $\beta$ is a weighting factor that describes the dimer's freedom of motion within the trap.
 By summing over each reference orientation $p(\hat{\bf n}_\alpha)$ is normalized.
-Implementation of Eq~\eqref{eq:boltz} helps to reduce the random elements seen in  Figure~\ref{fig:uniform} to give a better estimation, as shown in Figure~\ref{fig:biased}.
+Implementation of Eq.~\eqref{eq:boltz} helps to reduce the random elements seen in  Figure~\ref{fig:uniform} to give a better estimation, as shown in Figure~\ref{fig:biased}.
 
@@
 To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
-
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and  Discussion}
-\label{sec:Discussion}
+\section{Results and  Discussion
+\label{sec:Discussion}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first $10$ seconds of entering the optical trap.
-The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
-The dimer's position orientation was recorded every 1 ms for using as a test dataset for our model.
+The Brownian OT software was used to simulate the motion of a trapped
+dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first
+$10$ seconds of entering the optical trap.  The initial orientation
+was assumed as strictly vertical (in line with the beam propagation
+direction). The dimer's position and orientation was recorded every
+$1\,{\rm ms}$ for using as a test dataset for our model.
 
@@
 Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
-We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.\\
+We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\
+\section*{Acknowledgement}
 
-\noindent \textbf{Disclosures.} The authors declare no conflict of interest. \\
+The authors thank the support for this research from the funding
+provided by the Leverhulme Trust.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Disclosures}
+
+The authors declare no conflict of interest.
 

### 2022-12-21 Manuscript_rev2-LL.tex
more updates

@@
 The neural network was trained using the stochastic optimizer Adam
-\cite{Kingma_Ba_014} on a data set that was generated by randomly
+\cite{Kingma_Ba_2014} on a data set that was generated by randomly
 choosing $10^4$ orientations and calculating the corresponding scaled
@@
 \end{align}
-Where $\eta$ is the percentage error associated with our measurement.
-We ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and the vertical detector at $75^{\circ}$, with $\beta$ set to $1$:
+where $\eta$ is the percentage uncertainty associated with the light
+scattering measurement.  We tested the method for varying values of
+$\eta$.  Figure~\ref{fig:epsilon} shows the performance of the method
+with a selection of angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$
+and the vertical detector at $75^{\circ}$, with $\beta=1$:
 
@@
 
-As seen from Figure~\ref{fig:beta} $\beta$ helps to significantly reduce the random fluctuations in our model, eventually $\beta$ will begin to over correct our estimation.
-The exact value of $\beta$ that will best optimise our performance varies based on the expected signal error, therefore we can use it as variable for fine tuning our results to suit our expected signal error.
+As seen from Figure~\ref{fig:beta} $\beta$ helps to significantly
+reduce the random fluctuations in our model, eventually $\beta$ will
+begin to over correct our estimation.  The value of $\beta$ that will
+best optimise our performance varies based on the expected signal
+error, therefore we can use it as variable for fine tuning our results
+to suit our expected signal error.
 
-In future, we want to investigate the possibility of extending our prior $p(\hat{\bf n})$ so we account for not just the change in orientation after one time step but after 'n' time steps to investigate the auto correlation function of our dimer's motion.
+In future, we plan to investigate the possibility of extending our
+prior $p(\hat{\bf n})$ so we account for not just the change in
+orientation after one time step but after 'n' time steps to
+investigate the auto correlation function of the dimer's motion.
 

### 2022-12-21 Manuscript_rev2-LL.tex
cleaning up figure captions

@@
 
-We consider our dimer from Section~\ref{sec:brownian} in an optical trap; at any point in time we can draw a unit vector $\hat{s}$ that points from the centre of larger sphere, to the centre of the smaller sphere.
-A plane wave 'probe' laser, directly perpendicular to the trapping laser, is incident on the dimer generating a scattering pattern that is dependent upon the dimer's orientation $I(\hat{\bf s}, \theta)$, computed using MSTM.
-To represent the experimental set up, four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) were chosen and the calculated intensity at each angle $\theta_k$ was recorded as $I(\hat{s}, \theta_k)$.
+We consider our dimer from Section~\ref{sec:brownian} in an optical
+trap; at any point in time we can draw a unit vector $\hat{\bf s}$
+that points from the centre of larger sphere, to the centre of the
+smaller sphere.  A plane wave 'probe' laser, directly perpendicular to
+the trapping laser, is incident on the dimer generating a scattering
+pattern that is dependent upon the dimer's orientation
+$I(\hat{\bf s}, \theta)$, computed using MSTM. To represent the
+experimental set up, four angles
+$(\theta_1, \theta_2, \theta_3, \theta_4)$ were chosen and the
+calculated intensity at each angle $\theta_k$ was recorded as
+$I(\hat{s}, \theta_k)$.
 
@@
 \includegraphics[width=0.5\textwidth]{./Images/fig3.png}
-	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.
-Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
-	\label{fig:uniform}
+\caption{\label{fig:uniform}
+%
+Estimation of the dimer orientation, assuming $p(\hat{\bf n}_\alpha)$
+to be uniform. Blue line denotes the best result we can achieve,
+orange line denotes the result provided by eq~\ref{eq:bayes}. Dotted
+line denotes spacing between two neighbouring
+reference orientations $\hat{\bf n}_\alpha$ (see Section~\ref{sec:Bayes}).
+%
+}
 \end{figure}
@@
 \end{align}
-where $\beta$ is a weighting factor that describes the dimer's freedom of motion within the trap.
-By summing over each reference orientation $p(\hat{\bf n}_\alpha)$ is normalized.
-Implementation of Eq.~\eqref{eq:boltz} helps to reduce the random elements seen in  Figure~\ref{fig:uniform} to give a better estimation, as shown in Figure~\ref{fig:biased}.
+where $\beta$ is a weighting factor that describes the dimer's freedom
+of motion within the trap.  By summing over each reference orientation
+$p(\hat{\bf n}_\alpha)$ is normalized. Implementation of
+Eq.~\eqref{eq:boltz} helps to reduce the random elements seen in
+Figure~\ref{fig:uniform} to give a better estimation, as shown in
+Figure~\ref{fig:biased}.
 
@@
 \includegraphics[width=0.5\textwidth]{./Images/fig4.png}
-	\caption{Model's estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ being defined by Eq~\eqref{eq:boltz}.
-Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-Dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$, see Sec~\ref{sec:Bayes}.}
-	\label{fig:biased}
+\caption{\label{fig:biased}
+%
+Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
+by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
+achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+Dotted line denotes spacing between two neighbouring reference
+orientations $\hat{\bf n}_\alpha$ (see Section~\ref{sec:Bayes}).
+  %
+}
 \end{figure}
@@
 \end{subfigure}
-	\caption{Simulation results of: (a) the dimers orientation vector with time, (b) the dimer's [x,y,z] position with time.}
-	\label{fig:motion}
+\caption{\label{fig:motion}
+%
+Simulation results of: (a) the dimers orientation vector with time,
+(b) the dimer's [x,y,z] position with time.
+%
+}
 \end{figure}
@@
 \end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=1.174]$, (b) $15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~  $[F(K_l)=0.457]$.}
-	\label{fig:epsilon}
+\caption{\label{fig:epsilon}
+%
+Model prediction for signal error of (a) $1\%$ ~ $[F(K_l)=1.174]$, (b)
+$15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~ $[F(K_l)=0.457]$.
+%
+}
 \end{figure}
@@
 \end{subfigure}
-	\caption{Model prediction for signal error of $1\%$, $15\%$, and $25\%$ with (a) $\beta=4$ ~ $[F(K_l)=1.914]$, (b) $\beta=15$ ~  $[F(K_l)=0.734]$, and (c) $\beta=22$ ~ $[F(K_l)=0.621]$.} 
-	\label{fig:beta}
+\caption{\label{fig:beta}
+%
+  Model prediction for signal error of $1\%$, $15\%$, and $25\%$ with (a) $\beta=4$ ~ $[F(K_l)=1.914]$, (b) $\beta=15$ ~  $[F(K_l)=0.734]$, and (c) $\beta=22$ ~ $[F(K_l)=0.621]$.
+%
+} 
 \end{figure}

### 2022-12-22 Manuscript_rev2-LL.tex
more minor edits
more minor updates

@@
   While optical trapping is a well understood method for force
-  transduction and detection,characterisation of trapped entities
-  poses a two-fold challenge --- one experimental concerning the
+  transduction and detection, characterisation of trapped entities
+  poses a two-fold challenge --- one experimental, concerning the
   arrangement of light detectors and the other, theoretical involving
@@
   engineering challenges due to the space constraints in a
-  conventional optical trapping setup.  We propose here a plausible
+  conventional optical trapping setup.  We propose a plausible
   scenario of detecting scattered light from an optically trapped
@@
 \begin{cases}
-		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
+1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$} \\
 0 & \text{anywhere else}
@@
 \begin{align}
-K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
-K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right] \\
+  K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#}
+  \\
+  K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right]
+  \\
 F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
@@
 
-The worst case scenario is akin to randomly choosing a reference orientation at each time step.
-The greater the value of $F(K_l)$ the better our model's confidence is in characterising the dimer's motion.
-Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.
+The worst case scenario is akin to randomly choosing a reference
+orientation at each time step. The greater the value of $F(K_l)$, the
+better our model's confidence is in characterising the dimer's
+motion. Because our model is dependent on several parameters we need
+to a sophisticated method for understanding how these parameters
+correlate with $F(K_l)$.
 
@@
 %
-Simulation results of: (a) the dimers orientation vector with time,
-(b) the dimer's [x,y,z] position with time.
+Simulation results for the dimer (a) orientation and
+(b) position with time.
 %
@@
 %
-In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
-Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.
-Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.
+In the simulations of Vigilante~\emph{et~al}.\
+\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
+investigated; their findings showed that the optical torque on the
+dimer goes to zero while aligned vertically and is at its maximum in a
+horizontal alignment. Therefore, the inversion of an asymmetric dimer
+suggests that if the size difference is significant, the optical
+torque is minimal for a dimer in both horizontal and vertical
+orientations. Once we have our experimental set up complete we can
+confirm this by trapping an asymmetric dimer and monitoring its
+orientation.
 
@@
 %
-Model prediction for signal error of (a) $1\%$ ~ $[F(K_l)=1.174]$, (b)
-$15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~ $[F(K_l)=0.457]$.
+Model prediction for signal error of (a) $\eta=1\%$ ~
+$[F(K_l)=1.174]$, (b) $\eta=15\%$ ~ [$[F(K_l)=0.606]$, and (c)
+$\eta=25\%$ ~ $[F(K_l)=0.457]$.
 %
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise quickly leads to a decrease in our model's performance.
-This is due to an inherent feature of the inverse scattering problem; that being that a two distinct regions of orientation space will become heavily intertwined in the intensity space despite remaining continuous.
-As the signal error increases our neural network will map the signal to the incorrect reference orientation.
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of noise
+quickly leads to a decrease in the method's performance.  This is due
+to an inherent feature of the inverse scattering problem: two distinct
+regions of orientation space will become highly intertwined in the
+intensity space despite each remaining continuously connected.  As the
+signal error increases our neural network will map the signal to the
+incorrect reference orientation.
 %
@@
 
-We have developed a method for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
-In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
-The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy.
-This has potential to be applied to characterise particle size and shape that are changing while in the optical trap.
-We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.
+We have developed a method for interpreting the dynamics of a trapped
+particle based purely on the light scattering pattern. In theory the
+model can be applied to any characteristic that impacts the light
+scattering pattern produced by the trapped entity. The MSTM package is
+flexible in calculating the light scattering for a variety of
+micro-particles; so long as there is a distinguishable difference in
+the light scattering pattern we can apply this method to characterise
+the particle to some degree of accuracy.  This has potential to be
+applied to characterise particle size and shape that are changing
+while in the optical trap. We hope to in the future apply this to our
+own experimental work by charactering the orientation of a trapped
+asymmetric particle.
 
-Additionally, we found that the due to the inverse light scattering problem that physical space regions can be continuously mapped to a space in the intensity region but with a loss of shape and size.
-In the future, we believe that we can improve our estimation by utilising a greater number of reference orientations and by optimising our choice of detection angles to better discretise our intensity regions.
+Additionally, we found that the due to the inverse light scattering
+problem that physical space regions can be continuously mapped to a
+space in the intensity region but with a loss of shape and size.  In
+the future, we believe that we can improve our estimation by utilising
+a greater number of reference orientations and by optimising our
+choice of detection angles to better discretise our intensity regions.
 

### 2023-02-23 python files/fitnest.py
Add files via upload

@@ -1,0 +1,45 @@
+import numpy as np
+import ultranest
+import ultranest.stepsampler
+from ultranest.plot import cornerplot
+
+class fit(object):
+    def __init__(self, params, x_data, y_data):
+        self.param_names = params
+        self.x = x_data
+        self.y = y_data
+            
+    def prior_transfrom(self, cube):
+        self.params = cube.copy()
+        for i in range(len(cube)):
+            self.params[i][0] = 10**(cube[i][0]*(8-6)+6)
+            self.params[i][1] = 10**(cube[i][1]*(8-6)+6)
+
+        return self.params
+
+    def log_likelihood(self, params):
+        likelihood = np.ndarray(shape=(len(params)))
+
+        for i in range(len(params)):
+            A = self.params[i][0]
+            B = self.params[i][1]
+            y_model = [1/(A+B*x**2) for x in self.x]
+            likelihood[i] = sum([((y-y_m)/0.05)**2 for y, y_m in zip(self.y, y_model)])
+
+        return likelihood
+
+    def run(self):
+        sampler = ultranest.ReactiveNestedSampler(self.param_names,
+        self.log_likelihood,
+        self.prior_transfrom,
+        resume = 'overwrite',
+        log_dir = './results',
+        vectorized = True)
+
+        print('Start')
+        sampler.stepsampler=ultranest.stepsampler.SliceSampler(nsteps=20, 
+        generate_direction=ultranest.stepsampler.generate_mixture_random_direction)
+        results=sampler.run(Lepsilon = 0.1)
+        sampler.print_results()
+        sampler.plot_run()
+        sampler.plot_corner()

### 2023-02-23 python files/plot.py

@@ -1,0 +1,60 @@
+import csv
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from fitnest import fit
+
+def row_reader(row):
+    for num in row:
+        if len(num) == 37:
+            item1, item2 =[], []
+            item1 = num[0:18]
+            item2 = num[19:37]
+        else:
+            item3, item4= [],[]
+            item3 = num[0:19]
+            item4 = num[20:38]
+
+    new_row =[float(item1), float(item2), float(item3), float(item4)]
+    return new_row
+        
+with open('11pnt6micron3pnt4amp.txt', 'r') as in_file:
+    reader = csv.reader(in_file.readlines()[4:])
+    
+    with open('log.csv', 'w') as out_file:
+        writer = csv.writer(out_file)
+        writer.writerows(row_reader(row) for row in reader)
+
+with open('log.csv', 'r') as file:
+    reader = csv.reader(file.readlines()[0::2])
+    freq, Px, Py =[], [], []
+    for row in reader:
+        freq.append(float(row[0]))
+        Px.append(float(row[1]))
+        Py.append(float(row[2]))
+
+params = ['A', 'B']
+lorentz = fit(params, freq[4:300], Px[4:300])
+lorentz.run()
+
+
+with open(r'C:\Users\Administrator\programs\python files\results\info\post_summary.csv') as csvfile:
+    reader = csv.reader(csvfile)
+    headings = next(reader)
+    row = next(reader)
+    A = float(row[0])
+    B = float(row[1])
+
+    fc = (A/B)**0.5
+    D = (2*np.pi**2/(24*B))
+
+y_model = [D/(f**2+fc**2) for f in freq]
+
+fig  = plt.figure() 
+ax = fig.add_subplot()  
+ax.plot(freq, Px)
+ax.plot(freq, y_model)
+ax.set_yscale('log')
+ax.set_xscale('log')
+plt.show()
+

### 2023-02-23 python files/PSD_fit.py
update

@@ -1,0 +1,98 @@
+import csv
+import numpy as np
+import matplotlib.pyplot as plt;
+from scipy.optimize import curve_fit;
+from fitnest import fit;
+
+#Code for reading and fitting PSD data from FDdat files
+
+#Function reads FDdat file and formats rows into csv cells [Frequency, Px, Py, Sum]
+def row_reader(row):
+    for num in row:
+        lst = row[0].split()
+        item1= lst[0]
+        item2=lst[1]
+        item3=lst[2]
+        item4=lst[3]
+
+    new_row =[float(item1), float(item2), float(item3), float(item4)]
+    return new_row
+
+#Simplified Lorentzian fit of power spectrum with constants A and B
+def lorentz(x,A,B):
+    return (1/(A+B*x**2))
+
+#Function reads the name of the file and stores the graph in appropriate folder with useful name.  
+def get_label_path(label):
+    label = label.split('_')
+    if label[0] == 'Sillica':
+        material = 'Sillica'
+    elif label[0] == 'Vaterite':
+        material = 'Vaterite'
+    elif label[0] == 'PS':
+        material = 'PS'
+    else:
+        material = 'LC'
+    
+    power = label[1]
+    size = label[2][:-4]
+    return material, power, size
+
+#lead path is just the neccesarray jargon to locate directory 
+lead_path = '/home/daniel/Documents/dmaciver97/python files/' 
+label = 'Sillica_2.5A_1.57mu.txt'
+material, power, size= get_label_path(label)
+
+path = lead_path+label
+        
+#convert FDdat file to csv file and then store first 3 coloums as [freq, Px, Py]
+with open(path, 'r') as in_file:
+    reader = csv.reader(in_file.readlines()[4:])
+    
+    with open('log.csv', 'w') as out_file:
+        writer = csv.writer(out_file)
+        writer.writerows(row_reader(row) for row in reader)
+
+with open('log.csv', 'r') as file:
+    reader = csv.reader(file.readlines())
+    freq, Px, Py =[], [], []
+    for row in reader:
[…]

### 2023-02-23 Experiment3.py
git config pull.rebase false

@@ -1,648 +1,0 @@
-#!/usr/bin/python3
-
-import heapq
-import math
-import csv
-import numpy as np
-import pickle
-import pylab as plt
-from statistics import mean, stdev
-from scipy.interpolate import splrep, splev, splint
-from scipy.integrate import tplquad
-from itertools import product
-from numpy.linalg import norm
-import random
-import argparse
-import itertools
-import xml.etree.cElementTree as ET
-from xml.dom import minidom
-import multiprocessing
-
-
-from Beam import *
-from MSTM import *
-from Particle import *
-
-neighbor_list = [np.array(vec) for vec in product([-1, 0, 1], repeat=3)]
-
-
-
-
-################################################################################
-################################################################################
-################################################################################
-class Experiment3(object):
-
-
-    def __init__(self, pklfile='fit.pkl', angle_list=None):
-        
-        self.theta_list = None
-        self.n_list = self.set_measurement_angles        
-        self.sigma = 0.1e1
-        self.fit_dict = self.read_pickle(pklfile)
-        self.N = 10000
-        self.mstm = MSTM()
-
-        if (angle_list is None):
-            angle_list = [15.0, 45.0, 90.0]
-        self.set_measurement_angles(angle_list)
-
-        
-################################################################################
-    def read_pickle(self, filename):
-        pklfile = open(filename, 'rb')
-        fit_dict = pickle.load(pklfile)
-        pklfile.close()
-        return fit_dict
-
-
-################################################################################
-    def set_measurement_angles(self, theta_list):
-
-        self.theta_list = theta_list
-        
-        n_list = []
-        Iraw_measurements = [[] for theta in theta_list]
-        Deriv_measurements =[[] for theta in theta_list]
-        sum_prob = 0.0
-        for n, fit in self.fit_dict.items():
-            Iraw = [ float(splev(theta, fit) ) for theta in theta_list]
-            Deriv = [float(splev(theta, fit, der =1)) for theta in theta_list]
-            for index, I in enumerate(Iraw):
-                Iraw_measurements[index].append(I)
-            prob = 1.0
-            sum_prob += prob
[…]

### 2023-02-23 Scoring2.py

@@ -1,325 +1,0 @@
-#!/usr/bin/python3.9
-
-import numpy as np
-from statistics import mean, stdev
-from scipy.interpolate import splrep, splev, splint
-import quaternion
-from subprocess import check_call
-import random as rand
-from quaternion import as_rotation_matrix
-from multiprocessing import Process, Queue, Lock
-import matplotlib.pyplot as plt
-import Experiment3
-import xml.etree.ElementTree as ET
-from datetime import datetime
- 
-class Scoring2(object):
-
-    def __init__(self):
-        self.traj = np.load('traj_test.csv.npy')
-        
-        self.time_steps = 1000
-        self.dt = 1e-3
-        self.pkl = 'fit.pkl'
-        self.sim = Experiment3.Experiment3(self.pkl, angle_list=None)
-        self.angle_list = [15,55,90]
-
-    def test(self, angle_list, epsilon):
-        #Main function for iterating through the main traj list and assigning an estimtated orienation
-        self.angle_list = angle_list
-        self.sim.set_measurement_angles(self.angle_list)
-        self.t_list, self.dist_list, self.est_list = [], [], []
-        self.prob_max_list =[]
-        s_list, est, true=[], [], []
-        #print(self.sim.n_list)
-        
-
-        for count, frame in enumerate(self.traj[0:10]):
-            #print(count)
-            q = np.quaternion(frame[3], frame[4], frame[5], frame[6])
-            R = as_rotation_matrix(q)
-            s = np.array([R[0,2], R[1,2], R[2,2]])
-            s_list.append(s)
-            
-            fit = self.sim.get_scatter_fit(s)
-            I_raw = [splev(angle, fit) for angle in self.angle_list]
-
-            index  = self.sim.get_distances(s)
-            n = self.sim.n_list[index]['n']
-            true.append(n)
-            #s_prior is used to inform our prior estimate of the particle's orienation
-            if count != 0:
-                s_prior = n_est
-                
-            else:
-                s_prior = s
-
-            index_est, self.prob_list = self.sim.estimate_orientation(I_raw,
-                                                                      s_prior,
-                                                                      epsilon)
-        
-            n_est =  self.sim.n_list[index_est]['n']
-            est.append(n_est)
[…]

### 2023-02-23 python files/fitnest.py

@@ -12,8 +12,8 @@ class fit(object):
     def prior_transfrom(self, cube):
         self.params = cube.copy()
         for i in range(len(cube)):
-            self.params[i][0] = 10**(cube[i][0]*(8-6)+6)
-            self.params[i][1] = 10**(cube[i][1]*(8-6)+6)
+            self.params[i][0] = 10**(cube[i][0]*8+4)
+            self.params[i][1] = 10**(cube[i][1]*8+4)
 
         return self.params
 
@@ -24,7 +24,7 @@ class fit(object):
             A = self.params[i][0]
             B = self.params[i][1]
             y_model = [1/(A+B*x**2) for x in self.x]
-            likelihood[i] = sum([((y-y_m)/0.05)**2 for y, y_m in zip(self.y, y_model)])
+            likelihood[i] = (((sum([((y-y_m)/1)**2 for y, y_m in zip(self.y, y_model)]))))
 
         return likelihood
 

### 2023-02-23 python files/plot.py

@@ -1,60 +1,0 @@
-import csv
-import numpy as np
-import matplotlib.pyplot as plt
-from scipy.optimize import curve_fit
-from fitnest import fit
-
-def row_reader(row):
-    for num in row:
-        if len(num) == 37:
-            item1, item2 =[], []
-            item1 = num[0:18]
-            item2 = num[19:37]
-        else:
-            item3, item4= [],[]
-            item3 = num[0:19]
-            item4 = num[20:38]
-
-    new_row =[float(item1), float(item2), float(item3), float(item4)]
-    return new_row
-        
-with open('11pnt6micron3pnt4amp.txt', 'r') as in_file:
-    reader = csv.reader(in_file.readlines()[4:])
-    
-    with open('log.csv', 'w') as out_file:
-        writer = csv.writer(out_file)
-        writer.writerows(row_reader(row) for row in reader)
-
-with open('log.csv', 'r') as file:
-    reader = csv.reader(file.readlines()[0::2])
-    freq, Px, Py =[], [], []
-    for row in reader:
-        freq.append(float(row[0]))
-        Px.append(float(row[1]))
-        Py.append(float(row[2]))
-
-params = ['A', 'B']
-lorentz = fit(params, freq[4:300], Px[4:300])
-lorentz.run()
-
-
-with open(r'C:\Users\Administrator\programs\python files\results\info\post_summary.csv') as csvfile:
-    reader = csv.reader(csvfile)
-    headings = next(reader)
-    row = next(reader)
-    A = float(row[0])
-    B = float(row[1])
-
-    fc = (A/B)**0.5
-    D = (2*np.pi**2/(24*B))
-
-y_model = [D/(f**2+fc**2) for f in freq]
-
-fig  = plt.figure() 
-ax = fig.add_subplot()  
-ax.plot(freq, Px)
-ax.plot(freq, y_model)
-ax.set_yscale('log')
-ax.set_xscale('log')
-plt.show()
-

### 2023-02-24 python files/PSD_fit.py
git config pull.rebase false

@@ -2,25 +2,25 @@ import csv
 import numpy as np
 import matplotlib.pyplot as plt;
 from scipy.optimize import curve_fit;
-from fitnest import fit;
+from fitting import fit
 
 #Code for reading and fitting PSD data from FDdat files
 
 #Function reads FDdat file and formats rows into csv cells [Frequency, Px, Py, Sum]
 def row_reader(row):
     for num in row:
-        lst = row[0].split()
-        item1= lst[0]
-        item2=lst[1]
-        item3=lst[2]
-        item4=lst[3]
-
+        lst = (num.split(' '))
+        if len(num)==37:
+            item1= lst[0]
+            item2=lst[1]
+        else:
+            item3=lst[1]
+            item4=lst[2]
+    
     new_row =[float(item1), float(item2), float(item3), float(item4)]
     return new_row
 
-#Simplified Lorentzian fit of power spectrum with constants A and B
-def lorentz(x,A,B):
-    return (1/(A+B*x**2))
+
 
 #Function reads the name of the file and stores the graph in appropriate folder with useful name.  
 def get_label_path(label):
@@ -40,7 +40,7 @@ def get_label_path(label):
 
 #lead path is just the neccesarray jargon to locate directory 
 lead_path = '/home/daniel/Documents/dmaciver97/python files/' 
-label = 'Sillica_2.5A_1.57mu.txt'
+label = 'Vaterite_2.00A_3mu.txt'
 material, power, size= get_label_path(label)
 
 path = lead_path+label
@@ -61,12 +61,15 @@ with open('log.csv', 'r') as file:
         Px.append(float(row[1]))
         Py.append(float(row[2]))
 
-#curve_fit uses lorentzian to fit to data 
-popt, pcov = curve_fit(lorentz, freq[1:300], Px[1:300], p0=[1000,1])
-A,B= popt
+#Simplified Lorentzian fit of power spectrum with constants A and B
+def lorentz(x,A,B):
+    return (1/(A+B*x**2))
+
+popt, pcov  = curve_fit(lorentz, freq[5:300], Px[5:300])
+A,B = popt
 fc = (A/B)**0.5
-D = (2*np.pi**2/(15*B))
[…]

### 2023-02-24 python files/Scoring2.py

@@ -1,0 +1,325 @@
+#!/usr/bin/python3.9
+
+import numpy as np
+from statistics import mean, stdev
+from scipy.interpolate import splrep, splev, splint
+import quaternion
+from subprocess import check_call
+import random as rand
+from quaternion import as_rotation_matrix
+from multiprocessing import Process, Queue, Lock
+import matplotlib.pyplot as plt
+import Experiment3
+import xml.etree.ElementTree as ET
+from datetime import datetime
+ 
+class Scoring2(object):
+
+    def __init__(self):
+        self.traj = np.load('traj_test.csv.npy')
+        
+        self.time_steps = 1000
+        self.dt = 1e-3
+        self.pkl = 'fit.pkl'
+        self.sim = Experiment3.Experiment3(self.pkl, angle_list=None)
+        self.angle_list = [15,55,90]
+
+    def test(self, angle_list, epsilon):
+        #Main function for iterating through the main traj list and assigning an estimtated orienation
+        self.angle_list = angle_list
+        self.sim.set_measurement_angles(self.angle_list)
+        self.t_list, self.dist_list, self.est_list = [], [], []
+        self.prob_max_list =[]
+        s_list, est, true=[], [], []
+        #print(self.sim.n_list)
+        
+
+        for count, frame in enumerate(self.traj[0:10]):
+            #print(count)
+            q = np.quaternion(frame[3], frame[4], frame[5], frame[6])
+            R = as_rotation_matrix(q)
+            s = np.array([R[0,2], R[1,2], R[2,2]])
+            s_list.append(s)
+            
+            fit = self.sim.get_scatter_fit(s)
+            I_raw = [splev(angle, fit) for angle in self.angle_list]
+
+            index  = self.sim.get_distances(s)
+            n = self.sim.n_list[index]['n']
+            true.append(n)
+            #s_prior is used to inform our prior estimate of the particle's orienation
+            if count != 0:
+                s_prior = n_est
+                
+            else:
+                s_prior = s
+
+            index_est, self.prob_list = self.sim.estimate_orientation(I_raw,
+                                                                      s_prior,
+                                                                      epsilon)
+        
+            n_est =  self.sim.n_list[index_est]['n']
+            est.append(n_est)
[…]

### 2023-02-24 python files/fitting.py

@@ -1,0 +1,57 @@
+from scipy.optimize import curve_fit
+
+
+#Simplified Lorentzian fit of power spectrum with constants A and B
+def lorentz(x,A,B):
+    return (1/(A+B*x**2))
+
+def fit(x_data, y_data):
+    index = -1
+    error = 1e10
+    for i in range(20):
+        index += 1
+        popt, pcov = curve_fit(lorentz, x_data[index:], y_data[index:], p0 = [10000,10])
+        A, B = popt
+        y_model = [1/(A+B*x**2) for x in x_data]
+        new_error = sum([np.log(((y_m-y)/5)**2) for y_m, y in zip(y_model, y_data)])
+        print(new_error)
+        if new_error < error:
+            error = new_error
+            A_best =  A
+            B_best = B
+            print(index)
+    
+
+    print(A_best, B_best)
+    return A_best, B_best
+
+
+if __name__ == '__main__':
+    import csv
+    import numpy as np
+    import matplotlib.pyplot as plt
+    with open('log.csv', 'r') as file:
+        reader = csv.reader(file.readlines())
+        freq, Px, Py =[], [], []
+        for row in reader:
+            freq.append(float(row[0]))
+            Px.append(float(row[1]))
+            Py.append(float(row[2]))
+
+    A, B  = fit(freq, Px)
+    fc = (A/B)**0.5
+    D = (2*np.pi**2/(15*B))
+    Px_model = [D/(f**2+fc**2) for f in freq]
+
+    fig  = plt.figure()
+    ax1 = fig.add_subplot() 
+    ax1.plot(freq[1:], Px[1:])
+    ax1.plot(freq[1:], Px_model[1:], label= f'${np.round(D,3)}/(f^2+{np.round(fc,2)}^2)$')
+    ax1.legend()
+    ax1.set_ylabel('Power $[V^2/Hz]$')
+    ax1.set_xlabel('Frequency [Hz]')
+    ax1.set_yscale('log')
+    ax1.set_xscale('log')
+    ax1.set_xlim(1,2.0e03)
+    plt.show()
+

### 2023-02-24 python files/fitting.py
tidying

@@ -1,57 +1,0 @@
-from scipy.optimize import curve_fit
-
-
-#Simplified Lorentzian fit of power spectrum with constants A and B
-def lorentz(x,A,B):
-    return (1/(A+B*x**2))
-
-def fit(x_data, y_data):
-    index = -1
-    error = 1e10
-    for i in range(20):
-        index += 1
-        popt, pcov = curve_fit(lorentz, x_data[index:], y_data[index:], p0 = [10000,10])
-        A, B = popt
-        y_model = [1/(A+B*x**2) for x in x_data]
-        new_error = sum([np.log(((y_m-y)/5)**2) for y_m, y in zip(y_model, y_data)])
-        print(new_error)
-        if new_error < error:
-            error = new_error
-            A_best =  A
-            B_best = B
-            print(index)
-    
-
-    print(A_best, B_best)
-    return A_best, B_best
-
-
-if __name__ == '__main__':
-    import csv
-    import numpy as np
-    import matplotlib.pyplot as plt
-    with open('log.csv', 'r') as file:
-        reader = csv.reader(file.readlines())
-        freq, Px, Py =[], [], []
-        for row in reader:
-            freq.append(float(row[0]))
-            Px.append(float(row[1]))
-            Py.append(float(row[2]))
-
-    A, B  = fit(freq, Px)
-    fc = (A/B)**0.5
-    D = (2*np.pi**2/(15*B))
-    Px_model = [D/(f**2+fc**2) for f in freq]
-
-    fig  = plt.figure()
-    ax1 = fig.add_subplot() 
-    ax1.plot(freq[1:], Px[1:])
-    ax1.plot(freq[1:], Px_model[1:], label= f'${np.round(D,3)}/(f^2+{np.round(fc,2)}^2)$')
-    ax1.legend()
-    ax1.set_ylabel('Power $[V^2/Hz]$')
-    ax1.set_xlabel('Frequency [Hz]')
-    ax1.set_yscale('log')
-    ax1.set_xscale('log')
-    ax1.set_xlim(1,2.0e03)
-    plt.show()
-

### 2023-02-24 Latex files/Manuscript.tex
Tidying up repo

--- 2022-11-14 Manuscript.tex
@@
 where $\epsilon$ is the percent error expected from our detection
-fibres.  Equation~\eqref{eq:conditional} assigns a probability to each
-reference orientation based on how similar the ``reference orientation
-signal'' $y_{k}(\hat{\bf n}_\alpha)$ is to the actual signal
-$y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation
-$\bar{\bf s}$.  However, if our values of $y_k(\hat{\bf n}_\alpha)$
-are similar to one another, the probability distribution from
-Eq.~\eqref{eq:conditional} will therefore assign similar probability
-for multiple reference orientations.  Bayesian inference allows one to
-update a probability distribution by accounting for information known
-prior to making a measurement.  We apply this by accounting for our
-distribution of reference orientations and instantaneous signals.
+fibres.
+
+Equation~\eqref{eq:conditional} gives the probability that a signal
+($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference
+orientation $\hat{\rm n}_\alpha$.
+%
+Using Bayes' theorem, the probability that the dimer is in reference
+orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
+$y_3$) is measured, can be estimated:
+%
+%
+%
+%update a probability distribution by accounting for information known
+%prior to making a measurement.  We apply this by accounting for our
+%distribution of reference orientations and instantaneous signals.
 \begin{align}
@@
 estimates of the distributions of particle orientations and
-instantaneous signals, respectively.  The former can be estimated by
-some Boltzmann distribution based on how far apart reference
-orientations are in physical space, covered in Section~\ref{sec:2.3}.
-The latter prior $p(y)$ is the probability of our dimer producing a
-signal $y_k(\bar{\bf s})$ that falls with our distribution of
-reference signals $y_k(\hat{\bf n})$.  This is given by taking the
+instantaneous signals, respectively.
+%
+The former can be estimated by some Boltzmann distribution based on
+how far apart reference orientations are in physical space, covered in
+Section~\ref{sec:2.3}.  The latter prior $p(y)$ is the probability of
+measuring a signal ($y_1$, $y_2$, $y_3$).  This is given by taking the
 discrete integral over collection of reference orientations.
@@
 simulation of the dimer (see Section~\ref{sec:2.1}).
+%
+% based on how similar the ``reference orientation
+%signal'' $y_{k}(\hat{\bf n}_\alpha)$ is to the actual signal
+%$y_{k}(\bar{\bf s})$ generated by the dimer in its real orientation
+%$\bar{\bf s}$.
+%
+However, if our values of $y_k(\hat{\bf n}_\alpha)$
+are similar to one another, the probability distribution from
+Eq.~\eqref{eq:conditional} will therefore assign similar probability
+for multiple reference orientations.
 

### 2023-02-24 Latex files/Optimizationnotes.tex
Tidying up repo

--- 2022-10-27 Optimizationnotes.tex
@@
 
-The experiemental set up that this is based on is a inverted microscope optical trap, a suspension of microparticles are loaded onto a plastic cover slide.
-Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm HeNe laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
-The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microcopes camera.
-The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lense of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by []'s paper.
+We are currently developing an experimental set up that will allow us to light scattering
 
@@
 
-\subsection*{Further improvements}
-While our model provides a satisfactory prediction of the dimer's motion; ideally this model can be exteneded to any particle characteristic that has a clear effect on the scattering from the trapped entity.
-For example using the mstm package we can determine how differently sized particles in similar orientations will produce different scattering patterns.
-
 \section*{Conclusion}

### 2023-02-24 python files/PSD_fit.py
Formatted PSD.py

@@ -1,8 +1,14 @@
 import csv
+import argparse
 import numpy as np
 import matplotlib.pyplot as plt;
 from scipy.optimize import curve_fit;
-from fitting import fit
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--label', type=str, default='Vaterite_2.00A_3mu.txt', help='FDdat text file')
+args = parser.parse_args()
+
+label = args.label
 
 #Code for reading and fitting PSD data from FDdat files
 
@@ -36,12 +42,12 @@ def get_label_path(label):
     
     power = label[1]
     size = label[2][:-4]
-    return material, power, size
+    radius  = float(size[:-2])*1e-6
+    return material, power, size, radius
 
 #lead path is just the neccesarray jargon to locate directory 
-lead_path = '/home/daniel/Documents/dmaciver97/python files/' 
-label = 'Vaterite_2.00A_3mu.txt'
-material, power, size= get_label_path(label)
+lead_path = '/home/daniel/Documents/dmaciver97/python files/FDdat files/' 
+material, power, size, radius= get_label_path(label)
 
 path = lead_path+label
         
@@ -65,18 +71,27 @@ with open('log.csv', 'r') as file:
 def lorentz(x,A,B):
     return (1/(A+B*x**2))
 
+#Fitting the data to a Lorentzian and then plotting the plots for the X and Y QPD axis
+# Calculate conversion factor Beta
+
+gamma0 = 6*np.pi*1.0013e-3*radius 
+D_einstein = 1.38e-23*290/gamma0  #units m^2/s
+
 popt, pcov  = curve_fit(lorentz, freq[5:300], Px[5:300])
 A,B = popt
 fc = (A/B)**0.5
-D = (2*np.pi**2/(B))
-Px_model = [D/(2*np.pi**2*(f**2+fc**2))  for f in freq]
+D = (2*np.pi**2/(B)) #units V^2/s
+Beta = D/D_einstein #V^2/m^2
+Px_model = [D/(Beta*2*np.pi**2*(f**2+fc**2))  for f in freq]
+Px = [x/Beta for x in Px]
 
[…]

### 2023-02-24 python files/PSD_fit.py
Mod lorentz

@@ -60,7 +60,7 @@ with open(path, 'r') as in_file:
         writer.writerows(row_reader(row) for row in reader)
 
 with open('log.csv', 'r') as file:
-    reader = csv.reader(file.readlines())
+    reader = csv.reader(file.readlines()[:300])
     freq, Px, Py =[], [], []
     for row in reader:
         freq.append(float(row[0]))
@@ -70,6 +70,9 @@ with open('log.csv', 'r') as file:
 #Simplified Lorentzian fit of power spectrum with constants A and B
 def lorentz(x,A,B):
     return (1/(A+B*x**2))
+
+def mod_lorentz(x,a,b,c,fc):
+    return (a+(b/((x**2+fc**2)+c**2)))
 
 #Fitting the data to a Lorentzian and then plotting the plots for the X and Y QPD axis
 # Calculate conversion factor Beta
@@ -117,3 +120,17 @@ ax2.set_xscale('log')
 ax2.set_xlim(1,2.0e03)
 fig.tight_layout()
 plt.savefig(f'/home/daniel/Documents/dmaciver97/python files/figures/{material}/{size}@{power}.png')
+
+#popt, pcov =  curve_fit(mod_lorentz, freq[6:300], Px[6:300])
+#a,b,c,fc = popt
+
+#model = [a+(b/((x**2+fc**2)+c**2)) for x in Px]
+
+#plt.close('all')
+#fig = plt.figure()
+#ax = fig.add_subplot()
+#ax.plot(freq[1:], Px[1:])
+#ax.plot(freq[1:], model[1:])
+#ax.set_yscale('log')
+#ax.set_xscale('log')
+#plt.show()

### 2023-02-24 python files/PSD_fit.py
update

@@ -72,7 +72,7 @@ def lorentz(x,A,B):
     return (1/(A+B*x**2))
 
 def mod_lorentz(x,a,b,c,fc):
-    return (a+(b/((x**2+fc**2)+c**2)))
+    return (a+(b/((x-fc)**2+c**2)))
 
 #Fitting the data to a Lorentzian and then plotting the plots for the X and Y QPD axis
 # Calculate conversion factor Beta
@@ -86,7 +86,7 @@ fc = (A/B)**0.5
 D = (2*np.pi**2/(B)) #units V^2/s
 Beta = D/D_einstein #V^2/m^2
 Px_model = [D/(Beta*2*np.pi**2*(f**2+fc**2))  for f in freq]
-Px = [x/Beta for x in Px]
+#Px = [x/Beta for x in Px]
 
 D_label = str('{:.2e}'.format(D/Beta))
 fig  = plt.figure()
@@ -124,7 +124,7 @@ plt.savefig(f'/home/daniel/Documents/dmaciver97/python files/figures/{material}/
 #popt, pcov =  curve_fit(mod_lorentz, freq[6:300], Px[6:300])
 #a,b,c,fc = popt
 
-#model = [a+(b/((x**2+fc**2)+c**2)) for x in Px]
+#model = [a+(b/((x-fc)**2+c**2)) for x in Px]
 
 #plt.close('all')
 #fig = plt.figure()

### 2023-03-27 python files/PSD_fit.py
update tp gitignore

@@ -15,6 +15,7 @@ label = args.label
 #Function reads FDdat file and formats rows into csv cells [Frequency, Px, Py, Sum]
 def row_reader(row):
     for num in row:
+
         lst = (num.split(' '))
         if len(num)==37:
             item1= lst[0]
@@ -46,7 +47,7 @@ def get_label_path(label):
     return material, power, size, radius
 
 #lead path is just the neccesarray jargon to locate directory 
-lead_path = 'C:\\Users\\Administrator\\programs\\git repos\\dmaciver97\\python files\\FDdat files\\'
+lead_path = "FDdat files\\"
 material, power, size, radius= get_label_path(label)
 
 path = lead_path+label
@@ -60,7 +61,7 @@ with open(path, 'r') as in_file:
         writer.writerows(row_reader(row) for row in reader)
 
 with open('log.csv', 'r') as file:
-    reader = csv.reader(file.readlines()[:300])
+    reader = csv.reader(file.readlines()[:300:2])
     freq, Px, Py =[], [], []
     for row in reader:
         freq.append(float(row[0]))
@@ -80,13 +81,13 @@ def mod_lorentz(x,a,b,c,fc):
 gamma0 = 6*np.pi*1.0013e-3*radius 
 D_einstein = 1.38e-23*290/gamma0  #units m^2/s
 
-popt, pcov  = curve_fit(lorentz, freq[5:300], Px[5:300])
+popt, pcov  = curve_fit(lorentz, freq[10:300], Px[10:300])
 A,B = popt
 fc = (A/B)**0.5
 D = (2*np.pi**2/(B)) #units V^2/s
 Beta = D/D_einstein #V^2/m^2
 Px_model = [D/(Beta*2*np.pi**2*(f**2+fc**2))  for f in freq]
-#Px = [x/Beta for x in Px]
+Px = [x/Beta for x in Px]
 
 D_label = str('{:.2e}'.format(D/Beta))
 fig  = plt.figure()
@@ -100,7 +101,7 @@ ax1.set_yscale('log')
 ax1.set_xscale('log')
 ax1.set_xlim(1,2.0e03)
 
-popt, pcov  = curve_fit(lorentz, freq[5:300], Py[5:300])
+popt, pcov  = curve_fit(lorentz, freq[10:300], Py[10:300])
 A,B = popt
 fc = (A/B)**0.5
 D = (2*np.pi**2/(B))
[…]

### 2023-03-29 Latex files/Manuscript_rev2 mh.tex
Tidying up Latex files

\documentclass[final,  3p]{elsarticle}

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomencl}
\makenomenclature
\renewcommand{\nomlabel}[1]{\hfil #1\hfil}

\journal{Journal of Quantitative Spectroscopy \& Radiative Transfer}

\begin{document}

\begin{frontmatter}

\title{Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles}

\author[aff1]{Dan Maciver\corref{cor1}} 
\ead{Daniel.Maciver.2016@uni.strath.ac.uk}

\author[aff1]{Praveen Parthasarathi}

\author[aff1]{Leo Lue}

\author[aff1]{Jan Sefcik}

\author[aff1]{Mark Haw}

\cortext[cor1]{Corresponding author}
\affiliation[aff1]{organization={Department of Chemical Engineering,
			University of Strathclyde},
            addressline={75 Montrose Street}, 
            city={Glasgow},
            postcode={G1 1XL}, 
            country={Scotland}}

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection, further characterisation of trapped entities using light-scattering
  poses a two-fold challenge --- one experimental, concerning the optimal
  arrangement of light detectors to gather data, and the other theoretical, involving
  solving of the inverse light scattering problem in order to interpret this data. Experimentally, combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose here a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data, combined with a neural-network trained on data simulated to mimic light scattering
  detection signals in such scenarios, may be used for solving the
  inverse light scattering problem and characterising complex trapped
  entities.  To demonstrate the method we discuss its application to
  measuring the instantaneous orientations of a trapped asymmetric microsphere
  dimer. We argue that the method can be extended to
  determine any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

\begin{highlights}
\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.  
\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
\end{highlights}

\begin{keyword}
	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
\end{keyword}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}

Since their invention in the late 1980s, optical tweezers have found application 
in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to
 testing the fundamental assumptions of quantum mechanics  \cite{yin2013large}, thanks 
mainly to the ability of the tweezer to transduce and detect forces on the order of a 
few pico newtons. Going beyond forces, further structural, dynamic and chemical 
characterisation of complex trapped entities could provide useful information, as 
demonstrated in   areas such as metrology \cite{arita2020coherent} and colloidal
 aggregation \cite{burns1990optical}. Spectroscopic techniques such as Raman scattering 
\cite{gupta2014raman} have been used for chemical characterisation of trapped objects, 
while dynamical characterisation has been demonstrated using data from the tweezer's 
Quadrant Photo Detector by following the centre-of-mass Brownian motion of the trapped 
entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass 
\cite{yifat2021facile}.

The possibility of integrating an \textit{additional} light scattering detection scheme 
within an optical trapping setup was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998}, 
where a single-mode optical fibre was aligned to detect the scattered
light from a trapped bead and study its Brownian motion.  While this provided dynamical 
information, structural information about the trapped bead was precluded as the measurement 
was obtained only at a single angle.  In this work, we propose a scheme that expands on the 
technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of 
angles (Figure~\ref{fig:setup}), combined with a novel Bayesian inference-based analysis
 technique to enable interpretation of the resulting multi-angle data as well as 
optimisation to provide maximal information from the signal.

To demonstrate the analysis, we study a simple, illustrative example of a trapped entity 
more complex than a single sphere \textit{i.e.} a trapped asymmetric dimer.  
As a paradigmatic example of extracting structural/dynamic information from the 
resulting scattering data, we explore how to estimate the dimer's instantaneous 
orientation from the scattering signals using Bayesian inference, as well as how 
to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable 
estimate. As an example of the relatively sparse literature on measuring orientations of
 complex trapped objects, ref.~\cite{raudsepp2022estimating} employs imaging to study the 
orientation of trapped dimers: scattering can give more quantitative and potentially more
 rapid time-resolved information, but only of course if the scattering signals can be 
interpreted. Here, we first train a neural network to effectively identify the mapping
 between scattering signals and dimer orientation, by calculating the scattering signal 
from a  simulated asymmetric dimer undergoing Brownian motion in an optical trap and 
mapping to the known instantaneous orientation of the simulated dimer. We then show 
how Bayesian inference can be used to optimise extraction of the true dimer orientation
 from the light scattering signals. Furthermore we demonstrate how the model can be
 fine tuned in situations where measurement uncertainty becomes significant.

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
\end{subfigure}
\caption{\label{fig:setup}
  %
  Proposed experimental set up for scattering measurements from an object in an optical trap. The probe beam for scattering measurements is incident perpendicular to the trapping laser propagation direction. a) Side view. b) Top view. Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see text).
%
}
\end{figure}

\printnomenclature[0.55in]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\bar{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation estimation from scattering measurements}
\label{sec:Bayes}

For our simulated dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere to the centre of the smaller sphere. A plane wave 'probe' laser, perpendicular to the trapping laser, is incident on the dimer, generating a scattering pattern  dependent on the dimer's orientation $I(\hat{\bf s}, \theta)$ which is computed using MSTM. To represent the experimental set up consisting of a set of optical fibres recording scattered light, we choose four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) and record the calculated intensity at each angle $\theta_k$, $I(\hat{s}, \theta_k)$.

Our goal is to determine the orientation of the trapped dimer based on the scattering data $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space  into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} \ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a unit sphere \cite{Rey2006} (Figure~\ref{fig:dimer}b) leading to a maximum nearest-neighbour spacing between two neighbouring reference orientations of 0.895 radians. Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. (Note that if all measurement angles were in the same plane we would obtain identical signals for certain sets of different reference orientations, hence $\theta_4$ projects out of the horizontal plane). The raw intensities are normalized according to:
\begin{align}
\label{eq:scale}
  y_k(\hat{\bf n}_\alpha)
  = 
  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}
where the denominator is simply the standard deviation across the set of values $I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex. Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks. We thus utilised the Python machine learning program \textit{scikit-learn} to build a neural network for identifying the dimer's orientation from its light scattering signal. The network was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then using the network to estimate the probability of a given signal coming from a dimer in a given reference orientation. The network's loss function was evaluated and used to improve the estimation, the network being trained until the improvement in the loss function was less than 0.0001. 
%
Importantly, the estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer,  utilising Bayesian inference to update the neural network's estimation:

\begin{align}
  \label{eq:bayes}
  p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
  &=
    \frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
    p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
\end{align}
where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of the distributions of particle orientations and instantaneous signals, respectively.
\textit{Without} any prior evidence we must assume that the orientation prior of the dimer
$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the dimer's possible current orientation from knowledge of previous
measurements can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$ (see Section~\ref{sec:test}).  The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).  This is given by taking the discrete integral over the collection of reference orientations:
\begin{align}
  p(y_1, y_2, y_3, y_4)
  =
  \sum_{\alpha=1}^{n_{\rm ref}}
  p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
  p(\hat{\bf n}_\alpha)
\end{align}
From \eqref{eq:bayes} we obtain the key result, a mass probability distribution denoting the
probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a measured
signal ($y_1$, $y_2$, $y_3$), \textit{i.e.} an estimated mapping from scattering measurement to orientation estimate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimate's actual accuracy. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and  Discussion}
\label{sec:Discussion}\subsection{Testing the Model}
\label{sec:test}
Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap. The trapping laser is 1064nm NIR focused through a 1.25 NA objective. The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively. We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.

We applied Eq.~\eqref{eq:bayes}, taking the reference orientation with the highest probability  as our estimate of the dimer's instantaneous orientation $\hat{\bf n}_{est}$. To visualise the model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time. For comparison, we also plotted the radian distance between the  dimer's instantaneous orientation and the closest reference orientation, denoted $\hat{\bf n}_{best}$. The dotted line indicates the maximum radian distance ($0.896$ radians) between two \textit{neighbouring} reference orientations: if we are under this line then we know our estimate is at least neighbouring the best result. Assuming a uniform prior of the reference orientations $p(\hat{\bf n}_\alpha)$  the neural network's predictions ($\hat{\bf n}_{est}$ from Eq.~\eqref{eq:bayes}) are at times reasonable, but there are significant large and random jumps away from the correct result (Fig.~\ref{fig:uniform}).

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/fig3.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Horizontal dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$.}
	\label{fig:uniform}
\end{figure}

One reason we observe such large jumps in  orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space : even a large change in orientation can involve a small change in scattering. Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step. To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior,  accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness. Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step according to the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$, \textit{i.e.} taking into account the different likelihoods of the estimated orientation jumping between nearby or distant reference orientations:

\begin{align}
  p(\hat{\bf n}_\alpha)
  &=\frac{e^{\beta (\hat{\bf n}_\alpha 
  	\cdot \hat{\bf n}_{est}(t-\Delta t))}}
  {\sum_{\alpha=1}^{n_{\rm ref}}
	e^{\beta (\hat{\bf n}_\alpha 
	\cdot \hat{\bf n}_{est}(t-\Delta t)}}
	\label{eq:boltz}
\end{align}
Here $\beta$ is a weighting factor describing the dimer's freedom of motion within the trap. As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{./Images/fig4.png}
\caption{\label{fig:biased}
%
Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
Dotted line denotes spacing between two neighbouring reference
orientations $\hat{\bf n}_\alpha$ (see Section~\ref{sec:Bayes}).
  %
}
\end{figure}

The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}. By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation. To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
\begin{align}
  K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#}
  \\
  K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right]
  \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

The worst case scenario is akin to randomly choosing a reference
orientation at each time step. The greater the value of $F(K_l)$, the
better our model's confidence is in characterising the dimer's
motion. Because our model is dependent on several parameters we need
to a sophisticated method for understanding how these parameters
correlate with $F(K_l)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Asymmetric dimer dynamics}
\label{sec:motion}

The Brownian OT software was used to simulate the motion of a trapped
dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first
$10$ seconds of entering the optical trap.  The initial orientation
was assumed as strictly vertical (in line with the beam propagation
direction). The dimer's position and orientation was recorded every
$1\,{\rm ms}$ for using as a test dataset for our model.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig5a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig5b.png}
	\end{subfigure}
	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it reaches a horizontal orientation before drawing closer again as it inverts completely. Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratis $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Impact of measurement noise on model predictions}
\label{sec:epsilon}

So far a key assumption of the neural network implementation is that the detected scattering signal  has no uncertainty associated with it. In reality of course detected scattering signals  will have some non-zero level of measurement  noise. To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
\begin{align}
	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
\end{align}
where $\epsilon$ is the percentage error associated with the scattering signal.  Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig6a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig6b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig6c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=1.174]$, (b) $15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~  $[F(K_l)=0.457]$.}
	\label{fig:epsilon}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, the inclusion of signal noise quickly leads to a decrease in the model's performance. This is due to an inherent feature of the inverse scattering problem: two distinct regions in orientation space can become heavily intertwined and thus no longer well separated when mapped to intensity space (even though the mapping remains  continuous): so even small uncertainties in the scattering data can lead to large 'mistakes' in the choice of orientation by the neural network. (Indeed if this was not the case the inverse scattering problem would be quite simple.) Fortunately, we can alleviate the impact of signal noise by increasing the weighting factor $\beta$ reflecting the degree of freedom the dimer has within the trap, once again because we know that for a given trap strength, large rapid changes in the dimer orientation are physically limited. We show model results for the same signal noise as Fig.~\ref{fig:epsilon} but with increasing $\beta$ in Figure~\ref{fig:beta}.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig7a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig7b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig7c.png}
	\end{subfigure}
	\caption{Model prediction for signal noise $\epsilon$ of $1\%$, $15\%$, and $25\%$ with (a) $\beta=4$ ~ $[F(K_l)=1.914]$, (b) $\beta=15$ ~  $[F(K_l)=0.734]$, and (c) $\beta=22$ ~ $[F(K_l)=0.621]$.} 
	\label{fig:beta}
\end{figure}

Increasing $\beta$ helps significantly reduce the random fluctuations in the model. Of course, at too high a $\beta$ the orientation estimation will be 'over-corrected' \textit{i.e.} real changes in orientation indicated by real changes in scattering, as opposed to noise, will be suppressed. The exact value of $\beta$ that will best optimise model performance varies based on the expected signal error and the actual trap strength and trapped object dynamics, therefore  it can be used as a variable for 'intelligently' fine tuning results from the neural network and Bayesian inference, using the details of the trap setup and measurement system as essentially 'prior knowledge'.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}

We have developed a method for measuring the dynamics of an optically-trapped 'complex object' based purely on limited measurements of the object's light scattering at a small number of detection angles. We demonstrate the method using the orientation of an asymmetric dimer as the dynamical variable and object of interest, but in principle the model can be applied to any characteristic that impacts the light scattering pattern produced by a trapped entity. The MSTM package is a flexible tool for calculating the light scattering of complex objects using a representation of the object as a set of micro-particles, enabling training of a neural network to enable categorisation of the mapping between scattering and trapped object characteristics. By taking account of knowledge about the physically realistic behaviour of the trapped object and the characteristics of the trap (which impact the dynamics of the object), the Bayesian inference method can be refined to provide a reliable estimation of object characteristics of interest, even in the presence of measurement noise. Fundamentally the inverse scattering problem is difficult to solve since the mapping between object characteristics and scattering can be highly complex: but Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.\\

Here we simplify the problem somewhat by employing a relatively small finite number of 'reference orientations' to map between scattering and dimer orientation: the precision of estimation could be improved by utilising a greater number of reference orientations, although there remains a balance between the realisable precision of orientation estimate and the noise level of the scattering measurement. Another avenue to further explore would be using the method to optimise the choice of detection angles, essentially to find the region in the mapping between measured scattering and orientation that offers the best degree of confidence through optimal separation of scattering signals for distinct orientations. For \textit{sequences} of data such as dynamic measurements, a further potential enhancement would be to consider more complex correlations based on prior expectations of the dynamics. Here already we improve the method using a non-uniform prior based on only the immediately previous measurement in time (see Section \ref{sec:Bayes}): considering more detailed correlations such as multiple previous timesteps is likely to further enhance the reliability of the estimation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{Acknowledgement.} The authors thank the support for this research from the funding provided by the Leverhulme Trust. \\

\noindent \textbf{Disclosures.} The authors declare no conflict of interest. \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{bib} 
\bibliographystyle{ieeetr}

\newpage
\appendix
\onecolumn
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appendix}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\begin{table}[h]
\begin{center}
\caption{\label{tab:A1}
%
Reference Orientations vector components$^*$}
\begin{tabular}{|c|c|c|c|}
\hline\hline
$\alpha$ & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
\hline
$1$  & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$2$  & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$3$  & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$4$  & $ 1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$5$  & $ 0.00000$ &  $ 1.00000$ & $ 0.00000$ \\
$6$  & $ 0.00000$ &  $ 0.00000$ & $ 1.00000$ \\
$7$  & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$8$  & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$9$  & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$10$ & $ 0.00000$ &  $ 0.00000$ & $-1.00000$ \\
$11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$12$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$14$ & $ 0.00000$ &  $ -1.0000$ & $ 0.00000$ \\
$15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
$16$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
$17$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
$18$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$19$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$21$ & $-1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$23$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$24$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$25$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$26$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$27$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$28$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
$28$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
$30$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
\hline\hline
\multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
		\end{tabular}
	\end{center}
\label{sec:App1}
\end{table}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \begin{center}
    \caption{\label{tab:A2}
      % 
      Raw intensities $I_k^*$ and scaled intensities $y_k$}
    \begin{tabular}{|c|c|c|c|c|c|c|}
      \hline\hline
      $\alpha$  & $I(\hat{\bf n}_\alpha\ ,\ 15^{\circ})$ &  $I(\hat{\bf n}_\alpha \ , \ 55^{\circ})$ &  $I(\hat{\bf n}_\alpha \ ,\ 90^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 15^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 55^{\circ})$ & $y(\hat{\bf n}_\alpha \ , \ 90^{\circ})$ \\
\hline
$1$  & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
$2$  & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
$3$  & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
$4$  & $4.5187$ & $0.0459$ & $0.0221$ & $-0.9770$ & $ 0.2062$ & $ 0.9695$ \\
$5$  & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
$6$  & $7.1799$ & $0.0377$ & $0.0142$ & $ 0.5458$ & $-0.0383$ & $-0.1886$ \\
$7$  & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
$8$  & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
$9$  & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
$10$ & $4.6351$ & $0.1536$ & $0.0221$ & $-0.9103$ & $ 3.4040$ & $ 0.9641$ \\
$11$ & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
$12$ & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
$13$ & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
$14$ & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
$15$ & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
$16$ & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
$17$ & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
$18$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
$19$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
$20$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
$21$ & $4.2902$ & $0.0103$ & $0.0142$ & $-1.1077$ & $-0.8510$ & $-0.1890$ \\
$22$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 1.8281$ & $-0.7835$ \\
$23$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
$24$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
$25$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
$26$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
$27$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
$28$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 0.1005$ & $ 0.0102$ \\
$29$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
$30$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
\hline\hline
\multicolumn{7}{l}{\small *$I_k$ values are calculated using MSTM package.}
\end{tabular}
\end{center}
\label{sec:App2}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
\endinput

### 2023-03-29 python files/autocorr_fit.py
Tidying up bib file

@@ -5,7 +5,6 @@ from scipy.optimize import curve_fit
 import scipy
 from scipy.signal import correlate
 import matplotlib.pyplot as plt
-import statsmodels.api as sm
 
 class autocorr_fit(object):
     def __init__(self, path):

### 2023-03-31 Latex files/Manuscript_submission.tex
Update to Manuscript_submission

--- 2023-03-29 Latex files/Manuscript_rev2 mh.tex
@@
 
+\subsection{Brownian Simulation}
+\label{sec:brownian}
+
+We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
+Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+shaped sphere clusters.
+We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating its diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
+For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
+\begin{align}
+	\begin{pmatrix}
+		F_x \\ F_y \\ F_z \\ T_x \\ T_y \\ T_z 
+	\end{pmatrix}
+	= \begin{pmatrix}
+	a_1 &   0  &     0     &   0  & d_1 &     0 \\
+	 0  &  a_1 &     0     & -d_1 &  0  &     0 \\
+	 0  &   0  & a_1 + a_2 &   0  &  0  &     0 \\
+	 0  & -d_1 &     0     &  b_1 &  0  &     0 \\
+	d_1 &   0  &     0     &   0  & b_1 &     0 \\
+	 0  &   0  &     0     &   0  &  0  & b_1 + b_2     
+	\end{pmatrix}
+	\begin{pmatrix}
+		v_{k0} - V_k \\ v_{k0} - V_k \\ v_{k0} - V_k \\
+		\omega_x - \Omega_x \\ \omega_y - \Omega_y \\ \omega_z - \Omega_z
+	\end{pmatrix}
+\end{align}
+
+where each matrix component is scaled by a factor of $k_B T/\pi \eta$ When both sphere's are identically sized - as in Fung~\textit{et.
+al}'s work - $d_1$ goes to zero; however for our work with asymmetric spheres these components are significant enough to be considered.
+We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.
+
 \begin{figure}[h]
@@
 Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$.
-(Note that if all measurement angles were in the same plane we would obtain identical signals for certain sets of different reference orientations, hence $\theta_4$ projects out of the horizontal plane).
+While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:Discussion} for a detailed breakdown of our choice of detection angles.
 The raw intensities are normalized according to:

### 2023-04-20 Latex files/Manuscript_rev3.tex
Updated Manuscript_rev3

--- 2023-03-31 Latex files/Manuscript_submission.tex
@@
 
-\journal{Journal of Quantitative Spectroscopy \& Radiative Transfer}
+\journal{}
 
@@
 
-Since their invention in the late 1980s, optical tweezers have found application 
-in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to
- testing the fundamental assumptions of quantum mechanics  \cite{yin2013large}, thanks 
-mainly to the ability of the tweezer to transduce and detect forces on the order of a 
-few pico newtons. Going beyond forces, further structural, dynamic and chemical 
-characterisation of complex trapped entities could provide useful information, as 
-demonstrated in   areas such as metrology \cite{arita2020coherent} and colloidal
- aggregation \cite{burns1990optical}. Spectroscopic techniques such as Raman scattering 
-\cite{gupta2014raman} have been used for chemical characterisation of trapped objects, 
-while dynamical characterisation has been demonstrated using data from the tweezer's 
-Quadrant Photo Detector by following the centre-of-mass Brownian motion of the trapped 
-entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass 
-\cite{yifat2021facile}.
+Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico-newtons.
+Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
+Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass\cite{yifat2021facile}.
+A recent work aimed at characterising trapped entities demonstrated how neural networks maybe trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}.
+A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle.
+Thus, past studies on optical trapping have focused rather heavily on particle-tracking but not particle-characterisation.
 
-The possibility of integrating an \textit{additional} light scattering detection scheme 
-within an optical trapping setup was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998}, 
-where a single-mode optical fibre was aligned to detect the scattered
-light from a trapped bead and study its Brownian motion.  While this provided dynamical 
-information, structural information about the trapped bead was precluded as the measurement 
-was obtained only at a single angle.  In this work, we propose a scheme that expands on the 
-technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of 
-angles (Figure~\ref{fig:setup}), combined with a novel Bayesian inference-based analysis
- technique to enable interpretation of the resulting multi-angle data as well as 
-optimisation to provide maximal information from the signal.
+While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics.
+A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion.
+While this provided dynamical information, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
+Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise isotropically scattering particles.
+This limits its application to biological matter as the internal structure of many cells makes them inherently isotropic scatterers \cite{Watson_2023}.
+In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation to provide maximal information from the signal.
 
-To demonstrate the analysis, we study a simple, illustrative example of a trapped entity 
-more complex than a single sphere \textit{i.e.} a trapped asymmetric dimer.  
-As a paradigmatic example of extracting structural/dynamic information from the 
-resulting scattering data, we explore how to estimate the dimer's instantaneous 
-orientation from the scattering signals using Bayesian inference, as well as how 
-to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable 
-estimate. As an example of the relatively sparse literature on measuring orientations of
- complex trapped objects, ref.~\cite{raudsepp2022estimating} employs imaging to study the 
-orientation of trapped dimers: scattering can give more quantitative and potentially more
- rapid time-resolved information, but only of course if the scattering signals can be 
-interpreted. Here, we first train a neural network to effectively identify the mapping
- between scattering signals and dimer orientation, by calculating the scattering signal 
-from a  simulated asymmetric dimer undergoing Brownian motion in an optical trap and 
-mapping to the known instantaneous orientation of the simulated dimer. We then show 
-how Bayesian inference can be used to optimise extraction of the true dimer orientation
- from the light scattering signals. Furthermore we demonstrate how the model can be
- fine tuned in situations where measurement uncertainty becomes significant.
+To demonstrate the analysis, we study a simple, illustrative example of an isotropic scattering entity \emph{i.e.} an asymmetric dimer.
+As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate.
+As an example of the relatively sparse literature on measuring orientations of complex trapped objects, ref.
+employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted.
+Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer.
+We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals.
+Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
@@
 Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
-shaped sphere clusters.
-We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating its diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
+shaped sphere clusters. We simulate the motion of a dimer trapped in a 
+highly focused Gaussian beam by calculating the optical forces imparted by 
+the laser, and the Brownian forces imparted by the surrounding fluid.
+MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}.
+The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
 For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Orientation estimation from scattering measurements}
@@
 Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$.
-While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:Discussion} for a detailed breakdown of our choice of detection angles.
+While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown of our choice of detection angles.
 The raw intensities are normalized according to:
@@
 Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex.
+Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle.
 Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Calculation of error}
@@
 but it does not directly indicate our estimate's actual accuracy. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results and  Discussion}
-\label{sec:Discussion}\subsection{Testing the Model}
+\label{sec:Discussion}
+\subsection{Minimal number of detectors}
+\label{sec:detectors}
+The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200.
+For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres around a single point.
+
+Initially we tried capturing scattered light only in the x-y plane as to simulate the detectors all being placed on a glass cover slip.
+The results of this estimate are shown below.
+As can be seen from Figure~ the neural network performs particularly poorly in this situation, seemingly jumping back and forth between different reference orientation at random.
+We found that when all three detectors are in the same plane the expected signal can appear identical despite the dimer being in completely different orientations.
+This is shown in Figure~ which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels shows that the dimer's scattering is indistinguishable in these two reference orientation.
+More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
+This is illustrative of the difficulty behind the inverse light scattering problem; as attributing the measured signal will not reveal any new information about the entity unless we can remove uncertainty around other aspects of the scatterer.
+
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig2b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/intensity_space.png}
+	\end{subfigure}
+	\caption{(a) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point (b) Expected signal for each reference orientation, some signals map to multiple reference orientation and are labelled as such.}
+	\label{fig:angles}
+\end{figure}
+
+To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique.
+With only three detectors the difference in expected signals can appear insignificant; by adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
+In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing the Model}
 \label{sec:test}
@@
 
-One reason we observe such large jumps in  orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space : even a large change in orientation can involve a small change in scattering.
+One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering.
 Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step.
 To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness.
-Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step according to the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$, \textit{i.e.} taking into account the different likelihoods of the estimated orientation jumping between nearby or distant reference orientations:
+Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$.
+Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Asymmetric dimer dynamics}
-\label{sec:motion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Asymmetric dimer dynamics}
+%\label{sec:motion}
 
-The Brownian OT software was used to simulate the motion of a trapped
-dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first
-$10$ seconds of entering the optical trap.  The initial orientation
-was assumed as strictly vertical (in line with the beam propagation
-direction). The dimer's position and orientation was recorded every
-$1\,{\rm ms}$ for using as a test dataset for our model.
+%The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $10$ seconds of entering the optical trap.
+The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
+The dimer's position and orientation was recorded every $1\,{\rm ms}$ for using as a test dataset for our model.
 
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/fig5a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig5b.png}
-	\end{subfigure}
-	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
-	\label{fig:motion}
-\end{figure}
+%\begin{figure}[h]
+%	\centering
+%	\begin{subfigure}{0.45\textwidth}
+	%	\subcaption{}
+	%	\includegraphics[width =\textwidth]{./Images/fig5a.png}
+%	\end{subfigure}
+%	\begin{subfigure}{0.45\textwidth}
+%		\subcaption{}
+%		\includegraphics[width=\textwidth]{./Images/fig5b.png}
+%	\end{subfigure}
+%	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
+%	\label{fig:motion}
+%\end{figure}
 
-As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
+%As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
 Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
 It is interesting to note that the dimer is furthest from the trap centre as it reaches a horizontal orientation before drawing closer again as it inverts completely.
-Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratis $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
+Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
 %
-In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+%In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
 Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Impact of measurement noise on model predictions}
@@
 So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it.
-In reality of course detected scattering signals  will have some non-zero level of measurement  noise.
+In reality of course scattering signals will always have some non-zero measurement noise.
+This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself.
 To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
@@
 This is due to an inherent feature of the inverse scattering problem: two distinct regions in orientation space can become heavily intertwined and thus no longer well separated when mapped to intensity space (even though the mapping remains  continuous): so even small uncertainties in the scattering data can lead to large 'mistakes' in the choice of orientation by the neural network.
-(Indeed if this was not the case the inverse scattering problem would be quite simple.) Fortunately, we can alleviate the impact of signal noise by increasing the weighting factor $\beta$ reflecting the degree of freedom the dimer has within the trap, once again because we know that for a given trap strength, large rapid changes in the dimer orientation are physically limited.
-We show model results for the same signal noise as Fig.~\ref{fig:epsilon} but with increasing $\beta$ in Figure~\ref{fig:beta}.
+(Indeed if this was not the case the inverse scattering problem would be quite simple.)
 
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig7a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig7b.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig7c.png}
-	\end{subfigure}
-	\caption{Model prediction for signal noise $\epsilon$ of $1\%$, $15\%$, and $25\%$ with (a) $\beta=4$ ~ $[F(K_l)=1.914]$, (b) $\beta=15$ ~  $[F(K_l)=0.734]$, and (c) $\beta=22$ ~ $[F(K_l)=0.621]$.} 
-	\label{fig:beta}
-\end{figure}
+To reduce the effects of the signal noise we took the time average of the expected signal over several time steps and then had our neural network estimate the orientation based on the average signal.
+This allowed us to mitigate the effects of any signal error that may persist in light scattering measurements, as shown below.
 
-Increasing $\beta$ helps significantly reduce the random fluctuations in the model.
-Of course, at too high a $\beta$ the orientation estimation will be 'over-corrected' \textit{i.e.} real changes in orientation indicated by real changes in scattering, as opposed to noise, will be suppressed.
-The exact value of $\beta$ that will best optimise model performance varies based on the expected signal error and the actual trap strength and trapped object dynamics, therefore  it can be used as a variable for 'intelligently' fine tuning results from the neural network and Bayesian inference, using the details of the trap setup and measurement system as essentially 'prior knowledge'.
+Diagnostics of optically trapped biological cells often have to contend with large fluctuations in scattering signals due to the fact that their light scattering ability is isotropic.
+Understanding the dynamics of these cells within an optical trap are essential for better characterisation methods.
+Our method demonstrates that we can extract orientational dynamics from a trapped entity via modelling its scattering in MSTM and then applying it to a raw scattering signal.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}

### 2023-06-06 Manuscript_Colloids.tex
Add files via upload

--- 2023-04-20 Latex files/Manuscript_rev3.tex
@@
 
+\usepackage{csvsimple}
 \usepackage{amssymb, amsthm}
@@
 \usepackage{nomencl}
+\usepackage{tabularx}
 \makenomenclature
@@
 
+\begin{flushright}
+	Daniel Maciver \\
+	daniel.maciver.2016@uni.strath.ac.uk \\
+	University of Strathclyde, \\
+	75 Montrose Street, \\
+	Glasgow, \\
+	G1 1XL \\
+	07793445441 
+\end{flushright}
+
 \begin{abstract}
@@
   poses a two-fold challenge --- one experimental, concerning the optimal
-  arrangement of light detectors to gather data, and the other theoretical, involving
+  arrangement of light detectors to gather data and minimising signal noise, and the other theoretical, involving
   solving of the inverse light scattering problem in order to interpret this data. Experimentally, combining static
@@
 
-\begin{highlights}
-\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.
-\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
-\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
-\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
-\end{highlights}
+%\begin{highlights}
+%\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.
+%\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
+%\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
+%\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
+%\end{highlights}
 
@@
 \end{frontmatter}
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle.
-Thus, past studies on optical trapping have focused rather heavily on particle-tracking but not particle-characterisation.
+Thus, past studies on optical trapping have focused on either on particle trapping to study trapping dynamics, or on the characterisation of particles with relatively simple complex trapping dynamics.
 
-While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics.
-A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion.
-While this provided dynamical information, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
-Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise isotropically scattering particles.
-This limits its application to biological matter as the internal structure of many cells makes them inherently isotropic scatterers \cite{Watson_2023}.
-In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation to provide maximal information from the signal.
+While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles with simple trapping dynamics, and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics.
+A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion - commonly now refereed to as Localised dynamic light scattering (LDLS).
+This was later expanded upon by \cite{PhysRevE.65.041921} by collecting back scattered light to characterise the Stokes friction coefficient as a function of trapping depth.
+While both papers provided dynamical information, structural information about the trapped bead was precluded as the scattered wave was only measured across a small angular range.
+Furthermore, the main drawback to a LDLS is that it struggles to characterise asymmetric particles such as nano rods or dimers; in which case the Brownian motion is no longer simply translational, but rotational about the centre of mass \cite{yin2013large}.
 
-To demonstrate the analysis, we study a simple, illustrative example of an isotropic scattering entity \emph{i.e.} an asymmetric dimer.
+One way of mitigating this is to remove translational fluctuations from the analysis; by monitoring its the scattering, Cang and co-workers were able to spatially fix a gold nanorod at the centre of trapping laser by moving the sample plane accordingly, to an accuracy of 200 nm \cite{Cang2006}.
+However, this has limited application's to anisotropic scatters such as biological matter where the internal structure of many cells makes them inherently anisotropic \cite{Watson_2023}.
+Thus even if the Cartesian coordinates are relatively fixed characterisation of the scattering pattern will have to separate the contributions of internal structure and orientation.
+In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique, to enable interpretation of the resulting multi-angle data from an anisotropic, asymmetric scatterer, as well as optimisation to provide maximal information from the signal.
+
+To demonstrate the analysis, we study a simple, illustrative example of an anisotropic scattering entity \emph{i.e.} an asymmetric dimer.
 As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate.
-As an example of the relatively sparse literature on measuring orientations of complex trapped objects, ref.
-employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted.
+As an example of the relatively sparse literature on measuring orientations of complex trapped objects, \cite{raudsepp2022estimating} employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted.
 Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer.
 We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals.
-Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.
+Furthermore we demonstrate how the model's performance when dealing with signal noise, a common problem when analysing scattering behaviour.
 
-\begin{figure}
+\begin{figure} [b]
 \centering
@@
 b) Top view.
-Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see text).
+Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see Sec~\ref{sec:detectors}).
 %
@@
 
-\printnomenclature[0.55in]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
 \label{sec:Method}
+\subsection{Orientation estimation from scattering measurements}
+\label{sec:Bayes}
 
-\subsection{Brownian Simulation}
-\label{sec:brownian}
-
-We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
-Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
-shaped sphere clusters. We simulate the motion of a dimer trapped in a 
-highly focused Gaussian beam by calculating the optical forces imparted by 
-the laser, and the Brownian forces imparted by the surrounding fluid.
-MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}.
-The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
-For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
-\begin{align}
-	\begin{pmatrix}
-		F_x \\ F_y \\ F_z \\ T_x \\ T_y \\ T_z 
-	\end{pmatrix}
-	= \begin{pmatrix}
-	a_1 &   0  &     0     &   0  & d_1 &     0 \\
-	 0  &  a_1 &     0     & -d_1 &  0  &     0 \\
-	 0  &   0  & a_1 + a_2 &   0  &  0  &     0 \\
-	 0  & -d_1 &     0     &  b_1 &  0  &     0 \\
-	d_1 &   0  &     0     &   0  & b_1 &     0 \\
-	 0  &   0  &     0     &   0  &  0  & b_1 + b_2     
-	\end{pmatrix}
-	\begin{pmatrix}
-		v_{k0} - V_k \\ v_{k0} - V_k \\ v_{k0} - V_k \\
-		\omega_x - \Omega_x \\ \omega_y - \Omega_y \\ \omega_z - \Omega_z
-	\end{pmatrix}
-\end{align}
-
-where each matrix component is scaled by a factor of $k_B T/\pi \eta$ When both sphere's are identically sized - as in Fung~\textit{et.
-al}'s work - $d_1$ goes to zero; however for our work with asymmetric spheres these components are significant enough to be considered.
-We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.
+Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere to the centre of the smaller sphere.
+A plane wave 'probe' laser, perpendicular to the trapping laser, is incident on the dimer, generating a scattering pattern  dependent on the dimer's orientation $I(\hat{\bf s}, \theta)$ which can be computed using  software such as MSTM \cite{Mishchenko1996MSTM}.
+To represent the experimental set up consisting of a set of optical fibres recording scattered light, we choose four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) and record the calculated intensity at each angle $\theta_k$, $I(\hat{s}, \theta_k)$.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Orientation estimation from scattering measurements}
-\label{sec:Bayes}
-
-For our simulated dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere to the centre of the smaller sphere.
-A plane wave 'probe' laser, perpendicular to the trapping laser, is incident on the dimer, generating a scattering pattern  dependent on the dimer's orientation $I(\hat{\bf s}, \theta)$ which is computed using MSTM.
-To represent the experimental set up consisting of a set of optical fibres recording scattered light, we choose four angles  ($\theta_1, \ \theta_2, \ \theta_3, \ \theta_4$) and record the calculated intensity at each angle $\theta_k$, $I(\hat{s}, \theta_k)$.
-
-Our goal is to determine the orientation of the trapped dimer based on the scattering data $I(\hat{n}, \ \theta_k)$.
+Our goal is to determine the orientation of the trapped dimer based on the measured intensity $I(\hat{n}, \ \theta_k)$.
 Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).
@@
 
-Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex.
+Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem.
+While it is trivial to compute the light scattering pattern for any given particle with any particular characteristic (i.e.
+size, shape, or orientation), inferring the light scattering from a unknown particle to determine said characteristic is incredibly difficult due to complex mapping between scattering and said characteristic.
 Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle.
+Shown below is two clusters of orientation vectors and there respective measured scattering signals - the points have been coloured based on their proximity to the centre of their respective cluster.
+While the orientation space appears tightly packed and ordered the signal space quickly spreads out in an asymmetric fashion.
+Furthermore as seen in Fig~\ref{fig:mixing}b the signal mapping can intersect itself which only further increases the complexity.
+While in some instances the mapping between one reference orientation and another is discrete, in other instances the mapping becomes far more complex to discern.
+
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig3a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig3b.png}
+	\end{subfigure}
+	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals.
+Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
+	\label{fig:mixing}
+\end{figure}
+
 Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Calculation of error}
@@
 where the scattering measurement could imply a wide range of possible orientations --- 
-but it does not directly indicate our estimate's actual accuracy. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Brownian Simulation}
+\label{sec:brownian}
+
+We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
+Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+shaped sphere clusters.
+We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid.
+MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}.
+The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
+We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-3} \ s$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results and  Discussion}
@@
 The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200.
-For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres around a single point.
-
-Initially we tried capturing scattered light only in the x-y plane as to simulate the detectors all being placed on a glass cover slip.
-The results of this estimate are shown below.
-As can be seen from Figure~ the neural network performs particularly poorly in this situation, seemingly jumping back and forth between different reference orientation at random.
-We found that when all three detectors are in the same plane the expected signal can appear identical despite the dimer being in completely different orientations.
-This is shown in Figure~ which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels shows that the dimer's scattering is indistinguishable in these two reference orientation.
-More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
-This is illustrative of the difficulty behind the inverse light scattering problem; as attributing the measured signal will not reveal any new information about the entity unless we can remove uncertainty around other aspects of the scatterer.
+For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.
 
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig2b.png}
+		\includegraphics[width=\textwidth]{./Images/fig4a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/intensity_space.png}
+		\includegraphics[width=\textwidth]{./Images/fig4b.png}
 	\end{subfigure}
-	\caption{(a) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point (b) Expected signal for each reference orientation, some signals map to multiple reference orientation and are labelled as such.}
-	\label{fig:angles}
+	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
+	\label{fig:detectors}
 \end{figure}
 
+ When all of the detectors lie in the same plane the expected signal can appear identical despite the dimer being in completely different orientations.
+This is shown in Figure~\ref{fig:detectors} which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels is because the dimer's scattering is indistinguishable in these two reference orientation.
+It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors.
+More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
+This is illustrative of the difficulty behind the inverse light scattering problem; as one cannot always map a given signal to a particular parameter value.
+
 To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique.
-With only three detectors the difference in expected signals can appear insignificant; by adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
+As seen between Figures \ref{fig:detectors}a \& b each reference orientation now has a unique scattering signal, though with only three detectors the difference in expected signals can appear insignificant.
+By adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
 In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Testing the Model}
 \label{sec:test}
-Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
+Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.45$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
 The trapping laser is 1064nm NIR focused through a 1.25 NA objective.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{./Images/fig3.png}
-	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$.
+	\includegraphics[width=0.5\textwidth]{./Images/fig5.png}
+	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity.
 Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
-Horizontal dotted line denotes maximum spacing between two neighbouring $\hat{\bf n}_\alpha$.}
+Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
 	\label{fig:uniform}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{./Images/fig4.png}
+\includegraphics[width=0.5\textwidth]{./Images/fig8a.png}
 \caption{\label{fig:biased}
@@
 achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-Dotted line denotes spacing between two neighbouring reference
-orientations $\hat{\bf n}_\alpha$ (see Section~\ref{sec:Bayes}).
+Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
   %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Asymmetric dimer dynamics}
-%\label{sec:motion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Asymmetric dimer dynamics}
+\label{sec:motion}
 
-%The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $10$ seconds of entering the optical trap.
+The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $10$ seconds of entering the optical trap.
 The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
@@
 
-%\begin{figure}[h]
-%	\centering
-%	\begin{subfigure}{0.45\textwidth}
-	%	\subcaption{}
-	%	\includegraphics[width =\textwidth]{./Images/fig5a.png}
-%	\end{subfigure}
-%	\begin{subfigure}{0.45\textwidth}
-%		\subcaption{}
-%		\includegraphics[width=\textwidth]{./Images/fig5b.png}
-%	\end{subfigure}
-%	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
-%	\label{fig:motion}
-%\end{figure}
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{./Images/fig7a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig7b.png}
+	\end{subfigure}
+	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
+\end{figure}
 
-%As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.
-Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.
-It is interesting to note that the dimer is furthest from the trap centre as it reaches a horizontal orientation before drawing closer again as it inverts completely.
-Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
-%
-%In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
-Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Impact of measurement noise on model predictions}
+In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state.
+We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer.
+In the future we hope to further investigate the motion of asymmetric dimers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Accounting for sources of error in light scattering measurements}
+When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor.
+Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle.
+However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance.
+We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors.
+We highlight how we address these areas below.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subparagraph{Impact of incorrect dimer sizing}
+\label{sec:lam}
+
+One of the main limitations of our model is that we assume that the dimer being modelled in MSTM is accurate to the dimer being trapped in the optical tweezer.
+Sizing molecules accurately is a significant challenge for single particle analysis so there is bound to be some uncertainty with the measurements.
+We ran our model 3 times with the neural net being trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
+
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.33\textwidth}
+			\subcaption{}
+			\includegraphics[width =\textwidth]{./Images/fig8a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig8b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig8c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
+	\label{fig:size}
+\end{figure}
+As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model.
+This amounts to just over $100 nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
+This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
+We addressed this by increasing the number of available reference points from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
+While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{./Images/fig9a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig9b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig9c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
+	\label{fig:refs}
+\end{figure}
+
+Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer.
+This suggests that overshooting our size estimate will be less detrimental to our estimation.
+Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations.
+This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction.
+Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subparagraph{Impact of measurement noise on model predictions}
 \label{sec:epsilon}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig6a.png}
+		\includegraphics[width=\textwidth]{./Images/fig10a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig6b.png}
+		\includegraphics[width=\textwidth]{./Images/fig10b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig6c.png}
+		\includegraphics[width=\textwidth]{./Images/fig10c.png}
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=1.174]$, (b) $15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~  $[F(K_l)=0.457]$.}
+	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
 	\label{fig:epsilon}
@@
 
-To reduce the effects of the signal noise we took the time average of the expected signal over several time steps and then had our neural network estimate the orientation based on the average signal.
-This allowed us to mitigate the effects of any signal error that may persist in light scattering measurements, as shown below.
+To reduce the effects of the signal noise we took the time average of the expected signal over 0.02s and then had our neural network estimate the orientation based on the average signal.
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig11a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig11b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{./Images/fig11c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $20\,{\rm ms}$}
+	\label{fig:time average}
+\end{figure}
+This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model.
+There appears to be no clear correlation between the length over which we time average and the performance of our model.
+Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network.
+Fortunately, time averaging even over $20\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
 
-Diagnostics of optically trapped biological cells often have to contend with large fluctuations in scattering signals due to the fact that their light scattering ability is isotropic.
-Understanding the dynamics of these cells within an optical trap are essential for better characterisation methods.
-Our method demonstrates that we can extract orientational dynamics from a trapped entity via modelling its scattering in MSTM and then applying it to a raw scattering signal.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
 \onecolumn
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Appendix}
@@
 		%
-Reference Orientations vector components$^*$}
-\begin{tabular}{|c|c|c|c|}
+		Reference Orientations vector components, for $n_{refs} = 30 \ ^*$}
+		\begin{tabularx}{0.65\textwidth}{
+				| >{\centering\arraybackslash}X
+				| >{\centering\arraybackslash}X
+				| >{\centering\arraybackslash}X
+				| >{\centering\arraybackslash}X | }
 			\hline\hline
-$\alpha$ & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
-\hline
-$1$  & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
-$2$  & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
-$3$  & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-$4$  & $ 1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
-$5$  & $ 0.00000$ &  $ 1.00000$ & $ 0.00000$ \\
-$6$  & $ 0.00000$ &  $ 0.00000$ & $ 1.00000$ \\
-$7$  & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
-$8$  & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
-$9$  & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
-$10$ & $ 0.00000$ &  $ 0.00000$ & $-1.00000$ \\
-$11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
-$12$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
-$13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
-$14$ & $ 0.00000$ &  $ -1.0000$ & $ 0.00000$ \\
-$15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
-$16$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
-$17$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
-$18$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
-$19$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
-$20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
-$21$ & $-1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
-$22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
-$23$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
-$24$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
-$25$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
-$26$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
-$27$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
-$28$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
-$28$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
-$30$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
+			$\alpha$ & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ 
+			\\ \hline
+			\csvreader[
+			head to column names
+			]
+			{./data/n_list.csv}{}{
+			\index\ & \nx\ & \ny\ & \nz\ \\
+			} \\
 			\hline\hline 
 		\multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
-		\end{tabular}
+		\end{tabularx}
 \end{center}
-\label{sec:App1}
 \end{table}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}[h]
@@
 		    \hline
-$1$  & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
-$2$  & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
-$3$  & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
-$4$  & $4.5187$ & $0.0459$ & $0.0221$ & $-0.9770$ & $ 0.2062$ & $ 0.9695$ \\
-$5$  & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
-$6$  & $7.1799$ & $0.0377$ & $0.0142$ & $ 0.5458$ & $-0.0383$ & $-0.1886$ \\
-$7$  & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
-$8$  & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
-$9$  & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
-$10$ & $4.6351$ & $0.1536$ & $0.0221$ & $-0.9103$ & $ 3.4040$ & $ 0.9641$ \\
-$11$ & $5.6777$ & $0.0180$ & $0.0122$ & $-0.3137$ & $-0.6242$ & $-0.4838$ \\
-$12$ & $5.2364$ & $0.0088$ & $0.0102$ & $-0.5663$ & $-0.8951$ & $-0.7826$ \\
-$13$ & $9.0297$ & $0.0141$ & $0.0234$ & $ 1.6044$ & $-0.7378$ & $ 1.1622$ \\
-$14$ & $8.5891$ & $0.0392$ & $0.0244$ & $ 1.3523$ & $ 0.0060$ & $ 1.3031$ \\
-$15$ & $5.0015$ & $0.0071$ & $0.0095$ & $-0.7007$ & $-0.9468$ & $-0.8792$ \\
-$16$ & $4.8573$ & $0.0578$ & $0.0095$ & $-0.7832$ & $ 0.5604$ & $-0.8794$ \\
-$17$ & $9.0184$ & $0.0618$ & $0.0273$ & $ 1.5979$ & $ 0.6774$ & $ 1.7262$ \\
-$18$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
-$19$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
-$20$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
-$21$ & $4.2902$ & $0.0103$ & $0.0142$ & $-1.1077$ & $-0.8510$ & $-0.1890$ \\
-$22$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 1.8281$ & $-0.7835$ \\
-$23$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
-$24$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
-$25$ & $7.1549$ & $0.0408$ & $0.0076$ & $ 0.5315$ & $ 0.0532$ & $-1.1580$ \\
-$26$ & $4.0546$ & $0.0085$ & $0.0076$ & $-1.2425$ & $-0.9035$ & $-1.1576$ \\
-$27$ & $5.9164$ & $0.0122$ & $0.0188$ & $-0.1771$ & $-0.7945$ & $ 0.4765$ \\
-$28$ & $7.9307$ & $0.1005$ & $0.0102$ & $ 0.9754$ & $ 0.1005$ & $ 0.0102$ \\
-$29$ & $4.0693$ & $0.0414$ & $0.0122$ & $-1.2342$ & $ 0.0718$ & $-0.4840$ \\
-$30$ & $6.5422$ & $0.0506$ & $0.0234$ & $ 0.1809$ & $ 0.3446$ & $ 1.1622$ \\
+			\csvreader[
+			head to column names
+			]
+			{./data/y_list.csv}{}{
+				\index\ & \Ix\ & \Iy\ & \Iz\ & \yx\ & \yy\ & \yz\ \\
+			} \\
 			\hline\hline 
-\multicolumn{7}{l}{\small *$I_k$ values are calculated using MSTM package.}
+			\multicolumn{7}{l}{\small *$I_k$ values are calculated using MSTM package.} \\
 		\end{tabular}
 	\end{center}
-\label{sec:App2}
 \end{table}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2023-06-06 Latex files/Manuscript_Colloids.tex
cleaning up

--- 2023-06-06 Manuscript_Colloids.tex
@@
 
-\author[aff1]{Dan Maciver\corref{cor1}} 
+\author[aff1]{Daniel Maciver\corref{cor1}} 
 \ead{Daniel.Maciver.2016@uni.strath.ac.uk}
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
-We simulated a silica dimer ($n = 1.45$) in a suspension of water ($n_{med} = 1.33$) over the course of $10 \ s$ with a simulation time step of $1 \times 10^{-3} \ s$.
+We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.
+We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}.
+We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.
 
@@
 
-The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $10$ seconds of entering the optical trap.
-The initial orientation was assumed as strictly vertical (in line with the beam propagation direction).
-The dimer's position and orientation was recorded every $1\,{\rm ms}$ for using as a test dataset for our model.
+The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $1$ seconds of entering the optical trap.
+The initial orientation was set at $s = (0.923, 0.0, 0.385)$.
+The dimer's position and orientation was recorded every $10\,{\rm \mu s}$ for using as a test dataset for our model.
 

### 2023-06-06 Latex files/archive/Investigation into low light scattering.tex
cleaning up

--- 2022-11-04 Investigation into low light scattering.tex
@@
 	\section*{Introduction}
-	Optical trapping has been a vital tool in the study of single molecule properties, the ability to deliver a stable and precise force vector has proved invaluable for studying biological groups in particular \cite{Biophysics_trapping}.
-Recent years there has been a growing interest in the study of colloid aggregation and growth; where the optical binding effect resulted in aggregation of particles around the trapping focus \cite{Colloids_and_tweezers}.
-Raman scattering techniques can be used to determine particle characteristics in a bulk or individual instance and have been applied to optical trapping methods to determine compositional information of the trapped entity \cite{Raman_Trapping,Raman_trapping_Biophysics}.
-However, the particle's dynamics in the trap can not be elucidated based purely on spectroscopy.
-For simple near spherical particles the motion can be largely attributed to Brownian motion which can be readily measured using a quadrant photo detector.
-It is not unusual however, that two microspheres to become attached to one another; in the presence of a focused Gaussian beam the microsphere dimer will reorient itself to align along the propagation axis of the beam.
-If the dimer is symmetrical then this reorientation motion is instantaneous, in the case of an asymmetrical dimer the alignment rapidly fluctuates about the propagation axis before eventually achieving an equilibrated orientation.
-Developing a method to model a dimer's motion within an optical trap based on information gathered in situ will prove useful for understanding the dynamics of more complexly shaped particle's that cannot be directly imaged.
+	Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics~\cite{Wang 2021} to those that test the fundamental assumptions of Quantum Mechanics ~\cite{Duan 2013} thanks mainly to the ability of the tweezer to transduce and detect forces in the order of a few piconewton.
+Recent applications of the Optical Tweezers in metrological measurements ~\cite{Scott Waitukaitis} and colloidal aggregation ~\cite{Optical Binding} demand the tweezer to also be able to characterise  the trapped entity.
+To this end, chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering ~\cite{P K Gupta} and dynamical characterisation has been achieved mainly by following the centre-of-mass brownian motion of the trapped entity using a Quadrant Photo Detector~\cite{Alexander Rohrbach}.
+While this technique has allowed for detection of linear motion of trapped entities  with sub-nanometre precision~\cite{Alexander Rohrbach} and  rotation~\cite{recent paper} of the centre-of-mass, there is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in ~\cite{New Zealand paper} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved.
 
-	This has already been achieved with single microsphere using dynamic light scattering measurements of the trapping beam; where Bar-Ziv et al utilised dynamic light scattering to characterise the trap's stiffness and anisotropy based on the Brownian motion of a trapped silica microsphere \cite{Dynamic_trap_scattering}.
-They note that their method assumes that the silica scatters isotropically which allows them to place a single detection fibre at $75^\circ$ from the propagation axis.
-In the case where the trapped object scatters anisotropically - such as for an asymmetrical dimer - it is necessary to utilise multiple detection fibres to build a better picture of the trapped object's orientation based on the produced light scattering.
-However, for inverted microscope optical traps there is a significant engineering challenge of placing enough detection fibres at the correct angles within a close enough distance to the trapping focus to collect a clear signal.
-This means that any fibre based detection will have to rely on maximising the information provided by the light scattering signals collected at a limited number of detection fibres.
-Excluding the engineering challenge involved with ensuring proper coupling between the probe laser and the detection fibres, the primary concern is how much useful information can be extracted from the trapped particle.
-
-	The experimental set up that this is based on is a inverted microscope optical trap, a suspension of micro particles are loaded onto a plastic cover slide.
-Onto the slide four optical fibres are stuck down; one fibre walks a 633 nm He-Ne laser onto the cover slip pointed directly at the trap focus, and the 3 remaining fibres are spaced equidistantly from the trap focus.
-The distance between each fibre and the trap focus is roughly $200\mu m$, with the exact angles between the particle and the trap focus being measured via the microscopes camera.
-The trapping laser is a 1064 nm NIR laser that is introduced via a 1.25 NA objective; to increase the trapping depth of the objective a lens of focal length 100 cm was added into the beam path to transform the profile entering the objective from parallel to converging, as advised by \cite{Long_distance_trap}'s paper.
+Possibility of integrating light scattering with optical trapping was demonstrated by Saffran et.al.,~\cite{Saffran} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and it's brownian motion was studied.
+While this allowed for studying the dynamics, any information with regards structure of the trapped bead was precluded as the measurement was a single-angle scattering measurement.
+In this work, we propose a novel Bayesian Inference based technique to interpret light scattering data detected in a scheme that expands on the technique in ~\cite{Saffran} to detect scattered light simultaneously at 3 angles as shown in Figure 1.
 
@@
 
+	The issue with having a high signal error is that our model's prediction can be very far off from the correct answer and still be confident with it's results; fortunately our priori for the dimer's orientation can help to correct our estimate by limiting the probability in the physical space.
+We found that the inclusion of our priori estimate compared to a uniform priori imporved our results by around $20 \%$.
+We
 	\subsection*{Ultra-nest results}
@@
 	We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern.
-The model can successfully predict the dimer's motion using its prior estimation to restrain the estimate in the physical space.
+Compared to a random selection of the particle's orientation we find that our model has a $150\%$ increase in it's accuracy.
+Even more interestingly, our model performs better under noisy signal conditions by allowing our estimate to self-correct itself, in the future we hope to remove the artifacts of the signal noise by combing multiple estimates using different detection angles.
+This can be achieved via the inclusion of additional signal detections, we refrained from collecting more signal point for two key reasons; firstly, we wanted to show the extent of the mathmatical model in the case of limited detection schemes, and secondly, we wanted to limit our model to what is possible in a real world experimental set up.
+Once we have a clearer understanding of the maximum limit of detection fibres that can be placed we can update our model to account for this and imporve the accuracy futher.
 
-	Analysis of the parameter space for our choice of detection angles shows a strong preference for angles far from the $0^{\circ}$ suggesting a more accurate estimation can be achieved by considering back scattering signals.
-We hope to validate our model by collecting light scattering signals from a trapped dimer and testing it against the video and QPD signals generated from our result.
+	Our simulations of asymetric dimers found that for large sphere ratios the dimer will prefer to be orientated against the beam propogation and will even move further from the trapping focus to invert itself before coming back to the laser focus.
+It is interesting that is only the case for dimers where the size ratio is greater than 1:2, understanding how the optical torque varies with size ratio can provide unique insight into the dynamics of non-spherical trapping targets.
+
+	Overall the model developed can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity.
+The MSTM package is increadibly flexible and allows us to calculate the light scattering for a variety of micro partilces, so long as we can calculate the light scattering we can apply our model to help characterise the trapped model.
+We hope to extened this work for characterising continous processes such as nucleation within an optical trap.
 

### 2023-06-06 Latex files/archive/Manuscript_rev2-LL.tex
cleaning up

--- 2023-03-29 Latex files/Manuscript_rev2 mh.tex
@@
 One reason we observe such large jumps in  orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space : even a large change in orientation can involve a small change in scattering.
-Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step.
+Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from timestep to timestep.
 To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior,  accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness.
-Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step according to the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$, \textit{i.e.} taking into account the different likelihoods of the estimated orientation jumping between nearby or distant reference orientations:
+Here the prior of the reference orientations $p(\hat{\bf n}_\alpha$ was redefined at each time step according to the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$, \textit{i.e.} taking into account the different likelihoods of the estimated orientation jumping between nearby or distant reference orientations:
 

### 2023-06-06 Latex files/archive/Manuscript_rev3.tex
cleaning up

--- 2023-04-20 Latex files/Manuscript_rev3.tex
@@
 \usepackage{nomencl}
+\usepackage{array}
+\usepackage{tabularx}
 \makenomenclature
@@
 While this provided dynamical information, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle.
-Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise isotropically scattering particles.
-This limits its application to biological matter as the internal structure of many cells makes them inherently isotropic scatterers \cite{Watson_2023}.
+Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise anisotropically scattering particles.
+This limits its application to biological matter as the internal structure of many cells makes them inherently anisotropic scatterers \cite{Watson_2023}.
 In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation to provide maximal information from the signal.
@@
 Reference Orientations vector components$^*$}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabularx}{0.65\textwidth}{
+		| >{\centering\arraybackslash}X 
+		| >{\centering\arraybackslash}X 
+		| >{\centering\arraybackslash}X 
+		| >{\centering\arraybackslash}X| }
 \hline\hline
@@
 \multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
-		\end{tabular}
+
+\end{tabularx}
 \end{center}

### 2023-08-22 python files/autocorr_fit.py
22/08/2023 update

@@ -9,7 +9,6 @@ import matplotlib.pyplot as plt
 class autocorr_fit(object):
     def __init__(self, path):
         self.path = path
-        self.label = self.path[self.path.rfind("/")+1:]
         #self.material, self.power, self.size, self.radius = self.get_label_path(self.label)
 
     def row_reader(self, row):

### 2023-08-22 python files/time_domain_fit.py

@@ -1,0 +1,135 @@
+import argparse
+import csv
+import numpy as np
+from scipy.optimize import curve_fit
+import scipy
+from scipy.signal import correlate
+import matplotlib.pyplot as plt
+from datetime import datetime
+
+class time_domain:
+    ''' 
+    Class for processing time domain data from QPD results,
+    main two ouputs from this class are the equipartition and 
+    auto-correlation fittings. The former providing a rough 
+    estimate of the trap stiffness, the latter an estimation
+    of the trapping medium's viscosity.
+
+    In order for either method to work we require a conversion
+    factor to convert our data from Voltz to m, this conversion 
+    factor is computed in the freq_domain.py file. By fitting
+    the PSD we can make a comparison between our experimental
+    value of D (diffusion constant) and the value of D given
+    by the Einstein relation. Both this file and the freq_domain.py
+    file should be run in tandem to get accurate results. 
+    
+    '''
+    def __init__(self, beta):
+        # Only necessary constant is the conversion factor
+        self.beta = beta
+        self.kT = 1.38e-23*290
+        current_time = datetime()
+        self.tag = current_time.now()
+    
+    def create_csv_file(self, path):
+        # 
+        with open(path, 'r') as input:
+            reader = csv.reader(input.readlines()[5:])
+
+            with open('tmp_tdat.csv', 'w') as output: 
+                writer = csv.writer(output)
+                writer.writerows(self.row_reader(row) for row in reader)
+
+        return self.get_data('tmp_tdat.csv')
+    
+    def get_data(self, file):
+        with open(file, 'r') as data:
+            reader = csv.reader(data.readlines())
+            time, x_QPD, y_QPD = [], [], []
+            for row in reader:
+                time.append(row[0])
+                x_QPD.append(row[1]/np.sqrt(self.beta))
+                y_QPD.append(row[2]/np.sqrt(self.beta))
+
+        return time, x_QPD, y_QPD
+
+    def row_reader(self, row):
+        line = list(row[0].split())
+        new_row = [float(line[0]), float(line[1]), float(line[2]), float(line[3])]
+
+        return new_row
+
+    def equipartition_fit(self, data):
+        data_eq = np.mean(data)
[…]

### 2023-08-22 Manuscript/Manuscript_Colloids.tex
linux pc push

--- 2023-06-06 Latex files/Manuscript_Colloids.tex
@@
 
+\begin{center}
+		{\bf{Fig 1a, 1b}}
+\end{center}
 \begin{figure} [b]
@@
 
+\begin{center}
+	\bf{Fig 2a, 2b}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 3a, 3b}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 4a, 4b}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 5}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 6}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 7a, 7b}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 8a, 8b, 8c}
+\end{center}
 \begin{figure}[h]
@@
 As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model.
-This amounts to just over $100 nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
+This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
 This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
-We addressed this by increasing the number of available reference points from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
+We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
 While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
+
+\begin{center}
+	\bf{Fig 9a, 9b, 9c}
+\end{center}
 \begin{figure}[h]
@@
 
+\begin{center}
+	\bf{Fig 10a, 10b, 10c}
+\end{center}
 \begin{figure}[h]
@@
 
-To reduce the effects of the signal noise we took the time average of the expected signal over 0.02s and then had our neural network estimate the orientation based on the average signal.
+To reduce the effects of the signal noise we took the time average of the expected signal over $0.001 s$ and then had our neural network estimate the orientation based on the average signal.
+
+\begin{center}
+	\bf{Fig 11a, 11b, 11c}
+\end{center}
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $20\,{\rm ms}$}
+	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $1\,{\rm ms}$}
 	\label{fig:time average}
@@
 Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network.
-Fortunately, time averaging even over $20\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
+Fortunately, time averaging even over $1\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
 

### 2023-10-25 Latex files/Elisevier/Cover_letter CSA.tex
Update to Latex files sub folder

\documentclass[]{letter}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000
\usepackage[absolute,overlay]{textpos}

\begin{document}

\begin{textblock*}{13cm}(4cm,2.54cm) % {block width} (coords) 
	\begin{letter}{Dear Editors,}
	\opening{}
	We are pleased to submit a manuscript titles \textbf{‘Investigation of Multi-Angle Detection Schemes for Light Scattering of Optically Trapped Asymmetric Colloidal Particles’} P Parthasarathi, M Haw, L Lue, J Sefcik,
	and myself for publication as an article in Optical Communications.

	This work introduces a new method to quantitatively characterize
	optically trapped colloidal particles, by measuring light scattered at
	multiple angles and analysing the data based on Bayesian inference.
	To demonstrate the method, we apply it estimate the instantaneous
	orientations of an optically trapped asymmetric colloidal dimer.  This
	technique can be easily extended to characterize the size and shape of
	optically trapped colloidal particles and offers a powerful tool to
	study colloidal systems. The ability to interpret light scattering signals from an optically trapped entity should be of interest to readers
	of your journal.

	We hope that you will consider our work for publication.  If you need
	any further information or materials, please let me know.  Thank you
	for your attention.

	Sincerely yours,

	Daniel Maciver \\

	Department of Chemical and Process Engineering \\
	University of Strathclyde \\
	James Weir Building, 75 Montrose Street \\
	Glasgow, G1 1XJ \\
	United Kingdom \\

	phone: +44-779-344-5441 \\
	fax:   +44-779-344-5441 \\
	email: daniel.maciver.2016@uni.strath.ac.uk
	\end{letter}
\end{textblock*}

%enclosure listing
%\encl{}

\end{document}

### 2023-10-25 Latex files/Manuscripts Scattering/Manuscript_Optical Communications.tex
Update to Latex files sub folder

--- 2023-08-22 Manuscript/Manuscript_Colloids.tex
@@
-\documentclass[final,  3p]{elsarticle}
+\documentclass[preprint,  3p]{elsarticle}
 
@@
 
-\title{Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles}
+\title{Investigation of Multi-Angle Detection Schemes for Light Scattering of Optically Trapped Asymmetric Colloidal Particles}
 
@@
             country={Scotland}}
-
-\begin{flushright}
-	Daniel Maciver \\
-	daniel.maciver.2016@uni.strath.ac.uk \\
-	University of Strathclyde, \\
-	75 Montrose Street, \\
-	Glasgow, \\
-	G1 1XL \\
-	07793445441 
-\end{flushright}
-
 \begin{abstract}
@@
   %
-  While optical trapping is a well understood method for force
-  transduction and detection, further characterisation of trapped entities using light-scattering
-  poses a two-fold challenge --- one experimental, concerning the optimal
-  arrangement of light detectors to gather data and minimising signal noise, and the other theoretical, involving
-  solving of the inverse light scattering problem in order to interpret this data. Experimentally, combining static
-  light scattering techniques with optical trapping poses significant
-  engineering challenges due to the space constraints in a
-  conventional optical trapping setup.  We propose here a plausible
-  scenario of detecting scattered light from an optically trapped
-  asymmetric microstructure using a novel, multi-angle, optical-fibre
-  based detection scheme and demonstrate how a Bayesian inference
-  based analysis of the data, combined with a neural-network trained on data simulated to mimic light scattering
-  detection signals in such scenarios, may be used for solving the
-  inverse light scattering problem and characterising complex trapped
-  entities.  To demonstrate the method we discuss its application to
+  Optical trapping is a well understood method for transduction and detection
+  of forces on trapped colloidal particles, these trapped entities can be
+  further characterised using light-scattering, posing a two-fold challenge:
+  one experimental, concerning the optimal arrangement of detectors to gather
+  data and minimise signal noise, and the other theoretical, involving solving
+  of the inverse scattering problem in order to interpret light scattering
+  data to determine size, shape, or orientation of the trapped object.
+  Experimentally, combining static light scattering techniques with optical
+  trapping poses significant engineering challenges due to the space
+  constraints in a conventional optical trapping setup. We investigated a
+  plausible scenario of detecting scattered light from an optically trapped
+  asymmetric particles using a novel, multi-angle, optical-fibre based
+  detection scheme, we show how a Bayesian inference-based analysis of the
+  data, combined with a neural-network trained on data simulated to mimic
+  light scattering detection signals in such scenarios, can be used for
+  solving the inverse light scattering problem and characterising colloidal
+  trapped entities. To demonstrate the method, we discuss its application to
   measuring the instantaneous orientations of a trapped asymmetric microsphere
-  dimer. We argue that the method can be extended to
-  determine any characteristics of the trapped microstructure that
-  influence the light scattering pattern.
+  dimer and determine the minimum number of detectors required for a reliable
+  estimation in the presence of signal noise. This approach can be extended to
+  determine any characteristics of the trapped microstructure that influence
+  the light scattering pattern, including size and shape of colloidal objects. 
 \end{abstract}
+
+%\begin{graphicalabstract}
+%\begin{figure} [h]
+%	\centering
+%	\begin{subfigure}{0.45\textwidth}
+%		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
+%	\end{subfigure}
+%	\begin{subfigure}{0.45\textwidth}
+%		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
+%	\end{subfigure}
+%\end{figure}
+%\end{graphicalabstract}
 
@@
 \label{sec:Intro}
-
-Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico-newtons.
-Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.
-Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass\cite{yifat2021facile}.
-A recent work aimed at characterising trapped entities demonstrated how neural networks maybe trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}.
-A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle.
+Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks to the Brownian dynamics of colloidal systems to the tweezer can transduce and detect forces down to the order of a few pico-newtons.
+Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal science \cite{burns1990optical}.
+Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass \cite{yifat2021facile}.
+A recent work aimed at characterising trapped entities demonstrated how neural networks can be trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}.
+A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells also was attempted \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle.
 Thus, past studies on optical trapping have focused on either on particle trapping to study trapping dynamics, or on the characterisation of particles with relatively simple complex trapping dynamics.
 
-While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles with simple trapping dynamics, and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics.
-A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion - commonly now refereed to as Localised dynamic light scattering (LDLS).
-This was later expanded upon by \cite{PhysRevE.65.041921} by collecting back scattered light to characterise the Stokes friction coefficient as a function of trapping depth.
-While both papers provided dynamical information, structural information about the trapped bead was precluded as the scattered wave was only measured across a small angular range.
-Furthermore, the main drawback to a LDLS is that it struggles to characterise asymmetric particles such as nano rods or dimers; in which case the Brownian motion is no longer simply translational, but rotational about the centre of mass \cite{yin2013large}.
+While both \cite{Carvalho_2023} and \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles with simple trapping dynamics, and \cite{Watson_2023} describes an experimental setup that is difficult to adopt and suffers from a low bandwidth that might not be best suited for monitoring dynamics.
+A light scattering detection scheme built around an optical trap that is easier to implement and has the advantage of high bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion, commonly now referred to as Localised Dynamic Light Scattering (LDLS).
+This was later expanded upon \cite{Viana2002} by collecting back scattered light to characterise the Stokes friction coefficient as a function of trapping depth.
+While both papers provided dynamical information, structural information about the trapped bead was not available as the scattered light was only measured across a small angular range.
+Furthermore, the main drawback to a LDLS is that it cannot be used to characterise asymmetric particles such as dimers or more complex aggregates; in which case the Brownian motion is no longer simply translational, but rotational about the centre of mass \cite{yin2013large}.
 
-One way of mitigating this is to remove translational fluctuations from the analysis; by monitoring its the scattering, Cang and co-workers were able to spatially fix a gold nanorod at the centre of trapping laser by moving the sample plane accordingly, to an accuracy of 200 nm \cite{Cang2006}.
-However, this has limited application's to anisotropic scatters such as biological matter where the internal structure of many cells makes them inherently anisotropic \cite{Watson_2023}.
-Thus even if the Cartesian coordinates are relatively fixed characterisation of the scattering pattern will have to separate the contributions of internal structure and orientation.
-In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique, to enable interpretation of the resulting multi-angle data from an anisotropic, asymmetric scatterer, as well as optimisation to provide maximal information from the signal.
+One way of mitigating this is to remove translational fluctuations from the analysis; by monitoring it’s scattering pattern, Cang and co-workers were able to spatially fix a gold nanorod at the centre of trapping laser by moving the sample plane accordingly, to an accuracy of 200 nm \cite{Hu2006}.
+However, this has limited applications to anisotropic scatterers such as biological matter where the internal structure of many cells makes them inherently anisotropic \cite{Watson_2023}.
+Thus, even if the Cartesian coordinates are fixed, characterisation of the scattering pattern will have to separate the contributions of size, structure, and orientation.
+As an example of the relatively sparse literature on measuring orientations of complex trapped objects, \cite{raudsepp2022estimating} employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted.
+In this work, we propose a novel approach that expands on the previous approaches \cite{Bar-Ziv_1998} \cite{Viana2002} to detect scattered light simultaneously at multiple angles (Figure 1), combined with a novel Bayesian inference-based analysis technique, to enable interpretation of the resulting multi-angle data from an anisotropic, asymmetric scatterer, and optimisation to provide maximal information from the detector signals.
 
-To demonstrate the analysis, we study a simple, illustrative example of an anisotropic scattering entity \emph{i.e.} an asymmetric dimer.
-As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate.
-As an example of the relatively sparse literature on measuring orientations of complex trapped objects, \cite{raudsepp2022estimating} employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted.
-Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer.
-We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals.
-Furthermore we demonstrate how the model's performance when dealing with signal noise, a common problem when analysing scattering behaviour.
+To demonstrate this approach, we study an anisotropic scattering entity i.e.
+an asymmetric dimer, to determine dynamic and structural information about the trapped entity.
+As a paradigmatic example of extracting information from the scatterer’s scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate We, first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the 2 known instantaneous orientation of the simulated dimer.
+We then show how Bayesian inference can be used to optimise our estimation of the true dimer orientation from the light scattering signals.
+Furthermore, we demonstrate how the model's performance when dealing with signal noise, a common problem when analysing scattering behaviour.
+This approach can be extended to determine any characteristic that influences the light scattering pattern of trapped colloidal objects including: size, shape, and orientation.
 
-\begin{center}
-		{\bf{Fig 1a, 1b}}
-\end{center}
-\begin{figure} [b]
+\begin{figure} [h]
 \centering
@@
 
-\begin{center}
-	\bf{Fig 2a, 2b}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 3a, 3b}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 4a, 4b}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 5}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 6}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 7a, 7b}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 8a, 8b, 8c}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 9a, 9b, 9c}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 10a, 10b, 10c}
-\end{center}
 \begin{figure}[h]
@@
 
-\begin{center}
-	\bf{Fig 11a, 11b, 11c}
-\end{center}
 \begin{figure}[h]
@@
 
+From the above discussion it’s clear that estimation of the dimer’s orientation is a problem that can be endlessly tuned to fully maximise our end result.
+Here we simplify the problem somewhat by employing a relatively small finite number of 'reference orientations' to map between scattering and dimer orientation: the precision of estimation could be improved by utilising a greater number of reference orientations, although there remains a balance between the realisable precision of orientation estimate and the noise level of the scattering measurement.
+Another avenue to further explore would be using the method to optimise the choice of detection angles, essentially to find the region in the mapping between measured scattering and orientation that others the best degree of confidence through optimal separation of scattering signals for distinct orientations.
+For sequences of data such as dynamic measurements, a further potential enhancement would be to consider more complex correlations based on prior expectations of the dynamics.
+Here already we improve the method using a non-uniform prior based on only the immediately previous measurement in time (see Section 2.1): considering a non-uniform grouping of reference orientations might result in a better estimation, if we have information regarding the dimer’s preferred axis of rotation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We have developed a method for measuring the dynamics of an optically-trapped 'complex object' based purely on limited measurements of the object's light scattering at a small number of detection angles.
-We demonstrate the method using the orientation of an asymmetric dimer as the dynamical variable and object of interest, but in principle the model can be applied to any characteristic that impacts the light scattering pattern produced by a trapped entity.
+We have developed a method for measuring the dynamics of an optically-trapped colloidal objects based purely on measurements of the object's light scattering at a small number of detection angles.
+We demonstrate the method using the orientation of an asymmetric dimer as the dynamic variable and object of interest respectively, but in principle the model can be applied to any characteristic that impacts the light scattering pattern produced by a trapped entity such as size and shape.
 The MSTM package is a flexible tool for calculating the light scattering of complex objects using a representation of the object as a set of micro-particles, enabling training of a neural network to enable categorisation of the mapping between scattering and trapped object characteristics.
-By taking account of knowledge about the physically realistic behaviour of the trapped object and the characteristics of the trap (which impact the dynamics of the object), the Bayesian inference method can be refined to provide a reliable estimation of object characteristics of interest, even in the presence of measurement noise.
-Fundamentally the inverse scattering problem is difficult to solve since the mapping between object characteristics and scattering can be highly complex: but Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.\\
-
-Here we simplify the problem somewhat by employing a relatively small finite number of 'reference orientations' to map between scattering and dimer orientation: the precision of estimation could be improved by utilising a greater number of reference orientations, although there remains a balance between the realisable precision of orientation estimate and the noise level of the scattering measurement.
-Another avenue to further explore would be using the method to optimise the choice of detection angles, essentially to find the region in the mapping between measured scattering and orientation that offers the best degree of confidence through optimal separation of scattering signals for distinct orientations.
-For \textit{sequences} of data such as dynamic measurements, a further potential enhancement would be to consider more complex correlations based on prior expectations of the dynamics.
-Here already we improve the method using a non-uniform prior based on only the immediately previous measurement in time (see Section \ref{sec:Bayes}): considering more detailed correlations such as multiple previous timesteps is likely to further enhance the reliability of the estimation.
+By taking account of the physically realistic behaviour of the trapped object and the characteristics of the trap (which impact the dynamics of the object), the Bayesian inference method can be re ned to provide a reliable estimation of object characteristics of interest, even in the presence of measurement noise.
+Fundamentally, the inverse scattering problem is difficult to solve, since the mapping between object characteristics and scattering can be highly complex.
+We determined the minimum number of detectors required for a reliable estimation in the presence of measurement noise; furthermore, we demonstrated that the arrangement of these detectors is critical for a reliable estimation of an objects orientation.
+However, Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.
 

### 2023-10-25 Images/freq_domain_fit.py

@@ -1,45 +1,0 @@
-import argparse
-import csv
-import numpy as np
-from scipy.optimize import curve_fit
-import scipy
-from scipy.signal import correlate
-import matplotlib.pyplot as plt
-from datetime import datetime
-
-class freq_domain:
-    '''
-    Class for running frequency domain analysis from QPD data, 
-    '''
-    def __init__(self):
-        self.kT = 1.38e-23*290
-        current_time = datetime()
-        self.tag = current_time.now()
-
-    def create_csv_file(self, path):
-
-        def row_reader(self, row):
-            line = list(row[0].split())
-            new_row = [float(line[0]), float(line[1]), float(line[2]), float(line[3])]
-
-            return new_row
-        
-        with open(path, 'r') as input:
-            reader = csv.reader(input.readlines()[5:])
-
-            with open('tmp_tdat.csv', 'w') as output: 
-                writer = csv.writer(output)
-                writer.writerows(self.row_reader(row) for row in reader)
-
-        return self.get_data('tmp_fdat.csv')
-    
-    def get_data(self, file):
-        with open(file, 'r') as data:
-            reader = csv.reader(data.readlines())
-            time, x_QPD, y_QPD = [], [], []
-            for row in reader:
-                time.append(row[0])
-                x_QPD.append(row[1])
-                y_QPD.append(row[2])
-
-        return time, x_QPD, y_QPD

### 2023-10-25 python files/freq_domain_fit.py

@@ -1,0 +1,87 @@
+import argparse
+import csv
+import numpy as np
+from scipy.optimize import curve_fit
+import scipy
+from scipy.signal import correlate
+import matplotlib.pyplot as plt
+from datetime import datetime
+
+class freq_domain:
+    '''
+    Class for running frequency domain analysis from QPD data, 
+    '''
+    def __init__(self):
+        self.kT = 1.38e-23*290
+        current_time = datetime.now()
+        self.tag = current_time.now()
+
+    def create_csv_file(self, file):
+
+        def row_reader(row):
+            line = list(row[0].split())
+            line += list(row[1].split())
+            new_row = [float(line[0]), float(line[1]), float(line[2]), float(line[3])]
+
+            return new_row
+        main_path = 'C:\\Users\\xbb16146\\dmaciver97\\data\\Trapping\\'
+        path = main_path+file
+        with open(path, 'r') as input:
+            reader = csv.reader(input.readlines()[5:])
+
+            with open('tmp_fdat.csv', 'w', newline='') as output: 
+                writer = csv.writer(output)
+                writer.writerows(row_reader(row) for row in reader)
+
+        return self.get_data('.\\tmp_fdat.csv')
+    
+    def get_data(self, file):
+        with open(file, 'r') as data:
+            reader = csv.reader(data.readlines())
+            freq, x_QPD, y_QPD = [], [], []
+            for row in reader:
+                freq.append(float(row[0]))
+                x_QPD.append(float(row[1]))
+                y_QPD.append(float(row[2]))
+
+        return freq, x_QPD, y_QPD
+    
+    def plot_psds(self, freq, x_fdat, y_fdat):
+        plt.plot(freq, x_fdat)
+        plt.plot(freq, y_fdat)
+        plt.xscale('log')
+        plt.yscale('log')
+        plt.show()
+
+    def fit_psd(self, freq, fdat):
+        
+        def simplified_psd(f, A, B):
+            return 1/(A+B*f**2)
+        
+        def evaluate_fit(fit, fdat):
+            SS_res = np.sum([(y-f)**2 for y,f in zip(fdat, fit)])
+            SS_tot = np.sum([(y-np.mean(fdat))**2 for y in fdat])
+            return 1-(SS_res/SS_tot)
+        
+        r2 = 0
+        for i in range(5):
[…]

### 2023-10-25 python files/maclaurin.py

@@ -1,0 +1,21 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+x_range = np.linspace(-5, 5, 50)
+
+
+for x in x_range:
+    macsum = 1
+    kterm = 1
+    k = 1
+    while kterm > 10e-11:
+        kterm *= (x/k)
+        macsum += kterm
+        k += 1
+    plt.scatter(x, macsum, c='black', s =3)
+
+
+plt.plot(x_range, [np.exp(x) for x in x_range])
+plt.vlines(0, -10, np.exp(5), color='black', linestyle ='--')
+plt.show()
+plt.close()

### 2023-10-25 python files/test.py

@@ -1,1 +1,10 @@
-print('hello world') 
+import csv
+import numpy as np
+import statsmodels
+import matplotlib.pyplot as plt
+
+raw = np.load('.\python files\tmp\Raw2023-07-05 13_16_32.454088.csv.npy')
+I0 = raw[::50, 0]
+
+plot_acf(I0, lags= 100)
+plt.show()

### 2023-10-25 python files/time_domain_fit.py

@@ -49,6 +49,7 @@ class time_domain:
 
         def row_reader(self, row):
             line = list(row[0].split())
+            
             new_row = [float(line[0]), float(line[1]), float(line[2]), float(line[3])]
 
             return new_row
@@ -58,7 +59,7 @@ class time_domain:
 
             with open('tmp_tdat.csv', 'w') as output: 
                 writer = csv.writer(output)
-                writer.writerows(self.row_reader(row) for row in reader)
+                writer.writerows(row_reader(row) for row in reader)
 
         return self.get_data('tmp_tdat.csv')
     

### 2023-10-25 python files/triangle.py

@@ -1,0 +1,80 @@
+import numpy as np
+import plotly.graph_objects as go
+import turtle
+import warnings
+
+while True:
+    try:
+        sides = np.zeros(3)
+        sides[0] = float(input('Side no 1: '))
+        sides[1] = float(input('Side no 2: '))
+        sides[2] = float(input('Side no 3: '))
+        rotation = float(input('Starting angle from horizontal: '))
+        sides = sorted(sides)
+        break
+    except ValueError:
+        print('Sides cannot be strings')
+
+a,b,c = sides[0], sides[1], sides[2]
+# If triangle sides are wrong then it will throw up a run time error
+# when calculating the angles
+
+with warnings.catch_warnings(record=True) as w:
+    
+    alpha = np.arccos((b**2+c**2-a**2)/(2*b*c))*(180/np.pi)
+    beta = np.arccos((a**2+c**2-b**2)/(2*a*c))*(180/np.pi)
+    gamma = np.arccos((a**2+b**2-c**2)/(2*a*b))*(180/np.pi)
+   
+    if len(w) > 0:
+        print('These side lengths cannot form a triangle')
+        exit(0)
+    else:
+        area = 0.5*a*b*np.sin(gamma*np.pi/180)
+        height = 2*area/a
+        
+
+# Rearrange angles for turtle drawing
+# each angle is opposite its corresponding side so needs to be shifted
+angles = [gamma, alpha, beta]
+
+# Define the Triangle Type
+if a == b and b == c:
+    tritype = 'equilateral triangle'
+elif (a == b and a != c) or (b == c and b != a) or (c == a and c != b):
+    tritype = 'isoceles triangle'
+elif a != b and a != c:
+    tritype = 'scalene triangle'
+
+# Last check involves angles so should be seperate so it isn't skipped
+if np.isclose(90, gamma, atol = 0.001):
+    tritype = 'Right Angled triangle'
+
+t = turtle.Turtle()
+screen = turtle.Screen()
+rootwindow = screen.getcanvas().winfo_toplevel()
+rootwindow.call('wm', 'attributes', '.', '-topmost', '1')
+rootwindow.call('wm', 'attributes', '.', '-topmost', '0')
+screen.setup(800,800)
+# Adjust starting position
+t.penup()
+t.goto(-10*a, -10*height)
+t.pendown()
+t.left(rotation)
+for i in range(3):
+    #distance = sides[i]
[…]

### 2024-02-06 Latex files/Manuscript Dimers/Manuscript_Optical Communications.tex
Adding lib folders for scattering and dimer manus

\documentclass[preprint,  3p]{elsarticle}

\usepackage{csvsimple}
\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomencl}
\usepackage{tabularx}
\makenomenclature
\renewcommand{\nomlabel}[1]{\hfil #1\hfil}

\journal{}

\begin{document}
	\begin{frontmatter}
		\title{Computer Simulations of The Dynamics of Asymmetric Dimers in Optical Traps of Varying Polarization}

		\author[aff1]{Praveen Parthasarathi\corref{cor1}}
		\ead{praveen.parthasarathi@strath.ac.uk}

		\author[aff1]{Daniel Maciver}

		\author[aff1]{Leo Lue}

		\author[aff1]{Jan Sefcik}

		\author[aff1]{Mark Haw}

		\cortext[cor1]{Corresponding author}
		\affiliation[aff1]{organization={Department of Chemical Engineering,
				University of Strathclyde},
			addressline={75 Montrose Street}, 
			city={Glasgow},
			postcode={G1 1XL}, 
			country={Scotland}}

		\begin{abstract}
			% \justifying
			%
			We report on computer simulations of the dynamics of trapping, equilibrium positions and orientations of asymmetric microsphere dimers of varying size ratios in a single Gaussian beam Optical Trap of varying laser-polarization. We show that in case of a dimer, the trapping force profile deviates very quickly from a simple linear sum of forces on individual spheres as the size of the second sphere is increased and shows multiple axial-equilibrium positions for the case of larger bead closer to the laser focus as against the case of a smaller bead closer to the focus. Furthermore, we also investigate sustained rotations of dimers about their long-axis when trapped in a circularly polarized laser and study the effect of size-asymmetry on rotations. We hope these studies might be useful to researchers investigating colloidal aggregation and nucleation in optical traps and we also make a case for optically trapped dimers as rotors for use in experiments such as colloidal mixing.  
		\end{abstract}

		%\begin{graphicalabstract}
		%\begin{figure} [h]
		%	\centering
		%	\begin{subfigure}{0.45\textwidth}
			%		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
			%	\end{subfigure}
		%	\begin{subfigure}{0.45\textwidth}
			%		\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
			%	\end{subfigure}
		%\end{figure}
		%\end{graphicalabstract}

		%\begin{highlights}
		%\item Asymmetric microsphere dimers can undergo complex dynamics including full inversions in the presence of a trapping potential, depending sensitively on the size ratio of the spheres making up the dimer.  
		%\item Orientations can be inferred from scattering signals through classification by a combination of neural networks and Bayesian inference. 
		%\item Signal error has a significant effect on the performance of the neural network, which can be countered via biasing of the prior distribution. 
		%\item Future applications include continuous processes where trapped particle characteristics such as size and shape are changing with time.  
		%\end{highlights}

		\begin{keyword}
			Optical Trapping \sep Brownian Dynamics \sep Micro-Rotors  \sep Colloids 
		\end{keyword}

	\end{frontmatter}

\section{Introduction}
Ever since their invention in 1986 by Arthur Ashkin \cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been used extensively in colloidal physics \cite{Leonardo_2008, Koehler_2011, Curran_1999}. Of particular interest has been the use of a single trapped microsphere as a tracer to probe the rheological properties of the fluid of suspension \cite{Atakhorrami_2006}. However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead in a Gaussian beam OT. Li and Arlt \cite{Li_2008} studied the case of two microspheres trapped in a single OT and opined that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness. Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu_2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Sheng-Hua et al \cite{ShengHua_2005} assumed that the two trapped beads experience different trap stiffnesses and went on to show that the trapped beads collide. Experiments in \cite{Praveen_2016} confirmed that the two trapped beads indeed experience different trap stiffnesses when trapped in a single Gaussian beam OT.

Besides multiple trapped beads, trapping and dynamics of asymmetric shapes have generated a lot of interest. Mihiretie et.al., \cite{Loudet_2014} studied the trapping of ellipsoidal microparticles and found that particles of aspect ratio ‘k’ greater than 3 and less than 0.3 exhibited oscillatory dynamics with a coupling between translation and rotation in the trap and those with aspect ratio outside this range exhibited less pronounced oscillations with the major axis pre-dominantly pointed along the laser propagation direction. Cao et.al., \cite{Curran_1999} performed T-Matrix based simulations of the trapping dynamics of nanowires and microcylinders in the Rayleigh regime and found that polarization dictated the final equilibrium orientation with the nanowires equilibrating with a large tilt angle about the laser propagation direction when the beam was linearly polarized and along the beam propagation direction in circularly polarized traps. Chetana D et, al., \cite{Chetana_2022} increased the complexity of the problem by considering an object with both shape as well as optical anisotropy in the form of a chicken Red Blood Cell (cRBC) and showed that the trapping dynamics and the final equilibrium orientation of the cRBC was decided by the beam polarization with the cell showing a dual reorientation behaviour in a linearly polarized trap and a partial second reorientation in a circularly polarized trap.  
In this work, we extend the Brownian dynamical simulations carried out by Vigilante et.al., \cite{Vigilante_2020} to the case of asymmetric dimers. We begin by showing that the optical force on a microsphere-dimer quickly deviates from a simple sum of forces on two spheres and thus underscore the relevance of the choice of mstm, a well-established package for calculating light scattering from a cluster of spheres \cite{Mackowski_2011} and within it ott, a well-established toolbox for computing forces and torques in an OT \cite{Nieminen_2007}.  We show validations of the dynamical simulations for the case of a single trapped bead in both linear and circularly polarized traps with the suspension medium being water at room temperature. We then reproduce the results for dynamics of a symmetric dimer in \cite{Vigilante_2020} as further validation of our codes. We organize our results into 3 sections with the focus on axial and orientation trapping equilibria for different size-ratios of the dimers in the first section, determinations of equilibrium orientations and equilibrium positions of various dimers with varied initial conditions of orientation and position in linearly and circularly polarised traps, Finally, we report interesting rotational dynamics of dimers seen only when the light is not linearly polarised. We hope that our results could inform studies on colloidal aggregation and help design rotors for use in applications such as lab-on-a-chip and micromachines.

\newpage
\bibliography{dimer_bib} 
\bibliographystyle{ieeetr}

\section{Methodology}
\paragraph{Simulating motion within an optical trap}
We assume the dimers to be low-Reynold’s number objects and write down their dynamics within the trap using the Langevin equation:
\begin{align}
	\label{eq:langevin}
	\frac{dx(t)}{dt}=\frac{F(x)}{\gamma}+\sqrt{2D}\eta_x(t)
\end{align}

Where $F(x)$ is the optical force directed towards the focal point of the laser However, this is only accurate when close to the equilibrium position, if we wish to probe the phase space we need a means of computing the optical forces for any position or orientation. Fortunately, Vigilante et al \cite{Vigilante_2020} have already built the ground work for simulating multi-sphere trapping. This work is an extension of the simulation work conducted by \cite{Vigilante_2020} into multi-sphere objects; their paper already has an excellent breakdown of the mathematics that go into calculating the optical and hydrodynamic forces, as such we will only refer to the mathematical methodology when necessary or when our work diverges from \cite{Vigilante_2020}. For simulating asymmetric objects, in which translation and rotation dynamics are coupled, it is useful to have two separate reference systems: The laboratory frame, which refers to the larger system centred at the focal point of the trapping laser; and the particle reference frame which defines the positions of individual spheres, in the case of a dimer the spheres are positioned so that the centre of diffusion is at the origin of the particle reference frame. The axis of the particle reference frame is used to define the orientation of the dimer; with Uz intersecting the centres of the two spheres; any rotations are applied to the reference axis so that the orientation is preserved.

The code combines the optical force calculation code from Optical Tweezer Toolbox (ott) and the T-matrix calculations from Multi-Sphere T-Matrix (mstm); after initializing the target object mstm calculates its T-matrix before returning the T-matrix object to ott to calculate the optical forces and torques as well as the hydrodynamic forces due to the Brownian force. This circumvents the limitations of both ott and mstm; mainly that ott struggles to model the T-matrix of multi-shape objects that experience internal scattering, and that mstm cannot model scattering from lasers with a high NA commonly used in optical trapping. After calculating the total force and torque in the laboratory frame they are scaled down to the particle reference frame, where the reference axis is rotated appropriately before then displacing the object in the laboratory frame to update its position.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.75\textwidth]{./Images/Lab_frame.png}
	\caption{\label{fig:lab_frame}
		%
		Render of trapping beam and a dimer (size ratio 2) with respective laboratory frame and particle reference frame. The laboratory frame origin is set at the focal point of the Gaussian beam. The particle reference axes are used to set the relative positions of each sphere, the axes are centred on dimer's centre of diffusion and used to describe the dimer’s orientation. Scale bar in bottom right corner represents 1 $\mu$m
		%
	}
\end{figure}

The Brownian OT code from Vigilante has code for simulating both dimers, and sphere clusters; the former is only capable of processing symmetric dimers with no ability to freely tune the size of each sphere. To simulate the behaviour of asymmetric dimers we modelled the entity as a sphere cluster consisting of two spheres tangentially attached, the larger sphere is defined as sphere 1 with radius ‘$r_1$’ the second sphere is then defined by the size ratio $\lambda$ so that $r_2 = \lambda r_1$.  Knowing the size ratio is critical for defining the dimer’s centre of diffusion and also for appropriately scaling its diffusion tensor. The diffusion tensor is calculated from Nir and Acrivos \cite{Nir_1973} work into arbitrary shaped spherical objects; where they write the diffusion tensor as a 6x6 matrix, when the dimer is symmetric only the central diagonal is non-zero. However, for asymmetric dimers the forces on the spheres are unequal resulting in stress build up along the dimer’s long axis, the diffusion tensor is then given as:
\begin{align}
	\label{eq:diffusion}
	D=\frac{k_BT}{\pi\mu}\left[
	\begin{matrix}
		a_1 & 0 & 0 & 0 & d_1 & 0\\
		0 & a_1 & 0 &-d_1 & 0 & 0\\
		0 & 0 & a_1+a_2 & 0 & 0 & 0\\
		0 & -d_1 & 0 & b_1 & 0 & 0\\
		d_1 & 0 & 0 & 0 & b_1 & 0\\
		0 & 0 & 0 & 0 & 0 & b_1+b_2\\
	\end{matrix}\right]
\end{align}
Where the constants are spline fitted from the results of [14], due to the dimer’s asymmetry the friction coefficient – which is related to the diffusion coefficient – is different depending on the axis the fluid is flowing parallel to. The Brownian displacements and rotations are generated at random in proportion to the diffusion tensor. 
\paragraph{Finding equilibrium positions}
In order to determine the equilibrium trapping locations of an arbitrary multi-sphere cluster we need to be able to compute the optical force at any position and orientation. The optical forces can be computed by a summation of the incident and scattered beam expansion coefficients; the z-axis force is given below:

\section{Results and Discussion}
Trapping forces on a sphere are often written as the rate of change in momenta of incident and scattered beams \cite{Nieminen_2007}. We show in Fig. [1], the axial-force on a single trapped polystyrene micro bead of varying radii as the bead is moved along the z-axis computed using the Optical Tweezer Toolbox ott under a linearly polarised trap generated with an objective lens of NA 1.2 and wavelength 1064nm. For the case of a dimer, we use mstm to compute the scattered fields and ott to compute the forces and torques consequent of the scattering processes. It may seem reasonable to write down the force on a dimer as the sum of forces on two beads that make up the dimer with positions of the centre of each bead measured from the beam focus. We put this approximation to test in Fig [2] where we compare it with a rigorous calculation of the force using mstm and ott and see it fail as soon as the second bead starts being a 10th in radius of the first bead. This makes the case for use of packages like mstm which are dedicated to computing multi-sphere scattering. We adopt the python codes in \cite{Vigilante_2020} to simulate Brownian dynamics of a dimer made of spheres of diameter 0.8micron trapped in a left-circularly polarised light of power 5mW and the spheres are assumed to be made of polystyrene of refractive index 1.58 with water of index 1.33 at 295K as the medium of suspension. Optical forces and torques on this dimer are shown in Fig [3].  and rotation frequency of the dimers as a function of the ratio of the Stokes vector element $S_3/S_0$ is shown in Fig. [4]. We see very good agreement of these results with those reported in \cite{Vigilante_2020}.

\subsection{Axial forces and Equilibrium positions}
We explore the trapping landscape by plotting out the trapping force as the dimer is moved along the z-axis and see that for the case of large-bead-over-small, there are multiple trapping equilibria compared to the case of small-bead-over-large as seen in Fig [4] and Fig [5]. In all these calculations, we maintain the radius of one of the beads making up the dimer at 1micron and vary the size of the other bead to 500nm, 250nm, and 100nm respectively in the case of size-ratios 2, 4, and 10. In Fig. [6] we plot out the force curve for a symmetric dimer with the size of each sphere equal to 1 $\mu m$.

\end{document}

### 2024-02-06 python files/psd_test.py

@@ -1,0 +1,26 @@
+import numpy as np
+from scipy.signal import periodogram
+import random
+import matplotlib.pyplot as plt
+
+dt = 1e-5
+w_i = [0]
+N = 100000
+for i in range(N):
+    w_i.append(random.normalvariate(0,1)+w_i[i])
+
+step_sizes = np.logspace(0, 5, 10000)
+msd_w = []
+tau = [step*dt for step in step_sizes]
+for step in step_sizes:
+	n = int(N/step)
+	di = int(step)
+	tmp_w = []
+	for i in range(0, n-1):
+		tmp_w.append((w_i[i+di]-w_i[i])**2)
+	msd_w.append(np.mean(tmp_w))
+
+plt.plot(tau, msd_w)
+plt.xscale('log')
+plt.yscale('log')
+plt.show()

### 2024-02-06 python files/test.py

@@ -1,10 +1,33 @@
 import csv
+import random
 import numpy as np
-import statsmodels
 import matplotlib.pyplot as plt
 
-raw = np.load('.\python files\tmp\Raw2023-07-05 13_16_32.454088.csv.npy')
-I0 = raw[::50, 0]
+import numpy as np
+import matplotlib.pyplot as plt
+vec = [random.normalvariate(0,1),random.normalvariate(0,1),random.normalvariate(0,1)]
+mag = np.dot(vec,vec)
+S = [v/mag for v in vec]
+def get_circle():
+	L = np.hypot(S[0], np.hypot(S[1], S[2]))
+	if S[0] > 0: sig = L
+	else: sig = -L
+	h = S[0]+sig
+	beta = -1/(sig*h)
+	f = beta*S[1]
+	g = beta*S[2]
+	return [f*h, 1+f*S[1], f*S[2]], [g*h, g*S[1], 1+g*S[2]]
 
-plot_acf(I0, lags= 100)
+fig = plt.figure()
+ax = fig.add_subplot(projection ='3d')
+ax.quiver(0,0,0, S[0], S[1], S[2])
+ax.set_xlim(-1,1)
+ax.set_ylim(-1,1)
+ax.set_zlim(-1,1)
+U, V = get_circle()
+theta_list = np.linspace(0,2*np.pi, 100)
+for theta in theta_list: 
+	pos = [(s+np.cos(theta)*u+np.sin(theta)*v) for s,u,v in zip(S,U,V)]
+	ax.scatter(pos[0], pos[1], pos[2])
+
 plt.show()

### 2024-02-06 Latex files/Manuscript Dimers/Manuscript_Optical Communications.tex
Adding rotation papers and updating dimer manu

@@
 \usepackage{tabularx}
+\usepackage{graphicx}
+\usepackage[skip=0.333\baselineskip]{caption}
 \makenomenclature
@@
 \newpage
-\bibliography{dimer_bib} 
-\bibliographystyle{ieeetr}
 
@@
 [6] we plot out the force curve for a symmetric dimer with the size of each sphere equal to 1 $\mu m$.
+\begin{figure}
+\begin{subfigure}{.475\linewidth}
+	\includegraphics[width=\linewidth]{Images/lam=2_theta=0.png}
+	\caption{}
+	\label{lam=2}
+\end{subfigure}\hfill % <-- "\hfill"
+\begin{subfigure}{.475\linewidth}
+	\includegraphics[width=\linewidth]{Images/lam=4_theta=0.png}
+	\caption{}
+	\label{lam=4}
+\end{subfigure}
 
+\medskip % create some *vertical* separation between the graphs
+\begin{subfigure}{.475\linewidth}
+	\includegraphics[width=\linewidth]{Images/lam=10_theta=0.png}
+	\caption{}
+	\label{lam=10}
+\end{subfigure}\hfill % <-- "\hfill"
+\begin{subfigure}{.3\linewidth}
+	\includegraphics[width=0.45\linewidth, keepaspectratio]{Images/theta=0.png}
+	\label{large over small}
+\end{subfigure}
+\caption{Plots of force vs displacement of the point of the contact of the spheres(µm) for the case of a dimer with larger bead above the smaller bead when viewed along the positive z-direction.
+Size ratios in (a), (b), and (c) are 1:2, 1:4, and 1:10 respectively.
+The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}{.475\linewidth}
+		\includegraphics[width=\linewidth]{Images/lam=2_theta=180.png}
+		\caption{}
+		\label{lam=2_inverted}
+	\end{subfigure}\hfill % <-- "\hfill"
+	\begin{subfigure}{.475\linewidth}
+		\includegraphics[width=\linewidth]{Images/lam=4_theta=180.png}
+		\caption{}
+		\label{lam=4_inverted}
+	\end{subfigure}
+
+	\medskip % create some *vertical* separation between the graphs
+	\begin{subfigure}{.475\linewidth}
+		\includegraphics[width=\linewidth]{Images/lam=10_theta=180.png}
+		\caption{}
+		\label{lam=10_inverted}
+	\end{subfigure}\hfill % <-- "\hfill"
+	\begin{subfigure}{.29\linewidth}
+		\includegraphics[width=0.35\linewidth, keepaspectratio]{Images/theta=180.png}
+		\label{small_over_large}
+	\end{subfigure}
+	\caption{Plots of force vs displacement of the point of the contact of the spheres(µm) for the case of a dimer with smaller bead above the larger bead when viewed along the positive z-direction.
+Size ratios in (a), (b), and (c) are 1:2, 1:4, and 1:10 respectively.
+The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
+\end{figure}
+
+Figure 5 shows that there exists a second trap below the focus of the trapping beam, this trap is only stable when the dimer is orientated against the beam propagation direction.
+We extend this analysis for size ratios between 1 to 10.
+Not only is the second trap only possible within the range 1.1 to 4, after which the second trap stiffness goes to 0, the trapping positions as a function of size ratio are shown below.
+
+\subsection{Torques and equilibrium orientations:}
+Having determined the axial-equilibrium positions of the dimers, we go about determining the steady-state orientations of the dimers assuming they’ve been trapped at their axial-equilibrium positions.
+The torques reported here are about the X and Y directions of the dimer as discussed in the section on Methods.
+Fig.
+[7] shows plots of Torque as a function of angles made with the X and Y axes of a reference frame whose origin coincides with the point-of-contact of the dimer as described in the section on Methods and the orientation of the dimer is small-over-large looking along the positive z-axis.
+In Fig 7(a) and Fig 7(b), we show plots of torque vs orientation for a dimer of size-ratio 1:2 at axial positions -1.63µm and 1.2µm respectively.
+In Fig 7(c) and Fig 7(d), we show torque vs orientation plots for a dimer of size ratio 1:4 at axial positions 0.11µm and -0.84µm respectively and in Fig 7(e) we show torque vs orientation for a dimer of size ratio 1:10 trapped at the focus.
+In Fig [8] we show torque curves for a symmetric dimer where the microspheres are both assumed to be of radius 1micron.
+We change the polarisation of light to linear in plots shown in Fig [9].
+We assume the light to be polarised along x-axis and see no significant torque along this direction and therefore we only show torque vs angular displacement about y-axis.
+In Fig.
+[10], we show these results for a symmetric dimer where the microsphere radius is 1micron.
+
+Simpson and Hanna have reported a strong coupling between translational and rotational motion of rod-shaped microscopic particles in an OT [] and we feel it’s reasonable to expect a similar effect in case of dimers.
+Therefore, we probe the stability of the axial-equilibrium to rotations when dimers are rotated after being trapped at their axial equilibrium positions reported in Fig [5].
+We therefore plot axial force as a function of displacement about the y-axis when the trap is linearly polarised along x-direction.
+These are shown in Fig [11].
+
+\subsection{Rotation Dynamics}
+Dimers have been shown to rotate within a circularly polarised trap [] meaning they have the potential to be used in micro rotor applications.
+Interestingly, while the rotation speed is linearly proportional to the laser power it is also dependent on both the orientation and size of the dimer, with the orientation having a greater impact on the rotation speed.
+We suspect that since the dimer’s trapping position is also dependent on the orientation there is a lower transfer of angular momentum when the dimer is closer to the trap focus.
+This is reflected in Fig 12b where we see the impact of polarization angle on the rotation speed, maximising at 45o and seeing no rotation at 0¬o and 90o, we simulated dimers of different size ratios with the solid lines representing the rotation speed when the dimer is orientated with the sphere II towards the beam focus and individual points representing the rotation speed when sphere II is orientated away from the beam focus.
+Despite having the same overall shape and size a 1:4 dimer will not rotate when oriented towards the beam focus, where as the reverse case the dimer rotates at a frequency of roughly 45 Hz.
+
+This is also the case for our second trap below the focus of the beam, in this case we see an even greater rotation rate than above the focus.
+Interestingly the direction of rotation is flipped upon crossing the focus of the trap, this suggests that the rotations are not induced by a spin-curl effects as these effects are shown to only be caused by the polarisation of the electromagnetic field.
+We also note that the relationship is also somewhat related to the dimer’s displacement from the trap focus, where a greater displacement results in a faster rotation rate, where in each case as soon as the dimer is trapped at the focus the rotation frequency goes to 0.
+
+\section{Conclusions and Outlook}
+
+	We study by means of computer simulations, the orientation dependent trapping landscapes for asymmetric microsphere dimers in both linearly and circularly polarized traps and show that trapping with circularly polarized light can lead to sustained rotations that go roughly as 0.25 Hz/mW for the orientation of small sphere over large when viewed opposite to the beam propagation direction.
+These results should provide insights with regards development of experiments probing the trapping behaviour of dimers and also make for interesting comparisons with other complex shapes such as ellipsoids and rods.
+Furthermore, we hope our results might be useful for development of micro-rotors for lab-on-a-chip assays.
+
+\newpage
+\bibliography{dimer_bib} 
+\bibliographystyle{ieeetr}
 \end{document}

### 2024-05-17 python files/tmp/testfile.py
Files for meeting

@@ -1,0 +1,183 @@
+#!/usr/bin/python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from statsmodels.tsa.stattools import ccf
+
+
+def make_ccf(pair1, pair2):
+    fwr_diag_cross_product = ccf(pair1, pair2, adjusted=True, fft=True)
+    fwr_diag_cross_product /= np.max(abs(fwr_diag_cross_product))
+    bck_diag_cross_product = ccf(pair1[::-1], pair2[::-1], adjusted=True, fft=True)[::-1]
+    bck_diag_cross_product /= np.max(abs(bck_diag_cross_product))
+    return np.r_[bck_diag_cross_product[:-1], fwr_diag_cross_product]
+
+q1_dat, q2_dat, q3_dat, q4_dat = [], [], [], []
+t_list = []
+
+fig, (ax1, ax2, ax3) = plt.subplots(nrows=1, ncols=3, sharey=True)
+plt.suptitle('Quadrant Cross Correlation')
+nlags = 50
+step_size = 1
+
+tau_ot1 = 3.75e-4
+tau_ot2 = 9e-4
+with open('./QPD_output/data_14-May-2024 13:58:16.txt', 'r') as input:
+    lines = input.readlines()
+    for line in lines[1:-1:step_size]:
+        #print(line)
+        items =  line.split(',')
+        q1_dat.append(float(items[0]))
+        q2_dat.append(float(items[1]))
+        q3_dat.append(float(items[2]))
+        q4_dat.append(float(items[3]))
+        t_list.append(float(items[4]))
+
+lag = t_list[2]-t_list[1]
+#print('dt = ' + str(lag), 'fs = ' + str(1/lag))
+
+#from statsmodels.graphics.tsaplots import plot_acf, acf
+
+#from test_signal_corr import get_ACF, get_acf
+
+pair1 = [(q1+q2)-(q3+q4) for q1,q2, q3, q4 in zip(q1_dat, q2_dat, q3_dat, q4_dat)]
+pair2 = [q1+q2 for q1,q2 in zip(q2_dat, q3_dat)]
+
+ccf_output_diag1 = make_ccf(q1_dat, q4_dat)
+ccf_output_diag2 = make_ccf(q2_dat, q3_dat)
+
[…]

### 2024-05-20 Latex_files/Manuscript Dimers/Manuscript_Optical Communications.tex
09:31 20-05-24

--- 2024-02-06 Latex files/Manuscript Dimers/Manuscript_Optical Communications.tex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
@@
 \begin{keyword}
-			Optical Trapping \sep Brownian Dynamics \sep Micro-Rotors  \sep Colloids 
+  Optical Trapping \sep Brownian Dynamics \sep Micro-Rotors \sep
+  Colloids
 \end{keyword}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%        
 \section{Introduction}
+
 Ever since their invention in 1986 by Arthur Ashkin \cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been used extensively in colloidal physics \cite{Leonardo_2008, Koehler_2011, Curran_1999}.
@@
 Chetana D et, al., \cite{Chetana_2022} increased the complexity of the problem by considering an object with both shape as well as optical anisotropy in the form of a chicken Red Blood Cell (cRBC) and showed that the trapping dynamics and the final equilibrium orientation of the cRBC was decided by the beam polarization with the cell showing a dual reorientation behaviour in a linearly polarized trap and a partial second reorientation in a circularly polarized trap.
+
 In this work, we extend the Brownian dynamical simulations carried out by Vigilante et.al., \cite{Vigilante_2020} to the case of asymmetric dimers.
@@
 We then reproduce the results for dynamics of a symmetric dimer in \cite{Vigilante_2020} as further validation of our codes.
-We organize our results into 3 sections with the focus on axial and orientation trapping equilibria for different size-ratios of the dimers in the first section, determinations of equilibrium orientations and equilibrium positions of various dimers with varied initial conditions of orientation and position in linearly and circularly polarised traps, Finally, we report interesting rotational dynamics of dimers seen only when the light is not linearly polarised.
+We organize our results into 3 sections with the focus on axial and orientation trapping equilibria for different size-ratios of the dimers in the first section, determinations of equilibrium orientations and equilibrium positions of various dimers with varied initial conditions of orientation and position in linearly and circularly polarised traps.
+Finally, we report interesting rotational dynamics of dimers seen only when the light is not linearly polarised.
 We hope that our results could inform studies on colloidal aggregation and help design rotors for use in applications such as lab-on-a-chip and micromachines.
 
-\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%        
+\section{Methodology}
 
-\section{Methodology}
 \paragraph{Simulating motion within an optical trap}
@@
 \end{align}
-
-Where $F(x)$ is the optical force directed towards the focal point of the laser However, this is only accurate when close to the equilibrium position, if we wish to probe the phase space we need a means of computing the optical forces for any position or orientation.
-Fortunately, Vigilante et al \cite{Vigilante_2020} have already built the ground work for simulating multi-sphere trapping.
+where $F(x)$ is the optical force directed towards the focal point of the laser However, this is only accurate when close to the equilibrium position, if we wish to probe the phase space we need a means of computing the optical forces for any position or orientation.
+Fortunately, Vigilante et~al.\ \cite{Vigilante_2020} have already built the ground work for simulating multi-sphere trapping.
 This work is an extension of the simulation work conducted by \cite{Vigilante_2020} into multi-sphere objects; their paper already has an excellent breakdown of the mathematics that go into calculating the optical and hydrodynamic forces, as such we will only refer to the mathematical methodology when necessary or when our work diverges from \cite{Vigilante_2020}.
@@
 \end{align}
-Where the constants are spline fitted from the results of [14], due to the dimer’s asymmetry the friction coefficient – which is related to the diffusion coefficient – is different depending on the axis the fluid is flowing parallel to.
+where the constants are spline fitted from the results of [14], due to the dimer’s asymmetry the friction coefficient – which is related to the diffusion coefficient – is different depending on the axis the fluid is flowing parallel to.
 The Brownian displacements and rotations are generated at random in proportion to the diffusion tensor.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%        
 \section{Results and Discussion}
+
 Trapping forces on a sphere are often written as the rate of change in momenta of incident and scattered beams \cite{Nieminen_2007}.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Axial forces and Equilibrium positions}
+
 We explore the trapping landscape by plotting out the trapping force as the dimer is moved along the z-axis and see that for the case of large-bead-over-small, there are multiple trapping equilibria compared to the case of small-bead-over-large as seen in Fig [4] and Fig [5].
@@
 [6] we plot out the force curve for a symmetric dimer with the size of each sphere equal to 1 $\mu m$.
+
 \begin{figure}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Torques and equilibrium orientations:}
+
 Having determined the axial-equilibrium positions of the dimers, we go about determining the steady-state orientations of the dimers assuming they’ve been trapped at their axial-equilibrium positions.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Rotation Dynamics}
+
 Dimers have been shown to rotate within a circularly polarised trap [] meaning they have the potential to be used in micro rotor applications.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%        
 \section{Conclusions and Outlook}
@@
 \newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%        
 \bibliography{dimer_bib} 
 \bibliographystyle{ieeetr}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2024-05-20 Latex_files/Manuscripts Scattering/Manuscript_Optica.tex
09:31 20-05-24

--- 2023-10-25 Latex files/Manuscripts Scattering/Manuscript_Optical Communications.tex
@@
-\documentclass[preprint,  3p]{elsarticle}
+\documentclass[12pt]{optica-article}
 
 \usepackage{csvsimple}
-\usepackage{amssymb, amsthm}
+\usepackage{lineno}
+
 \usepackage{amsmath}
@@
 \renewcommand{\nomlabel}[1]{\hfil #1\hfil}
+\linenumbers
 
-\journal{}
+\journal{opticajournal}
 
@@
 
-\begin{frontmatter}
+%\begin{frontmatter}
 
@@
 
-\author[aff1]{Daniel Maciver\corref{cor1}} 
-\ead{Daniel.Maciver.2016@uni.strath.ac.uk}
+\author{Daniel Maciver, \authormark{1,*}, Praveen Parthasarathi, \authormark{1}, Leo Lue, \authormark{1}, Jan Sefcik, \authormark{1},
+Mark Haw, \authormark{1}}
 
-\author[aff1]{Praveen Parthasarathi}
+\address{\authormark{1}Department of Chemical Engineering,
+	University of Strathclyde, 75 Montrose Street, Glasgow,
+	G1 1XL, Scotland}
+\email{\authormark{*}daniel.maciver.2016@uni.strath.ac.uk}
 
-\author[aff1]{Leo Lue}
-
-\author[aff1]{Jan Sefcik}
-
-\author[aff1]{Mark Haw}
-
-\cortext[cor1]{Corresponding author}
-\affiliation[aff1]{organization={Department of Chemical Engineering,
-			University of Strathclyde},
-            addressline={75 Montrose Street}, 
-            city={Glasgow},
-            postcode={G1 1XL}, 
-            country={Scotland}}
 \begin{abstract}
@@
 
-\begin{keyword}
-	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
-\end{keyword}
+%\begin{keyword}
+%	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
+%\end{keyword}
 
-\end{frontmatter}
+%\end{frontmatter}
 \newpage
@@
 		Reference Orientations vector components, for $n_{refs} = 30 \ ^*$}
-		\begin{tabularx}{0.65\textwidth}{
+		\begin{tabularx}{0.9\textwidth}{
 				| >{\centering\arraybackslash}X
@@
 			\\ \hline
-			\csvreader[
-			head to column names
-			]
-			{./data/n_list.csv}{}{
-			\index\ & \nx\ & \ny\ & \nz\ \\
-			} \\
+			$1$ & $0.2958759$ & $0.2958759$ & $0.9082483$ \\
+			$2$ & $0.9082483$ & $0.2958759$ & $0.2958759$ \\
+			$3$ & $0.2958759$ & $0.9082483$ & $0.2958759$ \\
+			$4$ & $0.2958759$ & $0.2958759$ & $-0.9082483$ \\
+			$5$ & $0.9082483$ & $0.2958759$ & $-0.2958759$ \\
+			$6$ & $0.2958759$ & $0.9082483$ & $-0.2958759$ \\
+			$7$ & $0.2958759$ & $-0.2958759$ & $0.9082483$ \\
+			$8$ & $0.9082483$ & $-0.2958759$ & $0.2958759$ \\
+			$9$ & $0.2958759$ & $-0.9082483$ & $0.2958759$ \\
+			$10$ & $0.2958759$ & $-0.2958759$ & $-0.9082483$ \\
+			$11$ & $0.9082483$ & $-0.2958759$ & $-0.2958759$ \\
+			$12$ & $0.2958759$ & $-0.9082483$ & $-0.2958759$ \\
+			$13$ & $-0.2958759$ & $0.2958759$ & $0.9082483$ \\
+			$14$ & $-0.9082483$ & $0.2958759$ & $0.2958759$ \\
+			$15$ & $-0.2958759$ & $0.9082483$ & $0.2958759$ \\
+			$16$ & $-0.2958759$ & $0.2958759$ & $-0.9082483$ \\
+			$17$ & $-0.9082483$ & $0.2958759$ & $-0.2958759$ \\
+			$18$ & $-0.2958759$ & $0.9082483$ & $-0.2958759$ \\
+			$19$ & $-0.2958759$ & $-0.2958759$ & $0.9082483$ \\
+			$20$ & $-0.9082483$ & $-0.2958759$ & $0.2958759$ \\
+			$21$ & $-0.2958759$ & $-0.9082483$ & $0.2958759$ \\
+			$22$ & $-0.2958759$ & $-0.2958759$ & $-0.9082483$ \\
+			$23$ & $-0.9082483$ & $-0.2958759$ & $-0.2958759$ \\
+			$24$ & $-0.2958759$ & $-0.9082483$ & $-0.2958759$ \\
+			$25$ & $1.0000$ & $0.0000$ & $0.0000$ \\
+			$26$ & $0.0000$ & $1.0000$ & $0.0000$ \\
+			$27$ & $0.0000$ & $0.0000$ & $1.0000$ \\
+			$28$ & $-1.000$ & $0.0000$ & $0.0000$ \\
+			$29$ & $0.0000$ & $-1.000$ & $0.0000$ \\
+			$30$ & $0.0000$ & $0.0000$ & $-1.000$ \\
+
 			\hline\hline 
@@
 		    \hline
-			\csvreader[
-			head to column names
-			]
-			{./data/y_list.csv}{}{
-				\index\ & \Ix\ & \Iy\ & \Iz\ & \yx\ & \yy\ & \yz\ \\
-			} \\
+			$1$ & $5.236437793$ & $0.008879799$ & $0.01023413$ & $-0.566323866$ & $-0.895169311$ & $-0.782655503$ \\
+			$2$ & $9.029762808$ & $0.014176754$ & $0.023474524$ & $1.604434643$ & $-0.737872411$ & $1.16224444$ \\
+			$3$ & $5.677784222$ & $0.018003042$ & $0.012268563$ & $-0.313760083$ & $-0.624248031$ & $-0.48381477$ \\
+			$4$ & $4.054681384$ & $0.008596164$ & $0.007681417$ & $-1.242592777$ & $-0.903592052$ & $-1.157627171$ \\
+			$5$ & $5.916429873$ & $0.012267124$ & $0.018806686$ & $-0.17719333$ & $-0.794580264$ & $0.47657917$ \\
+			$6$ & $7.154962253$ & $0.040816852$ & $0.007678353$ & $0.531566054$ & $0.053224459$ & $-1.158077234$ \\
+			$7$ & $4.857371303$ & $0.057897419$ & $0.009575087$ & $-0.78324749$ & $0.560444156$ & $-0.87946335$ \\
+			$8$ & $9.018457316$ & $0.061837715$ & $0.027314068$ & $1.597964991$ & $0.677454091$ & $1.726240514$ \\
+			$9$ & $5.001594138$ & $0.007138369$ & $0.009576248$ & $-0.700714897$ & $-0.946882332$ & $-0.879292708$ \\
+			$10$ & $4.069312021$ & $0.041444603$ & $0.012267124$ & $-1.234220286$ & $0.071865963$ & $-0.484026249$ \\
+			$11$ & $6.542222096$ & $0.050631978$ & $0.023474524$ & $0.180920933$ & $0.34469166$ & $1.16224444$ \\
+			$12$ & $7.930714067$ & $0.10058846$ & $0.010228239$ & $0.975495811$ & $1.828185342$ & $-0.78352083$ \\
+			$13$ & $5.236437793$ & $0.008879799$ & $0.01023413$ & $-0.566323866$ & $-0.895169311$ & $-0.782655503$ \\
+			$14$ & $9.029762808$ & $0.014176754$ & $0.023474524$ & $1.604434643$ & $-0.737872411$ & $1.16224444$ \\
+			$15$ & $5.677784222$ & $0.018003042$ & $0.012268563$ & $-0.313760083$ & $-0.624248031$ & $-0.48381477$ \\
+			$16$ & $4.054681384$ & $0.008596164$ & $0.007681417$ & $-1.242592777$ & $-0.903592052$ & $-1.157627171$ \\
+			$17$ & $5.916429873$ & $0.012267124$ & $0.018806686$ & $-0.17719333$ & $-0.794580264$ & $0.47657917$ \\
+			$18$ & $7.154962253$ & $0.040816852$ & $0.007678353$ & $0.531566054$ & $0.053224459$ & $-1.158077234$ \\
+			$19$ & $4.857371303$ & $0.057897419$ & $0.009575087$ & $-0.78324749$ & $0.560444156$ & $-0.87946335$ \\
+			$20$ & $9.018457316$ & $0.061837715$ & $0.027314068$ & $1.597964991$ & $0.677454091$ & $1.726240514$ \\
+			$21$ & $5.001594138$ & $0.007138369$ & $0.009576248$ & $-0.700714897$ & $-0.946882332$ & $-0.879292708$ \\
+			$22$ & $4.069312021$ & $0.041444603$ & $0.012267124$ & $-1.234220286$ & $0.071865963$ & $-0.484026249$ \\
+			$23$ & $6.542222096$ & $0.050631978$ & $0.023474524$ & $0.180920933$ & $0.34469166$ & $1.16224444$ \\
+			$24$ & $7.930714067$ & $0.10058846$ & $0.010228239$ & $0.975495811$ & $1.828185342$ & $-0.78352083$ \\
+			$25$ & $8.589197415$ & $0.039227387$ & $0.024433841$ & $1.352317814$ & $0.006024147$ & $1.303159911$ \\
+			$26$ & $7.179988381$ & $0.037734463$ & $0.014278111$ & $0.545887442$ & $-0.038309302$ & $-0.188629493$ \\
+			$27$ & $4.518783843$ & $0.045969647$ & $0.022162922$ & $-0.977006687$ & $0.206240387$ & $0.9695815$ \\
+			$28$ & $8.589197415$ & $0.039227387$ & $0.024433841$ & $1.352317814$ & $0.006024147$ & $1.303159911$ \\
+			$29$ & $4.635190763$ & $0.153655662$ & $0.02212562$ & $-0.910391959$ & $3.404054059$ & $0.964102078$ \\
+			$30$ & $4.290255329$ & $0.010364637$ & $0.014275348$ & $-1.107783832$ & $-0.851075979$ & $-0.189035408$ \\
 			\hline\hline 

### 2024-05-20 Latex_files/Manuscripts Scattering/Manuscript_Optical Communications.tex
09:31 20-05-24

--- 2024-05-20 Latex_files/Manuscripts Scattering/Manuscript_Optica.tex
@@
-\documentclass[12pt]{optica-article}
+\documentclass[preprint,  3p]{elsarticle}
 
 \usepackage{csvsimple}
-\usepackage{lineno}
-
+\usepackage{amssymb, amsthm}
 \usepackage{amsmath}
@@
 \renewcommand{\nomlabel}[1]{\hfil #1\hfil}
-\linenumbers
 
-\journal{opticajournal}
+\journal{}
 
@@
 
-%\begin{frontmatter}
+\begin{frontmatter}
 
@@
 
-\author{Daniel Maciver, \authormark{1,*}, Praveen Parthasarathi, \authormark{1}, Leo Lue, \authormark{1}, Jan Sefcik, \authormark{1},
-Mark Haw, \authormark{1}}
+\author[aff1]{Daniel Maciver\corref{cor1}} 
+\ead{Daniel.Maciver.2016@uni.strath.ac.uk}
 
-\address{\authormark{1}Department of Chemical Engineering,
-	University of Strathclyde, 75 Montrose Street, Glasgow,
-	G1 1XL, Scotland}
-\email{\authormark{*}daniel.maciver.2016@uni.strath.ac.uk}
+\author[aff1]{Praveen Parthasarathi}
 
+\author[aff1]{Leo Lue}
+
+\author[aff1]{Jan Sefcik}
+
+\author[aff1]{Mark Haw}
+
+\cortext[cor1]{Corresponding author}
+\affiliation[aff1]{organization={Department of Chemical Engineering,
+			University of Strathclyde},
+            addressline={75 Montrose Street}, 
+            city={Glasgow},
+            postcode={G1 1XL}, 
+            country={Scotland}}
 \begin{abstract}
@@
 
-%\begin{keyword}
-%	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
-%\end{keyword}
+\begin{keyword}
+	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
+\end{keyword}
 
-%\end{frontmatter}
+\end{frontmatter}
 \newpage
@@
 Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle.
-Shown below is two clusters of orientation vectors and there respective measured scattering signals - the points have been coloured based on their proximity to the centre of their respective cluster.
+Fig~\ref{fig:motion} shows the mapping between two clusters of orientation vectors and there respective measured scattering signals, the colouring represents the distance in orientation space from the centre of each cluster.
 While the orientation space appears tightly packed and ordered the signal space quickly spreads out in an asymmetric fashion.
@@
 	\end{subfigure}
-	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals.
+	\caption{Mapping between (a) orientation vectors and (b) their respective scattering signals.
 Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Calculation of error}
+\subsection{Evaluating model performance}
 \label{sec:divergence}
@@
 For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.
+Typically for an inverted microscope set up the most obvious plane to place your fibres is in the x-y plane as it is both easier to align with the trapping beam, and to measure the exact angle from the centre of the beam to each fibre.
+Ideally with a given number of fibres it should be possible to produce a unique signal at each point in time.
 
@@
 This is shown in Figure~\ref{fig:detectors} which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels is because the dimer's scattering is indistinguishable in these two reference orientation.
-It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors.
+It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors, if the fibres were instead placed in say the z-x plane we would see a similar problem with vertical reference orientations.
 More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
@@
 By adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
+The difficulty with placing fibres out of the x-y plane is ensuring the fibre remains at a fixed angle throughout the measurement, and how to actually measure the position of the out of plane fibre accurately.
 In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.
@@
 We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer.
-In the future we hope to further investigate the motion of asymmetric dimers.
+In the future we hope to perform a more exhaustive discussion of the behaviour dimers demonstrate within an optical trap, accounting for factors such as polarisation, size, orientation, position, and so on.
+For now we are more concerned with our model's performance in determining the orientation for an example trajectory, one who's motion is not easily understood based on current literature.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{align}
+	\label{eq:var}
 	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
@@
 where $\epsilon$ is the percentage error associated with the scattering signal.
-Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:
+This assumption that the signal noise is a simple percent error of our collected signal only holds true if we can be sure that the variation across the entire surface of our detector is smaller than that.
+The variation over for a detector k located by angles $\theta_k$ and $\phi_k$ is given by:
+
+\begin{align}
+	\sigma_k = {\bar{I}_k} \ ^{-1}(I(\theta_k, \phi_k)-\bar{I}_k)
+\end{align}
+
+We know that as we move further away from the centre of the trap this variation decreases with distance.
+We can use mstm to compute the variation over the surface of our detector as a function of distance for each detector used.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.75\textwidth]{./Images/Detector vs distance_modified.png}
+	\caption{$\%$ Variation over the surface of our 4 detectors as a function of their distance from the centre of the trap; legend shows the spherical coordinates of each detector ($\theta$ = azimuthal angle, $\phi$ = polar angle).
+Dotted lines mark 15 $\%$ and 1 $\%$ which correspond to signal error used below.}
+	\label{fig:detector}
+\end{figure}
+As shown in Fig~\ref{fig:detector} the variation over the detector drops off with distance, the parrallel nature of the curves indicates that the variance accross the detector follows an inverse square law.
+Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$.
+We moved each detector back to a distance of 500 $\mu m$ so that Eq~\ref{eq:var} holds true for a $1 \%$ signal noise.
 

### 2024-05-20 Latex_files/Optica/Optica-template.tex
09:31 20-05-24

\documentclass{optica-article}

\journal{opticajournal} % for journals or Optica Open

\articletype{Research Article}

\usepackage{lineno}
\linenumbers % Turn off line numbering for Optica Open preprint submissions.

\begin{document}

\title{Universal manuscript template for Optica Publishing Group journals}

\author{Author One,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}

\address{\authormark{1}Peer Review, Publications Department, Optica Publishing Group, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optica Publishing Group, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}Currently with the Department of Electronic Journals, Optica Publishing Group, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}

\email{\authormark{*}opex@optica.org} %% email address is required; see note below about the corresponding author designation

% use {asbstract*} to suppress the copyright line. Copyright information will be added in production

\begin{abstract*} 
\LaTeX{} manuscripts submitted to Optica Publishing Group journals may use these instructions and this universal template format. The template simplifies manuscript preparation and eases transfer between journals. \emph{Applied Optics}, JOSA A, JOSA B, \emph{Journal of Optical Communications and Networking}, \emph{Optica Quantum}, and \emph{Photonics Research} authors should use the length-check template if a precise length estimate is needed. \emph{Optics Letters} authors and authors of short \emph{Optica} articles are encouraged to use the length-check template. Authors using this universal template will still need to adhere to article-length restrictions based on the final, published format. Authors may also \href{https://opticaopen.org}{submit articles} prepared using this template to the Optica Publishing Group preprint server, \href{https://preprints.opticaopen.org}{Optica Open}. However, doing so is optional. Please refer to the submission guidelines found there. Note that copyright and licensing information should no longer be added to your Journal or Optica Open manuscript.

\end{abstract*}

%%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Adherence to the specifications listed in this template is essential for efficient review and publication of submissions. Proper reference format is especially important (see Section \ref{sec:refs}). For submissions to the Optica Open preprint server, many of the style and format guidelines described in this template are not applicable, although adhering to the instructions in this template will ease the process to convert a preprint to an Optica Publishing Group journal submission.

\section{Corresponding author}

We require manuscripts to identify a single corresponding author. The corresponding author typically is the person who submits the manuscript and handles correspondence throughout the peer review and publication process.

\begin{verbatim}
\author{Author One\authormark{1} and Author Two\authormark{2,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optica Publishing Group, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optica Publishing Group,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\

\email{\authormark{*}xyz@optica.org}
\end{verbatim}

This format will generate the following appearance:

\bigskip

\author{Author One\authormark{1} and Author Two\authormark{2,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optica Publishing Group, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optica Publishing Group,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
%\authormark{3}xyz@optica.org}

\email{\authormark{*}xyz@optica.org}}

\medskip

If other statements about author contribution and contact are needed, they can be added in addition to the corresponding author designation.
\begin{verbatim}
\author{Author One\authormark{1,$\dag$} and Author Two\authormark{2,$\dag$,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optica Publishing Group, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optica Publishing Group,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{$\dag$}The authors contributed equally to this work.\\
\authormark{*}xyz@optica.org}}
\end{verbatim}

This format will generate the following appearance:

\medskip

\author{Author One\authormark{1,\dag} and Author Two\authormark{2,\dag,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optica Publishing Group, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optica Publishing Group, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{\dag}The authors contributed equally to this work.\\
\authormark{*}xyz@optica.org}
\medskip

\section{Abstract}
The abstract should be limited to approximately 100 words. If the work of another author is cited in the abstract, that citation should be written out without a number, (e.g., journal, volume, first page, and year in square brackets [Opt. Express {\bfseries 22}, 1234 (2014)]), and a separate citation should be included in the body of the text. The first reference cited in the main text must be [1]. Do not include numbers, bullets, or lists inside the abstract. Note that no licensing or copyright statement should appear on the manuscript itself. Such information will be added to the Accepted Manuscript, the published version, or the Optica Open preprint as part of production to ensure accuracy and consistency of appearance.

\begin{figure}[ht!]
\centering\includegraphics[width=7cm]{opticafig1}
\caption{Sample caption (Fig. 2, \cite{Yelin:03}).}
\end{figure}

\section{Assessing final manuscript length}
The Universal Manuscript Template is based on the Express journal layout and will provide an accurate length estimate for \emph{Optics Express}, \emph{Biomedical Optics Express},  \emph{Optical Materials Express}, \emph{Optica Quantum}, and \emph{Optics Continuum}. \emph{Applied Optics}, JOSAA, JOSAB, \emph{Optics Letters}, \emph{Optica}, \emph{Optica Quantum}, \emph{Optics Continuum}, and \emph{Photonics Research} publish articles in a two-column layout. To estimate the final page count in a two-column layout, multiply the manuscript page count (in increments of 1/4 page) by 60\%. For example, 11.5 pages in the Universal Manuscript Template are roughly equivalent to 7 composed two-column pages. Note that the estimate is only an approximation, as treatment of figure sizing, equation display, and other aspects can vary greatly across manuscripts. Authors of Letters may use the legacy template for a more accurate length estimate.

\section{Figures, tables, and supplementary materials}

\subsection{Figures and tables}
Figures and tables should be placed in the body of the manuscript. Standard \LaTeX{} environments should be used to place tables and figures:
\begin{verbatim}
\begin{figure}[htbp]
\centering\includegraphics[width=7cm]{opticafig1}
\caption{Sample caption (Fig. 2, \cite{Yelin:03}).}
\end{figure}
\end{verbatim}

\subsection{Supplementary materials in Optica Publishing Group journals}
Our journals allow authors to include supplementary materials as integral parts of a manuscript. Such materials are subject to peer-review procedures along with the rest of the paper and should be uploaded and described using our Prism manuscript system. Please refer to the \href{https://opg.optica.org/submit/style/supplementary_materials.cfm}{Author Guidelines for Supplementary Materials in Optica Publishing Group Journals} for more detailed instructions on labeling supplementary materials and your manuscript. For preprints submitted to Optica Open a link to supplemental material should be included in the submission, but do not upload the material.

\textbf{Authors may also include Supplemental Documents} (PDF documents with expanded descriptions or methods) with the primary manuscript. At this time, supplemental PDF files are not accepted for partner titles, JOCN and \emph{Photonics Research}. To reference the supplementary document, the statement ``See Supplement 1 for supporting content.'' should appear at the bottom of the manuscript (above the References heading). Supplemental documents are not accepted for Optica Open preprints.

\subsection{Sample Dataset Citation}

1. M. Partridge, "Spectra evolution during coating," figshare (2014), http://dx.doi.org/10.6084/m9.figshare.1004612.

\subsection{Sample Code Citation}

2. C. Rivers, "Epipy: Python tools for epidemiology," figshare (2014) [retrieved 13 May 2015], http://dx.doi.org/10.6084/m9.figshare.1005064.

\section{Mathematical and scientific notation}

\subsection{Displayed equations} Displayed equations should be centered.
Equation numbers should appear at the right-hand margin, in
parentheses:
\begin{equation}
J(\rho) =
 \frac{\gamma^2}{2} \; \sum_{k({\rm even}) = -\infty}^{\infty}
	\frac{(1 + k \tau)}{ \left[ (1 + k \tau)^2 + (\gamma  \rho)^2  \right]^{3/2} }.
\end{equation}

All equations should be numbered in the order in which they appear
and should be referenced  from within the main text as Eq. (1),
Eq. (2), and so on [or as inequality (1), etc., as appropriate].

\section{Backmatter}

Backmatter sections should be listed in the order Funding/Acknowledgment/Disclosures/Data Availability Statement/Supplemental Document section. An example of backmatter with each of these sections included is shown below.

\begin{backmatter}
\bmsection{Funding}
Content in the funding section will be generated entirely from details submitted to Prism. Authors may add placeholder text in the manuscript to assess length, but any text added to this section in the manuscript will be replaced during production and will display official funder names along with any grant numbers provided. If additional details about a funder are required, they may be added to the Acknowledgments, even if this duplicates information in the funding section. See the example below in Acknowledgements. For preprint submissions, please include funder names and grant numbers in the manuscript.

\bmsection{Acknowledgments}
The section title should not follow the numbering scheme of the body of the paper. Additional information crediting individuals who contributed to the work being reported, clarifying who received funding from a particular source, or other information that does not fit the criteria for the funding block may also be included; for example, ``K. Flockhart thanks the National Science Foundation for help identifying collaborators for this work.''

\bmsection{Disclosures}
Disclosures should be listed in a separate nonnumbered section at the end of the manuscript. List the Disclosures codes identified on the \href{https://opg.optica.org/submit/review/conflicts-interest-policy.cfm}{Conflict of Interest policy page}, as shown in the examples below:

\medskip

\noindent ABC: 123 Corporation (I,E,P), DEF: 456 Corporation (R,S). GHI: 789 Corporation (C).

\medskip

\noindent If there are no disclosures, then list ``The authors declare no conflicts of interest.''

\bmsection{Data Availability Statement}
A Data Availability Statement (DAS) will be required for all submissions beginning 1 March 2021. The DAS should be an unnumbered separate section titled ``Data Availability'' that
immediately follows the Disclosures section. See the \href{https://opg.optica.org/submit/review/data-availability-policy.cfm}{Data Availability Statement policy page} for more information.

Optica has identified four common (sometimes overlapping) situations that authors should use as guidance. These are provided as minimal models, and authors should feel free to
include any additional details that may be relevant.

\begin{enumerate}
\item When datasets are included as integral supplementary material in the paper, they must be declared (e.g., as "Dataset 1" following our current supplementary materials policy) and cited in the DAS, and should appear in the references.

\bmsection{Data availability} Data underlying the results presented in this paper are available in Dataset 1, Ref. [3].

\bigskip

\item When datasets are cited but not submitted as integral supplementary material, they must be cited in the DAS and should appear in the references.

\bmsection{Data availability} Data underlying the results presented in this paper are available in Ref. [3].

\bigskip

\item If the data generated or analyzed as part of the research are not publicly available, that should be stated. Authors are encouraged to explain why (e.g.~the data may be restricted for privacy reasons), and how the data might be obtained or accessed in the future.

\bmsection{Data availability} Data underlying the results presented in this paper are not publicly available at this time but may be obtained from the authors upon reasonable request.

\bigskip

\item If no data were generated or analyzed in the presented research, that should be stated.

\bmsection{Data availability} No data were generated or analyzed in the presented research.
\end{enumerate}

\bigskip

\noindent Data availability statements are not required for preprint submissions.

\bmsection{Supplemental document}
See Supplement 1 for supporting content.

\end{backmatter}

\section{References}
\label{sec:refs}
Proper formatting of references is extremely important, not only for consistent appearance but also for accurate electronic tagging. Please follow the guidelines provided below on formatting, callouts, and use of Bib\TeX.

\subsection{Formatting reference items}
Each source must have its own reference number. Footnotes (notes at the bottom of text pages) are not used in our journals. List up to three authors, and if there are more than three use \emph{et al.} after that. Examples of common reference types can be found in the  \href{https://opg.optica.org/jot/submit/style/oestyleguide.cfm} {style guide}.

The commands \verb+\begin{thebibliography}{}+ and \verb+\end{thebibliography}+ format the section according to standard style, showing the title {\bfseries References}.  Use the \verb+\bibitem{label}+ command to start each reference.

\subsection{Formatting reference citations}
References should be numbered consecutively in the order in which they are referenced in the body of the paper. Set reference callouts with standard \verb+\cite{}+ command or set manually inside square brackets [1].

To reference multiple articles at once, simply use a comma to separate the reference labels, e.g. \verb+\cite{Yelin:03,Masajada:13,Zhang:14}+, produces \cite{Yelin:03,Masajada:13,Zhang:14}.
%Using the \texttt{cite.sty} package will make these citations appear like so: [2--4].

\subsection{Bib\TeX}
\label{sec:bibtex}
Bib\TeX{} may be used to create a file containing the references, whose contents (i.e., contents of \texttt{.bbl} file) can then be pasted into the bibliography section of the \texttt{.tex} file. A Bib\TeX{} style file, \texttt{opticajnl.bst}, is provided.

If your manuscript already contains a manually formatted \verb|\begin{thebibliography}|... \verb|\end{thebibliography}| list, then delete the \texttt{latexmkrc} file (if present) from your submission files. However you should ensure that your manually-formatted reference list adheres to style accurately.

\section{Conclusion}
After proofreading the manuscript, compress your .tex manuscript file and all figures (which should be in EPS or PDF format) in a ZIP, TAR or TAR-GZIP package. All files must be referenced at the root level (e.g., file \texttt{figure-1.eps}, not \texttt{/myfigs/figure-1.eps}). If there are supplementary materials, the associated files should not be included in your manuscript archive but be uploaded separately through the Prism interface.

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%

Add references with BibTeX or manually.
\cite{Zhang:14,OPTICA,FORSTER2007,Dean2006,testthesis,Yelin:03,Masajada:13,codeexample}

%%%%%%%%%% If using BibTeX:
\bibliography{sample}

%%%%%%%%%% If preparing manually:
% \begin{thebibliography}{1}
% \newcommand{\enquote}[1]{``#1''}

% \bibitem{Zhang:14}
% Y.~Zhang, S.~Qiao, L.~Sun, Q.~W. Shi, W.~Huang, L.~Li, and Z.~Yang,
%   \enquote{Photoinduced active terahertz metamaterials with nanostructured
%   vanadium dioxide film deposited by sol-gel method,}
%   {\protect\JournalTitle{Optics Express}} \textbf{22}, 11070--11078 (2014).

% \bibitem{Optica}
% {Optica}, \enquote{{Optica Publishing Group},}
%   \url{http://www.opg.optica.org}.

% \bibitem{FORSTER2007}
% P.~Forster, V.~Ramaswamy, P.~Artaxo, T.~Bernsten, R.~Betts, D.~Fahey,
%   J.~Haywood, J.~Lean, D.~Lowe, G.~Myhre, J.~Nganga, R.~Prinn, G.~Raga,
%   M.~Schulz, and R.~V. Dorland, \enquote{Changes in atmospheric consituents and
%   in radiative forcing,} in \enquote{Climate Change 2007: The Physical Science
%   Basis. Contribution of Working Group 1 to the Fourth assesment report of
%   Intergovernmental Panel on Climate Change,}  S.~Solomon, D.~Qin, M.~Manning,
%   Z.~Chen, M.~Marquis, K.~B. Averyt, M.~Tignor, and H.~L. Miler, eds.
%   (Cambridge University Press, 2007).

% \end{thebibliography}

\end{document}

### 2024-05-20 Latex_files/Spie/article.tex
09:31 20-05-24

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tocloft}

\title{SPIE journal papers: sample manuscript showing style and formatting specifications}

\author[a]{First Author}
\author[a]{Second Author}
\author[b]{Third Author}
\author[a,b,*]{Fourth Author}
\affil[a]{University Name, Faculty Group, Department, Street Address, City, Country, Postal Code}
\affil[b]{Company Name, Street Address, City, Country, Postal Code}

\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table} 
\begin{document} 
\maketitle

\begin{abstract}
This document shows the required format and appearance of a manuscript prepared for SPIE journals. It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the following journals require the use of structured abstracts in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether structured or single-paragraph, the abstract should be a summary of the paper and not an introduction. Because the abstract may be used in abstracting and indexing databases, it should be self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the objectives, methodology used, results obtained, and their significance. A list of up to six keywords should immediately follow. 
\end{abstract}

% Include a list of up to six keywords after the abstract
\keywords{optics, photonics, light, lasers, journal manuscripts, LaTeX template}

% Include email contact information for corresponding author
{\noindent \footnotesize\textbf{*}Fourth author name,  \linkable{myemail@university.edu} }

\begin{spacing}{2}   % use double spacing for rest of manuscript

\section{Introduction}
\label{sect:intro}  % \label{} allows reference to this section
This document shows the format and appearance of a manuscript prepared for submission to SPIE journals. Note that this template is only intended to be used as a guideline for author convenience. It is designed for optimum clarity and ease of reading for editors and reviewers, but the template does not reflect the final page layout of a published journal paper. Accepted papers are professionally typeset in XML according to the layout and design of the journal.

\subsection{Use of This Document}

This document is prepared using LaTeX2e\cite{Lamport94,Goossens97} with the class file {\ttfamily spieman.cls}.  The LaTeX source file used to create this document is {\ttfamily article.tex}, which contains important formatting information embedded in it. Authors may use it as a template to create their own manuscript. While LaTeX properly handles most formatting issues, the author may occasionally need to intervene to obtain a satisfactorily  formatted manuscript.

\subsection{English}

Authors are strongly encouraged to follow the principles of sound technical writing, as found in Refs.~\citenum{Alred03} and \citenum{Perelman97}, for example. In addition, good English usage is essential. Authors whose native language is not English may wish to collaborate with a colleague whose English skills are more advanced. Alternatively, you may wish to have your manuscript professionally edited prior to submission by Editage, our recommended independent editorial service: \linkable{https://www.editage.com/spie/}. SPIE authors will receive a 15\% discount off their services. A spell checker can be helpful to discover misspelled words, but authors should also proofread their papers carefully prior to submission. Manuscripts that do not meet acceptable English standards or lack clarity may be rejected.

\subsection{Page Setup and Fonts}

All text and figures, including footnotes, must fit inside a text area 6.5 in.\ wide by 9 in.\ high (16.51 by 22.86 cm). Manuscripts must be formatted for US letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1 in.\ on the bottom, and 1 in.\ on the left and right.

The Times New Roman font is used throughout the manuscript, in the sizes and styles shown in Table~\ref{tab:fonts}. If this font is not available, use a similar serif font. The manuscript should not contain headers or footers. Pages should be numbered.

\begin{table}[ht]
\caption{Fonts sizes and styles.} 
\label{tab:fonts}
\begin{center}       
\begin{tabular}{|l|l|} %% this creates two columns
%% |l|l| to left justify each column entry
%% |c|c| to center each column entry
%% use of \rule[]{}{} below opens up each row
\hline
\rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
\hline\hline
\rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
\hline 
\end{tabular}
\end{center}
\end{table}

\section{Parts of Manuscript}

This section describes the normal structure of a manuscript and how each part should be handled. The appropriate vertical spacing between various parts of this document is achieved in LaTeX through the proper use of defined constructs, such as \verb|\section{}|.

\subsection{Title and Author Information}
\label{sect:title}
The article title appears left justified at the top of the first page. The title font is 16 pt., bold. The rules for capitalizing the title are the same as for sentences; only the first word, proper nouns, and acronyms should be capitalized. Do not begin titles with articles (for example, a, an, the) or prepositions (for example, on, by, etc.). The word ``novel'' should not appear in the title, as publication will imply novelty. Avoid the use of acronyms in the title, unless they are widely understood.

The list of authors immediately follows the title, 18 points below. The font is 12 pt., bold and the author names are left justified. The author affiliations and addresses follow the names, in 10-pt., normal font and left justified. For multiple affiliations, each affiliation should appear on a separate line. Superscript letters (a, b, c, etc.) should be used to associate multiple authors with their respective affiliations. The corresponding author should be identified with an asterisk, and that person's email address should be provided below the keywords.

\subsection{Abstract}
The abstract should be a summary of the paper and not an introduction. Because the abstract may be used in abstracting journals, it should be self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the objectives, methodology used, results obtained, and their significance. Please note that the following journals require the use of structured abstracts in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on the website of the journal.

\subsection{Subject terms/Keywords}
Keywords are required. Please provide 3-6 keywords related to your paper.

\subsection{Body of Paper}
The body of the paper consists of numbered sections that present the main findings. These sections should be organized to best present the material.

To provide transition elements in your paper, it is important to refer back (or forward) to specific sections. Such references are made by indicating the section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1 contained a description...'' If the word Section, Reference, Equation, or Figure starts a sentence, it is spelled out. When occurring in the middle of a sentence, these words are abbreviated Sec., Ref., Eq., and Fig.

At the first occurrence of an acronym, spell it out followed by the acronym in parentheses, for example, charge-coupled diode (CCD).

\subsection{Footnotes}
Textual footnotes should be used rarely to present important documentary or explanatory material whose inclusion in the text would be distracting.\footnote{Example of a footnote.} Due to problems with HTML display, use of footnotes should generally be avoided. If absolutely necessary, the footnote mark must come at the end of a sentence. To insert a footnote, use the {\verb|\footnote{}|} command.

\subsection{Appendices}
Brief appendices may be included when necessary, such as derivations of equations, proofs of theorems, and details of algorithms. Equations and figures appearing in appendices should continue sequential numbering from earlier in the paper.

\subsection{Disclosures}
Conflicts of interest should be declared under a separate header. If the authors have no competing interests to declare, then a statement should be included declaring no conflicts of interest. For assistance generating a disclosure statement, see the form available from  the International Committee of Medical Journal Editors website: \linkable{http://www.icmje.org/conflicts-of-interest/}

\subsection{Code, Data, and Materials Availability}
In support of open scientific exchange, SPIE journals require Data and Code Availability Statements in all accepted papers. This requirement went into effect on 1 May 2023. These statements should describe how to access any data that would be required to replicate or interpret the findings reported in the paper.

\subsection{Acknowledgments}
Acknowledgments and funding information should be added after the conclusion, and before references. Include grant numbers and the full name of the funding body. The acknowledgments section does not have a section number.

\subsection{References}
The References section lists books, articles, and reports that are cited in the paper. This section does not have a section number. The references are numbered in the order in which they are cited. Examples of the format to be followed are given at the end of this document.

The reference list at the end of this document is created using BibTeX, which looks through the file {\ttfamily report.bib} for the entries cited in the LaTeX source file.  The format of the reference list is determined by the bibliography style file {\ttfamily spiejour.bst}, as specified in the \\ \verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may be directly formatted in the LaTeX source file.

For books\cite{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.  Journal-article references \cite{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.  A reference to a proceedings paper or a chapter in an edited book\cite{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.  References to an article in the SPIE Proceedings may include the conference name, as shown in Ref.~\citenum{Hanson93c}.

The references are numbered in the order of their first citation. Citations to the references are made using superscripts, as demonstrated in the preceding paragraph. One may also directly refer to a reference within the text, for example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more references should be separated by a comma with no space between them. Multiple sequential references should be displayed with a dash between the first and last numbers \cite{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.

\subsubsection{Reference linking and DOIs}
A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a digital object, such as a journal article or a book chapter, that provides a persistent link to its location on the internet. The use of DOIs allows readers to easily access cited articles. Authors should include the DOI at the end of each reference in brackets if a DOI is available. See examples at the end of this manuscript. A free DOI lookup service is available from CrossRef at \\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will facilitate reference linking and is highly recommended.

In the present LaTeX template, the author needs to add the DOI reference by including it in a ``note'' in the bibliography file, as shown in the file {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}. The DOI may be used by the reader to locate that document with the link: {\verb+http://dx.doi.org10.1117/12.154577+}.

\subsection{Biographies}
A brief professional biography of approximately 75 words may be provided for each author, if available. Biographies should be placed at the end of the paper, after the references. Personal information such as hobbies or birthplace/birthdate should not be included. Author photographs are not published.

\section{Section Formatting}
\label{sect:sections}
In LaTeX, a new section is created with the \verb|\section{}| command, which automatically numbers the sections. Sections will be numbered sequentially, starting with the first section after the abstract, except for the acknowledgments and references. (Note that numbering of section headings is not required, but the numbering must be consistent if used.) All section headings should be left justified.

Main section headings are in 12-pt. bold font, left-justified and in title case, where important words are capitalized.

Paragraphs that immediately follow a section heading are leading paragraphs and should not be indented, according to standard publishing style. The same goes for leading paragraphs of subsections and sub-subsections. Subsequent paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no additional space between paragraphs. In LaTeX, paragraphs are separated by blank lines in the source file. Indentation of the first line of a paragraph may be avoided by starting it with \verb|\noindent|.

\subsection{Subsection Headings}
All important words in a subsection (level 1) header are capitalized. Subsection numbers consist of the section number, followed by a period, and the subsection number within that section, without a period at the end. The heading is left justified and its font is 12 pt. italic.

\subsubsection{Sub-subsection headings}
The first word of a sub-subsection is capitalized. The rest of the text is not capitalized, except for proper names and acronyms (the latter should only be used if well known). The heading is left justified and its font is 11 pt. italic.

\section{Figures and Tables}

\subsection{Figures}

Figures are numbered in the order in which they are called out in the text. They should appear in the document in numerical order and as close as possible to their first reference in the text. It may be necessary to move figures or tables around to enhance readability. LaTeX will attempt to place figures at the top or bottom of a page in which they are first referenced.

Figures, along with their captions, should be separated from the main text by  0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or graph. Figure captions start with the abbreviation ``Fig'' in front of the figure number, followed by a period, and the text in 10-pt. font. See Fig.~\ref{fig:example} for an example.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[height=5.5cm]{mcr3b.eps}
\end{tabular}
\end{center}
\caption 
{ \label{fig:example}
Example of a figure caption. } 
\end{figure}

Authors may wish to create figures consisting of two or more images, in which case, they should be neatly arranged in a rectangular array.  In no case, should the article's text be wrapped around a figure. Figure~\ref{fig:example2} shows two side-by-side images. When a figure contains more than one image, the author must submit them as a single image file. Further details about figure formatting can be found in the author guidelines for each specific SPIE journal: \\
\linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images 
\\
(a) \hspace{5.1cm} (b)
\end{tabular}
\end{center}
\caption 
{ \label{fig:example2}
Example of a figure containing multiple images: (a) sun and (b) blob. Figures containing multiple images must be submitted to SPIE as a single image file.} 
\end{figure}

\subsection{Tables}
Tables are numbered in the order in which they are referenced. They should appear in the document in numerical order and as close as possible to their first reference in the text. It is preferable to have tables appear at the top or bottom of the page, if possible. Table captions are handled identically to those for figures, except that they appear above the table. See Table~\ref{tab:fonts} for an example.

\subsection{Video}
Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player. The recommended maximum size for each video file is 10-12 MB. Authors may insert a representative still image from the video file in the manuscript as a figure. The caption label will be linked by the publisher to the actual video file. The video may also be mentioned in an existing figure caption. Multimedia files are treated in the same manner as figures and they will be numbered sequentially with normal figures.  The video number, file type, and file size should be included in parentheses at the end of the figure caption. See Figure \ref{vid:satellite} for an example.

\begin{video}
\begin{center}
{\includegraphics[height=5cm]{satellite.eps}}
\\
\end{center}
\caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
\end{video}

\appendix    % this command starts appendixes

\section{Miscellaneous Formatting Details}
\label{sect:misc}
At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.

\subsection{Formatting Equations}
Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.  Important equations appear on their own line.  Such equations are centered.  For example, ``The expression for the field of view is
\begin{equation}
\label{eq:fov}
2 a = \frac{(b + 1)}{3c} \, ,
\end{equation}
where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.

Equations are considered to be part of a sentence and should be punctuated accordingly. In the above example, a comma appears after the equation because the next line is a subordinate clause. If the equation ends the sentence, a period should follow the equation. The line following an equation should not be indented unless it is meant to start a new paragraph. Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.

References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise. Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..

\subsection{Formatting Theorems}

To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.  The text of the theorem continues on the same line in normal, 10-pt. font, achieved in LaTeX using \verb|\footnotesize|.  For example,

\vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}

% \disclosures 
\subsection*{Disclosures}
Conflicts of interest should be declared under a "Disclosures" header. If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.

\subsection* {Code, Data, and Materials Availability} 
In support of open scientific exchange, SPIE journals require Code, Data, and Materials Availability Statements in all accepted papers. This requirement went into effect on 1 May 2023. These statements should describe how to access any data that would be required to replicate or interpret the findings reported in the paper. Authors are encouraged to make the data and code related to the manuscript publicly available whenever possible, and utilize repositories that are well-known to the field (FigShare, Github, CodeOcean, etc.). If the data or code cannot be made publicly available, the authors should state the reason and explain how it can be obtained. Likewise, if data sharing is not applicable, the statement must say so. Example statements may be found in the Author Guidelines for the journal.

\subsection* {Acknowledgments}
This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

%%%%% References %%%%%

\bibliography{report}   % bibliography data in report.bib
\bibliographystyle{spiejour}   % makes bibtex use spiejour.bst

%%%%% Biographies of authors %%%%%

\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering. He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.  He is the author of more than 50 journal papers and has written three book chapters. His current research interests include optical interconnects, holography, and optoelectronic systems. He is a member of SPIE.

\vspace{1ex}
\noindent Biographies and photographs of the other authors are not available.

\listoffigures
\listoftables

\end{spacing}
\end{document}

### 2024-05-20 python files/create_psd.py

@@ -1,182 +1,0 @@
-#!/usr/bin/python3
-
-from ast import alias
-from matplotlib.mlab import psd
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-from scipy.fft import fft
-from scipy.signal import periodogram
-from scipy.optimize import curve_fit, least_squares
-
-def create_psd(data_file):
-        
-    x_dat = []
-    y_dat = []
-    z_dat = []
-    t_list = []
-
-    if data_file[-3:] == 'txt':
-        print('Loading QPD data')
-        test = data_file.split('/')
-        tag = test[2][:-3]
-        with open(data_file, 'r') as input:
-            lines = input.readlines()
-            for line in lines[1:-1]:
-                items =  line.split(',')
-                x_dat.append(float(items[0]))
-                y_dat.append(float(items[1]))
-                z_dat.append(float(items[2]))
-                t_list.append(float(items[3]))
-        
-        dt = t_list[10]-t_list[9]
-        fs = 1/dt
-
-    elif data_file[-3:] == 'npy':
-        print('Loading trajectory file')
-        test = data_file.split('/')
-        tag = test[2][:-8]
-        traj = np.load(data_file)
-        dt = 1e-5
-        for i, frame in enumerate(traj[1000:]):
-            x_dat.append(frame[0])
-            y_dat.append(frame[1])
-            t_list.append(dt*i)
-        
-        fs = 1/dt
-
-    def loretzian (x, A, B):
-            return 1/(A+B*x**2)
-    
-    def alias_loretzian(k, dx, c, i):
-            N = len(freq_block[i:])
-            return (dt*dx**2)/(1+c**2-2*c*np.cos(np.pi*(k+1)/N))
-    
-    def res_lorentz(x, k, y, i):
-        y_ex = alias_loretzian(k, x[0], x[1], i)
-        sig = y_ex**-1/(nb)**0.5
-        return ((y**-1 - y_ex**-1)/sig)
-
-    def itereative_fitting(freq, psd):
-        '''
[…]

### 2024-05-20 python files/test.py

@@ -1,27 +1,0 @@
-import csv
-import random
-from tkinter import X
-import numpy as np
-import matplotlib.pyplot as plt
-
-x_list = np.linspace(-1,1,10)
-y_list = np.linspace(-1,1,10)
-
-xx, yy = np.meshgrid(x_list, y_list)
-
-fig = plt.figure()
-ax = fig.add_subplot(projection='3d')
-ax.set_xlim(-1,1)
-ax.set_ylim(-1,1)
-ax.set_zlim(-1,1)
-
-for x, y in zip(xx, yy):
-    phi = np.arctan2(x,y)
-    r = (x**2+y**2)**0.5
-    theta = np.arcsin(r)
-    s = [r*np.sin(theta)*np.cos(phi),
-         r*np.sin(theta)*np.sin(phi),
-         r*np.cos(theta)]
-    ax.scatter(s[0], s[1], s[2])
-
-plt.show()

### 2024-05-20 python files/tmp/testfile.py

@@ -1,183 +1,0 @@
-#!/usr/bin/python3
-
-import numpy as np
-import matplotlib.pyplot as plt
-from statsmodels.tsa.stattools import ccf
-
-
-def make_ccf(pair1, pair2):
-    fwr_diag_cross_product = ccf(pair1, pair2, adjusted=True, fft=True)
-    fwr_diag_cross_product /= np.max(abs(fwr_diag_cross_product))
-    bck_diag_cross_product = ccf(pair1[::-1], pair2[::-1], adjusted=True, fft=True)[::-1]
-    bck_diag_cross_product /= np.max(abs(bck_diag_cross_product))
-    return np.r_[bck_diag_cross_product[:-1], fwr_diag_cross_product]
-
-q1_dat, q2_dat, q3_dat, q4_dat = [], [], [], []
-t_list = []
-
-fig, (ax1, ax2, ax3) = plt.subplots(nrows=1, ncols=3, sharey=True)
-plt.suptitle('Quadrant Cross Correlation')
-nlags = 50
-step_size = 1
-
-tau_ot1 = 3.75e-4
-tau_ot2 = 9e-4
-with open('./QPD_output/data_14-May-2024 13:58:16.txt', 'r') as input:
-    lines = input.readlines()
-    for line in lines[1:-1:step_size]:
-        #print(line)
-        items =  line.split(',')
-        q1_dat.append(float(items[0]))
-        q2_dat.append(float(items[1]))
-        q3_dat.append(float(items[2]))
-        q4_dat.append(float(items[3]))
-        t_list.append(float(items[4]))
-
-lag = t_list[2]-t_list[1]
-#print('dt = ' + str(lag), 'fs = ' + str(1/lag))
-
-#from statsmodels.graphics.tsaplots import plot_acf, acf
-
-#from test_signal_corr import get_ACF, get_acf
-
-pair1 = [(q1+q2)-(q3+q4) for q1,q2, q3, q4 in zip(q1_dat, q2_dat, q3_dat, q4_dat)]
-pair2 = [q1+q2 for q1,q2 in zip(q2_dat, q3_dat)]
-
-ccf_output_diag1 = make_ccf(q1_dat, q4_dat)
-ccf_output_diag2 = make_ccf(q2_dat, q3_dat)
-
[…]

### 2024-05-20 python_files/create_psd.py

@@ -1,0 +1,185 @@
+#!/usr/bin/python3
+
+from matplotlib.mlab import psd
+from pathlib import PurePosixPath, PureWindowsPath
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.fft import fft
+from scipy.signal import periodogram
+from scipy.optimize import curve_fit, least_squares
+
+def create_psd(data_file):
+    
+    x_dat = []
+    y_dat = []
+    z_dat = []
+    t_list = []
+
+    if data_file[-3:] == 'txt':
+        print('Loading QPD data')
+        test = data_file.split('/')
+        tag = test[-1][:-3]
+        with open(data_file, 'r') as input:
+            lines = input.readlines()
+            for line in lines[1:-1]:
+                items =  line.split(',')
+                x_dat.append(float(items[0]))
+                y_dat.append(float(items[1]))
+                z_dat.append(float(items[2]))
+                t_list.append(float(items[3]))
+        
+        dt = t_list[10]-t_list[9]
+        fs = 1/dt
+
+    elif data_file[-3:] == 'npy':
+        print('Loading trajectory file')
+        test = data_file.split('\\')
+        print(test)
+        tag = test[-1][:-8]
+        print(tag)
+        name = str(PurePosixPath(PureWindowsPath(data_file)))
+        traj = np.load('./'+name[:])
+        dt = 1e-5
+        for i, frame in enumerate(traj[1000:]):
+            x_dat.append(frame[0])
+            y_dat.append(frame[1])
+            t_list.append(dt*i)
+        
+        fs = 1/dt
+
+    def loretzian (x, A, B):
+            return 1/(A+B*x**2)
+    
+    def alias_loretzian(k, dx, c, i):
+            N = len(freq_block[i:])
+            return (dt*dx**2)/(1+c**2-2*c*np.cos(np.pi*(k+1)/N))
+    
+    def res_lorentz(x, k, y, i):
+        y_ex = alias_loretzian(k, x[0], x[1], i)
+        sig = y_ex**-1/(nb)**0.5
[…]

### 2024-06-05 Latex_files/Thesis/chapters/chapter 2/chapter2.tex
adding thesis and chapter dir

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Strath PhD Thesis Template
%	by Jethro Browell [jethro.browell@strath.ac.uk]
%
%	Guidelines for thesis format, submission and content are found in
%	General and Course Regulations for Graduate and Postgraduate
%	Awards and Degrees, section 20.6.
%
%	Using .eps or .pdf is recomended to prduce high quality figures etc.
%
%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}

% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

% Page Headers
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % optional
%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
\fancyhead[L]{\nouppercase{\leftmark}}
\cfoot{\thepage}
\pagestyle{fancy}

% Draft Watermark
\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{chapter}{1}
\chapter{Theory and methods}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the work being done with 
optical tweezers in particle classification, and how tweezing has been 
used to investigate nucleation events.

\section{Working Principle}
Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects. If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself. While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory. Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}. This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0. However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.65\linewidth]{ot_working_principal.png}
	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards. Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
\end{figure}

\section{Electromagnetism and optical tweezing}
Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
	&\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
	&\nabla \cdot \bold{B} = 0 \\
	&\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
	&\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
\end{align}
These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported. The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point). The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion. The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.

\subsection{Ray-Optics Regime}
The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law. Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam. Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge. As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:

\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray. For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam. The ray optics model is ideal for

\subsection{Lornez-Mie Theory}
The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

\begin{align}
	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
\end{align}

Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user. However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on. From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.

Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory. The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.

\subsection{Rayleigh Regime}
The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size. The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}. The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:

\begin{align}
	F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
	F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
	F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
	& \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
	\text{where:} \nonumber \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle. Scattering force however is dependent on the effective scattering cross sectional area.

\begin{align}
	F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
	\text{where:} \nonumber \\
	C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}). However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field. For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.

The T matrix method was first developed by Peter Waterman with his 
research into acoustic wave scattering, this would later be extended to 
electromagnetic waves. Put simply, the method sets a boundary condition 
where the incident field within the interior field is cancelled out via 
interference; these boundary conditions allows us to solve the Maxwell 
equations by expanding the incident and scattered fields using spherical 
vector waves. Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:

\begin{align}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle. However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.

\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation. Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour. The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour. Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out. This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule. The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}
Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force. The $F(t)$ is there for convention which accounts for any external forces acting on the particle. We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:

\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$. Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy. If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). The full Langevin equation for an optically trapped particle is therefore given as:
\begin{align}
	\label{eq:langevin}
	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
\end{align}

Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}. Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms. If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}
Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression. If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}
By reversing the time step (i.e. use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e. $x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for white noise as no information is known about $W(t)$ at any time. We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range. Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}
Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes. We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time. We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$. Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.

\section{Calibration Techniques}
\label{sec:calibration_techniques}
There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized. The basis for each of these methods stems from the analytical solution of the Langevin equation:
\begin{align}
	\label{eq:anylitical_lang}
	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
\end{align}

\subsection{Potential Well Analysis}
The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center. Therefore a simple analysis method is to understand the height and width of said potential well.

Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap. This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations. With this in mind we can write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}
By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long. It also provides no information on the particle itself (i.e. the friction coefficient $\gamma$ and diffusion coefficient $D$).

\subsection{Equipartition method}
The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials. This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences. Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
\begin{align}
	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
\end{align}

By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength. Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects. Likewise with the potential-analysis route, no information is gleaned about the particle itself.

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles). This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available. As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}
Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid. If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for. If we wish to understand the effects of a given external factor (i.e. an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.

For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus. We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e. $t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series. Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
\begin{align}
	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
\end{align}

If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
\begin{align}
	\label{eq:MSD}
	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
\end{align}
From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
\begin{figure}
	\centering
	\includegraphics[width=0.67\linewidth]{MSD.png}
	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well. The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
\end{figure}
The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus. Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength. The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.

\subsubsection{Angular Mean Square Displacement (MSAD)}
It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
\begin{align}
	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
	2\left[1-\frac{1}{4\beta\kappa_r} 
	\left(\frac{exp(\beta\kappa_r)-1}
	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
	}\right)^2\right]
\end{align}  
Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles. This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources. Taking the Fourier transform of a particle's trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}
where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align} 
We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
\begin{align}
	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
\end{align}

We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.

The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased. Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
\begin{align}
	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
\end{align}
Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}. Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary. Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length. To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units. Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.

\begin{align}
	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
\end{align}

With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation. PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.

\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

\end{document}

### 2024-06-05 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
adding thesis and chapter dir

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Strath PhD Thesis Template
%	by Jethro Browell [jethro.browell@strath.ac.uk]
%
%	Guidelines for thesis format, submission and content are found in
%	General and Course Regulations for Graduate and Postgraduate
%	Awards and Degrees, section 20.6.
%
%	Using .eps or .pdf is recomended to prduce high quality figures etc.
%
%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}

% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

% Page Headers
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % optional
%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
\fancyhead[L]{\nouppercase{\leftmark}}
\cfoot{\thepage}
\pagestyle{fancy}

% Draft Watermark
\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{chapter}{3}
\chapter{Complex Langevin dynamics of simple spherical aggregates}
Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed. However, while working with dense colloidal suspensions, one often ends up trapping more than one sphere. Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness. Theoretical studies on the case of two trapped microspheres by Xu \textit{et.al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well. There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.

In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.

\section{Positional and Orientational dependence of Trapping forces}
If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus. The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser. We can do so because the optical force is only dependent on the particle's relative position. If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=0.67\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap. As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus. Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}. Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}. This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio. More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled. This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap. And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.

Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
\end{figure}

Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap. 
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

\end{document}

### 2024-06-05 Latex_files/Thesis/chapters/chapter 5/chapter5.tex
adding thesis and chapter dir

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Strath PhD Thesis Template
%	by Jethro Browell [jethro.browell@strath.ac.uk]
%
%	Guidelines for thesis format, submission and content are found in
%	General and Course Regulations for Graduate and Postgraduate
%	Awards and Degrees, section 20.6.
%
%	Using .eps or .pdf is recomended to prduce high quality figures etc.
%
%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}

% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

% Page Headers
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % optional
%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
\fancyhead[L]{\nouppercase{\leftmark}}
\cfoot{\thepage}
\pagestyle{fancy}

% Draft Watermark
\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{chapter}{4}
\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\section{Simulative QPD}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.

\subsection{Incident beam}
\label{sec:scattering}
The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). *Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:

\begin{align}
	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
\end{align}

Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD. For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully. In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.

We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers. If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:

\begin{align}
	\begin{pmatrix}
		p_{nm}\\
		q_{nm}
	\end{pmatrix} =T 
	\begin{pmatrix}
		a_{nm}\\
		b_{nm}
	\end{pmatrix} = 
	\begin{pmatrix}
		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
		bT^{TM}_{nm} & bT^{TE}_{nm}
	\end{pmatrix}
	\begin{pmatrix}
		a_{nm}\\
		b_{nm}
	\end{pmatrix}
\end{align}

For *mstm* the T-matrix is packed as a column vector:

\begin{align}
	T_{MSTM} = 
	\begin{bmatrix} 
		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
		... & ...\\ 
		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
		----&----\\ 
		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
	\end{bmatrix}
\end{align}

Where as *ott* packs packs the T-matrix with sub matrices:

\begin{align}
	T_{Ott} = 
	\begin{bmatrix} 
		\begin{pmatrix}
			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
		\end{pmatrix} \\ 
		\begin{pmatrix}
			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
		\end{pmatrix}\\
		....\\ 
		\begin{pmatrix}
			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
			bT^{TM}_{nm} & bT^{TE}_{nm}
		\end{pmatrix}
	\end{bmatrix}
\end{align}

\subsection{Scattered and Total Fields}
With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field. Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam. Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):

The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure). This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field. I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses. The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$

Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam. This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$. The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.

Where the points are Rohrbach's results and the solid line's are our QPD's own replication. The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus. Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes. Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.

Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction. We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk. Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up. This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.  First we checked for this by measuring the QPD response for random positions within the XY plane:

Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk. This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.

\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects. For two correlated power spectra, the cross correlation is given as:

\begin{align}
	P_{x} = |\hat{S}_x(f)|^2 \\ 
	P_{y} = |\hat{S}_y(f)|^2 \\
	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
\end{align}

Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies. They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies. They found these transformed positions by minimising the sum cross-correlation:

\begin{align}
	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
\end{align}

Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap. Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:

\begin{align}
	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
\end{align}

Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum. With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.

\section{Rotations and Asymmetric particles}
Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging. Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions. This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.

Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same). Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic. This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer. As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.

Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer. The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.

\section{Power Spectra of single sphere vs spherical aggregates}
In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}. In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2. As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields. Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.

\begin{figure}[h]
	\label{fig:Rohrbach}
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
	\end{subfigure}
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
	\end{subfigure}
	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
\end{figure}
Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin. The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.

In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser. Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case. We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.

Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum. Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra. However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5. Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2. This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion. Therefore one should expect that for aggregates containing multiple sized particles (e.g. cellular samples) typical characterisation techniques may over estimate the strength of the trap.

Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light. Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally. However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis. This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres. They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis. This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate. We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios. The red line is for the case where the larger sphere is above the smaller sphere. The blue line is the inverted case, while the initial position is above the focus of the trap. And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
\end{figure}

Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion. These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction. The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}. This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector. Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.

\newpage

\section{Monitoring Stochastic rotational motion using static light scattering}
Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects. As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}. While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence

\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page. The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
\subsection{Dimer}
The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam. We scale the position of each sphere by the factor L:
\begin{align}
	L = \frac{1}{k} = \frac{\lambda}{2\pi}
\end{align}

Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
\subsection{Beam}
Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.

\subsection{Detectors and Pixels}

Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
\begin{align}
	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
\end{align}
So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components. The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.

Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector. The pixels are distributed using a sunflower distribution given by:

Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system. Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector. From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate. In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.

This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's. For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.

\section{Interpretation of scattering data into orientation estimates}
MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation. While the latter method is satisfactory for simulation purposes it

\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

\end{document}

### 2024-06-05 Latex_files/Thesis/main.tex
adding thesis and chapter dir

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Strath PhD Thesis Template
%	by Jethro Browell [jethro.browell@strath.ac.uk]
%
%	Guidelines for thesis format, submission and content are found in
%	General and Course Regulations for Graduate and Postgraduate
%	Awards and Degrees, section 20.6.
%
%	Using .eps or .pdf is recomended to prduce high quality figures etc.
%
%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}

% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

% Page Headers
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % optional
%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
\fancyhead[L]{\nouppercase{\leftmark}}
\cfoot{\thepage}
\pagestyle{fancy}

% Draft Watermark
\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}
\pretitle{
  \begin{flushright}
  \vspace{-9.5cm}
%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
  \includegraphics[width=5cm]{logo} \\[6cm]
  \end{flushright}
  \begin{center}
  \LARGE
  \posttitle{\end{center}}
  }

\title{PhD Title \\ PhD Thesis}
\author{D. Maciver\\[-0.8ex]
\small Chemical and Process Engineering Department\\[-0.8ex]
\small University of Strathclyde, Glasgow\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frontmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Declaration
\topskip0pt
\vspace*{\fill}
\noindent
\begin{quote}
	\centering
	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
	%
	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50. Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis. \\[5pt]
	%
	%Signed: \\
	%Date:
\end{quote}
\vspace*{\fill}

\chapter{Abstract}

\tableofcontents

\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{Preface/Acknowledgements}
I would like to acknowledge...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
number of industries such as pharmaceuticals, food production, and 
electronics; where the extraction of dilute materials can help improve 
product quality while keeping production costs low. The industrialisation 
of crystallisation has allowed engineers to reliably and efficiently 
induce the crystal formation within a bulk liquid. However, while on a 
large scale crystallisation is seemingly an understood physical process, 
just a small amount of investigation into the literature reveals that at 
a micro level there is yet to be unifying theory that can accurately 
explain the process of crystallisation, more specifically nucleation.

Nucleation is the initial formation of the crystalline phase and can be 
broadly divided into two main processes. Firstly there is primary 
nucleation, in which the nucleus forms from the bulk liquid; and 
secondary nucleation, where the presence of already existing crystals 
promotes the formation and growth of new nuclei

\section{Nucleation Theories}
Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully. While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now. In particular there is a growing desire for a concrete theory to describe the process of nucleation.

\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}. While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation. The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute. Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus. Therefore the free energy change of a nucleus of size $r$ is given by:
\begin{align}
	\Delta G = \Phi r - \gamma r^{2/3}
\end{align}
From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive. This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution. With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}
Where A is a prefactor that accounts for kinetic contributions. One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple. Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation). In addition the theory implies that any collection of solute molecules below the critical size are unstable.

Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process. However, in practice there are a number of challenges to applying this to experimental measurements.

Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
\begin{align}
	\Delta G = \gamma n^{2/3}-\phi n \\
	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
\end{align}
So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy. The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius. This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}

Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows. Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.

There are several results that indicate that the CNT is not a good fit for nucleation rates:
\textcolor{red}{
	\begin{itemize}
		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass. We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure. To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ]. 
		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled. Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases. [ ]
	\end{itemize}
}

CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate. However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case. Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.  It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards. This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state. Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation. The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'. The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis. Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.

\subsection{Literature related to laser induced nucleation}
From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one. Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}. Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}. Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly. In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.

High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN. The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}. There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}. It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.

Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused. When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}. Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products. When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}. Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events. In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}. The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}. Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum. Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components. The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}. However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}. Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}. This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.

For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM. Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus. This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam. Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM. However, by far the best method for inducing rotation is using birefringent materials.

Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation. For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}. Calculating the optical torque applied to a birefringent material is given via:

\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.  The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle. Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion. Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.

\chapter{Theory and methods}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the work being done with 
optical tweezers in particle classification, and how tweezing has been 
used to investigate nucleation events.

\section{Working Principle}
Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects. If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself. While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory. Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}. This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0. However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards. Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
\end{figure}

\section{Electromagnetism and optical tweezing}
Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
    &\nabla \cdot \bold{B} = 0 \\
    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
\end{align}
These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported. The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point). The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion. The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.

\subsection{Ray-Optics Regime}
The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law. Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam. Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge. As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:

\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray. For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam. The ray optics model is ideal for

\subsection{Lornez-Mie Theory}
The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

\begin{align}
	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
\end{align}

Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user. However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on. From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.

Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory. The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.

\subsection{Rayleigh Regime}
The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size. The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}. The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:

\begin{align}
    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
    \text{where:} \nonumber \\
    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle. Scattering force however is dependent on the effective scattering cross sectional area.

\begin{align}
    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
    \text{where:} \nonumber \\
    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}). However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field. For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.

The T matrix method was first developed by Peter Waterman with his 
research into acoustic wave scattering, this would later be extended to 
electromagnetic waves. Put simply, the method sets a boundary condition 
where the incident field within the interior field is cancelled out via 
interference; these boundary conditions allows us to solve the Maxwell 
equations by expanding the incident and scattered fields using spherical 
vector waves. Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:

\begin{align}
    \begin{pmatrix}
    q_{mn} \\
    p_{mn} 
    \end{pmatrix}
    = \bold{T} 
    \begin{pmatrix}
        a_{mn} \\
        b_{mn}
    \end{pmatrix}
    = \begin{pmatrix}
        T_{11} \ T_{12} \\
        T_{21} \ T_{22}
    \end{pmatrix}
    \begin{pmatrix}
        a_{mn} \\
        b_{mn}
    \end{pmatrix}
\end{align}

The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle. However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.

\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation. Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour. The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour. Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out. This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule. The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}
Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force. The $F(t)$ is there for convention which accounts for any external forces acting on the particle. We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:

\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$. Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy. If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). The full Langevin equation for an optically trapped particle is therefore given as:
\begin{align}
	\label{eq:langevin}
	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
\end{align}

Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}. Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms. If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}
Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression. If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}
By reversing the time step (i.e. use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e. $x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for white noise as no information is known about $W(t)$ at any time. We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range. Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}
Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes. We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time. We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$. Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.

\section{Calibration Techniques}
\label{sec:calibration_techniques}
There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized. The basis for each of these methods stems from the analytical solution of the Langevin equation:
\begin{align}
	\label{eq:anylitical_lang}
	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
\end{align}

\subsection{Potential Well Analysis}
The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center. Therefore a simple analysis method is to understand the height and width of said potential well.

Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap. This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations. With this in mind we can write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}
By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long. It also provides no information on the particle itself (i.e. the friction coefficient $\gamma$ and diffusion coefficient $D$).

\subsection{Equipartition method}
The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials. This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences. Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
\begin{align}
	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
\end{align}

By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength. Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects. Likewise with the potential-analysis route, no information is gleaned about the particle itself.

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles). This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available. As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}
Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid. If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for. If we wish to understand the effects of a given external factor (i.e. an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.

For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus. We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e. $t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series. Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
\begin{align}
	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
\end{align}

If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
\begin{align}
	\label{eq:MSD}
	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
\end{align}
From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
\begin{figure}
	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well. The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
\end{figure}
The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus. Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength. The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.

\subsubsection{Angular Mean Square Displacement (MSAD)}
It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
\begin{align}
	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
	2\left[1-\frac{1}{4\beta\kappa_r} 
	\left(\frac{exp(\beta\kappa_r)-1}
	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
	}\right)^2\right]
\end{align}  
Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles. This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources. Taking the Fourier transform of a particle's trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}
where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align} 
We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
\begin{align}
	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
\end{align}

We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.

The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased. Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
\begin{align}
	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
\end{align}
Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}. Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary. Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length. To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units. Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.

\begin{align}
	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
\end{align}

With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation. PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution. The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results. It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}. Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}. These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.

\section{Optical Tweezer Equipment}
In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals. The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser. Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture. To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications. In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification. Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging. By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations. While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement. The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane. Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm. To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.

\subsection{Position detection methods}
There are several methods for monitoring the position of 
A QPD is a typical example of a position detection system commonly used 
in optical tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is directed onto the QPD, with the maximum 
intensity mapping to the particle's centre of mass. By summing the 
horizontal and vertical quadrants together the particle's centre of mass 
is tracked in the x-y plane. While the tracking is accurate the QPD 
needs to be calibrated in order to convert the voltage signals into 
actual distance reading.

\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
Generation of fluid shear can be achieved via two avenues: Firstly, by 
utilising circularly polarised light it is possible to transfer angular 
momentum from the laser to the trapped entity. Secondly, one can 
directly move the trap within the imaging plane by steering the beam 
using either a galvanometric mirror or gimbal mirror. The following 
chapter outlines the work done with shear generated by circularly 
polarised light and the challenges of applying this to localising 
nucleation.

There are several options for particles that can be rotated using 
optical tweezers []. Over the course of this project two different micro 
spheres where investigated, vaterite and liquid crystal droplets. Both 
can be readily synthesised in the lab and are will rotate at a variety 
of sizes. While silver nano particles were considered their high cost 
and small size meant they were disregarded as an option for optical 
rotation.

\subsection{Rotation of Vaterite micro spheres}
Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping. 
Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.

\chapter{Complex Langevin dynamics of simple spherical aggregates}
Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised. However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead. Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness. Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well. There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.

In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.

\section{Positional and Orientational dependence of Trapping forces}
If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus. The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser. We can do so because the optical force is only dependent on the particle's relative position. If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}

	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap. As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus. Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}. Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}. This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.

\begin{figure}[t]
	\centering
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
\medskip
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio. More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled. This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap. And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.

Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
\end{figure}

Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.

\section{Simulative QPD}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.

\subsection{Incident beam}
\label{sec:scattering}
The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). *Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:

\begin{align}
	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
\end{align}

Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD. For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully. In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.

We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers. If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:

\begin{align}
	\begin{pmatrix}
		p_{nm}\\
		q_{nm}
	\end{pmatrix} =T 
	\begin{pmatrix}
		a_{nm}\\
		b_{nm}
	\end{pmatrix} = 
	\begin{pmatrix}
		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
		bT^{TM}_{nm} & bT^{TE}_{nm}
	\end{pmatrix}
	\begin{pmatrix}
		a_{nm}\\
		b_{nm}
	\end{pmatrix}
\end{align}

For *mstm* the T-matrix is packed as a column vector:

\begin{align}
	T_{MSTM} = 
	\begin{bmatrix} 
		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
		... & ...\\ 
		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
		----&----\\ 
		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
	\end{bmatrix}
\end{align}

Where as *ott* packs packs the T-matrix with sub matrices:

\begin{align}
	T_{Ott} = 
	\begin{bmatrix} 
		\begin{pmatrix}
			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
		\end{pmatrix} \\ 
		\begin{pmatrix}
			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
		\end{pmatrix}\\
		....\\ 
		\begin{pmatrix}
			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
			bT^{TM}_{nm} & bT^{TE}_{nm}
		\end{pmatrix}
	\end{bmatrix}
\end{align}

\subsection{Scattered and Total Fields}
With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field. Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam. Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):

The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure). This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field. I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses. The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$

Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam. This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$. The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.

Where the points are Rohrbach's results and the solid line's are our QPD's own replication. The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus. Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes. Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.

Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction. We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk. Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up. This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.  First we checked for this by measuring the QPD response for random positions within the XY plane:

Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk. This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.

\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects. For two correlated power spectra, the cross correlation is given as:

\begin{align}
	P_{x} = |\hat{S}_x(f)|^2 \\ 
	P_{y} = |\hat{S}_y(f)|^2 \\
	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
\end{align}

Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies. They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies. They found these transformed positions by minimising the sum cross-correlation:

\begin{align}
	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
\end{align}

Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap. Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:

\begin{align}
	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
\end{align}

Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum. With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.

\section{Rotations and Asymmetric particles}
Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging. Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions. This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.

Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same). Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic. This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer. As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.

Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer. The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.

\section{Power Spectra of single sphere vs spherical aggregates}
In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}. In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2. As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields. Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.

\begin{figure}[h]
	\label{fig:Rohrbach}
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
	\end{subfigure}
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
	\end{subfigure}
	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
\end{figure}
Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin. The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.

In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser. Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case. We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.

Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum. Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra. However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5. Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2. This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion. Therefore one should expect that for aggregates containing multiple sized particles (e.g. cellular samples) typical characterisation techniques may over estimate the strength of the trap.

Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light. Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally. However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis. This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres. They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis. This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate. We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios. The red line is for the case where the larger sphere is above the smaller sphere. The blue line is the inverted case, while the initial position is above the focus of the trap. And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
\end{figure}

Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion. These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction. The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}. This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector. Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.

\newpage

\section{Monitoring Stochastic rotational motion using static light scattering}
Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects. As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}. While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence

\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page. The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
\subsection{Dimer}
The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam. We scale the position of each sphere by the factor L:
\begin{align}
	L = \frac{1}{k} = \frac{\lambda}{2\pi}
\end{align}

Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
\subsection{Beam}
Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.

\subsection{Detectors and Pixels}

Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
\begin{align}
	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
\end{align}
So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components. The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.

Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector. The pixels are distributed using a sunflower distribution given by:

Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system. Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector. From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate. In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.

This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's. For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.

\section{Interpretation of scattering data into orientation estimates}
MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation. While the latter method is satisfactory for simulation purposes it

\chapter{Effects of shear flow on the nucleation of crystals}
\section{Shear induced Nucleation}
It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e. propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event. Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off. This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.  There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions. In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate. They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus. There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised. However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius. In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing. This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.

Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium. Typically one would use a birefringent particle (i.e. vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam. Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere. Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field? This requires that we both know the fluid velocity, and the shear rate applied by our probe particle. Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.

Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}
Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively. In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle. There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.

Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution. To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW. The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis. With the particle size and rotation frequency, the tangential rotation speed is calculated via:

\begin{align}
	\label{eq:birefringent_speed}
	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
\end{align}

Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre. Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere. The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):

\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}

\begin{figure}[h]
	\label{fig:vaterite_shear}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere. (b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line. Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW. Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
\end{figure}

From Fig.\ref{fig:vaterite_shear} it's clear that

Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised. One can describe the position parameter of an ellipse as such:

\begin{align}
	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
\end{align}

where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit. Subbing this in and then taking the partial derivative of position gives:

\begin{align}
	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
\end{align}

In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.

\begin{align}
	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point. Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:

\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
\end{align}

Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path

\chapter{Closing Remarks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\chapter{Stuff That Didn't Fit Anywhere Else}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

\end{document}

### 2024-06-06 Latex_files/Thesis/chapters/chapter 2/chapter2.tex
06/06/2024 18:10

@@
 \begin{align}
-	&\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-	&\nabla \cdot \bold{B} = 0 \\
-	&\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-	&\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
+	&\nabla \cdot \mathbf{E} = \frac{\rho_ v}{\epsilon_0} \\
+	&\nabla \cdot \mathbf{B} = 0 \\
+	&\nabla \times \mathbf{E}  = \frac{\delta \mathbf{B}}{\delta t} \\
+	&\nabla \times \mathbf{B} = j_0 +\frac{\delta \mathbf{E}}{\delta t}    
 \end{align}
@@
 \begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
+	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
 \end{align} 
@@
 
+\section{Scattering methods}
+\label{sec:scattering}
+\subsection{Tmatrix}
 The T matrix method was first developed by Peter Waterman with his 
@@
 However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
+
+\subsection{Discrete Dipole Approximation}
 
@@
 \begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
+	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{B}T}} 
 \end{align}
@@
 	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
+	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
 \end{align}
@@
 \end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
+Where $\gamma$ is the objects friction coefficient which for spheres is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
 If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
@@
 \begin{align}
+	\label{eq:lorentzian}
 	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
 \end{align}
-
 We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
 Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.75\linewidth]{PSD.png}
+	\caption{Example PSD fitted using Eq.~\ref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water.
+The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector.
+Using a correction factor (see Eq.~\ref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
+\end{figure}
 The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
@@
 \begin{align}
+	\label{eq:correction_factor}
 	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
@@
 
+\section{Simulation of spherical aggregates}
+Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}.
+As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
+\begin{align}
+	\frac{\mathbf{d}\vec{r}(t)}{\mathbf{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
+\end{align}
+Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case.
+Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation.
+The rotational form of the Langevin equation for a dipole within an external potential is given as:
+\begin{align}
+	\frac{\mathbf{d}\vec{u}(t)}{\mathbf{dt}} = \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector.
+$\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
+\begin{align}
+	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
+\end{align}
+For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
+\begin{align}
+	\label{eq:full_langevin}
+		\frac{\bold{d}\vec{r}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t) \\
+		\frac{\bold{d}\vec{u}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+Fortunately, we do not need to solve these directly as the latter two random variables can be easily approximated if the thermal energy of the system is know, and the rate of change can be assumed as linear if we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll \kappa_u/\gamma_R$.
+In doing so we now only need to compute the optical force and torque applied to our dimer, this has already been compiled in a MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007}.
+In which they use the results from \cite{Farsund1996} to compute both the optical force and torque using the beam coefficients - the form given by \cite{Crichton2000THEMD} is an easier form to compute:
+\begin{equation}
+\begin{split}
+	\bold{F_z}=&-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right. \\ 
+	&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+	& \left. \times \Im[b_{n,m}b*_{n,m}+a_{n,m}a*_{n+1,m} - q_{n,m}q*_{n,m}+p_{n,m}p*_{n+1,m}] \right)
+\end{split}
+\end{equation}
+\begin{equation}
+\begin{split}
+	\bold{T_z}=&-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right. \\ 
+	&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+	& \left. \times \Re[b_{n,m}a*_{n,m}+a_{n,m}b*_{n+1,m} - p_{n,m}q*_{n,m}+q_{n,m}p*_{n+1,m}] \right)
+\end{split}
+\end{equation}
+Where $a_{n,m}$, $b_{n,m}$, $p_{n,m}$ $q_{n,m}$ are the beam coefficients of the incident and scattered fields respectively.
+We can get the transverse force and torque components in a similar form by applying a simple rotation transformation.
+In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate).
+Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.
+
+Throughout this PhD we use the work of Vigilante to perform a systematic study of the dynamics demonstrated by asymmetric dimers in both plane and circularly polarised light.
+But furthermore we expand upon their work to create a fully functional quadrant photo-diode to simulate the response from a calibration test.
 \bibliographystyle{ieeetran}
 \bibliography{thesis_bib}
-
 \end{document}

### 2024-06-06 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
06/06/2024 18:10

@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
@@
 This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
+\begin{figure}[h!]
 	\centering
@@
 
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
 Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
@@
 
-\begin{figure}
+\begin{figure}[h!]
 	\centering
@@
 Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
+
+\section{Continuous rotational motion due to second-order scattering}
+One aspect that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
+Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
+However, theoretical and experimental work by \cite{Yevick2017} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
+This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
+They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
+This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
+We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
+	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
+The red line is for the case where the larger sphere is above the smaller sphere.
+The blue line is the inverted case, while the initial position is above the focus of the trap.
+And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
+\end{figure}
+
+Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
+We used \textit{mstsm} to look at the stokes parameters from the scattered field from a simple plane wave incident on our dimer, the proportion of circularly polarised light is minimal compared to the proportion of plane polarised light, which indicates that this rotational motion is not due to any inhomogeneity in the dimer that might impart angular momentum to the scattered beam - as compared to a anisotropic scatterer like Vaterite.
+
+These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
+The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
+This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
+Furthermore, we see a non-linear increase in the rotational speed of our dimers with size, the drag torque from the surrounding fluid is $~\propto r^3$ so the expectation is that the rotation frequency should fall off with increasing size.
+This indicates that the rotational motion is due to the shape asymmetry of the dimer and not solely due to the beam's angular momentum.
+Measurement of this photo-kinetic force is difficult to achieve due to the fact that previous analysis was conducted in the Rayleigh regime, where the polarizability of our dimer can be approximated as:
+\begin{align}
+	\bold{p}(r,t) = \alpha_xE_x(r,t)\hat{x}_n +\alpha_yE_y(r,t)\hat{y}_n +\alpha_zE_z(r,t)\hat{z}_n
+\end{align}
+Where the polarizability is given as a 3D vector for the three principle Cartesian directions.
+In order to measure the magnitude of second order contributions we would need to construct a dipole array that fully captures the scattering of a dimer.
+
+\section{Conclusions}
+
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Manuscript Dimers/Manuscript_Optical Communications.tex

@@
 
-Ever since their invention in 1986 by Arthur Ashkin \cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been used extensively in colloidal physics \cite{Leonardo_2008, Koehler_2011, Curran_1999}.
-Of particular interest has been the use of a single trapped microsphere as a tracer to probe the rheological properties of the fluid of suspension \cite{Atakhorrami_2006}.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead in a Gaussian beam OT.
-Li and Arlt \cite{Li_2008} studied the case of two microspheres trapped in a single OT and opined that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu_2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Sheng-Hua et al \cite{ShengHua_2005} assumed that the two trapped beads experience different trap stiffnesses and went on to show that the trapped beads collide.
-Experiments in \cite{Praveen_2016} confirmed that the two trapped beads indeed experience different trap stiffnesses when trapped in a single Gaussian beam OT.
+Ever since their invention in 1986 by Arthur Ashkin
+\cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been
+used extensively in colloidal physics
+\cite{Leonardo_2008,Koehler_2011, Curran_1999}.  Of particular
+interest has been the use of a single trapped microsphere as a tracer
+to probe the rheological properties of the fluid of suspension
+\cite{Atakhorrami_2006}.  However, while working with dense colloidal
+suspensions, one often ends up trapping more than one microbead in a
+Gaussian beam OT.  Li and Arlt \cite{Li_2008} studied the case of two
+microspheres trapped in a single OT and opined that multiple trapped
+beads could be mistaken for a single trapped bead with altered trap
+stiffness. Theoretical studies on the case of two trapped microspheres
+by Xu et~al.\ \cite{Xu_2005} employed a ray-optics based model to show
+that the two trapped beads are brought into physical contact with each
+other by optical forces and they also calculated the axial equilibrium
+positions of the two trapped beads as a function of their size.
+Sheng-Hua et~al.\ \cite{ShengHua_2005} assumed that the two trapped
+beads experience different trap stiffnesses and went on to show that
+the trapped beads collide.  Experiments in \cite{Praveen_2016}
+confirmed that the two trapped beads indeed experience different trap
+stiffnesses when trapped in a single Gaussian beam OT.
 
-Besides multiple trapped beads, trapping and dynamics of asymmetric shapes have generated a lot of interest.
-Mihiretie et.al., \cite{Loudet_2014} studied the trapping of ellipsoidal microparticles and found that particles of aspect ratio ‘k’ greater than 3 and less than 0.3 exhibited oscillatory dynamics with a coupling between translation and rotation in the trap and those with aspect ratio outside this range exhibited less pronounced oscillations with the major axis pre-dominantly pointed along the laser propagation direction.
-Cao et.al., \cite{Curran_1999} performed T-Matrix based simulations of the trapping dynamics of nanowires and microcylinders in the Rayleigh regime and found that polarization dictated the final equilibrium orientation with the nanowires equilibrating with a large tilt angle about the laser propagation direction when the beam was linearly polarized and along the beam propagation direction in circularly polarized traps.
-Chetana D et, al., \cite{Chetana_2022} increased the complexity of the problem by considering an object with both shape as well as optical anisotropy in the form of a chicken Red Blood Cell (cRBC) and showed that the trapping dynamics and the final equilibrium orientation of the cRBC was decided by the beam polarization with the cell showing a dual reorientation behaviour in a linearly polarized trap and a partial second reorientation in a circularly polarized trap.
+Besides multiple trapped beads, trapping and dynamics of asymmetric
+shapes have generated a lot of interest.  Mihiretie et~al.\
+\cite{Loudet_2014} studied the trapping of ellipsoidal microparticles
+and found that particles of aspect ratio ‘k’ greater than $3$ and less
+than $0.3$ exhibited oscillatory dynamics with a coupling between
+translation and rotation in the trap and those with aspect ratio
+outside this range exhibited less pronounced oscillations with the
+major axis pre-dominantly pointed along the laser propagation
+direction.  Cao et~al.\ \cite{Curran_1999} performed T-Matrix based
+simulations of the trapping dynamics of nanowires and microcylinders
+in the Rayleigh regime and found that polarization dictated the final
+equilibrium orientation with the nanowires equilibrating with a large
+tilt angle about the laser propagation direction when the beam was
+linearly polarized and along the beam propagation direction in
+circularly polarized traps.  Chetana et~al.\ \cite{Chetana_2022}
+increased the complexity of the problem by considering an object with
+both shape as well as optical anisotropy in the form of a chicken red
+bloodc Cell (cRBC) and showed that the trapping dynamics and the final
+equilibrium orientation of the cRBC was decided by the beam
+polarization with the cell showing a dual reorientation behaviour in a
+linearly polarized trap and a partial second reorientation in a
+circularly polarized trap.
 
@@
 \paragraph{Simulating motion within an optical trap}
-We assume the dimers to be low-Reynold’s number objects and write down their dynamics within the trap using the Langevin equation:
+We assume the dimers to be low-Reynolds number objects and write down
+their dynamics within the trap using the Langevin equation:
 \begin{align}

### 2024-06-07 Latex_files/Manuscripts Scattering/Manuscript_Spie.tex

--- 2024-05-20 Latex_files/Manuscripts Scattering/Manuscript_Optica.tex
@@
-\documentclass[12pt]{optica-article}
+\documentclass[12pt]{spieman}
 
@@
 \usepackage{lineno}
-
 \usepackage{amsmath}
@@
 \usepackage{tabularx}
+\usepackage{graphicx}
 \makenomenclature
@@
 \linenumbers
-
-\journal{opticajournal}
-
-\begin{document}
 
@@
 
-\author{Daniel Maciver, \authormark{1,*}, Praveen Parthasarathi, \authormark{1}, Leo Lue, \authormark{1}, Jan Sefcik, \authormark{1},
-Mark Haw, \authormark{1}}
+\author[a*]{Daniel Maciver}
+\author[a]{Praveen Parthasarathi}
+\author[a]{Leo Lue} 
+\author[a]{Jan Sefcik}
+\author[a]{Mark Haw}
 
-\address{\authormark{1}Department of Chemical Engineering,
+\affil[a]{Department of Chemical Engineering,
 	University of Strathclyde, 75 Montrose Street, Glasgow,
 	G1 1XL, Scotland}
-\email{\authormark{*}daniel.maciver.2016@uni.strath.ac.uk}
 
+\begin{document}
+	\maketitle
 \begin{abstract}
@@
 \section{Introduction}
+\begin{spacing}{1.5}
+
 \label{sec:Intro}
@@
 \begin{subfigure}{0.45\textwidth}
-	\subcaption{}
-	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
+	\includegraphics[width=5.5cm, height=5.5cm]{./Images/fig1a.png}
 \end{subfigure}
 \begin{subfigure}{0.45\textwidth}
-	\subcaption{}
-	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
+	\includegraphics[width=5.5cm, height=5.5cm]{./Images/fig1b.png}
 \end{subfigure}
@@
 	\centering
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig2a.png}
+		\includegraphics[width=5.5cm]{./Images/fig2a.png}
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig2b.png}
+		\includegraphics[width=5.5cm]{./Images/fig2b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig3a.png}
+		\includegraphics[width=5.5cm]{./Images/fig3a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig3b.png}
+		\includegraphics[width=5.5cm]{./Images/fig3b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig4a.png}
+		\includegraphics[width=5.5cm]{./Images/fig4a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig4b.png}
+		\includegraphics[width=5.5cm]{./Images/fig4b.png}
 	\end{subfigure}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{./Images/fig8a.png}
+\includegraphics[width=5.5cm]{./Images/fig8a.png}
 \caption{\label{fig:biased}
@@
 		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/fig7a.png}
+		\includegraphics[width =5.5cm]{./Images/fig7a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig7b.png}
+		\includegraphics[width=5.5cm]{./Images/fig7b.png}
 	\end{subfigure}
@@
 			\subcaption{}
-			\includegraphics[width =\textwidth]{./Images/fig8a.png}
+			\includegraphics[width =5.5cm]{./Images/fig8a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig8b.png}
+		\includegraphics[width=5.5cm]{./Images/fig8b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig8c.png}
+		\includegraphics[width=5.5cm]{./Images/fig8c.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width =\textwidth]{./Images/fig9a.png}
+		\includegraphics[width =5.5cm]{./Images/fig9a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig9b.png}
+		\includegraphics[width=5.5cm]{./Images/fig9b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig9c.png}
+		\includegraphics[width=5.5cm]{./Images/fig9c.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig10a.png}
+		\includegraphics[width=5.5cm]{./Images/fig10a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig10b.png}
+		\includegraphics[width=5.5cm]{./Images/fig10b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig10c.png}
+		\includegraphics[width=5.5cm]{./Images/fig10c.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig11a.png}
+		\includegraphics[width=5.5cm]{./Images/fig11a.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig11b.png}
+		\includegraphics[width=5.5cm]{./Images/fig11b.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\textwidth]{./Images/fig11c.png}
+		\includegraphics[width=5.5cm]{./Images/fig11c.png}
 	\end{subfigure}
@@
 However, Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.
+\end{spacing}
 
@@
 		Reference Orientations vector components, for $n_{refs} = 30 \ ^*$}
-		\begin{tabularx}{0.9\textwidth}{
+		\begin{tabularx}{0.7\textwidth}{
 				| >{\centering\arraybackslash}X

### 2024-06-07 Latex_files/Thesis/chapters/chapter 1/chapter1.tex

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
 \usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
 
@@
 \begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/chapter2.tex

@@
 But furthermore we expand upon their work to create a fully functional quadrant photo-diode to simulate the response from a calibration test.
+This builds upon the work from \cite{Rohrbach2002} which applied the fundamental principles of Lorenz-Mie theory to replicate the response signal of a QPD being used in back focal-plane interferometry.
+
+\subsection{Simulated Quadrant Photodiode}
+In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface.
+While trapping a micro-particle, the scattered and incident fields combine together and interfere with one another.
+These fields are collected by a condenser lens in the far field limit and are focused onto the QPD surface, the total intensity can be evaluated as:
+\begin{align}
+I(x,y) = \epsilon_0c\left|
+\begin{bmatrix} 
+	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
+	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
+	E_{i,z}(x,y)+E_{s,z}(x,y)
+\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
+\end{align}
+The last term is simply a representative step term that defines the outer limit by which we evaluate the electric field, this is analogous to our condenser lens removing noise from other light sources by only accepting light at a specific acceptance angle defined by its numerical aperture $NA_c$.
+Depending on the relative size of our particle we can adjust the acceptance angle, this has very little effect on the transverse signals, but for axial evaluations of a trapped particle the numerical aperture should be tuned so that the resultant response curve has negative slope in order to allow for axial position detection, the method for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
+
+The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
+\textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
+
+\begin{align}
+	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
+\end{align}
+
+Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
+For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
+In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
+
+We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers.
+The scattered and incident fields are then combined
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
+\end{figure}
+
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 3/chapter3.tex

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
 \begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
+\setcounter{chapter}{2}
 \chapter{Effects of localised shearing on crystal growth and nucleation}
@@
 
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
 \section{Shear induced Nucleation}
@@
 	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
+		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
 		\subcaption{}
@@
 	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
+		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
 		\subcaption{}
@@
 
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 5/chapter5.tex

@@
 
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
 \section{Monitoring Stochastic rotational motion using static light scattering}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/chapter2.tex
moving text

--- 2024-06-05 Latex_files/Thesis/chapters/chapter 2/chapter2.tex
@@
 \setcounter{chapter}{1}
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-	&\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-	&\nabla \cdot \bold{B} = 0 \\
-	&\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-	&\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-	F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-	F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-	F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-	& \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-	\text{where:} \nonumber \\
-	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-	exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-	F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-	\text{where:} \nonumber \\
-	C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-	\begin{pmatrix}
-		q_{mn} \\
-		p_{mn} 
-	\end{pmatrix}
-	= \bold{T} 
-	\begin{pmatrix}
-		a_{mn} \\
-		b_{mn}
-	\end{pmatrix}
-	= \begin{pmatrix}
-		T_{11} \ T_{12} \\
-		T_{21} \ T_{22}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{mn} \\
-		b_{mn}
-	\end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.67\linewidth]{MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
+\input{main}
 
@@
 \bibliography{thesis_bib}
-
 \end{document}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
moving text

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Strath PhD Thesis Template
-%	by Jethro Browell [jethro.browell@strath.ac.uk]
-%
-%	Guidelines for thesis format, submission and content are found in
-%	General and Course Regulations for Graduate and Postgraduate
-%	Awards and Degrees, section 20.6.
-%
-%	Using .eps or .pdf is recomended to prduce high quality figures etc.
-%
-%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper,oneside,11pt]{book}
-\setcounter{secnumdepth}{3}
-\usepackage{amsbsy}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{mathrsfs}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[strict]{changepage}
-
-% Page Margins - Strath Requirement
-\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
-
-% Page Headers
-\usepackage{fancyhdr}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt} % optional
-%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
-\fancyhead[L]{\nouppercase{\leftmark}}
-\cfoot{\thepage}
-\pagestyle{fancy}
-
-% Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
-
-% Line Spacing
-%\def\baselinestretch{1.5} 
-\usepackage{setspace}
-\setstretch{1.5}
-
-% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
-\usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
 \chapter{Theory and methods}
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
+	\includegraphics[width=0.65\linewidth]{ot_working_principal.png}
 	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
@@
 \begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
+	&\nabla \cdot \mathbf{E} = \frac{\rho_ v}{\epsilon_0} \\
+	&\nabla \cdot \mathbf{B} = 0 \\
+	&\nabla \times \mathbf{E}  = \frac{\delta \mathbf{B}}{\delta t} \\
+	&\nabla \times \mathbf{B} = j_0 +\frac{\delta \mathbf{E}}{\delta t}    
 \end{align}
@@
 \begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
+	\centering
+	\includegraphics[width=0.67\linewidth]{MSD.png}
 	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
@@
 
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
+\section{Simulation of spherical aggregates}
+Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}.
+As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
+\begin{align}
+	\frac{\mathbf{d}\vec{r}(t)}{\mathbf{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
+\end{align}
+Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case.
+Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation.
+The rotational form of the Langevin equation for a dipole within an external potential is given as:
+\begin{align}
+	\frac{\mathbf{d}\vec{u}(t)}{\mathbf{dt}} = \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector.
+$\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
+\begin{align}
+	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
+\end{align}
+For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
+\begin{align}
+	\label{eq:full_langevin}
+		\frac{\bold{d}\vec{r}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t) \\
+		\frac{\bold{d}\vec{u}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+Fortunately, we do not need to solve these directly as the latter two random variables can be easily approximated if the thermal energy of the system is know, and the rate of change can be assumed as linear if we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll \kappa_u/\gamma_R$.
+In doing so we now only need to compute the optical force and torque applied to our dimer, this has already been compiled in a MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007}.
+In which they use the results from \cite{Farsund1996} to compute both the optical force and torque using the beam coefficients - the form given by \cite{Crichton2000THEMD} is an easier form to compute:
+\begin{equation}
+\begin{split}
+	\bold{F_z}=&-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right. \\ 
+	&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+	& \left. \times \Im[b_{n,m}b*_{n,m}+a_{n,m}a*_{n+1,m} - q_{n,m}q*_{n,m}+p_{n,m}p*_{n+1,m}] \right)
+\end{split}
+\end{equation}
+\begin{equation}
+\begin{split}
+	\bold{T_z}=&-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right. \\ 
+	&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+	& \left. \times \Re[b_{n,m}a*_{n,m}+a_{n,m}b*_{n+1,m} - p_{n,m}q*_{n,m}+q_{n,m}p*_{n+1,m}] \right)
+\end{split}
+\end{equation}
+Where $a_{n,m}$, $b_{n,m}$, $p_{n,m}$ $q_{n,m}$ are the beam coefficients of the incident and scattered fields respectively.
+We can get the transverse force and torque components in a similar form by applying a simple rotation transformation.
+In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate).
+Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.
 
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
+Throughout this PhD we use the work of Vigilante to perform a systematic study of the dynamics demonstrated by asymmetric dimers in both plane and circularly polarised light.
+But furthermore we expand upon their work to create a fully functional quadrant photo-diode to simulate the response from a calibration test.
+This builds upon the work from \cite{Rohrbach2002} which applied the fundamental principles of Lorenz-Mie theory to replicate the response signal of a QPD being used in back focal-plane interferometry.
 
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
+\subsection{Simulated Quadrant Photodiode}
+In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface.
+While trapping a micro-particle, the scattered and incident fields combine together and interfere with one another.
+These fields are collected by a condenser lens in the far field limit and are focused onto the QPD surface, the total intensity can be evaluated as:
+\begin{align}
+I(x,y) = \epsilon_0c\left|
+\begin{bmatrix} 
+	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
+	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
+	E_{i,z}(x,y)+E_{s,z}(x,y)
+\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
+\end{align}
+The last term is simply a representative step term that defines the outer limit by which we evaluate the electric field, this is analogous to our condenser lens removing noise from other light sources by only accepting light at a specific acceptance angle defined by its numerical aperture $NA_c$.
+Depending on the relative size of our particle we can adjust the acceptance angle, this has very little effect on the transverse signals, but for axial evaluations of a trapped particle the numerical aperture should be tuned so that the resultant response curve has negative slope in order to allow for axial position detection, the method for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
 
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
 The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
+\textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
 
@@
 
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
+Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
 For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
@@
 
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
+We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers.
+The scattered and incident fields are then combined
 
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
 \end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
-
-\end{document}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 1/chapter1.tex
updating main and chapters using

@@
 
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
+\include{main}
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 1/main.tex
updating main and chapters using

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Strath PhD Thesis Template
-%	by Jethro Browell [jethro.browell@strath.ac.uk]
-%
-%	Guidelines for thesis format, submission and content are found in
-%	General and Course Regulations for Graduate and Postgraduate
-%	Awards and Degrees, section 20.6.
-%
-%	Using .eps or .pdf is recomended to prduce high quality figures etc.
-%
-%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper,oneside,11pt]{book}
-\setcounter{secnumdepth}{3}
-\usepackage{amsbsy}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{mathrsfs}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[strict]{changepage}
-
-% Page Margins - Strath Requirement
-\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
-
-% Page Headers
-\usepackage{fancyhdr}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt} % optional
-%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
-\fancyhead[L]{\nouppercase{\leftmark}}
-\cfoot{\thepage}
-\pagestyle{fancy}
-
-% Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
-
-% Line Spacing
-%\def\baselinestretch{1.5} 
-\usepackage{setspace}
-\setstretch{1.5}
-
-% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
-\usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \chapter{Introduction}
@@
 Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
-
-\end{document}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 3/chapter3.tex
updating main and chapters using

@@
 \setcounter{chapter}{2}
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
+\include{main}
 

### 2024-06-07 Latex_files/Thesis/chapters/chapter 3/main.tex
updating main and chapters using

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Strath PhD Thesis Template
-%	by Jethro Browell [jethro.browell@strath.ac.uk]
-%
-%	Guidelines for thesis format, submission and content are found in
-%	General and Course Regulations for Graduate and Postgraduate
-%	Awards and Degrees, section 20.6.
-%
-%	Using .eps or .pdf is recomended to prduce high quality figures etc.
-%
-%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper,oneside,11pt]{book}
-\setcounter{secnumdepth}{3}
-\usepackage{amsbsy}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{mathrsfs}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[strict]{changepage}
-
-% Page Margins - Strath Requirement
-\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
-
-% Page Headers
-\usepackage{fancyhdr}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt} % optional
-%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
-\fancyhead[L]{\nouppercase{\leftmark}}
-\cfoot{\thepage}
-\pagestyle{fancy}
-
-% Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
-
-% Line Spacing
-%\def\baselinestretch{1.5} 
-\usepackage{setspace}
-\setstretch{1.5}
-
-% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
-\usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
 \chapter{Effects of localised shearing on crystal growth and nucleation}
@@
 Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
-
-\end{document}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
updating main and chapters using

--- 2024-06-05 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
@@
 \setcounter{chapter}{3}
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one sphere.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu \textit{et.al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
+\include{main}
 
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{transverse_force.png}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{0.85\linewidth}
-		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{0.85\linewidth}
-		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 4/main.tex
updating main and chapters using

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Strath PhD Thesis Template
-%	by Jethro Browell [jethro.browell@strath.ac.uk]
-%
-%	Guidelines for thesis format, submission and content are found in
-%	General and Course Regulations for Graduate and Postgraduate
-%	Awards and Degrees, section 20.6.
-%
-%	Using .eps or .pdf is recomended to prduce high quality figures etc.
-%
-%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper,oneside,11pt]{book}
-\setcounter{secnumdepth}{3}
-\usepackage{amsbsy}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{mathrsfs}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[strict]{changepage}
-
-% Page Margins - Strath Requirement
-\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
-
-% Page Headers
-\usepackage{fancyhdr}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt} % optional
-%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
-\fancyhead[L]{\nouppercase{\leftmark}}
-\cfoot{\thepage}
-\pagestyle{fancy}
-
-% Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
-
-% Line Spacing
-%\def\baselinestretch{1.5} 
-\usepackage{setspace}
-\setstretch{1.5}
-
-% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
-\usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
 \chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
+Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
+However, while working with dense colloidal suspensions, one often ends up trapping more than one sphere.
 Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
+Theoretical studies on the case of two trapped microspheres by Xu \textit{et.al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
 Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
 	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
+		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
 		\caption{}
@@
 	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
+		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
 		\caption{}
@@
 		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
+		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
 		\label{large over small}
@@
 		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
+		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
 		\label{small_over_large}
@@
 We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
+As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
 This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
 This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
+\begin{figure}[h!]
 	\centering
 	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
+	\includegraphics[width=0.67\linewidth]{transverse_force.png}
 	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
+	\begin{subfigure}{0.85\linewidth}
+		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
 		\caption{}
@@
 	\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
+	\begin{subfigure}{0.85\linewidth}
+		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
 		\caption{}
@@
 
-\begin{figure}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
+	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
 	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
@@
 
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
+\section{Continuous rotational motion due to second-order scattering}
+One aspect that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
 Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
+However, theoretical and experimental work by \cite{Yevick2017} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
 This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
 They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
+This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
 We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
+	\includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
 	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
@@
 Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
+We used \textit{mstsm} to look at the stokes parameters from the scattered field from a simple plane wave incident on our dimer, the proportion of circularly polarised light is minimal compared to the proportion of plane polarised light, which indicates that this rotational motion is not due to any inhomogeneity in the dimer that might impart angular momentum to the scattered beam - as compared to a anisotropic scatterer like Vaterite.
+
+These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
 The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
 This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
+Furthermore, we see a non-linear increase in the rotational speed of our dimers with size, the drag torque from the surrounding fluid is $~\propto r^3$ so the expectation is that the rotation frequency should fall off with increasing size.
+This indicates that the rotational motion is due to the shape asymmetry of the dimer and not solely due to the beam's angular momentum.
+Measurement of this photo-kinetic force is difficult to achieve due to the fact that previous analysis was conducted in the Rayleigh regime, where the polarizability of our dimer can be approximated as:
+\begin{align}
+	\bold{p}(r,t) = \alpha_xE_x(r,t)\hat{x}_n +\alpha_yE_y(r,t)\hat{y}_n +\alpha_zE_z(r,t)\hat{z}_n
+\end{align}
+Where the polarizability is given as a 3D vector for the three principle Cartesian directions.
+In order to measure the magnitude of second order contributions we would need to construct a dipole array that fully captures the scattering of a dimer.
 
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
-
-\end{document}
+\section{Conclusions}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 5/chapter5.tex
updating main and chapters using

@@
 \setcounter{chapter}{4}
-\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
+\include{main}
 \bibliographystyle{ieeetran}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 5/main.tex
updating main and chapters using

--- 2024-06-05 Latex_files/Thesis/main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Strath PhD Thesis Template
-%	by Jethro Browell [jethro.browell@strath.ac.uk]
-%
-%	Guidelines for thesis format, submission and content are found in
-%	General and Course Regulations for Graduate and Postgraduate
-%	Awards and Degrees, section 20.6.
-%
-%	Using .eps or .pdf is recomended to prduce high quality figures etc.
-%
-%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper,oneside,11pt]{book}
-\setcounter{secnumdepth}{3}
-\usepackage{amsbsy}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{mathrsfs}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{epsfig}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[strict]{changepage}
-
-% Page Margins - Strath Requirement
-\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
-
-% Page Headers
-\usepackage{fancyhdr}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt} % optional
-%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
-\fancyhead[L]{\nouppercase{\leftmark}}
-\cfoot{\thepage}
-\pagestyle{fancy}
-
-% Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
-
-% Line Spacing
-%\def\baselinestretch{1.5} 
-\usepackage{setspace}
-\setstretch{1.5}
-
-% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
-\usepackage{titling}
-\pretitle{
-  \begin{flushright}
-  \vspace{-9.5cm}
-%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
-  \end{flushright}
-  \begin{center}
-  \LARGE
-  \posttitle{\end{center}}
-  }
-
-\title{PhD Title \\ PhD Thesis}
-\author{D. Maciver\\[-0.8ex]
-\small Chemical and Process Engineering Department\\[-0.8ex]
-\small University of Strathclyde, Glasgow\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frontmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Declaration
-\topskip0pt
-\vspace*{\fill}
-\noindent
-\begin{quote}
-	\centering
-	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
-	%
-	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50.
-Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis.
-\\[5pt]
-	%
-	%Signed: \\
-	%Date:
-\end{quote}
-\vspace*{\fill}
-
-\chapter{Abstract}
-
-\tableofcontents
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-
-\chapter{Preface/Acknowledgements}
-I would like to acknowledge...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
-
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
-
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
-
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
-
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
+\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
 \section{Simulative QPD}
@@
 		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
+		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
 	\end{subfigure}
@@
 		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
+		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
 	\end{subfigure}
@@
 cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
 
@@
 While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
-
-\chapter{Closing Remarks}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
-
-\end{document}

### 2024-06-07 Latex_files/Thesis/main.tex
updating main and chapters using

@@
 \usepackage[strict]{changepage}
+\usepackage{graphicx}
+\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
 
@@
 
-\chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
+\input{chapters/chapter 1/main}
 
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
+\input{chapters/chapter 2/main}
 
-\section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
+\input{chapters/chapter 3/main}
 
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
-\begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
-\end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
+\input{chapters/chapter 4/main}
 
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
-
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
-
-\subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
-
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
-
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
-
-\subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
-
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
-
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
-
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
-
-\chapter{Theory and methods}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-\section{Working Principle}
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/ot_working_principal.png}
-	\caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-\end{figure}
-
-\section{Electromagnetism and optical tweezing}
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
-\begin{align}
-    &\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
-    &\nabla \cdot \bold{B} = 0 \\
-    &\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
-    &\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
-\end{align}
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
-
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
-
-\subsection{Ray-Optics Regime}
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
-
-\subsection{Lornez-Mie Theory}
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
-\end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
-
-\begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
-\end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
-
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
-
-\subsection{Rayleigh Regime}
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
-\begin{align}
-    F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-     F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-    F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-    & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-    \text{where:} \nonumber \\
-    I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-    exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
-
-\begin{align}
-    F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-    \text{where:} \nonumber \\
-    C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
-
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
-
-\begin{align}
-    \begin{pmatrix}
-    q_{mn} \\
-    p_{mn} 
-    \end{pmatrix}
-    = \bold{T} 
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-    = \begin{pmatrix}
-        T_{11} \ T_{12} \\
-        T_{21} \ T_{22}
-    \end{pmatrix}
-    \begin{pmatrix}
-        a_{mn} \\
-        b_{mn}
-    \end{pmatrix}
-\end{align}
-
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
-
-\section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
-\end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
-
-\begin{align}
-	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
-	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
-\end{align}
-
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
-The full Langevin equation for an optically trapped particle is therefore given as:
-\begin{align}
-	\label{eq:langevin}
-	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
-\end{align}
-
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
-
-\subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
-\begin{align}
-	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
-\end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
-\begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
-\end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
-\begin{align}
-	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
-\end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
-\begin{align}
-	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
-\end{align}
-
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\section{Calibration Techniques}
-\label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
-\begin{align}
-	\label{eq:anylitical_lang}
-	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
-\end{align}
-
-\subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
-\begin{align}
-	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
-\end{align}
-which therefore means we can write the potential well as:
-\begin{align}
-	\label{eq:potential_well}
-	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
-\end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
-\begin{align}
-	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
-
-\subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
-\begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
-	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
-\end{align}
-
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
-
-\subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
-\begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
-\end{align}
-Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
-
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
-\begin{align}
-	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
-\end{align}
-
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
-\begin{align}
-	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
-\end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
-	\includegraphics[width=0.67\linewidth]{figs/MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
-\end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
-
-\subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
-\begin{align}
-	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
-	2\left[1-\frac{1}{4\beta\kappa_r} 
-	\left(\frac{exp(\beta\kappa_r)-1}
-	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
-	}\right)^2\right]
-\end{align}  
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
-
-\subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
-\begin{align}
-	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
-\end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align} 
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
-\begin{align}
-	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
-\end{align}
-
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
-
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
-\begin{align}
-	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
-\end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
-\begin{align}
-	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
-\end{align}
-
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
-
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
-
-\subsection{Position detection methods}
-There are several methods for monitoring the position of 
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
-
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
-
-There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
-
-\subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
-
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
-Much of the theory discussed in Chapter 2 assumes that the target particle in question will be a single sphere, one who's scattering and motion is easily generalised.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
-
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
-\section{Positional and Orientational dependence of Trapping forces}
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
-
-\begin{figure}[h]
-	\centering
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
-		\includegraphics[width=\linewidth]{figs/lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{figs/theta=180.png}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
-\end{figure}
-
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
-\begin{figure}
-	\centering
-	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{figs/transverse_force.png}
-
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
-\end{figure}
-
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-\medskip
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
-	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
-\end{figure}
-
-This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio.
-More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled.
-This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap.
-And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{figs/off_axis_trap.png}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
-\end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
-
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{figs/QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
-
-Now one of the things that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \textit{Grier et al} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.65\linewidth]{figs/rotation_rate_vs_size.png}
-	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
-\end{figure}
-
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore because most of these experiments operate in the Rayleigh regime orientational effects are not usually considered whereas our own results make it clear that the initial orientation and position of spherical aggregates drastically effects their steady state behaviour.
-
-\newpage
-
-\section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
-
-\subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
-\subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
-
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
-
-\subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
-\begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
-\end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
-
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
-
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
-
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
-
-\section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
-
-\chapter{Effects of shear flow on the nucleation of crystals}
-\section{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
-
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
-
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-\begin{figure}[h]
-	\label{fig:vaterite_shear}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_fluid_flow.png}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
-		\includegraphics[width=\linewidth]{figs/vaterite_shear_rate.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} it's clear that
-
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
-
-\begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
-\end{align}
-
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
-
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
-
-\begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
-
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-\end{align}
-
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
+\input{chapters/chapter 5/main}
 

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
updating main and chapters using
reformatting chapter2

@@
 \chapter{Theory and methods}
+
 The two key areas of this project cover two widely different fields; 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Working Principle}
+
 Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Electromagnetism and optical tweezing}
+
 Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
@@
 \begin{align}
-	&\nabla \cdot \mathbf{E} = \frac{\rho_ v}{\epsilon_0} \\
-	&\nabla \cdot \mathbf{B} = 0 \\
-	&\nabla \times \mathbf{E}  = \frac{\delta \mathbf{B}}{\delta t} \\
-	&\nabla \times \mathbf{B} = j_0 +\frac{\delta \mathbf{E}}{\delta t}    
+  \nabla \cdot \mathbf{E}
+  &= \frac{\rho_ v}{\epsilon_0}
+  \\
+  \nabla \cdot \mathbf{B}
+  &= 0
+  \\
+  \nabla \times \mathbf{E}
+  &= \frac{\partial \mathbf{B}}{\partial t}
+  \\
+  \nabla \times \mathbf{B}
+  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
 \end{align}
+where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
+current, $\epsilon_0$ is the permittivity of free space.
+
 These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Ray-Optics Regime}
+
 The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
@@
 As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
-
 \begin{align}
@@
 \end{align}
+where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
+in refractive indices between the solution and the target particle,
+and $P_i$ is the power of the individual ray. For a beam with a
+Gaussian intensity distribution $P_i$ will fall off as you move from
+the centre of the beam.  The ray optics model is ideal for
 
-Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray.
-For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam.
-The ray optics model is ideal for
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Lornez-Mie Theory}
 
-\subsection{Lornez-Mie Theory}
 The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
+	\includegraphics[width=0.75\linewidth]{./figs/fixed_polarisation.png}
 \end{figure}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
more updates...

@@
 
-Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects.
-If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself.
-While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory.
-Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0.
-However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.
+Optical tweezing operates on the principle that light carries both
+linear and angular momentum, which can be transferred to other
+objects.  If light is reflected or refracted by a medium part of the
+light's momentum is transferred to the medium itself.  While at large
+scales this effect is insignificant in most cases, at micron scales
+the forces imparted by individual photons begin to become significant
+enough to influence its trajectory.  Ashkin's work into optical
+tweezing found that micron sized entities could be spatially locked by
+aiming a laser with a Gaussian intensity profile directly upward at a
+the cell while being suspended in a liquid medium \cite{Ashkin1970}.
+This is demonstrated diagram below where we have a simple sphere
+trapped by a Gaussian beam, if we consider two rays of light hitting
+our sphere we see that while in the centre each ray is refracted in an
+equal opposite direction to one another and so the net force imparted
+is 0.  However when displaced the net force is unbalanced and points
+towards the centre of the trap; this principle can be generalised to 3
+degrees of freedom if we consider a focused beam instead of a
+para-axial beam, typically the trap will be weaker along the beam axis
+compared to the transverse trap strength.
 
@@
 The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below:
+The differential forms of which are given below
+\cite{Jackson_1975}:
 \begin{align}
@@
 where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
-current, $\epsilon_0$ is the permittivity of free space.
+current, and $\epsilon_0$ is the permittivity of free space.
 
@@
 \begin{align}
-	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
-	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
-	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
+  E_{\rm inc}(r)
+  &=
+    E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r)
+  \\
+  E_{\rm int}(r)
+  &=
+    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right)
+  \\
+  E_{scat}(r)
+  &=
+    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
 \end{align}
-
-Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
+where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
 However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
@@
 
-Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
+Lorenz-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
 The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Simulation of spherical aggregates}
+
 Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}.
@@
 \begin{align}
-	\frac{\mathbf{d}\vec{r}(t)}{\mathbf{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
+\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
 \end{align}
@@
 \begin{align}
-	\frac{\mathbf{d}\vec{u}(t)}{\mathbf{dt}} = \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+  \frac{{d}\vec{u}(t)}{{dt}}
+  =
+  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
+  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
 \end{align}
@@
 	\label{eq:full_langevin}
-		\frac{\bold{d}\vec{r}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t) \\
-		\frac{\bold{d}\vec{u}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+  \frac{{d}\vec{r}(t)}{{dt}}
+  &=
+    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
+  \\
+  \frac{{d}\vec{u}(t)}{{dt}}
+  &=
+    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
+    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
 \end{align}
-Fortunately, we do not need to solve these directly as the latter two random variables can be easily approximated if the thermal energy of the system is know, and the rate of change can be assumed as linear if we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll \kappa_u/\gamma_R$.
-In doing so we now only need to compute the optical force and torque applied to our dimer, this has already been compiled in a MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007}.
-In which they use the results from \cite{Farsund1996} to compute both the optical force and torque using the beam coefficients - the form given by \cite{Crichton2000THEMD} is an easier form to compute:
+Fortunately, we do not need to solve these directly as the latter two
+random variables can be easily approximated if the thermal energy of
+the system is know, and the rate of change can be assumed as linear if
+we take a sufficiently small time step that
+$\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll
+\kappa_u/\gamma_R$. In doing so we now only need to compute the
+optical force and torque applied to our dimer, this has already been
+compiled in a MATLAB package called \textit{Optical Tweezer Toolbox}
+or \textit{ott} \cite{Nieminen2007}. In which they use the results
+from \cite{Farsund1996} to compute both the optical force and torque
+using the beam coefficients - the form given by
+\cite{Crichton2000THEMD} is an easier form to compute:
 \begin{equation}
 \begin{split}
-	\bold{F_z}=&-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right. \\ 
-	&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+  \bold{F_z}
+  &=
+    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right.
+  \\ 
+  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+  \\
   & \left. \times \Im[b_{n,m}b*_{n,m}+a_{n,m}a*_{n+1,m} - q_{n,m}q*_{n,m}+p_{n,m}p*_{n+1,m}] \right)
@@
 \begin{split}
-	\bold{T_z}=&-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right. \\ 
-	&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
+  \bold{T_z}
+  &=
+    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right.
+  \\ 
+  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+  \\
   & \left. \times \Re[b_{n,m}a*_{n,m}+a_{n,m}b*_{n+1,m} - p_{n,m}q*_{n,m}+q_{n,m}p*_{n+1,m}] \right)
@@
 \end{equation}
-Where $a_{n,m}$, $b_{n,m}$, $p_{n,m}$ $q_{n,m}$ are the beam coefficients of the incident and scattered fields respectively.
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients of the incident and scattered fields respectively.
 We can get the transverse force and torque components in a similar form by applying a simple rotation transformation.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulated Quadrant Photodiode}
+
 In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface.
@@
 \textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
 \begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
+	E_{\rm inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
 \end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
+where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
 For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
more updates

@@
 For a single plane wave the incident, internal, and scattered fields are given as:
-
 \begin{align}
-  E_{\rm inc}(r)
+  {\bf E}_{\rm inc}({\bf r})
   &=
-    E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r)
+    E_0 \sum_{nm}^\infty \left[
+    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
+    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
   \\
-  E_{\rm int}(r)
+  {\bf E}_{\rm int}({\bf r})
   &=
-    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right)
+    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
+    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
+    \right]
   \\
-  E_{scat}(r)
+  {\bf E}_{\rm scat}({\bf r})
   &=
-    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
+    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
+      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
+    \right]
 \end{align}
-where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user.
-However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on.
-From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.
+where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
+are the expansion coefficients of each of the fields, for the incident
+field computing its expansion coefficients is possible via analytical
+methods and are completely dependent on the beam conditions imposed by
+the user.  However, computing the expansion coefficients of the
+internal and external fields be far more tedious depending on the
+shape of the target and what properties of the scattered field are
+desired --- all of which is discussed later on.  From an optical
+trapping perspective the force exerted by a focused electric field can
+be found by computing the Maxwell stress tensor which only requires
+the total magnitude of the incident and scattered fields, essentially
+computing the momentum contained in the incident beam and how much of
+it has been transferred to the target particle.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Rayleigh Regime}
+
 The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
@@
 The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
-
 \begin{align}
-	F_{grad,x} &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-	F_{grad,y} &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r) \\
-	F_{grad,z} &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-	\left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \nonumber \times I(r) \\ 
-	& \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right] \\
-	\text{where:} \nonumber \\
+  F_{grad,x}
+  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
+  \\
+  F_{grad,y}
+  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
+  \\
+  F_{grad,z}
+  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
+    \nonumber \times I(r)
+  \\ 
+  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
+  \\
+  \text{where:}
+  \nonumber
+  \\
 	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-	exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
+	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
 \end{align}
@@
 Scattering force however is dependent on the effective scattering cross sectional area.
-
 \begin{align}
-	F_{scat} &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
+  F_{\rm scat}
+  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
   \text{where:} \nonumber \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Scattering methods}
 \label{sec:scattering}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Tmatrix}
-The T matrix method was first developed by Peter Waterman with his 
-research into acoustic wave scattering, this would later be extended to 
-electromagnetic waves. Put simply, the method sets a boundary condition 
-where the incident field within the interior field is cancelled out via 
-interference; these boundary conditions allows us to solve the Maxwell 
-equations by expanding the incident and scattered fields using spherical 
-vector waves.
-Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:
 
+The $T$-matrix method was first developed by Peter Waterman with his
+research into acoustic wave scattering, this would later be extended
+to electromagnetic waves. Put simply, the method sets a boundary
+condition where the incident field within the interior field is
+cancelled out via interference; these boundary conditions allows us to
+solve the Maxwell equations by expanding the incident and scattered
+fields using spherical vector waves. Because the boundary condition
+ensures that the incident field is halted at the interior surface of
+the object, the scattering field can be related to the incident field
+by the objects $T$-matrix:
 \begin{align}
@@
 
-The T-matrix method is exceptionally useful for computing the scattering from any arbitrary shaped particle.
-However, the T-matrix method by itself can be limited by computation power for aggregates of particles; while possible to compute the scattering of each individual particle the system of equations cannot be solved analytically and must be solved numerically to determine properties for the aggregate for one particular orientation.
+The $T$-matrix method is exceptionally useful for computing the
+scattering from any arbitrary shaped particle.  However, the
+$T$-matrix method by itself can be limited by computation power for
+aggregates of particles; while possible to compute the scattering of
+each individual particle the system of equations cannot be solved
+analytically and must be solved numerically to determine properties
+for the aggregate for one particular orientation.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Discrete Dipole Approximation}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Langevin Equation}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 \chapter{Complex Langevin dynamics of simple spherical aggregates}
+
+\comment{Do you actually examine aggregates larger than dimers?  If
+  not, then just change the title to reflect that you just examine dimers.}
+
 Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
@@
 Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu \textit{et.al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
+Theoretical studies on the case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
 Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Positional and Orientational dependence of Trapping forces}
+
 If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Continuous rotational motion due to second-order scattering}
-One aspect that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light.
-Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally.
-However, theoretical and experimental work by \cite{Yevick2017} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres.
-They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis.
-This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate.
-We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
+
+One aspect that has yet to be covered in depth with regards to
+spherical aggregates of any construction is their interaction with
+circularly polarised light.  Typically the spin density of an electric
+field cannot be reduced in homogenous medium due to the fact that the
+spin angular momentum is conserved locally.  However, theoretical and
+experimental work by \cite{Yevick2017} found that highly focused
+Gaussian beams could produce second order effects in the Rayleigh
+regime resulting in a photo-kinetic force that results in orbital
+motion about the beam's central axis. This effect is rather minimal
+for single sphere's, resulting in a orbital frequency on the order of
+$10^{-1}\,{\rm Hz}$, with an order of magnitude difference when
+trapping aggregates of spheres.  They computed the circulation rate by
+computing the time-average probability flux; however, when extended to
+the Mie regime we see a completely different behaviour, instead
+experiencing an optical torque about their long axis.  This rotation
+was first noted by Vigilante and co-workers who only considered this
+behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of
+simulations for differently sized dimers in a circularly polarised
+trap and looked at the rotation rate.  We found that not only is the
+rotation rate dependent on the size of the dimer, but also on its
+orientation and therefore their axial position.
 
@@
 
-Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion.
-We used \textit{mstsm} to look at the stokes parameters from the scattered field from a simple plane wave incident on our dimer, the proportion of circularly polarised light is minimal compared to the proportion of plane polarised light, which indicates that this rotational motion is not due to any inhomogeneity in the dimer that might impart angular momentum to the scattered beam - as compared to a anisotropic scatterer like Vaterite.
+It is difficult to see from the graph, but the rotation rate never
+truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which would
+imply that a second sphere of radius $200\,{\rm nm}$ is enough to
+induce rotational motion.  We used MSTM to look at the stokes
+parameters from the scattered field from a simple plane wave incident
+on our dimer, the proportion of circularly polarised light is minimal
+compared to the proportion of plane polarised light, which indicates
+that this rotational motion is not due to any inhomogeneity in the
+dimer that might impart angular momentum to the scattered beam - as
+compared to a anisotropic scatterer like Vaterite.
 
-These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction.
-The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}.
-This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector.
-Furthermore, we see a non-linear increase in the rotational speed of our dimers with size, the drag torque from the surrounding fluid is $~\propto r^3$ so the expectation is that the rotation frequency should fall off with increasing size.
-This indicates that the rotational motion is due to the shape asymmetry of the dimer and not solely due to the beam's angular momentum.
-Measurement of this photo-kinetic force is difficult to achieve due to the fact that previous analysis was conducted in the Rayleigh regime, where the polarizability of our dimer can be approximated as:
+These results are somewhat contrary to other work with silica dimers
+\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
+have trapped the dimer in an orientation perpendicular to the beam
+propagation direction.  The rotational motion is induced due to the
+asymmetric geometry creating an unbalanced polarisation susceptibility
+along its long axis as compared to its short axis; therefore its long
+axis is aligned with the polarisation vector and can rotate
+freely\cite{Ahn2018}.  This however cannot be the case with our
+simulations as the dimer rotates about its long axis, meaning there
+cannot be an asymmetric axis to align with the beam's polarisation
+vector. Furthermore, we see a non-linear increase in the rotational
+speed of our dimers with size, the drag torque from the surrounding
+fluid is $\propto r^3$ so the expectation is that the rotation
+frequency should fall off with increasing size.  This indicates that
+the rotational motion is due to the shape asymmetry of the dimer and
+not solely due to the beam's angular momentum.  Measurement of this
+photo-kinetic force is difficult to achieve due to the fact that
+previous analysis was conducted in the Rayleigh regime, where the
+polarizability of our dimer can be approximated as:
 \begin{align}
-	\bold{p}(r,t) = \alpha_xE_x(r,t)\hat{x}_n +\alpha_yE_y(r,t)\hat{y}_n +\alpha_zE_z(r,t)\hat{z}_n
+  \bold{p}({\bf r},t)
+  =
+  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
+  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
+  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
 \end{align}
-Where the polarizability is given as a 3D vector for the three principle Cartesian directions.
-In order to measure the magnitude of second order contributions we would need to construct a dipole array that fully captures the scattering of a dimer.
+where the polarizability is given as a 3D vector for the three
+principle Cartesian directions.  In order to measure the magnitude of
+second order contributions we would need to construct a dipole array
+that fully captures the scattering of a dimer.
 
+\comment{It would be good to have a plot of the torque in each of the
+  three axes as a function of the polarization of the trapping beam,
+  from linearly polarized to elliptically polarized to circularly polarized.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}

### 2024-06-08 Latex_files/Thesis/chapters/chapter 2/main.tex
addressing comments and added figures to chapter 4

@@
 
-These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.
+These 4 equations describe how the electric and magnetic fields behave and are related to one another at a microscopic level.
+Any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied at the macroscopic level, where one must account for the medium's interactions with the EM field.
+Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
+\begin{align}
+	\nabla \cdot \mathbf{D}
+	&= \rho_f
+	\\
+	\nabla \cdot \mathbf{B}
+	&= 0
+	\\
+	\nabla \times \mathbf{E}
+	&= \frac{\partial \mathbf{B}}{\partial t}
+	\\
+	\nabla \times \mathbf{H}
+	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
+\end{align}
+Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current caused by the presence of free charges in the medium, and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and magnetizing fields respectively.
+The latter two are defined by:
+\begin{align}
+	\nonumber
+	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
+	\\ 
+	\nonumber
+	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
+\end{align}
+Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the magnetisation field, these fields arise due to the bound charges throughout the medium interacting with the EM field.
 
@@
 \end{align} 
-This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
+Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$) This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
 For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
@@
     \right] 
+    \\
+	\mathbf{E}_{tot}({\bf r}) &= 
+	\begin{cases}
+		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
+		\\ 
+		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
+	\end{cases}
 \end{align}
@@
 \begin{align}
-	x'(t) &= \frac{x(t+\Delta t)-x(t)}{\Delta t}=\frac{dx(t))}{dt} \\
-	x^2(t) &= \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
+	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
+	\\
+	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
 \end{align}
@@
 \begin{align}
-	\left<U(x)\right> = \frac{1}{2}\kappa\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
+	\left<U(x)\right> = \frac{1}{2}\kappa_x\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
 	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
@@
 \cite{Crichton2000THEMD} is an easier form to compute:
+
 \begin{equation}
@@
   &=
-    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right.
+    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
   \\ 
@@
   \\
-  & \left. \times \Im[b_{n,m}b*_{n,m}+a_{n,m}a*_{n+1,m} - q_{n,m}q*_{n,m}+p_{n,m}p*_{n+1,m}] \right)
+  & \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
 \end{split}
@@
   &=
-    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right.
+    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
   \\ 
@@
   \\
-  & \left. \times \Re[b_{n,m}a*_{n,m}+a_{n,m}b*_{n+1,m} - p_{n,m}q*_{n,m}+q_{n,m}p*_{n+1,m}] \right)
+  & \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
 \end{split}
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{./figs/fixed_polarisation.png}
+	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
 \end{figure}

### 2024-06-08 Latex_files/Thesis/chapters/chapter 4/main.tex
addressing comments and added figures to chapter 4

@@
-\chapter{Complex Langevin dynamics of simple spherical aggregates}
+\chapter{Complex Langevin dynamics of spherical dimers}
 
@@
   not, then just change the title to reflect that you just examine dimers.}
-
 Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
@@
 second order contributions we would need to construct a dipole array
-that fully captures the scattering of a dimer.
+that fully captures the scattering of a dimer. Measuring the optical torque 
+makes it clear that the polarizability is a contributing factor to this optical 
+rotation phenomena. Rotating a symmetric dimer in the $x-z$ plane reveals 
+that while the dimer can be rotated in an orientation perpendicular 
+to the beam rotational torque is maximised when rotated while aligned with 
+the optical axis.
+
+\begin{figure}[t]
+	\begin{subfigure}{\linewidth}
+	\centering
+	\includegraphics[width=0.6\linewidth]{tx_different_polarisations.png}
+	\end{subfigure}
+	\begin{subfigure}{\linewidth}
+	\centering
+	\includegraphics[width=0.6\linewidth]{ty_different_polarisations.png}
+	\end{subfigure}
+	\begin{subfigure}{\linewidth}
+	\centering
+	\includegraphics[width=0.6\linewidth]{tz_different_polarisations.png}
+	\end{subfigure}
+\caption{Calculation of the optical torque about the three primary axis on a symmetric dimer in linear, elliptical, and circular polarisation beams.}
+\end{figure}
+\newpage
 

### 2024-06-10 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 \end{figure}
-
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
+\newpage
+\subsection{Non-trivial harmonic traps}\label{sec:off-axis}
+Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
 However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
@@
 the optical axis.
-
-\begin{figure}[t]
-	\begin{subfigure}{\linewidth}
-	\centering
-	\includegraphics[width=0.6\linewidth]{tx_different_polarisations.png}
-	\end{subfigure}
-	\begin{subfigure}{\linewidth}
-	\centering
-	\includegraphics[width=0.6\linewidth]{ty_different_polarisations.png}
-	\end{subfigure}
-	\begin{subfigure}{\linewidth}
-	\centering
-	\includegraphics[width=0.6\linewidth]{tz_different_polarisations.png}
-	\end{subfigure}
-\caption{Calculation of the optical torque about the three primary axis on a symmetric dimer in linear, elliptical, and circular polarisation beams.}
-\end{figure}
+\begin{SCfigure}
+	\includegraphics[width=0.7\linewidth,height= 22cm]{torque_different_polarisations.png}
+	\caption{Optical torque against polar angle $\theta$ about the three primary axis (top: torque about the x-axis; middle: torque about the y-axis; bottom: torque about the z-axis) on a symmetric dimer in linear, elliptical, and circular polarisation beams.
+Diagram to the right is for visual clarity about the direction of $\theta$.
+\vspace{7.5cm}}
+\end{SCfigure}
 \newpage
 
-\comment{It would be good to have a plot of the torque in each of the
-  three axes as a function of the polarization of the trapping beam,
-  from linearly polarized to elliptically polarized to circularly polarized.}
+\subsection{Gyroscopic Precession using asymmetric dimers}
+As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there is the potential for non-vertical trapping orientations in which the dimer is located in a harmonic trap.
+When trapped by a circularly polarised light these dimers exhibit gyroscopic precession, demonstrating rotational motion not just about its long axis, but also around the optical axis of the beam.
+This gyroscopic motion has been demonstrated previously in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed for micron scale aggregates, typically because it was assumed that an off-axis orientation would not be stable due to the magnitude of the electric field.
+
+\comment{Need to add figure here demonstrating the rotational behaviour but not sure how to best convey that motion.
+Perhaps just a simple trajectory diagram showing the particles motion (position in the first column, $u_x$ in the second, then $u_y$, then $u_z$).}
 

### 2024-06-10 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 \chapter{Complex Langevin dynamics of spherical dimers}
-
-\comment{Do you actually examine aggregates larger than dimers?  If
-  not, then just change the title to reflect that you just examine dimers.}
 Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
@@
 An example for a dimer of size ratio 2 is shown below:
-
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
+	\begin{subfigure}{0.67\linewidth}
+		\includegraphics[width=\linewidth]{off_axis_trap.png}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
+
+		\caption{}
+	\end{subfigure}
+	\captionsetup{hangindent=2.35cm}
 	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
+The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.
+Right hand render shows a dimer in its off-axis configuration}
 \end{figure}
-
-Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.
+Interestingly while the trap strength of these off-axis traps are 
+similar in magnitude to the vertically aligned dimers, but when 
+the laser power is lowered (around 5 mW) the traps become metastable 
+resulting in the dimer escaping from after some random time while 
+within the trap. If the overall potential depth could be characterised 
+then dimers placed into this orientation could be used as a 
+micro-scale temperature alarm, where by fine tuning of the dimer's 
+parameters would allow you to construct a potential well that can 
+only be escaped when the local fluid temperature exceeds a certain maximum value.
 
@@
 field cannot be reduced in homogenous medium due to the fact that the
-spin angular momentum is conserved locally.  However, theoretical and
+spin angular momentum is conserved locally. In which case the only way
+to transfer angular momentum to use an object that is anisotropic.
+This is seen most clearly with birefringent materials, due to fact that
+the crystal lattice itself is anisotropic, a circularly polarised beam
+will transfer an optical torque that is proportional to the difference in 
+the refractive index of the particle. The optical torque is given as:
+\begin{equation}
+	\label{eq:opt_torque}
+	\begin{aligned}
+		\tau_{opt} = \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
+	\end{aligned}
+\end{equation}
+Where $\Delta n$ is the difference in refracive indices of the 
+particle, and $\phi$ indicates the phase shift in the electric field
+(for circular polarised light $\phi= \pi/4$) However, theoretical and
 experimental work by \cite{Yevick2017} found that highly focused
@@
 the Mie regime we see a completely different behaviour, instead
-experiencing an optical torque about their long axis.  This rotation
+experiencing an optical torque about their long axis.S This rotation
 was first noted by Vigilante and co-workers who only considered this
@@
 \subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there is the potential for non-vertical trapping orientations in which the dimer is located in a harmonic trap.
-When trapped by a circularly polarised light these dimers exhibit gyroscopic precession, demonstrating rotational motion not just about its long axis, but also around the optical axis of the beam.
-This gyroscopic motion has been demonstrated previously in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed for micron scale aggregates, typically because it was assumed that an off-axis orientation would not be stable due to the magnitude of the electric field.
+As mentioned in section~\ref{sec:off-axis} for specificity sized dimers 
+there is the potential for non-vertical trapping orientations in which 
+the dimer is located in a harmonic trap. When trapped by a circularly 
+polarised light these dimers exhibit gyroscopic precession, demonstrating 
+rotational motion not just about its long axis, but also around the 
+optical axis of the beam. This gyroscopic motion has been demonstrated 
+previously in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
+but has not been observed for micron scale aggregates. Since the torque is 
+computed by evaluating the beam coefficients it can be difficult to 
+apply this result to micro-rheology experiments as one would need to know
+the exact magnitude of the optical torque ahead of time in order to make
+estimations about the local fluid viscosity. This is trivial for a birefringent
+spherical particle, less so for spherical aggregates whose final position and 
+orientation are unknown.
 
@@
 
+\section{Characterisation of asymmetric dimers via PSD analysis}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-06-10 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers.
-The scattered and incident fields are then combined
+The scattered and incident fields are then combined together in the far field to get $I(x,y)$, the quadrant and overall signals are calculated via:
+\begin{align}
+	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
+	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
+	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
+	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
+\end{align}
+Where the denominator is the total intensity on the QPD while there is no particle within the trap.
+The QPD sensitivity is dependent on both the polarisation of the incident beam and the displacement of the target sphere, this is shown in fig.~\ref{fig:totalfield}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.85\linewidth]{fixed_polarisation.png}
+	\captionsetup{margin=0.5cm}
+	\caption{Total Field incident on the quadrant photo diode for a several different displacements.
+Top Left: $2\mu m $ diameter silica sphere placed at the focus of a  beam.
+Top Right: silica sphere is now displaced by $1 \mu m$ along the x-axis.
+Bottom Right: silica sphere is now displaced by $1 \mu m$ along the y-axis.
+Bottom Left: silica sphere is now displaced by $0.707 \mu m$ along the x and y axis.
+The trapping beam is a $TEM_00$ Gaussian, polarised along the along the x-axis.
+$S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
+	\label{fig:totalfield}
+\end{figure}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
-\end{figure}
+Rather than displacing and rotating the particle itself, we instead rotate and displace the incident beam and recalculate the beam coefficients; the total beam is then inversely rotated in order to record the QPD signal.
+With this any trajectory can be collected from the QPD by displacing and rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss how to extract information from the QPD's outputted signal.

### 2024-06-10 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 \chapter{Complex Langevin dynamics of spherical dimers}
+\label{chapter:langevin_dynamics}
 Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.

### 2024-06-11 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 \chapter{Theory and methods}
-
+\label{chapter:theory}
 The two key areas of this project cover two widely different fields; 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Discrete Dipole Approximation}
+%\subsection{Discrete Dipole Approximation}
 
@@
 \begin{align}
+	\label{eq:alaised_lorentzian}
 	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
@@
 \subsection{Simulated Quadrant Photodiode}
-
+\label{sec:simulated_QPD}
 In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface.

### 2024-06-11 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 		\centering
-		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
+		\raisebox{-45pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
 		\label{large over small}
@@
 		\centering
-		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
+		\includegraphics[width=0.3\linewidth,
+		keepaspectratio]{theta=180.png}
 		\label{small_over_large}
@@
 	\label{fig:transverse_force}
-	\includegraphics[width=0.67\linewidth]{transverse_force.png}
+	\includegraphics[width=\linewidth]{transverse_force.png}
 	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
@@
 		\includegraphics[width=\linewidth]{off_axis_trap.png}
-		\caption{}
 	\end{subfigure}
 	\begin{subfigure}{0.32\linewidth}
-		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
+		\raisebox{-110pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
 
@@
 \end{figure}
+
 Interestingly while the trap strength of these off-axis traps are 
@@
 \end{equation}
-Where $\Delta n$ is the difference in refracive indices of the 
+Where $\Delta n$ is the difference in refractive indices of the 
 particle, and $\phi$ indicates the phase shift in the electric field
@@
   \centering
-  \includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
+  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
   \caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
@@
 \begin{SCfigure}
-	\includegraphics[width=0.7\linewidth,height= 22cm]{torque_different_polarisations.png}
+	\includegraphics[width=0.65\linewidth,height= 15cm]{torque_different_polarisations.png}
 	\caption{Optical torque against polar angle $\theta$ about the three primary axis (top: torque about the x-axis; middle: torque about the y-axis; bottom: torque about the z-axis) on a symmetric dimer in linear, elliptical, and circular polarisation beams.
-Diagram to the right is for visual clarity about the direction of $\theta$.
-\vspace{7.5cm}}
+Diagram to the right is for visual clarity about the direction of $\theta$.\vspace{4.5cm}}
 \end{SCfigure}
@@
 \subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity sized dimers 
-there is the potential for non-vertical trapping orientations in which 
-the dimer is located in a harmonic trap. When trapped by a circularly 
-polarised light these dimers exhibit gyroscopic precession, demonstrating 
-rotational motion not just about its long axis, but also around the 
-optical axis of the beam. This gyroscopic motion has been demonstrated 
-previously in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
-but has not been observed for micron scale aggregates. Since the torque is 
-computed by evaluating the beam coefficients it can be difficult to 
-apply this result to micro-rheology experiments as one would need to know
-the exact magnitude of the optical torque ahead of time in order to make
-estimations about the local fluid viscosity. This is trivial for a birefringent
-spherical particle, less so for spherical aggregates whose final position and 
+As mentioned in section~\ref{sec:off-axis} for specificity 
+sized dimers there is the potential for non-vertical trapping 
+orientations in which the dimer is located in a harmonic 
+trap. When trapped by a circularly polarised light these 
+dimers exhibit gyroscopic precession. As shown in fig~\ref{fig:gyro} 
+the dimer's is rotating about it's long axis (as seen by 
+the periodic behaviour of $u_x$ and $u_y$), while also precessing 
+around the optical axis of the beam, resulting in a second 
+order oscillation in $s_z$.
+\begin{figure}[h]
+	\label{fig:gyro}
+	\centering
+	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
+	\caption{3 second trajectory of a dimer trapped in an off axis orientation, laser power was set to 100 mW.
+The far left column depicts the dimer's centre of mass position with time; middle two columns are the $x$, $y$, and $z$ components of the vectors $u_x$ and $u_y$; last column depicts the components of the vector $u_z$ which defines the dimer's orientation.}
+\end{figure}
+
+Applying a Fourier analysis to the above trajectory reveals the 3
+fundamental frequencies typically associated with precession; 
+the $u_{z,1}$ and $x(t)$ series show a precession frequency of
+$~1.33\ Hz$ whereas the series $u_{x,1}$ and $u_{y,1}$ show a 
+combined periodic signal - a rotational frequency of $~23\ Hz$
+and a nutation frequency of $~20\ Hz$. Previous studies into 
+amorphous silica nanoparticles found a linear relationship 
+with the rotational frequency and the laser power, but no such
+relationship existed with the precession frequency. Our own 
+results shows a similar linear relationship with vertically
+aligned dimers.
+\begin{figure}
+	\includegraphics[width=\linewidth]{RotationvsPower.png}
+	\caption{Rotational frequency vs laser power for a dimer in both vertical orientations.}
+\end{figure}
+
+The linear relationship could partly be due to fact that we do not
+account for the change in viscous forces with increasing
+laser power. It is far more likely that the rotational frequency
+reaches a maximum value assuming that the bulk fluid can readily 
+absorb the laser.
+
+This gyroscopic motion has been demonstrated previously 
+in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
+but has not been observed for micron scale aggregates. 
+Since the torque is computed by evaluating the beam coefficients 
+it is be difficult to apply this result to micro-rheology 
+experiments as one would need to know the exact magnitude 
+of the optical torque ahead of time in order to make 
+estimations about the local fluid viscosity. This is 
+trivial for a birefringent spherical particle, less 
+so for spherical aggregates whose equilibrium position and 
 orientation are unknown.
 
-\comment{Need to add figure here demonstrating the rotational behaviour but not sure how to best convey that motion.
-Perhaps just a simple trajectory diagram showing the particles motion (position in the first column, $u_x$ in the second, then $u_y$, then $u_z$).}
+\section{Characterisation of asymmetric dimers via PSD analysis}
+As discussed in \ref{sec:simulated_QPD}, one of the methods
+developed to work in conjunction with \cite{Vigilante2020} is 
+a simulated quadrant photo diode for as a position detection
+system. While it is possible to extract all of the relevant 
+dynamical information from a simulation, confirming the same 
+behaviour in an experimental setting can be challenging.
 
-\section{Characterisation of asymmetric dimers via PSD analysis}
+As a benchmark we start by considering a single sphere within 
+an optical trap. After a 3 second trajectory (the typical time
+necessary to collect a power spectrum) the sphere's power 
+spectra was recorded and a fitted to eq.~\ref{eq:alaised_lorentzian}.
+\begin{figure}[h]
+	\label{fig:psd_sphere}
+	\centering
+	\includegraphics[width=\linewidth]{PSD_sphere.png}
+	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian}, scattered points represents
+	the blocked data ($n_b=100$). Corner frequency for the 
+	Lorentzians are reported in the legend.}
+\end{figure}
+
+As shown in fig.~\ref{fig:psd_sphere}, the two power spectra
+report different corner frequencies which would be indicate 
+that the trap is not perfectly circular. Using \textit{ott} 
+we can compare the expected trap strength to what is reported
+by a quadrant photo diode:
+\begin{center}
+	\begin{tabular}{ |c|c|c|c|c| } 
+		\hline
+		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
+		\hline
+		$f_c\ [Hz]$ & 447 & 450 & 439 & 474 \\
+		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 \\
+		\hline
+		Ellipticity &
+		\multicolumn{2}{|c|}{8.16 \%} &
+		\multicolumn{2}{|c|}{27.17 \%} \\
+		\hline
+	\end{tabular}
+\end{center}
+
+Where the ellipticity of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
+the beam wavefront. Its clear from these initial results that the 
+QPD is more sensitive to changes along the y-axis than the x-axis 
+when compared to the direct \textit{ott} calculations. Typically,
+even an industrial Gaussian beam will produce an elliptical diffraction
+limit spot when heavily focused; in their tutorial for optimizing 
+the PSD analysis, Berg and Sorensen reported a ellipticity of
+around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. If we now consider a symmetric the 
+discrepancy between \textit{ott} and the simulated QPD grows.
+
+\begin{center}
+	\begin{tabular}{ |c|c|c|c|c| } 
+		\hline
+		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
+		\hline
+		$f_c\ [Hz]$ & 409 & 334 & 431 & 424 \\
+		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 \\
+		\hline
+		Ellipticity &
+		\multicolumn{2}{|c|}{42.8 \%} &
+		\multicolumn{2}{|c|}{12.7 \%} \\
+		\hline
+	\end{tabular}
+\end{center}
+
+Now we see that the \textit{ott} predicts a more elliptical 
+trap compared to the QPD model which says the trap is far
+more symmetrical while trapping a symmetric dimer. This 
+change can be partially explained by the fact that rotational
+effects are not accounted for in the Lorentzian power spectra,
+only translational motion. Typically, rotational motion is only
+ever detected when it is periodic (take for example a power spectra
+from a rotating birefringent sphere), when the motion is stochastic
+the entire power spectra is effected making it near impossible
+to separate the translational and rotational contributions from a 
+single power spectra.
 
@@
 \section{Conclusions}
+\comment{Conclusions still need to be written out}

### 2024-06-13 Latex_files/Thesis/chapters/chapter 1/main.tex
updates chapter 1

@@
 \chapter{Introduction}
-Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food production, and 
-electronics; where the extraction of dilute materials can help improve 
-product quality while keeping production costs low. The industrialisation 
-of crystallisation has allowed engineers to reliably and efficiently 
-induce the crystal formation within a bulk liquid. However, while on a 
-large scale crystallisation is seemingly an understood physical process, 
-just a small amount of investigation into the literature reveals that at 
-a micro level there is yet to be unifying theory that can accurately 
-explain the process of crystallisation, more specifically nucleation.
+Crystallisation is one of the oldest forms of phase separation 
+used by humanity\cite{Schoen1956}, put simply, it is the 
+formation and growth of a new structured phase within 
+a disordered bulk phase. This has applications in a 
+number of industries such as pharmaceuticals, food 
+production, and electronics \cite{Myerson2002}; where 
+the extraction of dilute materials can help improve 
+product quality while keeping production costs low. 
+The industrialisation of crystallisation has allowed 
+engineers to reliably and efficiently induce the crystal 
+formation within a bulk phase. However, while on a 
+large scale crystallisation is seemingly an understood 
+physical process, just a small amount of investigation 
+into the literature reveals that at a micro scale there 
+is yet to be unifying theory that can accurately explain 
+the process of crystallisation \cite{Fu2021}. Of which 
+there two primary areas of focus, nucleation and crystal 
+growth. The latter focusing on how an already stable 
+crystal grows and how it takes on its final shape; 
+whereas the former is more concerned with which factors 
+contribute to its formation in order to control crystal growth.
 
-Nucleation is the initial formation of the crystalline phase and can be 
-broadly divided into two main processes. Firstly there is primary 
-nucleation, in which the nucleus forms from the bulk liquid; and 
-secondary nucleation, where the presence of already existing crystals 
-promotes the formation and growth of new nuclei
+The focus of this research is partly based around investigating
+the impact of a focused electromagnetic field can influence the 
+nucleation process by inducing shear flow within the bulk phase. 
+This in hope that the localised EM field can be used to study 
+different phenomena in nucleation in order to develop a better
+understanding of the underlying mechanisms.
+
+\section{Nucleation}
+Nucleation is an example of a binary phase separation, 
+where a dilute phase is miscible in a bulk phase, more 
+often called the solute and solvent respectively. Because 
+of thermodynamics, the two can only remain miscible while 
+below a certain threshold concentration ($C_{eq}$) 
+- the chemical potential $\mu$ for a miscible solution 
+is greater than the potential required to separate 
+the two phases. Once $C_{eq}$ is exceeded the there is
+a potential difference driving the solution to separate the 
+two phases. Since different combinations of solute and solvent
+will have different threshold concentrations, researchers often 
+instead measure the ratio between the solute and solvent by using 
+'supersaturation', one form of measuring supersaturation is shown 
+below \cite{Mullin2001}:
+\begin{align}
+	S(T) = \frac{C_{solute}}{C_{eq}(T)}
+\end{align}
+Where $C_{solute}$ is just the raw solute concentration, $C_{eq}(T)$
+is the threshold concentration at temperature $T$.
+While the solution remains supersaturated there is a 
+chemical potential driving force for the solute 
+to coalesce and separate from the solution as an ordered solid, 
+the first formation of the crystal is referred to as the 
+nucleus and understanding its formation of has been the 
+focus of researchers for decades now. In particular 
+there is a growing desire for a unified theory to 
+describe the multiple paths that nucleation can take.
+
+\subsection{Primary vs Secondary nucleation}
+From an industrial perspective, the nucleation process can
+be broadly categorised into either primary of secondary 
+nucleation. The former being described as the formation of an
+initial nucleus within the bulk phase; absent of any external
+stimuli primary nucleation is considered stochastic and is typically
+undesirable for industrial applications where consistent 
+residence times are desirable.
+
+In contrast, secondary nucleation is the result of a initial 
+seed crystal inducing further nucleation within the bulk solution
+\cite{Botsaris1976}, in absence of a seed crystal nucleation is
+near impossible. The research of Secondary Nucleation is fascinating
+for its own merits, not only is it speculated that the supersaturation 
+barrier is non-existent \cite{Cashmore2022}
 
 \section{Nucleation Theories}
-Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully.
-While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now.
-In particular there is a growing desire for a concrete theory to describe the process of nucleation.
+\subsection{Classical Nucleation Theory (CNT)}
+Sometimes referred to as 'Gibs Nucleation Theory' the 
+original theory was first formed from the works of Volmer 
+and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
+While initially it was more focused into describing droplet 
+formation in condensing vapours it was extrapolated to describe
+crystallisation. The central premise of classical theory 
+is that nucleation occurs stochastically throughout the bulk
+phase due to collisions between individual solute molecules,
+ions, or atoms; at the same time the bulk phase is resistant
+to the formation of a new phase. The competition between
+these random collisions and the bulk solution can be used to 
+predict the probability of a newly formed nucleus.
 
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}.
-While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation.
-The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute.
-Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus.
-Therefore the free energy change of a nucleus of size $r$ is given by:
+Consider a supersaturated solution, after some time enough individual
+sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
+The newly formed phase has a lower chemical potential than the 
+surrounding solution, reducing the free energy of the system; 
+at the same time, the formation of a new interface is resited 
+by the bulk phase due to surface tension. The net free energy 
+of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
 \begin{align}
-	\Delta G = \Phi r - \gamma r^{2/3}
+	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
 \end{align}
-From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive.
-This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution.
-With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
+Where $v$ is the approximate volume of an individual molecule, $k_B$
+is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
+tension of the bulk solution. Plotting the free energy of the system
+against nucleus size reveals a critical size above which the gain 
+in free energy exceeds the interfacial tension. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
+	\caption{Free energy diagram of a newly formed nucleus according 
+		     to the Classical Nucleation Theory. The total free energy (blue)
+		     is due to the competition between the volume free energy gain
+		     (green) and the interfacial free energy cost (orange). Dotted
+		     lines are for a higher supersaturation than the solid lines,
+		     the interfacial energy cost is independent of supersaturation.}
+\end{figure}
+
+The maximum value of $\Delta G_{tot}$ is the free energy barrier 
+that any newly formed nucleus needs to overcome in order to stabilise.
+The nucleation rate (the volume of new crystalline material formed per
+unit time), is therefore commonly defined as being dependent on the 
+energy barrier $\Delta G^*$:
 \begin{align}
@@
 \end{align}
-Where A is a prefactor that accounts for kinetic contributions.
-One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple.
-Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation).
-In addition the theory implies that any collection of solute molecules below the critical size are unstable.
+Where $A$ is a pre-factor that can be fine tuned to the exact demands
+of the system (typically involving Zeldovich Factor, $Z$) the free 
+energy barrier can be found by finding the turning point of $\Delta G_{tot}$.
 
-Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process.
-However, in practice there are a number of challenges to applying this to experimental measurements.
-
-Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
-\begin{align}
-	\Delta G = \gamma n^{2/3}-\phi n \\
-	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
-	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
-	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
-\end{align}
 So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
@@
 	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
+		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 \cite. 
 		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.

### 2024-06-13 python_files/test.py

@@ -1,25 +1,36 @@
-import csv
-import random
-from tkinter import X
 import numpy as np
 import matplotlib.pyplot as plt
 
-step_size = 1
-q1_dat, q2_dat, q3_dat, q4_dat = [], [], [], []
-t_list = []
-with open('.\\python_files\\tmp\\data_14-May-202413_58_16.txt', 'r') as input:
-    lines = input.readlines()
-    for line in lines[1:-1:step_size]:
-        print(line)
-        items =  line.split(',')
-        q1_dat.append(float(items[0]))
-        q2_dat.append(float(items[1]))
-        q3_dat.append(float(items[2]))
-        q4_dat.append(float(items[3]))
-        t_list.append(float(items[4]))
-print(q1_dat)
-Q = [[q1, q2, q3, q4] for q1, q2, q3, q4 in zip(q1_dat, q2_dat, q3_dat, q4_dat)]
+gamma = 79.97e-3
+v_mol = 1e-30
+kT = 1.38e-23*298
+r_list = np.linspace(0, 1.6e-10, 1000)
+r_star1 = 2*v_mol*gamma/(kT*np.log(1.5))
+r_star2 = 2*v_mol*gamma/(kT*np.log(1.8)) 
+def free_energy(r, S):
+	surface = 4*np.pi*gamma*r**2
+	volume = (4*np.pi*r**3/(3*v_mol))*kT*np.log(S) 
+	return (surface - volume)/kT, surface/kT, volume/kT
 
-print(Q)
-test = [q.index(max(q))+1 for q in Q]
-print(test)
+plt.plot(r_list, [free_energy(r,1.5)[0] for r in r_list],
+		 c='blue', label = r'$\Delta G_{tot}$ S = 1.5' )
+plt.plot(r_list, [free_energy(r,1.8)[0] for r in r_list],
+		 linestyle= '--', c='blue',label =r'$\Delta G_{tot}$, S = 1.8' )
+plt.plot(r_list, [free_energy(r,1.5)[1] for r in r_list], 
+		 c='orange', label =r'$\Delta G_{inf}$')
+plt.plot(r_list, [-free_energy(r,1.5)[2] for r in r_list],
+		 linestyle= '--',c='green', label =r'$\Delta G_{vol}$, S = 1.5')
+plt.plot(r_list, [-free_energy(r,1.8)[2] for r in r_list], 
[…]

### 2024-06-14 Latex_files/Thesis/chapters/chapter 2/main.tex
updates chapter 1
updates

@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.67\linewidth]{finite_differences.png}
+	\caption{Example trajectory created using Finite Differences method
+			 for a $2\mu m$ diameter sphere.
+Trap stiffness's were estimated using \textit{ott} at $\kappa_x = \kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$.
+The particle's motion can be localised around the shaded ellipsoid.}
+\end{figure}
+
 \section{Calibration Techniques}
@@
 \end{align}
-Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector.
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a sphere, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector.
 $\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
@@
 Bottom Left: silica sphere is now displaced by $0.707 \mu m$ along the x and y axis.
-The trapping beam is a $TEM_00$ Gaussian, polarised along the along the x-axis.
+The trapping beam is a $TEM_{00}$ Gaussian, polarised along the along the x-axis.
 $S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}

### 2024-06-14 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 the PSD analysis, Berg and Sorensen reported a ellipticity of
-around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. If we now consider a symmetric the 
+around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. The reason for this discrepancy
+can be explained partly by the fact that our estimation of the 
+trap geometry via \textit{ott} is based on the the actual 
+interactions with the beam near close to the focus whereas
+a QPD is estimating the trap geometry by extrapolating from
+far field scattering signals. If we now consider a symmetric the 
 discrepancy between \textit{ott} and the simulated QPD grows.
@@
 trap compared to the QPD model which says the trap is far
-more symmetrical while trapping a symmetric dimer. This 
-change can be partially explained by the fact that rotational
+more symmetrical while trapping a symmetric dimer. The reason that
+\textit{ott} no longer expects a circular trap could be due to
+how it computes the beam shape coefficients; by point matching in 
+the far field before the focus means a loss in accuracy for objects
+that trap above the focus. The change in the QPD estimation  
+can be partially explained by the fact that rotational
 effects are not accounted for in the Lorentzian power spectra,

### 2024-06-25 Latex_files/Thesis/chapters/chapter 1/main.tex
updates to ch1 and ch3

@@
 often called the solute and solvent respectively. Because 
-of thermodynamics, the two can only remain miscible while 
-below a certain threshold concentration ($C_{eq}$) 
-- the chemical potential $\mu$ for a miscible solution 
+of thermodynamics, the two can only remain in equilibrium while 
+below a specific concentration ($C_{eq}$) - below which 
+the chemical potential $\mu$ for a miscible solution 
 is greater than the potential required to separate 
-the two phases. Once $C_{eq}$ is exceeded the there is
-a potential difference driving the solution to separate the 
+the two phases. Once $C_{eq}$ is exceeded there is a chemical
+ potential difference driving the solution to separate the 
 two phases. Since different combinations of solute and solvent
-will have different threshold concentrations, researchers often 
+will have different equilibrium concentrations, researchers often 
 instead measure the ratio between the solute and solvent by using 
@@
 \begin{align}
-	S(T) = \frac{C_{solute}}{C_{eq}(T)}
+	\label{eq:supersaturation}
+	S(T) = \frac{C_{sol}}{C_{eq}(T)}
 \end{align}
-Where $C_{solute}$ is just the raw solute concentration, $C_{eq}(T)$
-is the threshold concentration at temperature $T$.
+Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
+is the equilibrium concentration at temperature $T$.
 While the solution remains supersaturated there is a 
@@
 nucleus and understanding its formation of has been the 
-focus of researchers for decades now. In particular 
-there is a growing desire for a unified theory to 
-describe the multiple paths that nucleation can take.
+focus of researchers for decades now. Typically, for an 
+industrial crystallisation process the working principle 
+is based on controlling and manipulating the supersaturation
+of the system.
 
@@
 stimuli primary nucleation is considered stochastic and is typically
-undesirable for industrial applications where consistent 
-residence times are desirable.
+undesirable for industrial applications where short and consistent 
+residence times are desirable. At a small scale one can estimate the 
+nucleation rate by making repeated measurements of sample solutions 
+and seeing how many have nucleated after a given time, giving us a 
+Poisson distribution.
+
+\begin{align}
+	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
+\end{align}
+
+Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$
+is the volume of the individual samples, and $M^*(t)$ \& $M$ are the 
+number of nucleated samples and the total number of samples used respectively.
+While this is useful for studying the effects of different parameters
+at a small scale, for industrial applications there are so many
+additional factors at play that can throw off this estimation.
 
@@
 near impossible. The research of Secondary Nucleation is fascinating
-for its own merits, not only is it speculated that the supersaturation 
-barrier is non-existent \cite{Cashmore2022}
+for its own merits; is it speculated that the supersaturation 
+barrier is non-existent \cite{Cashmore2022}, and furthermore there 
+are a multitude of mechanisms that can induce secondary 
+nucleation making it imperative that researchers can control the
+nucleation rate at an industrial scale.
+
+\begin{figure}[h!]
+	\includegraphics[width=\linewidth]{secondary_nucleation.jpg}
+	\caption{Secondary Nucleation mechanisms, classified by Agrawal and Paterson \cite{Agrawal2015}}
+\end{figure}
 
@@
 \begin{figure}[h!]
+	\label{fig:free_energy}
 	\centering
@@
 \subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.
-It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards.
-This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state.
-Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation.
-The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.
+The two step nucleation theory is an extension to the CNT, colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}.
+\textit{in situ} techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.
+It can be understood by Oswald's rule \cite{Ostwald1897}, which says that any crystallising system does not
+take the path to the lowest possible energy state but instead first transitions to the state with the smallest free energy barrier, further transitions can still occur but this method minimises the overall free energy cost.
+If we consider the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider a case where the supersaturation is low ($S\approx1.01$) then a dense liquid droplet will have a lower chemical potential than the surrounding fluid
+
+\section{Crystallisation methods}
+\subsection{Cooling Crystallisation}
+For some binary mixtures the supersaturation is heavily dependent on 
+the solution temperature, therefore a simple method of producing crystals 
+is by cooling the mixture to induce crystal formation. At ambient 
+temperatures the solute concentration is too high to be fully incorporated
+into the solution ($S\gg 1$), after heating however the solute is fully dissolved
+into the solution ($S<1$). Now as the mixture is allowed to cool to room
+temperature the supersaturation will increase until crystal formation begins,
+the rate of cooling drastically influencing the size and number of crystals
+produced.
+
+If $dT/dt$ is high then the final product will consist of large 
+crystal and be low in number, as the nucleation rate is directly related
+to the supersaturation only a handful of nuclei can form before the remaining
+solute grows onto the surface. If $dT/dt$ is low then the final product will 
+consist of smaller crystals and be far more numerous, as the supersaturation 
+is so large that the new nuclei are forming continuously. Between these two 
+extremes, one can define the meta-stable zone width, a region in the 
+temperature-concentration phase space where both nucleation and crystal growth
+can be reliably controlled. The lower limit being given by the solubility curve
+of the binary mixture, and the upper limit being defined by the metastable zone.
+
+\begin{figure}[h!]
+	\label{fig:MSZW}
+	\includegraphics[width=\linewidth]{MSZW.png}
+	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red).
+3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
+\end{figure}
+
+The viability of cooling crystallisation is dependent on the meta-stable 
+zone width, too narrow and the process is difficult to control, too wide 
+and the crystal growth rate may be insufficient for the desired outcome.
 

### 2024-06-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch1 and ch3

@@
 These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
+In this chapter I outline the optical tweezer equipment used during this PhD, the initial attempts made to induce nucleation via optical rotation, and the impact of a moving beam on the crystal growth of a newly formed nucleus.
 
 \section{Optical Tweezer Equipment}
-In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals.
-The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser.
-Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture.
-To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications.
-In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification.
-Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging.
-By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations.
-While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement.
-The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane.
-Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm.
-To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.
+In general, all optical tweezers require a laser driver, two
+microscope objectives (one for trapping and one for imaging),
+and a means of controlling the position of the loaded sample. 
+While there are other pieces of equipment used in modern tweezer
+experiments these are the bare minimum requirements for any 
+optical tweezer. The laser used for this project was a 1064 
+nm near infrared laser - provided by CNI Lasers – that has 
+an adjustable power supply to vary the energy output of the 
+laser - the remaining optics where supplied by Thor Labs.
+Experimental work has shown that the trapping efficiency 
+increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
+\cite{kim2003dependence} where $D_{obj}$ is the diameter
+of the objective aperture. To expand the beam front we 
+utilise a Galilean beam expansion arrangement (indicated 
+by $L_1$, and $L_2$) as recommended for high power laser 
+applications. In our initial experiments the beam expansion 
+provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then 
+the 4f correlator provides a further $1.25\times$ magnification.
+It should be noted that the galvano-mirror requires the use 
+of a Keplerian beam expansion arrangement which reduces
+the transmitted laser power due to localised heating of the air.
+Afterwards the laser is passed through a dichroic mirror that 
+separates incoming infrared and visible light, this is to 
+prevent the laser from damaging the CCD camera used for imaging
+the trapping plane. The laser is then focused to a diffraction
+limited spot by a 1.25 NA objective. By increasing the numerical
+aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth 
+is reduced due to spherical aberrations. While it is possible 
+to increase the trapping depth \cite{Reihani2006} by adjusting 
+the objective's tube length this approach is incompatible with 
+our trapping arrangement. A 0.25 NA condenser objective refocuses
+the scattered laser light and also provide an aperture for an 
+imaging LED to illuminate the focal plane. Samples are loaded 
+onto a piezo driven table to that is inserted between the trapping
+and condensing objectives; the piezo drivers allow for sub-micron
+control of the sample position to a degree as small as a 10 nm. 
+To detect and monitor the position of a trapped particle a 
+quadrant photo diode (QPD) was utilised.
+
+\begin{figure}
+	\centering
+	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
+	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively.}
+\end{figure}
 
 \subsection{Position detection methods}
-There are several methods for monitoring the position of 
+In order to accurately capture the dynamics of a trapped particle, 
+a position detection system must be utilised. As shown 
 A QPD is a typical example of a position detection system commonly used 

### 2024-06-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 utilise a Galilean beam expansion arrangement (indicated 
-by $L_1$, and $L_2$) as recommended for high power laser 
+by $f_1$, and $f_2$) as recommended for high power laser 
 applications. In our initial experiments the beam expansion 
-provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then 
-the 4f correlator provides a further $1.25\times$ magnification.
+provides a $4\times$ magnification; however, in later experiments 
+where we utilise a galvano-mirror the beam expansion is $3\times$ 
+and then the 4f correlator provides a further $1.25\times$ magnification 
+(using $f_3$ and $f_4$) - where the magnification is given by.
+\begin{align}
+	\frac{D_2}{D_1} = \frac{f_2}{f_1}
+\end{align}
 It should be noted that the galvano-mirror requires the use 
@@
 limited spot by a 1.25 NA objective. By increasing the numerical
-aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth 
-is reduced due to spherical aberrations. While it is possible 
-to increase the trapping depth \cite{Reihani2006} by adjusting 
-the objective's tube length this approach is incompatible with 
-our trapping arrangement. A 0.25 NA condenser objective refocuses
+aperture of the objective, the gradient force at the focal point 
+is increased; the trade-off being that the for higher NA values 
+the trapping depth is reduced due to spherical aberrations.
+While it is possible to increase the trapping depth \cite{Reihani2006} 
+by adjusting the objective's tube length this approach is incompatible 
+with our trapping arrangement. A 0.25 NA condenser objective refocuses
 the scattered laser light and also provide an aperture for an 
@@
 \begin{figure}
+	\label{fig:setup}
 	\centering
 	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
-	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively.}
+	\caption{Optical tweezer set up used for the majority of the PhD.
+The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively.
+Diagram not drawn to scale.}
 \end{figure}
@@
 In order to accurately capture the dynamics of a trapped particle, 
-a position detection system must be utilised. As shown 
+a position detection system must be utilised. While possible to characterise
+an optical trap using a linear position detector
 A QPD is a typical example of a position detection system commonly used 

### 2024-06-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 In order to accurately capture the dynamics of a trapped particle, 
-a position detection system must be utilised. While possible to characterise
-an optical trap using a linear position detector
-A QPD is a typical example of a position detection system commonly used 
-in optical tweezers due to their high sampling rate, high degree of 
+a position detection system must be utilised. There are 3 possible
+methods of position detection: video-analysis, lateral-effect
+position sensing, and quadrant photodiodes. The former being 
+ideally suited for multiple traps or situations where precision
+is not the top priority. In order to match the force measurements
+of back-focal plane interferometry requires the camera's frame
+rate to exceed $1\ kHz$ which can be difficult to achieve while 
+maintaining a decent resolution \cite{Gibson2008}. In comparison
+off the shelf back-focal plane detectors can achieve temporal
+resolutions anywhere from $10-100\ kHz$.
+
+A QPD is frequently used position detection system for optical 
+tweezers due to their high sampling rate, high degree of 
 precision, and ease of set up. The QPD is constructed of four photo 
 diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is directed onto the QPD, with the maximum 
-intensity mapping to the particle's centre of mass. By summing the 
-horizontal and vertical quadrants together the particle's centre of mass 
-is tracked in the x-y plane. While the tracking is accurate the QPD 
-needs to be calibrated in order to convert the voltage signals into 
-actual distance reading.
+interference pattern produced is focused onto the QPD, with 
+the maximum intensity mapping to the particle's centre of mass. 
+By summing the voltages of the horizontal and vertical quadrants 
+together the particle's centre of mass is tracked in the x-y 
+plane. Axial displacement can be estimated by observing the change
+in the total voltage of the QPD. The outputted signal gives an
+indication of the particle's relative displacement from the beam 
+focus, but in order to convert the signal to distance units the
+trap needs to be calibrated (assuming a linear response curve).
+
+A lateral-effect sensor has a similar output but works using a 
+the entire photodiode as a single cell. The four corners of 
+the sensor act as anodes connected to a base plate cathode, as 
+the beam moves across the surface of the detector the 4 corners
+detect a photocurrent and
 

### 2024-06-26 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
+	\caption{Comparison between QPD and Lateral effect photodiodes.
+	The four quadrants of a QPD (left) experience different photocurrents
+	based on the total intensity of light incident on each section 
+	(labelled A, B, C, D). 
+	Whereas a Lateral effect sensor (right) uses the resistive properties
+	of the photodiode surface to vary the create different photocurrents 
+	passing through the anodes A, B, C, and D.}
+\end{figure}
+
 A lateral-effect sensor has a similar output but works using a 
-the entire photodiode as a single cell. The four corners of 
-the sensor act as anodes connected to a base plate cathode, as 
-the beam moves across the surface of the detector the 4 corners
-detect a photocurrent and
+the entire sensor as a single cell analogous to the focal plane of
+the trapping beam. The four corners of the sensor act as anodes 
+connected to a base plate cathode, as the beam moves across the
+surface of the detector each anode will experience a different 
+photocurrent depending on how close the centre of the 
+interference pattern is to each anode. The advantage of a 
+lateral effect detector is that the linear regime is much larger 
+than a QPD making it much better for monitoring the position of a
+trapped particle. However, Lateral-effect sensors are often limited
+in their spacial resolution due to high signal-to-noise ratios, 
+requiring a high intensity of light on the sensor in order to 
+get a clean signal. As a result, most optical trapping experiments
+are conducted using a QPD as opposed to a lateral-effect sensor, 
+as often the displacement is small enough that the QPD response 
+curve can be considered linear.
+
+\section{Calibration of Tweezer Setup}
+Prior to testing optical rotation the trapping laser was calibrated 
+using power spectra analysis
 
@@
 \section{Shear induced Nucleation}
+\comment{This belongs in the introduction chapter me thinks}
+
 It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
@@
 
-Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium.
-Typically one would use a birefringent particle (i.e.
-vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam.
-Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere.
-Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field?
-This requires that we both know the fluid velocity, and the shear rate applied by our probe particle.
-Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.
+Optical tweezing has often been used for micro-rheology, by computing the
+exact forces being exerted on the trapped sphere, one can determine the
+local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
+Typically one would use a birefringent particle (i.e. vaterite, liquid
+crystals, etc) and rotating it within the fluid, the maximum rotation
+rate being a product of the fluid drag resisting the torque of the 
+trapping beam \cite{RodriguezSevilla2018}. Or if you want to measure 
+fluid flow you can instead use a micro-rotor to see how fluid flow 
+propagates in the medium \cite{Knoener2005}. Likewise, one can use 
+a galvanometric mirror to probe the drag force of the fluid, by 
+understanding the trap strength (calibrating using a low frequency 
+signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I
 
-Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
+Understanding the fluid velocity around our trapped object is determined 
+mostly by the Reynold's number of our system, for a sphere submersed 
+in a moving fluid of velocity $U$ this is given by:
 \begin{align}
@@
 With the particle size and rotation frequency, the tangential rotation speed is calculated via:
-
 \begin{align}
@@
 
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre.
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere.
-The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):
-
+Where $d$ is the particle diameter, $\omega$ is the rotation frequency
+reported by the QPD, and $r$ is the distance from the particle's centre. 
+Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating
+outward from the centre of the sphere. The shear rate can then be computed
+as the partial derivative fluid flow (assuming shearing is generated purely
+by the flow field):
 \begin{align}
@@
 
-\begin{figure}[h]
+Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
+(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of supersaturated solutions of Glycine and water. After locating a singular microsphere the
+vaterite was trapped in a circularly polarised light and brought close
+to the droplet edge. Due to the increased viscosity of the supersaturated 
+solution and the proximity to the droplet edge, the observed rotation rate 
+was rather low or non-existent, the results are tabulated below:
+
+\begin{table}
+	\centering
+	\caption{Results from rotating vaterite within supersaturated solution}
+	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
+		\hline
+		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
+		\hline
+		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
+		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
+		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
+		\hline
+		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
+		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
+		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
+		\hline
+		\multirow{3}*{1.49} & 4.76 & $~0$ & $\times$ \\
+		\cline{2-4} & 7.27 & $~0$ & $\times$ \\
+		\cline{2-4} & 1.52 & $~0$ & $\times$ \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{figure}[h!]
 	\label{fig:vaterite_shear}
@@
 
-From Fig.\ref{fig:vaterite_shear} it's clear that
+From Fig.\ref{fig:vaterite_shear}~(b) it's clear that the rotational speeds
+achieved both in the lab and from experimental literature are well below the 
+theoretical maximum shear rate reported by \cite{Debuysschere2023} as you move 
+away from the surface of your micro-rotor.
+
+This would explain partly why we could never seen nucleation even in the case where our vaterite could rotate freely.
+The closest we could trap a microsphere to the droplet edge was in the range of $5-10\mu\ m$, at that distance the fluid flow would be so low that not even using a liquid droplet rotor would achieve the rotational speeds necessary to localise nucleation.
 

### 2024-06-26 Latex_files/Thesis/main_win.tex
updates

--- 2024-06-07 Latex_files/Thesis/main.tex
@@
 \usepackage{graphicx}
-\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
 
+\graphicspath{{C:/Users/xbb16146//dmaciver97/Latex_files/Thesis/figs}}
+\usepackage{sidecap}
 % Page Margins - Strath Requirement
@@
 % Draft Watermark
-\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
+%\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}
+
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+\newcommand{\comment}[1]{{\bf\textcolor{red}{#1}}}
+% \newcommand{\comment}[1]{{{#1}}}
 

### 2024-06-28 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 There are several options for particles that can be rotated using 
-optical tweezers []. Over the course of this project two different micro 
+optical tweezers \cite{Parkin2009, Saito2022}. Over the course of 
+the PhD two different micro 
 spheres where investigated, vaterite and liquid crystal droplets. Both 
@@
 rotation.
+
+Vaterite samples where made by the first preparing equal amounts of 
+$CaCl_2$ and $NaCO_3$ at a concentration of $0.1M$, at the same time
+a vial of $0.03M\ MgSO_4$ was prepared and set aside for later. First
+a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
+and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
+Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
+$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
+After 10 minutes of slow but continuous mixing a few drops of Agepon was added
+to halt the reaction, the solution was filtered and washed 3 times with 
+distilled water before being suspended in water.
 
@@
 Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.
+This has been shown theoretically for both simple colloidal \cite{Mura2016,Debuysschere2023,Richard2015}  and ice crystal formation \cite{Goswami2020}; however, no experimental work into these systems has been conducted to prove this is the case.
 There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
+In \cite{Debuysschere2023} it was found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
+Using the theoretical model established in \cite{Mura2016,2001} which modifies the CNT to account for the effects of a nucleus undergoing shearing, accounting for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
 There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
@@
 
-Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
-(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of supersaturated solutions of Glycine and water. After locating a singular microsphere the
-vaterite was trapped in a circularly polarised light and brought close
-to the droplet edge. Due to the increased viscosity of the supersaturated 
-solution and the proximity to the droplet edge, the observed rotation rate 
-was rather low or non-existent, the results are tabulated below:
-
-\begin{table}
-	\centering
-	\caption{Results from rotating vaterite within supersaturated solution}
-	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
-		\hline
-		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
-		\hline
-		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
-		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
-		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
-		\hline
-		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
-		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
-		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
-		\hline
-		\multirow{3}*{1.49} & 4.76 & $~0$ & $\times$ \\
-		\cline{2-4} & 7.27 & $~0$ & $\times$ \\
-		\cline{2-4} & 1.52 & $~0$ & $\times$ \\
-		\hline
-	\end{tabular}
-\end{table}
+First we tested the rotational behaviour of vaterite in distilled water, 
+samples of vaterite where diluted down and $200\mu L$ was pipetted onto
+the sample stage. Due to Van der Waal's forces some of the microspheres 
+were stuck together, fortunately individual sphere's were still
 
@@
 away from the surface of your micro-rotor.
+
+Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
+(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of supersaturated solutions of Glycine and water. After locating a singular microsphere the
+vaterite was trapped in a circularly polarised light and brought close
+to the droplet edge, after a period of ten minutes if no nucleation event
+was observed the particle was released and another particle was trapped. 
+Due to the increased viscosity of the supersaturated solution and the 
+proximity to the droplet edge, the observed rotation rate 
+was rather low or non-existent, the results are tabulated below:
+
+\begin{table}
+	\centering
+	\caption{Results from rotating vaterite within supersaturated solution. Solubility concentration for Glycine was at $16^\circ$, $C^*=0.2016g/g$}
+	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
+		\hline
+		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
+		\hline
+		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
+		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
+		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
+		\hline
+		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
+		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
+		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
+		\hline
+		\multirow{3}*{1.29} & 3.47 & 0.00 & $\times$ \\
+		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
+		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
+		\hline
+		\multirow{3}*{1.40} & 6.32 & 0.00 & $\times$ \\
+		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
+		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
+		\hline
+		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
+		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
+		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
+		\hline
+	\end{tabular}
+\end{table}
 

### 2024-07-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 
+Vaterite is a polymorph of calcium carbonate that is rarely seen in 
+nature due to its low stability. However unlike its other polymorphs of 
+calcite and aragonite, when synthesised vaterite will typically form 
+small spherical particles making them ideal for optical trapping. 
+Synthesis of vaterite micro spheres requires fine control of the nucleation
+process in order to maintain polymorphic control.
+
 Vaterite samples where made by the first preparing equal amounts of 
-$CaCl_2$ and $NaCO_3$ at a concentration of $0.1M$, at the same time
-a vial of $0.03M\ MgSO_4$ was prepared and set aside for later. First
+$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
+a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
 a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
@@
 
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.4\linewidth}
+		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
+		\subcaption{}
+	\end{subfigure}
+	\begin{subfigure}{0.55\linewidth}
+		\includegraphics[width=\linewidth]{rotating_psd.png}
+		\subcaption{}
+	\end{subfigure}
+	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap.
+(b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
+\end{figure}
+
 \subsection{Rotation of Vaterite micro spheres}
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.
 

### 2024-07-04 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 		\raisebox{-110pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
-
 		\caption{}
@@
 by a quadrant photo diode:
+
 \begin{center}
-	\begin{tabular}{ |c|c|c|c|c| } 
+	\begin{tabular}{ |c|c|c|c|c|c|c| } 
 		\hline
-		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
 		\hline
-		$f_c\ [Hz]$ & 447 & 450 & 439 & 474 \\
-		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 \\
+		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
+		& \parbox{1.25cm}{\centering 523} & 513 \\
+		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
 		\hline
@@
 		\multicolumn{2}{|c|}{8.16 \%} &
-		\multicolumn{2}{|c|}{27.17 \%} \\
+		\multicolumn{2}{|c|}{27.17 \%} &
+		\multicolumn{2}{|c|}{13.8 \%} \\
 		\hline
@@
 
-Where the ellipticity of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
+Where the ellipticity of the beam is given by 
+$e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
 the beam wavefront. Its clear from these initial results that the 
@@
 the PSD analysis, Berg and Sorensen reported a ellipticity of
-around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. The reason for this discrepancy
+around 15 \% after a total calibration time of 81 seconds 
+\cite{BergSoerensen2004}. The reason for this discrepancy
 can be explained partly by the fact that our estimation of the 
@@
 \begin{center}
-	\begin{tabular}{ |c|c|c|c|c| } 
+	\begin{tabular}{ |c|c|c|c|c|c|c| } 
 		\hline
-		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
 		\hline
-		$f_c\ [Hz]$ & 409 & 334 & 431 & 424 \\
-		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 \\
+		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
+		& \parbox{1.25cm}{\centering 274} & 285 \\
+		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
 		\hline
@@
 		\multicolumn{2}{|c|}{42.8 \%} &
-		\multicolumn{2}{|c|}{12.7 \%} \\
+		\multicolumn{2}{|c|}{12.7 \%} & 
+		\multicolumn{2}{|c|}{13.8 \%} \\
 		\hline

### 2024-07-04 python_files/test.py

@@ -1,36 +1,34 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import quaternion
+from quaternion import as_spherical_coords
 
-gamma = 79.97e-3
-v_mol = 1e-30
-kT = 1.38e-23*298
-r_list = np.linspace(0, 1.6e-10, 1000)
-r_star1 = 2*v_mol*gamma/(kT*np.log(1.5))
-r_star2 = 2*v_mol*gamma/(kT*np.log(1.8)) 
-def free_energy(r, S):
-	surface = 4*np.pi*gamma*r**2
-	volume = (4*np.pi*r**3/(3*v_mol))*kT*np.log(S) 
-	return (surface - volume)/kT, surface/kT, volume/kT
+with open('C:\\Users\\xbb16146\\dmaciver97\\python_files\\sphere_in_LCP_Quadrants.txt', 'r') as in_file:
+    lines = in_file.readlines()
+    x_dat, y_dat, z_dat, = [], [], []
+    q1, q2, q3, q4 = [], [], [], []
+    index = []
+    for line in lines[1:-1]:
+        items = line.split(',')
+        x_dat.append(float(items[0])+ float(items[1])- float(items[2]) - float(items[3])) 
+        y_dat.append(float(items[0])+ float(items[2])- float(items[1]) - float(items[3]))
+        z_dat.append(float(items[0])+ float(items[1])+ float(items[2]) + float(items[3]))
+        q1.append(float(items[0]))
+        q2.append(float(items[1]))
+        q3.append(float(items[2]))
+        q4.append(float(items[3]))
+        index.append(int(float(items[4])/1e-5))
 
-plt.plot(r_list, [free_energy(r,1.5)[0] for r in r_list],
-		 c='blue', label = r'$\Delta G_{tot}$ S = 1.5' )
-plt.plot(r_list, [free_energy(r,1.8)[0] for r in r_list],
-		 linestyle= '--', c='blue',label =r'$\Delta G_{tot}$, S = 1.8' )
-plt.plot(r_list, [free_energy(r,1.5)[1] for r in r_list], 
-		 c='orange', label =r'$\Delta G_{inf}$')
-plt.plot(r_list, [-free_energy(r,1.5)[2] for r in r_list],
-		 linestyle= '--',c='green', label =r'$\Delta G_{vol}$, S = 1.5')
[…]

### 2024-07-04 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

@@
 Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
+
+\subsection{Shear induced Nucleation}
+It has long been known that solution nucleation is influenced by shearing 
+of the fluid, from stirring agents in the container (i.e. propellers) to 
+the container boundary layers, however there has yet to be a clear mechanism 
+by which said shearing effects the nucleation event. Theoretical research 
+into shear induced nucleation suggests that there should be a slight increase 
+in the nucleation rate at low shear rates, reaching a maximum increase in 
+nucleation rate, and then at higher shear rates the nucleation rate begins 
+to drop off. This has been shown theoretically for both simple colloidal
+\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
+\cite{Goswami2020}; however, no experimental work into these systems has 
+been conducted to prove this is the case.  There is some experimental 
+evidence for this phenomena in simple salt and protein solutions - though 
+the authors emphasise that mechanical agitation cannot be ruled out - 
+there has not been a exhaustive study into the shearing effects apart from 
+in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
+rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
+the highest nucleation rate. Using the theoretical model established in 
+\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
+a nucleus undergoing shearing, accounting for the fact that a nucleus' 
+growth is undergoing competition between flow-mediated molecular transport 
+and the strain applied by the flow field which inhibits the growth of the 
+nucleus. There central conclusion (from both the theoretical and experimental 
+results) is that there is an optimal shear rate in which the nucleation 
+rate is maximised. However, a question that arises from this result, if 
+there is a optimal shear rate in which molecular transport is maximised 
+and strain is minimised, then surely there should also be a shear rate 
+in which the molecular transport and strain are equal - allowing one to 
+suspend a nucleus at a constant radius. In this scenario, the molecular 
+transport would prevent the nucleus from dissolving, but the strain would 
+prevent the nucleus from growing. This however would require one to be 
+able to apply a continuous shear rate to a targeted nucleus with high 
+precision, there is also no model for an individual nucleus undergoing growth.
+
+\section{Aims and Objectives}
+Overall the aim of the PhD is to study viability of using micro-rotors to
+generate localised fluid flow around the beam focus. The results are reported
+in chapter 3, this is then succeeded by experimental work where we 
+instead using a galvano-mirror to generate a moving beam focus within an 
+undersaturated solution to control the crystal growth front. In addition,
+computer simulations into asymmetric dimers and their applications as complex
+micro-rotors are discussed in chapter 4. Lastly two novel methods of measuring
+rotational motion are discussed in chapter 5; firstly via a novel detection
+fibre method that allows for instantaneous measurements of the orientational
+behaviour of optically trapped ellipsoids/dimers; and secondly we create 
+a simulative quadrant photo diode that replicates laboratory results, utilising 
+linear regression techniques we measure the change in orientation in order to
+measure the optical torque applied to a non-birefringent particle.

### 2024-07-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 optical tweezers \cite{Parkin2009, Saito2022}. Over the course of 
-the PhD two different micro 
-spheres where investigated, vaterite and liquid crystal droplets. Both 
-can be readily synthesised in the lab and are will rotate at a variety 
-of sizes. While silver nano particles were considered their high cost 
-and small size meant they were disregarded as an option for optical 
-rotation.
+the PhD two different micro spheres where investigated, vaterite 
+and liquid crystal droplets. Both can be readily synthesised in 
+the lab and are will rotate at a variety of sizes. While silver 
+nano particles were considered their high cost and small size 
+meant they were disregarded as an option for optical rotation.
 
@@
 calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping. 
+small spherical particles making them ideal for optical trapping and rotation. 
 Synthesis of vaterite micro spheres requires fine control of the nucleation
-process in order to maintain polymorphic control.
+process in order to maintain polymorphic stability, though for the purposes
+of optical rotation the exact polymorph is not as important as its morphology
+as all 3 polymorphs are inherently birefringent.
 
@@
 
+Liquid crystals are an intriguing example of materials with mixed phase
+properties, unlike typical solutes such as Glycine, a liquid crystal
+can still maintain some degree of order between its individual molecules
+while in the liquid state. This is due to the fact that liquid crystals
+are constructed of long ordered molecules that demonstrate a long range
+ordering. There are three main types of liquid crystal transition methods:
+thermotropic crystals will transition to their liquid crystal phase when 
+sufficiently heated; Lyotropic materials can undergo this transition due 
+to changes in temperature and concentration; and lastly Metallotropic 
+materials - which are composed of both organic and inorganic molecules - 
+change phase according to the ratio of organic to inorganic molecules present.
+Liquid crystal rotors are rather simple in their production, 
+4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
+and a small amount was added to a vial of distilled water. The solution 
+was then heated in a water bath to $25^\circ$ in order to transition the
+solid crystal into its liquid crystal state. The individual droplets 
+are inherently birefringent and would rotate readily while suspended in 
+distilled water.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\linewidth]{LC_sample.png}
+	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
+\end{figure}
 \subsection{Rotation of Vaterite micro spheres} 
-
-\section{Shear induced Nucleation}
-\comment{This belongs in the introduction chapter me thinks}
-
-It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e.
-propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event.
-Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off.
-This has been shown theoretically for both simple colloidal \cite{Mura2016,Debuysschere2023,Richard2015}  and ice crystal formation \cite{Goswami2020}; however, no experimental work into these systems has been conducted to prove this is the case.
-There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \cite{Debuysschere2023} it was found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-Using the theoretical model established in \cite{Mura2016,2001} which modifies the CNT to account for the effects of a nucleus undergoing shearing, accounting for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus.
-There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised.
-However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius.
-In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing.
-This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.
-
 Optical tweezing has often been used for micro-rheology, by computing the
@@
 \end{align}
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively.
-In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle.
-There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.
+Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
+density and viscosity respectively. In our case we do not have a fluid
+moving around a sphere but a sphere moving through the fluid at some 
+velocity $U$, assuming a no-slip boundary condition we can model the 
+fluid velocity profile based on the velocity of the particle. There are 
+two possible avenues for generating shear flow with a trapped particle; 
+rotation of birefringent particles, and fluid flow induced by particle motion.
 
-Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution.
-To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW.
-The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis.
-With the particle size and rotation frequency, the tangential rotation speed is calculated via:
+Rotating birefringent particles are by far the most common method for 
+generating and measuring fluid flow in a solution. To see if we can 
+even achieve the theoretical maximum shear rate, vaterite spheres 
+were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and 
+trapped with the 1064 nm laser at set to 450 mW. The rotation frequency 
+was determined using the QPD, and the particle sizes were computed 
+by image analysis. With the particle size and rotation frequency, the 
+tangential rotation speed is calculated via:
 \begin{align}
@@
 the sample stage. Due to Van der Waal's forces some of the microspheres 
-were stuck together, fortunately individual sphere's were still
+were stuck together, fortunately individual sphere's were still present.
+Over the course of a few weeks multiple sphere's where trapped 
+and their rotation speeds were measured via the QPD.
 
@@
 
+\section{Rotation of Vaterite in supersaturated solutions}
+
 Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
-(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of supersaturated solutions of Glycine and water. After locating a singular microsphere the
-vaterite was trapped in a circularly polarised light and brought close
-to the droplet edge, after a period of ten minutes if no nucleation event
-was observed the particle was released and another particle was trapped. 
-Due to the increased viscosity of the supersaturated solution and the 
-proximity to the droplet edge, the observed rotation rate 
+(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of 
+supersaturated solutions of Glycine and water. After locating a 
+singular microsphere the vaterite was trapped in a circularly 
+polarised light and brought close to the droplet edge, after a 
+period of ten minutes if no nucleation event was observed the 
+particle was released and another particle was trapped. Due to 
+the increased viscosity of the supersaturated solution and the 
+proximity to the droplet edge, this was done as previous work with
+optical tweezers have only seen nucleation near the solution-air interface
+\cite{Gowayed2021, Liao2022, Walton2018}. The observed rotation rate 
 was rather low or non-existent, the results are tabulated below:
-
 \begin{table}
 	\centering
-	\caption{Results from rotating vaterite within supersaturated solution. Solubility concentration for Glycine was at $16^\circ$, $C^*=0.2016g/g$}
+	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
 	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
@@
 
-This would explain partly why we could never seen nucleation even in the case where our vaterite could rotate freely.
-The closest we could trap a microsphere to the droplet edge was in the range of $5-10\mu\ m$, at that distance the fluid flow would be so low that not even using a liquid droplet rotor would achieve the rotational speeds necessary to localise nucleation.
+The calculations from fig.\ref{fig:vaterite_shear} indicate that even
+in pure water the rotational speed achieved by a microsphere is
+insufficient to achieve a significant shear rate more than a few microns
+from its surface. The closest we could trap a microsphere to the droplet
+edge was in the range of $5-10\mu\ m$, at that distance the fluid flow
+would be so low that not even using a liquid droplet rotor would achieve
+the rotational speeds necessary to localise nucleation. While in theory
+a sufficiently focused laser could rotate any microsphere to a fast enough 
+to match the speeds expected by \cite{Debuysschere2023} the localised intensity
+would be so large that even using $D_2O$ would see a significant increase in 
+temperature.
+
+\section{Shearing via a Galvano-mirror}
 

### 2024-07-11 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 
-Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser.
-From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.
-The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations.
-The differential forms of which are given below
-\cite{Jackson_1975}:
+Proper understanding of optical tweezing requires an understanding of
+how the trapped particle interacts with the trapping laser. From a 
+electromagnetism perspective the laser creates a spatially and 
+temporally coherent electric field that scatters light off of a 
+trapped particle.  The laws governing electric and magnetic fields 
+are summarised most succinctly via the Maxwell equations. The 
+differential forms of which are given below \cite{Jackson_1975}:
 \begin{align}
@@
 
-These 4 equations describe how the electric and magnetic fields behave and are related to one another at a microscopic level.
-Any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied at the macroscopic level, where one must account for the medium's interactions with the EM field.
+These 4 equations describe how the electric and magnetic fields behave
+and are related to one another at a microscopic level. Any discussion 
+of optical trapping is underpinned by the fact that in all cases the 
+Maxwell Equations must be satisfied at the macroscopic level, where 
+one must account for the medium's interactions with the EM field. 
 Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
@@
 \end{align}
-Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current caused by the presence of free charges in the medium, and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and magnetizing fields respectively.
+Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
+caused by the presence of free charges in the medium, and $\mathbf{D}$ 
+and $\mathbf{H}$ are the displacement and magnetizing fields respectively. 
 The latter two are defined by:
@@
 \end{align}
-Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the magnetisation field, these fields arise due to the bound charges throughout the medium interacting with the EM field.
+Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
+magnetisation field, these fields arise due to the bound charges 
+throughout the medium interacting with the EM field.
 
-The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported.
-The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point).
-The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion.
-The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.
+The force exerted by an optical tweezer can be subdivided into the 
+gradient and scattering components, for most modelling research this 
+is how the force fields are reported. The gradient force is a conservative 
+force brought about by the polarisation of dielectric materials, 
+anything around 10\% of the trapping wavelength will be attracted along 
+the gradient of the electric field to the region of greatest intensity 
+(for a simple Gaussian Beam this would be at the focal point). The 
+scattering force arises from the scattered field of the trapping beam 
+pushing the target object away from the focal point; unlike the gradient \\
+force, the scattering force is non-conservative and is loosely 
+proportional to the particle's Brownian motion. The equilibrium position 
+(where the mean squared displacement is minimised) is found when the 
+gradient force far exceeds the scattering force, interpreting and 
+calculating these forces is dependent on the ratio between the particle 
+size and the trapping wavelength.
 
@@
 
-The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law.
-Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam.
-Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge.
-As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
+The Ray-Optics model is the simplest to understand, the theory models 
+the laser as a collection of individual 'rays' that propagate and are 
+refracted according to Snell's Law. Based on the change in direction 
+momentum is transferred to the target particle; with rays closest to 
+the centre of the beam having greater intensity than those rays at 
+the very edge of the beam. Consider a particle struck by two rays in 
+a Gaussian beam (see Fig below), one coming close to the centre, and 
+the other ray coming from the edge. As each ray is refracted by the 
+target sphere a force is imparted onto it, the total force imparted 
+is given by:
 \begin{align}
@@
 
-The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere.
-The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves.
-For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
+The Lorenz Mie theory provides an exact solution to the maxwell equations 
+for the scattering caused by an isotropic sphere. The theory describes the 
+scattered wave given off by a dielectric sphere when incident by a plane 
+wave as a summation of partial spherical waves. For any spherical wave the 
+vector fields must solve the Helmholtz wave equation given by:
 \begin{align}
@@
 \end{align}
-Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$) This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease.
-For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another.
-For a single plane wave the incident, internal, and scattered fields are given as:
+
+Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
+ This combined with the constraints of Maxwell's Equations leaves 
+ very few exact solutions apart from spherical or planar waves, 
+ both of which can be converted between with relative ease. For 
+ example, a plane wave electric field can be expanded into spherical 
+ harmonics and likewise any spherical wave can be described as 
+ combination of plane waves offset from one another. For a single 
+ plane wave the incident, internal, and scattered fields are given as:
 \begin{align}
+	\label{eq:lornez_mie}
   {\bf E}_{\rm inc}({\bf r})
@@
   \\
+  \label{eq:internal}
   {\bf E}_{\rm int}({\bf r})
@@
 where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
-are the expansion coefficients of each of the fields, for the incident
+are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
+and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
 field computing its expansion coefficients is possible via analytical
@@
 
-Lorenz-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory.
-The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions.
-In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.
+Lorenz-Mie theory can be applied to describe the scattering from any 
+particle regardless of size, however as we either increase or decrease 
+the size of the target particles the infinite series converges allowing 
+one to ignore much of the tedium of Lorenz-Mie theory. The ray optics 
+model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. 
+In the latter case we can simplify the scattering problem by approximating 
+our target as a single dipole, focusing only its interactions with the 
+incident field.
 
@@
 
-The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size.
-The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}.
-The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
+The Rayleigh approximation is the limiting approximation for describing 
+a particles motion in an electromagnetic field who's wavelength is several 
+times greater than the particle's size. The underlying theory is that a \\
+dielectric sphere can be treated as a dipole while in the presence of the 
+electromagnetic field; in which case the scattering force is given simply 
+by the scattering of the induced dipole, and the gradient force is due 
+to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
+principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
+in MKS units as a restorative rectangular force field:
 \begin{align}
@@
 
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is the beam waist, and $a$ is the radius of the particle.
-Scattering force however is dependent on the effective scattering cross sectional area.
+Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
+the beam waist, and $a$ is the radius of the particle. Scattering force 
+however is dependent on the effective scattering cross sectional area. 
 \begin{align}
@@
 
-The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}).
-However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field.
-For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.
+The Rayleigh regime allows for easy computation of the gradient and 
+scattering forces due to the assumption that the particle is a point 
+dipole, so much so that higher order scattering problems can be 
+simplified by subdividing the particle into discrete dipoles (see 
+Sec~\ref{sec:scattering}). However as the target particle gets larger 
+this assumption fails to accurately describe the trapping force due 
+to the complexity in the gradient field. For particle sizes close to 
+the laser wavelength the scattered field is best described by Lorenz-Mie theory.
 
@@
 \label{sec:scattering}
-
+There are several methods available to calculate the scattered field 
+produced by a particle, while one can compute this from directly applying
+Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
+tedious for some applications.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Tmatrix}
-
+\subsection{T-matrix Method}
 The $T$-matrix method was first developed by Peter Waterman with his
-research into acoustic wave scattering, this would later be extended
-to electromagnetic waves. Put simply, the method sets a boundary
-condition where the incident field within the interior field is
-cancelled out via interference; these boundary conditions allows us to
-solve the Maxwell equations by expanding the incident and scattered
-fields using spherical vector waves. Because the boundary condition
-ensures that the incident field is halted at the interior surface of
-the object, the scattering field can be related to the incident field
-by the objects $T$-matrix:
+research into acoustic wave scattering \cite{Waterman1969}, this 
+would later be extended to electromagnetic waves. Sometimes being 
+referred to as the extended boundary condition method (ECBM), 
+the method replaces the scatterer with a series of surface currents
+over the targets surface, these currents are chosen so that the
+electric field outside is identical to the original problem \cite{Wriedt1998}. 
+This choice of currents negates the need to compute the internal fields
+which reduces the scattering problem to a system of linear equations 
+relating the incident beam coefficients to the scattered beam coefficients
+the relationship between each can be summarised as:
 \begin{align}
@@
 
-The $T$-matrix method is exceptionally useful for computing the
-scattering from any arbitrary shaped particle.  However, the
-$T$-matrix method by itself can be limited by computation power for
-aggregates of particles; while possible to compute the scattering of
-each individual particle the system of equations cannot be solved
-analytically and must be solved numerically to determine properties
-for the aggregate for one particular orientation.
+Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
+$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
+this PhD in particular we utilise an extension of the ECBM known 
+as multi-sphere $T$-matrix method (\textit{mstm}), developed by Mackowski
+\cite{Mackowski2011} the computational code computes the scattered field 
+from each sphere within the target cluster and the incident field itself, 
+this scattering between spheres converges to a final result but modern codes 
+truncate the calculations to fit a desired accuracy and computational time.
+
+The $T$-matrix method is exceptionally useful for computing the scattering 
+from any arbitrary spherical aggregate. However, the $T$-matrix method by 
+itself can be computationally taxing as the number of spheres increases 
+\cite{Mackowski2011}. While it is possible to solve for the scattering of
+each individual sphere this is only applicable for a single orientation and 
+can be even slower for large aggregates \cite{Mackowski1996, Xu1995}. The 
+benefit of \textit{mstm} is that the major scattering properties (scattering 
+and extinction cross sections, scattering matrices, etc) can all be computed 
+both for single orientations, or averaged over multiple orientations to determine
+the average scattering from the target particle.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Discrete Dipole Approximation}
+\subsection{Discrete Dipole Approximation}
+The discrete dipole approximation (DDA) is a general method that can 
+be applied to the scattering from particles of arbitrary composition
+and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
+the DDA method approximates the particle as being constructed of dipoles.
+Each dipole interacts with both the incident field and the scattered fields
+from every other dipole surrounding it. The resulting scattered field is 
+identical to the scattered field produced by direct integration of 
+Eq.~\ref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
+The integral form for the internal electric field inside a scatterer is 
+given as:
+\begin{align}
+	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
+	\chi(\mathbf{r'})\mathbf{E(r')}
+\end{align}
+
+Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
+defines the impulse response between two separate dipoles; and $\chi$ is the 
+susceptibility of the medium, which describes the degree of polarisation of
+the medium in the presence of an electric field. With the internal field 
+calculated the scattered field can be computed, one of the primary advantages
+of DDA over the T-matrix method is that the the composition of the target can 
+be changed freely. When comparing different computational scattering methods,
+the ECBM was found to be better suited for simulating the scattering of symmetric
+targets as the ECBM can use the targets symmetry to speed up calculations \cite{Wriedt1998},
+however when dealing with inhomogeneous media DDA is more efficient compared to
+ECBM.
 
@@
 \section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation.
-Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour.
-Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out.
-This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule.
-The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
+Describing any microscopic motion requires an understanding of a molecules
+diffusive behaviour, for the case of optical tweezers the most complete model 
+of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
+derivations are sufficient for macroscopic behaviours the Langevin equation 
+better describes the microscopic characteristics of any diffusive behaviour. 
+The Fickian model is not suitable for the applications of optical tweezers as 
+it assumes that all diffusive motion is driven by a gradient of molecular density
+\cite{Gillespie2012}, not only is this often not the case but it fails to describe 
+the motion of any single molecule, only the collective behaviour. Einstein's model 
+however considers the collisions experienced by individual molecules \cite{Gillespie2012a}, 
+if we assume that all collisions are random and only consider the behaviour 
+after a finite time step $\delta t$ then the individual collisions should 
+cancel out, of course trying to understand the dynamics over a shorter period
+of time is not described by Einstein's model as you would have to consider
+each collision in turn \cite{Gillespie2012, Gillespie2012}. This is insufficient 
+for our interests as an optical tweezers relaxation time (given by 
+$\tau = \kappa/\gamma$) is often so short that we must consider the individual 
+collisions experienced by any one molecule. The Langevin model of diffusion 
+therefore assumes that the net force on a particular particle is described 
+fully by these individual collisions \cite{Gillespie2012c}:
 \begin{align}
@@
 \end{align}
-Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force.
-The $F(t)$ is there for convention which accounts for any external forces acting on the particle.
-We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:
 
+Where the first term accounts for inertial forces, the second term accounts 
+for friction forces which counteract the particles current motion ($\gamma_0$ 
+is the friction coefficient), and the final term accounts for the random 
+Brownian force. The $F(t)$ is there for convention which accounts for any 
+external forces acting on the particle. We can say that the noise term 
+$W(t)$ has a Gaussian probability, with a correlation function of:
 \begin{align}
@@
 
-The Langevin model can be extrapolated to describe the diffusive behaviour of an overall system, but for this project we can instead consider the behaviour of some particle with a diffusion tensor $D$ suspended in a viscous fluid and spacially localised by an optical potential with trap strength $\kappa$.
-Assuming the only external force acting on our particle is the laser the net force should be exactly equal to force of the stochastic collisions due to the fluids thermal energy.
-If we focus our analysis when the particle is stably trapped and assume that the trap is harmonic we can model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$).
+The Langevin model can be extrapolated to describe the diffusive behaviour 
+of an overall system, but for this project we can instead consider the 
+behaviour of some particle with a diffusion tensor $D$ suspended in a 
+viscous fluid and spacially localised by an optical potential with trap 
+strength $\kappa$. Assuming the only external force acting on our particle 
+is the laser the net force should be exactly equal to force of the stochastic
+collisions due to the fluids thermal energy. If we focus our analysis when 
+the particle is stably trapped and assume that the trap is harmonic we can 
+model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
 The full Langevin equation for an optically trapped particle is therefore given as:
@@
 
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped particles, however the analytical solution of the Langevin equation requires integration of the white noise term making it difficult to simulate the trajectory of a given particle \cite{Volpe2013}.
-Instead it is often far easier to solve the equation numerically and apply use the analytical solution to calibrate and extract information about the particle and fluid, and how the two interact with the optical trap.
+Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped 
+particles, however the analytical solution of the Langevin equation requires 
+integration of the white noise term making it difficult to simulate the 
+trajectory of a given particle \cite{Volpe2013}. Instead it is often far 
+easier to solve the equation numerically and apply use the analytical 
+solution to calibrate and extract information about the particle and fluid, 
+and how the two interact with the optical trap.
 
 \subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms.
-If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
+The Finite Difference approach involves discretizing the time and spatial 
+elements in order to approximate the higher order terms. If we assume that 
+$x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then 
+we can use the Taylor series expansion to get:
 \begin{align}
@@
 \end{align}
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression.
-If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
+Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
+term n and the actual expression. If we limit our approach to the first 
+derivative only we find that for sufficiently small values of $R_1$ the 
+velocity and acceleration can be approximated as:
 \begin{align}
@@
 \end{align}
-By reversing the time step (i.e.
-use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e.
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$).
-The same cannot be done for white noise as no information is known about $W(t)$ at any time.
-We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range.
-Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
+By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
+velocity and acceleration based on the previous time steps we can 
+discretize the position by taking finitely small  time steps (i.e. 
+$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
+white noise as no information is known about $W(t)$ at any time. We 
+can instead say that the velocity of a Brownian particle should 
+approximate our white noise as a random walk, where at each new time 
+step the position changes randomly within a given range. Constricting 
+the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white 
+noise using the finite-difference approach as:
 \begin{align}
@@
 \end{align}
-Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes.
-We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$.
-Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
+Where $w_i$ is a random real number between -1 and 1, we can say that 
+it is normally distributed around 0 for simulation purposes. We can 
+rearrange this for $x_i$ to approximate the Brownian motion of a 
+particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
+Now in the case of an optical trap the restoration time scale is 
+given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
+is far greater than the characteristic time, therefore for simulation 
+purposes we can neglect the particle's inertia which allows us to 
+write the motion of an optically trapped particle as:
 \begin{align}
@@
 
-This result can be generalised for a 3-dimensional description of an optically trapped particle, where each Cartesian direction has its own unique characteristic restoration time.
-We see from the result that trajectory is dependent on only a handful of factors, the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal energy of the system $k_BT$ - with the latter two being related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$.
-Therefore by calculating these parameters to a high degree of precision allows one to get precise description of the forces experienced by a target particle, which in the past has been used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
+This result can be generalised for a 3-dimensional description of an 
+optically trapped particle, where each Cartesian direction has its 
+own unique characteristic restoration time. We see from the result 
+that trajectory is dependent on only a handful of factors, the trap 
+stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal 
+energy of the system $k_BT$ - with the latter two being related by 
+Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$. 
+Therefore by calculating these parameters to a high degree of precision 
+allows one to get precise description of the forces experienced by a 
+target particle, which in the past has been used for highly accurate 
+force transduction \cite{BergSoerensen2004, Smith2003}.
 
@@
 \label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the optical trap, each approach has its drawbacks and benefits so each option should be chosen based on what elements want to be characterized.
-The basis for each of these methods stems from the analytical solution of the Langevin equation:
+There are several approaches for calibrating and characterizing the 
+optical trap, each approach has its drawbacks and benefits so each 
+option should be chosen based on what elements want to be characterized. 
+The basis for each of these methods stems from the analytical solution 
+of the Langevin equation:
 \begin{align}
@@
 \subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center.
-Therefore a simple analysis method is to understand the height and width of said potential well.
+The Langevin equation for an optically trapped assumes that the trap 
+acts similar to a Hookean spring that creates a potential well about 
+its centre. Therefore a simple analysis method is to understand the 
+height and width of said potential well.
 
-Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap.
-This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations.
-With this in mind we can write the probability of finding the particle at position $x$ as:
+Potential analysis is a useful technique for estimating the strength 
+of an optical trap; this method assumes that the force acting on the 
+particle is purely conservative, an accurate presupposition if we 
+ignore the motion of the particle as it enters the trap. This is 
+because the scattering force is far more significant far away from 
+the potential well and is negligible if the trap strength is much 
+greater than the thermal fluctuations. With this in mind we can 
+write the probability of finding the particle at position $x$ as:
 \begin{align}
@@
 \end{align}
-Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
+Now assuming our laser acts as a Gaussian beam we should be able 
+to describe the probability distribution $\rho(x)$ centred at 
+some equilibrium position $x_0$:
 \begin{align}
@@
 \end{align}
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$.
-This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long.
-It also provides no information on the particle itself (i.e.
-the friction coefficient $\gamma$ and diffusion coefficient $D$).
+
+By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} 
+we can fit the potential well in order to determine the trap strength 
+$\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. 
+This has some limitations in that the large fluctuations can throw 
+off the fit meaning a longer acquisition time is necessary to properly 
+fit the potential well, making it difficult to characterise weakly 
+trapped particles who may not remain trapped for long. It also 
+provides no information on the particle itself (i.e. the friction 
+coefficient $\gamma$ and diffusion coefficient $D$).
 
 \subsection{Equipartition method}
-The Equipartition method is by far the fastest and simplest means for estimating the trap strength but unlike Potential Analysis is limited strictly to harmonic potentials.
-This can be often not the case for highly focused beams, as the trap strength can vary due to polarisation differences.
-Simply put we can use the equipartition theorem to relate the potential well to the particle's thermal energy using eq.~\ref{eq:prob_dist}:
+The Equipartition method is by far the fastest and simplest means 
+for estimating the trap strength but unlike Potential Analysis is 
+limited strictly to harmonic potentials. This can be often not the 
+case for highly focused beams, as the trap strength can vary due to 
+polarisation differences. Simply put we can use the equipartition 
+theorem to relate the potential well to the particle's thermal 
+energy using eq.~\ref{eq:prob_dist}:
 \begin{align}
@@
 
-By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength.
-Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects.
-Likewise with the potential-analysis route, no information is gleaned about the particle itself.
+By taking a time average over multiple trajectories to get 
+$\left<x-x_{eq}\right>$ we can get a fairly accurate estimation 
+of the trap strength. Because of this requires a time average of 
+the particle's displacement any large errors in the position 
+measurement can have knock-on effects. Likewise with the potential 
+analysis route, no information is gleaned about the particle itself.
 
 \subsection{Mean Square Displacement}
-Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles).
-This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available.
-As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
+Mean square displacement (MSD) is a common means of describing 
+the random motion of a given particle (or group of particles). 
+This is useful information if say for example we want to understand 
+reaction kinetics on the surface of a catalyst, if we know how 
+far its likely to move from the surface we can tell if its likely 
+to react when a catalytic site becomes available. As it pertains 
+to colloids, consider a suspension of silica spheres immersed in 
+a fluid undergoing Brownian motion (as described by the Langevin 
+Equation) so that:
 \begin{align}
@@
 \end{align}
-Where $\gamma$ is the objects friction coefficient which for spheres is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid.
-If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for.
-If we wish to understand the effects of a given external factor (i.e.
-an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.
 
-For each sphere we can record its position in the $x-y$ plane and measure its displacement from a set reference point; for example with an optical tweezer this could be the beam focus.
-We can measure the MSD by forming a 'window' between two points in time of the trajectory (i.e.
-$t \&  t+\tau$) and sliding this window along the entire trajectory length - to eliminate -ve displacements we take the square - we can then take the average of this series.
-Repeating over a range of time lags allows us to describe the MSD as a function of $\tau$:
+Where $\gamma$ is the objects friction coefficient which for spheres 
+is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white 
+noise variable that is directly related to the thermal energy of 
+the surrounding fluid. If the motion is truly random then we should 
+see an average displacement of 0 regardless of how long we measure 
+for. If we wish to understand the effects of a given external factor 
+(i.e. an electric field or localised heating), simply looking at 
+displacement will reveal nothing of value as its difficult to 
+differentiate between diffusive and a biased motion.
+
+For each sphere we can record its position in the $x-y$ plane and 
+measure its displacement from a set reference point; for example 
+with an optical tweezer this could be the beam focus. We can measure 
+the MSD by forming a 'window' between two points in time of the 
+trajectory (i.e. $t \&  t+\tau$) and sliding this window along the 
+entire trajectory length - to eliminate -ve displacements we take 
+the square - we can then take the average of this series. Repeating 
+over a range of time lags allows us to describe the MSD as a function 
+of $\tau$:
 \begin{align}
@@
 
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can expand out the squared term to get an analytical expression for the MSD as a function of time lags:
+If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can 
+expand out the squared term to get an analytical expression for the 
+MSD as a function of time lags:
 \begin{align}
@@
 \end{align}
-From this expression its evident that the mean squared displacement increases with larger values of $\tau$ until it reaches a maximum value as shown below by the dotted line.
-\begin{figure}
+
+From this expression its evident that the mean squared displacement 
+increases with larger values of $\tau$ until it reaches a maximum 
+value as shown below by the dotted line.
+\begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.67\linewidth]{MSD.png}
+	\includegraphics[width=\linewidth]{MSD.png}
 	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
@@
 \end{figure}
-The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ the particle is experiencing the harmonic potential described by the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle is said to be freely diffusing within the trap focus.
-Of course for a freely diffusing object the MSD will never reach a plateau value, comparing MSD's for different particles provides a simple visual indicator of the difference in trapping strength.
-The MSD method is an already very versatile analytical tool for diffusive motion, however it is rather slow in computing time meaning it is only really beneficial when a high degree of accuracy is required and shorter time resolutions are unavailable - such as using a quadrant photo diode instead or a high speed CCD.
+
+The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ 
+the particle is experiencing the harmonic potential described by 
+the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle 
+is said to be freely diffusing within the trap focus. Of course for 
+a freely diffusing object the MSD will never reach a plateau value, 
+comparing MSD's for different particles provides a simple visual 
+indicator of the difference in trapping strength. The MSD method is 
+an already very versatile analytical tool for diffusive motion, however 
+it is rather slow in computing time meaning it is only really beneficial 
+when a high degree of accuracy is required and shorter time resolutions 
+are unavailable - such as using a quadrant photo diode instead or a high 
+speed CCD.
 
 \subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, however there is yet to be a analytical expression for the MSAD of a freely diffusing particle.
-Vigilante \textit{et al} \cite{Vigilante2020} derived the upper limit of a dimer's MSAD along its long axis by assuming it was strongly trapped and so had limited angular motion, there expression gives:
+It is also possible to plot the angular MSD (MSAD) using simulative data, 
+however there is yet to be a analytical expression for the MSAD of a 
+freely diffusing particle. Vigilante \textit{et al} \cite{Vigilante2020} 
+derived the upper limit of a dimer's MSAD along its long axis by assuming 
+it was strongly trapped and so had limited angular motion, there expression 
+gives:
 \begin{align}
@@
 \end{align}
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because they couldn't solve the Einstein-Smoluchowski equation which describes the diffusion constant for dielectric particles.
-This would require a full description of a particle's electrical mobility and charge distribution - the latter could be achieved via a discrete dipole approximation, the former would be dependent on both the particle's position but relative orientation to the electric field.
+
+Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because 
+they couldn't solve the Einstein-Smoluchowski equation which describes the 
+diffusion constant for dielectric particles. This would require a full 
+description of a particle's electrical mobility and charge distribution 
+- the latter could be achieved via a discrete dipole approximation, the 
+former would be dependent on both the particle's position but relative 
+orientation to the electric field.
 
 \subsection{Power Spectrum Density (PSD)}
-The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources.
-Taking the Fourier transform of a particle's trajectory yields:
+The power spectral density (PSD) method is by far the most versatile 
+method for observing the dynamics of any object within an optical trap, 
+allowing for fast calibration times while also quickly filtering out 
+typical noise sources. Taking the Fourier transform of a particle's 
+trajectory yields:
 \begin{align}
@@
 \end{align}
-where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
+
+where $\hat{\eta}$ is the Fourier transform of the white noise, 
+where the values are exponentially distributed as opposed to being 
+normally distributed in the time domain \cite{BergSoerensen2004}, 
+the correlation function is given as:
 \begin{align}
@@
 \end{align}
-We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
+
+We can therefore ignore the white noise from our analysis by looking 
+at the spectral density of $\hat{x}(f)$: 
 \begin{align}
@@
 \end{align}
-We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$.
-Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
+
+We can fit the Lorentzian via a simplified geometric series 
+$S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion 
+coefficient (in arbitrary units) and the corner frequency $f_c$ 
+which is directly related to the trap strength via 
+$f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression 
+of the mean squared displacement we see two distinct regions, when 
+$f\ll f_c$ the PSD reaches a plateau value that when converted to 
+length units represents the maximum displacement the particle can 
+move beyond the focus, however when $f\gg f_c$ the PSD falls off 
+exponentially which denotes the particle is freely diffusing within 
+the beam focus.
 \begin{figure}[t]
 	\centering
-	\includegraphics[width=0.75\linewidth]{PSD.png}
+	\includegraphics[width=\linewidth]{PSD.png}
 	\caption{Example PSD fitted using Eq.~\ref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water.
@@
 \end{figure}
-The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased.
-Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
+
+The Lorentzian relationship is only valid for frequency terms up to 
+the Nyquist frequency (half of our sampling rate), this is because we 
+are only taking a finite sampling of the particle's trajectory meaning 
+the signal is aliased. Berg and Sorensen provide a suitable modified 
+Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
 \begin{align}
@@
 \end{align}
-Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}.
-Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary.
-Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length.
-To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units.
-Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.
-
+Where N is the total number of samples taken, $\Delta x \ \& \ c$ 
+have no direct physical interpretation and are defined in 
+\cite{BergSoerensen2004}. Further modifications can be made to the 
+power spectrum model but this is only useful when a high degree of 
+accuracy is necessary. Typically power spectra are recorded using 
+a Quadrant Photo Diode (QPD), which records motion in voltage units
+, not in units of length. To convert between the two we have a few 
+methods: Firstly if we have a mono-disperse suspension of particles, 
+the laser can be scanned across the surface of our target particle; 
+so long as the particle's size is known a linear conversion factor 
+can be used to convert from voltage to length units. Secondly, if 
+the size distribution is very wide but each particle can be accurately 
+sized, then a conversion factor can be approximated by comparing the 
+fitted value of the diffusion coefficient, and the reported value 
+given by the Stokes-Einstein relation.
 \begin{align}
@@
 
-With the latter method, the local fluid viscosity must be known to a high degree of accuracy, depending on the local heating effect this may be as trivial increase or it may be significant enough to drastically alter the characterisation.
-PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$ \cite{BergSoerensen2004}, it captures all of the information acquired by other calibration techniques while filtering out noise and requiring a relatively small amount of data collected.
+With the latter method, the local fluid viscosity must be known to a 
+high degree of accuracy, depending on the local heating effect this 
+may be as trivial increase or it may be significant enough to drastically 
+alter the characterisation. PSD analysis is often seen as the gold 
+standard for calibration as it can be fine tuned to the point that 
+optical forces can be computed on the order of $10^{-15} N$ 
+\cite{BergSoerensen2004}, it captures all of the information acquired 
+by other calibration techniques while filtering out noise and requiring 
+a relatively small amount of data collected.
 
@@
 
-Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}.
-As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
+Later chapters cover the dynamics of spherical aggregates and anisotropic 
+scatterers, these subjects are particularly difficult to characterise 
+using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
+As an example consider a symmetric dimer as a paradigmatic aggregate; 
+if we consider the Langevin equation for such a aggregate within an optical 
+trap we have:
 \begin{align}
@@
 \end{align}
-Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case.
-Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation.
-The rotational form of the Langevin equation for a dipole within an external potential is given as:
+
+Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
+motion is generalised to a 3 dimensional case. Except now, the dimer 
+is undergoing random rotational motion in addition to its Brownian 
+translational motion the first term on the right hand side is no longer 
+purely a function of the dimer's position but also on its orientation. 
+The rotational form of the Langevin equation for a dipole within an external 
+potential is given as:
 \begin{align}
@@
 \end{align}
-Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a sphere, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector.
-$\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
+
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of the 
+two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational 
+friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a 
+sphere, if the dimer is within a harmonic potential we can write the first 
+term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times 
+\vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. 
+$\lambda(t)$ is the Brownian rotations from the surrounding fluid, 
+like in the translational case the Brownian rotations are normally 
+distributed and are also uncorrelated so that:
 \begin{align}
@@
 \end{align}
-For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
+
+For an asymmetric scatterer whose radius is comparable to that of the 
+electric field's wavelength we now have a system of simultaneous equations:
 \begin{align}
@@
 \end{align}
+
 Fortunately, we do not need to solve these directly as the latter two
 random variables can be easily approximated if the thermal energy of
-the system is know, and the rate of change can be assumed as linear if
+the system is known, and the rate of change can be assumed as linear if
 we take a sufficiently small time step that
@@
 \cite{Crichton2000THEMD} is an easier form to compute:
-
 \begin{equation}
@@
 \end{equation}
-where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients of the incident and scattered fields respectively.
-We can get the transverse force and torque components in a similar form by applying a simple rotation transformation.
-In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate).
-Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.
 
-Throughout this PhD we use the work of Vigilante to perform a systematic study of the dynamics demonstrated by asymmetric dimers in both plane and circularly polarised light.
-But furthermore we expand upon their work to create a fully functional quadrant photo-diode to simulate the response from a calibration test.
-This builds upon the work from \cite{Rohrbach2002} which applied the fundamental principles of Lorenz-Mie theory to replicate the response signal of a QPD being used in back focal-plane interferometry.
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients
+of the incident and scattered fields respectively. We can get the transverse 
+force and torque components in a similar form by applying a simple rotation 
+transformation. In order to get the scattered beam coefficients we can use 
+\textit{mstm} \cite{Mackowski2011} to calculate the T-matrix of our dimer 
+(or any spherical aggregate). Vigilante et al compiled together a python 
+package that combines both \textit{ott} and \textit{MSTM} to simulate the 
+behaviour of spherical aggregates within a predefined optical trap.
 
+Throughout this PhD we use the work of Vigilante to perform a systematic 
+study of the dynamics demonstrated by asymmetric dimers in both plane and 
+circularly polarised light. But furthermore we expand upon their work to 
+create a fully functional quadrant photo-diode to simulate the response 
+from a calibration test. This builds upon the work from \cite{Rohrbach2002} 
+which applied the fundamental principles of Lorenz-Mie theory to replicate 
+the response signal of a QPD being used in back focal-plane interferometry.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:simulated_QPD}
-In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface.
-While trapping a micro-particle, the scattered and incident fields combine together and interfere with one another.
-These fields are collected by a condenser lens in the far field limit and are focused onto the QPD surface, the total intensity can be evaluated as:
+In order to simulate a typical experimental set up with a QPD installed 
+as a position detection system we need to evaluate the total magnitude of 
+the electric field incident on the photo-diode surface. While trapping a 
+micro-particle, the scattered and incident fields combine together and 
+interfere with one another. These fields are collected by a condenser 
+lens in the far field limit and are focused onto the QPD surface, the 
+total intensity can be evaluated as:
 \begin{align}
@@
 \end{align}
-The last term is simply a representative step term that defines the outer limit by which we evaluate the electric field, this is analogous to our condenser lens removing noise from other light sources by only accepting light at a specific acceptance angle defined by its numerical aperture $NA_c$.
-Depending on the relative size of our particle we can adjust the acceptance angle, this has very little effect on the transverse signals, but for axial evaluations of a trapped particle the numerical aperture should be tuned so that the resultant response curve has negative slope in order to allow for axial position detection, the method for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
 
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-\textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
+The last term is simply a representative step term that defines the outer 
+limit by which we evaluate the electric field, this is analogous to our 
+condenser lens removing noise from other light sources by only accepting 
+light at a specific acceptance angle defined by its numerical aperture 
+$NA_c$. Depending on the relative size of our particle we can adjust 
+the acceptance angle, this has very little effect on the transverse 
+signals, but for axial evaluations of a trapped particle the numerical 
+aperture should be tuned so that the resultant response curve has negative 
+slope in order to allow for axial position detection, the method for 
+finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
+
+The incident beam is simple enough to define given our set up parameters, 
+for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
+beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
+\textit{Ott} uses a point matching approach to approximate the beam shape 
+coefficients of the incident field by fitting it to the far field estimate, 
+the beam is of the form:
 \begin{align}
@@
 \end{align}
-where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
 
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers.
-The scattered and incident fields are then combined together in the far field to get $I(x,y)$, the quadrant and overall signals are calculated via:
+where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave 
+functions, these differ from Eq.~\ref{eq:lornez_mie} in the fact that the 
+field can either be expressed as an incoming/outgoing wave (with a 
+singularity at the origin) or as a regular wave around the origin; for 
+incoming/outgoing waves the wave functions use the first/second forms of 
+the Hankel function respectfully. In order to compute the regular spherical 
+wave at the origin we replace the Hankel function with the Bessel function 
+which is simply the average of the first and second forms of the Hankel 
+function, so at the origin we avoid a singularity of the EM field.
+
+We can if we want further restrict the incident beam by applying setting 
+the truncation angle to match our microscope object, this essentially 
+applies a cut off point to the In order to compute the scattering from 
+the target particle \textit{ott} uses the $T$-matrix method, this is not 
+essential for a simple sphere but is essential for complex shaped particles 
+such as dimers. The matrix from mstm needs to be repackaged to work with the
+\textit{ott} software, converting it from a column vector into a system of 
+sub matrices.
+
+The scattered and incident fields are then combined together 
+in the far field to get $I(x,y)$, the quadrant and overall signals are 
+calculated via:
 \begin{align}
@@
 \end{align}
-Where the denominator is the total intensity on the QPD while there is no particle within the trap.
-The QPD sensitivity is dependent on both the polarisation of the incident beam and the displacement of the target sphere, this is shown in fig.~\ref{fig:totalfield}.
-\begin{figure}[h]
+
+Where the denominator is the total intensity on the QPD while there is 
+no particle within the trap. The QPD sensitivity is dependent on both 
+the polarisation of the incident beam and the displacement of the target 
+sphere, this is shown in fig.~\ref{fig:totalfield}.
+\begin{figure}[h!]
 	\centering
@@
 
-Rather than displacing and rotating the particle itself, we instead rotate and displace the incident beam and recalculate the beam coefficients; the total beam is then inversely rotated in order to record the QPD signal.
-With this any trajectory can be collected from the QPD by displacing and rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss how to extract information from the QPD's outputted signal.
+Rather than displacing and rotating the particle itself, we instead rotate 
+and displace the incident beam and recalculate the beam coefficients; the 
+total beam is then inversely rotated in order to record the QPD signal. This
+ensures that for a homogenous sphere, even if the sphere itself undergoes
+rotation the QPD signal is unaffected by the change in orientation.
+
+To confirm that our method is producing accurate results, we ran a comparison
+between our simulative QPD and the results from \cite{Rohrbach2002}. Where a 
+$300\ nm$ diameter sphere is scanned across the path of a focused Gaussian beam
+($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is
+suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
+set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
+produced the following response curve.
+\begin{figure}[h!]
+	\label{fig:Rohrbach}
+	\begin{subfigure}{0.475 \linewidth}
+		\subcaption{}
+		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
+	\end{subfigure}
+	\begin{subfigure}{0.475 \linewidth}
+		\subcaption{}
+		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
+	\end{subfigure}
+	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
+Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from \cite{Rohrbach2002}.}
+\end{figure}
+
+With this any trajectory can be collected from the QPD by displacing and 
+rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} 
+and \ref{chapter:simulated_detection} we discuss how to extract information 
+from the QPD's signal.

### 2024-07-11 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 \chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution.
-The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results.
-It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}.
-Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}.
-These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.
-In this chapter I outline the optical tweezer equipment used during this PhD, the initial attempts made to induce nucleation via optical rotation, and the impact of a moving beam on the crystal growth of a newly formed nucleus.
+As outlined in Chapter 1, the original outline of the PhD was to investigate 
+the possibility of using optical tweezing as a means of initiating and 
+controlling nucleation by generating fluid flow within a small droplet of 
+supersaturated solution. The goal of which would be to understand the 
+influence of shearing on nucleation at a micro level as compared to larger 
+scale results. It has been shown that for macro-scale systems, the likelihood 
+of nucleation increases to a maximum value under increased shearing 
+\cite{Debuysschere2023, Mura2016}. Theoretical research into the matter 
+identified two competing effects that effect a crystal in moving fluid fields; 
+firstly, nucleation is enhanced due to the increased mass transfer of solute 
+material; and secondly, shear flow against the crystal surface leading to a 
+decrease in growth \cite{Mura2016}. These two competing effects are validated 
+by experimental work using glycine solutions, showing that beyond a certain 
+shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this 
+chapter I outline the optical tweezer equipment used during this PhD, the 
+initial attempts made to induce nucleation via optical rotation, and the 
+impact of a moving beam on the crystal growth of a newly formed nucleus.
 
@@
 
+\subsection{Fourier Optics and 4f correlators}
+A 4f correlator is an example of Fourier optics in practice, understanding
+that a focused lens takes a Fourier transform of the light profile. 
+Consider a laser with a circular Gaussian profile, if you were to place 
+a detector there you would pick up the intensity as a function of its 
+position within the beam. If however you focused the light into a single 
+point (using a +ve focal lens) you are actually seeing a measurement of 
+the phase of your laser with position, in which you would see a diffraction
+limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
+collimated. In imaging systems, a series of focal lenses can be used to 
+filter out unwanted scattering from an image (or in an inverse case 
+differentiate between different images), the placement of each lens is shown below.
+For our applications a 4f correlator is utilised to ensure that the motion
+of the galvano-mirrors does not move the focal point of the laser, allowing 
+for a stable trap even while in motion.
+
 \section{Calibration of Tweezer Setup}
@@
 to halt the reaction, the solution was filtered and washed 3 times with 
-distilled water before being suspended in water.
+distilled water before being suspended in water. When trapped in circularly 
+polarised light, the anisotropic scattering of the sphere results in a periodic 
+signal on the QPD. Therefore, the resulting power spectrum is not a Lorentzian
+but now also displays peaks that appear at integer multiples of the particles
+rotational frequency.
 
 \begin{figure}[h!]
+	\label{fig:vaterite}
 	\centering
@@
 solid crystal into its liquid crystal state. The individual droplets 
-are inherently birefringent and would rotate readily while suspended in 
-distilled water.
-
+are inherently birefringent and rotate in a circularly polarised trap. 
 \begin{figure}[h!]
@@
 \end{figure}
-\subsection{Rotation of Vaterite micro spheres} 
+
+The liquid crystal droplets had a much faster rotation rate than comparable 
+vaterite spheres, due to their higher degree of birefringence and the fact
+that the droplets are far closer to perfect spheres making angular momentum
+transfer more efficient.
+
+\subsection{Rotation of birefringent micro spheres} 
 Optical tweezing has often been used for micro-rheology, by computing the
@@
 local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
-Typically one would use a birefringent particle (i.e. vaterite, liquid
-crystals, etc) and rotating it within the fluid, the maximum rotation
-rate being a product of the fluid drag resisting the torque of the 
-trapping beam \cite{RodriguezSevilla2018}. Or if you want to measure 
+Using a birefringent particle and rotating it within the fluid, the maximum 
+rotation rate is due to the fluid drag resisting the torque of the 
+trapping beam \cite{RodriguezSevilla2018}. If you want to measure 
 fluid flow you can instead use a micro-rotor to see how fluid flow 
@@
 
-From Fig.\ref{fig:vaterite_shear}~(b) it's clear that the rotational speeds
-achieved both in the lab and from experimental literature are well below the 
-theoretical maximum shear rate reported by \cite{Debuysschere2023} as you move 
-away from the surface of your micro-rotor.
+From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
+particle size and rotation rate, this is contrary to much of the theoretical
+predictions that predict an exponential decay with particle size. This can
+be in part due to the fact that synthesising perfectly spherical spheres that
+have uniform birefringence is difficult over multiple experiments and tests. 
+Despite our best efforts at controlling the growth rate the vaterite spheres
+would often combine together while suspended in water after a short period of 
+time. The fastest reported rotation rate found during this PhD was by \cite{Arita2016}
+that achieved a rotation rate of $5\ kHz$, this is plotted on Fig~\ref{fig:vaterite_shear}
+as the dotted line. Even at that extreme a rotation rate the region in which 
+nucleation is at its optimal likelihood is only $20\ nm$ wide meaning that 
+localising nucleation around a rotating sphere is vanishingly small.
 
-\section{Rotation of Vaterite in supersaturated solutions}
-
+\section{Rotation of micro-rotors in supersaturated solutions}
 Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
@@
 was rather low or non-existent, the results are tabulated below:
-\begin{table}
+\begin{table}[h!]
 	\centering
@@
 in pure water the rotational speed achieved by a microsphere is
-insufficient to achieve a significant shear rate more than a few microns
+insufficient to achieve a significant shear rate more than a few tenths of a microns
 from its surface. The closest we could trap a microsphere to the droplet
@@
 \section{Shearing via a Galvano-mirror}
+Rather than utilising birefringence rotation, which is subject to issues
+during synthesis and achieving an ideal particle size, a galvano-mirror
+was installed to induce controlled particle motion. While typically 
+galvano and gimble mirrors are used to trap multiple particles in a 
+regular pattern, in a sufficiently dilute solution a single silica 
+micro sphere can be moved quickly through a fluid along a preset path.
+Calculating the shear rate around an individual particle is difficult to
+do precisely but for low Reynolds numbers we can get an adequate approximation.
 
-Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised.
-One can describe the position parameter of an ellipse as such:
+For a simple circular path one can estimate the sphere's speed by the 
+radius of its path and the frequency of its orbit $U = R\omega$; however 
+for a more complex path, such as an elliptical orbit the curve needs 
+to be parametrised. One can describe the position parameter of an ellipse 
+as such:
 
@@
 
-where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit.
-Subbing this in and then taking the partial derivative of position gives:
+where a and b are the different characteristic radii of an ellipse, if 
+we say that $u$ describes time from some initial point we can say $u=t\omega$ 
+where omega is the frequency of orbit. Substituting this in and then taking 
+the partial derivative of position gives:
 
@@
 
-In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.
+In order to compute U we simply take the magnitude of our velocity. 
+For low velocities the fluid flow at the sphere's surface can be computed 
+based on its velocity.
 
@@
 
-Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point.
-Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:
+Where $\theta$ is the angle from the direction of movement to the point
+you wish to measure, and $r$ is the radial distance to that point. Again
+taking the partial derivative we can get the shear rate for a particle moving 
+through the fluid:
 
@@
 Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
+
+\section{Nucleation using a moving beam}
+As mentioned previously, shearing via optical rotation and particle displacement
+did not result in

### 2024-07-11 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		\centering
-		\raisebox{-45pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
+		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
 		\label{large over small}
@@
 \end{figure}
-\newpage
-\subsection{Non-trivial harmonic traps}\label{sec:off-axis}
+
+\subsection{Non-trivial harmonic traps}
+\label{sec:off-axis}
 Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
@@
 	\begin{subfigure}{0.32\linewidth}
-		\raisebox{-110pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
+		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
 		\caption{}
@@
 spherical aggregates of any construction is their interaction with
-circularly polarised light. Typically the spin density of an electric
-field cannot be reduced in homogenous medium due to the fact that the
-spin angular momentum is conserved locally. In which case the only way
-to transfer angular momentum to use an object that is anisotropic.
-This is seen most clearly with birefringent materials, due to fact that
-the crystal lattice itself is anisotropic, a circularly polarised beam
-will transfer an optical torque that is proportional to the difference in 
+circularly polarised light. While most homogenous objects are trapped
+no differently in different polarisations, dimers have been shown to
+experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018}. 
+Typically the spin density of an electric field cannot be reduced in 
+homogenous medium due to the fact that the spin angular momentum is 
+conserved locally. In which case the only way to transfer angular 
+momentum to use an object that is anisotropic. This is seen most 
+clearly with birefringent materials, due to fact that the crystal 
+lattice itself is anisotropic, a circularly polarised beam will 
+transfer an optical torque that is proportional to the difference in 
 the refractive index of the particle. The optical torque is given as:
@@
 \end{equation}
+
 Where $\Delta n$ is the difference in refractive indices of the 
 particle, and $\phi$ indicates the phase shift in the electric field
-(for circular polarised light $\phi= \pi/4$) However, theoretical and
-experimental work by \cite{Yevick2017} found that highly focused
-Gaussian beams could produce second order effects in the Rayleigh
-regime resulting in a photo-kinetic force that results in orbital
-motion about the beam's central axis. This effect is rather minimal
-for single sphere's, resulting in a orbital frequency on the order of
-$10^{-1}\,{\rm Hz}$, with an order of magnitude difference when
-trapping aggregates of spheres.  They computed the circulation rate by
-computing the time-average probability flux; however, when extended to
-the Mie regime we see a completely different behaviour, instead
-experiencing an optical torque about their long axis.S This rotation
-was first noted by Vigilante and co-workers who only considered this
+(for circular polarised light $\phi= \pi/4$). The optical rotation seen
+with dimers can be attributed to the polarisation of the trapping beam as
+taking a Fourier transform of the transverse components of the dimers 
+orientation time series reveals a clean rotation rate. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.9\linewidth]{rotation_vs_pol.png}
+	\caption{Rotation frequency vs component phase difference for differently sized dimers.
+The solid lines represent the rotation rate experienced while the dimer is in its standard orientation, whereas the solid points are for the case where the orientation is inverted.
+Laser power $= 100\ mW$}
+\end{figure}
+
+A possible explanation of this phenomena was devised by by \cite{Yevick2017}, 
+who found that highly focused Gaussian beams could produce second 
+order effects in the Rayleigh regime resulting in a photo-kinetic 
+force that results in orbital motion about the beam's central axis. 
+This effect is rather minimal for single sphere's, resulting in a 
+orbital frequency on the order of $10^{-1}\,{\rm Hz}$, with an 
+order of magnitude difference when trapping aggregates of spheres. 
+They computed the circulation rate by computing the time-average 
+probability flux; however, when extended to the Mie regime we 
+see a completely different behaviour, instead experiencing an 
+optical torque about their long axis.S This rotation was first 
+noted by Vigilante and co-workers who only considered this
 behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of
@@
 orientation and therefore their axial position.
-
-\begin{figure}[h]
+\begin{figure}[h!]
   \centering
@@
 imply that a second sphere of radius $200\,{\rm nm}$ is enough to
-induce rotational motion.  We used MSTM to look at the stokes
+induce rotational motion. We used \textit{mstm} to look at the stokes
 parameters from the scattered field from a simple plane wave incident
@@
 dimer that might impart angular momentum to the scattered beam - as
-compared to a anisotropic scatterer like Vaterite.
+compared to a anisotropic scatterer like vaterite.
 
@@
 \end{align}
+
 where the polarizability is given as a 3D vector for the three
@@
 	\includegraphics[width=\linewidth]{PSD_sphere.png}
-	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian}, scattered points represents
-	the blocked data ($n_b=100$). Corner frequency for the 
-	Lorentzians are reported in the legend.}
+	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian}, scattered points represents the blocked data ($n_b=100$). Corner frequency for the 
+	Lorentzian curves are reported in the legend.}
 \end{figure}
@@
 can be explained partly by the fact that our estimation of the 
-trap geometry via \textit{ott} is based on the the actual 
-interactions with the beam near close to the focus whereas
-a QPD is estimating the trap geometry by extrapolating from
-far field scattering signals. If we now consider a symmetric the 
-discrepancy between \textit{ott} and the simulated QPD grows.
+trap geometry via \textit{ott} is based on the force-displacement
+curve while the sphere is moving in only one direction, whereas the
+QPD is estimating the trap geometry by extrapolating from far field 
+scattering signals. We now consider a symmetric dimer in identical 
+simulative conditions.
 
@@
 trap compared to the QPD model which says the trap is far
-more symmetrical while trapping a symmetric dimer. The reason that
-\textit{ott} no longer expects a circular trap could be due to
-how it computes the beam shape coefficients; by point matching in 
-the far field before the focus means a loss in accuracy for objects
-that trap above the focus. The change in the QPD estimation  
-can be partially explained by the fact that rotational
-effects are not accounted for in the Lorentzian power spectra,
-only translational motion. Typically, rotational motion is only
-ever detected when it is periodic (take for example a power spectra
-from a rotating birefringent sphere), when the motion is stochastic
-the entire power spectra is effected making it near impossible
-to separate the translational and rotational contributions from a 
-single power spectra.
+more symmetrical while trapping a symmetric dimer. A potential 
+reason that \textit{ott} no longer expects a circular trap 
+could be due to how it computes the beam shape coefficients; 
+by point matching in the far field before the focus means a 
+loss in accuracy for objects that trap above the focus. 
+The change in the QPD estimation can be partially explained 
+by the fact that rotational effects are not accounted for in 
+the Lorentzian power spectra, only translational motion. 
+Typically, rotational motion is only ever detected when it 
+is periodic (take for example Fig.~\ref{fig:vaterite}), when 
+the motion is stochastic the entire power spectra is effected 
+making it near impossible to separate the translational and 
+rotational contributions from a single power spectra.
 
@@
 \section{Conclusions}
-\comment{Conclusions still need to be written out}
+Considering the simplicity of a scatterer such as a dimer,
+one would assume that the dynamics of such an object would be
+relatively easy to predict. Simulations of dimers in the Mie regime
+show that not only do they have multiple positions and orientations 
+in which they can be trapped but also that their interaction
+with circularly polarised light is heavily dependent on the axial 
+position and trapping orientation. Dimer's have the potential to be used
+as adaptive micro-rotors, being simple to synthesise and can be made 
+out of any material of choice, but in order for these particles to be
+used as such one needs a means of characterising the optical torque
+applied. While estimations can be made in the Rayleigh regime due to 
+the dimer being treated as point dipole, Mie regime micro rotors require 
+an in-depth theoretical description of the mechanism that is creating 
+this optical torque.

### 2024-07-11 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 \chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
+\label{chapter:simulated_detection}
 \section{Simulative QPD}
@@
 \subsection{Incident beam}
-\label{sec:scattering}
+\label{sec:fibre_scattering}
 The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
@@
 In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
+In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity
+of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
 As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
@@
 
-\begin{figure}[h]
-	\label{fig:Rohrbach}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from [].}
-\end{figure}
 Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.

### 2024-07-16 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 impact of a moving beam on the crystal growth of a newly formed nucleus.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \section{Optical Tweezer Equipment}
@@
 \end{align}
+
 It should be noted that the galvano-mirror requires the use 
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Position detection methods}
@@
 trap needs to be calibrated (assuming a linear response curve).
-
 \begin{figure}[h!]
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Fourier Optics and 4f correlators}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Calibration of Tweezer Setup}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Synthesis of Birefringent Micro spheres}
@@
 rotational frequency. 
-
 \begin{figure}[h!]
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Rotation of birefringent micro spheres} 
@@
 \end{align}
+
 Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
@@
 and their rotation speeds were measured via the QPD. 
-
 \begin{figure}[h!]
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Rotation of micro-rotors in supersaturated solutions}
 Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
-(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of 
-supersaturated solutions of Glycine and water. After locating a 
-singular microsphere the vaterite was trapped in a circularly 
-polarised light and brought close to the droplet edge, after a 
-period of ten minutes if no nucleation event was observed the 
-particle was released and another particle was trapped. Due to 
-the increased viscosity of the supersaturated solution and the 
-proximity to the droplet edge, this was done as previous work with
-optical tweezers have only seen nucleation near the solution-air interface
-\cite{Gowayed2021, Liao2022, Walton2018}. The observed rotation rate 
-was rather low or non-existent, the results are tabulated below:
+(see sec.~\ref{sec:vaterite}), and then suspended in distilled 
+water, at the same time a supersaturated solution of Glycine 
+and water was prepared and $50\ \mu L$ was pipetted onto a 
+glass cover slip. A single microsphere the vaterite was trapped 
+in a circularly polarised light and brought close to the droplet 
+edge.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
+	\caption{Diagram of optical trapping set up for rotating birefringent 
+		particles in a supersaturated solution. Left: side view of the 
+		trapping set up showing the location of the trap focus at the 
+		edge of the droplet of a supersaturated solution. Right: top down 
+		view of the glycine droplet with a trapped birefringent particle 
+		shown close to edge of the trap. As the particle rotates the drag 
+		force from the surrounding fluid generates a flow field around 
+		itself (see Eq.~\ref{eq:birefringent_speed}).}
+\end{figure}
+
+After a period of ten minutes if no nucleation event was 
+observed the particle was released and another particle was 
+trapped. Due to the increased viscosity of the supersaturated 
+solution and the proximity to the droplet edge, as previous works
+have only seen nucleation near the solution-air interface \cite{Gowayed2021, 
+Liao2022, Walton2018}, the observed rotation rate was rather low or 
+non-existent, the results are tabulated below:
 \begin{table}[h!]
@@
 		\hline
-		\multirow{3}*{1.29} & 3.47 & 0.00 & $\times$ \\
+		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
 		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
@@
 		\hline
-		\multirow{3}*{1.40} & 6.32 & 0.00 & $\times$ \\
+		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
 		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Shearing via a Galvano-mirror}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Nucleation using a moving beam}
 As mentioned previously, shearing via optical rotation and particle displacement
-did not result in
+did not result in any localised nucleation events even while in the proximity of 
+the droplet edge. During the experiments with the galvano-mirror, it was found 
+that when no particle was present in the optical trap nucleation events would 
+occur while the beam was close to the edge of the droplet, even though the solution
+was undersaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
+but was more interesting is how the beam's motion influenced the growth of the nucleus.
+
+Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
+event in supersaturated glycine solution (S=1.03), the beam is a stationary being 
+$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
+nucleus forms at the trap focus, growing quickly until after roughly $6$ seconds 
+the crystal escapes the trap. The key take away to remember is that the beam has no 
+real influence over the crystal front,instead the crystal grows outward from the 
+centre of the focus.
+\begin{figure}[h!]
+	\label{fig:stationary_beam}
+	\centering
+	\includegraphics[width=0.7\linewidth]{frames_no_beam_movement.pdf}
+	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
+	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
+	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the
+	trap.}
+\end{figure}
+
+Now while trying to induce fluid flow by optically trapping a silica particle and 
+using a galvano-mirror to move the particle through the solution, nucleation events
+were observed in which the crystal front follows the path of the laser. See below 
+in \textcolor{red}{Fig.XX} where we have the laser focus moving in a small elliptical 
+pattern,

### 2024-07-16 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 \label{chapter:simulated_detection}
+As outlined in the end of Chapter 4, one of the difficulties in characterising 
+interactions with asymmetric objects is the coupled motion between translation 
+and rotation. This can be somewhat ignored if the rotational motion is periodic,
+like when the power spectra of elliptical polystyrene particles was fitted by 
+Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
+rotational motion is stochastic the problem is more complex. For example, when a
+novel trap characterisation technique was implemented by Saffron \textit{et al}
+\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
+characterise both the axial and lateral trap stiffness acting on microspheres. The
+only drawback admitted to in their work was that the technique was constrained to
+isotropic scatters as their theoretical model for describing the auto-correlation
+function was predicated on the fact that any variations in the signal are due to
+the particles translational motion within the confines of a cylindrical trap.
+
 \section{Simulative QPD}

### 2024-07-24 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

@@
 \end{align}
+
 Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
@@
 
-\subsection{Primary vs Secondary nucleation}
+\subsection{Primary \& Secondary nucleation}
 From an industrial perspective, the nucleation process can
@@
 \end{align}
+
 Where $v$ is the approximate volume of an individual molecule, $k_B$
@@
 \begin{figure}[h!]
-	\label{fig:free_energy}
 	\centering
@@
 		     the interfacial energy cost is independent of supersaturation.}
+	     \label{fig:free_energy}
 \end{figure}
@@
 \end{align}
+
 Where $A$ is a pre-factor that can be fine tuned to the exact demands
@@
 
-So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy.
-The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius.
-This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}
-
-Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows.
-Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.
-
-There are several results that indicate that the CNT is not a good fit for nucleation rates:
-\textcolor{red}{
-	\begin{itemize}
-		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 \cite. 
-		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass.
-We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure.
-To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ].
-		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled.
-Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases.
-[ ]
-	\end{itemize}
-}
-
-CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate.
-However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case.
-Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.
+CNT is often regarded as a good description of the macro system, its obvious
+that for all crystallization systems there is an inherent energy barrier that 
+dictates the nucleation rate. Where it falters is in its predictive ability, 
+both in estimating nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in 
+the structure of newly formed nuclei \cite{Lee1999, Yau2001}. Recent studies 
+suggest classical nucleation is merely one of many possible pathways that can 
+be taken to produce a structured crystalline phase. Prompting the development of alternative theories to better describe the nucleation process.
 
 \subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}.
-\textit{in situ} techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.
-It can be understood by Oswald's rule \cite{Ostwald1897}, which says that any crystallising system does not
-take the path to the lowest possible energy state but instead first transitions to the state with the smallest free energy barrier, further transitions can still occur but this method minimises the overall free energy cost.
-If we consider the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider a case where the supersaturation is low ($S\approx1.01$) then a dense liquid droplet will have a lower chemical potential than the surrounding fluid
+The two step nucleation theory is an extension to the CNT, colloid simulations 
+showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) 
+interactions allow for the formation of a liquid-liquid metastable phase from 
+which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
+\textit{in situ} techniques for studying nucleation several papers reported 
+the presence of stable liquid-like clusters that formed prior to nucleation
+\cite{Savage2009, Wolde1997, Soga1999}. It can be understood by Oswald's rule
+\cite{Ostwald1897}, which says that any crystallising system does not immediately 
+take the path to the lowest possible energy state but instead first transitions 
+to the state with the smallest free energy barrier, further transitions can still 
+occur but this pathway minimises the overall free energy cost. If we consider 
+the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider 
+a case where the supersaturation is low ($S\approx1.01$) then a dense liquid 
+droplet will have a lower chemical potential than the surrounding fluid but 
+the crystalline phase will have a subsequently lower chemical potential than 
+the dense liquid phase. Initially it was suspected that the formation of the 
+dense liquid phase was a result of stochastic density fluctuations in the system; 
+further simulations suggest that the the local bond order is a stronger driving 
+force than the local density \cite{Tan2013}.
 
+The current research into two step nucleation (or now more commonly referred 
+to as multi-step nucleation theory) is developing a robust framework to 
+describe what nucleation pathway will occur given the initial conditions. 
+Reviews of all currently documented nucleation pathways identified highlighted 
+the fact the need for a the development of \textit{in situ} techniques that 
+can induce nucleation locally but can also reliably identify the nucleation 
+pathway across a broad range of experimental conditions \cite{Karthika2016, Fu2021}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Crystallisation methods}
@@
 \begin{figure}[h!]
-	\label{fig:MSZW}
 	\includegraphics[width=\linewidth]{MSZW.png}
@@
 3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
+	\label{fig:MSZW}
 \end{figure}
@@
 
+\subsection{Evaporative Crystallisation}
+In situations where control of the final product size or shape is not the main
+focus, often the cheapest method of producing a crystalline product is simply
+to allow the solvent evaporate and separate from the solute. Depending on the 
+total volume of solvent to evaporate this process can take on time scale of 
+several days to complete.
+
 \section{Optical Tweezers}
 \subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'.
-The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis.
-Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}.
-This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}.
-Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.
+Optical tweezing has been a field of applied optics ever since the 1970s
+when Ashkin \cite{Ashkin1970} first showed that focused light was capable 
+of trapping micron sized particles due to light exerting 'radiation pressure'. 
+The working principle was that a light source such as a laser could trap 
+small objects within a 2D plane, as long as the light source had an 
+approximately Gaussian profile. Soon after, Ashkin showed that the introduction 
+of a microscope objective would allow one to focus the light source to a 
+diffraction limited point that would stably trap small objects within a 
+confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study 
+biological material and would later be used to probe microscopic properties 
+such as the formation of colloidal aggregates \cite{Yi2021} to the drag 
+forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the 
+predictable behaviour of light, optical tweezers have become essential for 
+measuring and exerting precise forces on the magnitude of pico-newtons 
+allowing one to probe the material properties of the smallest materials.
 
 \subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one.
-Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}.
-Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}.
-Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly.
-In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.
+From as early as 1996 it has been known that laser irradiation is a viable
+method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, 
+the first reported case was notable as it used a 1.064 $\mu m$ laser meaning 
+it was unlikely to be a photo-chemical reaction but rather a physical one. 
+Future research has found nucleation can be induced by 1 of 3 routes involving 
+direct laser induction: firstly, Non-photochemical laser induced nucleation 
+(NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996,
+Garetz2002,Sun2006}, several papers have debated the exact mechanism that 
+induces NPLIN \cite{Garetz2002, Knott2011}. Two suggested hypothesis are: 
+an optical Kerr effect, where the solute molecules are aligned to lower the 
+nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in 
+which solute clusters are stabilised within the electric field which drastically 
+increases the likelihood of nucleation\cite{Alexander2008}. Both hypothesise 
+have their limitations and there has yet to be a single theory that explains 
+NLPIN thoroughly. In either case the mean pulse intensity needs to be kept 
+relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses 
+lead to a completely different nucleation mechanism.
 
-High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN.
-The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}.
-There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}.
-It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.
+High intensity laser induced nucleation (HILIN), where the pulse intensity 
+is on the order of several $PW/cm^2$ is far simpler a mechanism to explain 
+in comparison to NPLIN. The production of nuclei can be wholly associated 
+to a cavitation process within the target solution, where the laser focus 
+results in thermo-cavitation and the subsequent pressure change leads to
+a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}. 
+There is still not a general consensus on how cavitation influences the 
+local supersaturation \textcolor{red}{[ , ]}, nor is there a clear 
+understanding of what properties of the crystal can be controlled 
+\textcolor{red}{[ , ]}. It has been suggested that in theory any solution 
+can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a 
+strong understanding of the phenomena both before and after cavitation 
+occurs.
 
-Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products.
-When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}.
-Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.
+Lastly, there is trapping induced nucleation, this is where optical tweezers 
+come into play, the optical trap has been shown to have different effects 
+on supersaturated solutions depending on where it has been focused. When 
+focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ 
+were shown to create a dense liquid droplet of glycine and water 
+\textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid 
+region was populated by clusters that would consolidate together upon being 
+focused by the optical trap \textcolor{red}{[ ]}. Molecular simulations of 
+glycine solutions showed that these clusters are unstable when using pure 
+glycine below the saturation point \textcolor{red}{[ ]} suggesting that the 
+clusters are formed due to glycine reaction products. When the optical trap 
+was moved from the cover slip to the air-solution interface nucleation would 
+occur before a dense liquid region could form \textcolor{red}{[ ]}. Repeated 
+experiments where the laser is focused on the air-solution interface have 
+lead to a variety of different nucleation events. In some instances the 
+nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; 
+whereas allowing a solution to age results in the formation of amorphous 
+precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}. 
+The precursors are only seen when the solution is irradiated by an optical tweezer 
+and the growth rate can be controlled somewhat by varying the laser power 
+\textcolor{red}{[ ]}. Notably the only work has been done with simply 
+irradiating the solution with a trapping potential, there has not been an 
+attempt to introduce a trappable object into the solution, the trapping 
+potential has been used to influence the growth of a crystal front 
+\textcolor{red}{[ ]}.
 
 \subsection{Optical Rotation}
-For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components.
-The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}.
-However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}.
-Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}.
-This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.
+For any electromagnetic field it is possible to transfer both linear and 
+angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
+is said to have both orbital and spin momentum. Though there is some 
+debate on how to decompose the total momentum into these two components 
+\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
+the exact quantities and will instead look at the broader effects of both 
+components. Orbital angular momentum arises from the shape of the wavefront 
+of the particular field in question; for simple Gaussian beams the wavefronts 
+are uniform and equally spaced resulting in the typical radiation pressure 
+that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
+modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
+non-uniform wave fronts meaning the orbital momentum has both angular and
+linear components; depending on the relative size of the target particle 
+one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.
 
-For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM.
-Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus.
-This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam.
-Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM.
-However, by far the best method for inducing rotation is using birefringent materials.
+Spin angular momentum (SAM) is attributed to the spin density of the field, 
+early research has shown that the spin density is non-zero for any beam despite
+the fact that the total SAM transferred to a medium is 0 \cite{Svak2018, 
+Bliokh2014}. This has sparked debate if SAM is even a physical quantity as 
+it does not aid in the transport of energy directly \cite{Bliokh2014} and 
+so cannot be directly observed in some cases despite being non-zero. This 
+paradox is resolved by representing the wave as an array of spin momentum 
+loops that all together cancel one-another out when the medium is homogeneous.
+Spacial inhomogeneities cause these spin loops to non longer be equal, resulting
+in non-zero spin density, anisotropic mediums (such as birefringent crystal 
+lattices) experience a transfer of spin angular momentum allowing for optical
+torque transfer. It has even been shown that in the case of homogeneous spheres, 
+the spin component of a focused beam can result in additional orbital angular
+momentum via spin-to-orbital conversion \cite{Yevick2017}.
 
-Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation.
-For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}.
-Calculating the optical torque applied to a birefringent material is given via:
+Birefringence is a material property often seen in crystalline materials, 
+if the crystal lattice has a different structures when viewed at different 
+orientations then light will be refracted differently depending on its 
+polarisation. For circularly polarised light the inhomogeneity results in 
+a high degree of SAM being transferred to the target object \cite{Parkin2009,
+Arita2016}, this has been exploited to rotate microspheres as fast as 1000 Hz 
+while suspended in a bulk medium \cite{Arita2016} as well as a means of 
+measuring the local temperature and shear response of said medium 
+\cite{Millen2014, RodriguezSevilla2018}. Calculating the optical torque 
+applied to a birefringent material is given via:
 
@@
 
-Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.
-The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation.
-The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle.
-Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion.
-Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.
+Where $\theta$ is the angle between the particle's long axis and the Poynting
+vector of the local EM field, and $\phi$ is the phase shift in the EM field.  
+The first term represents the 'orientational' torque which is due to the target 
+particle being aligned along the EM field, when aligned $\theta=0$ meaning 
+the entire term is negligible for particle's with a stable orientation. The 
+second term is due purely to the polarisation of the optical trap, for 
+circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred 
+to the target particle. Birefringence can also be induced if the target 
+particle has an anisotropic shape, in particular if the particle shape is 
+elongated along one major axis; the go to particle shape is a silica dimer 
+(two spheres tangentially attached) due to silica's stability and strong 
+adhesion. Using a silica dimer research groups have achieved a rotation 
+frequency in the realm of several GHz \cite{Ahn2018} in a vacuum.
 
 \subsection{Shear induced Nucleation}
-It has long been known that solution nucleation is influenced by shearing 
-of the fluid, from stirring agents in the container (i.e. propellers) to 
-the container boundary layers, however there has yet to be a clear mechanism 
-by which said shearing effects the nucleation event. Theoretical research 
+It has long been known that fluid shear rate plays a role in influencing
+nucleation; however, the exact relationship between shear rate and nucleation
+rate has only been recently understood for specific solutions. Theoretical research 
 into shear induced nucleation suggests that there should be a slight increase 
@@
 in chapter 3, this is then succeeded by experimental work where we 
-instead using a galvano-mirror to generate a moving beam focus within an 
-undersaturated solution to control the crystal growth front. In addition,
-computer simulations into asymmetric dimers and their applications as complex
-micro-rotors are discussed in chapter 4. Lastly two novel methods of measuring
-rotational motion are discussed in chapter 5; firstly via a novel detection
-fibre method that allows for instantaneous measurements of the orientational
-behaviour of optically trapped ellipsoids/dimers; and secondly we create 
-a simulative quadrant photo diode that replicates laboratory results, utilising 
-linear regression techniques we measure the change in orientation in order to
-measure the optical torque applied to a non-birefringent particle.
+instead using a galvano-mirror to generate shear flow. While overall unsuccessful
+the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus.
+
+In addition, computer simulations studying the dynamics of asymmetric dimers 
+and their applications as complex micro-rotors are discussed in chapter 4. 
+Lastly two novel methods of measuring rotational motion are discussed in 
+chapter 5; firstly via a novel detection fibre method that allows for 
+instantaneous measurements of the orientational behaviour of optically 
+trapped ellipsoids/dimers; and secondly we create a simulative quadrant 
+photo diode that replicates laboratory results, utilising linear regression 
+techniques we measure the change in orientation in order to measure the 
+optical torque applied to a non-birefringent particle.

### 2024-07-24 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 the forces imparted by individual photons begin to become significant
-enough to influence its trajectory. Ashkin's work into optical
-tweezing found that micron sized entities could be spatially locked by
-aiming a laser with a Gaussian intensity profile directly upward at a
-the cell while being suspended in a liquid medium \cite{Ashkin1970}.
-This is demonstrated diagram below where we have a simple sphere
-trapped by a Gaussian beam, if we consider two rays of light hitting
-our sphere we see that while in the centre each ray is refracted in an
-equal opposite direction to one another and so the net force imparted
-is 0.  However when displaced the net force is unbalanced and points
-towards the centre of the trap; this principle can be generalised to 3
-degrees of freedom if we consider a focused beam instead of a
-para-axial beam, typically the trap will be weaker along the beam axis
-compared to the transverse trap strength.
+enough to influence the trajectory of individual particles. Ashkin's 
+work into optical tweezing found that micron sized entities could be 
+spatially locked by aiming a laser with a Gaussian intensity profile 
+directly upward at a the cell while being suspended in a liquid medium 
+\cite{Ashkin1970}.
 
-\begin{figure}[h]
+This is demonstrated in fig.~\ref{fig:working_principle} where we have 
+a simple sphere trapped by a circularly symmetrical beam with the intensity
+falling off from the centre in a normal distribution (commonly referred to
+as a Gaussian beam). If we consider two rays of light hitting our sphere 
+we see that while in the centre each ray is refracted in an equal opposite 
+direction to one another and so the net force imparted is 0. When displaced 
+in the x-y plane the net force is unbalanced and points towards the centre 
+of the trap; this principle can be generalised to 3 degrees of freedom if we 
+consider a focused beam instead of a para-axial beam, typically the trap will be
+weaker along the beam axis compared to the transverse trap strength.
+
+\begin{figure}[h!]
   \centering
@@
 Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
+  \label{fig:working_principle}
 \end{figure}
@@
 \end{align}
+
 Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
@@
 is how the force fields are reported. The gradient force is a conservative 
-force brought about by the polarisation of dielectric materials, 
-anything around 10\% of the trapping wavelength will be attracted along 
-the gradient of the electric field to the region of greatest intensity 
-(for a simple Gaussian Beam this would be at the focal point). The 
-scattering force arises from the scattered field of the trapping beam 
-pushing the target object away from the focal point; unlike the gradient \\
-force, the scattering force is non-conservative and is loosely 
-proportional to the particle's Brownian motion. The equilibrium position 
+force brought about by the polarisation of dielectric materials which 
+is directed towards the point of maximum intensity (for a simple Gaussian 
+beam this would be at it's focal point) \cite{YasuhiroHarada1996}. 
+Typically, optical tweezers will utilise higher numerical aperture lenses 
+in order to increase the intensity, thus allowing for stronger 
+gradient forces.
+
+The scattering force arises from the scattered field of the trapping beam 
+pushing the target object away from the focal point. The equilibrium position 
 (where the mean squared displacement is minimised) is found when the 
-gradient force far exceeds the scattering force, interpreting and 
-calculating these forces is dependent on the ratio between the particle 
-size and the trapping wavelength.
+gradient force far exceeds the both the scattering force, and the thermal 
+energy of the particle. Describing and calculating the optical forces is 
+dependent on the ratio between the particle size and the trapping wavelength.
 
@@
 \end{align}
+
 where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
 in refractive indices between the solution and the target particle,
-and $P_i$ is the power of the individual ray. For a beam with a
+and $P_i$ is the power of the individual ray. The net force can be 
+subdivided into its gradient and scattering components, where the latter
+directs the particle to the centre of the beam, and the latter acts on 
+the particle in the direction of beam propagation. For a beam with a
 Gaussian intensity distribution $P_i$ will fall off as you move from
-the centre of the beam.  The ray optics model is ideal for
+the centre of the beam.
 
@@
 \begin{align}
-	\label{eq:lornez_mie}
+	\label{eq:incident}
 	{\bf E}_{\rm inc}({\bf r})
@@
   \\
+  \label{eq:scattered}
   {\bf E}_{\rm scat}({\bf r})
@@
 \end{align}
+
 where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
@@
 the user. However, computing the expansion coefficients of the
-internal and external fields be far more tedious depending on the
+internal and scattered fields is far more tedious depending on the
 shape of the target and what properties of the scattered field are
-desired --- all of which is discussed later on.  From an optical
+desired --- see sec.~\ref{sec:scattering}.  From an optical
 trapping perspective the force exerted by a focused electric field can
@@
 Lorenz-Mie theory can be applied to describe the scattering from any 
-particle regardless of size, however as we either increase or decrease 
-the size of the target particles the infinite series converges allowing 
-one to ignore much of the tedium of Lorenz-Mie theory. The ray optics 
-model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. 
-In the latter case we can simplify the scattering problem by approximating 
-our target as a single dipole, focusing only its interactions with the 
-incident field.
+particle regardless of size, though as the size of the target particle 
+changes - relative to the light being scattered - the theory describing
+the scattered light is better approximated by alternative theories. For 
+particles that are far larger than the wavelength of light the scattering
+is best described by the Ray-Optics theory. In the opposite case the 
+target particle is better approximated as a single electric dipole as 
+described by the Rayleigh scattering field.
 
@@
 a particles motion in an electromagnetic field who's wavelength is several 
-times greater than the particle's size. The underlying theory is that a \\
+times greater than the particle's size. The underlying theory is that a 
 dielectric sphere can be treated as a dipole while in the presence of the 
@@
 the method replaces the scatterer with a series of surface currents
-over the targets surface, these currents are chosen so that the
+over the targets surface. These currents are chosen so that the
 electric field outside is identical to the original problem \cite{Wriedt1998}. 
@@
 \begin{align}
+	\label{eq:t_matrix}
 	\begin{pmatrix}
@@
 $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
-this PhD in particular we utilise an extension of the ECBM known 
-as multi-sphere $T$-matrix method (\textit{mstm}), developed by Mackowski
+this PhD in particular we utilise an extension of the ECBM known as 
+multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
 \cite{Mackowski2011} the computational code computes the scattered field 
-from each sphere within the target cluster and the incident field itself, 
-this scattering between spheres converges to a final result but modern codes 
-truncate the calculations to fit a desired accuracy and computational time.
+from each sphere within the target cluster and the incident field itself. 
+The scattering between spheres slowly converges to a final result but modern 
+codes truncate the calculations to an upper number of iterations (denoted by
+$N_{max}$) to fit a desired accuracy and computational time.
 
@@
 itself can be computationally taxing as the number of spheres increases 
-\cite{Mackowski2011}. While it is possible to solve for the scattering of
-each individual sphere this is only applicable for a single orientation and 
-can be even slower for large aggregates \cite{Mackowski1996, Xu1995}. The 
-benefit of \textit{mstm} is that the major scattering properties (scattering 
-and extinction cross sections, scattering matrices, etc) can all be computed 
-both for single orientations, or averaged over multiple orientations to determine
-the average scattering from the target particle.
+\cite{Mackowski2011}. While it is possible to solve for the full electromagnetic 
+field of each the entire cluster this is only applicable for a single 
+orientation and can be even slower for large aggregates \cite{Mackowski1996, 
+Xu1995}. The benefit of \textit{mstm} is that the major scattering properties 
+(scattering and extinction cross sections, scattering matrices, etc) can 
+all be computed both for single orientations, or averaged over multiple 
+orientations to determine the average scattering from the target particle.
 
@@
 identical to the scattered field produced by direct integration of 
-Eq.~\ref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
+Eq.~\eqref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
 The integral form for the internal electric field inside a scatterer is 
@@
 better describes the microscopic characteristics of any diffusive behaviour.
-The Fickian model is not suitable for the applications of optical tweezers as 
-it assumes that all diffusive motion is driven by a gradient of molecular density
-\cite{Gillespie2012}, not only is this often not the case but it fails to describe 
-the motion of any single molecule, only the collective behaviour. Einstein's model 
-however considers the collisions experienced by individual molecules \cite{Gillespie2012a}, 
-if we assume that all collisions are random and only consider the behaviour 
-after a finite time step $\delta t$ then the individual collisions should 
-cancel out, of course trying to understand the dynamics over a shorter period
-of time is not described by Einstein's model as you would have to consider
-each collision in turn \cite{Gillespie2012, Gillespie2012}. This is insufficient 
-for our interests as an optical tweezers relaxation time (given by 
-$\tau = \kappa/\gamma$) is often so short that we must consider the individual 
-collisions experienced by any one molecule. The Langevin model of diffusion 
-therefore assumes that the net force on a particular particle is described 
-fully by these individual collisions \cite{Gillespie2012c}:
+
+The Fickian model describes the net flux of solute molecules into a finite
+volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$
+\cite{Gillespie2012}. It assumes that the solute molecules do not collide with 
+one another or with molecules in the solution; overall, the Fickian model is
+used to describe how solute molecules disperse over long periods of time. It 
+does not provide any incite into the forces acting on individual molecules, 
+whereas the Langevin model captures the physical interactions between an 
+individual molecule and the surrounding fluid over a wide range of time scales.
+
+The Einstein model of diffusion expands upon the Fickian model by considering
+the collisions between individual molecules. If we consider a single particle 
+suspended in a solution it will experience multiple collisions with other 
+molecules \cite{Gillespie2012a} over a given time interval $\delta t$. The 
+Einstein molecule allows us to consider how individual molecules behave while
+suspended in a surrounding fluid. Where it begins to falter is when we consider
+bringing $\delta t$ to $0$; in this scenario the molecule experiences not many
+but only individual collisions. The Einstein model does not consider the inertia 
+of the molecule in question and so for very short time frames the model implies 
+that the molecule's velocity changes instantly after each collision \cite{Gillespie2012a,
+Gillespie2012b}. Furthermore, the kinetic energy of each collision is not limited
+by the thermal energy of the system, meaning that using the Einstein model to 
+predict a particle's trajectory to an finite degree of accuracy implies that 
+the particle has infinite kinetic energy \cite{Gillespie2012b}. This failure to 
+describe motion over smaller time frames is addressed in the Langevin model by 
+accounting for the fluid drag of the system, where any sudden change in velocity 
+must result in a proportionally opposed drag force \cite{Gillespie2012c}.
+
+The Langevin model of diffusion assumes that the net force on a particular 
+particle is described fully by these individual collisions \cite{Gillespie2012c}. 
+Unlike the Fickian model it provides a full description of the interactions 
+between the target molecule and the surrounding fluid; while also being able
+to describe its motion over any time scale, unlike the Einstein model:
 \begin{align}
@@
 
-Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped 
+\eqref{eq:langevin} provides an accurate description of strongly trapped 
 particles, however the analytical solution of the Langevin equation requires 
@@
 \end{align}
+
 Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
@@
 \end{align}
+
 By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
@@
 \end{align}
+
 Where $w_i$ is a random real number between -1 and 1, we can say that 
@@
 	\label{eq:sim_langevin}
-	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
+	x_i = x_{i-1} - \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
 \end{align}
@@
 force transduction \cite{BergSoerensen2004, Smith2003}.
-
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
@@
 
+Positional data is often acquired in an experimental setting either using
+image analysis or photodiodes to infer the particles position relative to
+the beam focus. The former is often used in cases when precision is not a 
+key concern, as often a standard CCD camera will have limited spacial and 
+temporal resolution. The latter method often requires the use of a quadrant
+photo diode(QPD), the data provided by the QPD will often need to be converted
+to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.
+
 \subsection{Potential Well Analysis}
@@
 \end{align}
+
 which therefore means we can write the potential well as:
@@
 \end{align}
+
 Now assuming our laser acts as a Gaussian beam we should be able 
@@
 
-By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} 
+By inserting \eqref{eq:prob_dist} into \eqref{eq:potential_well} 
 we can fit the potential well in order to determine the trap strength 
@@
 theorem to relate the potential well to the particle's thermal 
-energy using eq.~\ref{eq:prob_dist}:
+energy using \eqref{eq:prob_dist}:
 \begin{align}
@@
 
-If we use eq.~\ref{eq:anylitical_lang} for an optical tweezer we can 
+If we use \eqref{eq:anylitical_lang} for an optical tweezer we can 
 expand out the squared term to get an analytical expression for the 
@@
 	\includegraphics[width=\linewidth]{MSD.png}
-	\caption{Example mean squared displacement using eq.~\ref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
+	\caption{Example mean squared displacement using \eqref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
 The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
@@
 It is also possible to plot the angular MSD (MSAD) using simulative data, 
-however there is yet to be a analytical expression for the MSAD of a 
-freely diffusing particle. Vigilante \textit{et al} \cite{Vigilante2020} 
-derived the upper limit of a dimer's MSAD along its long axis by assuming 
-it was strongly trapped and so had limited angular motion, there expression 
-gives:
+however using this to describe the torque acting on an asymmetric object 
+is difficult due to fact there is no simple approximation of the rotational
+stiffness. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper 
+limit of a dimer's MSAD along its long axis by assuming it was strongly 
+trapped and so had limited angular motion, there expression gives:
 \begin{align}
@@
 
-Vigilante's paper expressed that they couldn't compute $MSD(\tau)$ because 
-they couldn't solve the Einstein-Smoluchowski equation which describes the 
-diffusion constant for dielectric particles. This would require a full 
-description of a particle's electrical mobility and charge distribution 
-- the latter could be achieved via a discrete dipole approximation, the 
-former would be dependent on both the particle's position but relative 
-orientation to the electric field.
+Where, $u_z$ is the unit vector connecting the two spheres of the dimer, 
+$\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness of the trap, and $F$ 
+is Dawson's integral \cite{Oldham2008}. Vigilante's paper expressed that 
+they couldn't compute $MSD(\tau)$ because they couldn't solve the 
+Einstein-Smoluchowski equation which describes the diffusion constant for 
+dielectric particles. This would require a full description of a 
+particle's electrical mobility and charge distribution - the latter could 
+be achieved via a discrete dipole approximation, the former would be 
+dependent on both the particle's position and relative orientation to the 
+electric field.
 
@@
 
-We can fit the Lorentzian via a simplified geometric series 
-$S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion 
-coefficient (in arbitrary units) and the corner frequency $f_c$ 
-which is directly related to the trap strength via 
-$f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression 
-of the mean squared displacement we see two distinct regions, when 
-$f\ll f_c$ the PSD reaches a plateau value that when converted to 
-length units represents the maximum displacement the particle can 
-move beyond the focus, however when $f\gg f_c$ the PSD falls off 
-exponentially which denotes the particle is freely diffusing within 
-the beam focus.
-\begin{figure}[t]
+Where $t_msr$ is the period over which we collected data. The Lorentzian 
+can be fitted via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ 
+which allows us to compute both the diffusion coefficient (in arbitrary 
+units) and the corner frequency $f_c$ which is directly related to 
+the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. The Lorentzian shape
+implies that the trap is harmonic - but not symmetric - which assumes that 
+the particle itself is an isotropic scatterer. Anisotropic scatterers will
+not produce a Lorentzian curve due to the coupled rotation and translation, 
+though depending on the nature of that rotation (i.e. whether it is stochastic
+\cite{Bang2020} or periodic \cite{Yogesha2012}) will effect how drastically it 
+departs from a typical Lorentzian curve.
+
+Like with the analytical expression of the mean squared displacement we 
+see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value 
+that when converted to length units represents the maximum displacement 
+the particle can move beyond the focus, however when $f\gg f_c$ the 
+PSD falls off exponentially which denotes the particle is freely diffusing 
+within the beam focus.
+\begin{figure}[h!]
 	\centering
 	\includegraphics[width=\linewidth]{PSD.png}
-	\caption{Example PSD fitted using Eq.~\ref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water.
+	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water.
 The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector.
-Using a correction factor (see Eq.~\ref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
+Using a correction factor (see \eqref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
 \end{figure}
@@
 \end{align}
+
 Where N is the total number of samples taken, $\Delta x \ \& \ c$ 
@@
 accuracy is necessary. Typically power spectra are recorded using 
-a Quadrant Photo Diode (QPD), which records motion in voltage units
-, not in units of length. To convert between the two we have a few 
-methods: Firstly if we have a mono-disperse suspension of particles, 
-the laser can be scanned across the surface of our target particle; 
-so long as the particle's size is known a linear conversion factor 
-can be used to convert from voltage to length units. Secondly, if 
-the size distribution is very wide but each particle can be accurately 
-sized, then a conversion factor can be approximated by comparing the 
-fitted value of the diffusion coefficient, and the reported value 
+a Quadrant Photo Diode (QPD), which records motion in voltage units, 
+not in units of length. There are two main methods for converting to
+physical units: If multiple photodetectors are available then a 
+differential interference contrast (DIC) system can be used to compute
+the linear relationship between the beads displacement and the photodiodes
+signal \cite{Capitanio2002}. While this is useful when high force precision 
+is necessary it also allows you to collect information about the particle's 
+motion along a single direction, making it ideal for less focused beams. 
+Alternatively, if the size distribution is very wide but each particle can be 
+accurately sized, then a conversion factor can be approximated by comparing 
+the fitted value of the diffusion coefficient, and the reported value 
 given by the Stokes-Einstein relation.
 \begin{align}
-	\label{eq:correction_factor}
+	\label{eq:conversion_factor}
 	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
@@
 the system is known, and the rate of change can be assumed as linear if
-we take a sufficiently small time step that
-$\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll
-\kappa_u/\gamma_R$. In doing so we now only need to compute the
-optical force and torque applied to our dimer, this has already been
-compiled in a MATLAB package called \textit{Optical Tweezer Toolbox}
-or \textit{ott} \cite{Nieminen2007}. In which they use the results
-from \cite{Farsund1996} to compute both the optical force and torque
-using the beam coefficients - the form given by
-\cite{Crichton2000THEMD} is an easier form to compute:
+we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
+\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
+to compute the optical force and torque applied to th dimer.
+
+Using a  MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007} we can compute the beam shape coefficients ($a_{nm}$ \& $b_{nm}$)
+for any desired beam type. Using the results from \cite{Farsund1996} we can then 
+compute both the optical force and torque using the beam coefficients and the 
+scattering coefficients ($q_{nm}$ \& $p_{nm}$) which are found by calculating
+the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The form given by \cite{Crichton2000THEMD} is an easier 
+form to compute for higher orders of $N_{max}$:
 \begin{equation}
@@
 force and torque components in a similar form by applying a simple rotation 
-transformation. In order to get the scattered beam coefficients we can use 
-\textit{mstm} \cite{Mackowski2011} to calculate the T-matrix of our dimer 
-(or any spherical aggregate). Vigilante et al compiled together a python 
-package that combines both \textit{ott} and \textit{MSTM} to simulate the 
-behaviour of spherical aggregates within a predefined optical trap.
+transformation. With the optical forces and torques computed all that remains
+is to compute the Brownian forces and torques which are constrained by the relation.
+\begin{align}
+	\langle q_iq_j\rangle =2D_{ij}\Delta t
+\end{align}
+
+Vigilante et al compiled together a python package that combines both \textit{ott} 
+and \textit{MSTM} to simulate the behaviour of spherical aggregates within 
+a predefined optical trap \cite{Vigilante2020}.
 
@@
 where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave 
-functions, these differ from Eq.~\ref{eq:lornez_mie} in the fact that the 
+functions, these differ from \eqref{eq:lornez_mie} in the fact that the 
 field can either be expressed as an incoming/outgoing wave (with a 
@@
 	\centering
-	\includegraphics[width=0.85\linewidth]{fixed_polarisation.png}
+	\includegraphics[width=\linewidth]{fixed_polarisation.png}
 	\captionsetup{margin=0.5cm}
-	\caption{Total Field incident on the quadrant photo diode for a several different displacements.
-Top Left: $2\mu m $ diameter silica sphere placed at the focus of a  beam.
-Top Right: silica sphere is now displaced by $1 \mu m$ along the x-axis.
-Bottom Right: silica sphere is now displaced by $1 \mu m$ along the y-axis.
-Bottom Left: silica sphere is now displaced by $0.707 \mu m$ along the x and y axis.
-The trapping beam is a $TEM_{00}$ Gaussian, polarised along the along the x-axis.
-$S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
+	\caption{Total Field incident on the quadrant photo diode for a several
+		different displacements. Top Left: $2\mu m $ diameter silica sphere
+		placed at the focus of a  beam. Top Right: silica sphere is now
+		displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere 
+		is now displaced by $1 \mu m$ along the y-axis. Bottom Left: silica 
+		sphere is now displaced by $0.707 \mu m$ along the x and y axis. 
+		The trapping beam is a $TEM_{00}$ Gaussian, polarised along the 
+		along the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of 
+		each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
 	\label{fig:totalfield}
@@
 \begin{figure}[h!]
-	\label{fig:Rohrbach}
 	\begin{subfigure}{0.475 \linewidth}
@@
 Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from \cite{Rohrbach2002}.}
+	\label{fig:Rohrbach}
 \end{figure}

### 2024-07-24 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 quadrant photo diode (QPD) was utilised. 
-
-\begin{figure}
-	\label{fig:setup}
+\begin{figure}[h!]
 	\centering
@@
 Diagram not drawn to scale.}
+	\label{fig:setup}
 \end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In order to accurately capture the dynamics of a trapped particle, 
-a position detection system must be utilised. There are 3 possible
+a position detection system is utilised. There are 3 possible
 methods of position detection: video-analysis, lateral-effect
-position sensing, and quadrant photodiodes. The former being 
+position sensing, and photodiodes. The former being 
 ideally suited for multiple traps or situations where precision
@@
 off the shelf back-focal plane detectors can achieve temporal
-resolutions anywhere from $10-100\ kHz$.
+resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
 
@@
 differentiate between different images), the placement of each lens is shown below.
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.475\linewidth}
+		\includegraphics[width=\linewidth]{signal_generator.jpg}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.475\linewidth}
+		\includegraphics[width=\linewidth, height=0.7\linewidth]
+		{galvano_mirror_controllers.jpg}
+		\caption{}
+	\end{subfigure}
+
+	\caption{Signal generator galvano mirror controller, channel 1 controls the x-axis 
+		mirror, while channel 2 controls the y-axis mirror. Both channels can be manipulated
+		independently.}
+\end{figure}
+
 For our applications a 4f correlator is utilised to ensure that the motion
 of the galvano-mirrors does not move the focal point of the laser, allowing 
-for a stable trap even while in motion.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Calibration of Tweezer Setup}
-Prior to testing optical rotation the trapping laser was calibrated 
-using power spectra analysis
+for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
+the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
+installed $150\ mm$ from the second mirror of the galvano, and the latter being 
+installed $200\ mm$ from the back focal plane of the trapping objective. The 
+signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
+dual channel signal control. This allowed us to precisely control the alignment,
+amplitude, phase, and frequency of both mirrors making alignment much easier. For
+basic trapping calibration the galvano-mirrors were set to a simple DC output, 
+providing a fixed spot which operates like a typical optical trap.
 
@@
 \begin{figure}[h!]
-	\label{fig:vaterite}
 	\centering
@@
 (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
+	\label{fig:vaterite}
 \end{figure}
 
+As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
+demonstrates a Lorentzian curve but modified with these periodic peaks, 
+while the Lorentzian can be loosely fitted to the end tail there exists 
+no current model for describing the power spectra. The closest approximation
+to this was conducted by \cite{Yogesha2012} where they describe the 
+rotational motion of ellipsoidal polystyrene particles. The critical assumption
+being that the particle perfectly rotates in the $x-y$ plane; however it
+has long been suspected that birefringent microspheres experience random
+rotations outside of the $x-y$ plane \cite{Bang2020} making it very difficult 
+to characterise the forces on rotating microspheres without a proper understanding 
+of the optical torque being applied to it.
+
+\subsection{Liquid Crystal Rotors}
 Liquid crystals are an intriguing example of materials with mixed phase
@@
 \begin{figure}[h!]
-	\label{fig:vaterite_shear}
 	\begin{subfigure}{0.5\linewidth}
@@
 Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
+	\label{fig:vaterite_shear}
 \end{figure}
@@
 
+It is not impossible that fluid shearing could be used in the future to localise
+nucleation; but from these results, using individual micro-rotors do not seem like 
+an appropriate method. If multiple micro-rotors could be trapped in close proximity 
+to one another they could create a large region of fluid where nucleation is more 
+likely than the bulk fluid. Micro-rotors have been created that allow for precise
+control of suspended micro-particles \cite{Butaite2019} and could potentially be
+used to generate sufficient shearing, however these could not be used in this PhD 
+as we lacked the necessary hardware to form multiple gradient traps.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 for a more complex path, such as an elliptical orbit the curve needs 
-to be parametrised. One can describe the position parameter of an ellipse 
-as such:
+to be parametrised. One can describe the position parameter of a circular 
+path as such:
 
 \begin{align}
-	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
+	r(u) = \left[rcos(2\pi u), rsin(2\pi u), 0 \right]
 \end{align}
 
-where a and b are the different characteristic radii of an ellipse, if 
-we say that $u$ describes time from some initial point we can say $u=t\omega$ 
+If we say that $u$ describes time from some initial point we can say $u=t\omega$ 
 where omega is the frequency of orbit. Substituting this in and then taking 
@@
 \begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
+	v(t) = \frac{dr(t)}{dt} = \left[-2\pi r\omega \ sin(2\pi t\omega),
+	\ 2\pi r\omega \ cos(2\pi t\omega), 0 \right]
 \end{align}
@@
 \begin{align}
-	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
+	u_r(r)=-|v(t)|^2cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
 \end{align}
@@
 \begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
+	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
 \end{align}
 
-Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path
+Moving a silica bead along an circular path can generate significant fluid 
+flow around a larger volume compared to comparable microspheres.
 
@@
 Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
-event in supersaturated glycine solution (S=1.03), the beam is a stationary being 
+event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
 $\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
-nucleus forms at the trap focus, growing quickly until after roughly $6$ seconds 
-the crystal escapes the trap. The key take away to remember is that the beam has no 
-real influence over the crystal front,instead the crystal grows outward from the 
-centre of the focus.
+nucleus forms at the trap focus, growing quickly (growth rate was approximated using
+imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
+until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
+using optical tweezers shows that the growth rate is only loosely connected to the 
+solutions supersaturation, local conditions play a much larger role in the growth 
+rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
+is escaped is due to the fact that crystal is far too large to be held in place and 
+is in fact still growing as the solution is supersaturated. The key take away to 
+remember is that the beam has no real influence over the crystal shape, instead 
+it grows outward from the trap focus.
 \begin{figure}[h!]
-	\label{fig:stationary_beam}
 	\centering
@@
 	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
-	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the
-	trap.}
+	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the trap.}
+	\label{fig:stationary_beam}
 \end{figure}
@@
 using a galvano-mirror to move the particle through the solution, nucleation events
-were observed in which the crystal front follows the path of the laser. See below 
-in \textcolor{red}{Fig.XX} where we have the laser focus moving in a small elliptical 
-pattern,
+were observed in which crystal growth follows the laser focus. Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus moving in a small elliptical 
+pattern; interestingly, due to the addition of the silica suspension we see the presence
+of small spheres, these appear too small to be silica beads which have a uniform radius
+of $1.57\ \mu m$, we surmise these could be small clusters of Glycine that had previously
+been shown to form when the aqueous solutions where irradiated with a focused laser
+\cite{Tsuboi2009, Gowayed2021}. While no droplets are seen directly entering the focus a 
+nucleus forms close to the droplet edge, unlike in Fig.~\ref{fig:stationary_beam} the 
+crystal does not grow out from the focal point evenly; due to the galvano mirror, 
+the crystal is simultaneously being moved by and growing around the focal point of 
+the trap. Because of this the crystal nucleus lacks a clear morphology at first, 
+until roughly $20\ s$ the crystal reaches a almost prismatic structure, with further 
+irradiation increasing the size. Interestingly the galvano-mirror allows the trap to 
+impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) 
+and (d), where even though the crystal is not directly in the trap focus it rotates in 
+the $x-y$ plane and gets trapped again at a corner. The rotation could not be due to fluid
+flow close to the surface of the crystal as trap should have no influence on fluid molecules
+smaller than a few hundred nano-meters. In figs.\ref{fig:eliptical_beam_1}(e) and (f), 
+the crystal growth is localised to the corner, even though previously the crystal growth 
+was roughly uniform (compare figs. \ref{fig:eliptical_beam_1}(b) and (c)). The area growth
+rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at
+$45.03\ \mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and (f) the 
+growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.9\linewidth]{frames_eliptical_beam.pdf}
+	\caption{Laser induced nucleation at the edge of a droplet of 
+		supersaturated glycine solution. (b) shows the first instance 
+		of a crystal nucleus, growing quickly through (c)-(e) until 
+		after $6.665\ s$ the crystal begins to escape the trap.}
+	\label{fig:eliptical_beam_1}
+\end{figure}
+
+In this instance $20\ \mu L$ of glycine and water ($S=1.03$) was added to $10\ \mu L$ 
+of a dilute water-silica mixture making the solution undersaturated ($S\approx0.7$). 
+Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
+\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both cases the
+crystal morphology is not influenced by the location of the laser focus, it has been
+shown that polarisation can influence the polymorphic control for simple salts 
+\cite{Garetz1996, Garetz2002} but no research has been done into the influence of a 
+moving beam front on the growth of a newly formed nucleus.
+
+\section{Influence of a moving beam front on seed crystals}
+To see if the above phenomena had any direct influence on a

### 2024-07-24 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 function was predicated on the fact that any variations in the signal are due to
-the particles translational motion within the confines of a cylindrical trap.
+the particles translational motion within the confines of a cylindrical trap 
+\cite{BarZiv1997}. However, it is somewhat naive to assume that the orientational 
+change
 
@@
 \section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to characterise due to the coupling of rotational and translational effects.
-As shown in the previous chapter the dynamics of even simple dimers are heavily dependent on the particle's position and orientation, this is reflected in literature where several engineering solutions have been devised to decouple translational and orientational motion \textcolor{red}{[ , ]}.
-While a majority of latter work has been focused on nano-particles (falling into the Rayleigh Regime) and utilising florescence
+Orientational dynamics to an anisotropically scattering shape are difficult to 
+characterise due to the coupling of rotational and translational effects. As 
+shown in the previous chapter the dynamics of even simple dimers are heavily 
+dependent on the particle's position and orientation, this is reflected in 
+literature where several engineering solutions have been devised to decouple 
+translational and orientational motion \textcolor{red}{[ , ]}. While a majority 
+of latter work has been focused on nano-particles (falling into the Rayleigh 
+Regime) and utilising florescence
 
 \subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page.
-The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
+In the case of our probe beam the +z direction points from the surface of the 
+probe directly to the centre of diffusion in our dimer, with the +y direction 
+pointing towards what would be our trapping beam, and the +x direction pointing 
+into the page. The origin of our coordinate system is fixed on the centre of 
+the focus of our trapping beam, meaning as our dimer's centre of diffusion moves 
+the origin is kept constant.
+
+\subsection{Beam}
+The probe is a defined as a plane wave by setting the Gaussian beam parameter 
+$C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector 
+as $[1,0,0,0]$. The plane wave in \textit{mstm} is given by \eqref{eq:incident}.
+
 \subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam.
-We scale the position of each sphere by the factor L:
+The dimer is defined by two spheres, in the trapping frame this would be orientated 
+by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as 
+the large sphere is furthest from the trapping beam. We scale the position of each 
+sphere by the factor L:
 \begin{align}
@@
 
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
-\subsection{Beam}
-Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0.
-We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.
+Where $\lambda$ is the wavelength of light and is given in nanometres, this means 
+every position must be also provided in nanometres.
 
 \subsection{Detectors and Pixels}
-
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
+Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted 
+later for testing), with the position of the detector being defined by the polar 
+and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
 \begin{align}
@@
 \end{align}
-So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components.
-The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.
+So if you were to place the fibres in the trapping frames x-y plane (i.e. flat against
+the cover slip in a covenantal trapping context), in the probe beam frame will only 
+have x and z components. The orientation of the detector is just the -ve of its 
+position and not scaled by the distance term r, the orientation being the vector from 
+the origin of the detector to the center of the dimer.
 
-Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector.
-The pixels are distributed using a sunflower distribution given by:
+Pixels can be thought of as points lying on the surface of the detector, by getting 
+the spherical coordinates of each pixel we can compute the intensity at each point on 
+the detector, allowing us to determine the average intensity on a given detector.
 
-Where each point is first scaled to the size of the detector - for now we assume a radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the surface of the detector to the main coordinate system.
-Their orientation is treated much in the same as the detector's orientation, being its -ve position and scaled down by the radial distance from the dimer to the pixel, making it a unit vector.
-From their position we can also grab their spherical coordinates to create a list of points for mstm to evaluate.
-In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
+Where each point is first scaled to the size of the detector - for now we assume a 
+radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the 
+surface of the detector to the main coordinate system. Their orientation is treated 
+much in the same as the detector's orientation, being its -ve position and scaled down 
+by the radial distance from the dimer to the pixel, making it a unit vector. From 
+their position we can also grab their spherical coordinates to create a list of points 
+for mstm to evaluate. In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
 
-This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's.
-For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.
+This allows us to define the surface of any detector perpendicular to the origin of 
+our coordinate system, we can freely choose the size and resolution of our fibre's. 
+For every pixel we define the spherical coordinates relative to our origin, this is 
+added to a list of $\theta, \phi$ pairs that are evaluated by \textit{mstm}.
 
 \section{Interpretation of scattering data into orientation estimates}
-MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation.
-While the latter method is satisfactory for simulation purposes it
+Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any 
+point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere 
+to the centre of the smaller sphere. A plane wave probe beam is incident on 
+the trapping laser, is incident on the dimer, generating a scattering pattern
+dependent on the dimer's orientation $I(\hat{\bf s}, \theta, \phi)$ which can be 
+computed using \textit{mstm} \cite{I.Mishchenko1996}. To represent the experimental 
+set up consisting of a set of optical fibres recording scattered light, we choose 
+four sets of spherical angles [$(\theta_1,\ \phi_1), \ (\theta_2,\ \phi_2), \ 
+(\theta_3, \ \phi_3), \ (\theta_4,\ \phi_4)$] and record the calculated intensity 
+at each angle $I(\hat{s},\ \theta_k,\ \phi_k)$.
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig2a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig2b.png}
+	\end{subfigure}
+	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
+	\label{fig:dimer}
+\end{figure}
+
+Our goal is to determine the orientation of the trapped dimer based on the measured 
+intensity $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate 
+of the dimer's orientation, for the purposes of interpretation of the scattering and 
+optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then 
+use as 'classification categories' in a neural network methodology to map scattering 
+data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} 
+\ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a 
+unit sphere \cite{Reyuthor2006} (Figure~\ref{fig:dimer}b) leading to a maximum 
+nearest-neighbour spacing between two neighbouring reference orientations of 0.895 
+radians. Using MSTM we compute the raw intensities at each of the measurement angles 
+that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary 
+there are several constraining factors that limit our ability to infer useful information 
+from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown 
+of our choice of detection angles. The raw intensities are normalized according to:
+\begin{align}
+	\label{eq:scale}
+	y_k(\hat{\bf n}_\alpha)
+	= 
+	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
+	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
+\end{align}
+
+where the denominator is simply the standard deviation across the set of values
+$I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled 
+signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
+
+Note that the collected scattering signals are not necessarily simply related to 
+their associated reference orientations: as is well known from such examples of 
+the inverse scattering problem. While it is trivial to compute the light scattering 
+pattern for any given particle with any particular characteristic (i.e. size, 
+shape, or orientation), inferring the light scattering from a unknown particle 
+to determine said characteristic is incredibly difficult due to complex mapping 
+between scattering and said characteristic. Even if the orientation space is divided 
+evenly between reference orientation the subsequent signal space ends up being 
+appearing mixed making simple comparisons of signals useless for inferring information 
+on the particle. Shown below is two clusters of orientation vectors and there 
+respective measured scattering signals - the points have been coloured based on 
+their proximity to the centre of their respective cluster. While the orientation 
+space appears tightly packed and ordered the signal space quickly spreads out in 
+an asymmetric fashion. Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
+mapping can intersect itself which only further increases the complexity. While 
+in some instances the mapping between one reference orientation and another is 
+discrete, in other instances the mapping becomes far more complex to discern.
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig3a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig3b.png}
+	\end{subfigure}
+	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals.
+Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
+	\label{fig:mixing}
+\end{figure}
+
+Nevertheless, at least where the uncertainty in signal measurements is low (see 
+below), we can predict the orientation from the scattering by utilising 
+computational techniques such as neural networks. We thus utilised the Python 
+machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a 
+neural network for identifying the dimer's orientation from its light scattering 
+signal. The network was trained by generating a database of random orientation 
+vectors, calculating the corresponding light scattering signals, and then using 
+the network to estimate the probability of a given signal coming from a dimer in 
+a given reference orientation. The network's loss function was evaluated and used 
+to improve the estimation, the network being trained until the improvement in the 
+loss function was less than 0.0001.
+
+Importantly, the estimation provided by the neural network can be improved further 
+by accounting for any prior information we know about the dimer, utilising Bayesian 
+inference to update the neural network's estimation: 
+\begin{align}
+	\label{eq:bayes}
+	p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
+	&=
+	\frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
+		p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
+\end{align}
+
+where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of 
+the distributions of particle orientations and instantaneous signals, respectively.
+\textit{Without} any prior evidence we must assume that the orientation prior of 
+the dimer $p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
+dimer's possible current orientation from knowledge of previous measurements can 
+be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$. The latter prior 
+$p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$). This is 
+given by taking the discrete integral over the collection of reference orientations:
+\begin{align}
+	p(y_1, y_2, y_3, y_4)
+	=
+	\sum_{\alpha=1}^{n_{\rm ref}}
+	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
+	p(\hat{\bf n}_\alpha)
+\end{align}
+
+From \eqref{eq:bayes} we obtain the key result, a mass probability distribution 
+denoting the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ 
+given a measured signal ($y_1$, $y_2$, $y_3$), \textit{i.e.} an estimated mapping 
+from scattering measurement to orientation estimate.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Calculation of error}
+\label{sec:divergence}
+To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$.
+MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$.
+Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}.
+An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
+\begin{align}
+	\label{eq:best}
+	p_{best} = 
+	\begin{cases}
+		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
+		0 & \text{anywhere else}
+	\end{cases}
+\end{align}
+In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
+probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
+\begin{align}
+	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
+	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
+	\right]
+	\label{eq;kullback}
+\end{align}
+where a larger value of $K_l$ indicates that our model is less confident in its
+prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
+in the estimated dimer orientation probability --- a distribution strongly peaked at 
+some value would give us more confidence in that value than a near-uniform distribution 
+where the scattering measurement could imply a wide range of possible orientations --- 
+but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Brownian Simulation}
+\label{sec:brownian}
+
+We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
+Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+shaped sphere clusters.
+We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid.
+MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}.
+The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
+We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.
+We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}.
+We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.
+
+The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200.
+For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig4a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.4\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig4b.png}
+	\end{subfigure}
+	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
+	\label{fig:detectors}
+\end{figure}
+
+When all of the detectors lie in the same plane the expected signal can appear identical despite the dimer being in completely different orientations.
+This is shown in Figure~\ref{fig:detectors} which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels is because the dimer's scattering is indistinguishable in these two reference orientation.
+It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors.
+More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
+This is illustrative of the difficulty behind the inverse light scattering problem; as one cannot always map a given signal to a particular parameter value.
+
+To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique.
+As seen between Figures \ref{fig:detectors}a \& b each reference orientation now has a unique scattering signal, though with only three detectors the difference in expected signals can appear insignificant.
+By adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
+In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing the Model}
+\label{sec:test}
+Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.45$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
+The trapping laser is 1064nm NIR focused through a 1.25 NA objective.
+The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively.
+We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.
+
+We applied Eq.~\eqref{eq:bayes}, taking the reference orientation with the highest probability  as our estimate of the dimer's instantaneous orientation $\hat{\bf n}_{est}$.
+To visualise the model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time.
+For comparison, we also plotted the radian distance between the  dimer's instantaneous orientation and the closest reference orientation, denoted $\hat{\bf n}_{best}$.
+The dotted line indicates the maximum radian distance ($0.896$ radians) between two \textit{neighbouring} reference orientations: if we are under this line then we know our estimate is at least neighbouring the best result.
+Assuming a uniform prior of the reference orientations $p(\hat{\bf n}_\alpha)$  the neural network's predictions ($\hat{\bf n}_{est}$ from Eq.~\eqref{eq:bayes}) are at times reasonable, but there are significant large and random jumps away from the correct result (Fig.~\ref{fig:uniform}).
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{fig5.png}
+	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity.
+Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
+Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
+	\label{fig:uniform}
+\end{figure}
+
+One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering.
+Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step.
+To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness.
+Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$.
+Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:
+
+\begin{align}
+	p(\hat{\bf n}_\alpha)
+	&=\frac{e^{\beta (\hat{\bf n}_\alpha 
+			\cdot \hat{\bf n}_{est}(t-\Delta t))}}
+	{\sum_{\alpha=1}^{n_{\rm ref}}
+		e^{\beta (\hat{\bf n}_\alpha 
+			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
+	\label{eq:boltz}
+\end{align}
+Here $\beta$ is a weighting factor describing the dimer's freedom of motion within the trap.
+As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{fig8a.png}
+	\caption{\label{fig:biased}
+		%
+		Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
+		by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
+		achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+		Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
+		%
+	}
+\end{figure}
+
+The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}.
+By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
+To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
+\begin{align}
+	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#}
+	\\
+	K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right]
+	\\
+	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
+\end{align}
+
+The worst case scenario is akin to randomly choosing a reference
+orientation at each time step. The greater the value of $F(K_l)$, the
+better our model's confidence is in characterising the dimer's
+motion. Because our model is dependent on several parameters we need
+to a sophisticated method for understanding how these parameters
+correlate with $F(K_l)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Asymmetric dimer dynamics}
+\label{sec:motion}
+A simulation of a asymmetric dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) 
+trapped in an off-axis trapping orientation was used as a test case for our 
+model, since it was previously shown that dimers can be optically rotated
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig7a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig7b.png}
+	\end{subfigure}
+	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
+\end{figure}
+
+In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state.
+We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer.
+In the future we hope to further investigate the motion of asymmetric dimers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Accounting for sources of error in light scattering measurements}
+When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor.
+Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle.
+However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance.
+We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors.
+We highlight how we address these areas below.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subparagraph{Impact of incorrect dimer sizing}
+\label{sec:lam}
+
+One of the main limitations of our model is that we assume that the dimer being modelled in MSTM is accurate to the dimer being trapped in the optical tweezer.
+Sizing molecules accurately is a significant challenge for single particle analysis so there is bound to be some uncertainty with the measurements.
+We ran our model 3 times with the neural net being trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.33\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig8a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig8b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig8c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
+	\label{fig:size}
+\end{figure}
+
+As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model.
+This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
+This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
+We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
+While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig9a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig9b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig9c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
+	\label{fig:refs}
+\end{figure}
+
+Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer.
+This suggests that overshooting our size estimate will be less detrimental to our estimation.
+Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations.
+This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction.
+Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subparagraph{Impact of measurement noise on model predictions}
+\label{sec:epsilon}
+
+So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it.
+In reality of course scattering signals will always have some non-zero measurement noise.
+This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself.
+To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
+\begin{align}
+	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
+\end{align}
+where $\epsilon$ is the percentage error associated with the scattering signal.
+Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
+	\label{fig:epsilon}
+\end{figure}
+
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of signal noise quickly leads to a decrease in the model's performance.
+This is due to an inherent feature of the inverse scattering problem: two distinct regions in orientation space can become heavily intertwined and thus no longer well separated when mapped to intensity space (even though the mapping remains continuous): so even small uncertainties in the scattering data can lead to large 'mistakes' in the choice of orientation by the neural network.
+(Indeed if this was not the case the inverse scattering problem would be quite simple.)
+
+To reduce the effects of the signal noise we took the time average of the expected signal over $0.001 s$ and then had our neural network estimate the orientation based on the average signal.
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $1\,{\rm ms}$}
+	\label{fig:time average}
+\end{figure}
+This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model.
+There appears to be no clear correlation between the length over which we time average and the performance of our model.
+Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network.
+Fortunately, time averaging even over $1\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
+
+From the above discussion it’s clear that estimation of the dimer’s orientation is a problem that can be endlessly tuned to fully maximise our end result.
+Here we simplify the problem somewhat by employing a relatively small finite number of 'reference orientations' to map between scattering and dimer orientation: the precision of estimation could be improved by utilising a greater number of reference orientations, although there remains a balance between the realisable precision of orientation estimate and the noise level of the scattering measurement.
+Another avenue to further explore would be using the method to optimise the choice of detection angles, essentially to find the region in the mapping between measured scattering and orientation that others the best degree of confidence through optimal separation of scattering signals for distinct orientations.
+For sequences of data such as dynamic measurements, a further potential enhancement would be to consider more complex correlations based on prior expectations of the dynamics.
+Here already we improve the method using a non-uniform prior based on only the immediately previous measurement in time (see Section 2.1): considering a non-uniform grouping of reference orientations might result in a better estimation, if we have information regarding the dimer’s preferred axis of rotation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+\label{sec:Conclusion}
+
+We have developed a method for measuring the dynamics of an optically-trapped colloidal objects based purely on measurements of the object's light scattering at a small number of detection angles.
+We demonstrate the method using the orientation of an asymmetric dimer as the dynamic variable and object of interest respectively, but in principle the model can be applied to any characteristic that impacts the light scattering pattern produced by a trapped entity such as size and shape.
+The MSTM package is a flexible tool for calculating the light scattering of complex objects using a representation of the object as a set of micro-particles, enabling training of a neural network to enable categorisation of the mapping between scattering and trapped object characteristics.
+By taking account of the physically realistic behaviour of the trapped object and the characteristics of the trap (which impact the dynamics of the object), the Bayesian inference method can be re ned to provide a reliable estimation of object characteristics of interest, even in the presence of measurement noise.
+Fundamentally, the inverse scattering problem is difficult to solve, since the mapping between object characteristics and scattering can be highly complex.
+We determined the minimum number of detectors required for a reliable estimation in the presence of measurement noise; furthermore, we demonstrated that the arrangement of these detectors is critical for a reliable estimation of an objects orientation.
+However, Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.

### 2024-07-24 python_files/test.py

@@ -1,34 +1,41 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import quaternion
-from quaternion import as_spherical_coords
+import scipy.misc as misc
 
-with open('C:\\Users\\xbb16146\\dmaciver97\\python_files\\sphere_in_LCP_Quadrants.txt', 'r') as in_file:
-    lines = in_file.readlines()
-    x_dat, y_dat, z_dat, = [], [], []
-    q1, q2, q3, q4 = [], [], [], []
-    index = []
-    for line in lines[1:-1]:
-        items = line.split(',')
-        x_dat.append(float(items[0])+ float(items[1])- float(items[2]) - float(items[3])) 
-        y_dat.append(float(items[0])+ float(items[2])- float(items[1]) - float(items[3]))
-        z_dat.append(float(items[0])+ float(items[1])+ float(items[2]) + float(items[3]))
-        q1.append(float(items[0]))
-        q2.append(float(items[1]))
-        q3.append(float(items[2]))
-        q4.append(float(items[3]))
-        index.append(int(float(items[4])/1e-5))
+p = 1;                  # Degree of LG mode
+l = 2;                  # Order of LG mode
+w0 = 2.0;               # Beam waist
+k = 2*np.pi/532.0e-9;   # Wavenumber of light
 
-traj = np.load('Sphere_in_LCP_Trajectory.npy')
-x_disp, y_disp, z_disp = [], [], []
-for i in index:
-    x_disp.append(traj[i,0])
-    y_disp.append(traj[i,1])
-    z_disp.append(traj[i,2])
-    q = np.quaternion(traj[i,3],traj[i,4],traj[i,5],traj[i,6])
-    
-plt.scatter(x_disp, x_dat)
+zR = k*w0**2.0/2;       # Calculate the Rayleigh range
+
+# Setup the cartesian grid for the plot at plane z
+z = 0.0;
+xx, yy = np.meshgrid(np.linspace(-5, 5), np.linspace(-5, 5));
+
+# Calculate the cylindrical coordinates
+r = np.sqrt(xx**2 + yy**2);
+phi = np.arctan2(yy, xx);
+
[…]

### 2024-07-31 Latex_files/Thesis/chapters/chapter 2/main.tex
corrections made to chapters 2 and 4 of thesis doc

@@
 spatially locked by aiming a laser with a Gaussian intensity profile 
-directly upward at a the cell while being suspended in a liquid medium 
-\cite{Ashkin1970}.
+directly upward at a particle while being suspended in a liquid 
+medium \cite{Ashkin1970}.
 
@@
 in the x-y plane the net force is unbalanced and points towards the centre 
-of the trap; this principle can be generalised to 3 degrees of freedom if we 
+of the trap. This principle can be generalised to 3 degrees of freedom if we 
 consider a focused beam instead of a para-axial beam, typically the trap will be
@@
 Proper understanding of optical tweezing requires an understanding of
-how the trapped particle interacts with the trapping laser. From a 
+how the trapped particle interacts with the trapping laser. From an 
 electromagnetism perspective the laser creates a spatially and 
@@
 
-These 4 equations describe how the electric and magnetic fields behave
-and are related to one another at a microscopic level. Any discussion 
+These four equations describe how the electric and magnetic fields behave
+and relate to one another at a microscopic level. Any discussion 
 of optical trapping is underpinned by the fact that in all cases the 
-Maxwell Equations must be satisfied at the macroscopic level, where 
+Maxwell equations must be satisfied at the macroscopic level, where 
 one must account for the medium's interactions with the EM field. 
-Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
+Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
 \begin{align}
@@
 caused by the presence of free charges in the medium, and $\mathbf{D}$ 
-and $\mathbf{H}$ are the displacement and magnetizing fields respectively. 
+and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
 The latter two are defined by:
@@
 
-The Ray-Optics model is the simplest to understand, the theory models 
+The Ray-Optics model is the simplest to understand, this theory models 
 the laser as a collection of individual 'rays' that propagate and are 
@@
 Gaussian intensity distribution $P_i$ will fall off as you move from
-the centre of the beam.
+the centre of the beam. The ray optics model is ideal when dealing 
+with larger particles whose diameter far exceed the wavelength of light
+being scattered. While it can be useful in predicting the forces 
+experienced by said particle's it does not fully capture the behaviour
+of light when considering interference between different rays.
 
@@
 
-The Lorenz Mie theory provides an exact solution to the maxwell equations 
-for the scattering caused by an isotropic sphere. The theory describes the 
+The Lorenz-Mie theory provides an exact solution to the Maxwell equations 
+for the scattering caused by an isotropic sphere. This theory describes the 
 scattered wave given off by a dielectric sphere when incident by a plane 
@@
 
-Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
-This combined with the constraints of Maxwell's Equations leaves 
-very few exact solutions apart from spherical or planar waves, 
-both of which can be converted between with relative ease. For 
-example, a plane wave electric field can be expanded into spherical 
-harmonics and likewise any spherical wave can be described as 
-combination of plane waves offset from one another. For a single 
-plane wave the incident, internal, and scattered fields are given as: 
+Where $k$ is the wave number of the electromagnetic radiation 
+($k = 2\pi/\lambda$). This combined with the constraints of 
+Maxwell's equations leaves very few exact solutions apart from 
+spherical or planar waves, both of which can be converted between 
+with relative ease. For example, a plane wave electric field can 
+be expanded into spherical harmonics and likewise any spherical 
+wave can be described as a combination of plane waves offset from 
+one another. For a single plane wave the incident, internal, and 
+scattered fields are given as: 
 \begin{align}
@@
 be found by computing the Maxwell stress tensor which only requires
-the total magnitude of the incident and scattered fields, essentially
-computing the momentum contained in the incident beam and how much of
-it has been transferred to the target particle.
+the total magnitude of the incident and scattered fields. By computing 
+the momentum contained in the incident and scattered fields we can measure
+the total momentum transferred to the target particle.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Rayleigh Regime}
+\subsection{Rayleigh Scattering}
 
 The Rayleigh approximation is the limiting approximation for describing 
-a particles motion in an electromagnetic field who's wavelength is several 
+a particles motion in an electromagnetic field whose wavelength is several 
 times greater than the particle's size. The underlying theory is that a 
@@
 Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
-the beam waist, and $a$ is the radius of the particle. Scattering force 
-however is dependent on the effective scattering cross sectional area. 
+the beam waist, and $a$ is the radius of the particle. The scattering force 
+however, is dependent on the effective scattering cross-sectional area. 
 \begin{align}
@@
 scattering forces due to the assumption that the particle is a point 
-dipole, so much so that higher order scattering problems can be 
+dipole. So much so that higher order scattering problems can be 
 simplified by subdividing the particle into discrete dipoles (see 
-Sec~\ref{sec:scattering}). However as the target particle gets larger 
+Sec~\ref{sec:scattering}). However, as the target particle gets larger 
 this assumption fails to accurately describe the trapping force due 
 to the complexity in the gradient field. For particle sizes close to 
-the laser wavelength the scattered field is best described by Lorenz-Mie theory.
+the laser wavelength the scattered field is best described by Lorenz-Mie 
+theory.
 
@@
 There are several methods available to calculate the scattered field 
-produced by a particle, while one can compute this from directly applying
-Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
-tedious for some applications.  
+produced by a particle. While one can compute this by directly applying
+Lorenz-Mie theory, describing the interactions between focused beams and 
+complicated shapes can be computationally slow and memory intensive. As a
+result, several alternative scattering approaches have been developed that
+allow for faster computation times simplifying the scattering problem.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 research into acoustic wave scattering \cite{Waterman1969}, this 
-would later be extended to electromagnetic waves. Sometimes being 
+would later be extended to electromagnetic waves. Sometimes  
 referred to as the extended boundary condition method (ECBM), 
@@
 over the targets surface. These currents are chosen so that the
-electric field outside is identical to the original problem \cite{Wriedt1998}. 
-This choice of currents negates the need to compute the internal fields
-which reduces the scattering problem to a system of linear equations 
-relating the incident beam coefficients to the scattered beam coefficients
-the relationship between each can be summarised as:
+electromagnetic field outside is identical to the original problem 
+\cite{Wriedt1998}. This choice of currents negates the need to 
+compute the internal fields which reduces the scattering problem 
+to a system of linear equations relating the incident beam coefficients 
+to the scattered beam coefficients. The relationship between each can 
+be summarised as:
 \begin{align}
@@
 $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
-this PhD in particular we utilise an extension of the ECBM known as 
+this project we utilised an extension of the ECBM known as 
 multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
@@
 the medium in the presence of an electric field. With the internal field 
-calculated the scattered field can be computed, one of the primary advantages
-of DDA over the T-matrix method is that the the composition of the target can 
+calculated, the scattered field can be computed, one of the primary advantages
+of DDA over the T-matrix method is that the composition of the target can 
 be changed freely. When comparing different computational scattering methods,
 the ECBM was found to be better suited for simulating the scattering of symmetric
-targets as the ECBM can use the targets symmetry to speed up calculations \cite{Wriedt1998},
-however when dealing with inhomogeneous media DDA is more efficient compared to
-ECBM.
+targets as the ECBM can use the target's symmetry to speed up calculations \cite{Wriedt1998}. But when dealing with inhomogeneous media DDA is 
+more efficient compared to ECBM.
 
@@
 strength $\kappa$. Assuming the only external force acting on our particle 
-is the laser the net force should be exactly equal to force of the stochastic
+is the laser, the net force should be exactly equal to force of the stochastic
 collisions due to the fluids thermal energy. If we focus our analysis when 
-the particle is stably trapped and assume that the trap is harmonic we can 
+the particle is stably trapped and assume that the trap is harmonic, we can 
 model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
@@
 \eqref{eq:langevin} provides an accurate description of strongly trapped 
-particles, however the analytical solution of the Langevin equation requires 
-integration of the white noise term making it difficult to simulate the 
-trajectory of a given particle \cite{Volpe2013}. Instead it is often far 
-easier to solve the equation numerically and apply use the analytical 
+particles. Despite this the analytical solution of the Langevin equation requires 
+integration of the noise term making it difficult to simulate the 
+trajectory of a given particle \cite{Volpe2013}. Instead, it is often far 
+easier to solve the equation numerically and use the analytical 
 solution to calibrate and extract information about the particle and fluid, 
@@
 term n and the actual expression. If we limit our approach to the first 
-derivative only we find that for sufficiently small values of $R_1$ the 
+derivative only, we find that for sufficiently small values of $R_1$ the 
 velocity and acceleration can be approximated as:
@@
 By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
-velocity and acceleration based on the previous time steps we can 
-discretize the position by taking finitely small  time steps (i.e. 
+velocity and acceleration based on the previous time steps, we can 
+discretise the position by taking finitely small  time steps (i.e. 
 $x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
-white noise as no information is known about $W(t)$ at any time. We 
+noise, as no information is known about $W(t)$ at any time. We 
 can instead say that the velocity of a Brownian particle should 
-approximate our white noise as a random walk, where at each new time 
-step the position changes randomly within a given range. Constricting 
-the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white 
+approximate our noise as a random walk, where at each new time 
+step, the position changes randomly within a given range. Constricting 
+the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
 noise using the finite-difference approach as:
@@
 particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
-Now in the case of an optical trap the restoration time scale is 
+Now in the case of an optical trap, the restoration time scale is 
 given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
-is far greater than the characteristic time, therefore for simulation 
-purposes we can neglect the particle's inertia which allows us to 
-write the motion of an optically trapped particle as:
+is far greater than the characteristic time. Therefore for simulation 
+purposes, we can neglect the particle's inertia which allows the motion 
+of an optically trapped particle to be written as:
 \begin{align}
@@
 stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal 
-energy of the system $k_BT$ - with the latter two being related by 
-Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$. 
-Therefore by calculating these parameters to a high degree of precision 
-allows one to get precise description of the forces experienced by a 
+energy of the system $k_BT$ (with the latter two being related by 
+Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$). 
+Therefore, by calculating these parameters to a high degree of precision 
+allows one to get a precise description of the forces experienced by a 
 target particle, which in the past has been used for highly accurate 
@@
 \begin{equation}
+	\label{eq:optical_force}
 \begin{split}
@@
 where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave 
-functions, these differ from \eqref{eq:lornez_mie} in the fact that the 
+functions, these differ from \eqref{eq:incident} in the fact that the 
 field can either be expressed as an incoming/outgoing wave (with a 
@@
 
-We can if we want further restrict the incident beam by applying setting 
-the truncation angle to match our microscope object, this essentially 
-applies a cut off point to the In order to compute the scattering from 
+We can if we want further restrict the incident beam by setting 
+the truncation angle to match our microscope object, this applies a 
+hard cut off to the incident beam to better represent experimental 
+conditions. In order to compute the scattering from 
 the target particle \textit{ott} uses the $T$-matrix method, this is not 
@@
 
-Rather than displacing and rotating the particle itself, we instead rotate 
-and displace the incident beam and recalculate the beam coefficients; the 
-total beam is then inversely rotated in order to record the QPD signal. This
-ensures that for a homogenous sphere, even if the sphere itself undergoes
-rotation the QPD signal is unaffected by the change in orientation.
+While translational motion have no knock on effects to the QPD signal, 
+rotational changes are between the incident and total fields. This results
+in rotational motion being biased in the QPD signal - even when collecting
+signals from isotropic scatterers. To remedy this an inverse rotation is
+applied to the total field.
 
@@
 set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
-produced the following response curve.
+produced the following response curve:
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
-Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from \cite{Rohrbach2002}.}
+	\caption{Comparison between QPD response signal versus work conducted by 
+		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal 
+		produced by QPD using \textit{ott} and points represent the signal 
+		response collected from \cite{Rohrbach2002}.}
 	\label{fig:Rohrbach}
@@
 
-With this any trajectory can be collected from the QPD by displacing and 
+The discrepancy between our simulated QPD and the results from 
+Rohrbach can be attributed to the fact that the QPD signal  With this any trajectory can be collected from the QPD by displacing and 
 rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} 

### 2024-07-31 Latex_files/Thesis/chapters/chapter 4/main.tex
corrections made to chapters 2 and 4 of thesis doc

@@
 \label{chapter:langevin_dynamics}
-Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed.
-However, while working with dense colloidal suspensions, one often ends up trapping more than one sphere.
-Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness.
-Theoretical studies on the case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size.
-Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
-There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
+Much of the calibration theory discussed in Chapter 2 assumes that the
+target particle in question is a single sphere, one who's scattering and 
+motion is easily computed. However, while working with dense colloidal 
+suspensions, one often ends up trapping more than one sphere. Li and 
+Arlt \cite{Li2008} studied the case of two microspheres trapped in a single 
+OT and found that multiple trapped beads could be mistaken for a single 
+trapped bead with altered trap stiffness. Theoretical studies on the case 
+of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a 
+ray-optics based model to show that the two trapped beads are brought into 
+physical contact with each other by optical forces and they also calculated 
+the axial equilibrium positions of the two trapped beads as a function of 
+their size. Experiments in \cite{Praveen2016} confirmed that the two trapped 
+beads indeed experience different trap stiffnesses in the vicinity of the 
+same potential well. There are further discussions looking into the dynamics 
+of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, 
+Chetana2022}, their results all showing that predicting the behaviour an 
+arbitrary shaped particle comes with great difficulty due to the fact that 
+the optical force is dependent on a greater number of variables such as 
+orientation and size factors.
 
-In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
+With the initial goal of the PhD being to induce nucleation events via a 
+spherical micro-rotor the goal of this chapter was to - in a limited capacity
+simulate and investigate the influence of a second particle being bound to our 
+target sphere. The choice of a dimer, instead of an amorphous solid that 
+might better represent a growing crystalline solid, allows us to consider 
+how the dynamics of the aggregate change by varying the size factor. We build 
+upon the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
+asymmetric dimers and how varying size parameters alters the dynamics and 
+additionally makes characterising their interactions within an optical trap 
+more cumbersome. Attempting to simulate an amorphous aggregate is rather
+difficult as calculating the optical force and torque is computationally 
+slow and orientation specific.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Positional and Orientational dependence of Trapping forces}
+\label{sec:eq_positions}
+If we wanted to start from first principles and determine the trap strength 
+on our target particle the first step would be to locate the equilibrium position
+relative to the trap focus. For a single sphere it is easy to enough to understand 
+that its centre of mass will be drawn to focal point of the laser due to gradient
+forces, once there the force is analogous to a harmonic spring with a fixed trap
+stiffness. Now, if we consider instead a dimer, we now have two spheres both being
+drawn to the focus along by the same gradient force; in addition the scattering force
+is significantly more complex due to both spheres scattering the electromagnetic 
+fields. This mutual scattering between individual spheres is what makes simulating
+spherical aggregates far more difficult compared to a single sphere, and even harder
+still to predict the position where the dimer's centre of mass is in equilibrium.
 
-If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus.
-The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser.
-We can do so because the optical force is only dependent on the particle's relative position.
-If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.
+Because the scattering force is only significant in the direction of beam 
+propagation the potential well in the transverse plane can still be assumed 
+to be harmonic around the central beam axis. The axial optical force cannot 
+be assumed to behave as a simple harmonic trap. The methodology for computing 
+optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the 
+trapping force and determine where a simple sphere would be located relative 
+to focal point of the laser by finding the position that minimises the net 
+optical force in a negative feed back loop ($\delta F/\delta x < 0$) - we can 
+assume that for a dielectric sphere the optical torque is negligible. For a 
+dimer (or any arbitrary spherical aggregate), we now must consider both its 
+position and orientation and find where the net optical force and torque are 
+minimised.
 
+\paragraph{Simulation Parameters}
+As a paradigmatic example, consider a dimer suspended in water ($n_p = 1.59, 
+n_m = 1.33$) located at the focus of a Gaussian beam (more specifically a Laguerre-Gaussian beam of mode $[0,0]$), the beam is focused by a objective 
+with numerical aperture of 1.2 and is x polarised. The size ratio of the two 
+sphere's is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ 
+unless specified otherwise; the dimer's orientation is given by a unit vector connecting the centres of both spheres, we define the 'standard' orientation 
+as being aligned with the direction of beam propagation direction - and 
+therefore the 'inverted' orientation is defined when the dimer is orientated 
+against the direction of beam orientation (see figures~\ref{fig:paradigmatic}(b) 
+and (d)).
+
+After computing its $T$-matrix via \textit{mstm} and supplying that 
+to \textit{ott} we compute the optical force via \eqref{eq:optical_force} 
+in the axial direction while the dimer is in its 'standard' orientation. 
+As expected we see a single point where the dimer will be in equilibrium, 
+the linear fit in fig.~\ref{fig:paradigmatic}(a) indicates a that the force 
+can be modelled as a harmonic potential close to the equilibrium position ($F_z\approx-\kappa_z z$). The second point where the axial 
+force goes to 0 cannot be considered as equilibrium position as the positive 
+gradient indicates that the trap is unstable unless Brownian motion is ignored.
+
+We repeated the same calculation but now while the dimer is in its 'inverted' orientation, instead of a single point where the optical force is minimised we 
+see that there are instead two separate equilibrium positions, one above the focus
+and one below the focus. In this particular example the two positions are far enough
+apart that both can be considered as separate harmonic traps.    
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{.475\linewidth}
+	\begin{subfigure}{.75\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
@@
 	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.475\linewidth}
+	\begin{subfigure}{.25\linewidth}
+		\centering
+		\raisebox{50pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
+		\caption{}
+		\label{large over small}
+	\end{subfigure}
+	\medskip
+	\begin{subfigure}{.75\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
@@
 	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{.475\linewidth}
+	\begin{subfigure}{.25\linewidth}
 		\centering
-		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\begin{subfigure}{.475\linewidth}
-		\centering
-		\includegraphics[width=0.3\linewidth,
-		keepaspectratio]{theta=180.png}
+		\raisebox{30pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
+		\caption{}
 		\label{small_over_large}
 	\end{subfigure}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2.
-(a) is the case where the smaller sphere is orientated with the beam propagation direction.
-(b) is the inverted case, smaller sphere oriented against the propagation direction.
-Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
+	\caption{Plots of force vs displacement of the centre of mass of 
+		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
+		the case where the dimer is in its' 'standard' orientation. 
+		(c) is the case where the dimer is in its' 'inverted' orientation 
+		(b) and (d) are renders to visualise the dimer orientation are shown 
+		below each plot. The black lines on each force-curve is a linear 
+		fit with the slope being reported as the trap stiffness in the legend.}
+		\label{fig:paradigmatic}
 \end{figure}
 
-We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable.
-As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased.
-This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap.
-This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements.
+We can see that both equilibrium positions have comparable axial trap stiffness 
+($\kappa_z$), however the difference in the transverse trap stiffness ($\kappa_x$) 
+is far more noticeable. As shown in Fig~\ref{fig:transverse_force}, 
+the dimer's orientation and relative position significantly changes the force 
+curve; not only is the trap wider when inverted but the trap stiffness is 
+increased. This highlights one of the challenges involved with studying 
+asymmetric particles, even though its a simple enough process to trap them 
+they maybe characterised very differently depending on their relative position 
+and orientation towards the focus. This can have a significant impact on 
+rheological studies - or attempting to probe any local property - as the 
+variance in trap strength can result in large errors over repeated measurements. 
 \begin{figure}[h!]
 	\centering
+	\includegraphics[width=\linewidth]{transverse_force.png}
+	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
+		where a positive force indicates the dimer is directed right on the x-axis, 
+		and vice versa for a negative force.
+The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c).
+The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
 	\label{fig:transverse_force}
-	\includegraphics[width=\linewidth]{transverse_force.png}
-	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane.
-With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
 \end{figure}
 
-For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap.
-As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength.
-However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus.
-Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}.
-Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}.
-This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.
+\newpage
+For completeness the harmonic traps were located for dimers across a range 
+of size ratios - from $a_{I}/a_{II} = 1$ to $a_{I}/a_{II}=10$ - while also 
+recording the trap stiffness for each trap. The same simulation parameters are 
+used here as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
+As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease the dimer begins to
+approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
+to $a_{I}/a_{II}=4$), a second equilibrium position is found below the trapping 
+focus. Previous work using the ray-optics model have confirmed even in the case 
+that two spheres begin separated the electric field will align the particles as 
+such that they make contact and are trapped together about a single trapping 
+position \cite{Xu2005}. Furthermore it has been shown through proper manipulation 
+of the Gaussian or Bessel beam modes that any number of trapping potentials can 
+be developed \cite{Shahabadi2020} for nanoparticles. This result however, is the 
+first example of an orientation dependent trapping situation using only a 
+$TEM_00$ beam. Typical experimental arrangements cannot determine much information 
+on the axial position of a trapped particle relative to the trap focus; this 
+result indicates not only that dimers can be trapped in multiple axial positions 
+but also their trapping behaviour is heavily dependent on said axial position. 
+As such it is necessary that positional information in the z-axis can be elucidated
+if multiple spheres are trapped simultaneously.
 
@@
 	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions.
-(a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position.
-(b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
+	\caption{Equilibrium positions of optically trapped dimers with varying 
+		size ratio, dashed lines represent unstable traps whereas solid lines 
+		are for stable equilibrium positions. (a) shows that dimers while in their
+		'standard' orientation will always have a single equilibrium position. 
+		(b) shows that when the same dimer is in its' 'inverted' orientation 
+		can be trapped in two axial positions, one below the focus and one above 
+		the focus.}
+	\label{fig:eq_positions}
 \end{figure}
 
-\subsection{Non-trivial harmonic traps}
+\newpage
+\subsection{Non-trivial equilibrium configurations}
 \label{sec:off-axis}
-Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap.
-However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}.
-Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations.
-An example for a dimer of size ratio 2 is shown below:
+Computing the equilibrium positions when a dimer is aligned with the electric 
+field is relatively simple as the orientational torque is minimised (see
+Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change
+orientation enough to escape the trap. However, that does not rule out the 
+possibility that there is a stable configuration where the orientation not 
+strictly vertical, in fact most experimental work with symmetric nano-dimers 
+will trap them lying perpendicular to the beam direction \cite{Ahn2018, 
+Reimann2018}. Unlike in Sec.~\ref{sec:eq_positions} we cannot simply measure 
+the optical force and torque as the parameter space is too large and determining 
+if a particular position and orientation is stable is not clear based solely 
+on force and torque measurements \cite{Bui2017}. Using the same simulation 
+parameters as before we ran a number short simulations (total simulation time 
+was $0.005$ s) with the laser power increased to $500\ mW$. Each simulation 
+started with the dimer in a different starting position and orientation, due 
+to the high laser power the dimers either escaped the trap or were stably 
+trapped. The $z-\theta$ phase space - where $\theta$ is the angle between the 
+direction of beam propagation and the dimer's orientation vector ($\theta=0^
+\circ$ is the 'standard' orientation) - can be divided into different regions 
+depending on which equilibrium configuration is reached.
+
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\captionsetup{hangindent=2.35cm}
-	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW.
-The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.
-Right hand render shows a dimer in its off-axis configuration}
+	%\captionsetup{hangindent=2.35cm}
+	\caption{Map of $z-\theta$ phase space using a dimer of size ratio 2 with 
+		a laser power of 500 mW ($\theta=0\circ$ is the 'standard' orientation 
+		and $\theta=180^\circ$ is the 'inverted' orientation). The stable 
+		configurations are indicated by the larger circles and the starting 
+		conditions are colour coded to match the stable point they end up in. 
+		Right hand render shows a dimer in its off-axis configuration.}
+	\label{fig:off_axis}
 \end{figure}
 
-Interestingly while the trap strength of these off-axis traps are 
-similar in magnitude to the vertically aligned dimers, but when 
-the laser power is lowered (around 5 mW) the traps become metastable 
-resulting in the dimer escaping from after some random time while 
-within the trap. If the overall potential depth could be characterised 
-then dimers placed into this orientation could be used as a 
-micro-scale temperature alarm, where by fine tuning of the dimer's 
-parameters would allow you to construct a potential well that can 
-only be escaped when the local fluid temperature exceeds a certain maximum value.
+Interestingly while the trap strength of these off-axis traps are similar 
+in magnitude to the vertically aligned dimers, but when the laser power 
+is lowered (around $5\ mW$) the traps become metastable; after reaching its'
+equilibrium configuration the particle behaves similarly to a typically
+trapped dimer but the trapping potential is small enough that the dimer
+escapes in as little time as less than a second or after nearly a full 
+3 seconds. Running similar simulations but for vertical configurations 
+sees the dimer remaining trapped, even after 30 seconds of run time, 
+indicating that the trapping potential is far greater than the thermal 
+energy. This suggests that the reason this off-axis configuration is due 
+to the rotational motion more than translational motion. If the overall 
+potential depth could be characterised then dimers placed into this 
+orientation could be used as a micro-scale temperature alarm, where 
+by fine tuning of the dimer's parameters would allow you to construct a 
+potential well that can only be escaped when the local fluid temperature 
+exceeds a certain maximum value.
 
@@
 transfer an optical torque that is proportional to the difference in 
-the refractive index of the particle. The optical torque is given as:
+the refractive index of the particle's ordinary and extraordinary axis. 
+The total optical torque is given as:
 \begin{equation}
@@
 	\begin{aligned}
-		\tau_{opt} = \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
+		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
 	\end{aligned}
@@
 Where $\Delta n$ is the difference in refractive indices of the 
-particle, and $\phi$ indicates the phase shift in the electric field
-(for circular polarised light $\phi= \pi/4$). The optical rotation seen
-with dimers can be attributed to the polarisation of the trapping beam as
-taking a Fourier transform of the transverse components of the dimers 
-orientation time series reveals a clean rotation rate. 
+particle, $\phi$ indicates the phase shift in the electric field
+(for circular polarised light $\phi= \pi/4$), and $\theta$ is the angle
+between the particle's main axis and the direction of the electric field.
+The optical torque of non-birefringent spherical aggregates has no such 
+relationship as the mechanism behind their rotation is not as clearly defined.
+
+To visualise the relationship between the polarisation state of the 
+trapping beam, and the rotation rate we simulated the motion of an 
+optically trapped dimer in beams of varying polarisation ($NA=1.2$, 
+$P=100\ mW$), and the dimer is composed of silica ($n_p=1.59$, 
+$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t 
+=10^-5$) and at the end we looked at the orientational time series; 
+the dimer's orientation is recorded as a quaternion which can be easily 
+converted to a 3-dimensional rotation matrix. By considering only the 
+transverse components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) 
+of the rotation matrix and taking the Fourier transformation of their 
+time series reveals the rotational frequency. The laser power is set to 
+$100\ mW$ to avoid large rotational fluctuations and so that the Fourier 
+series of the transverse components approximates $\delta(\omega_{rot}-f)$ 
+- the Dirac delta function centred at the rotational frequency $\omega_{rot}$.
+
+%% Example plot of rotating beam trajectory and then the same thing but 
+%% Fourier transformed
+\begin{equation}
+\begin{split}
+	q(t) \rightarrow R(t) =& 
+	\begin{pmatrix}
+		U_{1,x}(t) && U_{2,x}(t) && U_{3,x}(t) \\
+		U_{1,y}(t) && U_{2,y}(t) && U_{3,y}(t) \\
+		U_{1,z}(t) && U_{2,z}(t) && U_{3,z}(t) 
+	\end{pmatrix} \\
+	\rightarrow
+	&\int^\infty_{-\infty}R(t)e^{-i2\pi ft} dt = 
+	\begin{pmatrix}
+		\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f) \\
+		\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f)\\
+		\delta(f) && \delta(f) && \delta(f)
+	\end{pmatrix}
+\end{split}
+\end{equation}
+
+If the rotational frequency was not immediately obvious the simulation was
+repeated but over a longer simulation time. Four different size ratio of 
+dimers were studied, both in their 'standard' and 'inverted' orientations. 
+The results of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.9\linewidth]{rotation_vs_pol.png}
-	\caption{Rotation frequency vs component phase difference for differently sized dimers.
-The solid lines represent the rotation rate experienced while the dimer is in its standard orientation, whereas the solid points are for the case where the orientation is inverted.
-Laser power $= 100\ mW$}
+	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
+	\caption{Rotation frequency vs component phase difference for differently 
+		sized dimers. The solid lines represent the rotation rate experienced 
+		while the dimer is in its standard orientation, whereas the solid points 
+		are for the case where the orientation is inverted. Laser power $= 100\ mW$.}
+	\label{fig:rotation_vs_pol}
 \end{figure}
@@
 force that results in orbital motion about the beam's central axis.
-This effect is rather minimal for single sphere's, resulting in a 
-orbital frequency on the order of $10^{-1}\,{\rm Hz}$, with an 
-order of magnitude difference when trapping aggregates of spheres. 
-They computed the circulation rate by computing the time-average 
-probability flux; however, when extended to the Mie regime we 
-see a completely different behaviour, instead experiencing an 
-optical torque about their long axis.S This rotation was first 
-noted by Vigilante and co-workers who only considered this
-behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of
-simulations for differently sized dimers in a circularly polarised
-trap and looked at the rotation rate.  We found that not only is the
-rotation rate dependent on the size of the dimer, but also on its
+This was verified by \cite{Ruffner2012} who found that by trapping 
+a silica bead and recording its position over the course of several 
+minutes; while not immediately evident from the trajectory by measuring 
+the probability flux of finding a particle in a given position 
+revealed a helicity to the sphere's motion which matched the helicity 
+of the trapping beam. This would hardly be considered similar to 
+the behaviour shown by dimers; firstly because the probability flux 
+is an indicator of the sphere's position not its orientation - the 
+sphere can be thought to 'orbit' around the central beam axis with a 
+radius of orbit on the order of $10^{-9}\ m$. Secondly, while 
+\cite{Yevick2017} does compute a rotational frequency, for a single 
+sphere this rotation rate is at most $0.01\ Hz$ whereas the lowest 
+rotation rate recorded for a dimer ($a_{I}/a_{II}$) was around $2\ Hz$, 
+a full two orders of magnitude greater. As such the rotation demonstrated
+cannot be explained due to photo-kinetic forces.
+
+This rotation was first noted by Vigilante and co-workers who only 
+considered this behaviour for a symmetric dimer \cite{Vigilante2020}; 
+to expand upon this work we looked at how the rotation rate changed
+with the dimers' size ratio ($a_{I}/a_{II}$) and if the orientation
+of the dimer plays any impact on the rotational frequency. By repeating
+the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
+a circularly polarised beam $\phi=90\circ$ it was found that not only is 
+the rotation rate dependent on the size of the dimer, but also on its 
 orientation and therefore their axial position.
@@
   \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
-  \caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios.
-The red line is for the case where the larger sphere is above the smaller sphere.
-The blue line is the inverted case, while the initial position is above the focus of the trap.
-And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
+  \caption{Rotation rate plotted against dimer size ratio while trapped in a
+  	circularly polarised beam; a positive rotation rate indicates clockwise 
+  	rotation, whereas a negative rotation rate indicates counter-clockwise 
+  	rotation. The red line is for the case of a dimer in its 'standard' 
+  	orientation. The blue line is for the case when the dimer is in its 
+  	'inverted' orientation while trapped above the focus of the beam. And 
+  	lastly the green line is for the case when the dimer is in its 
+  	'inverted' orientation, but when it is trapped below the focus of the 
+  	beam.}
 \end{figure}
@@
 imply that a second sphere of radius $200\,{\rm nm}$ is enough to
-induce rotational motion. We used \textit{mstm} to look at the stokes
+induce rotational motion. This brings into question what mechanism is
+generating the optical torque. We used \textit{mstm} to look at the stokes
 parameters from the scattered field from a simple plane wave incident
-on our dimer, the proportion of circularly polarised light is minimal
-compared to the proportion of plane polarised light, which indicates
-that this rotational motion is not due to any inhomogeneity in the
-dimer that might impart angular momentum to the scattered beam - as
-compared to a anisotropic scatterer like vaterite.
+on a symmetric dimer, the proportion of circularly polarised light is 
+minimal compared to the proportion of plane polarised light, which 
+indicates that this rotational motion is not due to any inhomogeneity 
+in the dimer that might impart angular momentum to the scattered beam - 
+as compared to a anisotropic scatterer like vaterite. Therefore the 
+rotational motion must be due to
 
-These results are somewhat contrary to other work with silica dimers
+These results are somewhat contradictory to other work with silica dimers
 \cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
 have trapped the dimer in an orientation perpendicular to the beam
-propagation direction.  The rotational motion is induced due to the
-asymmetric geometry creating an unbalanced polarisation susceptibility
+propagation direction. The rotational motion is attributed to the
+dimers asymmetric geometry creating an unbalanced polarisation susceptibility
 along its long axis as compared to its short axis; therefore its long
 axis is aligned with the polarisation vector and can rotate
-freely\cite{Ahn2018}.  This however cannot be the case with our
+freely \cite{Ahn2018}. This however cannot be the case with our
 simulations as the dimer rotates about its long axis, meaning there
@@
 the dimer's is rotating about it's long axis (as seen by 
-the periodic behaviour of $u_x$ and $u_y$), while also precessing 
+the periodic behaviour of $U_x$ and $U_y$), while also precessing 
 around the optical axis of the beam, resulting in a second 
@@
 \begin{figure}[h]
-	\label{fig:gyro}
 	\centering
 	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
-	\caption{3 second trajectory of a dimer trapped in an off axis orientation, laser power was set to 100 mW.
-The far left column depicts the dimer's centre of mass position with time; middle two columns are the $x$, $y$, and $z$ components of the vectors $u_x$ and $u_y$; last column depicts the components of the vector $u_z$ which defines the dimer's orientation.}
+	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
+		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
+		The far left column depicts the dimer's centre of mass position with 
+		time; middle two columns are the $x$, $y$, and $z$ components of the 
+		vectors $u_x$ and $u_y$; last column depicts the components of the 
+		vector $u_z$ which defines the dimer's orientation.}
+	\label{fig:gyro}
 \end{figure}
 
-Applying a Fourier analysis to the above trajectory reveals the 3
-fundamental frequencies typically associated with precession; 
-the $u_{z,1}$ and $x(t)$ series show a precession frequency of
-$~1.33\ Hz$ whereas the series $u_{x,1}$ and $u_{y,1}$ show a 
-combined periodic signal - a rotational frequency of $~23\ Hz$
-and a nutation frequency of $~20\ Hz$. Previous studies into 
-amorphous silica nanoparticles found a linear relationship 
-with the rotational frequency and the laser power, but no such
-relationship existed with the precession frequency. Our own 
-results shows a similar linear relationship with vertically
-aligned dimers.
+Applying a Fourier analysis to the above trajectory reveals the 3 fundamental
+frequencies typically associated with precession; the $u_{z,1}$ and $x(t)$ 
+series show a precession frequency of $~1.33\ Hz$ whereas the series $u_{x,1}$ 
+and $u_{y,1}$ show a combined periodic signal - a rotational frequency of 
+$~23\ Hz$ and a nutation frequency of $~20\ Hz$. Previous studies into 
+amorphous silica nanoparticles found a linear relationship with the 
+rotational frequency and the laser power, but no such relationship existed 
+with the precession frequency. Our own results shows a similar linear 
+relationship with vertically aligned dimers.
 \begin{figure}
@@
 The linear relationship could partly be due to fact that we do not
-account for the change in viscous forces with increasing
-laser power. It is far more likely that the rotational frequency
-reaches a maximum value assuming that the bulk fluid can readily 
-absorb the laser.
+account for the change in viscous forces with increasing laser power. 
+Due to the localised heating effect 
+It is far more likely that the rotational frequency reaches a maximum 
+value assuming that the bulk fluid can readily absorb the laser.
 
-This gyroscopic motion has been demonstrated previously 
-in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
-but has not been observed for micron scale aggregates. 
-Since the torque is computed by evaluating the beam coefficients 
-it is be difficult to apply this result to micro-rheology 
-experiments as one would need to know the exact magnitude 
-of the optical torque ahead of time in order to make 
-estimations about the local fluid viscosity. This is 
-trivial for a birefringent spherical particle, less 
-so for spherical aggregates whose equilibrium position and 
-orientation are unknown.
+This gyroscopic motion has been demonstrated previously in nanoparticles 
+\cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed 
+for micron scale aggregates. Since the torque applied to the dimer is 
+computed by evaluating the beam coefficients of the scattered field it is 
+difficult to apply this result to micro-rheology experiments as one would 
+need to know the exact magnitude of the optical torque ahead of time in 
+order to make estimations about the local fluid viscosity. This is trivial 
+for a birefringent spherical particle, less so for spherical aggregates 
+whose equilibrium position and orientation are unknown. However, further 
+analysis of the mechanism behind the precessive motion of off-axis dimers
+may provide insights into controlling Brownian motion. An experimental work
+trying to 'cool' nano-dimers by controlling the motion in all 6 degrees of 
+freedom found that even while the rotation about the short axis' could be 
+controlled the free rotation around the dimers' long axis resulted in an 
+unpredictable torsional vibration \cite{Bang2020}. Understanding how 
+rotational motion arises in the Mie-Regieme could allow researchers to 
+build a robust theoretical framework to construct beam structures that 
+eliminate any unwanted rotational motion from a target particle. Conversely, 
+the same framework could allow for precise measurements of the optical 
+torque applied to a target particle, allowing for characterisation of 
+complex shaped particles' interactions with an optical trap.
 
 \section{Characterisation of asymmetric dimers via PSD analysis}
-As discussed in \ref{sec:simulated_QPD}, one of the methods
-developed to work in conjunction with \cite{Vigilante2020} is 
-a simulated quadrant photo diode for as a position detection
-system. While it is possible to extract all of the relevant 
-dynamical information from a simulation, confirming the same 
-behaviour in an experimental setting can be challenging.
+As discussed in \ref{sec:simulated_QPD}, one of the methods developed to 
+work in conjunction with \cite{Vigilante2020} is a simulated quadrant photo 
+diode for as a position detection system. While it is possible to extract 
+all of the relevant dynamical information from a simulation, confirming 
+the same behaviour in an experimental setting can be challenging if dealing
+with a non-birefringent particle.
 
-As a benchmark we start by considering a single sphere within 
-an optical trap. After a 3 second trajectory (the typical time
-necessary to collect a power spectrum) the sphere's power 
-spectra was recorded and a fitted to eq.~\ref{eq:alaised_lorentzian}.
+As a benchmark we start by considering a single sphere within an optical trap. 
+A single silica sphere suspended in water ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$)
+is trapped by a focus Gaussian beam and its trajectory was recorded. A 3 second 
+is trajectory is a typical measurement time for collecting a power spectrum, 
+the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
 \begin{figure}[h]
-	\label{fig:psd_sphere}
 	\centering
 	\includegraphics[width=\linewidth]{PSD_sphere.png}
-	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian}, scattered points represents the blocked data ($n_b=100$). Corner frequency for the 
-	Lorentzian curves are reported in the legend.}
+	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
+		 scattered points represents the blocked data ($n_b=100$). Corner 
+		 frequency for the Lorentzian curves are reported in the legend.}
+	\label{fig:psd_sphere}
 \end{figure}
 
-As shown in fig.~\ref{fig:psd_sphere}, the two power spectra
-report different corner frequencies which would be indicate 
-that the trap is not perfectly circular. Using \textit{ott} 
-we can compare the expected trap strength to what is reported
-by a quadrant photo diode:
+As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
+different corner frequencies which would be indicate that the trap 
+is not perfectly circular. Using \textit{ott} we can compare the 
+expected trap strength to what is reported by a quadrant photo diode:
 
@@
 
-Where the ellipticity of the beam is given by 
-$e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
-the beam wavefront. Its clear from these initial results that the 
-QPD is more sensitive to changes along the y-axis than the x-axis 
+Where the ellipticity of the beam is given by $e = (1-\kappa_y/
+\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
+wavefront. Its clear from these initial results that the QPD 
+is more sensitive to changes along the y-axis than the x-axis 
 when compared to the direct \textit{ott} calculations. Typically, 
-even an industrial Gaussian beam will produce an elliptical diffraction
-limit spot when heavily focused; in their tutorial for optimizing 
-the PSD analysis, Berg and Sorensen reported a ellipticity of
-around 15 \% after a total calibration time of 81 seconds 
-\cite{BergSoerensen2004}. The reason for this discrepancy
-can be explained partly by the fact that our estimation of the 
-trap geometry via \textit{ott} is based on the force-displacement
-curve while the sphere is moving in only one direction, whereas the
-QPD is estimating the trap geometry by extrapolating from far field 
-scattering signals. We now consider a symmetric dimer in identical 
-simulative conditions.
+even an industrial Gaussian beam will produce an elliptical 
+diffraction limit spot when heavily focused; in their tutorial 
+for optimizing the PSD analysis, Berg and Sorensen reported a 
+ellipticity of around 15 \% after a total calibration time of 
+81 seconds \cite{BergSoerensen2004}. The reason for this 
+discrepancy can be explained partly by the fact that our 
+estimation of the trap geometry via \textit{ott} is based on 
+the force-displacement curve while the sphere is moving in only 
+one direction, whereas the QPD is estimating the trap geometry 
+by extrapolating from far field scattering signals. We now 
+consider a symmetric dimer in identical simulative conditions.
 

### 2024-08-02 Latex_files/Thesis/chapters/chapter 2/main.tex
tidying up chapter 3

@@
 
+\subsubsection{Power spectral analysis for rotating objects}
+The scattering from rotating objects can be partially characterised via
+the power spectrum density method. Typically papers reporting on the 
+rotation rate of a trapped object will collect a power spectrum and look
+at the maximum frequency term to determine its rotation rate. However this
+often neglects any information on the trapping forces acting on particle
+and only looks at the torque applying around one of the particle's primary 
+axis. There has only been one notable effort to fully characterise the 
+full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
+a theoretical model for that relates the power spectrum to both the 
+rotational and translational motion simultaneously.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsection{Simulation parameters}
 Throughout this PhD we use the work of Vigilante to perform a systematic 
 study of the dynamics demonstrated by asymmetric dimers in both plane and 
-circularly polarised light. But furthermore we expand upon their work to 
-create a fully functional quadrant photo-diode to simulate the response 
-from a calibration test. This builds upon the work from \cite{Rohrbach2002} 
-which applied the fundamental principles of Lorenz-Mie theory to replicate 
-the response signal of a QPD being used in back focal-plane interferometry.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simulated Quadrant Photodiode}
+circularly polarised light. The simulations have two primary frames of 
+reference that require explanation. Firstly, there is the laboratory frame 
+which describes the coordinate system, we define the origin of our simulation
+as the focus of our trapping beam. While the choice in direction for the $x$ 
+and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
+aligned with the direction of beam propagation. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.6\linewidth]{lab_frame.png}
+	\caption{Ray scattering render of both the laboratory frame and particle 
+		frame for a dimer and a Gaussian beam. The origin of the laboratory 
+		frame is defined as the focus of the trapping beam, with the positive 
+		$z$-axis being aligned with the direction of beam propagation. The 
+		particle frame is centered on the dimer's centre of mass with the 
+		primary axis $U_z$ being aligned with the centres of the two spheres.
+		Black line is scaled to represent $1\ \mu m$.}
+	\label{fig:lab_frame} 
+\end{figure}
+
+Secondly there is the particle frame which describes the orientation state of 
+the particle, orientations are reported as quaternions which can be readily 
+transformed into rotation matrices. The matrix describes the direction of the 
+particles 3 orientation axis $U_x$, $U_y$, and $U_z$; the latter is oriented 
+so that it is collinear with the centres of the two spheres. The origin of 
+the 3 axis' is set on the particle's centre of mass, this is also used to 
+define it's displacement from the origin of the laboratory frame.
+
+Furthermore we expand upon their work to create a fully functional quadrant 
+photo-diode to simulate the response from a calibration test. This builds 
+upon the work from \cite{Rohrbach2002} which applied the fundamental 
+principles of Lorenz-Mie theory to replicate the response signal of a QPD 
+being used in back focal-plane interferometry.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulated Quadrant Photodiode}
 \label{sec:simulated_QPD}

### 2024-08-02 Latex_files/Thesis/chapters/chapter 3/main.tex
tidying up chapter 3

@@
 \chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the PhD was to investigate 
-the possibility of using optical tweezing as a means of initiating and 
-controlling nucleation by generating fluid flow within a small droplet of 
-supersaturated solution. The goal of which would be to understand the 
-influence of shearing on nucleation at a micro level as compared to larger 
-scale results. It has been shown that for macro-scale systems, the likelihood 
-of nucleation increases to a maximum value under increased shearing 
-\cite{Debuysschere2023, Mura2016}. Theoretical research into the matter 
-identified two competing effects that effect a crystal in moving fluid fields; 
-firstly, nucleation is enhanced due to the increased mass transfer of solute 
-material; and secondly, shear flow against the crystal surface leading to a 
-decrease in growth \cite{Mura2016}. These two competing effects are validated 
-by experimental work using glycine solutions, showing that beyond a certain 
-shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this 
-chapter I outline the optical tweezer equipment used during this PhD, the 
-initial attempts made to induce nucleation via optical rotation, and the 
-impact of a moving beam on the crystal growth of a newly formed nucleus.
+As outlined in Chapter 1, the original outline of the project is to 
+investigate the possibility of using optical tweezing to induce nucleation 
+by generating fluid flow within a supersaturated solution. The intent 
+of which would be twofold: Firstly to have a repeatable means of 
+inducing nucleation under different solution conditions. And secondly, 
+to understand the influence of shearing on nucleation at a micro 
+level as compared to results in bulk fluid. It has been shown that 
+for macro-scale systems, the likelihood of nucleation increases to 
+a maximum value under increased shearing \cite{Debuysschere2023, 
+Mura2016}. Mura and Zaccone developed a theoretical framework to 
+describe how the a newly formed nucleus experiences two additional 
+growth factors when placed in a moving fluid. Firstly, due to increased 
+molecular transport of solute molecules the nucleation rate is enhanced 
+in low to moderate fluid flows. But in addition, due to shear flow the 
+crystal surface undergoes deformation which suppresses the nucleation 
+rate undergoing faster fluid flow \cite{Mura2016}. The theory was 
+confirmed for glycine solution by Debuysschere \textit{et al} who
+identified that the nucleation rate of supersaturated glycine was 
+enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ \cite{Debuysschere2023}.
+After which the nucleation rate began to decrease but was 
+still greater compared to the case where fluid flow was minimal.
+
+Optical tweezers can been used to rotate a whole host of micro-rotors, 
+with the fastest reported results exceeding $1000\ Hz$ in heavy water
+\cite{Arita2016}. Therefore, it stands to reason that a micro-rotor
+rotating in a supersaturated solution could generate sufficient fluid
+flow such that the nucleation rate around said particle is enhanced 
+compared to the bulk fluid. We focused on two primary candidates for
+micro-rotors, Vaterite and 4-Heptyl-4-biphenylcarbonitrile (7CB). The
+former being a polymorph of calcium carbonate and the latter an example
+of nematic liquid crystals, both of which have been used repeatedly in
+previous micro-rotor. In addition, we also consider the application of
+using techniques beam steering to generate fluid flow by trapping silica
+micro-beads. In this instance the fluid flow is generated not due to the
+transfer of angular momentum, but due to shearing caused by a moving 
+sphere through stagnant fluid.
+
+To begin with, the discussion of the necessary optical equipment is 
+covered, drawing attention to specialised components and techniques 
+that are not standard in optical trapping set ups.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Optical Tweezer Equipment}
-In general, all optical tweezers require a laser driver, two
-microscope objectives (one for trapping and one for imaging),
-and a means of controlling the position of the loaded sample. 
-While there are other pieces of equipment used in modern tweezer
-experiments these are the bare minimum requirements for any 
-optical tweezer. The laser used for this project was a 1064 
-nm near infrared laser - provided by CNI Lasers – that has 
-an adjustable power supply to vary the energy output of the 
-laser - the remaining optics where supplied by Thor Labs.
-Experimental work has shown that the trapping efficiency 
-increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
-\cite{kim2003dependence} where $D_{obj}$ is the diameter
-of the objective aperture. To expand the beam front we 
-utilise a Galilean beam expansion arrangement (indicated 
-by $f_1$, and $f_2$) as recommended for high power laser 
-applications. In our initial experiments the beam expansion 
-provides a $4\times$ magnification; however, in later experiments 
-where we utilise a galvano-mirror the beam expansion is $3\times$ 
-and then the 4f correlator provides a further $1.25\times$ magnification 
-(using $f_3$ and $f_4$) - where the magnification is given by.
+In general, all optical tweezers require a laser driver, a focusing 
+microscope objectives, a position controller, and position detector.
+The laser used for this project was a $1064\ nm$ near infrared laser 
+- provided by CNI Lasers – that was focused by a Nikon 100x oil 
+immersion lens. The choice of an oil immersion lens is important as 
+the optical oil used prevents a loss of focus when used on a glass 
+cover slip. Now, experimental work has shown that the trapping 
+efficiency increases with beam diameter up until it exceeds 
+$\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where $D_{obj}$ is 
+the diameter of the objective aperture. To expand the beam front we 
+utilise a Galilean beam expansion arrangement (indicated by $f_1$, 
+and $f_2$ in Fig.~\ref{fig:setup}) as recommended for high power laser 
+applications. In our initial experiments the beam expansion provides a 
+$4\times$ magnification. Whereas in later experiments we utilised a 
+galvano-mirror the beam expansion is $3\times$ and then the 4f 
+correlator provides a further $1.25\times$ magnification (using $f_3$ 
+and $f_4$) - the magnification is given by.
 \begin{align}
@@
 
-It should be noted that the galvano-mirror requires the use 
-of a Keplerian beam expansion arrangement which reduces
-the transmitted laser power due to localised heating of the air.
-Afterwards the laser is passed through a dichroic mirror that 
-separates incoming infrared and visible light, this is to 
-prevent the laser from damaging the CCD camera used for imaging
-the trapping plane. The laser is then focused to a diffraction
-limited spot by a 1.25 NA objective. By increasing the numerical
-aperture of the objective, the gradient force at the focal point 
-is increased; the trade-off being that the for higher NA values 
-the trapping depth is reduced due to spherical aberrations.
+It should be noted that the galvano-mirror requires the use of a 
+Keplerian beam expansion arrangement which reduces the transmitted 
+laser power due to localised heating of the air. Afterwards the 
+laser is passed through a dichroic mirror that separates incoming 
+infrared and visible light, this is to prevent the laser from 
+damaging the CCD camera used for imaging the trapping plane. The 
+laser is then focused to a diffraction limited spot by the objective. 
+Utilizing a high numerical aperture objective enhances the gradient 
+force at the focal point; the trade-off being that the for higher NA 
+objectives the trapping depth is reduced due to spherical aberrations. 
 While it is possible to increase the trapping depth \cite{Reihani2006} 
@@
 with our trapping arrangement. A 0.25 NA condenser objective refocuses 
-the scattered laser light and also provide an aperture for an 
-imaging LED to illuminate the focal plane. Samples are loaded 
-onto a piezo driven table to that is inserted between the trapping
-and condensing objectives; the piezo drivers allow for sub-micron
-control of the sample position to a degree as small as a 10 nm. 
-To detect and monitor the position of a trapped particle a 
-quadrant photo diode (QPD) was utilised. 
+the scattered laser light and also provide an aperture for an imaging 
+LED to illuminate the focal plane. Samples are loaded onto a piezo 
+driven table to that is inserted between the trapping and condensing 
+objectives; the piezo drivers allow for sub-micron control of the 
+beam focus position to a degree as small as a 10 nm. To detect and 
+monitor the position of a trapped particle a quadrant photo diode 
+(QPD) was utilised. 
 \begin{figure}[h!]
@@
 \subsection{Position detection methods}
-In order to accurately capture the dynamics of a trapped particle, 
-a position detection system is utilised. There are 3 possible
-methods of position detection: video-analysis, lateral-effect
-position sensing, and photodiodes. The former being 
-ideally suited for multiple traps or situations where precision
-is not the top priority. In order to match the force measurements
-of back-focal plane interferometry requires the camera's frame
-rate to exceed $1\ kHz$ which can be difficult to achieve while 
-maintaining a decent resolution \cite{Gibson2008}. In comparison
-off the shelf back-focal plane detectors can achieve temporal
-resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
+In order to accurately capture the dynamics of a trapped particle, a 
+position detection system is required. There are 3 possible methods 
+of position detection: video-analysis, lateral-effect position sensing, 
+and photodiodes. The former being ideally suited for multiple traps 
+or situations where precision is not the top priority. In order to 
+match the force measurements of back-focal plane interferometry requires 
+the camera's frame rate to exceed $1\ kHz$ which can be difficult 
+to achieve while maintaining a decent resolution \cite{Gibson2008}. 
+In comparison off the shelf back-focal plane detectors can achieve 
+temporal resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
 
-A QPD is frequently used position detection system for optical 
-tweezers due to their high sampling rate, high degree of 
-precision, and ease of set up. The QPD is constructed of four photo 
-diodes assembled in a quadrant formation, when a particle is trapped the 
-interference pattern produced is focused onto the QPD, with 
-the maximum intensity mapping to the particle's centre of mass. 
-By summing the voltages of the horizontal and vertical quadrants 
-together the particle's centre of mass is tracked in the x-y 
-plane. Axial displacement can be estimated by observing the change
-in the total voltage of the QPD. The outputted signal gives an
-indication of the particle's relative displacement from the beam 
-focus, but in order to convert the signal to distance units the
-trap needs to be calibrated (assuming a linear response curve).
+A quadrant photo diode (QPD) is a frequently used position detection 
+system for optical tweezers due to their high sampling rate, high 
+degree of precision, and ease of set up. The QPD is constructed of 
+four photo diodes assembled in a quadrant formation, when a particle 
+is trapped the interference pattern produced is focused onto the QPD, 
+with the maximum intensity mapping to the particle's centre of mass. 
+By summing the voltages of the horizontal and vertical quadrants together 
+the particle's centre of mass is tracked in the x-y plane. Axial 
+displacement can be estimated by observing the change in the total 
+voltage of the QPD. The outputted signal gives an indication of the 
+particle's relative displacement from the beam focus, but in order to 
+convert the signal to distance units the trap needs to be calibrated 
+(assuming a linear response curve).
 \begin{figure}[h!]
@@
 
-A lateral-effect sensor has a similar output but works using a 
-the entire sensor as a single cell analogous to the focal plane of
-the trapping beam. The four corners of the sensor act as anodes 
-connected to a base plate cathode, as the beam moves across the
-surface of the detector each anode will experience a different 
-photocurrent depending on how close the centre of the 
-interference pattern is to each anode. The advantage of a 
-lateral effect detector is that the linear regime is much larger 
-than a QPD making it much better for monitoring the position of a
-trapped particle. However, Lateral-effect sensors are often limited
-in their spacial resolution due to high signal-to-noise ratios, 
-requiring a high intensity of light on the sensor in order to 
-get a clean signal. As a result, most optical trapping experiments
-are conducted using a QPD as opposed to a lateral-effect sensor, 
-as often the displacement is small enough that the QPD response 
-curve can be considered linear.
+A lateral-effect sensor has a similar output but works using a the 
+entire sensor as a single cell analogous to the focal plane of the 
+trapping beam. The four corners of the sensor act as anodes connected 
+to a base plate cathode, as the beam moves across the surface of the 
+detector each anode will experience a different photocurrent depending 
+on how close the centre of the interference pattern is to each anode. 
+The advantage of a lateral effect detector is that the linear regime 
+is much larger than a QPD making it much better for monitoring the 
+position of a trapped particle. However, Lateral-effect sensors are 
+often limited in their spacial resolution due to high signal-to-noise 
+ratios, requiring a high intensity of light on the sensor in order to 
+get a clean signal. As a result, most optical force measurements are 
+conducted using a QPD as opposed to a lateral-effect sensor, as often 
+the displacement is small enough that the signal-displacement curve 
+can be considered linear.
 
@@
 filter out unwanted scattering from an image (or in an inverse case 
-differentiate between different images), the placement of each lens is shown below.
+differentiate between different images), the placement of each lens is 
+shown below.
 \begin{figure}[h!]
@@
 \label{sec:vaterite}
-Generation of fluid shear can be achieved via two avenues: Firstly, by 
-utilising circularly polarised light it is possible to transfer angular 
-momentum from the laser to the trapped entity. Secondly, one can 
-directly move the trap within the imaging plane by steering the beam 
-using either a galvanometric mirror or gimbal mirror. The following 
-chapter outlines the work done with shear generated by circularly 
-polarised light and the challenges of applying this to localising 
-nucleation.
+There are several options for particles that can be rotated using optical 
+tweezers \cite{Parkin2009, Saito2022}. Over the course of the project two 
+different micro spheres where investigated, vaterite and liquid crystal 
+droplets. Both can be readily synthesised in the lab and are will rotate 
+at a variety of sizes.
 
-There are several options for particles that can be rotated using 
-optical tweezers \cite{Parkin2009, Saito2022}. Over the course of 
-the PhD two different micro spheres where investigated, vaterite 
-and liquid crystal droplets. Both can be readily synthesised in 
-the lab and are will rotate at a variety of sizes. While silver 
-nano particles were considered their high cost and small size 
-meant they were disregarded as an option for optical rotation.
-
-Vaterite is a polymorph of calcium carbonate that is rarely seen in 
-nature due to its low stability. However unlike its other polymorphs of 
-calcite and aragonite, when synthesised vaterite will typically form 
-small spherical particles making them ideal for optical trapping and rotation. 
-Synthesis of vaterite micro spheres requires fine control of the nucleation
-process in order to maintain polymorphic stability, though for the purposes
-of optical rotation the exact polymorph is not as important as its morphology
-as all 3 polymorphs are inherently birefringent.
+Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
+due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
+other polymorphs of calcite and aragonite, when synthesised vaterite will 
+typically form small spherical particles making them ideal for optical 
+trapping and rotation. Synthesis of vaterite micro spheres requires fine 
+control of the nucleation process in order to maintain polymorphic stability, 
+though for the purposes of optical rotation the exact polymorph is not as 
+important as its morphology as all 3 polymorphs are inherently birefringent.
 
@@
 to halt the reaction, the solution was filtered and washed 3 times with 
-distilled water before being suspended in water. When trapped in circularly 
-polarised light, the anisotropic scattering of the sphere results in a periodic 
-signal on the QPD. Therefore, the resulting power spectrum is not a Lorentzian
-but now also displays peaks that appear at integer multiples of the particles
-rotational frequency. 
+distilled water before being suspended in water.
+
+When trapped in circularly polarised light, the anisotropic crystal lattice
+allows spin angular momentum to be transferred to the vaterite particle, 
+resulting in a rotation about the beam axis. In addition, the anisotropic 
+scattering causes the QPD signal to vary with a constant periodicity that
+is attributed to its rotation rate. Therefore, the resulting power spectrum 
+is not a Lorentzian but now also displays peaks that appear at integer 
+multiples of the particles rotational frequency. 
 \begin{figure}[h!]
@@
 to this was conducted by \cite{Yogesha2012} where they describe the 
-rotational motion of ellipsoidal polystyrene particles. The critical assumption
-being that the particle perfectly rotates in the $x-y$ plane; however it
-has long been suspected that birefringent microspheres experience random
-rotations outside of the $x-y$ plane \cite{Bang2020} making it very difficult 
-to characterise the forces on rotating microspheres without a proper understanding 
-of the optical torque being applied to it.
+rotational motion of ellipsoidal polystyrene particles. The critical 
+assumption being that the particle perfectly rotates in the $x-y$ plane; 
+however it has long been suspected that birefringent microspheres 
+experience torques outside of the $x-y$ plane \cite{Volpe2023} making 
+it very difficult to characterise the behaviour of rotating birefringent 
+microspheres without a proper understanding of the full optical torque 
+being applied to it.
 
@@
 Liquid crystals are an intriguing example of materials with mixed phase
-properties, unlike typical solutes such as Glycine, a liquid crystal
+properties. Unlike typical solutes such as Glycine, a liquid crystal
 can still maintain some degree of order between its individual molecules
 while in the liquid state. This is due to the fact that liquid crystals
-are constructed of long ordered molecules that demonstrate a long range
+are constructed of ordered molecules that demonstrate a long range
 ordering. There are three main types of liquid crystal transition methods:
-thermotropic crystals will transition to their liquid crystal phase when 
-sufficiently heated; Lyotropic materials can undergo this transition due 
-to changes in temperature and concentration; and lastly Metallotropic 
+Thermotropic crystals will transition to their liquid crystal phase when 
+sufficiently heated. Lyotropic materials can undergo this transition due 
+to changes in temperature and concentration. And lastly, Metallotropic 
 materials - which are composed of both organic and inorganic molecules - 
 change phase according to the ratio of organic to inorganic molecules present.
-Liquid crystal rotors are rather simple in their production, 
-4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
+Liquid crystal rotors are rather simple in their production, 4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
 and a small amount was added to a vial of distilled water. The solution 
 was then heated in a water bath to $25^\circ$ in order to transition the
-solid crystal into its liquid crystal state. The individual droplets 
-are inherently birefringent and rotate in a circularly polarised trap. 
+solid crystal into its liquid crystal state. The solution can then be 
+loaded onto a sample cover slip and the individual droplets visualised.
+The molecules of 7CB will align with a strong electric field, and due to
+the spherical droplet geometry the droplets are inherently birefringent. 
 \begin{figure}[h!]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Rotation of birefringent micro spheres} 
+\section{Rotation of birefringent micro spheres} 
 Optical tweezing has often been used for micro-rheology, by computing the
@@
 Using a birefringent particle and rotating it within the fluid, the maximum 
-rotation rate is due to the fluid drag resisting the torque of the 
-trapping beam \cite{RodriguezSevilla2018}. If you want to measure 
-fluid flow you can instead use a micro-rotor to see how fluid flow 
-propagates in the medium \cite{Knoener2005}. Likewise, one can use 
-a galvanometric mirror to probe the drag force of the fluid, by 
-understanding the trap strength (calibrating using a low frequency 
-signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I
+rotation rate is due to the fluid drag resisting the torque of the trapping 
+beam \cite{RodriguezSevilla2018}. If you want to measure fluid flow you can 
+instead use a micro-rotor to see how fluid flow propagates in the medium \cite{Knoener2005}. Likewise, one can use a beam steering arrangement to 
+probe the drag force of the fluid, by understanding the trap strength 
+(calibrating using a low frequency signal) one can measure the drag force 
+experienced by the local fluid \cite{RobertsonAnderson2018}. I
 
 Understanding the fluid velocity around our trapped object is determined 
-mostly by the Reynold's number of our system, for a sphere submersed 
-in a moving fluid of velocity $U$ this is given by:
+mostly by the Reynold's number of the system, for a sphere submersed in a 
+moving fluid of velocity $U$ this is given by:
 \begin{align}
@@
 \end{align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Estimation of fluid flow around micro-rotors}
+First we determined the upper rotation rate that could be achieved 
+using both Vaterite and liquid crystal spheres. Vaterite samples were 
+synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
+\ref{sec:vaterite}), and then suspended in distilled water. A sample
+of $200\ \mu L$ was pipptted and a single microsphere was captured via
+a circular polarised beam.
 
-First we tested the rotational behaviour of vaterite in distilled water, 
-samples of vaterite where diluted down and $200\mu L$ was pipetted onto
-the sample stage. Due to Van der Waal's forces some of the microspheres 
-were stuck together, fortunately individual sphere's were still present.
-Over the course of a few weeks multiple sphere's where trapped 
-and their rotation speeds were measured via the QPD. 
+Due to Van der Waal's forces some of the microspheres were stuck together, 
+fortunately individual sphere's were still present. Multiple microspheres 
+were trapped and their rotation rate was determined by looking at the 
+peak frequency component of the collected power spectrum. The shear flow
+was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
+were operating in the bulk fluid and away from any boundaries.
 \begin{figure}[h!]
-	\begin{subfigure}{0.5\linewidth}
+	\centering
+	\begin{subfigure}{0.82\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
-		\subcaption{}
 	\end{subfigure}
-	\begin{subfigure}{0.5\linewidth}
+	\begin{subfigure}{0.82\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
-		\subcaption{}
 	\end{subfigure}
-	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere.
-(b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line.
-Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW.
-Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
+	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
+		vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
+		is indicated by the dotted line. Vaterite radii and rotation frequencies 
+		are shown, the laser power was kept constant at 450 mW. Reported rotation
+		rates, and their corresponding fluid flow and shear rates, for vaterite 
+		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
 	\label{fig:vaterite_shear}
@@
 be in part due to the fact that synthesising perfectly spherical spheres that
-have uniform birefringence is difficult over multiple experiments and tests. 
-Despite our best efforts at controlling the growth rate the vaterite spheres
-would often combine together while suspended in water after a short period of 
-time. The fastest reported rotation rate found during this PhD was by \cite{Arita2016}
-that achieved a rotation rate of $5\ kHz$, this is plotted on Fig~\ref{fig:vaterite_shear}
-as the dotted line. Even at that extreme a rotation rate the region in which 
-nucleation is at its optimal likelihood is only $20\ nm$ wide meaning that 
-localising nucleation around a rotating sphere is vanishingly small.
+have uniform birefringence across the whole population is difficult. Despite 
+our best efforts at controlling the growth rate the smallest particle ever 
+synthesised was around $3\ \mu m$ in diameter. The vaterite spheres would 
+often stick together while suspended in water after a short period of time. 
+The fastest reported rotation rate found during this project was by 
+\cite{Arita2016} that achieved a rotation rate of $5\ kHz$, this is plotted 
+on Fig~\ref{fig:vaterite_shear} as the dotted line. Even at that extreme a 
+rotation rate the region in which nucleation is at its optimal likelihood is 
+only $20\ nm$ wide. This suggests that a more complicated fluid flow 
+arrangement is required in order to achieve the necessary shear rate and over 
+a larger volume of fluid.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rotation of micro-rotors in supersaturated solutions}
-Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
-(see sec.~\ref{sec:vaterite}), and then suspended in distilled 
-water, at the same time a supersaturated solution of Glycine 
-and water was prepared and $50\ \mu L$ was pipetted onto a 
-glass cover slip. A single microsphere the vaterite was trapped 
-in a circularly polarised light and brought close to the droplet 
-edge.
 \begin{figure}[h!]

### 2024-08-02 Latex_files/Thesis/chapters/chapter 4/main.tex
tidying up chapter 3

@@
 
-\newpage
 For completeness the harmonic traps were located for dimers across a range 
@@
 	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
-		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
+		off axis orientation with a circular polarised beam ($P= 100\ mW$). 
 		The far left column depicts the dimer's centre of mass position with 

### 2024-08-07 Latex_files/Thesis/chapters/chapter 4/main.tex

@@
 After computing its $T$-matrix via \textit{mstm} and supplying that 
-to \textit{ott} we compute the optical force via \eqref{eq:optical_force} 
-in the axial direction while the dimer is in its 'standard' orientation. 
-As expected we see a single point where the dimer will be in equilibrium, 
-the linear fit in fig.~\ref{fig:paradigmatic}(a) indicates a that the force 
-can be modelled as a harmonic potential close to the equilibrium position ($F_z\approx-\kappa_z z$). The second point where the axial 
-force goes to 0 cannot be considered as equilibrium position as the positive 
-gradient indicates that the trap is unstable unless Brownian motion is ignored.
+to \textit{ott} we compute the optical force exerted by a $50\ mW$ laser
+via \eqref{eq:optical_force} in the axial direction while the dimer is 
+in its 'standard' orientation. As expected we see a single point where 
+the dimer will be in equilibrium, the linear fit in fig.~\ref{fig:paradigmatic}
+(a) indicates a that the force can be modelled as a harmonic potential close 
+to the equilibrium position ($F_z\approx-\kappa_z z$). The second point 
+where the axial force goes to 0 cannot be considered as equilibrium position 
+as the positive gradient indicates that the trap is unstable unless Brownian 
+motion is ignored.
 
@@
 	\centering
-	\begin{subfigure}{.75\linewidth}
+	\begin{subfigure}{.65\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
@@
 	\medskip
-	\begin{subfigure}{.75\linewidth}
+	\begin{subfigure}{.65\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
@@
 		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
-		the case where the dimer is in its' 'standard' orientation. 
-		(c) is the case where the dimer is in its' 'inverted' orientation 
+		the case where the dimer is in its' 'standard' orientation, where
+		the dimer is trapped at $z = 1.71\ \mu m$. (c) is the case where 
+		the dimer is in its' 'inverted' orientation, the dimer is trapped 
+		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. 
 		(b) and (d) are renders to visualise the dimer orientation are shown 
@@
 
-We can see that both equilibrium positions have comparable axial trap stiffness 
-($\kappa_z$), however the difference in the transverse trap stiffness ($\kappa_x$) 
-is far more noticeable. As shown in Fig~\ref{fig:transverse_force}, 
-the dimer's orientation and relative position significantly changes the force 
-curve; not only is the trap wider when inverted but the trap stiffness is 
-increased. This highlights one of the challenges involved with studying 
+We can see that both equilibrium positions have comparable axial trap 
+stiffness ($\kappa_z$), however the difference in the transverse trap 
+stiffness ($\kappa_x$) is far more noticeable. The same dimer was trapped
+at each of the axial equilibrium positions and the transverse force was 
+evaluated. While in all three cases the dimer can be trapped the linear
+range where 
+This highlights one of the challenges involved with studying 
 asymmetric particles, even though its a simple enough process to trap them 
@@
 	\centering
-	\begin{subfigure}{0.85\linewidth}
+	\begin{subfigure}{0.75\linewidth}
 		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
@@
 	\medskip
-	\begin{subfigure}{0.85\linewidth}
+	\begin{subfigure}{0.75\linewidth}
 		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
@@
 optically trapped dimer in beams of varying polarisation ($NA=1.2$, 
-$P=100\ mW$), and the dimer is composed of silica ($n_p=1.59$, 
+$P=100\ mW$), and the dimer is composed of polystyrene ($n_p=1.59$, 
 $n_m=1.33$). Each simulation was run for $1$ second ($\Delta t 
@@
 As a benchmark we start by considering a single sphere within an optical trap. 
-A single silica sphere suspended in water ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$)
+A single polystyrene sphere suspended in water ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$)
 is trapped by a focus Gaussian beam and its trajectory was recorded. A 3 second 

### 2024-08-07 Latex_files/Thesis/chapters/chapter 4/main.tex
chp 4

@@
 
+\newpage
 For completeness the harmonic traps were located for dimers across a range 
@@
 	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
-		off axis orientation with a circular polarised beam ($P= 100\ mW$). 
+		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
 		The far left column depicts the dimer's centre of mass position with 

### 2024-08-12 Latex_files/Thesis/chapters/chapter 1/main.tex
updates to ch3

@@
 \chapter{Introduction}
-Crystallisation is one of the oldest forms of phase separation 
-used by humanity\cite{Schoen1956}, put simply, it is the 
-formation and growth of a new structured phase within 
-a disordered bulk phase. This has applications in a 
-number of industries such as pharmaceuticals, food 
-production, and electronics \cite{Myerson2002}; where 
-the extraction of dilute materials can help improve 
-product quality while keeping production costs low. 
-The industrialisation of crystallisation has allowed 
-engineers to reliably and efficiently induce the crystal 
-formation within a bulk phase. However, while on a 
-large scale crystallisation is seemingly an understood 
-physical process, just a small amount of investigation 
-into the literature reveals that at a micro scale there 
-is yet to be unifying theory that can accurately explain 
-the process of crystallisation \cite{Fu2021}. Of which 
-there two primary areas of focus, nucleation and crystal 
-growth. The latter focusing on how an already stable 
-crystal grows and how it takes on its final shape; 
-whereas the former is more concerned with which factors 
+Crystallisation is one of the oldest forms of phase separation used by humanity\cite{Schoen1956}, put simply, it is the formation and growth 
+of a new structured phase within a disordered bulk phase. This has 
+applications in a number of industries such as pharmaceuticals, food 
+production, and electronics \cite{Myerson2002}. Where the extraction 
+of dilute materials can help improve product quality while keeping 
+production costs low. The industrialisation of crystallisation has 
+allowed engineers to reliably and efficiently induce the crystal 
+formation within a bulk phase. However, while on a large scale 
+crystallisation is seemingly an understood physical process, just a 
+small amount of investigation into the literature reveals that at a 
+micro scale there is yet to be unifying theory that can accurately 
+explain the process of crystallisation \cite{Fu2021}. Of which there 
+two primary areas of focus, nucleation and crystal growth. The latter 
+focusing on how an already stable crystal grows and how it takes on its 
+final shape. Whereas the former is more concerned with which factors 
 contribute to its formation in order to control crystal growth.
 
-The focus of this research is partly based around investigating
-the impact of a focused electromagnetic field can influence the 
-nucleation process by inducing shear flow within the bulk phase. 
-This in hope that the localised EM field can be used to study 
-different phenomena in nucleation in order to develop a better
-understanding of the underlying mechanisms.
+The focus of this research is partly based around investigating the impact 
+of a focused electromagnetic (EM) field can influence the nucleation process 
+by inducing shear flow within the bulk phase. This in hope that the 
+localised EM field can be used to study different phenomena in nucleation 
+in order to develop a better understanding of the underlying mechanisms.
 
 \section{Nucleation}
-Nucleation is an example of a binary phase separation, 
-where a dilute phase is miscible in a bulk phase, more 
-often called the solute and solvent respectively. Because 
-of thermodynamics, the two can only remain in equilibrium while 
-below a specific concentration ($C_{eq}$) - below which 
-the chemical potential $\mu$ for a miscible solution 
-is greater than the potential required to separate 
-the two phases. Once $C_{eq}$ is exceeded there is a chemical
- potential difference driving the solution to separate the 
-two phases. Since different combinations of solute and solvent
-will have different equilibrium concentrations, researchers often 
-instead measure the ratio between the solute and solvent by using 
-'supersaturation', one form of measuring supersaturation is shown 
-below \cite{Mullin2001}:
+Nucleation is an example of a binary phase separation, where a dilute phase 
+is miscible in a bulk phase, more often called the solute and solvent 
+respectively. Because of thermodynamics, the two can only remain in 
+equilibrium while below a specific concentration ($C_{eq}$) - below which 
+the chemical potential $\mu$ for a miscible solution is greater than the 
+potential required to separate the two phases. Once $C_{eq}$ is exceeded 
+there is a chemical potential difference driving the solution to separate 
+the two phases. Since different combinations of solute and solvent will 
+have different equilibrium concentrations, researchers often instead measure 
+the ratio between the solute and solvent by using 'supersaturation', one 
+form of measuring supersaturation is shown below \cite{Mullin2001}:
 \begin{align}
@@
 
-Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
-is the equilibrium concentration at temperature $T$.
-While the solution remains supersaturated there is a 
-chemical potential driving force for the solute 
-to coalesce and separate from the solution as an ordered solid, 
-the first formation of the crystal is referred to as the 
-nucleus and understanding its formation of has been the 
-focus of researchers for decades now. Typically, for an 
-industrial crystallisation process the working principle 
-is based on controlling and manipulating the supersaturation
-of the system.
+Where $C_{sol}$ is just the solute concentration, and $C_{eq}(T)$is the 
+equilibrium concentration at temperature $T$. While the solution remains 
+supersaturated there is a chemical potential driving force for the solute 
+to coalesce and separate from the solution as an ordered solid, the first 
+formation of the crystal is referred to as the nucleus and understanding its 
+formation of has been the focus of researchers for decades now. Typically, 
+for an industrial crystallisation process the working principle is based on 
+controlling and manipulating the supersaturation of the system. Of course 
+nucleation events in a variety of different situations which each require 
+their own classification.
 
 \subsection{Primary \& Secondary nucleation}
-From an industrial perspective, the nucleation process can
-be broadly categorised into either primary of secondary 
-nucleation. The former being described as the formation of an
-initial nucleus within the bulk phase; absent of any external
-stimuli primary nucleation is considered stochastic and is typically
-undesirable for industrial applications where short and consistent 
-residence times are desirable. At a small scale one can estimate the 
-nucleation rate by making repeated measurements of sample solutions 
-and seeing how many have nucleated after a given time, giving us a 
-Poisson distribution.
+From an industrial perspective, the nucleation process can be broadly 
+categorised into either primary of secondary nucleation. The former 
+describes the formation of an initial nucleus within the bulk phase, 
+absent of any external stimuli. Primary nucleation is therefore considered 
+stochastic as there is no reliable means of predicting where a nucleus
+may form, or how long it will take. The only reliable indicator being 
+that higher supersaturations will result in faster nucleation rates. At 
+a small scale one can estimate the nucleation rate by making repeated 
+measurements of sample solutions and seeing how many have nucleated 
+after a given time, giving us a Poisson probability distribution.
 
@@
 
-Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$
-is the volume of the individual samples, and $M^*(t)$ \& $M$ are the 
-number of nucleated samples and the total number of samples used respectively.
-While this is useful for studying the effects of different parameters
-at a small scale, for industrial applications there are so many
-additional factors at play that can throw off this estimation.
+Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$ is 
+the volume of the individual samples, and $M^*(t)$ \& $M$ are the number 
+of nucleated samples and the total number of samples used respectively.
+While this is useful for studying the effects of different parameters at 
+a small scale, for industrial applications there are to many external 
+factors for primary nucleation to be the only
 
@@
 \subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the 
-original theory was first formed from the works of Volmer 
-and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
-While initially it was more focused into describing droplet 
-formation in condensing vapours it was extrapolated to describe
-crystallisation. The central premise of classical theory 
-is that nucleation occurs stochastically throughout the bulk
-phase due to collisions between individual solute molecules,
-ions, or atoms; at the same time the bulk phase is resistant
-to the formation of a new phase. The competition between
-these random collisions and the bulk solution can be used to 
-predict the probability of a newly formed nucleus.
+Sometimes referred to as 'Gibs Nucleation Theory' the original theory 
+was first formed from the works of Volmer and Weber, and Frenkel 
+\cite{Frenkel1939, Volmer1926}. While initially it was more focused 
+into describing droplet formation in condensing vapours it was 
+extrapolated to describe crystallisation. The central premise of 
+classical theory is that nucleation occurs stochastically due to 
+collisions between individual solute molecules, ions, or atoms. At the 
+same time the bulk phase is resistant to the formation of a new phase. 
+The competition between these random collisions and the bulk solution 
+can be used to predict the probability of a newly formed nucleus.
 
 Consider a supersaturated solution, after some time enough individual
-sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
-The newly formed phase has a lower chemical potential than the 
-surrounding solution, reducing the free energy of the system; 
-at the same time, the formation of a new interface is resited 
-by the bulk phase due to surface tension. The net free energy 
-of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
+sub units collide, forming a nucleus of volume $4\pi r^3/3$. The newly 
+formed phase has a lower chemical potential than the surrounding solution, 
+reducing the free energy of the system. Simultaneously, the formation of 
+a new interface is resited by the bulk phase due to surface tension. 
+The net free energy of the system for a nucleus of radius $r$ is given 
+as \cite{Karthika2016}:
 \begin{align}
 	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
+	\label{eq:CNT} 
 \end{align}
 
-Where $v$ is the approximate volume of an individual molecule, $k_B$
-is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
-tension of the bulk solution. Plotting the free energy of the system
-against nucleus size reveals a critical size above which the gain 
-in free energy exceeds the interfacial tension. 
+Where $v$ is the approximate volume of an individual molecule, $k_B$ is 
+the Boltzmann constant, and $\sigma_{inf}$ is the interfacial tension of 
+the bulk solution. This assumes that the nucleus will have a spherical 
+morphology so that the surface tension $\sigma_{inf}$ is a scalar value. 
+Looking at \eqref{eq:CNT} suggests that there must be some critical size 
+$r$ where the free energy gain from the nucleus exceeds the surface tension 
+of the surrounding fluid. Plotting the free energy of the system against 
+nucleus size reveals a critical size above which the gain in free energy 
+exceeds the interfacial tension. Furthermore it shows how increasing the 
+supersaturation of the system reduces said barrier. 
 \begin{figure}[h!]
@@
 
-The maximum value of $\Delta G_{tot}$ is the free energy barrier 
-that any newly formed nucleus needs to overcome in order to stabilise.
-The nucleation rate (the volume of new crystalline material formed per
-unit time), is therefore commonly defined as being dependent on the 
-energy barrier $\Delta G^*$:
+The maximum value of $\Delta G_{tot}$ is the free energy barrier that any 
+newly formed nucleus needs to overcome in order to stabilise. The nucleation 
+rate (the volume of new crystalline material formed per unit time), is 
+therefore commonly defined as being dependent on the energy barrier $\Delta G^*$:
 \begin{align}
@@
 
-\subsection{Optical Rotation}
+\subsection{Optical Torque}
 For any electromagnetic field it is possible to transfer both linear and 
@@
 
-\subsection{Shear induced Nucleation}
+\section{Shear induced Nucleation}
 It has long been known that fluid shear rate plays a role in influencing

### 2024-08-12 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

@@
 trapping and rotation. Synthesis of vaterite micro spheres requires fine 
-control of the nucleation process in order to maintain polymorphic stability, 
-though for the purposes of optical rotation the exact polymorph is not as 
+control of the crystal growth process in order to maintain polymorphic stability. 
+Though for the purposes of optical rotation the exact polymorph is not as 
 important as its morphology as all 3 polymorphs are inherently birefringent.
@@
 rotational motion of ellipsoidal polystyrene particles. The critical 
-assumption being that the particle perfectly rotates in the $x-y$ plane; 
-however it has long been suspected that birefringent microspheres 
-experience torques outside of the $x-y$ plane \cite{Volpe2023} making 
-it very difficult to characterise the behaviour of rotating birefringent 
-microspheres without a proper understanding of the full optical torque 
-being applied to it.
+assumption being that the particle perfectly rotates in the $x-y$ plane. 
+It has long been suspected that birefringent microspheres experience torques 
+outside of the $x-y$ plane \cite{Volpe2023} making it very difficult to 
+characterise the behaviour of rotating birefringent microspheres without a 
+proper understanding of the full optical torque being applied to it.
 
@@
 reported by the QPD, and $r$ is the distance from the particle's centre. 
-Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating
+Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
 outward from the centre of the sphere. The shear rate can then be computed
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Estimation of fluid flow around micro-rotors}
+\subsection{Estimation of fluid flow around micro-rotors in bulk fluid}
 First we determined the upper rotation rate that could be achieved 
@@
 	\centering
-	\begin{subfigure}{0.82\linewidth}
+	\begin{subfigure}{0.75\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
 	\end{subfigure}
-	\begin{subfigure}{0.82\linewidth}
+	\begin{subfigure}{0.75\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
@@
 rotation rate the region in which nucleation is at its optimal likelihood is 
-only $20\ nm$ wide. This suggests that a more complicated fluid flow 
-arrangement is required in order to achieve the necessary shear rate and over 
-a larger volume of fluid.
+only $20\ nm$ wide. If instead the micro-rotor was within the vicinity of a
+solid boundary, the shear rate would be enhanced due to the no-slip boundary
+condition.
 
+\section{Micro-rotors in Supersaturated solution}
+If rotation rates in bulk solution are insufficient then a micro-rotor rotating
+close to an artificial barrier may be able to improve the shear rate of the 
+surrounding fluid. Of course placing a solid barrier in a supersaturated fluid
+may well encourage nucleation somewhere on the surface outside of our control.
+Instead we chose to use the droplet edge of the supersaturated solution, while
+not a hard barrier per say, the molecular mobility close to the droplet edge is
+reduced due to surface tension. Furthermore, it has been shown through multiple 
+results that nucleation is enhanced at the air-solution interface \cite{Liao2022,
+Yuyama2010, Sugiyama2009}. 
 \begin{figure}[h!]
@@
 
-After a period of ten minutes if no nucleation event was 
-observed the particle was released and another particle was 
-trapped. Due to the increased viscosity of the supersaturated 
-solution and the proximity to the droplet edge, as previous works
-have only seen nucleation near the solution-air interface \cite{Gowayed2021, 
-Liao2022, Walton2018}, the observed rotation rate was rather low or 
-non-existent, the results are tabulated below:
+Supersaturated solutions of glycine and water were prepared and stored in
+an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
+into the solution and $20\ \mu L$ was pipetted onto the cover slip. A vaterite
+sphere was located, trapped, and moved as close to the droplet edge as possible.
+After measuring the microsphere's rotational frequency the sphere was left to
+rotate for a period of ten minutes after which, if no nucleation event was 
+observed the particle was released. The overall results are catalogued in 
+table~\ref{tab:Nucleation}
 \begin{table}[h!]
@@
 	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
+	\label{tab:Nucleation}
 	\begin{tabular}[width=\textwidth]{|c|c|c|c|}

### 2024-08-12 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to ch3

@@
 	\centering
-	\begin{subfigure}{.65\linewidth}
+	\begin{subfigure}{.75\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
@@
 	\medskip
-	\begin{subfigure}{.65\linewidth}
+	\begin{subfigure}{.75\linewidth}
 		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
@@
 evaluated. While in all three cases the dimer can be trapped the linear
-range where 
+range where that would typically associated with a stable trap is far 
+narrower in the 'standard' orientation compared to the 'inverted' cases. 
 This highlights one of the challenges involved with studying 
@@
 	\centering
-	\begin{subfigure}{0.75\linewidth}
+	\begin{subfigure}{0.85\linewidth}
 		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
@@
 	\medskip
-	\begin{subfigure}{0.75\linewidth}
+	\begin{subfigure}{0.85\linewidth}
 		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
@@
 
-One aspect that has yet to be covered in depth with regards to
-spherical aggregates of any construction is their interaction with
-circularly polarised light. While most homogenous objects are trapped
-no differently in different polarisations, dimers have been shown to
-experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018}. 
-Typically the spin density of an electric field cannot be reduced in 
-homogenous medium due to the fact that the spin angular momentum is 
-conserved locally. In which case the only way to transfer angular 
-momentum to use an object that is anisotropic. This is seen most 
-clearly with birefringent materials, due to fact that the crystal 
-lattice itself is anisotropic, a circularly polarised beam will 
-transfer an optical torque that is proportional to the difference in 
-the refractive index of the particle's ordinary and extraordinary axis. 
-The total optical torque is given as:
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
+One aspect that has yet to be covered in depth with regards to spherical 
+aggregates of any construction is their interaction with circularly 
+polarised light. For homogenous spheres the optical torque is regarded as
+being negligible as the spin density cannot impart angular momentum while 
+propagating in a homogenous medium. Dimers however, have been shown to
+experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
+while trapped in a circular polarised beam.
 
-Where $\Delta n$ is the difference in refractive indices of the 
-particle, $\phi$ indicates the phase shift in the electric field
-(for circular polarised light $\phi= \pi/4$), and $\theta$ is the angle
-between the particle's main axis and the direction of the electric field.
-The optical torque of non-birefringent spherical aggregates has no such 
-relationship as the mechanism behind their rotation is not as clearly defined.
-
-To visualise the relationship between the polarisation state of the 
-trapping beam, and the rotation rate we simulated the motion of an 
-optically trapped dimer in beams of varying polarisation ($NA=1.2$, 
-$P=100\ mW$), and the dimer is composed of polystyrene ($n_p=1.59$, 
-$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t 
-=10^-5$) and at the end we looked at the orientational time series; 
-the dimer's orientation is recorded as a quaternion which can be easily 
-converted to a 3-dimensional rotation matrix. By considering only the 
-transverse components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) 
-of the rotation matrix and taking the Fourier transformation of their 
-time series reveals the rotational frequency. The laser power is set to 
-$100\ mW$ to avoid large rotational fluctuations and so that the Fourier 
-series of the transverse components approximates $\delta(\omega_{rot}-f)$ 
-- the Dirac delta function centred at the rotational frequency $\omega_{rot}$.
+First we wanted to confirm that the rotation seen in our simulations is
+based on the polarisation of the trapping beam. To that end, we simulated 
+the motion of an optically trapped dimer in beams of varying polarisation 
+($NA=1.2$, $P=100\ mW$), the dimer is composed of polystyrene ($n_p=1.59$, 
+$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t =10^-5$) 
+and at the end we looked at the orientational time series; the dimer's 
+orientation is recorded as a quaternion which can be easily converted to 
+a 3-dimensional rotation matrix. By considering only the transverse 
+components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
+matrix and taking the Fourier transformation of their time series reveals 
+the rotational frequency. The laser power is set to $100\ mW$ to avoid 
+large thermal fluctuations and so that the Fourier series of the transverse 
+components approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function 
+centred at the rotational frequency $\omega_{rot}$.
 
@@
 
-A possible explanation of this phenomena was devised by by \cite{Yevick2017}, 
-who found that highly focused Gaussian beams could produce second 
-order effects in the Rayleigh regime resulting in a photo-kinetic 
-force that results in orbital motion about the beam's central axis.
-This was verified by \cite{Ruffner2012} who found that by trapping 
-a silica bead and recording its position over the course of several 
-minutes; while not immediately evident from the trajectory by measuring 
-the probability flux of finding a particle in a given position 
-revealed a helicity to the sphere's motion which matched the helicity 
-of the trapping beam. This would hardly be considered similar to 
-the behaviour shown by dimers; firstly because the probability flux 
-is an indicator of the sphere's position not its orientation - the 
-sphere can be thought to 'orbit' around the central beam axis with a 
-radius of orbit on the order of $10^{-9}\ m$. Secondly, while 
-\cite{Yevick2017} does compute a rotational frequency, for a single 
-sphere this rotation rate is at most $0.01\ Hz$ whereas the lowest 
-rotation rate recorded for a dimer ($a_{I}/a_{II}$) was around $2\ Hz$, 
-a full two orders of magnitude greater. As such the rotation demonstrated
-cannot be explained due to photo-kinetic forces.
+This shows us that these optical rotations are polarisation dependent 
+and not merely an example of the dimer scattering light asymmetrically. 
+The question then becomes, by what mechanism is the angular momentum of
+the trapping beam being transferred to the dimer.
 
-This rotation was first noted by Vigilante and co-workers who only 
-considered this behaviour for a symmetric dimer \cite{Vigilante2020}; 
-to expand upon this work we looked at how the rotation rate changed
-with the dimers' size ratio ($a_{I}/a_{II}$) and if the orientation
-of the dimer plays any impact on the rotational frequency. By repeating
+The rotation was first noted by Vigilante and co-workers who only 
+considered this behaviour for a symmetric dimer \cite{Vigilante2020}.
+In their work they attribute this to spin-curl effects, in which the 
+curl of the spin density leads to a second order optical force that 
+orbits around the beams central axis \cite{Yevick2017}. While several 
+papers have demonstrated this phenomena \cite{Zhao2007,Zhao2009, Wang2010} 
+it was only properly formalised by by \cite{Ruffner2012}. In which they 
+showed that the seemingly random trajectory of a trapped sphere was biased
+by the polarisation state of the trapping beam. While not immediately evident 
+from the trajectory the helicity of the trapping beam was revealed by 
+computing the particle's probability flux using.
+\begin{align}
+	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
+	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
+	\label{eq:prob_flux}
+\end{align}
+
+where $\delta_{\sigma_j}$ is the kernel of an adaptive density estimator
+\cite{Silverman1986}. \eqref{eq:prob_flux} describes the direction a trapped
+sphere is most likely move in given our statistical knowledge of the 
+trajectories probability density function. A finite estimation of the density
+function $p(r)$ is used in \cite{Ruffner2012}.
+\begin{align}
+	p(r) = \frac{1}{N}\sum_{j=1}^N \delta_{\sigma_j}(r-r_j)
+	\label{eq:prob_density}
+\end{align}
+
+The probability flux reveals a biased motion in the trajectory of a single 
+sphere (see Fig.~\ref{fig:Ruffner_Grier}). This biased motion results in a
+slight orbital motion about the central axis of the trapping beam, the orbital
+frequency is shown to be proportional to the polarisation state of the trapping
+beam. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
+	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how the momentum 
+		density of a Gaussian beam is twisted while using circularly polarised 
+		light. The top row (figures (b)-(d)) shows a 7 sphere cluster trapped 
+		in a circularly polarised beam. Due to the clusters asymmetric 
+		susceptibility to polarization the cluster rotates in the $x-y$ plane.
+Whereas the bottom row (figures (e) - (g)) show the similar results for a single sphere.
+In this instance the sphere does not rotate but instead
+		orbits the beam axis. In both instances the motion is proportional to 
+		the degree of polarisation (see figures (d) and (g)) but for the single
+		sphere this motion is only revealed when using \eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused with permission from author}
+	\label{fig:Ruffner_Grier}
+\end{figure}
+
+While the results from \cite{Ruffner2012} suggest that the optical rotation 
+seen in asymmetric dimers is attributed to the same spin-curl forces there 
+are several discrepancies that cannot be explained purely by the spin-curl 
+force. Firstly, there is the question of how the spin-curl force results in
+an optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) and (e) we 
+can see that the behaviour of the 7 sphere cluster rotates about the central 
+axis whereas the single sphere has an only slight orbital bias in its motion. 
+The authors attributed this to the fact that the cluster is scattering light 
+where the spin-curl force is more substantial due to it having a wider profile. 
+This would suggest that particles' whose longer axis lies in the plane perpendicular
+to the direction of propagation should experience any notable degree of 
+torque. And indeed this appears to be the case for a number of experiments 
+involving nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
+
+It does raise the question of how come our simulations show that despite the 
+dimer being orientated so their long axis' align with the parallel to the 
+direction of propagation they still readily rotate. You would not expect that
+the addition of a single additional sphere should drastically adjust the torque
+especially if said sphere is relatively small. However when we measured the 
+optical torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found 
+the exact opposite. In both cases we used the same trapping beam as used for
+figure~\ref{fig:rotation_vs_pol} but with a circularly polarised beam. Both 
+the sphere and dimer were rotated in the $x-z$ plane and the all three components 
+of the optical torque were recorded.
+\begin{figure}[h!]
+	\centering	
+	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
+	\caption{Optical toque experienced by a dimer ($a_{I}/a_{II}=10$) and a 
+	single isotropic sphere. Both were rotated in the $x-z$ plane and the angle 
+	between $U_z$ and the beam axis gives the polar angle.The solid lines denote 
+	the torque experienced by the dimer whereas the dashed lines represent the 
+	torque experienced by the sphere.}
+	\label{fig:sphere_dimer_torque}
+\end{figure}
+
+It should be noted that the difference in magnitude is not clearly shown in 
+figure~\ref{fig:sphere_dimer_torque}. In truth, the optical torque acting on 
+the dimer $10^{13}$ times greater than the torque applied to the single sphere. 
+The torques about the $x$ and $y$ axis can be somewhat understood as the second 
+sphere is being drawn back towards the centre of the trap by the gradient forces. 
+The same cannot be said for the $z$component of the optical torque; while this 
+effect could still be attributed to the spin-curl force, but it is clear that the internal scattering between the two spheres has some unintended effects.
+
+ By repeating
 the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
@@
 \subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity 
-sized dimers there is the potential for non-vertical trapping 
-orientations in which the dimer is located in a harmonic 
-trap. When trapped by a circularly polarised light these 
-dimers exhibit gyroscopic precession. As shown in fig~\ref{fig:gyro} 
-the dimer's is rotating about it's long axis (as seen by 
-the periodic behaviour of $U_x$ and $U_y$), while also precessing 
-around the optical axis of the beam, resulting in a second 
-order oscillation in $s_z$.
+As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there 
+is the potential for non-vertical trapping orientations in which the dimer is 
+still stably trapped. When a circularly polarised beam is used the dimer exhibits gyroscopic precession. As shown in fig~\ref{fig:gyro} the dimer's trajectory
+exhibits periodic rotation about its long axis $U_z$, precession motion as the
+dimer spins around the central beam axis, and nutation motion as the dimer's
+orientation rocks backwards and forwards. 
 \begin{figure}[h]
@@
 	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
-		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
+		off axis orientation with a circularly polarised beam ($P= 100\ mW$). 
 		The far left column depicts the dimer's centre of mass position with 

### 2024-08-12 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

@@
 tweezers \cite{Parkin2009, Saito2022}. Over the course of the project two 
-different micro spheres where investigated, vaterite and liquid crystal 
+different micro spheres where investigated, Vaterite and liquid crystal 
 droplets. Both can be readily synthesised in the lab and are will rotate 
@@
 due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
-other polymorphs of calcite and aragonite, when synthesised vaterite will 
+other polymorphs of calcite and aragonite, when synthesised Vaterite will 
 typically form small spherical particles making them ideal for optical 
-trapping and rotation. Synthesis of vaterite micro spheres requires fine 
+trapping and rotation. Synthesis of Vaterite micro spheres requires fine 
 control of the crystal growth process in order to maintain polymorphic stability. 
@@
 When trapped in circularly polarised light, the anisotropic crystal lattice
-allows spin angular momentum to be transferred to the vaterite particle, 
+allows spin angular momentum to be transferred to the Vaterite particle, 
 resulting in a rotation about the beam axis. In addition, the anisotropic 
@@
 	\end{subfigure}
-	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap.
-(b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
+	\caption{(a)Sample Vaterite sphere suspended in water and trapped by circular polarised trap.
+(b) Collected power spectrum from rotating Vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
 	\label{fig:vaterite}
@@
 The liquid crystal droplets had a much faster rotation rate than comparable 
-vaterite spheres, due to their higher degree of birefringence and the fact
+Vaterite spheres, due to their higher degree of birefringence and the fact
 that the droplets are far closer to perfect spheres making angular momentum
@@
 generating and measuring fluid flow in a solution. To see if we can 
-even achieve the theoretical maximum shear rate, vaterite spheres 
+even achieve the theoretical maximum shear rate, Vaterite spheres 
 were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and 
@@
 \ref{sec:vaterite}), and then suspended in distilled water. A sample
-of $200\ \mu L$ was pipptted and a single microsphere was captured via
+of $200\ \mu L$ was pipetted and a single microsphere was captured via
 a circular polarised beam.
@@
 	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
-		vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
+		Vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
 		is indicated by the dotted line. Vaterite radii and rotation frequencies 
 		are shown, the laser power was kept constant at 450 mW. Reported rotation
-		rates, and their corresponding fluid flow and shear rates, for vaterite 
+		rates, and their corresponding fluid flow and shear rates, for Vaterite 
 		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
@@
 our best efforts at controlling the growth rate the smallest particle ever 
-synthesised was around $3\ \mu m$ in diameter. The vaterite spheres would 
+synthesised was around $3\ \mu m$ in diameter. The Vaterite spheres would 
 often stick together while suspended in water after a short period of time. 
@@
 an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
-into the solution and $20\ \mu L$ was pipetted onto the cover slip. A vaterite
+into the solution and $20\ \mu L$ was pipetted onto the cover slip. A Vaterite
 sphere was located, trapped, and moved as close to the droplet edge as possible.
@@
 	\centering
-	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
+	\caption{Results from rotating Vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
 	\label{tab:Nucleation}
@@
 
-The calculations from fig.\ref{fig:vaterite_shear} indicate that even
-in pure water the rotational speed achieved by a microsphere is
-insufficient to achieve a significant shear rate more than a few tenths of a microns
-from its surface. The closest we could trap a microsphere to the droplet
-edge was in the range of $5-10\mu\ m$, at that distance the fluid flow
-would be so low that not even using a liquid droplet rotor would achieve
-the rotational speeds necessary to localise nucleation. While in theory
-a sufficiently focused laser could rotate any microsphere to a fast enough 
-to match the speeds expected by \cite{Debuysschere2023} the localised intensity
-would be so large that even using $D_2O$ would see a significant increase in 
-temperature.
+Trying to trap a particle close to the edge proved more challenging than 
+expected. Unlike in previous reports where the beam is focused at the upper
+edge of the droplet \cite{Liao2022, Yuyama2010, Sugiyama2022}, we attempted 
+trapping into the crook of the droplet. It is suspected that trapping is 
+much harder at the interface due to increased surface tension and unpredictable
+scattering forces. The closest we could trap a microsphere to the droplet edge 
+was in the range of $5-10\mu\ m$, at that distance the fluid flow is so low that 
+even the presence of a hard boundary would be insufficient for shearing the 
+fluid. Furthermore, as is evident in Table~\ref{tab:Nucleation}, the rotation 
+rate drops off significantly with increased supersaturation, due to higher 
+fluid viscosities.
+While in theory a sufficiently focused laser could rotate any microsphere to a fast enough to reach the shear rate predicted by \cite{Debuysschere2023} the localised intensity would be so large that even
+using $D_2O$ would see a significant increase in temperature.
 
-It is not impossible that fluid shearing could be used in the future to localise
-nucleation; but from these results, using individual micro-rotors do not seem like 
-an appropriate method. If multiple micro-rotors could be trapped in close proximity 
-to one another they could create a large region of fluid where nucleation is more 
-likely than the bulk fluid. Micro-rotors have been created that allow for precise
+It is not impossible that fluid shearing could be used in the future to 
+localise nucleation; but from these results, using individual micro-rotors 
+is not an appropriate method. Firstly, the area of influence is far too small 
+to see any noticeable increase in the nucleation rate. And secondly, increased 
+fluid viscosity significantly reduces the limits the maximum rotation rate 
+possible. If multiple micro-rotors could be trapped in close proximity to one 
+another they could create a large region of fluid where nucleation is more likely than the bulk fluid. Micro-rotors have been created that allow for precise
 control of suspended micro-particles \cite{Butaite2019} and could potentially be
-used to generate sufficient shearing, however these could not be used in this PhD 
-as we lacked the necessary hardware to form multiple gradient traps.
+used to generate sufficient shearing. However these could not be used in this 
+project as we lacked the necessary hardware to form multiple gradient traps.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shearing via a Galvano-mirror}
-Rather than utilising birefringence rotation, which is subject to issues
-during synthesis and achieving an ideal particle size, a galvano-mirror
-was installed to induce controlled particle motion. While typically 
-galvano and gimble mirrors are used to trap multiple particles in a 
-regular pattern, in a sufficiently dilute solution a single silica 
-micro sphere can be moved quickly through a fluid along a preset path.
-Calculating the shear rate around an individual particle is difficult to
-do precisely but for low Reynolds numbers we can get an adequate approximation.
+\section{Shearing via Galvano-mirror manipulation}
+An alternative approach to generating fluid shear is to use a galvano-mirror
+to rapidly move a trapped particle in a bulk fluid.
+While typically galvano and gimble mirrors are used to trap multiple particles simultaneously, a single micro sphere can be moved quickly through a fluid along a preset path.
+The only
+limitation on the particle's speed being the ratio of the trap stiffness to
+the drag force. Calculating the shear rate around an individual particle is 
+difficult to do precisely but for low Reynolds numbers we can get an adequate approximation.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nucleation using a moving beam}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Nucleation with a Stationary and Moving Beam}
 As mentioned previously, shearing via optical rotation and particle displacement
@@
 
+\subsection{Stationary beam}
 Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
@@
 using optical tweezers shows that the growth rate is only loosely connected to the 
-solutions supersaturation, local conditions play a much larger role in the growth 
+solutions supersaturation. Local conditions play a much larger role in the growth 
 rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
@@
 
-Now while trying to induce fluid flow by optically trapping a silica particle and 
-using a galvano-mirror to move the particle through the solution, nucleation events
-were observed in which crystal growth follows the laser focus. Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus moving in a small elliptical 
-pattern; interestingly, due to the addition of the silica suspension we see the presence
-of small spheres, these appear too small to be silica beads which have a uniform radius
-of $1.57\ \mu m$, we surmise these could be small clusters of Glycine that had previously
-been shown to form when the aqueous solutions where irradiated with a focused laser
-\cite{Tsuboi2009, Gowayed2021}. While no droplets are seen directly entering the focus a 
-nucleus forms close to the droplet edge, unlike in Fig.~\ref{fig:stationary_beam} the 
-crystal does not grow out from the focal point evenly; due to the galvano mirror, 
-the crystal is simultaneously being moved by and growing around the focal point of 
-the trap. Because of this the crystal nucleus lacks a clear morphology at first, 
-until roughly $20\ s$ the crystal reaches a almost prismatic structure, with further 
-irradiation increasing the size. Interestingly the galvano-mirror allows the trap to 
-impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) 
-and (d), where even though the crystal is not directly in the trap focus it rotates in 
-the $x-y$ plane and gets trapped again at a corner. The rotation could not be due to fluid
-flow close to the surface of the crystal as trap should have no influence on fluid molecules
-smaller than a few hundred nano-meters. In figs.\ref{fig:eliptical_beam_1}(e) and (f), 
-the crystal growth is localised to the corner, even though previously the crystal growth 
-was roughly uniform (compare figs. \ref{fig:eliptical_beam_1}(b) and (c)). The area growth
-rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at
-$45.03\ \mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and (f) the 
-growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
+\subsection{Moving Beam}
+\label{sec:stationary}
+To test if a rapidly moving silica bead could generate the necessary shear rate 
+for crystal nucleation we wanted to see if a trapped silica bead could be trapped
+and moved in an aqueous solution.
+$20\ \mu L$ of glycine and water ($S=1.03$) was added to $10\ \mu L$ of a dilute water-silica mixture making the solution undersaturated ($S\approx0.7$).
+However, due to the beam's motion we instead
+encountered several unexpected results.
+
+Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
+moving in a small elliptical pattern; interestingly. After a few minutes we 
+noted the appearance of several small droplets, these appear too small to be 
+silica beads which have a uniform radius of $1.57\ \mu m$. We surmise these 
+could be small clusters of Glycine that had previously been shown to form 
+when the aqueous solutions where irradiated with a focused laser \cite{Tsuboi2009,
+Gowayed2021}. While no droplets are seen directly entering the focus a nucleus 
+forms close to the droplet edge, unlike in Fig.~\ref{fig:stationary_beam} the 
+crystal does not grow out from the focal point evenly. Due to the galvano mirror, 
+the crystal is simultaneously being moved by and growing around the focal point 
+of the trap. Because of this the crystal nucleus lacks a clear morphology at 
+first. Until roughly $20\ s$ the crystal reaches a almost prismatic structure, 
+with further irradiation increasing the size.
+
+Interestingly the galvano-mirror allows the trap to impart a slight torque 
+on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
+where even though the crystal is not directly in the trap focus it rotates 
+in the $x-y$ plane and gets trapped again at a corner. The rotation could 
+not be due to fluid flow close to the surface of the crystal as trap should 
+have no influence on fluid molecules smaller than a few hundred nano-meters. 
+In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal growth becomes 
+localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
+\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
+(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
 \begin{figure}[h!]
@@
 
-In this instance $20\ \mu L$ of glycine and water ($S=1.03$) was added to $10\ \mu L$ 
-of a dilute water-silica mixture making the solution undersaturated ($S\approx0.7$). 
 Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
-\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both cases the
-crystal morphology is not influenced by the location of the laser focus, it has been
-shown that polarisation can influence the polymorphic control for simple salts 
-\cite{Garetz1996, Garetz2002} but no research has been done into the influence of a 
-moving beam front on the growth of a newly formed nucleus.
+\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both 
+cases the crystal morphology is not influenced by the location of the laser focus, nor do they report the appearance of glycine droplets. Reports of these droplets
+has been found only in supersaturated conditions, often using a lower NA objective
+to provide a wider trap focus \cite{Liao2022, Yuyama2010, Yuyama2012}.
+
+\subsection{Direct trapping of Glycine clusters}
+One common aspect involving these droplets is the fact when brought to the laser 
+focus the droplets would nucleate immediately \cite{Liao2022}. We made up a 
+solution similar to Sec~\ref{sec:stationary} but without any silica droplets, 
+we once again focused the beam close to the droplet edge. Again after a few 
+minutes of irradiation we noted the presence of droplets, because no silica had 
+been added these droplets had to be from the glycine solution. Trapping these
+
+It has been shown that polarisation can influence the polymorphic control for 
+simple salts \cite{Garetz1996, Garetz2002} but no research has been done into 
+the influence of a moving beam front on the growth of a newly formed nucleus.
 
@@
 To see if the above phenomena had any direct influence on a
+
+\section{Summary of Moving Beam Phenomena}
+To summarise, the introduction of a moving beam helps to accelerate the local
+growth of a newly formed crystal. This is due to the presence of glycine droplets
+that accumulate near the interface between the liquid solution and air.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{galvano_diagram.pdf}
+	\caption{Diagram outlining how a moving beam assists in the growth of a crystal
+		nucleus. (a) a crystal nucleus is partially trapped by a moving beam with 
+		solute droplets close to its surface. (b) droplets close to the laser
+		focus are drawn in by gradient forces and moved towards the crystal surface.
+		(c) these droplets provide material to the main crystal, resulting in
+		localised growth around the laser focus. (d) eventually the crystal area 
+		either fully surrounds the laser focus or the solution surrounding the laser
+		is depleted of solute material.}
+\end{figure}
+
+There are still several factors that need to be investigated, but due to time
+constraints it was not possible to properly study this phenomena. Firstly, 
+there is the question of what conditions result in the production of concentrated
+droplets, it is not clear if the presence of a laser is required or if these
+droplets naturally occurring. Prior literature would suggest that the laser is 
+required \cite{Liao2022, Tsuboi2009}, but this would not explain why in many 
+cases the droplets are found far outside the influence of the optical trap. It
+has been suggested that optical traps can attract microspheres over a wider 
+area, but it is hard to say that this would result in the creation of liquid 
+droplets \cite{Yi2021}.
+
+Secondly, there is the question of how these droplets supply material to the 
+bulk crystal. In some instances it is clear that the droplets are being drawn 
+into the trap, however, in other instances while there are no droplets close to
+the vicinity of the optical trap the crystal continues to grow. If these droplets
+are a necessary precursor to induce crystal nucleation then understanding how
+they are added to the bulk crystal may help with our understanding of the kinetics
+of multi-step nucleation.
+
+\section{Conclusion}

### 2024-08-12 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to ch3

@@
 therefore the 'inverted' orientation is defined when the dimer is orientated 
-against the direction of beam orientation (see figures~\ref{fig:paradigmatic}(b) 
-and (d)).
+against the direction of beam orientation (see left hand side of figure~\ref{fig:paradigmatic}).
 
@@
 		\centering
-		\raisebox{50pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
-		\caption{}
+		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
 		\label{large over small}
@@
 		\centering
-		\raisebox{30pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
-		\caption{}
+		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
 		\label{small_over_large}
@@
 		the case where the dimer is in its' 'standard' orientation, where
-		the dimer is trapped at $z = 1.71\ \mu m$. (c) is the case where 
+		the dimer is trapped at $z = 1.71\ \mu m$. (b) is the case where 
 		the dimer is in its' 'inverted' orientation, the dimer is trapped 
 		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. 
-		(b) and (d) are renders to visualise the dimer orientation are shown 
+		On the left are renders to visualise the dimer orientation are shown 
 		below each plot. The black lines on each force-curve is a linear 

### 2024-08-13 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

@@
 \subsection{Stationary beam}
+\label{sec:stationary}
 Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
@@
 \subsection{Moving Beam}
-\label{sec:stationary}
+\label{sec:moving}
 To test if a rapidly moving silica bead could generate the necessary shear rate 
@@
 \subsection{Direct trapping of Glycine clusters}
+\label{sec:clusters}
 One common aspect involving these droplets is the fact when brought to the laser 
@@
 \section{Summary of Moving Beam Phenomena}
-To summarise, the introduction of a moving beam helps to accelerate the local
-growth of a newly formed crystal. This is due to the presence of glycine droplets
-that accumulate near the interface between the liquid solution and air.
+To summarise, the introduction of a moving beam helps to accelerate the 
+local growth of a newly formed crystal. This is due to the presence of 
+glycine droplets that accumulate near the interface between the liquid 
+solution and air. The theory behind the localised growth can be summarised 
+thusly.
 
+Initial nucleation is similar to typical optical trapping induced nucleation,
+with the air solution interface limiting the molecular mobility of the solute
+molecules \cite{Liao2022, Sugiyama2009, Gowayed2021}. The moving beam front 
+can influence the motion of the nucleus initially, but eventually the drag 
+force means the crystal is not moved by the optical trap. Localised crystal
+growth occurs when the trap is close to or partially over the interface of 
+the crystal (see figure~\ref{fig:local_nucleation}(a)). As shown in \ref{sec:clusters}, the optical trap can manipulate these droplets similar 
+to microspheres. When in close proximity to the trap these droplets are 
+brought towards the crystal surface (see figure~\ref{fig:local_nucleation}(b)). 
+These provide material that grows the crystal around that region (see figure~\ref{fig:local_nucleation}(c)). Eventually the local solution is either 
+depleted of solute material or the crystal front has grown to fully encompass 
+the trap, preventing further growth (see figure~\ref{fig:local_nucleation}(d)). 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{galvano_diagram.pdf}
+	\includegraphics[width=0.8\linewidth]{galvano_diagram.pdf}
 	\caption{Diagram outlining how a moving beam assists in the growth of a crystal
@@
 		is depleted of solute material.}
+	\label{fig:local_nucleation}
 \end{figure}
+
+The reason the seed crystals saw no further crystal growth when irradiated 
+with the trap is due to the fact that these experiments were carried out 
+in a bulk solution, absent of any interfaces. As such the clusters seen in
+\ref{sec:moving} \& \ref{sec:clusters} are not present and cannot provide
+material to accelerate crystal growth.
 

### 2024-08-13 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

@@
 		\hline
-		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
+		Super Saturation & Particle radius [$\mu m$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
 		\hline
-		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
-		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
-		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
+		\multirow{3}*{1.01} & 2.34 & 10.4 & $\times$ \\
+		\cline{2-4} & 5.67 & 9.63 & $\times$ \\
+		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
 		\hline
@@
 \begin{align}
-	u_r(r)=-|v(t)|^2cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
+	u_r(r)=-|v(t)|^2\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
 \end{align}
 
-Where $\theta$ is the angle from the direction of movement to the point
-you wish to measure, and $r$ is the radial distance to that point. Again
-taking the partial derivative we can get the shear rate for a particle moving 
-through the fluid:
-
+Where $r$ is the radial distance to that point. Again taking the partial 
+derivative we can get the shear rate for a particle moving through the fluid:
 \begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
+	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
+	\label{eq:galvo_shear}
 \end{align}
@@
 Moving a silica bead along an circular path can generate significant fluid 
-flow around a larger volume compared to comparable microspheres.
+flow around a larger volume compared to comparable micro-rotors. Using \eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
+would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
+equates to moving the silica bead around a circular path with a frequency of 
+$100\ Hz$.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{galvano_shear_rate.png}
+	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
+		in bulk fluid moving at different speeds. The shear rate is calculated 
+		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
+		in a circular path and so the speed is constant through out its path.}
+	\label{fig:galvano_shear}
+\end{figure}
+
+From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
+mirror a better option for generating high shear rates but also over a larger 
+volume.
 
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{frames_no_beam_movement.pdf}
+	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
 	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{frames_eliptical_beam.pdf}
-	\caption{Laser induced nucleation at the edge of a droplet of 
-		supersaturated glycine solution. (b) shows the first instance 
-		of a crystal nucleus, growing quickly through (c)-(e) until 
-		after $6.665\ s$ the crystal begins to escape the trap.}
+	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
+	\caption{Frames from a longer video depicting the growth of a nucleus using 
+		a moving beam. Initially the crystal shape is amorphous (a) but eventually
+		reaches a more regular shape (b). This crystal is still influenced by the
+		optical trap as even when not directly irradiated by the laser the crystal
+		rotates between (c) and (d). When the laser is focused on a corner the 
+		crystal growth is localised to that region, resulting in an elongated 
+		section forming between frames (e) and (f).}
 	\label{fig:eliptical_beam_1}
@@
 \cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both 
-cases the crystal morphology is not influenced by the location of the laser focus, nor do they report the appearance of glycine droplets. Reports of these droplets
+cases the crystal morphology is not influenced by the location of the laser focus, 
+nor do they report the appearance of glycine droplets. Reports of these droplets
 has been found only in supersaturated conditions, often using a lower NA objective
@@
 One common aspect involving these droplets is the fact when brought to the laser 
-focus the droplets would nucleate immediately \cite{Liao2022}. We made up a 
-solution similar to Sec~\ref{sec:stationary} but without any silica droplets, 
-we once again focused the beam close to the droplet edge. Again after a few 
-minutes of irradiation we noted the presence of droplets, because no silica had 
-been added these droplets had to be from the glycine solution. Trapping these
+focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
+to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
+again the beam was focused close to the droplet edge, this time the galvano mirror
+was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
+After a few minutes of irradiation droplets were seen entering the camera frame, 
+because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
+several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
+\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
+crystal structure becomes clearer.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.9\linewidth]{cluster_trapping.pdf}
+	\caption{Frames from a longer video demonstrating the trapping of a glycine 
+		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
+		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
+		black) being brought into contact with a larger droplet. (b) upon contact 
+		a nucleus can be seen between the two droplets. The growth is rather slow 
+		with the crystal having no clear defined morphology through (c) and (d). 
+		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
+	\label{fig:cluster_trapping}
+\end{figure}
 
-It has been shown that polarisation can influence the polymorphic control for 
-simple salts \cite{Garetz1996, Garetz2002} but no research has been done into 
-the influence of a moving beam front on the growth of a newly formed nucleus.
+The fact that these droplets can be trapped indicates they must have a higher 
+refractive index than the surrounding solution. However, getting an exact 
+measurement of the
 
 \section{Influence of a moving beam front on seed crystals}
-To see if the above phenomena had any direct influence on a
+\textcolor{red}{Need to write up these results. To summarise the crystal 
+	dissolved around the laser focus instead of growing as the seed crystals 
+	are in the solution and not close to the air-solution interface}
 

### 2024-08-15 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 occur while the beam was close to the edge of the droplet, even though the solution
-was undersaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
-but was more interesting is how the beam's motion influenced the growth of the nucleus.
+was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
+but was more interesting is how the beam's motion influenced the growth of the 
+nucleus.
 
@@
 is in fact still growing as the solution is supersaturated. The key take away to 
-remember is that the beam has no real influence over the crystal shape, instead 
-it grows outward from the trap focus.
+remember is that the beam has no influence over the crystal shape, instead it 
+grows outward from the trap focus. Furthermore, due to the fact that the solution
+is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
+the crystal escapes as its size exceeds the trap focus. 
 \begin{figure}[h!]
@@
 	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
-	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the trap.}
+	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
+	the trap.}
 	\label{fig:stationary_beam}
@@
 for crystal nucleation we wanted to see if a trapped silica bead could be trapped
-and moved in an aqueous solution.
-$20\ \mu L$ of glycine and water ($S=1.03$) was added to $10\ \mu L$ of a dilute water-silica mixture making the solution undersaturated ($S\approx0.7$).
-However, due to the beam's motion we instead
+and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
+added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
+unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
 encountered several unexpected results.
@@
 moving in a small elliptical pattern; interestingly. After a few minutes we 
-noted the appearance of several small droplets, these appear too small to be 
-silica beads which have a uniform radius of $1.57\ \mu m$. We surmise these 
-could be small clusters of Glycine that had previously been shown to form 
-when the aqueous solutions where irradiated with a focused laser \cite{Tsuboi2009,
-Gowayed2021}. While no droplets are seen directly entering the focus a nucleus 
-forms close to the droplet edge, unlike in Fig.~\ref{fig:stationary_beam} the 
-crystal does not grow out from the focal point evenly. Due to the galvano mirror, 
-the crystal is simultaneously being moved by and growing around the focal point 
-of the trap. Because of this the crystal nucleus lacks a clear morphology at 
-first. Until roughly $20\ s$ the crystal reaches a almost prismatic structure, 
-with further irradiation increasing the size.
+noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
+for a larger example of these droplets). These appear too have a wide distribution
+of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
+\mu m$. We surmise these could be small clusters of Glycine that had previously 
+been shown to form when the aqueous solutions where irradiated with a 
+focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
+directly entering the focus a nucleus forms close to the droplet edge, 
+unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
+the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
+being moved by and growing around the focal point of the trap. Because of 
+this the crystal nucleus lacks a clear morphology at first. Until roughly 
+$20\ s$ the crystal reaches a almost prismatic structure, with further 
+irradiation increasing the size.
 
@@
 in the $x-y$ plane and gets trapped again at a corner. The rotation could 
-not be due to fluid flow close to the surface of the crystal as trap should 
-have no influence on fluid molecules smaller than a few hundred nano-meters. 
-In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal growth becomes 
-localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
+not be due to fluid flow close to the surface of the crystal as the dipole 
+moment of individual water molecules is too small to be influenced by an 
+optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
+growth becomes localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
 \mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
@@
 Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
-\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both 
-cases the crystal morphology is not influenced by the location of the laser focus, 
-nor do they report the appearance of glycine droplets. Reports of these droplets
-has been found only in supersaturated conditions, often using a lower NA objective
-to provide a wider trap focus \cite{Liao2022, Yuyama2010, Yuyama2012}.
+\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
+moving beam. This modification allows for the crystal growth to be localised to
+a specific region of the bulk crystal whereas with a stationary beam there is no
+control over the crystal morphology. In fact this allows for a much finer control
+over the exact shape of the crystal nucleus, in some cases allowing for growth out
+of the viewing plane as shown in
+
+\begin{figure}
+	\centering
+\end{figure}
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{cluster_trapping.pdf}
+	\includegraphics[width=0.7\linewidth]{cluster_trapping.pdf}
 	\caption{Frames from a longer video demonstrating the trapping of a glycine 
@@
 The fact that these droplets can be trapped indicates they must have a higher 
-refractive index than the surrounding solution. However, getting an exact 
-measurement of the
+refractive index than the surrounding solution. It has been shown that the 
+concentration of glycine solution is correlated with the refractive indices 
+of the liquid \cite{Gowayed2021, Orttung1963}. This suggests that these droplets 
+serve to provide solute material to the bulk crystal.
 

### 2024-09-06 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 06/09/2024

@@
 temperatures the solute concentration is too high to be fully incorporated
-into the solution ($S\gg 1$), after heating however the solute is fully dissolved
-into the solution ($S<1$). Now as the mixture is allowed to cool to room
-temperature the supersaturation will increase until crystal formation begins,
-the rate of cooling drastically influencing the size and number of crystals
-produced.
+into the solution ($S\gg 1$), after heating however the solute is fully 
+dissolved into the solution ($S<1$). Now as the mixture is allowed to cool 
+to room temperature the supersaturation will increase until crystal formation begins, the rate of cooling drastically influencing the size and number of crystals produced.
 
@@
 
+\subsection{Anti-solvent Crystallisation} 
+Anti-solvent crystallisation is a more involved separation method ,
+by adding a new solvent that is miscible with the old solvent but is 
+immiscible with the solute, the solute's effective solubility is decreased.
+As such, the solution becomes supersaturated. This has the trade off in that the overall mass ratio between the solvent and solute is now lower. Careful
+measurement of the anti-solvent is required in order to ensure that the 
+decrease in solubility is not out matched by the decrease in mass ratio.
+Anti-solvent crystallisation is not as common due to the fact that not all
+solutions will have an ideal anti-solvent candidate.
+
 \section{Optical Tweezers}
@@
 
+\subsection{Gaussian Beams}
+A Gaussian Beam is a simplistic approximation of a laser beam, in short
+it is an unbounded plane wave whose intensity of the electromagnetic 
+field falls off from the centre similar to a normal distribution. This 
+is perfectly fine for beams with a divergence angle of $0^{\circ}$, but 
+later studies into focused beams has shown that the assumption of a 
+Gaussian intensity profile is inaccurate. A study conducted by Lock and Gouesbet found that as the Gaussian beam was focused the fall off in 
+the electric field would consequentially induce additional electromagnetic fields \cite{Lock1994}.
+As a result the shape of the beam profile diverges from a Gaussian profile for focused beams.
+Regardless, for optical trapping experiments it is useful to approximate a Gaussian beam as many descriptions
+of lasers are simply combinations of plane waves (with varying amplitudes
+and wave vectors). More information on electromagnetic theory is covered in 
+chapter 2 but for now any mention of a focused laser can be thought of as a
+Gaussian beam, unless specified otherwise.
+
 \subsection{Literature related to laser induced nucleation}
 From as early as 1996 it has been known that laser irradiation is a viable
-method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, 
-the first reported case was notable as it used a 1.064 $\mu m$ laser meaning 
-it was unlikely to be a photo-chemical reaction but rather a physical one. 
-Future research has found nucleation can be induced by 1 of 3 routes involving 
-direct laser induction: firstly, Non-photochemical laser induced nucleation 
-(NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996,
-Garetz2002,Sun2006}, several papers have debated the exact mechanism that 
-induces NPLIN \cite{Garetz2002, Knott2011}. Two suggested hypothesis are: 
-an optical Kerr effect, where the solute molecules are aligned to lower the 
-nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in 
-which solute clusters are stabilised within the electric field which drastically 
-increases the likelihood of nucleation\cite{Alexander2008}. Both hypothesise 
-have their limitations and there has yet to be a single theory that explains 
-NLPIN thoroughly. In either case the mean pulse intensity needs to be kept 
-relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses 
-lead to a completely different nucleation mechanism.
+method of inducing nucleation within a supersaturated solution
+\cite{Garetz1996}. The first reported case was notable as it used a 
+1.064 $\mu m$ laser, the glycine solutions would appear transparent to
+such a laser which would suggest there was no photo-chemical reaction.
+Later studies into this phenomena found that the laser polarisation 
+can influence the polymorph produced. With circularly polarised light 
+producing $\alpha$-glycine and linearly polarised light forming
+$\gamma$-glycine \cite{Garetz2002}. Future research has found nucleation 
+can be induced by 1 of 3 routes.
 
+\subsubsection{Non-Photochemical Laser Induced Nucleation}
+Non-photochemical laser induced nucleation (NPLIN) involves irradiating 
+a solution with a pulsed laser \cite{Garetz1996,Garetz2002,Sun2006}. 
+The laser itself does not have to be heavily focused, instead irradiating
+a large region of the solution all at once. The choice of laser is of
+particular importance; with nucleation probability changing depending 
+on the wavelength \cite{Kacker2017}.
+In addition, the choice of solute will effect the setup, not only because some solute's are unaffected, but also because there is a minimum laser threshold before nucleation is observed
+\cite{Garetz2002}. Several papers have debated the exact mechanism that 
+induces NPLIN \cite{Garetz2002, Knott2011}. A suggested theory to this is 
+an optical Kerr effect, for anisotropically charged solute molecules the 
+electric field can reorient them to match the propagation direction 
+\cite{Garetz2002}. If enough molecules are co-aligned the free energy 
+barrier is reduced to allow for ambient nucleation \cite{Knott2011}. 
+An alternative theory is the dielectric polarisation effect, in 
+conditions that are unfavourable to cluster formation the polarising 
+effect can stabilise the clusters \cite{Alexander2008}. As the cluster  concentration rises so does the likelihood of nucleation \cite{Vekilov2010}. 
+Both theories are similar to one another but where the optical Kerr theory
+is limited to anisotropic solute molecules, the direct polarisation 
+theory is more flexible. Regardless both theories struggle to explain
+why the phenomena is not observed in all nucleation systems 
+\cite{Korede2023}, such as acetamide which is similar to urea which 
+does nucleate when irradiated \cite{Ward2016}. One of the benefits of 
+NPLIN is that since the pulses are relatively low in their intensity 
+they can be fired off quickly in succession, allowing for continuos 
+crystallisation set ups. Overall, the NPLIN phenomena needs further 
+research to properly describe its effects. The mean pulse intensity 
+needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), 
+as high intensity pulses lead to a completely different nucleation 
+mechanism.
+
+\subsubsection{High Intensity Laser Induced Nucleation}
 High intensity laser induced nucleation (HILIN), where the pulse intensity 
@@
 to a cavitation process within the target solution, where the laser focus 
-results in thermo-cavitation and the subsequent pressure change leads to
-a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}. 
-There is still not a general consensus on how cavitation influences the 
-local supersaturation \textcolor{red}{[ , ]}, nor is there a clear 
-understanding of what properties of the crystal can be controlled 
-\textcolor{red}{[ , ]}. It has been suggested that in theory any solution 
-can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a 
-strong understanding of the phenomena both before and after cavitation 
-occurs.
+results in thermo-cavitation and the subsequent pressure wave leads to
+a nucleation event around the focus of the laser \cite{Yoshikawa2005, 
+Soare2011, Barber2019}. What remains in question is both how the physical 
+properties (size, polymorph, etc) are influenced by the cavitation process, 
+and how the pressure change triggers nucleation. The former has already 
+been investigated; by adjusting the focal position Ikeda \textit{et al} 
+could control the polymorph of indomethacin \cite{Ikeda2015}, this is not 
+a universal method however, as it has also been shown that laser power can 
+influence the crystal polymorph \cite{Wang2010}. The latter is a tricky 
+task to address due to the fact that these cavitation bubbles form and collapse in less than $100\ \mu s$.
+Using fluorescence dyed proteins, researchers were able to observe a sudden spike in fluorescence just
+as the cavitation bubble began to collapse, they suggested that due 
+to the collapse of the cavitation bubble the protein clusters are brought together at the lasers focal point. However, while the fluorescence 
+imaging indicates a local concentration increase it is difficult to 
+quantify this change depending on the size of the bubble \cite{Korede2023}. It
+has been suggested that in theory any solution can undergo HILIN \cite{Korede2023}, but proving such a theory requires a clear 
+understanding of the phenomena both before and after cavitation occurs. Current research aims to combine experimental research with computer 
+simulations to develop a universal theory, with the hope that this could 
+also be related to NPLIN.
 
+\subsubsection{Trapping Induced Nucleation}
 Lastly, there is trapping induced nucleation, this is where optical tweezers 
-come into play, the optical trap has been shown to have different effects 
-on supersaturated solutions depending on where it has been focused. When 
-focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ 
-were shown to create a dense liquid droplet of glycine and water 
-\textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid 
-region was populated by clusters that would consolidate together upon being 
-focused by the optical trap \textcolor{red}{[ ]}. Molecular simulations of 
-glycine solutions showed that these clusters are unstable when using pure 
-glycine below the saturation point \textcolor{red}{[ ]} suggesting that the 
-clusters are formed due to glycine reaction products. When the optical trap 
-was moved from the cover slip to the air-solution interface nucleation would 
-occur before a dense liquid region could form \textcolor{red}{[ ]}. Repeated 
-experiments where the laser is focused on the air-solution interface have 
-lead to a variety of different nucleation events. In some instances the 
-nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; 
-whereas allowing a solution to age results in the formation of amorphous 
-precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}. 
-The precursors are only seen when the solution is irradiated by an optical tweezer 
-and the growth rate can be controlled somewhat by varying the laser power 
-\textcolor{red}{[ ]}. Notably the only work has been done with simply 
-irradiating the solution with a trapping potential, there has not been an 
-attempt to introduce a trappable object into the solution, the trapping 
-potential has been used to influence the growth of a crystal front 
-\textcolor{red}{[ ]}.
+come into play. Due to the radiation pressure created by the focused beam,
+it is possible to manipulate the solute, this was demonstrated with amino
+acids such as glycine \cite{Tsuboi2009}. Whether or not a crystal forms is 
+due to the location of the laser focus.
+When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \cite{Yuyama2010, Yuyama2012}.
+The dipole moment of the glycine molecules is too small to be influenced by the optical trap, as such it would suggest that larger aggregates are being manipulated.
+Applying DLS analysis to the dense liquid region showed that it was populated by clusters that would consolidate together upon being focused by the
+optical trap \cite{Gowayed2021}.
+Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point suggesting that the clusters are formed due to glycine reaction products \cite{Sweatman2022}.
+When the optical trap is moved from the cover slip to the air-solution interface, nucleation would occur before a dense liquid region could form \cite{Yuyama2010}.
+Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
+In some instances the nucleation occurs spontaneously after a short period of time \cite{Yuyama2010}.
+Whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \cite{Liao2022}.
+The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \cite{Liao2022}.
+The reason why nucleation is only seen at the air-solution interface is due to the limited molecular mobility close to the interface.
+Often tweezing experiments will use a hydrophilic
+coating to minimise the height of the solution droplet and further limit
+the molecular mobility \cite{Yuyama2012, Gowayed2021}.
+
+Walton and Wynne discussed a plausible model for how the tweezer focus
+could result in a nucleation event. Put simply, when the laser is focused
+at the solution the radiation pressure draws in solute material, creating 
+a concentrated region of solute. This also creates a depleted region 
+around the focus and raises the local temperature. When the laser is 
+turned off the depleted region around the focus quickly cools back to the 
+ambient temperature. This sudden cooling allows for nucleation to occur
+just outside the focus.
+
+Laser induced nucleation has the potential to be a viable method for 
+\textit{in-situ} studying of nucleation events. Using high numerical 
+aperture lenses one can localise the nucleation event to a specific 
+region of the solution. The current issue is that the mechanism behind
+laser induced nucleation is not fully understood, as such it is rather
+difficult to modify the laser for different solution parameters. Instead
+it may be more effective to manipulate the solution using trapped particles.
+An common method is by apply a torque to a trapped particle, creating 
+a micro-rotor that can can generate fluid flow.
 
 \subsection{Optical Torque}
-For any electromagnetic field it is possible to transfer both linear and 
+It has long been known that electromagnetic fields can transfer linear and
 angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
@@
 debate on how to decompose the total momentum into these two components 
-\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
+\cite{Bruce2020, Svak2018}, for this project we do not need to calculate 
 the exact quantities and will instead look at the broader effects of both 
@@
 loops that all together cancel one-another out when the medium is homogeneous.
-Spacial inhomogeneities cause these spin loops to non longer be equal, resulting
+Spacial inhomogeneities cause these spin loops to no longer be equal, resulting
 in non-zero spin density, anisotropic mediums (such as birefringent crystal 
-lattices) experience a transfer of spin angular momentum allowing for optical
-torque transfer. It has even been shown that in the case of homogeneous spheres, 
-the spin component of a focused beam can result in additional orbital angular
-momentum via spin-to-orbital conversion \cite{Yevick2017}.
+lattices) experience a transfer of spin angular momentum, imparting an optical
+torque.
 
 Birefringence is a material property often seen in crystalline materials, 
-if the crystal lattice has a different structures when viewed at different 
-orientations then light will be refracted differently depending on its 
-polarisation. For circularly polarised light the inhomogeneity results in 
-a high degree of SAM being transferred to the target object \cite{Parkin2009,
-Arita2016}, this has been exploited to rotate microspheres as fast as 1000 Hz 
-while suspended in a bulk medium \cite{Arita2016} as well as a means of 
-measuring the local temperature and shear response of said medium 
-\cite{Millen2014, RodriguezSevilla2018}. Calculating the optical torque 
-applied to a birefringent material is given via:
-
+if the crystal lattice has different refractive indices for varying 
+polarisations of light. For circularly polarised light this inhomogeneity 
+results in a high degree of SAM being transferred to the target object \cite{Parkin2009, Arita2016}, this has been exploited to rotate microspheres 
+as fast as 1000 Hz while suspended in a bulk medium \cite{Arita2016} as well 
+as a means of measuring the local temperature and shear response of said 
+medium \cite{Millen2014, RodriguezSevilla2018}. Calculating the optical 
+torque applied to a birefringent material is given via:
 \begin{equation}
@@
 	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
+		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi 
+		\\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
 	\end{aligned}
@@
 
-Where $\theta$ is the angle between the particle's long axis and the Poynting
-vector of the local EM field, and $\phi$ is the phase shift in the EM field.  
-The first term represents the 'orientational' torque which is due to the target 
-particle being aligned along the EM field, when aligned $\theta=0$ meaning 
-the entire term is negligible for particle's with a stable orientation. The 
-second term is due purely to the polarisation of the optical trap, for 
-circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred 
-to the target particle. Birefringence can also be induced if the target 
-particle has an anisotropic shape, in particular if the particle shape is 
-elongated along one major axis; the go to particle shape is a silica dimer 
-(two spheres tangentially attached) due to silica's stability and strong 
-adhesion. Using a silica dimer research groups have achieved a rotation 
-frequency in the realm of several GHz \cite{Ahn2018} in a vacuum.
+Where $\theta$ is the angle between the particle's long axis and the 
+polarisation vector of the local EM field, and $\phi$ is the phase shift 
+in the EM field. The first term represents the 'orientational' torque which 
+is due to the target particle being aligned with the EM amplitude, when 
+aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. For sphere's this is always the case as they lack a 
+distinctive major axis. The second term is due purely to the polarisation 
+of the optical trap, for circularly polarised light $\phi=\pi/4$ thus 
+maximising the torque transferred to the target particle. Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
+birefringence, but there are other mechanisms that result in optical torque.
+
+A common example is shape induced birefringence. If a particle has an 
+anisotropic shape, it is more susceptible to being polarised along its 
+longer axis than its shorter axis. Consider, for example, an ellipsoid 
+elongated along one of its primary axis' ($r_z > r_x = r_y$). In a 
+plane polarised beam such a particle will align with the polarisation
+vector. Therefore, the particle will rotate as angular momentum is 
+transferred along its long axis. The common feature for shape induced
+birefringence is that all of these particles are aligned in the plane
+of polarisation. Often the optical torque experienced is far greater 
+than similar spherical particles that are birefringent. Currently 
+spherical dimers are being rotated in vacuums to measure quantum forces
+and torques \cite{Ahn2018, Reimann2018}. There are some cases alternative 
+cases where particles are rotated while not being aligned in the plane 
+of the polarisation. However in these cases their shape is often 
+specifically engineered to scatter light either clockwise or 
+counter-clockwise \cite{Higurashi1994}.
+
+Other examples of optical torque is when an anisotropic particle is 
+aligned with the beam's direction of propagation (in which case $\theta
+=\pi/2$ and the first term disappears). This is analogous to an optically trapped sphere, where alongside a restoring force the particle also 
+experiences a restoring torque. This seemingly random rotational motion
+is referred to as libation \cite{Bruce2020}, often in typical suspension
+trapping situations (where the particle is suspended in a fluid) the 
+translational motion is washed out by the translational motion. As such, 
+many experiments elect to trap in low pressure environments to precisely 
+measure the optical torque being exerted by the optical trap \cite{Ahn2018}.
+This has lead to experiments to try and achieve '0 kelvin' motion, where
+by trapping a silica dimer they were able to restrict its motion using 
+3 optical traps simultaneously. Despite this, they found that the dimer's
+rotational motion about its long axis could not be controlled leading to
+the undesired rotational modes \cite{Bang2020}.
+
+The detection and measurement of optical torque is still a field of intense
+research, not only does it have potential to understand quantum fluctuations
+in a particle's motion but also allows for the creation of more effective
+micro-rotors. The latter being especially pertinent for understanding the 
+behaviour of fluids experiencing localised shearing.
 
@@
 nucleation; however, the exact relationship between shear rate and nucleation
-rate has only been recently understood for specific solutions. Theoretical research 
-into shear induced nucleation suggests that there should be a slight increase 
-in the nucleation rate at low shear rates, reaching a maximum increase in 
-nucleation rate, and then at higher shear rates the nucleation rate begins 
-to drop off. This has been shown theoretically for both simple colloidal
-\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
-\cite{Goswami2020}; however, no experimental work into these systems has 
-been conducted to prove this is the case.  There is some experimental 
-evidence for this phenomena in simple salt and protein solutions - though 
-the authors emphasise that mechanical agitation cannot be ruled out - 
-there has not been a exhaustive study into the shearing effects apart from 
-in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
-rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
-the highest nucleation rate. Using the theoretical model established in 
-\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
-a nucleus undergoing shearing, accounting for the fact that a nucleus' 
-growth is undergoing competition between flow-mediated molecular transport 
-and the strain applied by the flow field which inhibits the growth of the 
-nucleus. There central conclusion (from both the theoretical and experimental 
-results) is that there is an optimal shear rate in which the nucleation 
-rate is maximised. However, a question that arises from this result, if 
-there is a optimal shear rate in which molecular transport is maximised 
-and strain is minimised, then surely there should also be a shear rate 
-in which the molecular transport and strain are equal - allowing one to 
-suspend a nucleus at a constant radius. In this scenario, the molecular 
-transport would prevent the nucleus from dissolving, but the strain would 
-prevent the nucleus from growing. This however would require one to be 
-able to apply a continuous shear rate to a targeted nucleus with high 
-precision, there is also no model for an individual nucleus undergoing growth.
+rate has only been recently understood for specific solutions. Theoretical research into shear induced nucleation suggests that there should be a 
+slight increase in the nucleation rate at low shear rates, reaching a 
+maximum increase in nucleation rate, and then at higher shear rates the 
+nucleation rate begins to drop off.
 
-\section{Aims and Objectives}
+This has been shown theoretically for both simple colloidal \cite{Mura2016,
+Debuysschere2023,Richard2015} and ice crystal formation \cite{Goswami2020}; however, no experimental work into these systems has been conducted to prove 
+this is the case. There is some experimental evidence for this phenomena in 
+simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
+In \cite{Debuysschere2023} it was found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
+Using the theoretical model established in \cite{Mura2016,2001} which modifies the CNT to account for the effects of a nucleus undergoing shearing,
+accounting for the fact that a nucleus' growth is undergoing competition 
+between flow-mediated molecular transport and the strain applied by the 
+flow field which inhibits the growth of the nucleus. There central 
+conclusion (from both the theoretical and experimental results) is that 
+there is an optimal shear rate in which the nucleation rate is maximised.
+
+However, a question that arises from this result, if there is a optimal 
+shear rate in which molecular transport is maximised and strain is 
+minimised, then surely there should also be a shear rate in which the 
+molecular transport and strain are equal - allowing one to suspend a 
+nucleus at a constant radius. In this scenario, the molecular transport 
+would prevent the nucleus from dissolving, but the strain would prevent 
+the nucleus from growing. This however would require one to be able 
+to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus in a continuous fluid 
+field.
+
+\section{Significance of Thesis}
+As I have hoped to make clear in the above introduction, the current 
+state nucleation theory is rather cumbersome at a micro-level. Models
+such as CNT and multi-step nucleation are not sufficient for describing
+the myriad of potential pathways nucleation can go down. As suggested
+by some review articles, the best way to address this is by developing 
+\textit{in-situ} methods that can study the pre nucleation phase in 
+greater detail \cite{Fu2021}. Furthermore, the ability to localise 
+nucleation allows for better characterisation of the kinetics of crystal 
+growth. Laser induced nucleation stands to be an ideal method to study
+nucleation, as the laser output can be concentrated to a small area. 
+In this project we want to consider another application of optical 
+tweezers, their ability to rotate micro-rotors to generate fluid flow.
+Such a method has not been tried before and would allow for us to 
+consider how fluid flow impacts nucleation dynamics at the micro-level.
+
+\section{Overview}
 Overall the aim of the PhD is to study viability of using micro-rotors to
 generate localised fluid flow around the beam focus. The results are reported
-in chapter 3, this is then succeeded by experimental work where we 
-instead using a galvano-mirror to generate shear flow. While overall unsuccessful
-the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus.
+in chapter 3, this is then succeeded by experimental work where we instead 
+using a galvano-mirror to generate shear flow. While overall unsuccessful
+the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus. This presents a new insights for controlling
+and studying the growth of a newly formed nucleus by precise movement of the
+trapping focus.
 
@@
 and their applications as complex micro-rotors are discussed in chapter 4. 
+The latter results are interesting as a potential insight into how optical
+torque is transferred to anisotropic objects.
+
 Lastly two novel methods of measuring rotational motion are discussed in 

### 2024-09-06 Latex_files/Thesis/chapters/chapter 2/main.tex
updates 06/09/2024

@@
 gradient force far exceeds the both the scattering force, and the thermal 
-energy of the particle. Describing and calculating the optical forces is 
-dependent on the ratio between the particle size and the trapping wavelength.
-
+energy of the particle.
+
+When a sphere is located at the centre of a laser the gradient force will
+try to recentre the sphere if it is displaced. If the restoring force is
+proportional to the sphere's displacement this is referred to as a harmonic
+trap. The ratio of the restoring force to the displacement is called the 
+trap stiffness (denoted by $\kappa$), and is an important characteristic 
+of the trap. Shown below is a typical force displacement curve for a single
+sphere in a focused Gaussian beam.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.78\linewidth]{sphere_force_curve.png}
+	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
+			while moving along the x-axis, where $x=0$ is the centre of the
+			beam. The harmonic regime is highlight by the black linear fit.}
+	\label{fig:harmonic_trap}
+\end{figure}
+
+The strength of the optical trap is directly proportional to both the laser
+power and the contrast in refractive indices between the medium and the
+particle. Fully describing and calculating the optical forces is dependent on 
+the ratio between the particle size and the trapping wavelength. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Using a  MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007} we can compute the beam shape coefficients ($a_{nm}$ \& $b_{nm}$)
-for any desired beam type. Using the results from \cite{Farsund1996} we can then 
-compute both the optical force and torque using the beam coefficients and the 
-scattering coefficients ($q_{nm}$ \& $p_{nm}$) which are found by calculating
-the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The form given by \cite{Crichton2000THEMD} is an easier 
-form to compute for higher orders of $N_{max}$:
+Using a  MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007} we can compute the beam shape coefficients ($a_{nm}$ \& 
+$b_{nm}$) for any desired beam type. Using the results from \cite{Farsund1996} 
+we can then compute both the optical force and torque using the beam coefficients 
+and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) which are found by 
+calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and 
+then using \eqref{eq:t_matrix}. The form given by \cite{Crichton2000THEMD} is 
+an easier form to compute for higher orders of $N_{max}$:
 \begin{equation}
@@
 
-Vigilante et al compiled together a python package that combines both \textit{ott} 
-and \textit{MSTM} to simulate the behaviour of spherical aggregates within 
-a predefined optical trap \cite{Vigilante2020}.
+Vigilante et al compiled together a python package that combines both 
+\textit{ott} and \textit{mstm} to simulate the behaviour of spherical 
+aggregates within a predefined optical trap \cite{Vigilante2020}. Now 
+\textit{ott} does have inbuilt $T$-matrix modules that can be used to 
+model the scattering from different shaped particles; however, for 
+aggregates of spheres the resultant force is nonsensical (see Chapter 
+4 for further details). As such we rely on the \textit{mstm} for 
+simulating dimers.
 
 \subsection{Simulation parameters}
+\label{sec:sim_parameters}
 Throughout this PhD we use the work of Vigilante to perform a systematic 
@@
 transformed into rotation matrices. The matrix describes the direction of the 
-particles 3 orientation axis $U_x$, $U_y$, and $U_z$; the latter is oriented 
-so that it is collinear with the centres of the two spheres. The origin of 
-the 3 axis' is set on the particle's centre of mass, this is also used to 
-define it's displacement from the origin of the laboratory frame.
-
-Furthermore we expand upon their work to create a fully functional quadrant 
-photo-diode to simulate the response from a calibration test. This builds 
-upon the work from \cite{Rohrbach2002} which applied the fundamental 
-principles of Lorenz-Mie theory to replicate the response signal of a QPD 
-being used in back focal-plane interferometry.
-
+particles 3 orientation axis $U_x$, $U_y$, and $U_z$. The latter is oriented 
+so that it is collinear with the centres of the two spheres; the direction of
+vector $U_z$ is such that it goes from the centre of $a_{II}$ to the centre of
+$a_{I}$. The origin of the 3 axis' is set on the particle's centre of mass, 
+this is also used to define it's displacement from the origin of the laboratory 
+frame. The diffusion tensor of the dimer is more complex than a sphere's as we 
+must consider the rotational effects experienced by the dimer. We use the 
+analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided the 
+solution coefficients depending on the dimer's size ratio ($a_I/a_{II}$). Rather 
+than recalculate the coefficients for every possible configuration of dimer we 
+use a spline fitting function to approximate the diffusion tensor. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:simulated_QPD}
+
+Beyond merely simulating the forces experienced by a spherical aggregate we
+also developed a simulative quadrant photo diode to replicate the results
+from a typical calibration test. This builds upon the work from \cite{Rohrbach2002} which applied Lorenz-Mie theory to replicate the response signal of a QPD 
+being used in back focal-plane interferometry. However, rather than be 
+constrained to individual spheres we can now consider the expected response 
+from any type of spherical aggregate.
+
 In order to simulate a typical experimental set up with a QPD installed 

### 2024-09-06 Latex_files/Thesis/chapters/chapter 4/main.tex
updates 06/09/2024

@@
 With the initial goal of the PhD being to induce nucleation events via a 
-spherical micro-rotor the goal of this chapter was to - in a limited capacity
+spherical micro-rotor the goal of this chapter is to - in a limited capacity
 simulate and investigate the influence of a second particle being bound to our 
@@
 might better represent a growing crystalline solid, allows us to consider 
-how the dynamics of the aggregate change by varying the size factor. We build 
+how the dynamics of the aggregate change by varying the size factor. 
+Additionally, attempting to simulate an amorphous aggregate is rather difficult 
+as calculating the optical force and torque is computationally slow. We build 
 upon the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
@@
 additionally makes characterising their interactions within an optical trap 
-more cumbersome. Attempting to simulate an amorphous aggregate is rather
-difficult as calculating the optical force and torque is computationally 
-slow and orientation specific.
+more cumbersome.
+
+In this chapter we will consider how the addition of a second sphere changes 
+the trapping dynamics by introducing multiple equilibrium positions. These 
+positions are dependent on both the size ratio of the dimer and its orientation
+relative to the trap. This is important to address, for while the the dimers' 
+diffusive behaviour is the same, regardless of position and orientation, how it
+interacts with the trap may produce the wildly different results. Furthermore,
+we look at how our dimers interact with circularly polarised light, this is 
+especially pertinent as the original plan was to utilise circularly polarised 
+light to generate fluid flow in a supersaturated solution. The rotational motion 
+of dimers in circularly polarised light goes against much of the established 
+literature and suggests that the transfer of spin angular momentum is enabled 
+by combining spheres together. And lastly, we demonstrate how a quadrant photo 
+diode, performs in characterising the actual interactions between a dimer and 
+the optical trap. We do so to demonstrate that much of the dimer's trajectory
+information is either lost or poorly described by simple calibration techniques, 
+making their motion difficult to characterise.
+
+\subsection{Simulation Parameters}
+As a paradigmatic example, consider a dimer suspended in water ($n_p = 1.59, 
+n_m = 1.33$) located at the focus of a Gaussian beam (more specifically a 
+Laguerre-Gaussian beam of mode $[0,0]$), the beam is focused by a objective 
+with numerical aperture of 1.2 and is x polarised. The size ratio of the two 
+sphere's is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ 
+unless specified otherwise; the dimer's orientation is given by a unit vector 
+connecting the centres of both spheres (see \ref{sec:sim_parameters}). We 
+define the 'standard' orientation as being aligned with the direction of 
+beam propagation direction - and therefore the 'inverted' orientation is 
+defined when the dimer is orientated against the direction of beam 
+orientation (see left hand side of figure~\ref{fig:paradigmatic}).
 
@@
 forces, once there the force is analogous to a harmonic spring with a fixed trap
-stiffness. Now, if we consider instead a dimer, we now have two spheres both being
-drawn to the focus along by the same gradient force; in addition the scattering force
-is significantly more complex due to both spheres scattering the electromagnetic 
-fields. This mutual scattering between individual spheres is what makes simulating
-spherical aggregates far more difficult compared to a single sphere, and even harder
-still to predict the position where the dimer's centre of mass is in equilibrium.
+stiffness (see fig.\ref{fig:harmonic_trap}). Now, if we consider instead a dimer, 
+we now have two spheres both being drawn to the focus along by the same gradient 
+force; in addition the scattering force is significantly more complex due to both 
+spheres scattering the electromagnetic fields. This mutual scattering between individual 
+spheres is what makes simulating spherical aggregates far more difficult compared
+to a single sphere, and even harder still to predict the position where the dimer's 
+centre of mass is in equilibrium.
 
@@
 be assumed to behave as a simple harmonic trap. The methodology for computing 
-optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the 
+optical forces has been covered extensively for a number of different trapping 
+conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the 
 trapping force and determine where a simple sphere would be located relative 
@@
 minimised.
-
-\paragraph{Simulation Parameters}
-As a paradigmatic example, consider a dimer suspended in water ($n_p = 1.59, 
-n_m = 1.33$) located at the focus of a Gaussian beam (more specifically a Laguerre-Gaussian beam of mode $[0,0]$), the beam is focused by a objective 
-with numerical aperture of 1.2 and is x polarised. The size ratio of the two 
-sphere's is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ 
-unless specified otherwise; the dimer's orientation is given by a unit vector connecting the centres of both spheres, we define the 'standard' orientation 
-as being aligned with the direction of beam propagation direction - and 
-therefore the 'inverted' orientation is defined when the dimer is orientated 
-against the direction of beam orientation (see left hand side of figure~\ref{fig:paradigmatic}).
 
@@
 
-We repeated the same calculation but now while the dimer is in its 'inverted' orientation, instead of a single point where the optical force is minimised we 
+We repeated the same calculation but now while the dimer is in its 'inverted' 
+orientation, instead of a single point where the optical force is minimised we 
 see that there are instead two separate equilibrium positions, one above the focus
@@
 if multiple spheres are trapped simultaneously. 
-
 \begin{figure}[h!]
@@
 field is relatively simple as the orientational torque is minimised (see
-Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change
-orientation enough to escape the trap. However, that does not rule out the 
+Eq.\ref{eq:opt_torque}). Meaning once trapped the dimer is unlikely to change
+orientation enough to escape the trap. Regardless, that does not rule out the 
 possibility that there is a stable configuration where the orientation not 
@@
 depending on which equilibrium configuration is reached.
-
 \begin{figure}[h!]
@@
 indicating that the trapping potential is far greater than the thermal 
-energy. This suggests that the reason this off-axis configuration is due 
-to the rotational motion more than translational motion. If the overall 
-potential depth could be characterised then dimers placed into this 
-orientation could be used as a micro-scale temperature alarm, where 
-by fine tuning of the dimer's parameters would allow you to construct a 
-potential well that can only be escaped when the local fluid temperature 
-exceeds a certain maximum value.
+energy. This suggests that the reason this off-axis configuration is 
+metastable is due to the increased rotational freedom. Similar 
+configurations have been explored with ellipsoids; Zhu \textit{et al.} 
+looked at the dynamics of various elliptical particles and found that 
+regardless of shape or initial orientation the particle would tend towards
+either a purely vertical or purely horizontal orientation \cite{Zhu2021}. 
+It highlights an interesting aspect of the dimer's behaviour, namely that 
+the potential phase space $U(\textbf{r}, \theta, \phi)$ is not continuous.
 
@@
 \section{Continuous rotational motion due to second-order scattering}
-
 One aspect that has yet to be covered in depth with regards to spherical 
@@
 experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
-while trapped in a circular polarised beam.
+while trapped in a circular polarised beam. In our simulations we found 
+that dimers would rotate about their long axis when trapped in circularly
+polarised light. In this section we want to discuss how this behaviour 
+is influenced by size, position, and orientation; and furthermore, we wish
+to address possible explanations for this behaviour, as none of the current
+theories into optically induced rotation seem plausible.
 
+\subsection{Polarisation Dependency on Dimer trajectory}
 First we wanted to confirm that the rotation seen in our simulations is
@@
 		sized dimers. The solid lines represent the rotation rate experienced 
-		while the dimer is in its standard orientation, whereas the solid points 
-		are for the case where the orientation is inverted. Laser power $= 100\ mW$.}
+		while the dimer is in its standard orientation, whereas the solid points are for the case where the orientation is inverted. Laser power $= 100\ mW$.}
 	\label{fig:rotation_vs_pol}
@@
 showed that the seemingly random trajectory of a trapped sphere was biased
-by the polarisation state of the trapping beam. While not immediately evident 
-from the trajectory the helicity of the trapping beam was revealed by 
-computing the particle's probability flux using.
+by the polarisation state of the trapping beam. While not immediately 
+evident from the trajectory the helicity of the trapping beam was revealed 
+by computing the particle's probability flux using.
 \begin{align}
@@
 where the spin-curl force is more substantial due to it having a wider profile. 
-This would suggest that particles' whose longer axis lies in the plane perpendicular
-to the direction of propagation should experience any notable degree of 
-torque. And indeed this appears to be the case for a number of experiments 
-involving nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
+This would suggest that particles' whose longer axis lies in the plane perpendicular to the direction of propagation should experience any notable degree of torque.
+And indeed this appears to be the case for a number of experiments involving nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
 
+\subsection{Comparison of optical torques for Sphere and Dimers}
 It does raise the question of how come our simulations show that despite the 
@@
 figure~\ref{fig:rotation_vs_pol} but with a circularly polarised beam. Both 
-the sphere and dimer were rotated in the $x-z$ plane and the all three components 
-of the optical torque were recorded.
+the sphere and dimer were rotated in the $x-z$ plane and the all three components of the optical torque were recorded.
 \begin{figure}[h!]
@@
 
-It should be noted that the difference in magnitude is not clearly shown in 
-figure~\ref{fig:sphere_dimer_torque}. In truth, the optical torque acting on 
-the dimer $10^{13}$ times greater than the torque applied to the single sphere. 
 The torques about the $x$ and $y$ axis can be somewhat understood as the second 
 sphere is being drawn back towards the centre of the trap by the gradient forces.
-The same cannot be said for the $z$component of the optical torque; while this 
-effect could still be attributed to the spin-curl force, but it is clear that the internal scattering between the two spheres has some unintended effects.
+The same cannot be said for the $z$component of the optical torque; while this effect could still be attributed to the spin-curl force, but it is clear that the internal scattering between the two spheres has some unintended effects.
 
- By repeating
-the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
-a circularly polarised beam $\phi=90\circ$ it was found that not only is 
-the rotation rate dependent on the size of the dimer, but also on its 
-orientation and therefore their axial position.
+\subsection{Rotational frequency as a function of size ratio and orientation}
+Moving on, next we want to see how the rotational speed is influenced by 
+different sized dimers. Intuitively, you would expect that a larger would 
+experience a greater torque and therefore rotate faster.
+By repeating the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for a circularly polarised beam $\phi=90\circ$ it was found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
 \begin{figure}[h!]
@@
 in the dimer that might impart angular momentum to the scattered beam - 
-as compared to a anisotropic scatterer like vaterite. Therefore the 
-rotational motion must be due to
+as compared to a anisotropic scatterer like vaterite.
 
@@
 makes it clear that the polarizability is a contributing factor to this optical 
-rotation phenomena. Rotating a symmetric dimer in the $x-z$ plane reveals 
-that while the dimer can be rotated in an orientation perpendicular 
-to the beam rotational torque is maximised when rotated while aligned with 
-the optical axis.
-\begin{SCfigure}
-	\includegraphics[width=0.65\linewidth,height= 15cm]{torque_different_polarisations.png}
-	\caption{Optical torque against polar angle $\theta$ about the three primary axis (top: torque about the x-axis; middle: torque about the y-axis; bottom: torque about the z-axis) on a symmetric dimer in linear, elliptical, and circular polarisation beams.
-Diagram to the right is for visual clarity about the direction of $\theta$.\vspace{4.5cm}}
-\end{SCfigure}
-\newpage
+rotation phenomena.
 
@@
 is the potential for non-vertical trapping orientations in which the dimer is 
-still stably trapped. When a circularly polarised beam is used the dimer exhibits gyroscopic precession. As shown in fig~\ref{fig:gyro} the dimer's trajectory
-exhibits periodic rotation about its long axis $U_z$, precession motion as the
-dimer spins around the central beam axis, and nutation motion as the dimer's
-orientation rocks backwards and forwards. 
+still stably trapped.
+When a circularly polarised beam is used the dimer exhibits gyroscopic precession.
+As shown in fig~\ref{fig:gyro} the dimer's trajectory exhibits periodic rotation about its long axis $U_z$, precession motion as the dimer spins around the central beam axis, and nutation motion as the dimer's orientation rocks backwards and forwards.
 \begin{figure}[h]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}

### 2024-09-16 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the work being done with 
-optical tweezers in particle classification, and how tweezing has been 
-used to investigate nucleation events.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Working Principle}
-
-Optical tweezing operates on the principle that light carries both
-linear and angular momentum, which can be transferred to other
-objects. If light is reflected or refracted by a medium part of the
-light's momentum is transferred to the medium itself. While at large
-scales this effect is insignificant in most cases, at micron scales
-the forces imparted by individual photons begin to become significant
-enough to influence the trajectory of individual particles. Ashkin's 
-work into optical tweezing found that micron sized entities could be 
-spatially locked by aiming a laser with a Gaussian intensity profile 
-directly upward at a particle while being suspended in a liquid 
-medium \cite{Ashkin1970}.
-
-This is demonstrated in fig.~\ref{fig:working_principle} where we have 
-a simple sphere trapped by a circularly symmetrical beam with the intensity
-falling off from the centre in a normal distribution (commonly referred to
-as a Gaussian beam). If we consider two rays of light hitting our sphere 
-we see that while in the centre each ray is refracted in an equal opposite 
-direction to one another and so the net force imparted is 0. When displaced 
-in the x-y plane the net force is unbalanced and points towards the centre 
-of the trap. This principle can be generalised to 3 degrees of freedom if we 
-consider a focused beam instead of a para-axial beam, typically the trap will be
-weaker along the beam axis compared to the transverse trap strength.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.65\linewidth]{ot_working_principal.png}
-  \caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards.
-Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
-  \label{fig:working_principle}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+working principles behind optical tweezing, the differences in 
+scattering theories, and the broader theoretical calculations used
+throughout the rest of the Thesis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Electromagnetism and optical tweezing}
@@
 \begin{align}
-	\nonumber
 	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
 	\\ 
-	\nonumber
 	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
@@
 gradient and scattering components, for most modelling research this 
-is how the force fields are reported. The gradient force is a conservative 
-force brought about by the polarisation of dielectric materials which 
-is directed towards the point of maximum intensity (for a simple Gaussian 
-beam this would be at it's focal point) \cite{YasuhiroHarada1996}. 
-Typically, optical tweezers will utilise higher numerical aperture lenses 
-in order to increase the intensity, thus allowing for stronger 
-gradient forces.
-
-The scattering force arises from the scattered field of the trapping beam 
-pushing the target object away from the focal point. The equilibrium position 
-(where the mean squared displacement is minimised) is found when the 
-gradient force far exceeds the both the scattering force, and the thermal 
-energy of the particle.
-
-When a sphere is located at the centre of a laser the gradient force will
-try to recentre the sphere if it is displaced. If the restoring force is
-proportional to the sphere's displacement this is referred to as a harmonic
-trap. The ratio of the restoring force to the displacement is called the 
-trap stiffness (denoted by $\kappa$), and is an important characteristic 
-of the trap. Shown below is a typical force displacement curve for a single
-sphere in a focused Gaussian beam.
-
+is how the force fields are reported. The gradient force is a 
+conservative force brought about by the polarisation of dielectric 
+materials which is directed towards the point of maximum intensity 
+(for a simple Gaussian beam this would be at it's focal point) \cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
+higher numerical aperture lenses in order to increase the intensity, 
+thus allowing for stronger gradient forces at the trade off of 
+decreased trapping depth.
+
+The scattering force arises from the scattered field of the trapping 
+beam pushing the target object away from the focal point. While some 
+trapping arrangements can induce transverse scattering forces for 
+simple spheres the scattering force is only significant in the direction
+of propagation \cite{Capitanio2002} The equilibrium position is found 
+when the gradient force far exceeds the both the scattering force, 
+and the thermal energy of the particle.
+
+\subsubsection{Harmonic Traps}
+\label{sec:harmonic_traps}
+When a sphere is located at the centre of a laser the gradient force
+will try to recentre the sphere if it is displaced. If the restoring 
+force is proportional to the sphere's displacement this is referred 
+to as a harmonic trap. The ratio of the restoring force to the 
+displacement is called the trap stiffness (denoted by $\kappa$), and 
+is an important characteristic of the trap. Shown below is a typical 
+force displacement curve for a single sphere in a focused Gaussian beam. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.78\linewidth]{sphere_force_curve.png}
+	\includegraphics[width=\linewidth]{sphere_force_curve.png}
 	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
@@
 
-The strength of the optical trap is directly proportional to both the laser
-power and the contrast in refractive indices between the medium and the
-particle. Fully describing and calculating the optical forces is dependent on 
-the ratio between the particle size and the trapping wavelength. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The strength of the optical trap is directly proportional to both 
+the laser power and the contrast in refractive indices between the 
+medium and the particle. Fully describing and calculating the 
+optical forces is dependent on the ratio between the particle size 
+and the trapping wavelength. Depending on this ratio, the optical 
+forces can be described differently.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Ray-Optics Regime}
@@
 the very edge of the beam. Consider a particle struck by two rays in 
-a Gaussian beam (see Fig below), one coming close to the centre, and 
-the other ray coming from the edge. As each ray is refracted by the 
-target sphere a force is imparted onto it, the total force imparted 
-is given by:
+a Gaussian beam, one coming close to the centre, and the other ray 
+coming from the edge. As each ray is refracted by the target sphere a 
+force is imparted onto it, the total force imparted is given by:
 \begin{align}
@@
 
-The Lorenz-Mie theory provides an exact solution to the Maxwell equations 
-for the scattering caused by an isotropic sphere. This theory describes the 
-scattered wave given off by a dielectric sphere when incident by a plane 
-wave as a summation of partial spherical waves. For any spherical wave the 
-vector fields must solve the Helmholtz wave equation given by:
+The Lorenz-Mie theory provides an exact solution to the Maxwell 
+equations for the scattering caused by an isotropic sphere. This 
+theory describes the scattered wave given off by a dielectric 
+sphere when incident by a plane wave as a summation of partial 
+spherical waves. For any spherical wave the vector fields must 
+solve the Helmholtz wave equation given by:
 \begin{align}
@@
 be found by computing the Maxwell stress tensor which only requires
-the total magnitude of the incident and scattered fields. By computing 
-the momentum contained in the incident and scattered fields we can measure
-the total momentum transferred to the target particle.
-
-Lorenz-Mie theory can be applied to describe the scattering from any 
-particle regardless of size, though as the size of the target particle 
-changes - relative to the light being scattered - the theory describing
-the scattered light is better approximated by alternative theories. For 
-particles that are far larger than the wavelength of light the scattering
-is best described by the Ray-Optics theory. In the opposite case the 
-target particle is better approximated as a single electric dipole as 
-described by the Rayleigh scattering field.
+the total magnitude of the incident and scattered fields. An alternative
+method is to compute the momentum carried by the incident and scattered
+beams and taking the difference between to measure the momentum 
+transferred to the particle. This project utilises the results from 
+Farsund \cite{Farsund1996} but in the form used by Crichton 
+\cite{Crichton2000THEMD}. As an example the linear momentum carried 
+by any beam can be computed via:
+\begin{equation}
+	\label{eq:linear_momentum_in}
+	\begin{split}
+		\langle P_{z,\ in} \rangle
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+And the linear momentum carried by the scattered beam is given by:
+\begin{equation}
+	\label{eq:linear_momentum_out}
+	\begin{split}
+		\langle P_{z,\ out} \rangle
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+Therefore the total force imparted to the trapped particle is thus:
+\begin{align}
+	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
+\end{align}
+
+Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
+coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
+Similar expressions are shown in \ref{sec:simulations} for 
+calculating the optical torque. Lorenz-Mie theory can be applied 
+to describe the scattering from any particle regardless of size, 
+though as the size of the target particle changes - relative to 
+the light being scattered - the theory describing the scattered 
+light is better approximated by alternative theories. For 
+particles that are far larger than the wavelength of light the 
+scattering is best described by the Ray-Optics theory. In the 
+opposite case the target particle is better approximated as a 
+single electric dipole as described by the Rayleigh scattering 
+field.
 
@@
 
-The Rayleigh approximation is the limiting approximation for describing 
-a particles motion in an electromagnetic field whose wavelength is several 
-times greater than the particle's size. The underlying theory is that a 
-dielectric sphere can be treated as a dipole while in the presence of the 
-electromagnetic field; in which case the scattering force is given simply 
-by the scattering of the induced dipole, and the gradient force is due 
+The Rayleigh approximation is the limiting approximation for 
+describing a particles scattering in an electromagnetic field 
+whose wavelength is several times greater than the particle's 
+size. The underlying theory is that a dielectric sphere can be 
+treated as a dipole while in the presence of the electromagnetic 
+field. In which case the scattering force is given simply by the 
+scattering of the induced dipole, and the gradient force is due 
 to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
-principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
-in MKS units as a restorative rectangular force field:
+principle Cartesian axis are described by Harada et al 
+\cite{YasuhiroHarada1996} in MKS units as a restorative rectangular 
+force field:
 \begin{align}
@@
 scattering forces due to the assumption that the particle is a point 
-dipole. So much so that higher order scattering problems can be 
-simplified by subdividing the particle into discrete dipoles (see 
-Sec~\ref{sec:scattering}). However, as the target particle gets larger 
-this assumption fails to accurately describe the trapping force due 
-to the complexity in the gradient field. For particle sizes close to 
-the laser wavelength the scattered field is best described by Lorenz-Mie 
-theory.
+dipole. Unlike Lorenz-Mie scattering, there is no need to compute the 
+scattered field (see \eqref{eq:scattered}) for an individual sphere. 
+Higher complexity scattering problems can be simplified by subdividing 
+the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
+in which case the scattered field is now necessary. As the target 
+particle gets larger this assumption fails to accurately describe 
+the trapping force. For particle sizes close to the laser wavelength 
+the trapping force is best described by Lorenz-Mie theory.
 
@@
 
-Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
-$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
-this project we utilised an extension of the ECBM known as 
+Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, 
+and $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. 
+For this project we utilised an extension of the ECBM known as 
 multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
-\cite{Mackowski2011} the computational code computes the scattered field 
-from each sphere within the target cluster and the incident field itself. 
-The scattering between spheres slowly converges to a final result but modern 
-codes truncate the calculations to an upper number of iterations (denoted by
-$N_{max}$) to fit a desired accuracy and computational time.
-
-The $T$-matrix method is exceptionally useful for computing the scattering 
-from any arbitrary spherical aggregate. However, the $T$-matrix method by 
-itself can be computationally taxing as the number of spheres increases 
-\cite{Mackowski2011}. While it is possible to solve for the full electromagnetic 
-field of each the entire cluster this is only applicable for a single 
-orientation and can be even slower for large aggregates \cite{Mackowski1996, 
-Xu1995}. The benefit of \textit{mstm} is that the major scattering properties 
-(scattering and extinction cross sections, scattering matrices, etc) can 
-all be computed both for single orientations, or averaged over multiple 
-orientations to determine the average scattering from the target particle.
+\cite{Mackowski2011} the computational code computes the scattered 
+field from each sphere within the target cluster and the incident 
+field itself. The scattering between spheres slowly converges to a 
+final result but modern codes truncate the calculations to an upper 
+number of iterations (denoted by $N_{max}$) to fit a desired accuracy 
+and computational time.
+
+The $T$-matrix method is exceptionally useful for computing the 
+scattering from any arbitrary spherical aggregate. However, the 
+$T$-matrix method by itself can be computationally taxing as the 
+number of spheres increases \cite{Mackowski2011}. While it is 
+possible to solve for the full electromagnetic field of each the 
+entire cluster this is only applicable for a single orientation 
+and can be even slower for large aggregates \cite{Mackowski1996, 
+Xu1995}. The benefit of \textit{mstm} is that the major scattering 
+properties (scattering and extinction cross sections, scattering 
+matrices, etc) can all be computed both for single orientations, 
+or averaged over multiple orientations to determine the average 
+scattering from the target particle.
 
@@
 
-Where the first term accounts for inertial forces, the second term accounts 
-for friction forces which counteract the particles current motion ($\gamma_0$ 
-is the friction coefficient), and the final term accounts for the random 
-Brownian force. The $F(t)$ is there for convention which accounts for any 
-external forces acting on the particle. We can say that the noise term 
-$W(t)$ has a Gaussian probability, with a correlation function of:
+Where the first term accounts for inertial forces, the second term 
+accounts for friction forces which counteract the particles current 
+motion ($\gamma_0$ is the friction coefficient), and the final term 
+accounts for the random Brownian force. The $F(t)$ is there for 
+convention which accounts for any external forces acting on the 
+particle. These three terms are equal to the random Brownian motion 
+of the surrounding fluid. We can say that the noise term $W(t)$ has 
+a normal distribution, being scaled by the thermal energy of the 
+system, with a correlation function of:
 \begin{align}
@@
 \section{Simulation of spherical aggregates}
-
+\label{sec:simulations}
 Later chapters cover the dynamics of spherical aggregates and anisotropic 
@@
 
-Using a  MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007} we can compute the beam shape coefficients ($a_{nm}$ \& 
-$b_{nm}$) for any desired beam type. Using the results from \cite{Farsund1996} 
-we can then compute both the optical force and torque using the beam coefficients 
-and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) which are found by 
-calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and 
-then using \eqref{eq:t_matrix}. The form given by \cite{Crichton2000THEMD} is 
-an easier form to compute for higher orders of $N_{max}$:
+Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
+or \textit{ott} \cite{Nieminen2007} we can compute the beam 
+shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
+type. Using the results from \cite{Farsund1996} we can then 
+compute both the optical force and torque using the beam 
+coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
+which are found by calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
+total force in along the z-direction and the total torque about 
+the z-axis are provided
 \begin{equation}
@@
 
-\subsection{Simulation parameters}
+\subsection{Simulation reference frame}
 \label{sec:sim_parameters}
@@
 \begin{align}
-	E_{\rm inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave 
-functions, these differ from \eqref{eq:incident} in the fact that the 
-field can either be expressed as an incoming/outgoing wave (with a 
+	E_{\rm inc}(r)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(r)+b_{nm}RgN_{nm}(r)
+\end{align}
+
+where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical 
+wave functions, these differ from \eqref{eq:incident} and 
+\eqref{eq:scattered} in the fact that the field can either be 
+expressed as an incoming/outgoing wave (with a 
 singularity at the origin) or as a regular wave around the origin; for 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 	passing through the anodes A, B, C, and D.}
+	\label{fig:position_detectors}
 \end{figure}
@@
 \section{Nucleation with a Stationary and Moving Beam}
-As mentioned previously, shearing via optical rotation and particle displacement
-did not result in any localised nucleation events even while in the proximity of 
-the droplet edge. During the experiments with the galvano-mirror, it was found 
-that when no particle was present in the optical trap nucleation events would 
-occur while the beam was close to the edge of the droplet, even though the solution
-was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
-but was more interesting is how the beam's motion influenced the growth of the 
-nucleus.
+As mentioned previously, shearing via optical rotation and particle 
+displacement did not result in any localised nucleation events even 
+while in the proximity of the droplet edge. During the experiments 
+with the galvano-mirror, it was found that when no particle was 
+present in the optical trap nucleation events would occur while the 
+beam was close to the edge of the droplet, even though the solution
+was unsaturated. This has been reported prior \cite{Rungsimanon2010, 
+Liao2022}, but was more interesting is how the beam's motion 
+influenced the growth of the nucleus.
 
@@
 \label{sec:stationary}
-Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
-event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
-$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
-nucleus forms at the trap focus, growing quickly (growth rate was approximated using
-imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
-until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
-using optical tweezers shows that the growth rate is only loosely connected to the 
-solutions supersaturation. Local conditions play a much larger role in the growth 
-rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
-is escaped is due to the fact that crystal is far too large to be held in place and 
-is in fact still growing as the solution is supersaturated. The key take away to 
-remember is that the beam has no influence over the crystal shape, instead it 
-grows outward from the trap focus. Furthermore, due to the fact that the solution
-is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
-the crystal escapes as its size exceeds the trap focus. 
+As an example case, consider below in Fig.~\ref{fig:stationary_beam} 
+the frames taken from a nucleation event in supersaturated glycine 
+solution ($S=1.03$), the beam is a stationary being $\approx3.5 
+\mu m$ from the droplet edge. After a period of roughly 5 minutes a 
+nucleus forms at the trap focus, growing quickly from the focal point 
+of the trap until after roughly $6$ seconds the crystal escapes. 
+The growth rate was approximated using imageJ to be on the order of 
+$700\ \mu m^2/min$. Comparing to previous literature using optical 
+tweezers shows that the growth rate is only loosely connected to the 
+solutions supersaturation. Local conditions play a much larger role 
+in the growth rate than just the concentration \cite{Flannigan2023}. 
+A likely reason that the trap is escaped is due to the fact that 
+crystal is far too large to be held in place and is in fact still 
+growing as the solution is supersaturated. The key take away to 
+remember is that the beam has no influence over the crystal shape, 
+instead it grows outward from the trap focus. Furthermore, due to 
+the fact that the solution is supersaturated the crystal growth 
+cannot be contained to the trap focus. Instead the crystal 
+escapes as its size exceeds the trap focus. 
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
-	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
-	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
-	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
-	the trap.}
+	\caption{Laser induced nucleation at the edge of a droplet of supersaturated glycine solution. (b) shows the first instance 
+	of a crystal nucleus, growing quickly through (c)-(e) until 
+	after $6.665\ s$ the crystal begins to escape the trap.}
 	\label{fig:stationary_beam}
@@
 
+Fig.~\ref{fig:stationary_beam} makes it clear that in general, the
+trap focus can be used to induce nucleation. Previous reports
+have shown that this can occur over a range of supersaturations 
+\cite{Rungsimanon2010,Liao2022,Flannigan2023}. In every case they
+utilise a stationary beam as this is often easier to track. Using
+a galvno mirror we now consider the impact of a moving beam.
+
 \subsection{Moving Beam}
 \label{sec:moving}
-To test if a rapidly moving silica bead could generate the necessary shear rate 
-for crystal nucleation we wanted to see if a trapped silica bead could be trapped
-and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
-added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
-unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
-encountered several unexpected results.
+To test if a rapidly moving silica bead could generate the 
+necessary shear rate for crystal nucleation, a silica bead 
+be trapped in an aqueous solution. $20\ \mu L$ of glycine 
+and water ($S=1.03$) was added to $10\ \mu L$ of a dilute 
+water-silica mixture making the solution unsaturated ($S
+\approx0.7$). While trying to bring a silica bead towards
+the droplet edge several smaller droplets were seen within
+the fluid (see figure~\ref{fig:cluster_trapping} for a 
+larger example of these droplets).
 
-Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
-moving in a small elliptical pattern; interestingly. After a few minutes we 
-noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
-for a larger example of these droplets). These appear too have a wide distribution
-of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
-\mu m$. We surmise these could be small clusters of Glycine that had previously 
-been shown to form when the aqueous solutions where irradiated with a 
-focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
-directly entering the focus a nucleus forms close to the droplet edge, 
-unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
-the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
-being moved by and growing around the focal point of the trap. Because of 
-this the crystal nucleus lacks a clear morphology at first. Until roughly 
-$20\ s$ the crystal reaches a almost prismatic structure, with further 
-irradiation increasing the size.
+These appear too have a wide distribution of sizes, unlike 
+silica microspheres which have a uniform radius of $1.57\ 
+\mu m$. We surmise these could be small clusters of Glycine 
+that had previously been shown to form when the aqueous 
+solutions where irradiated with a focused laser \cite{Tsuboi2009, 
+Gowayed2021}. While no droplets are seen directly entering 
+the focus a nucleus forms close to the droplet edge, unlike 
+in Fig.~\ref{fig:stationary_beam} the crystal does not grow 
+out from the focal point evenly. Due to the galvano mirror, 
+the crystal is simultaneously being moved by and growing 
+around the focal point of the trap. Because of this the 
+crystal nucleus lacks a clear morphology at first. This is 
+in contrast to previous works with undersaturated solutions, 
+where even then the laser focus has little to no impact on 
+the crystal form \cite{Rungsimanon2010}.
 
-Interestingly the galvano-mirror allows the trap to impart a slight torque 
-on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
-where even though the crystal is not directly in the trap focus it rotates 
-in the $x-y$ plane and gets trapped again at a corner. The rotation could 
-not be due to fluid flow close to the surface of the crystal as the dipole 
-moment of individual water molecules is too small to be influenced by an 
-optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
-growth becomes localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
-\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
-(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
+Interestingly the galvano-mirror allows the trap to impart 
+a slight torque on the crystal, as shown in 
+fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even though 
+the crystal is not directly in the trap focus it rotates in 
+the $x-y$ plane and gets trapped again at a corner. The 
+rotation could not be due to fluid flow close to the surface 
+of the crystal as the dipole moment of individual water 
+molecules is too small to be influenced by an optical trap. 
+In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
+growth becomes localised to the corner. The area growth rate 
+between figures \ref{fig:eliptical_beam_1}(a) and (d) was 
+approximated using imageJ at $45.03\ \mu m^2 /min$, where as 
+between figures \ref{fig:eliptical_beam_1}(e) and (f) the 
+growth rate at that particular edge was estimated at $42.10\ 
+\mu m^2/min$. 
 \begin{figure}[h!]
@@
 
-Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
-\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
-moving beam. This modification allows for the crystal growth to be localised to
-a specific region of the bulk crystal whereas with a stationary beam there is no
-control over the crystal morphology. In fact this allows for a much finer control
-over the exact shape of the crystal nucleus, in some cases allowing for growth out
-of the viewing plane as shown in
-
-\begin{figure}
-	\centering
-\end{figure}
+Nucleation in undersaturated conditions has been reported 
+previously in $D_2O$ \cite{Rungsimanon2010} and $H_2O$ 
+\cite{Flannigan2023}, though not involving a moving beam. 
+This modification allows for the crystal growth to be 
+localised to a specific region of the bulk crystal whereas 
+with a stationary beam there is no control over the crystal 
+morphology. In fact this allows for a much finer control
+over the exact shape of the crystal nucleus, in some cases 
+allowing for growth out of the viewing plane as shown in
+fig.XX
 
@@
 \section{Influence of a moving beam front on seed crystals}
-\textcolor{red}{Need to write up these results. To summarise the crystal 
-	dissolved around the laser focus instead of growing as the seed crystals 
-	are in the solution and not close to the air-solution interface}
+Localised crystal growth could have useful applications for 
+tailoring seed crystals. To understand how the a moving beam
+front could be
 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 \label{chapter:simulated_detection}
-As outlined in the end of Chapter 4, one of the difficulties in characterising 
-interactions with asymmetric objects is the coupled motion between translation 
-and rotation. This can be somewhat ignored if the rotational motion is periodic,
-like when the power spectra of elliptical polystyrene particles was fitted by 
-Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
-rotational motion is stochastic the problem is more complex. For example, when a
-novel trap characterisation technique was implemented by Saffron \textit{et al}
-\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
-characterise both the axial and lateral trap stiffness acting on microspheres. The
-only drawback admitted to in their work was that the technique was constrained to
-isotropic scatters as their theoretical model for describing the auto-correlation
-function was predicated on the fact that any variations in the signal are due to
-the particles translational motion within the confines of a cylindrical trap 
-\cite{BarZiv1997}. However, it is somewhat naive to assume that the orientational 
-change
+As outlined in the end of Chapter 4, one of the difficulties 
+in characterising interactions with asymmetric objects is 
+the coupled motion between translation and rotation. In order
+to characterise the optical trap a position detection system 
+such as a lateral effect photodiode or a quadrant photo diode
+(see fig.\ref{fig:position_detectors}). Typically, a position 
+detection system assumes that there exists a linear relationship 
+between the a particle's displacement and the detected signal.
+However, this is not always the case as demonstrated in chapter
+4; as dimer's demonstrate non-harmonic trapping beyond a certain 
+displacement. In addition, the rotational effects of the trapped
+dimer become a more significant factor when characterising the 
+trap strength. This can have unintended effects when it comes to
+areas of research that require a precise understanding of the 
+hydrodynamic behaviour of particles. As such there has been a 
+concerted effort to either eliminate rotational motion, or describe
+how the rotational behaviour influences the trajectory reported 
+by the position detector.
 
-\section{Simulative QPD}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
+Rotational motion about a single axis is easiest to account for.  
+When the power spectra of elliptical polystyrene particles was 
+fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
+assumed that the rotational motion was purely in the transverse
+plane. As such they did not have to account for any variance in 
+the trapping strength due to orientation nor did they need to
+consider non-periodic rotational behaviour. In the case where 
+rotational motion is stochastic the problem is more complex. 
+For example, when an optical fibre trap characterisation 
+technique was implemented by Saffron \textit{et al} 
+\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
+light scattering to characterise both the axial and lateral 
+trap stiffness acting on microspheres. The only drawback 
+admitted to in their work was that the technique was constrained 
+to isotropic scatters as their theoretical model for describing 
+the auto-correlation function was predicated on the fact that 
+any variations in the signal are due to the particles 
+translational motion within the confines of a cylindrical trap 
+\cite{BarZiv1997}. Where the upper limit of the cylindrical 
+trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
+However, as demonstrated by the results from Chapter 4.1, 
+the axial traps of spherical aggregates is often situated 
+far beyond the Rayleigh range (for a 1.2 NA laser this is 
+$\pm5.985 \mu m$).
 
-\subsection{Incident beam}
-\label{sec:fibre_scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
+To accurately describe the full dynamics of an anisotropic
+scatterer its clear that additional characterisation 
+techniques are necessary. In this chapter we demonstrate 
+two such techniques, quadrant photo-diode simulation, and 
+an optical fibre static light scattering arrangement. The 
+former is based on the work of Rohrbach and Stelzer 
+\cite{Rohrbach2002}, where their work is limited to single 
+spheres, we extend the possible particle space to allow for 
+any theoretical combination of spheres. The results from
+chapter 4 demonstrate that the rotational motion causes 
+significant deviations between the computed trapping strength
+and the expected trapping strength according to Lorenz-Mie
+theory. By utilizing vector regression we can accurately 
+recreate a particle's trajectory simply from the QPD signal, 
+both in the transverse plane and along the beam axis.
 
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
+The latter method builds upon \cite{BarZiv1997} but now 
+accounts for the full rotational motion. This was initially
+developed in tandem with the work conducted in chapter 3,
+with the idea being that if a crystal nucleus formed on a 
+rotating sphere how would that nucleus influence the motion
+of the rotating sphere. Using static light scattering we can 
+map the outputted signal from a optically trapped dimer to 
+its expected orientation in real time. Thus allowing us to 
+estimate the optical torque being applied to the dimer by 
+the trapping beam. As demonstrated in chapter 4, the 
+influence of a circularly polarised beam can result in 
+gyroscopic precession for specifically sized dimers.
 
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity
-of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
+We start by outlining the working principle behind the
+optical fibre detection system, before demonstrating 
+how it can be used to instantaneously predict 
+orientational information on a trapped particle. We then
+consider common factors of error in the characterisation
+process, such as signal error, and incorrect particle 
+sizing. Both of which have a significant impact on the 
+performance of the model, but by utilising Bayesian 
+inference, and time averaging we can better refine our 
+model.
 
 \section{Monitoring Stochastic rotational motion using static light scattering}
-Orientational dynamics to an anisotropically scattering shape are difficult to 
-characterise due to the coupling of rotational and translational effects. As 
-shown in the previous chapter the dynamics of even simple dimers are heavily 
-dependent on the particle's position and orientation, this is reflected in 
-literature where several engineering solutions have been devised to decouple 
-translational and orientational motion \textcolor{red}{[ , ]}. While a majority 
-of latter work has been focused on nano-particles (falling into the Rayleigh 
-Regime) and utilising florescence
-
+Part of the optical fibre arrangement borrows code from the 
+optical tweezer simulations discussed in Chapter 2 and 4. 
+Rather than repeat the simulation details we will instead 
+only focus on the specifics of 
 \subsection{Coordinate System}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{QPD Simulations}
+A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
+
+\subsection{Incident beam}
+\label{sec:fibre_scattering}
+The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
+*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
+
+\begin{align}
+	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
+\end{align}
+
+Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
+For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
+In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
+
+We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
+If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
+
+\begin{align}
+	\begin{pmatrix}
+		p_{nm}\\
+		q_{nm}
+	\end{pmatrix} =T 
+	\begin{pmatrix}
+		a_{nm}\\
+		b_{nm}
+	\end{pmatrix} = 
+	\begin{pmatrix}
+		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
+		bT^{TM}_{nm} & bT^{TE}_{nm}
+	\end{pmatrix}
+	\begin{pmatrix}
+		a_{nm}\\
+		b_{nm}
+	\end{pmatrix}
+\end{align}
+
+For *mstm* the T-matrix is packed as a column vector:
+
+\begin{align}
+	T_{MSTM} = 
+	\begin{bmatrix} 
+		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
+		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
+		... & ...\\ 
+		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
+		----&----\\ 
+		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
+	\end{bmatrix}
+\end{align}
+
+Where as *ott* packs packs the T-matrix with sub matrices:
+
+\begin{align}
+	T_{Ott} = 
+	\begin{bmatrix} 
+		\begin{pmatrix}
+			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
+			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
+		\end{pmatrix} \\ 
+		\begin{pmatrix}
+			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
+			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
+		\end{pmatrix}\\
+		....\\ 
+		\begin{pmatrix}
+			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
+			bT^{TM}_{nm} & bT^{TE}_{nm}
+		\end{pmatrix}
+	\end{bmatrix}
+\end{align}
+
+\subsection{Scattered and Total Fields}
+With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
+Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
+Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
+
+The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
+This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
+I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
+The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
+
+Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
+This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
+The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
+
+Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
+The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
+Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
+Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
+
+Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
+We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
+Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
+This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
+First we checked for this by measuring the QPD response for random positions within the XY plane:
+
+Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
+This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
+
+\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
+For two correlated power spectra, the cross correlation is given as:
+
+\begin{align}
+	P_{x} = |\hat{S}_x(f)|^2 \\ 
+	P_{y} = |\hat{S}_y(f)|^2 \\
+	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
+\end{align}
+
+Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
+They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
+They found these transformed positions by minimising the sum cross-correlation:
+
+\begin{align}
+	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
+\end{align}
+
+Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
+Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
+
+\begin{align}
+	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
+\end{align}
+
+Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
+With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
+
+\section{Rotations and Asymmetric particles}
+Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
+Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
+This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
+
+Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
+Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
+This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
+As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
+
+Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
+The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
+For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
+
+\section{Power Spectra of single sphere vs spherical aggregates}
+In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
+In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity
+of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
+As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
+Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
+
+Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
+The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
+
+In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
+Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
+We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
+
+Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
+Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
+However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
+Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
+This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
+Therefore one should expect that for aggregates containing multiple sized particles (e.g.
+cellular samples) typical characterisation techniques may over estimate the strength of the trap.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-09-16 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

--- 2024-08-15 Latex_files/Thesis/chapters/chapter 3/main.tex
@@
 	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
-		in bulk fluid moving at different speeds. The shear rate is calculated 
+		in bulk water moving at different speeds. The shear rate is calculated 
 		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
@@
 mirror a better option for generating high shear rates but also over a larger 
-volume.
+volume. The only limitation to this method is the effective drag experienced 
+by the trapped particle.
 
@@
 
-\section{Influence of a moving beam front on seed crystals}
+\subsection{Influence of a moving beam front on seed crystals}
 \textcolor{red}{Need to write up these results. To summarise the crystal 
@@
 are a necessary precursor to induce crystal nucleation then understanding how
-they are added to the bulk crystal may help with our understanding of the kinetics
-of multi-step nucleation.
+they provide materials to the bulk crystal may help with our understanding of 
+the kinetics of multi-step nucleation.
 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 1/main.tex

@@
 \subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT, colloid simulations 
-showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) 
-interactions allow for the formation of a liquid-liquid metastable phase from 
-which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
-\textit{in situ} techniques for studying nucleation several papers reported 
-the presence of stable liquid-like clusters that formed prior to nucleation
-\cite{Savage2009, Wolde1997, Soga1999}. It can be understood by Oswald's rule
-\cite{Ostwald1897}, which says that any crystallising system does not immediately 
-take the path to the lowest possible energy state but instead first transitions 
-to the state with the smallest free energy barrier, further transitions can still 
-occur but this pathway minimises the overall free energy cost. If we consider 
-the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider 
-a case where the supersaturation is low ($S\approx1.01$) then a dense liquid 
-droplet will have a lower chemical potential than the surrounding fluid but 
-the crystalline phase will have a subsequently lower chemical potential than 
-the dense liquid phase. Initially it was suspected that the formation of the 
-dense liquid phase was a result of stochastic density fluctuations in the system; 
-further simulations suggest that the the local bond order is a stronger driving 
-force than the local density \cite{Tan2013}.
+The two step nucleation theory is an extension to the CNT that suggests 
+that prior to nucleation, the solute will cluster together. This was 
+shown via colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}.
+Several papers later reported the
+presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}. The formation of these clusters 
+can be understood by Oswald's rule; which says that any crystallising 
+system does not immediately take the path to the lowest possible energy 
+state but instead first transitions to the state with the smallest free 
+energy barrier \cite{Ostwald1897}. Further phase transitions can still 
+occur but this pathway minimises the overall free energy cost. Initially 
+it was suspected that the formation of the dense liquid phase was a 
+result of stochastic density fluctuations in the system. However, 
+simulations suggest that the the local bond order is a stronger driving 
+force than the local density \cite{Tan2013}. Studies into two step 
+nucleation in Glycine solutions have noted that the solutions need to be 
+aged before the presence of clusters was detected. This suggests that 
+there exist further barriers to the formation of clusters in some systems,
+this has lead to instead calling this phenomena as Multi-step Nucleation.
+
+\subsection{Non-classical Nucleation}
 
@@
 particular importance; with nucleation probability changing depending 
-on the wavelength \cite{Kacker2017}.
-In addition, the choice of solute will effect the setup, not only because some solute's are unaffected, but also because there is a minimum laser threshold before nucleation is observed
+on the wavelength \cite{Kacker2017}. In addition, the choice of solute will 
+effect the setup, not only because some solute's are unaffected, but also 
+because there is a minimum laser threshold before nucleation is observed 
 \cite{Garetz2002}. Several papers have debated the exact mechanism that 
@@
 conditions that are unfavourable to cluster formation the polarising 
-effect can stabilise the clusters \cite{Alexander2008}. As the cluster  concentration rises so does the likelihood of nucleation \cite{Vekilov2010}. 
+effect can stabilise the clusters \cite{Alexander2008}. As the cluster 
+concentration rises so does the likelihood of nucleation \cite{Vekilov2010}. 
 Both theories are similar to one another but where the optical Kerr theory
@@
 influence the crystal polymorph \cite{Wang2010}. The latter is a tricky 
-task to address due to the fact that these cavitation bubbles form and collapse in less than $100\ \mu s$.
-Using fluorescence dyed proteins, researchers were able to observe a sudden spike in fluorescence just
+task to address due to the fact that these cavitation bubbles form and 
+collapse in less than $100\ \mu s$. Using fluorescence dyed proteins, 
+researchers were able to observe a sudden spike in fluorescence just 
 as the cavitation bubble began to collapse, they suggested that due 
-to the collapse of the cavitation bubble the protein clusters are brought together at the lasers focal point. However, while the fluorescence 
+to the collapse of the cavitation bubble the protein clusters are 
+brought together at the lasers focal point. However, while the fluorescence 
 imaging indicates a local concentration increase it is difficult to 
-quantify this change depending on the size of the bubble \cite{Korede2023}. It
-has been suggested that in theory any solution can undergo HILIN \cite{Korede2023}, but proving such a theory requires a clear 
-understanding of the phenomena both before and after cavitation occurs. Current research aims to combine experimental research with computer 
+quantify this change depending on the size of the bubble \cite{Korede2023}. 
+It has been suggested that in theory any solution can undergo HILIN 
+\cite{Korede2023}, but proving such a theory requires a clear 
+understanding of the phenomena both before and after cavitation occurs. 
+Current research aims to combine experimental research with computer 
 simulations to develop a universal theory, with the hope that this could 
@@
 \subsubsection{Trapping Induced Nucleation}
-Lastly, there is trapping induced nucleation, this is where optical tweezers 
-come into play. Due to the radiation pressure created by the focused beam,
-it is possible to manipulate the solute, this was demonstrated with amino
-acids such as glycine \cite{Tsuboi2009}. Whether or not a crystal forms is 
-due to the location of the laser focus.
-When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \cite{Yuyama2010, Yuyama2012}.
-The dipole moment of the glycine molecules is too small to be influenced by the optical trap, as such it would suggest that larger aggregates are being manipulated.
-Applying DLS analysis to the dense liquid region showed that it was populated by clusters that would consolidate together upon being focused by the
-optical trap \cite{Gowayed2021}.
-Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point suggesting that the clusters are formed due to glycine reaction products \cite{Sweatman2022}.
-When the optical trap is moved from the cover slip to the air-solution interface, nucleation would occur before a dense liquid region could form \cite{Yuyama2010}.
-Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events.
-In some instances the nucleation occurs spontaneously after a short period of time \cite{Yuyama2010}.
-Whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \cite{Liao2022}.
-The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \cite{Liao2022}.
-The reason why nucleation is only seen at the air-solution interface is due to the limited molecular mobility close to the interface.
-Often tweezing experiments will use a hydrophilic
-coating to minimise the height of the solution droplet and further limit
+Lastly, there is trapping induced nucleation, this is where optical 
+tweezers come into play. Due to the radiation pressure created by 
+the focused beam, it is possible to manipulate the solute, this was 
+demonstrated with amino acids such as glycine \cite{Tsuboi2009}. 
+Whether or not a crystal forms is due to the location of the laser 
+focus. When focusing on the cover slip, supersaturated solutions 
+of glycine and $D_2O$ were shown to create a dense liquid droplet 
+of glycine and water \cite{Yuyama2010, Yuyama2012}. The dipole 
+moment of the glycine molecules is too small to be influenced by 
+the optical trap, as such it would suggest that larger aggregates 
+are being manipulated. Applying DLS analysis to the dense liquid 
+region showed that it was populated by clusters that would 
+consolidate together upon being focused by the optical trap 
+\cite{Gowayed2021}. Molecular simulations of glycine solutions 
+showed that these clusters are unstable when using pure glycine 
+below the saturation point suggesting that the clusters are 
+formed due to glycine reaction products \cite{Sweatman2022}. 
+When the optical trap is moved from the cover slip to the 
+air-solution interface, nucleation would occur before a dense 
+liquid region could form \cite{Yuyama2010}. Repeated experiments 
+where the laser is focused on the air-solution interface have 
+lead to a variety of different nucleation events. In some 
+instances the nucleation occurs spontaneously after a short 
+period of time \cite{Yuyama2010}. Whereas allowing a solution to 
+age results in the formation of amorphous precursors that when 
+irradiated will nucleate immediately \cite{Liao2022}. The 
+precursors are only seen when the solution is irradiated by an 
+optical tweezer and the growth rate can be controlled somewhat 
+by varying the laser power \cite{Liao2022}. The reason why 
+nucleation is only seen at the air-solution interface is due 
+to the limited molecular mobility close to the interface. 
+Often tweezing experiments will use a hydrophilic coating to 
+minimise the height of the solution droplet and further limit
 the molecular mobility \cite{Yuyama2012, Gowayed2021}.
 
-Walton and Wynne discussed a plausible model for how the tweezer focus
-could result in a nucleation event. Put simply, when the laser is focused
-at the solution the radiation pressure draws in solute material, creating 
-a concentrated region of solute. This also creates a depleted region 
-around the focus and raises the local temperature. When the laser is 
-turned off the depleted region around the focus quickly cools back to the 
-ambient temperature. This sudden cooling allows for nucleation to occur
+Walton and Wynne discussed a plausible model for how the tweezer 
+focus could result in a nucleation event. Put simply, when the 
+laser is focused at the solution the radiation pressure draws 
+in solute material, creating a concentrated region of solute. 
+This also creates a depleted region around the focus and raises 
+the local temperature. When the laser is turned off the depleted 
+region around the focus quickly cools back to the ambient 
+temperature. This sudden cooling allows for nucleation to occur
 just outside the focus.
 
-Laser induced nucleation has the potential to be a viable method for 
-\textit{in-situ} studying of nucleation events. Using high numerical 
-aperture lenses one can localise the nucleation event to a specific 
-region of the solution. The current issue is that the mechanism behind
-laser induced nucleation is not fully understood, as such it is rather
-difficult to modify the laser for different solution parameters. Instead
-it may be more effective to manipulate the solution using trapped particles.
-An common method is by apply a torque to a trapped particle, creating 
-a micro-rotor that can can generate fluid flow.
+Laser induced nucleation has the potential to be a viable method 
+for \textit{in-situ} studying of nucleation events. Using high 
+numerical aperture lenses one can localise the nucleation event 
+to a specific region of the solution. The current issue is that 
+the mechanism behind laser induced nucleation is not fully 
+understood, as such it is rather difficult to modify the laser 
+for different solution parameters. Instead it may be more 
+effective to manipulate the solution using trapped particles. 
+An common method is by apply a torque to a trapped particle, 
+creating a micro-rotor that can can generate fluid flow.
 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 2/main.tex

@@
     E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
-      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
+      -i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
     \right] 
@@
 where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
-are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
-and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
-field computing its expansion coefficients is possible via analytical
-methods and are completely dependent on the beam conditions imposed by
-the user. However, computing the expansion coefficients of the
-internal and scattered fields is far more tedious depending on the
-shape of the target and what properties of the scattered field are
-desired --- see sec.~\ref{sec:scattering}.  From an optical
-trapping perspective the force exerted by a focused electric field can
-be found by computing the Maxwell stress tensor which only requires
-the total magnitude of the incident and scattered fields. An alternative
-method is to compute the momentum carried by the incident and scattered
-beams and taking the difference between to measure the momentum 
-transferred to the particle. This project utilises the results from 
-Farsund \cite{Farsund1996} but in the form used by Crichton 
-\cite{Crichton2000THEMD}. As an example the linear momentum carried 
-by any beam can be computed via:
+are the expansion coefficients of each of the fields, and $M_{nm}$ 
+and $N_{nm}$ are the vector spherical harmonics. The superscripts
+denote whether the field is inward or outward propagating; so a 
+superscript (1) is inward propagating, and (2) is outward propagating.  
+For the incident field computing its expansion coefficients is possible 
+via analytical methods and are completely dependent on the beam 
+conditions imposed by the user. However, computing the expansion 
+coefficients of the internal and scattered fields is far more 
+tedious depending on the shape of the target and what properties 
+of the scattered field are desired --- see sec.~\ref{sec:scattering}.  
+From an optical trapping perspective the force exerted by a focused 
+electric field can be found by computing the Maxwell stress tensor 
+which only requires the total magnitude of the incident and scattered 
+fields. An alternative method is to compute the momentum carried by 
+the incident and scattered beams and taking the difference between 
+to measure the momentum transferred to the particle. This project 
+utilises the results from Farsund \cite{Farsund1996} but in the 
+form used by Crichton \cite{Crichton2000THEMD}. As an example the 
+linear momentum carried by any beam can be computed via:
 \begin{equation}
@@
 total force in along the z-direction and the total torque about 
-the z-axis are provided
+the z-axis are provided:
 \begin{equation}
@@
 \textit{Ott} uses a point matching approach to approximate the beam shape 
-coefficients of the incident field by fitting it to the far field estimate, 
-the beam is of the form:
-\begin{align}
-	E_{\rm inc}(r)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(r)+b_{nm}RgN_{nm}(r)
-\end{align}
-
-where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical 
-wave functions, these differ from \eqref{eq:incident} and 
-\eqref{eq:scattered} in the fact that the field can either be 
-expressed as an incoming/outgoing wave (with a 
-singularity at the origin) or as a regular wave around the origin; for 
-incoming/outgoing waves the wave functions use the first/second forms of 
-the Hankel function respectfully. In order to compute the regular spherical 
-wave at the origin we replace the Hankel function with the Bessel function 
-which is simply the average of the first and second forms of the Hankel 
-function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by setting 
+coefficients of the incident field by fitting it to the far field estimate. 
+From the QPD's perspective it is receiving light both from the incident 
+and scattered beam simultaneously, as such both fields must be expressed
+using outgoing vector spherical harmonics.
+\begin{align}
+	E_{\rm inc}(r)&=
+	E_0\sum^\infty_{nm}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
+	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
+	\\
+	E_{\rm scat}(r)&=E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
+	-i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
+	\right] 
+\end{align}
+
+We can, if we want, further restrict the incident beam by setting 
 the truncation angle to match our microscope object, this applies a 
@@
 conditions. In order to compute the scattering from 
-the target particle \textit{ott} uses the $T$-matrix method, this is not 
-essential for a simple sphere but is essential for complex shaped particles 
-such as dimers. The matrix from mstm needs to be repackaged to work with the
-\textit{ott} software, converting it from a column vector into a system of 
-sub matrices.
+the target particle \textit{ott} uses the $T$-matrix method, this is 
+not essential for a simple sphere but is essential for complex shaped 
+particles such as dimers. The matrix from mstm needs to be repackaged 
+to work with the \textit{ott} software, converting it from a column 
+vector into a system of sub matrices.
 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 4/main.tex

@@
 Much of the calibration theory discussed in Chapter 2 assumes that the
-target particle in question is a single sphere, one who's scattering and 
+target particle in question is a single sphere, one whose scattering and 
 motion is easily computed. However, while working with dense colloidal 
@@
 OT and found that multiple trapped beads could be mistaken for a single 
-trapped bead with altered trap stiffness. Theoretical studies on the case 
-of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a 
-ray-optics based model to show that the two trapped beads are brought into 
-physical contact with each other by optical forces and they also calculated 
-the axial equilibrium positions of the two trapped beads as a function of 
-their size. Experiments in \cite{Praveen2016} confirmed that the two trapped 
-beads indeed experience different trap stiffnesses in the vicinity of the 
-same potential well. There are further discussions looking into the dynamics 
-of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, 
-Chetana2022}, their results all showing that predicting the behaviour an 
-arbitrary shaped particle comes with great difficulty due to the fact that 
-the optical force is dependent on a greater number of variables such as 
-orientation and size factors.
+trapped bead with altered trap stiffness (see \ref{sec:harmonic_traps}).
+Theoretical studies on the case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics based model to show
+that the two trapped beads are brought into physical contact with each 
+other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Experiments 
+in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
+
+There are further discussions looking into the dynamics 
+of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to
+the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
 
 With the initial goal of the PhD being to induce nucleation events via a 
-spherical micro-rotor the goal of this chapter is to - in a limited capacity
+spherical micro-rotor, the aim of this chapter is to - in a limited capacity
 simulate and investigate the influence of a second particle being bound to our 
@@
 upon the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
-asymmetric dimers and how varying size parameters alters the dynamics and 
-additionally makes characterising their interactions within an optical trap 
-more cumbersome.
+asymmetric spherical dimers and how varying size parameters alters the 
+dynamics and additionally makes characterising their interactions within an optical trap more cumbersome.
 
-In this chapter we will consider how the addition of a second sphere changes 
-the trapping dynamics by introducing multiple equilibrium positions. These 
-positions are dependent on both the size ratio of the dimer and its orientation
-relative to the trap. This is important to address, for while the the dimers' 
-diffusive behaviour is the same, regardless of position and orientation, how it
-interacts with the trap may produce the wildly different results. Furthermore,
-we look at how our dimers interact with circularly polarised light, this is 
-especially pertinent as the original plan was to utilise circularly polarised 
-light to generate fluid flow in a supersaturated solution. The rotational motion 
-of dimers in circularly polarised light goes against much of the established 
-literature and suggests that the transfer of spin angular momentum is enabled 
-by combining spheres together. And lastly, we demonstrate how a quadrant photo 
-diode, performs in characterising the actual interactions between a dimer and 
-the optical trap. We do so to demonstrate that much of the dimer's trajectory
-information is either lost or poorly described by simple calibration techniques, 
+In this chapter we will consider how the addition of a second sphere 
+changes the trapping dynamics by introducing multiple equilibrium 
+positions. These positions are dependent on both the size ratio of 
+the dimer and its orientation relative to the trap. This is important 
+to address, for while the the dimers' diffusive behaviour is the same, regardless of position and orientation, how it interacts with the trap 
+may produce the wildly different results. Furthermore, we look at how 
+our dimers interact with circularly polarised light, this is especially pertinent as the original plan was to utilise circularly polarised 
+light to generate fluid flow in a supersaturated solution.
+The rotational motion of dimers in circularly polarised light goes against much of the established literature and suggests that the transfer of spin angular
+momentum is enabled by combining particles together. And lastly, we 
+demonstrate how a quadrant photo diode, performs in characterising 
+the actual interactions between a dimer and the optical trap. We do so 
+to demonstrate that much of the dimer's trajectory information is 
+either lost or poorly described by simple calibration techniques, 
 making their motion difficult to characterise.
@@
 \subsection{Simulation Parameters}
-As a paradigmatic example, consider a dimer suspended in water ($n_p = 1.59, 
-n_m = 1.33$) located at the focus of a Gaussian beam (more specifically a 
-Laguerre-Gaussian beam of mode $[0,0]$), the beam is focused by a objective 
-with numerical aperture of 1.2 and is x polarised. The size ratio of the two 
-sphere's is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ 
-unless specified otherwise; the dimer's orientation is given by a unit vector 
-connecting the centres of both spheres (see \ref{sec:sim_parameters}). We 
-define the 'standard' orientation as being aligned with the direction of 
-beam propagation direction - and therefore the 'inverted' orientation is 
+As a paradigmatic example, consider a dimer suspended in water 
+($n_p = 1.59, n_m = 1.33$) located at the focus of a Gaussian 
+beam (more specifically a Laguerre-Gaussian beam of mode $[0,0]$, 
+see \eqref{eq:incident}), the beam is focused by a objective with 
+numerical aperture of 1.2 and is x polarised. The size ratio of 
+the two spheres is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is 
+kept at $1\ \mu m$ unless specified otherwise; the dimer's 
+orientation is given by a unit vector connecting the centres of 
+both spheres (see \ref{sec:sim_parameters}). We define the 
+'standard' orientation as being aligned with the direction of beam 
+propagation direction - and therefore the 'inverted' orientation is 
 defined when the dimer is orientated against the direction of beam 
@@
 \label{sec:eq_positions}
-If we wanted to start from first principles and determine the trap strength 
-on our target particle the first step would be to locate the equilibrium position
-relative to the trap focus. For a single sphere it is easy to enough to understand 
-that its centre of mass will be drawn to focal point of the laser due to gradient
-forces, once there the force is analogous to a harmonic spring with a fixed trap
-stiffness (see fig.\ref{fig:harmonic_trap}). Now, if we consider instead a dimer, 
-we now have two spheres both being drawn to the focus along by the same gradient 
-force; in addition the scattering force is significantly more complex due to both 
-spheres scattering the electromagnetic fields. This mutual scattering between individual 
-spheres is what makes simulating spherical aggregates far more difficult compared
-to a single sphere, and even harder still to predict the position where the dimer's 
-centre of mass is in equilibrium.
+If we wanted to start from first principles and determine the trap 
+strength on our target particle the first step would be to locate the equilibrium position relative to the trap focus. For a single sphere 
+it is easy to enough to understand that its centre of mass will be 
+drawn to focal point of the laser due to gradient forces, once there 
+the force is analogous to a harmonic spring with a fixed trap stiffness 
+(see fig.\ref{fig:harmonic_trap}). Now, if we consider instead a dimer, 
+we now have two spheres both being drawn to the focus along by the same gradient force; in addition the scattering force is significantly more 
+complex due to both spheres scattering the electromagnetic fields. 
+This mutual scattering between individual spheres is what makes simulating spherical aggregates far more difficult compared to a single sphere, 
+and even harder still to predict the position where the dimer's centre 
+of mass is in equilibrium.
 
-Because the scattering force is only significant in the direction of beam 
-propagation the potential well in the transverse plane can still be assumed 
-to be harmonic around the central beam axis. The axial optical force cannot 
-be assumed to behave as a simple harmonic trap. The methodology for computing 
-optical forces has been covered extensively for a number of different trapping 
-conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the 
-trapping force and determine where a simple sphere would be located relative 
-to focal point of the laser by finding the position that minimises the net 
-optical force in a negative feed back loop ($\delta F/\delta x < 0$) - we can 
-assume that for a dielectric sphere the optical torque is negligible. For a 
-dimer (or any arbitrary spherical aggregate), we now must consider both its 
-position and orientation and find where the net optical force and torque are 
-minimised.
+Because the scattering force is only significant in the direction of 
+beam propagation \cite{Capitanio2002} we can assume there is only a 
+single potential well in the transverse plane. Along the beam axis, 
+such an assumption is no longer valid due to fact that spherical 
+aggregates experience inter-particle scattering. This is a key reason
+for using \textit{mstm} as it accounts for that behaviour. The 
+methodology for computing optical forces has been covered extensively 
+for a number of different trapping conditions \cite{RanhaNeves2019}. 
+So it is relatively easy to compute the trapping force and determine 
+the equilibrium position by finding the position that minimises the 
+net optical force and the local gradient is negative ($\delta F/
+\delta x < 0$) - we can assume that for a dielectric sphere the 
+optical torque is negligible. For a dimer (or any arbitrary spherical aggregate), we now must consider both its position and orientation 
+and find where the net optical force and torque are minimised.
 

### 2024-09-16 Latex_files/Thesis/chapters/chapter 5/main.tex

@@
 
-\section{Monitoring Stochastic rotational motion using static light scattering}
+\section{Monitoring Stochastic rotational motion using static 
+		light scattering}
 Part of the optical fibre arrangement borrows code from the 
@@
 Rather than repeat the simulation details we will instead 
-only focus on the specifics of 
+only focus on the specifics of the optical fibre scattering.
+
 \subsection{Coordinate System}
-In the case of our probe beam the +z direction points from the surface of the 
-probe directly to the centre of diffusion in our dimer, with the +y direction 
-pointing towards what would be our trapping beam, and the +x direction pointing 
-into the page. The origin of our coordinate system is fixed on the centre of 
-the focus of our trapping beam, meaning as our dimer's centre of diffusion moves 
-the origin is kept constant.
-
-\subsection{Beam}
-The probe is a defined as a plane wave by setting the Gaussian beam parameter 
-$C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector 
-as $[1,0,0,0]$. The plane wave in \textit{mstm} is given by \eqref{eq:incident}.
+Consider a particle (either a sphere or dimer) trapped by
+a Gaussian beam, with its focal point being set at [0,0,0]. 
+At the same time an optical fibre directs a plane wave that 
+is incident on the trapped particle scattering light in all 
+directions. The probe beam is assumed to be x-polarised and 
+propagating in the +z direction, this is a $90^{\circ}$ 
+rotation from the previous simulations (see fig\emph{X.X}).
 
 \subsection{Dimer}
-The dimer is defined by two spheres, in the trapping frame this would be orientated 
-by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as 
-the large sphere is furthest from the trapping beam. We scale the position of each 
-sphere by the factor L:
-\begin{align}
-	L = \frac{1}{k} = \frac{\lambda}{2\pi}
-\end{align}
+The dimer is defined by two spheres with refractive indices
+of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
+requires a target origin for setting the scattering expansion, as
+with our simulations we set it as the dimer's centre of diffusion.
+The near field scattering from the dimer is shown in 
+fig.~\ref{fig:nf_scattering}.
 
-Where $\lambda$ is the wavelength of light and is given in nanometres, this means 
-every position must be also provided in nanometres.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{nf_scattering_dimer.png}
+	\caption{Re ${\bf E_z}$ distribution from a symmetric dimer when 
+	irradiated by the probe beam (viewed in the x-y plane). Axis are 
+	given in dimensionless units of $k\hat{x}$ and $k\hat{y}$ respectively.}
+	\label{fig:nf_scattering}
+\end{figure}
 
 \subsection{Detectors and Pixels}
-Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted 
-later for testing), with the position of the detector being defined by the polar 
-and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
+Each detector is placed some distance $\bf r$ from the origin, 
+detector positions are defined by angles $\theta$ \& $\phi$. With 
+$\theta=\phi=0$ being directly in front of the probe beam. The 
+position of each detector is given by:
 \begin{align}
@@
 \end{align}
-So if you were to place the fibres in the trapping frames x-y plane (i.e. flat against
-the cover slip in a covenantal trapping context), in the probe beam frame will only 
-have x and z components. The orientation of the detector is just the -ve of its 
-position and not scaled by the distance term r, the orientation being the vector from 
-the origin of the detector to the center of the dimer.
+Where r is the distance between the detector and the target origin. 
+In an ideal situation the surface of the detector is oriented towards 
+the target origin perfectly. Pixels can be thought of as points lying 
+on the surface of the detector, each point has its own position in real
+space. Using \textit{mstm} we can find the components of the scattering 
+matrix at each point and compute the intensity of the electric field via:
 
-Pixels can be thought of as points lying on the surface of the detector, by getting 
-the spherical coordinates of each pixel we can compute the intensity at each point on 
-the detector, allowing us to determine the average intensity on a given detector.
+\begin{align}
+	\begin{pmatrix}
+		I_s \\ Q_s \\ U_s \\ V_s
+	\end{pmatrix} = 
+	\begin{pmatrix}
+		S_{11} & S_{12} & S_{13} & S_{14} \\
+		S_{21} & S_{22} & S_{23} & S_{24} \\
+		S_{31} & S_{32} & S_{33} & S_{34} \\
+		S_{41} & S_{42} & S_{43} & S_{44} \\
+	\end{pmatrix}
+	\begin{pmatrix}
+		I_i \\ Q_i \\ U_i \\ V_i
+	\end{pmatrix}
+\end{align}
+Where:
+\begin{align}
+	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2 \\
+	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
+\end{align}
 
-Where each point is first scaled to the size of the detector - for now we assume a 
-radius of $2.5 \times 10^4 \ nm$ - and then translated from its position on the 
-surface of the detector to the main coordinate system. Their orientation is treated 
-much in the same as the detector's orientation, being its -ve position and scaled down 
-by the radial distance from the dimer to the pixel, making it a unit vector. From 
-their position we can also grab their spherical coordinates to create a list of points 
-for mstm to evaluate. In order to accurately describe any detector surface we can use a Householder transformation to get the perpendicular circular surface of any vector.
+By computing the scattering at each point on the detectors surface 
+we can compute the average scattering over the detector. As the 
+detectors are moved further away from the target origin the intensity 
+distribution across the detector surface falls off until near uniform.
 
-This allows us to define the surface of any detector perpendicular to the origin of 
-our coordinate system, we can freely choose the size and resolution of our fibre's. 
-For every pixel we define the spherical coordinates relative to our origin, this is 
-added to a list of $\theta, \phi$ pairs that are evaluated by \textit{mstm}.
+By moving the detectors further back we can treat each detector as single
+point and compute the intensity at the centre of the detector rather than
+consider the full surface.
 

### 2024-09-23 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 \subsection{Influence of a moving beam front on seed crystals}
-\textcolor{red}{Need to write up these results. To summarise the crystal 
-	dissolved around the laser focus instead of growing as the seed crystals 
-	are in the solution and not close to the air-solution interface}
+One potential application of this phenomena would be in using a 
+moving the moving beam front as a method for shaping the final
+crystal morphology. Glycine seed crystals were grown via 
+evaporative crystallisation over night. The resultant crystals
+were as wide as $0.5 mm$ in some cases. Individual crystals were 
+collected and suspended in a water + glycine solution ($S = 1.001$)
+and the laser was switched on.
 

### 2024-09-23 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 with our simulations we set it as the dimer's centre of diffusion.
-The near field scattering from the dimer is shown in 
-fig.~\ref{fig:nf_scattering}.
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{nf_scattering_dimer.png}
-	\caption{Re ${\bf E_z}$ distribution from a symmetric dimer when 
-	irradiated by the probe beam (viewed in the x-y plane). Axis are 
-	given in dimensionless units of $k\hat{x}$ and $k\hat{y}$ respectively.}
-	\label{fig:nf_scattering}
-\end{figure}
-
-\subsection{Detectors and Pixels}
+\subsection{Scattering and Detection}
 Each detector is placed some distance $\bf r$ from the origin, 
@@
 \begin{align}
-	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
+	\begin{pmatrix}
+		x_{fiber} \\ y_{fiber} \\ z_{fiber}
+	\end{pmatrix} = 
+	\begin{pmatrix}
+		rcos(\phi)sin(\theta) \\ rsin(\phi)sin(\theta) \\ rcos(\theta)]
+	\end{pmatrix}
 \end{align}
@@
 \begin{align}
-	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2 \\
+	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2  \label{eq:intensity} 
+	\\
 	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
@@
 By computing the scattering at each point on the detectors surface 
-we can compute the average scattering over the detector. As the 
-detectors are moved further away from the target origin the intensity 
-distribution across the detector surface falls off until near uniform.
+we can compute the average scattering over the detector. From a
+experimental perspective the distance from the particle to the 
+detectors must be considered. The near field scattering can be 
+computed with relative ease using \textit{mstm} as shown in 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{near_field_intensity_side_profile.png}
+	\caption{Intensity distribution (as defined by \eqref{eq:intensity}) 
+		from a symmetric dimer when irradiated by the probe beam 
+		propagating in the $k\hat{z}$ direction (viewed in the 
+		z-y plane). Axis are given in dimensionless units of 
+		$k\hat{z}$ and $k\hat{y}$.}
+	\label{fig:nf_scattering}
+\end{figure}
 
-By moving the detectors further back we can treat each detector as single
-point and compute the intensity at the centre of the detector rather than
-consider the full surface.
+While \textit{mstm} does provide the amplitude matrix it is not
+scaled to account for distance from the target to the detector.
+We amend this by simply scaling according to the inverse square 
+law. As the detector is moved further from the target the 
+intensity distribution across the entire detector will decrease
+substantially. Because of the fall off it is more computationally 7
+efficient to instead consider only the central pixel of the 
+detector rather than across the entire surface. If we consider 
+the difference between the average intensity and the intensity 
+of the central point we can see that after not even $20 \mu$ 
+away from the target particle the average intensity across the 
+surface is within $0.1\%$ of the central pixel. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
+	\caption{Normalised difference between the central pixel intensity
+	$I(\theta, \phi)$ and the average pixel intensity over the surface
+	of the optical fibre $\langle I \rangle$ as a function of distance.
+	4 different optical detectors were tested (spherical coordinates 
+	are given in the legend). While the choice of coordinates is 
+	arbitrary they all demonstrate the same exponential decay.}
+\end{figure}
+
+By moving the detectors further back we can treat each detector 
+as single point and compute the intensity at the centre of the 
+detector rather than consider the full surface. For computing 
+the instantaneous orientation we only consider the intensity 
+at the centre of the detector.
 
 \section{Interpretation of scattering data into orientation estimates}
-Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any 
-point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere 
-to the centre of the smaller sphere. A plane wave probe beam is incident on 
-the trapping laser, is incident on the dimer, generating a scattering pattern
-dependent on the dimer's orientation $I(\hat{\bf s}, \theta, \phi)$ which can be 
-computed using \textit{mstm} \cite{I.Mishchenko1996}. To represent the experimental 
-set up consisting of a set of optical fibres recording scattered light, we choose 
-four sets of spherical angles [$(\theta_1,\ \phi_1), \ (\theta_2,\ \phi_2), \ 
-(\theta_3, \ \phi_3), \ (\theta_4,\ \phi_4)$] and record the calculated intensity 
-at each angle $I(\hat{s},\ \theta_k,\ \phi_k)$.
+We now discuss the methodology for computing the instantaneous 
+orientation of an optically bound particle based on scattering
+detected by a finite grouping of optical fibres.
 
+Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), 
+we can define at any point in time a unit vector $\hat{s}$ 
+pointing from the centre of the larger sphere to the centre 
+of the smaller sphere. A plane wave probe beam incident, 
+is incident on the dimer, generating a scattering pattern
+dependent on the dimer's orientation $I(\hat{\bf s}, 
+\theta, \phi)$ which is computed using \textit{mstm} 
+\cite{I.Mishchenko1996}. To represent the experimental set 
+up consisting of a set of optical fibres recording scattered 
+light, we choose four sets of spherical angles [$(\theta_1,\ 
+\phi_1), \ (\theta_2,\ \phi_2), \ (\theta_3, \ \phi_3), \ 
+(\theta_4,\ \phi_4)$] and record the calculated intensity at 
+each angle $I(\hat{s},\ \theta_k,\ \phi_k)$. 
 \begin{figure}[h!]
@@
 
-Our goal is to determine the orientation of the trapped dimer based on the measured 
-intensity $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate 
-of the dimer's orientation, for the purposes of interpretation of the scattering and 
-optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then 
-use as 'classification categories' in a neural network methodology to map scattering 
-data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} 
-\ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a 
-unit sphere \cite{Reyuthor2006} (Figure~\ref{fig:dimer}b) leading to a maximum 
-nearest-neighbour spacing between two neighbouring reference orientations of 0.895 
-radians. Using MSTM we compute the raw intensities at each of the measurement angles 
-that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary 
-there are several constraining factors that limit our ability to infer useful information 
-from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown 
-of our choice of detection angles. The raw intensities are normalized according to:
+Our goal is to determine the orientation of the trapped dimer 
+based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
+Rather than aim an exact estimate of the dimer's orientation, 
+for the purposes of interpretation of the scattering and 
+optimisation of the measurement setup it is more convenient 
+to discretize the possible orientation space into a number 
+of possible reference orientations. We can then use each as 
+'classification categories' in a neural network methodology 
+to map scattering data to orientation (see below for further 
+discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
+reference orientations $\hat{\bf n}_{\alpha}$  evenly 
+distributed on a unit sphere \cite{Reyuthor2006} 
+(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
+neighbour spacing between two neighbouring reference 
+orientations of 0.895 radians. Using \textit{mstm} we 
+compute the raw intensities at each of the measurement angles 
+that would be generated by a dimer in each reference 
+orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the 
+number and position of detection fibres is technically arbitrary 
+there are several constraining factors that limit our ability 
+to infer useful information from the trapped object, see Section~\ref{sec:brownian} for a detailed breakdown of our 
+choice of detection angles. In order for a neural network
+to properly weight each signal the raw intensities are 
+normalized according to:
 \begin{align}
@@
 
-where the denominator is simply the standard deviation across the set of values
-$I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled 
-signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.
+where the denominator is simply the standard deviation across 
+the set of values $I(\hat{\bf n}, \theta_k)$. Note that the 
+collected scattering signals are not necessarily simply 
+related to their associated reference orientations: as is well 
+known from such examples of the inverse scattering problem. 
+While it is trivial to compute the light scattering pattern 
+for any given particle with any particular characteristic (i.e. 
+size, shape, or orientation), inferring the light scattering 
+from a unknown particle to determine said characteristic is 
+incredibly difficult due to complex mapping between scattering 
+and said characteristic. Even if the orientation space is 
+divided evenly between reference orientations the subsequent 
+signal space ends up being appearing mixed making simple 
+comparisons of signals useless for inferring information on the 
+particle.
 
-Note that the collected scattering signals are not necessarily simply related to 
-their associated reference orientations: as is well known from such examples of 
-the inverse scattering problem. While it is trivial to compute the light scattering 
-pattern for any given particle with any particular characteristic (i.e. size, 
-shape, or orientation), inferring the light scattering from a unknown particle 
-to determine said characteristic is incredibly difficult due to complex mapping 
-between scattering and said characteristic. Even if the orientation space is divided 
-evenly between reference orientation the subsequent signal space ends up being 
-appearing mixed making simple comparisons of signals useless for inferring information 
-on the particle. Shown below is two clusters of orientation vectors and there 
-respective measured scattering signals - the points have been coloured based on 
-their proximity to the centre of their respective cluster. While the orientation 
-space appears tightly packed and ordered the signal space quickly spreads out in 
-an asymmetric fashion. Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
-mapping can intersect itself which only further increases the complexity. While 
-in some instances the mapping between one reference orientation and another is 
-discrete, in other instances the mapping becomes far more complex to discern.
-
+Fig~\ref{fig:mixing} shows two clusters of orientation 
+vectors and their respective signals collected from 3 detectors
+- the points have been coloured based on their proximity 
+to the centre of their respective cluster. While the 
+orientation space appears tightly packed and ordered the 
+signal space quickly spreads out in an asymmetric fashion. 
+Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
+mapping can intersect itself which only further increases 
+the complexity. While in some instances the mapping between 
+one reference orientation and another is discrete, in other 
+instances the mapping becomes far more complex to discern.
+\newpage 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.8\textwidth}
 		\subcaption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.8\textwidth}
 		\subcaption{}
@@
 \end{figure}
+\newpage
+Nevertheless, at least where the uncertainty in signal measurements 
+is low, we can predict the orientation from the scattering scattering 
+by utilising computational techniques such as neural networks. We 
+thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
+dimer's orientation from its light scattering signal. The network 
+was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then 
+using the network to estimate the probability of a given signal coming 
+from a dimer in a given reference orientation. The network's loss 
+function was evaluated and used to improve the estimation, the 
+network being trained until the improvement in the loss function 
+was less than 0.0001.
 
-Nevertheless, at least where the uncertainty in signal measurements is low (see 
-below), we can predict the orientation from the scattering by utilising 
-computational techniques such as neural networks. We thus utilised the Python 
-machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a 
-neural network for identifying the dimer's orientation from its light scattering 
-signal. The network was trained by generating a database of random orientation 
-vectors, calculating the corresponding light scattering signals, and then using 
-the network to estimate the probability of a given signal coming from a dimer in 
-a given reference orientation. The network's loss function was evaluated and used 
-to improve the estimation, the network being trained until the improvement in the 
-loss function was less than 0.0001.
-
-Importantly, the estimation provided by the neural network can be improved further 
-by accounting for any prior information we know about the dimer, utilising Bayesian 
-inference to update the neural network's estimation: 
+Importantly, the estimation provided by the neural network can be 
+improved further by accounting for any prior information we know 
+about the dimer, utilising Bayesian inference to update the neural 
+network's estimation: 
 \begin{align}
@@
 
-where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of 
-the distributions of particle orientations and instantaneous signals, respectively.
-\textit{Without} any prior evidence we must assume that the orientation prior of 
-the dimer $p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
-dimer's possible current orientation from knowledge of previous measurements can 
-be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$. The latter prior 
-$p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$). This is 
-given by taking the discrete integral over the collection of reference orientations:
+where $p(\hat{\bf n}_\alpha)$ and $p(y_k(\hat{\bf s}))$ are the prior 
+estimates of the distributions of particle orientations and 
+instantaneous signals, respectively. \textit{Without} any prior 
+evidence we must assume that the orientation prior of the dimer 
+$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
+dimer's possible current orientation from knowledge of previous 
+measurements can be used to inform our estimate of $p(\hat{\bf 
+n}_{\alpha})$. The latter prior $p(y_k(\hat{s}))$ is the probability of 
+measuring a signal ($y_1$, $y_2$, $y_3$). This is given by taking 
+the discrete integral over the collection of reference orientations:
 \begin{align}
-	p(y_1, y_2, y_3, y_4)
+	p(y_k(\hat{\bf s}))
 	=
@@
 
-From \eqref{eq:bayes} we obtain the key result, a mass probability distribution 
-denoting the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ 
-given a measured signal ($y_1$, $y_2$, $y_3$), \textit{i.e.} an estimated mapping 
-from scattering measurement to orientation estimate.
+From \eqref{eq:bayes} we obtain the key result, a mass probability distribution denoting the probability that our dimer is in 
+orientation $\hat{\bf n}_{\alpha}$ given a measured signal ($y_1$, 
+$y_2$, $y_3$), \textit{i.e.} an estimated mapping from scattering 
+measurement to orientation estimate.
 
@@
 
-We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
-Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
+We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
+Brownian OT combines MSTM \cite{Mackowski1996} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
 shaped sphere clusters.
@@
 
-One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering.
-Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step.
-To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness.
-Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$.
-Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:
+One reason we observe such large jumps in the estimated 
+orientation is simply because the neural network is assuming
+that the signals are unrelated from one another. As demonstrated
+by fig~\ref{fig:mixing} the orientation and signal spaces are 
+uncorrelated with one another which results in the neural network
+to assign orientations incorrectly. Combining this fact with 
+use of a uniform prior ($p(\hat{\bf n}_\alpha)$), there 
+is no constraint on how much estimated orientation can change 
+from time-step to time-step.
 
+To improve the estimation we can therefore use knowledge of the 
+physical limitations of the object in the trap and its dynamics. 
+Imposing a more physically grounded prior, accounting in this case 
+for the fact that the rotational motion of the dimer is limited by 
+the rotational trap stiffness ($\kappa_r$). Here the prior of the 
+reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at 
+each time step as a Boltzmann distribution of the physical distance 
+between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and 
+each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are 
+reweighing our estimation based on the size of rotation, with smaller movements being favoured over large movements:
 \begin{align}
 	p(\hat{\bf n}_\alpha)
-	&=\frac{e^{\beta (\hat{\bf n}_\alpha 
-			\cdot \hat{\bf n}_{est}(t-\Delta t))}}
+	&=\frac{e^{\hat{\bf n}_\alpha 
+			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
 	{\sum_{\alpha=1}^{n_{\rm ref}}
-		e^{\beta (\hat{\bf n}_\alpha 
+		e^{\hat{\bf n}_\alpha 
 			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
@@
 \end{align}
-Here $\beta$ is a weighting factor describing the dimer's freedom of motion within the trap.
+
 As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.

### 2024-09-23 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 dimer's orientation from its light scattering signal. The network 
-was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then 
+was trained by generating a database of random orientation vectors, 
+calculating the corresponding light scattering signals, and then 
 using the network to estimate the probability of a given signal coming 
@@
 
-From \eqref{eq:bayes} we obtain the key result, a mass probability distribution denoting the probability that our dimer is in 
+From \eqref{eq:bayes} we obtain the key result, a mass probability 
+distribution denoting the probability that our dimer is in 
 orientation $\hat{\bf n}_{\alpha}$ given a measured signal ($y_1$, 
@@
 \label{sec:divergence}
-To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$.
-MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$.
-Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}.
-An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
+Using \eqref{eq:bayes} we get a snapshot estimation of the instantaneous
+orientation of the trapped dimer based on a single measurement from a 
+collection of detector signals. Of course in reality these detectors 
+will be returning a discrete time series representative of the dimer's
+trajectory within the optical trap.
+
+Using the simulation code from \cite{Vigilante2020} we can replicate 
+this trajectory and furthermore use \textit{mstm}to calculate the far field 
+intensity produced by said dimer. At each time step we measure the signal
+produced at each measurement angle and use our neural network to calculate
+$p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha, t)$.
+
+Evaluating how well Eq.~\eqref{eq:bayes} performs in predicting $\hat{\bf n}$
+is difficult because we have discretised the orientation space meaning 
+there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method allows us to compare the divergence 
+between two probability, namely Eq.~\eqref{eq:bayes} and a case scenario.
+\begin{align}
+	D_{KL}(P\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)) = \sum P(t) \frac{P(t)}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)} 
+\end{align}
+
+Where $P(t)$ is the case scenario we are comparing against. In an ideal
+case our model would correctly predict the dimer's orientation with 
+100\% accuracy:
 \begin{align}
 	\label{eq:best}
-	p_{best} = 
+	p_{best} = P(t) =
 	\begin{cases}
 		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
-		0 & \text{anywhere else}
+		0 & \text{when $\hat{\bf n}_\alpha$ $\neq$ $\hat{\bf n}_{best}$}
 	\end{cases}
 \end{align}
-In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
-probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
+
+In reality the distribution from Eq.~\eqref{eq:bayes} will assign some 
+non-zero probability to every reference orientation. Therefore the 
+divergence of Eq.~\eqref{eq:bayes} at each time step is given as:
 \begin{align}
-	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
-	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
+	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
+	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
 	\right]
@@
 \end{align}
-where a larger value of $K_l$ indicates that our model is less confident in its
-prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
-in the estimated dimer orientation probability --- a distribution strongly peaked at 
-some value would give us more confidence in that value than a near-uniform distribution 
-where the scattering measurement could imply a wide range of possible orientations --- 
-but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.
+
+where a larger value of $D_{KL}$ indicates that our model is less confident 
+in its prediction of the dimer's orientation. The divergence $D_{KL}$ thus 
+illustrates the 'spread' in the estimated dimer orientation probability 
+-- a distribution strongly peaked at some value would give us more 
+confidence in that value than a near-uniform distribution where the 
+scattering measurement could imply a wide range of possible orientations 
+-- but it does not directly indicate our estimates actual accuracy. 
+In order to reflect the accuracy of each prediction by using a weighting
+function such as:
+\begin{align}
+	\omega_\alpha = \frac{1}{2} (\hat{\bf n}_{best} \cdot \hat{\bf n}_\alpha+1)
+\end{align}
+
+Which is just the normalised dot product between our estimate and the
+best possible estimate, so for highly inaccurate estimations $D_{KL}$
+is far greater than if we narrowly missed the $\hat{\bf n}_{best}$. With
+this in mind the divergence for a single measurement is:
+
+\begin{align}
+	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
+	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)\omega_\alpha}
+	\right]
+\end{align}
+
+Later we will show how we set a baseline for evaluating the performance 
+of \eqref{eq:bayes} across an entire trajectory and how it handles common
+errors in a static light scattering setup.
 
@@
 
-We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap.
-Brownian OT combines MSTM \cite{Mackowski1996} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
-shaped sphere clusters.
-We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid.
-MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}.
-The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}.
-We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$.
-We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}.
-We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.
+We use the Brownian OT package developed by Fung~\textit{et~al} 
+\cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. We simulate the 
+motion of a dimer trapped in a highly focused Gaussian beam by 
+calculating the optical forces imparted by the laser, and the 
+Brownian force due to the surrounding fluid. MSTM provides the 
+necessary T-matrix to compute the optical force via \textit{ott}. 
+The Brownian force is found by computing the dimer's diffusion 
+tensor according to the analytical solutions provided by Nir and 
+Acrivos \cite{Nir1973}. We simulated a polystyrene dimer ($n = 
+1.59$) in a suspension of water ($n_{med} = 1.33$) over the 
+course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} 
+\ s$. We placed the dimer 4 microns below the trap focus at an 
+angle $30^{\circ}$ from the horizontal, the resulting trajectory 
+is shown below in Sec~\ref{sec:motion}. We chose these initial 
+parameters because it demonstrates our model's performance in non 
+steady state conditions.
 
@@
 
-In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
+In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
 However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state.
@@
 This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
-We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
+We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Reyuthor2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
 While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.

### 2024-10-04 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 04-10-2024

@@
 final shape. Whereas the former is more concerned with which factors 
-contribute to its formation in order to control crystal growth.
+contribute to its formation in order to control crystal growth.SW
 
-The focus of this research is partly based around investigating the impact 
-of a focused electromagnetic (EM) field can influence the nucleation process 
-by inducing shear flow within the bulk phase. This in hope that the 
-localised EM field can be used to study different phenomena in nucleation 
-in order to develop a better understanding of the underlying mechanisms.
+In this chapter we will outline our current understanding of nucleation,
+why there is a gap in the literature, and why there is still a need for
+local control of nucleation events. Furthermore, we will also highlight
+recent developments involving the use of optical tweezers (and other 
+laser based methods) in order to develop our understanding of nucleation.
+Lastly we highlight the connection between optical tweezer's ability to 
+probe fluid viscosity via rotational motion, and the known link between
+nucleation and fluid flow, to propose a novel potential method of 
+creating localised nucleation events.
 
@@
 after a given time, giving us a Poisson probability distribution.
-
 \begin{align}
@@
 seed crystal inducing further nucleation within the bulk solution
-\cite{Botsaris1976}, in absence of a seed crystal nucleation is
-near impossible. The research of Secondary Nucleation is fascinating
-for its own merits; is it speculated that the supersaturation 
-barrier is non-existent \cite{Cashmore2022}, and furthermore there 
-are a multitude of mechanisms that can induce secondary 
-nucleation making it imperative that researchers can control the
-nucleation rate at an industrial scale.
-
+\cite{Botsaris1976}. The research of Secondary nucleation is 
+fascinating for its own merits; is it speculated that the 
+supersaturation barrier is non-existent \cite{Cashmore2022}, 
+Furthermore control of secondary nucleation events are crucial 
+for ensuring reliable industrial crystalliser performance, 
+fig.~\ref{fig:secondary} depicts an attempt of classifying  
+every possible mechanism that could lead to secondary nucleation.
 \begin{figure}[h!]
-	\includegraphics[width=\linewidth]{secondary_nucleation.jpg}
-	\caption{Secondary Nucleation mechanisms, classified by Agrawal and Paterson \cite{Agrawal2015}}
+	\centering
+	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
+	\caption{Secondary Nucleation mechanisms, classified by Agrawal 
+		and Paterson \cite{Agrawal2015}}
+	\label{fig:secondary}
 \end{figure}
 
+Its clear that preventing secondary nucleation is impossible
+with so many possible mechanism, modelling its contribution
+to the nucleation rate has been the area of research for some
+time now \cite{Cashmore2022, Flannigan2023}. In most cases 
+secondary nucleation is actually desirable for large scale 
+production as primary nucleation is a stochastic process.
+
 \section{Nucleation Theories}
+A number of potential theories have been proposed to explain 
+how nucleation occurs at a microscopic scale. In doing so 
+we could potentially predict the expected yield of a given
+crystalliser based on the initial conditions of the 
+solution. Outlined below are some of the most popular theories
+currently used to describe nucleation.
+
 \subsection{Classical Nucleation Theory (CNT)}
@@
 \subsection{Non-classical Nucleation} 
-
-The current research into two step nucleation (or now more commonly referred 
-to as multi-step nucleation theory) is developing a robust framework to 
-describe what nucleation pathway will occur given the initial conditions. 
-Reviews of all currently documented nucleation pathways identified highlighted 
-the fact the need for a the development of \textit{in situ} techniques that 
-can induce nucleation locally but can also reliably identify the nucleation 
-pathway across a broad range of experimental conditions \cite{Karthika2016, Fu2021}.
+The current research into two step nucleation (or now more commonly 
+referred to as multi-step nucleation theory) is developing a robust 
+framework to describe what nucleation pathway will occur given the 
+initial conditions. Reviews of all currently documented nucleation 
+pathways identified highlighted the fact the need for a the 
+development of \textit{in situ} techniques that can induce nucleation 
+locally but can also reliably identify the nucleation pathway across 
+a broad range of experimental conditions \cite{Karthika2016, Fu2021}.
 
@@
 \subsection{Gaussian Beams}
-A Gaussian Beam is a simplistic approximation of a laser beam, in short
-it is an unbounded plane wave whose intensity of the electromagnetic 
-field falls off from the centre similar to a normal distribution. This 
-is perfectly fine for beams with a divergence angle of $0^{\circ}$, but 
-later studies into focused beams has shown that the assumption of a 
-Gaussian intensity profile is inaccurate. A study conducted by Lock and Gouesbet found that as the Gaussian beam was focused the fall off in 
-the electric field would consequentially induce additional electromagnetic fields \cite{Lock1994}.
-As a result the shape of the beam profile diverges from a Gaussian profile for focused beams.
-Regardless, for optical trapping experiments it is useful to approximate a Gaussian beam as many descriptions
-of lasers are simply combinations of plane waves (with varying amplitudes
-and wave vectors). More information on electromagnetic theory is covered in 
-chapter 2 but for now any mention of a focused laser can be thought of as a
-Gaussian beam, unless specified otherwise.
+A Gaussian Beam is a simplistic approximation of a laser beam, in 
+short it is an unbounded plane wave whose intensity of the 
+electromagnetic field falls off from the centre similar to a normal distribution. This is perfectly fine for beams with a divergence 
+angle of $0^{\circ}$, but later studies into focused beams showed 
+that the assumption of a Gaussian intensity profile is inaccurate. 
+A study conducted by Lock and Gouesbet found that as the Gaussian 
+beam was focused the fall off in the electric field would 
+consequentially induce additional electromagnetic fields 
+\cite{Lock1994}. As a result the shape of the beam profile diverges 
+from a Gaussian profile for focused beams. Regardless, for optical 
+trapping experiments it is useful to assume a near Gaussian beam as 
+many descriptions of lasers are simply combinations of plane waves 
+(with varying amplitudes and wave vectors).  More information on 
+electromagnetic theory is covered in chapter 2 but for now any 
+mention of a focused laser can be thought of as a Gaussian beam, 
+unless specified otherwise.
 
@@
 \section{Significance of Thesis}
-As I have hoped to make clear in the above introduction, the current 
-state nucleation theory is rather cumbersome at a micro-level. Models
-such as CNT and multi-step nucleation are not sufficient for describing
-the myriad of potential pathways nucleation can go down. As suggested
-by some review articles, the best way to address this is by developing 
-\textit{in-situ} methods that can study the pre nucleation phase in 
-greater detail \cite{Fu2021}. Furthermore, the ability to localise 
-nucleation allows for better characterisation of the kinetics of crystal 
-growth. Laser induced nucleation stands to be an ideal method to study
-nucleation, as the laser output can be concentrated to a small area. 
-In this project we want to consider another application of optical 
-tweezers, their ability to rotate micro-rotors to generate fluid flow.
-Such a method has not been tried before and would allow for us to 
-consider how fluid flow impacts nucleation dynamics at the micro-level.
+As I have hoped to make clear in the above introduction, the 
+current state nucleation theory is rather cumbersome at a 
+micro-level. Models such as CNT and multi-step nucleation are 
+not sufficient for describing the myriad of potential pathways 
+nucleation can go down. As suggested by some review articles, 
+the best way to address this is by developing \textit{in-situ} 
+methods that can study the pre nucleation phase in greater detail \cite{Fu2021}. Furthermore, the ability to localise nucleation 
+allows for better characterisation of the kinetics of crystal 
+growth. Laser induced nucleation stands to be an ideal method to 
+study nucleation, as the laser output can be concentrated to a 
+small area \cite{Korede2023}. This fine control over the local 
+fluid would allow for individual nucleation events to be
+characterised as a factor of the local fluid properties. One 
+challenge lies in the fact that its clear that the local fluid 
+properties have a direct influence on the likelihood of laser
+induced nucleation from occurring \cite{Korede2023, Ward2016,
+Yuyama2012, Liao2022}.
+
+A way around this would be to try and induce the local fluid 
+without directly relying on the electromagnetic field. Optical 
+tweezers can reliably do so already by applying an optical 
+torque to a trapped particle in order to induce fluid flow 
+\cite{Bishop2004, RobertsonAnderson2018}. It's already well
+documented that shearing will enhance nucleation events at a
+macro-level \cite{Debuysschere2023}. Micro-rheological studies 
+using optical tweezing have been more interested in probing 
+the local viscosity rather than try and use it as a means of 
+shearing the fluid to induce crystal growth.
 
 \section{Overview}
-Overall the aim of the PhD is to study viability of using micro-rotors to
-generate localised fluid flow around the beam focus. The results are reported
-in chapter 3, this is then succeeded by experimental work where we instead 
-using a galvano-mirror to generate shear flow. While overall unsuccessful
-the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus. This presents a new insights for controlling
-and studying the growth of a newly formed nucleus by precise movement of the
-trapping focus.
+Overall the aim of the PhD is to study viability of using 
+micro-rotors to generate localised fluid flow around the 
+beam focus. The results are reported in chapter 3, this is 
+then succeeded by experimental work where we instead using 
+a galvano-mirror to generate shear flow. While overall 
+unsuccessful the addition of a moving beam focus showed 
+that the growth of a nucleus can be localised around the 
+trap focus. This presents a new insights for controlling
+and studying the growth of a newly formed nucleus by 
+precise movement of the trapping focus.
 
-In addition, computer simulations studying the dynamics of asymmetric dimers 
-and their applications as complex micro-rotors are discussed in chapter 4. 
-The latter results are interesting as a potential insight into how optical
-torque is transferred to anisotropic objects.
+The latter chapters cover computer simulations into the 
+behaviour of microscopic spherical dimers in an optical
+trap. Prior research into dimers using back focal plane
+interferometry has mostly considered their trapping 
+behaviour to be similar to a single sphere but with a 
+difference in trapping strength. Computer simulations
+reveal a host of new behaviours dependent on the dimer's
+size, orientation, proximity to the trapping focus, and 
+even the polarisation of the trapping beam. The latter in
+particular suggests that multi-spherical particles can act
+as sophisticated micro-rotors.
 
-Lastly two novel methods of measuring rotational motion are discussed in 
-chapter 5; firstly via a novel detection fibre method that allows for 
-instantaneous measurements of the orientational behaviour of optically 
-trapped ellipsoids/dimers; and secondly we create a simulative quadrant 
-photo diode that replicates laboratory results, utilising linear regression 
-techniques we measure the change in orientation in order to measure the 
-optical torque applied to a non-birefringent particle.
+However, this raises a its own host of experimental challenges,
+namely how do we characterise the behaviour of an arbitrary
+particle. Relying on current characterisation techniques is not
+possible as they are predicated on the trapping object to behave
+like an isolated sphere. Two novel methods of measuring rotational 
+motion are discussed in chapter 5; firstly via a novel detection 
+fibre method that allows for instantaneous measurements of the 
+orientational behaviour of optically trapped ellipsoids/dimers; 
+and secondly we create a simulative quadrant photo diode that 
+replicates laboratory results, utilising linear regression 
+techniques we measure the change in orientation in order to 
+measure the optical torque applied to a non-birefringent particle.

### 2024-10-04 Latex_files/Thesis/chapters/chapter 2/main.tex
updates 04-10-2024

@@
 materials which is directed towards the point of maximum intensity 
-(for a simple Gaussian beam this would be at it's focal point) \cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
+(for a simple Gaussian beam this would be at it's focal point) 
+\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
 higher numerical aperture lenses in order to increase the intensity, 
@@
 of propagation \cite{Capitanio2002} The equilibrium position is found 
-when the gradient force far exceeds the both the scattering force, 
+when the gradient force far exceeds both the the scattering force, 
 and the thermal energy of the particle.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Harmonic Traps}
@@
 	\centering
-	\includegraphics[width=\linewidth]{sphere_force_curve.png}
+	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
 	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
@@
 optical forces is dependent on the ratio between the particle size 
-and the trapping wavelength. Depending on this ratio, the optical 
-forces can be described differently.
+($a$) and the trapping wavelength ($\lambda$). The scattering theory 
+for computing the optical force is dependent on this ratio.  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Lornez-Mie Theory}
+
+The Lorenz-Mie theory provides an exact solution to the Maxwell 
+equations for the scattering caused by an isotropic sphere. This 
+theory describes the scattered wave given off by a dielectric 
+sphere when incident by a plane wave as a summation of partial 
+spherical waves. For any spherical wave the vector fields must 
+solve the Helmholtz wave equation given by:
+\begin{align}
+	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
+\end{align}
+
+Where $k$ is the wave number of the electromagnetic radiation 
+($k = 2\pi/\lambda$). This combined with the constraints of 
+Maxwell's equations leaves very few exact solutions apart from 
+spherical or planar waves. For a laser of arbitrary shape the 
+electric field can be described as a summation of partial 
+spherical waves multiplied by an expansion coefficient. Depending 
+on the beam in question the expansion coefficients can be tailored
+to match the desired beam shape. For a Gaussian beam interacting 
+with a particle of volume $V$, the incident, internal, and 
+scattered fields are given as \cite{RanhaNeves2019}: 
+\begin{align}
+	\label{eq:incident}
+	{\bf E}_{\rm inc}({\bf r})
+	&=
+	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
+	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
+	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
+	\\
+	\label{eq:internal}
+	{\bf E}_{\rm int}({\bf r})
+	&=
+	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
+	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
+	\right]
+	\\
+	\label{eq:scattered}
+	{\bf E}_{\rm scat}({\bf r})
+	&=
+	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
+	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
+	\right] 
+\end{align}
+
+where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
+are the expansion coefficients of each of the fields, and ${\bf M}_
+{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
+spherical harmonics respectively. The superscripts denote whether 
+the field is inward [(1)] or outward propagating [(3)]. By linearity 
+of Maxwell's equations, its possible to relate the expansion 
+coefficients of the scattered field directly to those of the incident 
+field. As such we can write the scattering coefficients in terms of 
+their incident components and the target's 'T-matrix'.
+\begin{align}
+	\label{eq:t_matrix}
+	\begin{pmatrix}
+		q_{mn} \\
+		p_{mn} 
+	\end{pmatrix}
+	= \bold{T} 
+	\begin{pmatrix}
+		a_{mn} \\
+		b_{mn}
+	\end{pmatrix}
+	= \begin{pmatrix}
+		T_{11} & T_{12} \\
+		T_{21} & T_{22}
+	\end{pmatrix}
+	\begin{pmatrix}
+		a_{mn} \\
+		b_{mn}
+	\end{pmatrix}
+\end{align}
+
+Where the matrix \textbf{T} accounts for the targets shape, and 
+refractive index. The $T$-matrix method can be easily 
+applied to any solution of Maxwell's equations, for example Eq.~\eqref{eq:scattered} can be instead be rewritten as:
+\begin{align}
+{\bf E}_{\rm scat}({\bf r})
+&=
+E_0 \sum_{n}^\infty \sum_{m=-n}^n 
+\begin{pmatrix}
+	T_{11} & T_{12} \\
+	T_{21} & T_{22}
+\end{pmatrix}
+\begin{pmatrix}
+	a_{mn} \\
+	b_{mn}
+\end{pmatrix}
+\begin{pmatrix}
+	{\bf N}_{mn}^{(3)} & {\bf M}_{mn}^{(3)}
+\end{pmatrix}
+\end{align}
+
+In order to compute the optical force and torque applied to a 
+given object one simply needs to know the momentum carried by 
+the incident and scattered fields. So long as we know the 
+expansion coefficients of the incident beam we can find the 
+scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
+compute the force. This project utilises the results from 
+Farsund \cite{Farsund1996} but in the form used by Crichton \cite{Crichton2000THEMD}. As an example the linear momentum 
+carried by any beam can be computed via:
+\begin{equation}
+	\label{eq:linear_momentum_in}
+	\begin{split}
+		\langle P_{z,\ in} \rangle
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+And the linear momentum carried by the scattered beam is given by:
+\begin{equation}
+	\label{eq:linear_momentum_out}
+	\begin{split}
+		\langle P_{z,\ out} \rangle
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+Therefore the total force imparted to the trapped particle is thus:
+\begin{align}
+	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
+	\label{eq:Fz}
+\end{align}
+
+\noindent
+Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
+coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
+Similar expressions are shown in Sec.~\ref{sec:simulations} for 
+calculating the optical torque. Lorenz-Mie theory can be applied 
+to describe the scattering from any particle regardless of size, 
+though as the size of the target particle changes - relative to 
+the wavelength of light - the theory describing the electromagnetic 
+fields are better easier to approximate by alternative theories. 
+When $a \gg \lambda$ the optical force is best described by the 
+Ray-Optics model. In the opposite case the target particle is 
+better approximated as a single electric dipole as described by 
+the Rayleigh approximation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The Ray-Optics model is the simplest to understand, this theory models 
-the laser as a collection of individual 'rays' that propagate and are 
-refracted according to Snell's Law. Based on the change in direction 
-momentum is transferred to the target particle; with rays closest to 
-the centre of the beam having greater intensity than those rays at 
-the very edge of the beam. Consider a particle struck by two rays in 
-a Gaussian beam, one coming close to the centre, and the other ray 
-coming from the edge. As each ray is refracted by the target sphere a 
-force is imparted onto it, the total force imparted is given by:
+the electromagnetic field as a collection of individual 'rays' that 
+propagate and are refracted according to Snell's Law. Based on the 
+change in direction momentum is transferred to the target particle; 
+with rays closest to the centre of the beam having greater intensity 
+than those rays at the very edge of the beam. Consider a particle 
+struck by two rays in a Gaussian beam, one coming close to the centre, 
+and the other ray coming from the edge. As each ray is refracted by 
+the target sphere a force is imparted onto it, the total force 
+imparted is given by:
 \begin{align}
@@
 
+\noindent 
 where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Lornez-Mie Theory}
-
-The Lorenz-Mie theory provides an exact solution to the Maxwell 
-equations for the scattering caused by an isotropic sphere. This 
-theory describes the scattered wave given off by a dielectric 
-sphere when incident by a plane wave as a summation of partial 
-spherical waves. For any spherical wave the vector fields must 
-solve the Helmholtz wave equation given by:
-\begin{align}
-	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
-\end{align}
-
-Where $k$ is the wave number of the electromagnetic radiation 
-($k = 2\pi/\lambda$). This combined with the constraints of 
-Maxwell's equations leaves very few exact solutions apart from 
-spherical or planar waves, both of which can be converted between 
-with relative ease. For example, a plane wave electric field can 
-be expanded into spherical harmonics and likewise any spherical 
-wave can be described as a combination of plane waves offset from 
-one another. For a single plane wave the incident, internal, and 
-scattered fields are given as: 
-\begin{align}
-	\label{eq:incident}
-	{\bf E}_{\rm inc}({\bf r})
-	&=
-    E_0 \sum_{nm}^\infty \left[
-    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
-    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
-  \\
-  \label{eq:internal}
-  {\bf E}_{\rm int}({\bf r})
-  &=
-    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
-    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
-    \right]
-  \\
-  \label{eq:scattered}
-  {\bf E}_{\rm scat}({\bf r})
-  &=
-    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
-      -i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
-    \right] 
-    \\
-	\mathbf{E}_{tot}({\bf r}) &= 
-	\begin{cases}
-		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
-		\\ 
-		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
-	\end{cases}
-\end{align}
-
-where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
-are the expansion coefficients of each of the fields, and $M_{nm}$ 
-and $N_{nm}$ are the vector spherical harmonics. The superscripts
-denote whether the field is inward or outward propagating; so a 
-superscript (1) is inward propagating, and (2) is outward propagating.  
-For the incident field computing its expansion coefficients is possible 
-via analytical methods and are completely dependent on the beam 
-conditions imposed by the user. However, computing the expansion 
-coefficients of the internal and scattered fields is far more 
-tedious depending on the shape of the target and what properties 
-of the scattered field are desired --- see sec.~\ref{sec:scattering}.  
-From an optical trapping perspective the force exerted by a focused 
-electric field can be found by computing the Maxwell stress tensor 
-which only requires the total magnitude of the incident and scattered 
-fields. An alternative method is to compute the momentum carried by 
-the incident and scattered beams and taking the difference between 
-to measure the momentum transferred to the particle. This project 
-utilises the results from Farsund \cite{Farsund1996} but in the 
-form used by Crichton \cite{Crichton2000THEMD}. As an example the 
-linear momentum carried by any beam can be computed via:
-\begin{equation}
-	\label{eq:linear_momentum_in}
-	\begin{split}
-		\langle P_{z,\ in} \rangle
-		&=
-		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
-		\\ 
-		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-		\\
-		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
-	\end{split}
-\end{equation}
-
-And the linear momentum carried by the scattered beam is given by:
-\begin{equation}
-	\label{eq:linear_momentum_out}
-	\begin{split}
-		\langle P_{z,\ out} \rangle
-		&=
-		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
-		\\ 
-		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-		\\
-		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
-	\end{split}
-\end{equation}
-
-Therefore the total force imparted to the trapped particle is thus:
-\begin{align}
-	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
-\end{align}
-
-Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
-coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
-Similar expressions are shown in \ref{sec:simulations} for 
-calculating the optical torque. Lorenz-Mie theory can be applied 
-to describe the scattering from any particle regardless of size, 
-though as the size of the target particle changes - relative to 
-the light being scattered - the theory describing the scattered 
-light is better approximated by alternative theories. For 
-particles that are far larger than the wavelength of light the 
-scattering is best described by the Ray-Optics theory. In the 
-opposite case the target particle is better approximated as a 
-single electric dipole as described by the Rayleigh scattering 
-field.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Rayleigh Scattering}
 
-The Rayleigh approximation is the limiting approximation for 
-describing a particles scattering in an electromagnetic field 
-whose wavelength is several times greater than the particle's 
-size. The underlying theory is that a dielectric sphere can be 
-treated as a dipole while in the presence of the electromagnetic 
-field. In which case the scattering force is given simply by the 
-scattering of the induced dipole, and the gradient force is due 
-to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
-principle Cartesian axis are described by Harada et al 
-\cite{YasuhiroHarada1996} in MKS units as a restorative rectangular 
-force field:
+The Rayleigh approximation is for describing a particle when 
+$a \ll \lambda$. The underlying theory is that a dielectric sphere 
+can be treated as a dipole while in the presence of the 
+electromagnetic field. In which case the scattering force is 
+given simply by the scattering of the induced dipole, and the 
+gradient force is due to the Lorentz force \cite{Gordon1973}. 
+The gradient forces in the principle Cartesian axis are 
+described by Harada et al \cite{YasuhiroHarada1996} in MKS units 
+as a restorative rectangular force field:
 \begin{align}
@@
     \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
-  \\
+  \label{eq:Rayleigh_start}\\
   F_{grad,y}
@@
   &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
     \nonumber \times I(r)
@@
   \\
+  \label{eq:Rayleigh_end}
   \text{where:}
@@
 
-Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
-the beam waist, and $a$ is the radius of the particle. The scattering force 
-however, is dependent on the effective scattering cross-sectional area. 
+\noindent
+Where $m$ is the relative refractive index ($n_1/n_2$), $n_2$ is 
+the particle's refractive index, $\omega_0$ is the beam waist, 
+$a$ is the radius of the particle, and $\tilde{x}, \tilde{y}, 
+\tilde{z}$ are simply the $x, y, z$ coordinates but scaled by the 
+beam radius. The scattering force however, is dependent on the 
+effective scattering cross-sectional area. 
 \begin{align}
@@
 
-The Rayleigh regime allows for easy computation of the gradient and 
-scattering forces due to the assumption that the particle is a point 
-dipole. Unlike Lorenz-Mie scattering, there is no need to compute the 
-scattered field (see \eqref{eq:scattered}) for an individual sphere. 
-Higher complexity scattering problems can be simplified by subdividing 
-the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
-in which case the scattered field is now necessary. As the target 
-particle gets larger this assumption fails to accurately describe 
-the trapping force. For particle sizes close to the laser wavelength 
-the trapping force is best described by Lorenz-Mie theory.
+There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
+\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
+Rayleigh regime is not concerned with the state of the scattered
+field. This is because if the particle is approximated by a point
+dipole we can ignore the momentum transfer calculation and simply
+compute the Lorenz force for a dipole in an electric field. This is 
+not to say that the scattered field is not influenced by the particle, 
+the scattered field can be related to the incident field like so:
+\begin{align}
+	\label{eq:Rayleigh_scattering}
+	\begin{pmatrix}
+		{\bf E}^s_\perp \\
+		{\bf E}^s_\parallel 
+	\end{pmatrix}
+	= \begin{pmatrix}
+		S_{11} & 0 \\
+		0 & S_{22}
+	\end{pmatrix}
+	\begin{pmatrix}
+		{\bf E}^i_\perp \\
+		{\bf E}^i_\parallel
+	\end{pmatrix}
+\end{align}
+
+\noindent
+Where $S_{11}$, and $S_{22}$ are elements of the amplitude scattering 
+matrix, this is analogous to a single plane wave scattering off a 
+sphere. The difference between Eq.~\eqref{eq:Rayleigh_scattering} and
+Eq.~\eqref{eq:t_matrix} is that the $T$-matrix accounts for the 
+scattering from multiple plane waves that make up the laser. Because the 
+particle can be approximated by a point dipole the scattering field is 
+linearly related to the scattered field. While not important for optical 
+force calculation higher complexity scattering problems can be simplified 
+by subdividing the particle into discrete dipoles (see 
+Sec~\ref{sec:scattering}), in which case the scattered field is now 
+significant. As the target particle gets larger this assumption fails 
+to accurately describe the trapping force \cite{Li2021}. For particle 
+sizes close to the laser wavelength the trapping force is best described 
+by Lorenz-Mie theory.
 
@@
 \subsection{T-matrix Method}
-The $T$-matrix method was first developed by Peter Waterman with his
-research into acoustic wave scattering \cite{Waterman1969}, this 
-would later be extended to electromagnetic waves. Sometimes  
-referred to as the extended boundary condition method (ECBM), 
-the method replaces the scatterer with a series of surface currents
+
+The $T$-matrix method was first developed by Peter Waterman with 
+his research into acoustic wave scattering \cite{Waterman1969}, 
+this would later be extended to electromagnetic waves. Sometimes 
+referred to as the extended boundary condition method (ECBM), the 
+method replaces the scatterer with a series of surface currents 
 over the targets surface. These currents are chosen so that the 
-electromagnetic field outside is identical to the original problem 
-\cite{Wriedt1998}. This choice of currents negates the need to 
-compute the internal fields which reduces the scattering problem 
-to a system of linear equations relating the incident beam coefficients 
-to the scattered beam coefficients. The relationship between each can 
-be summarised as:
-\begin{align}
-	\label{eq:t_matrix}
-	\begin{pmatrix}
-		q_{mn} \\
-		p_{mn} 
-	\end{pmatrix}
-	= \bold{T} 
-	\begin{pmatrix}
-		a_{mn} \\
-		b_{mn}
-	\end{pmatrix}
-	= \begin{pmatrix}
-		T_{11} \ T_{12} \\
-		T_{21} \ T_{22}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{mn} \\
-		b_{mn}
-	\end{pmatrix}
-\end{align}
-
-Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, 
-and $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. 
+electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the fields one can derive the 
+individual elements of \textbf{T} until a desired accuracy is 
+achieved.
+
 For this project we utilised an extension of the ECBM known as 
-multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
-\cite{Mackowski2011} the computational code computes the scattered 
-field from each sphere within the target cluster and the incident 
-field itself. The scattering between spheres slowly converges to a 
-final result but modern codes truncate the calculations to an upper 
-number of iterations (denoted by $N_{max}$) to fit a desired accuracy 
-and computational time.
+multi-sphere $T$-matrix method (\textit{mstm}). Developed by 
+Mackowski \cite{Mackowski2011} the computational code computes 
+the scattered field from each sphere within the target cluster 
+and the incident field itself. The scattering between spheres 
+will converge to a final result but modern codes truncate 
+the calculations to an upper number of iterations (denoted by 
+$N_{max}$) to fit a desired accuracy and computational time. 
+By centring the expansion at the origin of each sphere the 
+individual sphere $T$-matrices can be computed, then using the
+addition theorem, the target centred $T$-matrix is computed.
 
@@
 \subsection{Discrete Dipole Approximation}
-The discrete dipole approximation (DDA) is a general method that can 
-be applied to the scattering from particles of arbitrary composition
-and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
-the DDA method approximates the particle as being constructed of dipoles.
-Each dipole interacts with both the incident field and the scattered fields
-from every other dipole surrounding it. The resulting scattered field is 
-identical to the scattered field produced by direct integration of 
-Eq.~\eqref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
-The integral form for the internal electric field inside a scatterer is 
-given as:
-\begin{align}
-	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
+The discrete dipole approximation (DDA) is a general method that 
+can be applied to the scattering from particles of arbitrary 
+composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
+being constructed of discrete dipoles. Each dipole interacts with 
+both the incident field and the scattered fields from every other 
+dipole surrounding it. The resulting scattered field is identical 
+to the scattered field produced by direct integration of Eq.~\eqref{eq:internal} throughout the full particle volume 
+\cite{Goedecke1988}. The integral form for the total electric field 
+inside a scatterer is given as \cite{Wriedt1998}:
+\begin{align}
+	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  \mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
 	\chi(\mathbf{r'})\mathbf{E(r')}
-\end{align}
-
-Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
-defines the impulse response between two separate dipoles; and $\chi$ is the 
-susceptibility of the medium, which describes the degree of polarisation of
-the medium in the presence of an electric field. With the internal field 
-calculated, the scattered field can be computed, one of the primary advantages
-of DDA over the T-matrix method is that the composition of the target can 
-be changed freely. When comparing different computational scattering methods,
-the ECBM was found to be better suited for simulating the scattering of symmetric
-targets as the ECBM can use the target's symmetry to speed up calculations \cite{Wriedt1998}. But when dealing with inhomogeneous media DDA is 
-more efficient compared to ECBM.
+	\label{eq:DDA}
+\end{align}
+
+Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, 
+which defines the impulse response between two separate dipoles; and 
+$\chi$ is the susceptibility of the medium, which describes the 
+degree of polarisation of the medium in the presence of an electric 
+field. One of the primary advantages of DDA over the T-matrix method 
+is that the composition of the target can be changed freely. When 
+comparing different computational scattering methods, the ECBM was 
+found to be better suited for simulating the scattering of symmetric
+targets as the ECBM can use the target's symmetry to speed up 
+calculations \cite{Wriedt1998}. But when dealing with inhomogeneous 
+media DDA is more efficient compared to ECBM.
+
+The DDA method can be used to derive the target's $T$-matrix. After
+generating the far field expansion of the electric field (Eq.~
+\eqref{eq:DDA}), the vector spherical expansion far field 
+(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
+expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
+the T-matrix of the target.
 
@@
 \section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a molecules
-diffusive behaviour, for the case of optical tweezers the most complete model 
-of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
-derivations are sufficient for macroscopic behaviours the Langevin equation 
-better describes the microscopic characteristics of any diffusive behaviour.
-
-The Fickian model describes the net flux of solute molecules into a finite
-volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$
-\cite{Gillespie2012}. It assumes that the solute molecules do not collide with 
-one another or with molecules in the solution; overall, the Fickian model is
-used to describe how solute molecules disperse over long periods of time. It 
-does not provide any incite into the forces acting on individual molecules, 
-whereas the Langevin model captures the physical interactions between an 
-individual molecule and the surrounding fluid over a wide range of time scales.
-
-The Einstein model of diffusion expands upon the Fickian model by considering
-the collisions between individual molecules. If we consider a single particle 
-suspended in a solution it will experience multiple collisions with other 
-molecules \cite{Gillespie2012a} over a given time interval $\delta t$. The 
-Einstein molecule allows us to consider how individual molecules behave while
-suspended in a surrounding fluid. Where it begins to falter is when we consider
-bringing $\delta t$ to $0$; in this scenario the molecule experiences not many
-but only individual collisions. The Einstein model does not consider the inertia 
-of the molecule in question and so for very short time frames the model implies 
-that the molecule's velocity changes instantly after each collision \cite{Gillespie2012a,
-Gillespie2012b}. Furthermore, the kinetic energy of each collision is not limited
-by the thermal energy of the system, meaning that using the Einstein model to 
-predict a particle's trajectory to an finite degree of accuracy implies that 
-the particle has infinite kinetic energy \cite{Gillespie2012b}. This failure to 
-describe motion over smaller time frames is addressed in the Langevin model by 
-accounting for the fluid drag of the system, where any sudden change in velocity 
-must result in a proportionally opposed drag force \cite{Gillespie2012c}.
-
-The Langevin model of diffusion assumes that the net force on a particular 
-particle is described fully by these individual collisions \cite{Gillespie2012c}. 
-Unlike the Fickian model it provides a full description of the interactions 
-between the target molecule and the surrounding fluid; while also being able
-to describe its motion over any time scale, unlike the Einstein model:
+Describing any microscopic motion requires an understanding of a 
+molecules diffusive behaviour, for the case of optical tweezers 
+the most complete model of diffusion is the Langevin equation. 
+Models such as the Fickian, and Einstein derivations are sufficient 
+for macroscopic behaviours the Langevin equation better describes 
+the microscopic characteristics of any diffusive behaviour.
+
+The Fickian model describes the net flux of solute molecules into a 
+finite volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
+solute molecules do not collide with one another or with molecules in 
+the solution; overall, the Fickian model is used to describe how solute 
+molecules disperse over long periods of time. It does not provide any 
+incite into the forces acting on individual molecules, whereas the 
+Langevin model captures the physical interactions between an individual 
+molecule and the surrounding fluid over a wide range of time scales.
+
+The Einstein model of diffusion expands upon the Fickian model by 
+considering the collisions between individual molecules. If we 
+consider a single particle suspended in a solution it will experience 
+multiple collisions with other molecules \cite{Gillespie2012a} over a 
+given time interval $\delta t$. The Einstein molecule allows us to 
+consider how individual molecules behave while suspended in a 
+surrounding fluid. Where it begins to falter is when we consider 
+bringing $\delta t$ to $0$; in this scenario the molecule experiences 
+not many but only individual collisions. The Einstein model does not 
+consider the inertia of the molecule in question and so for very short 
+time frames the model implies that the molecule's velocity changes 
+instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
+Furthermore, the kinetic energy of each collision is not limited by 
+the thermal energy of the system, meaning that using the Einstein model 
+to predict a particle's trajectory to an finite degree of accuracy 
+implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
+This failure to describe motion over smaller time frames is addressed 
+in the Langevin model by accounting for the fluid drag of the system, 
+where any sudden change in velocity must result in a proportionally 
+opposed drag force \cite{Gillespie2012c}.
+
+The Langevin model of diffusion assumes that the net force on a 
+particular particle is described fully by these individual 
+collisions \cite{Gillespie2012c}. Unlike the Fickian model it 
+provides a full description of the interactions between the target 
+molecule and the surrounding fluid; while also being able to 
+describe its motion over any time scale, unlike the Einstein model:
 \begin{align}
@@
 force transduction \cite{BergSoerensen2004, Smith2003}.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.67\linewidth]{finite_differences.png}
-	\caption{Example trajectory created using Finite Differences method
+\begin{SCfigure}[][h!]
+		\includegraphics[height=0.68\linewidth]{finite_differences.png}
+	\caption{\rule{0ex}{20ex}Example trajectory created using Finite Differences method
 			 for a $2\mu m$ diameter sphere.
@@
 The particle's motion can be localised around the shaded ellipsoid.}
-\end{figure}
-
+\end{SCfigure}
+
+\newpage
 \section{Calibration Techniques}
@@
 allowing for fast calibration times while also quickly filtering out 
-typical noise sources. Taking the Fourier transform of a particle's 
-trajectory yields:
+typical noise sources. After recording a trajectory over some time 
+$t_{msr}$, taking the Fourier transform of a particle's trajectory 
+yields:
 \begin{align}
@@
 
-where $\hat{\eta}$ is the Fourier transform of the white noise, 
-where the values are exponentially distributed as opposed to being 
-normally distributed in the time domain \cite{BergSoerensen2004}, 
-the correlation function is given as:
-\begin{align}
-	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
-\end{align}
-
+where $x$ can either be the physical position (relative to the trap
+focus) or a recorded signal from a position-detector,and $\hat{\eta}$ 
+is the Fourier transform of the white noise (see Eq.~\eqref{eq:full_langevin}). Where the values are exponentially distributed as opposed to being 
+normally distributed in the time domain \cite{BergSoerensen2004}.
 We can therefore ignore the white noise from our analysis by looking 
-at the spectral density of $\hat{x}(f)$: 
+at the spectral density of $\hat{x}(f)$ that produces a Lorentzian
+curve: 
 \begin{align}
@@
 
-Where $t_msr$ is the period over which we collected data. The Lorentzian 
-can be fitted via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ 
-which allows us to compute both the diffusion coefficient (in arbitrary 
-units) and the corner frequency $f_c$ which is directly related to 
-the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. The Lorentzian shape
-implies that the trap is harmonic - but not symmetric - which assumes that 
-the particle itself is an isotropic scatterer. Anisotropic scatterers will
-not produce a Lorentzian curve due to the coupled rotation and translation, 
-though depending on the nature of that rotation (i.e. whether it is stochastic
-\cite{Bang2020} or periodic \cite{Yogesha2012}) will effect how drastically it 
-departs from a typical Lorentzian curve.
+Eq.~\eqref{eq:lorentzian} can be fitted via a simplified geometric 
+series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the 
+diffusion coefficient (in arbitrary units) and the corner frequency 
+$f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. The Lorentzian shape implies that the trap 
+is harmonic - but not symmetric - which assumes that the particle 
+itself is an isotropic scatterer. Anisotropic scatterers can 
+produce an Lorentzian curve if the angular component of the power
+spectra is insignificant. The magnitude of rotational motion 
+(whether it is stochastic \cite{Bang2020} or periodic 
+\cite{Yogesha2012}) will effect how drastically it departs from a 
+typical Lorentzian curve.
 
@@
 that when converted to length units represents the maximum displacement 
-the particle can move beyond the focus, however when $f\gg f_c$ the 
-PSD falls off exponentially which denotes the particle is freely diffusing 
-within the beam focus.
+the particle can move beyond the focus. When $f\gg f_c$ the PSD falls 
+off exponentially which denotes the particle is freely diffusing within 
+the beam focus.
 \begin{figure}[h!]
@@
 
-where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients
-of the incident and scattered fields respectively. We can get the transverse 
-force and torque components in a similar form by applying a simple rotation 
-transformation. With the optical forces and torques computed all that remains
-is to compute the Brownian forces and torques which are constrained by the relation.
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
+coefficients of the incident and scattered fields respectively. 
+We can get the $x$ and $y$ force and torque components in a 
+similar form by applying a simple rotation transformation. With 
+the optical forces and torques computed all that remains is to 
+compute the Brownian forces and torques which are constrained 
+by the relation.
 \begin{align}
@@
 
-Vigilante et al compiled together a python package that combines both 
-\textit{ott} and \textit{mstm} to simulate the behaviour of spherical 
-aggregates within a predefined optical trap \cite{Vigilante2020}. Now 
-\textit{ott} does have inbuilt $T$-matrix modules that can be used to 
-model the scattering from different shaped particles; however, for 
-aggregates of spheres the resultant force is nonsensical (see Chapter 
-4 for further details). As such we rely on the \textit{mstm} for 
-simulating dimers.
+Vigilante et al compiled together a python package that combines 
+both \textit{ott} and \textit{mstm} to simulate the behaviour of 
+spherical aggregates within a predefined optical trap 
+\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
+modules that can be used to model the scattering from different 
+shaped particles; however, for aggregates of spheres the resultant 
+force is nonsensical (see Chapter 4 for further details). As such 
+we rely on the \textit{mstm} for simulating dimers.
 
@@
 
-Beyond merely simulating the forces experienced by a spherical aggregate we
-also developed a simulative quadrant photo diode to replicate the results
-from a typical calibration test. This builds upon the work from \cite{Rohrbach2002} which applied Lorenz-Mie theory to replicate the response signal of a QPD 
-being used in back focal-plane interferometry. However, rather than be 
-constrained to individual spheres we can now consider the expected response 
-from any type of spherical aggregate.
-
-In order to simulate a typical experimental set up with a QPD installed 
-as a position detection system we need to evaluate the total magnitude of 
-the electric field incident on the photo-diode surface. While trapping a 
-micro-particle, the scattered and incident fields combine together and 
-interfere with one another. These fields are collected by a condenser 
-lens in the far field limit and are focused onto the QPD surface, the 
-total intensity can be evaluated as:
+Beyond merely simulating the forces experienced by a spherical 
+aggregate we also developed a simulative quadrant photo diode 
+to replicate the results from a typical calibration test. This 
+builds upon the work from \cite{Rohrbach2002} which applied 
+Lorenz-Mie theory to replicate the response signal of a QPD 
+being used in back focal-plane interferometry. Rather than be 
+constrained to individual spheres we can now consider the 
+expected response from any type of spherical aggregate.
+
+In order to simulate a typical experimental set up with a QPD 
+installed as a position detection system we need to evaluate 
+the total magnitude of the electric field incident on the 
+photo-diode surface. While trapping a micro-particle, the 
+scattered and incident fields combine together and interfere 
+with one another. These fields are collected by a condenser 
+lens in the far field limit and are focused onto the QPD 
+surface, the total intensity can be evaluated as:
 \begin{align}
@@
 
-The last term is simply a representative step term that defines the outer 
-limit by which we evaluate the electric field, this is analogous to our 
-condenser lens removing noise from other light sources by only accepting 
-light at a specific acceptance angle defined by its numerical aperture 
-$NA_c$. Depending on the relative size of our particle we can adjust 
-the acceptance angle, this has very little effect on the transverse 
-signals, but for axial evaluations of a trapped particle the numerical 
-aperture should be tuned so that the resultant response curve has negative 
-slope in order to allow for axial position detection, the method for 
-finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
-
-The incident beam is simple enough to define given our set up parameters, 
-for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
+The last term is simply a representative step term that defines 
+the outer limit by which we evaluate the electric field, this is 
+analogous to our condenser lens removing noise from other light 
+sources by only accepting light at a specific acceptance angle 
+defined by its numerical aperture $NA_c$. Depending on the 
+relative size of our particle we can adjust the acceptance angle, 
+this has very little effect on the transverse signals, but for 
+axial evaluations of a trapped particle the numerical aperture 
+should be tuned so that the resultant response curve has negative 
+slope in order to allow for axial position detection, the method 
+for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
+
+The incident beam is simple enough to define given our set up 
+parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
 beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
@@
 	\centering
-	\includegraphics[width=\linewidth]{fixed_polarisation.png}
-	\captionsetup{margin=0.5cm}
+	\includegraphics[width=0.7\linewidth]{fixed_polarisation.png}
+	%\captionsetup{margin=0.5cm}
 	\caption{Total Field incident on the quadrant photo diode for a several
-		different displacements. Top Left: $2\mu m $ diameter silica sphere
+		different displacements. Top Left: $2\mu m$ diameter silica sphere
 		placed at the focus of a  beam. Top Right: silica sphere is now
 		displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere 
-		is now displaced by $1 \mu m$ along the y-axis. Bottom Left: silica 
-		sphere is now displaced by $0.707 \mu m$ along the x and y axis. 
-		The trapping beam is a $TEM_{00}$ Gaussian, polarised along the 
-		along the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of 
-		each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
+		is now displaced along the y-axis. Bottom Left: silica sphere is 
+		now displaced by $0.707 \mu m$ along the x and y axis. The 
+		trapping beam is a $TEM_{00}$ Gaussian, polarised along the along 
+		the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
 	\label{fig:totalfield}
@@
 
+\newpage
 While translational motion have no knock on effects to the QPD signal, 
@@
 The discrepancy between our simulated QPD and the results from 
-Rohrbach can be attributed to the fact that the QPD signal  With this any trajectory can be collected from the QPD by displacing and 
-rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} 
-and \ref{chapter:simulated_detection} we discuss how to extract information 
-from the QPD's signal.
+Rohrbach can be attributed to the fact that the position to 
+signal error grows as you move further from trap focus. In most
+optical trapping experiments we usually are calibrating a strong
+trap where the mean square displacement falls well within the
+linear regime shown in fig.~\ref{fig:Rohrbach}. In which case the
+error between Rohrbach's model and our own is inconsequential for 
+calibration purposes. With this any trajectory can be collected 
+from the QPD by displacing and rotating the beam accordingly, 
+later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of accuracy
+that can be achieved using back focal plane interferometry to 
+characterise non-spherical targets.

### 2024-10-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates 04-10-2024

@@
 path as such:
-
 \begin{align}
@@
 the partial derivative of position gives:
-
 \begin{align}
@@
 based on its velocity.
-
 \begin{align}
@@
 	\centering
-	\includegraphics[width=\linewidth]{galvano_shear_rate.png}
+	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
 	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
@@
 by the trapped particle. 
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\centering
-	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
+	\includegraphics[width=0.75\linewidth, height=1.1\linewidth]{frames_no_beam_movement.pdf}
 	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
@@
 
+\newpage
 \subsection{Moving Beam}
@@
 \label{sec:clusters}
-One common aspect involving these droplets is the fact when brought to the laser 
-focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
-to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
-again the beam was focused close to the droplet edge, this time the galvano mirror
-was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
-After a few minutes of irradiation droplets were seen entering the camera frame, 
-because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
-several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
-\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
-crystal structure becomes clearer.
+One common aspect involving these droplets is the fact when brought 
+to the laser focus the droplets would nucleate immediately 
+\cite{Liao2022}. A solution similar to Sec~\ref{sec:stationary} was 
+made up, but without any silica droplets. Once again the beam was 
+focused close to the droplet edge, this time the galvano mirror was 
+scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
+After a few minutes of irradiation droplets were seen entering the 
+camera frame, because no silica had been added these droplets had 
+to be from the glycine solution. Trapping individual droplets did 
+not result in immediate nucleation even after several minutes being 
+trapped. Trying to bring two droplets together resulted in nucleation 
+between the two droplets, compared to \ref{sec:stationary} \& 
+\ref{sec:moving} the growth is much slower, taking nearly 40 seconds 
+before the crystal structure becomes clearer.
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.7\linewidth]{cluster_trapping.pdf}
+	\includegraphics[width=0.8\linewidth]{cluster_trapping.pdf}
 	\caption{Frames from a longer video demonstrating the trapping of a glycine 
@@
 
-The fact that these droplets can be trapped indicates they must have a higher 
-refractive index than the surrounding solution. It has been shown that the 
-concentration of glycine solution is correlated with the refractive indices 
-of the liquid \cite{Gowayed2021, Orttung1963}. This suggests that these droplets 
-serve to provide solute material to the bulk crystal.
+The fact that these droplets can be trapped indicates they must 
+have a higher refractive index than the surrounding solution. 
+Previous reports have shown that a focused trap will draw in 
+solute material from the bulk \cite{Tsuboi2009, Gowayed2021}. It 
+has been shown that the concentration of glycine solution is correlated 
+with the refractive indices of the liquid \cite{Gowayed2021, Orttung1963}. 
+This suggests that these droplets serve to provide solute material 
+to the bulk crystal.
+
+\newpage
 
@@
 collected and suspended in a water + glycine solution ($S = 1.001$)
-and the laser was switched on.
+and the laser was scanned along a narrow linear path.
+
+As shown by fig.~\ref{fig:seed_crystals} the laser does not in fact 
+promote crystal growth but instead forces the crystal to dissolve 
+into the bulk solution. The dissolution rate is not as substantial 
+as the growth rate seen in previous sections. It takes over 20 
+minutes for the for the crystal surface to dissolve more than a 
+few microns. Furthermore we do not report any sightings of droplets
+close to the surface of the crystal nor do we see anything enter the 
+optical trap.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{seed_crystals.pdf}
+	\label{fig:seed_crystals}
+	\caption{Frames from a seed crystal being irradiated by a scanning
+		laser. The laser path is shown below in red. Initially the 
+		crystal edge is nearly perpendicular to the laser path. However,
+		over the course of nearly twenty minutes the crystal around the 
+		laser has dissolved leaving a concave indent.}
+\end{figure}
+
+This has some implications: Firstly, the lack of supposed glycine 
+droplets suggests that the presence of an interface is crucial in 
+order to promote the formation of these droplets. This is consistent
+with previous reports on laser nucleation, where a air-solution or 
+solution-glass interface is necessary for the trap focus to have any 
+noticeable effect on the solute \cite{Gowayed2021, Liao2022, 
+Rungsimanon2010}. Secondly, the fact that the seed crystal dissolved 
+in a saturated solution suggests that this could be due to a localised
+heating effect. A commonly used heating model for optical tweezers is 
+the Peterman model \cite{Peterman2003} which accounts for the lateral 
+distance between the optical trap and a local heat sink (such as the 
+glass cover slip or microscope objective). Even with the heat sink 
+the predicted temperature rise in pure water would be around $5-10 K$.  
+The fact that the seed crystal dissolves so slowly could simply be due 
+to the fact that the local heating is lessened by the moving beam front.
 
@@
 \section{Conclusion}
+
+In conclusion, the experimental results from this chapter provide no 
+affirmative evidence of shear induced crystal nucleation. This is in
+part due to the limited rotational speed that could be achieved using
+vaterite microspheres. Simply adding a micro-rotor to a supersaturated
+solution will not yield a significant fluid flow to induce crystal 
+nucleation. Furthermore, the region in which the local shear rate is 
+significant enough is so small that any nucleation events that could 
+occur there would not grow large enough to be stable. This does not 
+however, suggest that shear induced nucleation is impossible at a 
+micro scale. In fact there have been several developments lately that
+allow for the precise rotation of multiple rotors to control the 
+precise motion of silica beads. The challenge lies in both building the
+optical equipment to form multiple strong traps (~ 3 per rotor). These 
+traps would have to be strong enough to not only hold the rotors but 
+also move them at a sufficiently fast rotational speed that they could
+induce a significant enough fluid flow.
+
+On the other hand, the results utilising the galvano mirror have some 
+interesting implications for future research into crystal nucleation
+using optical tweezers. We noted that in undersaturated conditions the
+presence of 'droplets' that could be trapped and later nucleated when 
+brought into contact with one another. The fact that these droplets 
+could be trapped suggests they have a higher refractive index to the 
+solution, given that the solutions were filtered prior to irradiation
+dismisses that they could be dust particles. Prior literature suggests
+that optical trapping of a supersaturated solution can lead to the 
+production of nano droplets \cite{Gowayed2021, Tsuboi2009}. Controlling
+the localised growth requires that the growth and trapping are handled
+by two separate lasers. Optical traps for amorphous solids has been 
+shown as a possibility but the algorithm would need to be adaptive 
+to the changing crystal shape.

### 2024-10-04 Latex_files/Thesis/chapters/chapter 4/main.tex
updates 04-10-2024

@@
 \label{chapter:langevin_dynamics}
-Much of the calibration theory discussed in Chapter 2 assumes that the
-target particle in question is a single sphere, one whose scattering and 
-motion is easily computed. However, while working with dense colloidal 
-suspensions, one often ends up trapping more than one sphere. Li and 
-Arlt \cite{Li2008} studied the case of two microspheres trapped in a single 
-OT and found that multiple trapped beads could be mistaken for a single 
-trapped bead with altered trap stiffness (see \ref{sec:harmonic_traps}).
-Theoretical studies on the case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics based model to show
-that the two trapped beads are brought into physical contact with each 
-other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Experiments 
-in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well.
+Much of the calibration theory discussed in Chapter 2 assumes that 
+the target particle in question is a single sphere, one whose 
+scattering and motion is easily computed. However, while working 
+with dense colloidal suspensions, one often ends up trapping more 
+than one sphere. Li and Arlt \cite{Li2008} studied the case of two 
+microspheres trapped in a single OT and found that multiple trapped 
+beads could be mistaken for a single trapped bead with altered trap 
+stiffness (see \ref{sec:harmonic_traps}). Theoretical studies on the 
+case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} 
+employed a ray-optics based model to show that the two trapped beads 
+are brought into physical contact with each other by optical forces 
+and they also calculated the axial equilibrium positions of the two 
+trapped beads as a function of their size. Experiments in 
+\cite{Praveen2016} confirmed that the two trapped beads indeed 
+experience different trap stiffnesses in the vicinity of the same 
+potential well.
 
-There are further discussions looking into the dynamics 
-of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to
-the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.
+There are further discussions looking into the dynamics of a whole 
+host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour 
+an arbitrary shaped particle comes with great difficulty due to the 
+fact that the optical force is dependent on a greater number of 
+variables such as orientation and size factors.
 
-With the initial goal of the PhD being to induce nucleation events via a 
-spherical micro-rotor, the aim of this chapter is to - in a limited capacity
-simulate and investigate the influence of a second particle being bound to our 
-target sphere. The choice of a dimer, instead of an amorphous solid that 
-might better represent a growing crystalline solid, allows us to consider 
-how the dynamics of the aggregate change by varying the size factor. 
-Additionally, attempting to simulate an amorphous aggregate is rather difficult 
-as calculating the optical force and torque is computationally slow. We build 
-upon the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
+With the initial goal of the PhD being to induce nucleation events 
+via a spherical micro-rotor, the aim of this chapter is to - in a 
+limited capacity simulate and investigate the influence of a second 
+particle being bound to our target sphere. The choice of a dimer, 
+instead of an amorphous solid that might better represent a growing 
+crystalline solid, allows us to consider how the dynamics of the 
+aggregate change by varying the size factor. Additionally, attempting 
+to simulate an amorphous aggregate is rather difficult as calculating 
+the optical force and torque is computationally slow. We build upon 
+the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
 asymmetric spherical dimers and how varying size parameters alters the 
-dynamics and additionally makes characterising their interactions within an optical trap more cumbersome.
+dynamics and additionally makes characterising their interactions within 
+an optical trap more cumbersome.
 
@@
 the dimer and its orientation relative to the trap. This is important 
-to address, for while the the dimers' diffusive behaviour is the same, regardless of position and orientation, how it interacts with the trap 
+to address, for while the the dimers' diffusive behaviour is the same, 
+regardless of position and orientation, how it interacts with the trap 
 may produce the wildly different results. Furthermore, we look at how 
-our dimers interact with circularly polarised light, this is especially pertinent as the original plan was to utilise circularly polarised 
-light to generate fluid flow in a supersaturated solution.
-The rotational motion of dimers in circularly polarised light goes against much of the established literature and suggests that the transfer of spin angular
-momentum is enabled by combining particles together. And lastly, we 
-demonstrate how a quadrant photo diode, performs in characterising 
-the actual interactions between a dimer and the optical trap. We do so 
-to demonstrate that much of the dimer's trajectory information is 
-either lost or poorly described by simple calibration techniques, 
-making their motion difficult to characterise.
+our dimers interact with circularly polarised light, this is especially 
+pertinent as the original plan was to utilise circularly polarised 
+light to generate fluid flow in a supersaturated solution. The 
+rotational motion of dimers in circularly polarised light goes against 
+much of the established literature and suggests that the transfer of 
+spin angular momentum is enabled by combining particles together. And 
+lastly, we demonstrate how a quadrant photo diode, performs in 
+characterising the actual interactions between a dimer and the optical 
+trap. We do so to demonstrate that much of the dimer's trajectory 
+information is either lost or poorly described by simple calibration 
+techniques, making their motion difficult to characterise.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Characterisation of asymmetric dimers via PSD analysis}
-As discussed in \ref{sec:simulated_QPD}, one of the methods developed to 
-work in conjunction with \cite{Vigilante2020} is a simulated quadrant photo 
-diode for as a position detection system. While it is possible to extract 
-all of the relevant dynamical information from a simulation, confirming 
-the same behaviour in an experimental setting can be challenging if dealing
-with a non-birefringent particle.
+As discussed in \ref{sec:simulated_QPD}, one of the methods 
+developed to work in conjunction with \cite{Vigilante2020} 
+is a simulated quadrant photo diode for as a position 
+detection system. While it is possible to extract all of 
+the relevant dynamical information from a simulation, 
+confirming the same behaviour in an experimental setting can 
+be challenging if dealing with a non-birefringent anisotropic 
+scatterer. The QPD is composed of 4 photodiodes that measure
+the intensity of light incident on their surfaces. Using 
+\textit{ott} we can define a region in the far-field that is 
+analogues to the surface of the QPD surface. Evaluating the 
+electric field across this surface gives us an approximation
+of the QPD signal outputted in an experimental situation. This
+does not provide a one-to-one result however as hardware errors
+(i.e. internal resistance, external light sources, and vibrations)
+would distort the signal somewhat. We can use this as a means 
+to examine the limitations of using back-focal plane interferometry
+as a means for characterising anisotropic scatterers.
 
-As a benchmark we start by considering a single sphere within an optical trap. 
-A single polystyrene sphere suspended in water ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$)
-is trapped by a focus Gaussian beam and its trajectory was recorded. A 3 second 
-is trajectory is a typical measurement time for collecting a power spectrum, 
-the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
+As a benchmark we start by considering a single sphere within an 
+optical trap. A single polystyrene sphere suspended in water 
+($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focus 
+Gaussian beam and its trajectory was recorded. A 3 second is 
+trajectory is a typical measurement time for collecting a power 
+spectrum, the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
 \begin{figure}[h]
@@
 expected trap strength to what is reported by a quadrant photo diode:
-
 \begin{center}

### 2024-10-04 Latex_files/Thesis/chapters/chapter 5/main.tex
updates 04-10-2024

@@
 	are given in the legend). While the choice of coordinates is 
-	arbitrary they all demonstrate the same exponential decay.}
+	arbitrary they all demonstrate the same exponential decay. Dashed
+	lines indicate the difference after $20\mu m$ from the trap focus,
+	in most practical cases getting an optical fibre any closer would
+	disrupt the trap stability.}
 \end{figure}
@@
 is difficult because we have discretised the orientation space meaning 
-there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method allows us to compare the divergence 
-between two probability, namely Eq.~\eqref{eq:bayes} and a case scenario.
+there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method provides a measure of the relative
+entropy change between two probability distributions:
 \begin{align}
-	D_{KL}(P\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)) = \sum P(t) \frac{P(t)}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)} 
+	D_{KL}(P(t)\parallel Q(t)) = \sum P(t) \log_2\left[\frac{P(t)}{Q(t)}\right] 
+	\label{eq:D_KL}
 \end{align}
 
-Where $P(t)$ is the case scenario we are comparing against. In an ideal
-case our model would correctly predict the dimer's orientation with 
-100\% accuracy:
+Where $P(t)$ and $Q(t)$ are two probability distributions we are
+comparing, often it is common to use a base 2 log as this converts 
+the entropy to base 2. By relative entropy we are talking about 
+the useful information gained/lost by switching from $P(t)$ to 
+$Q(t)$. In our case $Q(t)$ is simply the distribution given by
+Eq.~\eqref{eq:bayes} and $P(t)$ is an idealised distribution, 
+one where the model is accurate with 100\% certainty.
 \begin{align}
@@
 
-In reality the distribution from Eq.~\eqref{eq:bayes} will assign some 
-non-zero probability to every reference orientation. Therefore the 
-divergence of Eq.~\eqref{eq:bayes} at each time step is given as:
+In reality the distribution from Eq.~\eqref{eq:bayes} will assign 
+some non-zero probability to every reference orientation. Therefore 
+the relative entropy for a single measurement is given as:
 \begin{align}
 	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
-	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
+	= \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
 	\right]
-	\label{eq;kullback}
+	\label{eq:kullback}
 \end{align}
 
-where a larger value of $D_{KL}$ indicates that our model is less confident 
-in its prediction of the dimer's orientation. The divergence $D_{KL}$ thus 
-illustrates the 'spread' in the estimated dimer orientation probability 
--- a distribution strongly peaked at some value would give us more 
-confidence in that value than a near-uniform distribution where the 
-scattering measurement could imply a wide range of possible orientations 
--- but it does not directly indicate our estimates actual accuracy. 
-In order to reflect the accuracy of each prediction by using a weighting
-function such as:
-\begin{align}
-	\omega_\alpha = \frac{1}{2} (\hat{\bf n}_{best} \cdot \hat{\bf n}_\alpha+1)
-\end{align}
-
-Which is just the normalised dot product between our estimate and the
-best possible estimate, so for highly inaccurate estimations $D_{KL}$
-is far greater than if we narrowly missed the $\hat{\bf n}_{best}$. With
-this in mind the divergence for a single measurement is:
-
-\begin{align}
-	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
-	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)\omega_\alpha}
-	\right]
-\end{align}
-
-Later we will show how we set a baseline for evaluating the performance 
-of \eqref{eq:bayes} across an entire trajectory and how it handles common
-errors in a static light scattering setup.
-
+The relative entropy is a measure if how spread out our distribution
+is. With larger values of $D_{KL}$ indicating that our prediction is 
+spread out so the correct orientation is unsure about $\hat{\bf n}_
+{best}$. Whereas a low value of $D_{KL}$ means that our model is more 
+confident in $\hat{\bf n}_{best}$. The only draw back being that the 
+relative entropy does nothing to indicate accuracy, only how closely 
+our distribution matches our ideal result. This will be better 
+exemplified later in \ref{sec:test}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 calculating the optical forces imparted by the laser, and the 
-Brownian force due to the surrounding fluid. MSTM provides the 
-necessary T-matrix to compute the optical force via \textit{ott}. 
-The Brownian force is found by computing the dimer's diffusion 
-tensor according to the analytical solutions provided by Nir and 
-Acrivos \cite{Nir1973}. We simulated a polystyrene dimer ($n = 
-1.59$) in a suspension of water ($n_{med} = 1.33$) over the 
-course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} 
-\ s$. We placed the dimer 4 microns below the trap focus at an 
-angle $30^{\circ}$ from the horizontal, the resulting trajectory 
-is shown below in Sec~\ref{sec:motion}. We chose these initial 
-parameters because it demonstrates our model's performance in non 
-steady state conditions.
-
-The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200.
-For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.
-
+Brownian force due to the surrounding fluid. We simulated a 
+polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} 
+= 1.33$). As discussed in chapter 4 dimers can be stably trapped
+in off-axis orientations. We simulate a dimer (of size ratio 2) that
+is trapped in this off-axis orientation by a $5 mW$ laser, due to 
+the meta-stability of the trap the dimer is only trapped for second
+before it escapes the trap. The dimer's position and orientation 
+is shown below:
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\subcaption{}
-		\includegraphics[width=\textwidth]{fig4a.png}
+		\includegraphics[width =\textwidth]{fig7a.png}
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\subcaption{}
-		\includegraphics[width=\textwidth]{fig4b.png}
+		\includegraphics[width=\textwidth]{fig7b.png}
 	\end{subfigure}
-	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
-	\label{fig:detectors}
+	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
 \end{figure}
 
-When all of the detectors lie in the same plane the expected signal can appear identical despite the dimer being in completely different orientations.
-This is shown in Figure~\ref{fig:detectors} which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels is because the dimer's scattering is indistinguishable in these two reference orientation.
-It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors.
-More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique.
-This is illustrative of the difficulty behind the inverse light scattering problem; as one cannot always map a given signal to a particular parameter value.
+At the same we use \textit{mstm} to replicate the intensity of light
+incident on the surfaces of detectors placed around the focus of the
+optical trap. The exact number of detectors was initially assumed to 
+be arbitrary, in that it made no difference to our estimate whether 
+we used 2 angles or 200. For practical purposes it seemed beneficial 
+that we demonstrate our method works for a minimal number of detection 
+angles, as geometric constraints come into play when trying to install 
+a high number of detection fibres for any optical tweezer set up.
 
-To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique.
-As seen between Figures \ref{fig:detectors}a \& b each reference orientation now has a unique scattering signal, though with only three detectors the difference in expected signals can appear insignificant.
-By adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.
-In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.
+When all of the detectors lie in the same plane the expected signal 
+can appear identical despite the dimer being in completely different orientations. It should be noted that these pairs are reflected in 
+one or more axis which suggests that these are due to the arrangement 
+of our detectors. More specifically, if the detectors are placed say 
+in the x-y plane then only when the dimer is pointed nearly fully 
+upright will the expected signal be entirely unique. This is not an 
+issue with typical analytical static light scattering experiments 
+because the signal being collected is averaged over a population of 
+particles all in different orientations. Whereas in our set up we are concerned with the scattering of a single particle whose scattering is orientationally dependant.
+
+To remedy this we raise the third detector out of the x-y plane; 
+as such the expected signals from each reference orientation is 
+unique, though with limited resolution. By adding a $4^{th}$ 
+detector we can differentiate signals more reliably, improving the 
+neural networks performance. In line with our goal of making this 
+method viable in a laboratory setting we decided not to increase 
+the number of detectors further than 4.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Testing the Model}
+\subsection{Testing the Neural Network}
 \label{sec:test}
-Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.45$) trapped in water ($n = 1.33$) within a 5 mW optical trap.
-The trapping laser is 1064nm NIR focused through a 1.25 NA objective.
-The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively.
-We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.
-
-We applied Eq.~\eqref{eq:bayes}, taking the reference orientation with the highest probability  as our estimate of the dimer's instantaneous orientation $\hat{\bf n}_{est}$.
-To visualise the model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time.
-For comparison, we also plotted the radian distance between the  dimer's instantaneous orientation and the closest reference orientation, denoted $\hat{\bf n}_{best}$.
-The dotted line indicates the maximum radian distance ($0.896$ radians) between two \textit{neighbouring} reference orientations: if we are under this line then we know our estimate is at least neighbouring the best result.
-Assuming a uniform prior of the reference orientations $p(\hat{\bf n}_\alpha)$  the neural network's predictions ($\hat{\bf n}_{est}$ from Eq.~\eqref{eq:bayes}) are at times reasonable, but there are significant large and random jumps away from the correct result (Fig.~\ref{fig:uniform}).
-
+We applied Eq.~\eqref{eq:bayes}, taking the reference orientation 
+with the highest probability  as our estimate of the dimer's 
+instantaneous orientation $\hat{\bf n}_{est}$. To visualise the 
+model's performance we plotted the orientation components of our 
+estimation $\hat{\bf n}_{\alpha}$ and the dimer's \emph{actual} 
+instantaneous orientation $\hat{\bf s}$ versus time. For comparison, 
+we also plotted the orientation components of the closest reference orientation, $\hat{\bf n}_{best}$. Assuming a uniform prior of the 
+reference orientations $p(\hat{\bf n}_\alpha)$ the neural network's predictions ($\hat{\bf n}_{\alpha}$ from Eq.~\eqref{eq:bayes}) are 
+at times reasonable, but there are significant large and random jumps 
+away from the correct result (Fig.~\ref{fig:uniform}). 
 \begin{figure}[h!]
@@
 	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity.
-Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
+Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the blue line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
 Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
@@
 orientation is simply because the neural network is assuming
-that the signals are unrelated from one another. As demonstrated
-by fig~\ref{fig:mixing} the orientation and signal spaces are 
-uncorrelated with one another which results in the neural network
-to assign orientations incorrectly. Combining this fact with 
-use of a uniform prior ($p(\hat{\bf n}_\alpha)$), there 
-is no constraint on how much estimated orientation can change 
-from time-step to time-step.
+that the signals are unrelated from one another. We can 
+observe this if we take the average relative entropy change
+across the entire trajectory:
+
+\begin{align}
+	\langle D_{KL}\rangle = \frac{1}{N} \sum_{i=1}^{N} \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha\parallel y_k(\hat{s}),t_i)}\right]
+	\label{eq:score}
+\end{align}
+
+Where we get an average entropy change of $\approx 0.604$, in 
+base 2 this means that on average the model has narrowed down 
+its options from 1 in 30 to 1 in 1.52. This is not reflected
+though in fig.~\ref{fig:uniform} which indicates that while 
+the model is highly confident in its results it is not at all 
+accurate. As demonstrated by fig~\ref{fig:mixing} the 
+orientation and signal spaces are uncorrelated with one another 
+which results in the neural network to assign orientations 
+incorrectly. Combining this fact with use of a uniform prior 
+($p(\hat{\bf n}_\alpha)$), there is no constraint on how much 
+estimated orientation can change from time-step to time-step.
 
@@
 As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.
-
+However at the same
+time result is still subpar.
 \begin{figure}[h]
@@
 
-The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}.
-By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
-To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
-\begin{align}
-	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#}
-	\\
-	K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right]
-	\\
-	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
-\end{align}
+Once again using Eq.~\eqref{eq:score} we now get a result of 
+$\approx 0.204$ or a 1 in 1.2 result. This is only a slight 
+improvement in terms of confidence but has drastically reduced 
+the variation in our models predictions. There are other 
+improvements that can be made, which additionally also help with
+potential sources of error in our set up.
 
-The worst case scenario is akin to randomly choosing a reference
-orientation at each time step. The greater the value of $F(K_l)$, the
-better our model's confidence is in characterising the dimer's
-motion. Because our model is dependent on several parameters we need
-to a sophisticated method for understanding how these parameters
-correlate with $F(K_l)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Asymmetric dimer dynamics}
-\label{sec:motion}
-A simulation of a asymmetric dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) 
-trapped in an off-axis trapping orientation was used as a test case for our 
-model, since it was previously shown that dimers can be optically rotated
-
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width =\textwidth]{fig7a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig7b.png}
-	\end{subfigure}
-	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
-	\label{fig:motion}
-\end{figure}
-
-In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.
-However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state.
-We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer.
-In the future we hope to further investigate the motion of asymmetric dimers.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Accounting for sources of error in light scattering measurements}
-When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor.
-Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle.
-However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance.
-We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors.
-We highlight how we address these areas below.
+When it comes to analysing light scattering from any size particle,
+error analysis becomes a significant factor. Typically this can be 
+accounted for by averaging over long periods of time to get an 
+assessment of the steady state conditions of the target particle. 
+However in our case where we wish to know the instantaneous 
+orientation, we instead have to rely on our understanding of how 
+uncertainty can effect our model's performance. We identified two 
+areas which are likely sources of error in our estimation: firstly, 
+an incorrect modelling of the target particle, and secondly, signal 
+noise arising from experimental factors. We highlight how we address 
+these areas below. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-One of the main limitations of our model is that we assume that the dimer being modelled in MSTM is accurate to the dimer being trapped in the optical tweezer.
-Sizing molecules accurately is a significant challenge for single particle analysis so there is bound to be some uncertainty with the measurements.
-We ran our model 3 times with the neural net being trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
+One of the main limitations of our model is that we assume that 
+the dimer being modelled in \textit{mstm} is accurate to the 
+dimer being trapped in the optical tweezer. Sizing molecules 
+accurately is a significant challenge for single particle 
+analysis so there is bound to be some uncertainty with the 
+measurements. We ran our model 3 times with the neural net being 
+trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
+	\caption{Model estimates of orientation when neural net has 
+		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.519]$, (b) 1:2.05 $[\langle D_{KL}
+		\rangle=3.706]$, (c) 1:1.95 $[\langle D_{KL}\rangle=
+		3.705]$ ($n_{refs} = 30$)}
 	\label{fig:size}
@@
 	\end{subfigure}
-	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
+	\caption{Model estimates of orientation when neural net has 
+		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.593]$, (b) 1:2.05 $[\langle D_{KL}\rangle
+		=5.659]$, (c) 1:1.95 $[\langle D_{KL}\rangle=3.279]$, 
+		($n_{refs} = 126$)}
 	\label{fig:refs}
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
+	\caption{Model prediction for signal error of (a) $1\%$ ~
+		$[\langle D_{KL}\rangle=0.962]$, (b) $15\%$ $[\langle D_{KL}\rangle=13.654]$, and (c) $25\%$ ~  $[\langle D_{KL}\rangle=13.017]$.}
 	\label{fig:epsilon}
@@
 
-As can be seen from Figure~\ref{fig:epsilon}, the inclusion of signal noise quickly leads to a decrease in the model's performance.
-This is due to an inherent feature of the inverse scattering problem: two distinct regions in orientation space can become heavily intertwined and thus no longer well separated when mapped to intensity space (even though the mapping remains continuous): so even small uncertainties in the scattering data can lead to large 'mistakes' in the choice of orientation by the neural network.
-(Indeed if this was not the case the inverse scattering problem would be quite simple.)
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of 
+signal noise quickly leads to a decrease in the model's 
+performance. So much so that beyond 15\% the model is actually 
+worse than just randomly guessing.This is due to an inherent 
+feature of the inverse scattering problem: two distinct regions 
+in orientation space can become heavily intertwined and thus no 
+longer well separated when mapped to intensity space (even 
+though the mapping remains continuous): so even small 
+uncertainties in the scattering data can lead to large 'mistakes' 
+in the choice of orientation by the neural network. (Indeed if 
+this was not the case the inverse scattering problem would be 
+quite simple.)
 
-To reduce the effects of the signal noise we took the time average of the expected signal over $0.001 s$ and then had our neural network estimate the orientation based on the average signal.
+To reduce the effects of the signal noise we took the time 
+average of the expected signal over $0.001 s$ and then had our 
+neural network estimate the orientation based on the average 
+signal.
 
@@
 	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $1\,{\rm ms}$}
+	\caption{Model prediction for signal error of (a) $1\%$ 
+		$[\langle D_{KL}\rangle=1.447]$, (b) $15\%$ $[\langle D_{KL}\rangle=4.670]$, and (c) $25\%$ $[\langle D_{KL}
+		\rangle=7.911]$, time averaged over $1\,{\rm ms}$}
 	\label{fig:time average}
 \end{figure}
-This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model.
-There appears to be no clear correlation between the length over which we time average and the performance of our model.
-Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network.
-Fortunately, time averaging even over $1\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
+This resulted in a reduction in the overall signal noise and 
+provided a higher degree of accuracy for our model. There 
+appears to be no clear correlation between the length over 
+which we time average and the performance of our model. Time 
+averaging over every 0.05s resulted in a drastically worse 
+performance; this is due to the fact that over longer time 
+periods there is greater uncertainty regarding how the dimer's 
+orientation has changed, thus tracking the instantaneous 
+orientation becomes harder for the neural network. Fortunately, 
+time averaging even over $1\,{\rm ms}$ seems to provide a 
+satisfactory estimation of the dimer's angular dynamics within 
+the optical trap.
 
-From the above discussion it’s clear that estimation of the dimer’s orientation is a problem that can be endlessly tuned to fully maximise our end result.
-Here we simplify the problem somewhat by employing a relatively small finite number of 'reference orientations' to map between scattering and dimer orientation: the precision of estimation could be improved by utilising a greater number of reference orientations, although there remains a balance between the realisable precision of orientation estimate and the noise level of the scattering measurement.
-Another avenue to further explore would be using the method to optimise the choice of detection angles, essentially to find the region in the mapping between measured scattering and orientation that others the best degree of confidence through optimal separation of scattering signals for distinct orientations.
-For sequences of data such as dynamic measurements, a further potential enhancement would be to consider more complex correlations based on prior expectations of the dynamics.
-Here already we improve the method using a non-uniform prior based on only the immediately previous measurement in time (see Section 2.1): considering a non-uniform grouping of reference orientations might result in a better estimation, if we have information regarding the dimer’s preferred axis of rotation.
+From the above discussion it’s clear that estimation of the 
+dimer’s orientation is a problem that can be endlessly tuned 
+to fully maximise our end result. Here we simplify the problem 
+somewhat by employing a relatively small finite number of 
+'reference orientations' to map between scattering and dimer 
+orientation: the precision of estimation could be improved by 
+utilising a greater number of reference orientations, although 
+there remains a balance between the realisable precision of 
+orientation estimate and the noise level of the scattering 
+measurement. Another avenue to further explore would be using 
+the method to optimise the choice of detection angles, 
+essentially to find the region in the mapping between measured 
+scattering and orientation that others the best degree of 
+confidence through optimal separation of scattering signals 
+for distinct orientations. For sequences of data such as 
+dynamic measurements, a further potential enhancement would 
+be to consider more complex correlations based on prior 
+expectations of the dynamics. Here already we improve the 
+method using a non-uniform prior based on only the immediately 
+previous measurement in time (see Section 2.1): considering 
+a non-uniform grouping of reference orientations might result 
+in a better estimation, if we have information regarding the 
+dimer’s preferred axis of rotation.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{QPD Simulations}
@@
 
-We have developed a method for measuring the dynamics of an optically-trapped colloidal objects based purely on measurements of the object's light scattering at a small number of detection angles.
-We demonstrate the method using the orientation of an asymmetric dimer as the dynamic variable and object of interest respectively, but in principle the model can be applied to any characteristic that impacts the light scattering pattern produced by a trapped entity such as size and shape.
-The MSTM package is a flexible tool for calculating the light scattering of complex objects using a representation of the object as a set of micro-particles, enabling training of a neural network to enable categorisation of the mapping between scattering and trapped object characteristics.
-By taking account of the physically realistic behaviour of the trapped object and the characteristics of the trap (which impact the dynamics of the object), the Bayesian inference method can be re ned to provide a reliable estimation of object characteristics of interest, even in the presence of measurement noise.
-Fundamentally, the inverse scattering problem is difficult to solve, since the mapping between object characteristics and scattering can be highly complex.
-We determined the minimum number of detectors required for a reliable estimation in the presence of measurement noise; furthermore, we demonstrated that the arrangement of these detectors is critical for a reliable estimation of an objects orientation.
-However, Bayesian inference based on neural network estimation of the mapping provides a powerful method for practical applications, extending the use of optical trapping beyond measuring microscopic force response toward detailed structural and dynamic information about complex trapped entities.
+We have developed a method for measuring the dynamics of an 
+optically-trapped colloidal objects based purely on 
+measurements of the object's light scattering at a small 
+number of detection angles. We demonstrate the method using 
+the orientation of an asymmetric dimer as the dynamic variable 
+and object of interest respectively, but in principle the 
+model can be applied to any characteristic that impacts the 
+light scattering pattern produced by a trapped entity such 
+as size and shape. The MSTM package is a flexible tool for 
+calculating the light scattering of complex objects using a 
+representation of the object as a set of micro-particles, 
+enabling training of a neural network to enable categorisation 
+of the mapping between scattering and trapped object 
+characteristics. By taking account of the physically realistic 
+behaviour of the trapped object and the characteristics of 
+the trap (which impact the dynamics of the object), the 
+Bayesian inference method can be re ned to provide a reliable 
+estimation of object characteristics of interest, even in the 
+presence of measurement noise. Fundamentally, the inverse 
+scattering problem is difficult to solve, since the mapping 
+between object characteristics and scattering can be highly 
+complex. We determined the minimum number of detectors 
+required for a reliable estimation in the presence of 
+measurement noise; furthermore, we demonstrated that the 
+arrangement of these detectors is critical for a reliable 
+estimation of an objects orientation. However, Bayesian 
+inference based on neural network estimation of the mapping 
+provides a powerful method for practical applications, 
+extending the use of optical trapping beyond measuring 
+microscopic force response toward detailed structural and 
+dynamic information about complex trapped entities.

### 2024-10-04 Latex_files/Thesis/main_win.tex
updates
updates to ch3
updates 04-10-2024

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{cite}
+%\usepackage{cite}
 \usepackage{enumitem}
@@
 \usepackage{graphicx}
-
-\graphicspath{{C:/Users/xbb16146//dmaciver97/Latex_files/Thesis/figs}}
+\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
 \usepackage{sidecap}
@@
 \usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}
+
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+
+\usepackage[
+backend=biber,
+bibstyle=numeric,
+citestyle=numeric,
+minbibnames=3, maxbibnames=3,
+]{biblatex}
+\addbibresource{thesis_bib.bib}
 
@@
 %  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
-  \includegraphics[width=5cm]{logo} \\[6cm]
+  \includegraphics[width=5cm]{logo.pdf} \\[6cm]
   \end{flushright}
@@
 
-\title{PhD Title \\ PhD Thesis}
+\title{Application of Micro-Rotors to Induce Crystal Nucleation: 
+	   Challenges, Takeaways, and Future Prospects\\ PhD Thesis}
 \author{D. Maciver\\[-0.8ex]
@@
 \addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{ieeetran}
-\bibliography{thesis_bib}
+\printbibliography
 

### 2024-10-25 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

@@
 
-\subsection{Optical Torque}
+\subsection{Optical Torque and rotation}
+\label{sec:opt_torque}
 It has long been known that electromagnetic fields can transfer linear and
@@
 if the crystal lattice has different refractive indices for varying 
-polarisations of light. For circularly polarised light this inhomogeneity 
-results in a high degree of SAM being transferred to the target object \cite{Parkin2009, Arita2016}, this has been exploited to rotate microspheres 
-as fast as 1000 Hz while suspended in a bulk medium \cite{Arita2016} as well 
-as a means of measuring the local temperature and shear response of said 
-medium \cite{Millen2014, RodriguezSevilla2018}. Calculating the optical 
-torque applied to a birefringent material is given via:
+polarisations of light. Because of this the material has two different 
+refractive indices depending on the polarisation of light. For circularly 
+polarised light this inhomogeneity results in a high degree of SAM being 
+transferred to the target object \cite{Parkin2009, Arita2016}. The greater 
+the difference between the two refractive indces the greater the angular
+momentum transfer.
+
+The ability to transfer angular momentum has been exploited to rotate 
+microspheres as fast as 1000 Hz while suspended in a bulk medium 
+\cite{Arita2016} as well as a means of measuring the local temperature 
+and shear response of said medium \cite{Millen2014, RodriguezSevilla2018}. 
+Calculating the optical torque applied to a birefringent material is given via:
 \begin{equation}
@@
 
-Where $\theta$ is the angle between the particle's long axis and the 
-polarisation vector of the local EM field, and $\phi$ is the phase shift 
-in the EM field. The first term represents the 'orientational' torque which 
-is due to the target particle being aligned with the EM amplitude, when 
-aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. For sphere's this is always the case as they lack a 
-distinctive major axis. The second term is due purely to the polarisation 
-of the optical trap, for circularly polarised light $\phi=\pi/4$ thus 
-maximising the torque transferred to the target particle. Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
+Where $\Delta n$ is the difference between the two refractive indices, 
+$\theta$ is the angle between the particle's long axis and the polarisation 
+vector of the local EM field, and $\phi$ is the phase shift in the EM 
+field. The first term represents the 'orientational' torque which is 
+due to the target particle being aligned with the EM amplitude, when 
+aligned $\theta=0$ meaning the entire term is negligible for particle's 
+with a stable orientation. For sphere's this is always the case as they 
+lack a distinctive major axis. The second term is due purely to the 
+polarisation of the optical trap, for circularly polarised light 
+$\phi=\pi/4$ thus maximising the torque transferred to the target particle. 
+Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
 birefringence, but there are other mechanisms that result in optical torque.
@@
 aligned with the beam's direction of propagation (in which case $\theta
-=\pi/2$ and the first term disappears). This is analogous to an optically trapped sphere, where alongside a restoring force the particle also 
+=\pi/2$ and the first term disappears). This is analogous to an optically 
+trapped sphere, where alongside a restoring force the particle also 
 experiences a restoring torque. This seemingly random rotational motion
@@
 nucleation; however, the exact relationship between shear rate and nucleation
-rate has only been recently understood for specific solutions. Theoretical research into shear induced nucleation suggests that there should be a 
+rate has only been recently understood for specific solutions. Theoretical 
+research into shear induced nucleation suggests that there should be a 
 slight increase in the nucleation rate at low shear rates, reaching a 
@@
 This has been shown theoretically for both simple colloidal \cite{Mura2016,
-Debuysschere2023,Richard2015} and ice crystal formation \cite{Goswami2020}; however, no experimental work into these systems has been conducted to prove 
+Debuysschere2023,Richard2015} and ice crystal formation \cite{Goswami2020}; 
+however, no experimental work into these systems has been conducted to prove 
 this is the case. There is some experimental evidence for this phenomena in 
-simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions.
-In \cite{Debuysschere2023} it was found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate.
-Using the theoretical model established in \cite{Mura2016,2001} which modifies the CNT to account for the effects of a nucleus undergoing shearing,
+simple salt and protein solutions - though the authors emphasise that 
+mechanical agitation cannot be ruled out - there has not been a exhaustive 
+study into the shearing effects apart from in glycine solutions. In 
+\cite{Debuysschere2023} it was found that a shear rate of around $3000\ 
+s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate. 
+Using the theoretical model established in \cite{Mura2016,2001} which modifies 
+the CNT to account for the effects of a nucleus undergoing shearing, 
 accounting for the fact that a nucleus' growth is undergoing competition 
@@
 the nucleus from growing. This however would require one to be able 
-to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus in a continuous fluid 
+to apply a continuous shear rate to a targeted nucleus with high precision, 
+there is also no model for an individual nucleus in a continuous fluid 
 field.
@@
 the best way to address this is by developing \textit{in-situ} 
-methods that can study the pre nucleation phase in greater detail \cite{Fu2021}. Furthermore, the ability to localise nucleation 
+methods that can study the pre nucleation phase in greater detail 
+\cite{Fu2021}. Furthermore, the ability to localise nucleation 
 allows for better characterisation of the kinetics of crystal 

### 2024-10-25 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 dipole we can ignore the momentum transfer calculation and simply
-compute the Lorenz force for a dipole in an electric field. This is 
-not to say that the scattered field is not influenced by the particle, 
-the scattered field can be related to the incident field like so:
-\begin{align}
-	\label{eq:Rayleigh_scattering}
-	\begin{pmatrix}
-		{\bf E}^s_\perp \\
-		{\bf E}^s_\parallel 
-	\end{pmatrix}
-	= \begin{pmatrix}
-		S_{11} & 0 \\
-		0 & S_{22}
-	\end{pmatrix}
-	\begin{pmatrix}
-		{\bf E}^i_\perp \\
-		{\bf E}^i_\parallel
-	\end{pmatrix}
-\end{align}
-
-\noindent
-Where $S_{11}$, and $S_{22}$ are elements of the amplitude scattering 
-matrix, this is analogous to a single plane wave scattering off a 
-sphere. The difference between Eq.~\eqref{eq:Rayleigh_scattering} and
-Eq.~\eqref{eq:t_matrix} is that the $T$-matrix accounts for the 
-scattering from multiple plane waves that make up the laser. Because the 
-particle can be approximated by a point dipole the scattering field is 
-linearly related to the scattered field. While not important for optical 
-force calculation higher complexity scattering problems can be simplified 
-by subdividing the particle into discrete dipoles (see 
-Sec~\ref{sec:scattering}), in which case the scattered field is now 
-significant. As the target particle gets larger this assumption fails 
-to accurately describe the trapping force \cite{Li2021}. For particle 
-sizes close to the laser wavelength the trapping force is best described 
-by Lorenz-Mie theory.
+compute the Lorenz force for a dipole in an electric field. While 
+not important for optical force calculation higher complexity 
+scattering problems can be simplified by subdividing the particle 
+into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
+the scattered field is now significant. As the target particle gets 
+larger this assumption fails to accurately describe the trapping force \cite{Li2021}. For larger particles the optical force and scattering 
+is better described via the Lorenz-Mie theory.
 
@@
 \label{sec:scattering}
-There are several methods available to calculate the scattered field 
-produced by a particle. While one can compute this by directly applying
-Lorenz-Mie theory, describing the interactions between focused beams and 
-complicated shapes can be computationally slow and memory intensive. As a
-result, several alternative scattering approaches have been developed that
-allow for faster computation times simplifying the scattering problem.
+There are several methods available to calculate the scattered 
+field produced by a particle. While one can compute this by 
+directly applying Lorenz-Mie theory, describing the interactions 
+between focused beams and complicated shapes can be computationally 
+slow and memory intensive. Several alternative scattering approaches 
+have been developed that allow for faster computation times 
+simplifying the scattering problem. While all of these methods will
+produce identical results the key difference is in how one represents
+the scatterer.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 over the targets surface. These currents are chosen so that the 
-electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the fields one can derive the 
-individual elements of \textbf{T} until a desired accuracy is 
-achieved.
+electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the harmonic functions one can 
+derive the individual elements of \textbf{T} until a desired 
+accuracy is achieved. One can either use the ECBM to compute the 
+scattering for a fixed particle orientation and position. Or they
+can evaluate the scattering for any particle orientation by 
+computing its $T$-matrix.
 
 For this project we utilised an extension of the ECBM known as 
-multi-sphere $T$-matrix method (\textit{mstm}). Developed by 
-Mackowski \cite{Mackowski2011} the computational code computes 
-the scattered field from each sphere within the target cluster 
-and the incident field itself. The scattering between spheres 
-will converge to a final result but modern codes truncate 
-the calculations to an upper number of iterations (denoted by 
-$N_{max}$) to fit a desired accuracy and computational time. 
-By centring the expansion at the origin of each sphere the 
-individual sphere $T$-matrices can be computed, then using the
-addition theorem, the target centred $T$-matrix is computed.
+multi-sphere $T$-matrix method (\textit{mstm}) developed by 
+Mackowski \cite{Mackowski2011}. Where Lorenz-Mie Theory centres
+the spherical harmonics expansion about the sphere's origin it
+often struggles to adjust the expansion to any arbitrary origin.
+\textit{mstm} utilises an addition theorem to translate the 
+individual spherical expansions from the centres of each 
+individual sphere to a preset target origin. \textit{mstm}'s 
+primary function is to compute the averaged scattering from 
+large spherical aggregates, but for our purposes it can be
+also utilised to compute $T$ matrices for arbitrary shaped
+spherical aggregates.
 
@@
 composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
-being constructed of discrete dipoles. Each dipole interacts with 
-both the incident field and the scattered fields from every other 
-dipole surrounding it. The resulting scattered field is identical 
-to the scattered field produced by direct integration of Eq.~\eqref{eq:internal} throughout the full particle volume 
-\cite{Goedecke1988}. The integral form for the total electric field 
-inside a scatterer is given as \cite{Wriedt1998}:
+being constructed of discrete dipoles. This is essentially an 
+extension of the Rayleigh Regime, though in order to compute the 
+optical force one requires a full description of the electric 
+field. Each dipole interacts with both the incident field and the 
+scattered fields from every other dipole surrounding it. The 
+resulting scattered field is identical to the scattered field 
+produced by direct integration of Eq.~\eqref{eq:internal} 
+throughout the full particle volume \cite{Goedecke1988}. The 
+integral form for the total electric field inside a scatterer is 
+given as \cite{Wriedt1998}:
 \begin{align}
@@
 
-By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
+By reversing the time step (i.e. use $-\Delta t$) to approximate the
 velocity and acceleration based on the previous time steps, we can 
@@
 \label{sec:sim_parameters}
-Throughout this PhD we use the work of Vigilante to perform a systematic 
-study of the dynamics demonstrated by asymmetric dimers in both plane and 
-circularly polarised light. The simulations have two primary frames of 
-reference that require explanation. Firstly, there is the laboratory frame 
-which describes the coordinate system, we define the origin of our simulation
-as the focus of our trapping beam. While the choice in direction for the $x$ 
-and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
-aligned with the direction of beam propagation. 
+Throughout this project we use the work of Vigilante to perform a 
+systematic study of the dynamics demonstrated by asymmetric dimers 
+in both plane and circularly polarised light. The simulations have 
+two primary frames of reference that require explanation. Firstly, 
+there is the laboratory frame which describes the coordinate system, 
+we define the origin of our simulation as the focus of our trapping 
+beam. While the choice in direction for the $x$ and $y$ axis are 
+arbitrary the $z$ axis is specifically chosen so that it is aligned 
+with the direction of beam propagation. 
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.6\linewidth]{lab_frame.png}
-	\caption{Ray scattering render of both the laboratory frame and particle 
-		frame for a dimer and a Gaussian beam. The origin of the laboratory 
-		frame is defined as the focus of the trapping beam, with the positive 
-		$z$-axis being aligned with the direction of beam propagation. The 
-		particle frame is centered on the dimer's centre of mass with the 
-		primary axis $U_z$ being aligned with the centres of the two spheres.
-		Black line is scaled to represent $1\ \mu m$.}
+	\caption{Ray scattering render of both the laboratory frame and 
+		particle frame for a dimer and a Gaussian beam. The origin 
+		of the laboratory frame is defined as the focus of the 
+		trapping beam, with the positive $z$-axis being aligned with 
+		the direction of beam propagation. The particle frame is 
+		centred on the dimer's centre of mass with the primary axis 
+		$U_z$ being aligned with the centres of the two spheres. Black 
+		line is scaled to represent $1\ \mu m$.}
 	\label{fig:lab_frame} 
@@
 
-Secondly there is the particle frame which describes the orientation state of 
-the particle, orientations are reported as quaternions which can be readily 
-transformed into rotation matrices. The matrix describes the direction of the 
-particles 3 orientation axis $U_x$, $U_y$, and $U_z$. The latter is oriented 
-so that it is collinear with the centres of the two spheres; the direction of
-vector $U_z$ is such that it goes from the centre of $a_{II}$ to the centre of
-$a_{I}$. The origin of the 3 axis' is set on the particle's centre of mass, 
-this is also used to define it's displacement from the origin of the laboratory 
-frame. The diffusion tensor of the dimer is more complex than a sphere's as we 
-must consider the rotational effects experienced by the dimer. We use the 
-analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided the 
-solution coefficients depending on the dimer's size ratio ($a_I/a_{II}$). Rather 
-than recalculate the coefficients for every possible configuration of dimer we 
-use a spline fitting function to approximate the diffusion tensor. 
+Secondly there is the particle frame which describes the orientation 
+state of the particle, orientations are reported as quaternions which 
+can be readily transformed into rotation matrices. The matrix describes 
+the direction of the particles 3 orientation axis $U_x$, $U_y$, and 
+$U_z$. The latter is oriented so that it is collinear with the centres
+of the two spheres; the direction of vector $U_z$ is such that it goes 
+from the centre of $a_{II}$ to the centre of $a_{I}$. The origin of the 
+3 axis' is set on the particle's centre of mass, this is also used to 
+define it's displacement from the origin of the laboratory frame. The 
+diffusion tensor of the dimer is more complex than a sphere's as we must consider the rotational effects experienced by the dimer. We use the 
+analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided 
+the solution coefficients depending on the dimer's size ratio 
+($a_I/a_{II}$). Rather than recalculate the coefficients for every possible configuration of dimer we use a spline fitting function to approximate 
+the diffusion tensor. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-10-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

@@
 \chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, the original outline of the project is to 
-investigate the possibility of using optical tweezing to induce nucleation 
-by generating fluid flow within a supersaturated solution. The intent 
+As outlined in Chapter 1, part of the projects aim is to investigate 
+the possibility of using optical tweezing to induce nucleation by 
+generating fluid flow within a supersaturated solution. The intent 
 of which would be twofold: Firstly to have a repeatable means of 
@@
 to understand the influence of shearing on nucleation at a micro 
-level as compared to results in bulk fluid. It has been shown that 
-for macro-scale systems, the likelihood of nucleation increases to 
-a maximum value under increased shearing \cite{Debuysschere2023, 
-Mura2016}. Mura and Zaccone developed a theoretical framework to 
-describe how the a newly formed nucleus experiences two additional 
-growth factors when placed in a moving fluid. Firstly, due to increased 
-molecular transport of solute molecules the nucleation rate is enhanced 
-in low to moderate fluid flows. But in addition, due to shear flow the 
-crystal surface undergoes deformation which suppresses the nucleation 
-rate undergoing faster fluid flow \cite{Mura2016}. The theory was 
-confirmed for glycine solution by Debuysschere \textit{et al} who
-identified that the nucleation rate of supersaturated glycine was 
-enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ \cite{Debuysschere2023}.
-After which the nucleation rate began to decrease but was 
-still greater compared to the case where fluid flow was minimal.
+level as compared to results in bulk fluid.
 
-Optical tweezers can been used to rotate a whole host of micro-rotors, 
-with the fastest reported results exceeding $1000\ Hz$ in heavy water
-\cite{Arita2016}. Therefore, it stands to reason that a micro-rotor
-rotating in a supersaturated solution could generate sufficient fluid
-flow such that the nucleation rate around said particle is enhanced 
-compared to the bulk fluid. We focused on two primary candidates for
-micro-rotors, Vaterite and 4-Heptyl-4-biphenylcarbonitrile (7CB). The
-former being a polymorph of calcium carbonate and the latter an example
-of nematic liquid crystals, both of which have been used repeatedly in
-previous micro-rotor. In addition, we also consider the application of
-using techniques beam steering to generate fluid flow by trapping silica
-micro-beads. In this instance the fluid flow is generated not due to the
-transfer of angular momentum, but due to shearing caused by a moving 
-sphere through stagnant fluid.
+It has been shown that for macro-scale systems, the likelihood of 
+nucleation increases to a maximum value under increased shearing 
+\cite{Debuysschere2023, Mura2016}. Mura and Zaccone developed a 
+theoretical framework to describe how the a newly formed nucleus 
+experiences two additional growth factors when placed in a moving 
+fluid. Firstly, due to increased molecular transport of solute 
+molecules the nucleation rate is enhanced in low to moderate fluid 
+flows. But in addition, due to shear flow the crystal surface 
+undergoes deformation which suppresses the nucleation rate 
+undergoing faster fluid flow \cite{Mura2016}. Experimental results 
+with glycine solution support this theory; Debuysschere \textit{et 
+al} demonstrated that the nucleation rate of supersaturated glycine 
+was enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ 
+\cite{Debuysschere2023}. After which the nucleation rate began to 
+decrease but was still greater compared to the case where fluid 
+flow was minimal.
+
+Optical tweezers can been used to rotate a whole host of microscopic 
+objects, with the fastest reported results exceeding $1000\ Hz$ in 
+heavy water \cite{Arita2016}. If such a micro-rotor was suspended in
+a supersaturated solution the fluid flow around it could be fast enough
+that the nucleation rate is locally enhanced. We focused on two 
+primary candidates for rotation, Vaterite and 
+4-Heptyl-4-biphenylcarbonitrile (7CB). The former being a polymorph of 
+calcium carbonate and the latter an example of nematic liquid crystals, 
+both of which have been used repeatedly in previous micro-rotor research
+\cite{Saito2022, Arita2016, Parkin2009}. In addition, we also consider 
+the application of using techniques beam steering to generate fluid 
+flow by trapping silica micro-beads. In this instance the fluid flow is 
+generated not due to the transfer of angular momentum, but due to 
+shearing caused by a moving sphere through stagnant fluid.
 
@@
 In general, all optical tweezers require a laser driver, a focusing 
-microscope objectives, a position controller, and position detector.
-The laser used for this project was a $1064\ nm$ near infrared laser 
-- provided by CNI Lasers – that was focused by a Nikon 100x oil 
-immersion lens. The choice of an oil immersion lens is important as 
-the optical oil used prevents a loss of focus when used on a glass 
-cover slip. Now, experimental work has shown that the trapping 
-efficiency increases with beam diameter up until it exceeds 
-$\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where $D_{obj}$ is 
-the diameter of the objective aperture. To expand the beam front we 
-utilise a Galilean beam expansion arrangement (indicated by $f_1$, 
-and $f_2$ in Fig.~\ref{fig:setup}) as recommended for high power laser 
-applications. In our initial experiments the beam expansion provides a 
-$4\times$ magnification. Whereas in later experiments we utilised a 
-galvano-mirror the beam expansion is $3\times$ and then the 4f 
-correlator provides a further $1.25\times$ magnification (using $f_3$ 
-and $f_4$) - the magnification is given by.
+microscope objectives, a position controller, and position detector 
+\cite{Gieseler2020}. The laser used for this project was a $1064\ 
+nm$ near infrared laser - provided by CNI Lasers – that was focused 
+by a Nikon 100x oil immersion lens. The choice of an oil immersion 
+lens is important as the optical oil used prevents a loss of focus 
+when used on a glass cover slip. Now, experimental work has shown 
+that the trapping efficiency increases with beam diameter up until 
+it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
+$D_{obj}$ is the diameter of the objective aperture. To expand the 
+beam front we utilise a Galilean beam expansion arrangement 
+(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
+for high power laser applications. In our initial experiments the 
+beam expansion provides a $4\times$ magnification. Whereas in later 
+experiments we utilised a galvano-mirror the beam expansion is 
+$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
+provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
+- the magnification is given by.
 \begin{align}
@@
 	\caption{Optical tweezer set up used for the majority of the PhD. 
-The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively.
-Diagram not drawn to scale.}
+		The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, 
+		$60\ mm$, $150\ mm$, \& $200\ mm$ respectively [M = mirrors,
+		DM = dichroic mirrors, f = focal lenses]. Diagram not 
+		drawn to scale.}
 	\label{fig:setup}
@@
 of position detection: video-analysis, lateral-effect position sensing, 
-and photodiodes. The former being ideally suited for multiple traps 
-or situations where precision is not the top priority. In order to 
-match the force measurements of back-focal plane interferometry requires 
-the camera's frame rate to exceed $1\ kHz$ which can be difficult 
-to achieve while maintaining a decent resolution \cite{Gibson2008}. 
-In comparison off the shelf back-focal plane detectors can achieve 
-temporal resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
+and photodiodes.
+
+Video analysis is ideally suited for multiple traps or situations where 
+precision is not the top priority. Whereas lateral-effect and photodiode 
+position detection are two examples of back-focal plane interferometry, 
+where the interference pattern produced by the target particle is 
+extrapolated to determine its position. In order for video analysis to 
+match the accuracy of back-focal plane interferometry methods requires 
+the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
+achieve while maintaining a decent resolution \cite{Gibson2008}. In 
+comparison off the shelf back-focal plane detectors can achieve temporal 
+resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
 
@@
 \subsection{Fourier Optics and 4f correlators}
+\label{sec:4f_correlation}
+As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
+of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
+purpose. These are in fact crucial for the operation of the Galvano, the
+pair of them can be more accurately called a 4f correlator.
+
 A 4f correlator is an example of Fourier optics in practice, understanding
@@
 differentiate between different images), the placement of each lens is 
-shown below.
+shown via fig.~\ref{fig:signal_generator}.
 \begin{figure}[h!]
@@
 		independently.}
+	\label{fig:signal_generator}
 \end{figure}
@@
 dual channel signal control. This allowed us to precisely control the alignment,
-amplitude, phase, and frequency of both mirrors making alignment much easier. For
-basic trapping calibration the galvano-mirrors were set to a simple DC output, 
-providing a fixed spot which operates like a typical optical trap.
+amplitude, phase, and frequency of both mirrors making alignment much easier.
+
+While in theory the frequency can be increased until the mechanical limit of the 
+mirrors is reached ($\approx 1000\ Hz$), the practical upper limit is determined 
+by the trap strength. For a silica sphere this is on the order of $300\ Hz$ when
+suspended in water. Likewise while in theory the signal amplitude can be increased
+until the voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
+of the focal lenses limits the maximum amplitude to $0.5 V$ as any greater will
+move the laser beyond the lens. For basic trapping calibration the galvano-mirrors 
+were set to DC output, providing a fixed spot which operates like a typical optical 
+trap.
 
@@
 \label{sec:vaterite}
-There are several options for particles that can be rotated using optical 
-tweezers \cite{Parkin2009, Saito2022}. Over the course of the project two 
-different micro spheres where investigated, Vaterite and liquid crystal 
-droplets. Both can be readily synthesised in the lab and are will rotate 
-at a variety of sizes.
+There are several options for particles that can be rotated 
+using optical tweezers \cite{Parkin2009, Saito2022}. Over 
+the course of the project two different micro spheres where 
+investigated, Vaterite and liquid crystal droplets. Both can 
+be readily synthesised in the lab and are will rotate at a 
+variety of sizes (see \ref{sec:opt_torque}).
 
-Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
-due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
-other polymorphs of calcite and aragonite, when synthesised Vaterite will 
-typically form small spherical particles making them ideal for optical 
-trapping and rotation. Synthesis of Vaterite micro spheres requires fine 
-control of the crystal growth process in order to maintain polymorphic stability. 
-Though for the purposes of optical rotation the exact polymorph is not as 
-important as its morphology as all 3 polymorphs are inherently birefringent.
+Vaterite is a polymorph of calcium carbonate that is rarely 
+seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
+All three polymorphs are inherently birefringent meaning 
+that they can be rotated using circularly polarised light. 
+However unlike its other polymorphs of calcite and aragonite, 
+when synthesised vaterite will typically form small spherical 
+particles making them ideal for optical trapping and rotation. 
+Synthesis of Vaterite micro spheres requires fine control of 
+the crystal growth process in order to maintain polymorphic 
+stability. Though for the purposes of optical rotation the 
+exact polymorph is not as important as its morphology as all 
+3 polymorphs are inherently birefringent.
 
-Vaterite samples where made by the first preparing equal amounts of 
-$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
-a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
-a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
-and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
-Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
-$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
-After 10 minutes of slow but continuous mixing a few drops of Agepon was added
-to halt the reaction, the solution was filtered and washed 3 times with 
-distilled water before being suspended in water.
+Vaterite samples where made by the first preparing equal amounts 
+of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
+the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
+aside for later. Firsta small vial was filled with $1.5mL$ of 
+$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
+$NaCO_3$ respectively, forming a seed solution. Next, a larger 
+vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
+$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
+After 10 minutes of slow but continuous mixing a few drops of 
+Agepon was added to halt the reaction, the solution was filtered 
+and washed 3 times with distilled water before being suspended 
+in water.
 
-When trapped in circularly polarised light, the anisotropic crystal lattice
-allows spin angular momentum to be transferred to the Vaterite particle, 
-resulting in a rotation about the beam axis. In addition, the anisotropic 
-scattering causes the QPD signal to vary with a constant periodicity that
-is attributed to its rotation rate. Therefore, the resulting power spectrum 
-is not a Lorentzian but now also displays peaks that appear at integer 
-multiples of the particles rotational frequency. 
+When trapped in circularly polarised light, the anisotropic 
+crystal lattice allows spin angular momentum to be transferred 
+to the Vaterite particle, resulting in a rotation about the 
+beam axis. Because the particle scatters light anisotropically 
+the signal detected by the QPD shows periodic fluctuations 
+\cite{Yogesha2012}. In addition due to the change in polarisation 
+there is also a periodic variation that appears at twice the 
+rotational frequency \cite{Monteiro2018}. Therefore, the resulting 
+power spectrum is not a Lorentzian but now also displays peaks 
+that appear at integer multiples of the particles rotational 
+frequency. 
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{(a)Sample Vaterite sphere suspended in water and trapped by circular polarised trap.
-(b) Collected power spectrum from rotating Vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
+	\caption{(a)Sample Vaterite sphere suspended in water and trapped by
+		 circular polarised trap. (b) Collected power spectrum from 
+		 rotating Vaterite, peaks in the power spectrum appear at integer 
+		 multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
 	\label{fig:vaterite}
@@
 change phase according to the ratio of organic to inorganic molecules present.
-Liquid crystal rotors are rather simple in their production, 4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
+Liquid crystal rotors are rather simple in their production, 
+4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
 and a small amount was added to a vial of distilled water. The solution 
@@
 local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
-Using a birefringent particle and rotating it within the fluid, the maximum 
-rotation rate is due to the fluid drag resisting the torque of the trapping 
-beam \cite{RodriguezSevilla2018}. If you want to measure fluid flow you can 
-instead use a micro-rotor to see how fluid flow propagates in the medium \cite{Knoener2005}. Likewise, one can use a beam steering arrangement to 
-probe the drag force of the fluid, by understanding the trap strength 
-(calibrating using a low frequency signal) one can measure the drag force 
-experienced by the local fluid \cite{RobertsonAnderson2018}. I
+When it comes to optical rotation, a rotating particle will experience a
+fluid drag torque that is proportional to its rotation rate, the maximum 
+rotation rate is when the fluid drag is equal to the optical torque 
+of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
+to measure local temperature changes \cite{Millen2014}, or for understanding 
+how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
+one can use a beam steering arrangement to probe the drag force of the 
+fluid, by understanding the trap strength (calibrating using a low 
+frequency signal) one can measure the drag force experienced by the local 
+fluid \cite{RobertsonAnderson2018}. I
 
-Understanding the fluid velocity around our trapped object is determined 
-mostly by the Reynold's number of the system, for a sphere submersed in a 
+For any discussion of fluid flow its important that we know the fluid 
+regime which is given by the Reynolds number. For a sphere submersed in a 
 moving fluid of velocity $U$ this is given by:
@@
 velocity $U$, assuming a no-slip boundary condition we can model the 
-fluid velocity profile based on the velocity of the particle. There are 
-two possible avenues for generating shear flow with a trapped particle; 
-rotation of birefringent particles, and fluid flow induced by particle motion.
+fluid velocity profile based on the velocity of the particle. Assuming
+that for now the fluid properties are unaffected by the rotating particle
+we see that for an individual sphere the Reynolds number can only change 
+due to increasing/decreasing the fluid velocity. The fluid velocity is 
+directly proportional to the particles rotation rate for low Reynolds
+numbers \cite{Bruce2020}. Given the small size of the particles used 
+($1-10\mu\ m$ in diameter), and the relatively low rotation rates that 
+are possible with an optical tweezer, for the case of our analysis the
+Reynolds number will not never be large enough to consider factors such
+as turbulence. There are two possible avenues for generating shear flow 
+with a trapped particle; rotation of birefringent particles, and fluid 
+flow induced by particle motion.
 
-Rotating birefringent particles are by far the most common method for 
-generating and measuring fluid flow in a solution. To see if we can 
-even achieve the theoretical maximum shear rate, Vaterite spheres 
-were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and 
-trapped with the 1064 nm laser at set to 450 mW. The rotation frequency 
-was determined using the QPD, and the particle sizes were computed 
-by image analysis. With the particle size and rotation frequency, the 
-tangential rotation speed is calculated via:
+Rotating birefringent particles are the more common method for generating 
+and measuring fluid flow in a solution. To see if we can even achieve the 
+theoretical maximum shear rate, Vaterite spheres were synthesised (see 
+Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm 
+laser at set to 450 mW. The rotation frequency was determined using the 
+QPD, and the particle sizes were computed by image analysis. With the 
+particle size and rotation frequency, the tangential rotation speed is 
+calculated via:
 \begin{align}
@@
 as the partial derivative fluid flow (assuming shearing is generated purely
-by the flow field):
+by the flow field). This approach was used previously with liquid crystal 
+rotors \cite{Saito2022}:
 \begin{align}
 	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
+	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= 
+	\frac{\pi}{2}\frac{d^3}{r^3}\omega
 \end{align}
+
+One would expect that we set some reference distance where the fluid flow 
+would be considered zero. However, at a microscopic level we are more 
+concerned by the fluid shear for an individual nucleation event. We are
+essentially computing what the shear rate would be for a small volume 
+of fluid between two moving parallel plates. This is similar logic to Mura 
+and Zaccone's work where they considered the shear on a subcritical nucleus
+in a flow field \cite{Mura2016}. They go further in depth to discuss the 
+impact of material transport, overall coming to the conclusion that 
+increasing shear rate leads to an increase the nucleation rate up till some
+maximum. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Estimation of fluid flow around micro-rotors in bulk fluid}
+\subsection{Estimation of fluid flow around a rotating particle in bulk fluid}
 First we determined the upper rotation rate that could be achieved 
-using both Vaterite and liquid crystal spheres. Vaterite samples were 
-synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
-\ref{sec:vaterite}), and then suspended in distilled water. A sample
-of $200\ \mu L$ was pipetted and a single microsphere was captured via
-a circular polarised beam.
+using both Vaterite and liquid crystal spheres. Vaterite samples 
+were synthesised according to \cite{Parkin2009, Bishop2004} (see 
+sec.~\ref{sec:vaterite}), and then suspended in distilled water. 
+A sample of $200\ \mu L$ was pipetted and a single microsphere was 
+captured via a circular polarised beam.
 
-Due to Van der Waal's forces some of the microspheres were stuck together, 
-fortunately individual sphere's were still present. Multiple microspheres 
-were trapped and their rotation rate was determined by looking at the 
-peak frequency component of the collected power spectrum. The shear flow
-was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
-were operating in the bulk fluid and away from any boundaries.
+Due to Van der Waal's forces some of the microspheres were stuck 
+together, fortunately individual sphere's were still present. 
+Individual microspheres were trapped and their rotation rate was 
+determined by looking at the peak frequency component of the 
+collected power spectrum. Due to their anisotropic scattering the 
+QPD signal will oscillate periodically. After taking the Fourier
+transform of the QPD signal the power spectrum has distinctive
+peaks at integer multiples of the rotational frequency. The largest 
+peak is often twice the rotational frequency which is due to the 
+polarization modulation of the scattered field, the first peak 
+is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
+An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.
+
+\begin{figure}[h!]
+	\begin{subfigure}{0.49\linewidth}
+		\includegraphics[width=\linewidth]{time_domain_data.png}
+	\end{subfigure}
+	\begin{subfigure}{0.49\linewidth}
+		\includegraphics[width=\linewidth]{frequency_domain_data.png}
+	\end{subfigure}
+	\caption{Example of QPD signal being converted from the time 
+		domain to the frequency domain. (left) Raw signal collected
+		from the QPD over the 1st second of its trajectory showing 
+		a periodic spacing with a period of $1/f_{rot}$. (right) 
+		Power spectrum of the QPD signal, the y-axis is a linear 
+		scale to demonstrate the clear peaks in the power spectra. 
+		The rotational frequency is the first instance of a peak 
+		forming, and after that the peaks appear in integer 
+		multiples of $f_{rot}$}
+	\label{fig:tdat_fdat}
+\end{figure}
+
+The accuracy of the rotational frequency is dependent on the sample
+rate of the QPD and the sample duration \cite{BergSoerensen2004}. 
+With higher sampling rates resulting in a wider frequency window 
+(see Sec.\ref{sec:psd}) and longer sampling durations resulting in 
+more distinct rotational peaks. The default QPD sampling rate is 
+$2^{17}\ Hz$ which is more than adequate for detecting even the 
+slightest low frequency signals. And the default sampling duration 
+was $3\ s$ which for most cases gave us clear distinct peaks for 
+even relatively low rotational frequencies. After trapping a 
+vaterite microsphere the power spectrum was recorded and the 
+rotation frequency was acquired. This was repeated several times
+for each sphere trapped, to confirm that the rotational frequency
+was constant and that the sphere had reached its maximum rotation 
+rate.
+
+After computing the rotational frequency the radial fluid speed 
+could be estimated using Eq.~\eqref{eq:birefringent_speed}. From 
+there we can estimate the shear flow experienced by a small volume 
+of fluid within the flow field (see Eq.\eqref{eq:birefringent_shear}). 
+Here we only consider the fluid flow as it propagates outward from 
+the axis of rotation. While the cover slip will have some effect 
+fluid flow it is difficult to quantify the effects of a hard 
+boundary on a rotating object when the boundary is perpendicular to 
+the axis of rotation. 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.75\linewidth}
+	\begin{subfigure}{0.7\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
+		\caption{}
 	\end{subfigure}
-	\begin{subfigure}{0.75\linewidth}
+	\begin{subfigure}{0.7\linewidth}
 		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
+		\caption{}
 	\end{subfigure}
-	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
-		Vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
-		is indicated by the dotted line. Vaterite radii and rotation frequencies 
-		are shown, the laser power was kept constant at 450 mW. Reported rotation
-		rates, and their corresponding fluid flow and shear rates, for Vaterite 
-		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
+	\caption{(Top) Fluid flow radiating out from the surface of a
+		rotating Vaterite sphere. (Bottom) Shear rates computed 
+		using Eq.\ref{eq:birefringent_shear}, optimal shear rate is 
+		of $3000 s^{-1}$ is indicated by the dotted line. Vaterite 
+		radii and rotation frequencies are shown, the laser power 
+		was kept constant at 450 mW. Reported rotation rates, and 
+		their corresponding fluid flow and shear rates, for Vaterite 
+		are also plotted alongside lab results. Results from 
+		\cite{Parkin2009, Arita2016} are included as well.}
 	\label{fig:vaterite_shear}
@@
 
-From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
-particle size and rotation rate, this is contrary to much of the theoretical
-predictions that predict an exponential decay with particle size. This can
-be in part due to the fact that synthesising perfectly spherical spheres that
-have uniform birefringence across the whole population is difficult. Despite 
-our best efforts at controlling the growth rate the smallest particle ever 
-synthesised was around $3\ \mu m$ in diameter. The Vaterite spheres would 
-often stick together while suspended in water after a short period of time. 
-The fastest reported rotation rate found during this project was by 
-\cite{Arita2016} that achieved a rotation rate of $5\ kHz$, this is plotted 
-on Fig~\ref{fig:vaterite_shear} as the dotted line. Even at that extreme a 
-rotation rate the region in which nucleation is at its optimal likelihood is 
-only $20\ nm$ wide. If instead the micro-rotor was within the vicinity of a
-solid boundary, the shear rate would be enhanced due to the no-slip boundary
-condition.
+From Fig.\ref{fig:vaterite_shear} there is not a strong relationship 
+between particle size and rotation rate, this is contrary to much of 
+the theoretical predictions that predict an exponential decay with 
+particle size. This can be in part due to the fact that synthesising 
+perfectly spherical spheres that have uniform birefringence across the 
+whole population is difficult. Despite our best efforts at controlling 
+the growth rate the smallest particle ever synthesised was around $3\ 
+\mu m$ in diameter. The Vaterite spheres would often stick together 
+while suspended in water after a short period of time. The fastest 
+reported rotation rate within a fluid was by \cite{Arita2016} 
+that achieved a rotation rate of $5\ kHz$, this is plotted in Fig~\ref{fig:vaterite_shear}. In addition we added the results from 
+\cite{Parkin2009} as a more realistic example.
 
-\section{Micro-rotors in Supersaturated solution}
-If rotation rates in bulk solution are insufficient then a micro-rotor rotating
-close to an artificial barrier may be able to improve the shear rate of the 
-surrounding fluid. Of course placing a solid barrier in a supersaturated fluid
-may well encourage nucleation somewhere on the surface outside of our control.
-Instead we chose to use the droplet edge of the supersaturated solution, while
-not a hard barrier per say, the molecular mobility close to the droplet edge is
-reduced due to surface tension. Furthermore, it has been shown through multiple 
-results that nucleation is enhanced at the air-solution interface \cite{Liao2022,
-Yuyama2010, Sugiyama2009}. 
+The optimal shear rate predicted by \cite{Debuysschere2023} is 
+plotted on Fig.~\ref{fig:vaterite_shear}(b), the outer dotted 
+lines represent the point when the nucleation rate is less than 
+90\% of the maximum. We can visualise the fluid around the 
+rotating sphere by dividing it up into radial sections. Assuming 
+the sphere is rotating sufficiently quickly there are three 
+sections of concern, as demonstrated by fig.~\ref{fig:shear_diagram}. 
+The fluid close to the sphere experiences a shear rate that 
+is far greater than the optimal shear rate, in which case 
+nucleation is suppressed \cite{Mura2016}. Secondly, there is 
+a region of fluid that experiences a shear rate close to the 
+optimal value ($\pm 5\%$) where the nucleation rate is enhanced 
+significantly compared to the bulk fluid. Lastly, beyond the 
+optimal region the shear rate drops off exponentially in which 
+case the nucleation rate is barely different compared to the 
+bulk fluid.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.875\linewidth]{shear_rate_diagram.pdf}
+	\caption{Diagram demonstrating how fluid shear rate varies 
+	radially from a rotating sphere (rotation rate = $f_{rot}$). 
+	(a) the shear rate exceeds the optimal shear rate 
+	($\dot{\gamma}_{opt}$) and suppresses the nucleation rate. 
+	(b) shear rate is close to the optimal amount, enhancing the 
+	nucleation rate. (c) shear rate is far below the optimal 
+	amount and the nucleation rate is comparable to the bulk fluid.}
+	\label{fig:shear_diagram}
+\end{figure}
+
+\section{Rotating spheres in Supersaturated solution}
+If rotation rates in bulk solution are insufficient then a 
+rotating sphere close to an artificial barrier may be able 
+to improve the shear rate of the surrounding fluid. Of course 
+placing a solid barrier in a supersaturated fluid may well 
+encourage nucleation somewhere on the surface outside of our 
+control. Instead we chose to use the droplet edge of the 
+supersaturated solution, while not a hard barrier per say, 
+the molecular mobility close to the droplet edge is reduced 
+due to surface tension. Furthermore, it has been shown through 
+multiple results that nucleation is enhanced at the air-solution 
+interface \cite{Liao2022,Yuyama2010, Sugiyama2009}.
+
+Estimating the distance from the trap focus to the cover slip
+is difficult to measure directly, often requiring one fits a 
+complex Lorentzian curve to the recorded power spectra in order 
+to measure the fluid drag from being in close proximity to the 
+cover slip \cite{BergSoerensen2004}. We estimate that the 
+vaterite must be at least $15-20\mu\ m$ from the cover slip for 
+us to reliably trap them. The contact angle between the cover 
+slip and solution was previously measured in \cite{Flannigan2023},
+while being loosely dependent on the supersaturation the 
+measured contact angle for supersaturated solutions was between 
+$30^{\circ}$ and $35^{\circ}$. This is shown in fig.\ref{fig:3.9}.
 \begin{figure}[h!]
@@
 	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
-	\caption{Diagram of optical trapping set up for rotating birefringent 
-		particles in a supersaturated solution. Left: side view of the 
-		trapping set up showing the location of the trap focus at the 
-		edge of the droplet of a supersaturated solution. Right: top down 
-		view of the glycine droplet with a trapped birefringent particle 
-		shown close to edge of the trap. As the particle rotates the drag 
-		force from the surrounding fluid generates a flow field around 
-		itself (see Eq.~\ref{eq:birefringent_speed}).}
+	\caption{Diagram of optical trapping set up for rotating 
+		birefringent particles in a supersaturated solution. 
+		Left: side view of the trapping set up showing the 
+		location of the trap focus at the edge of the droplet 
+		of a supersaturated solution. The approximate contact 
+		angle (calculated by \cite{Flannigan2023}) is used to 
+		provide a scale of how close a trapped particle could 
+		be located. Right: top down view of the glycine droplet 
+		with a trapped birefringent particle shown close to 
+		edge of the trap. As the particle rotates the drag 
+		force from the surrounding fluid generates a flow field 
+		around itself (see Eq.~\ref{eq:birefringent_speed}).}
+	\label{fig:3.9}
 \end{figure}
 
-Supersaturated solutions of glycine and water were prepared and stored in
-an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
-into the solution and $20\ \mu L$ was pipetted onto the cover slip. A Vaterite
-sphere was located, trapped, and moved as close to the droplet edge as possible.
-After measuring the microsphere's rotational frequency the sphere was left to
-rotate for a period of ten minutes after which, if no nucleation event was 
-observed the particle was released. The overall results are catalogued in 
-table~\ref{tab:Nucleation}
+Supersaturated solutions of glycine and water were prepared 
+and stored in an incubator at $40^\circ C$ prior to use. 
+When ready to be studied $15\ \mu L$ of vaterite suspension 
+was pipetted into the solution and $20\ \mu L$ of the combined 
+solution was pipetted onto the cover slip. A Vaterite sphere 
+was located, trapped, and moved to the droplet edge where the 
+solution meets the cover slip. This removed the need for 
+hydrophilic coatings to achieve a flat fluid layer (as used by 
+\cite{Liao2022, Rungsimanon2010}).
+
+Particle sizing was achieved by converting the particle diameter 
+from pixels to physical units within an accuracy of $\pm0.05\mu 
+m$. The QPD was used to measure the frequency of fluctuations 
+in the QPD signal. After measuring the rotational frequency the 
+sphere was left to rotate for a period of ten minutes after which, 
+if no nucleation event was observed the particle was released. 
+The overall results are catalogued in table~\ref{tab:Nucleation}.
 \begin{table}[h!]
@@
 		\multirow{3}*{1.01} & 2.34 & 10.4 & $\times$ \\
+		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
 		\cline{2-4} & 5.67 & 9.63 & $\times$ \\
-		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
 		\hline
@@
 		\hline
-		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
-		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
+		\multirow{3}*{1.4} & 1.59 & 0.00 & $\times$ \\
+		\cline{2-4} & 3.47 & 0.00 & $\times$ \\
 		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
 		\hline
-		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
-		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
+		\multirow{3}*{1.45} & 3.68 & 0.00 & $\times$ \\
 		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
+		\cline{2-4} & 6.32 & 0.00 & $\times$ \\
 		\hline
-		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
+		\multirow{3}*{1.49} & 1.52 & $0.00$ & $\times$ \\
+		\cline{2-4} & 4.76 & $0.00$ & $\times$ \\
 		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
-		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
 		\hline
@@
 
-Trying to trap a particle close to the edge proved more challenging than 
-expected. Unlike in previous reports where the beam is focused at the upper
-edge of the droplet \cite{Liao2022, Yuyama2010, Sugiyama2022}, we attempted 
-trapping into the crook of the droplet. It is suspected that trapping is 
-much harder at the interface due to increased surface tension and unpredictable
-scattering forces. The closest we could trap a microsphere to the droplet edge 
-was in the range of $5-10\mu\ m$, at that distance the fluid flow is so low that 
-even the presence of a hard boundary would be insufficient for shearing the 
-fluid. Furthermore, as is evident in Table~\ref{tab:Nucleation}, the rotation 
-rate drops off significantly with increased supersaturation, due to higher 
-fluid viscosities.
-While in theory a sufficiently focused laser could rotate any microsphere to a fast enough to reach the shear rate predicted by \cite{Debuysschere2023} the localised intensity would be so large that even
-using $D_2O$ would see a significant increase in temperature.
+It should be noted that rotational frequencies equal to 
+$0.00\ Hz$ are not due to a rounding error. Simply looking
+at the live video shows no rotation of the vaterite particles,
+and even looking at the QPD signals shows no discernable 
+fluctuations that would arise due to rotation. Trying to trap 
+a particle close to the edge proved more challenging than 
+expected. Unlike in previous reports where the beam is 
+focused at the upper edge of the droplet \cite{Liao2022, 
+Yuyama2010, Sugiyama2022}, we attempted trapping in close 
+to the contact point of the droplet and the cover slip.
 
-It is not impossible that fluid shearing could be used in the future to 
-localise nucleation; but from these results, using individual micro-rotors 
-is not an appropriate method. Firstly, the area of influence is far too small 
-to see any noticeable increase in the nucleation rate. And secondly, increased 
-fluid viscosity significantly reduces the limits the maximum rotation rate 
-possible. If multiple micro-rotors could be trapped in close proximity to one 
-another they could create a large region of fluid where nucleation is more likely than the bulk fluid. Micro-rotors have been created that allow for precise
-control of suspended micro-particles \cite{Butaite2019} and could potentially be
-used to generate sufficient shearing. However these could not be used in this 
-project as we lacked the necessary hardware to form multiple gradient traps.
+It is suspected that trapping is much harder at the 
+interface due to increased surface tension and unpredictable 
+scattering forces. The closest we could trap a microsphere to 
+the droplet edge was in the range of $5-10\mu\ m$, at that 
+distance the fluid flow is so low that even the presence of a 
+hard boundary would be insufficient for shearing the fluid. 
+Furthermore, as is evident in Table~\ref{tab:Nucleation}, 
+the rotation rate drops off significantly with increased 
+supersaturation, due to higher fluid viscosities. While in 
+theory a sufficiently focused laser could rotate any 
+microsphere to a fast enough to reach the shear rate 
+predicted by \cite{Debuysschere2023} the localised intensity 
+would be so large that even using $D_2O$ would see a significant 
+increase in temperature.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Shearing via Galvano-mirror manipulation}
-An alternative approach to generating fluid shear is to use a galvano-mirror
-to rapidly move a trapped particle in a bulk fluid.
-While typically galvano and gimble mirrors are used to trap multiple particles simultaneously, a single micro sphere can be moved quickly through a fluid along a preset path.
-The only
-limitation on the particle's speed being the ratio of the trap stiffness to
-the drag force. Calculating the shear rate around an individual particle is 
-difficult to do precisely but for low Reynolds numbers we can get an adequate approximation.
+It is not impossible that fluid shearing could be used in the 
+future to localise nucleation; but from these results, using 
+individual micro-rotors is not an appropriate method due to
+two key factors. Firstly, the area of influence is far too 
+small to see any noticeable increase in the nucleation rate 
+this is demonstrated most clearly in fig.~\ref{fig:vaterite_shear}. 
+Due to the exponential decay in the expected shear rate it is 
+very difficult trying to shear a large volume of fluid. Even 
+if the optimal shear rate at a micro-level was an order of 
+magnitude less than what was predicted by \cite{Debuysschere2023}, 
+only a small volume of fluid would even experience that shear 
+rate. In order to properly understand the microscopic effects 
+of fluid shear a method of localised shearing over a larger 
+fluid volume is necessary.
 
-For a simple circular path one can estimate the sphere's speed by the 
-radius of its path and the frequency of its orbit $U = R\omega$; however 
-for a more complex path, such as an elliptical orbit the curve needs 
-to be parametrised. One can describe the position parameter of a circular 
-path as such:
-\begin{align}
-	r(u) = \left[rcos(2\pi u), rsin(2\pi u), 0 \right]
-\end{align}
+Secondly, increased fluid viscosity significantly reduces 
+the limits the maximum rotation rate possible. As indicated 
+by \ref{tab:Nucleation}, the rotation rate achievable by a
+vaterite sphere falls off significantly with increasing 
+supersaturation. Experimental estimations of relative viscosities
+showed a 12\% increase when looking at undersaturated glycine
+solutions ($S\approx0.34$) \cite{Patyar2020}. As far as we
+are aware there are no reported fluid viscosities for higher
+supersaturations. Extrapolating the results from \cite{Patyar2020}
+suggests that a saturated solution would have a 35\% increase 
+in fluid viscosity (compared to pure water). This also does 
+not account for any other factors such as the fact that close
+to a boundary a rotating particle will experience a reaction
+torque from the stationary surface \cite{Bruce2020}.
 
-If we say that $u$ describes time from some initial point we can say $u=t\omega$ 
-where omega is the frequency of orbit. Substituting this in and then taking 
-the partial derivative of position gives:
-\begin{align}
-	v(t) = \frac{dr(t)}{dt} = \left[-2\pi r\omega \ sin(2\pi t\omega),
-	\ 2\pi r\omega \ cos(2\pi t\omega), 0 \right]
-\end{align}
-
-In order to compute U we simply take the magnitude of our velocity. 
-For low velocities the fluid flow at the sphere's surface can be computed 
-based on its velocity.
-\begin{align}
-	u_r(r)=-|v(t)|^2\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
-\end{align}
-
-Where $r$ is the radial distance to that point. Again taking the partial 
-derivative we can get the shear rate for a particle moving through the fluid:
-\begin{align}
-	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
-	\label{eq:galvo_shear}
-\end{align}
-
-Moving a silica bead along an circular path can generate significant fluid 
-flow around a larger volume compared to comparable micro-rotors. Using \eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
-would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
-equates to moving the silica bead around a circular path with a frequency of 
-$100\ Hz$.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
-	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
-		in bulk water moving at different speeds. The shear rate is calculated 
-		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
-		in a circular path and so the speed is constant through out its path.}
-	\label{fig:galvano_shear}
-\end{figure}
-
-From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
-mirror a better option for generating high shear rates but also over a larger 
-volume. The only limitation to this method is the effective drag experienced 
-by the trapped particle. 
+If multiple micro-rotors could be trapped in close proximity 
+to one another they could create a large region of fluid where 
+nucleation is more likely than the bulk fluid. Because optical
+torque is not contingent on the fluid properties the only 
+limiting factor would be the total angular momentum transferred 
+to each particle. Micro-rotors have been created that allow for 
+precise control of suspended micro-particles \cite{Butaite2019} 
+and could potentially be used to generate sufficient shearing. 
+However these could not be used in this project as we lacked the 
+necessary hardware to form multiple gradient traps.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Nucleation with a Stationary and Moving Beam}
-As mentioned previously, shearing via optical rotation and particle displacement
-did not result in any localised nucleation events even while in the proximity of 
-the droplet edge. During the experiments with the galvano-mirror, it was found 
-that when no particle was present in the optical trap nucleation events would 
-occur while the beam was close to the edge of the droplet, even though the solution
-was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
-but was more interesting is how the beam's motion influenced the growth of the 
-nucleus.
+As mentioned previously, shearing via optical rotation did 
+not result in any localised nucleation events even while 
+in the proximity of the droplet edge. An alternative approach
+was suggested, using a galvano mirror to move a particle 
+quickly through a supersaturated solution. While a particle
+is within the optical trap the gradient forces are 
+significantly reduced meaning simply trapping a particle is
+insufficient for inducing nucleation \cite{Flannigan2023}. 
+This is why we did not see any nucleation events while 
+rotating the vaterite, there is not a large gradient force
+to draw in material. We did not consider a focus beam by 
+itself as this has already been demonstrated in previous
+experiments \cite{Flannigan2023, Liao2022, Rungsimanon2010}.
+
+Preliminary calculations suggested a particle moving rapidly
+through a fluid could produce significantly greater shear 
+rates (see \emph{Appendix X.X} for breakdown) than a rotating 
+particle. During initial testing using silica beads we found 
+that even in undersaturated solutions the optical trap would 
+produce a crystal nucleus when the trap was empty. This has 
+been reported prior \cite{Rungsimanon2010, Liao2022}, but 
+what is more interesting is how the beam's motion influenced 
+the growth of the nucleus.
 
@@
 \label{sec:stationary}
-Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
-event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
-$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
-nucleus forms at the trap focus, growing quickly (growth rate was approximated using
-imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
-until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
-using optical tweezers shows that the growth rate is only loosely connected to the 
-solutions supersaturation. Local conditions play a much larger role in the growth 
-rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
-is escaped is due to the fact that crystal is far too large to be held in place and 
-is in fact still growing as the solution is supersaturated. The key take away to 
-remember is that the beam has no influence over the crystal shape, instead it 
-grows outward from the trap focus. Furthermore, due to the fact that the solution
-is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
-the crystal escapes as its size exceeds the trap focus. 
+Its well known that a supersaturated solution can nucleate
+if irradiated by a focused beam. We confirmed that our 
+laser was of sufficient power and focus to induce crystal
+nucleation. The laser was focused at the edge of a $20\mu L$
+droplet consisting of water and dissolved glycine ($S=1.03$). 
+The solution was monitored for 10 minutes, if a nucleation 
+event was observed the event was recorded using a high speed 
+CCD. If no nucleation event was observed after 10 minutes 
+the solution was disposed of and a new sample was prepared. 
+After ten repeats only 30\% of the samples observed a 
+nucleation event at the focus. All of which displayed similar 
+behaviour to one another.
+
+Consider Fig.~\ref{fig:stationary_beam}, the frames taken from 
+a nucleation event, the beam is a stationary being $\approx3.5 
+\mu m$ from the droplet edge. After a period of roughly 5 minutes 
+a nucleus forms at the trap focus, growing quickly (growth rate 
+was approximated using imageJ to be on the order of $700\ \mu 
+m^2/min$) from the focal point of the trap until after roughly 
+$6$ seconds the crystal escapes. Comparing to previous literature
+using optical tweezers shows that the growth rate is only loosely 
+connected to the solutions supersaturation. Crystal growth rates 
+for a solution of glycine and water ($S\approx1.00$) was found to
+vary from as much as $3900 \mu\ m^2/min$ to as low as $675\mu 
+m^2/min$. Local conditions play a much larger role in the growth 
+rate than just the concentration \cite{Flannigan2023}. A likely 
+reason that the trap is escaped is due to the fact that crystal 
+is far too large to be held in place and is in fact still growing 
+as the solution is supersaturated. The key take away to remember 
+is that the beam has no influence over the crystal shape, instead 
+it grows outward from the trap focus. Furthermore, due to the fact 
+that the solution is supersaturated the crystal growth cannot be 
+contained to the trap focus. Instead the crystal escapes as its 
+size exceeds the trap focus. 
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
-	\caption{Frames from a longer video depicting the growth of a nucleus using 
-		a moving beam. Initially the crystal shape is amorphous (a) but eventually
-		reaches a more regular shape (b). This crystal is still influenced by the
-		optical trap as even when not directly irradiated by the laser the crystal
-		rotates between (c) and (d). When the laser is focused on a corner the 
-		crystal growth is localised to that region, resulting in an elongated 
-		section forming between frames (e) and (f).}
+	\caption{Frames from a longer video depicting the growth 
+		of a nucleus using a moving beam. Beam power is kept 
+		at 700 mW and the supersaturation was estimated at 
+		S=0.86. Initially the crystal shape is amorphous (a) 
+		but eventually reaches a more regular shape (b). This crystal is still influenced by the optical trap as 
+		even when not directly irradiated by the laser the 
+		crystal rotates between (c) and (d). When the laser 
+		is focused on a corner the crystal growth is localised 
+		to that region, resulting in an elongated section 
+		forming between frames (e) and (f).}
 	\label{fig:eliptical_beam_1}
@@
 growth occurs when the trap is close to or partially over the interface of 
-the crystal (see figure~\ref{fig:local_nucleation}(a)). As shown in \ref{sec:clusters}, the optical trap can manipulate these droplets similar 
+the crystal (see figure~\ref{fig:local_nucleation}(a)). As shown in \ref{sec:clusters}, 
+the optical trap can manipulate these droplets similar 
 to microspheres. When in close proximity to the trap these droplets are 

### 2024-10-25 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

@@
 optical trap. A single polystyrene sphere suspended in water 
-($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focus 
-Gaussian beam and its trajectory was recorded. A 3 second is 
-trajectory is a typical measurement time for collecting a power 
-spectrum, the spectra was fitted to eq.~\ref{eq:alaised_lorentzian}.
+($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
+Gaussian beam ($NA=1.25$) using circularly polarised light. For 
+the sake of time efficiency the trajectory was sampled every 10
+time steps, meaning the upper bound on the power spectra is
+$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
+we fitted the power spectra using the aliased Lorentzian 
+(Eq.~\eqref{eq:aliased_lorentzian}).
 \begin{figure}[h]
@@
 different corner frequencies which would be indicate that the trap 
-is not perfectly circular. Using \textit{ott} we can compare the 
-expected trap strength to what is reported by a quadrant photo diode:
+is not perfectly circular. We can use both \textit{ott} and the 
+trajectory itself to derive an estimation of the trap geometry. 
+The corner frequencies and corresponding trap stiffness are 
+reported below:
+\newpage
+
 \begin{center}
+	\captionof{table}{QPD fitting for single sphere}
 	\begin{tabular}{ |c|c|c|c|c|c|c| } 
 		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory fitting}\\
 		\hline
@@
 		& \parbox{1.25cm}{\centering 523} & 513 \\
+		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
 		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
@@
 		\hline
+
 	\end{tabular}
@@
 
-Where the ellipticity of the beam is given by $e = (1-\kappa_y/
-\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
-wavefront. Its clear from these initial results that the QPD 
-is more sensitive to changes along the y-axis than the x-axis 
-when compared to the direct \textit{ott} calculations. Typically, 
-even an industrial Gaussian beam will produce an elliptical 
-diffraction limit spot when heavily focused; in their tutorial 
-for optimizing the PSD analysis, Berg and Sorensen reported a 
-ellipticity of around 15 \% after a total calibration time of 
-81 seconds \cite{BergSoerensen2004}. The reason for this 
-discrepancy can be explained partly by the fact that our 
-estimation of the trap geometry via \textit{ott} is based on 
-the force-displacement curve while the sphere is moving in only 
-one direction, whereas the QPD is estimating the trap geometry 
-by extrapolating from far field scattering signals. We now 
-consider a symmetric dimer in identical simulative conditions.
+Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, where
+the variance is based upon our choice of frequency boundaries
+({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And the ellipticity 
+of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a 
+measure of the symmetry of the beam wavefront. Its clear from these 
+initial results that the QPD is more sensitive to changes along the 
+y-axis than the x-axis when compared to the direct \textit{ott} 
+calculations. This is somewhat reflected in the trajectory results. 
+Typically, even an industrial Gaussian beam will produce an 
+elliptical diffraction limit spot when heavily focused; in their 
+tutorial for optimizing the PSD analysis, Berg and Sorensen reported 
+a ellipticity of around 15 \% after a total calibration time of 80 
+seconds \cite{BergSoerensen2004}.
+
+The reason for the discrepancies between all 3 methods is due to what 
+is actually being measured. The \textit{ott} estimates are simply 
+looking at the differences in the trapping strength along the Cartesian 
+axis'. If calibrated over an long enough time frame you would expect 
+that the resulting power spectra would exactly mirror the \textit{ott} predictions. However over a short calibration time a QPD can predict
+large discrepancies in the trap strength. There is a clear trade off 
+in terms of accuracy and computation time as shorter calibration runs 
+are computationally more efficient but prone to errors. This is the case
+even if we know the exact positional data.
+
+With this in mind, let us consider a symmetric dimer that is optically
+trapped by the same Gaussian beam. Not only does the dimer's equilibrium
+position change but it is subjected to rotational motion due to the 
+beam's angular momentum. This is reflected
 
 \begin{center}
+	\captionof{table}{QPD fitting for symmetric dimer}
 	\begin{tabular}{ |c|c|c|c|c|c|c| } 
@@
 \section{Conclusions}
-Considering the simplicity of a scatterer such as a dimer,
-one would assume that the dynamics of such an object would be
-relatively easy to predict. Simulations of dimers in the Mie regime
-show that not only do they have multiple positions and orientations 
-in which they can be trapped but also that their interaction
-with circularly polarised light is heavily dependent on the axial 
-position and trapping orientation. Dimer's have the potential to be used
-as adaptive micro-rotors, being simple to synthesise and can be made 
-out of any material of choice, but in order for these particles to be
-used as such one needs a means of characterising the optical torque
-applied. While estimations can be made in the Rayleigh regime due to 
-the dimer being treated as point dipole, Mie regime micro rotors require 
-an in-depth theoretical description of the mechanism that is creating 
-this optical torque.
+Considering the simplicity of a scatterer such as a dimer, one would 
+assume that the dynamics of such an object would be relatively easy 
+to predict. Simulations of dimers in the Mie regime show that not 
+only do they have multiple positions and orientations in which they 
+can be trapped but also that their interaction with circularly 
+polarised light is heavily dependent on the axial position and 
+trapping orientation.
+
+Dimer's have the potential to be used as tunable micro-rotors, being 
+simple to synthesise and can be made out of any material of choice.
+The rotation demonstrated by dimer's is not accurately described in
+previous literature which raises questions on the interactions between
+circularly polarised light and spherical aggregates. Being made up of
+homogenous materials their is no change in polarisation as expected
+by birefringence. Shape birefringence is also not applicable as the 
+preferred orientation is perpendicular to the polarisation axis, 
+meaning there can be no asymmetry in the dimer's susceptibility.

### 2024-10-25 Latex_files/Thesis/main_win.tex
updates

@@
 backend=biber,
-bibstyle=numeric,
-citestyle=numeric,
+bibstyle=ieee,
+citestyle=ieee,
 minbibnames=3, maxbibnames=3,
@@
 \appendix
-\chapter{Stuff That Didn't Fit Anywhere Else}
+\chapter{Appendix}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Estimation of fluid shear rates via galvano mirror}
+An alternative approach to generating fluid shear is to use a 
+galvano-mirror to rapidly move a trapped particle in a bulk 
+fluid. While typically galvano and gimble mirrors are used to 
+trap multiple particles simultaneously, a single micro sphere 
+can be moved quickly through a fluid along a preset path. The 
+only limitation on the particle's speed being the ratio of the 
+trap stiffness to the drag force. For low Reynolds numbers we 
+can assume that the fluid flow is inversely proportional to the
+distance from the particle surface (see Eq.~\eqref{eq:birefringent_shear})
 
+For a simple circular path one can estimate the sphere's speed 
+by the radius of its path $R$ and the frequency of its orbit 
+$\omega$ so that $U = R\omega$. For a more complex path, such 
+as an elliptical orbit the curve needs to be parametrised. One 
+can describe the position of a sphere on an elliptical path as 
+such:
+\begin{align}
+	r(p) = \left[acos(2\pi p), bsin(2\pi p), 0 \right]
+\end{align}
+
+Where a and b are the characteristic radii of an ellipse. If we 
+say that $p$ is some dimensionless parameter that describes the 
+distance travelled along a trajectory (so $p=0$ is the start of 
+the circular orbit and $p=1$ is just as we return to the start).
+Then we can rewrite p as $p=t\omega$ where $\omega$ is the period
+of orbit. Taking the partial derivative of the position gives 
+us the sphere's velocity vector along that path. 
+\begin{align}
+	v(t) = \frac{\delta r(t)}{\delta t} = \left[-2\pi a\omega \ sin(2\pi t\omega),
+	\ 2\pi b\omega \ cos(2\pi t\omega), 0 \right]
+\end{align}
+
+In order to compute the particle's speed $U$ we simply take the 
+magnitude of our velocity. For low velocities the fluid flow at 
+the sphere's surface can be computed based on its velocity.
+\begin{align}
+	u_r(r)=-|v(t)|\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
+\end{align}
+
+Where $r$ is the radial distance to that point. Again taking the partial 
+derivative we can get the shear rate for a particle moving through the fluid:
+\begin{align}
+	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
+	\label{eq:galvo_shear}
+\end{align}
+
+Moving a silica bead along an circular path can generate significant fluid 
+flow around a larger volume compared to comparable micro-rotors. Using 
+\eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
+would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
+equates to moving the silica bead around a circular path with a frequency of 
+$100\ Hz$.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
+	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
+		in bulk water moving at different speeds. The shear rate is calculated 
+		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
+		in a circular path and so the speed is constant through out its path.}
+	\label{fig:galvano_shear}
+\end{figure}
+
+From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
+mirror a better option for generating high shear rates but also over a larger 
+volume. The only limitation to this method is the effective drag experienced 
+by the trapped particle. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-10-31 Latex_files/Thesis/chapters/chapter 3/main.tex
chapter 3 updates

@@
 was suggested, using a galvano mirror to move a particle 
-quickly through a supersaturated solution. While a particle
-is within the optical trap the gradient forces are 
-significantly reduced meaning simply trapping a particle is
-insufficient for inducing nucleation \cite{Flannigan2023}. 
-This is why we did not see any nucleation events while 
-rotating the vaterite, there is not a large gradient force
-to draw in material. We did not consider a focus beam by 
-itself as this has already been demonstrated in previous
-experiments \cite{Flannigan2023, Liao2022, Rungsimanon2010}.
-
-Preliminary calculations suggested a particle moving rapidly
-through a fluid could produce significantly greater shear 
-rates (see \emph{Appendix X.X} for breakdown) than a rotating 
-particle. During initial testing using silica beads we found 
-that even in undersaturated solutions the optical trap would 
-produce a crystal nucleus when the trap was empty. This has 
-been reported prior \cite{Rungsimanon2010, Liao2022}, but 
-what is more interesting is how the beam's motion influenced 
-the growth of the nucleus.
+quickly through a supersaturated solution. Preliminary 
+calculations suggested a particle moving rapidly through 
+a fluid could produce significantly greater shear rates 
+(see \emph{Appendix X.X} for breakdown) than a rotating 
+particle. While a particle is within the optical trap the 
+gradient forces are significantly reduced meaning simply 
+trapping a particle is insufficient for inducing nucleation \cite{Flannigan2023}. This is why we did not see any 
+nucleation events while rotating the vaterite, there is 
+not a large gradient force to draw in material.
+
+During initial testing using silica beads we found that 
+even in undersaturated solutions the optical trap would 
+produce a crystal nucleus when the trap was empty. This 
+has been reported prior \cite{Rungsimanon2010, Liao2022}, 
+but what is more interesting is how the beam's motion 
+influenced the growth of the nucleus.
+
+\subsubsection{Choice of Crystal growth rate units}
+The growth rate of a crystal can be described using different
+units depending on the situation. For example, seed crystals 
+often use growth units in the form of \textit{length per unit
+time}. This can either be used to describe the overall length
+of the longest edge of a crystal, or to describe how individual
+faces grow separately. This works well for crystals with a 
+clear morphology, in regards to laser induced nucleation the 
+crystal growth is heavily influenced by the local fluid 
+conditions meaning tracking the length of an individual face 
+is challenging. Instead we utilise image analysis software 
+such as imageJ to measure the area of the crystal after each 
+frame to get a rough estimate of the crystal growth rate in 
+units of \textit{length$^2$ per unit time}. With the addition 
+of the contact angle information from \cite{Flannigan2023} we 
+can approximate the crystal height based on how close it is to 
+the droplet edge.
 
@@
 Its well known that a supersaturated solution can nucleate
-if irradiated by a focused beam. We confirmed that our 
-laser was of sufficient power and focus to induce crystal
-nucleation. The laser was focused at the edge of a $20\mu L$
-droplet consisting of water and dissolved glycine ($S=1.03$). 
-The solution was monitored for 10 minutes, if a nucleation 
-event was observed the event was recorded using a high speed 
-CCD. If no nucleation event was observed after 10 minutes 
-the solution was disposed of and a new sample was prepared. 
-After ten repeats only 30\% of the samples observed a 
-nucleation event at the focus. All of which displayed similar 
+if irradiated by a focused beam with sufficient power 
+\cite{Rungsimanon2010}. First we look at how a laser induced
+nucleation event unfolds using a stationary focused beam. 
+The laser was focused at the edge of a $20\mu L$ droplet 
+consisting of water and dissolved glycine ($S=1.03$). The 
+solution was monitored for 10 minutes, if a nucleation event 
+was observed the event was recorded using a high speed CCD. 
+If no nucleation event was observed after 10 minutes the 
+solution was disposed of and a new sample was prepared. After 
+ten repeats only 30\% of the samples observed a nucleation 
+event at the focus. All of which displayed similar growth 
 behaviour to one another.
@@
 m^2/min$) from the focal point of the trap until after roughly 
-$6$ seconds the crystal escapes. Comparing to previous literature
-using optical tweezers shows that the growth rate is only loosely 
-connected to the solutions supersaturation. Crystal growth rates 
-for a solution of glycine and water ($S\approx1.00$) was found to
-vary from as much as $3900 \mu\ m^2/min$ to as low as $675\mu 
-m^2/min$. Local conditions play a much larger role in the growth 
-rate than just the concentration \cite{Flannigan2023}. A likely 
-reason that the trap is escaped is due to the fact that crystal 
-is far too large to be held in place and is in fact still growing 
-as the solution is supersaturated. The key take away to remember 
-is that the beam has no influence over the crystal shape, instead 
-it grows outward from the trap focus. Furthermore, due to the fact 
-that the solution is supersaturated the crystal growth cannot be 
-contained to the trap focus. Instead the crystal escapes as its 
-size exceeds the trap focus. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.75\linewidth, height=1.1\linewidth]{frames_no_beam_movement.pdf}
+$6$ seconds the crystal escapes. A likely reason that the trap 
+is escaped is due to the fact that crystal is far too large to 
+be held in place and is in fact still growing as the solution is supersaturated. Based on the contact angle measurements from \cite{Flannigan2023}, the droplet height could not have been 
+any more than $2-5\mu m$ at the trap focus.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth, height=1.3\linewidth]{frames_no_beam_movement.pdf}
 	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
@@
 
-\newpage
+Comparing to previous literature using optical tweezers shows 
+that the growth rate is only loosely connected to the solutions 
+supersaturation. Crystal growth rates for a solution of glycine 
+and water ($S\approx1.00$) was found to vary from as much as 
+$3900 \mu\ m^2/min$ to as low as $675\mu m^2/min$ 
+\cite{Flannigan2023}. The reason for such high crystal growth 
+rates is due primarily to the tweezer focus bringing in material 
+due to gradient forces. A study of glycine crystals in water 
+found that for a similarly supersaturated solution the growth 
+rate along the {011} face was around $0.1\ \mu m/min$ where as 
+the {010} face was found to have a growth rate of around $0.01\ 
+\mu m/min$. While its difficult to extrapolate an exact area 
+growth rate from these results we can approximate that if the 
+we looked directly perpendicular to the crystal growth the area 
+would increase at $\approx0.02\ \mu^2/min$ (see X.X for a 
+breakdown of this approximation).
+
+The key take away to remember is that the beam has no influence 
+over the crystal shape, instead it grows outward from the trap 
+focus. Furthermore, due to the fact that the solution is 
+supersaturated the crystal growth cannot be contained to the 
+trap focus. Instead the crystal escapes as its size exceeds the 
+trap focus.
+
 \subsection{Moving Beam}
 \label{sec:moving}
-To test if a rapidly moving silica bead could generate the necessary shear rate 
-for crystal nucleation we wanted to see if a trapped silica bead could be trapped
-and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
-added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
-unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
-encountered several unexpected results.
-
-Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
-moving in a small elliptical pattern; interestingly. After a few minutes we 
-noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
-for a larger example of these droplets). These appear too have a wide distribution
-of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
-\mu m$. We surmise these could be small clusters of Glycine that had previously 
-been shown to form when the aqueous solutions where irradiated with a 
-focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
-directly entering the focus a nucleus forms close to the droplet edge, 
-unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
-the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
-being moved by and growing around the focal point of the trap. Because of 
-this the crystal nucleus lacks a clear morphology at first. Until roughly 
-$20\ s$ the crystal reaches a almost prismatic structure, with further 
-irradiation increasing the size.
-
-Interestingly the galvano-mirror allows the trap to impart a slight torque 
-on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
-where even though the crystal is not directly in the trap focus it rotates 
-in the $x-y$ plane and gets trapped again at a corner. The rotation could 
-not be due to fluid flow close to the surface of the crystal as the dipole 
-moment of individual water molecules is too small to be influenced by an 
-optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
-growth becomes localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
-\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
-(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
+To test if a rapidly moving silica bead could generate the 
+necessary shear rate for crystal nucleation we wanted to 
+see if a trapped silica bead could be trapped and moved in 
+an aqueous solution. $20\ \mu L$ of glycine and water ($S=
+1.03$) was added to $10\ \mu L$ of a dilute water-silica 
+mixture making the solution unsaturated ($S\approx0.7$). 
+However, due to the beam's motion we instead encountered 
+unexpected growth behaviour.
+
+Shown below in Fig.~\ref{fig:eliptical_beam_1} where 
+we have the laser focus moving in a small elliptical 
+pattern. While nothing is seen directly entering the 
+focus a nucleus forms close to the droplet edge, 
+unlike in fig.~\ref{fig:stationary_beam} the crystal 
+does not grow out from the focal point evenly. Due to 
+the galvano mirror, the crystal is simultaneously 
+being moved by and growing around the focal point of 
+the trap. Because of this the crystal nucleus lacks 
+a clear morphology at first. Until roughly $20\ s$ 
+the crystal reaches a almost prismatic structure, 
+with further irradiation increasing the size.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{frames_eliptical_beam.pdf}
 	\caption{Frames from a longer video depicting the growth 
@@
 
-Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
-\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
-moving beam. This modification allows for the crystal growth to be localised to
-a specific region of the bulk crystal whereas with a stationary beam there is no
-control over the crystal morphology. In fact this allows for a much finer control
-over the exact shape of the crystal nucleus, in some cases allowing for growth out
-of the viewing plane as shown in
-
-\begin{figure}
-	\centering
-\end{figure}
+Interestingly the galvano-mirror allows the trap to 
+impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
+though the crystal is not directly in the trap focus 
+it rotates in the $x-y$ plane and gets trapped again 
+at a corner. The rotation could not be due to fluid 
+flow close to the surface of the crystal as the 
+dipole moment of individual water molecules is too 
+small to be influenced by an optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
+crystal growth becomes localised to the corner. The 
+area growth rate between figures \ref{fig:eliptical_beam_1}
+(a) and (d) was approximated using imageJ at $45.03\ 
+\mu m^2 /min$, where as between figures 
+\ref{fig:eliptical_beam_1}(e) and (f) the growth rate 
+at that particular edge was estimated at $42.10\ \mu m^2/min$.
+
+Nucleation in undersaturated conditions has been 
+reported previously in $D_2O$ \cite{Rungsimanon2010} 
+and $H_2O$ \cite{Flannigan2023}, though not involving 
+a moving beam. This modification allows for the crystal 
+growth to be localised to a specific region of the bulk 
+crystal whereas with a stationary beam there is no 
+control over the crystal morphology. In fact this allows 
+for a much finer control over the exact shape of the 
+crystal nucleus, in some cases allowing for growth out
+of the viewing plane as shown in fig.~\ref{fig:out_of_plane}.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{out_of_plane.pdf}
+	\caption{Interesting crystal growth on the surface of an
+	existing crystal. Initially the crystal surface seems 
+	flat (a) but eventually a different crystal structure 
+	grows out off of the surface (b).}
+	\label{fig:out_of_plane}
+\end{figure}
+
+Where a large outgrowth forms on the surface of a nucleus,
+over a period of $30\ s$ the outgrowth develops into a 
+rhombic crystal structure. The unique crystal growth 
+demonstrated is likely a factor of the moving beam focus,
+but this raises a question over how the laser can localise
+the growth only around the focus and why the crystal does 
+not grow or shrink when not directly within the trap. There
+should be some material that is being drawn into the trap
+as it scans across the camera frame.
 
@@
 \label{sec:clusters}
-One common aspect involving these droplets is the fact when brought 
-to the laser focus the droplets would nucleate immediately 
-\cite{Liao2022}. A solution similar to Sec~\ref{sec:stationary} was 
-made up, but without any silica droplets. Once again the beam was 
-focused close to the droplet edge, this time the galvano mirror was 
-scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
-After a few minutes of irradiation droplets were seen entering the 
-camera frame, because no silica had been added these droplets had 
-to be from the glycine solution. Trapping individual droplets did 
-not result in immediate nucleation even after several minutes being 
-trapped. Trying to bring two droplets together resulted in nucleation 
-between the two droplets, compared to \ref{sec:stationary} \& 
-\ref{sec:moving} the growth is much slower, taking nearly 40 seconds 
-before the crystal structure becomes clearer.
+In a repeat experiment a solution similar to sec.~
+\ref{sec:stationary} was made up, but without any 
+silica micro-spheres. Once again the beam was focused 
+close to the droplet edge, this time the galvano 
+mirror was scanning a circular path (as shown in fig.~\ref{fig:cluster_trapping}). After a few 
+minutes of irradiation droplets were seen entering 
+the camera frame. Because no silica had been added, 
+and that the solutions were filtered, these droplets 
+had to be from the glycine solution. Trapping 
+individual droplets did not result in immediate 
+nucleation even after several minutes being trapped. 
+Trying to bring two droplets together resulted in 
+nucleation between the two droplets, compared to 
+\ref{sec:stationary} \& \ref{sec:moving} the growth 
+is much slower, taking nearly 40 seconds before the 
+crystal structure becomes clearer.
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.8\linewidth]{cluster_trapping.pdf}
-	\caption{Frames from a longer video demonstrating the trapping of a glycine 
-		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
-		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
-		black) being brought into contact with a larger droplet. (b) upon contact 
-		a nucleus can be seen between the two droplets. The growth is rather slow 
-		with the crystal having no clear defined morphology through (c) and (d). 
-		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
+	\caption{Frames from a longer video demonstrating the trapping 
+		of a glycine droplet. Solution is undersaturated glycine 
+		and water (S = 0.86), with the laser power is set at $750\ 
+		mW$. (a) shows a trapped droplet (outlined in black) being 
+		brought into contact with a larger droplet. (b) upon contact 
+		a nucleus can be seen between the two droplets. The growth 
+		is rather slow with the crystal having no clear defined 
+		morphology through (c) and (d). Between frames (e) and (f) 
+		the larger droplet finally joins the main crystal.}
 	\label{fig:cluster_trapping}
@@
 
-The fact that these droplets can be trapped indicates they must 
-have a higher refractive index than the surrounding solution. 
-Previous reports have shown that a focused trap will draw in 
-solute material from the bulk \cite{Tsuboi2009, Gowayed2021}. It 
-has been shown that the concentration of glycine solution is correlated 
-with the refractive indices of the liquid \cite{Gowayed2021, Orttung1963}. 
-This suggests that these droplets serve to provide solute material 
-to the bulk crystal.
-
+The fact that these droplets can be trapped indicates 
+they must have a higher refractive index than the 
+surrounding solution. Previous reports have shown that 
+a focused trap will draw in solute material from the 
+bulk \cite{Tsuboi2009, Gowayed2021}. This suggests that 
+the droplets observed are concentrated spheres of glycine, 
+which also explains how nucleation was initiated when those
+two spheres were brought into contact with one another. The 
+droplets must be providing material for the crystal growth.
+It has been shown that the concentration of glycine solution 
+is correlated with the refractive indices of the liquid 
+\cite{Gowayed2021, Orttung1963}. This suggests that we could
+measure the concentration of individual droplets based on the
+trapping strength of each droplet. 
 \newpage
@@
 \subsection{Influence of a moving beam front on seed crystals}
-One potential application of this phenomena would be in using a 
-moving the moving beam front as a method for shaping the final
-crystal morphology. Glycine seed crystals were grown via 
-evaporative crystallisation over night. The resultant crystals
-were as wide as $0.5 mm$ in some cases. Individual crystals were 
-collected and suspended in a water + glycine solution ($S = 1.001$)
-and the laser was scanned along a narrow linear path.
-
-As shown by fig.~\ref{fig:seed_crystals} the laser does not in fact 
-promote crystal growth but instead forces the crystal to dissolve 
-into the bulk solution. The dissolution rate is not as substantial 
-as the growth rate seen in previous sections. It takes over 20 
-minutes for the for the crystal surface to dissolve more than a 
-few microns. Furthermore we do not report any sightings of droplets
-close to the surface of the crystal nor do we see anything enter the 
-optical trap.
+\label{sec:seed_crystals}
+One potential application of this phenomena would be in using 
+a moving the moving beam front as a method for shaping the 
+final crystal morphology. Therefore, we wanted to test if a 
+moving beam front could have any influence on the shape of a
+seed crystal submerged in a bulk solution. Glycine seed crystals 
+were grown via evaporative crystallisation over night. The 
+resultant crystals were as wide as $0.5 mm$ in some cases. 
+Individual crystals were collected and suspended in a water 
++ glycine solution ($S = 1.001$) and the laser was scanned 
+along a narrow linear path. It should be noted that the seed 
+crystal was fully submerged in the solution, meaning that there
+is no interface between the air and solution close to the seed.
+Five separate repeats were conducted, all demonstrating similar
+behaviour to one another.
+
+As shown by fig.~\ref{fig:seed_crystals} the laser does not 
+in fact promote crystal growth but instead forces the crystal 
+to dissolve into the bulk solution. The dissolution rate is not 
+as substantial as the growth rate seen in previous sections. 
+It takes over 20 minutes for the for the crystal surface to 
+dissolve more than a few microns. Furthermore we do not report 
+any sightings of droplets close to the surface of the crystal 
+nor do we see anything enter the optical trap.
 \begin{figure}[h!]
@@
 	\includegraphics[width=\linewidth]{seed_crystals.pdf}
+	\caption{Frames from a seed crystal being irradiated by a 
+		scanning laser. The laser path is shown below in red. 
+		Initially the crystal edge is nearly perpendicular to 
+		the laser path. Over the course of nearly twenty minutes 
+		the crystal around the laser has dissolved leaving a 
+		concave indent.}
 	\label{fig:seed_crystals}
-	\caption{Frames from a seed crystal being irradiated by a scanning
-		laser. The laser path is shown below in red. Initially the 
-		crystal edge is nearly perpendicular to the laser path. However,
-		over the course of nearly twenty minutes the crystal around the 
-		laser has dissolved leaving a concave indent.}
-\end{figure}
-
-This has some implications: Firstly, the lack of supposed glycine 
-droplets suggests that the presence of an interface is crucial in 
-order to promote the formation of these droplets. This is consistent
-with previous reports on laser nucleation, where a air-solution or 
-solution-glass interface is necessary for the trap focus to have any 
-noticeable effect on the solute \cite{Gowayed2021, Liao2022, 
-Rungsimanon2010}. Secondly, the fact that the seed crystal dissolved 
-in a saturated solution suggests that this could be due to a localised
-heating effect. A commonly used heating model for optical tweezers is 
-the Peterman model \cite{Peterman2003} which accounts for the lateral 
-distance between the optical trap and a local heat sink (such as the 
-glass cover slip or microscope objective). Even with the heat sink 
-the predicted temperature rise in pure water would be around $5-10 K$.  
-The fact that the seed crystal dissolves so slowly could simply be due 
-to the fact that the local heating is lessened by the moving beam front.
+\end{figure}
+
+This has some implications: Firstly, the lack of supposed 
+glycine droplets suggests that the presence of an interface 
+is crucial in order to promote the formation of these droplets. 
+This is consistent with previous reports on laser nucleation, 
+where a air-solution or solution-glass interface is necessary 
+for the trap focus to have any noticeable effect on the solute \cite{Gowayed2021, Liao2022, Rungsimanon2010}. Secondly, the 
+fact that the seed crystal dissolved in a saturated solution 
+suggests that this could be due to a localised heating effect. 
+A commonly used heating model for optical tweezers is the 
+Peterman model \cite{Peterman2003} which accounts for the lateral 
+distance between the optical trap and a local heat sink (such 
+as the glass cover slip or microscope objective). Even with 
+the heat sink the predicted temperature rise in pure water would 
+be around $5-10 K$.  The fact that the seed crystal dissolves 
+so slowly could simply be due to the fact that the local heating 
+is lessened by the moving beam front.
 
 \section{Summary of Moving Beam Phenomena}
-To summarise, the introduction of a moving beam helps to accelerate the 
-local growth of a newly formed crystal. This is due to the presence of 
-glycine droplets that accumulate near the interface between the liquid 
-solution and air. The theory behind the localised growth can be summarised 
+To summarise, the introduction of a moving beam helps to 
+accelerate the local growth of a newly formed crystal. It 
+seems that this phenomena will only occur when close to 
+interface between the solution and air. We note the presence
+of droplets that can sometimes be seen entering the trap, 
+though this is not always necessary for a nucleation event 
+to occur. A plausible description of the phenomena is described
 thusly.
 
-Initial nucleation is similar to typical optical trapping induced nucleation,
-with the air solution interface limiting the molecular mobility of the solute
-molecules \cite{Liao2022, Sugiyama2009, Gowayed2021}. The moving beam front 
-can influence the motion of the nucleus initially, but eventually the drag 
-force means the crystal is not moved by the optical trap. Localised crystal
-growth occurs when the trap is close to or partially over the interface of 
-the crystal (see figure~\ref{fig:local_nucleation}(a)). As shown in \ref{sec:clusters}, 
-the optical trap can manipulate these droplets similar 
-to microspheres. When in close proximity to the trap these droplets are 
-brought towards the crystal surface (see figure~\ref{fig:local_nucleation}(b)). 
-These provide material that grows the crystal around that region (see figure~\ref{fig:local_nucleation}(c)). Eventually the local solution is either 
-depleted of solute material or the crystal front has grown to fully encompass 
-the trap, preventing further growth (see figure~\ref{fig:local_nucleation}(d)). 
+Initial nucleation is similar to typical optical trapping 
+induced nucleation, with the air solution interface limiting 
+the molecular mobility of the solute molecules \cite{Liao2022, 
+Sugiyama2009, Gowayed2021}. The moving beam front can influence 
+the motion of the nucleus initially, but eventually the drag 
+force means the crystal is not moved by the optical trap. 
+Localised crystal growth occurs when the trap is close to or 
+partially over the interface of the crystal (see fig.~
+\ref{fig:local_nucleation}(a)). This suggests that the 
+laser itself is bringing in new solute material that can then
+adhere to the crystal surface. Interestingly even when the 
+crystal growth is localised to a small section of the crystal
+we do not observe the crystal dissolving. This is consistent 
+with other observed laser induced nucleation results, where 
+as long as the laser is active the crystal remains stable 
+within the solution \cite{Flannigan2023, Rungsimanon2010,
+Liao2022, Sugiyama2009}. It does raise further questions 
+however, mainly in an undersaturated solution is a moving 
+beam front able to increase the size of the formed nucleus
+compared to a stationary beam? Additionally, it raises the 
+question about what the role the 'glycine droplets' play in 
+the nucleation process.
+
+As shown in \ref{sec:clusters}, the optical trap can manipulate 
+these droplets similar to microspheres. When in close proximity 
+to the trap these droplets are brought towards the crystal 
+surface (see fig.~\ref{fig:local_nucleation}(b)). We have 
+already suggested that these droplets must contain glycine as 
+they are to large to be a silica microsphere, and the solutions 
+were filtered prior to being studied. The droplets provide 
+material that grows the crystal around that region (see fig.~\ref{fig:local_nucleation}(c)). Eventually the local 
+solution is either depleted of solute material or the crystal 
+front has grown to fully encompass the trap, preventing further 
+growth (see fig.~\ref{fig:local_nucleation}(d)). 
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.8\linewidth]{galvano_diagram.pdf}
-	\caption{Diagram outlining how a moving beam assists in the growth of a crystal
-		nucleus. (a) a crystal nucleus is partially trapped by a moving beam with 
-		solute droplets close to its surface. (b) droplets close to the laser
-		focus are drawn in by gradient forces and moved towards the crystal surface.
-		(c) these droplets provide material to the main crystal, resulting in
-		localised growth around the laser focus. (d) eventually the crystal area 
-		either fully surrounds the laser focus or the solution surrounding the laser
-		is depleted of solute material.}
+	\caption{Diagram outlining how a moving beam assists in the 
+		growth of a crystal nucleus. (a) a crystal nucleus is 
+		partially trapped by a moving beam with solute droplets 
+		close to its surface. (b) droplets close to the laser
+		focus can be drawn in by gradient forces and moved 
+		towards the crystal surface. (c) the laser brings in 
+		material that is then deposited on the surface of the 
+		crystal. (d) eventually the crystal area either fully 
+		surrounds the laser focus or the solution surrounding 
+		the laser is depleted of solute material.}
 	\label{fig:local_nucleation}
@@
 
-The reason the seed crystals saw no further crystal growth when irradiated 
-with the trap is due to the fact that these experiments were carried out 
-in a bulk solution, absent of any interfaces. As such the clusters seen in
-\ref{sec:moving} \& \ref{sec:clusters} are not present and cannot provide
-material to accelerate crystal growth.
-
-There are still several factors that need to be investigated, but due to time
-constraints it was not possible to properly study this phenomena. Firstly, 
-there is the question of what conditions result in the production of concentrated
-droplets, it is not clear if the presence of a laser is required or if these
-droplets naturally occurring. Prior literature would suggest that the laser is 
-required \cite{Liao2022, Tsuboi2009}, but this would not explain why in many 
-cases the droplets are found far outside the influence of the optical trap. It
-has been suggested that optical traps can attract microspheres over a wider 
-area, but it is hard to say that this would result in the creation of liquid 
-droplets \cite{Yi2021}.
-
-Secondly, there is the question of how these droplets supply material to the 
-bulk crystal. In some instances it is clear that the droplets are being drawn 
-into the trap, however, in other instances while there are no droplets close to
-the vicinity of the optical trap the crystal continues to grow. If these droplets
-are a necessary precursor to induce crystal nucleation then understanding how
-they provide materials to the bulk crystal may help with our understanding of 
-the kinetics of multi-step nucleation.
+What remains unclear is why the solution interface is necessary
+for localised crystal growth. We know that the glycine droplets 
+will form when the laser is close to the glass-solution interface
+\cite{Gowayed2021, Yuyama2010, Yuyama2012}. This is consistent 
+with our results as being able to image the droplet edge requires 
+that the laser is brought close to the cover slip. However it is 
+not clear how exactly these droplets contribute to localised 
+crystal growth as in sec.~\ref{sec:moving} no droplets are seen 
+entering the trap. It seems likely that the presence of droplets 
+is necessary as we see that the seed crystals dissolve while 
+there are no droplets present.
+
+There are still several factors that need to be investigated. 
+Firstly, there is the question of what conditions result in the 
+production of concentrated droplets, it is not clear if the 
+presence of a laser is required or if these droplets naturally 
+occurring. Prior literature would suggest that the laser is 
+required \cite{Liao2022, Tsuboi2009}, but this would not explain 
+why in many cases the droplets are found far outside the influence 
+of the optical trap. It has been shown that phase separation 
+occurs when a trap is focused close to the cover slip 
+\cite{Gowayed2021, Yuyama2010, Yuyama2012}, but there has been 
+no cases of nucleation occurring at the cover slip. Furthermore
+these experiments all occurred in supersaturated condition 
+($S\ge 1.50$) with a much wider focus ($NA \le 1.00$) whereas 
+our results occurred in undersaturated conditions with a 
+tighter focus. We did not not any occurrence of a phase 
+separation as reported by \cite{Gowayed2021, Yuyama2010, Yuyama2012}
+and yet we see the appearance of droplets that we surmise 
+are partially constructed of glycine.
+
+Secondly, there is the question of how these droplets supply 
+material to the bulk crystal. In some instances it is clear 
+that the droplets are being drawn into the trap, however, in 
+other instances while there are no droplets close to the 
+vicinity of the optical trap the crystal continues to grow. 
+If these droplets are a necessary precursor to induce crystal 
+nucleation then understanding how they provide materials to 
+the bulk crystal may help with our understanding of the 
+kinetics of multi-step nucleation.
 
 \section{Conclusion}
-
-In conclusion, the experimental results from this chapter provide no 
-affirmative evidence of shear induced crystal nucleation. This is in
-part due to the limited rotational speed that could be achieved using
-vaterite microspheres. Simply adding a micro-rotor to a supersaturated
-solution will not yield a significant fluid flow to induce crystal 
-nucleation. Furthermore, the region in which the local shear rate is 
-significant enough is so small that any nucleation events that could 
-occur there would not grow large enough to be stable. This does not 
-however, suggest that shear induced nucleation is impossible at a 
-micro scale. In fact there have been several developments lately that
-allow for the precise rotation of multiple rotors to control the 
-precise motion of silica beads. The challenge lies in both building the
-optical equipment to form multiple strong traps (~ 3 per rotor). These 
-traps would have to be strong enough to not only hold the rotors but 
-also move them at a sufficiently fast rotational speed that they could
-induce a significant enough fluid flow.
-
-On the other hand, the results utilising the galvano mirror have some 
-interesting implications for future research into crystal nucleation
-using optical tweezers. We noted that in undersaturated conditions the
-presence of 'droplets' that could be trapped and later nucleated when 
-brought into contact with one another. The fact that these droplets 
-could be trapped suggests they have a higher refractive index to the 
-solution, given that the solutions were filtered prior to irradiation
-dismisses that they could be dust particles. Prior literature suggests
-that optical trapping of a supersaturated solution can lead to the 
-production of nano droplets \cite{Gowayed2021, Tsuboi2009}. Controlling
-the localised growth requires that the growth and trapping are handled
-by two separate lasers. Optical traps for amorphous solids has been 
-shown as a possibility but the algorithm would need to be adaptive 
-to the changing crystal shape.
+In conclusion, the experimental results from this chapter 
+provide no affirmative evidence of shear induced crystal 
+nucleation. This is in part due to the limited rotational 
+speed that could be achieved using vaterite microspheres. 
+Simply adding a micro-rotor to a supersaturated solution 
+will not yield a significant fluid flow to induce crystal 
+nucleation. Furthermore, the region in which the local 
+shear rate is significant enough is so small that any 
+nucleation events that could occur there may not grow large 
+enough to be stable. This does not however, suggest that 
+shear induced nucleation is impossible at a micro scale. 
+In fact there have been several developments lately that
+allow for the precise rotation of multiple rotors to 
+control the precise motion of silica beads \cite{Butaite2019}. 
+This would allow for direct control of the shear rate 
+within a given volume of fluid, allowing one to study the 
+impact of fluid flow on a nucleus at a micro-scale.
+
+On the other hand, the results utilising the galvano mirror 
+have some interesting implications for future research into 
+crystal nucleation using optical tweezers. We noted that a 
+moving beam front is capable of localising the crystal growth
+around the beam focus in undersaturated conditions. Typically
+in laser induced nucleation the final crystal morphology is 
+influenced primarily by the local fluid conditions 
+\cite{Flannigan2023, Liao2022}. If this phenomena could be 
+fully understood we could have a means of directly manipulating 
+the morphology of any nucleating crystal. We noted that in 
+undersaturated conditions the presence of droplets that could 
+be trapped and later nucleated when brought into contact with 
+one another. There remains some uncertainty over the role these 
+droplets play in a nucleation event. Further research is also 
+necessary to understand how the fluid and laser parameters 
+(i.e. supersaturation and numerical aperture) impact the 
+formation of these droplets. The fact that these droplets could 
+be trapped suggests they have a higher refractive index to the 
+solution, given that the solutions were filtered prior to 
+irradiation dismisses that they could be dust particles. Prior 
+literature suggests that optical trapping of a supersaturated 
+solution can lead to the production of nano droplets 
+\cite{Gowayed2021, Tsuboi2009}.

### 2024-10-31 Latex_files/Thesis/main_win.tex
chapter 3 updates

@@
 \appendix
-\chapter{Appendix}}
+\chapter{Appendix}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 by the trapped particle. 
+\newpage
+
+\section{Approximation of Area growth rate}
+Data collected from \cite{Li1992} provides face growth rates 
+for glycine seed crystals in supersaturated solutions. The 
+reported growth rates are given in terms of $\mu m/sec$, in 
+order to compare to the laser induced nucleation events we 
+need some method of converting to area units. In \cite{Li1992}
+the seed crystals are viewed from a top down perspective with
+the edges marked $L_b$ and $L_c$. The authors calculated the 
+face growth rates by:
+\begin{align}
+	G_{011} = \frac{1}{2}sin(67.5^\circ)\frac{dL_c}{dt} \\
+	G_{010} = \frac{1}{2}sin(68.5^\circ)\frac{dL_b}{dt}
+\end{align}
+
+The face growth rates are provided in the final paper, its
+easy enough to rearrange the two equations to get $dL_c/dt$
+and $dL_b/dt$. The area growth rate can be computed via the
+chain rule, where we overestimate the area of the glycine 
+crystal to have a rectangular cross section of side lengths
+$L_b$ and $L_c$. Numerically computing the area gives us an
+approximate total growth rate of around $0.02 \mu m^2/min$.
+This is also an over estimation of the crystal growth rates 
+as in \cite{Li1992} they only consider crystals between 
+$30\mu m$ and $80\mu m$ and note that the growth rate of 
+smaller crystals is even slower.
+
+Overall this demonstrates clearly that the growth rate of a
+single crystal is accelerated by the prescience of an optical
+trap, at least just after nucleation. The presence of an 
+optical trap close to a larger seed crystal is shown to 
+inhibit the growth of the crystal front, both with a stationary
+beam \cite{Flannigan2023} and a moving beam (see sec.~
+\ref{sec:seed_crystals})
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-11-05 Latex_files/Thesis/chapters/chapter 1/main.tex
praveen commments plus chp 4 rework

@@
 \chapter{Introduction}
-Crystallisation is one of the oldest forms of phase separation used by humanity\cite{Schoen1956}, put simply, it is the formation and growth 
-of a new structured phase within a disordered bulk phase. This has 
-applications in a number of industries such as pharmaceuticals, food 
-production, and electronics \cite{Myerson2002}. Where the extraction 
-of dilute materials can help improve product quality while keeping 
-production costs low. The industrialisation of crystallisation has 
-allowed engineers to reliably and efficiently induce the crystal 
-formation within a bulk phase. However, while on a large scale 
-crystallisation is seemingly an understood physical process, just a 
-small amount of investigation into the literature reveals that at a 
-micro scale there is yet to be unifying theory that can accurately 
-explain the process of crystallisation \cite{Fu2021}. Of which there 
-two primary areas of focus, nucleation and crystal growth. The latter 
-focusing on how an already stable crystal grows and how it takes on its 
-final shape. Whereas the former is more concerned with which factors 
-contribute to its formation in order to control crystal growth.SW
+Crystallisation is one of the oldest forms of phase 
+separation used by humanity\cite{Schoen1956}, put 
+simply, it is the formation and growth of a new 
+structured phase within a disordered bulk phase. 
+This has applications in a number of industries such 
+as pharmaceuticals \cite{Gao2017}, food production 
+\cite{Hartel2002}, and electronics \cite{Myerson2002}. 
+Where the extraction of dilute materials can help 
+improve product quality while keeping production costs 
+low. The industrialisation of crystallisation has 
+allowed engineers to reliably and efficiently induce 
+the crystal formation within a bulk phase. However, 
+while on a large scale crystallisation is seemingly an 
+understood physical process, just a small amount of 
+investigation into the literature reveals that at a 
+micro scale there is yet to be unifying theory that can 
+accurately explain the process of crystallisation 
+\cite{Fu2021}. Crystallisation can be subdivided into 
+two steps: Nucleation and Crystal growth. The latter 
+focusing on how an already stable crystal grows and how 
+it takes on its final shape. Whereas the former is more 
+concerned with the first initial moments of a crystals 
+formation within a bulk phase.
 
-In this chapter we will outline our current understanding of nucleation,
-why there is a gap in the literature, and why there is still a need for
-local control of nucleation events. Furthermore, we will also highlight
-recent developments involving the use of optical tweezers (and other 
-laser based methods) in order to develop our understanding of nucleation.
-Lastly we highlight the connection between optical tweezer's ability to 
-probe fluid viscosity via rotational motion, and the known link between
-nucleation and fluid flow, to propose a novel potential method of 
-creating localised nucleation events.
+In this chapter we will outline our current understanding 
+of nucleation,why there is a gap in the literature, and why 
+there is still a need for local control of nucleation events. 
+Furthermore, we will also highlight recent developments 
+involving the use of optical tweezers (and other laser based 
+methods) in order to develop our understanding of nucleation. 
+Lastly we highlight the connection between optical tweezer's 
+ability to probe fluid viscosity via rotational motion, and 
+the known link between nucleation and fluid flow, to propose 
+a novel potential method of creating localised nucleation events.
 
 \section{Nucleation}
-Nucleation is an example of a binary phase separation, where a dilute phase 
-is miscible in a bulk phase, more often called the solute and solvent 
-respectively. Because of thermodynamics, the two can only remain in 
-equilibrium while below a specific concentration ($C_{eq}$) - below which 
-the chemical potential $\mu$ for a miscible solution is greater than the 
-potential required to separate the two phases. Once $C_{eq}$ is exceeded 
-there is a chemical potential difference driving the solution to separate 
-the two phases. Since different combinations of solute and solvent will 
-have different equilibrium concentrations, researchers often instead measure 
-the ratio between the solute and solvent by using 'supersaturation', one 
-form of measuring supersaturation is shown below \cite{Mullin2001}:
+Nucleation is an example of a binary phase separation, 
+where a dilute phase is miscible in a bulk phase, more 
+often called the solute and solvent respectively. Because 
+of thermodynamics, the two can only remain in equilibrium 
+while below a specific concentration ($C_{eq}$) - below 
+which the chemical potential $\mu$ for a miscible solution 
+is greater than the potential required to separate the two 
+phases. Once $C_{eq}$ is exceeded there is a chemical potential 
+difference driving the solution to separate the two phases. 
+Since different combinations of solute and solvent will 
+have different equilibrium concentrations, researchers often 
+instead measure the ratio between the solute and solvent by 
+using 'supersaturation' \cite{Mullin2001}:
 \begin{align}
@@
 
-Where $C_{sol}$ is just the solute concentration, and $C_{eq}(T)$is the 
-equilibrium concentration at temperature $T$. While the solution remains 
-supersaturated there is a chemical potential driving force for the solute 
-to coalesce and separate from the solution as an ordered solid, the first 
-formation of the crystal is referred to as the nucleus and understanding its 
-formation of has been the focus of researchers for decades now. Typically, 
-for an industrial crystallisation process the working principle is based on 
-controlling and manipulating the supersaturation of the system. Of course 
-nucleation events in a variety of different situations which each require 
-their own classification.
+Where $C_{sol}$ is just the solute concentration, and 
+$C_{eq}(T)$ is the equilibrium concentration at temperature 
+$T$. While the solution remains supersaturated there 
+is a chemical potential driving the solute to coalesce 
+and separate from the solution as an ordered solid, the 
+first formation of the crystal is referred to as the 
+nucleus and understanding its formation has been the 
+focus of researchers for decades now. Typically, for an 
+industrial crystallisation process the working principle 
+is based on controlling and manipulating the supersaturation 
+of the system.
 
 \subsection{Primary \& Secondary nucleation}
-From an industrial perspective, the nucleation process can be broadly 
-categorised into either primary of secondary nucleation. The former 
-describes the formation of an initial nucleus within the bulk phase, 
-absent of any external stimuli. Primary nucleation is therefore considered 
-stochastic as there is no reliable means of predicting where a nucleus
-may form, or how long it will take. The only reliable indicator being 
-that higher supersaturations will result in faster nucleation rates. At 
-a small scale one can estimate the nucleation rate by making repeated 
-measurements of sample solutions and seeing how many have nucleated 
-after a given time, giving us a Poisson probability distribution.
+From an industrial perspective, the nucleation process can 
+be broadly categorised into either primary of secondary
+nucleation. The former describes the formation of an initial
+nucleus within the bulk phase, absent of any external stimuli.
+Primary nucleation is therefore considered stochastic as there 
+is no reliable means of predicting where a nucleus may form, 
+or how long it will take. The only reliable indicator being 
+that higher supersaturations will result in faster nucleation
+rates. At a small scale one can estimate the nucleation rate 
+by making repeated measurements of sample solutions and seeing
+how many have nucleated after a given time, giving us a Poisson
+probability distribution.
 \begin{align}
@@
 
-Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$ is 
-the volume of the individual samples, and $M^*(t)$ \& $M$ are the number 
-of nucleated samples and the total number of samples used respectively.
-While this is useful for studying the effects of different parameters at 
-a small scale, for industrial applications there are to many external 
-factors for primary nucleation to be the only
+Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
+$V$ is the volume of the individual samples, and $M^*(t)$ \& 
+$M$ are the number of nucleated samples and the total number 
+of samples used respectively. While this is useful for studying
+the effects of different parameters at a small scale, for
+industrial applications there are to many external factors 
+for primary nucleation to be measured accurately.
 
-In contrast, secondary nucleation is the result of a initial 
-seed crystal inducing further nucleation within the bulk solution
-\cite{Botsaris1976}. The research of Secondary nucleation is 
-fascinating for its own merits; is it speculated that the 
-supersaturation barrier is non-existent \cite{Cashmore2022}, 
-Furthermore control of secondary nucleation events are crucial 
-for ensuring reliable industrial crystalliser performance, 
-fig.~\ref{fig:secondary} depicts an attempt of classifying  
-every possible mechanism that could lead to secondary nucleation.
+Secondary nucleation is the result of a initial seed crystal 
+inducing further nucleation within the bulk solution 
+\cite{Botsaris1976}. The seed crystals is often specially 
+prepared ahead of time and added to a supersaturate solution. 
+Due to interactions with the surrounding fluid and the container 
+walls it is possible for the seed crystals to act as a surface
+for further nucleation \cite{Anwar2015}. Control of secondary 
+nucleation events are crucial for ensuring reliable industrial 
+crystalliser performance. Fig.~\ref{fig:secondary} depicts an 
+attempt of classifying every possible mechanism that could lead 
+to secondary nucleation.
 \begin{figure}[h!]
@@
 	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
-	\caption{Secondary Nucleation mechanisms, classified by Agrawal 
-		and Paterson \cite{Agrawal2015}}
+	\caption{Secondary Nucleation mechanisms, classified by
+		Agrawal and Paterson \cite{Agrawal2015}}
 	\label{fig:secondary}
@@
 
-Its clear that preventing secondary nucleation is impossible
-with so many possible mechanism, modelling its contribution
-to the nucleation rate has been the area of research for some
-time now \cite{Cashmore2022, Flannigan2023}. In most cases 
-secondary nucleation is actually desirable for large scale 
-production as primary nucleation is a stochastic process.
+This is not a universal classification system, there are a 
+variety of opinions on how best to characterise different 
+phenomena. It is heavily dependent on the theory used to 
+describe nucleation events, this is accurate for both 
+secondary and primary nucleation.
 
 \section{Nucleation Theories}
-A number of potential theories have been proposed to explain 
+Several potential theories have been proposed to explain 
 how nucleation occurs at a microscopic scale. In doing so 
 we could potentially predict the expected yield of a given
-crystalliser based on the initial conditions of the 
-solution. Outlined below are some of the most popular theories
-currently used to describe nucleation.
+crystalliser based on the initial conditions of the solution. 
+Outlined below are some of the most popular theories currently 
+used to describe nucleation.
 
 \subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibs Nucleation Theory' the original theory 
-was first formed from the works of Volmer and Weber, and Frenkel 
-\cite{Frenkel1939, Volmer1926}. While initially it was more focused 
-into describing droplet formation in condensing vapours it was 
-extrapolated to describe crystallisation. The central premise of 
-classical theory is that nucleation occurs stochastically due to 
-collisions between individual solute molecules, ions, or atoms. At the 
-same time the bulk phase is resistant to the formation of a new phase. 
-The competition between these random collisions and the bulk solution 
-can be used to predict the probability of a newly formed nucleus.
+Sometimes referred to as 'Gibbs Nucleation Theory' the 
+original theory was first formed from the works of Volmer 
+and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
+While initially it was more focused on describing droplet
+formation in condensing vapours it was extrapolated to 
+describe crystallisation. The central premise of classical 
+theory is that nucleation occurs stochastically due to 
+collisions between individual solute molecules, ions, or 
+atoms. At the same time the bulk phase is resistant to the
+formation of a new phase. The competition between these 
+random collisions and the bulk solution can be used to 
+predict the probability of a newly formed nucleus.
 
-Consider a supersaturated solution, after some time enough individual
-sub units collide, forming a nucleus of volume $4\pi r^3/3$. The newly 
-formed phase has a lower chemical potential than the surrounding solution, 
-reducing the free energy of the system. Simultaneously, the formation of 
-a new interface is resited by the bulk phase due to surface tension. 
-The net free energy of the system for a nucleus of radius $r$ is given 
-as \cite{Karthika2016}:
+Consider a supersaturated solution, after some time 
+enough individual sub units collide, forming a nucleus 
+of volume $4\pi r^3/3$. The newly formed phase has a 
+lower chemical potential than the surrounding solution, 
+reducing the free energy of the system. Simultaneously, 
+the formation of a new interface is resited by the bulk 
+phase due to surface tension. The net free energy of the 
+system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
 \begin{align}
@@
 
-Where $v$ is the approximate volume of an individual molecule, $k_B$ is 
-the Boltzmann constant, and $\sigma_{inf}$ is the interfacial tension of 
-the bulk solution. This assumes that the nucleus will have a spherical 
-morphology so that the surface tension $\sigma_{inf}$ is a scalar value. 
-Looking at \eqref{eq:CNT} suggests that there must be some critical size 
-$r$ where the free energy gain from the nucleus exceeds the surface tension 
-of the surrounding fluid. Plotting the free energy of the system against 
-nucleus size reveals a critical size above which the gain in free energy 
-exceeds the interfacial tension. Furthermore it shows how increasing the 
+Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
+$v$ is the approximate volume of an individual molecule, 
+$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
+interfacial tension of the bulk solution. This assumes that 
+the nucleus will have a spherical morphology so that the 
+surface tension $\sigma_{inf}$ is a scalar value. Looking at 
+\eqref{eq:CNT} suggests that there must be some critical size 
+$r$ where the free energy gain from the nucleus exceeds the 
+surface tension of the surrounding fluid. This is reflected in 
+fig.~\ref{fig:free_energy} where we plot the free energy of the 
+system against nucleus size. This reveals a critical size above 
+which the gain in free energy exceeds the interfacial tension. 
+Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
 supersaturation of the system reduces said barrier. 
@@
 
-The maximum value of $\Delta G_{tot}$ is the free energy barrier that any 
-newly formed nucleus needs to overcome in order to stabilise. The nucleation 
-rate (the volume of new crystalline material formed per unit time), is 
-therefore commonly defined as being dependent on the energy barrier $\Delta G^*$:
+The maximum value of $\Delta G_{tot}$ is the free energy barrier 
+that any newly formed nucleus needs to overcome in order to stabilise. 
+The nucleation rate (the volume of new crystalline material formed per 
+unit time), is therefore commonly defined as being dependent on the 
+energy barrier $\Delta G^*$:
 \begin{align}
@@
 
-Where $A$ is a pre-factor that can be fine tuned to the exact demands
-of the system (typically involving Zeldovich Factor, $Z$) the free 
-energy barrier can be found by finding the turning point of $\Delta G_{tot}$.
+Where $A$ is a pre-factor that can be fine tuned to the exact 
+demands of the system, the free energy barrier can be found by finding 
+the stationary point of $\Delta G_{tot}$.
 
-CNT is often regarded as a good description of the macro system, its obvious
-that for all crystallization systems there is an inherent energy barrier that 
-dictates the nucleation rate. Where it falters is in its predictive ability, 
-both in estimating nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in 
-the structure of newly formed nuclei \cite{Lee1999, Yau2001}. Recent studies 
-suggest classical nucleation is merely one of many possible pathways that can 
-be taken to produce a structured crystalline phase. Prompting the development of alternative theories to better describe the nucleation process.
+CNT is often regarded as a good description of the macro system, its 
+obvious that for all crystallization systems there is an inherent energy 
+barrier that dictates the nucleation rate. Where it falters is in its 
+predictive ability, both in estimating nucleation rates \cite{Gharibeh2005, 
+Vekilov2010}, and in the structure of newly formed nuclei \cite{Lee1999, 
+Yau2001}. Recent studies suggest classical nucleation is merely one of many 
+possible pathways that can be taken to produce a structured crystalline phase. 
+Prompting the development of alternative theories to better describe the 
+nucleation process.
 
@@
 The two step nucleation theory is an extension to the CNT that suggests 
-that prior to nucleation, the solute will cluster together. This was 
-shown via colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}.
-Several papers later reported the
-presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}. The formation of these clusters 
-can be understood by Oswald's rule; which says that any crystallising 
-system does not immediately take the path to the lowest possible energy 
-state but instead first transitions to the state with the smallest free 
-energy barrier \cite{Ostwald1897}. Further phase transitions can still 
-occur but this pathway minimises the overall free energy cost. Initially 
-it was suspected that the formation of the dense liquid phase was a 
-result of stochastic density fluctuations in the system. However, 
-simulations suggest that the the local bond order is a stronger driving 
-force than the local density \cite{Tan2013}. Studies into two step 
+that prior to nucleation, the solute will cluster together. This has 
+been shown via colloid simulations using protein molecules that have 
+significant short range interactions \cite{Wolde1997, Gliko2005}. These 
+interactions allow for the formation of a liquid-liquid metastable phase 
+from which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
+Several papers later reported the presence of stable liquid-like clusters 
+that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}. 
+The formation of these clusters can be understood by Oswald's rule; which 
+says that any crystallising system does not immediately take the path to 
+the lowest possible energy state but instead first transitions to the state 
+with the smallest free energy barrier \cite{Ostwald1897}. Further phase 
+transitions can still occur but this pathway minimises the overall free 
+energy cost. Initially it was suspected that the formation of the dense 
+liquid phase was a result of stochastic density fluctuations in the system. 
+However, simulations suggest that the the local bond order is a stronger 
+driving force than the local density \cite{Tan2013}. Studies into two step 
 nucleation in Glycine solutions have noted that the solutions need to be 
-aged before the presence of clusters was detected. This suggests that 
-there exist further barriers to the formation of clusters in some systems,
-this has lead to instead calling this phenomena as Multi-step Nucleation.
+aged before the presence of clusters was detected \cite{Liao2022}. This 
+suggests that there exist further barriers to the formation of clusters 
+in some systems, this has lead to instead calling this phenomena as 
+Multi-step Nucleation.
 
@@
 initial conditions. Reviews of all currently documented nucleation 
-pathways identified highlighted the fact the need for a the 
-development of \textit{in situ} techniques that can induce nucleation 
-locally but can also reliably identify the nucleation pathway across 
-a broad range of experimental conditions \cite{Karthika2016, Fu2021}.
+pathways highlighted the need for the development of \textit{in situ} 
+techniques that can induce nucleation locally but can also reliably 
+identify the nucleation pathway across a broad range of experimental 
+conditions \cite{Karthika2016, Fu2021}.
 
@@
 \subsection{Cooling Crystallisation}
-For some binary mixtures the supersaturation is heavily dependent on 
-the solution temperature, therefore a simple method of producing crystals 
-is by cooling the mixture to induce crystal formation. At ambient 
-temperatures the solute concentration is too high to be fully incorporated
-into the solution ($S\gg 1$), after heating however the solute is fully 
-dissolved into the solution ($S<1$). Now as the mixture is allowed to cool 
-to room temperature the supersaturation will increase until crystal formation begins, the rate of cooling drastically influencing the size and number of crystals produced.
+For some binary mixtures the supersaturation is heavily dependent 
+on the solution temperature, therefore a simple method of producing 
+crystals is by cooling the mixture to induce crystal formation. At 
+ambient temperatures the solute concentration is too high to be 
+fully incorporated into the solution ($S\gg 1$), after heating however 
+the solute is fully dissolved into the solution ($S<1$). Now as the 
+mixture is allowed to cool to room temperature the supersaturation 
+will increase until crystal formation begins, the rate of cooling 
+drastically influences the size and number of crystals produced.
 
 If $dT/dt$ is high then the final product will consist of large 
-crystal and be low in number, as the nucleation rate is directly related
-to the supersaturation only a handful of nuclei can form before the remaining
-solute grows onto the surface. If $dT/dt$ is low then the final product will 
-consist of smaller crystals and be far more numerous, as the supersaturation 
-is so large that the new nuclei are forming continuously. Between these two 
-extremes, one can define the meta-stable zone width, a region in the 
-temperature-concentration phase space where both nucleation and crystal growth
-can be reliably controlled. The lower limit being given by the solubility curve
-of the binary mixture, and the upper limit being defined by the metastable zone.
+crystal and be low in number, as the nucleation rate is directly 
+related to the supersaturation only a handful of nuclei can form 
+before the remaining solute grows onto the surface. If $dT/dt$ is 
+low then the final product will consist of smaller crystals and be 
+far more numerous, as the supersaturation is so large that the new 
+nuclei are forming continuously. Between these two extremes, one 
+can define a meta-stable zone, a region in the temperature-
+concentration phase space where both nucleation and crystal growth
+can be reliably controlled. The lower limit being given by the 
+solubility curve of the binary mixture, and the upper limit being 
+defined by the point where crystals begin to spontaneously nucleate
+within the solution.
 
@@
 	\includegraphics[width=\linewidth]{MSZW.png}
-	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red).
-3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
+	\caption{Typical Concentration vs Temperature plot with the 
+		solubility curve (black) and metastable zone (red). 3 
+		different cooling curves are shown as well: a high rate 
+		of cooling (orange) shows the mixture quickly dropping 
+		below the solubility curve resulting in no new nuclei 
+		forming. A low rate of cooling (green) shows the mixture 
+		exceeding the metastable zone, where now nucleation occurs 
+		spontaneously and cannot be controlled without further 
+		cooling. A typical cooling rate (blue) shows how a 
+		typical cooling crystalliser will operate, sitting in the 
+		middle of the other two curves.}
 	\label{fig:MSZW}
@@
 
-The viability of cooling crystallisation is dependent on the meta-stable 
-zone width, too narrow and the process is difficult to control, too wide 
-and the crystal growth rate may be insufficient for the desired outcome.
+The viability of cooling crystallisation is dependent on the 
+meta-stable zone width, too narrow and the process is difficult to 
+control, too wide and the crystal growth rate may be insufficient 
+for the desired outcome.
 
 \subsection{Evaporative Crystallisation}
-In situations where control of the final product size or shape is not the main
-focus, often the cheapest method of producing a crystalline product is simply
-to allow the solvent evaporate and separate from the solute. Depending on the 
-total volume of solvent to evaporate this process can take on time scale of 
-several days to complete.
+In situations where control of the final product size or shape 
+is not the main focus, often the cheapest method of producing a 
+crystalline product is simply to allow the solvent evaporate and 
+separate from the solute. Depending on the total volume of solvent 
+to evaporate this process can take on time scale of several days 
+to complete. This is often used for the formation of seed crystals 
+when larger seed crystals are required \cite{Srinivasan2008, Li1992}.
 
 \subsection{Anti-solvent Crystallisation} 
-Anti-solvent crystallisation is a more involved separation method ,
-by adding a new solvent that is miscible with the old solvent but is 
-immiscible with the solute, the solute's effective solubility is decreased.
-As such, the solution becomes supersaturated. This has the trade off in that the overall mass ratio between the solvent and solute is now lower. Careful
-measurement of the anti-solvent is required in order to ensure that the 
-decrease in solubility is not out matched by the decrease in mass ratio.
-Anti-solvent crystallisation is not as common due to the fact that not all
-solutions will have an ideal anti-solvent candidate.
+Anti-solvent crystallisation is a more involved separation method,
+by adding a new solvent that is miscible with the old solvent but 
+is immiscible with the solute, the solute's effective solubility 
+is decreased. As such, the solution becomes supersaturated. This 
+has the trade off in that the overall mass ratio between the solvent 
+and solute is now lower. Careful measurement of the anti-solvent is 
+required in order to ensure that the decrease in solubility is not 
+out matched by the decrease in mass ratio.Anti-solvent crystallisation 
+is not as common due to the fact that not all solutions will have an 
+ideal anti-solvent candidate.
 
@@
 \subsection{Background}
-Optical tweezing has been a field of applied optics ever since the 1970s
-when Ashkin \cite{Ashkin1970} first showed that focused light was capable 
-of trapping micron sized particles due to light exerting 'radiation pressure'. 
-The working principle was that a light source such as a laser could trap 
-small objects within a 2D plane, as long as the light source had an 
-approximately Gaussian profile. Soon after, Ashkin showed that the introduction 
-of a microscope objective would allow one to focus the light source to a 
-diffraction limited point that would stably trap small objects within a 
-confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study 
-biological material and would later be used to probe microscopic properties 
-such as the formation of colloidal aggregates \cite{Yi2021} to the drag 
-forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the 
-predictable behaviour of light, optical tweezers have become essential for 
+Optical tweezing has been a field of applied optics ever since 
+the 1980s when Ashkin \cite{Ashkin1970} first showed that focused 
+light was capable of trapping micron sized particles due to light 
+exerting 'radiation pressure'. The working principle was that a 
+light source such as a laser could trap small objects within a 2D 
+plane, as long as the light source had an approximately Gaussian 
+profile (colloquially called a Gaussian beam). Soon after, Ashkin 
+showed that the introduction of a microscope objective would allow 
+one to focus the light source to a diffraction limited point that 
+would stably trap small objects within a confined volume 
+\cite{Ashkin1980}. This allowed Ashkin and others to study individual 
+biological cells as the overall force was on the order of $10^{-12}\ 
+N$ while being non-invasive to the internal structure. Later it would 
+be used to probe microscopic properties such as the formation of 
+colloidal aggregates \cite{Yi2021} to the drag forces exerted by a 
+pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
+behaviour of light, optical tweezers have become essential for 
 measuring and exerting precise forces on the magnitude of pico-newtons 
@@
 
-\subsection{Gaussian Beams}
-A Gaussian Beam is a simplistic approximation of a laser beam, in 
-short it is an unbounded plane wave whose intensity of the 
-electromagnetic field falls off from the centre similar to a normal distribution. This is perfectly fine for beams with a divergence 
-angle of $0^{\circ}$, but later studies into focused beams showed 
-that the assumption of a Gaussian intensity profile is inaccurate. 
-A study conducted by Lock and Gouesbet found that as the Gaussian 
-beam was focused the fall off in the electric field would 
-consequentially induce additional electromagnetic fields 
-\cite{Lock1994}. As a result the shape of the beam profile diverges 
-from a Gaussian profile for focused beams. Regardless, for optical 
-trapping experiments it is useful to assume a near Gaussian beam as 
-many descriptions of lasers are simply combinations of plane waves 
-(with varying amplitudes and wave vectors).  More information on 
-electromagnetic theory is covered in chapter 2 but for now any 
-mention of a focused laser can be thought of as a Gaussian beam, 
-unless specified otherwise.
-
 \subsection{Literature related to laser induced nucleation}
-From as early as 1996 it has been known that laser irradiation is a viable
-method of inducing nucleation within a supersaturated solution
-\cite{Garetz1996}. The first reported case was notable as it used a 
-1.064 $\mu m$ laser, the glycine solutions would appear transparent to
-such a laser which would suggest there was no photo-chemical reaction.
-Later studies into this phenomena found that the laser polarisation 
-can influence the polymorph produced. With circularly polarised light 
-producing $\alpha$-glycine and linearly polarised light forming
-$\gamma$-glycine \cite{Garetz2002}. Future research has found nucleation 
-can be induced by 1 of 3 routes.
+From as early as 1996 it has been known that laser irradiation 
+using a Gaussian beam is a viable method of inducing nucleation 
+within a supersaturated solution \cite{Garetz1996}. The first 
+reported case was notable as it used a 1.064 $\mu m$ laser, the 
+glycine solutions would appear transparent to such a laser which 
+would suggest there was no photo-chemical reaction. Later studies 
+into this phenomena found that the laser polarisation can influence 
+the polymorph produced. With circularly polarised light producing 
+$\alpha$-glycine and linearly polarised light forming $\gamma$-glycine 
+\cite{Garetz2002}. Future research has found nucleation can be 
+induced by 1 of 3 routes.
 
 \subsubsection{Non-Photochemical Laser Induced Nucleation}
-Non-photochemical laser induced nucleation (NPLIN) involves irradiating 
-a solution with a pulsed laser \cite{Garetz1996,Garetz2002,Sun2006}. 
-The laser itself does not have to be heavily focused, instead irradiating
-a large region of the solution all at once. The choice of laser is of
-particular importance; with nucleation probability changing depending 
-on the wavelength \cite{Kacker2017}. In addition, the choice of solute will 
-effect the setup, not only because some solute's are unaffected, but also 
-because there is a minimum laser threshold before nucleation is observed 
-\cite{Garetz2002}. Several papers have debated the exact mechanism that 
-induces NPLIN \cite{Garetz2002, Knott2011}. A suggested theory to this is 
-an optical Kerr effect, for anisotropically charged solute molecules the 
-electric field can reorient them to match the propagation direction 
-\cite{Garetz2002}. If enough molecules are co-aligned the free energy 
-barrier is reduced to allow for ambient nucleation \cite{Knott2011}. 
-An alternative theory is the dielectric polarisation effect, in 
-conditions that are unfavourable to cluster formation the polarising 
-effect can stabilise the clusters \cite{Alexander2008}. As the cluster 
-concentration rises so does the likelihood of nucleation \cite{Vekilov2010}. 
-Both theories are similar to one another but where the optical Kerr theory
-is limited to anisotropic solute molecules, the direct polarisation 
-theory is more flexible. Regardless both theories struggle to explain
-why the phenomena is not observed in all nucleation systems 
-\cite{Korede2023}, such as acetamide which is similar to urea which 
-does nucleate when irradiated \cite{Ward2016}. One of the benefits of 
-NPLIN is that since the pulses are relatively low in their intensity 
-they can be fired off quickly in succession, allowing for continuos 
-crystallisation set ups. Overall, the NPLIN phenomena needs further 
-research to properly describe its effects. The mean pulse intensity 
-needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), 
-as high intensity pulses lead to a completely different nucleation 
+Non-photochemical laser induced nucleation (NPLIN) involves 
+irradiating a solution with a pulsed laser \cite{Garetz1996,
+Garetz2002,Sun2006}. The laser itself does not have to be 
+heavily focused, instead irradiating a large region of the 
+solution all at once. The choice of laser is of particular 
+importance; with nucleation probability changing depending 
+on the wavelength. A study of KCl solutions found that for 
+lower intensities it was found that nucleation was favoured 
+for lower wavelengths but above a peak intensity of $5 MW/
+cm^2$ the wavelength independence disappeared \cite{Kacker2017}. 
+Measurements of the intensity prior and after irradiation 
+confirmed this wavelength dependence was not due to any 
+photo-chemical interactions \cite{Kacker2017}.
+
+Additionally, the choice of solute will effect the setup, not 
+only because some solute's are unaffected, but also because 
+there is a minimum laser threshold before nucleation is observed 
+\cite{Garetz2002}. Several papers have debated the exact mechanism 
+that induces NPLIN \cite{Garetz2002, Knott2011}. A suggested 
+theory to this is an optical Kerr effect: For anisotropically 
+charged solute molecules the electric field can reorient them 
+to match the propagation direction \cite{Garetz2002}. If enough 
+molecules are co-aligned the free energy barrier is reduced to 
+allow for ambient nucleation \cite{Knott2011}. An alternative 
+theory is the dielectric polarisation effect, in conditions that 
+are unfavourable to cluster formation the polarising effect can 
+stabilise the clusters \cite{Alexander2008}. As the cluster 
+concentration rises so does the likelihood of nucleation 
+\cite{Vekilov2010}.
+
+Both theories are similar to one another but where the optical 
+Kerr theory is limited to anisotropic solute molecules, the 
+direct polarisation theory is more flexible. Regardless both 
+theories struggle to explain why the phenomena is not observed 
+in all nucleation systems \cite{Korede2023}, such as acetamide 
+which is similar to urea which does nucleate when irradiated 
+\cite{Ward2016}. One of the benefits of NPLIN is that since the 
+pulses are relatively low in their intensity they can be fired 
+off quickly in succession, allowing for continuos crystallisation 
+set ups. Overall, the NPLIN phenomena needs further research to 
+properly describe its effects. The mean pulse intensity needs to 
+be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as 
+high intensity pulses lead to a completely different nucleation 
 mechanism.
@@
 \subsubsection{High Intensity Laser Induced Nucleation}
-High intensity laser induced nucleation (HILIN), where the pulse intensity 
-is on the order of several $PW/cm^2$ is far simpler a mechanism to explain 
-in comparison to NPLIN. The production of nuclei can be wholly associated 
-to a cavitation process within the target solution, where the laser focus 
-results in thermo-cavitation and the subsequent pressure wave leads to
-a nucleation event around the focus of the laser \cite{Yoshikawa2005, 
-Soare2011, Barber2019}. What remains in question is both how the physical 
-properties (size, polymorph, etc) are influenced by the cavitation process, 
-and how the pressure change triggers nucleation. The former has already 
-been investigated; by adjusting the focal position Ikeda \textit{et al} 
-could control the polymorph of indomethacin \cite{Ikeda2015}, this is not 
-a universal method however, as it has also been shown that laser power can 
-influence the crystal polymorph \cite{Wang2010}. The latter is a tricky 
-task to address due to the fact that these cavitation bubbles form and 
-collapse in less than $100\ \mu s$. Using fluorescence dyed proteins, 
-researchers were able to observe a sudden spike in fluorescence just 
-as the cavitation bubble began to collapse, they suggested that due 
-to the collapse of the cavitation bubble the protein clusters are 
-brought together at the lasers focal point. However, while the fluorescence 
-imaging indicates a local concentration increase it is difficult to 
-quantify this change depending on the size of the bubble \cite{Korede2023}. 
-It has been suggested that in theory any solution can undergo HILIN 
+High intensity laser induced nucleation (HILIN), where the pulse 
+intensity is on the order of several $PW/cm^2$ is far simpler a 
+mechanism to explain in comparison to NPLIN. The production of 
+nuclei can be wholly associated to a cavitation process within 
+the target solution, where the laser focus results in thermo-
+cavitation and the subsequent pressure wave leads to a nucleation 
+event around the focus of the laser \cite{Yoshikawa2005, Soare2011, 
+Barber2019}.
+
+What remains in question is both how the physical properties (size, 
+polymorph, etc) are influenced by the cavitation process, and how 
+the pressure change triggers nucleation. The former has already 
+been investigated; by adjusting the focal position Ikeda \textit{et 
+al} could control the polymorph of indomethacin \cite{Ikeda2015}, 
+this is not a universal method however, as it has also been shown 
+that laser power can influence the crystal polymorph \cite{Wang2010}. 
+The latter is a tricky task to address due to the fact that these 
+cavitation bubbles form and collapse in less than $100\ \mu s$. 
+Using fluorescence dyed proteins, researchers were able to observe 
+a sudden spike in fluorescence just as the cavitation bubble began 
+to collapse, they suggested that due to the collapse of the cavitation 
+bubble the protein clusters are brought together at the lasers focal 
+point. However, while the fluorescence imaging indicates a local 
+concentration increase it is difficult to quantify this change 
+depending on the size of the bubble \cite{Korede2023}. It has been 
+suggested that in theory any solution can undergo HILIN 
 \cite{Korede2023}, but proving such a theory requires a clear 
@@
 Current research aims to combine experimental research with computer 
-simulations to develop a universal theory, with the hope that this could 
-also be related to NPLIN.
+simulations to develop a universal theory, with the hope that this 
+could also be related to NPLIN.
 
@@
 for \textit{in-situ} studying of nucleation events. Using high 
-numerical aperture lenses one can localise the nucleation event 
+numerical aperture lens one can localise the nucleation event 
 to a specific region of the solution. The current issue is that 
@@
 effective to manipulate the solution using trapped particles. 
-An common method is by apply a torque to a trapped particle, 
-creating a micro-rotor that can can generate fluid flow.
+One way would to generate an optical torque on a trapped particle
+and therefore shear the surrounding fluid, a method that is 
+already in common use for micro-rheological studies \cite{Bishop2004, 
+RobertsonAnderson2018}
 
@@
 \label{sec:opt_torque}
-It has long been known that electromagnetic fields can transfer linear and
-angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
-is said to have both orbital and spin momentum. Though there is some 
-debate on how to decompose the total momentum into these two components 
-\cite{Bruce2020, Svak2018}, for this project we do not need to calculate 
-the exact quantities and will instead look at the broader effects of both 
-components. Orbital angular momentum arises from the shape of the wavefront 
-of the particular field in question; for simple Gaussian beams the wavefronts 
-are uniform and equally spaced resulting in the typical radiation pressure 
-that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
-modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
-non-uniform wave fronts meaning the orbital momentum has both angular and
-linear components; depending on the relative size of the target particle 
-one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.
+It has long been known that electromagnetic fields can transfer
+linear and angular momentum \cite{Beth1936MechanicalDA}; more 
+accurately the field is said to have both orbital and spin momentum. 
+Though there is some debate on how to decompose the total momentum 
+into these two components \cite{Bruce2020, Svak2018}, for this 
+project we do not need to calculate the exact quantities and will 
+instead look at the broader effects of both components. Orbital 
+angular momentum arises from the shape of the wavefront of the 
+particular field in question; for simple Gaussian beams the wavefronts 
+are uniform and equally spaced resulting in the typical radiation 
+pressure that Ashkin and co demonstrated \cite{Ashkin1980}. However, 
+higher order modes of a Gaussian beam (for example: Laguerre-Gaussian 
+modes) have non-uniform wave fronts meaning the orbital momentum has 
+both angular and linear components; depending on the relative size 
+of the target particle one can induce rotation, or orbiting 
+\cite{Bruce2020, Courtial2000}.
 
-Spin angular momentum (SAM) is attributed to the spin density of the field, 
-early research has shown that the spin density is non-zero for any beam despite
-the fact that the total SAM transferred to a medium is 0 \cite{Svak2018, 
-Bliokh2014}. This has sparked debate if SAM is even a physical quantity as 
-it does not aid in the transport of energy directly \cite{Bliokh2014} and 
-so cannot be directly observed in some cases despite being non-zero. This 
-paradox is resolved by representing the wave as an array of spin momentum 
-loops that all together cancel one-another out when the medium is homogeneous.
-Spacial inhomogeneities cause these spin loops to no longer be equal, resulting
-in non-zero spin density, anisotropic mediums (such as birefringent crystal 
-lattices) experience a transfer of spin angular momentum, imparting an optical
-torque.
+Spin angular momentum (SAM) is attributed to the spin density of 
+the field, early research has shown that the spin density is non-
+zero for any beam despite the fact that the total SAM transferred 
+to a medium is 0 \cite{Svak2018, Bliokh2014}. This has sparked 
+debate if SAM is even a physical quantity as it does not aid in 
+the transport of energy directly \cite{Bliokh2014} and so cannot 
+be directly observed in some cases despite being non-zero. This 
+paradox is resolved by representing the wave as an array of spin 
+momentum loops that cancel one-another out when the medium is 
+homogeneous.Spacial inhomogeneities cause these spin loops to no 
+longer be equal, resulting in non-zero spin density, anisotropic 
+mediums (such as birefringent crystal lattices) experience a 
+transfer of spin angular momentum, imparting an optical torque.
 
-Birefringence is a material property often seen in crystalline materials, 
-if the crystal lattice has different refractive indices for varying 
-polarisations of light. Because of this the material has two different 
-refractive indices depending on the polarisation of light. For circularly 
-polarised light this inhomogeneity results in a high degree of SAM being 
-transferred to the target object \cite{Parkin2009, Arita2016}. The greater 
-the difference between the two refractive indces the greater the angular
-momentum transfer.
+Birefringence is a material property often seen in crystalline 
+materials, where the crystal lattice has a different structure 
+dependent on its orientation. Since light is composed of waves 
+that propagate in orthogonally to one another, a birefringent 
+material will refract light differently depending on the light's 
+polarisation. Therefore it can be said that the material has 
+two separate refractive indices. For circularly polarised light 
+this inhomogeneity results in a high degree of SAM being 
+transferred to the target object \cite{Parkin2009, Arita2016}. 
+The greater the difference between the two refractive indices the 
+greater the angular momentum transfer.
 
-The ability to transfer angular momentum has been exploited to rotate 
-microspheres as fast as 1000 Hz while suspended in a bulk medium 
-\cite{Arita2016} as well as a means of measuring the local temperature 
-and shear response of said medium \cite{Millen2014, RodriguezSevilla2018}. 
-Calculating the optical torque applied to a birefringent material is given via:
+The ability to transfer angular momentum has been exploited 
+to rotate microspheres as fast as 1000 Hz while suspended 
+in a bulk medium \cite{Arita2016} as well as a means of 
+measuring the local temperature and shear response of said 
+medium \cite{Millen2014, RodriguezSevilla2018}. Calculating 
+the optical torque applied to a birefringent material is given 
+via:
 \begin{equation}
@@
 
-Where $\Delta n$ is the difference between the two refractive indices, 
-$\theta$ is the angle between the particle's long axis and the polarisation 
-vector of the local EM field, and $\phi$ is the phase shift in the EM 
-field. The first term represents the 'orientational' torque which is 
-due to the target particle being aligned with the EM amplitude, when 
-aligned $\theta=0$ meaning the entire term is negligible for particle's 
-with a stable orientation. For sphere's this is always the case as they 
-lack a distinctive major axis. The second term is due purely to the 
-polarisation of the optical trap, for circularly polarised light 
-$\phi=\pi/4$ thus maximising the torque transferred to the target particle. 
-Eq.~\eqref{eq:opt_torque} is only applicable for particles with a known
-birefringence, but there are other mechanisms that result in optical torque.
+Where $\Delta n$ is the difference between the two refractive 
+indices, $\theta$ is the angle between the particle's long 
+axis and the polarisation vector of the local EM field, and 
+$\phi$ is the phase shift in the EM field. The first term 
+represents the 'orientational' torque which aligns the long 
+axis of the particle with the electric field, when aligned 
+$\theta=0$ meaning the entire term is negligible for particle's 
+with a stable orientation. The second term is due purely to the 
+polarisation of the laser, for circularly polarised light 
+$\phi=\pi/4$ thus maximising the torque transferred to the 
+target particle. Eq.~\eqref{eq:opt_torque} is only applicable 
+for particles with a known birefringence, but there are other 
+mechanisms that result in optical torque.
 
-A common example is shape induced birefringence. If a particle has an 
-anisotropic shape, it is more susceptible to being polarised along its 
-longer axis than its shorter axis. Consider, for example, an ellipsoid 
-elongated along one of its primary axis' ($r_z > r_x = r_y$). In a 
-plane polarised beam such a particle will align with the polarisation
-vector. Therefore, the particle will rotate as angular momentum is 
-transferred along its long axis. The common feature for shape induced
-birefringence is that all of these particles are aligned in the plane
-of polarisation. Often the optical torque experienced is far greater 
-than similar spherical particles that are birefringent. Currently 
-spherical dimers are being rotated in vacuums to measure quantum forces
-and torques \cite{Ahn2018, Reimann2018}. There are some cases alternative 
-cases where particles are rotated while not being aligned in the plane 
-of the polarisation. However in these cases their shape is often 
-specifically engineered to scatter light either clockwise or 
-counter-clockwise \cite{Higurashi1994}.
+A common example is shape induced birefringence. If a particle 
+has an anisotropic shape, it is more susceptible to being 
+polarised along its longer axis than its shorter axis. Consider, 
+for example, an ellipsoid elongated along one of its primary 
+axis' ($r_z > r_x = r_y$). In a plane polarised beam such a 
+particle will align with the polarisation vector. Therefore, 
+the particle will rotate as angular momentum is transferred 
+along its long axis. One common feature, regardless of shape,
+is that a particle with shape birefringence will rotate when 
+it lies perpendicular to the direction of propagation. This 
+is seen most evidently with spherical dimers \cite{Ahn2018, 
+Reimann2018} but even for elliptical particles rotational 
+motion is only detected when their long axis is not aligned 
+with the direction of propagation \cite{Zhu2021, Mihiretie2014}.
 
-Other examples of optical torque is when an anisotropic particle is 
-aligned with the beam's direction of propagation (in which case $\theta
-=\pi/2$ and the first term disappears). This is analogous to an optically 
-trapped sphere, where alongside a restoring force the particle also 
-experiences a restoring torque. This seemingly random rotational motion
-is referred to as libation \cite{Bruce2020}, often in typical suspension
-trapping situations (where the particle is suspended in a fluid) the 
-translational motion is washed out by the translational motion. As such, 
-many experiments elect to trap in low pressure environments to precisely 
-measure the optical torque being exerted by the optical trap \cite{Ahn2018}.
-This has lead to experiments to try and achieve '0 kelvin' motion, where
-by trapping a silica dimer they were able to restrict its motion using 
-3 optical traps simultaneously. Despite this, they found that the dimer's
-rotational motion about its long axis could not be controlled leading to
-the undesired rotational modes \cite{Bang2020}.
+Often the optical torque experienced is far greater than similar 
+spherical particles that are birefringent \cite{Bruce2020}. 
+Currently spherical dimers are being rotated in vacuums to 
+measure quantum forces and torques \cite{Ahn2018, Reimann2018}. 
+There are some alternative cases where particles are rotated 
+while not being aligned in the plane of the polarisation. However 
+in these cases their shape is often specifically engineered to 
+scatter light in such a way that the net momentum change always
+occurs in one direction regardless of the laser polarisation 
+\cite{Higurashi1994}.
 
-The detection and measurement of optical torque is still a field of intense
-research, not only does it have potential to understand quantum fluctuations
-in a particle's motion but also allows for the creation of more effective
-micro-rotors. The latter being especially pertinent for understanding the 
-behaviour of fluids experiencing localised shearing.
+Other examples of optical torque is when an anisotropic 
+particle is aligned with the beam's direction of propagation 
+(in which case $\theta=\pi/2$ and the first term disappears). 
+This is analogous to an optically trapped sphere, where 
+alongside a restoring force the particle also experiences a 
+restoring torque. This seemingly random rotational motion is 
+referred to as libation \cite{Bruce2020},often in typical 
+suspension trapping situations (where the particle is suspended 
+in a fluid) the rotational motion is washed out by the 
+translational motion. As such, many experiments elect to trap 
+in low pressure environments to precisely measure the optical 
+torque being exerted by the optical trap \cite{Ahn2018}. This 
+has lead experiments to try and achieve '0 kelvin' motion, 
+where by trapping a silica dimer they were able to restrict its 
+motion using 3 optical traps simultaneously. Despite this, they 
+found that the dimer's rotational motion about its long axis 
+could not be controlled leading to the undesired rotational
+modes \cite{Bang2020}.
+
+The detection and measurement of optical torque is still a 
+field of intense research, not only does it have potential 
+to understand quantum fluctuations in a particle's motion 
+but also allows for the creation of more effective
+micro-rotors. The latter being especially pertinent for 
+understanding the behaviour of fluids experiencing localised 
+shearing.
 
 \section{Shear induced Nucleation}
-It has long been known that fluid shear rate plays a role in influencing
-nucleation; however, the exact relationship between shear rate and nucleation
-rate has only been recently understood for specific solutions. Theoretical 
-research into shear induced nucleation suggests that there should be a 
-slight increase in the nucleation rate at low shear rates, reaching a 
-maximum increase in nucleation rate, and then at higher shear rates the 
-nucleation rate begins to drop off.
+It has long been known that fluid shear rate plays a role in 
+influencing nucleation; however, the exact relationship between 
+shear rate and nucleation rate has only been recently understood 
+for specific solutions. Theoretical research into shear induced 
+nucleation suggests that there should be a slight increase in the 
+nucleation rate at low shear rates, reaching a maximum increase in 
+nucleation rate, and then at higher shear rates the nucleation rate 
+begins to drop off.
 
-This has been shown theoretically for both simple colloidal \cite{Mura2016,
-Debuysschere2023,Richard2015} and ice crystal formation \cite{Goswami2020}; 
-however, no experimental work into these systems has been conducted to prove 
-this is the case. There is some experimental evidence for this phenomena in 
-simple salt and protein solutions - though the authors emphasise that 
-mechanical agitation cannot be ruled out - there has not been a exhaustive 
-study into the shearing effects apart from in glycine solutions. In 
-\cite{Debuysschere2023} it was found that a shear rate of around $3000\ 
-s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate. 
-Using the theoretical model established in \cite{Mura2016,2001} which modifies 
-the CNT to account for the effects of a nucleus undergoing shearing, 
-accounting for the fact that a nucleus' growth is undergoing competition 
-between flow-mediated molecular transport and the strain applied by the 
-flow field which inhibits the growth of the nucleus. There central 
-conclusion (from both the theoretical and experimental results) is that 
-there is an optimal shear rate in which the nucleation rate is maximised.
+This has been shown theoretically for both simple colloidal 
+\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal 
+formation \cite{Goswami2020}; however, no experimental work 
+into these systems has been conducted to confirm these theories. 
+There is some experimental evidence for this phenomena in simple 
+salt and protein solutions - though the authors emphasise that 
+mechanical agitation cannot be ruled out - there has not been a 
+exhaustive study into the shearing effects apart from in glycine 
+solutions. In \cite{Debuysschere2023} it was found that a shear 
+rate of around $3000\ s^{-1}$ was the maximum shear rate that 
+would yield the highest nucleation rate. Using the theoretical 
+model established in \cite{Mura2016,2001} which modifies the CNT 
+to account for the effects of a nucleus undergoing shearing, 
+accounting for the fact that a nucleus' growth is undergoing 
+competition between flow-mediated molecular transport and the 
+strain applied by the flow field which inhibits the growth of 
+the nucleus. There central conclusion (from both the theoretical 
+and experimental results) is that there is an optimal shear rate 
+in which the nucleation rate is maximised.
 
-However, a question that arises from this result, if there is a optimal 
-shear rate in which molecular transport is maximised and strain is 
-minimised, then surely there should also be a shear rate in which the 
-molecular transport and strain are equal - allowing one to suspend a 
-nucleus at a constant radius. In this scenario, the molecular transport 
-would prevent the nucleus from dissolving, but the strain would prevent 
-the nucleus from growing. This however would require one to be able 
-to apply a continuous shear rate to a targeted nucleus with high precision, 
-there is also no model for an individual nucleus in a continuous fluid 
+However, a question that arises from this result, if there is 
+a optimal shear rate in which molecular transport is maximised 
+and strain is minimised, then surely there should also be a shear 
+rate in which the molecular transport and strain are equal - 
+allowing one to suspend a nucleus at a constant radius. In this 
+scenario, the molecular transport would prevent the nucleus from 
+dissolving, but the strain would prevent the nucleus from growing. 
+This however would require one to be able to apply a continuous 
+shear rate to a targeted nucleus with high precision, there is 
+also no model for an individual nucleus in a continuous fluid 
 field.
@@
 As I have hoped to make clear in the above introduction, the 
-current state nucleation theory is rather cumbersome at a 
+current state of nucleation theory is rather cumbersome at a 
 micro-level. Models such as CNT and multi-step nucleation are 
@@
 beam focus. The results are reported in chapter 3, this is 
-then succeeded by experimental work where we instead using 
-a galvano-mirror to generate shear flow. While overall 
-unsuccessful the addition of a moving beam focus showed 
-that the growth of a nucleus can be localised around the 
-trap focus. This presents a new insights for controlling
-and studying the growth of a newly formed nucleus by 
-precise movement of the trapping focus.
+then followed by experimental work where we use a galvano-
+mirror to move the beam and hence generate shear flow. While 
+overall unsuccessful the addition of a moving beam focus 
+showed that the growth of a nucleus can be localised around 
+the trap focus. This presents a new insights for controlling 
+and studying the growth of a newly formed nucleus by precise 
+movement of the trapping focus.
 

### 2024-11-05 Latex_files/Thesis/chapters/chapter 2/main.tex
praveen commments plus chp 4 rework

@@
 (for a simple Gaussian beam this would be at it's focal point) 
-\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
-higher numerical aperture lenses in order to increase the intensity, 
+\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise a
+higher numerical aperture lens in order to increase the intensity, 
 thus allowing for stronger gradient forces at the trade off of 
@@
 of propagation \cite{Capitanio2002} The equilibrium position is found 
-when the gradient force far exceeds both the the scattering force, 
+when the gradient force far cancels out the the scattering force, 
 and the thermal energy of the particle.
@@
 displacement is called the trap stiffness (denoted by $\kappa$), and 
-is an important characteristic of the trap. Shown below is a typical 
-force displacement curve for a single sphere in a focused Gaussian beam. 
+is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap}
+demonstrates an example case of the force-displacement curve for a 
+simple sphere in an optical trap. Laser power and numerical aperture
+are omitted as the exact shape is not of importance, only the linear 
+region about the origin.
 \begin{figure}[h!]
@@
 
-Where the matrix \textbf{T} accounts for the targets shape, and 
-refractive index. The $T$-matrix method can be easily 
-applied to any solution of Maxwell's equations, for example Eq.~\eqref{eq:scattered} can be instead be rewritten as:
+Where the matrix \textbf{T} accounts for the targets shape, 
+and refractive index. The $T$-matrix method can be easily 
+applied to any solution of Maxwell's equations, for example 
+Eq.~\eqref{eq:scattered} can be instead be rewritten as:
 \begin{align}
@@
 compute the force. This project utilises the results from 
-Farsund \cite{Farsund1996} but in the form used by Crichton \cite{Crichton2000THEMD}. As an example the linear momentum 
+Farsund \cite{Farsund1996} but in the form used by Crichton 
+\cite{Crichton2000THEMD}. As an example the linear momentum 
 carried by any beam can be computed via:
@@
 subdivided into its gradient and scattering components, where the latter
-directs the particle to the centre of the beam, and the latter acts on 
+directs the particle to the centre of the beam, and the former acts on 
 the particle in the direction of beam propagation. For a beam with a
@@
 the scattered field is now significant. As the target particle gets 
-larger this assumption fails to accurately describe the trapping force \cite{Li2021}. For larger particles the optical force and scattering 
+larger this assumption fails to accurately describe the trapping force 
+\cite{Li2021}. For larger particles the optical force and scattering 
 is better described via the Lorenz-Mie theory.
@@
 over the targets surface. These currents are chosen so that the 
-electromagnetic field outside is identical to the original problem \cite{Wriedt1998}. By expanding the harmonic functions one can 
+electromagnetic field outside is identical to the original problem 
+\cite{Wriedt1998}. By expanding the harmonic functions one can 
 derive the individual elements of \textbf{T} until a desired 
@@
 can be applied to the scattering from particles of arbitrary 
-composition and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, the DDA method approximates the particle as 
+composition and geometry. Developed by Purcell and Pennypacker 
+\cite{Purcell1973}, the DDA method approximates the particle as 
 being constructed of discrete dipoles. This is essentially an 
@@
 \begin{align}
-	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  \mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
+	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
+	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
 	\chi(\mathbf{r'})\mathbf{E(r')}
@@
 The Fickian model describes the net flux of solute molecules into a 
-finite volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
+finite volume of fluid being proportional to the density gradient 
+$\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
 solute molecules do not collide with one another or with molecules in 
@@
 molecules disperse over long periods of time. It does not provide any 
-incite into the forces acting on individual molecules, whereas the 
+insight into the forces acting on individual molecules, whereas the 
 Langevin model captures the physical interactions between an individual 
@@
 multiple collisions with other molecules \cite{Gillespie2012a} over a 
-given time interval $\delta t$. The Einstein molecule allows us to 
+given time interval $\delta t$. The Einstein model allows us to 
 consider how individual molecules behave while suspended in a 
@@
 	\includegraphics[height=0.68\linewidth]{finite_differences.png}
-	\caption{\rule{0ex}{20ex}Example trajectory created using Finite Differences method
-			 for a $2\mu m$ diameter sphere.
-Trap stiffness's were estimated using \textit{ott} at $\kappa_x = \kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$.
+	\caption{\rule{0ex}{20ex}Example trajectory created using Finite 
+		Differences method for a $2\mu m$ diameter sphere. Trap 
+		stiffness's were estimated using \textit{ott} at $\kappa_x = 
+		\kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$. 
 		The particle's motion can be localised around the shaded ellipsoid.}
@@
 	\includegraphics[width=\linewidth]{MSD.png}
-	\caption{Example mean squared displacement using \eqref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well.
-The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
+	\caption{Example mean squared displacement using \eqref{eq:MSD}, 
+		for a $1\mu m$ sphere trapped by an optical potential well. 
+		The dotted line represents the upper limit of the sphere's 
+		displacement due to the optical trap.}
 \end{figure}
@@
 focus) or a recorded signal from a position-detector,and $\hat{\eta}$ 
-is the Fourier transform of the white noise (see Eq.~\eqref{eq:full_langevin}). Where the values are exponentially distributed as opposed to being 
+is the Fourier transform of the white noise (see Eq.~\eqref{eq:full_langevin}). 
+Where the values are exponentially distributed as opposed to being 
 normally distributed in the time domain \cite{BergSoerensen2004}.
@@
 diffusion coefficient (in arbitrary units) and the corner frequency 
-$f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. The Lorentzian shape implies that the trap 
+$f_c$ which is directly related to the trap strength via $f_c = 
+\kappa_x/(2\pi\gamma)$. The Lorentzian shape implies that the trap 
 is harmonic - but not symmetric - which assumes that the particle 
@@
 	\includegraphics[width=\linewidth]{PSD.png}
-	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water.
-The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector.
-Using a correction factor (see \eqref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
+	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power 
+		spectra is collected from an optically trapped silica sphere 
+		suspended in water. The difference in magnitude is due to 
+		the asymmetry of the quadrant photo diode having a stronger 
+		signal response in the direction of the polarisation vector. 
+		Using a conversion factor (see \eqref{eq:conversion_factor}) 
+		will adjust the power spectra to better describe the trap shape.}
 \end{figure}
@@
 full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
-a theoretical model for that relates the power spectrum to both the 
+a theoretical model that relates the power spectrum to both the 
 rotational and translational motion simultaneously.
@@
 \begin{align}
-\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
+	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
+	\sqrt{2\vec{D}_x}\eta(t)
 \end{align}
@@
 Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
-motion is generalised to a 3 dimensional case. Except now, the dimer 
+motion is generalised to a 3 dimensional case. Here we omit the inertia term 
+as the restoration time of the optical trap is far greater than the characteristic
+time of the particle ($\kappa/\gamma \gg m/\gamma$). Now, the dimer 
 is undergoing random rotational motion in addition to its Brownian 
 translational motion the first term on the right hand side is no longer 
-purely a function of the dimer's position but also on its orientation. 
+purely a function of the dimer's position but also of its orientation. 
 The rotational form of the Langevin equation for a dipole within an external 
@@
 coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
-which are found by calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
+which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
+\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
 total force in along the z-direction and the total torque about 
@@
 \label{sec:sim_parameters}
-Throughout this project we use the work of Vigilante to perform a 
-systematic study of the dynamics demonstrated by asymmetric dimers 
-in both plane and circularly polarised light. The simulations have 
-two primary frames of reference that require explanation. Firstly, 
-there is the laboratory frame which describes the coordinate system, 
-we define the origin of our simulation as the focus of our trapping 
-beam. While the choice in direction for the $x$ and $y$ axis are 
-arbitrary the $z$ axis is specifically chosen so that it is aligned 
-with the direction of beam propagation. 
+Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
+to perform a systematic study of the dynamics demonstrated by 
+asymmetric dimers in both plane and circularly polarised light. 
+The simulations have two primary frames of reference that require 
+explanation. Firstly, there is the laboratory frame which 
+describes the coordinate system, we define the origin of our 
+simulation as the focus of our trapping beam. While the choice in 
+direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
+specifically chosen so that it is aligned with the direction of 
+beam propagation. 
 \begin{figure}[h!]
@@
 	E_{i,z}(x,y)+E_{s,z}(x,y)
-\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
+\end{bmatrix} \right|^2 \times step(\theta_{NA}-\theta(x,y))
 \end{align}
@@
 	E_{\rm inc}(r)&=
-	E_0\sum^\infty_{nm}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
+	E_0\sum^\infty_{n}\sum^n_{m=-n}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
 	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
 	\\
-	E_{\rm scat}(r)&=E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
-	-i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
+	E_{\rm scat}(r)&=E_0 \sum^\infty_n \sum^n_{m=-n} \left[
+	p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
 	\right] 
@@
 
-We can, if we want, further restrict the incident beam by setting 
-the truncation angle to match our microscope object, this applies a 
-hard cut off to the incident beam to better represent experimental 
-conditions. In order to compute the scattering from 
-the target particle \textit{ott} uses the $T$-matrix method, this is 
-not essential for a simple sphere but is essential for complex shaped 
-particles such as dimers. The matrix from mstm needs to be repackaged 
-to work with the \textit{ott} software, converting it from a column 
-vector into a system of sub matrices.
-
-The scattered and incident fields are then combined together 
-in the far field to get $I(x,y)$, the quadrant and overall signals are 
-calculated via:
+Where the superscript (2) denotes an outgoing spherical harmonic
+function. In order to compute the scattering from the target particle \textit{ott} uses the $T$-matrix method, this is not essential 
+for a simple sphere but is essential for complex shaped particles 
+such as dimers. The scattered and incident fields are then combined 
+together in the far field to get $I(x,y)$, the quadrant and overall 
+signals are calculated via:
 \begin{align}
@@
 
-\newpage
 While translational motion have no knock on effects to the QPD signal, 
@@
 
-The discrepancy between our simulated QPD and the results from 
-Rohrbach can be attributed to the fact that the position to 
-signal error grows as you move further from trap focus. In most
-optical trapping experiments we usually are calibrating a strong
-trap where the mean square displacement falls well within the
-linear regime shown in fig.~\ref{fig:Rohrbach}. In which case the
-error between Rohrbach's model and our own is inconsequential for 
-calibration purposes. With this any trajectory can be collected 
+The discrepancy between our simulated QPD and the results 
+from Rohrbach can be attributed to the fact that the position 
+to signal error grows as you move further from trap focus. 
+As the particle moves further from the trap focus the error 
+in the attributed signal grows substantially \cite{Rohrbach2002}.
+In most optical trapping experiments we usually are calibrating 
+a strong trap where the mean square displacement falls well within 
+the linear regime shown in fig.~\ref{fig:Rohrbach}. In which case 
+the error between Rohrbach's model and our own is inconsequential 
+for calibration purposes. With this any trajectory can be collected 
 from the QPD by displacing and rotating the beam accordingly, 
-later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of accuracy
-that can be achieved using back focal plane interferometry to 
-characterise non-spherical targets.
+later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of 
+accuracy that can be achieved using back focal plane interferometry 
+to characterise non-spherical targets.

### 2024-11-05 Latex_files/Thesis/chapters/chapter 4/main.tex
praveen commments plus chp 4 rework

@@
 \label{chapter:langevin_dynamics}
-Much of the calibration theory discussed in Chapter 2 assumes that 
-the target particle in question is a single sphere, one whose 
-scattering and motion is easily computed. However, while working 
-with dense colloidal suspensions, one often ends up trapping more 
-than one sphere. Li and Arlt \cite{Li2008} studied the case of two 
-microspheres trapped in a single OT and found that multiple trapped 
-beads could be mistaken for a single trapped bead with altered trap 
-stiffness (see \ref{sec:harmonic_traps}). Theoretical studies on the 
-case of two trapped microspheres by Xu~\textit{et~al.} \cite{Xu2005} 
-employed a ray-optics based model to show that the two trapped beads 
-are brought into physical contact with each other by optical forces 
-and they also calculated the axial equilibrium positions of the two 
-trapped beads as a function of their size. Experiments in 
-\cite{Praveen2016} confirmed that the two trapped beads indeed 
-experience different trap stiffnesses in the vicinity of the same 
-potential well.
+Much of the calibration theory discussed in Chapter 2 
+assumes that the target particle in question is a single 
+sphere, one whose scattering and motion is easily computed. 
+However, while working with dense colloidal suspensions, 
+one often ends up trapping more than one sphere. Li and 
+Arlt \cite{Li2008} studied the case of two microspheres 
+trapped in a single OT and found that multiple trapped 
+beads could be mistaken for a single trapped bead with 
+altered trap stiffness (see \ref{sec:harmonic_traps}). 
+Theoretical studies on the case of two trapped microspheres 
+by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
+based model to show that the two trapped beads are brought 
+into physical contact with each other by optical forces 
+and they also calculated the axial equilibrium positions 
+of the two trapped beads as a function of their size. 
+Experiments in \cite{Praveen2016} confirmed that the two 
+trapped beads indeed experience different trap stiffnesses 
+in the vicinity of the same potential well.
 
-There are further discussions looking into the dynamics of a whole 
-host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour 
-an arbitrary shaped particle comes with great difficulty due to the 
-fact that the optical force is dependent on a greater number of 
-variables such as orientation and size factors.
+There are further discussions looking into the dynamics of 
+a whole host of asymmetrically shaped particles \cite{Loudet2014, 
+ShengHua2005, Chetana2022}, their results all showing that 
+predicting the behaviour an arbitrary shaped particle comes 
+with great difficulty due to the fact that the optical force 
+is dependent on a greater number of variables such as 
+orientation and size factors.
 
-With the initial goal of the PhD being to induce nucleation events 
-via a spherical micro-rotor, the aim of this chapter is to - in a 
-limited capacity simulate and investigate the influence of a second 
-particle being bound to our target sphere. The choice of a dimer, 
-instead of an amorphous solid that might better represent a growing 
-crystalline solid, allows us to consider how the dynamics of the 
-aggregate change by varying the size factor. Additionally, attempting 
-to simulate an amorphous aggregate is rather difficult as calculating 
-the optical force and torque is computationally slow. We build upon 
-the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
-asymmetric spherical dimers and how varying size parameters alters the 
-dynamics and additionally makes characterising their interactions within 
-an optical trap more cumbersome.
+With the initial goal of the project being to induce 
+nucleation events via a rotating sphere, the aim of this 
+chapter is to - in a limited capacity simulate and investigate 
+the influence of a second particle being bound to our target 
+sphere. The choice of a dimer, instead of an amorphous solid 
+that might better represent a growing crystalline solid, 
+allows us to consider how the dynamics of the aggregate change 
+by varying a single parameter, namely the size ratio of the 
+two spheres. Additionally, attempting to simulate an amorphous 
+aggregate is rather difficult as calculating the optical force 
+and torque is computationally slow. We build upon the works of 
+Vigilante \textit{et al} \cite{Vigilante2020} to consider 
+asymmetric spherical dimers and how varying size parameters 
+alters the dynamics and additionally makes characterising their 
+interactions within an optical trap more cumbersome.
 
@@
 \label{sec:eq_positions}
-If we wanted to start from first principles and determine the trap 
-strength on our target particle the first step would be to locate the equilibrium position relative to the trap focus. For a single sphere 
-it is easy to enough to understand that its centre of mass will be 
-drawn to focal point of the laser due to gradient forces, once there 
-the force is analogous to a harmonic spring with a fixed trap stiffness 
-(see fig.\ref{fig:harmonic_trap}). Now, if we consider instead a dimer, 
-we now have two spheres both being drawn to the focus along by the same gradient force; in addition the scattering force is significantly more 
-complex due to both spheres scattering the electromagnetic fields. 
-This mutual scattering between individual spheres is what makes simulating spherical aggregates far more difficult compared to a single sphere, 
-and even harder still to predict the position where the dimer's centre 
-of mass is in equilibrium.
+If we wanted to start from first principles and determine 
+the trap strength on our target particle the first step 
+would be to locate the equilibrium position relative to 
+the trap focus. For a single sphere it is easy to enough 
+to understand that its centre of mass will be drawn to 
+focal point of the laser due to gradient forces, once there 
+the force is analogous to a harmonic spring with a fixed 
+trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
+we consider instead a dimer, we now have two spheres both 
+being drawn to the focus along by the same gradient force; 
+in addition the scattering force is significantly more complex 
+due to both spheres scattering the electromagnetic fields. 
+This mutual scattering between individual spheres is what 
+makes simulating spherical aggregates far more difficult 
+compared to a single sphere, and even harder still to 
+predict the position where the dimer's centre of mass is 
+in equilibrium.
 
-Because the scattering force is only significant in the direction of 
-beam propagation \cite{Capitanio2002} we can assume there is only a 
-single potential well in the transverse plane. Along the beam axis, 
-such an assumption is no longer valid due to fact that spherical 
-aggregates experience inter-particle scattering. This is a key reason
-for using \textit{mstm} as it accounts for that behaviour. The 
-methodology for computing optical forces has been covered extensively 
-for a number of different trapping conditions \cite{RanhaNeves2019}. 
-So it is relatively easy to compute the trapping force and determine 
-the equilibrium position by finding the position that minimises the 
-net optical force and the local gradient is negative ($\delta F/
-\delta x < 0$) - we can assume that for a dielectric sphere the 
-optical torque is negligible. For a dimer (or any arbitrary spherical aggregate), we now must consider both its position and orientation 
-and find where the net optical force and torque are minimised.
+Because the scattering force is only significant in the 
+direction of beam propagation \cite{Capitanio2002} we 
+can assume there will only be a single potential well 
+lying at the centre of the beam. Along the beam axis, 
+such an assumption is no longer valid due to fact that 
+spherical aggregates experience inter-particle scattering. 
+This is a key reason for using \textit{mstm} as it 
+accounts for that behaviour. The methodology for computing 
+optical forces has been covered extensively for a number 
+of different trapping conditions \cite{RanhaNeves2019}. 
+So it is relatively easy to compute the trapping force 
+and determine the equilibrium position by finding the 
+position that minimises the net optical force and the 
+local gradient is negative ($\delta F/\delta x < 0$) - 
+we can assume that for a dielectric sphere the optical 
+torque is negligible. For a dimer (or any arbitrary 
+spherical aggregate), we now must consider both its 
+position and orientation and find where the net optical 
+force and torque are minimised.
 
@@
 		\centering
-		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
+		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
 		\label{large over small}
@@
 		\centering
-		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
+		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
 		\label{small_over_large}
@@
 either a purely vertical or purely horizontal orientation \cite{Zhu2021}.
-It highlights an interesting aspect of the dimer's behaviour, namely that 
-the potential phase space $U(\textbf{r}, \theta, \phi)$ is not continuous.
 
@@
 	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
-	\caption{Rotation frequency vs component phase difference for differently 
-		sized dimers. The solid lines represent the rotation rate experienced 
-		while the dimer is in its standard orientation, whereas the solid points are for the case where the orientation is inverted. Laser power $= 100\ mW$.}
+	\caption{Rotation frequency vs component phase difference for 
+		differently sized dimers. The solid lines represent the 
+		rotation rate experienced while the dimer is in its standard 
+		orientation, whereas the solid points are for the case 
+		where the orientation is inverted. Laser power = $100\ mW$.}
 	\label{fig:rotation_vs_pol}
@@
 	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
-	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how the momentum 
-		density of a Gaussian beam is twisted while using circularly polarised 
-		light. The top row (figures (b)-(d)) shows a 7 sphere cluster trapped 
-		in a circularly polarised beam. Due to the clusters asymmetric 
-		susceptibility to polarization the cluster rotates in the $x-y$ plane.
-Whereas the bottom row (figures (e) - (g)) show the similar results for a single sphere.
-In this instance the sphere does not rotate but instead
-		orbits the beam axis. In both instances the motion is proportional to 
-		the degree of polarisation (see figures (d) and (g)) but for the single
-		sphere this motion is only revealed when using \eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused with permission from author}
+	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how 
+		the momentum density of a Gaussian beam is twisted while 
+		using circularly polarised light. The top row (figures 
+		(b)-(d)) shows a 7 sphere cluster trapped in a circularly 
+		polarised beam. Due to the clusters asymmetric susceptibility 
+		to polarization the cluster rotates in the $x-y$ plane. 
+		Whereas the bottom row (figures (e) - (g)) show the similar 
+		results for a single sphere. In this instance the sphere 
+		does not rotate but instead orbits the beam axis. In both 
+		instances the motion is proportional to the degree of 
+		polarisation (see figures (d) and (g)) but for the single 
+		sphere this motion is only revealed when using 
+		\eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused 
+		with permission from author}
 	\label{fig:Ruffner_Grier}
@@
 
-While the results from \cite{Ruffner2012} suggest that the optical rotation 
-seen in asymmetric dimers is attributed to the same spin-curl forces there 
-are several discrepancies that cannot be explained purely by the spin-curl 
-force. Firstly, there is the question of how the spin-curl force results in
-an optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) and (e) we 
-can see that the behaviour of the 7 sphere cluster rotates about the central 
-axis whereas the single sphere has an only slight orbital bias in its motion. 
-The authors attributed this to the fact that the cluster is scattering light 
-where the spin-curl force is more substantial due to it having a wider profile. 
-This would suggest that particles' whose longer axis lies in the plane perpendicular to the direction of propagation should experience any notable degree of torque.
-And indeed this appears to be the case for a number of experiments involving nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
+While the results from \cite{Ruffner2012} suggest that the 
+optical rotation seen in asymmetric dimers is attributed to 
+the same spin-curl forces there are several discrepancies that 
+cannot be explained purely by the spin-curl force. Firstly, 
+there is the question of how the spin-curl force results in an 
+optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) 
+and (e) we can see that the behaviour of the 7 sphere cluster 
+rotates about the central axis whereas the single sphere has 
+an only slight orbital bias in its motion. The authors attributed 
+this to the fact that the cluster is scattering light where the 
+spin-curl force is more substantial due to it having a wider 
+profile. This would suggest that particles' whose longer axis 
+lies in the plane perpendicular to the direction of propagation 
+should experience any notable degree of torque. And indeed this 
+appears to be the case for a number of experiments involving 
+nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
 
 \subsection{Comparison of optical torques for Sphere and Dimers}
-It does raise the question of how come our simulations show that despite the 
-dimer being orientated so their long axis' align with the parallel to the 
-direction of propagation they still readily rotate. You would not expect that
-the addition of a single additional sphere should drastically adjust the torque
-especially if said sphere is relatively small. However when we measured the 
-optical torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found 
-the exact opposite. In both cases we used the same trapping beam as used for
-figure~\ref{fig:rotation_vs_pol} but with a circularly polarised beam. Both 
-the sphere and dimer were rotated in the $x-z$ plane and the all three components of the optical torque were recorded.
+It does raise the question of how come our simulations show that 
+despite the dimer being orientated so their long axis' align with 
+the parallel to the direction of propagation they still readily 
+rotate. You would not expect that the addition of a single additional 
+sphere should drastically adjust the torque especially if said sphere 
+is relatively small. However when we measured the optical torque 
+of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found the 
+exact opposite. In both cases we used the same trapping beam as used 
+for fig.~\ref{fig:rotation_vs_pol} but with a circularly polarised 
+beam. Both the sphere and dimer were rotated in the $x-z$ plane and 
+the all three components of the optical torque were recorded.
 \begin{figure}[h!]
@@
 
-The torques about the $x$ and $y$ axis can be somewhat understood as the second 
-sphere is being drawn back towards the centre of the trap by the gradient forces.
-The same cannot be said for the $z$component of the optical torque; while this effect could still be attributed to the spin-curl force, but it is clear that the internal scattering between the two spheres has some unintended effects.
+The torques about the $x$ and $y$ axis can be somewhat understood 
+as the second sphere is being drawn back towards the centre of the 
+trap by the gradient forces. The same cannot be said for the $z$-
+component of the optical torque; while this effect could still be 
+attributed to the spin-curl force, but it is clear that the internal 
+scattering between the two spheres has some unintended effects.
 
 \subsection{Rotational frequency as a function of size ratio and orientation}
-Moving on, next we want to see how the rotational speed is influenced by 
-different sized dimers. Intuitively, you would expect that a larger would 
-experience a greater torque and therefore rotate faster.
-By repeating the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for a circularly polarised beam $\phi=90\circ$ it was found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.
+Next we want to see how the rotational speed is influenced by 
+different sized dimers. Intuitively, you would expect that a 
+larger would experience a greater torque and therefore rotate 
+faster. By repeating the same kinds of simulation as used in 
+\ref{fig:rotation_vs_pol} but for a circularly polarised beam 
+$\phi=90\circ$ it was found that not only is the rotation rate 
+dependent on the size of the dimer, but also on its orientation 
+and therefore their axial position.
 \begin{figure}[h!]
@@
 second order contributions we would need to construct a dipole array
-that fully captures the scattering of a dimer. Measuring the optical torque 
-makes it clear that the polarizability is a contributing factor to this optical 
-rotation phenomena.
+that fully captures the scattering of a dimer. Measuring the optical 
+torque makes it clear that the polarizability is a contributing factor 
+to this optical rotation phenomena.
 
 \subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there 
-is the potential for non-vertical trapping orientations in which the dimer is 
-still stably trapped.
-When a circularly polarised beam is used the dimer exhibits gyroscopic precession.
-As shown in fig~\ref{fig:gyro} the dimer's trajectory exhibits periodic rotation about its long axis $U_z$, precession motion as the dimer spins around the central beam axis, and nutation motion as the dimer's orientation rocks backwards and forwards.
+As mentioned in section~\ref{sec:off-axis} for specificity sized dimers
+there is the potential for non-vertical trapping orientations in which 
+the dimer is still stably trapped. When a circularly polarised beam is 
+used the dimer exhibits gyroscopic precession. As shown in fig.
+~\ref{fig:gyro} the dimer's trajectory exhibits periodic rotation about 
+its long axis $U_z$, precession motion as the dimer spins around the 
+central beam axis, and nutation motion as the dimer's orientation rocks 
+backwards and forwards. 
 \begin{figure}[h]
@@
 rotational frequency and the laser power, but no such relationship existed 
-with the precession frequency. Our own results shows a similar linear 
-relationship with vertically aligned dimers.
-\begin{figure}
-	\includegraphics[width=\linewidth]{RotationvsPower.png}
-	\caption{Rotational frequency vs laser power for a dimer in both vertical orientations.}
-\end{figure}
-
-The linear relationship could partly be due to fact that we do not
-account for the change in viscous forces with increasing laser power. 
-Due to the localised heating effect 
-It is far more likely that the rotational frequency reaches a maximum 
-value assuming that the bulk fluid can readily absorb the laser.
+with the precession frequency.
 
@@
 		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory fitting}\\
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
+		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
+		fitting}\\
 		\hline
@@
 axis'. If calibrated over an long enough time frame you would expect 
-that the resulting power spectra would exactly mirror the \textit{ott} predictions. However over a short calibration time a QPD can predict
+that the resulting power spectra would exactly mirror the \textit{ott} 
+predictions. However over a short calibration time a QPD can predict
 large discrepancies in the trap strength. There is a clear trade off 
@@
 position change but it is subjected to rotational motion due to the 
-beam's angular momentum. This is reflected
+beam's angular momentum. This is reflected in the calibration results
+using the simulated QPD, where we see a drastically different estimation
+between the \textit{ott} estimate and the QPD estimate.
 
@@
 		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
+		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
 		\hline

### 2024-11-18 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to chapter 4, among others

@@
 
-In this chapter we will consider how the addition of a second sphere 
-changes the trapping dynamics by introducing multiple equilibrium 
-positions. These positions are dependent on both the size ratio of 
-the dimer and its orientation relative to the trap. This is important 
-to address, for while the the dimers' diffusive behaviour is the same, 
-regardless of position and orientation, how it interacts with the trap 
-may produce the wildly different results. Furthermore, we look at how 
-our dimers interact with circularly polarised light, this is especially 
-pertinent as the original plan was to utilise circularly polarised 
-light to generate fluid flow in a supersaturated solution. The 
-rotational motion of dimers in circularly polarised light goes against 
-much of the established literature and suggests that the transfer of 
-spin angular momentum is enabled by combining particles together. And 
-lastly, we demonstrate how a quadrant photo diode, performs in 
-characterising the actual interactions between a dimer and the optical 
-trap. We do so to demonstrate that much of the dimer's trajectory 
-information is either lost or poorly described by simple calibration 
-techniques, making their motion difficult to characterise.
+In this chapter we will consider how the addition of a 
+second sphere changes the trapping dynamics by introducing 
+multiple equilibrium positions. These positions are 
+dependent on both the size ratio of the dimer and its 
+orientation relative to the trap. This is important to 
+address, for while the the dimers' diffusive behaviour 
+is the same, regardless of position and orientation, 
+how it interacts with the trap may produce the wildly 
+different results. Furthermore, we look at how our dimers 
+interact with circularly polarised light, this is especially 
+pertinent as the original plan was to utilise circularly 
+polarised light to generate fluid flow in a supersaturated 
+solution. The rotational motion of dimers in circularly 
+polarised light goes against much of the established 
+literature and suggests that the transfer of spin angular 
+momentum is enabled by combining particles together. And 
+lastly, we demonstrate how a quadrant photo diode, performs 
+in characterising the actual interactions between a dimer 
+and the optical trap. We do so to demonstrate that much 
+of the dimer's trajectory information is either lost or 
+poorly described by simple calibration techniques, making 
+their motion difficult to characterise.
 
 \subsection{Simulation Parameters}
-As a paradigmatic example, consider a dimer suspended in water 
-($n_p = 1.59, n_m = 1.33$) located at the focus of a Gaussian 
-beam (more specifically a Laguerre-Gaussian beam of mode $[0,0]$, 
-see \eqref{eq:incident}), the beam is focused by a objective with 
-numerical aperture of 1.2 and is x polarised. The size ratio of 
-the two spheres is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is 
-kept at $1\ \mu m$ unless specified otherwise; the dimer's 
-orientation is given by a unit vector connecting the centres of 
-both spheres (see \ref{sec:sim_parameters}). We define the 
-'standard' orientation as being aligned with the direction of beam 
-propagation direction - and therefore the 'inverted' orientation is 
-defined when the dimer is orientated against the direction of beam 
-orientation (see left hand side of figure~\ref{fig:paradigmatic}).
+As a paradigmatic example, consider a dimer suspended in 
+water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
+a Gaussian beam (more specifically a Laguerre-Gaussian 
+beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
+is focused by a objective with numerical aperture of 1.2 
+and is x polarised. The size ratio of the two spheres is 
+given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
+\mu m$ unless specified otherwise; the dimer's orientation 
+is given by a unit vector connecting the centres of both 
+spheres (see \ref{sec:sim_parameters}). We define the 
+'standard' orientation as being aligned with the direction 
+of beam propagation direction - and therefore the 'inverted' 
+orientation is defined when the dimer is orientated against 
+the direction of beam orientation (see left hand side of 
+figure~\ref{fig:paradigmatic}).
 
@@
 
-After computing its $T$-matrix via \textit{mstm} and supplying that 
-to \textit{ott} we compute the optical force exerted by a $50\ mW$ laser
-via \eqref{eq:optical_force} in the axial direction while the dimer is 
-in its 'standard' orientation. As expected we see a single point where 
-the dimer will be in equilibrium, the linear fit in fig.~\ref{fig:paradigmatic}
-(a) indicates a that the force can be modelled as a harmonic potential close 
-to the equilibrium position ($F_z\approx-\kappa_z z$). The second point 
-where the axial force goes to 0 cannot be considered as equilibrium position 
-as the positive gradient indicates that the trap is unstable unless Brownian 
-motion is ignored.
+After computing its $T$-matrix via \textit{mstm} and 
+supplying that to \textit{ott} we compute the optical 
+force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
+in the axial direction while the dimer is in its 'standard' 
+orientation. As expected we see a single point where 
+the dimer will be in equilibrium, the linear fit in 
+fig.~\ref{fig:paradigmatic}(a) indicates a that the 
+force can be modelled as a harmonic potential close to 
+the equilibrium position ($F_z\approx-\kappa_z z$). 
+The second point where the axial force goes to 0 cannot 
+be considered as equilibrium position as the positive 
+gradient indicates that the trap is unstable unless 
+Brownian motion is ignored.
 
-We repeated the same calculation but now while the dimer is in its 'inverted' 
-orientation, instead of a single point where the optical force is minimised we 
-see that there are instead two separate equilibrium positions, one above the focus
-and one below the focus. In this particular example the two positions are far enough
-apart that both can be considered as separate harmonic traps.    
+We repeated the same calculation but now while the 
+dimer is in its 'inverted' orientation, instead of 
+a single point where the optical force is minimised 
+we see that there are instead two separate equilibrium 
+positions, one above the focus and one below the 
+focus. In this particular example the two positions 
+are far enough apart that both can be considered as 
+separate harmonic traps.    
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Plots of force vs displacement of the centre of mass of 
-		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
-		the case where the dimer is in its' 'standard' orientation, where
-		the dimer is trapped at $z = 1.71\ \mu m$. (b) is the case where 
-		the dimer is in its' 'inverted' orientation, the dimer is trapped 
-		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. 
-		On the left are renders to visualise the dimer orientation are shown 
-		below each plot. The black lines on each force-curve is a linear 
-		fit with the slope being reported as the trap stiffness in the legend.}
+	\caption{Plots of force vs displacement of the 
+		centre of mass of the dimer (µm) for the 
+		case of a dimer of size ratio 2. (a) is the 
+		case where the dimer is in its' 'standard' 
+		orientation, where the dimer is trapped at 
+		$z = 1.71\ \mu m$. (b) is the case where the 
+		dimer is in its' 'inverted' orientation, the 
+		dimer is trapped at two positions: $z = 1.20\ 
+		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
+		renders to visualise the dimer orientation 
+		are shown below each plot. The black lines 
+		on each force-curve is a linear fit with the 
+		slope being reported as the trap stiffness in 
+		the legend.}
 	\label{fig:paradigmatic}
@@
 
-We can see that both equilibrium positions have comparable axial trap 
-stiffness ($\kappa_z$), however the difference in the transverse trap 
-stiffness ($\kappa_x$) is far more noticeable. The same dimer was trapped
-at each of the axial equilibrium positions and the transverse force was 
-evaluated. While in all three cases the dimer can be trapped the linear
-range where that would typically associated with a stable trap is far 
-narrower in the 'standard' orientation compared to the 'inverted' cases. 
-This highlights one of the challenges involved with studying 
-asymmetric particles, even though its a simple enough process to trap them 
-they maybe characterised very differently depending on their relative position 
-and orientation towards the focus. This can have a significant impact on 
-rheological studies - or attempting to probe any local property - as the 
-variance in trap strength can result in large errors over repeated measurements. 
+We can see that both equilibrium positions have comparable 
+axial trap stiffness ($\kappa_z$), however the difference 
+in the transverse trap stiffness ($\kappa_x$) is far more 
+noticeable. The same dimer was trapped at each of the axial 
+equilibrium positions and the transverse force was evaluated. 
+While in all three cases the dimer can be trapped the linear
+range where that would typically associated with a stable 
+trap is far narrower in the 'standard' orientation compared 
+to the 'inverted' cases. This highlights one of the challenges 
+involved with studying asymmetric particles, even though its 
+a simple enough process to trap them they maybe characterised 
+very differently depending on their relative position and 
+orientation towards the focus. This can have a significant 
+impact on rheological studies - or attempting to probe any 
+local property - as the variance in trap strength can result 
+in large errors over repeated measurements. 
 \begin{figure}[h!]
@@
 	\includegraphics[width=\linewidth]{transverse_force.png}
-	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
-		where a positive force indicates the dimer is directed right on the x-axis, 
-		and vice versa for a negative force.
-The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c).
-The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
+	\caption{Plots of force vs displacement of the dimer's 
+		centre of mass spheres, where a positive force 
+		indicates the dimer is directed right on the x-axis, 
+		and vice versa for a negative force. The same simulation 
+		parameters are used here as in fig~\ref{fig:paradigmatic}
+		(a) and (c). The blue curve representing the force 
+		response for a dimer in its standard orientation, orange 
+		being the inverted case, and green the same case but 
+		placed below the focus.}
 	\label{fig:transverse_force}
@@
 \newpage
-For completeness the harmonic traps were located for dimers across a range 
-of size ratios - from $a_{I}/a_{II} = 1$ to $a_{I}/a_{II}=10$ - while also 
-recording the trap stiffness for each trap. The same simulation parameters are 
-used here as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
-As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease the dimer begins to
-approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
-to $a_{I}/a_{II}=4$), a second equilibrium position is found below the trapping 
-focus. Previous work using the ray-optics model have confirmed even in the case 
-that two spheres begin separated the electric field will align the particles as 
-such that they make contact and are trapped together about a single trapping 
-position \cite{Xu2005}. Furthermore it has been shown through proper manipulation 
-of the Gaussian or Bessel beam modes that any number of trapping potentials can 
-be developed \cite{Shahabadi2020} for nanoparticles. This result however, is the 
-first example of an orientation dependent trapping situation using only a 
-$TEM_00$ beam. Typical experimental arrangements cannot determine much information 
-on the axial position of a trapped particle relative to the trap focus; this 
-result indicates not only that dimers can be trapped in multiple axial positions 
-but also their trapping behaviour is heavily dependent on said axial position. 
-As such it is necessary that positional information in the z-axis can be elucidated
-if multiple spheres are trapped simultaneously. 
+For completeness the harmonic traps were located for dimers
+across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
+$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
+for each trap. The same simulation parameters are used here 
+as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
+As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
+the dimer begins to approximate a single homogenous sphere 
+- at least in terms of location and trap strength. However, 
+for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
+to $a_{I}/a_{II}=4$), a second equilibrium position is found 
+below the trapping focus. Previous work using the ray-optics 
+model have confirmed even in the case that two spheres begin 
+separated the electric field will align the particles as such 
+that they make contact and are trapped together about a single 
+trapping position \cite{Xu2005}. Furthermore it has been shown 
+through proper manipulation of the Gaussian or Bessel beam 
+modes that any number of trapping potentials can be developed 
+\cite{Shahabadi2020} for nanoparticles. This result however, 
+is the first example of an orientation dependent trapping 
+situation using only a $TEM_00$ beam. Typical experimental 
+arrangements cannot determine much information on the axial 
+position of a trapped particle relative to the trap focus; 
+this result indicates not only that dimers can be trapped 
+in multiple axial positions but also their trapping behaviour 
+is heavily dependent on said axial position. As such it is 
+necessary that positional information in the z-axis can be 
+elucidated if multiple spheres are trapped simultaneously. 
 \begin{figure}[h!]
@@
 \label{sec:off-axis}
-Computing the equilibrium positions when a dimer is aligned with the electric 
-field is relatively simple as the orientational torque is minimised (see
-Eq.\ref{eq:opt_torque}). Meaning once trapped the dimer is unlikely to change
-orientation enough to escape the trap. Regardless, that does not rule out the 
-possibility that there is a stable configuration where the orientation not 
-strictly vertical, in fact most experimental work with symmetric nano-dimers 
-will trap them lying perpendicular to the beam direction \cite{Ahn2018, 
-Reimann2018}. Unlike in Sec.~\ref{sec:eq_positions} we cannot simply measure 
-the optical force and torque as the parameter space is too large and determining 
-if a particular position and orientation is stable is not clear based solely 
-on force and torque measurements \cite{Bui2017}. Using the same simulation 
-parameters as before we ran a number short simulations (total simulation time 
-was $0.005$ s) with the laser power increased to $500\ mW$. Each simulation 
-started with the dimer in a different starting position and orientation, due 
-to the high laser power the dimers either escaped the trap or were stably 
-trapped. The $z-\theta$ phase space - where $\theta$ is the angle between the 
-direction of beam propagation and the dimer's orientation vector ($\theta=0^
-\circ$ is the 'standard' orientation) - can be divided into different regions 
-depending on which equilibrium configuration is reached.
+Computing the equilibrium positions when a dimer is aligned 
+with the electric field is relatively simple as the orientational 
+torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
+trapped the dimer is unlikely to change orientation enough to 
+escape the trap. Regardless, that does not rule out the 
+possibility that there is a stable configuration where the 
+orientation not strictly vertical, in fact most experimental 
+work with symmetric nano-dimers will trap them lying perpendicular 
+to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
+\ref{sec:eq_positions} we cannot simply measure the optical force 
+and torque as the parameter space is too large and determining if 
+a particular position and orientation is stable is not clear based 
+solely on force and torque measurements \cite{Bui2017}. Using the 
+same simulation parameters as before we ran a number short 
+simulations (total simulation time was $0.005$ s) with the laser 
+power increased to $500\ mW$. Each simulation started with the 
+dimer in a different starting position and orientation, due to the 
+high laser power the dimers either escaped the trap or were stably 
+trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
+between the direction of beam propagation and the dimer's orientation 
+vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
+divided into different regions depending on which equilibrium 
+configuration is reached.
 \begin{figure}[h!]
@@
 	%\captionsetup{hangindent=2.35cm}
-	\caption{Map of $z-\theta$ phase space using a dimer of size ratio 2 with 
-		a laser power of 500 mW ($\theta=0\circ$ is the 'standard' orientation 
-		and $\theta=180^\circ$ is the 'inverted' orientation). The stable 
-		configurations are indicated by the larger circles and the starting 
-		conditions are colour coded to match the stable point they end up in. 
-		Right hand render shows a dimer in its off-axis configuration.}
+	\caption{Map of $z-\theta$ phase space using a dimer of
+		size ratio 2 with a laser power of 500 mW ($\theta=0
+		\circ$ is the 'standard' orientation and $\theta=180
+		^\circ$ is the 'inverted' orientation). The stable 
+		configurations are indicated by the larger circles 
+		and the starting conditions are colour coded to match 
+		the stable point they end up in. Right hand render 
+		shows a dimer in its off-axis configuration.}
 	\label{fig:off_axis}
@@
 
-Interestingly while the trap strength of these off-axis traps are similar 
-in magnitude to the vertically aligned dimers, but when the laser power 
-is lowered (around $5\ mW$) the traps become metastable; after reaching its'
-equilibrium configuration the particle behaves similarly to a typically
-trapped dimer but the trapping potential is small enough that the dimer
-escapes in as little time as less than a second or after nearly a full 
-3 seconds. Running similar simulations but for vertical configurations 
-sees the dimer remaining trapped, even after 30 seconds of run time, 
-indicating that the trapping potential is far greater than the thermal 
-energy. This suggests that the reason this off-axis configuration is 
-metastable is due to the increased rotational freedom. Similar 
-configurations have been explored with ellipsoids; Zhu \textit{et al.} 
-looked at the dynamics of various elliptical particles and found that 
-regardless of shape or initial orientation the particle would tend towards
-either a purely vertical or purely horizontal orientation \cite{Zhu2021}.
+Interestingly while the trap strength of these off-axis traps 
+are similar in magnitude to the vertically aligned dimers, 
+but when the laser power is lowered (around $5\ mW$) the traps 
+become metastable; after reaching its equilibrium configuration 
+the particle behaves similarly to a typically trapped dimer but 
+the trapping potential is small enough that the dimer escapes 
+in as little time as less than a second or after nearly a full 
+3 seconds. Running similar simulations but for vertical 
+configurations sees the dimer remaining trapped, even after 30 
+seconds of run time, indicating that the trapping potential is 
+far greater than the thermal energy. This suggests that the 
+reason this off-axis configuration is metastable is due to the 
+increased rotational freedom. Similar configurations have been 
+explored with ellipsoids; Zhu \textit{et al.} looked at the 
+dynamics of various elliptical particles and found that regardless 
+of shape or initial orientation the particle would tend towards
+either a purely vertical or purely horizontal orientation 
+\cite{Zhu2021}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Continuous rotational motion due to second-order scattering}
-One aspect that has yet to be covered in depth with regards to spherical 
-aggregates of any construction is their interaction with circularly 
-polarised light. For homogenous spheres the optical torque is regarded as
-being negligible as the spin density cannot impart angular momentum while 
-propagating in a homogenous medium. Dimers however, have been shown to
-experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
-while trapped in a circular polarised beam. In our simulations we found 
-that dimers would rotate about their long axis when trapped in circularly
-polarised light. In this section we want to discuss how this behaviour 
-is influenced by size, position, and orientation; and furthermore, we wish
-to address possible explanations for this behaviour, as none of the current
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Continuous rotational motion in circularly polarised light}
+One aspect that has yet to be covered in depth with regards 
+to spherical aggregates of any construction is their interaction 
+with circularly polarised light. For homogenous spheres the 
+optical torque is regarded as being negligible as the spin 
+density cannot impart angular momentum while propagating in a 
+homogenous medium. Dimers however, have been shown to experience 
+an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
+while trapped in a circular polarised beam. In our simulations 
+we found that dimers would rotate about their long axis when 
+trapped in circularly polarised light. In this section we want 
+to discuss how this behaviour is influenced by size, position, 
+and orientation; and furthermore, we wish to address possible 
+explanations for this behaviour, as none of the current
 theories into optically induced rotation seem plausible.
@@
 \subsection{Polarisation Dependency on Dimer trajectory}
-First we wanted to confirm that the rotation seen in our simulations is
-based on the polarisation of the trapping beam. To that end, we simulated 
-the motion of an optically trapped dimer in beams of varying polarisation 
-($NA=1.2$, $P=100\ mW$), the dimer is composed of polystyrene ($n_p=1.59$, 
-$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t =10^-5$) 
-and at the end we looked at the orientational time series; the dimer's 
-orientation is recorded as a quaternion which can be easily converted to 
-a 3-dimensional rotation matrix. By considering only the transverse 
-components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
-matrix and taking the Fourier transformation of their time series reveals 
-the rotational frequency. The laser power is set to $100\ mW$ to avoid 
-large thermal fluctuations and so that the Fourier series of the transverse 
-components approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function 
-centred at the rotational frequency $\omega_{rot}$.
+\label{sec:rot_pol}
+In their paper Vigilante \textit{et al} attribute the rotational motion 
+to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
+scattering is a viable theory for describing optical rotation, however
+there is usually an optical axis about which rotation occurs. For a dimer
+this is typically the orientation vector that passes through the centres 
+of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
+cases where the particle's cross sectional shape is engineered to scatter
+light in one particular direction \cite{Higurashi1994}, but if this was
+the case then the dimer should rotate when illuminated by any polarisation
+of light.
+
+To that end, we simulated the motion of an optically trapped dimer in 
+beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
+composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
+run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
+orientational time series; the dimer's orientation is recorded as a 
+quaternion which can be easily converted to a 3-dimensional rotation 
+matrix. By considering only the transverse components ($U_{1,x}$, 
+$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
+the Fourier transformation of their time series reveals the rotational 
+frequency. The laser power is set to $100\ mW$ to avoid large thermal 
+fluctuations and so that the Fourier series of the transverse components 
+approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
+at the rotational frequency $\omega_{rot}$.
 
@@
 
-This shows us that these optical rotations are polarisation dependent 
-and not merely an example of the dimer scattering light asymmetrically. 
-The question then becomes, by what mechanism is the angular momentum of
-the trapping beam being transferred to the dimer.
+This shows us that these optical rotations are polarisation 
+dependent and not merely an example of the dimer scattering 
+light asymmetrically. We also know that this cannot be a 
+similar mechanism to that of experimental works \cite{Ahn2018, 
+Reimann2018} as in our simulations the dimer's long axis
+is parallel to the electric field. In this orientation 
+the dimer's optical axis is aligned perpendicular to the
+polarisation vector meaning that there should be no angular
+momentum transferred along the optical axis. This is consistent 
+with other experiments involving elongated particles. A dual
+beam trap was used to study the dynamics of "disk like" particles. 
+They found that these particles had one of two stable
+orientations: Either orienting with the 'flat' side perpendicular
+to the beam trap, with no rotation being observed \cite{Brzobohaty2015}. 
+Or orientating with the 'flat' side parallel to the trap, in which 
+case the long axis was aligned with the polarisation vector and 
+thus rotational motion was detected \cite{Brzobohaty2015}.
 
-The rotation was first noted by Vigilante and co-workers who only 
-considered this behaviour for a symmetric dimer \cite{Vigilante2020}.
-In their work they attribute this to spin-curl effects, in which the 
-curl of the spin density leads to a second order optical force that 
-orbits around the beams central axis \cite{Yevick2017}. While several 
-papers have demonstrated this phenomena \cite{Zhao2007,Zhao2009, Wang2010} 
-it was only properly formalised by by \cite{Ruffner2012}. In which they 
-showed that the seemingly random trajectory of a trapped sphere was biased
-by the polarisation state of the trapping beam. While not immediately 
-evident from the trajectory the helicity of the trapping beam was revealed 
+\subsection{Brownian Vortex via Curl of Spin momentum}
+In their work Vigilante refers to the spin-curl effects 
+demonstrated by Grier \textit{et al} \cite{Ruffner2012}, 
+in which the curl of the spin density leads to a second 
+order optical force that orbits around the beams central 
+axis \cite{Yevick2017}. While several papers have 
+demonstrated this phenomena \cite{Zhao2007, Zhao2009, 
+Wang2010} it was only properly formalised by \cite{Ruffner2012}. 
+In which they showed that the seemingly random trajectory 
+of a trapped sphere was biased by the polarisation state 
+of the trapping beam. While not immediately evident from 
+the trajectory the helicity of the trapping beam was revealed 
 by computing the particle's probability flux using.
@@
 
-where $\delta_{\sigma_j}$ is the kernel of an adaptive density estimator
-\cite{Silverman1986}. \eqref{eq:prob_flux} describes the direction a trapped
-sphere is most likely move in given our statistical knowledge of the 
-trajectories probability density function. A finite estimation of the density
-function $p(r)$ is used in \cite{Ruffner2012}.
+where $\delta_{\sigma_j}$ is the kernel of an adaptive 
+density estimator \cite{Silverman1986}. \eqref{eq:prob_flux} 
+describes the direction a trapped sphere is most likely 
+move in given our statistical knowledge of the 
+trajectories probability density function. A finite 
+estimation of the density function $p(r)$ is used in 
+\cite{Ruffner2012}.
 \begin{align}
@@
 
-The probability flux reveals a biased motion in the trajectory of a single 
-sphere (see Fig.~\ref{fig:Ruffner_Grier}). This biased motion results in a
-slight orbital motion about the central axis of the trapping beam, the orbital
-frequency is shown to be proportional to the polarisation state of the trapping
-beam. 
+The probability flux reveals a biased motion in the trajectory 
+of a single sphere (see Fig.~\ref{fig:Ruffner_Grier}). This 
+biased motion results in a slight orbital motion about the 
+central axis of the trapping beam, the orbital frequency is 
+shown to be proportional to the polarisation state of the 
+trapping beam. 
 \begin{figure}[h!]
@@
 While the results from \cite{Ruffner2012} suggest that the 
-optical rotation seen in asymmetric dimers is attributed to 
-the same spin-curl forces there are several discrepancies that 
-cannot be explained purely by the spin-curl force. Firstly, 
-there is the question of how the spin-curl force results in an 
-optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) 
-and (e) we can see that the behaviour of the 7 sphere cluster 
-rotates about the central axis whereas the single sphere has 
-an only slight orbital bias in its motion. The authors attributed 
-this to the fact that the cluster is scattering light where the 
-spin-curl force is more substantial due to it having a wider 
-profile. This would suggest that particles' whose longer axis 
-lies in the plane perpendicular to the direction of propagation 
-should experience any notable degree of torque. And indeed this 
-appears to be the case for a number of experiments involving 
-nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.
+optical rotation seen in asymmetric dimers can be attributed to 
+the same spin-curl forces there are several questions that 
+cannot be explained purely by the spin-curl force.
 
-\subsection{Comparison of optical torques for Sphere and Dimers}
-It does raise the question of how come our simulations show that 
-despite the dimer being orientated so their long axis' align with 
-the parallel to the direction of propagation they still readily 
-rotate. You would not expect that the addition of a single additional 
-sphere should drastically adjust the torque especially if said sphere 
-is relatively small. However when we measured the optical torque 
-of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found the 
-exact opposite. In both cases we used the same trapping beam as used 
-for fig.~\ref{fig:rotation_vs_pol} but with a circularly polarised 
-beam. Both the sphere and dimer were rotated in the $x-z$ plane and 
-the all three components of the optical torque were recorded.
+\subsection{Optical torque differences}
+In their paper Grier \textit{et al} show that a 7 sphere cluster
+will rotate when trapped in circularly polarised light but a 
+homogenous sphere will have a slight curl to its trajectory. 
+They attribute this to the curl of the spin angular momentum
+generating a "Brownian vortex". But what is more interesting, 
+and something that appears unaddressed, is how the effects of 
+this vortex change based on the overall shape of the target
+particle.
+
+You would not expect that the addition of a single additional 
+sphere should drastically adjust the torque especially if said 
+sphere is relatively small. However when we measured the optical 
+torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we 
+found the exact opposite. In both cases we used the same trapping 
+beam as used for fig.~\ref{fig:rotation_vs_pol} but with a 
+circularly polarised beam. Both the sphere and dimer were rotated 
+in the $x-z$ plane and the all three components of the optical 
+torque were recorded.
 \begin{figure}[h!]
@@
 	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
-	\caption{Optical toque experienced by a dimer ($a_{I}/a_{II}=10$) and a 
-	single isotropic sphere. Both were rotated in the $x-z$ plane and the angle 
-	between $U_z$ and the beam axis gives the polar angle.The solid lines denote 
-	the torque experienced by the dimer whereas the dashed lines represent the 
-	torque experienced by the sphere.}
+	\caption{Optical toque experienced by a dimer ($a_{I}/a
+	{II}=10$) and a single isotropic sphere. Both were rotated 
+	in the $x-z$ plane and the angle between $U_z$ and the beam 
+	axis gives the polar angle.The solid lines denote the torque 
+	experienced by the dimer whereas the dashed lines represent 
+	the torque experienced by the sphere.}
 	\label{fig:sphere_dimer_torque}
@@
 trap by the gradient forces. The same cannot be said for the $z$-
-component of the optical torque; while this effect could still be 
-attributed to the spin-curl force, but it is clear that the internal 
-scattering between the two spheres has some unintended effects.
+component of the optical torque, which is non-zero even at a 
+$80^\circ$ angle. This cannot be simply explained via spin-curl 
+effects, as the 'Brownian vortex' should instead be driving the
+dimer around the beam axis - similar to how the 7 sphere cluster 
+behaved. What is clear that the mere combination of two spheres
+results in a unique behaviour that has not been previously 
+investigated.
 
-\subsection{Rotational frequency as a function of size ratio and orientation}
-Next we want to see how the rotational speed is influenced by 
-different sized dimers. Intuitively, you would expect that a 
-larger would experience a greater torque and therefore rotate 
-faster. By repeating the same kinds of simulation as used in 
-\ref{fig:rotation_vs_pol} but for a circularly polarised beam 
+\subsection{Rotational frequency as a function of size ratio and 
+	orientation}
+\label{sec:rot_size}	
+One factor t Intuitively, you would expect that a 
+larger particle would experience a greater torque and therefore 
+rotate faster. By repeating the same kinds of simulation as used 
+in \ref{fig:rotation_vs_pol} but for a circularly polarised beam 
 $\phi=90\circ$ it was found that not only is the rotation rate 
@@
   \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
-  \caption{Rotation rate plotted against dimer size ratio while trapped in a
-  	circularly polarised beam; a positive rotation rate indicates clockwise 
-  	rotation, whereas a negative rotation rate indicates counter-clockwise 
-  	rotation. The red line is for the case of a dimer in its 'standard' 
-  	orientation. The blue line is for the case when the dimer is in its 
-  	'inverted' orientation while trapped above the focus of the beam. And 
-  	lastly the green line is for the case when the dimer is in its 
-  	'inverted' orientation, but when it is trapped below the focus of the 
-  	beam.}
+  \caption{Rotation rate plotted against dimer size ratio while 
+  	trapped in a circularly polarised beam; a positive rotation 
+  	rate indicates clockwise rotation, whereas a negative 
+  	rotation rate indicates counter-clockwise rotation. The red 
+  	line is for the case of a dimer in its 'standard' orientation. 
+  	The blue line is for the case when the dimer is in its 
+  	'inverted' orientation while trapped above the focus of the 
+  	beam. And lastly the green line is for the case when the 
+  	dimer is in its 'inverted' orientation, but when it is trapped 
+  	below the focus of the beam.}
 \end{figure}
@@
 It is difficult to see from the graph, but the rotation rate never
-truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which would
-imply that a second sphere of radius $200\,{\rm nm}$ is enough to
-induce rotational motion. This brings into question what mechanism is
-generating the optical torque. We used \textit{mstm} to look at the stokes
-parameters from the scattered field from a simple plane wave incident
-on a symmetric dimer, the proportion of circularly polarised light is 
-minimal compared to the proportion of plane polarised light, which 
-indicates that this rotational motion is not due to any inhomogeneity 
-in the dimer that might impart angular momentum to the scattered beam - 
-as compared to a anisotropic scatterer like vaterite.
-
-These results are somewhat contradictory to other work with silica dimers
-\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
-have trapped the dimer in an orientation perpendicular to the beam
-propagation direction. The rotational motion is attributed to the
-dimers asymmetric geometry creating an unbalanced polarisation susceptibility
-along its long axis as compared to its short axis; therefore its long
-axis is aligned with the polarisation vector and can rotate
-freely \cite{Ahn2018}. This however cannot be the case with our
-simulations as the dimer rotates about its long axis, meaning there
-cannot be an asymmetric axis to align with the beam's polarisation
-vector. Furthermore, we see a non-linear increase in the rotational
-speed of our dimers with size, the drag torque from the surrounding
-fluid is $\propto r^3$ so the expectation is that the rotation
-frequency should fall off with increasing size. This indicates that
-the rotational motion is due to the shape asymmetry of the dimer and
-not solely due to the beam's angular momentum. Measurement of this
-photo-kinetic force is difficult to achieve due to the fact that
-previous analysis was conducted in the Rayleigh regime, where the
-polarizability of our dimer can be approximated as:
-\begin{align}
-  \bold{p}({\bf r},t)
-  =
-  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
-  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
-  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
-\end{align}
-
-where the polarizability is given as a 3D vector for the three
-principle Cartesian directions. In order to measure the magnitude of
-second order contributions we would need to construct a dipole array
-that fully captures the scattering of a dimer. Measuring the optical 
-torque makes it clear that the polarizability is a contributing factor 
-to this optical rotation phenomena.
+truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which 
+would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
+to induce rotational motion. What is also interesting is that the 
+rotation rate is not correlated directly with either the particle 
+size or the equilibrium position. This is in stark contrast to previous
+reports of dimer optical rotation; Ahn \textit{et al.} reported that
+the rotational frequency is maximised when the dimer is symmetric 
+\cite{Ahn2018}. Their work was conducted in a vacuum so the only 
+limiting factor to the optical rotational motion is the structural 
+stability of the particle itself.
 
 \subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity sized dimers
-there is the potential for non-vertical trapping orientations in which 
-the dimer is still stably trapped. When a circularly polarised beam is 
-used the dimer exhibits gyroscopic precession. As shown in fig.
-~\ref{fig:gyro} the dimer's trajectory exhibits periodic rotation about 
-its long axis $U_z$, precession motion as the dimer spins around the 
-central beam axis, and nutation motion as the dimer's orientation rocks 
-backwards and forwards. 
+As mentioned in section~\ref{sec:off-axis} for specificity sized 
+dimers there is the potential for non-vertical trapping orientations 
+in which the dimer is still stably trapped. When a circularly 
+polarised beam is used the dimer exhibits gyroscopic precession. 
+As shown in fig.~\ref{fig:gyro} the dimer's trajectory exhibits 
+periodic rotation, both around its long axis and about the beam 
+axis.
 \begin{figure}[h]
@@
 	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
-	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
-		off axis orientation with a circularly polarised beam ($P= 100\ mW$). 
-		The far left column depicts the dimer's centre of mass position with 
-		time; middle two columns are the $x$, $y$, and $z$ components of the 
-		vectors $u_x$ and $u_y$; last column depicts the components of the 
-		vector $u_z$ which defines the dimer's orientation.}
+	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped 
+		in an off axis orientation with a circularly polarised beam 
+		($P= 100\ mW$). The far left column depicts the dimer's centre 
+		of mass position with time; the remaining 3 columns show the 9
+		components of the dimers' rotation matrix, with each column being
+		associated with one of its three principal axis. }
 	\label{fig:gyro}
@@
 
-Applying a Fourier analysis to the above trajectory reveals the 3 fundamental
-frequencies typically associated with precession; the $u_{z,1}$ and $x(t)$ 
-series show a precession frequency of $~1.33\ Hz$ whereas the series $u_{x,1}$ 
-and $u_{y,1}$ show a combined periodic signal - a rotational frequency of 
-$~23\ Hz$ and a nutation frequency of $~20\ Hz$. Previous studies into 
-amorphous silica nanoparticles found a linear relationship with the 
-rotational frequency and the laser power, but no such relationship existed 
-with the precession frequency.
+Applying a Fourier analysis to the above trajectory reveals 
+the 3 fundamental frequencies typically associated with 
+gyroscopic precession. Fig.~\ref{fig:gyro_diagram} gives 
+a representative idea of the motion seen in fig.~\ref{fig:gyro}.
+Firstly there is precession, is denoted by $\psi$, this
+causes the dimer to rotate in the $XY$ plane while maintaining
+a constant polar angle. The precession can be due to the dimer 
+having inhomogeneous polarisation susceptibility. Since its 
+long axis is more susceptible to being polarised the dimer will
+try to align its long axis with the polarisation vector 
+\cite{Bruce2020}. Second there is nutation, denoted by $\theta$, 
+which is due to the dimer's Brownian motion being influenced by 
+the 'Brownian vortex', as predicted by \cite{Ruffner2012}. And 
+lastly there is rotation, denoted by $\omega$, where the dimer
+spins around its long axis as it did in \ref{sec:rot_pol} \&
+\ref{sec:rot_size}.
 
-This gyroscopic motion has been demonstrated previously in nanoparticles 
-\cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed 
-for micron scale aggregates. Since the torque applied to the dimer is 
-computed by evaluating the beam coefficients of the scattered field it is 
-difficult to apply this result to micro-rheology experiments as one would 
-need to know the exact magnitude of the optical torque ahead of time in 
-order to make estimations about the local fluid viscosity. This is trivial 
-for a birefringent spherical particle, less so for spherical aggregates 
-whose equilibrium position and orientation are unknown. However, further 
-analysis of the mechanism behind the precessive motion of off-axis dimers
-may provide insights into controlling Brownian motion. An experimental work
-trying to 'cool' nano-dimers by controlling the motion in all 6 degrees of 
-freedom found that even while the rotation about the short axis' could be 
-controlled the free rotation around the dimers' long axis resulted in an 
-unpredictable torsional vibration \cite{Bang2020}. Understanding how 
-rotational motion arises in the Mie-Regieme could allow researchers to 
-build a robust theoretical framework to construct beam structures that 
-eliminate any unwanted rotational motion from a target particle. Conversely, 
-the same framework could allow for precise measurements of the optical 
-torque applied to a target particle, allowing for characterisation of 
-complex shaped particles' interactions with an optical trap.
+What is interesting is that while we see the precession and 
+nutation also contribute to the rotation frequency. As can be 
+seen in both fig.~\ref{fig:gyro} and fig.~\ref{fig:gyro_diagram}
+the rotation matrix components $U_{x,x}$, $U_{x,y}$, $U_{y,x}$
+and $U_{y,y}$
 
+\begin{figure}
+	\hspace{-2.3cm}
+	\includegraphics[width=1.2\linewidth]{gyroscopic_diagram.pdf}
+	\caption{Representative diagram of the gyroscopic precession 
+	from fig.~\ref{fig:gyro}. The dimer has three principal angular
+	frequencies: The rotation ($\omega$) occurs around the 
+	the dimer's long axis. Precession ($\psi$) is seen where the dimer 
+	rotates around the beam axis. Nutation ($\theta$) is due to the dimer's 
+	centre of diffusion orbiting the beam axis. Shown on the right
+	is the power spectra from fig.~\ref{fig:gyro} with the associated 
+	frequencies labelled. The power spectrum have been zoomed in on
+	the relevant frequencies to highlight the precise values.}
+	\label{fig:gyro_diagram}
+\end{figure}
+
+This gyroscopic motion has been demonstrated previously in 
+nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
+but has not been observed for micron scale aggregates. What
+is interesting that the precession is seen around the particle's
+optical axis (in the case of a dimer this would be its long 
+axis). Fig.~\ref{fig:gyro} instead shows the precession 
+occurring perpendicular to the optical axis. In the former 
+case the precession can be explained as simply occurring due 
+to the reaction torque from the surrounding fluid \cite{Zhu2021}.
+In their analysis of ellipsoidal particles, Zhu \textit{et al.}
+found that when the particle rotated around their minimal inertial 
+axis (similar to the observed behaviour in dimers), the particle 
+would quickly 'stabilise' by aligning its' optical axis with the 
+beam axis. This highlights that a spherical dimer is capable of 
+rotating about its minimal inertial axis while remaining stably
+trapped. This has potential implications beyond simply 
+understanding particle dynamics.
+
+Further analysis of the mechanism behind the precession of off-axis 
+dimers may provide insights into controlling Brownian motion. An 
+experimental work trying to 'cool' nano-dimers by controlling the 
+motion in all 6 degrees of freedom found that even while the 
+rotation about the short axis' could be controlled the free rotation 
+around the dimers' long axis resulted in an unpredictable torsional 
+vibration \cite{Bang2020}. Understanding how rotational motion 
+arises in the Mie-regime could allow researchers to build a robust 
+theoretical framework to construct beam structures that eliminate any 
+unwanted rotational motion from a target particle. Conversely, the 
+same framework could allow for precise measurements of the optical 
+torque applied to a target particle, allowing for characterisation 
+of complex shaped particles' interactions with an optical trap. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Characterisation of asymmetric dimers via PSD analysis}
+\section{Characterisation of asymmetric dimers via PSD \\
+	analysis}
 As discussed in \ref{sec:simulated_QPD}, one of the methods 
@@
 the relevant dynamical information from a simulation, 
-confirming the same behaviour in an experimental setting can 
-be challenging if dealing with a non-birefringent anisotropic 
-scatterer. The QPD is composed of 4 photodiodes that measure
-the intensity of light incident on their surfaces. Using 
-\textit{ott} we can define a region in the far-field that is 
-analogues to the surface of the QPD surface. Evaluating the 
-electric field across this surface gives us an approximation
-of the QPD signal outputted in an experimental situation. This
-does not provide a one-to-one result however as hardware errors
-(i.e. internal resistance, external light sources, and vibrations)
-would distort the signal somewhat. We can use this as a means 
-to examine the limitations of using back-focal plane interferometry
-as a means for characterising anisotropic scatterers.
+confirming the same behaviour in an experimental setting 
+can be challenging if dealing with a non-birefringent 
+anisotropic scatterer. The QPD is composed of 4 photodiodes 
+that measure the intensity of light incident on their 
+surfaces. Using \textit{ott} we can define a region in the 
+far-field that is analogues to the surface of the QPD surface. 
+Evaluating the electric field across this surface gives us 
+an approximation of the QPD signal outputted in an experimental 
+situation. This does not provide a one-to-one result however 
+as hardware errors (i.e. internal resistance, external light 
+sources, and vibrations) would distort the signal somewhat. We 
+can use this as a means to examine the limitations of using 
+back-focal plane interferometry as a means for characterising 
+anisotropic scatterers.
 
-As a benchmark we start by considering a single sphere within an 
-optical trap. A single polystyrene sphere suspended in water 
+As a benchmark we start by considering a single sphere within 
+an optical trap. A single polystyrene sphere suspended in water 
 ($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
@@
 (Eq.~\eqref{eq:aliased_lorentzian}).
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
 	\includegraphics[width=\linewidth]{PSD_sphere.png}
-	\caption{Recorded power spectra fitted to eq.~\ref{eq:alaised_lorentzian},
-		 scattered points represents the blocked data ($n_b=100$). Corner 
-		 frequency for the Lorentzian curves are reported in the legend.}
+	\caption{Recorded power spectra fitted to eq.~
+		\ref{eq:alaised_lorentzian}, scattered points represents 
+		the blocked data ($n_b=100$). Corner frequency for the 
+		Lorentzian curves are reported in the legend.}
 	\label{fig:psd_sphere}
@@
 
-As shown in fig.~\ref{fig:psd_sphere}, the two power spectra report 
-different corner frequencies which would be indicate that the trap 
-is not perfectly circular. We can use both \textit{ott} and the 
-trajectory itself to derive an estimation of the trap geometry. 
-The corner frequencies and corresponding trap stiffness are 
-reported below:
-\newpage
+As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
+report different corner frequencies which would be indicate 
+that the trap is not perfectly circular. We can use both 
+\textit{ott} and the trajectory itself to derive an estimation 
+of the trap geometry. The corner frequencies and corresponding 
+trap stiffness are reported below:
 
@@
 
-Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, where
-the variance is based upon our choice of frequency boundaries
-({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And the ellipticity 
-of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a 
-measure of the symmetry of the beam wavefront. Its clear from these 
-initial results that the QPD is more sensitive to changes along the 
-y-axis than the x-axis when compared to the direct \textit{ott} 
-calculations. This is somewhat reflected in the trajectory results. 
-Typically, even an industrial Gaussian beam will produce an 
-elliptical diffraction limit spot when heavily focused; in their 
-tutorial for optimizing the PSD analysis, Berg and Sorensen reported 
-a ellipticity of around 15 \% after a total calibration time of 80 
-seconds \cite{BergSoerensen2004}.
+Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
+where the variance is based upon our choice of frequency 
+boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
+the ellipticity of the beam is given by $e = (1-\kappa_y/
+\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
+wavefront. Its clear from these initial results that the QPD 
+is more sensitive to changes along the y-axis than the x-axis 
+when compared to the direct \textit{ott} calculations. This is 
+somewhat reflected in the trajectory results. Typically, even 
+an industrial Gaussian beam will produce an elliptical diffraction 
+limit spot when heavily focused; in their tutorial for optimizing 
+the PSD analysis, Berg and Sorensen reported a ellipticity of 
+around 15 \% after a total calibration time of 80 seconds 
+\cite{BergSoerensen2004}.
 
-The reason for the discrepancies between all 3 methods is due to what 
-is actually being measured. The \textit{ott} estimates are simply 
-looking at the differences in the trapping strength along the Cartesian 
-axis'. If calibrated over an long enough time frame you would expect 
-that the resulting power spectra would exactly mirror the \textit{ott} 
-predictions. However over a short calibration time a QPD can predict
-large discrepancies in the trap strength. There is a clear trade off 
-in terms of accuracy and computation time as shorter calibration runs 
-are computationally more efficient but prone to errors. This is the case
-even if we know the exact positional data.
+The reason for the discrepancies between all 3 methods is due 
+to what is actually being measured. The \textit{ott} estimates 
+are simply looking at the differences in the trapping strength 
+along the Cartesian axis'. If calibrated over an long enough 
+time frame you would expect that the resulting power spectra 
+would exactly mirror the \textit{ott} predictions. However over 
+a short calibration time a QPD will only sample a small trajectory 
+that does not fully explore the trapping plane. There is a clear 
+trade off in terms of accuracy and computation time as shorter 
+calibration runs are computationally more efficient but prone to 
+errors. This is the case even if we know the exact positional data.
 
-With this in mind, let us consider a symmetric dimer that is optically
-trapped by the same Gaussian beam. Not only does the dimer's equilibrium
-position change but it is subjected to rotational motion due to the 
-beam's angular momentum. This is reflected in the calibration results
-using the simulated QPD, where we see a drastically different estimation
-between the \textit{ott} estimate and the QPD estimate.
+With this in mind, let us consider a symmetric dimer that is 
+optically trapped by the same Gaussian beam. Not only does 
+the dimer's equilibrium position change but it is subjected 
+to rotational motion due to its unequal moments of inertia. 
+This is reflected in the calibration results using the simulated 
+QPD, where we see a drastically different estimation between the 
+\textit{ott} estimate and the QPD estimate.
 
@@
 previous literature which raises questions on the interactions between
-circularly polarised light and spherical aggregates. Being made up of
-homogenous materials their is no change in polarisation as expected
-by birefringence. Shape birefringence is also not applicable as the 
-preferred orientation is perpendicular to the polarisation axis, 
-meaning there can be no asymmetry in the dimer's susceptibility.
+circularly polarised light and spherical aggregates. Until now, our 
+understanding of angular momentum transfer has been either focused on 
+single particles (where momentum transfer is easily described by 
+Lorenz-Mie theory) or for large aggregates of particles, where the 
+angular momentum transfer between individual particles is not considered. 
+This could provide a greater understanding of angular momentum and help in 
+the development of better torque sensing or torque preventing methods.
+
+Conventional calibration methods will not be adequate however, while
+angular motion can be detected using complicated set ups \cite{Bang2020},
+this can only provide an estimate of the magnitude of a particle's 
+angular motion. This is perfectly fine in cases where the rotational
+motion is stochastic and intuitively predictable (i.e. a dimer in a
+vertical orientation undergoing stochastic Brownian motion), but in 
+the case the the motion is instead periodic characterising it requires
+instantaneous measurements of the particle's motion.

### 2024-11-26 Latex_files/Thesis/chapters/chapter 2/main.tex
updates to chp 4 & 2

@@
 \label{chapter:theory}
-The two key areas of this project cover two widely different fields; 
-crystallisation theory, which is covered in the previous introduction 
-section, and tweezing theory. The following chapter summarises the 
-working principles behind optical tweezing, the differences in 
-scattering theories, and the broader theoretical calculations used
-throughout the rest of the Thesis.
+The two key areas of this project cover two widely different
+fields; crystallisation theory, which is covered in the 
+previous introduction section, and tweezing theory. The 
+following chapter summarises the working principles behind 
+optical tweezing, the differences in scattering theories, 
+and the broader theoretical calculations used throughout the 
+rest of the thesis.
 
@@
 
-Proper understanding of optical tweezing requires an understanding of
-how the trapped particle interacts with the trapping laser. From an 
-electromagnetism perspective the laser creates a spatially and 
-temporally coherent electric field that scatters light off of a 
-trapped particle.  The laws governing electric and magnetic fields 
-are summarised most succinctly via the Maxwell equations. The 
-differential forms of which are given below \cite{Jackson_1975}:
+Proper understanding of optical tweezing requires a description
+of how the trapped particle interacts with the trapping laser. 
+From an electromagnetism perspective the laser creates a 
+spatially and temporally coherent electric field that scatters 
+light off of a trapped particle. The laws governing electric 
+and magnetic fields are summarised most succinctly via the 
+Maxwell equations. The differential forms of which are given 
+below \cite{Jackson_1975}:
 \begin{align}
@@
 \end{align}
-where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
-current, and $\epsilon_0$ is the permittivity of free space.
-
-These four equations describe how the electric and magnetic fields behave
-and relate to one another at a microscopic level. Any discussion 
-of optical trapping is underpinned by the fact that in all cases the 
-Maxwell equations must be satisfied at the macroscopic level, where 
-one must account for the medium's interactions with the EM field. 
-Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
+where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
+charge current, and $\epsilon_0$ is the permittivity of free 
+space.
+
+These four equations describe how the electric and magnetic fields 
+behave and relate to one another at a microscopic level. Any 
+discussion of optical trapping is underpinned by the fact that in 
+all cases the Maxwell equations must be satisfied at the macroscopic 
+level, where one must account for the medium's interactions with the 
+EM field. Macroscopic forms of the Maxwell equations are given below
+\cite{Jackson_1975}:
 \begin{align}
@@
 \end{align}
-Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
-caused by the presence of free charges in the medium, and $\mathbf{D}$ 
-and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
-The latter two are defined by:
+Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and 
+current caused by the presence of free charges in the medium, 
+and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and 
+magnetising fields respectively. The latter two are defined 
+by:
 \begin{align}
@@
 
-Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
-magnetisation field, these fields arise due to the bound charges 
-throughout the medium interacting with the EM field.
-
-The force exerted by an optical tweezer can be subdivided into the 
-gradient and scattering components, for most modelling research this 
-is how the force fields are reported. The gradient force is a 
-conservative force brought about by the polarisation of dielectric 
-materials which is directed towards the point of maximum intensity 
-(for a simple Gaussian beam this would be at it's focal point) 
-\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise a
-higher numerical aperture lens in order to increase the intensity, 
-thus allowing for stronger gradient forces at the trade off of 
-decreased trapping depth.
-
-The scattering force arises from the scattered field of the trapping 
-beam pushing the target object away from the focal point. While some 
-trapping arrangements can induce transverse scattering forces for 
-simple spheres the scattering force is only significant in the direction
-of propagation \cite{Capitanio2002} The equilibrium position is found 
-when the gradient force far cancels out the the scattering force, 
-and the thermal energy of the particle.
+Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ 
+is the magnetisation field, these fields arise due to the 
+bound charges throughout the medium interacting with the EM 
+field.
+
+The force exerted by an optical tweezer can be subdivided into 
+the gradient and scattering components, for most modelling 
+research this is how the force fields are reported. The gradient 
+force is a conservative force brought about by the polarisation 
+of dielectric materials which is directed towards the point of 
+maximum intensity (for a simple Gaussian beam this would be at 
+it's focal point) \cite{YasuhiroHarada1996}. Typically, optical 
+tweezers will utilise a higher numerical aperture lens in order 
+to increase the intensity, thus allowing for stronger gradient 
+forces at the trade off of decreased trapping depth.
+
+The scattering force arises from the scattered field of the 
+trapping beam pushing the target object away from the focal 
+point. While some trapping arrangements can induce transverse 
+scattering forces for simple spheres the scattering force is 
+only significant in the direction of propagation \cite{Capitanio2002}.
+The equilibrium position is found when the gradient force 
+cancels out the the scattering force, and the thermal energy 
+of the particle.
 
@@
 \label{sec:harmonic_traps}
-When a sphere is located at the centre of a laser the gradient force
-will try to recentre the sphere if it is displaced. If the restoring 
-force is proportional to the sphere's displacement this is referred 
-to as a harmonic trap. The ratio of the restoring force to the 
-displacement is called the trap stiffness (denoted by $\kappa$), and 
-is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap}
-demonstrates an example case of the force-displacement curve for a 
-simple sphere in an optical trap. Laser power and numerical aperture
-are omitted as the exact shape is not of importance, only the linear 
-region about the origin.
+When a sphere is located at the centre of a laser the gradient 
+force will try to recentre the sphere if it is displaced. If 
+the restoring force is proportional to the sphere's displacement 
+this is referred to as a harmonic trap. The ratio of the restoring 
+force to the displacement is called the trap stiffness (denoted 
+by $\kappa$), and is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap} demonstrates an example case of the
+force-displacement curve for a simple sphere in an optical trap. 
+Laser power and numerical aperture are omitted as the exact shape 
+is not of importance, only the linear region about the origin.
 \begin{figure}[h!]
@@
 
-The strength of the optical trap is directly proportional to both 
-the laser power and the contrast in refractive indices between the 
-medium and the particle. Fully describing and calculating the 
-optical forces is dependent on the ratio between the particle size 
-($a$) and the trapping wavelength ($\lambda$). The scattering theory 
-for computing the optical force is dependent on this ratio.  
+The strength of the optical trap is directly proportional to 
+both the laser power and the contrast in refractive indices 
+between the medium and the particle. Fully describing the 
+optical forces is dependent on the ratio between the particle 
+size ($a$) and the trapping wavelength ($\lambda$). The 
+scattering theory that describes the optical force is dependent 
+on this ratio.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 on the beam in question the expansion coefficients can be tailored
-to match the desired beam shape. For a Gaussian beam interacting 
-with a particle of volume $V$, the incident, internal, and 
-scattered fields are given as \cite{RanhaNeves2019}: 
+to match the desired beam shape. For a approximate Gaussian beam 
+interacting with a particle of volume $V$, the incident, internal, 
+and scattered fields are given as \cite{RanhaNeves2019}: 
 \begin{align}
@@
 Farsund \cite{Farsund1996} but in the form used by Crichton 
-\cite{Crichton2000THEMD}. As an example the linear momentum 
-carried by any beam can be computed via:
+\cite{Crichton2000THEMD}. As an example the time averaged linear 
+momentum carried by any beam can be computed via:
 \begin{equation}
@@
 
-And the linear momentum carried by the scattered beam is given by:
+And the time averaged linear momentum carried by the scattered 
+beam is given by:
 \begin{equation}
@@
 \begin{align}
-	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
+	\boldsymbol{F_z} =\frac{dP}{dt} = \langle P_{z,\ in}\rangle 
+	-\langle P_{z,\ out}\rangle
 	\label{eq:Fz}
@@
 the wavelength of light - the theory describing the electromagnetic 
-fields are better easier to approximate by alternative theories. 
-When $a \gg \lambda$ the optical force is best described by the 
-Ray-Optics model. In the opposite case the target particle is 
+fields are easier to approximate by alternative theories. When the 
+particle radius is $\gg \lambda$ the optical force is best described 
+by the Ray-Optics model. In the opposite case the target particle is 
 better approximated as a single electric dipole as described by 
@@
 The Rayleigh approximation is for describing a particle when 
-$a \ll \lambda$. The underlying theory is that a dielectric sphere 
-can be treated as a dipole while in the presence of the 
-electromagnetic field. In which case the scattering force is 
-given simply by the scattering of the induced dipole, and the 
-gradient force is due to the Lorentz force \cite{Gordon1973}. 
-The gradient forces in the principle Cartesian axis are 
-described by Harada et al \cite{YasuhiroHarada1996} in MKS units 
-as a restorative rectangular force field:
+the particle radius is $\ll \lambda$. The underlying theory 
+is that a dielectric sphere can be treated as a dipole while 
+in the presence of the electromagnetic field. In which case 
+the scattering force is given simply by the scattering of 
+the induced dipole, and the gradient force is due to the 
+Lorentz force \cite{Gordon1973}. The gradient forces in the 
+principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
+rectangular force field:
 \begin{align}
@@
 the scattered field is now significant. As the target particle gets 
-larger this assumption fails to accurately describe the trapping force 
+larger this approximation fails to accurately describe the trapping force 
 \cite{Li2021}. For larger particles the optical force and scattering 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Scattering methods}
@@
 simplifying the scattering problem. While all of these methods will
-produce identical results the key difference is in how one represents
+produce similar results the key difference is in how one represents
 the scatterer.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{T-matrix Method}
-
 The $T$-matrix method was first developed by Peter Waterman with 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Discrete Dipole Approximation}
@@
 being constructed of discrete dipoles. This is essentially an 
-extension of the Rayleigh Regime, though in order to compute the 
-optical force one requires a full description of the electric 
+extension of the Rayleigh approximation, though in order to compute 
+the optical force one requires a full description of the electric 
 field. Each dipole interacts with both the incident field and the 
@@
 the T-matrix of the target. 
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Describing the trajectory of a trapped particle}
+While describing the optical forces is sufficient for 
+theoretical characterisation of an optical trap, trying 
+to replicate that model in an experimental setting is 
+near impossible. As such it is often far simpler to 
+characterise the optical trap by recording the particle's
+trajectory (see \ref{sec:position_detection} for 
+specifics regarding the equipment), and then fitting it 
+to our presumed description of the beam. The trajectory 
+is easily understood if we understand the particle's 
+diffusive behaviour.
+
 \section{Langevin Equation}
-Describing any microscopic motion requires an understanding of a 
-molecules diffusive behaviour, for the case of optical tweezers 
-the most complete model of diffusion is the Langevin equation. 
-Models such as the Fickian, and Einstein derivations are sufficient 
-for macroscopic behaviours the Langevin equation better describes 
-the microscopic characteristics of any diffusive behaviour.
-
-The Fickian model describes the net flux of solute molecules into a 
-finite volume of fluid being proportional to the density gradient 
-$\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
-solute molecules do not collide with one another or with molecules in 
-the solution; overall, the Fickian model is used to describe how solute 
-molecules disperse over long periods of time. It does not provide any 
-insight into the forces acting on individual molecules, whereas the 
-Langevin model captures the physical interactions between an individual 
-molecule and the surrounding fluid over a wide range of time scales.
-
-The Einstein model of diffusion expands upon the Fickian model by 
-considering the collisions between individual molecules. If we 
-consider a single particle suspended in a solution it will experience 
-multiple collisions with other molecules \cite{Gillespie2012a} over a 
-given time interval $\delta t$. The Einstein model allows us to 
-consider how individual molecules behave while suspended in a 
-surrounding fluid. Where it begins to falter is when we consider 
-bringing $\delta t$ to $0$; in this scenario the molecule experiences 
-not many but only individual collisions. The Einstein model does not 
-consider the inertia of the molecule in question and so for very short 
-time frames the model implies that the molecule's velocity changes 
-instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
-Furthermore, the kinetic energy of each collision is not limited by 
-the thermal energy of the system, meaning that using the Einstein model 
-to predict a particle's trajectory to an finite degree of accuracy 
-implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
-This failure to describe motion over smaller time frames is addressed 
-in the Langevin model by accounting for the fluid drag of the system, 
-where any sudden change in velocity must result in a proportionally 
-opposed drag force \cite{Gillespie2012c}.
-
-The Langevin model of diffusion assumes that the net force on a 
-particular particle is described fully by these individual 
+Describing any microscopic motion requires an understanding 
+of a molecules diffusive behaviour, for the case of optical 
+tweezers the most complete model of diffusion is the Langevin 
+equation. Models such as the Fickian, and Einstein derivations 
+are sufficient for macroscopic behaviours the Langevin 
+equation better describes the microscopic characteristics of 
+any diffusive behaviour.
+
+The Fickian model describes the net flux of solute molecules 
+into a finite volume of fluid being proportional to the 
+density gradient $\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. 
+It assumes that the solute molecules do not collide with one 
+another or with molecules in the solution; overall, the Fickian 
+model is used to describe how solute molecules disperse over 
+long periods of time. It does not provide any insight into 
+the forces acting on individual molecules, nor does it prove 
+particularly useful over smaller time scales. Whereas the Langevin 
+model captures the physical interactions between an individual 
+molecule and the surrounding fluid over a wide range of time 
+scales.
+
+The Einstein model of diffusion expands upon the Fickian model 
+by considering the collisions between individual molecules. If 
+we consider a single particle suspended in a solution it will 
+experience multiple collisions with other molecules 
+\cite{Gillespie2012a} over a given time interval $\delta t$. 
+Applying statistical analysis to an observed trajectory allows
+one to make measurements of a particle's diffusive behaviour 
+within a fluid. Where it begins to falter is when we consider 
+bringing $\delta t$ to $0$; in this scenario the molecule 
+experiences not many but only individual collisions. The 
+Einstein model does not consider the inertia of the molecule 
+in question and so for very short time frames the model 
+implies that the molecule's velocity changes instantly after 
+each collision \cite{Gillespie2012a, Gillespie2012b}. 
+Furthermore, the kinetic energy of each collision is not 
+limited by the thermal energy of the system, meaning that 
+using the Einstein model to predict a particle's trajectory 
+to an finite degree of accuracy implies that the particle 
+has infinite kinetic energy \cite{Gillespie2012b}. This 
+failure to describe motion over smaller time frames is 
+addressed in the Langevin model by accounting for the fluid 
+drag of the system, where any sudden change in velocity must 
+result in a proportionally opposed drag force \cite{Gillespie2012c}.
+
+The Langevin model of diffusion assumes that the net force on 
+a particular particle is described fully by these individual 
 collisions \cite{Gillespie2012c}. Unlike the Fickian model it 
-provides a full description of the interactions between the target 
-molecule and the surrounding fluid; while also being able to 
-describe its motion over any time scale, unlike the Einstein model:
+provides a full description of the interactions between the 
+target molecule and the surrounding fluid; while also being 
+able to describe its motion over any time scale, unlike the 
+Einstein model:
 \begin{align}
@@
 
-Where the first term accounts for inertial forces, the second term 
-accounts for friction forces which counteract the particles current 
-motion ($\gamma_0$ is the friction coefficient), and the final term 
-accounts for the random Brownian force. The $F(t)$ is there for 
-convention which accounts for any external forces acting on the 
-particle. These three terms are equal to the random Brownian motion 
-of the surrounding fluid. We can say that the noise term $W(t)$ has 
-a normal distribution, being scaled by the thermal energy of the 
-system, with a correlation function of:
+Where the first term accounts for inertial forces, the second 
+term accounts for friction forces which counteract the particles 
+current motion ($\gamma_0$ is the friction coefficient), and 
+the final term accounts for the random Brownian force. The 
+$F(t)$ is there for convention which accounts for any external 
+forces acting on the particle. These three terms are equal to 
+the random Brownian motion of the surrounding fluid. We can 
+say that the noise term $W(t)$ has a normal distribution, 
+being scaled by the thermal energy of the system, with a 
+correlation function of:
 \begin{align}
@@
 
-The Langevin model can be extrapolated to describe the diffusive behaviour 
-of an overall system, but for this project we can instead consider the 
-behaviour of some particle with a diffusion tensor $D$ suspended in a 
-viscous fluid and spacially localised by an optical potential with trap 
-strength $\kappa$. Assuming the only external force acting on our particle 
-is the laser, the net force should be exactly equal to force of the stochastic
-collisions due to the fluids thermal energy. If we focus our analysis when 
-the particle is stably trapped and assume that the trap is harmonic, we can 
-model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
-The full Langevin equation for an optically trapped particle is therefore given as:
+The Langevin model can be extrapolated to describe the diffusive 
+behaviour of an overall system, but for this project we can 
+instead consider the behaviour of some particle with a diffusion 
+tensor $D$ suspended in a viscous fluid and spacially localised 
+by an optical potential with trap strength $\kappa$. Assuming the 
+only external force acting on our particle is the laser, the net 
+force should be exactly equal to force of the stochastic collisions 
+due to the fluids thermal energy. If we focus our analysis when 
+the particle is stably trapped and assume that the trap is harmonic, 
+we can model the trapping force as a Hookean spring ($F(t) \approx 
+\kappa x(t)$). The full Langevin equation for an optically trapped 
+particle is therefore given as:
 \begin{align}
@@
 
-\eqref{eq:langevin} provides an accurate description of strongly trapped 
-particles. Despite this the analytical solution of the Langevin equation requires 
-integration of the noise term making it difficult to simulate the 
-trajectory of a given particle \cite{Volpe2013}. Instead, it is often far 
-easier to solve the equation numerically and use the analytical 
-solution to calibrate and extract information about the particle and fluid, 
-and how the two interact with the optical trap.
+\eqref{eq:langevin} provides an accurate description of strongly 
+trapped particles. Despite this the analytical solution of the 
+Langevin equation requires integration of the noise term making 
+it difficult to simulate the trajectory of a given particle 
+\cite{Volpe2013}. Instead, it is often far easier to solve the 
+equation numerically and use the analytical solution to calibrate 
+and extract information about the particle and fluid, and how the 
+two interact with the optical trap.
 
 \subsection{Finite Difference}
-The Finite Difference approach involves discretizing the time and spatial 
-elements in order to approximate the higher order terms. If we assume that 
-$x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then 
-we can use the Taylor series expansion to get:
+The Finite Difference approach involves discretizing the time and 
+spatial elements in order to approximate the higher order terms. 
+If we assume that $x(t)$ is differentiable to n (we can find its 
+$n^{th}$ derivative) then we can use the Taylor series expansion 
+to get:
 \begin{align}
@@
 
-Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
-term n and the actual expression. If we limit our approach to the first 
-derivative only, we find that for sufficiently small values of $R_1$ the 
-velocity and acceleration can be approximated as:
+Where $R_n(x(t))$ is the remainder term between the Taylor 
+expansion to term n and the actual expression. If we limit our 
+approach to the first derivative only, we find that for 
+sufficiently small values of $R_1$ the velocity and acceleration 
+can be approximated as:
 \begin{align}
@@
 
-By reversing the time step (i.e. use $-\Delta t$) to approximate the
-velocity and acceleration based on the previous time steps, we can 
-discretise the position by taking finitely small  time steps (i.e. 
-$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
-noise, as no information is known about $W(t)$ at any time. We 
-can instead say that the velocity of a Brownian particle should 
-approximate our noise as a random walk, where at each new time 
-step, the position changes randomly within a given range. Constricting 
-the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
-noise using the finite-difference approach as:
+By reversing the time step (i.e. use $-\Delta t$) to approximate 
+the velocity and acceleration based on the previous time steps, 
+we can discretise the position by taking finitely small  time 
+steps (i.e. $x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same 
+cannot be done for noise, as no information is known about $W(t)$ 
+at any time. We can instead say that the velocity of a Brownian 
+particle should approximate our noise as a random walk, where at 
+each new time step, the position changes randomly within a given 
+range. Constricting the variance to $\sqrt{2D}/\Delta t$ allows 
+us to represent the noise using the finite-difference approach as:
 \begin{align}
@@
 
-Where $w_i$ is a random real number between -1 and 1, we can say that 
-it is normally distributed around 0 for simulation purposes. We can 
-rearrange this for $x_i$ to approximate the Brownian motion of a 
-particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
-Now in the case of an optical trap, the restoration time scale is 
-given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
-is far greater than the characteristic time. Therefore for simulation 
-purposes, we can neglect the particle's inertia which allows the motion 
-of an optically trapped particle to be written as:
+Where $w_i$ is a random real number between -1 and 1, we can say 
+that it is normally distributed around 0 for simulation purposes. 
+We can rearrange this for $x_i$ to approximate the Brownian motion 
+of a particle (setting $x_0=0$), where the characteristic time is 
+$\tau = m/\gamma$. Now in the case of an optical trap, the 
+restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ 
+which for strongly trapped particles is far greater than the 
+characteristic time. Therefore for simulation purposes, we can 
+neglect the particle's inertia which allows the motion of an 
+optically trapped particle to be written as:
 \begin{align}
@@
 
-This result can be generalised for a 3-dimensional description of an 
-optically trapped particle, where each Cartesian direction has its 
-own unique characteristic restoration time. We see from the result 
-that trajectory is dependent on only a handful of factors, the trap 
-stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal 
-energy of the system $k_BT$ (with the latter two being related by 
-Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$). 
-Therefore, by calculating these parameters to a high degree of precision 
-allows one to get a precise description of the forces experienced by a 
-target particle, which in the past has been used for highly accurate 
-force transduction \cite{BergSoerensen2004, Smith2003}.
+This result can be generalised for a 3-dimensional description of 
+an optically trapped particle, where each Cartesian direction has 
+its own unique characteristic restoration time. We see from the 
+result that trajectory is dependent on only a handful of factors, 
+the trap stiffness $\kappa_x$, the fluid viscosity $\gamma$, and 
+the thermal energy of the system $k_BT$ (with the latter two being 
+related by Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$). Therefore, by calculating these parameters to a high 
+degree of precision allows one to get a precise description of the 
+forces experienced by a target particle, which in the past has been 
+used for highly accurate force transduction \cite{BergSoerensen2004, Smith2003}.
 \begin{SCfigure}[][h!]
@@
 \label{sec:calibration_techniques}
-There are several approaches for calibrating and characterizing the 
-optical trap, each approach has its drawbacks and benefits so each 
-option should be chosen based on what elements want to be characterized. 
-The basis for each of these methods stems from the analytical solution 
-of the Langevin equation:
+There are several approaches for calibrating and characterizing 
+the optical trap, each approach has its drawbacks and benefits 
+so each option should be chosen based on what elements want to 
+be characterized. The basis for each of these methods stems from 
+the analytical solution of the Langevin equation:
 \begin{align}
@@
 
-Positional data is often acquired in an experimental setting either using
-image analysis or photodiodes to infer the particles position relative to
-the beam focus. The former is often used in cases when precision is not a 
-key concern, as often a standard CCD camera will have limited spacial and 
-temporal resolution. The latter method often requires the use of a quadrant
-photo diode(QPD), the data provided by the QPD will often need to be converted
-to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.
+Positional data is often acquired in an experimental setting 
+either using image analysis or photodiodes to infer the particles 
+position relative tothe beam focus. The former is often used in 
+cases when precision is not a key concern, as often a standard 
+CCD camera will have limited spacial and temporal resolution. 
+The latter method often requires the use of a quadrant photo 
+diode(QPD), the data provided by the QPD will often need to be 
+converted to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.
 
 \subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the trap 
-acts similar to a Hookean spring that creates a potential well about 
-its centre. Therefore a simple analysis method is to understand the 
-height and width of said potential well.
-
-Potential analysis is a useful technique for estimating the strength 
-of an optical trap; this method assumes that the force acting on the 
-particle is purely conservative, an accurate presupposition if we 
-ignore the motion of the particle as it enters the trap. This is 
-because the scattering force is far more significant far away from 
-the potential well and is negligible if the trap strength is much 
-greater than the thermal fluctuations. With this in mind we can 
-write the probability of finding the particle at position $x$ as:
+The Langevin equation for an optically trapped assumes that the 
+trap acts similar to a Hookean spring that creates a potential 
+well about its centre. Therefore a simple analysis method is to 
+understand the height and width of said potential well.
+
+Potential analysis is a useful technique for estimating the 
+strength of an optical trap; this method assumes that the force 
+acting on the particle is purely conservative, an accurate 
+presupposition if we ignore the motion of the particle as it 
+enters the trap. This is because the scattering force is far 
+more significant far away from the potential well and is 
+negligible if the trap strength is much greater than the thermal 
+fluctuations. With this in mind we can write the probability of 
+finding the particle at position $x$ as:
 \begin{align}
@@
 \subsubsection{Angular Mean Square Displacement (MSAD)}
-It is also possible to plot the angular MSD (MSAD) using simulative data, 
-however using this to describe the torque acting on an asymmetric object 
-is difficult due to fact there is no simple approximation of the rotational
-stiffness. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper 
-limit of a dimer's MSAD along its long axis by assuming it was strongly 
-trapped and so had limited angular motion, there expression gives:
+It is also possible to plot the angular MSD (MSAD) using 
+simulative or experimental data, but there is no complementary 
+expression for \eqref{eq:MSD}. This is because while the 
+particle is close to the focus the optical force is easily 
+modelled as a Hookean spring; conversely, the optical torque
+is far more complex even if the particle is strongly trapped. 
+Vigilante \textit{et al} \cite{Vigilante2020} derived the 
+upper limit of a dimer's MSAD along its long axis by assuming 
+it was strongly trapped and so had limited angular motion, 
+there expression gives:
 \begin{align}
@@
 
-Where, $u_z$ is the unit vector connecting the two spheres of the dimer, 
-$\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness of the trap, and $F$ 
-is Dawson's integral \cite{Oldham2008}. Vigilante's paper expressed that 
-they couldn't compute $MSD(\tau)$ because they couldn't solve the 
-Einstein-Smoluchowski equation which describes the diffusion constant for 
-dielectric particles. This would require a full description of a 
-particle's electrical mobility and charge distribution - the latter could 
-be achieved via a discrete dipole approximation, the former would be 
-dependent on both the particle's position and relative orientation to the 
-electric field.
+Where, $u_z$ is the unit vector connecting the two spheres of 
+the dimer, $\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness 
+of the trap, and $F$ is Dawson's integral \cite{Oldham2008}. 
+Vigilante's paper expressed that they couldn't compute 
+$MSAD(\tau)$ because they couldn't solve the Einstein-Smoluchowski 
+equation which describes the diffusion constant for dielectric 
+particles. This would require a full description of a particle's 
+electrical mobility and charge distribution - the latter could be 
+achieved via a discrete dipole approximation, the former would be 
+dependent on both the particle's position and relative orientation 
+to the electric field.
 
@@
 physical units: If multiple photodetectors are available then a 
-differential interference contrast (DIC) system can be used to compute
-the linear relationship between the beads displacement and the photodiodes
-signal \cite{Capitanio2002}. While this is useful when high force precision 
-is necessary it also allows you to collect information about the particle's 
-motion along a single direction, making it ideal for less focused beams. 
-Alternatively, if the size distribution is very wide but each particle can be 
-accurately sized, then a conversion factor can be approximated by comparing 
-the fitted value of the diffusion coefficient, and the reported value 
-given by the Stokes-Einstein relation.
+differential interference contrast (DIC) system can be used to 
+compute the linear relationship between the beads displacement and 
+the photodiodes signal \cite{Capitanio2002}. While this is useful 
+when high force precision is necessary it also allows you to collect information about the particle's motion along a single direction, 
+making it ideal for less focused beams. Alternatively, if the size distribution is very wide but each particle can be accurately sized, 
+then a conversion factor can be approximated by comparing the fitted 
+value of the diffusion coefficient, and the reported value given by 
+the Stokes-Einstein relation.
 \begin{align}
@@
 The scattering from rotating objects can be partially characterised via
-the power spectrum density method. Typically papers reporting on the 
-rotation rate of a trapped object will collect a power spectrum and look
-at the maximum frequency term to determine its rotation rate. However this
-often neglects any information on the trapping forces acting on particle
-and only looks at the torque applying around one of the particle's primary 
-axis. There has only been one notable effort to fully characterise the 
-full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
-a theoretical model that relates the power spectrum to both the 
-rotational and translational motion simultaneously.
+the power spectrum density method. If the particle is birefringent then
+as it rotates the detected signal will oscillate periodically. When a 
+Fourier transform is taken the power spectra will have a peak signal at 
+that periodic signal.Typically papers reporting on the rotation rate of 
+a trapped object will collect a power spectrum and look at the maximum frequency term to determine its rotation rate. However this often neglects 
+any information on the trapping forces acting on particle and only looks 
+at the torque applying around one of the particle's primary axis. There 
+has only been one notable effort to fully characterise the full trapping dynamics on a rotating body, \cite{Yogesha2012} developed a theoretical 
+model that relates the power spectrum to both the rotational and 
+translational motion simultaneously. This is applicable to particles with
+a polarizable long axis but is insufficient for birefringent particles.
 
@@
 motion is generalised to a 3 dimensional case. Here we omit the inertia term 
-as the restoration time of the optical trap is far greater than the characteristic
-time of the particle ($\kappa/\gamma \gg m/\gamma$). Now, the dimer 
-is undergoing random rotational motion in addition to its Brownian 
+as the restoration time of the optical trap is far greater than the characteristic time of the particle ($\kappa/\gamma \gg m/\gamma$).
+Now, the dimer is undergoing random rotational motion in addition to its Brownian
 translational motion the first term on the right hand side is no longer 
@@
 The rotational form of the Langevin equation for a dipole within an external 
-potential is given as:
+potential is given as \cite{Mazo2008}:
 \begin{align}
@@
   \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
-  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
-\end{align}
-
-Where $\vec{u}(t)$ is the unit vector aligned along the centres of the 
-two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational 
-friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a 
-sphere, if the dimer is within a harmonic potential we can write the first 
-term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times 
-\vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. 
-$\lambda(t)$ is the Brownian rotations from the surrounding fluid, 
-like in the translational case the Brownian rotations are normally 
-distributed and are also uncorrelated so that:
+  + \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of 
+the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the 
+rotational friction coefficient which is given as $\gamma_R = 8\pi
+\eta r^3$ for a sphere. If we assume that over the course of a 
+trajectory the dimer has a 'equilibrium' orientation in which the 
+gradient force is maximised we can say that the first right hand 
+term can be replaced with $\vec{\kappa}_u(\vec{r}(t))$ to represent
+a harmonic potential. $\lambda(t)$ is the Brownian rotations from the 
+surrounding fluid, like in the translational case the Brownian 
+rotations are normally distributed and are also uncorrelated so that:
 \begin{align}
@@
   &=
-    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
+    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2k_BT\vec{D}}\eta(t)
   \\
@@
     \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
-    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
+    + \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
 \end{align}
@@
 slope in order to allow for axial position detection, the method 
-for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.
+for finding this angle $\theta_\Theta$ is discussed in 
+\cite{Friedrich2012}.
 
@@
 
-While translational motion have no knock on effects to the QPD signal, 
-rotational changes are between the incident and total fields. This results
-in rotational motion being biased in the QPD signal - even when collecting
-signals from isotropic scatterers. To remedy this an inverse rotation is
-applied to the total field.
+While translational motion has no knock on effects to the QPD signal, 
+rotational changes are double counted by both the incident and total 
+field. This results in rotational motion being biased in the QPD signal 
+- even when collecting signals from isotropic scatterers. To prevent 
+this we need to rotate the total field via the inverse rotation matrix
+of the dimer. Shown in fig, is a flow diagram outlining the full work
+flow of a given trajectory, and how the QPD data is acquired.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.75\linewidth]{Simulation_workflow.pdf}
+	\caption{Block flow diagram of how a trajectory and QPD data is 
+		calculated and collected. Relevant equations are mentioned 
+		at each step.}
+\end{figure}
+\newpage
 

### 2024-11-26 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to chp 4 & 2

@@
 
+\subsection{QPD for angular displacement detection}
+
+The next logical step is to consider whether or not a QPD can at 
+all be utilised for detection of rotational motion. This has been
+shown for nano-particles that exhibit periodic rotational motion
+using by considering the difference in diagonal quadrants. The 
+benefit of detecting rotational motion is that one can begin to 
+build a better understanding about how angular momentum is 
+transferred to a dimer which could be extended to more complex
+particles.
+
+At just a cursory glance it would appear that there is a clear
+relationship between the QPD signal and the orientation of the 
+dimer. If we look at the QPD signal produced by a dimer is 
+rotated
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-01-26 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

@@
 Crystallisation is one of the oldest forms of phase 
-separation used by humanity\cite{Schoen1956}, put 
+separation used by humanity \cite{Schoen1956}, put 
 simply, it is the formation and growth of a new 
@@
 In this chapter we will outline our current understanding 
-of nucleation,why there is a gap in the literature, and why 
+of nucleation, why there is a gap in the literature, and why 
 there is still a need for local control of nucleation events. 
@@
 \subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT that suggests 
-that prior to nucleation, the solute will cluster together. This has 
-been shown via colloid simulations using protein molecules that have 
-significant short range interactions \cite{Wolde1997, Gliko2005}. These 
-interactions allow for the formation of a liquid-liquid metastable phase 
-from which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
-Several papers later reported the presence of stable liquid-like clusters 
-that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}. 
-The formation of these clusters can be understood by Oswald's rule; which 
-says that any crystallising system does not immediately take the path to 
-the lowest possible energy state but instead first transitions to the state 
-with the smallest free energy barrier \cite{Ostwald1897}. Further phase 
-transitions can still occur but this pathway minimises the overall free 
-energy cost. Initially it was suspected that the formation of the dense 
-liquid phase was a result of stochastic density fluctuations in the system. 
-However, simulations suggest that the the local bond order is a stronger 
-driving force than the local density \cite{Tan2013}. Studies into two step 
-nucleation in Glycine solutions have noted that the solutions need to be 
-aged before the presence of clusters was detected \cite{Liao2022}. This 
-suggests that there exist further barriers to the formation of clusters 
-in some systems, this has lead to instead calling this phenomena as 
-Multi-step Nucleation.
+The two step nucleation theory is an extension to the CNT 
+that suggests that prior to nucleation, the solute will 
+form precursor structures. The CNT assumes that solute 
+structures below the critical radius should be unstable 
+within the solution due to interfacial forces trying to 
+dissolve the solute. This assumption is based on the idea 
+that any interactions between solute molecules are much
+smaller than the surface tension due to the bulk solution.
+If instead the solute molecule interactions are significant
+then it stands to reason that subcritical liquid-like 
+structures are stable.
+
+The first observation of a liquid-like structure was in 
+$CaCO_3$ solutions, initially it was assumed that polymers
+were a necessary additive to create a precursor structure 
+\cite{Driessche2017, Karthika2016}. This is not the case when 
+the solution is above its critical temperature, allowing the 
+solution to be in the binodal region of the phase diagram. 
+This would result in the formation of two liquid phases, 
+one being rich in solute, the other being solute poor 
+\cite{Karthika2016, Fu2021}. Several papers later reported 
+the presence of stable liquid-like clusters that formed 
+prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{oswalds_rule.jpg}
+	\caption{Free energy diagram of Oswald's rule for a 
+	crystal with three possible polymorphs. The diagram 
+	shows that there exist local minima in the free energy 
+	that represent the different polymorphic forms. A similar
+	free energy diagram can be used to describe the two-step
+	nucleation mechanism.}
+\end{figure}
+
+The formation of these clusters can be understood by Oswald's 
+rule; which says that any crystallising system does not 
+immediately take the path to the lowest possible energy state 
+but instead first transitions to the state with the smallest 
+free energy barrier \cite{Ostwald1897}. Oswald's rule was 
+assumed to be applicable for crystal polymorphs but can be
+generalised to describe the nucleation process also. Further phase 
+transitions can still occur but the pathway taken should 
+always minimises the overall free energy cost. It is assumed
+that between the amorphous phase and the crystalline phase 
+the solute undergoes dehydration while changing structurally 
+into a more ordered phase, though there is some discussion 
+whether the process merely undergoes dehydration, or if there 
+is a more complex chemical reaction
+\cite{Karthika2016}.
+
+With regards to liquid-precursors, there is still a lot of
+discussion about the relationship between the solution 
+conditions and the the precursors. The size and rate at 
+which these precursors agglomerate based on the degree of 
+saturation would provide invaluable information on the 
+kinetics of two-step nucleation \cite{Fu2021}. In some 
+cases it has even been shown that 'ageing' solutions is 
+a necessary step in order to detect the presence of 
+stable amorphous clusters \cite{Liao2022}. This was in 
+solutions that were supersaturated and yet after the 
+formation of precursors spontaneous nucleation was not
+seen even after several days of observation \cite{Liao2022}.
+This suggests that in some cases the precursors are
+metastable and require an external trigger to induce
+crystallisation. The presence of metastable structures
+and other phenomena has lead to development of a more 
+universal theory, namely non-classical nucleation.
 
 \subsection{Non-classical Nucleation} 
-The current research into two step nucleation (or now more commonly 
-referred to as multi-step nucleation theory) is developing a robust 
-framework to describe what nucleation pathway will occur given the 
-initial conditions. Reviews of all currently documented nucleation 
-pathways highlighted the need for the development of \textit{in situ} 
-techniques that can induce nucleation locally but can also reliably 
-identify the nucleation pathway across a broad range of experimental 
-conditions \cite{Karthika2016, Fu2021}.
+Two-step nucleation is based on the idea that between
+the bulk solution and the final crystalline phase there
+exists and intermediate stage. While this is broadly an
+accepted concept, research has shown that in some cases
+the intermediate stage can be subdivided into multiple 
+states. This has been shown to be the case with metals
+in both solid and liquid conditions \cite{Cao2020, Ye2023}
+and colloidal simulations have indicated that the 
+precursors undergo structural change over time \cite{Tan2013}. 
+While in some cases it has been possible to directly
+monitor the structural change in the amorphous clusters, 
+many organic compounds have not received such treatment
+due to the difficulty of characterising and monitoring 
+the pre nucleation stage \cite{Fu2021}. This is often 
+why two-step nucleation is simply referred to as 
+multi-step nucleation.
+
+There are also a myriad of other unique nucleation events
+that deviate significantly from both the classical and 
+two-step nucleation theories. A review paper discussing
+recent advances in non-classical crystallisation highlighted
+that depending on the collision kinetics of the system 
+and the molecular interactions could result in drastically
+different reaction kinetics. In the former case, unstable
+subcritical nuclei have been shown to collide and coalesce
+into a supercritical nucleus, but only in conditions where 
+subcritical nucleation are plentiful despite a low degree
+of supersaturation \cite{Baumgartner2013}. The latter case 
+is more prevalent for dipole solute molecules, the anisotropic 
+attraction results in nuclei with non-spherical geometries, 
+this results in varying kinetics for agglomeration dependent 
+on the orientation of each molecule \cite{Yau2001}. Phenomena 
+such as the above examples has been collectively refereed to 
+as non classical nucleation, it is known that they must be 
+related to both CNT and multi-step nucleation but the connection 
+is not clear based on the current literature.
+
+As such, much of the research into nucleation theory is 
+focused on developing \textit{in situ} techniques that can 
+monitor and characterise the pre nucleation stage, this
+would glean information on cases when both classical 
+and non-classical nucleation is a possibility 
+\cite{Karthika2016}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Crystallisation methods}
-\subsection{Cooling Crystallisation}
-For some binary mixtures the supersaturation is heavily dependent 
-on the solution temperature, therefore a simple method of producing 
-crystals is by cooling the mixture to induce crystal formation. At 
-ambient temperatures the solute concentration is too high to be 
-fully incorporated into the solution ($S\gg 1$), after heating however 
-the solute is fully dissolved into the solution ($S<1$). Now as the 
-mixture is allowed to cool to room temperature the supersaturation 
-will increase until crystal formation begins, the rate of cooling 
-drastically influences the size and number of crystals produced.
+\section{In situ techniques for studying nucleation}
+There are several different phenomena \cite{Fu2021, Karthika2016} 
+that are not well described by any individual theoretical framework 
+(CNT or multi-step nucleation for example). There is an increasing 
+interest in probing the pre-nucleated solution.
 
-If $dT/dt$ is high then the final product will consist of large 
-crystal and be low in number, as the nucleation rate is directly 
-related to the supersaturation only a handful of nuclei can form 
-before the remaining solute grows onto the surface. If $dT/dt$ is 
-low then the final product will consist of smaller crystals and be 
-far more numerous, as the supersaturation is so large that the new 
-nuclei are forming continuously. Between these two extremes, one 
-can define a meta-stable zone, a region in the temperature-
-concentration phase space where both nucleation and crystal growth
-can be reliably controlled. The lower limit being given by the 
-solubility curve of the binary mixture, and the upper limit being 
-defined by the point where crystals begin to spontaneously nucleate
-within the solution.
+One of the key conclusions is the need for the development of 
+better \textit{in situ} techniques for characterizing and 
+imaging the nucleation pathways. Understanding which specific 
+pathway is unfolding under a certain set of experimental 
+parameters would allow one to develop a map of the thermodynamic 
+landscape available to a pre-nucleation system. The challenge 
+therein lies in developing experimental methods by which one can 
+reproduce similar nucleation events and in turn use 
+characterisation techniques to identify the pathway taken. The 
+method not only needs to be repeatable fora single set of system 
+parameters, but also needs to be flexible enough to replicate 
+results regardless of the system parameters (i.e. varying 
+supersaturation, temperature, and solute choice). There have 
+been several different methods deployed to try and observe 
+crystallisation at a microscopic scale.
 
-\begin{figure}[h!]
-	\includegraphics[width=\linewidth]{MSZW.png}
-	\caption{Typical Concentration vs Temperature plot with the 
-		solubility curve (black) and metastable zone (red). 3 
-		different cooling curves are shown as well: a high rate 
-		of cooling (orange) shows the mixture quickly dropping 
-		below the solubility curve resulting in no new nuclei 
-		forming. A low rate of cooling (green) shows the mixture 
-		exceeding the metastable zone, where now nucleation occurs 
-		spontaneously and cannot be controlled without further 
-		cooling. A typical cooling rate (blue) shows how a 
-		typical cooling crystalliser will operate, sitting in the 
-		middle of the other two curves.}
-	\label{fig:MSZW}
-\end{figure}
+\subsection{Computer Simulations}
+While not a direct observation of nucleation events, computer
+simulations have proved an invaluable tool for studying and 
+testing predictions about the pre-nucleated solution. By 
+comparing the simulative nucleation rate to the expected 
+rate given by current models, researchers can test newly 
+developed theories and glean information about the microscopic
+parameters of a system.
 
-The viability of cooling crystallisation is dependent on the 
-meta-stable zone width, too narrow and the process is difficult to 
-control, too wide and the crystal growth rate may be insufficient 
-for the desired outcome.
+The main challenge facing computer simulations is the issue
+of limited scope. Molecular dynamic simulations are constrained
+by their choice of time and length scale, for example simulations
+in a canonical ensemble face issues where the chemical potential
+is depleted to the point that the driving force for crystallisation
+is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
+addressed but requires an extensive increase in computational 
+resources to maintain the chemical potential equilibria. Like wise
+in most cases the choice of time scale is a crucial factor, shorter
+time scales provide a more accurate evolution of the solution; 
+however, in some cases the time scale magnitude for crystallisation
+is often far greater than the time scale used in the simulation. This
+has led to more advanced sampling techniques that allow for longer 
+time scales which has provided insights into the pre-nucleation 
+solution stage and formation of pre-nucleated clusters \cite{Finney2023}.
 
-\subsection{Evaporative Crystallisation}
-In situations where control of the final product size or shape 
-is not the main focus, often the cheapest method of producing a 
-crystalline product is simply to allow the solvent evaporate and 
-separate from the solute. Depending on the total volume of solvent 
-to evaporate this process can take on time scale of several days 
-to complete. This is often used for the formation of seed crystals 
-when larger seed crystals are required \cite{Srinivasan2008, Li1992}.
+The study of the pre-nucleated solution and the evolution of 
+precursors has been aided greatly by the use of simulative 
+studies. Early on it was shown in colloidal crystallisation 
+that the structure of the precursor was time dependent; 
+colloidal simulations showed that regardless of the molecular
+interactions between colloids the atomic packing pattern was 
+predominately hexagonal centred, over time the packing
+became dominated by body centred packing in the case 
+where the interactions were longer reaching \cite{Tan2013}. 
+They also confirmed that precursor formation was driven not
+by local density variations, but instead was driven by the 
+local bond order of the colloidal precursors \cite{Tan2013}, 
+providing strong evidence that the colloidal pre-nucleation 
+clusters undergo rearrangement prior to nucleation. Further
+developments in $NaCl$ simulations have shown that by varying
+the density fluctuations the precursor structure could either
+be amorphous, the classical rock-salt structure, or a new 
+structure similar to wurtzite. While this result is difficult
+to confirm it provides a important conclusion, that simulations
+can hint at new nucleation pathways that have not been confirmed
+experimentally. This means that computer simulations can not
+be merely used to confirm theories, but can be used to examine
+and search for novel results.
 
-\subsection{Anti-solvent Crystallisation} 
-Anti-solvent crystallisation is a more involved separation method,
-by adding a new solvent that is miscible with the old solvent but 
-is immiscible with the solute, the solute's effective solubility 
-is decreased. As such, the solution becomes supersaturated. This 
-has the trade off in that the overall mass ratio between the solvent 
-and solute is now lower. Careful measurement of the anti-solvent is 
-required in order to ensure that the decrease in solubility is not 
-out matched by the decrease in mass ratio.Anti-solvent crystallisation 
-is not as common due to the fact that not all solutions will have an 
-ideal anti-solvent candidate.
+Overall computer simulations make up a core area of research 
+in the study of nucleation. The technique is however limited,
+often struggling to examine nucleation over longer time scales
+or larger number of particles. While sampling methods and 
+enhanced simulation techniques has alleviated this somewhat 
+they are still far from being able to describe the behaviour 
+of larger molecules undergoing nucleation.
+
+\subsection{Transmission Electron Microscope (TEM)}
+One of the more well known techniques used for direct observations 
+of crystallisation dynamics is using transmission electron 
+microscopy (TEM). The basic working principle involves creating a 
+focused beam of electrons that are directed onto a target sample.
+Due to the wave-particle duality, the electrons can be treated 
+as a unified beam of light, whose wavelength is dependent on the 
+kinetic energy of the electrons. The wavelength of an individual 
+electron is given by \cite{Williams2009}:
+\begin{align}
+	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
+\end{align}
+
+\noindent
+When incident on a sample the wave is scattered by the sample 
+and can then be detected by an imaging screen that measures 
+the difference in intensity between the scattered and 
+transmitted beams. The advantage of TEM is that one can bypass
+the diffraction limit imposed by visible light, the maximum 
+resolution that can be achieved by a microscope is given by
+\cite{Champness2020}:
+\begin{align}
+	d = \frac{\lambda}{2NA}
+\end{align}
+
+\noindent
+where NA is the numerical aperture of the objective lens, and
+$\lambda$ is the wavelength of light used to illuminate a sample.
+For a visible light LED with an average wavelength of $550~nm$, 
+the upper limit for an objective lens (NA of 1.2) is around 
+$230~nm$ whereas a similar lens focusing an electron beam 
+with an average kinetic energy of $100~keV$ will have a maximum 
+resolution of just over $4~pm$ \cite{Champness2020}. There is
+an upper limit to the maximum resolution possible as electron 
+lens are far inferior to what modern optics can achieve 
+\cite{Champness2020} so there is a trade off between quality and 
+resolution.
+
+TEM technology has proven invaluable to the study of nucleation
+at an atomic level. The samples are often only a few hundred 
+atoms thick in order to optimise the imaging results. Much of 
+the supporting evidence for non-classical nucleation has been 
+demonstrated via TEM \cite{Ye2023}. For example when Cao 
+\textit{et al} observed the formation of amorphous atomic 
+cluster using $\gamma-$Fe,Au, and Re; all three of which 
+underwent continuous transformation before eventually 
+reaching an final ordered state \cite{Cao2020}. The amorphous 
+cluster did not immediately transition to an ordered structure, 
+instead fluctuating somewhat randomly until eventually settling 
+into a crystalline structure \cite{Cao2020}. This was confirmed 
+both in solid growth and liquid growth conditions \cite{Cao2020, 
+Loh2016}, suggesting that the non-classical nucleation path 
+dominates in all cases of metal crystal formation \cite{Ye2023}. 
+Confirmation is necessary for different solute choices, as in 
+the case of organic solids it is believed that non-classical 
+nucleation only dominates in low supersaturations whereas 
+classical nucleation is dominate at higher supersaturations.
+
+TEM analysis has not been extended to the study of organic 
+crystals, proteins, or simple salts at an atomic level 
+however. The main reason being that organic materials can be 
+damaged by the electron beam and therefore need to be frozen 
+prior to being imaged \cite{Champness2020}. This was used to 
+study the polymorphic time dependence of glycine crystals, by 
+varying the time prior to freezing Broadhurst \textit{et al} 
+showed that $\beta$ glycine dominates initially but $\alpha$-
+glycine is more stable after a longer period, $\gamma$-glycine 
+was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
+the time scales with which polymorphs form, it doesn't provide 
+much information on the kinetics of crystal growth as these 
+were snap shots and not direct observations of the nucleation 
+process.
+
+Overall TEM analysis is invaluable for the study of atomic 
+crystal growth, the secondary scattering from electron 
+beams also allows for chemical and structural information 
+to be gleaned from the target. This has given strong evidence
+for the non-classical nucleation theory, particularly in 
+metal crystal formation. Due to the high energy of the 
+electrons, organic materials have not received the same
+focus outside of cryoTEM analysis. Non-invasive methods
+are therefore necessary to gain an understanding of the
+crystal growth and formation of organic crystals, 
+particularly when both classical and multi-step nucleation 
+can occur.
 
@@
 measuring and exerting precise forces on the magnitude of pico-newtons 
-allowing one to probe the material properties of the smallest materials.
+allowing one to probe the material properties of the smallest 
+materials.
 
@@
 
+\subsection{Characterisation of rotational motion}
+
+Rotational motion about a single axis is easiest to account for.  
+When the power spectra of elliptical polystyrene particles was 
+fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
+assumed that the rotational motion was purely in the transverse
+plane. As such they did not have to account for any variance in 
+the trapping strength due to orientation nor did they need to
+consider non-periodic rotational behaviour. In the case where 
+rotational motion is stochastic the problem is more complex. 
+For example, when an optical fibre trap characterisation 
+technique was implemented by Saffron \textit{et al} 
+\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
+light scattering to characterise both the axial and lateral 
+trap stiffness acting on microspheres. The only drawback 
+admitted to in their work was that the technique was constrained 
+to isotropic scatters as their theoretical model for describing 
+the auto-correlation function was predicated on the fact that 
+any variations in the signal are due to the particles 
+translational motion within the confines of a cylindrical trap 
+\cite{BarZiv1997}. Where the upper limit of the cylindrical 
+trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
+However, as demonstrated by the results from Chapter 4.1, 
+the axial traps of spherical aggregates is often situated 
+far beyond the Rayleigh range (for a 1.2 NA laser this is 
+$\pm5.985 \mu m$).
+
 \section{Shear induced Nucleation}

### 2025-01-26 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

@@
 \label{chapter:theory}
-The two key areas of this project cover two widely different
+The two key areas of this project cover two somewhat unconnected
 fields; crystallisation theory, which is covered in the 
@@
 Proper understanding of optical tweezing requires a description
-of how the trapped particle interacts with the trapping laser. 
+of how the trapped particle interacts with the laser. 
 From an electromagnetism perspective the laser creates a 
 spatially and temporally coherent electric field that scatters 
-light off of a trapped particle. The laws governing electric 
+light when passing though a medium. The laws governing electric 
 and magnetic fields are summarised most succinctly via the 
@@
   &= \frac{\rho_ v}{\epsilon_0}
+  \label{eq:max_start}
   \\
@@
   &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}
+  \label{eq:max_end}    
 \end{align}
@@
 
-Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ 
-is the magnetisation field, these fields arise due to the 
-bound charges throughout the medium interacting with the EM 
-field.
+Where $\mu_0$ is the vacuum permeability $\mathbf{P}$ 
+is the polarization field and $\mathbf{M}$ is the 
+magnetisation field, these fields arise due to the 
+bound charges throughout the medium interacting with 
+the EM field.
 
@@
 
-The scattering force arises from the scattered field of the 
-trapping beam pushing the target object away from the focal 
-point. While some trapping arrangements can induce transverse 
-scattering forces for simple spheres the scattering force is 
-only significant in the direction of propagation \cite{Capitanio2002}.
-The equilibrium position is found when the gradient force 
-cancels out the the scattering force, and the thermal energy 
-of the particle.
+The scattering force arises from the motion of bound charges 
+generating secondary scattering as they move through the 
+electric field \cite{YasuhiroHarada1996}. Even if the particle
+experiences no external forces the scattering force is still 
+significant as the electric field oscillates, though for most
+practical applications the time dependence is ignored. While 
+some trapping arrangements can induce transverse scattering 
+forces for simple spheres the scattering force is only 
+significant in the direction of propagation \cite{Capitanio2002}.
+In order to determine where a particle will remain at rest 
+in an optical trap requires determining where the potential 
+well formed by the gradient force is equal to the thermal 
+energy of the system.
 
@@
 	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
-	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
-			while moving along the x-axis, where $x=0$ is the centre of the
-			beam. The harmonic regime is highlight by the black linear fit.}
+	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement
+		curve while moving along the x-axis, where $x=0$ is the 
+		centre of the beam. The harmonic regime is highlight by the 
+		linear fit given in black.}
 	\label{fig:harmonic_trap}
@@
 sphere when incident by a plane wave as a summation of partial 
-spherical waves. For any spherical wave the vector fields must 
-solve the Helmholtz wave equation given by:
+spherical waves. For any wave equation the vector fields must 
+solve the Helmholtz equation given by:
 \begin{align}
@@
 to match the desired beam shape. For a approximate Gaussian beam 
-interacting with a particle of volume $V$, the incident, internal, 
+interacting with a sphere of radius $a$, the incident, internal, 
 and scattered fields are given as \cite{RanhaNeves2019}: 
@@
 {nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
-spherical harmonics respectively. The superscripts denote whether 
+spherical harmonics respectively. The superscript denote whether 
 the field is inward [(1)] or outward propagating [(3)]. By linearity 
@@
 \cite{Crichton2000THEMD}. As an example the time averaged linear 
-momentum carried by any beam can be computed via:
+momentum carried by any incident beam can be computed via:
 \begin{equation}
@@
 \noindent
-Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
+Where $a_{nm}, b_{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
 coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
@@
 calculating the optical torque. Lorenz-Mie theory can be applied 
-to describe the scattering from any particle regardless of size, 
-though as the size of the target particle changes - relative to 
+to describe the scattering from any particle regardless of size.
+Though as the size of the target particle changes - relative to 
 the wavelength of light - the theory describing the electromagnetic 
@@
 experienced by said particle's it does not fully capture the behaviour
-of light when considering interference between different rays.
+of light when considering interference between different rays. And 
+as such no information is learned about the scattered field.
 
@@
   \label{eq:Rayleigh_end}
-  \text{where:}
+\end{align}
+where, for a approximate Gaussian beam the intensity $I(r)$ 
+close to the beam focus can be calculated as such.
+\begin{align}
   \nonumber
-  \\
 	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
@@
 There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
-\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
-Rayleigh regime is not concerned with the state of the scattered
-field. This is because if the particle is approximated by a point
-dipole we can ignore the momentum transfer calculation and simply
-compute the Lorenz force for a dipole in an electric field. While 
-not important for optical force calculation higher complexity 
-scattering problems can be simplified by subdividing the particle 
-into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
-the scattered field is now significant. As the target particle gets 
-larger this approximation fails to accurately describe the trapping force 
-\cite{Li2021}. For larger particles the optical force and scattering 
-is better described via the Lorenz-Mie theory.
+\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- \eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
+the state of the scattered field. This is because if the particle 
+is approximated by a point dipole we can ignore the momentum 
+transfer calculation and simply compute the Lorenz force for a 
+dipole in an electric field. While not important for optical force calculation higher complexity scattering problems can be simplified 
+by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), in which case the scattered field is now significant. As the target particle gets larger this approximation 
+fails to accurately describe the trapping force \cite{Li2021}. For 
+larger particles the optical force and scattering is better described 
+via the Lorenz-Mie theory.
 
@@
 this would later be extended to electromagnetic waves. Sometimes 
-referred to as the extended boundary condition method (ECBM), the 
+referred to as the extended boundary condition method (EBCM), the 
 method replaces the scatterer with a series of surface currents 
@@
 derive the individual elements of \textbf{T} until a desired 
-accuracy is achieved. One can either use the ECBM to compute the 
+accuracy is achieved. One can either use the EBCM to compute the 
 scattering for a fixed particle orientation and position. Or they
 can evaluate the scattering for any particle orientation by 
-computing its $T$-matrix.
-
-For this project we utilised an extension of the ECBM known as 
+computing its $T$-matrix. Because the T-matrix is based only 
+on the properties of the scatterer it can be reused in 
+calculations regardless of the incident field in question.
+
+For this project we utilised an extension of the EBCM known as 
 multi-sphere $T$-matrix method (\textit{mstm}) developed by 
@@
 number of spheres increases \cite{Mackowski2011}. While it is 
-possible to solve for the full electromagnetic field of each the 
+possible to solve for the full electromagnetic field for the 
 entire cluster this is only applicable for a single orientation 
@@
 extension of the Rayleigh approximation, though in order to compute 
-the optical force one requires a full description of the electric 
+the optical force one now requires a full description of the electric 
 field. Each dipole interacts with both the incident field and the 
@@
 is that the composition of the target can be changed freely. When 
-comparing different computational scattering methods, the ECBM was 
+comparing different computational scattering methods, the EBCM was 
 found to be better suited for simulating the scattering of symmetric
-targets as the ECBM can use the target's symmetry to speed up 
+targets as the EBCM can use the target's symmetry to speed up 
 calculations \cite{Wriedt1998}. But when dealing with inhomogeneous 
-media DDA is more efficient compared to ECBM.
+media DDA is more efficient compared to EBCM.
 
@@
 While describing the optical forces is sufficient for 
-theoretical characterisation of an optical trap, trying 
-to replicate that model in an experimental setting is 
-near impossible. As such it is often far simpler to 
+theoretical characterisation of an optical trap, in an 
+experimental setting the observed motion is not as 
+simple. As such it is often far simpler to 
 characterise the optical trap by recording the particle's
@@
 
-\section{Langevin Equation}
+\subsection{Langevin Equation}
 Describing any microscopic motion requires an understanding 
@@
 equation. Models such as the Fickian, and Einstein derivations 
-are sufficient for macroscopic behaviours the Langevin 
-equation better describes the microscopic characteristics of 
+are sufficient for macroscopic behaviours but the Langevin 
+model better describes the microscopic characteristics of 
 any diffusive behaviour.
@@
 able to describe its motion over any time scale, unlike the 
-Einstein model:
+Einstein model. The standard form of the Langevin model for 
+a particle undergoing Brownian motion is given as:
 \begin{align}
@@
 behaviour of an overall system, but for this project we can 
-instead consider the behaviour of some particle with a diffusion 
+instead consider the behaviour of a singular particle with a diffusion 
 tensor $D$ suspended in a viscous fluid and spacially localised 
@@
 
-\eqref{eq:langevin} provides an accurate description of strongly 
+Eq.~\eqref{eq:langevin} provides an accurate description of strongly 
 trapped particles. Despite this the analytical solution of the 
@@
 and extract information about the particle and fluid, and how the 
-two interact with the optical trap.
+two interact with the optical trap. Solving the Langevin equation
+numerically is far easier when considering small time steps.
 
@@
 which for strongly trapped particles is far greater than the 
-characteristic time. Therefore for simulation purposes, we can 
-neglect the particle's inertia which allows the motion of an 
-optically trapped particle to be written as:
+characteristic time. As such we can neglect the particle's inertia 
+which allows the motion of an optically trapped particle to be written 
+as:
 \begin{align}
@@
 
+\subsection{Beam shapes}
+A quick aside is necessary prior to the discussion 
+of calibration techniques, namely on the assumption 
+of the beams shape. It has been previously mentioned 
+that the focused beam can be approximated as a 
+Gaussian beam, what this implies is that the beam has 
+a circularly symmetrical intensity profile that can be
+described by the Gaussian function. There are alternative 
+beam formations that also preserve the circularly 
+symmetric profile, such as Bessel beams that have an 
+intensity profile that is described by the Bessel 
+function. 
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{Gaussian_Bessel.png}
+	\caption{Intensity distribution from a Gaussian and Bessel
+	beam.}
+\end{figure}
+
+An idealised form of a Gaussian and Bessel beam would be 
+unbounded, meaning the intensity of the electric field 
+never goes to 0 - approaching it asymptomatically instead. 
+This however breaks Maxwell's Equations (see Eq.
+\eqref{eq:max_start} - \eqref{eq:max_end}), what this 
+means is that in reality is that a true Gaussian or Bessel 
+beam cannot exist, only a close approximation \cite{Gouesbet1988}. 
+Due to the converging fields, a focused beam will experience 
+fluctuations in its electric field that distort the intensity 
+profile \cite{Lock1994}. This effect is insignificant 
+for slowly converging beams but for trapping objectives 
+used in optical tweezing there is a notable effect.
+
+When a Gaussian beam is linearly polarised and 
+focused to a point the focal point will not be 
+perfectly symmetrical but instead show some 
+degree of ellipticity even if the initial beam 
+itself is kept parallel. In practical terms this
+means that regardless of the calibration 
+technique used the trap stiffness along the x and
+y axis will not be equal to one another. Proper
+alignment can bring them close together and longer
+data acquisition times can reduce the uncertainty of
+the measured values, but there will always be a 
+slight variance in the two values. We can measure
+the ellipticity of the beam using the acquired values 
+from our calibration.
+\begin{align}
+	e = \left(1-\frac{\kappa_x}{\kappa_y}\right)^{1/2}
+\end{align}
+Generally the concern is more over the uncertainty in 
+the trap stiffness values and not the overall shape. 
+Even laser manufacturers will often state a maximum
+ellipticity, assuming perfect alignment and focusing, 
+in most cases a ellipticity of 10-20\%  is still 
+considered acceptable for most calibration results
+\cite{BergSoerensen2004}.
+
 \subsection{Potential Well Analysis}
-The Langevin equation for an optically trapped assumes that the 
-trap acts similar to a Hookean spring that creates a potential 
-well about its centre. Therefore a simple analysis method is to 
-understand the height and width of said potential well.
+The Langevin equation for an optically trapped particle assumes 
+that the trap acts similar to a Hookean spring that creates a 
+potential well about its centre. Therefore a simple analysis 
+method is to understand the height and width of said potential well.
 
@@
 
-Now assuming our laser acts as a Gaussian beam we should be able 
-to describe the probability distribution $\rho(x)$ centred at 
-some equilibrium position $x_0$:
+Now assuming our laser approximates a Gaussian beam we 
+should be able to describe the probability distribution 
+$\rho(x)$ centred at some equilibrium position $x_0$:
 \begin{align}
 	\label{eq:prob_dist}
-	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
-\end{align}
-
-By inserting \eqref{eq:prob_dist} into \eqref{eq:potential_well} 
-we can fit the potential well in order to determine the trap strength 
-$\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. 
-This has some limitations in that the large fluctuations can throw 
-off the fit meaning a longer acquisition time is necessary to properly 
-fit the potential well, making it difficult to characterise weakly 
-trapped particles who may not remain trapped for long. It also 
-provides no information on the particle itself (i.e. the friction 
-coefficient $\gamma$ and diffusion coefficient $D$).
+	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}x^2\right)
+\end{align}
+
+After observing the particles position over a long enough
+period of time we can construct and experimental estimation
+of $\rho(x)$. By inserting \eqref{eq:prob_dist} into \eqref{eq:potential_well} we can fit the potential well in 
+order to determine the trap strength $\kappa_x$. This has 
+some limitations in that the large fluctuations can throw 
+off the fit meaning a longer acquisition time is necessary 
+to properly fit the potential well, making it difficult to 
+characterise weakly trapped particles who may not remain 
+trapped for long. It also provides no information on the 
+particle itself (i.e. the friction coefficient $\gamma$ and 
+diffusion tensor $D$).
 
@@
 $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation 
-of the trap strength. Because of this requires a time average of 
+of the trap strength. Because this requires a time average of 
 the particle's displacement any large errors in the position 
@@
 \begin{align}
-	mx''(t) + \gamma x'(t) = \eta(t)
+	m\frac{\delta^2 x}{\delta t^2}+ 
+	\gamma\frac{\delta x}{\delta t} = \eta(t)
 \end{align}
@@
 the surrounding fluid. If the motion is truly random then we should 
-see an average displacement of 0 regardless of how long we measure 
-for. If we wish to understand the effects of a given external factor 
-(i.e. an electric field or localised heating), simply looking at 
-displacement will reveal nothing of value as its difficult to 
-differentiate between diffusive and a biased motion.
+see the average displacement increase over longer observation periods. 
+This can be useful for measuring characteristic times of a system, or
+in the case of optical trapping, characterise the optical field.
 
@@
 the MSD by forming a 'window' between two points in time of the 
-trajectory (i.e. $t \&  t+\tau$) and sliding this window along the 
-entire trajectory length - to eliminate -ve displacements we take 
-the square - we can then take the average of this series. Repeating 
-over a range of time lags allows us to describe the MSD as a function 
-of $\tau$:
+trajectory (i.e. $t\ \&  t+\tau$) and sliding this window along the 
+entire trajectory length and then take the average of this series. Repeating over a range of time lags ($\tau$) allows us to describe 
+the MSD as a function of $\tau$:
 \begin{align}
@@
 increases with larger values of $\tau$ until it reaches a maximum 
-value as shown below by the dotted line.
+value as shown in fig.~\ref{fig:MSD} by the dotted line.
 \begin{figure}[h!]
@@
 		displacement due to the optical trap.}
+	\label{fig:MSD}
 \end{figure}
@@
 the particle is experiencing the harmonic potential described by 
-the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle 
+the equipartition theorem, and when $\tau \ll \tau_{OT}$ the particle 
 is said to be freely diffusing within the trap focus. Of course for 
@@
 
-\subsubsection{Angular Mean Square Displacement (MSAD)}
+\subsubsection{Mean Square Angular Displacement (MSAD)}
 It is also possible to plot the angular MSD (MSAD) using 
 simulative or experimental data, but there is no complementary 
-expression for \eqref{eq:MSD}. This is because while the 
+expression for Eq.~\eqref{eq:MSD}. This is because while the 
 particle is close to the focus the optical force is easily 
-modelled as a Hookean spring; conversely, the optical torque
-is far more complex even if the particle is strongly trapped. 
+modelled as a Hookean spring, even if the particle undergoes
+some rotation. Conversely, the optical torque applied to a 
+particle is far more sensitive to changes in position meaning 
+a simple linear approximation is not satisfactory. 
 Vigilante \textit{et al} \cite{Vigilante2020} derived the 
@@
 where $x$ can either be the physical position (relative to the trap
-focus) or a recorded signal from a position-detector,and $\hat{\eta}$ 
-is the Fourier transform of the white noise (see Eq.~\eqref{eq:full_langevin}). 
-Where the values are exponentially distributed as opposed to being 
-normally distributed in the time domain \cite{BergSoerensen2004}.
-We can therefore ignore the white noise from our analysis by looking 
-at the spectral density of $\hat{x}(f)$ that produces a Lorentzian
+focus) or a recorded signal from a position-detector, and $\hat{\eta}$ 
+is the Fourier transform of the white noise (see Eq.~\eqref{eq:full_langevin}). Where the values are exponentially distributed as opposed to being normally distributed in the time 
+domain \cite{BergSoerensen2004}. We can therefore ignore the white 
+noise from our analysis by looking at the spectral density of $\hat{x}(f)$ that produces a Lorentzian
 curve: 
@@
 high degree of accuracy, depending on the local heating effect this 
-may be as trivial increase or it may be significant enough to drastically 
-alter the characterisation. PSD analysis is often seen as the gold 
-standard for calibration as it can be fine tuned to the point that 
-optical forces can be computed on the order of $10^{-15} N$ 
+may be a trivial increase or it may be significant enough to drastically 
+alter the characterisation.
+This also assumes that the particle shape is known, and by implication its diffusion tensor.
+PSD analysis is often seen as the gold standard for calibration as it can be fine tuned to the point that optical forces can be computed on the order of $10^{-15} N$
 \cite{BergSoerensen2004}, it captures all of the information acquired 
 by other calibration techniques while filtering out noise and requiring 
-a relatively small amount of data collected.
+a relatively small amount of data collection.
 
@@
 as it rotates the detected signal will oscillate periodically. When a 
-Fourier transform is taken the power spectra will have a peak signal at 
-that periodic signal.Typically papers reporting on the rotation rate of 
-a trapped object will collect a power spectrum and look at the maximum frequency term to determine its rotation rate. However this often neglects 
-any information on the trapping forces acting on particle and only looks 
-at the torque applying around one of the particle's primary axis. There 
-has only been one notable effort to fully characterise the full trapping dynamics on a rotating body, \cite{Yogesha2012} developed a theoretical 
-model that relates the power spectrum to both the rotational and 
-translational motion simultaneously. This is applicable to particles with
-a polarizable long axis but is insufficient for birefringent particles.
+Fourier transform of the time series is taken the power spectra will 
+have a peak signal at that periodic frequency. Typically papers 
+reporting on the rotation rate of a trapped object will collect a power spectrum and look at the maximum frequency term to determine its rotation rate.
+However this often neglects any information on the trapping forces acting on particle and only looks at the torque applying around one of the particle's primary axis.
+There has only been one notable effort to fully characterise the full trapping dynamics on a rotating body, \cite{Yogesha2012} developed a theoretical model that relates the power spectrum to both the rotational and translational motion simultaneously.
+This is applicable to particles with a polarizable long axis but is insufficient for birefringent particles.
 
@@
 trajectory the dimer has a 'equilibrium' orientation in which the 
-gradient force is maximised we can say that the first right hand 
+optical torque is minimised we can say that the first right hand 
 term can be replaced with $\vec{\kappa}_u(\vec{r}(t))$ to represent
-a harmonic potential. $\lambda(t)$ is the Brownian rotations from the 
-surrounding fluid, like in the translational case the Brownian 
-rotations are normally distributed and are also uncorrelated so that:
+a harmonic potential due to the torque. $\lambda(t)$ is the Brownian 
+rotations from the surrounding fluid, like in the translational case 
+the Brownian rotations are normally distributed and are also uncorrelated 
+so that:
 \begin{align}
@@
 \ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
-to compute the optical force and torque applied to th dimer.
+to compute the optical force and torque applied to the dimer.
 
@@
 constrained to individual spheres we can now consider the 
-expected response from any type of spherical aggregate.
+expected response from any type of spherical aggregate. The
+goal of such simulation software would be to simulate the 
+expected experimental results from a complex particle whose 
+exact shape and size may be difficult to determine.
 
@@
 Where the denominator is the total intensity on the QPD while there is 
-no particle within the trap. The QPD sensitivity is dependent on both 
-the polarisation of the incident beam and the displacement of the target 
-sphere, this is shown in fig.~\ref{fig:totalfield}.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.7\linewidth]{fixed_polarisation.png}
-	%\captionsetup{margin=0.5cm}
-	\caption{Total Field incident on the quadrant photo diode for a several
-		different displacements. Top Left: $2\mu m$ diameter silica sphere
-		placed at the focus of a  beam. Top Right: silica sphere is now
-		displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere 
-		is now displaced along the y-axis. Bottom Left: silica sphere is 
-		now displaced by $0.707 \mu m$ along the x and y axis. The 
-		trapping beam is a $TEM_{00}$ Gaussian, polarised along the along 
-		the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
-	\label{fig:totalfield}
-\end{figure}
+no particle within the trap. You would expect that $S_x$ and $S_y$ are
+near identical for equal particle displacements, however due to the 
+polarisation of the beam there will be a slight bias to the signals 
+for motion along the direction of the polarisation vector. This is 
+generally not an issue if the particle trajectory is sampled for a 
+long enough time period. By converting from signal units to length
+units (see \eqref{eq:conversion_factor}) the trap shape can be 
+discerned from the QPD signals, assuming the fluid properties are
+known to a high degree of accuracy.
 
@@
 this we need to rotate the total field via the inverse rotation matrix
-of the dimer. Shown in fig, is a flow diagram outlining the full work
-flow of a given trajectory, and how the QPD data is acquired.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.75\linewidth]{Simulation_workflow.pdf}
-	\caption{Block flow diagram of how a trajectory and QPD data is 
-		calculated and collected. Relevant equations are mentioned 
-		at each step.}
-\end{figure}
-\newpage
+of the dimer.
 

### 2025-01-26 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

--- 2024-11-18 Latex_files/Thesis/chapters/chapter 4/main.tex
@@
 trapped in a single OT and found that multiple trapped 
-beads could be mistaken for a single trapped bead with 
+beads could be mistaken for a single trapped bead with an 
 altered trap stiffness (see \ref{sec:harmonic_traps}). 
@@
 There are further discussions looking into the dynamics of 
-a whole host of asymmetrically shaped particles \cite{Loudet2014, 
-ShengHua2005, Chetana2022}, their results all showing that 
-predicting the behaviour an arbitrary shaped particle comes 
-with great difficulty due to the fact that the optical force 
-is dependent on a greater number of variables such as 
-orientation and size factors.
+a whole host of asymmetrically shaped particles (e.g. 
+ellipsoids, cylinders, amorphous solids, disks, etc) 
+\cite{Loudet2014, ShengHua2005, Chetana2022}, their results 
+all showing that predicting the behaviour an arbitrary shaped 
+particle comes with great difficulty due to the fact that the 
+optical force is dependent on a greater number of variables 
+such as orientation and size factors.
 
-With the initial goal of the project being to induce 
-nucleation events via a rotating sphere, the aim of this 
-chapter is to - in a limited capacity simulate and investigate 
-the influence of a second particle being bound to our target 
-sphere. The choice of a dimer, instead of an amorphous solid 
-that might better represent a growing crystalline solid, 
-allows us to consider how the dynamics of the aggregate change 
-by varying a single parameter, namely the size ratio of the 
-two spheres. Additionally, attempting to simulate an amorphous 
-aggregate is rather difficult as calculating the optical force 
-and torque is computationally slow. We build upon the works of 
-Vigilante \textit{et al} \cite{Vigilante2020} to consider 
-asymmetric spherical dimers and how varying size parameters 
-alters the dynamics and additionally makes characterising their 
-interactions within an optical trap more cumbersome.
-
-In this chapter we will consider how the addition of a 
-second sphere changes the trapping dynamics by introducing 
-multiple equilibrium positions. These positions are 
-dependent on both the size ratio of the dimer and its 
-orientation relative to the trap. This is important to 
-address, for while the the dimers' diffusive behaviour 
-is the same, regardless of position and orientation, 
-how it interacts with the trap may produce the wildly 
-different results. Furthermore, we look at how our dimers 
-interact with circularly polarised light, this is especially 
-pertinent as the original plan was to utilise circularly 
-polarised light to generate fluid flow in a supersaturated 
-solution. The rotational motion of dimers in circularly 
-polarised light goes against much of the established 
-literature and suggests that the transfer of spin angular 
-momentum is enabled by combining particles together. And 
-lastly, we demonstrate how a quadrant photo diode, performs 
-in characterising the actual interactions between a dimer 
-and the optical trap. We do so to demonstrate that much 
-of the dimer's trajectory information is either lost or 
-poorly described by simple calibration techniques, making 
-their motion difficult to characterise.
-
-\subsection{Simulation Parameters}
-As a paradigmatic example, consider a dimer suspended in 
+As a paradigmatic example, we consider a dimer suspended in 
 water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+With the initial goal of the project being to induce 
+nucleation events via a rotating sphere, the aim of this 
+chapter is to - in a limited capacity simulate and investigate 
+the influence of a second particle being bound to our target 
+sphere. The choice of a dimer, instead of an amorphous solid 
+that might better represent a growing crystalline solid, 
+allows us to consider how the dynamics of the aggregate change 
+by varying a single parameter, namely the size ratio of the 
+two spheres. Additionally, attempting to simulate an amorphous 
+aggregate is rather difficult as calculating the optical force 
+and torque is computationally slow.
+
+We utilise a python library that utilises \textit{mstm} and
+\textit{ott} to simulate the dynamical behaviour of spherical
+aggregates. A dimer is defined by its refractive index and the 
+size ratio of the two spheres.  By specifying the optical 
+properties of the target \textit{mstm} generates the targets 
+T-matrix which can then be used by \textit{ott} to compute the 
+optical forces generated by a focused beam.
+
+In this chapter we will consider how the addition of a 
+second sphere changes the trapping dynamics by introducing 
+multiple equilibrium positions. Furthermore, we look at how 
+dimers interact with circularly polarised light, this is 
+especially pertinent as the original plan was to utilise 
+circularly polarised light to generate fluid flow in a 
+supersaturated solution. And lastly, we demonstrate how a 
+quadrant photo diode, performs in characterising the actual 
+interactions between a dimer and the optical trap. We do so 
+to demonstrate that much of the dimer's trajectory information 
+is either lost or poorly described by simple calibration 
+techniques, making their motion difficult to characterise.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation Details}
+As a paradigmatic example, consider a dimer suspended in 
+water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
+a Gaussian beam (more specifically a Laguerre-Gaussian 
+beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
+is focused by a objective with numerical aperture of 1.2.
+The polarisation is defined using the Jones vector (see 
+section \ref{sec:rot_pol} for further details) unless 
+stated otherwise the Jones vector is kept at $[1.0, 0.0i]$. 
+The size ratio of the two spheres is given by $a_{I}/
+a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ unless 
+specified otherwise. The beam focus is used as the 
+origin of the coordinate system, with positive z going 
+in the direction of beam propagation. For simulations 
+we assume that the dimer is suspended in water at room 
+temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).
+
+In order to compute the magnitude of the Brownian forces we
+need to know the dimer's diffusion tensor \textbf{D}. Unlike
+for spheres where the translational and rotational diffusion 
+coefficients are equal along all three Cartesian axis', a
+dimer's translational and rotational diffusion is dependent
+on whether we consider the its long or short axis. We use
+the results from \cite{Nir1973} to compute the dimensionless 
+diffusion tensor and then scale it according to the radius of
+the largest sphere ($a_I$). The dimer's centre of diffusion
+is located a distance of $a_I\zeta$ from the point of contact
+between the two spheres, the coefficient $\zeta$ is based on
+the size ratio of the two spheres. We use the centre of 
+diffusion as the target origin that is provided to \textit{mstm}.
+
+The dimer's orientation is given by a rotation vector in 
+the form: 
+\begin{align}
+	\begin{pmatrix}
+		U_{1,x} & U_{2,x} & U_{3,x} \\
+		U_{1,y} & U_{2,y} & U_{3,y} \\
+		U_{1,z} & U_{2,z} & U_{3,z}
+	\end{pmatrix}
+\end{align}
+Where $U_3$ represents the orientation vector that passes
+through the centres of both sphere's (see \ref{sec:sim} 
+for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
+$U_3$ but have no physical representation for a dimer. We 
+can represent $U_3$ by it's spherical coordinates ($\theta$ 
+and $\phi$), where $\theta$ is the angle between the 
+orientation vector and the direction of propagation and 
+$\phi$ as the angle between the orientation vector and 
+the x-axis. We define the 'standard' orientation as 
+being aligned with the direction of beam propagation 
+direction ($\theta=0^\circ$) - and therefore the 'inverted' 
+orientation is defined when the dimer is orientated against 
+the direction of beam orientation ($\theta=180^\circ$) - see
+fig~\ref{fig:paradigmatic} for a visual of the laboratory and
+particle frame.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Positional and Orientational dependence of Trapping forces}
@@
 
-We can see that both equilibrium positions have comparable 
-axial trap stiffness ($\kappa_z$), however the difference 
-in the transverse trap stiffness ($\kappa_x$) is far more 
-noticeable. The same dimer was trapped at each of the axial 
-equilibrium positions and the transverse force was evaluated. 
-While in all three cases the dimer can be trapped the linear
-range where that would typically associated with a stable 
-trap is far narrower in the 'standard' orientation compared 
-to the 'inverted' cases. This highlights one of the challenges 
-involved with studying asymmetric particles, even though its 
-a simple enough process to trap them they maybe characterised 
-very differently depending on their relative position and 
-orientation towards the focus. This can have a significant 
-impact on rheological studies - or attempting to probe any 
-local property - as the variance in trap strength can result 
-in large errors over repeated measurements. 
+This is different to a single sphere, namely in that a
+single sphere should only have one equilibrium position
+and that it is much closer to the trap focus ($\approx
+0.05\mu\ m$ from the focus). We can see that both 
+equilibrium positions have comparable axial trap 
+stiffness ($\kappa_z$), however the difference in the 
+transverse trap stiffness ($\kappa_x$) is far more 
+noticeable. The same dimer was trapped at each of the 
+axial equilibrium positions and the transverse force was 
+evaluated. While in all three cases the dimer can be 
+trapped the linear range where that would typically 
+associated with a stable trap is far narrower in the 
+'standard' orientation compared to the 'inverted' cases. 
+This highlights one of the challenges involved with 
+studying asymmetric particles, even though its a 
+simple enough process to trap them they maybe 
+characterised very differently depending on their 
+relative position and orientation towards the focus. 
+This can have a significant impact on rheological 
+studies - or attempting to probe any local property 
+- as the variance in trap strength can result in 
+large errors over repeated measurements. 
 \begin{figure}[h!]
@@
 to $a_{I}/a_{II}=4$), a second equilibrium position is found 
-below the trapping focus. Previous work using the ray-optics 
-model have confirmed even in the case that two spheres begin 
-separated the electric field will align the particles as such 
-that they make contact and are trapped together about a single 
-trapping position \cite{Xu2005}. Furthermore it has been shown 
-through proper manipulation of the Gaussian or Bessel beam 
-modes that any number of trapping potentials can be developed 
-\cite{Shahabadi2020} for nanoparticles. This result however, 
-is the first example of an orientation dependent trapping 
-situation using only a $TEM_00$ beam. Typical experimental 
-arrangements cannot determine much information on the axial 
-position of a trapped particle relative to the trap focus; 
-this result indicates not only that dimers can be trapped 
-in multiple axial positions but also their trapping behaviour 
-is heavily dependent on said axial position. As such it is 
-necessary that positional information in the z-axis can be 
-elucidated if multiple spheres are trapped simultaneously. 
+below the trapping focus. We should note that there are also
+unstable traps that exist, these are also depicted in 
+fig.~\ref{fig:eq_positions}. While the optical force
+is zero at these positions the gradient $\delta F_z/\delta z$
+is positive and so any Brownian motion would displace the 
+dimer, either ejecting it from the optical trap or letting
+it be drawn towards another harmonic trap. 
 \begin{figure}[h!]
@@
 	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying 
-		size ratio, dashed lines represent unstable traps whereas solid lines 
-		are for stable equilibrium positions. (a) shows that dimers while in their
-		'standard' orientation will always have a single equilibrium position. 
-		(b) shows that when the same dimer is in its' 'inverted' orientation 
-		can be trapped in two axial positions, one below the focus and one above 
-		the focus.}
+	\caption{Equilibrium positions of optically trapped dimers 
+		with varying size ratio, dashed lines represent unstable 
+		traps whereas solid lines are for stable equilibrium 
+		positions. (a) shows that dimers while in their 
+		'standard' orientation will always have a single 
+		equilibrium position. (b) shows that when the same dimer 
+		is in its' 'inverted' orientation can be trapped in two 
+		axial positions, one below the focus and one above the 
+		focus.}
 	\label{fig:eq_positions}
@@
 
+Previous work using the ray-optics model have confirmed even 
+in the case that two spheres begin separated the electric field 
+will align the particles as such that they make contact and 
+are trapped together about a single trapping position 
+\cite{Xu2005}. Furthermore it has been shown through proper 
+manipulation of the Gaussian or Bessel beam modes that any 
+number of trapping potentials can be developed 
+\cite{Shahabadi2020} for nanoparticles. This result however, 
+is the first example of an orientation dependent trapping 
+situation using only a $TEM_00$ beam. Typical experimental 
+arrangements cannot determine much information on the axial 
+position of a trapped particle relative to the trap focus; 
+this result indicates not only that dimers can be trapped 
+in multiple axial positions but also their trapping behaviour 
+is heavily dependent on said axial position. As such it is 
+necessary that positional information in the z-axis can be 
+elucidated if multiple spheres are trapped simultaneously.
+
 \newpage
-\subsection{Non-trivial equilibrium configurations}
+\section{Non-trivial equilibrium configurations}
 \label{sec:off-axis}
-Computing the equilibrium positions when a dimer is aligned 
-with the electric field is relatively simple as the orientational 
-torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
-trapped the dimer is unlikely to change orientation enough to 
-escape the trap. Regardless, that does not rule out the 
-possibility that there is a stable configuration where the 
-orientation not strictly vertical, in fact most experimental 
-work with symmetric nano-dimers will trap them lying perpendicular 
-to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
-\ref{sec:eq_positions} we cannot simply measure the optical force 
-and torque as the parameter space is too large and determining if 
-a particular position and orientation is stable is not clear based 
-solely on force and torque measurements \cite{Bui2017}. Using the 
-same simulation parameters as before we ran a number short 
-simulations (total simulation time was $0.005$ s) with the laser 
-power increased to $500\ mW$. Each simulation started with the 
-dimer in a different starting position and orientation, due to the 
-high laser power the dimers either escaped the trap or were stably 
-trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
-between the direction of beam propagation and the dimer's orientation 
-vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
-divided into different regions depending on which equilibrium 
-configuration is reached.
+As mentioned previously in \ref{sec:opt_torque} the 
+torque applied to an anisotropic scatterer can be 
+broken down into two categories: torque generated 
+from the field polarisation and torque generated 
+by the gradient force of the electric field. The 
+former is generally only applicable for objects 
+that are birefringent whereas the latter is 
+applicable to dimers due to their elongated shape.
+Since both spheres are attracted towards the centre
+of intensity via the gradient force a dimer will 
+experience a restoring torque that keeps the it
+in an equilibrium orientation. For any elongated 
+particle the torque is minimised when the long 
+axis is aligned either parallel or perpendicular
+with the direction of the electric field. In this 
+section we try to find equilibrium configurations 
+(positions and orientations where both the optical
+force and torque are minimised).
+
+Computing the equilibrium positions when a dimer is 
+aligned with the electric field is relatively simple 
+as the orientational torque is minimised (see Eq.
+\ref{eq:opt_torque}). Meaning once trapped the 
+dimer is unlikely to change orientation enough to 
+escape the trap. Regardless, that does not rule out 
+the possibility that there is a stable configuration 
+where the orientation not strictly vertical, in fact 
+most experimental work with symmetric nano-dimers will 
+trap them lying perpendicular to the beam direction 
+\cite{Ahn2018, Reimann2018}. Unlike in Sec.~
+\ref{sec:eq_positions} we cannot simply compute the 
+optical force and torque as the parameter space is too 
+large and determining if a particular position and 
+orientation is stable is not clear based solely on 
+force and torque measurements \cite{Bui2017}. Using 
+the same simulation parameters as before we ran a 
+number short simulations (total simulation time was 
+$0.005$ s) with the laser power increased to $500\ mW$. 
+Each simulation started with the dimer in a different 
+starting position and orientation, due to the high 
+laser power the dimers either escaped the trap or 
+were stably trapped. We chose to only consider the $z-
+\theta$ phase space in order to simplify the explanation,
+considering all possible parameters would increase the 
+simulation time significantly. The $z-\theta$ phase space 
+can be divided into different regions depending on which 
+equilibrium configuration is reached.
 \begin{figure}[h!]
@@
 	\begin{subfigure}{0.32\linewidth}
-		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
-		\caption{}
+		\raisebox{0pt}[0pt][-10pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
 	\end{subfigure}
 	%\captionsetup{hangindent=2.35cm}
-	\caption{Map of $z-\theta$ phase space using a dimer of
-		size ratio 2 with a laser power of 500 mW ($\theta=0
-		\circ$ is the 'standard' orientation and $\theta=180
-		^\circ$ is the 'inverted' orientation). The stable 
-		configurations are indicated by the larger circles 
-		and the starting conditions are colour coded to match 
-		the stable point they end up in. Right hand render 
-		shows a dimer in its off-axis configuration.}
+	\caption{Map of $z-\theta$ phase space using a dimer 
+		of size ratio 2 with a laser power of 500 mW 
+		($\theta=0^\circ$ is the 'standard' orientation 
+		and $\theta=180^\circ$ is the 'inverted' 
+		orientation). The stable configurations are 
+		indicated by the larger circles and the 
+		starting conditions are colour coded to match 
+		the stable point they end up in. Right hand 
+		render shows a dimer in its off-axis configuration.}
 	\label{fig:off_axis}
@@
 
-Interestingly while the trap strength of these off-axis traps 
-are similar in magnitude to the vertically aligned dimers, 
-but when the laser power is lowered (around $5\ mW$) the traps 
-become metastable; after reaching its equilibrium configuration 
-the particle behaves similarly to a typically trapped dimer but 
-the trapping potential is small enough that the dimer escapes 
-in as little time as less than a second or after nearly a full 
-3 seconds. Running similar simulations but for vertical 
-configurations sees the dimer remaining trapped, even after 30 
-seconds of run time, indicating that the trapping potential is 
-far greater than the thermal energy. This suggests that the 
-reason this off-axis configuration is metastable is due to the 
-increased rotational freedom. Similar configurations have been 
-explored with ellipsoids; Zhu \textit{et al.} looked at the 
-dynamics of various elliptical particles and found that regardless 
-of shape or initial orientation the particle would tend towards
-either a purely vertical or purely horizontal orientation 
-\cite{Zhu2021}.
+Where the each starting location is colour coded to
+match the final equilibrium position reached, empty
+regions indicate that the trap was escaped. As 
+expected we see that the dimer has three separate 
+traps when the dimer is close to its vertical 
+orientation, similar to what we see from figs.~
+\ref{fig:paradigmatic} and \ref{fig:eq_positions}. 
+When the beginning orientation is close to horizontal 
+there exists a $4^th$ harmonic trap. Each of these 
+traps, has its own 'basin' of attraction around 
+which the dimer is drawn into, the strange shape 
+of these basins is likely due to how the asymmetric 
+shape interacts with the optical trap.
 
+Interestingly while the trap strength of these off-axis 
+configurations are similar in magnitude to the vertically 
+aligned dimers, but when the laser power is lowered 
+(around $5\ mW$) the traps appear metastable. With the 
+vertical configurations even after $30\ s$ of simulation
+time the dimer is still trapped whereas the off-axis 
+configuration can be escaped after only a few seconds of
+simulation time. The suggested that the configuration is 
+metastable is due to the increased rotational freedom. 
+Similar configurations have been explored with ellipsoids; 
+Zhu \textit{et al.} looked at the dynamics of various 
+elliptical particles and found that regardless of shape or 
+initial orientation the particle would tend towards either 
+a purely vertical or purely horizontal orientation 
+\cite{Zhu2021}. However, in our case the dimer never returns
+to a vertical orientation, suggesting their is a potential 
+barrier separating the different configurations. In order
+to build a better understanding of the potential landscape 
+we need to consider how the force and torque vary for a 
+dimer in the $z-\theta$
+
+\subsection{Force-Torque landscape of an arbitrary dimer}
+We use \textit{ott} to compute the optical force 
+and torque for the same dimer as used in fig.~
+\ref{fig:off_axis} in different positions or 
+orientations, this is depicted in fig.\ref{fig:force_torque}. 
+The magnitude of the force and torque are indicated
+by the colour intensity, with white indicating either
+zero force or torque. This helps show where a
+dimer may end up in equilibrium by finding 
+where both the torque and force go to 0.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{force_torque.png}
+	\caption{Force-Torque landscape over the $z-\theta$ 
+		phase space with a dimer is a size ratio of 2, 
+		laser power of 100mW. Left: 2D plot of the 
+		optical force in the axial direction as a 
+		function of position and orientation. Right: 
+		2D plot of the optical torque about the x-axis 
+		as a function of position and orientation.}
+	\label{fig:force_torque}
+\end{figure}
+
+The force and torque landscape provides some insight 
+into how the dimer can inhabit an off axis equilibrium
+orientation. Firstly, lets consider a typical trap, 
+while the dimer is close to vertical; in all of these 
+cases the optical force and torque directs the dimer 
+back to the configuration. In this case the potential 
+well is positive whether the dimer moves along the 
+z-axis or is rotated. We can assume that in both cases
+the translational and rotational stiffnesses are close
+enough to be modelled as harmonic potentials.
+
+However when in an off-axis configuration ($z\approx0$ 
+\& $90^\circ\le\theta\le115^\circ$) the optical torque 
+is working against the optical force, trying to move it 
+out of equilibrium while the optical force returns it to 
+equilibrium. This means that the potential well is more 
+akin to a saddle point, the only reason that the optical
+trap is not escaped immediately is because the the 
+translational trap stiffness is significantly greater than
+the gradient of the optical torque ($\kappa_z\gg\kappa_
+\theta$). Therefore it is more accurate to say that the 
+trap is not metastable, as when the trap is escaped it 
+is not because the Brownian displacement is greater than
+the potential well, rather the dimer is randomly 
+traversing along the x-axis (aka random Brownian 
+rotation) and has escaped the translational potential 
+well.
+
+The force-torque landscape also provides an explanation
+to the boundaries between the different trapping 'basins'
+in fig.~\ref{fig:off_axis}. The first point of interest 
+is the 'dead zone' around the orange basin. Looking at 
+fig.~\ref{fig:force_torque} we can see that around that
+region ($z \approx 2.5\mu$ \& $\theta\approx 110^\circ$)
+the optical torque changes sign as you move closer to 
+the trap focus. There is no clear physical reason why
+the optical torque change direction so abruptly but a 
+consequence of it is that dimers that are too close to 
+this region of the torque landscape are ejected from 
+the optical trap as the are drawn down towards the trap
+focus.
+
+The only pellicular result is the observation that the 
+blue basin can still be reached well below the trap 
+focus. As shown by fig.~\ref{fig:off_axis} there are a 
+handful of starting configurations that end up in the 
+upper trap while in the inverted orientation. Even 
+consulting fig.~\ref{fig:force_torque} does not indicate
+any reason why this occurs. The only conclusion we can 
+draw is that the orange and green basins of attraction 
+are small enough that it is possible for a dimer to 
+avoid both. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:rot_pol}
-In their paper Vigilante \textit{et al} attribute the rotational motion 
-to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
-scattering is a viable theory for describing optical rotation, however
-there is usually an optical axis about which rotation occurs. For a dimer
-this is typically the orientation vector that passes through the centres 
-of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
-cases where the particle's cross sectional shape is engineered to scatter
-light in one particular direction \cite{Higurashi1994}, but if this was
-the case then the dimer should rotate when illuminated by any polarisation
-of light.
+In their paper Vigilante \textit{et al} attribute the 
+rotational motion to the anisotropic shape of the dimer
+\cite{Vigilante2020}. Anisotropic scattering is a viable 
+theory for describing optical rotation, however there is 
+usually an optical axis about which rotation occurs. For 
+a dimer this is typically the orientation vector that 
+passes through the centres of both spheres \cite{Ahn2018, 
+Reimann2018, Bruce2020}. There have been cases where the 
+particle's cross sectional shape is engineered to scatter
+light in one particular direction \cite{Higurashi1994}, 
+but if this was the case then the dimer should rotate 
+when illuminated by any polarisation of light.
 
-To that end, we simulated the motion of an optically trapped dimer in 
-beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
-composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
-run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
-orientational time series; the dimer's orientation is recorded as a 
-quaternion which can be easily converted to a 3-dimensional rotation 
-matrix. By considering only the transverse components ($U_{1,x}$, 
-$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
-the Fourier transformation of their time series reveals the rotational 
-frequency. The laser power is set to $100\ mW$ to avoid large thermal 
-fluctuations and so that the Fourier series of the transverse components 
-approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
-at the rotational frequency $\omega_{rot}$.
+To that end, we simulated the motion of an optically 
+trapped dimer in beams of varying polarisation ($NA=1.2$, 
+$P=100\ mW$). As mentioned previously the beam polarisation
+is given using the Jones vector which describes the magnitude
+of the x and y components of the electric field (so the vector
+$[1.0, 0.0i]$ describes an electric field that only oscillates
+in the x direction). For a linearly polarised beam, both 
+components are in phase with one another, for an elliptical or
+circularly polarised beam we adjust the phase of each component. 
+So a Jones vector of $[1.0, 1.0i]$ describes a beam where the
+y component is $\pi/2$ radians out of phase with the x-component 
+resulting in an electric field that rotates clockwise about
+its direction of propagation. For each simulation we vary
+the phase in increments of $\pi/24$.
 
-%% Example plot of rotating beam trajectory and then the same thing but 
-%% Fourier transformed
-\begin{equation}
-	\begin{split}
-		q(t) \rightarrow R(t) =& 
-		\begin{pmatrix}
-			U_{1,x}(t) && U_{2,x}(t) && U_{3,x}(t) \\
-			U_{1,y}(t) && U_{2,y}(t) && U_{3,y}(t) \\
-			U_{1,z}(t) && U_{2,z}(t) && U_{3,z}(t) 
-		\end{pmatrix} \\
-		\rightarrow
-		&\int^\infty_{-\infty}R(t)e^{-i2\pi ft} dt = 
-		\begin{pmatrix}
-			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f) \\
-			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f)\\
-			\delta(f) && \delta(f) && \delta(f)
-		\end{pmatrix}
-	\end{split}
-\end{equation}
+\begin{figure}[h!]
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{rotating_traj.png}
+	\end{subfigure}
+	$\downarrow$
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{rotating_traj_fourier.png}
+	\end{subfigure}
+	\caption{Top: 3 second trajectory of a dimer trapped by a 
+		circularly polarised Gaussian beam. The far left 
+		column depicts the dimer's centre of mass position 
+		with time; the remaining 3 columns show the 9 components 
+		of the dimers' rotation matrix, with each column being
+		associated with one of its three principal axis. Bottom: 
+		The same trajectory but each time series has been 
+		replaced with its Fourier transform. The only non-zero 
+		elements are for the $U_{1,x}$, $U_{1,y}$, $U_{2,x}$, 
+		and $U_{2,y}$ which show a single peak at the rotation frequency.} 
+	\label{fig:fourier_transform}
+\end{figure}
 
-If the rotational frequency was not immediately obvious the simulation was
-repeated but over a longer simulation time. Four different size ratio of 
-dimers were studied, both in their 'standard' and 'inverted' orientations. 
-The results of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
+Each simulation was run for $1$ second ($\Delta t =10^{-5}\ 
+s$) and at the end we looked at the orientational time series; 
+the dimer's orientation is recorded as a quaternion which can 
+be easily converted to a 3-dimensional rotation matrix. 
+By considering only the transverse components ($U_{1,x}$, 
+$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
+matrix and taking the Fourier transformation of their 
+time series reveals the rotational frequency. The 
+laser power is set to $100\ mW$ to avoid large thermal 
+fluctuations and so that the Fourier series of the 
+transverse components approximates $\delta(\omega_
+{rot}-f)$ - the Dirac delta function centred at the 
+rotational frequency $\omega_{rot}$. This is 
+demonstrated in fig.~\ref{fig:fourier_transform}
+
+If the rotational frequency was not immediately obvious the 
+simulation was repeated but over a longer simulation time. 
+Four different size ratio of dimers were studied, both in 
+their 'standard' and 'inverted' orientations. The results 
+of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
-	\caption{Rotation frequency vs component phase difference for 
-		differently sized dimers. The solid lines represent the 
-		rotation rate experienced while the dimer is in its standard 
-		orientation, whereas the solid points are for the case 
-		where the orientation is inverted. Laser power = $100\ mW$.}
+	\includegraphics[width=0.9\linewidth]{rotation_rate_vs_pol.png}
+	\caption{Rotation frequency vs electric field phase difference 
+		for differently sized dimers. The solid lines represent 
+		the rotation rate experienced while the dimer is in its 
+		standard orientation, whereas the solid points are for 
+		the case where the orientation is inverted. Laser power = 
+		$100\ mW$.}
 	\label{fig:rotation_vs_pol}
@@
 	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
-	\caption{Optical toque experienced by a dimer ($a_{I}/a
+	\caption{Optical toque experienced by a dimer ($a_{I}/a_
 	{II}=10$) and a single isotropic sphere. Both were rotated 
@@
 would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
-to induce rotational motion. What is also interesting is that the 
-rotation rate is not correlated directly with either the particle 
-size or the equilibrium position. This is in stark contrast to previous
-reports of dimer optical rotation; Ahn \textit{et al.} reported that
-the rotational frequency is maximised when the dimer is symmetric 
-\cite{Ahn2018}. Their work was conducted in a vacuum so the only 
-limiting factor to the optical rotational motion is the structural 
-stability of the particle itself.
+to induce rotational motion. It is not clear if the optical torque 
+calculated for the larger size ratios is accurate as you would not 
+expect the rotation rate to plateau like this. From fig.~
+\ref{fig:sphere_dimer_torque} it seems clear that the torque is not
+merely an error of the numerical calculations as it would not have 
+such a strong angular and positional dependence. Most likely with 
+longer simulation times would reveal the real relationship behind
+the rotation rate and the size ratio.
+
+What is also interesting is that the rotation rate is not correlated 
+directly with either the particle size or the equilibrium position. 
+This is in stark contrast to previous reports of dimer optical 
+rotation; Ahn \textit{et al.} reported that the rotational frequency 
+is maximised when the dimer is symmetric \cite{Ahn2018}. Their work 
+was conducted experimentally in a vacuum so the only limiting factor 
+to the optical rotational motion is the structural stability of the 
+particle itself.
 
@@
 		components of the dimers' rotation matrix, with each column being
-		associated with one of its three principal axis. }
+		associated with one of its three principal axis.}
 	\label{fig:gyro}
@@
 spins around its long axis as it did in \ref{sec:rot_pol} \&
-\ref{sec:rot_size}.
+\ref{sec:rot_size}. If the rotation was due to the curl of the 
+spin angular momentum then you would only expect to see 
+precession and nutation, the addition of rotational motion 
+suggests that having multiple particles in close proximity 
+results in a transfer of angular momentum along the long 
+axis.
 
@@
 and $U_{y,y}$
-
-\begin{figure}
+\begin{figure}[h!]
 	\hspace{-2.3cm}
@@
 occurring perpendicular to the optical axis. In the former 
-case the precession can be explained as simply occurring due 
+case the precession was explained as simply occurring due 
 to the reaction torque from the surrounding fluid \cite{Zhu2021}.
@@
 of complex shaped particles' interactions with an optical trap.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Characterisation of asymmetric dimers via PSD \\
-	analysis}
-As discussed in \ref{sec:simulated_QPD}, one of the methods 
-developed to work in conjunction with \cite{Vigilante2020} 
-is a simulated quadrant photo diode for as a position 
-detection system. While it is possible to extract all of 
-the relevant dynamical information from a simulation, 
-confirming the same behaviour in an experimental setting 
-can be challenging if dealing with a non-birefringent 
-anisotropic scatterer. The QPD is composed of 4 photodiodes 
-that measure the intensity of light incident on their 
-surfaces. Using \textit{ott} we can define a region in the 
-far-field that is analogues to the surface of the QPD surface. 
-Evaluating the electric field across this surface gives us 
-an approximation of the QPD signal outputted in an experimental 
-situation. This does not provide a one-to-one result however 
-as hardware errors (i.e. internal resistance, external light 
-sources, and vibrations) would distort the signal somewhat. We 
-can use this as a means to examine the limitations of using 
-back-focal plane interferometry as a means for characterising 
-anisotropic scatterers.
 
-As a benchmark we start by considering a single sphere within 
-an optical trap. A single polystyrene sphere suspended in water 
-($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
-Gaussian beam ($NA=1.25$) using circularly polarised light. For 
-the sake of time efficiency the trajectory was sampled every 10
-time steps, meaning the upper bound on the power spectra is
-$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
-we fitted the power spectra using the aliased Lorentzian 
-(Eq.~\eqref{eq:aliased_lorentzian}).
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{PSD_sphere.png}
-	\caption{Recorded power spectra fitted to eq.~
-		\ref{eq:alaised_lorentzian}, scattered points represents 
-		the blocked data ($n_b=100$). Corner frequency for the 
-		Lorentzian curves are reported in the legend.}
-	\label{fig:psd_sphere}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
-report different corner frequencies which would be indicate 
-that the trap is not perfectly circular. We can use both 
-\textit{ott} and the trajectory itself to derive an estimation 
-of the trap geometry. The corner frequencies and corresponding 
-trap stiffness are reported below:
+\section{Conclusions}
+Considering the simplicity of a scatterer such as a dimer, 
+one would assume that the dynamics of such an object would 
+be relatively easy to predict. Simulations of dimers in 
+the Mie regime show that not only do they have multiple 
+positions and orientations in which they can be trapped but 
+also that their interaction with circularly polarised light 
+is heavily dependent on the axial position and trapping 
+orientation.
 
-\begin{center}
-	\captionof{table}{QPD fitting for single sphere}
-	\begin{tabular}{ |c|c|c|c|c|c|c| } 
-		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
-		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
-		fitting}\\
-		\hline
-		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
-		& \parbox{1.25cm}{\centering 523} & 513 \\
-		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
-		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
-		\hline
-		Ellipticity &
-		\multicolumn{2}{|c|}{8.16 \%} &
-		\multicolumn{2}{|c|}{27.17 \%} &
-		\multicolumn{2}{|c|}{13.8 \%} \\
-		\hline
-
-	\end{tabular}
-\end{center}
-
-Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
-where the variance is based upon our choice of frequency 
-boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
-the ellipticity of the beam is given by $e = (1-\kappa_y/
-\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
-wavefront. Its clear from these initial results that the QPD 
-is more sensitive to changes along the y-axis than the x-axis 
-when compared to the direct \textit{ott} calculations. This is 
-somewhat reflected in the trajectory results. Typically, even 
-an industrial Gaussian beam will produce an elliptical diffraction 
-limit spot when heavily focused; in their tutorial for optimizing 
-the PSD analysis, Berg and Sorensen reported a ellipticity of 
-around 15 \% after a total calibration time of 80 seconds 
-\cite{BergSoerensen2004}.
-
-The reason for the discrepancies between all 3 methods is due 
-to what is actually being measured. The \textit{ott} estimates 
-are simply looking at the differences in the trapping strength 
-along the Cartesian axis'. If calibrated over an long enough 
-time frame you would expect that the resulting power spectra 
-would exactly mirror the \textit{ott} predictions. However over 
-a short calibration time a QPD will only sample a small trajectory 
-that does not fully explore the trapping plane. There is a clear 
-trade off in terms of accuracy and computation time as shorter 
-calibration runs are computationally more efficient but prone to 
-errors. This is the case even if we know the exact positional data.
-
-With this in mind, let us consider a symmetric dimer that is 
-optically trapped by the same Gaussian beam. Not only does 
-the dimer's equilibrium position change but it is subjected 
-to rotational motion due to its unequal moments of inertia. 
-This is reflected in the calibration results using the simulated 
-QPD, where we see a drastically different estimation between the 
-\textit{ott} estimate and the QPD estimate.
-
-\begin{center}
-	\captionof{table}{QPD fitting for symmetric dimer}
-	\begin{tabular}{ |c|c|c|c|c|c|c| } 
-		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
-		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
-		\hline
-		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
-		& \parbox{1.25cm}{\centering 274} & 285 \\
-		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
-		\hline
-		Ellipticity &
-		\multicolumn{2}{|c|}{42.8 \%} &
-		\multicolumn{2}{|c|}{12.7 \%} & 
-		\multicolumn{2}{|c|}{13.8 \%} \\
-		\hline
-	\end{tabular}
-\end{center}
-
-Now we see that the \textit{ott} predicts a more elliptical 
-trap compared to the QPD model which says the trap is far
-more symmetrical while trapping a symmetric dimer. A potential 
-reason that \textit{ott} no longer expects a circular trap 
-could be due to how it computes the beam shape coefficients; 
-by point matching in the far field before the focus means a 
-loss in accuracy for objects that trap above the focus. 
-The change in the QPD estimation can be partially explained 
-by the fact that rotational effects are not accounted for in 
-the Lorentzian power spectra, only translational motion. 
-Typically, rotational motion is only ever detected when it 
-is periodic (take for example Fig.~\ref{fig:vaterite}), when 
-the motion is stochastic the entire power spectra is effected 
-making it near impossible to separate the translational and 
-rotational contributions from a single power spectra.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-Considering the simplicity of a scatterer such as a dimer, one would 
-assume that the dynamics of such an object would be relatively easy 
-to predict. Simulations of dimers in the Mie regime show that not 
-only do they have multiple positions and orientations in which they 
-can be trapped but also that their interaction with circularly 
-polarised light is heavily dependent on the axial position and 
-trapping orientation.
-
-Dimer's have the potential to be used as tunable micro-rotors, being 
-simple to synthesise and can be made out of any material of choice.
-The rotation demonstrated by dimer's is not accurately described in
-previous literature which raises questions on the interactions between
-circularly polarised light and spherical aggregates. Until now, our 
-understanding of angular momentum transfer has been either focused on 
-single particles (where momentum transfer is easily described by 
-Lorenz-Mie theory) or for large aggregates of particles, where the 
-angular momentum transfer between individual particles is not considered. 
-This could provide a greater understanding of angular momentum and help in 
-the development of better torque sensing or torque preventing methods.
-
-Conventional calibration methods will not be adequate however, while
-angular motion can be detected using complicated set ups \cite{Bang2020},
-this can only provide an estimate of the magnitude of a particle's 
-angular motion. This is perfectly fine in cases where the rotational
-motion is stochastic and intuitively predictable (i.e. a dimer in a
-vertical orientation undergoing stochastic Brownian motion), but in 
-the case the the motion is instead periodic characterising it requires
-instantaneous measurements of the particle's motion.
+Dimer's have the potential to be used as tunable micro-rotors, 
+being simple to synthesise and can be made out of any material 
+of choice. The rotation demonstrated by dimer's is not accurately
+described in previous literature which raises questions on the
+interactions between circularly polarised light and spherical 
+aggregates. Until now, our understanding of angular momentum 
+transfer has been either focused on single particles (where 
+momentum transfer is easily described by Lorenz-Mie theory) or 
+for large aggregates of particles, where the angular momentum 
+transfer between individual particles is not considered. This 
+could provide a greater understanding of angular momentum and 
+help in the development of better torque sensing or torque 
+preventing methods.

### 2025-01-26 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

@@
 
-Rotational motion about a single axis is easiest to account for.  
-When the power spectra of elliptical polystyrene particles was 
-fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
-assumed that the rotational motion was purely in the transverse
-plane. As such they did not have to account for any variance in 
-the trapping strength due to orientation nor did they need to
-consider non-periodic rotational behaviour. In the case where 
-rotational motion is stochastic the problem is more complex. 
-For example, when an optical fibre trap characterisation 
-technique was implemented by Saffron \textit{et al} 
-\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
-light scattering to characterise both the axial and lateral 
-trap stiffness acting on microspheres. The only drawback 
-admitted to in their work was that the technique was constrained 
-to isotropic scatters as their theoretical model for describing 
-the auto-correlation function was predicated on the fact that 
-any variations in the signal are due to the particles 
-translational motion within the confines of a cylindrical trap 
-\cite{BarZiv1997}. Where the upper limit of the cylindrical 
-trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
-However, as demonstrated by the results from Chapter 4.1, 
-the axial traps of spherical aggregates is often situated 
-far beyond the Rayleigh range (for a 1.2 NA laser this is 
-$\pm5.985 \mu m$).
+In chapter 4 we showed that while machine learning can be used 
+to accurately predict the translational displacements of an 
+aggregate particle, it fails when trying to account for the 
+rotational behaviour of said particle. In this chapter we 
+build upon \cite{BarZiv1997} but now account for the full 
+rotational motion. This was initially developed in tandem with 
+the work conducted in chapter 3, with the idea being that if a 
+crystal nucleus formed on a rotating sphere how would that 
+nucleus influence the motion of the rotating sphere. Using 
+static light scattering we can map the outputted signal from a 
+optically trapped dimer to its expected orientation in real 
+time. Thus allowing us to estimate the optical torque being 
+applied to the dimer by the trapping beam. As demonstrated 
+in chapter 4, the influence of a circularly polarised beam 
+can result in gyroscopic precession for specifically sized 
+dimers.
 
-To accurately describe the full dynamics of an anisotropic
-scatterer its clear that additional characterisation 
-techniques are necessary. In this chapter we demonstrate 
-two such techniques, quadrant photo-diode simulation, and 
-an optical fibre static light scattering arrangement. The 
-former is based on the work of Rohrbach and Stelzer 
-\cite{Rohrbach2002}, where their work is limited to single 
-spheres, we extend the possible particle space to allow for 
-any theoretical combination of spheres. The results from
-chapter 4 demonstrate that the rotational motion causes 
-significant deviations between the computed trapping strength
-and the expected trapping strength according to Lorenz-Mie
-theory. By utilizing vector regression we can accurately 
-recreate a particle's trajectory simply from the QPD signal, 
-both in the transverse plane and along the beam axis.
+We start by describing a quadrant photo diode system 
+simulated the power spectrum derived from a particle 
+trajectory. We demonstrate that even in idealised 
+circumstances the rotational motion is not revealed 
+from the collected power spectrum. We then go about 
+defining an optical fibre detection system from a 
+mathematical perspective. Before demonstrating how it 
+can be used to instantaneously predict orientational 
+information on a trapped particle. We then consider 
+common factors of error in the characterisation process, 
+such as signal error, and incorrect particle sizing. 
+Both of which have a significant impact on the performance 
+of the model, but by utilising Bayesian inference, and 
+time averaging we can better refine our model.
 
-The latter method builds upon \cite{BarZiv1997} but now 
-accounts for the full rotational motion. This was initially
-developed in tandem with the work conducted in chapter 3,
-with the idea being that if a crystal nucleus formed on a 
-rotating sphere how would that nucleus influence the motion
-of the rotating sphere. Using static light scattering we can 
-map the outputted signal from a optically trapped dimer to 
-its expected orientation in real time. Thus allowing us to 
-estimate the optical torque being applied to the dimer by 
-the trapping beam. As demonstrated in chapter 4, the 
-influence of a circularly polarised beam can result in 
-gyroscopic precession for specifically sized dimers.
+\section{Characterisation of asymmetric dimer dynamics via PSD analysis}
+As discussed in \ref{sec:simulated_QPD}, one of the methods 
+developed to work in conjunction with \cite{Vigilante2020} 
+is a simulated quadrant photo diode for as a position 
+detection system. While it is possible to extract all of 
+the relevant dynamical information from a simulation, 
+confirming the same behaviour in an experimental setting 
+can be challenging if dealing with a non-birefringent 
+anisotropic scatterer. The QPD is composed of 4 photodiodes 
+that measure the intensity of light incident on their 
+surfaces. Using \textit{ott} we can define a region in the 
+far-field that is analogues to the surface of the QPD surface. 
+Evaluating the electric field across this surface gives us 
+an approximation of the QPD signal outputted in an experimental 
+situation. This does not provide a one-to-one result however 
+as hardware errors (i.e. internal resistance, external light 
+sources, and vibrations) would distort the signal somewhat. We 
+can use this as a means to examine the limitations of using 
+back-focal plane interferometry as a means for characterising 
+anisotropic scatterers.
 
-We start by outlining the working principle behind the
-optical fibre detection system, before demonstrating 
-how it can be used to instantaneously predict 
-orientational information on a trapped particle. We then
-consider common factors of error in the characterisation
-process, such as signal error, and incorrect particle 
-sizing. Both of which have a significant impact on the 
-performance of the model, but by utilising Bayesian 
-inference, and time averaging we can better refine our 
-model.
+As a benchmark we start by considering a single sphere within 
+an optical trap. A single polystyrene sphere suspended in water 
+($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
+Gaussian beam ($NA=1.25$) using circularly polarised light. For 
+the sake of time efficiency the trajectory was sampled every 10
+time steps, meaning the upper bound on the power spectra is
+$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
+we fitted the power spectra using the aliased Lorentzian 
+(Eq.~\eqref{eq:aliased_lorentzian}). 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{PSD_sphere.png}
+	\caption{Recorded power spectra fitted to eq.~
+		\ref{eq:alaised_lorentzian}, scattered points represents 
+		the blocked data ($n_b=100$). Corner frequency for the 
+		Lorentzian curves are reported in the legend.}
+	\label{fig:psd_sphere}
+\end{figure}
+
+As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
+report different corner frequencies which would be indicate 
+that the trap is not perfectly circular. We can use both 
+\textit{ott} and the trajectory itself to derive an estimation 
+of the trap geometry. Additionally, we fitted the 
+same Lorentzian to the sphere's positional data, this would be a
+situation when the QPD signals are completely uncorrelated with 
+one another and there is a constant ratio between the QPD signal 
+and the sphere's position. The corner frequencies and corresponding 
+trap stiffness are reported below:
+
+\begin{center}
+	\captionof{table}{QPD fitting for single sphere}
+	\begin{tabular}{ |c|c|c|c|c|c|c| } 
+		\hline
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
+		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
+			fitting}\\
+		\hline
+		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
+		& \parbox{1.25cm}{\centering 523} & 513 \\
+		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
+		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
+		\hline
+		Ellipticity &
+		\multicolumn{2}{|c|}{8.16 \%} &
+		\multicolumn{2}{|c|}{27.17 \%} &
+		\multicolumn{2}{|c|}{13.8 \%} \\
+		\hline
+
+	\end{tabular}
+\end{center}
+
+Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
+where the variance is based upon our choice of frequency 
+boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
+the ellipticity of the beam is given by $e = (1-\kappa_y/
+\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
+wavefront. Its clear from these initial results that the QPD 
+is more sensitive to changes along the y-axis than the x-axis 
+when compared to the direct \textit{ott} calculations. This is 
+somewhat reflected in the trajectory results. Typically, even 
+an industrial Gaussian beam will produce an elliptical diffraction 
+limit spot when heavily focused; in their tutorial for optimizing 
+the PSD analysis, Berg and Sorensen reported a ellipticity of 
+around 15 \% after a total calibration time of 80 seconds 
+\cite{BergSoerensen2004}.
+
+The reason for the discrepancies between all 3 methods is due 
+to what is actually being measured. The \textit{ott} estimates 
+are simply looking at the differences in the trapping strength 
+along the Cartesian axis'. If calibrated over an long enough 
+time frame you would expect that the resulting power spectra 
+would exactly mirror the \textit{ott} predictions. However over 
+a short calibration time a QPD will only sample a small trajectory 
+that does not fully explore the trapping plane. There is a clear 
+trade off in terms of accuracy and computation time as shorter 
+calibration runs are computationally more efficient but prone to 
+errors.
+
+With this in mind, let us consider a symmetric dimer that is 
+optically trapped by the same Gaussian beam. Not only does 
+the dimer's equilibrium position change but it is subjected 
+to rotational motion due to its unequal moments of inertia. 
+This is reflected in the calibration results using the simulated 
+QPD, where we see a drastically different estimation between the 
+\textit{ott} estimate and the QPD estimate.
+
+\begin{center}
+	\captionof{table}{QPD fitting for symmetric dimer}
+	\label{tab:dimer}
+	\begin{tabular}{ |c|c|c|c|c|c|c| } 
+		\hline
+		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
+		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory fitting} \\
+		\hline
+		$f_c\ [Hz]$ & 445 & 409 & \parbox{1cm}{\centering 431} & 424 
+		& \parbox{1.25cm}{\centering 274} & 285 \\
+		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.22 & 9.16 & 7.82 & 7.91  \\
+		$\kappa\ [pN/\mu m]$ & 52.82 & 48.54 & 51.13 & 50.26 & 32.45 & 33.75 \\
+		\hline
+		Ellipticity &
+		\multicolumn{2}{|c|}{28.5 \%} &
+		\multicolumn{2}{|c|}{12.7 \%} & 
+		\multicolumn{2}{|c|}{13.8 \%} \\
+		\hline
+	\end{tabular}
+\end{center}
+
+Now we see that the \textit{ott} predicts a more elliptical 
+trap compared to the QPD model which says the trap is far
+more symmetrical while trapping a symmetric dimer. A potential 
+reason that \textit{ott} no longer expects a circular trap 
+is because that unlike a sphere, the force displacement 
+curve is not strictly harmonic. If we consider a dimer that 
+is some distance from the beam axis, the sphere closest to 
+the trap focus will experience a slightly greater force 
+compared to the sphere further from the trap. As such its not
+accurate to assume that the external force $F(x) = \kappa x$
+instead we must consider that it is a function of both the 
+position and orientation simultaneously.
+
+This coupling of the translational and orientational motion 
+has been highlighted previously \cite{Vigilante2020}, but 
+there effects have not been demonstrated in the context of
+an experimental situation. We can see from the difference in
+trajectory and QPD fittings that while our description of the
+trap shape is very similar (both methods give similar ellipticity
+values) but the magnitude of the trap strength are significantly
+different. This is partially explained by the fact that the
+QPD is not actually measuring the position of the particle
+but instead the intensity distribution of the total field 
+incident on the QPD surface. In which case if the dimer is 
+rotated slightly the QPD signal will change even if the dimers
+centre of diffusion remains stationary.
+
+\subsection{QPD for angular displacement detection}
+The next logical step is to consider whether or not a QPD can at 
+all be utilised for detection of rotational motion. This has been
+shown for nano-particles that exhibit periodic rotational motion
+using by considering the difference in diagonal quadrants. The 
+benefit of detecting rotational motion is that one can begin to 
+build a better understanding about how angular momentum is 
+transferred to a dimer which could be extended to more complex
+particles.
+
+At just a cursory glance it would appear that there is a clear
+relationship between the QPD signal and the orientation of the 
+dimer. If we look at the QPD signal produced by a dimer as it is 
+rotated we can see that it is equally sensitive to rotations in
+either the X-Z or Y-Z plane but there is no change in the signal 
+when rotated by in the X-Y plane.
+
+However if we compare this to fig.~\ref{fig:rohrbach} we can see 
+that the the signal change is dwarfed by the translational motion. 
+So while their is a clear relationship between the two trying to
+discern between translational and rotational contributions to the
+QPD signal is not a simple task. This is significant because it 
+means that with regards to the power spectrum fittings in table 
+\ref{tab:dimer} there is no way of ensuring that the measured 
+trap strengths are accurate. The variance $\sigma(f_c)$ is only 
+the variance based upon our fitting parameters and not on the 
+actual dynamics of the dimer. In the case where the particle's
+shape is not known prior its not possible to determine whether 
+any inaccuracies in the power spectra are due to the fit or due
+to the particle's motion.
+
+The dependence of $S_x$ and $S_y$ and the dimer's orientation
+(defined by the spherical angles $\theta$, $\phi$) is therefore
+not clear. To that end, we utilised machine vector regression,
+which takes as an input the voltage from the 4 quadrants and 
+tries to fit that to the dimer's trajectory. Fitting the QPD
+signal to the dimer's position shows promising results, shown in
+fig.~X.X is a prediction of a symmetric dimer's position based
+on the QPD signal. With the actual displacement on the x-axis 
+and the predicted result y-axis, and the dashed line represents 
+the y=x line - indicating an ideal prediction.
+
+Fig~X.X shows that its relatively trivial to predict a particle's
+position based off the QPD signal, even in axial direction which
+has not been done previously. It should be noted that these 
+displacements are relative to the trap focus and so accurate 
+tracking of the beam movements needs to be taken. This result is
+partially due to the fact that displacement and signal units are
+linearly related by \eqref{eq:conversion_factor}. Interestingly 
+when we run the same protocol for different sized dimer's there 
+seems to be a cut off point when the fit begins to fail. Plotting
+the $R^2$ for each fit vs its size ratio shows that beyond a size
+ratio of 5 the machine learning begins to fall off significantly.
+
+This is surprising because we would expect that as the second 
+sphere shrinks the dynamics should more closely approximate that 
+of a single sphere. We can compare our
 
@@
 of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
-requires a target origin for setting the scattering expansion, as
-with our simulations we set it as the dimer's centre of diffusion.
+requires a target origin for setting the scattering expansion, 
+there is no agreed upon solution for where the target origin 
+should be relative to the sphere positions. In chapter 4 we showed
+that altering the target origin would drastically alter the 
+tmatrix produced
 
@@
 intensity distribution across the entire detector will decrease
-substantially. Because of the fall off it is more computationally 7
+substantially. Because of the fall off it is more computationally
 efficient to instead consider only the central pixel of the 
@@
 
-By moving the detectors further back we can treat each detector 
-as single point and compute the intensity at the centre of the 
-detector rather than consider the full surface. For computing 
-the instantaneous orientation we only consider the intensity 
-at the centre of the detector.
+Given the trapping beam's beam waist is approximately $0.54 \mu m$,
+a optical fibre would need to be a minimum of $150 \mu m$ from the 
+trap centre in order to prevent noise from the trapping laser itself.
+At such a distance the variation in the signal is so small that it is 
+easier to evaluate the intensity at the centre of the detector instead
+of across the entire surface. This is how the instantaneous orientation
+is predicted later on in the chapter.
 
@@
 We now discuss the methodology for computing the instantaneous 
-orientation of an optically bound particle based on scattering
+orientation of an optically trapped particle based on scattering
 detected by a finite grouping of optical fibres.
@@
 neighbour spacing between two neighbouring reference 
-orientations of 0.895 radians. Using \textit{mstm} we 
-compute the raw intensities at each of the measurement angles 
+orientations of 0.895 radians.
+
+Using \textit{mstm} we compute the raw intensities (see 
+\eqref{eq:intensity}) at each of the measurement angles 
 that would be generated by a dimer in each reference 
-orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the 
-number and position of detection fibres is technically arbitrary 
-there are several constraining factors that limit our ability 
-to infer useful information from the trapped object, see Section~\ref{sec:brownian} for a detailed breakdown of our 
-choice of detection angles. In order for a neural network
-to properly weight each signal the raw intensities are 
-normalized according to:
+orientation, $I_s(\hat{\bf n}_{\alpha}, \theta_k)$. While 
+the number and position of detection fibres is technically 
+arbitrary there are several constraining factors that limit 
+our ability to infer useful information from the trapped 
+object, see Section~\ref{sec:brownian} for a detailed 
+breakdown of our choice of detection angles. In order for 
+a neural network to properly weight each signal the raw 
+intensities are normalized according to:
 \begin{align}
@@
 instances the mapping becomes far more complex to discern.
-\newpage 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.8\textwidth}
+	\begin{subfigure}{0.74\textwidth}
 		\subcaption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.8\textwidth}
+	\begin{subfigure}{0.74\textwidth}
 		\subcaption{}
@@
 \end{figure}
-\newpage
+
 Nevertheless, at least where the uncertainty in signal measurements 
@@
 is. With larger values of $D_{KL}$ indicating that our prediction is 
-spread out so the correct orientation is unsure about $\hat{\bf n}_
-{best}$. Whereas a low value of $D_{KL}$ means that our model is more 
-confident in $\hat{\bf n}_{best}$. The only draw back being that the 
-relative entropy does nothing to indicate accuracy, only how closely 
-our distribution matches our ideal result. This will be better 
+spread out so the neural net is unsure about $\hat{\bf n}_{best}$. 
+Whereas a low value of $D_{KL}$ means that our model is more confident 
+in $\hat{\bf n}_{best}$. The only draw back being that the relative 
+entropy does nothing to indicate accuracy, only how closely our 
+distribution matches our ideal result. This will be better 
 exemplified later in \ref{sec:test}.
@@
 
+\subsection{Choice on the number of detectors}
 When all of the detectors lie in the same plane the expected signal 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Testing the Neural Network}
+\section{Testing the Neural Network}
 \label{sec:test}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Accounting for sources of error in light scattering measurements}
+\section{Accounting for sources of error in light scattering measurements}
 When it comes to analysing light scattering from any size particle,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subparagraph{Impact of incorrect dimer sizing}
+\subsection{Impact of incorrect dimer sizing}
 \label{sec:lam}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subparagraph{Impact of measurement noise on model predictions}
+\subsection{Impact of measurement noise on model predictions}
 \label{sec:epsilon}
@@
 
-From the above discussion it’s clear that estimation of the 
-dimer’s orientation is a problem that can be endlessly tuned 
-to fully maximise our end result. Here we simplify the problem 
-somewhat by employing a relatively small finite number of 
-'reference orientations' to map between scattering and dimer 
-orientation: the precision of estimation could be improved by 
-utilising a greater number of reference orientations, although 
-there remains a balance between the realisable precision of 
-orientation estimate and the noise level of the scattering 
-measurement. Another avenue to further explore would be using 
+Overall it’s clear that estimation of the dimer’s orientation 
+is a problem that can be endlessly tuned to optimise the 
+result. Here we simplify the problem somewhat by employing 
+a relatively small finite number of 'reference orientations' 
+to map between scattering and dimer orientation: the precision 
+of estimation could be improved by utilising a greater number 
+of reference orientations, although there remains a balance 
+between the realisable precision of orientation estimate and 
+the noise level of the scattering measurement. 
+Another avenue to further explore would be using 
 the method to optimise the choice of detection angles, 
@@
 be to consider more complex correlations based on prior 
-expectations of the dynamics. Here already we improve the 
+expectations of the dynamics. The Debye equation for  Here already we improve the 
 method using a non-uniform prior based on only the immediately 
@@
 dimer’s preferred axis of rotation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{QPD Simulations}
-A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.
-
-\subsection{Incident beam}
-\label{sec:fibre_scattering}
-The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam).
-*Ott* uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
-
-\begin{align}
-	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
-\end{align}
-
-Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, *ott* allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD.
-For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully.
-In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
-
-We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle *ott* uses the t-matrix method, this is not essential for a simple sphere but is far more important for complex shaped particles such as our dimers.
-If the T-matrix is loaded in from *mstm* we need to convert the *mstm* t-matrix to a form more suitable for *ott*:
-
-\begin{align}
-	\begin{pmatrix}
-		p_{nm}\\
-		q_{nm}
-	\end{pmatrix} =T 
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-		bT^{TM}_{nm} & bT^{TE}_{nm}
-	\end{pmatrix}
-	\begin{pmatrix}
-		a_{nm}\\
-		b_{nm}
-	\end{pmatrix}
-\end{align}
-
-For *mstm* the T-matrix is packed as a column vector:
-
-\begin{align}
-	T_{MSTM} = 
-	\begin{bmatrix} 
-		aT^{TE}_{n,-n} & bT^{TE}_{n,-n} \\ 
-		aT^{TE}_{n, -n+1} & bT^{TE}_{n, -n+1} \\ 
-		... & ...\\ 
-		aT^{TE}_{n,n} & bT^{TE}_{n,n} \\ 
-		----&----\\ 
-		bT^{TM}_{n,-n} &bT^{TM}_{n,-n} 
-	\end{bmatrix}
-\end{align}
-
-Where as *ott* packs packs the T-matrix with sub matrices:
-
-\begin{align}
-	T_{Ott} = 
-	\begin{bmatrix} 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n} & aT^{TE}_{n,-n}\\ 
-			bT^{TM}_{n,-n}&bT^{TE}_{n,-n}
-		\end{pmatrix} \\ 
-		\begin{pmatrix}
-			aT^{TM}_{n,-n+1} & aT^{TE}_{n,-n+1}\\ 
-			bT^{TM}_{n,-n+1} & bT^{TE}_{n,-n+1}
-		\end{pmatrix}\\
-		....\\ 
-		\begin{pmatrix}
-			aT^{TM}_{nm} & aT^{TE}_{nm}\\ 
-			bT^{TM}_{nm} & bT^{TE}_{nm}
-		\end{pmatrix}
-	\end{bmatrix}
-\end{align}
-
-\subsection{Scattered and Total Fields}
-With the T-matrix in hand we can compute the scattered beam by multiplying our beam shape coefficients with the T-matrix to get out the scattered field.
-Now in order to simulate a real QPD we need to account for the motion of our target particle within the trap, taking a typical trajectory file we read off each line in order to translate and rotate the beam.
-Translation is a rather simple process, simply involving us to shift the beam laterally, small deflections are generally unnoticeable for the incident beam but are much more noticeable in the scattered field (the below figure shows the result of shifting the incident beam $1\mu m$ to the right):
-
-The QPD does not just pick up the scattered light however, it instead is receiving a combined signal from both the incident field and scattered field simultaneously, as mentioned by \textcolor{red}{Rohrbach} the total intensity can be computed by taking the magnitude of both the incident field *focused at the origin $[0,0,0]$* and the scattered field originating from $[\delta x, \delta y, \delta z]$ (top left and bottom right plots in the above figure).
-This means we do not need to worry about any translation effects being 'double-counted' in the QPD's signal as we are only shifting the scattered field meaning the QPD signal is only picking up the interference due to the shifted scattered field.
-I conducted some unitary tests where I scanned the beam position laterally along the x-axis and measured the QPD's 'x-signal' for x, circular and y-polarized light which yielded the following QPD responses.
-The target particle was $1.57\ \mu m$ and the scan range is set to $[-4, 4 \mu m]$
-
-Where $S_x$ is given in blue, and $S_y$ is given in orange, the plots make it clear that the polarisation of the beam have a minimal effect on the QPD signals if the particle is traversing in one direction, its clear that the displacement from the beam centre is far more important than the polarization of said beam.
-This is backed up when we look at \textcolor{red}{Rohrbach's} results, who studied a $150\ nm$ sphere ($n=1.57$) submerged in water with a focusing lens of numerical aperture 1.2, and beam power of $3\ mW$.
-The condensing lens' numerical aperture was not set to a particular value and was instead varied between 0.13-1.2, as a compromise we selected a condenser numerical aperture of 0.525, corresponding to a acceptance angle of $31.6^\circ$.
-
-Where the points are Rohrbach's results and the solid line's are our QPD's own replication.
-The dashed horizontal lines on the left represent the maximum displacement in the lateral displacement which is given by the combined beam radius and particle radius (assuming a beam radius of $0.54\ \mu m$); we can say that any displacements beyond this distance, while non-zero these displacements are unlikely to occur while a particle is trapped at the focus.
-Whereas the right most plot's dashed horizontal lines represents the Rayleigh range of the Gaussian beam, this represents the transition between plane wave and spherical wave regimes.
-Interestingly while our results close to Rohrbach's while close to the focus we see it begins to diverge beyond the first peak, this shouldn't be an issue for a typical optical tweezer calibration as we can assume that the maximum displacement will be within this linear regime.
-
-Now the above plots only consider the QPD response to movements along the cartesian axis, however obviously for any Brownian motion the movements are a combination of displacements in each cartesian direction.
-We might assume naively that any displacement $\Delta r$ will result in a linear combination of QPD responses; for low precision force measurements this assumption is adequate, however when high precision is required we find that this assumption is longer adequate due to something referred to as cross-talk.
-Cross talk arises when movement in one direction results in a QPD response change in the other orthogonal direction, there is no one reason for this effect, it could be a result of differing sensitivity in the photo diodes, it could be because the scattering is slightly asymmetric meaning the scattering falls outside the QPD, or it could be a result of mis-alignment in the set up.
-This can have unintended effects, for example it may lead to the apparent rise of a curved trajectory rather than a straight path: Consider a particle moving purely along the x-axis, with 0 cross-talk the QPD response should perfectly match the above curve $S_x(x,0,z_0)$, with $S_y$ being flat in comparison; if however there is cross talk between the channels then $S_y$ will have some significant non-zero value (or it may even grow with increasing displacement), implying that the particle is actually moving in both the x and y directions simultaneously.
-First we checked for this by measuring the QPD response for random positions within the XY plane:
-
-Where $S_x$ is plotted on top and $S_y$ is plotted below, as shown by the above plot we see that they still possess similar shapes to the previous plots but now with additional noise terms, making it clear that for any trajectory there will be cross talk.
-This means that trying to get a one-to-one measurement of the particle's displacement is not possible by simply looking at the QPD response, to do that we can need to calibrate the trap.
-
-\textcolor{red}{Berg and Flyvbjerg} have an excellent breakdown for accurately calibrating an optical tweezer, in addition they discuss how to minimise cross-talk effects.
-For two correlated power spectra, the cross correlation is given as:
-
-\begin{align}
-	P_{x} = |\hat{S}_x(f)|^2 \\ 
-	P_{y} = |\hat{S}_y(f)|^2 \\
-	\rightarrow P_{xy} = Re(\hat{S}_x\hat{S}^*_y)
-\end{align}
-
-Now if the two directions are correlated then $|P_{xy}|^2/P_xP_y$  should be non-zero for all frequencies, in order to eliminate cross-talk effects we need to minimise the cross-correlation for all frequencies.
-They showed that it is possible to find a transformation of the time series $(x(t), y(t))$  to one that possesses the property that $P_{x'y'}(f)=0$ for all frequencies.
-They found these transformed positions by minimising the sum cross-correlation:
-
-\begin{align}
-	\sum\frac{P_{x'y'}}{P_{x'}P_{y'}} = \sum\frac{(1+bc)P_{xy}+cP_x+bP_y}{(P_x+2bP_{xy}+b^2P_y)(Py+2cP_{xy}+c^2P_y)}
-\end{align}
-
-Where b and c are fitting parameters, by minimising this function one can adjust each spectra in order to eliminate cross talk effects and provide a more accurate calibration of the optical trap.
-Furthermore with the fitting completed, the time series can be then transformed in order to eliminate the cross talk effects:
-
-\begin{align}
-	x'(t) = S_x(t) + bS_y(t),\ y'(t) = S_y(t) + cS_x(t)
-\end{align}
-
-Where $x'$ and $y'$ are now uncoupled coordinates that when Fourier transformed provide the uncorrelated power spectrum.
-With the fitting complete we can now adjust our time series in order to get a replication of the lateral trajectory.
-
-\section{Rotations and Asymmetric particles}
-Rotational motion is something that is not often necessarily considered when characterising Brownian motion, often because separating the contributions from rotational and translational motion is challenging.
-Depending on the rotational and translational trap stiffness the collected QPD signal will be aliased and typical calibration techniques cannot characterise the particle and trap interactions.
-This is often why most of the research into trapping asymmetric objects does not delve into power spectra analysis, as there is no real way of modelling the QPD response from an asymmetric object.
-
-Now for isotropic scatterers any rotation is often not an issue when it comes to QPD analysis, even if a sphere rotates $180^{\circ}$the scattering should be identical (assuming its relative position is the same).
-Now *ott* deals with both translations and rotations by moving the beam itself and computing the scattering from by once again expanding the spherical wave functions, the problem that arises from this is that the portion of the spherical wave that is evaluated by the QPD will pick up said rotation and produce a non-zero signal even if the scatterer in question is isotropic.
-This means we need to counter rotate the total field prior to collecting the QPD signal, this captures the effects of an anisotropic scatterer but has no effect for an isotropic scatterer.
-As a test the QPD response from an isotropic sphere of $1.57 \ \mu m$ was collected at multiple angles, between 0 and $2\pi$, then as a comparison a symmetrical and 1:2 dimer were also subjected to the same rotations and their QPD responses evaluated.
-
-Where on the left we have an isotropic sphere, then a symmetrical dimer, and lastly a 1:2 dimer.
-The left most plot shows that simply using the inverted rotation matrix is sufficient to prevent rotation effects being double counted.
-For dimers however, rotating about a given axis gives us a clear change in the signal for even a slight rotation about any axis.
-
-\section{Power Spectra of single sphere vs spherical aggregates}
-In order to verify that the QPD is outputting correct signals we first compared the module to the only other known instance of simulating a QPD response, that being the work of \textcolor{red}{Rorhbach}.
-In their paper they looked at the signal response produced by a single sphere ($r = 150\ nm$, $n = 1.57$) being scanned along the three Cartesian axis in the proximity
-of a $1064\ nm$ NIR beam with a numerical aperture of 1.2.
-As shown by fig~\ref{fig:Rohrbach} the two responses are remarkably accurate close to the origin; with only slight deviations as you move past the harmonic region, the difference can be explained due to how the two evaluate the total fields.
-Where in our case the incident and scattered fields are reduced to 0 at the origin, whereas Rohrbach's work evaluates the fields the fields as being non-zero at all points, as you move beyond the origin this discrepancy grows.
-
-Since we are only interested in the dynamics of spherical aggregates upon reaching equilibrium we do not need to be concerned about the deviation from Rohrbach's results so long as the particle's displacement does not exceed the linear region at the origin.
-The axial response curve is slightly more involved, requiring that the condenser lens numerical aperture is adjusted until a harmonic response curve is achieved \cite{Friedrich2012}.
-
-In order to verify that the QPD can accurately capture the dynamics of the target particle we first simulated a single sphere ($a = 1\mu\ m$, $n=1.59$) in the focus of a $1064\ nm$ laser.
-Using \textit{ott} the trap stiffness is computed by applying a linear fit to the force-displacement curve along the x and y axis'; for a linearly polarised trap the gradient force is greater in the polarisation direction than transverse case.
-We see this reflected in the simulated QPD, the fitted Lorentzian curves have different corner frequencies indicating that the beam's polarisation is influencing the dynamics of the trap.
-
-Next we consider a dimer composed of two sphere's similar to the previous sphere, we again apply the QPD module to reproduce the power spectrum.
-Previous experimental work found that a pair of trapped beads would half the corner frequency from the reported power spectra.
-However, in our own simulations we see instead an increase in the expected trapping force from our QPD method, with the maximum trap stiffness expected for a dimer of size ratio 5.
-Compared to \textit{ott} in which the maximum trapping strength is expected at a size ratio of 2.
-This can be somewhat explained by the fact that as the second sphere shrinks the centre of diffusion approaches the centre of the larger sphere and thus the dimer's orientational behaviour is similar to a single sphere, only subject to slight Brownian motion.
-Therefore one should expect that for aggregates containing multiple sized particles (e.g.
-cellular samples) typical characterisation techniques may over estimate the strength of the trap.
 
@@
 dynamic information about complex trapped entities.
+
+Conventional calibration methods will not be adequate however, while
+angular motion can be detected using complicated set ups \cite{Bang2020},
+this can only provide an estimate of the magnitude of a particle's 
+angular motion. This is perfectly fine in cases where the rotational
+motion is stochastic and intuitively predictable (i.e. a dimer in a
+vertical orientation undergoing stochastic Brownian motion), but in 
+the case the the motion is instead periodic characterising it requires
+instantaneous measurements of the particle's motion.

### 2025-04-15 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 15/04/25

@@
 where a dilute phase is miscible in a bulk phase, more 
-often called the solute and solvent respectively. Because 
-of thermodynamics, the two can only remain in equilibrium 
-while below a specific concentration ($C_{eq}$) - below 
-which the chemical potential $\mu$ for a miscible solution 
-is greater than the potential required to separate the two 
-phases. Once $C_{eq}$ is exceeded there is a chemical potential 
+often called the solute and solvent respectively. When 
+a small amount of solute is added to the solvent the 
+solute will diffuse throughout the solvent, reaching
+an equilibrium concentration throughout the solution. 
+Because of molecular interactions, the two can only remain 
+in equilibrium while below a specific concentration ($C_{eq}$) 
+- below which the chemical potential $\mu$ for a single phase 
+is greater than the potential required to separate the two. 
+Once $C_{eq}$ is exceeded there is a chemical potential 
 difference driving the solution to separate the two phases. 
@@
 Where $C_{sol}$ is just the solute concentration, and 
-$C_{eq}(T)$ is the equilibrium concentration at temperature 
-$T$. While the solution remains supersaturated there 
-is a chemical potential driving the solute to coalesce 
-and separate from the solution as an ordered solid, the 
-first formation of the crystal is referred to as the 
+$C_{eq}(T)$ is the equilibrium concentration of the solute  
+at temperature $T$ based on the main solvent. While the 
+solution remains supersaturated there is a chemical 
+potential driving the solute to coalesce and separate 
+from the solution as an ordered solid, the first 
+formation of the crystal is referred to as the 
 nucleus and understanding its formation has been the 
@@
 Primary nucleation is therefore considered stochastic as there 
-is no reliable means of predicting where a nucleus may form, 
-or how long it will take. The only reliable indicator being 
-that higher supersaturations will result in faster nucleation
-rates. At a small scale one can estimate the nucleation rate 
-by making repeated measurements of sample solutions and seeing
-how many have nucleated after a given time, giving us a Poisson
-probability distribution.
+is no means of predicting where a nucleus may form, or how long 
+it will take. The only reliable indicator being that higher 
+supersaturations will result in faster nucleation rates. At a 
+small scale one can estimate the nucleation rate by making 
+repeated measurements of sample solutions and seeing how many 
+have nucleated after a given time, giving us a Poisson probability distribution.
 \begin{align}
 	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
+	\label{eq:poisson}
 \end{align}
@@
 $M$ are the number of nucleated samples and the total number 
-of samples used respectively. While this is useful for studying
-the effects of different parameters at a small scale, for
-industrial applications there are to many external factors 
+of samples used respectively. The nucleation rate is one of 
+the most important factors used in evaluating a crystallisers 
+performance, providing a rough estimation of the final product 
+mass over a given time frame. While \eqref{eq:poisson} is useful 
+for studying the effects of different parameters at a small scale, 
+for industrial applications there are to many external factors 
 for primary nucleation to be measured accurately.
@@
 \section{Nucleation Theories}
-Several potential theories have been proposed to explain 
-how nucleation occurs at a microscopic scale. In doing so 
-we could potentially predict the expected yield of a given
-crystalliser based on the initial conditions of the solution. 
-Outlined below are some of the most popular theories currently 
-used to describe nucleation.
+In order to predict the nucleation rate of a given solution
+we need a solid understanding of how the individual molecules
+interact such that they transition from an unordered phase to
+a ordered phase. Over the years several theories have been 
+proposed to explain how nucleation events occur, so far there
+has yet to be a standout candidate that reliably predicts 
+the systems behaviour. Either they fail to accurately predict
+nucleation rates or more generally cannot describe new 
+experimental results.
 
@@
 
-Consider a supersaturated solution, after some time 
-enough individual sub units collide, forming a nucleus 
-of volume $4\pi r^3/3$. The newly formed phase has a 
+Consider a supersaturated solution, at some time individual 
+sub units collide rapidly, one after another, forming a 
+nucleus of volume $4\pi r^3/3$. The newly formed phase has a 
 lower chemical potential than the surrounding solution, 
@@
 Where $A$ is a pre-factor that can be fine tuned to the exact 
-demands of the system, the free energy barrier can be found by finding 
-the stationary point of $\Delta G_{tot}$.
+demands of the system, the free energy barrier can be found by 
+finding the stationary point of $\Delta G_{tot}$.
 
-CNT is often regarded as a good description of the macro system, its 
-obvious that for all crystallization systems there is an inherent energy 
-barrier that dictates the nucleation rate. Where it falters is in its 
-predictive ability, both in estimating nucleation rates \cite{Gharibeh2005, 
-Vekilov2010}, and in the structure of newly formed nuclei \cite{Lee1999, 
-Yau2001}. Recent studies suggest classical nucleation is merely one of many 
-possible pathways that can be taken to produce a structured crystalline phase. 
-Prompting the development of alternative theories to better describe the 
-nucleation process.
+CNT is often regarded as a good description of the macro system, 
+its obvious that for all crystallization systems there is an 
+inherent energy barrier that dictates the nucleation rate. 
+Where it falters is in its predictive ability, both in estimating 
+nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
+structure of newly formed nuclei \cite{Lee1999, Yau2001}.
+Recent studies suggest classical nucleation is merely one of 
+many possible pathways that can be taken to produce an ordered 
+crystalline phase. Prompting the development of alternative 
+theories to better describe the nucleation process.
 
@@
 that any interactions between solute molecules are much
-smaller than the surface tension due to the bulk solution.
+weaker than that between solute and solvent molecules.
 If instead the solute molecule interactions are significant
@@
 characterisation techniques to identify the pathway taken. The 
-method not only needs to be repeatable fora single set of system 
+method not only needs to be repeatable for a single set of system 
 parameters, but also needs to be flexible enough to replicate 
@@
 
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since 
-the 1980s when Ashkin \cite{Ashkin1970} first showed that focused 
-light was capable of trapping micron sized particles due to light 
-exerting 'radiation pressure'. The working principle was that a 
-light source such as a laser could trap small objects within a 2D 
-plane, as long as the light source had an approximately Gaussian 
-profile (colloquially called a Gaussian beam). Soon after, Ashkin 
-showed that the introduction of a microscope objective would allow 
-one to focus the light source to a diffraction limited point that 
-would stably trap small objects within a confined volume 
-\cite{Ashkin1980}. This allowed Ashkin and others to study individual 
-biological cells as the overall force was on the order of $10^{-12}\ 
-N$ while being non-invasive to the internal structure. Later it would 
-be used to probe microscopic properties such as the formation of 
-colloidal aggregates \cite{Yi2021} to the drag forces exerted by a 
-pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
-behaviour of light, optical tweezers have become essential for 
-measuring and exerting precise forces on the magnitude of pico-newtons 
-allowing one to probe the material properties of the smallest 
-materials.
-
-\subsection{Literature related to laser induced nucleation}
+\subsection{Laser induced nucleation}
 From as early as 1996 it has been known that laser irradiation 
@@
 Lastly, there is trapping induced nucleation, this is where optical 
-tweezers come into play. Due to the radiation pressure created by 
-the focused beam, it is possible to manipulate the solute, this was 
-demonstrated with amino acids such as glycine \cite{Tsuboi2009}. 
+tweezers come into play (see below). Due to the radiation pressure 
+created by the focused beam, it is possible to manipulate the solute, 
+this was demonstrated with amino acids such as glycine \cite{Tsuboi2009}. 
 Whether or not a crystal forms is due to the location of the laser 
@@
 the optical trap, as such it would suggest that larger aggregates 
-are being manipulated. Applying DLS analysis to the dense liquid 
-region showed that it was populated by clusters that would 
-consolidate together upon being focused by the optical trap 
+are being manipulated. Applying dynamic light scattering analysis to 
+the dense liquid region showed that it was populated by clusters that 
+would consolidate together upon being focused by the optical trap 
 \cite{Gowayed2021}. Molecular simulations of glycine solutions 
@@
 already in common use for micro-rheological studies \cite{Bishop2004, 
-RobertsonAnderson2018}
+RobertsonAnderson2018}.
+
+\section{Optical Tweezers}
+\subsection{Background}
+Optical tweezing has been a field of applied optics ever since 
+the 1980s when Ashkin \cite{Ashkin1970} first showed that focused 
+light was capable of trapping micron sized particles due to light 
+exerting 'radiation pressure'. The working principle was that a 
+light source such as a laser could trap small objects within a 2D 
+plane, as long as the light source had an approximately Gaussian 
+profile (colloquially called a Gaussian beam). Soon after, Ashkin 
+showed that the introduction of a microscope objective would allow 
+one to focus the light source to a diffraction limited point that 
+would stably trap small objects within a confined volume 
+\cite{Ashkin1980}. This allowed Ashkin and others to study individual 
+biological cells as the overall force was on the order of $10^{-12}\ 
+N$ while being non-invasive to the internal structure. Later it would 
+be used to probe microscopic properties such as the formation of 
+colloidal aggregates \cite{Yi2021} to the drag forces exerted by a 
+pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
+behaviour of light, optical tweezers have become essential for 
+measuring and exerting precise forces on the magnitude of pico-newtons 
+allowing one to probe the material properties of the smallest 
+materials.
 
@@
 \label{sec:opt_torque}
-It has long been known that electromagnetic fields can transfer
-linear and angular momentum \cite{Beth1936MechanicalDA}; more 
-accurately the field is said to have both orbital and spin momentum. 
-Though there is some debate on how to decompose the total momentum 
-into these two components \cite{Bruce2020, Svak2018}, for this 
-project we do not need to calculate the exact quantities and will 
-instead look at the broader effects of both components. Orbital 
-angular momentum arises from the shape of the wavefront of the 
-particular field in question; for simple Gaussian beams the wavefronts 
-are uniform and equally spaced resulting in the typical radiation 
-pressure that Ashkin and co demonstrated \cite{Ashkin1980}. However, 
-higher order modes of a Gaussian beam (for example: Laguerre-Gaussian 
-modes) have non-uniform wave fronts meaning the orbital momentum has 
-both angular and linear components; depending on the relative size 
-of the target particle one can induce rotation, or orbiting 
-\cite{Bruce2020, Courtial2000}.
+Electromagnetic fields can transfer both linear and angular momentum \cite{Beth1936MechanicalDA}; more accurately the field is said to have 
+both orbital and spin momentum.
+Though there is some debate on how to decompose the total momentum into these two components \cite{Bruce2020, Svak2018}, for this project we do not need to calculate the exact
+quantities and will instead look at the broader effects of both 
+components. Orbital angular momentum arises from the shape of the 
+wavefront of the particular field in question; for simple Gaussian 
+beams the wavefronts are uniform and equally spaced resulting in the 
+typical radiation pressure that Ashkin and co demonstrated 
+\cite{Ashkin1980}. However, higher order modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have non-uniform wave fronts meaning 
+the orbital momentum has both angular and linear components; depending 
+on the relative size of the target particle one can induce rotation, 
+or orbiting \cite{Bruce2020, Courtial2000}.
 
@@
 \subsection{Characterisation of rotational motion}
-
 Rotational motion about a single axis is easiest to account for.  
@@
 growth. Laser induced nucleation stands to be an ideal method to 
-study nucleation, as the laser output can be concentrated to a 
-small area \cite{Korede2023}. This fine control over the local 
-fluid would allow for individual nucleation events to be
-characterised as a factor of the local fluid properties. One 
-challenge lies in the fact that its clear that the local fluid 
+study the nucleation of organic compounds, as the laser output 
+can be concentrated to a small area \cite{Korede2023} while not 
+altering the organic compound - as is the case with TEM. However,
+laser induced nucleation in itself is also poorly understood, 
+meaning trying to study the nucleation kinetics is difficult if the
+effect of the laser. With this in mind we suggest that we instead 
+utilise another application of optical tweezers, manipulation 
+of micro-scale particles to induce nucleation around a small 
+localised area. The most common method for manipulating a fluid is 
+by localised fluid shearing, which has been shown to directly 
+influence nucleation rates \cite{Debuysschere2023}. 
+One challenge lies in the fact that its clear that the local fluid 
 properties have a direct influence on the likelihood of laser

### 2025-04-15 Latex_files/Thesis/chapters/chapter 5/main.tex
updates 15/04/25

@@
-\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
+\chapter{Detection of rotational spherical aggregate rotational dynamics}
 \label{chapter:simulated_detection}
@@
 
-In chapter 4 we showed that while machine learning can be used 
-to accurately predict the translational displacements of an 
-aggregate particle, it fails when trying to account for the 
-rotational behaviour of said particle. In this chapter we 
-build upon \cite{BarZiv1997} but now account for the full 
-rotational motion. This was initially developed in tandem with 
-the work conducted in chapter 3, with the idea being that if a 
-crystal nucleus formed on a rotating sphere how would that 
-nucleus influence the motion of the rotating sphere. Using 
-static light scattering we can map the outputted signal from a 
-optically trapped dimer to its expected orientation in real 
-time. Thus allowing us to estimate the optical torque being 
-applied to the dimer by the trapping beam. As demonstrated 
-in chapter 4, the influence of a circularly polarised beam 
-can result in gyroscopic precession for specifically sized 
+In this chapter we build upon the work of \cite{Rohrbach2002} and
+\cite{BarZiv1997}; the former used to measure the relative position
+of the trapped aggregate, and the latter to measure the orientation
+of the aggregate. This was initially developed in tandem with the 
+work conducted in chapter 3, with the idea being that if a crystal 
+nucleus formed on a rotating sphere how would that nucleus influence 
+the motion of the rotating sphere. Using static light scattering we 
+can map the outputted signal from a optically trapped dimer to its 
+expected orientation in real time. Thus allowing us to estimate the 
+optical torque being applied to the dimer by the trapping beam. As demonstrated in chapter 4, the influence of a circularly polarised 
+beam can result in gyroscopic precession for specifically sized 
 dimers.
 
-We start by describing a quadrant photo diode system 
-simulated the power spectrum derived from a particle 
-trajectory. We demonstrate that even in idealised 
-circumstances the rotational motion is not revealed 
-from the collected power spectrum. We then go about 
-defining an optical fibre detection system from a 
-mathematical perspective. Before demonstrating how it 
-can be used to instantaneously predict orientational 
-information on a trapped particle. We then consider 
-common factors of error in the characterisation process, 
-such as signal error, and incorrect particle sizing. 
-Both of which have a significant impact on the performance 
-of the model, but by utilising Bayesian inference, and 
-time averaging we can better refine our model.
+We start by describing a quadrant photo diode system simulating the 
+power spectrum derived from a particle trajectory. We demonstrate 
+that even in idealised circumstances the rotational motion is not 
+revealed from the collected power spectrum. We then go about 
+defining an optical fibre detection system from a mathematical 
+perspective. Before demonstrating how it can be used to 
+instantaneously predict orientational information on a trapped 
+particle. We then consider common factors of error in the 
+characterisation process, such as signal error, and incorrect particle 
+sizing. Both of which have a significant impact on the performance 
+of the model, but by utilising Bayesian inference, and time averaging 
+we can better refine our model.
 
@@
 detection system. While it is possible to extract all of 
-the relevant dynamical information from a simulation, 
-confirming the same behaviour in an experimental setting 
-can be challenging if dealing with a non-birefringent 
-anisotropic scatterer. The QPD is composed of 4 photodiodes 
-that measure the intensity of light incident on their 
-surfaces. Using \textit{ott} we can define a region in the 
-far-field that is analogues to the surface of the QPD surface. 
-Evaluating the electric field across this surface gives us 
-an approximation of the QPD signal outputted in an experimental 
-situation. This does not provide a one-to-one result however 
-as hardware errors (i.e. internal resistance, external light 
-sources, and vibrations) would distort the signal somewhat. We 
-can use this as a means to examine the limitations of using 
-back-focal plane interferometry as a means for characterising 
-anisotropic scatterers.
+the relevant dynamics from a simulation, from an experimental
+standpoint characterising those dynamics is dependent on what
+experimental techniques are used. Translational motion can be
+characterised via a position detection system but angular 
+motion is far more difficult to detect, let along characterise.
 
-As a benchmark we start by considering a single sphere within 
-an optical trap. A single polystyrene sphere suspended in water 
-($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
-Gaussian beam ($NA=1.25$) using circularly polarised light. For 
-the sake of time efficiency the trajectory was sampled every 10
-time steps, meaning the upper bound on the power spectra is
-$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
-we fitted the power spectra using the aliased Lorentzian 
+The simulative QPD is composed of 4 photodiodes that measure the 
+intensity of light incident on their surfaces. Given that the 
+condensing lens will have a maximum acceptance angle given by 
+its numerical aperture ($\theta_{max} = \sin^{-1}\left(NA_c/n_
+{med}\right)$) we define the electric field incident on the QPD 
+surface as the elements of the scattered and incident field that 
+propagate within that angular range of $\pm \theta_{max}$. Using 
+\textit{ott} we can evaluate the total electric field to an 
+arbitrary spacial resolution across the surface of the QPD. The 
+reported signal is found by taking the difference between pairs 
+of quadrant signals (see \ref{sec:simulated_QPD} for further 
+detail). This does not provide a one-to-one result however as 
+hardware errors (i.e. internal resistance, external light sources, 
+and vibrations) would distort the signal. It can be used however to 
+show how if the reported power spectra is an accurate representation 
+of the true particle dynamics or if the additional degrees of freedom 
+have an adverse effect on the scattering. See chapter 2 for a breakdown
+of power spectral analysis.
+
+As a benchmark we start by considering a single sphere within an optical 
+trap. A single polystyrene sphere suspended in water ($a=1\mu m$, 
+$n_p=1.59$, $n_m=1.33$) was trapped by a focused Gaussian beam ($NA=1.25$) using circularly polarised light.
+For the sake of time efficiency the trajectory was sampled every 10 time steps, meaning the upper bound on
+the power spectra is $f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the 
+frequency window we fitted the power spectra using the aliased Lorentzian.
 (Eq.~\eqref{eq:aliased_lorentzian}). 
@@
 to what is actually being measured. The \textit{ott} estimates 
-are simply looking at the differences in the trapping strength 
-along the Cartesian axis'. If calibrated over an long enough 
-time frame you would expect that the resulting power spectra 
-would exactly mirror the \textit{ott} predictions. However over 
-a short calibration time a QPD will only sample a small trajectory 
-that does not fully explore the trapping plane. There is a clear 
-trade off in terms of accuracy and computation time as shorter 
-calibration runs are computationally more efficient but prone to 
-errors.
+are simply looking at the change in the trapping force as you
+move along a given axis'. In reality, a particle will be freely
+diffusing within the trap focus meaning over a short calibration 
+time the QPD is only estimating a weighted average of the trapping
+strength. If calibrated over an long enough time frame you would 
+expect that the resulting power spectra would exactly mirror the 
+\textit{ott} predictions. There is a clear trade off in terms of 
+accuracy and computation time as shorter calibration runs are 
+computationally more efficient but prone to errors. However, it
+should be noted that the estimation made by the QPD is still less
+than the results from \cite{BergSoerensen2004}, though this could
+be due to a lack of external noise signals.
 
@@
 \textit{ott} estimate and the QPD estimate.
-
 \begin{center}
@@
 shown for nano-particles that exhibit periodic rotational motion
-using by considering the difference in diagonal quadrants. The 
-benefit of detecting rotational motion is that one can begin to 
-build a better understanding about how angular momentum is 
-transferred to a dimer which could be extended to more complex
-particles.
+using by considering the difference in diagonal quadrants 
+\cite{Yifat2021}. The only difference between these results is 
+the fact that rotation occurs perpendicular to the particle's long 
+axis, whereas dimers rotate about their long axis. The benefit of 
+detecting rotational motion is that one can begin to build a better understanding about how angular momentum is transferred to a dimer 
+which could be extended to more complex particles. This could give 
+some understanding to the results of Chapter 4.
 
@@
 when rotated by in the X-Y plane. 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{rotation_figure}
+	\caption{QPD signals detected by a dimer being rotated while
+			centred at the focus of the optical trap. Top right:
+			rotation in the X-Z plane. Top left: rotation in the
+			Y-Z plane. Bottom: Rotation in the X-Y plane.}
+	\label{fig:QPD_rotation}
+\end{figure}
 
@@
 signal to the dimer's position shows promising results, shown in
-fig.~X.X is a prediction of a symmetric dimer's position based
-on the QPD signal. With the actual displacement on the x-axis 
-and the predicted result y-axis, and the dashed line represents 
-the y=x line - indicating an ideal prediction.
+fig~\ref{fig:QPD_rotation} is a prediction of a symmetric dimer's 
+position based on the QPD signal. With the actual displacement 
+on the x-axis and the predicted result y-axis, and the dashed 
+line represents the y=x line - indicating an ideal prediction.
 
-Fig~X.X shows that its relatively trivial to predict a particle's
-position based off the QPD signal, even in axial direction which
-has not been done previously. It should be noted that these 
-displacements are relative to the trap focus and so accurate 
-tracking of the beam movements needs to be taken. This result is
-partially due to the fact that displacement and signal units are
-linearly related by \eqref{eq:conversion_factor}. Interestingly 
-when we run the same protocol for different sized dimer's there 
-seems to be a cut off point when the fit begins to fail. Plotting
-the $R^2$ for each fit vs its size ratio shows that beyond a size
-ratio of 5 the machine learning begins to fall off significantly.
+Fig~\ref{fig:QPD_rotation} shows that its relatively trivial 
+to predict a particle's position based off the QPD signal, even 
+in axial direction which has not been done previously. It should 
+be noted that these displacements are relative to the trap focus 
+and so accurate tracking of the beam movements needs to be taken. 
+This result is partially due to the fact that displacement and 
+signal units are linearly related by \eqref{eq:conversion_factor}. Interestingly when we run the same protocol for different sized 
+dimer's there seems to be a cut off point when the fit begins to 
+fail. Plotting the $R^2$ for each fit vs its size ratio shows 
+that beyond a size ratio of 5 the machine learning begins to fall 
+off significantly.
 
@@
 sphere shrinks the dynamics should more closely approximate that 
-of a single sphere. We can compare our
+of a single sphere. This should be reflected in the total field
+incident on the QPD and therefore the machine learning program 
+should be able to easily detect the dimer's translational motion.
+We can therefore only conclude that the dimer's rotational motion
+is contributing to the scattering to the scattered field, even if
+it is miniscule the contribution is enough to throw off a 
+position detection system. Detection of rotational motion was 
+attempted using machine learning, the model was trained on the 
+same trajectory mentioned in \ref{sec:off-axis}, where the dimer 
+is trapped in an off-axis orientation, from a top down perspective
+it should be easy enough to note that the dimer is not vertically 
+trapped even if the exact angle is difficult to determine.
+
+Overall, the reliance on machine learning to perform vector 
+regression on the collected scattered field is not a viable 
+method for detecting let along characterising the angular 
+displacement of an isotropic scatterer. While there is a 
+clear correlation in the scattered signal and angular 
+displacement the translational motion makes up the majority 
+of the expected signal detected by the QPD. As of now, there
+is no optical arrangement that would allow rotational motion 
+while restricting translational motion.
 
@@
 that altering the target origin would drastically alter the 
-tmatrix produced
+tmatrix produced. We choose (somewhat arbitrary) to use 
+the dimer's centre of diffusion which is located $\zeta a_I$ from 
+the centre of the larger sphere. For a discussion see \emph{X.X}.
 
@@
 be arbitrary, in that it made no difference to our estimate whether 
-we used 2 angles or 200. For practical purposes it seemed beneficial 
-that we demonstrate our method works for a minimal number of detection 
-angles, as geometric constraints come into play when trying to install 
-a high number of detection fibres for any optical tweezer set up.
+we used 2 angles or 200.
 

### 2025-04-15 python_files/autocorr_fit.py

@@ -1,162 +1,0 @@
-import argparse
-import csv
-import numpy as np
-from scipy.optimize import curve_fit
-import scipy
-from scipy.fft import fft
-from scipy.signal import correlate
-import matplotlib.pyplot as plt
-
-class autocorr_fit(object):
-    def __init__(self, path):
-        self.path = path
-        #self.material, self.power, self.size, self.radius = self.get_label_path(self.label)
-
-    def row_reader(self, row):
-        lst=(row[0].split())
-        #print(lst)
-        
-        new_row = [float(lst[0]), float(lst[1]), float(lst[2]), float(lst[3])]
-        #print(new_row)
-        return new_row
-
-    def get_label_path(self, label):
-        label = label.split('_')
-        if label[0] == 'Silica':
-            self.material = 'Silica'
-        elif label[0] == 'Vaterite':
-            self.material = 'Vaterite'
-        elif label[0] == 'PS':
-            self.material = 'PS'
-        else:
-            self.material = 'LC'
-        
-        self.power = label[1]
-        self.size = label[2][:-4]
-        self.radius  = float(self.size[:-2])*1e-6
-        return self.material, self.power, self.size, self.radius
-    
-    def get_line_count(self, file):
-        reader = csv.reader(file.readlines())
-        lines = len(list(reader))
-        return lines
-
-    def create_csv_file(self):
-        with open(self.path, 'r') as in_file:
-            reader = csv.reader(in_file.readlines()[5:])
-            
-            with open('tdat.csv', 'w') as out_file:
-                writer = csv.writer(out_file)
-                writer.writerows(self.row_reader(row) for row in reader)
-
-        with open('tdat.csv', 'r') as data_file:
-            reader = csv.reader(data_file.readlines()[::2])
-            #t, x_data, y_data = 0, 0, 0
-            t, x_data, y_data = [], [], []
-            
-            for row in reader:
-                t.append(row[0])
-                x_data.append(row[1])
-                y_data.append(row[2])
-
-        return t, x_data, y_data
-    
-    def get_FDat(self, x_data, y_data):
-        window = 100
-
-        N = len(x_data)
[…]

### 2025-04-15 python_files/create_psd.py

@@ -1,185 +1,0 @@
-#!/usr/bin/python3
-
-from matplotlib.mlab import psd
-from pathlib import PurePosixPath, PureWindowsPath
-import numpy as np
-import matplotlib.pyplot as plt
-from scipy.fft import fft
-from scipy.signal import periodogram
-from scipy.optimize import curve_fit, least_squares
-
-def create_psd(data_file):
-    
-    x_dat = []
-    y_dat = []
-    z_dat = []
-    t_list = []
-
-    if data_file[-3:] == 'txt':
-        print('Loading QPD data')
-        test = data_file.split('/')
-        tag = test[-1][:-3]
-        with open(data_file, 'r') as input:
-            lines = input.readlines()
-            for line in lines[1:-1]:
-                items =  line.split(',')
-                x_dat.append(float(items[0]))
-                y_dat.append(float(items[1]))
-                z_dat.append(float(items[2]))
-                t_list.append(float(items[3]))
-        
-        dt = t_list[10]-t_list[9]
-        fs = 1/dt
-
-    elif data_file[-3:] == 'npy':
-        print('Loading trajectory file')
-        test = data_file.split('\\')
-        print(test)
-        tag = test[-1][:-8]
-        print(tag)
-        name = str(PurePosixPath(PureWindowsPath(data_file)))
-        traj = np.load('./'+name[:])
-        dt = 1e-5
-        for i, frame in enumerate(traj[1000:]):
-            x_dat.append(frame[0])
-            y_dat.append(frame[1])
-            t_list.append(dt*i)
-        
-        fs = 1/dt
-
-    def loretzian (x, A, B):
-            return 1/(A+B*x**2)
-    
-    def alias_loretzian(k, dx, c, i):
-            N = len(freq_block[i:])
-            return (dt*dx**2)/(1+c**2-2*c*np.cos(np.pi*(k+1)/N))
-    
-    def res_lorentz(x, k, y, i):
-        y_ex = alias_loretzian(k, x[0], x[1], i)
-        sig = y_ex**-1/(nb)**0.5
[…]

### 2025-04-15 python_files/fitnest.py

@@ -1,45 +1,0 @@
-import numpy as np
-import ultranest
-import ultranest.stepsampler
-from ultranest.plot import cornerplot
-
-class fit(object):
-    def __init__(self, params, x_data, y_data):
-        self.param_names = params
-        self.x = x_data
-        self.y = y_data
-            
-    def prior_transfrom(self, cube):
-        self.params = cube.copy()
-        for i in range(len(cube)):
-            self.params[i][0] = 10**(cube[i][0]*4+3)
-            self.params[i][1] = 10**(cube[i][1]*3+2)
-
-        return self.params
-
-    def log_likelihood(self, params):
-        likelihood = np.ndarray(shape=(len(params)))
-
-        for i in range(len(params)):
-            A = self.params[i][0]
-            B = self.params[i][1]
-            y_model = [1/(A+B*x**2) for x in self.x]
-            likelihood[i] = (((sum([((y-y_m)/1)**2 for y, y_m in zip(self.y, y_model)]))))
-
-        return likelihood
-
-    def run(self):
-        sampler = ultranest.ReactiveNestedSampler(self.param_names,
-        self.log_likelihood,
-        self.prior_transfrom,
-        resume = 'overwrite',
-        log_dir = './results',
-        vectorized = True)
-
-        print('Start')
-        sampler.stepsampler=ultranest.stepsampler.SliceSampler(nsteps=20, 
-        generate_direction=ultranest.stepsampler.generate_mixture_random_direction)
-        results=sampler.run(Lepsilon = 0.1)
-        sampler.print_results()
-        sampler.plot_run()
-        sampler.plot_corner()

### 2025-04-15 python_files/fluid_flow.py

@@ -1,36 +1,0 @@
-import numpy as np
-import scipy as sci
-import matplotlib.pyplot as plt
-from matplotlib.animation import FuncAnimation
-
-#^ define particle path and parameters
-radius_p = 1.57e-6#float(input("Particle radius [microns]: "))*1e-6
-path_radius = 4.71e-6#float(input("Orbit radius [microns]: "))*1e-6
-omega_0 = 1#float(input("Orbital frequency [Hz]: "))
-period = 1/omega_0
-
-#^ compute velocity 'U' along path
-fig, ax  = plt.subplots()
-scat, = ax.plot([],[], 'ro')
-vec = ax.arrow([],[], [], [])
-
-def init():
-    ax.set_xlim(-path_radius-radius_p, path_radius+radius_p)
-    ax.set_ylim(-path_radius-radius_p, path_radius+radius_p)
-    return scat,
-
-def update(frame):
-    x_data=path_radius*np.cos(2*np.pi*frame)
-    y_data=path_radius*np.sin(2*np.pi*frame)
-    vx_data =-path_radius*2*np.pi/period*np.sin(2*np.pi*frame)
-    vy_data =path_radius*2*np.pi/period*np.cos(2*np.pi*frame) 
-    scat.set_data(x_data, y_data)
-    vec.set_data(x = x_data, y = y_data, dx = vx_data*1e-2, dy= vy_data*1e-2,
-                 width = 1e-8, head_width=0.1e-6, head_length=0.1e-6)
-
-    return scat, vec
-
-ani = FuncAnimation(fig, update, frames=np.linspace(0, 1, 100),
-                    init_func=init, blit=True, interval = 10)
-
-plt.show()

### 2025-04-15 python_files/maclaurin.py

@@ -1,21 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-x_range = np.linspace(-5, 5, 50)
-
-
-for x in x_range:
-    macsum = 1
-    kterm = 1
-    k = 1
-    while kterm > 10e-11:
-        kterm *= (x/k)
-        macsum += kterm
-        k += 1
-    plt.scatter(x, macsum, c='black', s =3)
-
-
-plt.plot(x_range, [np.exp(x) for x in x_range])
-plt.vlines(0, -10, np.exp(5), color='black', linestyle ='--')
-plt.show()
-plt.close()

### 2025-04-15 python_files/psd_test.py

@@ -1,26 +1,0 @@
-import numpy as np
-from scipy.signal import periodogram
-import random
-import matplotlib.pyplot as plt
-
-dt = 1e-5
-w_i = [0]
-N = 100000
-for i in range(N):
-    w_i.append(random.normalvariate(0,1)+w_i[i])
-
-step_sizes = np.logspace(0, 5, 10000)
-msd_w = []
-tau = [step*dt for step in step_sizes]
-for step in step_sizes:
-	n = int(N/step)
-	di = int(step)
-	tmp_w = []
-	for i in range(0, n-1):
-		tmp_w.append((w_i[i+di]-w_i[i])**2)
-	msd_w.append(np.mean(tmp_w))
-
-plt.plot(tau, msd_w)
-plt.xscale('log')
-plt.yscale('log')
-plt.show()

### 2025-04-15 python_files/test.py

@@ -1,41 +1,91 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import scipy.misc as misc
+import pandas as pd
+from scipy.optimize import curve_fit
+from scipy.signal import periodogram
 
-p = 1;                  # Degree of LG mode
-l = 2;                  # Order of LG mode
-w0 = 2.0;               # Beam waist
-k = 2*np.pi/532.0e-9;   # Wavenumber of light
+with open('C:\\Users\\xbb16146\\dmaciver97\\python_files\\dimer_files\\data_07-May-2024 1to2QPD.txt', 'r') as infile:
+    lines = infile.readlines()
+    data = {'Time': [], 'X-dat': [], 'Y-dat': [], 'Z-dat': []}
+    for line in lines[1:-1]:
+        row = line.split(',')
+        data['Time'].append(float(row[3][:-1]))
+        data['X-dat'].append(float(row[0]))
+        data['Y-dat'].append(float(row[1]))
+        data['Z-dat'].append(float(row[2]))
+    df = pd.DataFrame(data)
+    pd.to_pickle(df, 'test.pkl')
 
-zR = k*w0**2.0/2;       # Calculate the Rayleigh range
-
-# Setup the cartesian grid for the plot at plane z
-z = 0.0;
-xx, yy = np.meshgrid(np.linspace(-5, 5), np.linspace(-5, 5));
-
-# Calculate the cylindrical coordinates
-r = np.sqrt(xx**2 + yy**2);
-phi = np.arctan2(yy, xx);
-
-U00 = 1.0/(1 + 1j*z/zR) * np.exp(-r**2.0/w0**2/(1 + 1j*z/zR));
-w = w0 * np.sqrt(1.0 + z**2/zR**2);
-R = np.sqrt(2.0)*r/w;
-
-# Lpl from OT toolbox (Nieminen et al., 2004)
-Lpl = misc.comb(p+l,p) * np.ones(np.shape(R));   # x = R(r, z).^2
-for m in range(1, p+1):
-    Lpl = Lpl + (-1.0)**m/misc.factorial(m) *misc.comb(p+l,p-m) * R**(2.0*m);
-
-U = U00*R**l*Lpl*np.exp(1j*l*phi)*np.exp(-1j*(2*p + l + 1)*np.arctan(z/zR));
-
-plt.figure()
-plt.title('Intensity')
-plt.pcolor(abs(U)**2);
[…]

### 2025-04-15 python_files/triangle.py

@@ -1,80 +1,0 @@
-import numpy as np
-import plotly.graph_objects as go
-import turtle
-import warnings
-
-while True:
-    try:
-        sides = np.zeros(3)
-        sides[0] = float(input('Side no 1: '))
-        sides[1] = float(input('Side no 2: '))
-        sides[2] = float(input('Side no 3: '))
-        rotation = float(input('Starting angle from horizontal: '))
-        sides = sorted(sides)
-        break
-    except ValueError:
-        print('Sides cannot be strings')
-
-a,b,c = sides[0], sides[1], sides[2]
-# If triangle sides are wrong then it will throw up a run time error
-# when calculating the angles
-
-with warnings.catch_warnings(record=True) as w:
-    
-    alpha = np.arccos((b**2+c**2-a**2)/(2*b*c))*(180/np.pi)
-    beta = np.arccos((a**2+c**2-b**2)/(2*a*c))*(180/np.pi)
-    gamma = np.arccos((a**2+b**2-c**2)/(2*a*b))*(180/np.pi)
-   
-    if len(w) > 0:
-        print('These side lengths cannot form a triangle')
-        exit(0)
-    else:
-        area = 0.5*a*b*np.sin(gamma*np.pi/180)
-        height = 2*area/a
-        
-
-# Rearrange angles for turtle drawing
-# each angle is opposite its corresponding side so needs to be shifted
-angles = [gamma, alpha, beta]
-
-# Define the Triangle Type
-if a == b and b == c:
-    tritype = 'equilateral triangle'
-elif (a == b and a != c) or (b == c and b != a) or (c == a and c != b):
-    tritype = 'isoceles triangle'
-elif a != b and a != c:
-    tritype = 'scalene triangle'
-
-# Last check involves angles so should be seperate so it isn't skipped
-if np.isclose(90, gamma, atol = 0.001):
-    tritype = 'Right Angled triangle'
-
-t = turtle.Turtle()
-screen = turtle.Screen()
-rootwindow = screen.getcanvas().winfo_toplevel()
-rootwindow.call('wm', 'attributes', '.', '-topmost', '1')
-rootwindow.call('wm', 'attributes', '.', '-topmost', '0')
-screen.setup(800,800)
-# Adjust starting position
-t.penup()
-t.goto(-10*a, -10*height)
-t.pendown()
-t.left(rotation)
-for i in range(3):
-    #distance = sides[i]
[…]

### 2025-04-29 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 25/04/25

@@
 improve product quality while keeping production costs 
-low. The industrialisation of crystallisation has 
-allowed engineers to reliably and efficiently induce 
-the crystal formation within a bulk phase. However, 
-while on a large scale crystallisation is seemingly an 
-understood physical process, just a small amount of 
-investigation into the literature reveals that at a 
-micro scale there is yet to be unifying theory that can 
-accurately explain the process of crystallisation 
-\cite{Fu2021}. Crystallisation can be subdivided into 
-two steps: Nucleation and Crystal growth. The latter 
-focusing on how an already stable crystal grows and how 
-it takes on its final shape. Whereas the former is more 
-concerned with the first initial moments of a crystals 
-formation within a bulk phase.
+low. Modern crystallisation methods involve manipulating 
+the solubility of the target chemical (aka the solute) 
+in order to encourage the formation of a separate phase
+that can be easily extracted \cite{Gao2017}. This is 
+simple enough to understand on a macro scale when the 
+concentration can be assumed as a time-dependent 
+variable; but at a micro level the crystallisation process
+is not merely subject to solubility but also due to 
+the inter-molecular forces that can inhibit or enhance 
+the formation of a new phase. Theories attempting to describe
+the phenomena but none have proven viable to explain the 
+various pathways available \cite{Fu2021}.
 
-In this chapter we will outline our current understanding 
-of nucleation, why there is a gap in the literature, and why 
-there is still a need for local control of nucleation events. 
-Furthermore, we will also highlight recent developments 
-involving the use of optical tweezers (and other laser based 
-methods) in order to develop our understanding of nucleation. 
-Lastly we highlight the connection between optical tweezer's 
-ability to probe fluid viscosity via rotational motion, and 
-the known link between nucleation and fluid flow, to propose 
-a novel potential method of creating localised nucleation events.
+Crystallisation can be subdivided into two stages: The first 
+stage - nucleation - is the formation of a brand new phase 
+within the bulk solution, while nucleation can occur readily 
+in most solutions it is difficult to predict if a nucleus 
+will remain stable within the solution or if it will be 
+dissolve. The latter stage - crystal growth - is focused 
+instead on how an already stable nucleus grows into its 
+final shape depending on the conditions of the surrounding
+fluid, understanding how and where deformities occur within
+the crystal structure are of extreme importance to several 
+industries \cite{Gao2017,Myerson2002}.
+
+This chapter will outline our current understanding of nucleation, 
+where there is a gap in the literature, and why there is still a 
+need for local control of nucleation events. Furthermore, it will 
+also highlight recent developments involving the use of optical 
+tweezers (and other \textit{in situ } methods) in order to develop 
+our understanding of nucleation. Lastly we highlight the connection 
+between optical tweezer's ability to probe fluid viscosity via 
+rotational motion, and the known link between nucleation and fluid 
+flow, to propose a novel potential method of creating localised 
+nucleation events.
 
@@
 difference driving the solution to separate the two phases.
+This does not mean that nucleation will occur immediately 
+once $C_{eq}$ is exceeded, only that there is a chemical
+potential supporting nucleation.
+
 Since different combinations of solute and solvent will 
@@
 \subsection{Primary \& Secondary nucleation}
-From an industrial perspective, the nucleation process can 
-be broadly categorised into either primary of secondary
-nucleation. The former describes the formation of an initial
-nucleus within the bulk phase, absent of any external stimuli.
-Primary nucleation is therefore considered stochastic as there 
-is no means of predicting where a nucleus may form, or how long 
-it will take. The only reliable indicator being that higher 
-supersaturations will result in faster nucleation rates. At a 
-small scale one can estimate the nucleation rate by making 
-repeated measurements of sample solutions and seeing how many 
-have nucleated after a given time, giving us a Poisson probability distribution.
+nucleation process be broadly categorised into either primary 
+or secondary nucleation. The former describes the formation 
+of an initial nucleus within the bulk phase, absent of any 
+external surfaces. Primary nucleation is dependent only on 
+the local concentrations of the solute which can fluctuate 
+unpredictably, as such it is often modelled at a large scale 
+as a stochastic process. The only reliable factor being 
+that higher supersaturations will result in more nucleation 
+events to occur sooner.
+
+An important metric for crystallisation procedures is the 
+nucleation rate of the solution, a measure of the total 
+volume or mass of crystalline material produced for a given 
+set of initial conditions. At an industrial scale this 
+metric is a contribution of both primary and secondary 
+nucleation events. At a small scale one can estimate the 
+contribution of the primary nucleation pathway by making 
+repeated measurements of sample solutions and seeing how 
+many have nucleated after a given time, giving us a Poisson 
+probability distribution.
 \begin{align}
@@
 
-Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
-$V$ is the volume of the individual samples, and $M^*(t)$ \& 
-$M$ are the number of nucleated samples and the total number 
-of samples used respectively. The nucleation rate is one of 
-the most important factors used in evaluating a crystallisers 
-performance, providing a rough estimation of the final product 
-mass over a given time frame. While \eqref{eq:poisson} is useful 
-for studying the effects of different parameters at a small scale, 
-for industrial applications there are to many external factors 
-for primary nucleation to be measured accurately.
+Where $J$ is the nucleation rate, $t_g$ is the 'growth time' 
+which accounts for the initial delay due to the stochastic 
+nature of primary nucleation, $V$ is the volume of the individual 
+samples, and $M^*(t)$ \& $M$ are the number of nucleated samples 
+and the total number of samples used respectively. While 
+\eqref{eq:poisson} is useful for studying the effects of different 
+parameters at a small scale, for industrial applications there are 
+to many external factors for primary nucleation to be measured 
+accurately. Overall while the primary nucleation rate is significant 
+when studying solutions for small volumes, the contributions of 
+secondary nucleation is far more significant when considering 
+nucleation rate for an industrial crystalliser.
 
@@
 inducing further nucleation within the bulk solution 
-\cite{Botsaris1976}. The seed crystals is often specially 
-prepared ahead of time and added to a supersaturate solution. 
-Due to interactions with the surrounding fluid and the container 
-walls it is possible for the seed crystals to act as a surface
-for further nucleation \cite{Anwar2015}. Control of secondary 
-nucleation events are crucial for ensuring reliable industrial 
-crystalliser performance. Fig.~\ref{fig:secondary} depicts an 
-attempt of classifying every possible mechanism that could lead 
-to secondary nucleation.
+\cite{Botsaris1976}. The seed crystals are often prepared ahead 
+of time and added to a supersaturate solution. Due to interactions 
+with the surrounding fluid and the container walls it is possible 
+for the seed crystals to act as a surface for further nucleation \cite{Anwar2015}. However this is just one possible way in which 
+a seed crystal can enhance the nucleation rate, depending on the 
+conditions of the crystalliser there are multiple mechanism by 
+which the seed crystal can contribute to the nucleation rate of 
+the system. Control of secondary nucleation events are crucial 
+for ensuring reliable industrial crystalliser performance. Fig.~\ref{fig:secondary} depicts an attempt of classifying every 
+possible mechanism that could lead to secondary nucleation.
 \begin{figure}[h!]
@@
 		     lines are for a higher supersaturation than the solid lines,
-		     the interfacial energy cost is independent of supersaturation.}
+		     the interfacial energy cost is independent of supersaturation.
+		     Diagram was constructed in python using a naive assumption 
+		     of $v=1\times 10^-12\ m^3$ and $\sigma_{inf}=0.0001\ J/m^2$}
 	\label{fig:free_energy}
@@
 its obvious that for all crystallization systems there is an 
-inherent energy barrier that dictates the nucleation rate. 
+inherent energy barrier that must be overcome prior to nucleation. 
 Where it falters is in its predictive ability, both in estimating 
@@
 many possible pathways that can be taken to produce an ordered 
-crystalline phase. Prompting the development of alternative 
-theories to better describe the nucleation process.
+crystalline phase. In situations where the solution is heavily 
+saturated with solute the driving force promoting nucleation is 
+so large that a nucleus can form in the bulk solution in a single 
+step.
+
+This is however not accurate at lower supersaturations, closer
+to the saturation point the solvent's surface tension is so 
+significant that the addition of individual solute molecules
+should not be enough to overcome the dissolution of the nucleus. 
+And yet nucleation is still possible, occurring far more reliably
+than predicted by the CNT. This suggests that the formation of 
+a nucleus is due to larger aggregates of solute that are stable 
+within the bulk solution. This has been confirmed experimentally 
+and lead to the development of Two Step Nucleation.
 

### 2025-04-29 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to chp 4 & 2
updates 25/04/25

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Position detection methods}
+\subsection{Position detection}
+\label{sec:position_detection}
 In order to accurately capture the dynamics of a trapped particle, a 
@@
 resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
-
-A quadrant photo diode (QPD) is a frequently used position detection 
-system for optical tweezers due to their high sampling rate, high 
-degree of precision, and ease of set up. The QPD is constructed of 
-four photo diodes assembled in a quadrant formation, when a particle 
-is trapped the interference pattern produced is focused onto the QPD, 
-with the maximum intensity mapping to the particle's centre of mass. 
-By summing the voltages of the horizontal and vertical quadrants together 
-the particle's centre of mass is tracked in the x-y plane. Axial 
-displacement can be estimated by observing the change in the total 
-voltage of the QPD. The outputted signal gives an indication of the 
-particle's relative displacement from the beam focus, but in order to 
-convert the signal to distance units the trap needs to be calibrated 
-(assuming a linear response curve).
-\begin{figure}[h!]
-	\centering
-	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
-	\caption{Comparison between QPD and Lateral effect photodiodes.
-	The four quadrants of a QPD (left) experience different photocurrents
-	based on the total intensity of light incident on each section 
-	(labelled A, B, C, D). 
-	Whereas a Lateral effect sensor (right) uses the resistive properties
-	of the photodiode surface to vary the create different photocurrents 
-	passing through the anodes A, B, C, and D.}
-\end{figure}
 
@@
 can be considered linear.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h!]
+	\centering
+	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
+	\caption{Comparison between QPD and Lateral effect photodiodes.
+	The four quadrants of a QPD (left) experience different photocurrents
+	based on the total intensity of light incident on each section 
+	(labelled A, B, C, D). 
+	Whereas a Lateral effect sensor (right) uses the resistive properties
+	of the photodiode surface to vary the create different photocurrents 
+	passing through the anodes A, B, C, and D.}
+\end{figure}
+
+A quadrant photo diode (QPD) is a frequently used position detection 
+system for optical tweezers due to their high sampling rate, high 
+degree of precision, and ease of set up. The QPD is constructed of 
+four photo diodes assembled in a quadrant formation, when a particle 
+is trapped the interference pattern produced is focused onto the QPD, 
+with the maximum intensity mapping to the particle's centre of mass. 
+By summing the voltages of the horizontal and vertical quadrants together 
+the particle's centre of mass is tracked in the x-y plane. Axial 
+displacement can be estimated by observing the change in the total 
+voltage of the QPD. The outputted signal gives an indication of the 
+particle's relative displacement from the beam focus, but in order to 
+convert the signal to distance units the trap needs to be calibrated 
+(assuming a linear response curve).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Fourier Optics and 4f correlators}
@@
 
-For our applications a 4f correlator is utilised to ensure that the motion
-of the galvano-mirrors does not move the focal point of the laser, allowing 
-for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
-the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
-installed $150\ mm$ from the second mirror of the galvano, and the latter being 
-installed $200\ mm$ from the back focal plane of the trapping objective. The 
-signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
-dual channel signal control. This allowed us to precisely control the alignment,
-amplitude, phase, and frequency of both mirrors making alignment much easier.
-
-While in theory the frequency can be increased until the mechanical limit of the 
-mirrors is reached ($\approx 1000\ Hz$), the practical upper limit is determined 
-by the trap strength. For a silica sphere this is on the order of $300\ Hz$ when
-suspended in water. Likewise while in theory the signal amplitude can be increased
-until the voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
-of the focal lenses limits the maximum amplitude to $0.5 V$ as any greater will
-move the laser beyond the lens. For basic trapping calibration the galvano-mirrors 
-were set to DC output, providing a fixed spot which operates like a typical optical 
-trap.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For our applications a 4f correlator is utilised to ensure that the 
+motion of the galvano-mirrors does not move the focal point of the 
+laser, allowing for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after the galvano-mirror we have our two lenses 
+- $f_3$ and $f_4$ - the former being installed $150\ mm$ from the 
+second mirror of the galvano, and the latter being installed $200\ mm$ 
+from the back focal plane of the trapping objective. The signal 
+generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
+dual channel signal control. This allowed us to precisely control the
+alignment, amplitude, phase, and frequency of both mirrors making 
+alignment much easier.
+
+While in theory the frequency can be increased until the mechanical 
+limit of the mirrors is reached ($\approx 1000\ Hz$), the practical 
+upper limit is determined by the trap strength. For a silica sphere 
+this is on the order of $300\ Hz$ when suspended in water. Likewise 
+while in theory the signal amplitude can be increased until the 
+voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
+of the focal lenses limits the maximum amplitude to $0.5 V$ as any 
+greater will move the laser beyond the lens. For basic trapping 
+calibration the galvano-mirrors were set to DC output, providing a 
+fixed spot which operates like a typical optical trap.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Synthesis of Birefringent Micro spheres}
@@
 stability. Though for the purposes of optical rotation the 
-exact polymorph is not as important as its morphology as all 
+exact polymorph is not as important as the morphology as all 
 3 polymorphs are inherently birefringent.
@@
 the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
-aside for later. Firsta small vial was filled with $1.5mL$ of 
+aside for later. First a small vial was filled with $1.5mL$ of 
 $CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Rotation of birefringent micro spheres} 
@@
 local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
+
 When it comes to optical rotation, a rotating particle will experience a

### 2025-06-27 Latex_files/Thesis/chapters/chapter 1/main.tex
27/06/2025 chapters 1,3,4 and structural changes

--- 2025-04-15 Latex_files/Thesis/chapters/chapter 1/main.tex
@@
 \chapter{Introduction}
-Crystallisation is one of the oldest forms of phase 
-separation used by humanity \cite{Schoen1956}, put 
-simply, it is the formation and growth of a new 
-structured phase within a disordered bulk phase. 
-This has applications in a number of industries such 
-as pharmaceuticals \cite{Gao2017}, food production 
-\cite{Hartel2002}, and electronics \cite{Myerson2002}. 
-Where the extraction of dilute materials can help 
-improve product quality while keeping production costs 
-low. The industrialisation of crystallisation has 
-allowed engineers to reliably and efficiently induce 
-the crystal formation within a bulk phase. However, 
-while on a large scale crystallisation is seemingly an 
-understood physical process, just a small amount of 
-investigation into the literature reveals that at a 
-micro scale there is yet to be unifying theory that can 
-accurately explain the process of crystallisation 
-\cite{Fu2021}. Crystallisation can be subdivided into 
-two steps: Nucleation and Crystal growth. The latter 
-focusing on how an already stable crystal grows and how 
-it takes on its final shape. Whereas the former is more 
-concerned with the first initial moments of a crystals 
-formation within a bulk phase.
+\section{Optical Tweezers}
+Precise manipulation of micro-particles and measurements of 
+microscopic forces has been development in scientific research.
+It has long been known that the motion of microscopic particles 
+is due to multiple random collisions, measuring the average 
+momentum transferred is often done by generating a potential 
+well constructed using light - more commonly referred to as 
+optical tweezing.
 
-In this chapter we will outline our current understanding 
-of nucleation, why there is a gap in the literature, and why 
-there is still a need for local control of nucleation events. 
-Furthermore, we will also highlight recent developments 
-involving the use of optical tweezers (and other laser based 
-methods) in order to develop our understanding of nucleation. 
-Lastly we highlight the connection between optical tweezer's 
-ability to probe fluid viscosity via rotational motion, and 
-the known link between nucleation and fluid flow, to propose 
-a novel potential method of creating localised nucleation events.
+Since the mid 1900's it was known that light can transfer momentum 
+to a medium, being referred to as 'radiation pressure', initially 
+this was applied by transferring both linear and angular momentum 
+to suspended mirrors \cite{Beth1936MechanicalDA}. This allowed 
+researchers to measure the momentum of light carried by a single 
+photon \cite{Beth1936MechanicalDA}. It was Ashkin who showed that 
+by constructing a precise wavefront one could confine biological 
+cells as small as a few microns \cite{Ashkin1970}. This was 
+especially useful as the process was non-invasive, preserving the 
+internal organelles while keeping the cell trapped.
 
-\section{Nucleation}
-Nucleation is an example of a binary phase separation, 
-where a dilute phase is miscible in a bulk phase, more 
-often called the solute and solvent respectively. When 
-a small amount of solute is added to the solvent the 
-solute will diffuse throughout the solvent, reaching
-an equilibrium concentration throughout the solution. 
-Because of molecular interactions, the two can only remain 
-in equilibrium while below a specific concentration ($C_{eq}$) 
-- below which the chemical potential $\mu$ for a single phase 
-is greater than the potential required to separate the two. 
-Once $C_{eq}$ is exceeded there is a chemical potential 
-difference driving the solution to separate the two phases. 
-Since different combinations of solute and solvent will 
-have different equilibrium concentrations, researchers often 
-instead measure the ratio between the solute and solvent by 
-using 'supersaturation' \cite{Mullin2001}:
-\begin{align}
-	\label{eq:supersaturation}
-	S(T) = \frac{C_{sol}}{C_{eq}(T)}
-\end{align}
+The working principle is that a collimated light source - such as 
+a laser - can confine a particle in the plane perpendicular to the 
+beam \cite{Ashkin1970}. The radiation pressure previously observed 
+is due to the momentum transferred to the target, when directed at 
+an isotropic scatterer the light the momentum transfer is minimised
+when the target particle is centred on the laser beam. Initially, 
+confinement in the direction of the laser was achieved by matching 
+the laser power so that the radiation pressure transferred would equal 
+the mass of the target \cite{Ashkin1970}. Soon after, Ashkin showed 
+that the by focusing the light source through a lens would confine 
+the target particle in all 3 Cartesian directions regardless of the
+laser power \cite{Ashkin1980}. The use of lenses to create a focal
+point allowed for the development of much stronger optical traps that
+could trap complex particles and biological materials.
 
-Where $C_{sol}$ is just the solute concentration, and 
-$C_{eq}(T)$ is the equilibrium concentration of the solute  
-at temperature $T$ based on the main solvent. While the 
-solution remains supersaturated there is a chemical 
-potential driving the solute to coalesce and separate 
-from the solution as an ordered solid, the first 
-formation of the crystal is referred to as the 
-nucleus and understanding its formation has been the 
-focus of researchers for decades now. Typically, for an 
-industrial crystallisation process the working principle 
-is based on controlling and manipulating the supersaturation 
-of the system.
+Beyond simply observations, optical tweezing allowed for precise 
+measurements of microscopic forces. Later it would be used to probe microscopic properties such as the Brownian forces exerted within a 
+pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
+behaviour of light, optical tweezers have become essential for 
+measuring and exerting precise forces on the magnitude of pico-newtons.
+However, these measurements are based on the idea that the trapped 
+particle in question is an isotropic scatterer, meaning that particle 
+is can be approximated as a sphere. This is an apt approximation if 
+we only care about measuring the force exerted due to translational 
+motion. In reality, rotational motion is a significant factor to the 
+trajectory of non-spherical particles, in which case our understanding
+of how light interacts with particles is incomplete without considering
+the optical torque exerted by a focused beam.
 
-\subsection{Primary \& Secondary nucleation}
-From an industrial perspective, the nucleation process can 
-be broadly categorised into either primary of secondary
-nucleation. The former describes the formation of an initial
-nucleus within the bulk phase, absent of any external stimuli.
-Primary nucleation is therefore considered stochastic as there 
-is no means of predicting where a nucleus may form, or how long 
-it will take. The only reliable indicator being that higher 
-supersaturations will result in faster nucleation rates. At a 
-small scale one can estimate the nucleation rate by making 
-repeated measurements of sample solutions and seeing how many 
-have nucleated after a given time, giving us a Poisson probability distribution.
-\begin{align}
-	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
-	\label{eq:poisson}
-\end{align}
+\subsection{Optical Torque and rotation}
+\label{sec:opt_torque}
+Electromagnetic fields can transfer both linear and angular momentum \cite{Beth1936MechanicalDA}; more accurately the field is said to have 
+both orbital and spin momentum.
+Though there is some debate on how to decompose the total momentum into these two components \cite{Bruce2020, Svak2018}, for this project we do not need to calculate the exact
+quantities and will instead look at the broader effects of both 
+components. Orbital angular momentum arises from the shape of the 
+wavefront of the particular field in question; for simple Gaussian 
+beams the wavefronts are uniform and equally spaced resulting in the 
+typical radiation pressure that Ashkin and co demonstrated 
+\cite{Ashkin1980}. However, higher order modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have non-uniform wave fronts meaning 
+the orbital momentum has both angular and linear components; depending 
+on the relative size of the target particle one can induce rotation, 
+or orbiting \cite{Bruce2020, Courtial2000}.
 
-Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
-$V$ is the volume of the individual samples, and $M^*(t)$ \& 
-$M$ are the number of nucleated samples and the total number 
-of samples used respectively. The nucleation rate is one of 
-the most important factors used in evaluating a crystallisers 
-performance, providing a rough estimation of the final product 
-mass over a given time frame. While \eqref{eq:poisson} is useful 
-for studying the effects of different parameters at a small scale, 
-for industrial applications there are to many external factors 
-for primary nucleation to be measured accurately.
+Spin angular momentum (SAM) is attributed to the spin density of 
+the field, early research has shown that the spin density is non-
+zero for any beam despite the fact that the total SAM transferred 
+to a medium is 0 \cite{Svak2018, Bliokh2014}. This has sparked 
+debate if SAM is even a physical quantity as it does not aid in 
+the transport of energy directly \cite{Bliokh2014} and so cannot 
+be directly observed in some cases despite being non-zero. This 
+paradox is resolved by representing the wave as an array of spin 
+momentum loops that cancel one-another out when the medium is 
+homogeneous.Spacial inhomogeneities cause these spin loops to no 
+longer be equal, resulting in non-zero spin density, anisotropic 
+mediums (such as birefringent crystal lattices) experience a 
+transfer of spin angular momentum, imparting an optical torque.
 
-Secondary nucleation is the result of a initial seed crystal 
-inducing further nucleation within the bulk solution 
-\cite{Botsaris1976}. The seed crystals is often specially 
-prepared ahead of time and added to a supersaturate solution. 
-Due to interactions with the surrounding fluid and the container 
-walls it is possible for the seed crystals to act as a surface
-for further nucleation \cite{Anwar2015}. Control of secondary 
-nucleation events are crucial for ensuring reliable industrial 
-crystalliser performance. Fig.~\ref{fig:secondary} depicts an 
-attempt of classifying every possible mechanism that could lead 
-to secondary nucleation.
+Birefringence is a material property often seen in crystalline 
+materials, where the crystal lattice has a different structure 
+dependent on its orientation. Since light is composed of waves 
+that propagate in orthogonally to one another, a birefringent 
+material will refract light differently depending on the light's 
+polarisation. Therefore it can be said that the material has 
+two separate refractive indices. For circularly polarised light 
+this inhomogeneity results in a high degree of SAM being 
+transferred to the target object \cite{Parkin2009, Arita2016}. 
+The greater the difference between the two refractive indices the 
+greater the angular momentum transfer.
+
+The ability to transfer angular momentum has been exploited 
+to rotate microspheres as fast as 1000 Hz while suspended 
+in a bulk medium \cite{Arita2016} as well as a means of 
+measuring the local temperature and shear response of said 
+medium \cite{Millen2014, RodriguezSevilla2018}. Calculating 
+the optical torque applied to a birefringent material is given 
+via:
+\begin{equation}
+	\label{eq:opt_torque}
+	\begin{aligned}
+		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi 
+		\\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
+	\end{aligned}
+\end{equation}
+
+Where $\Delta n$ is the difference between the two refractive 
+indices, $\theta$ is the angle between the particle's long 
+axis and the polarisation vector of the local EM field, and 
+$\phi$ is the phase shift in the EM field. The first term 
+represents the 'orientational' torque which aligns the long 
+axis of the particle with the electric field, when aligned 
+$\theta=0$ meaning the entire term is negligible for particle's 
+with a stable orientation. The second term is due purely to the 
+polarisation of the laser, for circularly polarised light 
+$\phi=\pi/4$ thus maximising the torque transferred to the 
+target particle. Eq.~\eqref{eq:opt_torque} is only applicable 
+for particles with a known birefringence, but there are other 
+mechanisms that result in optical torque.
+
+A common example is shape induced birefringence. If a particle 
+has an anisotropic shape, it is more susceptible to being 
+polarised along its longer axis than its shorter axis. Consider, 
+for example, an ellipsoid elongated along one of its primary 
+axis' ($r_z > r_x = r_y$). In a plane polarised beam such a 
+particle will align with the polarisation vector. Therefore, 
+the particle will rotate as angular momentum is transferred 
+along its long axis. One common feature, regardless of shape,
+is that a particle with shape birefringence will rotate when 
+it lies perpendicular to the direction of propagation. This 
+is seen most evidently with spherical dimers \cite{Ahn2018, 
+Reimann2018} but even for elliptical particles rotational 
+motion is only detected when their long axis is not aligned 
+with the direction of propagation \cite{Zhu2021, Mihiretie2014}.
+
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
-	\caption{Secondary Nucleation mechanisms, classified by
-		Agrawal and Paterson \cite{Agrawal2015}}
-	\label{fig:secondary}
+	\includegraphics[width=\linewidth]{dimer_rotation.jpeg}
+	\caption{Example of nano-dimer undergoing optical rotation 
+		in a circularly polarised beam. Due to the dimer's 
+		anisotropic susceptibility it is polarised along its 
+		long axis, the dimer will therefore align its long axis 
+		with the polarization vector. Reproduced with permission 
+		from \cite{Reimann2018}.}
 \end{figure}
 
-This is not a universal classification system, there are a 
-variety of opinions on how best to characterise different 
-phenomena. It is heavily dependent on the theory used to 
-describe nucleation events, this is accurate for both 
-secondary and primary nucleation.
+Often the optical torque experienced is far greater than similar 
+spherical particles that are birefringent \cite{Bruce2020}. 
+Currently spherical dimers are being rotated in vacuums to 
+measure quantum forces and torques \cite{Ahn2018, Reimann2018}. 
+There are some alternative cases where particles are rotated 
+while not being aligned in the plane of the polarisation. However 
+in these cases their shape is often specifically engineered to 
+scatter light in such a way that the net momentum change always
+occurs in one direction regardless of the laser polarisation 
+\cite{Higurashi1994}.
 
-\section{Nucleation Theories}
-In order to predict the nucleation rate of a given solution
-we need a solid understanding of how the individual molecules
-interact such that they transition from an unordered phase to
-a ordered phase. Over the years several theories have been 
-proposed to explain how nucleation events occur, so far there
-has yet to be a standout candidate that reliably predicts 
-the systems behaviour. Either they fail to accurately predict
-nucleation rates or more generally cannot describe new 
-experimental results.
+Other examples of optical torque is when an anisotropic 
+particle is aligned with the beam's direction of propagation 
+(in which case $\theta=\pi/2$ and the first term disappears). 
+This is analogous to an optically trapped sphere, where 
+alongside a restoring force the particle also experiences a 
+restoring torque. This seemingly random rotational motion is 
+referred to as libation \cite{Bruce2020},often in typical 
+suspension trapping situations (where the particle is suspended 
+in a fluid) the rotational motion is washed out by the 
+translational motion. As such, many experiments elect to trap 
+in low pressure environments to precisely measure the optical 
+torque being exerted by the optical trap \cite{Ahn2018}. This 
+has lead experiments to try and achieve '0 kelvin' motion, 
+where by trapping a silica dimer they were able to restrict its 
+motion using 3 optical traps simultaneously. Despite this, they 
+found that the dimer's rotational motion about its long axis 
+could not be controlled leading to the undesired rotational
+modes \cite{Bang2020}.
 
-\subsection{Classical Nucleation Theory (CNT)}
-Sometimes referred to as 'Gibbs Nucleation Theory' the 
-original theory was first formed from the works of Volmer 
-and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
-While initially it was more focused on describing droplet
-formation in condensing vapours it was extrapolated to 
-describe crystallisation. The central premise of classical 
-theory is that nucleation occurs stochastically due to 
-collisions between individual solute molecules, ions, or 
-atoms. At the same time the bulk phase is resistant to the
-formation of a new phase. The competition between these 
-random collisions and the bulk solution can be used to 
-predict the probability of a newly formed nucleus.
+The detection and measurement of optical torque is still a 
+field of intense research, not only does it have potential 
+to understand quantum fluctuations in a particle's motion 
+but also allows for the creation of more effective
+micro-rotors. The latter being especially pertinent for 
+understanding the behaviour of fluids experiencing localised 
+shearing.
 
-Consider a supersaturated solution, at some time individual 
-sub units collide rapidly, one after another, forming a 
-nucleus of volume $4\pi r^3/3$. The newly formed phase has a 
-lower chemical potential than the surrounding solution, 
-reducing the free energy of the system. Simultaneously, 
-the formation of a new interface is resited by the bulk 
-phase due to surface tension. The net free energy of the 
-system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
-\begin{align}
-	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
-	\label{eq:CNT} 
-\end{align}
+\subsection{Characterisation of rotational motion}
+Rotational motion about a single axis is easiest to account for.  
+When the power spectra of elliptical polystyrene particles was 
+fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
+assumed that the rotational motion was purely in the transverse
+plane. As such they did not have to account for any variance in 
+the trapping strength due to orientation nor did they need to
+consider non-periodic rotational behaviour. In the case where 
+rotational motion is stochastic the problem is more complex. 
+For example, when an optical fibre trap characterisation 
+technique was implemented by Saffron \textit{et al} 
+\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
+light scattering to characterise both the axial and lateral 
+trap stiffness acting on microspheres.
 
-Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
-$v$ is the approximate volume of an individual molecule, 
-$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
-interfacial tension of the bulk solution. This assumes that 
-the nucleus will have a spherical morphology so that the 
-surface tension $\sigma_{inf}$ is a scalar value. Looking at 
-\eqref{eq:CNT} suggests that there must be some critical size 
-$r$ where the free energy gain from the nucleus exceeds the 
-surface tension of the surrounding fluid. This is reflected in 
-fig.~\ref{fig:free_energy} where we plot the free energy of the 
-system against nucleus size. This reveals a critical size above 
-which the gain in free energy exceeds the interfacial tension. 
-Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
-supersaturation of the system reduces said barrier. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
-	\caption{Free energy diagram of a newly formed nucleus according 
-		     to the Classical Nucleation Theory. The total free energy (blue)
-		     is due to the competition between the volume free energy gain
-		     (green) and the interfacial free energy cost (orange). Dotted
-		     lines are for a higher supersaturation than the solid lines,
-		     the interfacial energy cost is independent of supersaturation.}
-	\label{fig:free_energy}
+	\includegraphics[width=\linewidth]{BarZivDiagram.png}
+	\caption{Diagram depicting the setup used by \cite{BarZiv1997}, 
+	a microsphere is trapped by a focused beam. The scattered beam
+	is collected by an optical fibre situated close to the trapped
+	sphere. Using DLS they are able to relate the autocorrelation 
+	function of the collated signal to characterise the trap stiffness 
+	of the optical trap.}
 \end{figure}
 
-The maximum value of $\Delta G_{tot}$ is the free energy barrier 
-that any newly formed nucleus needs to overcome in order to stabilise. 
-The nucleation rate (the volume of new crystalline material formed per 
-unit time), is therefore commonly defined as being dependent on the 
-energy barrier $\Delta G^*$:
-\begin{align}
-	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
-\end{align}
+The only drawback admitted to in their work was that the technique 
+was constrained to isotropic scatters as their theoretical model 
+for describing the auto-correlation function was predicated on the 
+fact that any variations in the signal are due to the particles 
+translational motion within the confines of a cylindrical trap 
+\cite{BarZiv1997}. Where the upper limit of the cylindrical trap 
+is given by the Rayleigh range ($z_R = n\omega_0/NA$). This is 
+discussed further in chapter 3 in the discussion of dimer dynamics, 
+the axial traps of spherical aggregates is often situated far beyond the Rayleigh range (for a 1.2 NA laser this is $\pm5.985 \mu m$).
 
-Where $A$ is a pre-factor that can be fine tuned to the exact 
-demands of the system, the free energy barrier can be found by 
-finding the stationary point of $\Delta G_{tot}$.
-
-CNT is often regarded as a good description of the macro system, 
-its obvious that for all crystallization systems there is an 
-inherent energy barrier that dictates the nucleation rate. 
-Where it falters is in its predictive ability, both in estimating 
-nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
-structure of newly formed nuclei \cite{Lee1999, Yau2001}.
-Recent studies suggest classical nucleation is merely one of 
-many possible pathways that can be taken to produce an ordered 
-crystalline phase. Prompting the development of alternative 
-theories to better describe the nucleation process.
-
-\subsection{Two Step Nucleation}
-The two step nucleation theory is an extension to the CNT 
-that suggests that prior to nucleation, the solute will 
-form precursor structures. The CNT assumes that solute 
-structures below the critical radius should be unstable 
-within the solution due to interfacial forces trying to 
-dissolve the solute. This assumption is based on the idea 
-that any interactions between solute molecules are much
-weaker than that between solute and solvent molecules.
-If instead the solute molecule interactions are significant
-then it stands to reason that subcritical liquid-like 
-structures are stable.
-
-The first observation of a liquid-like structure was in 
-$CaCO_3$ solutions, initially it was assumed that polymers
-were a necessary additive to create a precursor structure 
-\cite{Driessche2017, Karthika2016}. This is not the case when 
-the solution is above its critical temperature, allowing the 
-solution to be in the binodal region of the phase diagram. 
-This would result in the formation of two liquid phases, 
-one being rich in solute, the other being solute poor 
-\cite{Karthika2016, Fu2021}. Several papers later reported 
-the presence of stable liquid-like clusters that formed 
-prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{oswalds_rule.jpg}
-	\caption{Free energy diagram of Oswald's rule for a 
-	crystal with three possible polymorphs. The diagram 
-	shows that there exist local minima in the free energy 
-	that represent the different polymorphic forms. A similar
-	free energy diagram can be used to describe the two-step
-	nucleation mechanism.}
-\end{figure}
-
-The formation of these clusters can be understood by Oswald's 
-rule; which says that any crystallising system does not 
-immediately take the path to the lowest possible energy state 
-but instead first transitions to the state with the smallest 
-free energy barrier \cite{Ostwald1897}. Oswald's rule was 
-assumed to be applicable for crystal polymorphs but can be
-generalised to describe the nucleation process also. Further phase 
-transitions can still occur but the pathway taken should 
-always minimises the overall free energy cost. It is assumed
-that between the amorphous phase and the crystalline phase 
-the solute undergoes dehydration while changing structurally 
-into a more ordered phase, though there is some discussion 
-whether the process merely undergoes dehydration, or if there 
-is a more complex chemical reaction
-\cite{Karthika2016}.
-
-With regards to liquid-precursors, there is still a lot of
-discussion about the relationship between the solution 
-conditions and the the precursors. The size and rate at 
-which these precursors agglomerate based on the degree of 
-saturation would provide invaluable information on the 
-kinetics of two-step nucleation \cite{Fu2021}. In some 
-cases it has even been shown that 'ageing' solutions is 
-a necessary step in order to detect the presence of 
-stable amorphous clusters \cite{Liao2022}. This was in 
-solutions that were supersaturated and yet after the 
-formation of precursors spontaneous nucleation was not
-seen even after several days of observation \cite{Liao2022}.
-This suggests that in some cases the precursors are
-metastable and require an external trigger to induce
-crystallisation. The presence of metastable structures
-and other phenomena has lead to development of a more 
-universal theory, namely non-classical nucleation.
-
-\subsection{Non-classical Nucleation} 
-Two-step nucleation is based on the idea that between
-the bulk solution and the final crystalline phase there
-exists and intermediate stage. While this is broadly an
-accepted concept, research has shown that in some cases
-the intermediate stage can be subdivided into multiple 
-states. This has been shown to be the case with metals
-in both solid and liquid conditions \cite{Cao2020, Ye2023}
-and colloidal simulations have indicated that the 
-precursors undergo structural change over time \cite{Tan2013}. 
-While in some cases it has been possible to directly
-monitor the structural change in the amorphous clusters, 
-many organic compounds have not received such treatment
-due to the difficulty of characterising and monitoring 
-the pre nucleation stage \cite{Fu2021}. This is often 
-why two-step nucleation is simply referred to as 
-multi-step nucleation.
-
-There are also a myriad of other unique nucleation events
-that deviate significantly from both the classical and 
-two-step nucleation theories. A review paper discussing
-recent advances in non-classical crystallisation highlighted
-that depending on the collision kinetics of the system 
-and the molecular interactions could result in drastically
-different reaction kinetics. In the former case, unstable
-subcritical nuclei have been shown to collide and coalesce
-into a supercritical nucleus, but only in conditions where 
-subcritical nucleation are plentiful despite a low degree
-of supersaturation \cite{Baumgartner2013}. The latter case 
-is more prevalent for dipole solute molecules, the anisotropic 
-attraction results in nuclei with non-spherical geometries, 
-this results in varying kinetics for agglomeration dependent 
-on the orientation of each molecule \cite{Yau2001}. Phenomena 
-such as the above examples has been collectively refereed to 
-as non classical nucleation, it is known that they must be 
-related to both CNT and multi-step nucleation but the connection 
-is not clear based on the current literature.
-
-As such, much of the research into nucleation theory is 
-focused on developing \textit{in situ} techniques that can 
-monitor and characterise the pre nucleation stage, this
-would glean information on cases when both classical 
-and non-classical nucleation is a possibility 
-\cite{Karthika2016}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{In situ techniques for studying nucleation}
-There are several different phenomena \cite{Fu2021, Karthika2016} 
-that are not well described by any individual theoretical framework 
-(CNT or multi-step nucleation for example). There is an increasing 
-interest in probing the pre-nucleated solution.
-
-One of the key conclusions is the need for the development of 
-better \textit{in situ} techniques for characterizing and 
-imaging the nucleation pathways. Understanding which specific 
-pathway is unfolding under a certain set of experimental 
-parameters would allow one to develop a map of the thermodynamic 
-landscape available to a pre-nucleation system. The challenge 
-therein lies in developing experimental methods by which one can 
-reproduce similar nucleation events and in turn use 
-characterisation techniques to identify the pathway taken. The 
-method not only needs to be repeatable for a single set of system 
-parameters, but also needs to be flexible enough to replicate 
-results regardless of the system parameters (i.e. varying 
-supersaturation, temperature, and solute choice). There have 
-been several different methods deployed to try and observe 
-crystallisation at a microscopic scale.
-
-\subsection{Computer Simulations}
-While not a direct observation of nucleation events, computer
-simulations have proved an invaluable tool for studying and 
-testing predictions about the pre-nucleated solution. By 
-comparing the simulative nucleation rate to the expected 
-rate given by current models, researchers can test newly 
-developed theories and glean information about the microscopic
-parameters of a system.
-
-The main challenge facing computer simulations is the issue
-of limited scope. Molecular dynamic simulations are constrained
-by their choice of time and length scale, for example simulations
-in a canonical ensemble face issues where the chemical potential
-is depleted to the point that the driving force for crystallisation
-is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
-addressed but requires an extensive increase in computational 
-resources to maintain the chemical potential equilibria. Like wise
-in most cases the choice of time scale is a crucial factor, shorter
-time scales provide a more accurate evolution of the solution; 
-however, in some cases the time scale magnitude for crystallisation
-is often far greater than the time scale used in the simulation. This
-has led to more advanced sampling techniques that allow for longer 
-time scales which has provided insights into the pre-nucleation 
-solution stage and formation of pre-nucleated clusters \cite{Finney2023}.
-
-The study of the pre-nucleated solution and the evolution of 
-precursors has been aided greatly by the use of simulative 
-studies. Early on it was shown in colloidal crystallisation 
-that the structure of the precursor was time dependent; 
-colloidal simulations showed that regardless of the molecular
-interactions between colloids the atomic packing pattern was 
-predominately hexagonal centred, over time the packing
-became dominated by body centred packing in the case 
-where the interactions were longer reaching \cite{Tan2013}. 
-They also confirmed that precursor formation was driven not
-by local density variations, but instead was driven by the 
-local bond order of the colloidal precursors \cite{Tan2013}, 
-providing strong evidence that the colloidal pre-nucleation 
-clusters undergo rearrangement prior to nucleation. Further
-developments in $NaCl$ simulations have shown that by varying
-the density fluctuations the precursor structure could either
-be amorphous, the classical rock-salt structure, or a new 
-structure similar to wurtzite. While this result is difficult
-to confirm it provides a important conclusion, that simulations
-can hint at new nucleation pathways that have not been confirmed
-experimentally. This means that computer simulations can not
-be merely used to confirm theories, but can be used to examine
-and search for novel results.
-
-Overall computer simulations make up a core area of research 
-in the study of nucleation. The technique is however limited,
-often struggling to examine nucleation over longer time scales
-or larger number of particles. While sampling methods and 
-enhanced simulation techniques has alleviated this somewhat 
-they are still far from being able to describe the behaviour 
-of larger molecules undergoing nucleation.
-
-\subsection{Transmission Electron Microscope (TEM)}
-One of the more well known techniques used for direct observations 
-of crystallisation dynamics is using transmission electron 
-microscopy (TEM). The basic working principle involves creating a 
-focused beam of electrons that are directed onto a target sample.
-Due to the wave-particle duality, the electrons can be treated 
-as a unified beam of light, whose wavelength is dependent on the 
-kinetic energy of the electrons. The wavelength of an individual 
-electron is given by \cite{Williams2009}:
-\begin{align}
-	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
-\end{align}
-
-\noindent
-When incident on a sample the wave is scattered by the sample 
-and can then be detected by an imaging screen that measures 
-the difference in intensity between the scattered and 
-transmitted beams. The advantage of TEM is that one can bypass
-the diffraction limit imposed by visible light, the maximum 
-resolution that can be achieved by a microscope is given by
-\cite{Champness2020}:
-\begin{align}
-	d = \frac{\lambda}{2NA}
-\end{align}
-
-\noindent
-where NA is the numerical aperture of the objective lens, and
-$\lambda$ is the wavelength of light used to illuminate a sample.
-For a visible light LED with an average wavelength of $550~nm$, 
-the upper limit for an objective lens (NA of 1.2) is around 
-$230~nm$ whereas a similar lens focusing an electron beam 
-with an average kinetic energy of $100~keV$ will have a maximum 
-resolution of just over $4~pm$ \cite{Champness2020}. There is
-an upper limit to the maximum resolution possible as electron 
-lens are far inferior to what modern optics can achieve 
-\cite{Champness2020} so there is a trade off between quality and 
-resolution.
-
-TEM technology has proven invaluable to the study of nucleation
-at an atomic level. The samples are often only a few hundred 
-atoms thick in order to optimise the imaging results. Much of 
-the supporting evidence for non-classical nucleation has been 
-demonstrated via TEM \cite{Ye2023}. For example when Cao 
-\textit{et al} observed the formation of amorphous atomic 
-cluster using $\gamma-$Fe,Au, and Re; all three of which 
-underwent continuous transformation before eventually 
-reaching an final ordered state \cite{Cao2020}. The amorphous 
-cluster did not immediately transition to an ordered structure, 
-instead fluctuating somewhat randomly until eventually settling 
-into a crystalline structure \cite{Cao2020}. This was confirmed 
-both in solid growth and liquid growth conditions \cite{Cao2020, 
-Loh2016}, suggesting that the non-classical nucleation path 
-dominates in all cases of metal crystal formation \cite{Ye2023}. 
-Confirmation is necessary for different solute choices, as in 
-the case of organic solids it is believed that non-classical 
-nucleation only dominates in low supersaturations whereas 
-classical nucleation is dominate at higher supersaturations.
-
-TEM analysis has not been extended to the study of organic 
-crystals, proteins, or simple salts at an atomic level 
-however. The main reason being that organic materials can be 
-damaged by the electron beam and therefore need to be frozen 
-prior to being imaged \cite{Champness2020}. This was used to 
-study the polymorphic time dependence of glycine crystals, by 
-varying the time prior to freezing Broadhurst \textit{et al} 
-showed that $\beta$ glycine dominates initially but $\alpha$-
-glycine is more stable after a longer period, $\gamma$-glycine 
-was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
-the time scales with which polymorphs form, it doesn't provide 
-much information on the kinetics of crystal growth as these 
-were snap shots and not direct observations of the nucleation 
-process.
-
-Overall TEM analysis is invaluable for the study of atomic 
-crystal growth, the secondary scattering from electron 
-beams also allows for chemical and structural information 
-to be gleaned from the target. This has given strong evidence
-for the non-classical nucleation theory, particularly in 
-metal crystal formation. Due to the high energy of the 
-electrons, organic materials have not received the same
-focus outside of cryoTEM analysis. Non-invasive methods
-are therefore necessary to gain an understanding of the
-crystal growth and formation of organic crystals, 
-particularly when both classical and multi-step nucleation 
-can occur.
-
-\subsection{Laser induced nucleation}
+\section{Laser induced nucleation}
 From as early as 1996 it has been known that laser irradiation 
@@
 
-\section{Optical Tweezers}
-\subsection{Background}
-Optical tweezing has been a field of applied optics ever since 
-the 1980s when Ashkin \cite{Ashkin1970} first showed that focused 
-light was capable of trapping micron sized particles due to light 
-exerting 'radiation pressure'. The working principle was that a 
-light source such as a laser could trap small objects within a 2D 
-plane, as long as the light source had an approximately Gaussian 
-profile (colloquially called a Gaussian beam). Soon after, Ashkin 
-showed that the introduction of a microscope objective would allow 
-one to focus the light source to a diffraction limited point that 
-would stably trap small objects within a confined volume 
-\cite{Ashkin1980}. This allowed Ashkin and others to study individual 
-biological cells as the overall force was on the order of $10^{-12}\ 
-N$ while being non-invasive to the internal structure. Later it would 
-be used to probe microscopic properties such as the formation of 
-colloidal aggregates \cite{Yi2021} to the drag forces exerted by a 
-pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
-behaviour of light, optical tweezers have become essential for 
-measuring and exerting precise forces on the magnitude of pico-newtons 
-allowing one to probe the material properties of the smallest 
-materials.
-
-\subsection{Optical Torque and rotation}
-\label{sec:opt_torque}
-Electromagnetic fields can transfer both linear and angular momentum \cite{Beth1936MechanicalDA}; more accurately the field is said to have 
-both orbital and spin momentum.
-Though there is some debate on how to decompose the total momentum into these two components \cite{Bruce2020, Svak2018}, for this project we do not need to calculate the exact
-quantities and will instead look at the broader effects of both 
-components. Orbital angular momentum arises from the shape of the 
-wavefront of the particular field in question; for simple Gaussian 
-beams the wavefronts are uniform and equally spaced resulting in the 
-typical radiation pressure that Ashkin and co demonstrated 
-\cite{Ashkin1980}. However, higher order modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have non-uniform wave fronts meaning 
-the orbital momentum has both angular and linear components; depending 
-on the relative size of the target particle one can induce rotation, 
-or orbiting \cite{Bruce2020, Courtial2000}.
-
-Spin angular momentum (SAM) is attributed to the spin density of 
-the field, early research has shown that the spin density is non-
-zero for any beam despite the fact that the total SAM transferred 
-to a medium is 0 \cite{Svak2018, Bliokh2014}. This has sparked 
-debate if SAM is even a physical quantity as it does not aid in 
-the transport of energy directly \cite{Bliokh2014} and so cannot 
-be directly observed in some cases despite being non-zero. This 
-paradox is resolved by representing the wave as an array of spin 
-momentum loops that cancel one-another out when the medium is 
-homogeneous.Spacial inhomogeneities cause these spin loops to no 
-longer be equal, resulting in non-zero spin density, anisotropic 
-mediums (such as birefringent crystal lattices) experience a 
-transfer of spin angular momentum, imparting an optical torque.
-
-Birefringence is a material property often seen in crystalline 
-materials, where the crystal lattice has a different structure 
-dependent on its orientation. Since light is composed of waves 
-that propagate in orthogonally to one another, a birefringent 
-material will refract light differently depending on the light's 
-polarisation. Therefore it can be said that the material has 
-two separate refractive indices. For circularly polarised light 
-this inhomogeneity results in a high degree of SAM being 
-transferred to the target object \cite{Parkin2009, Arita2016}. 
-The greater the difference between the two refractive indices the 
-greater the angular momentum transfer.
-
-The ability to transfer angular momentum has been exploited 
-to rotate microspheres as fast as 1000 Hz while suspended 
-in a bulk medium \cite{Arita2016} as well as a means of 
-measuring the local temperature and shear response of said 
-medium \cite{Millen2014, RodriguezSevilla2018}. Calculating 
-the optical torque applied to a birefringent material is given 
-via:
-\begin{equation}
-	\label{eq:opt_torque}
-	\begin{aligned}
-		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi 
-		\\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
-	\end{aligned}
-\end{equation}
-
-Where $\Delta n$ is the difference between the two refractive 
-indices, $\theta$ is the angle between the particle's long 
-axis and the polarisation vector of the local EM field, and 
-$\phi$ is the phase shift in the EM field. The first term 
-represents the 'orientational' torque which aligns the long 
-axis of the particle with the electric field, when aligned 
-$\theta=0$ meaning the entire term is negligible for particle's 
-with a stable orientation. The second term is due purely to the 
-polarisation of the laser, for circularly polarised light 
-$\phi=\pi/4$ thus maximising the torque transferred to the 
-target particle. Eq.~\eqref{eq:opt_torque} is only applicable 
-for particles with a known birefringence, but there are other 
-mechanisms that result in optical torque.
-
-A common example is shape induced birefringence. If a particle 
-has an anisotropic shape, it is more susceptible to being 
-polarised along its longer axis than its shorter axis. Consider, 
-for example, an ellipsoid elongated along one of its primary 
-axis' ($r_z > r_x = r_y$). In a plane polarised beam such a 
-particle will align with the polarisation vector. Therefore, 
-the particle will rotate as angular momentum is transferred 
-along its long axis. One common feature, regardless of shape,
-is that a particle with shape birefringence will rotate when 
-it lies perpendicular to the direction of propagation. This 
-is seen most evidently with spherical dimers \cite{Ahn2018, 
-Reimann2018} but even for elliptical particles rotational 
-motion is only detected when their long axis is not aligned 
-with the direction of propagation \cite{Zhu2021, Mihiretie2014}.
-
-Often the optical torque experienced is far greater than similar 
-spherical particles that are birefringent \cite{Bruce2020}. 
-Currently spherical dimers are being rotated in vacuums to 
-measure quantum forces and torques \cite{Ahn2018, Reimann2018}. 
-There are some alternative cases where particles are rotated 
-while not being aligned in the plane of the polarisation. However 
-in these cases their shape is often specifically engineered to 
-scatter light in such a way that the net momentum change always
-occurs in one direction regardless of the laser polarisation 
-\cite{Higurashi1994}.
-
-Other examples of optical torque is when an anisotropic 
-particle is aligned with the beam's direction of propagation 
-(in which case $\theta=\pi/2$ and the first term disappears). 
-This is analogous to an optically trapped sphere, where 
-alongside a restoring force the particle also experiences a 
-restoring torque. This seemingly random rotational motion is 
-referred to as libation \cite{Bruce2020},often in typical 
-suspension trapping situations (where the particle is suspended 
-in a fluid) the rotational motion is washed out by the 
-translational motion. As such, many experiments elect to trap 
-in low pressure environments to precisely measure the optical 
-torque being exerted by the optical trap \cite{Ahn2018}. This 
-has lead experiments to try and achieve '0 kelvin' motion, 
-where by trapping a silica dimer they were able to restrict its 
-motion using 3 optical traps simultaneously. Despite this, they 
-found that the dimer's rotational motion about its long axis 
-could not be controlled leading to the undesired rotational
-modes \cite{Bang2020}.
-
-The detection and measurement of optical torque is still a 
-field of intense research, not only does it have potential 
-to understand quantum fluctuations in a particle's motion 
-but also allows for the creation of more effective
-micro-rotors. The latter being especially pertinent for 
-understanding the behaviour of fluids experiencing localised 
-shearing.
-
-\subsection{Characterisation of rotational motion}
-Rotational motion about a single axis is easiest to account for.  
-When the power spectra of elliptical polystyrene particles was 
-fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
-assumed that the rotational motion was purely in the transverse
-plane. As such they did not have to account for any variance in 
-the trapping strength due to orientation nor did they need to
-consider non-periodic rotational behaviour. In the case where 
-rotational motion is stochastic the problem is more complex. 
-For example, when an optical fibre trap characterisation 
-technique was implemented by Saffron \textit{et al} 
-\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
-light scattering to characterise both the axial and lateral 
-trap stiffness acting on microspheres. The only drawback 
-admitted to in their work was that the technique was constrained 
-to isotropic scatters as their theoretical model for describing 
-the auto-correlation function was predicated on the fact that 
-any variations in the signal are due to the particles 
-translational motion within the confines of a cylindrical trap 
-\cite{BarZiv1997}. Where the upper limit of the cylindrical 
-trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
-However, as demonstrated by the results from Chapter 4.1, 
-the axial traps of spherical aggregates is often situated 
-far beyond the Rayleigh range (for a 1.2 NA laser this is 
-$\pm5.985 \mu m$).
-
 \section{Shear induced Nucleation}
@@
 measure the optical torque applied to a non-birefringent particle. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-06-27 Latex_files/Thesis/chapters/chapter 2/main.tex
27/06/2025 chapters 1,3,4 and structural changes

@@
 \label{chapter:theory}
-The two key areas of this project cover two somewhat unconnected
-fields; crystallisation theory, which is covered in the 
-previous introduction section, and tweezing theory. The 
-following chapter summarises the working principles behind 
-optical tweezing, the differences in scattering theories, 
-and the broader theoretical calculations used throughout the 
-rest of the thesis.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textit{\textbf{Add overview here}} \\
+This chapter covers the theory behind light and particle interactions, 
+and how the exact theory varies for differently sized particles. After 
+that we show how this can be applied to existing diffusion models to 
+accurately simulate the behaviour of a particle within the vicinity of 
+an optical trap. Lastly, we highlight the techniques utilised in 
+experiments to both calibrate and characterise an optical trap based 
+on our understanding of the trap.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Electromagnetism and optical tweezing}
@@
 Proper understanding of optical tweezing requires a description
-of how the trapped particle interacts with the laser. 
-From an electromagnetism perspective the laser creates a 
-spatially and temporally coherent electric field that scatters 
-light when passing though a medium. The laws governing electric 
-and magnetic fields are summarised most succinctly via the 
-Maxwell equations. The differential forms of which are given 
-below \cite{Jackson_1975}:
-\begin{align}
-  \nabla \cdot \mathbf{E}
-  &= \frac{\rho_ v}{\epsilon_0}
-  \label{eq:max_start}
+of how the trapped particle interacts with the laser. From an
+electromagnetism perspective the laser creates a spatially and 
+temporally coherent electric field that scatters light when 
+passing though a medium. The laws governing electric and magnetic 
+fields are summarised most succinctly via the Maxwell equations. 
+These four equations describe how the electric and magnetic fields 
+behave and relate to one another. Any discussion of optical trapping 
+is underpinned by the fact that in all cases the Maxwell equations 
+must be satisfied. Macroscopic forms of the Maxwell equations are 
+given below
+\cite{Jackson_1975}:
+\begin{align}
+	\nabla \cdot \mathbf{D}
+	&= \rho_f
 	\\
@@
 	\\
-  \nabla \times \mathbf{B}
-  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}
-  \label{eq:max_end}    
-\end{align}
-where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
-charge current, and $\epsilon_0$ is the permittivity of free 
-space.
-
-These four equations describe how the electric and magnetic fields 
-behave and relate to one another at a microscopic level. Any 
-discussion of optical trapping is underpinned by the fact that in 
-all cases the Maxwell equations must be satisfied at the macroscopic 
-level, where one must account for the medium's interactions with the 
-EM field. Macroscopic forms of the Maxwell equations are given below
-\cite{Jackson_1975}:
-\begin{align}
-	\nabla \cdot \mathbf{D}
-	&= \rho_f
-	\\
-	\nabla \cdot \mathbf{B}
-	&= 0
-	\\
-	\nabla \times \mathbf{E}
-	&= \frac{\partial \mathbf{B}}{\partial t}
-	\\
 	\nabla \times \mathbf{H}
@@
 \end{align}
-Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and 
+where $\rho_f$ and $\mathbf{J}_f$ are the charge density and 
 current caused by the presence of free charges in the medium, 
 and $\mathbf{D}$ and $\mathbf{H}$ are the displacement and 
-magnetising fields respectively. The latter two are defined 
-by:
-\begin{align}
-	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
-	\\ 
-	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
-\end{align}
-
-Where $\mu_0$ is the vacuum permeability $\mathbf{P}$ 
-is the polarization field and $\mathbf{M}$ is the 
-magnetisation field, these fields arise due to the 
-bound charges throughout the medium interacting with 
-the EM field.
+magnetising fields respectively. The latter two are defined by:
+\begin{align}
+	\mathbf{D} &= \epsilon\mathbf{E}+\mathbf{P} 
+	\\ 
+	\mathbf{H} &= \frac{1}{\mu}\mathbf{B}-\mathbf{M}
+\end{align}
+where $\mu$ is the vacuum permeability,, $\epsilon$ is the 
+permittivity of free space, $\mathbf{P}$ is the polarization 
+field and $\mathbf{M}$ is the magnetisation field, these 
+fields arise due to the bound charges throughout the medium 
+interacting with the EM field.
 
@@
 \end{align} 
-
-Where $k$ is the wave number of the electromagnetic radiation 
-($k = 2\pi/\lambda$). This combined with the constraints of 
+where $k$ is the wave number of the electromagnetic radiation 
+($k = \omega^2 \mu \epsilon/c^2$). This combined with the constraints of 
 Maxwell's equations leaves very few exact solutions apart from 
@@
 	\\
-	\label{eq:internal}
-	{\bf E}_{\rm int}({\bf r})
-	&=
-	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
-	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
-	\right]
-	\\
 	\label{eq:scattered}
@@
 	\right] 
-\end{align}
-
-where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
+	\\
+	\label{eq:internal}
+	{\bf E}_{\rm int}({\bf r})
+	&=
+	{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r})
+\end{align}
+
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$
 are the expansion coefficients of each of the fields, and ${\bf M}_
@@
 and refractive index. The $T$-matrix method can be easily 
-applied to any solution of Maxwell's equations, for example 
-Eq.~\eqref{eq:scattered} can be instead be rewritten as:
-\begin{align}
-{\bf E}_{\rm scat}({\bf r})
-&=
-E_0 \sum_{n}^\infty \sum_{m=-n}^n 
-\begin{pmatrix}
-	T_{11} & T_{12} \\
-	T_{21} & T_{22}
-\end{pmatrix}
-\begin{pmatrix}
-	a_{mn} \\
-	b_{mn}
-\end{pmatrix}
-\begin{pmatrix}
-	{\bf N}_{mn}^{(3)} & {\bf M}_{mn}^{(3)}
-\end{pmatrix}
-\end{align}
-
-In order to compute the optical force and torque applied to a 
-given object one simply needs to know the momentum carried by 
-the incident and scattered fields. So long as we know the 
-expansion coefficients of the incident beam we can find the 
-scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
-compute the force. This project utilises the results from 
-Farsund \cite{Farsund1996} but in the form used by Crichton 
-\cite{Crichton2000THEMD}. As an example the time averaged linear 
-momentum carried by any incident beam can be computed via:
-\begin{equation}
-	\label{eq:linear_momentum_in}
-	\begin{split}
-		\langle P_{z,\ in} \rangle
-		&=
-		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
-		\\ 
-		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-		\\
-		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
-	\end{split}
-\end{equation}
+applied to any solution of Maxwell's equations. In order to compute 
+the optical force and torque applied to a given object one simply 
+needs to know the momentum carried by the incident and scattered 
+fields. So long as we know the expansion coefficients of the 
+incident beam we can find the scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to compute the force.
 
@@
 the Rayleigh approximation.  
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ray-Optics Regime}
-
-The Ray-Optics model is the simplest to understand, this theory models 
-the electromagnetic field as a collection of individual 'rays' that 
-propagate and are refracted according to Snell's Law. Based on the 
-change in direction momentum is transferred to the target particle; 
-with rays closest to the centre of the beam having greater intensity 
-than those rays at the very edge of the beam. Consider a particle 
-struck by two rays in a Gaussian beam, one coming close to the centre, 
-and the other ray coming from the edge. As each ray is refracted by 
-the target sphere a force is imparted onto it, the total force 
-imparted is given by:
-\begin{align}
-	F_i = Q_i\frac{\Delta n P_i}{c}
-\end{align}
-
-\noindent 
-where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
-in refractive indices between the solution and the target particle,
-and $P_i$ is the power of the individual ray. The net force can be 
-subdivided into its gradient and scattering components, where the latter
-directs the particle to the centre of the beam, and the former acts on 
-the particle in the direction of beam propagation. For a beam with a
-Gaussian intensity distribution $P_i$ will fall off as you move from
-the centre of the beam. The ray optics model is ideal when dealing 
-with larger particles whose diameter far exceed the wavelength of light
-being scattered. While it can be useful in predicting the forces 
-experienced by said particle's it does not fully capture the behaviour
-of light when considering interference between different rays. And 
-as such no information is learned about the scattered field.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Rayleigh Scattering}
-
-The Rayleigh approximation is for describing a particle when 
-the particle radius is $\ll \lambda$. The underlying theory 
-is that a dielectric sphere can be treated as a dipole while 
-in the presence of the electromagnetic field. In which case 
-the scattering force is given simply by the scattering of 
-the induced dipole, and the gradient force is due to the 
-Lorentz force \cite{Gordon1973}. The gradient forces in the 
-principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
-rectangular force field:
-\begin{align}
-  F_{grad,x}
-  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
-  \label{eq:Rayleigh_start}\\
-  F_{grad,y}
-  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
-  \\
-  F_{grad,z}
-  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
-    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
-    \nonumber \times I(r)
-  \\ 
-  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-  \\
-  \label{eq:Rayleigh_end}
-\end{align}
-where, for a approximate Gaussian beam the intensity $I(r)$ 
-close to the beam focus can be calculated as such.
-\begin{align}
-  \nonumber
-	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
-	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
-\end{align}
-
-\noindent
-Where $m$ is the relative refractive index ($n_1/n_2$), $n_2$ is 
-the particle's refractive index, $\omega_0$ is the beam waist, 
-$a$ is the radius of the particle, and $\tilde{x}, \tilde{y}, 
-\tilde{z}$ are simply the $x, y, z$ coordinates but scaled by the 
-beam radius. The scattering force however, is dependent on the 
-effective scattering cross-sectional area. 
-\begin{align}
-  F_{\rm scat}
-  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
-  \text{where:} \nonumber \\
-  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
-\end{align}
-
-There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
-\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- \eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
-the state of the scattered field. This is because if the particle 
-is approximated by a point dipole we can ignore the momentum 
-transfer calculation and simply compute the Lorenz force for a 
-dipole in an electric field. While not important for optical force calculation higher complexity scattering problems can be simplified 
-by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), in which case the scattered field is now significant. As the target particle gets larger this approximation 
-fails to accurately describe the trapping force \cite{Li2021}. For 
-larger particles the optical force and scattering is better described 
-via the Lorenz-Mie theory.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Scattering methods}
-\label{sec:scattering}
-There are several methods available to calculate the scattered 
-field produced by a particle. While one can compute this by 
-directly applying Lorenz-Mie theory, describing the interactions 
-between focused beams and complicated shapes can be computationally 
-slow and memory intensive. Several alternative scattering approaches 
-have been developed that allow for faster computation times 
-simplifying the scattering problem. While all of these methods will
-produce similar results the key difference is in how one represents
-the scatterer.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{T-matrix Method}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Discrete Dipole Approximation}
@@
 \newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Ray-Optics Regime}
+
+The Ray-Optics model is the simplest to understand, this theory models 
+the electromagnetic field as a collection of individual 'rays' that 
+propagate and are refracted according to Snell's Law. Based on the 
+change in direction momentum is transferred to the target particle; 
+with rays closest to the centre of the beam having greater intensity 
+than those rays at the very edge of the beam. Consider a particle 
+struck by two rays in a Gaussian beam, one coming close to the centre, 
+and the other ray coming from the edge. As each ray is refracted by 
+the target sphere a force is imparted onto it, the total force 
+imparted is given by:
+\begin{align}
+	F_i = Q_i\frac{\Delta n P_i}{c}
+\end{align}
+
+\noindent 
+where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
+in refractive indices between the solution and the target particle,
+and $P_i$ is the power of the individual ray. The net force can be 
+subdivided into its gradient and scattering components, where the latter
+directs the particle to the centre of the beam, and the former acts on 
+the particle in the direction of beam propagation. For a beam with a
+Gaussian intensity distribution $P_i$ will fall off as you move from
+the centre of the beam. The ray optics model is ideal when dealing 
+with larger particles whose diameter far exceed the wavelength of light
+being scattered. While it can be useful in predicting the forces 
+experienced by said particle's it does not fully capture the behaviour
+of light when considering interference between different rays. And 
+as such no information is learned about the scattered field.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Rayleigh Scattering}
+
+The Rayleigh approximation is for describing a particle when 
+the particle radius is $\ll \lambda$. The underlying theory 
+is that a dielectric sphere can be treated as a dipole while 
+in the presence of the electromagnetic field. In which case 
+the scattering force is given simply by the scattering of 
+the induced dipole, and the gradient force is due to the 
+Lorentz force \cite{Gordon1973}. The gradient forces in the 
+principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
+rectangular force field:
+\begin{align}
+  F_{grad,x}
+  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
+  \label{eq:Rayleigh_start}\\
+  F_{grad,y}
+  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
+  \\
+  F_{grad,z}
+  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
+    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
+    \nonumber \times I(r)
+  \\ 
+  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
+  \\
+  \label{eq:Rayleigh_end}
+\end{align}
+where, for a approximate Gaussian beam the intensity $I(r)$ 
+close to the beam focus can be calculated as such.
+\begin{align}
+  \nonumber
+	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
+	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
+\end{align}
+
+\noindent
+Where $m$ is the relative refractive index ($n_1/n_2$), $n_2$ is 
+the particle's refractive index, $\omega_0$ is the beam waist, 
+$a$ is the radius of the particle, and $\tilde{x}, \tilde{y}, 
+\tilde{z}$ are simply the $x, y, z$ coordinates but scaled by the 
+beam radius. The scattering force however, is dependent on the 
+effective scattering cross-sectional area. 
+\begin{align}
+  F_{\rm scat}
+  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
+  \text{where:} \nonumber \\
+  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
+\end{align}
+
+There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
+\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- \eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
+the state of the scattered field. This is because if the particle 
+is approximated by a point dipole we can ignore the momentum 
+transfer calculation and simply compute the Lorenz force for a 
+dipole in an electric field. While not important for optical force calculation higher complexity scattering problems can be simplified 
+by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), in which case the scattered field is now significant. As the target particle gets larger this approximation 
+fails to accurately describe the trapping force \cite{Li2021}. For 
+larger particles the optical force and scattering is better described 
+via the Lorenz-Mie theory.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Describing the trajectory of a trapped particle}
@@
 
-\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation of spherical aggregates}
+\label{sec:simulations}
+Later chapters cover the dynamics of spherical aggregates and anisotropic 
+scatterers, these subjects are particularly difficult to characterise 
+using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
+As an example consider a symmetric dimer as a paradigmatic aggregate; 
+if we consider the Langevin equation for such a aggregate within an optical 
+trap we have:
+\begin{align}
+	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
+	\sqrt{2\vec{D}_x}\eta(t)
+\end{align}
+
+Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
+motion is generalised to a 3 dimensional case. Here we omit the inertia term 
+as the restoration time of the optical trap is far greater than the characteristic time of the particle ($\kappa/\gamma \gg m/\gamma$).
+Now, the dimer is undergoing random rotational motion in addition to its Brownian
+translational motion the first term on the right hand side is no longer 
+purely a function of the dimer's position but also of its orientation. 
+The rotational form of the Langevin equation for a dipole within an external 
+potential is given as \cite{Mazo2008}:
+\begin{align}
+	\frac{{d}\vec{u}(t)}{{dt}}
+	=
+	\frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
+	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of 
+the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the 
+rotational friction coefficient which is given as $\gamma_R = 8\pi
+\eta r^3$ for a sphere. If we assume that over the course of a 
+trajectory the dimer has a 'equilibrium' orientation in which the 
+optical torque is minimised we can say that the first right hand 
+term can be replaced with $\vec{\kappa}_u(\vec{r}(t))$ to represent
+a harmonic potential due to the torque. $\lambda(t)$ is the Brownian 
+rotations from the surrounding fluid, like in the translational case 
+the Brownian rotations are normally distributed and are also uncorrelated 
+so that:
+\begin{align}
+	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
+\end{align}
+
+For an asymmetric scatterer whose radius is comparable to that of the 
+electric field's wavelength we now have a system of simultaneous equations:
+\begin{align}
+	\label{eq:full_langevin}
+	\frac{{d}\vec{r}(t)}{{dt}}
+	&=
+	\frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2k_BT\vec{D}}\eta(t)
+	\\
+	\frac{{d}\vec{u}(t)}{{dt}}
+	&=
+	\frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
+	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+
+Fortunately, we do not need to solve these directly as the latter two
+random variables can be easily approximated if the thermal energy of
+the system is known, and the rate of change can be assumed as linear if
+we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
+\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
+to compute the optical force and torque applied to the dimer.
+
+Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
+or \textit{ott} \cite{Nieminen2007} we can compute the beam 
+shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
+type. Using the results from \cite{Farsund1996} we can then 
+compute both the optical force and torque using the beam 
+coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
+which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
+\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
+total force in along the z-direction and the total torque about 
+the z-axis are provided:
+\begin{equation}
+	\label{eq:optical_force}
+	\begin{split}
+		\bold{F_z}
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+\begin{equation}
+	\begin{split}
+		\bold{T_z}
+		&=
+		-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
+		\\ 
+		&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
+coefficients of the incident and scattered fields respectively. 
+We can get the $x$ and $y$ force and torque components in a 
+similar form by applying a simple rotation transformation. With 
+the optical forces and torques computed all that remains is to 
+compute the Brownian forces and torques which are constrained 
+by the relation.
+\begin{align}
+	\langle q_iq_j\rangle =2D_{ij}\Delta t
+\end{align}
+
+Vigilante et al compiled together a python package that combines 
+both \textit{ott} and \textit{mstm} to simulate the behaviour of 
+spherical aggregates within a predefined optical trap 
+\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
+modules that can be used to model the scattering from different 
+shaped particles; however, for aggregates of spheres the resultant 
+force is nonsensical (see Chapter 4 for further details). As such 
+we rely on the \textit{mstm} for simulating dimers.  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Calibration Techniques}
@@
 This is applicable to particles with a polarizable long axis but is insufficient for birefringent particles.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulation of spherical aggregates}
-\label{sec:simulations}
-Later chapters cover the dynamics of spherical aggregates and anisotropic 
-scatterers, these subjects are particularly difficult to characterise 
-using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
-As an example consider a symmetric dimer as a paradigmatic aggregate; 
-if we consider the Langevin equation for such a aggregate within an optical 
-trap we have:
-\begin{align}
-	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
-	\sqrt{2\vec{D}_x}\eta(t)
-\end{align}
-
-Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
-motion is generalised to a 3 dimensional case. Here we omit the inertia term 
-as the restoration time of the optical trap is far greater than the characteristic time of the particle ($\kappa/\gamma \gg m/\gamma$).
-Now, the dimer is undergoing random rotational motion in addition to its Brownian
-translational motion the first term on the right hand side is no longer 
-purely a function of the dimer's position but also of its orientation. 
-The rotational form of the Langevin equation for a dipole within an external 
-potential is given as \cite{Mazo2008}:
-\begin{align}
-  \frac{{d}\vec{u}(t)}{{dt}}
-  =
-  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
-  + \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
-\end{align}
-
-Where $\vec{u}(t)$ is the unit vector aligned along the centres of 
-the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the 
-rotational friction coefficient which is given as $\gamma_R = 8\pi
-\eta r^3$ for a sphere. If we assume that over the course of a 
-trajectory the dimer has a 'equilibrium' orientation in which the 
-optical torque is minimised we can say that the first right hand 
-term can be replaced with $\vec{\kappa}_u(\vec{r}(t))$ to represent
-a harmonic potential due to the torque. $\lambda(t)$ is the Brownian 
-rotations from the surrounding fluid, like in the translational case 
-the Brownian rotations are normally distributed and are also uncorrelated 
-so that:
-\begin{align}
-  \left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
-\end{align}
-
-For an asymmetric scatterer whose radius is comparable to that of the 
-electric field's wavelength we now have a system of simultaneous equations:
-\begin{align}
-	\label{eq:full_langevin}
-  \frac{{d}\vec{r}(t)}{{dt}}
-  &=
-    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2k_BT\vec{D}}\eta(t)
-  \\
-  \frac{{d}\vec{u}(t)}{{dt}}
-  &=
-    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
-    + \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
-\end{align}
-
-Fortunately, we do not need to solve these directly as the latter two
-random variables can be easily approximated if the thermal energy of
-the system is known, and the rate of change can be assumed as linear if
-we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
-\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
-to compute the optical force and torque applied to the dimer.
-
-Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
-or \textit{ott} \cite{Nieminen2007} we can compute the beam 
-shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
-type. Using the results from \cite{Farsund1996} we can then 
-compute both the optical force and torque using the beam 
-coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
-which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
-\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
-total force in along the z-direction and the total torque about 
-the z-axis are provided:
-\begin{equation}
-	\label{eq:optical_force}
-\begin{split}
-  \bold{F_z}
-  &=
-    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
-  \\ 
-  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-  \\
-  & \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
-\end{split}
-\end{equation}
-\begin{equation}
-\begin{split}
-  \bold{T_z}
-  &=
-    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
-  \\ 
-  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-  \\
-  & \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
-\end{split}
-\end{equation}
-
-where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
-coefficients of the incident and scattered fields respectively. 
-We can get the $x$ and $y$ force and torque components in a 
-similar form by applying a simple rotation transformation. With 
-the optical forces and torques computed all that remains is to 
-compute the Brownian forces and torques which are constrained 
-by the relation.
-\begin{align}
-	\langle q_iq_j\rangle =2D_{ij}\Delta t
-\end{align}
-
-Vigilante et al compiled together a python package that combines 
-both \textit{ott} and \textit{mstm} to simulate the behaviour of 
-spherical aggregates within a predefined optical trap 
-\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
-modules that can be used to model the scattering from different 
-shaped particles; however, for aggregates of spheres the resultant 
-force is nonsensical (see Chapter 4 for further details). As such 
-we rely on the \textit{mstm} for simulating dimers.
-
-\subsection{Simulation reference frame}
-\label{sec:sim_parameters}
-Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
-to perform a systematic study of the dynamics demonstrated by 
-asymmetric dimers in both plane and circularly polarised light. 
-The simulations have two primary frames of reference that require 
-explanation. Firstly, there is the laboratory frame which 
-describes the coordinate system, we define the origin of our 
-simulation as the focus of our trapping beam. While the choice in 
-direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
-specifically chosen so that it is aligned with the direction of 
-beam propagation. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.6\linewidth]{lab_frame.png}
-	\caption{Ray scattering render of both the laboratory frame and 
-		particle frame for a dimer and a Gaussian beam. The origin 
-		of the laboratory frame is defined as the focus of the 
-		trapping beam, with the positive $z$-axis being aligned with 
-		the direction of beam propagation. The particle frame is 
-		centred on the dimer's centre of mass with the primary axis 
-		$U_z$ being aligned with the centres of the two spheres. Black 
-		line is scaled to represent $1\ \mu m$.}
-	\label{fig:lab_frame} 
-\end{figure}
-
-Secondly there is the particle frame which describes the orientation 
-state of the particle, orientations are reported as quaternions which 
-can be readily transformed into rotation matrices. The matrix describes 
-the direction of the particles 3 orientation axis $U_x$, $U_y$, and 
-$U_z$. The latter is oriented so that it is collinear with the centres
-of the two spheres; the direction of vector $U_z$ is such that it goes 
-from the centre of $a_{II}$ to the centre of $a_{I}$. The origin of the 
-3 axis' is set on the particle's centre of mass, this is also used to 
-define it's displacement from the origin of the laboratory frame. The 
-diffusion tensor of the dimer is more complex than a sphere's as we must consider the rotational effects experienced by the dimer. We use the 
-analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided 
-the solution coefficients depending on the dimer's size ratio 
-($a_I/a_{II}$). Rather than recalculate the coefficients for every possible configuration of dimer we use a spline fitting function to approximate 
-the diffusion tensor. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulated Quadrant Photodiode}
-\label{sec:simulated_QPD}
-
-Beyond merely simulating the forces experienced by a spherical 
-aggregate we also developed a simulative quadrant photo diode 
-to replicate the results from a typical calibration test. This 
-builds upon the work from \cite{Rohrbach2002} which applied 
-Lorenz-Mie theory to replicate the response signal of a QPD 
-being used in back focal-plane interferometry. Rather than be 
-constrained to individual spheres we can now consider the 
-expected response from any type of spherical aggregate. The
-goal of such simulation software would be to simulate the 
-expected experimental results from a complex particle whose 
-exact shape and size may be difficult to determine.
-
-In order to simulate a typical experimental set up with a QPD 
-installed as a position detection system we need to evaluate 
-the total magnitude of the electric field incident on the 
-photo-diode surface. While trapping a micro-particle, the 
-scattered and incident fields combine together and interfere 
-with one another. These fields are collected by a condenser 
-lens in the far field limit and are focused onto the QPD 
-surface, the total intensity can be evaluated as:
-\begin{align}
-I(x,y) = \epsilon_0c\left|
-\begin{bmatrix} 
-	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
-	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
-	E_{i,z}(x,y)+E_{s,z}(x,y)
-\end{bmatrix} \right|^2 \times step(\theta_{NA}-\theta(x,y))
-\end{align}
-
-The last term is simply a representative step term that defines 
-the outer limit by which we evaluate the electric field, this is 
-analogous to our condenser lens removing noise from other light 
-sources by only accepting light at a specific acceptance angle 
-defined by its numerical aperture $NA_c$. Depending on the 
-relative size of our particle we can adjust the acceptance angle, 
-this has very little effect on the transverse signals, but for 
-axial evaluations of a trapped particle the numerical aperture 
-should be tuned so that the resultant response curve has negative 
-slope in order to allow for axial position detection, the method 
-for finding this angle $\theta_\Theta$ is discussed in 
-\cite{Friedrich2012}.
-
-The incident beam is simple enough to define given our set up 
-parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
-beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
-\textit{Ott} uses a point matching approach to approximate the beam shape 
-coefficients of the incident field by fitting it to the far field estimate. 
-From the QPD's perspective it is receiving light both from the incident 
-and scattered beam simultaneously, as such both fields must be expressed
-using outgoing vector spherical harmonics.
-\begin{align}
-	E_{\rm inc}(r)&=
-	E_0\sum^\infty_{n}\sum^n_{m=-n}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
-	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
-	\\
-	E_{\rm scat}(r)&=E_0 \sum^\infty_n \sum^n_{m=-n} \left[
-	p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
-	\right] 
-\end{align}
-
-Where the superscript (2) denotes an outgoing spherical harmonic
-function. In order to compute the scattering from the target particle \textit{ott} uses the $T$-matrix method, this is not essential 
-for a simple sphere but is essential for complex shaped particles 
-such as dimers. The scattered and incident fields are then combined 
-together in the far field to get $I(x,y)$, the quadrant and overall 
-signals are calculated via:
-\begin{align}
-	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
-	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
-	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
-	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
-\end{align}
-
-Where the denominator is the total intensity on the QPD while there is 
-no particle within the trap. You would expect that $S_x$ and $S_y$ are
-near identical for equal particle displacements, however due to the 
-polarisation of the beam there will be a slight bias to the signals 
-for motion along the direction of the polarisation vector. This is 
-generally not an issue if the particle trajectory is sampled for a 
-long enough time period. By converting from signal units to length
-units (see \eqref{eq:conversion_factor}) the trap shape can be 
-discerned from the QPD signals, assuming the fluid properties are
-known to a high degree of accuracy.
-
-While translational motion has no knock on effects to the QPD signal, 
-rotational changes are double counted by both the incident and total 
-field. This results in rotational motion being biased in the QPD signal 
-- even when collecting signals from isotropic scatterers. To prevent 
-this we need to rotate the total field via the inverse rotation matrix
-of the dimer.
-
-To confirm that our method is producing accurate results, we ran a comparison
-between our simulative QPD and the results from \cite{Rohrbach2002}. Where a 
-$300\ nm$ diameter sphere is scanned across the path of a focused Gaussian beam
-($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is
-suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
-set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
-produced the following response curve:
-\begin{figure}[h!]
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
-	\end{subfigure}
-	\begin{subfigure}{0.475 \linewidth}
-		\subcaption{}
-		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
-	\end{subfigure}
-	\caption{Comparison between QPD response signal versus work conducted by 
-		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal 
-		produced by QPD using \textit{ott} and points represent the signal 
-		response collected from \cite{Rohrbach2002}.}
-	\label{fig:Rohrbach}
-\end{figure}
-
-The discrepancy between our simulated QPD and the results 
-from Rohrbach can be attributed to the fact that the position 
-to signal error grows as you move further from trap focus. 
-As the particle moves further from the trap focus the error 
-in the attributed signal grows substantially \cite{Rohrbach2002}.
-In most optical trapping experiments we usually are calibrating 
-a strong trap where the mean square displacement falls well within 
-the linear regime shown in fig.~\ref{fig:Rohrbach}. In which case 
-the error between Rohrbach's model and our own is inconsequential 
-for calibration purposes. With this any trajectory can be collected 
-from the QPD by displacing and rotating the beam accordingly, 
-later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of 
-accuracy that can be achieved using back focal plane interferometry 
-to characterise non-spherical targets.

### 2025-06-27 Latex_files/Thesis/chapters/chapter 3/main.tex
27/06/2025 chapters 1,3,4 and structural changes

--- 2024-10-25 Latex_files/Thesis/chapters/chapter 3/main.tex
@@
-\chapter{Effects of localised shearing on crystal growth and nucleation}
-As outlined in Chapter 1, part of the projects aim is to investigate 
-the possibility of using optical tweezing to induce nucleation by 
-generating fluid flow within a supersaturated solution. The intent 
-of which would be twofold: Firstly to have a repeatable means of 
-inducing nucleation under different solution conditions. And secondly, 
-to understand the influence of shearing on nucleation at a micro 
-level as compared to results in bulk fluid.
+\chapter{Complex Langevin dynamics of spherical dimers}
+\label{chapter:langevin_dynamics}
+\section{Overview}
+Often in calculation of optical forces and scattering the
+choice of target particle is often either a single sphere 
+or a large agglomeration of spheres. In optical tweezing 
+research the target particle is not guaranteed to be a 
+sphere, instead either being an asymmetrical shape, or being 
+a collection of smaller spheres.
 
-It has been shown that for macro-scale systems, the likelihood of 
-nucleation increases to a maximum value under increased shearing 
-\cite{Debuysschere2023, Mura2016}. Mura and Zaccone developed a 
-theoretical framework to describe how the a newly formed nucleus 
-experiences two additional growth factors when placed in a moving 
-fluid. Firstly, due to increased molecular transport of solute 
-molecules the nucleation rate is enhanced in low to moderate fluid 
-flows. But in addition, due to shear flow the crystal surface 
-undergoes deformation which suppresses the nucleation rate 
-undergoing faster fluid flow \cite{Mura2016}. Experimental results 
-with glycine solution support this theory; Debuysschere \textit{et 
-al} demonstrated that the nucleation rate of supersaturated glycine 
-was enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ 
-\cite{Debuysschere2023}. After which the nucleation rate began to 
-decrease but was still greater compared to the case where fluid 
-flow was minimal.
+As an example, while working with dense colloidal suspensions, 
+one often ends up trapping more than one sphere. Li and 
+Arlt \cite{Li2008} studied the case of two microspheres 
+trapped in a single OT and found that multiple trapped 
+beads could be mistaken for a single trapped bead with an  
+altered trap stiffness (see \ref{sec:harmonic_traps}). 
+Theoretical studies on the case of two trapped microspheres 
+by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
+based model to show that the two trapped beads are brought 
+into physical contact with each other by optical forces 
+and they also calculated the axial equilibrium positions 
+of the two trapped beads as a function of their size. 
+Experiments in \cite{Praveen2016} confirmed that the two 
+trapped beads indeed experience different trap stiffnesses 
+in the vicinity of the same potential well.
 
-Optical tweezers can been used to rotate a whole host of microscopic 
-objects, with the fastest reported results exceeding $1000\ Hz$ in 
-heavy water \cite{Arita2016}. If such a micro-rotor was suspended in
-a supersaturated solution the fluid flow around it could be fast enough
-that the nucleation rate is locally enhanced. We focused on two 
-primary candidates for rotation, Vaterite and 
-4-Heptyl-4-biphenylcarbonitrile (7CB). The former being a polymorph of 
-calcium carbonate and the latter an example of nematic liquid crystals, 
-both of which have been used repeatedly in previous micro-rotor research
-\cite{Saito2022, Arita2016, Parkin2009}. In addition, we also consider 
-the application of using techniques beam steering to generate fluid 
-flow by trapping silica micro-beads. In this instance the fluid flow is 
-generated not due to the transfer of angular momentum, but due to 
-shearing caused by a moving sphere through stagnant fluid.
+There are further discussions looking into the dynamics of 
+a whole host of asymmetrically shaped particles (e.g. 
+ellipsoids, cylinders, amorphous solids, disks, etc) 
+\cite{Loudet2014, ShengHua2005, Chetana2022}, their results 
+all showing that predicting the behaviour an arbitrary shaped 
+particle comes with great difficulty due to the fact that the 
+optical force is dependent on a greater number of variables 
+such as orientation and size factors. While these have been 
+noted as important factors, there has not been a concentrated
+effort to discuss their impact over the course of a typical 
+trajectory.
 
-To begin with, the discussion of the necessary optical equipment is 
-covered, drawing attention to specialised components and techniques 
-that are not standard in optical trapping set ups.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Optical Tweezer Equipment}
-In general, all optical tweezers require a laser driver, a focusing 
-microscope objectives, a position controller, and position detector 
-\cite{Gieseler2020}. The laser used for this project was a $1064\ 
-nm$ near infrared laser - provided by CNI Lasers – that was focused 
-by a Nikon 100x oil immersion lens. The choice of an oil immersion 
-lens is important as the optical oil used prevents a loss of focus 
-when used on a glass cover slip. Now, experimental work has shown 
-that the trapping efficiency increases with beam diameter up until 
-it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
-$D_{obj}$ is the diameter of the objective aperture. To expand the 
-beam front we utilise a Galilean beam expansion arrangement 
-(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
-for high power laser applications. In our initial experiments the 
-beam expansion provides a $4\times$ magnification. Whereas in later 
-experiments we utilised a galvano-mirror the beam expansion is 
-$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
-provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
-- the magnification is given by.
-\begin{align}
-	\frac{D_2}{D_1} = \frac{f_2}{f_1}
-\end{align}
+With the initial goal of the project being to induce 
+nucleation events via a rotating sphere, the aim of this 
+chapter is to - in a limited capacity simulate and investigate 
+the influence of a second particle being bound to our target 
+sphere. The choice of a dimer, instead of an amorphous solid 
+that might better represent a growing crystalline solid, 
+allows us to consider how the dynamics of the aggregate change 
+by varying a single parameter, namely the size ratio of the 
+two spheres. Additionally, attempting to simulate an amorphous 
+aggregate is rather difficult as calculating the optical force 
+and torque is computationally slow.
 
-It should be noted that the galvano-mirror requires the use of a 
-Keplerian beam expansion arrangement which reduces the transmitted 
-laser power due to localised heating of the air. Afterwards the 
-laser is passed through a dichroic mirror that separates incoming 
-infrared and visible light, this is to prevent the laser from 
-damaging the CCD camera used for imaging the trapping plane. The 
-laser is then focused to a diffraction limited spot by the objective. 
-Utilizing a high numerical aperture objective enhances the gradient 
-force at the focal point; the trade-off being that the for higher NA 
-objectives the trapping depth is reduced due to spherical aberrations. 
-While it is possible to increase the trapping depth \cite{Reihani2006} 
-by adjusting the objective's tube length this approach is incompatible 
-with our trapping arrangement. A 0.25 NA condenser objective refocuses 
-the scattered laser light and also provide an aperture for an imaging 
-LED to illuminate the focal plane. Samples are loaded onto a piezo 
-driven table to that is inserted between the trapping and condensing 
-objectives; the piezo drivers allow for sub-micron control of the 
-beam focus position to a degree as small as a 10 nm. To detect and 
-monitor the position of a trapped particle a quadrant photo diode 
-(QPD) was utilised. 
+ In this chapter we will consider how the addition of a 
+ second sphere changes the trapping dynamics by introducing 
+ multiple equilibrium positions. In addition we demonstrate 
+ that the angular dependency of the optical force allows for 
+ multiple equilibrium orientations for a single particle. 
+ Furthermore, we look at how dimers interact with circularly 
+ polarised light, noting how our current understanding of 
+ light and particle interactions does not accurately explain 
+ the observed behaviour. We show that a dimer can undergo 
+ gyroscopic precession by utilising this unique interaction 
+ with circularly polarised light.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation Details}
+As a paradigmatic example, consider a dimer suspended in 
+water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
+a Gaussian beam (more specifically a Laguerre-Gaussian 
+beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
+is focused by a objective with numerical aperture of 1.2.
+The polarisation is defined using the Jones vector which 
+describes the magnitude of the x and y components of the 
+electric field (so the vector $[1.0, 0.0i]$ describes an 
+electric field that only oscillates in the x direction). 
+Unless stated otherwise the Jones vector is kept at $[1.0, 
+0.0i]$.
+
+As a paradigmatic example, we study a simple spherical 
+dimer composed of polystyrene, while there is the 
+possibility of considering absorption we choose to keep 
+the refractive index fixed at 1.59. The ratio of the two 
+sphere radii is given by $a_{I}/a_{II}$ where $a_{I}$ is 
+kept at $1\ \mu m$ unless specified otherwise. For 
+simulations we assume that the dimer is suspended in water 
+at room temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).
+
+In order to compute the magnitude of the Brownian forces we
+need to know the dimer's diffusion tensor \textbf{D}. Unlike
+for spheres where the translational and rotational diffusion 
+coefficients are equal along all three Cartesian axis', a
+dimer's translational and rotational diffusion is dependent
+on whether we consider the its long or short axis. We use
+the results from \cite{Nir1973} to compute the dimensionless 
+diffusion tensor and then scale it according to the radius of
+the largest sphere ($a_I$). The dimer's centre of diffusion
+is located a distance of $a_I\zeta$ from the point of contact
+between the two spheres, the coefficient $\zeta$ is based on
+the size ratio of the two spheres. We use the centre of 
+diffusion as the target origin that is provided to \textit{mstm}
+
+Where $U_3$ represents the orientation vector that passes
+through the centres of both sphere's (see \ref{sec:sim} 
+for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
+$U_3$ but have no physical representation for a dimer. We 
+can represent $U_3$ by it's spherical coordinates ($\theta$ 
+and $\phi$), where $\theta$ is the angle between the 
+orientation vector and the direction of propagation and 
+$\phi$ as the angle between the orientation vector and 
+the x-axis. We define the 'standard' orientation as 
+being aligned with the direction of beam propagation 
+direction ($\theta=0^\circ$) - and therefore the 'inverted' 
+orientation is defined when the dimer is orientated against 
+the direction of beam orientation ($\theta=180^\circ$) - see
+fig~\ref{fig:paradigmatic} for a visual of the laboratory and
+particle frame.
+
+Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
+to perform a systematic study of the dynamics demonstrated by 
+asymmetric dimers in both plane and circularly polarised light. 
+The simulations have two primary frames of reference that require 
+explanation. Firstly, there is the laboratory frame which 
+describes the coordinate system, we define the origin of our 
+simulation as the focus of our trapping beam. While the choice in 
+direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
+specifically chosen so that it is aligned with the direction of 
+beam propagation. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
-	\caption{Optical tweezer set up used for the majority of the PhD. 
-		The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, 
-		$60\ mm$, $150\ mm$, \& $200\ mm$ respectively [M = mirrors,
-		DM = dichroic mirrors, f = focal lenses]. Diagram not 
-		drawn to scale.}
-	\label{fig:setup}
+	\includegraphics[width=0.6\linewidth]{lab_frame.png}
+	\caption{Ray scattering render of both the laboratory frame and 
+		particle frame for a dimer and a Gaussian beam. The origin 
+		of the laboratory frame is defined as the focus of the 
+		trapping beam, with the positive $z$-axis being aligned with 
+		the direction of beam propagation. The particle frame is 
+		centred on the dimer's centre of mass with the primary axis 
+		$U_z$ being aligned with the centres of the two spheres. Black 
+		line is scaled to represent $1\ \mu m$.}
+	\label{fig:lab_frame} 
 \end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Position detection methods}
-In order to accurately capture the dynamics of a trapped particle, a 
-position detection system is required. There are 3 possible methods 
-of position detection: video-analysis, lateral-effect position sensing, 
-and photodiodes.
+Secondly there is the particle frame which describes the orientation 
+state of the particle, orientations are reported as quaternions which 
+can be readily transformed into rotation matrices. The dimer's 
+orientation is given by a rotation matrix in the form: 
+\begin{align}
+	\begin{pmatrix}
+		U_{1,x} & U_{2,x} & U_{3,x} \\
+		U_{1,y} & U_{2,y} & U_{3,y} \\
+		U_{1,z} & U_{2,z} & U_{3,z}
+	\end{pmatrix}
+\end{align}
+The matrix describes the direction of the particles 3 orientation 
+axis $U_1$, $U_2$, and $U_3$. The latter is oriented so that it 
+is collinear with the centres of the two spheres; the direction 
+of vector $U_3$ is such that it goes from the centre of $a_{II}$ 
+to the centre of $a_{I}$. The origin of the 3 axis' is set on the 
+particle's centre of mass, this is also used to define it's 
+displacement from the origin of the laboratory frame. The diffusion 
+tensor of the dimer is more complex than a sphere's as we must 
+consider the rotational effects experienced by the dimer. We use 
+the analytical solution of Nir \& Acrivos \cite{Nir1973}, where 
+they provided the solution coefficients depending on the dimer's 
+size ratio ($a_I/a_{II}$). Rather than recalculate the coefficients 
+for every possible configuration of dimer we use a spline fitting 
+function to approximate the diffusion tensor.
 
-Video analysis is ideally suited for multiple traps or situations where 
-precision is not the top priority. Whereas lateral-effect and photodiode 
-position detection are two examples of back-focal plane interferometry, 
-where the interference pattern produced by the target particle is 
-extrapolated to determine its position. In order for video analysis to 
-match the accuracy of back-focal plane interferometry methods requires 
-the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
-achieve while maintaining a decent resolution \cite{Gibson2008}. In 
-comparison off the shelf back-focal plane detectors can achieve temporal 
-resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
+We utilise a python library that utilises \textit{mstm} and
+\textit{ott} to simulate the dynamical behaviour of spherical
+aggregates. By specifying the optical properties of the target 
+\textit{mstm} generates the targets T-matrix which can then be 
+used by \textit{ott} to compute the optical forces generated by 
+a focused beam. Using the Finite Differences approach we can solve
+the Langevin Equation to take discrete time steps and update the 
+dimer's position and orientation. A time step of $10 \mu s$ is 
+used in line with previous works \cite{Vigilante2020}.
 
-A quadrant photo diode (QPD) is a frequently used position detection 
-system for optical tweezers due to their high sampling rate, high 
-degree of precision, and ease of set up. The QPD is constructed of 
-four photo diodes assembled in a quadrant formation, when a particle 
-is trapped the interference pattern produced is focused onto the QPD, 
-with the maximum intensity mapping to the particle's centre of mass. 
-By summing the voltages of the horizontal and vertical quadrants together 
-the particle's centre of mass is tracked in the x-y plane. Axial 
-displacement can be estimated by observing the change in the total 
-voltage of the QPD. The outputted signal gives an indication of the 
-particle's relative displacement from the beam focus, but in order to 
-convert the signal to distance units the trap needs to be calibrated 
-(assuming a linear response curve).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Positional and Orientational dependence of Trapping forces}
+\label{sec:eq_positions}
+First, let us look at how the addition of a second sphere
+changes how we go about looking for equilibrium positions
+relative to the beam focus. Typically when trapping a target
+particle we do not care where it is spacially in relation 
+to the trap focus, we assume that it will end up in the same 
+equilibrium position regardless. This assumption can not 
+be taken when considering a dimer, or any aggregate of spheres.
+
+If we wanted to start from first principles and determine 
+the trap strength on our target particle the first step 
+would be to locate the equilibrium position relative to 
+the trap focus. For a single sphere it is easy to enough 
+to understand that its centre of mass will be drawn to 
+focal point of the laser due to gradient forces, once there 
+the force is analogous to a harmonic spring with a fixed 
+trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
+we consider instead a dimer, we now have two spheres both 
+being drawn to the focus along by the same gradient force; 
+in addition the scattering force is significantly more complex 
+due to both spheres scattering the electromagnetic fields. 
+This mutual scattering between individual spheres is what 
+makes simulating spherical aggregates far more difficult 
+compared to a single sphere, and even harder still to 
+predict the position where the dimer's centre of mass is 
+in equilibrium.
+
+Because the scattering force is only significant in the 
+direction of beam propagation \cite{Capitanio2002} we 
+can assume there will only be a single potential well 
+lying at the centre of the beam. Along the beam axis, 
+such an assumption is no longer valid due to fact that 
+spherical aggregates experience inter-particle scattering. 
+This is a key reason for using \textit{mstm} as it 
+accounts for that behaviour. The methodology for computing 
+optical forces has been covered extensively for a number 
+of different trapping conditions \cite{RanhaNeves2019}. 
+So it is relatively easy to compute the trapping force 
+and determine the equilibrium position by finding the 
+position that minimises the net optical force and the 
+local gradient is negative ($\delta F/\delta x < 0$) - 
+we can assume that for a dielectric sphere the optical 
+torque is negligible. For a dimer (or any arbitrary 
+spherical aggregate), we now must consider both its 
+position and orientation and find where the net optical 
+force and torque are minimised.
+
+After computing its $T$-matrix via \textit{mstm} and 
+supplying that to \textit{ott} we compute the optical 
+force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
+in the axial direction while the dimer is in its 'standard' 
+orientation. As expected we see a single point where 
+the dimer will be in equilibrium, the linear fit in 
+fig.~\ref{fig:paradigmatic}(a) indicates a that the 
+force can be modelled as a harmonic potential close to 
+the equilibrium position ($F_z\approx-\kappa_z z$). 
+The second point where the axial force goes to 0 cannot 
+be considered as equilibrium position as the positive 
+gradient indicates that the trap is unstable unless 
+Brownian motion is ignored.
+
+We repeated the same calculation but now while the 
+dimer is in its 'inverted' orientation, instead of 
+a single point where the optical force is minimised 
+we see that there are instead two separate equilibrium 
+positions, one above the focus and one below the 
+focus. In this particular example the two positions 
+are far enough apart that both can be considered as 
+separate harmonic traps.    
 \begin{figure}[h!]
 	\centering
-	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
-	\caption{Comparison between QPD and Lateral effect photodiodes.
-	The four quadrants of a QPD (left) experience different photocurrents
-	based on the total intensity of light incident on each section 
-	(labelled A, B, C, D). 
-	Whereas a Lateral effect sensor (right) uses the resistive properties
-	of the photodiode surface to vary the create different photocurrents 
-	passing through the anodes A, B, C, and D.}
+	\begin{subfigure}{.65\linewidth}
+		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
+		\caption{}
+		\label{lam=2}
+	\end{subfigure}\hfill % <-- "\hfill"
+	\begin{subfigure}{.25\linewidth}
+		\centering
+		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
+		\label{large over small}
+	\end{subfigure}
+	\medskip
+	\begin{subfigure}{.65\linewidth}
+		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
+		\caption{}
+		\label{lam=2_inverted}
+	\end{subfigure}\hfill % <-- "\hfill"
+	\begin{subfigure}{.25\linewidth}
+		\centering
+		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
+		\label{small_over_large}
+	\end{subfigure}
+	\caption{Plots of force vs displacement of the 
+		centre of mass of the dimer (µm) for the 
+		case of a dimer of size ratio 2. (a) is the 
+		case where the dimer is in its' 'standard' 
+		orientation, where the dimer is trapped at 
+		$z = 1.71\ \mu m$. (b) is the case where the 
+		dimer is in its' 'inverted' orientation, the 
+		dimer is trapped at two positions: $z = 1.20\ 
+		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
+		renders to visualise the dimer orientation 
+		are shown below each plot. The black lines 
+		on each force-curve is a linear fit with the 
+		slope being reported as the trap stiffness in 
+		the legend.}
+	\label{fig:paradigmatic}
 \end{figure}
 
-A lateral-effect sensor has a similar output but works using a the 
-entire sensor as a single cell analogous to the focal plane of the 
-trapping beam. The four corners of the sensor act as anodes connected 
-to a base plate cathode, as the beam moves across the surface of the 
-detector each anode will experience a different photocurrent depending 
-on how close the centre of the interference pattern is to each anode. 
-The advantage of a lateral effect detector is that the linear regime 
-is much larger than a QPD making it much better for monitoring the 
-position of a trapped particle. However, Lateral-effect sensors are 
-often limited in their spacial resolution due to high signal-to-noise 
-ratios, requiring a high intensity of light on the sensor in order to 
-get a clean signal. As a result, most optical force measurements are 
-conducted using a QPD as opposed to a lateral-effect sensor, as often 
-the displacement is small enough that the signal-displacement curve 
-can be considered linear.
+This is different to a single sphere, namely in that a
+single sphere should only have one equilibrium position
+and that it is much closer to the trap focus ($\approx
+0.05\mu\ m$ from the focus). This is also not similar 
+to other works where random clusters of spheres have 
+been shown to oscillate between stable points 
+\cite{Vigilante2020}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Fourier Optics and 4f correlators}
-\label{sec:4f_correlation}
-As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
-of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
-purpose. These are in fact crucial for the operation of the Galvano, the
-pair of them can be more accurately called a 4f correlator.
-
-A 4f correlator is an example of Fourier optics in practice, understanding
-that a focused lens takes a Fourier transform of the light profile. 
-Consider a laser with a circular Gaussian profile, if you were to place 
-a detector there you would pick up the intensity as a function of its 
-position within the beam. If however you focused the light into a single 
-point (using a +ve focal lens) you are actually seeing a measurement of 
-the phase of your laser with position, in which you would see a diffraction
-limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
-collimated. In imaging systems, a series of focal lenses can be used to 
-filter out unwanted scattering from an image (or in an inverse case 
-differentiate between different images), the placement of each lens is 
-shown via fig.~\ref{fig:signal_generator}.
+We can see that both equilibrium positions have 
+comparable axial trap stiffness ($\kappa_z$), 
+however the difference in the transverse trap 
+stiffness ($\kappa_x$) is far more noticeable. 
+The same dimer was trapped at each of the axial 
+equilibrium positions and the transverse force was 
+evaluated. While in all three cases the dimer can be 
+trapped the linear range where that would typically 
+associated with a stable trap is far narrower in the 
+'standard' orientation compared to the 'inverted' cases. 
+This highlights one of the challenges involved with 
+studying asymmetric particles, even though its a 
+simple enough process to trap them they maybe 
+characterised very differently depending on their 
+relative position and orientation towards the focus. 
+This can have a significant impact on rheological 
+studies - or attempting to probe any local property 
+- as the variance in trap strength can result in 
+large errors over repeated measurements. 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.475\linewidth}
-		\includegraphics[width=\linewidth]{signal_generator.jpg}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.475\linewidth}
-		\includegraphics[width=\linewidth, height=0.7\linewidth]
-		{galvano_mirror_controllers.jpg}
-		\caption{}
-	\end{subfigure}
-
-	\caption{Signal generator galvano mirror controller, channel 1 controls the x-axis 
-		mirror, while channel 2 controls the y-axis mirror. Both channels can be manipulated
-		independently.}
-	\label{fig:signal_generator}
-\end{figure}
-
-For our applications a 4f correlator is utilised to ensure that the motion
-of the galvano-mirrors does not move the focal point of the laser, allowing 
-for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
-the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
-installed $150\ mm$ from the second mirror of the galvano, and the latter being 
-installed $200\ mm$ from the back focal plane of the trapping objective. The 
-signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
-dual channel signal control. This allowed us to precisely control the alignment,
-amplitude, phase, and frequency of both mirrors making alignment much easier.
-
-While in theory the frequency can be increased until the mechanical limit of the 
-mirrors is reached ($\approx 1000\ Hz$), the practical upper limit is determined 
-by the trap strength. For a silica sphere this is on the order of $300\ Hz$ when
-suspended in water. Likewise while in theory the signal amplitude can be increased
-until the voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
-of the focal lenses limits the maximum amplitude to $0.5 V$ as any greater will
-move the laser beyond the lens. For basic trapping calibration the galvano-mirrors 
-were set to DC output, providing a fixed spot which operates like a typical optical 
-trap.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Synthesis of Birefringent Micro spheres}
-\label{sec:vaterite}
-There are several options for particles that can be rotated 
-using optical tweezers \cite{Parkin2009, Saito2022}. Over 
-the course of the project two different micro spheres where 
-investigated, Vaterite and liquid crystal droplets. Both can 
-be readily synthesised in the lab and are will rotate at a 
-variety of sizes (see \ref{sec:opt_torque}).
-
-Vaterite is a polymorph of calcium carbonate that is rarely 
-seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
-All three polymorphs are inherently birefringent meaning 
-that they can be rotated using circularly polarised light. 
-However unlike its other polymorphs of calcite and aragonite, 
-when synthesised vaterite will typically form small spherical 
-particles making them ideal for optical trapping and rotation. 
-Synthesis of Vaterite micro spheres requires fine control of 
-the crystal growth process in order to maintain polymorphic 
-stability. Though for the purposes of optical rotation the 
-exact polymorph is not as important as its morphology as all 
-3 polymorphs are inherently birefringent.
-
-Vaterite samples where made by the first preparing equal amounts 
-of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
-the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
-aside for later. Firsta small vial was filled with $1.5mL$ of 
-$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
-$NaCO_3$ respectively, forming a seed solution. Next, a larger 
-vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
-$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
-After 10 minutes of slow but continuous mixing a few drops of 
-Agepon was added to halt the reaction, the solution was filtered 
-and washed 3 times with distilled water before being suspended 
-in water.
-
-When trapped in circularly polarised light, the anisotropic 
-crystal lattice allows spin angular momentum to be transferred 
-to the Vaterite particle, resulting in a rotation about the 
-beam axis. Because the particle scatters light anisotropically 
-the signal detected by the QPD shows periodic fluctuations 
-\cite{Yogesha2012}. In addition due to the change in polarisation 
-there is also a periodic variation that appears at twice the 
-rotational frequency \cite{Monteiro2018}. Therefore, the resulting 
-power spectrum is not a Lorentzian but now also displays peaks 
-that appear at integer multiples of the particles rotational 
-frequency. 
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.4\linewidth}
-		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
-		\subcaption{}
-	\end{subfigure}
-	\begin{subfigure}{0.55\linewidth}
-		\includegraphics[width=\linewidth]{rotating_psd.png}
-		\subcaption{}
-	\end{subfigure}
-	\caption{(a)Sample Vaterite sphere suspended in water and trapped by
-		 circular polarised trap. (b) Collected power spectrum from 
-		 rotating Vaterite, peaks in the power spectrum appear at integer 
-		 multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
-	\label{fig:vaterite}
-\end{figure}
-
-As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
-demonstrates a Lorentzian curve but modified with these periodic peaks, 
-while the Lorentzian can be loosely fitted to the end tail there exists 
-no current model for describing the power spectra. The closest approximation
-to this was conducted by \cite{Yogesha2012} where they describe the 
-rotational motion of ellipsoidal polystyrene particles. The critical 
-assumption being that the particle perfectly rotates in the $x-y$ plane. 
-It has long been suspected that birefringent microspheres experience torques 
-outside of the $x-y$ plane \cite{Volpe2023} making it very difficult to 
-characterise the behaviour of rotating birefringent microspheres without a 
-proper understanding of the full optical torque being applied to it.
-
-\subsection{Liquid Crystal Rotors}
-Liquid crystals are an intriguing example of materials with mixed phase
-properties. Unlike typical solutes such as Glycine, a liquid crystal
-can still maintain some degree of order between its individual molecules
-while in the liquid state. This is due to the fact that liquid crystals
-are constructed of ordered molecules that demonstrate a long range
-ordering. There are three main types of liquid crystal transition methods:
-Thermotropic crystals will transition to their liquid crystal phase when 
-sufficiently heated. Lyotropic materials can undergo this transition due 
-to changes in temperature and concentration. And lastly, Metallotropic 
-materials - which are composed of both organic and inorganic molecules - 
-change phase according to the ratio of organic to inorganic molecules present.
-Liquid crystal rotors are rather simple in their production, 
-4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
-and a small amount was added to a vial of distilled water. The solution 
-was then heated in a water bath to $25^\circ$ in order to transition the
-solid crystal into its liquid crystal state. The solution can then be 
-loaded onto a sample cover slip and the individual droplets visualised.
-The molecules of 7CB will align with a strong electric field, and due to
-the spherical droplet geometry the droplets are inherently birefringent. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.5\linewidth]{LC_sample.png}
-	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
-\end{figure}
-
-The liquid crystal droplets had a much faster rotation rate than comparable 
-Vaterite spheres, due to their higher degree of birefringence and the fact
-that the droplets are far closer to perfect spheres making angular momentum
-transfer more efficient.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rotation of birefringent micro spheres} 
-Optical tweezing has often been used for micro-rheology, by computing the
-exact forces being exerted on the trapped sphere, one can determine the
-local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
-When it comes to optical rotation, a rotating particle will experience a
-fluid drag torque that is proportional to its rotation rate, the maximum 
-rotation rate is when the fluid drag is equal to the optical torque 
-of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
-to measure local temperature changes \cite{Millen2014}, or for understanding 
-how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
-one can use a beam steering arrangement to probe the drag force of the 
-fluid, by understanding the trap strength (calibrating using a low 
-frequency signal) one can measure the drag force experienced by the local 
-fluid \cite{RobertsonAnderson2018}. I
-
-For any discussion of fluid flow its important that we know the fluid 
-regime which is given by the Reynolds number. For a sphere submersed in a 
-moving fluid of velocity $U$ this is given by:
-\begin{align}
-	Re = \frac{\rho UD}{\mu}
-\end{align}
-
-Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
-density and viscosity respectively. In our case we do not have a fluid
-moving around a sphere but a sphere moving through the fluid at some 
-velocity $U$, assuming a no-slip boundary condition we can model the 
-fluid velocity profile based on the velocity of the particle. Assuming
-that for now the fluid properties are unaffected by the rotating particle
-we see that for an individual sphere the Reynolds number can only change 
-due to increasing/decreasing the fluid velocity. The fluid velocity is 
-directly proportional to the particles rotation rate for low Reynolds
-numbers \cite{Bruce2020}. Given the small size of the particles used 
-($1-10\mu\ m$ in diameter), and the relatively low rotation rates that 
-are possible with an optical tweezer, for the case of our analysis the
-Reynolds number will not never be large enough to consider factors such
-as turbulence. There are two possible avenues for generating shear flow 
-with a trapped particle; rotation of birefringent particles, and fluid 
-flow induced by particle motion.
-
-Rotating birefringent particles are the more common method for generating 
-and measuring fluid flow in a solution. To see if we can even achieve the 
-theoretical maximum shear rate, Vaterite spheres were synthesised (see 
-Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm 
-laser at set to 450 mW. The rotation frequency was determined using the 
-QPD, and the particle sizes were computed by image analysis. With the 
-particle size and rotation frequency, the tangential rotation speed is 
-calculated via:
-\begin{align}
-	\label{eq:birefringent_speed}
-	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
-\end{align}
-
-Where $d$ is the particle diameter, $\omega$ is the rotation frequency
-reported by the QPD, and $r$ is the distance from the particle's centre. 
-Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
-outward from the centre of the sphere. The shear rate can then be computed
-as the partial derivative fluid flow (assuming shearing is generated purely
-by the flow field). This approach was used previously with liquid crystal 
-rotors \cite{Saito2022}:
-\begin{align}
-	\label{eq:birefringent_shear}
-	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= 
-	\frac{\pi}{2}\frac{d^3}{r^3}\omega
-\end{align}
-
-One would expect that we set some reference distance where the fluid flow 
-would be considered zero. However, at a microscopic level we are more 
-concerned by the fluid shear for an individual nucleation event. We are
-essentially computing what the shear rate would be for a small volume 
-of fluid between two moving parallel plates. This is similar logic to Mura 
-and Zaccone's work where they considered the shear on a subcritical nucleus
-in a flow field \cite{Mura2016}. They go further in depth to discuss the 
-impact of material transport, overall coming to the conclusion that 
-increasing shear rate leads to an increase the nucleation rate up till some
-maximum. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Estimation of fluid flow around a rotating particle in bulk fluid}
-First we determined the upper rotation rate that could be achieved 
-using both Vaterite and liquid crystal spheres. Vaterite samples 
-were synthesised according to \cite{Parkin2009, Bishop2004} (see 
-sec.~\ref{sec:vaterite}), and then suspended in distilled water. 
-A sample of $200\ \mu L$ was pipetted and a single microsphere was 
-captured via a circular polarised beam.
-
-Due to Van der Waal's forces some of the microspheres were stuck 
-together, fortunately individual sphere's were still present. 
-Individual microspheres were trapped and their rotation rate was 
-determined by looking at the peak frequency component of the 
-collected power spectrum. Due to their anisotropic scattering the 
-QPD signal will oscillate periodically. After taking the Fourier
-transform of the QPD signal the power spectrum has distinctive
-peaks at integer multiples of the rotational frequency. The largest 
-peak is often twice the rotational frequency which is due to the 
-polarization modulation of the scattered field, the first peak 
-is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
-An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.
-
-\begin{figure}[h!]
-	\begin{subfigure}{0.49\linewidth}
-		\includegraphics[width=\linewidth]{time_domain_data.png}
-	\end{subfigure}
-	\begin{subfigure}{0.49\linewidth}
-		\includegraphics[width=\linewidth]{frequency_domain_data.png}
-	\end{subfigure}
-	\caption{Example of QPD signal being converted from the time 
-		domain to the frequency domain. (left) Raw signal collected
-		from the QPD over the 1st second of its trajectory showing 
-		a periodic spacing with a period of $1/f_{rot}$. (right) 
-		Power spectrum of the QPD signal, the y-axis is a linear 
-		scale to demonstrate the clear peaks in the power spectra. 
-		The rotational frequency is the first instance of a peak 
-		forming, and after that the peaks appear in integer 
-		multiples of $f_{rot}$}
-	\label{fig:tdat_fdat}
-\end{figure}
-
-The accuracy of the rotational frequency is dependent on the sample
-rate of the QPD and the sample duration \cite{BergSoerensen2004}. 
-With higher sampling rates resulting in a wider frequency window 
-(see Sec.\ref{sec:psd}) and longer sampling durations resulting in 
-more distinct rotational peaks. The default QPD sampling rate is 
-$2^{17}\ Hz$ which is more than adequate for detecting even the 
-slightest low frequency signals. And the default sampling duration 
-was $3\ s$ which for most cases gave us clear distinct peaks for 
-even relatively low rotational frequencies. After trapping a 
-vaterite microsphere the power spectrum was recorded and the 
-rotation frequency was acquired. This was repeated several times
-for each sphere trapped, to confirm that the rotational frequency
-was constant and that the sphere had reached its maximum rotation 
-rate.
-
-After computing the rotational frequency the radial fluid speed 
-could be estimated using Eq.~\eqref{eq:birefringent_speed}. From 
-there we can estimate the shear flow experienced by a small volume 
-of fluid within the flow field (see Eq.\eqref{eq:birefringent_shear}). 
-Here we only consider the fluid flow as it propagates outward from 
-the axis of rotation. While the cover slip will have some effect 
-fluid flow it is difficult to quantify the effects of a hard 
-boundary on a rotating object when the boundary is perpendicular to 
-the axis of rotation. 
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.7\linewidth}
-		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.7\linewidth}
-		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
-		\caption{}
-	\end{subfigure}
-	\caption{(Top) Fluid flow radiating out from the surface of a
-		rotating Vaterite sphere. (Bottom) Shear rates computed 
-		using Eq.\ref{eq:birefringent_shear}, optimal shear rate is 
-		of $3000 s^{-1}$ is indicated by the dotted line. Vaterite 
-		radii and rotation frequencies are shown, the laser power 
-		was kept constant at 450 mW. Reported rotation rates, and 
-		their corresponding fluid flow and shear rates, for Vaterite 
-		are also plotted alongside lab results. Results from 
-		\cite{Parkin2009, Arita2016} are included as well.}
-	\label{fig:vaterite_shear}
-\end{figure}
-
-From Fig.\ref{fig:vaterite_shear} there is not a strong relationship 
-between particle size and rotation rate, this is contrary to much of 
-the theoretical predictions that predict an exponential decay with 
-particle size. This can be in part due to the fact that synthesising 
-perfectly spherical spheres that have uniform birefringence across the 
-whole population is difficult. Despite our best efforts at controlling 
-the growth rate the smallest particle ever synthesised was around $3\ 
-\mu m$ in diameter. The Vaterite spheres would often stick together 
-while suspended in water after a short period of time. The fastest 
-reported rotation rate within a fluid was by \cite{Arita2016} 
-that achieved a rotation rate of $5\ kHz$, this is plotted in Fig~\ref{fig:vaterite_shear}. In addition we added the results from 
-\cite{Parkin2009} as a more realistic example.
-
-The optimal shear rate predicted by \cite{Debuysschere2023} is 
-plotted on Fig.~\ref{fig:vaterite_shear}(b), the outer dotted 
-lines represent the point when the nucleation rate is less than 
-90\% of the maximum. We can visualise the fluid around the 
-rotating sphere by dividing it up into radial sections. Assuming 
-the sphere is rotating sufficiently quickly there are three 
-sections of concern, as demonstrated by fig.~\ref{fig:shear_diagram}. 
-The fluid close to the sphere experiences a shear rate that 
-is far greater than the optimal shear rate, in which case 
-nucleation is suppressed \cite{Mura2016}. Secondly, there is 
-a region of fluid that experiences a shear rate close to the 
-optimal value ($\pm 5\%$) where the nucleation rate is enhanced 
-significantly compared to the bulk fluid. Lastly, beyond the 
-optimal region the shear rate drops off exponentially in which 
-case the nucleation rate is barely different compared to the 
-bulk fluid.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.875\linewidth]{shear_rate_diagram.pdf}
-	\caption{Diagram demonstrating how fluid shear rate varies 
-	radially from a rotating sphere (rotation rate = $f_{rot}$). 
-	(a) the shear rate exceeds the optimal shear rate 
-	($\dot{\gamma}_{opt}$) and suppresses the nucleation rate. 
-	(b) shear rate is close to the optimal amount, enhancing the 
-	nucleation rate. (c) shear rate is far below the optimal 
-	amount and the nucleation rate is comparable to the bulk fluid.}
-	\label{fig:shear_diagram}
-\end{figure}
-
-\section{Rotating spheres in Supersaturated solution}
-If rotation rates in bulk solution are insufficient then a 
-rotating sphere close to an artificial barrier may be able 
-to improve the shear rate of the surrounding fluid. Of course 
-placing a solid barrier in a supersaturated fluid may well 
-encourage nucleation somewhere on the surface outside of our 
-control. Instead we chose to use the droplet edge of the 
-supersaturated solution, while not a hard barrier per say, 
-the molecular mobility close to the droplet edge is reduced 
-due to surface tension. Furthermore, it has been shown through 
-multiple results that nucleation is enhanced at the air-solution 
-interface \cite{Liao2022,Yuyama2010, Sugiyama2009}.
-
-Estimating the distance from the trap focus to the cover slip
-is difficult to measure directly, often requiring one fits a 
-complex Lorentzian curve to the recorded power spectra in order 
-to measure the fluid drag from being in close proximity to the 
-cover slip \cite{BergSoerensen2004}. We estimate that the 
-vaterite must be at least $15-20\mu\ m$ from the cover slip for 
-us to reliably trap them. The contact angle between the cover 
-slip and solution was previously measured in \cite{Flannigan2023},
-while being loosely dependent on the supersaturation the 
-measured contact angle for supersaturated solutions was between 
-$30^{\circ}$ and $35^{\circ}$. This is shown in fig.\ref{fig:3.9}.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
-	\caption{Diagram of optical trapping set up for rotating 
-		birefringent particles in a supersaturated solution. 
-		Left: side view of the trapping set up showing the 
-		location of the trap focus at the edge of the droplet 
-		of a supersaturated solution. The approximate contact 
-		angle (calculated by \cite{Flannigan2023}) is used to 
-		provide a scale of how close a trapped particle could 
-		be located. Right: top down view of the glycine droplet 
-		with a trapped birefringent particle shown close to 
-		edge of the trap. As the particle rotates the drag 
-		force from the surrounding fluid generates a flow field 
-		around itself (see Eq.~\ref{eq:birefringent_speed}).}
-	\label{fig:3.9}
-\end{figure}
-
-Supersaturated solutions of glycine and water were prepared 
-and stored in an incubator at $40^\circ C$ prior to use. 
-When ready to be studied $15\ \mu L$ of vaterite suspension 
-was pipetted into the solution and $20\ \mu L$ of the combined 
-solution was pipetted onto the cover slip. A Vaterite sphere 
-was located, trapped, and moved to the droplet edge where the 
-solution meets the cover slip. This removed the need for 
-hydrophilic coatings to achieve a flat fluid layer (as used by 
-\cite{Liao2022, Rungsimanon2010}).
-
-Particle sizing was achieved by converting the particle diameter 
-from pixels to physical units within an accuracy of $\pm0.05\mu 
-m$. The QPD was used to measure the frequency of fluctuations 
-in the QPD signal. After measuring the rotational frequency the 
-sphere was left to rotate for a period of ten minutes after which, 
-if no nucleation event was observed the particle was released. 
-The overall results are catalogued in table~\ref{tab:Nucleation}.
-\begin{table}[h!]
-	\centering
-	\caption{Results from rotating Vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
-	\label{tab:Nucleation}
-	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
-		\hline
-		Super Saturation & Particle radius [$\mu m$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
-		\hline
-		\multirow{3}*{1.01} & 2.34 & 10.4 & $\times$ \\
-		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
-		\cline{2-4} & 5.67 & 9.63 & $\times$ \\
-		\hline
-		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
-		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
-		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
-		\hline
-		\multirow{3}*{1.4} & 1.59 & 0.00 & $\times$ \\
-		\cline{2-4} & 3.47 & 0.00 & $\times$ \\
-		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
-		\hline
-		\multirow{3}*{1.45} & 3.68 & 0.00 & $\times$ \\
-		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
-		\cline{2-4} & 6.32 & 0.00 & $\times$ \\
-		\hline
-		\multirow{3}*{1.49} & 1.52 & $0.00$ & $\times$ \\
-		\cline{2-4} & 4.76 & $0.00$ & $\times$ \\
-		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
-		\hline
-	\end{tabular}
-\end{table}
-
-It should be noted that rotational frequencies equal to 
-$0.00\ Hz$ are not due to a rounding error. Simply looking
-at the live video shows no rotation of the vaterite particles,
-and even looking at the QPD signals shows no discernable 
-fluctuations that would arise due to rotation. Trying to trap 
-a particle close to the edge proved more challenging than 
-expected. Unlike in previous reports where the beam is 
-focused at the upper edge of the droplet \cite{Liao2022, 
-Yuyama2010, Sugiyama2022}, we attempted trapping in close 
-to the contact point of the droplet and the cover slip.
-
-It is suspected that trapping is much harder at the 
-interface due to increased surface tension and unpredictable 
-scattering forces. The closest we could trap a microsphere to 
-the droplet edge was in the range of $5-10\mu\ m$, at that 
-distance the fluid flow is so low that even the presence of a 
-hard boundary would be insufficient for shearing the fluid. 
-Furthermore, as is evident in Table~\ref{tab:Nucleation}, 
-the rotation rate drops off significantly with increased 
-supersaturation, due to higher fluid viscosities. While in 
-theory a sufficiently focused laser could rotate any 
-microsphere to a fast enough to reach the shear rate 
-predicted by \cite{Debuysschere2023} the localised intensity 
-would be so large that even using $D_2O$ would see a significant 
-increase in temperature.
-
-It is not impossible that fluid shearing could be used in the 
-future to localise nucleation; but from these results, using 
-individual micro-rotors is not an appropriate method due to
-two key factors. Firstly, the area of influence is far too 
-small to see any noticeable increase in the nucleation rate 
-this is demonstrated most clearly in fig.~\ref{fig:vaterite_shear}. 
-Due to the exponential decay in the expected shear rate it is 
-very difficult trying to shear a large volume of fluid. Even 
-if the optimal shear rate at a micro-level was an order of 
-magnitude less than what was predicted by \cite{Debuysschere2023}, 
-only a small volume of fluid would even experience that shear 
-rate. In order to properly understand the microscopic effects 
-of fluid shear a method of localised shearing over a larger 
-fluid volume is necessary.
-
-Secondly, increased fluid viscosity significantly reduces 
-the limits the maximum rotation rate possible. As indicated 
-by \ref{tab:Nucleation}, the rotation rate achievable by a
-vaterite sphere falls off significantly with increasing 
-supersaturation. Experimental estimations of relative viscosities
-showed a 12\% increase when looking at undersaturated glycine
-solutions ($S\approx0.34$) \cite{Patyar2020}. As far as we
-are aware there are no reported fluid viscosities for higher
-supersaturations. Extrapolating the results from \cite{Patyar2020}
-suggests that a saturated solution would have a 35\% increase 
-in fluid viscosity (compared to pure water). This also does 
-not account for any other factors such as the fact that close
-to a boundary a rotating particle will experience a reaction
-torque from the stationary surface \cite{Bruce2020}.
-
-If multiple micro-rotors could be trapped in close proximity 
-to one another they could create a large region of fluid where 
-nucleation is more likely than the bulk fluid. Because optical
-torque is not contingent on the fluid properties the only 
-limiting factor would be the total angular momentum transferred 
-to each particle. Micro-rotors have been created that allow for 
-precise control of suspended micro-particles \cite{Butaite2019} 
-and could potentially be used to generate sufficient shearing. 
-However these could not be used in this project as we lacked the 
-necessary hardware to form multiple gradient traps.  
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nucleation with a Stationary and Moving Beam}
-As mentioned previously, shearing via optical rotation did 
-not result in any localised nucleation events even while 
-in the proximity of the droplet edge. An alternative approach
-was suggested, using a galvano mirror to move a particle 
-quickly through a supersaturated solution. While a particle
-is within the optical trap the gradient forces are 
-significantly reduced meaning simply trapping a particle is
-insufficient for inducing nucleation \cite{Flannigan2023}. 
-This is why we did not see any nucleation events while 
-rotating the vaterite, there is not a large gradient force
-to draw in material. We did not consider a focus beam by 
-itself as this has already been demonstrated in previous
-experiments \cite{Flannigan2023, Liao2022, Rungsimanon2010}.
-
-Preliminary calculations suggested a particle moving rapidly
-through a fluid could produce significantly greater shear 
-rates (see \emph{Appendix X.X} for breakdown) than a rotating 
-particle. During initial testing using silica beads we found 
-that even in undersaturated solutions the optical trap would 
-produce a crystal nucleus when the trap was empty. This has 
-been reported prior \cite{Rungsimanon2010, Liao2022}, but 
-what is more interesting is how the beam's motion influenced 
-the growth of the nucleus.
-
-\subsection{Stationary beam}
-\label{sec:stationary}
-Its well known that a supersaturated solution can nucleate
-if irradiated by a focused beam. We confirmed that our 
-laser was of sufficient power and focus to induce crystal
-nucleation. The laser was focused at the edge of a $20\mu L$
-droplet consisting of water and dissolved glycine ($S=1.03$). 
-The solution was monitored for 10 minutes, if a nucleation 
-event was observed the event was recorded using a high speed 
-CCD. If no nucleation event was observed after 10 minutes 
-the solution was disposed of and a new sample was prepared. 
-After ten repeats only 30\% of the samples observed a 
-nucleation event at the focus. All of which displayed similar 
-behaviour to one another.
-
-Consider Fig.~\ref{fig:stationary_beam}, the frames taken from 
-a nucleation event, the beam is a stationary being $\approx3.5 
-\mu m$ from the droplet edge. After a period of roughly 5 minutes 
-a nucleus forms at the trap focus, growing quickly (growth rate 
-was approximated using imageJ to be on the order of $700\ \mu 
-m^2/min$) from the focal point of the trap until after roughly 
-$6$ seconds the crystal escapes. Comparing to previous literature
-using optical tweezers shows that the growth rate is only loosely 
-connected to the solutions supersaturation. Crystal growth rates 
-for a solution of glycine and water ($S\approx1.00$) was found to
-vary from as much as $3900 \mu\ m^2/min$ to as low as $675\mu 
-m^2/min$. Local conditions play a much larger role in the growth 
-rate than just the concentration \cite{Flannigan2023}. A likely 
-reason that the trap is escaped is due to the fact that crystal 
-is far too large to be held in place and is in fact still growing 
-as the solution is supersaturated. The key take away to remember 
-is that the beam has no influence over the crystal shape, instead 
-it grows outward from the trap focus. Furthermore, due to the fact 
-that the solution is supersaturated the crystal growth cannot be 
-contained to the trap focus. Instead the crystal escapes as its 
-size exceeds the trap focus. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.75\linewidth, height=1.1\linewidth]{frames_no_beam_movement.pdf}
-	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
-	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
-	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
-	the trap.}
-	\label{fig:stationary_beam}
+	\includegraphics[width=\linewidth]{transverse_force.png}
+	\caption{Plots of force vs displacement of the dimer's 
+		centre of mass spheres, where a positive force 
+		indicates the dimer is directed right on the x-axis, 
+		and vice versa for a negative force. The same simulation 
+		parameters are used here as in fig~\ref{fig:paradigmatic}
+		(a) and (c). The blue curve representing the force 
+		response for a dimer in its standard orientation, orange 
+		being the inverted case, and green the same case but 
+		placed below the focus.}
+	\label{fig:transverse_force}
 \end{figure}
@@
 \newpage
-\subsection{Moving Beam}
-\label{sec:moving}
-To test if a rapidly moving silica bead could generate the necessary shear rate 
-for crystal nucleation we wanted to see if a trapped silica bead could be trapped
-and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
-added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
-unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
-encountered several unexpected results.
-
-Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
-moving in a small elliptical pattern; interestingly. After a few minutes we 
-noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
-for a larger example of these droplets). These appear too have a wide distribution
-of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
-\mu m$. We surmise these could be small clusters of Glycine that had previously 
-been shown to form when the aqueous solutions where irradiated with a 
-focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
-directly entering the focus a nucleus forms close to the droplet edge, 
-unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
-the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
-being moved by and growing around the focal point of the trap. Because of 
-this the crystal nucleus lacks a clear morphology at first. Until roughly 
-$20\ s$ the crystal reaches a almost prismatic structure, with further 
-irradiation increasing the size.
-
-Interestingly the galvano-mirror allows the trap to impart a slight torque 
-on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
-where even though the crystal is not directly in the trap focus it rotates 
-in the $x-y$ plane and gets trapped again at a corner. The rotation could 
-not be due to fluid flow close to the surface of the crystal as the dipole 
-moment of individual water molecules is too small to be influenced by an 
-optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
-growth becomes localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
-\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
-(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
+For completeness the harmonic traps were located for dimers
+across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
+$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
+for each trap. The same simulation parameters are used here 
+as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
+As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
+the dimer begins to approximate a single homogenous sphere 
+- at least in terms of location and trap strength. However, 
+for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
+to $a_{I}/a_{II}=4$), a second equilibrium position is found 
+below the trapping focus. We should note that there are also
+unstable traps that exist, these are also depicted in 
+fig.~\ref{fig:eq_positions}. While the optical force
+is zero at these positions the gradient $\delta F_z/\delta z$
+is positive and so any Brownian motion would displace the 
+dimer, either ejecting it from the optical trap or letting
+it be drawn towards another harmonic trap. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
-	\caption{Frames from a longer video depicting the growth 
-		of a nucleus using a moving beam. Beam power is kept 
-		at 700 mW and the supersaturation was estimated at 
-		S=0.86. Initially the crystal shape is amorphous (a) 
-		but eventually reaches a more regular shape (b). This crystal is still influenced by the optical trap as 
-		even when not directly irradiated by the laser the 
-		crystal rotates between (c) and (d). When the laser 
-		is focused on a corner the crystal growth is localised 
-		to that region, resulting in an elongated section 
-		forming between frames (e) and (f).}
-	\label{fig:eliptical_beam_1}
+	\begin{subfigure}{0.85\linewidth}
+		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
+		\caption{}
+		\label{eq_pos}
+	\end{subfigure}\hfill % <-- "\hfill"
+	\medskip
+	\begin{subfigure}{0.85\linewidth}
+		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
+		\caption{}
+		\label{eq_pos_inverted}
+	\end{subfigure}
+	\caption{Equilibrium positions of optically trapped dimers 
+		with varying size ratio, dashed lines represent unstable 
+		traps whereas solid lines are for stable equilibrium 
+		positions. (a) shows that dimers while in their 
+		'standard' orientation will always have a single 
+		equilibrium position. (b) shows that when the same dimer 
+		is in its' 'inverted' orientation can be trapped in two 
+		axial positions, one below the focus and one above the 
+		focus.}
+	\label{fig:eq_positions}
 \end{figure}
 
-Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
-\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
-moving beam. This modification allows for the crystal growth to be localised to
-a specific region of the bulk crystal whereas with a stationary beam there is no
-control over the crystal morphology. In fact this allows for a much finer control
-over the exact shape of the crystal nucleus, in some cases allowing for growth out
-of the viewing plane as shown in
+Previous work using the ray-optics model have confirmed even 
+in the case that two spheres begin separated the electric field 
+will align the particles as such that they make contact and 
+are trapped together about a single trapping position 
+\cite{Xu2005}. Furthermore it has been shown through proper 
+manipulation of the Gaussian or Bessel beam modes that any 
+number of trapping potentials can be developed 
+\cite{Shahabadi2020} for nanoparticles. This result however, 
+is the first example of an orientation dependent trapping 
+situation using only a $TEM_00$ beam. Typical experimental 
+arrangements cannot determine much information on the axial 
+position of a trapped particle relative to the trap focus; 
+this result indicates not only that dimers can be trapped 
+in multiple axial positions but also their trapping behaviour 
+is heavily dependent on the axial position. As such it is 
+necessary that positional information in the z-axis can be 
+elucidated if multiple spheres are trapped simultaneously.
 
-\begin{figure}
+\newpage
+\section{Non-trivial basins of attraction}
+\label{sec:off-axis}
+As mentioned previously in \ref{sec:opt_torque} the 
+torque applied to an anisotropic scatterer can be 
+broken down into two categories: torque generated 
+from the field polarisation and torque generated 
+by the gradient force of the electric field. The 
+former is generally only applicable for objects 
+that are birefringent whereas the latter is 
+applicable to dimers due to their elongated shape.
+Since both spheres are attracted towards the centre
+of intensity via the gradient force a dimer will 
+experience a restoring torque that keeps the it
+in an equilibrium orientation. For any elongated 
+particle the torque is minimised when the long 
+axis is aligned either parallel or perpendicular
+with the direction of the electric field. In this 
+section we try to find equilibrium configurations 
+(positions and orientations where both the optical
+force and torque are minimised).
+
+Computing the equilibrium positions when a dimer is 
+aligned with the electric field is relatively simple 
+as the orientational torque is minimised (see Eq.
+\ref{eq:opt_torque}). Meaning once trapped the 
+dimer is unlikely to change orientation enough to 
+escape the trap. Regardless, that does not rule out 
+the possibility that there is a stable configuration 
+where the orientation not strictly vertical, in fact 
+most experimental work with symmetric nano-dimers will 
+trap them lying perpendicular to the beam direction 
+\cite{Ahn2018, Reimann2018}. Unlike in Sec.~
+\ref{sec:eq_positions} we cannot simply compute the 
+optical force and torque as the parameter space is too 
+large and determining if a particular position and 
+orientation is stable is not clear based solely on 
+force and torque measurements \cite{Bui2017}. Using 
+the same simulation parameters as before we ran a 
+number short simulations (total simulation time was 
+$0.005$ s) with the laser power increased to $500\ mW$. 
+Each simulation started with the dimer in a different 
+starting position and orientation, due to the high 
+laser power the dimers either escaped the trap or 
+were stably trapped. We chose to only consider the $z-
+\theta$ phase space in order to simplify the explanation,
+considering all possible parameters would increase the 
+simulation time significantly. The $z-\theta$ phase space 
+can be divided into different regions depending on which 
+equilibrium configuration is reached.
+\begin{figure}[h!]
 	\centering
+	\begin{subfigure}{0.67\linewidth}
+		\includegraphics[width=\linewidth]{off_axis_trap.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\raisebox{0pt}[0pt][-10pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
+	\end{subfigure}
+	%\captionsetup{hangindent=2.35cm}
+	\caption{Map of $z-\theta$ phase space using a dimer 
+		of size ratio 2 with a laser power of 500 mW 
+		($\theta=0^\circ$ is the 'standard' orientation 
+		and $\theta=180^\circ$ is the 'inverted' 
+		orientation). The stable configurations are 
+		indicated by the larger circles and the 
+		starting conditions are colour coded to match 
+		the stable point they end up in. Right hand 
+		render shows a dimer in its off-axis configuration.}
+	\label{fig:off_axis}
 \end{figure}
 
-\subsection{Direct trapping of Glycine clusters}
-\label{sec:clusters}
-One common aspect involving these droplets is the fact when brought 
-to the laser focus the droplets would nucleate immediately 
-\cite{Liao2022}. A solution similar to Sec~\ref{sec:stationary} was 
-made up, but without any silica droplets. Once again the beam was 
-focused close to the droplet edge, this time the galvano mirror was 
-scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
-After a few minutes of irradiation droplets were seen entering the 
-camera frame, because no silica had been added these droplets had 
-to be from the glycine solution. Trapping individual droplets did 
-not result in immediate nucleation even after several minutes being 
-trapped. Trying to bring two droplets together resulted in nucleation 
-between the two droplets, compared to \ref{sec:stationary} \& 
-\ref{sec:moving} the growth is much slower, taking nearly 40 seconds 
-before the crystal structure becomes clearer.
+Where the each starting location is colour coded to
+match the final equilibrium position reached, empty
+regions indicate that the trap was escaped. As 
+expected we see that the dimer has three separate 
+traps when the dimer is close to its vertical 
+orientation, similar to what we see from figs.~
+\ref{fig:paradigmatic} and \ref{fig:eq_positions}. 
+When the beginning orientation is close to horizontal 
+there exists a $4^th$ harmonic trap. Each of these 
+traps, has its own 'basin' of attraction around 
+which the dimer is drawn into, the strange shape 
+of these basins is likely due to how the asymmetric 
+shape interacts with the optical trap.
+
+Interestingly while the trap strength of these off-axis 
+configurations are similar in magnitude to the vertically 
+aligned dimers, but when the laser power is lowered 
+(around $5\ mW$) the traps appear metastable. With the 
+vertical configurations even after $30\ s$ of simulation
+time the dimer is still trapped whereas the off-axis 
+configuration can be escaped after only a few seconds of
+simulation time. The suggested that the configuration is 
+metastable is due to the increased rotational freedom. 
+Similar configurations have been explored with ellipsoids; 
+Zhu \textit{et al.} looked at the dynamics of various 
+elliptical particles and found that regardless of shape or 
+initial orientation the particle would tend towards either 
+a purely vertical or purely horizontal orientation 
+\cite{Zhu2021}. However, in our case the dimer never returns
+to a vertical orientation, suggesting their is a potential 
+barrier separating the different configurations. In order
+to build a better understanding of the potential landscape 
+we need to consider how the force and torque vary for a 
+dimer in the $z-\theta$ phase space.
+
+\subsection{Force-Torque landscape of an arbitrary dimer}
+We use \textit{ott} to compute the optical force 
+and torque for the same dimer as used in fig.~
+\ref{fig:off_axis} in different positions or 
+orientations, this is depicted in fig.\ref{fig:force_torque}.
+
+The optical torques reported by \textit{ott} are
+about their respective axis - so $T_x$ is the torque
+that rotates the dimer about the x-axis in the particle
+frame.
+
+The magnitude of the force and torque are indicated
+by the colour intensity, with white indicating either
+zero force or torque. This helps show where a
+dimer may end up in equilibrium by finding 
+where both the torque and force go to 0.
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.8\linewidth]{cluster_trapping.pdf}
-	\caption{Frames from a longer video demonstrating the trapping of a glycine 
-		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
-		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
-		black) being brought into contact with a larger droplet. (b) upon contact 
-		a nucleus can be seen between the two droplets. The growth is rather slow 
-		with the crystal having no clear defined morphology through (c) and (d). 
-		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
-	\label{fig:cluster_trapping}
+	\includegraphics[width=\linewidth]{force_torque.png}
+	\caption{Force-Torque landscape over the $z-\theta$ 
+		phase space with a dimer is a size ratio of 2, 
+		laser power of 100mW. Left: 2D plot of the 
+		optical force in the axial direction as a 
+		function of position and orientation. Right: 
+		2D plot of the optical torque about the x-axis 
+		as a function of position and orientation.}
+	\label{fig:force_torque}
 \end{figure}
 
-The fact that these droplets can be trapped indicates they must 
-have a higher refractive index than the surrounding solution. 
-Previous reports have shown that a focused trap will draw in 
-solute material from the bulk \cite{Tsuboi2009, Gowayed2021}. It 
-has been shown that the concentration of glycine solution is correlated 
-with the refractive indices of the liquid \cite{Gowayed2021, Orttung1963}. 
-This suggests that these droplets serve to provide solute material 
-to the bulk crystal.
+The force and torque landscape provides some insight 
+into how the dimer can inhabit an off axis equilibrium
+orientation. Firstly, lets consider a typical trap, 
+while the dimer is close to vertical; in all of these 
+cases the optical force and torque directs the dimer 
+back to the configuration. In this case the potential 
+well is positive whether the dimer moves along the 
+z-axis or is rotated. We can assume that in both cases
+the translational and rotational stiffnesses are close
+enough to be modelled as harmonic potentials.
 
-\newpage
+However when in an off-axis configuration ($z\approx0$ 
+\& $90^\circ\le\theta\le115^\circ$) the optical torque 
+is working against the optical force, trying to move it 
+out of equilibrium while the optical force returns it to 
+equilibrium. This means that the potential well is more 
+akin to a saddle point, the only reason that the optical
+trap is not escaped immediately is because the the 
+translational trap stiffness is significantly greater than
+the gradient of the optical torque ($\kappa_z\gg\kappa_
+\theta$). Therefore it is more accurate to say that the 
+trap is not metastable, as when the trap is escaped it 
+is not because the Brownian displacement is greater than
+the potential well, rather the dimer is randomly 
+traversing along the x-axis (aka random Brownian 
+rotation) and has escaped the translational potential 
+well.
 
-\subsection{Influence of a moving beam front on seed crystals}
-One potential application of this phenomena would be in using a 
-moving the moving beam front as a method for shaping the final
-crystal morphology. Glycine seed crystals were grown via 
-evaporative crystallisation over night. The resultant crystals
-were as wide as $0.5 mm$ in some cases. Individual crystals were 
-collected and suspended in a water + glycine solution ($S = 1.001$)
-and the laser was scanned along a narrow linear path.
+The force-torque landscape also provides an explanation
+to the boundaries between the different trapping 'basins'
+in fig.~\ref{fig:off_axis}. The first point of interest 
+is the 'dead zone' around the orange basin. Looking at 
+fig.~\ref{fig:force_torque} we can see that around that
+region ($z \approx 2.5\mu$ \& $\theta\approx 110^\circ$)
+the optical torque changes sign as you move closer to 
+the trap focus. There is no clear physical reason why
+the optical torque change direction so abruptly but a 
+consequence of it is that dimers that are too close to 
+this region of the torque landscape are ejected from 
+the optical trap as the are drawn down towards the trap
+focus.
 
-As shown by fig.~\ref{fig:seed_crystals} the laser does not in fact 
-promote crystal growth but instead forces the crystal to dissolve 
-into the bulk solution. The dissolution rate is not as substantial 
-as the growth rate seen in previous sections. It takes over 20 
-minutes for the for the crystal surface to dissolve more than a 
-few microns. Furthermore we do not report any sightings of droplets
-close to the surface of the crystal nor do we see anything enter the 
-optical trap.
+The only pellicular result is the observation that the 
+blue basin can still be reached well below the trap 
+focus. As shown by fig.~\ref{fig:off_axis} there are a 
+handful of starting configurations that end up in the 
+upper trap while in the inverted orientation. Even 
+consulting fig.~\ref{fig:force_torque} does not indicate
+any reason why this occurs. The only conclusion we can 
+draw is that the orange and green basins of attraction 
+are small enough that it is possible for a dimer to 
+avoid both. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Continuous rotational motion in circularly polarised light}
+One aspect that has yet to be covered in depth with regards 
+to spherical aggregates of any construction is their interaction 
+with circularly polarised light. For homogenous spheres the 
+optical torque is regarded as being negligible as the spin 
+density cannot impart angular momentum while propagating in a 
+homogenous medium. Dimers however, have been shown to experience 
+an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
+while trapped in a circular polarised beam. In our simulations 
+we found that dimers would rotate about their long axis when 
+trapped in circularly polarised light. In this section we want 
+to discuss how this behaviour is influenced by size, position, 
+and orientation; and furthermore, we wish to address possible 
+explanations for this behaviour, as none of the current
+theories into optically induced rotation seem plausible.
+
+\subsection{Polarisation Dependency on Dimer trajectory}
+\label{sec:rot_pol}
+In their paper Vigilante \textit{et al} attribute the 
+rotational motion to the anisotropic shape of the dimer
+\cite{Vigilante2020}. Anisotropic scattering is a viable 
+theory for describing optical rotation, however there is 
+usually an optical axis about which rotation occurs. For 
+a dimer this is typically the orientation vector that 
+passes through the centres of both spheres \cite{Ahn2018, 
+Reimann2018, Bruce2020}. There have been cases where the 
+particle's cross sectional shape is engineered to scatter
+light in one particular direction \cite{Higurashi1994}, 
+but if this was the case then the dimer should rotate 
+when illuminated by any polarisation of light.
+
+To that end, we simulated the motion of an optically 
+trapped dimer in beams of varying polarisation ($NA=1.2$, 
+$P=100\ mW$). As mentioned previously the beam polarisation
+is given using the Jones vector, for each simulation we vary
+the phase in increments of $\pi/24$.
+
+\begin{figure}[h!]
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{rotating_traj.png}
+	\end{subfigure}
+	$\downarrow$
+	\begin{subfigure}{\linewidth}
+		\includegraphics[width=\linewidth]{rotating_traj_fourier.png}
+	\end{subfigure}
+	\caption{Top: 3 second trajectory of a dimer trapped by a 
+		circularly polarised Gaussian beam. The far left 
+		column depicts the dimer's centre of mass position 
+		with time; the remaining 3 columns show the 9 components 
+		of the dimers' rotation matrix, with each column being
+		associated with one of its three principal axis. Bottom: 
+		The same trajectory but each time series has been 
+		replaced with its Fourier transform. The only non-zero 
+		elements are for the $U_{1,x}$, $U_{1,y}$, $U_{2,x}$, 
+		and $U_{2,y}$ which show a single peak at the rotation frequency.} 
+	\label{fig:fourier_transform}
+\end{figure}
+
+Each simulation was run for $1$ second ($\Delta t =10^{-5}\ 
+s$) and at the end we looked at the orientational time series; 
+the dimer's orientation is recorded as a quaternion which can 
+be easily converted to a 3-dimensional rotation matrix. 
+By considering only the transverse components ($U_{1,x}$, 
+$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
+matrix and taking the Fourier transformation of their 
+time series reveals the rotational frequency. The 
+laser power is set to $100\ mW$ to avoid large thermal 
+fluctuations and so that the Fourier series of the 
+transverse components approximates $\delta(\omega_
+{rot}-f)$ - the Dirac delta function centred at the 
+rotational frequency $\omega_{rot}$. This is 
+demonstrated in fig.~\ref{fig:fourier_transform}
+
+If the rotational frequency was not immediately obvious the 
+simulation was repeated but over a longer simulation time. 
+Four different size ratio of dimers were studied, both in 
+their 'standard' and 'inverted' orientations. The results 
+of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{seed_crystals.pdf}
-	\label{fig:seed_crystals}
-	\caption{Frames from a seed crystal being irradiated by a scanning
-		laser. The laser path is shown below in red. Initially the 
-		crystal edge is nearly perpendicular to the laser path. However,
-		over the course of nearly twenty minutes the crystal around the 
-		laser has dissolved leaving a concave indent.}
+	\includegraphics[width=0.9\linewidth]{rotation_rate_vs_pol.png}
+	\caption{Rotation frequency vs electric field phase difference 
+		for differently sized dimers. The solid lines represent 
+		the rotation rate experienced while the dimer is in its 
+		standard orientation, whereas the solid points are for 
+		the case where the orientation is inverted. Laser power = 
+		$100\ mW$.}
+	\label{fig:rotation_vs_pol}
 \end{figure}
 
-This has some implications: Firstly, the lack of supposed glycine 
-droplets suggests that the presence of an interface is crucial in 
-order to promote the formation of these droplets. This is consistent
-with previous reports on laser nucleation, where a air-solution or 
-solution-glass interface is necessary for the trap focus to have any 
-noticeable effect on the solute \cite{Gowayed2021, Liao2022, 
-Rungsimanon2010}. Secondly, the fact that the seed crystal dissolved 
-in a saturated solution suggests that this could be due to a localised
-heating effect. A commonly used heating model for optical tweezers is 
-the Peterman model \cite{Peterman2003} which accounts for the lateral 
-distance between the optical trap and a local heat sink (such as the 
-glass cover slip or microscope objective). Even with the heat sink 
-the predicted temperature rise in pure water would be around $5-10 K$.  
-The fact that the seed crystal dissolves so slowly could simply be due 
-to the fact that the local heating is lessened by the moving beam front.
+This shows us that these optical rotations are polarisation 
+dependent and not merely an example of the dimer scattering 
+light asymmetrically. We also know that this cannot be a 
+similar mechanism to that of experimental works \cite{Ahn2018, 
+Reimann2018} as in our simulations the dimer's long axis
+is parallel to the electric field. In this orientation 
+the dimer's optical axis is aligned perpendicular to the
+polarisation vector meaning that there should be no angular
+momentum transferred along the optical axis. This is consistent 
+with other experiments involving elongated particles. A dual
+beam trap was used to study the dynamics of "disk like" particles. 
+They found that these particles had one of two stable
+orientations: Either orienting with the 'flat' side perpendicular
+to the beam trap, with no rotation being observed \cite{Brzobohaty2015}. 
+Or orientating with the 'flat' side parallel to the trap, in which 
+case the long axis was aligned with the polarisation vector and 
+thus rotational motion was detected \cite{Brzobohaty2015}.
 
-\section{Summary of Moving Beam Phenomena}
-To summarise, the introduction of a moving beam helps to accelerate the 
-local growth of a newly formed crystal. This is due to the presence of 
-glycine droplets that accumulate near the interface between the liquid 
-solution and air. The theory behind the localised growth can be summarised 
-thusly.
+\subsection{Brownian Vortex via Curl of Spin momentum}
+In their work Vigilante refers to the spin-curl effects 
+demonstrated by Grier \textit{et al} \cite{Ruffner2012}, 
+in which the curl of the spin density leads to a second 
+order optical force that orbits around the beams central 
+axis \cite{Yevick2017}. While several papers have 
+demonstrated this phenomena \cite{Zhao2007, Zhao2009, 
+Wang2010} it was only properly formalised by \cite{Ruffner2012}. 
+In which they showed that the seemingly random trajectory 
+of a trapped sphere was biased by the polarisation state 
+of the trapping beam. While not immediately evident from 
+the trajectory the helicity of the trapping beam was revealed 
+by computing the particle's probability flux using.
+\begin{align}
+	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
+	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
+	\label{eq:prob_flux}
+\end{align}
 
-Initial nucleation is similar to typical optical trapping induced nucleation,
-with the air solution interface limiting the molecular mobility of the solute
-molecules \cite{Liao2022, Sugiyama2009, Gowayed2021}. The moving beam front 
-can influence the motion of the nucleus initially, but eventually the drag 
-force means the crystal is not moved by the optical trap. Localised crystal
-growth occurs when the trap is close to or partially over the interface of 
-the crystal (see figure~\ref{fig:local_nucleation}(a)). As shown in \ref{sec:clusters}, 
-the optical trap can manipulate these droplets similar 
-to microspheres. When in close proximity to the trap these droplets are 
-brought towards the crystal surface (see figure~\ref{fig:local_nucleation}(b)). 
-These provide material that grows the crystal around that region (see figure~\ref{fig:local_nucleation}(c)). Eventually the local solution is either 
-depleted of solute material or the crystal front has grown to fully encompass 
-the trap, preventing further growth (see figure~\ref{fig:local_nucleation}(d)). 
+where $\delta_{\sigma_j}$ is the kernel of an adaptive 
+density estimator \cite{Silverman1986}. \eqref{eq:prob_flux} 
+describes the direction a trapped sphere is most likely 
+move in given our statistical knowledge of the 
+trajectories probability density function. A finite 
+estimation of the density function $p(r)$ is used in 
+\cite{Ruffner2012}.
+\begin{align}
+	p(r) = \frac{1}{N}\sum_{j=1}^N \delta_{\sigma_j}(r-r_j)
+	\label{eq:prob_density}
+\end{align}
+
+The probability flux reveals a biased motion in the trajectory 
+of a single sphere (see Fig.~\ref{fig:Ruffner_Grier}). This 
+biased motion results in a slight orbital motion about the 
+central axis of the trapping beam, the orbital frequency is 
+shown to be proportional to the polarisation state of the 
+trapping beam. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.8\linewidth]{galvano_diagram.pdf}
-	\caption{Diagram outlining how a moving beam assists in the growth of a crystal
-		nucleus. (a) a crystal nucleus is partially trapped by a moving beam with 
-		solute droplets close to its surface. (b) droplets close to the laser
-		focus are drawn in by gradient forces and moved towards the crystal surface.
-		(c) these droplets provide material to the main crystal, resulting in
-		localised growth around the laser focus. (d) eventually the crystal area 
-		either fully surrounds the laser focus or the solution surrounding the laser
-		is depleted of solute material.}
-	\label{fig:local_nucleation}
+	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
+	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how 
+		the momentum density of a Gaussian beam is twisted while 
+		using circularly polarised light. The top row (figures 
+		(b)-(d)) shows a 7 sphere cluster trapped in a circularly 
+		polarised beam. Due to the clusters asymmetric susceptibility 
+		to polarization the cluster rotates in the $x-y$ plane. 
+		Whereas the bottom row (figures (e) - (g)) show the similar 
+		results for a single sphere. In this instance the sphere 
+		does not rotate but instead orbits the beam axis. In both 
+		instances the motion is proportional to the degree of 
+		polarisation (see figures (d) and (g)) but for the single 
+		sphere this motion is only revealed when using 
+		\eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused 
+		with permission from author}
+	\label{fig:Ruffner_Grier}
 \end{figure}
 
-The reason the seed crystals saw no further crystal growth when irradiated 
-with the trap is due to the fact that these experiments were carried out 
-in a bulk solution, absent of any interfaces. As such the clusters seen in
-\ref{sec:moving} \& \ref{sec:clusters} are not present and cannot provide
-material to accelerate crystal growth.
+While the results from \cite{Ruffner2012} suggest that the 
+optical rotation seen in asymmetric dimers can be attributed to 
+the same spin-curl forces there are several questions that 
+cannot be explained purely by the spin-curl force.
 
-There are still several factors that need to be investigated, but due to time
-constraints it was not possible to properly study this phenomena. Firstly, 
-there is the question of what conditions result in the production of concentrated
-droplets, it is not clear if the presence of a laser is required or if these
-droplets naturally occurring. Prior literature would suggest that the laser is 
-required \cite{Liao2022, Tsuboi2009}, but this would not explain why in many 
-cases the droplets are found far outside the influence of the optical trap. It
-has been suggested that optical traps can attract microspheres over a wider 
-area, but it is hard to say that this would result in the creation of liquid 
-droplets \cite{Yi2021}.
+\subsection{Optical torque differences}
+In their paper Grier \textit{et al} show that a 7 sphere cluster
+will rotate when trapped in circularly polarised light but a 
+homogenous sphere will have a slight curl to its trajectory. 
+They attribute this to the curl of the spin angular momentum
+generating a "Brownian vortex". But what is more interesting, 
+and something that appears unaddressed, is how the effects of 
+this vortex change based on the overall shape of the target
+particle.
 
-Secondly, there is the question of how these droplets supply material to the 
-bulk crystal. In some instances it is clear that the droplets are being drawn 
-into the trap, however, in other instances while there are no droplets close to
-the vicinity of the optical trap the crystal continues to grow. If these droplets
-are a necessary precursor to induce crystal nucleation then understanding how
-they provide materials to the bulk crystal may help with our understanding of 
-the kinetics of multi-step nucleation.
+You would not expect that the addition of a single additional 
+sphere should drastically adjust the torque especially if said 
+sphere is relatively small. However when we measured the optical 
+torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we 
+found the exact opposite. In both cases we used the same trapping 
+beam as used for fig.~\ref{fig:rotation_vs_pol} but with a 
+circularly polarised beam. Both the sphere and dimer were rotated 
+in the $x-z$ plane and the all three components of the optical 
+torque were recorded.
+\begin{figure}[h!]
+	\centering	
+	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
+	\caption{Optical toque experienced by a dimer ($a_{I}/a_
+	{II}=10$) and a single isotropic sphere. Both were rotated 
+	in the $x-z$ plane and the angle between $U_z$ and the beam 
+	axis gives the polar angle.The solid lines denote the torque 
+	experienced by the dimer whereas the dashed lines represent 
+	the torque experienced by the sphere.}
+	\label{fig:sphere_dimer_torque}
+\end{figure}
 
-\section{Conclusion}
+The torques about the $x$ and $y$ axis can be somewhat understood 
+as the second sphere is being drawn back towards the centre of the 
+trap by the gradient forces. The same cannot be said for the $z$-
+component of the optical torque, which is non-zero even at a 
+$80^\circ$ angle. This cannot be simply explained via spin-curl 
+effects, as the 'Brownian vortex' should instead be driving the
+dimer around the beam axis - similar to how the 7 sphere cluster 
+behaved. What is clear that the mere combination of two spheres
+results in a unique behaviour that has not been previously 
+investigated.
 
-In conclusion, the experimental results from this chapter provide no 
-affirmative evidence of shear induced crystal nucleation. This is in
-part due to the limited rotational speed that could be achieved using
-vaterite microspheres. Simply adding a micro-rotor to a supersaturated
-solution will not yield a significant fluid flow to induce crystal 
-nucleation. Furthermore, the region in which the local shear rate is 
-significant enough is so small that any nucleation events that could 
-occur there would not grow large enough to be stable. This does not 
-however, suggest that shear induced nucleation is impossible at a 
-micro scale. In fact there have been several developments lately that
-allow for the precise rotation of multiple rotors to control the 
-precise motion of silica beads. The challenge lies in both building the
-optical equipment to form multiple strong traps (~ 3 per rotor). These 
-traps would have to be strong enough to not only hold the rotors but 
-also move them at a sufficiently fast rotational speed that they could
-induce a significant enough fluid flow.
+\subsection{Rotational frequency as a function of size ratio and 
+	orientation}
+\label{sec:rot_size}	
+One factor t Intuitively, you would expect that a 
+larger particle would experience a greater torque and therefore 
+rotate faster. By repeating the same kinds of simulation as used 
+in \ref{fig:rotation_vs_pol} but for a circularly polarised beam 
+$\phi=90\circ$ it was found that not only is the rotation rate 
+dependent on the size of the dimer, but also on its orientation 
+and therefore their axial position.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
+  \caption{Rotation rate plotted against dimer size ratio while 
+  	trapped in a circularly polarised beam; a positive rotation 
+  	rate indicates clockwise rotation, whereas a negative 
+  	rotation rate indicates counter-clockwise rotation. The red 
+  	line is for the case of a dimer in its 'standard' orientation. 
+  	The blue line is for the case when the dimer is in its 
+  	'inverted' orientation while trapped above the focus of the 
+  	beam. And lastly the green line is for the case when the 
+  	dimer is in its 'inverted' orientation, but when it is trapped 
+  	below the focus of the beam.}
+\end{figure}
 
-On the other hand, the results utilising the galvano mirror have some 
-interesting implications for future research into crystal nucleation
-using optical tweezers. We noted that in undersaturated conditions the
-presence of 'droplets' that could be trapped and later nucleated when 
-brought into contact with one another. The fact that these droplets 
-could be trapped suggests they have a higher refractive index to the 
-solution, given that the solutions were filtered prior to irradiation
-dismisses that they could be dust particles. Prior literature suggests
-that optical trapping of a supersaturated solution can lead to the 
-production of nano droplets \cite{Gowayed2021, Tsuboi2009}. Controlling
-the localised growth requires that the growth and trapping are handled
-by two separate lasers. Optical traps for amorphous solids has been 
-shown as a possibility but the algorithm would need to be adaptive 
-to the changing crystal shape.
+It is difficult to see from the graph, but the rotation rate never
+truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which 
+would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
+to induce rotational motion. It is not clear if the optical torque 
+calculated for the larger size ratios is accurate as you would not 
+expect the rotation rate to plateau like this. From fig.~
+\ref{fig:sphere_dimer_torque} it seems clear that the torque is not
+merely an error of the numerical calculations as it would not have 
+such a strong angular and positional dependence. Most likely with 
+longer simulation times would reveal the real relationship behind
+the rotation rate and the size ratio.
+
+What is also interesting is that the rotation rate is not correlated 
+directly with either the particle size or the equilibrium position. 
+This is in stark contrast to previous reports of dimer optical 
+rotation; Ahn \textit{et al.} reported that the rotational frequency 
+is maximised when the dimer is symmetric \cite{Ahn2018}. Their work 
+was conducted experimentally in a vacuum so the only limiting factor 
+to the optical rotational motion is the structural stability of the 
+particle itself.
+
+\subsection{Gyroscopic Precession using asymmetric dimers}
+As mentioned in section~\ref{sec:off-axis} for specificity sized 
+dimers there is the potential for non-vertical trapping orientations 
+in which the dimer is still stably trapped. When a circularly 
+polarised beam is used the dimer exhibits gyroscopic precession. 
+As shown in fig.~\ref{fig:gyro} the dimer's trajectory exhibits 
+periodic rotation, both around its long axis and about the beam 
+axis.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
+	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped 
+		in an off axis orientation with a circularly polarised beam 
+		($P= 100\ mW$). The far left column depicts the dimer's centre 
+		of mass position with time; the remaining 3 columns show the 9
+		components of the dimers' rotation matrix, with each column being
+		associated with one of its three principal axis.}
+	\label{fig:gyro}
+\end{figure}
+
+Applying a Fourier analysis to the above trajectory reveals 
+the 3 fundamental frequencies typically associated with 
+gyroscopic precession. Fig.~\ref{fig:gyro_diagram} gives 
+a representative idea of the motion seen in fig.~\ref{fig:gyro}.
+Firstly there is precession, is denoted by $\psi$, this
+causes the dimer to rotate in the $XY$ plane while maintaining
+a constant polar angle. The precession can be due to the dimer 
+having inhomogeneous polarisation susceptibility. Since its 
+long axis is more susceptible to being polarised the dimer will
+try to align its long axis with the polarisation vector 
+\cite{Bruce2020}. Second there is nutation, denoted by $\theta$, 
+which is due to the dimer's Brownian motion being influenced by 
+the 'Brownian vortex', as predicted by \cite{Ruffner2012}. And 
+lastly there is rotation, denoted by $\omega$, where the dimer
+spins around its long axis as it did in \ref{sec:rot_pol} \&
+\ref{sec:rot_size}. If the rotation was due to the curl of the 
+spin angular momentum then you would only expect to see 
+precession and nutation, the addition of rotational motion 
+suggests that having multiple particles in close proximity 
+results in a transfer of angular momentum along the long 
+axis.
+
+What is interesting is that while we see the precession and 
+nutation also contribute to the rotation frequency. As can be 
+seen in both fig.~\ref{fig:gyro} and fig.~\ref{fig:gyro_diagram}
+the rotation matrix components $U_{x,x}$, $U_{x,y}$, $U_{y,x}$
+and $U_{y,y}$
+\begin{figure}[h!]
+	\hspace{-2.3cm}
+	\includegraphics[width=1.2\linewidth]{gyroscopic_diagram.pdf}
+	\caption{Representative diagram of the gyroscopic precession 
+	from fig.~\ref{fig:gyro}. The dimer has three principal angular
+	frequencies: The rotation ($\omega$) occurs around the 
+	the dimer's long axis. Precession ($\psi$) is seen where the dimer 
+	rotates around the beam axis. Nutation ($\theta$) is due to the dimer's 
+	centre of diffusion orbiting the beam axis. Shown on the right
+	is the power spectra from fig.~\ref{fig:gyro} with the associated 
+	frequencies labelled. The power spectrum have been zoomed in on
+	the relevant frequencies to highlight the precise values.}
+	\label{fig:gyro_diagram}
+\end{figure}
+
+This gyroscopic motion has been demonstrated previously in 
+nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
+but has not been observed for micron scale aggregates. What
+is interesting that the precession is seen around the particle's
+optical axis (in the case of a dimer this would be its long 
+axis). Fig.~\ref{fig:gyro} instead shows the precession 
+occurring perpendicular to the optical axis. In the former 
+case the precession was explained as simply occurring due 
+to the reaction torque from the surrounding fluid \cite{Zhu2021}.
+In their analysis of ellipsoidal particles, Zhu \textit{et al.}
+found that when the particle rotated around their minimal inertial 
+axis (similar to the observed behaviour in dimers), the particle 
+would quickly 'stabilise' by aligning its' optical axis with the 
+beam axis. This highlights that a spherical dimer is capable of 
+rotating about its minimal inertial axis while remaining stably
+trapped. This has potential implications beyond simply 
+understanding particle dynamics.
+
+Further analysis of the mechanism behind the precession of off-axis 
+dimers may provide insights into controlling Brownian motion. An 
+experimental work trying to 'cool' nano-dimers by controlling the 
+motion in all 6 degrees of freedom found that even while the 
+rotation about the short axis' could be controlled the free rotation 
+around the dimers' long axis resulted in an unpredictable torsional 
+vibration \cite{Bang2020}. Understanding how rotational motion 
+arises in the Mie-regime could allow researchers to build a robust 
+theoretical framework to construct beam structures that eliminate any 
+unwanted rotational motion from a target particle. Conversely, the 
+same framework could allow for precise measurements of the optical 
+torque applied to a target particle, allowing for characterisation 
+of complex shaped particles' interactions with an optical trap.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusions}
+Considering the simplicity of a scatterer such as a dimer, 
+one would assume that the dynamics of such an object would 
+be relatively easy to predict. Simulations of dimers in 
+the Mie regime show that not only do they have multiple 
+positions and orientations in which they can be trapped but 
+also that their interaction with circularly polarised light 
+is heavily dependent on the axial position and trapping 
+orientation.
+
+Dimer's have the potential to be used as tunable micro-rotors, 
+being simple to synthesise and can be made out of any material 
+of choice. The rotation demonstrated by dimer's is not accurately
+described in previous literature which raises questions on the
+interactions between circularly polarised light and spherical 
+aggregates. Until now, our understanding of angular momentum 
+transfer has been either focused on single particles (where 
+momentum transfer is easily described by Lorenz-Mie theory) or 
+for large aggregates of particles, where the angular momentum 
+transfer between individual particles is not considered. This 
+could provide a greater understanding of angular momentum and 
+help in the development of better torque sensing or torque 
+preventing methods.

### 2025-06-27 Latex_files/Thesis/chapters/chapter 4/main.tex
updates 15/04/25
27/06/2025 chapters 1,3,4 and structural changes

--- 2024-11-18 Latex_files/Thesis/chapters/chapter 4/main.tex
@@
-\chapter{Complex Langevin dynamics of spherical dimers}
-\label{chapter:langevin_dynamics}
-Much of the calibration theory discussed in Chapter 2 
-assumes that the target particle in question is a single 
-sphere, one whose scattering and motion is easily computed. 
-However, while working with dense colloidal suspensions, 
-one often ends up trapping more than one sphere. Li and 
-Arlt \cite{Li2008} studied the case of two microspheres 
-trapped in a single OT and found that multiple trapped 
-beads could be mistaken for a single trapped bead with 
-altered trap stiffness (see \ref{sec:harmonic_traps}). 
-Theoretical studies on the case of two trapped microspheres 
-by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
-based model to show that the two trapped beads are brought 
-into physical contact with each other by optical forces 
-and they also calculated the axial equilibrium positions 
-of the two trapped beads as a function of their size. 
-Experiments in \cite{Praveen2016} confirmed that the two 
-trapped beads indeed experience different trap stiffnesses 
-in the vicinity of the same potential well.
+\chapter{Detection of rotational spherical aggregate rotational dynamics}
+\textit{\textbf{Add overview here}} \\
+\label{chapter:simulated_detection}
 
-There are further discussions looking into the dynamics of 
-a whole host of asymmetrically shaped particles \cite{Loudet2014, 
-ShengHua2005, Chetana2022}, their results all showing that 
-predicting the behaviour an arbitrary shaped particle comes 
-with great difficulty due to the fact that the optical force 
-is dependent on a greater number of variables such as 
-orientation and size factors.
+\section{Overview}
+In this chapter we go about defining an optical fibre detection 
+system from a mathematical perspective. Before demonstrating how 
+it can be used to instantaneously predict orientational information 
+on a trapped particle. We then consider common factors of error in the 
+characterisation process, such as signal error, and incorrect particle 
+sizing. Both of which have a significant impact on the performance 
+of the model, but by utilising Bayesian inference, and time averaging 
+we can better refine our prediction to a satisfactory degree.
 
-With the initial goal of the project being to induce 
-nucleation events via a rotating sphere, the aim of this 
-chapter is to - in a limited capacity simulate and investigate 
-the influence of a second particle being bound to our target 
-sphere. The choice of a dimer, instead of an amorphous solid 
-that might better represent a growing crystalline solid, 
-allows us to consider how the dynamics of the aggregate change 
-by varying a single parameter, namely the size ratio of the 
-two spheres. Additionally, attempting to simulate an amorphous 
-aggregate is rather difficult as calculating the optical force 
-and torque is computationally slow. We build upon the works of 
-Vigilante \textit{et al} \cite{Vigilante2020} to consider 
-asymmetric spherical dimers and how varying size parameters 
-alters the dynamics and additionally makes characterising their 
-interactions within an optical trap more cumbersome.
+\section{Introduction}
+As outlined in the end of Chapter 3, one of the difficulties 
+in characterising interactions with asymmetric objects is 
+the coupled motion between translation and rotation. In order
+to characterise the optical trap a position detection system 
+such as a lateral effect photodiode or a quadrant photo diode
+(see fig.\ref{fig:position_detectors}). Typically, a position 
+detection system assumes that there exists a linear relationship 
+between the a particle's displacement and the detected signal.
+However, this is not always the case as demonstrated in chapter
+4; as dimer's demonstrate non-harmonic trapping beyond a certain 
+displacement. In addition, the rotational effects of the trapped
+dimer become a more significant factor when characterising the 
+trap strength. This can have unintended effects when it comes to
+areas of research that require a precise understanding of the 
+hydrodynamic behaviour of particles. As such there has been a 
+concerted effort to either eliminate rotational motion, or describe
+how the rotational behaviour influences the trajectory reported 
+by the position detector.
 
-In this chapter we will consider how the addition of a 
-second sphere changes the trapping dynamics by introducing 
-multiple equilibrium positions. These positions are 
-dependent on both the size ratio of the dimer and its 
-orientation relative to the trap. This is important to 
-address, for while the the dimers' diffusive behaviour 
-is the same, regardless of position and orientation, 
-how it interacts with the trap may produce the wildly 
-different results. Furthermore, we look at how our dimers 
-interact with circularly polarised light, this is especially 
-pertinent as the original plan was to utilise circularly 
-polarised light to generate fluid flow in a supersaturated 
-solution. The rotational motion of dimers in circularly 
-polarised light goes against much of the established 
-literature and suggests that the transfer of spin angular 
-momentum is enabled by combining particles together. And 
-lastly, we demonstrate how a quadrant photo diode, performs 
-in characterising the actual interactions between a dimer 
-and the optical trap. We do so to demonstrate that much 
-of the dimer's trajectory information is either lost or 
-poorly described by simple calibration techniques, making 
-their motion difficult to characterise.
+In this chapter we build upon the work of \cite{BarZiv1997} to 
+determine the instantaneous orientation of the aggregate. Using 
+static light scattering we can map the outputted signal from a 
+optically trapped dimer to its expected orientation in real time. 
+Thus allowing us to estimate the optical torque being applied to 
+the dimer by the trapping beam. As demonstrated in chapter 3, 
+the influence of a circularly polarised beam can result in 
+gyroscopic precession for specifically sized dimers.
 
-\subsection{Simulation Parameters}
-As a paradigmatic example, consider a dimer suspended in 
-water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
-a Gaussian beam (more specifically a Laguerre-Gaussian 
-beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
-is focused by a objective with numerical aperture of 1.2 
-and is x polarised. The size ratio of the two spheres is 
-given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
-\mu m$ unless specified otherwise; the dimer's orientation 
-is given by a unit vector connecting the centres of both 
-spheres (see \ref{sec:sim_parameters}). We define the 
-'standard' orientation as being aligned with the direction 
-of beam propagation direction - and therefore the 'inverted' 
-orientation is defined when the dimer is orientated against 
-the direction of beam orientation (see left hand side of 
-figure~\ref{fig:paradigmatic}).
+\section{Monitoring Stochastic rotational motion using static 
+		light scattering}
+Part of the optical fibre arrangement borrows code from the 
+optical tweezer simulations discussed in Chapter 2 and 4. 
+Rather than repeat the simulation details we will instead 
+only focus on the specifics of the optical fibre scattering.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Positional and Orientational dependence of Trapping forces}
-\label{sec:eq_positions}
-If we wanted to start from first principles and determine 
-the trap strength on our target particle the first step 
-would be to locate the equilibrium position relative to 
-the trap focus. For a single sphere it is easy to enough 
-to understand that its centre of mass will be drawn to 
-focal point of the laser due to gradient forces, once there 
-the force is analogous to a harmonic spring with a fixed 
-trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
-we consider instead a dimer, we now have two spheres both 
-being drawn to the focus along by the same gradient force; 
-in addition the scattering force is significantly more complex 
-due to both spheres scattering the electromagnetic fields. 
-This mutual scattering between individual spheres is what 
-makes simulating spherical aggregates far more difficult 
-compared to a single sphere, and even harder still to 
-predict the position where the dimer's centre of mass is 
-in equilibrium.
+\subsection{Coordinate System}
+Consider a particle (either a sphere or dimer) trapped by
+a Gaussian beam, with its focal point being set at [0,0,0]. 
+At the same time an optical fibre directs a plane wave that 
+is incident on the trapped particle scattering light in all 
+directions. The probe beam is assumed to be x-polarised and 
+propagating in the +z direction, this is a $90^{\circ}$ 
+rotation from the previous simulations (see fig\emph{X.X}).
 
-Because the scattering force is only significant in the 
-direction of beam propagation \cite{Capitanio2002} we 
-can assume there will only be a single potential well 
-lying at the centre of the beam. Along the beam axis, 
-such an assumption is no longer valid due to fact that 
-spherical aggregates experience inter-particle scattering. 
-This is a key reason for using \textit{mstm} as it 
-accounts for that behaviour. The methodology for computing 
-optical forces has been covered extensively for a number 
-of different trapping conditions \cite{RanhaNeves2019}. 
-So it is relatively easy to compute the trapping force 
-and determine the equilibrium position by finding the 
-position that minimises the net optical force and the 
-local gradient is negative ($\delta F/\delta x < 0$) - 
-we can assume that for a dielectric sphere the optical 
-torque is negligible. For a dimer (or any arbitrary 
-spherical aggregate), we now must consider both its 
-position and orientation and find where the net optical 
-force and torque are minimised.
+\subsection{Dimer}
+The dimer is defined by two spheres with refractive indices
+of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
+requires a target origin for setting the scattering expansion, 
+there is no agreed upon solution for where the target origin 
+should be relative to the sphere positions. In chapter 4 we showed
+that altering the target origin would drastically alter the 
+tmatrix produced. We choose (somewhat arbitrary) to use 
+the dimer's centre of diffusion which is located $\zeta a_I$ from 
+the centre of the larger sphere. For a discussion see \emph{X.X}.
 
-After computing its $T$-matrix via \textit{mstm} and 
-supplying that to \textit{ott} we compute the optical 
-force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
-in the axial direction while the dimer is in its 'standard' 
-orientation. As expected we see a single point where 
-the dimer will be in equilibrium, the linear fit in 
-fig.~\ref{fig:paradigmatic}(a) indicates a that the 
-force can be modelled as a harmonic potential close to 
-the equilibrium position ($F_z\approx-\kappa_z z$). 
-The second point where the axial force goes to 0 cannot 
-be considered as equilibrium position as the positive 
-gradient indicates that the trap is unstable unless 
-Brownian motion is ignored.
+\subsection{Scattering and Detection}
+Each detector is placed some distance $\bf r$ from the origin, 
+detector positions are defined by angles $\theta$ \& $\phi$. With 
+$\theta=\phi=0$ being directly in front of the probe beam. The 
+position of each detector is given by:
+\begin{align}
+	\begin{pmatrix}
+		x_{fiber} \\ y_{fiber} \\ z_{fiber}
+	\end{pmatrix} = 
+	\begin{pmatrix}
+		rcos(\phi)sin(\theta) \\ rsin(\phi)sin(\theta) \\ rcos(\theta)]
+	\end{pmatrix}
+\end{align}
+Where r is the distance between the detector and the target origin. 
+In an ideal situation the surface of the detector is oriented towards 
+the target origin perfectly. Pixels can be thought of as points lying 
+on the surface of the detector, each point has its own position in real
+space. Using \textit{mstm} we can find the components of the scattering 
+matrix at each point and compute the intensity of the electric field via:
 
-We repeated the same calculation but now while the 
-dimer is in its 'inverted' orientation, instead of 
-a single point where the optical force is minimised 
-we see that there are instead two separate equilibrium 
-positions, one above the focus and one below the 
-focus. In this particular example the two positions 
-are far enough apart that both can be considered as 
-separate harmonic traps.    
+\begin{align}
+	\begin{pmatrix}
+		I_s \\ Q_s \\ U_s \\ V_s
+	\end{pmatrix} = 
+	\begin{pmatrix}
+		S_{11} & S_{12} & S_{13} & S_{14} \\
+		S_{21} & S_{22} & S_{23} & S_{24} \\
+		S_{31} & S_{32} & S_{33} & S_{34} \\
+		S_{41} & S_{42} & S_{43} & S_{44} \\
+	\end{pmatrix}
+	\begin{pmatrix}
+		I_i \\ Q_i \\ U_i \\ V_i
+	\end{pmatrix}
+\end{align}
+Where:
+\begin{align}
+	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2  \label{eq:intensity} 
+	\\
+	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
+\end{align}
+
+By computing the scattering at each point on the detectors surface 
+we can compute the average scattering over the detector. From a
+experimental perspective the distance from the particle to the 
+detectors must be considered. The near field scattering can be 
+computed with relative ease using \textit{mstm} as shown in 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{.65\linewidth}
-		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
-		\caption{}
-		\label{lam=2}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.25\linewidth}
-		\centering
-		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
-		\label{large over small}
-	\end{subfigure}
-	\medskip
-	\begin{subfigure}{.65\linewidth}
-		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
-		\caption{}
-		\label{lam=2_inverted}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\begin{subfigure}{.25\linewidth}
-		\centering
-		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
-		\label{small_over_large}
-	\end{subfigure}
-	\caption{Plots of force vs displacement of the 
-		centre of mass of the dimer (µm) for the 
-		case of a dimer of size ratio 2. (a) is the 
-		case where the dimer is in its' 'standard' 
-		orientation, where the dimer is trapped at 
-		$z = 1.71\ \mu m$. (b) is the case where the 
-		dimer is in its' 'inverted' orientation, the 
-		dimer is trapped at two positions: $z = 1.20\ 
-		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
-		renders to visualise the dimer orientation 
-		are shown below each plot. The black lines 
-		on each force-curve is a linear fit with the 
-		slope being reported as the trap stiffness in 
-		the legend.}
-	\label{fig:paradigmatic}
+	\includegraphics[width=\linewidth]{near_field_intensity_side_profile.png}
+	\caption{Intensity distribution (as defined by \eqref{eq:intensity}) 
+		from a symmetric dimer when irradiated by the probe beam 
+		propagating in the $k\hat{z}$ direction (viewed in the 
+		z-y plane). Axis are given in dimensionless units of 
+		$k\hat{z}$ and $k\hat{y}$.}
+	\label{fig:nf_scattering}
 \end{figure}
 
-We can see that both equilibrium positions have comparable 
-axial trap stiffness ($\kappa_z$), however the difference 
-in the transverse trap stiffness ($\kappa_x$) is far more 
-noticeable. The same dimer was trapped at each of the axial 
-equilibrium positions and the transverse force was evaluated. 
-While in all three cases the dimer can be trapped the linear
-range where that would typically associated with a stable 
-trap is far narrower in the 'standard' orientation compared 
-to the 'inverted' cases. This highlights one of the challenges 
-involved with studying asymmetric particles, even though its 
-a simple enough process to trap them they maybe characterised 
-very differently depending on their relative position and 
-orientation towards the focus. This can have a significant 
-impact on rheological studies - or attempting to probe any 
-local property - as the variance in trap strength can result 
-in large errors over repeated measurements. 
+While \textit{mstm} does provide the amplitude matrix it is not
+scaled to account for distance from the target to the detector.
+We amend this by simply scaling according to the inverse square 
+law. As the detector is moved further from the target the 
+intensity distribution across the entire detector will decrease
+substantially. Because of the fall off it is more computationally
+efficient to instead consider only the central pixel of the 
+detector rather than across the entire surface. If we consider 
+the difference between the average intensity and the intensity 
+of the central point we can see that after not even $20 \mu$ 
+away from the target particle the average intensity across the 
+surface is within $0.1\%$ of the central pixel. 
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{transverse_force.png}
-	\caption{Plots of force vs displacement of the dimer's 
-		centre of mass spheres, where a positive force 
-		indicates the dimer is directed right on the x-axis, 
-		and vice versa for a negative force. The same simulation 
-		parameters are used here as in fig~\ref{fig:paradigmatic}
-		(a) and (c). The blue curve representing the force 
-		response for a dimer in its standard orientation, orange 
-		being the inverted case, and green the same case but 
-		placed below the focus.}
-	\label{fig:transverse_force}
+	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
+	\caption{Normalised difference between the central pixel intensity
+	$I(\theta, \phi)$ and the average pixel intensity over the surface
+	of the optical fibre $\langle I \rangle$ as a function of distance.
+	4 different optical detectors were tested (spherical coordinates 
+	are given in the legend). While the choice of coordinates is 
+	arbitrary they all demonstrate the same exponential decay. Dashed
+	lines indicate the difference after $20\mu m$ from the trap focus,
+	in most practical cases getting an optical fibre any closer would
+	disrupt the trap stability.}
 \end{figure}
 
-\newpage
-For completeness the harmonic traps were located for dimers
-across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
-$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
-for each trap. The same simulation parameters are used here 
-as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
-As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
-the dimer begins to approximate a single homogenous sphere 
-- at least in terms of location and trap strength. However, 
-for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
-to $a_{I}/a_{II}=4$), a second equilibrium position is found 
-below the trapping focus. Previous work using the ray-optics 
-model have confirmed even in the case that two spheres begin 
-separated the electric field will align the particles as such 
-that they make contact and are trapped together about a single 
-trapping position \cite{Xu2005}. Furthermore it has been shown 
-through proper manipulation of the Gaussian or Bessel beam 
-modes that any number of trapping potentials can be developed 
-\cite{Shahabadi2020} for nanoparticles. This result however, 
-is the first example of an orientation dependent trapping 
-situation using only a $TEM_00$ beam. Typical experimental 
-arrangements cannot determine much information on the axial 
-position of a trapped particle relative to the trap focus; 
-this result indicates not only that dimers can be trapped 
-in multiple axial positions but also their trapping behaviour 
-is heavily dependent on said axial position. As such it is 
-necessary that positional information in the z-axis can be 
-elucidated if multiple spheres are trapped simultaneously. 
+Given the trapping beam's beam waist is approximately $0.54 \mu m$,
+a optical fibre would need to be a minimum of $150 \mu m$ from the 
+trap centre in order to prevent noise from the trapping laser itself.
+At such a distance the variation in the signal is so small that it is 
+easier to evaluate the intensity at the centre of the detector instead
+of across the entire surface. This is how the instantaneous orientation
+is predicted later on in the chapter.
+
+\section{Interpretation of scattering data into orientation estimates}
+We now discuss the methodology for computing the instantaneous 
+orientation of an optically trapped particle based on scattering
+detected by a finite grouping of optical fibres.
+
+Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), 
+we can define at any point in time a unit vector $\hat{s}$ 
+pointing from the centre of the larger sphere to the centre 
+of the smaller sphere. A plane wave probe beam incident, 
+is incident on the dimer, generating a scattering pattern
+dependent on the dimer's orientation $I(\hat{\bf s}, 
+\theta, \phi)$ which is computed using \textit{mstm} 
+\cite{I.Mishchenko1996}. To represent the experimental set 
+up consisting of a set of optical fibres recording scattered 
+light, we choose four sets of spherical angles [$(\theta_1,\ 
+\phi_1), \ (\theta_2,\ \phi_2), \ (\theta_3, \ \phi_3), \ 
+(\theta_4,\ \phi_4)$] and record the calculated intensity at 
+each angle $I(\hat{s},\ \theta_k,\ \phi_k)$. 
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.85\linewidth}
-		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
-		\caption{}
-		\label{eq_pos}
-	\end{subfigure}\hfill % <-- "\hfill"
-	\medskip
-	\begin{subfigure}{0.85\linewidth}
-		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
-		\caption{}
-		\label{eq_pos_inverted}
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig2a.png}
 	\end{subfigure}
-	\caption{Equilibrium positions of optically trapped dimers with varying 
-		size ratio, dashed lines represent unstable traps whereas solid lines 
-		are for stable equilibrium positions. (a) shows that dimers while in their
-		'standard' orientation will always have a single equilibrium position. 
-		(b) shows that when the same dimer is in its' 'inverted' orientation 
-		can be trapped in two axial positions, one below the focus and one above 
-		the focus.}
-	\label{fig:eq_positions}
+	\begin{subfigure}{0.49\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig2b.png}
+	\end{subfigure}
+	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
+	\label{fig:dimer}
 \end{figure}
 
-\newpage
-\subsection{Non-trivial equilibrium configurations}
-\label{sec:off-axis}
-Computing the equilibrium positions when a dimer is aligned 
-with the electric field is relatively simple as the orientational 
-torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
-trapped the dimer is unlikely to change orientation enough to 
-escape the trap. Regardless, that does not rule out the 
-possibility that there is a stable configuration where the 
-orientation not strictly vertical, in fact most experimental 
-work with symmetric nano-dimers will trap them lying perpendicular 
-to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
-\ref{sec:eq_positions} we cannot simply measure the optical force 
-and torque as the parameter space is too large and determining if 
-a particular position and orientation is stable is not clear based 
-solely on force and torque measurements \cite{Bui2017}. Using the 
-same simulation parameters as before we ran a number short 
-simulations (total simulation time was $0.005$ s) with the laser 
-power increased to $500\ mW$. Each simulation started with the 
-dimer in a different starting position and orientation, due to the 
-high laser power the dimers either escaped the trap or were stably 
-trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
-between the direction of beam propagation and the dimer's orientation 
-vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
-divided into different regions depending on which equilibrium 
-configuration is reached.
+Our goal is to determine the orientation of the trapped dimer 
+based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
+Rather than aim an exact estimate of the dimer's orientation, 
+for the purposes of interpretation of the scattering and 
+optimisation of the measurement setup it is more convenient 
+to discretize the possible orientation space into a number 
+of possible reference orientations. We can then use each as 
+'classification categories' in a neural network methodology 
+to map scattering data to orientation (see below for further 
+discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
+reference orientations $\hat{\bf n}_{\alpha}$  evenly 
+distributed on a unit sphere \cite{Reyuthor2006} 
+(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
+neighbour spacing between two neighbouring reference 
+orientations of 0.895 radians.
+
+Using \textit{mstm} we compute the raw intensities (see 
+\eqref{eq:intensity}) at each of the measurement angles 
+that would be generated by a dimer in each reference 
+orientation, $I_s(\hat{\bf n}_{\alpha}, \theta_k)$. While 
+the number and position of detection fibres is technically 
+arbitrary there are several constraining factors that limit 
+our ability to infer useful information from the trapped 
+object, see Section~\ref{sec:brownian} for a detailed 
+breakdown of our choice of detection angles. In order for 
+a neural network to properly weight each signal the raw 
+intensities are normalized according to:
+\begin{align}
+	\label{eq:scale}
+	y_k(\hat{\bf n}_\alpha)
+	= 
+	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
+	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
+\end{align}
+
+where the denominator is simply the standard deviation across 
+the set of values $I(\hat{\bf n}, \theta_k)$. Note that the 
+collected scattering signals are not necessarily simply 
+related to their associated reference orientations: as is well 
+known from such examples of the inverse scattering problem. 
+While it is trivial to compute the light scattering pattern 
+for any given particle with any particular characteristic (i.e. 
+size, shape, or orientation), inferring the light scattering 
+from a unknown particle to determine said characteristic is 
+incredibly difficult due to complex mapping between scattering 
+and said characteristic. Even if the orientation space is 
+divided evenly between reference orientations the subsequent 
+signal space ends up being appearing mixed making simple 
+comparisons of signals useless for inferring information on the 
+particle.
+
+Fig~\ref{fig:mixing} shows two clusters of orientation 
+vectors and their respective signals collected from 3 detectors
+- the points have been coloured based on their proximity 
+to the centre of their respective cluster. While the 
+orientation space appears tightly packed and ordered the 
+signal space quickly spreads out in an asymmetric fashion. 
+Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
+mapping can intersect itself which only further increases 
+the complexity. While in some instances the mapping between 
+one reference orientation and another is discrete, in other 
+instances the mapping becomes far more complex to discern.
 \begin{figure}[h!]
 	\centering
-	\begin{subfigure}{0.67\linewidth}
-		\includegraphics[width=\linewidth]{off_axis_trap.png}
+	\begin{subfigure}{0.74\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig3a.png}
 	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
-		\caption{}
+	\begin{subfigure}{0.74\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig3b.png}
 	\end{subfigure}
-	%\captionsetup{hangindent=2.35cm}
-	\caption{Map of $z-\theta$ phase space using a dimer of
-		size ratio 2 with a laser power of 500 mW ($\theta=0
-		\circ$ is the 'standard' orientation and $\theta=180
-		^\circ$ is the 'inverted' orientation). The stable 
-		configurations are indicated by the larger circles 
-		and the starting conditions are colour coded to match 
-		the stable point they end up in. Right hand render 
-		shows a dimer in its off-axis configuration.}
-	\label{fig:off_axis}
+	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals.
+Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
+	\label{fig:mixing}
 \end{figure}
 
-Interestingly while the trap strength of these off-axis traps 
-are similar in magnitude to the vertically aligned dimers, 
-but when the laser power is lowered (around $5\ mW$) the traps 
-become metastable; after reaching its equilibrium configuration 
-the particle behaves similarly to a typically trapped dimer but 
-the trapping potential is small enough that the dimer escapes 
-in as little time as less than a second or after nearly a full 
-3 seconds. Running similar simulations but for vertical 
-configurations sees the dimer remaining trapped, even after 30 
-seconds of run time, indicating that the trapping potential is 
-far greater than the thermal energy. This suggests that the 
-reason this off-axis configuration is metastable is due to the 
-increased rotational freedom. Similar configurations have been 
-explored with ellipsoids; Zhu \textit{et al.} looked at the 
-dynamics of various elliptical particles and found that regardless 
-of shape or initial orientation the particle would tend towards
-either a purely vertical or purely horizontal orientation 
-\cite{Zhu2021}.
+Nevertheless, at least where the uncertainty in signal measurements 
+is low, we can predict the orientation from the scattering scattering 
+by utilising computational techniques such as neural networks. We 
+thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
+dimer's orientation from its light scattering signal. The network 
+was trained by generating a database of random orientation vectors, 
+calculating the corresponding light scattering signals, and then 
+using the network to estimate the probability of a given signal coming 
+from a dimer in a given reference orientation. The network's loss 
+function was evaluated and used to improve the estimation, the 
+network being trained until the improvement in the loss function 
+was less than 0.0001.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Continuous rotational motion in circularly polarised light}
-One aspect that has yet to be covered in depth with regards 
-to spherical aggregates of any construction is their interaction 
-with circularly polarised light. For homogenous spheres the 
-optical torque is regarded as being negligible as the spin 
-density cannot impart angular momentum while propagating in a 
-homogenous medium. Dimers however, have been shown to experience 
-an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
-while trapped in a circular polarised beam. In our simulations 
-we found that dimers would rotate about their long axis when 
-trapped in circularly polarised light. In this section we want 
-to discuss how this behaviour is influenced by size, position, 
-and orientation; and furthermore, we wish to address possible 
-explanations for this behaviour, as none of the current
-theories into optically induced rotation seem plausible.
+Importantly, the estimation provided by the neural network can be 
+improved further by accounting for any prior information we know 
+about the dimer, utilising Bayesian inference to update the neural 
+network's estimation: 
+\begin{align}
+	\label{eq:bayes}
+	p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
+	&=
+	\frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
+		p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
+\end{align}
 
-\subsection{Polarisation Dependency on Dimer trajectory}
-\label{sec:rot_pol}
-In their paper Vigilante \textit{et al} attribute the rotational motion 
-to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
-scattering is a viable theory for describing optical rotation, however
-there is usually an optical axis about which rotation occurs. For a dimer
-this is typically the orientation vector that passes through the centres 
-of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
-cases where the particle's cross sectional shape is engineered to scatter
-light in one particular direction \cite{Higurashi1994}, but if this was
-the case then the dimer should rotate when illuminated by any polarisation
-of light.
+where $p(\hat{\bf n}_\alpha)$ and $p(y_k(\hat{\bf s}))$ are the prior 
+estimates of the distributions of particle orientations and 
+instantaneous signals, respectively. \textit{Without} any prior 
+evidence we must assume that the orientation prior of the dimer 
+$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
+dimer's possible current orientation from knowledge of previous 
+measurements can be used to inform our estimate of $p(\hat{\bf 
+n}_{\alpha})$. The latter prior $p(y_k(\hat{s}))$ is the probability of 
+measuring a signal ($y_1$, $y_2$, $y_3$). This is given by taking 
+the discrete integral over the collection of reference orientations:
+\begin{align}
+	p(y_k(\hat{\bf s}))
+	=
+	\sum_{\alpha=1}^{n_{\rm ref}}
+	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
+	p(\hat{\bf n}_\alpha)
+\end{align}
 
-To that end, we simulated the motion of an optically trapped dimer in 
-beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
-composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
-run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
-orientational time series; the dimer's orientation is recorded as a 
-quaternion which can be easily converted to a 3-dimensional rotation 
-matrix. By considering only the transverse components ($U_{1,x}$, 
-$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
-the Fourier transformation of their time series reveals the rotational 
-frequency. The laser power is set to $100\ mW$ to avoid large thermal 
-fluctuations and so that the Fourier series of the transverse components 
-approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
-at the rotational frequency $\omega_{rot}$.
+From \eqref{eq:bayes} we obtain the key result, a mass probability 
+distribution denoting the probability that our dimer is in 
+orientation $\hat{\bf n}_{\alpha}$ given a measured signal ($y_1$, 
+$y_2$, $y_3$), \textit{i.e.} an estimated mapping from scattering 
+measurement to orientation estimate.
 
-%% Example plot of rotating beam trajectory and then the same thing but 
-%% Fourier transformed
-\begin{equation}
-	\begin{split}
-		q(t) \rightarrow R(t) =& 
-		\begin{pmatrix}
-			U_{1,x}(t) && U_{2,x}(t) && U_{3,x}(t) \\
-			U_{1,y}(t) && U_{2,y}(t) && U_{3,y}(t) \\
-			U_{1,z}(t) && U_{2,z}(t) && U_{3,z}(t) 
-		\end{pmatrix} \\
-		\rightarrow
-		&\int^\infty_{-\infty}R(t)e^{-i2\pi ft} dt = 
-		\begin{pmatrix}
-			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f) \\
-			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f)\\
-			\delta(f) && \delta(f) && \delta(f)
-		\end{pmatrix}
-	\end{split}
-\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Calculation of error}
+\label{sec:divergence}
+Using \eqref{eq:bayes} we get a snapshot estimation of the instantaneous
+orientation of the trapped dimer based on a single measurement from a 
+collection of detector signals. Of course in reality these detectors 
+will be returning a discrete time series representative of the dimer's
+trajectory within the optical trap.
 
-If the rotational frequency was not immediately obvious the simulation was
-repeated but over a longer simulation time. Four different size ratio of 
-dimers were studied, both in their 'standard' and 'inverted' orientations. 
-The results of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
+Using the simulation code from \cite{Vigilante2020} we can replicate 
+this trajectory and furthermore use \textit{mstm}to calculate the far field 
+intensity produced by said dimer. At each time step we measure the signal
+produced at each measurement angle and use our neural network to calculate
+$p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha, t)$.
+
+Evaluating how well Eq.~\eqref{eq:bayes} performs in predicting $\hat{\bf n}$
+is difficult because we have discretised the orientation space meaning 
+there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method provides a measure of the relative
+entropy change between two probability distributions:
+\begin{align}
+	D_{KL}(P(t)\parallel Q(t)) = \sum P(t) \log_2\left[\frac{P(t)}{Q(t)}\right] 
+	\label{eq:D_KL}
+\end{align}
+
+Where $P(t)$ and $Q(t)$ are two probability distributions we are
+comparing, often it is common to use a base 2 log as this converts 
+the entropy to base 2. By relative entropy we are talking about 
+the useful information gained/lost by switching from $P(t)$ to 
+$Q(t)$. In our case $Q(t)$ is simply the distribution given by
+Eq.~\eqref{eq:bayes} and $P(t)$ is an idealised distribution, 
+one where the model is accurate with 100\% certainty.
+\begin{align}
+	\label{eq:best}
+	p_{best} = P(t) =
+	\begin{cases}
+		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
+		0 & \text{when $\hat{\bf n}_\alpha$ $\neq$ $\hat{\bf n}_{best}$}
+	\end{cases}
+\end{align}
+
+In reality the distribution from Eq.~\eqref{eq:bayes} will assign 
+some non-zero probability to every reference orientation. Therefore 
+the relative entropy for a single measurement is given as:
+\begin{align}
+	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
+	= \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
+	\right]
+	\label{eq:kullback}
+\end{align}
+
+The relative entropy is a measure if how spread out our distribution
+is. With larger values of $D_{KL}$ indicating that our prediction is 
+spread out so the neural net is unsure about $\hat{\bf n}_{best}$. 
+Whereas a low value of $D_{KL}$ means that our model is more confident 
+in $\hat{\bf n}_{best}$. The only draw back being that the relative 
+entropy does nothing to indicate accuracy, only how closely our 
+distribution matches our ideal result. This will be better 
+exemplified later in \ref{sec:test}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Brownian Simulation}
+\label{sec:brownian}
+
+We use the Brownian OT package developed by Fung~\textit{et~al} 
+\cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. We simulate the 
+motion of a dimer trapped in a highly focused Gaussian beam by 
+calculating the optical forces imparted by the laser, and the 
+Brownian force due to the surrounding fluid. We simulated a 
+polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} 
+= 1.33$). As discussed in chapter 4 dimers can be stably trapped
+in off-axis orientations. We simulate a dimer (of size ratio 2) that
+is trapped in this off-axis orientation by a $5 mW$ laser, due to 
+the meta-stability of the trap the dimer is only trapped for second
+before it escapes the trap. The dimer's position and orientation 
+is shown below:
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
-	\caption{Rotation frequency vs component phase difference for 
-		differently sized dimers. The solid lines represent the 
-		rotation rate experienced while the dimer is in its standard 
-		orientation, whereas the solid points are for the case 
-		where the orientation is inverted. Laser power = $100\ mW$.}
-	\label{fig:rotation_vs_pol}
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig7a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.45\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig7b.png}
+	\end{subfigure}
+	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
+	\label{fig:motion}
 \end{figure}
 
-This shows us that these optical rotations are polarisation 
-dependent and not merely an example of the dimer scattering 
-light asymmetrically. We also know that this cannot be a 
-similar mechanism to that of experimental works \cite{Ahn2018, 
-Reimann2018} as in our simulations the dimer's long axis
-is parallel to the electric field. In this orientation 
-the dimer's optical axis is aligned perpendicular to the
-polarisation vector meaning that there should be no angular
-momentum transferred along the optical axis. This is consistent 
-with other experiments involving elongated particles. A dual
-beam trap was used to study the dynamics of "disk like" particles. 
-They found that these particles had one of two stable
-orientations: Either orienting with the 'flat' side perpendicular
-to the beam trap, with no rotation being observed \cite{Brzobohaty2015}. 
-Or orientating with the 'flat' side parallel to the trap, in which 
-case the long axis was aligned with the polarisation vector and 
-thus rotational motion was detected \cite{Brzobohaty2015}.
+At the same we use \textit{mstm} to replicate the intensity of light
+incident on the surfaces of detectors placed around the focus of the
+optical trap. The exact number of detectors was initially assumed to 
+be arbitrary, in that it made no difference to our estimate whether 
+we used 2 angles or 200.
 
-\subsection{Brownian Vortex via Curl of Spin momentum}
-In their work Vigilante refers to the spin-curl effects 
-demonstrated by Grier \textit{et al} \cite{Ruffner2012}, 
-in which the curl of the spin density leads to a second 
-order optical force that orbits around the beams central 
-axis \cite{Yevick2017}. While several papers have 
-demonstrated this phenomena \cite{Zhao2007, Zhao2009, 
-Wang2010} it was only properly formalised by \cite{Ruffner2012}. 
-In which they showed that the seemingly random trajectory 
-of a trapped sphere was biased by the polarisation state 
-of the trapping beam. While not immediately evident from 
-the trajectory the helicity of the trapping beam was revealed 
-by computing the particle's probability flux using.
+\subsection{Choice on the number of detectors}
+When all of the detectors lie in the same plane the expected signal 
+can appear identical despite the dimer being in completely different orientations. It should be noted that these pairs are reflected in 
+one or more axis which suggests that these are due to the arrangement 
+of our detectors. More specifically, if the detectors are placed say 
+in the x-y plane then only when the dimer is pointed nearly fully 
+upright will the expected signal be entirely unique. This is not an 
+issue with typical analytical static light scattering experiments 
+because the signal being collected is averaged over a population of 
+particles all in different orientations. Whereas in our set up we are concerned with the scattering of a single particle whose scattering is orientationally dependant.
+
+To remedy this we raise the third detector out of the x-y plane; 
+as such the expected signals from each reference orientation is 
+unique, though with limited resolution. By adding a $4^{th}$ 
+detector we can differentiate signals more reliably, improving the 
+neural networks performance. In line with our goal of making this 
+method viable in a laboratory setting we decided not to increase 
+the number of detectors further than 4.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Testing the Neural Network}
+\label{sec:test}
+We applied Eq.~\eqref{eq:bayes}, taking the reference orientation 
+with the highest probability  as our estimate of the dimer's 
+instantaneous orientation $\hat{\bf n}_{est}$. To visualise the 
+model's performance we plotted the orientation components of our 
+estimation $\hat{\bf n}_{\alpha}$ and the dimer's \emph{actual} 
+instantaneous orientation $\hat{\bf s}$ versus time. For comparison, 
+we also plotted the orientation components of the closest reference orientation, $\hat{\bf n}_{best}$. Assuming a uniform prior of the 
+reference orientations $p(\hat{\bf n}_\alpha)$ the neural network's predictions ($\hat{\bf n}_{\alpha}$ from Eq.~\eqref{eq:bayes}) are 
+at times reasonable, but there are significant large and random jumps 
+away from the correct result (Fig.~\ref{fig:uniform}). 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\textwidth]{fig5.png}
+	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity.
+Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the blue line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
+Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
+	\label{fig:uniform}
+\end{figure}
+
+One reason we observe such large jumps in the estimated 
+orientation is simply because the neural network is assuming
+that the signals are unrelated from one another. We can 
+observe this if we take the average relative entropy change
+across the entire trajectory:
+
 \begin{align}
-	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
-	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
-	\label{eq:prob_flux}
+	\langle D_{KL}\rangle = \frac{1}{N} \sum_{i=1}^{N} \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha\parallel y_k(\hat{s}),t_i)}\right]
+	\label{eq:score}
 \end{align}
 
-where $\delta_{\sigma_j}$ is the kernel of an adaptive 
-density estimator \cite{Silverman1986}. \eqref{eq:prob_flux} 
-describes the direction a trapped sphere is most likely 
-move in given our statistical knowledge of the 
-trajectories probability density function. A finite 
-estimation of the density function $p(r)$ is used in 
-\cite{Ruffner2012}.
+Where we get an average entropy change of $\approx 0.604$, in 
+base 2 this means that on average the model has narrowed down 
+its options from 1 in 30 to 1 in 1.52. This is not reflected
+though in fig.~\ref{fig:uniform} which indicates that while 
+the model is highly confident in its results it is not at all 
+accurate. As demonstrated by fig~\ref{fig:mixing} the 
+orientation and signal spaces are uncorrelated with one another 
+which results in the neural network to assign orientations 
+incorrectly. Combining this fact with use of a uniform prior 
+($p(\hat{\bf n}_\alpha)$), there is no constraint on how much 
+estimated orientation can change from time-step to time-step.
+
+To improve the estimation we can therefore use knowledge of the 
+physical limitations of the object in the trap and its dynamics. 
+Imposing a more physically grounded prior, accounting in this case 
+for the fact that the rotational motion of the dimer is limited by 
+the rotational trap stiffness ($\kappa_r$). Here the prior of the 
+reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at 
+each time step as a Boltzmann distribution of the physical distance 
+between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and 
+each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are 
+reweighing our estimation based on the size of rotation, with smaller movements being favoured over large movements:
 \begin{align}
-	p(r) = \frac{1}{N}\sum_{j=1}^N \delta_{\sigma_j}(r-r_j)
-	\label{eq:prob_density}
+	p(\hat{\bf n}_\alpha)
+	&=\frac{e^{\hat{\bf n}_\alpha 
+			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
+	{\sum_{\alpha=1}^{n_{\rm ref}}
+		e^{\hat{\bf n}_\alpha 
+			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
+	\label{eq:boltz}
 \end{align}
 
-The probability flux reveals a biased motion in the trajectory 
-of a single sphere (see Fig.~\ref{fig:Ruffner_Grier}). This 
-biased motion results in a slight orbital motion about the 
-central axis of the trapping beam, the orbital frequency is 
-shown to be proportional to the polarisation state of the 
-trapping beam. 
+As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.
+However at the same
+time result is still subpar.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{fig8a.png}
+	\caption{\label{fig:biased}
+		%
+		Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
+		by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
+		achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
+		Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
+		%
+	}
+\end{figure}
+
+Once again using Eq.~\eqref{eq:score} we now get a result of 
+$\approx 0.204$ or a 1 in 1.2 result. This is only a slight 
+improvement in terms of confidence but has drastically reduced 
+the variation in our models predictions. There are other 
+improvements that can be made, which additionally also help with
+potential sources of error in our set up.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Accounting for sources of error in light scattering measurements}
+When it comes to analysing light scattering from any size particle,
+error analysis becomes a significant factor. Typically this can be 
+accounted for by averaging over long periods of time to get an 
+assessment of the steady state conditions of the target particle. 
+However in our case where we wish to know the instantaneous 
+orientation, we instead have to rely on our understanding of how 
+uncertainty can effect our model's performance. We identified two 
+areas which are likely sources of error in our estimation: firstly, 
+an incorrect modelling of the target particle, and secondly, signal 
+noise arising from experimental factors. We highlight how we address 
+these areas below. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Impact of incorrect dimer sizing}
+\label{sec:lam}
+
+One of the main limitations of our model is that we assume that 
+the dimer being modelled in \textit{mstm} is accurate to the 
+dimer being trapped in the optical tweezer. Sizing molecules 
+accurately is a significant challenge for single particle 
+analysis so there is bound to be some uncertainty with the 
+measurements. We ran our model 3 times with the neural net being 
+trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
-	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how 
-		the momentum density of a Gaussian beam is twisted while 
-		using circularly polarised light. The top row (figures 
-		(b)-(d)) shows a 7 sphere cluster trapped in a circularly 
-		polarised beam. Due to the clusters asymmetric susceptibility 
-		to polarization the cluster rotates in the $x-y$ plane. 
-		Whereas the bottom row (figures (e) - (g)) show the similar 
-		results for a single sphere. In this instance the sphere 
-		does not rotate but instead orbits the beam axis. In both 
-		instances the motion is proportional to the degree of 
-		polarisation (see figures (d) and (g)) but for the single 
-		sphere this motion is only revealed when using 
-		\eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused 
-		with permission from author}
-	\label{fig:Ruffner_Grier}
+	\begin{subfigure}{0.33\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig8a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig8b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig8c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has 
+		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.519]$, (b) 1:2.05 $[\langle D_{KL}
+		\rangle=3.706]$, (c) 1:1.95 $[\langle D_{KL}\rangle=
+		3.705]$ ($n_{refs} = 30$)}
+	\label{fig:size}
 \end{figure}
 
-While the results from \cite{Ruffner2012} suggest that the 
-optical rotation seen in asymmetric dimers can be attributed to 
-the same spin-curl forces there are several questions that 
-cannot be explained purely by the spin-curl force.
-
-\subsection{Optical torque differences}
-In their paper Grier \textit{et al} show that a 7 sphere cluster
-will rotate when trapped in circularly polarised light but a 
-homogenous sphere will have a slight curl to its trajectory. 
-They attribute this to the curl of the spin angular momentum
-generating a "Brownian vortex". But what is more interesting, 
-and something that appears unaddressed, is how the effects of 
-this vortex change based on the overall shape of the target
-particle.
-
-You would not expect that the addition of a single additional 
-sphere should drastically adjust the torque especially if said 
-sphere is relatively small. However when we measured the optical 
-torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we 
-found the exact opposite. In both cases we used the same trapping 
-beam as used for fig.~\ref{fig:rotation_vs_pol} but with a 
-circularly polarised beam. Both the sphere and dimer were rotated 
-in the $x-z$ plane and the all three components of the optical 
-torque were recorded.
+As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model.
+This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
+This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
+We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Reyuthor2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
+While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
-	\caption{Optical toque experienced by a dimer ($a_{I}/a
-	{II}=10$) and a single isotropic sphere. Both were rotated 
-	in the $x-z$ plane and the angle between $U_z$ and the beam 
-	axis gives the polar angle.The solid lines denote the torque 
-	experienced by the dimer whereas the dashed lines represent 
-	the torque experienced by the sphere.}
-	\label{fig:sphere_dimer_torque}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width =\textwidth]{fig9a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig9b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.31\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig9c.png}
+	\end{subfigure}
+	\caption{Model estimates of orientation when neural net has 
+		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.593]$, (b) 1:2.05 $[\langle D_{KL}\rangle
+		=5.659]$, (c) 1:1.95 $[\langle D_{KL}\rangle=3.279]$, 
+		($n_{refs} = 126$)}
+	\label{fig:refs}
 \end{figure}
 
-The torques about the $x$ and $y$ axis can be somewhat understood 
-as the second sphere is being drawn back towards the centre of the 
-trap by the gradient forces. The same cannot be said for the $z$-
-component of the optical torque, which is non-zero even at a 
-$80^\circ$ angle. This cannot be simply explained via spin-curl 
-effects, as the 'Brownian vortex' should instead be driving the
-dimer around the beam axis - similar to how the 7 sphere cluster 
-behaved. What is clear that the mere combination of two spheres
-results in a unique behaviour that has not been previously 
-investigated.
+Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer.
+This suggests that overshooting our size estimate will be less detrimental to our estimation.
+Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations.
+This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction.
+Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Impact of measurement noise on model predictions}
+\label{sec:epsilon}
 
-\subsection{Rotational frequency as a function of size ratio and 
-	orientation}
-\label{sec:rot_size}	
-One factor t Intuitively, you would expect that a 
-larger particle would experience a greater torque and therefore 
-rotate faster. By repeating the same kinds of simulation as used 
-in \ref{fig:rotation_vs_pol} but for a circularly polarised beam 
-$\phi=90\circ$ it was found that not only is the rotation rate 
-dependent on the size of the dimer, but also on its orientation 
-and therefore their axial position.
+So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it.
+In reality of course scattering signals will always have some non-zero measurement noise.
+This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself.
+To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
+\begin{align}
+	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
+\end{align}
+where $\epsilon$ is the percentage error associated with the scattering signal.
+Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:
+
 \begin{figure}[h!]
 	\centering
-  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
-  \caption{Rotation rate plotted against dimer size ratio while 
-  	trapped in a circularly polarised beam; a positive rotation 
-  	rate indicates clockwise rotation, whereas a negative 
-  	rotation rate indicates counter-clockwise rotation. The red 
-  	line is for the case of a dimer in its 'standard' orientation. 
-  	The blue line is for the case when the dimer is in its 
-  	'inverted' orientation while trapped above the focus of the 
-  	beam. And lastly the green line is for the case when the 
-  	dimer is in its 'inverted' orientation, but when it is trapped 
-  	below the focus of the beam.}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig10c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of (a) $1\%$ ~
+		$[\langle D_{KL}\rangle=0.962]$, (b) $15\%$ $[\langle D_{KL}\rangle=13.654]$, and (c) $25\%$ ~  $[\langle D_{KL}\rangle=13.017]$.}
+	\label{fig:epsilon}
 \end{figure}
 
-It is difficult to see from the graph, but the rotation rate never
-truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which 
-would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
-to induce rotational motion. What is also interesting is that the 
-rotation rate is not correlated directly with either the particle 
-size or the equilibrium position. This is in stark contrast to previous
-reports of dimer optical rotation; Ahn \textit{et al.} reported that
-the rotational frequency is maximised when the dimer is symmetric 
-\cite{Ahn2018}. Their work was conducted in a vacuum so the only 
-limiting factor to the optical rotational motion is the structural 
-stability of the particle itself.
+As can be seen from Figure~\ref{fig:epsilon}, the inclusion of 
+signal noise quickly leads to a decrease in the model's 
+performance. So much so that beyond 15\% the model is actually 
+worse than just randomly guessing.This is due to an inherent 
+feature of the inverse scattering problem: two distinct regions 
+in orientation space can become heavily intertwined and thus no 
+longer well separated when mapped to intensity space (even 
+though the mapping remains continuous): so even small 
+uncertainties in the scattering data can lead to large 'mistakes' 
+in the choice of orientation by the neural network. (Indeed if 
+this was not the case the inverse scattering problem would be 
+quite simple.)
 
-\subsection{Gyroscopic Precession using asymmetric dimers}
-As mentioned in section~\ref{sec:off-axis} for specificity sized 
-dimers there is the potential for non-vertical trapping orientations 
-in which the dimer is still stably trapped. When a circularly 
-polarised beam is used the dimer exhibits gyroscopic precession. 
-As shown in fig.~\ref{fig:gyro} the dimer's trajectory exhibits 
-periodic rotation, both around its long axis and about the beam 
-axis.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
-	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped 
-		in an off axis orientation with a circularly polarised beam 
-		($P= 100\ mW$). The far left column depicts the dimer's centre 
-		of mass position with time; the remaining 3 columns show the 9
-		components of the dimers' rotation matrix, with each column being
-		associated with one of its three principal axis. }
-	\label{fig:gyro}
-\end{figure}
+To reduce the effects of the signal noise we took the time 
+average of the expected signal over $0.001 s$ and then had our 
+neural network estimate the orientation based on the average 
+signal.
 
-Applying a Fourier analysis to the above trajectory reveals 
-the 3 fundamental frequencies typically associated with 
-gyroscopic precession. Fig.~\ref{fig:gyro_diagram} gives 
-a representative idea of the motion seen in fig.~\ref{fig:gyro}.
-Firstly there is precession, is denoted by $\psi$, this
-causes the dimer to rotate in the $XY$ plane while maintaining
-a constant polar angle. The precession can be due to the dimer 
-having inhomogeneous polarisation susceptibility. Since its 
-long axis is more susceptible to being polarised the dimer will
-try to align its long axis with the polarisation vector 
-\cite{Bruce2020}. Second there is nutation, denoted by $\theta$, 
-which is due to the dimer's Brownian motion being influenced by 
-the 'Brownian vortex', as predicted by \cite{Ruffner2012}. And 
-lastly there is rotation, denoted by $\omega$, where the dimer
-spins around its long axis as it did in \ref{sec:rot_pol} \&
-\ref{sec:rot_size}.
-
-What is interesting is that while we see the precession and 
-nutation also contribute to the rotation frequency. As can be 
-seen in both fig.~\ref{fig:gyro} and fig.~\ref{fig:gyro_diagram}
-the rotation matrix components $U_{x,x}$, $U_{x,y}$, $U_{y,x}$
-and $U_{y,y}$
-
-\begin{figure}
-	\hspace{-2.3cm}
-	\includegraphics[width=1.2\linewidth]{gyroscopic_diagram.pdf}
-	\caption{Representative diagram of the gyroscopic precession 
-	from fig.~\ref{fig:gyro}. The dimer has three principal angular
-	frequencies: The rotation ($\omega$) occurs around the 
-	the dimer's long axis. Precession ($\psi$) is seen where the dimer 
-	rotates around the beam axis. Nutation ($\theta$) is due to the dimer's 
-	centre of diffusion orbiting the beam axis. Shown on the right
-	is the power spectra from fig.~\ref{fig:gyro} with the associated 
-	frequencies labelled. The power spectrum have been zoomed in on
-	the relevant frequencies to highlight the precise values.}
-	\label{fig:gyro_diagram}
-\end{figure}
-
-This gyroscopic motion has been demonstrated previously in 
-nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
-but has not been observed for micron scale aggregates. What
-is interesting that the precession is seen around the particle's
-optical axis (in the case of a dimer this would be its long 
-axis). Fig.~\ref{fig:gyro} instead shows the precession 
-occurring perpendicular to the optical axis. In the former 
-case the precession can be explained as simply occurring due 
-to the reaction torque from the surrounding fluid \cite{Zhu2021}.
-In their analysis of ellipsoidal particles, Zhu \textit{et al.}
-found that when the particle rotated around their minimal inertial 
-axis (similar to the observed behaviour in dimers), the particle 
-would quickly 'stabilise' by aligning its' optical axis with the 
-beam axis. This highlights that a spherical dimer is capable of 
-rotating about its minimal inertial axis while remaining stably
-trapped. This has potential implications beyond simply 
-understanding particle dynamics.
-
-Further analysis of the mechanism behind the precession of off-axis 
-dimers may provide insights into controlling Brownian motion. An 
-experimental work trying to 'cool' nano-dimers by controlling the 
-motion in all 6 degrees of freedom found that even while the 
-rotation about the short axis' could be controlled the free rotation 
-around the dimers' long axis resulted in an unpredictable torsional 
-vibration \cite{Bang2020}. Understanding how rotational motion 
-arises in the Mie-regime could allow researchers to build a robust 
-theoretical framework to construct beam structures that eliminate any 
-unwanted rotational motion from a target particle. Conversely, the 
-same framework could allow for precise measurements of the optical 
-torque applied to a target particle, allowing for characterisation 
-of complex shaped particles' interactions with an optical trap. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Characterisation of asymmetric dimers via PSD \\
-	analysis}
-As discussed in \ref{sec:simulated_QPD}, one of the methods 
-developed to work in conjunction with \cite{Vigilante2020} 
-is a simulated quadrant photo diode for as a position 
-detection system. While it is possible to extract all of 
-the relevant dynamical information from a simulation, 
-confirming the same behaviour in an experimental setting 
-can be challenging if dealing with a non-birefringent 
-anisotropic scatterer. The QPD is composed of 4 photodiodes 
-that measure the intensity of light incident on their 
-surfaces. Using \textit{ott} we can define a region in the 
-far-field that is analogues to the surface of the QPD surface. 
-Evaluating the electric field across this surface gives us 
-an approximation of the QPD signal outputted in an experimental 
-situation. This does not provide a one-to-one result however 
-as hardware errors (i.e. internal resistance, external light 
-sources, and vibrations) would distort the signal somewhat. We 
-can use this as a means to examine the limitations of using 
-back-focal plane interferometry as a means for characterising 
-anisotropic scatterers.
-
-As a benchmark we start by considering a single sphere within 
-an optical trap. A single polystyrene sphere suspended in water 
-($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
-Gaussian beam ($NA=1.25$) using circularly polarised light. For 
-the sake of time efficiency the trajectory was sampled every 10
-time steps, meaning the upper bound on the power spectra is
-$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
-we fitted the power spectra using the aliased Lorentzian 
-(Eq.~\eqref{eq:aliased_lorentzian}).
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=\linewidth]{PSD_sphere.png}
-	\caption{Recorded power spectra fitted to eq.~
-		\ref{eq:alaised_lorentzian}, scattered points represents 
-		the blocked data ($n_b=100$). Corner frequency for the 
-		Lorentzian curves are reported in the legend.}
-	\label{fig:psd_sphere}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11a.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11b.png}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\subcaption{}
+		\includegraphics[width=\textwidth]{fig11c.png}
+	\end{subfigure}
+	\caption{Model prediction for signal error of (a) $1\%$ 
+		$[\langle D_{KL}\rangle=1.447]$, (b) $15\%$ $[\langle D_{KL}\rangle=4.670]$, and (c) $25\%$ $[\langle D_{KL}
+		\rangle=7.911]$, time averaged over $1\,{\rm ms}$}
+	\label{fig:time average}
 \end{figure}
+This resulted in a reduction in the overall signal noise and 
+provided a higher degree of accuracy for our model. There 
+appears to be no clear correlation between the length over 
+which we time average and the performance of our model. Time 
+averaging over every 0.05s resulted in a drastically worse 
+performance; this is due to the fact that over longer time 
+periods there is greater uncertainty regarding how the dimer's 
+orientation has changed, thus tracking the instantaneous 
+orientation becomes harder for the neural network. Fortunately, 
+time averaging even over $1\,{\rm ms}$ seems to provide a 
+satisfactory estimation of the dimer's angular dynamics within 
+the optical trap.
 
-As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
-report different corner frequencies which would be indicate 
-that the trap is not perfectly circular. We can use both 
-\textit{ott} and the trajectory itself to derive an estimation 
-of the trap geometry. The corner frequencies and corresponding 
-trap stiffness are reported below:
+Overall it’s clear that estimation of the dimer’s orientation 
+is a problem that can be endlessly tuned to optimise the 
+result. Here we simplify the problem somewhat by employing 
+a relatively small finite number of 'reference orientations' 
+to map between scattering and dimer orientation: the precision 
+of estimation could be improved by utilising a greater number 
+of reference orientations, although there remains a balance 
+between the realisable precision of orientation estimate and 
+the noise level of the scattering measurement. 
+Another avenue to further explore would be using 
+the method to optimise the choice of detection angles, 
+essentially to find the region in the mapping between measured 
+scattering and orientation that others the best degree of 
+confidence through optimal separation of scattering signals 
+for distinct orientations. For sequences of data such as 
+dynamic measurements, a further potential enhancement would 
+be to consider more complex correlations based on prior 
+expectations of the dynamics. The Debye equation for  Here already we improve the 
+method using a non-uniform prior based on only the immediately 
+previous measurement in time (see Section 2.1): considering 
+a non-uniform grouping of reference orientations might result 
+in a better estimation, if we have information regarding the 
+dimer’s preferred axis of rotation.
 
-\begin{center}
-	\captionof{table}{QPD fitting for single sphere}
-	\begin{tabular}{ |c|c|c|c|c|c|c| } 
-		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
-		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
-		fitting}\\
-		\hline
-		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
-		& \parbox{1.25cm}{\centering 523} & 513 \\
-		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
-		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
-		\hline
-		Ellipticity &
-		\multicolumn{2}{|c|}{8.16 \%} &
-		\multicolumn{2}{|c|}{27.17 \%} &
-		\multicolumn{2}{|c|}{13.8 \%} \\
-		\hline
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+\label{sec:Conclusion}
 
-	\end{tabular}
-\end{center}
-
-Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
-where the variance is based upon our choice of frequency 
-boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
-the ellipticity of the beam is given by $e = (1-\kappa_y/
-\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
-wavefront. Its clear from these initial results that the QPD 
-is more sensitive to changes along the y-axis than the x-axis 
-when compared to the direct \textit{ott} calculations. This is 
-somewhat reflected in the trajectory results. Typically, even 
-an industrial Gaussian beam will produce an elliptical diffraction 
-limit spot when heavily focused; in their tutorial for optimizing 
-the PSD analysis, Berg and Sorensen reported a ellipticity of 
-around 15 \% after a total calibration time of 80 seconds 
-\cite{BergSoerensen2004}.
-
-The reason for the discrepancies between all 3 methods is due 
-to what is actually being measured. The \textit{ott} estimates 
-are simply looking at the differences in the trapping strength 
-along the Cartesian axis'. If calibrated over an long enough 
-time frame you would expect that the resulting power spectra 
-would exactly mirror the \textit{ott} predictions. However over 
-a short calibration time a QPD will only sample a small trajectory 
-that does not fully explore the trapping plane. There is a clear 
-trade off in terms of accuracy and computation time as shorter 
-calibration runs are computationally more efficient but prone to 
-errors. This is the case even if we know the exact positional data.
-
-With this in mind, let us consider a symmetric dimer that is 
-optically trapped by the same Gaussian beam. Not only does 
-the dimer's equilibrium position change but it is subjected 
-to rotational motion due to its unequal moments of inertia. 
-This is reflected in the calibration results using the simulated 
-QPD, where we see a drastically different estimation between the 
-\textit{ott} estimate and the QPD estimate.
-
-\begin{center}
-	\captionof{table}{QPD fitting for symmetric dimer}
-	\begin{tabular}{ |c|c|c|c|c|c|c| } 
-		\hline
-		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
-		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
-		\hline
-		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
-		& \parbox{1.25cm}{\centering 274} & 285 \\
-		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
-		\hline
-		Ellipticity &
-		\multicolumn{2}{|c|}{42.8 \%} &
-		\multicolumn{2}{|c|}{12.7 \%} & 
-		\multicolumn{2}{|c|}{13.8 \%} \\
-		\hline
-	\end{tabular}
-\end{center}
-
-Now we see that the \textit{ott} predicts a more elliptical 
-trap compared to the QPD model which says the trap is far
-more symmetrical while trapping a symmetric dimer. A potential 
-reason that \textit{ott} no longer expects a circular trap 
-could be due to how it computes the beam shape coefficients; 
-by point matching in the far field before the focus means a 
-loss in accuracy for objects that trap above the focus. 
-The change in the QPD estimation can be partially explained 
-by the fact that rotational effects are not accounted for in 
-the Lorentzian power spectra, only translational motion. 
-Typically, rotational motion is only ever detected when it 
-is periodic (take for example Fig.~\ref{fig:vaterite}), when 
-the motion is stochastic the entire power spectra is effected 
-making it near impossible to separate the translational and 
-rotational contributions from a single power spectra.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-Considering the simplicity of a scatterer such as a dimer, one would 
-assume that the dynamics of such an object would be relatively easy 
-to predict. Simulations of dimers in the Mie regime show that not 
-only do they have multiple positions and orientations in which they 
-can be trapped but also that their interaction with circularly 
-polarised light is heavily dependent on the axial position and 
-trapping orientation.
-
-Dimer's have the potential to be used as tunable micro-rotors, being 
-simple to synthesise and can be made out of any material of choice.
-The rotation demonstrated by dimer's is not accurately described in
-previous literature which raises questions on the interactions between
-circularly polarised light and spherical aggregates. Until now, our 
-understanding of angular momentum transfer has been either focused on 
-single particles (where momentum transfer is easily described by 
-Lorenz-Mie theory) or for large aggregates of particles, where the 
-angular momentum transfer between individual particles is not considered. 
-This could provide a greater understanding of angular momentum and help in 
-the development of better torque sensing or torque preventing methods.
+We have developed a method for measuring the dynamics of an 
+optically-trapped colloidal objects based purely on 
+measurements of the object's light scattering at a small 
+number of detection angles. We demonstrate the method using 
+the orientation of an asymmetric dimer as the dynamic variable 
+and object of interest respectively, but in principle the 
+model can be applied to any characteristic that impacts the 
+light scattering pattern produced by a trapped entity such 
+as size and shape. The MSTM package is a flexible tool for 
+calculating the light scattering of complex objects using a 
+representation of the object as a set of micro-particles, 
+enabling training of a neural network to enable categorisation 
+of the mapping between scattering and trapped object 
+characteristics. By taking account of the physically realistic 
+behaviour of the trapped object and the characteristics of 
+the trap (which impact the dynamics of the object), the 
+Bayesian inference method can be re ned to provide a reliable 
+estimation of object characteristics of interest, even in the 
+presence of measurement noise. Fundamentally, the inverse 
+scattering problem is difficult to solve, since the mapping 
+between object characteristics and scattering can be highly 
+complex. We determined the minimum number of detectors 
+required for a reliable estimation in the presence of 
+measurement noise; furthermore, we demonstrated that the 
+arrangement of these detectors is critical for a reliable 
+estimation of an objects orientation. However, Bayesian 
+inference based on neural network estimation of the mapping 
+provides a powerful method for practical applications, 
+extending the use of optical trapping beyond measuring 
+microscopic force response toward detailed structural and 
+dynamic information about complex trapped entities.
 

### 2025-06-27 Latex_files/Thesis/chapters/chapter 5/main.tex
27/06/2025 chapters 1,3,4 and structural changes

@@
-\chapter{Detection of rotational spherical aggregate rotational dynamics}
-\label{chapter:simulated_detection}
-As outlined in the end of Chapter 4, one of the difficulties 
-in characterising interactions with asymmetric objects is 
-the coupled motion between translation and rotation. In order
-to characterise the optical trap a position detection system 
-such as a lateral effect photodiode or a quadrant photo diode
-(see fig.\ref{fig:position_detectors}). Typically, a position 
-detection system assumes that there exists a linear relationship 
-between the a particle's displacement and the detected signal.
-However, this is not always the case as demonstrated in chapter
-4; as dimer's demonstrate non-harmonic trapping beyond a certain 
-displacement. In addition, the rotational effects of the trapped
-dimer become a more significant factor when characterising the 
-trap strength. This can have unintended effects when it comes to
-areas of research that require a precise understanding of the 
-hydrodynamic behaviour of particles. As such there has been a 
-concerted effort to either eliminate rotational motion, or describe
-how the rotational behaviour influences the trajectory reported 
-by the position detector.
+\chapter{Simulated Quadrant Photodiode}
+\label{chapter:simulated_QPD}
 
-In this chapter we build upon the work of \cite{Rohrbach2002} and
-\cite{BarZiv1997}; the former used to measure the relative position
-of the trapped aggregate, and the latter to measure the orientation
-of the aggregate. This was initially developed in tandem with the 
-work conducted in chapter 3, with the idea being that if a crystal 
-nucleus formed on a rotating sphere how would that nucleus influence 
-the motion of the rotating sphere. Using static light scattering we 
-can map the outputted signal from a optically trapped dimer to its 
-expected orientation in real time. Thus allowing us to estimate the 
-optical torque being applied to the dimer by the trapping beam. As demonstrated in chapter 4, the influence of a circularly polarised 
-beam can result in gyroscopic precession for specifically sized 
-dimers.
+\section{Overview}
+In this chapter, we begin with a brief discussion of the working 
+principle behind a position detection device. We then provide a 
+framework for replicating the scattering signal recorded by a 
+quadrant photo diode. We show that extending the framework to 
+capture information from an arbitrary shaped particle is simple 
+in principle but difficult to interpret, we demonstrate as such 
+by comparing the expected signal from a trapped sphere, to that 
+of a trapped dimer.
 
-We start by describing a quadrant photo diode system simulating the 
-power spectrum derived from a particle trajectory. We demonstrate 
-that even in idealised circumstances the rotational motion is not 
-revealed from the collected power spectrum. We then go about 
-defining an optical fibre detection system from a mathematical 
-perspective. Before demonstrating how it can be used to 
-instantaneously predict orientational information on a trapped 
-particle. We then consider common factors of error in the 
-characterisation process, such as signal error, and incorrect particle 
-sizing. Both of which have a significant impact on the performance 
-of the model, but by utilising Bayesian inference, and time averaging 
-we can better refine our model.
+\section{Introduction}
+
+Beyond merely simulating the forces experienced by a spherical 
+aggregate we also developed a simulative quadrant photo diode 
+to replicate the results from a typical calibration test. This 
+builds upon the work from \cite{Rohrbach2002} which applied 
+Lorenz-Mie theory to replicate the response signal of a QPD 
+being used in back focal-plane interferometry. Rather than be 
+constrained to individual spheres we can now consider the 
+expected response from any type of spherical aggregate. The
+goal of such simulation software would be to simulate the 
+expected experimental results from a complex particle whose 
+exact shape and size may be difficult to determine.
+
+In order to simulate a typical experimental set up with a QPD 
+installed as a position detection system we need to evaluate 
+the total magnitude of the electric field incident on the 
+photo-diode surface. While trapping a micro-particle, the 
+scattered and incident fields combine together and interfere 
+with one another. These fields are collected by a condenser 
+lens in the far field limit and are focused onto the QPD 
+surface, the total intensity can be evaluated as:
+\begin{align}
+	I(x,y) = \epsilon_0c\left|
+	\begin{bmatrix} 
+		E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
+		E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
+		E_{i,z}(x,y)+E_{s,z}(x,y)
+	\end{bmatrix} \right|^2 \times step(\theta_{NA}-\theta(x,y))
+\end{align}
+
+The last term is simply a representative step term that defines 
+the outer limit by which we evaluate the electric field, this is 
+analogous to our condenser lens removing noise from other light 
+sources by only accepting light at a specific acceptance angle 
+defined by its numerical aperture $NA_c$. Depending on the 
+relative size of our particle we can adjust the acceptance angle, 
+this has very little effect on the transverse signals, but for 
+axial evaluations of a trapped particle the numerical aperture 
+should be tuned so that the resultant response curve has negative 
+slope in order to allow for axial position detection, the method 
+for finding this angle $\theta_\Theta$ is discussed in 
+\cite{Friedrich2012}.
+
+The incident beam is simple enough to define given our set up 
+parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
+beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
+\textit{Ott} uses a point matching approach to approximate the beam shape 
+coefficients of the incident field by fitting it to the far field estimate. 
+From the QPD's perspective it is receiving light both from the incident 
+and scattered beam simultaneously, as such both fields must be expressed
+using outgoing vector spherical harmonics.
+\begin{align}
+	E_{\rm inc}(r)&=
+	E_0\sum^\infty_{n}\sum^n_{m=-n}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
+	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
+	\\
+	E_{\rm scat}(r)&=E_0 \sum^\infty_n \sum^n_{m=-n} \left[
+	p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
+	\right] 
+\end{align}
+
+Where the superscript (2) denotes an outgoing spherical harmonic
+function. In order to compute the scattering from the target particle \textit{ott} uses the $T$-matrix method, this is not essential 
+for a simple sphere but is essential for complex shaped particles 
+such as dimers. The scattered and incident fields are then combined 
+together in the far field to get $I(x,y)$, the quadrant and overall 
+signals are calculated via:
+\begin{align}
+	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
+	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
+	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
+	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
+\end{align}
+
+Where the denominator is the total intensity on the QPD while there is 
+no particle within the trap. You would expect that $S_x$ and $S_y$ are
+near identical for equal particle displacements, however due to the 
+polarisation of the beam there will be a slight bias to the signals 
+for motion along the direction of the polarisation vector. This is 
+generally not an issue if the particle trajectory is sampled for a 
+long enough time period. By converting from signal units to length
+units (see \eqref{eq:conversion_factor}) the trap shape can be 
+discerned from the QPD signals, assuming the fluid properties are
+known to a high degree of accuracy.
+
+While translational motion has no knock on effects to the QPD signal, 
+rotational changes are double counted by both the incident and total 
+field. This results in rotational motion being biased in the QPD signal 
+- even when collecting signals from isotropic scatterers. To prevent 
+this we need to rotate the total field via the inverse rotation matrix
+of the dimer.
+
+To confirm that our method is producing accurate results, we ran a comparison
+between our simulative QPD and the results from \cite{Rohrbach2002}. Where a 
+$300\ nm$ diameter sphere is scanned across the path of a focused Gaussian beam
+($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is
+suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
+set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
+produced the following response curve:
+\begin{figure}[h!]
+	\begin{subfigure}{0.475 \linewidth}
+		\subcaption{}
+		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
+	\end{subfigure}
+	\begin{subfigure}{0.475 \linewidth}
+		\subcaption{}
+		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
+	\end{subfigure}
+	\caption{Comparison between QPD response signal versus work conducted by 
+		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal 
+		produced by QPD using \textit{ott} and points represent the signal 
+		response collected from \cite{Rohrbach2002}.}
+	\label{fig:Rohrbach}
+\end{figure}
+
+The discrepancy between our simulated QPD and the results 
+from Rohrbach can be attributed to the fact that the position 
+to signal error grows as you move further from trap focus. 
+As the particle moves further from the trap focus the error 
+in the attributed signal grows substantially \cite{Rohrbach2002}.
+In most optical trapping experiments we usually are calibrating 
+a strong trap where the mean square displacement falls well within 
+the linear regime shown in fig.~\ref{fig:Rohrbach}. In which case 
+the error between Rohrbach's model and our own is inconsequential 
+for calibration purposes. With this any trajectory can be collected 
+from the QPD by displacing and rotating the beam accordingly, 
+later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of 
+accuracy that can be achieved using back focal plane interferometry 
+to characterise non-spherical targets.
+
+\subsection{Position detection}
+\label{sec:position_detection}
+In order to accurately capture the dynamics of a trapped particle, a 
+position detection system is required. There are 3 possible methods 
+of position detection: video-analysis, lateral-effect position sensing, 
+and photodiodes.
+
+Video analysis is ideally suited for multiple traps or situations where 
+precision is not the top priority. Whereas lateral-effect and photodiode 
+position detection are two examples of back-focal plane interferometry, 
+where the interference pattern produced by the target particle is 
+extrapolated to determine its position. In order for video analysis to 
+match the accuracy of back-focal plane interferometry methods requires 
+the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
+achieve while maintaining a decent resolution \cite{Gibson2008}. In 
+comparison off the shelf back-focal plane detectors can achieve temporal 
+resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
+
+A lateral-effect sensor has a similar output but works using a the 
+entire sensor as a single cell analogous to the focal plane of the 
+trapping beam. The four corners of the sensor act as anodes connected 
+to a base plate cathode, as the beam moves across the surface of the 
+detector each anode will experience a different photocurrent depending 
+on how close the centre of the interference pattern is to each anode. 
+The advantage of a lateral effect detector is that the linear regime 
+is much larger than a QPD making it much better for monitoring the 
+position of a trapped particle. However, Lateral-effect sensors are 
+often limited in their spacial resolution due to high signal-to-noise 
+ratios, requiring a high intensity of light on the sensor in order to 
+get a clean signal. As a result, most optical force measurements are 
+conducted using a QPD as opposed to a lateral-effect sensor, as often 
+the displacement is small enough that the signal-displacement curve 
+can be considered linear.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
+	\caption{Comparison between QPD and Lateral effect photodiodes.
+		The four quadrants of a QPD (left) experience different photocurrents
+		based on the total intensity of light incident on each section 
+		(labelled A, B, C, D). 
+		Whereas a Lateral effect sensor (right) uses the resistive properties
+		of the photodiode surface to vary the create different photocurrents 
+		passing through the anodes A, B, C, and D.}
+\end{figure}
+
+A quadrant photo diode (QPD) is a frequently used position detection 
+system for optical tweezers due to their high sampling rate, high 
+degree of precision, and ease of set up. The QPD is constructed of 
+four photo diodes assembled in a quadrant formation, when a particle 
+is trapped the interference pattern produced is focused onto the QPD, 
+with the maximum intensity mapping to the particle's centre of mass. 
+By summing the voltages of the horizontal and vertical quadrants together 
+the particle's centre of mass is tracked in the $x-y$ plane. Axial 
+displacement can be estimated by observing the change in the total 
+voltage of the QPD. The outputted signal gives an indication of the 
+particle's relative displacement from the beam focus, but in order to 
+convert the signal to distance units the trap needs to be calibrated 
+(assuming a linear response curve).
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{rotation_figure}
+	\includegraphics[width=0.9\linewidth]{rotation_figure.png}
 	\caption{QPD signals detected by a dimer being rotated while
@@
 while restricting translational motion.
-
-\section{Monitoring Stochastic rotational motion using static 
-		light scattering}
-Part of the optical fibre arrangement borrows code from the 
-optical tweezer simulations discussed in Chapter 2 and 4. 
-Rather than repeat the simulation details we will instead 
-only focus on the specifics of the optical fibre scattering.
-
-\subsection{Coordinate System}
-Consider a particle (either a sphere or dimer) trapped by
-a Gaussian beam, with its focal point being set at [0,0,0]. 
-At the same time an optical fibre directs a plane wave that 
-is incident on the trapped particle scattering light in all 
-directions. The probe beam is assumed to be x-polarised and 
-propagating in the +z direction, this is a $90^{\circ}$ 
-rotation from the previous simulations (see fig\emph{X.X}).
-
-\subsection{Dimer}
-The dimer is defined by two spheres with refractive indices
-of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
-requires a target origin for setting the scattering expansion, 
-there is no agreed upon solution for where the target origin 
-should be relative to the sphere positions. In chapter 4 we showed
-that altering the target origin would drastically alter the 
-tmatrix produced. We choose (somewhat arbitrary) to use 
-the dimer's centre of diffusion which is located $\zeta a_I$ from 
-the centre of the larger sphere. For a discussion see \emph{X.X}.
-
-\subsection{Scattering and Detection}
-Each detector is placed some distance $\bf r$ from the origin, 
-detector positions are defined by angles $\theta$ \& $\phi$. With 
-$\theta=\phi=0$ being directly in front of the probe beam. The 
-position of each detector is given by:
-\begin{align}
-	\begin{pmatrix}
-		x_{fiber} \\ y_{fiber} \\ z_{fiber}
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		rcos(\phi)sin(\theta) \\ rsin(\phi)sin(\theta) \\ rcos(\theta)]
-	\end{pmatrix}
-\end{align}
-Where r is the distance between the detector and the target origin. 
-In an ideal situation the surface of the detector is oriented towards 
-the target origin perfectly. Pixels can be thought of as points lying 
-on the surface of the detector, each point has its own position in real
-space. Using \textit{mstm} we can find the components of the scattering 
-matrix at each point and compute the intensity of the electric field via:
-
-\begin{align}
-	\begin{pmatrix}
-		I_s \\ Q_s \\ U_s \\ V_s
-	\end{pmatrix} = 
-	\begin{pmatrix}
-		S_{11} & S_{12} & S_{13} & S_{14} \\
-		S_{21} & S_{22} & S_{23} & S_{24} \\
-		S_{31} & S_{32} & S_{33} & S_{34} \\
-		S_{41} & S_{42} & S_{43} & S_{44} \\
-	\end{pmatrix}
-	\begin{pmatrix}
-		I_i \\ Q_i \\ U_i \\ V_i
-	\end{pmatrix}
-\end{align}
-Where:
-\begin{align}
-	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2  \label{eq:intensity} 
-	\\
-	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
-\end{align}
-
-By computing the scattering at each point on the detectors surface 
-we can compute the average scattering over the detector. From a
-experimental perspective the distance from the particle to the 
-detectors must be considered. The near field scattering can be 
-computed with relative ease using \textit{mstm} as shown in 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{near_field_intensity_side_profile.png}
-	\caption{Intensity distribution (as defined by \eqref{eq:intensity}) 
-		from a symmetric dimer when irradiated by the probe beam 
-		propagating in the $k\hat{z}$ direction (viewed in the 
-		z-y plane). Axis are given in dimensionless units of 
-		$k\hat{z}$ and $k\hat{y}$.}
-	\label{fig:nf_scattering}
-\end{figure}
-
-While \textit{mstm} does provide the amplitude matrix it is not
-scaled to account for distance from the target to the detector.
-We amend this by simply scaling according to the inverse square 
-law. As the detector is moved further from the target the 
-intensity distribution across the entire detector will decrease
-substantially. Because of the fall off it is more computationally
-efficient to instead consider only the central pixel of the 
-detector rather than across the entire surface. If we consider 
-the difference between the average intensity and the intensity 
-of the central point we can see that after not even $20 \mu$ 
-away from the target particle the average intensity across the 
-surface is within $0.1\%$ of the central pixel. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
-	\caption{Normalised difference between the central pixel intensity
-	$I(\theta, \phi)$ and the average pixel intensity over the surface
-	of the optical fibre $\langle I \rangle$ as a function of distance.
-	4 different optical detectors were tested (spherical coordinates 
-	are given in the legend). While the choice of coordinates is 
-	arbitrary they all demonstrate the same exponential decay. Dashed
-	lines indicate the difference after $20\mu m$ from the trap focus,
-	in most practical cases getting an optical fibre any closer would
-	disrupt the trap stability.}
-\end{figure}
-
-Given the trapping beam's beam waist is approximately $0.54 \mu m$,
-a optical fibre would need to be a minimum of $150 \mu m$ from the 
-trap centre in order to prevent noise from the trapping laser itself.
-At such a distance the variation in the signal is so small that it is 
-easier to evaluate the intensity at the centre of the detector instead
-of across the entire surface. This is how the instantaneous orientation
-is predicted later on in the chapter.
-
-\section{Interpretation of scattering data into orientation estimates}
-We now discuss the methodology for computing the instantaneous 
-orientation of an optically trapped particle based on scattering
-detected by a finite grouping of optical fibres.
-
-Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), 
-we can define at any point in time a unit vector $\hat{s}$ 
-pointing from the centre of the larger sphere to the centre 
-of the smaller sphere. A plane wave probe beam incident, 
-is incident on the dimer, generating a scattering pattern
-dependent on the dimer's orientation $I(\hat{\bf s}, 
-\theta, \phi)$ which is computed using \textit{mstm} 
-\cite{I.Mishchenko1996}. To represent the experimental set 
-up consisting of a set of optical fibres recording scattered 
-light, we choose four sets of spherical angles [$(\theta_1,\ 
-\phi_1), \ (\theta_2,\ \phi_2), \ (\theta_3, \ \phi_3), \ 
-(\theta_4,\ \phi_4)$] and record the calculated intensity at 
-each angle $I(\hat{s},\ \theta_k,\ \phi_k)$. 
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.49\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig2a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig2b.png}
-	\end{subfigure}
-	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
-	\label{fig:dimer}
-\end{figure}
-
-Our goal is to determine the orientation of the trapped dimer 
-based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
-Rather than aim an exact estimate of the dimer's orientation, 
-for the purposes of interpretation of the scattering and 
-optimisation of the measurement setup it is more convenient 
-to discretize the possible orientation space into a number 
-of possible reference orientations. We can then use each as 
-'classification categories' in a neural network methodology 
-to map scattering data to orientation (see below for further 
-discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
-reference orientations $\hat{\bf n}_{\alpha}$  evenly 
-distributed on a unit sphere \cite{Reyuthor2006} 
-(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
-neighbour spacing between two neighbouring reference 
-orientations of 0.895 radians.
-
-Using \textit{mstm} we compute the raw intensities (see 
-\eqref{eq:intensity}) at each of the measurement angles 
-that would be generated by a dimer in each reference 
-orientation, $I_s(\hat{\bf n}_{\alpha}, \theta_k)$. While 
-the number and position of detection fibres is technically 
-arbitrary there are several constraining factors that limit 
-our ability to infer useful information from the trapped 
-object, see Section~\ref{sec:brownian} for a detailed 
-breakdown of our choice of detection angles. In order for 
-a neural network to properly weight each signal the raw 
-intensities are normalized according to:
-\begin{align}
-	\label{eq:scale}
-	y_k(\hat{\bf n}_\alpha)
-	= 
-	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
-	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
-\end{align}
-
-where the denominator is simply the standard deviation across 
-the set of values $I(\hat{\bf n}, \theta_k)$. Note that the 
-collected scattering signals are not necessarily simply 
-related to their associated reference orientations: as is well 
-known from such examples of the inverse scattering problem. 
-While it is trivial to compute the light scattering pattern 
-for any given particle with any particular characteristic (i.e. 
-size, shape, or orientation), inferring the light scattering 
-from a unknown particle to determine said characteristic is 
-incredibly difficult due to complex mapping between scattering 
-and said characteristic. Even if the orientation space is 
-divided evenly between reference orientations the subsequent 
-signal space ends up being appearing mixed making simple 
-comparisons of signals useless for inferring information on the 
-particle.
-
-Fig~\ref{fig:mixing} shows two clusters of orientation 
-vectors and their respective signals collected from 3 detectors
-- the points have been coloured based on their proximity 
-to the centre of their respective cluster. While the 
-orientation space appears tightly packed and ordered the 
-signal space quickly spreads out in an asymmetric fashion. 
-Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
-mapping can intersect itself which only further increases 
-the complexity. While in some instances the mapping between 
-one reference orientation and another is discrete, in other 
-instances the mapping becomes far more complex to discern.
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.74\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig3a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.74\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig3b.png}
-	\end{subfigure}
-	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals.
-Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
-	\label{fig:mixing}
-\end{figure}
-
-Nevertheless, at least where the uncertainty in signal measurements 
-is low, we can predict the orientation from the scattering scattering 
-by utilising computational techniques such as neural networks. We 
-thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
-dimer's orientation from its light scattering signal. The network 
-was trained by generating a database of random orientation vectors, 
-calculating the corresponding light scattering signals, and then 
-using the network to estimate the probability of a given signal coming 
-from a dimer in a given reference orientation. The network's loss 
-function was evaluated and used to improve the estimation, the 
-network being trained until the improvement in the loss function 
-was less than 0.0001.
-
-Importantly, the estimation provided by the neural network can be 
-improved further by accounting for any prior information we know 
-about the dimer, utilising Bayesian inference to update the neural 
-network's estimation: 
-\begin{align}
-	\label{eq:bayes}
-	p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
-	&=
-	\frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
-		p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
-\end{align}
-
-where $p(\hat{\bf n}_\alpha)$ and $p(y_k(\hat{\bf s}))$ are the prior 
-estimates of the distributions of particle orientations and 
-instantaneous signals, respectively. \textit{Without} any prior 
-evidence we must assume that the orientation prior of the dimer 
-$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
-dimer's possible current orientation from knowledge of previous 
-measurements can be used to inform our estimate of $p(\hat{\bf 
-n}_{\alpha})$. The latter prior $p(y_k(\hat{s}))$ is the probability of 
-measuring a signal ($y_1$, $y_2$, $y_3$). This is given by taking 
-the discrete integral over the collection of reference orientations:
-\begin{align}
-	p(y_k(\hat{\bf s}))
-	=
-	\sum_{\alpha=1}^{n_{\rm ref}}
-	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
-	p(\hat{\bf n}_\alpha)
-\end{align}
-
-From \eqref{eq:bayes} we obtain the key result, a mass probability 
-distribution denoting the probability that our dimer is in 
-orientation $\hat{\bf n}_{\alpha}$ given a measured signal ($y_1$, 
-$y_2$, $y_3$), \textit{i.e.} an estimated mapping from scattering 
-measurement to orientation estimate.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Calculation of error}
-\label{sec:divergence}
-Using \eqref{eq:bayes} we get a snapshot estimation of the instantaneous
-orientation of the trapped dimer based on a single measurement from a 
-collection of detector signals. Of course in reality these detectors 
-will be returning a discrete time series representative of the dimer's
-trajectory within the optical trap.
-
-Using the simulation code from \cite{Vigilante2020} we can replicate 
-this trajectory and furthermore use \textit{mstm}to calculate the far field 
-intensity produced by said dimer. At each time step we measure the signal
-produced at each measurement angle and use our neural network to calculate
-$p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha, t)$.
-
-Evaluating how well Eq.~\eqref{eq:bayes} performs in predicting $\hat{\bf n}$
-is difficult because we have discretised the orientation space meaning 
-there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method provides a measure of the relative
-entropy change between two probability distributions:
-\begin{align}
-	D_{KL}(P(t)\parallel Q(t)) = \sum P(t) \log_2\left[\frac{P(t)}{Q(t)}\right] 
-	\label{eq:D_KL}
-\end{align}
-
-Where $P(t)$ and $Q(t)$ are two probability distributions we are
-comparing, often it is common to use a base 2 log as this converts 
-the entropy to base 2. By relative entropy we are talking about 
-the useful information gained/lost by switching from $P(t)$ to 
-$Q(t)$. In our case $Q(t)$ is simply the distribution given by
-Eq.~\eqref{eq:bayes} and $P(t)$ is an idealised distribution, 
-one where the model is accurate with 100\% certainty.
-\begin{align}
-	\label{eq:best}
-	p_{best} = P(t) =
-	\begin{cases}
-		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
-		0 & \text{when $\hat{\bf n}_\alpha$ $\neq$ $\hat{\bf n}_{best}$}
-	\end{cases}
-\end{align}
-
-In reality the distribution from Eq.~\eqref{eq:bayes} will assign 
-some non-zero probability to every reference orientation. Therefore 
-the relative entropy for a single measurement is given as:
-\begin{align}
-	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
-	= \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
-	\right]
-	\label{eq:kullback}
-\end{align}
-
-The relative entropy is a measure if how spread out our distribution
-is. With larger values of $D_{KL}$ indicating that our prediction is 
-spread out so the neural net is unsure about $\hat{\bf n}_{best}$. 
-Whereas a low value of $D_{KL}$ means that our model is more confident 
-in $\hat{\bf n}_{best}$. The only draw back being that the relative 
-entropy does nothing to indicate accuracy, only how closely our 
-distribution matches our ideal result. This will be better 
-exemplified later in \ref{sec:test}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Brownian Simulation}
-\label{sec:brownian}
-
-We use the Brownian OT package developed by Fung~\textit{et~al} 
-\cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. We simulate the 
-motion of a dimer trapped in a highly focused Gaussian beam by 
-calculating the optical forces imparted by the laser, and the 
-Brownian force due to the surrounding fluid. We simulated a 
-polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} 
-= 1.33$). As discussed in chapter 4 dimers can be stably trapped
-in off-axis orientations. We simulate a dimer (of size ratio 2) that
-is trapped in this off-axis orientation by a $5 mW$ laser, due to 
-the meta-stability of the trap the dimer is only trapped for second
-before it escapes the trap. The dimer's position and orientation 
-is shown below:
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width =\textwidth]{fig7a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.45\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig7b.png}
-	\end{subfigure}
-	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
-	\label{fig:motion}
-\end{figure}
-
-At the same we use \textit{mstm} to replicate the intensity of light
-incident on the surfaces of detectors placed around the focus of the
-optical trap. The exact number of detectors was initially assumed to 
-be arbitrary, in that it made no difference to our estimate whether 
-we used 2 angles or 200.
-
-\subsection{Choice on the number of detectors}
-When all of the detectors lie in the same plane the expected signal 
-can appear identical despite the dimer being in completely different orientations. It should be noted that these pairs are reflected in 
-one or more axis which suggests that these are due to the arrangement 
-of our detectors. More specifically, if the detectors are placed say 
-in the x-y plane then only when the dimer is pointed nearly fully 
-upright will the expected signal be entirely unique. This is not an 
-issue with typical analytical static light scattering experiments 
-because the signal being collected is averaged over a population of 
-particles all in different orientations. Whereas in our set up we are concerned with the scattering of a single particle whose scattering is orientationally dependant.
-
-To remedy this we raise the third detector out of the x-y plane; 
-as such the expected signals from each reference orientation is 
-unique, though with limited resolution. By adding a $4^{th}$ 
-detector we can differentiate signals more reliably, improving the 
-neural networks performance. In line with our goal of making this 
-method viable in a laboratory setting we decided not to increase 
-the number of detectors further than 4.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Testing the Neural Network}
-\label{sec:test}
-We applied Eq.~\eqref{eq:bayes}, taking the reference orientation 
-with the highest probability  as our estimate of the dimer's 
-instantaneous orientation $\hat{\bf n}_{est}$. To visualise the 
-model's performance we plotted the orientation components of our 
-estimation $\hat{\bf n}_{\alpha}$ and the dimer's \emph{actual} 
-instantaneous orientation $\hat{\bf s}$ versus time. For comparison, 
-we also plotted the orientation components of the closest reference orientation, $\hat{\bf n}_{best}$. Assuming a uniform prior of the 
-reference orientations $p(\hat{\bf n}_\alpha)$ the neural network's predictions ($\hat{\bf n}_{\alpha}$ from Eq.~\eqref{eq:bayes}) are 
-at times reasonable, but there are significant large and random jumps 
-away from the correct result (Fig.~\ref{fig:uniform}). 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\textwidth]{fig5.png}
-	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity.
-Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the blue line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$.
-Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
-	\label{fig:uniform}
-\end{figure}
-
-One reason we observe such large jumps in the estimated 
-orientation is simply because the neural network is assuming
-that the signals are unrelated from one another. We can 
-observe this if we take the average relative entropy change
-across the entire trajectory:
-
-\begin{align}
-	\langle D_{KL}\rangle = \frac{1}{N} \sum_{i=1}^{N} \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha\parallel y_k(\hat{s}),t_i)}\right]
-	\label{eq:score}
-\end{align}
-
-Where we get an average entropy change of $\approx 0.604$, in 
-base 2 this means that on average the model has narrowed down 
-its options from 1 in 30 to 1 in 1.52. This is not reflected
-though in fig.~\ref{fig:uniform} which indicates that while 
-the model is highly confident in its results it is not at all 
-accurate. As demonstrated by fig~\ref{fig:mixing} the 
-orientation and signal spaces are uncorrelated with one another 
-which results in the neural network to assign orientations 
-incorrectly. Combining this fact with use of a uniform prior 
-($p(\hat{\bf n}_\alpha)$), there is no constraint on how much 
-estimated orientation can change from time-step to time-step.
-
-To improve the estimation we can therefore use knowledge of the 
-physical limitations of the object in the trap and its dynamics. 
-Imposing a more physically grounded prior, accounting in this case 
-for the fact that the rotational motion of the dimer is limited by 
-the rotational trap stiffness ($\kappa_r$). Here the prior of the 
-reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at 
-each time step as a Boltzmann distribution of the physical distance 
-between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and 
-each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are 
-reweighing our estimation based on the size of rotation, with smaller movements being favoured over large movements:
-\begin{align}
-	p(\hat{\bf n}_\alpha)
-	&=\frac{e^{\hat{\bf n}_\alpha 
-			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
-	{\sum_{\alpha=1}^{n_{\rm ref}}
-		e^{\hat{\bf n}_\alpha 
-			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
-	\label{eq:boltz}
-\end{align}
-
-As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.
-However at the same
-time result is still subpar.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth]{fig8a.png}
-	\caption{\label{fig:biased}
-		%
-		Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
-		by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
-		achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
-		Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
-		%
-	}
-\end{figure}
-
-Once again using Eq.~\eqref{eq:score} we now get a result of 
-$\approx 0.204$ or a 1 in 1.2 result. This is only a slight 
-improvement in terms of confidence but has drastically reduced 
-the variation in our models predictions. There are other 
-improvements that can be made, which additionally also help with
-potential sources of error in our set up.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Accounting for sources of error in light scattering measurements}
-When it comes to analysing light scattering from any size particle,
-error analysis becomes a significant factor. Typically this can be 
-accounted for by averaging over long periods of time to get an 
-assessment of the steady state conditions of the target particle. 
-However in our case where we wish to know the instantaneous 
-orientation, we instead have to rely on our understanding of how 
-uncertainty can effect our model's performance. We identified two 
-areas which are likely sources of error in our estimation: firstly, 
-an incorrect modelling of the target particle, and secondly, signal 
-noise arising from experimental factors. We highlight how we address 
-these areas below. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Impact of incorrect dimer sizing}
-\label{sec:lam}
-
-One of the main limitations of our model is that we assume that 
-the dimer being modelled in \textit{mstm} is accurate to the 
-dimer being trapped in the optical tweezer. Sizing molecules 
-accurately is a significant challenge for single particle 
-analysis so there is bound to be some uncertainty with the 
-measurements. We ran our model 3 times with the neural net being 
-trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.33\textwidth}
-		\subcaption{}
-		\includegraphics[width =\textwidth]{fig8a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.31\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig8b.png}
-	\end{subfigure}
-	\begin{subfigure}{0.31\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig8c.png}
-	\end{subfigure}
-	\caption{Model estimates of orientation when neural net has 
-		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.519]$, (b) 1:2.05 $[\langle D_{KL}
-		\rangle=3.706]$, (c) 1:1.95 $[\langle D_{KL}\rangle=
-		3.705]$ ($n_{refs} = 30$)}
-	\label{fig:size}
-\end{figure}
-
-As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model.
-This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%.
-This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact.
-We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Reyuthor2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}.
-While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.31\textwidth}
-		\subcaption{}
-		\includegraphics[width =\textwidth]{fig9a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.31\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig9b.png}
-	\end{subfigure}
-	\begin{subfigure}{0.31\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig9c.png}
-	\end{subfigure}
-	\caption{Model estimates of orientation when neural net has 
-		been trained on dimer of size ratio: (a) 1:2 $[\langle D_{KL}\rangle=0.593]$, (b) 1:2.05 $[\langle D_{KL}\rangle
-		=5.659]$, (c) 1:1.95 $[\langle D_{KL}\rangle=3.279]$, 
-		($n_{refs} = 126$)}
-	\label{fig:refs}
-\end{figure}
-
-Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer.
-This suggests that overshooting our size estimate will be less detrimental to our estimation.
-Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations.
-This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction.
-Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Impact of measurement noise on model predictions}
-\label{sec:epsilon}
-
-So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it.
-In reality of course scattering signals will always have some non-zero measurement noise.
-This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself.
-To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
-\begin{align}
-	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
-\end{align}
-where $\epsilon$ is the percentage error associated with the scattering signal.
-Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:
-
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig10a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig10b.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig10c.png}
-	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ ~
-		$[\langle D_{KL}\rangle=0.962]$, (b) $15\%$ $[\langle D_{KL}\rangle=13.654]$, and (c) $25\%$ ~  $[\langle D_{KL}\rangle=13.017]$.}
-	\label{fig:epsilon}
-\end{figure}
-
-As can be seen from Figure~\ref{fig:epsilon}, the inclusion of 
-signal noise quickly leads to a decrease in the model's 
-performance. So much so that beyond 15\% the model is actually 
-worse than just randomly guessing.This is due to an inherent 
-feature of the inverse scattering problem: two distinct regions 
-in orientation space can become heavily intertwined and thus no 
-longer well separated when mapped to intensity space (even 
-though the mapping remains continuous): so even small 
-uncertainties in the scattering data can lead to large 'mistakes' 
-in the choice of orientation by the neural network. (Indeed if 
-this was not the case the inverse scattering problem would be 
-quite simple.)
-
-To reduce the effects of the signal noise we took the time 
-average of the expected signal over $0.001 s$ and then had our 
-neural network estimate the orientation based on the average 
-signal.
-
-\begin{figure}[h!]
-	\centering
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig11a.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig11b.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\subcaption{}
-		\includegraphics[width=\textwidth]{fig11c.png}
-	\end{subfigure}
-	\caption{Model prediction for signal error of (a) $1\%$ 
-		$[\langle D_{KL}\rangle=1.447]$, (b) $15\%$ $[\langle D_{KL}\rangle=4.670]$, and (c) $25\%$ $[\langle D_{KL}
-		\rangle=7.911]$, time averaged over $1\,{\rm ms}$}
-	\label{fig:time average}
-\end{figure}
-This resulted in a reduction in the overall signal noise and 
-provided a higher degree of accuracy for our model. There 
-appears to be no clear correlation between the length over 
-which we time average and the performance of our model. Time 
-averaging over every 0.05s resulted in a drastically worse 
-performance; this is due to the fact that over longer time 
-periods there is greater uncertainty regarding how the dimer's 
-orientation has changed, thus tracking the instantaneous 
-orientation becomes harder for the neural network. Fortunately, 
-time averaging even over $1\,{\rm ms}$ seems to provide a 
-satisfactory estimation of the dimer's angular dynamics within 
-the optical trap.
-
-Overall it’s clear that estimation of the dimer’s orientation 
-is a problem that can be endlessly tuned to optimise the 
-result. Here we simplify the problem somewhat by employing 
-a relatively small finite number of 'reference orientations' 
-to map between scattering and dimer orientation: the precision 
-of estimation could be improved by utilising a greater number 
-of reference orientations, although there remains a balance 
-between the realisable precision of orientation estimate and 
-the noise level of the scattering measurement. 
-Another avenue to further explore would be using 
-the method to optimise the choice of detection angles, 
-essentially to find the region in the mapping between measured 
-scattering and orientation that others the best degree of 
-confidence through optimal separation of scattering signals 
-for distinct orientations. For sequences of data such as 
-dynamic measurements, a further potential enhancement would 
-be to consider more complex correlations based on prior 
-expectations of the dynamics. The Debye equation for  Here already we improve the 
-method using a non-uniform prior based on only the immediately 
-previous measurement in time (see Section 2.1): considering 
-a non-uniform grouping of reference orientations might result 
-in a better estimation, if we have information regarding the 
-dimer’s preferred axis of rotation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-\label{sec:Conclusion}
-
-We have developed a method for measuring the dynamics of an 
-optically-trapped colloidal objects based purely on 
-measurements of the object's light scattering at a small 
-number of detection angles. We demonstrate the method using 
-the orientation of an asymmetric dimer as the dynamic variable 
-and object of interest respectively, but in principle the 
-model can be applied to any characteristic that impacts the 
-light scattering pattern produced by a trapped entity such 
-as size and shape. The MSTM package is a flexible tool for 
-calculating the light scattering of complex objects using a 
-representation of the object as a set of micro-particles, 
-enabling training of a neural network to enable categorisation 
-of the mapping between scattering and trapped object 
-characteristics. By taking account of the physically realistic 
-behaviour of the trapped object and the characteristics of 
-the trap (which impact the dynamics of the object), the 
-Bayesian inference method can be re ned to provide a reliable 
-estimation of object characteristics of interest, even in the 
-presence of measurement noise. Fundamentally, the inverse 
-scattering problem is difficult to solve, since the mapping 
-between object characteristics and scattering can be highly 
-complex. We determined the minimum number of detectors 
-required for a reliable estimation in the presence of 
-measurement noise; furthermore, we demonstrated that the 
-arrangement of these detectors is critical for a reliable 
-estimation of an objects orientation. However, Bayesian 
-inference based on neural network estimation of the mapping 
-provides a powerful method for practical applications, 
-extending the use of optical trapping beyond measuring 
-microscopic force response toward detailed structural and 
-dynamic information about complex trapped entities.
-
-Conventional calibration methods will not be adequate however, while
-angular motion can be detected using complicated set ups \cite{Bang2020},
-this can only provide an estimate of the magnitude of a particle's 
-angular motion. This is perfectly fine in cases where the rotational
-motion is stochastic and intuitively predictable (i.e. a dimer in a
-vertical orientation undergoing stochastic Brownian motion), but in 
-the case the the motion is instead periodic characterising it requires
-instantaneous measurements of the particle's motion.

### 2025-06-27 Latex_files/Thesis/chapters/chapter 6/main.tex
27/06/2025 chapters 1,3,4 and structural changes

\chapter{Shear Induced Nucleation}
\section{Introduction}
One of the project aims is to investigate the possibility of using
optical tweezing to induce nucleation by generating fluid flow 
within a supersaturated solution. The intent of which would be 
twofold: Firstly to have a repeatable means of inducing nucleation
under different solution conditions. And secondly, to understand 
the influence of shearing on nucleation at a micro level as 
compared to results in bulk fluid.

It has been shown that for macro-scale systems, the likelihood of 
nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Mura and Zaccone developed a 
theoretical framework to describe how the nucleation process when
placed in a moving fluid, in which they highlighted two additional 
factors compared to a stationary fluid. Firstly, due to increased 
molecular transport of solute molecules the nucleation rate is 
enhanced in low to moderate fluid flows. But in addition, due to 
shear flow the crystal surface undergoes deformation which suppresses 
the nucleation rate undergoing faster fluid flow \cite{Mura2016}. Experimental results with glycine solution support this theory; 
Debuysschere \textit{et al} demonstrated that the nucleation rate 
of supersaturated glycine was enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ \cite{Debuysschere2023}, after which the nucleation rate 
began to decrease but was still greater compared to the case where 
fluid flow was minimal.

Optical tweezers can been used to rotate a whole host of microscopic 
objects, with the fastest reported results exceeding $1000\ Hz$ in 
heavy water \cite{Arita2016}. If such a micro-rotor was suspended in
a supersaturated solution the fluid flow around it could be fast enough
that the nucleation rate is locally enhanced. We focused on two 
primary candidates for rotation, Vaterite and 
4-Heptyl-4-biphenylcarbonitrile (7CB). The former being a polymorph of 
calcium carbonate and the latter an example of nematic liquid crystals, 
both of which have been used repeatedly in previous micro-rotor research
\cite{Saito2022, Arita2016, Parkin2009}. In addition, we also consider 
the application of using techniques beam steering to generate fluid 
flow by trapping silica micro-beads. In this instance the fluid flow is 
generated not due to the transfer of angular momentum, but due to 
shearing caused by a moving sphere through stagnant fluid.

In this chapter we cover the basics of nucleation theory, in particular
the development from the classical nucleation theory to the current 
non-classical theory. From there we discuss the experimental set up 
used during this project, highlighting some of the more notable 
optic techniques, before describing the synthesis and performance 
of different birefringent particles as optical rotors within 
supersaturated solutions. Lastly, we highlight some last minute 
results involving a galvano mirror and how the rapid movement of a 
optical trap can influence the growth and behaviour of a nucleus.

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. When 
a small amount of solute is added to the solvent the 
solute will diffuse throughout the solvent, reaching
an equilibrium concentration throughout the solution. 
Because of molecular interactions, the two can only remain 
in equilibrium while below a specific concentration ($C_{eq}$) 
- below which the chemical potential $\mu$ for a single phase 
is greater than the potential required to separate the two. 
Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases.
This does not mean that nucleation will occur immediately 
once $C_{eq}$ is exceeded, only that there is a chemical
potential supporting nucleation.

Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, and 
$C_{eq}(T)$ is the equilibrium concentration of the solute  
at temperature $T$ based on the main solvent. While the 
solution remains supersaturated there is a chemical 
potential driving the solute to coalesce and separate 
from the solution as an ordered solid, the first 
formation of the crystal is referred to as the 
nucleus and understanding its formation has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation 
of the system.

\subsection{Primary \& Secondary nucleation}
nucleation process be broadly categorised into either primary 
or secondary nucleation. The former describes the formation 
of an initial nucleus within the bulk phase, absent of any 
external surfaces. Primary nucleation is dependent only on 
the local concentrations of the solute which can fluctuate 
unpredictably, as such it is often modelled at a large scale 
as a stochastic process. The only reliable factor being 
that higher supersaturations will result in more nucleation 
events to occur sooner.

An important metric for crystallisation procedures is the 
nucleation rate of the solution, a measure of the total 
volume or mass of crystalline material produced for a given 
set of initial conditions. At an industrial scale this 
metric is a contribution of both primary and secondary 
nucleation events. At a small scale one can estimate the 
contribution of the primary nucleation pathway by making 
repeated measurements of sample solutions and seeing how 
many have nucleated after a given time, giving us a Poisson 
probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
	\label{eq:poisson}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time' 
which accounts for the initial delay due to the stochastic 
nature of primary nucleation, $V$ is the volume of the individual 
samples, and $M^*(t)$ \& $M$ are the number of nucleated samples 
and the total number of samples used respectively. While 
\eqref{eq:poisson} is useful for studying the effects of different 
parameters at a small scale, for industrial applications there are 
to many external factors for primary nucleation to be measured 
accurately. Overall while the primary nucleation rate is significant 
when studying solutions for small volumes, the contributions of 
secondary nucleation is far more significant when considering 
nucleation rate for an industrial crystalliser.

Secondary nucleation is the result of a initial seed crystal 
inducing further nucleation within the bulk solution 
\cite{Botsaris1976}. The seed crystals are often prepared ahead 
of time and added to a supersaturate solution. Due to interactions 
with the surrounding fluid and the container walls it is possible 
for the seed crystals to act as a surface for further nucleation \cite{Anwar2015}. However this is just one possible way in which 
a seed crystal can enhance the nucleation rate, depending on the 
conditions of the crystalliser there are multiple mechanism by 
which the seed crystal can contribute to the nucleation rate of 
the system. Control of secondary nucleation events are crucial 
for ensuring reliable industrial crystalliser performance. Fig.~\ref{fig:secondary} depicts an attempt of classifying every 
possible mechanism that could lead to secondary nucleation.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by
		Agrawal and Paterson \cite{Agrawal2015}}
	\label{fig:secondary}
\end{figure}

This is not a universal classification system, there are a 
variety of opinions on how best to characterise different 
phenomena. It is heavily dependent on the theory used to 
describe nucleation events, this is accurate for both 
secondary and primary nucleation.

\section{Nucleation Theories}
In order to predict the nucleation rate of a given solution
we need a solid understanding of how the individual molecules
interact such that they transition from an unordered phase to
a ordered phase. Over the years several theories have been 
proposed to explain how nucleation events occur, so far there
has yet to be a standout candidate that reliably predicts 
the systems behaviour. Either they fail to accurately predict
nucleation rates or more generally cannot describe new 
experimental results.

\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibbs Nucleation Theory' the 
original theory was first formed from the works of Volmer 
and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
While initially it was more focused on describing droplet
formation in condensing vapours it was extrapolated to 
describe crystallisation. The central premise of classical 
theory is that nucleation occurs stochastically due to 
collisions between individual solute molecules, ions, or 
atoms. At the same time the bulk phase is resistant to the
formation of a new phase. The competition between these 
random collisions and the bulk solution can be used to 
predict the probability of a newly formed nucleus.

Consider a supersaturated solution, at some time individual 
sub units collide rapidly, one after another, forming a 
nucleus of volume $4\pi r^3/3$. The newly formed phase has a 
lower chemical potential than the surrounding solution, 
reducing the free energy of the system. Simultaneously, 
the formation of a new interface is resited by the bulk 
phase due to surface tension. The net free energy of the 
system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
	\label{eq:CNT} 
\end{align}

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		to the Classical Nucleation Theory. The total free energy (blue)
		is due to the competition between the volume free energy gain
		(green) and the interfacial free energy cost (orange). Dotted
		lines are for a higher supersaturation than the solid lines,
		the interfacial energy cost is independent of supersaturation.
		Diagram was constructed in python using a naive assumption 
		of $v=1\times 10^-12\ m^3$ and $\sigma_{inf}=0.0001\ J/m^2$}
	\label{fig:free_energy}
\end{figure}

The maximum value of $\Delta G_{tot}$ is the free energy barrier 
that any newly formed nucleus needs to overcome in order to stabilise. 
The nucleation rate (the volume of new crystalline material formed per 
unit time), is therefore commonly defined as being dependent on the 
energy barrier $\Delta G^*$:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}

Where $A$ is a pre-factor that can be fine tuned to the exact 
demands of the system, the free energy barrier can be found by 
finding the stationary point of $\Delta G_{tot}$.

CNT is often regarded as a good description of the macro system, 
its obvious that for all crystallization systems there is an 
inherent energy barrier that must be overcome prior to nucleation. 
Where it falters is in its predictive ability, both in estimating 
nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
structure of newly formed nuclei \cite{Lee1999, Yau2001}.
Recent studies suggest classical nucleation is merely one of 
many possible pathways that can be taken to produce an ordered 
crystalline phase. In situations where the solution is heavily 
saturated with solute the driving force promoting nucleation is 
so large that a nucleus can form in the bulk solution in a single 
step.

This is however not accurate at lower supersaturations, closer
to the saturation point the solvent's surface tension is so 
significant that the addition of individual solute molecules
should not be enough to overcome the dissolution of the nucleus. 
And yet nucleation is still possible, occurring far more reliably
than predicted by the CNT. This suggests that the formation of 
a nucleus is due to larger aggregates of solute that are stable 
within the bulk solution. This has been confirmed experimentally 
and lead to the development of Two Step Nucleation.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT 
that suggests that prior to nucleation, the solute will 
form precursor structures. The CNT assumes that solute 
structures below the critical radius should be unstable 
within the solution due to interfacial forces trying to 
dissolve the solute. This assumption is based on the idea 
that any interactions between solute molecules are much
weaker than that between solute and solvent molecules.
If instead the solute molecule interactions are significant
then it stands to reason that subcritical liquid-like 
structures are stable.

The first observation of a liquid-like structure was in 
$CaCO_3$ solutions, initially it was assumed that polymers
were a necessary additive to create a precursor structure 
\cite{Driessche2017, Karthika2016}. This is not the case when 
the solution is above its critical temperature, allowing the 
solution to be in the binodal region of the phase diagram. 
This would result in the formation of two liquid phases, 
one being rich in solute, the other being solute poor 
\cite{Karthika2016, Fu2021}. Several papers later reported 
the presence of stable liquid-like clusters that formed 
prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{oswalds_rule.jpg}
	\caption{Free energy diagram of Oswald's rule for a 
		crystal with three possible polymorphs. The diagram 
		shows that there exist local minima in the free energy 
		that represent the different polymorphic forms. A similar
		free energy diagram can be used to describe the two-step
		nucleation mechanism.}
\end{figure}

The formation of these clusters can be understood by Oswald's 
rule; which says that any crystallising system does not 
immediately take the path to the lowest possible energy state 
but instead first transitions to the state with the smallest 
free energy barrier \cite{Ostwald1897}. Oswald's rule was 
assumed to be applicable for crystal polymorphs but can be
generalised to describe the nucleation process also. Further phase 
transitions can still occur but the pathway taken should 
always minimises the overall free energy cost. It is assumed
that between the amorphous phase and the crystalline phase 
the solute undergoes dehydration while changing structurally 
into a more ordered phase, though there is some discussion 
whether the process merely undergoes dehydration, or if there 
is a more complex chemical reaction
\cite{Karthika2016}.

With regards to liquid-precursors, there is still a lot of
discussion about the relationship between the solution 
conditions and the the precursors. The size and rate at 
which these precursors agglomerate based on the degree of 
saturation would provide invaluable information on the 
kinetics of two-step nucleation \cite{Fu2021}. In some 
cases it has even been shown that 'ageing' solutions is 
a necessary step in order to detect the presence of 
stable amorphous clusters \cite{Liao2022}. This was in 
solutions that were supersaturated and yet after the 
formation of precursors spontaneous nucleation was not
seen even after several days of observation \cite{Liao2022}.
This suggests that in some cases the precursors are
metastable and require an external trigger to induce
crystallisation. The presence of metastable structures
and other phenomena has lead to development of a more 
universal theory, namely non-classical nucleation.

\subsection{Non-classical Nucleation} 
Two-step nucleation is based on the idea that between
the bulk solution and the final crystalline phase there
exists and intermediate stage. While this is broadly an
accepted concept, research has shown that in some cases
the intermediate stage can be subdivided into multiple 
states. This has been shown to be the case with metals
in both solid and liquid conditions \cite{Cao2020, Ye2023}
and colloidal simulations have indicated that the 
precursors undergo structural change over time \cite{Tan2013}. 
While in some cases it has been possible to directly
monitor the structural change in the amorphous clusters, 
many organic compounds have not received such treatment
due to the difficulty of characterising and monitoring 
the pre nucleation stage \cite{Fu2021}. This is often 
why two-step nucleation is simply referred to as 
multi-step nucleation.

There are also a myriad of other unique nucleation events
that deviate significantly from both the classical and 
two-step nucleation theories. A review paper discussing
recent advances in non-classical crystallisation highlighted
that depending on the collision kinetics of the system 
and the molecular interactions could result in drastically
different reaction kinetics. In the former case, unstable
subcritical nuclei have been shown to collide and coalesce
into a supercritical nucleus, but only in conditions where 
subcritical nucleation are plentiful despite a low degree
of supersaturation \cite{Baumgartner2013}. The latter case 
is more prevalent for dipole solute molecules, the anisotropic 
attraction results in nuclei with non-spherical geometries, 
this results in varying kinetics for agglomeration dependent 
on the orientation of each molecule \cite{Yau2001}. Phenomena 
such as the above examples has been collectively refereed to 
as non classical nucleation, it is known that they must be 
related to both CNT and multi-step nucleation but the connection 
is not clear based on the current literature.

As such, much of the research into nucleation theory is 
focused on developing \textit{in situ} techniques that can 
monitor and characterise the pre nucleation stage, this
would glean information on cases when both classical 
and non-classical nucleation is a possibility 
\cite{Karthika2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{In situ techniques for studying nucleation}
There are several different phenomena \cite{Fu2021, Karthika2016} 
that are not well described by any individual theoretical framework 
(CNT or multi-step nucleation for example). There is an increasing 
interest in probing the pre-nucleated solution.

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable for a single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice). There have 
been several different methods deployed to try and observe 
crystallisation at a microscopic scale.

\subsection{Computer Simulations}
While not a direct observation of nucleation events, computer
simulations have proved an invaluable tool for studying and 
testing predictions about the pre-nucleated solution. By 
comparing the simulative nucleation rate to the expected 
rate given by current models, researchers can test newly 
developed theories and glean information about the microscopic
parameters of a system.

The main challenge facing computer simulations is the issue
of limited scope. Molecular dynamic simulations are constrained
by their choice of time and length scale, for example simulations
in a canonical ensemble face issues where the chemical potential
is depleted to the point that the driving force for crystallisation
is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
addressed but requires an extensive increase in computational 
resources to maintain the chemical potential equilibria. Like wise
in most cases the choice of time scale is a crucial factor, shorter
time scales provide a more accurate evolution of the solution; 
however, in some cases the time scale magnitude for crystallisation
is often far greater than the time scale used in the simulation. This
has led to more advanced sampling techniques that allow for longer 
time scales which has provided insights into the pre-nucleation 
solution stage and formation of pre-nucleated clusters \cite{Finney2023}.

The study of the pre-nucleated solution and the evolution of 
precursors has been aided greatly by the use of simulative 
studies. Early on it was shown in colloidal crystallisation 
that the structure of the precursor was time dependent; 
colloidal simulations showed that regardless of the molecular
interactions between colloids the atomic packing pattern was 
predominately hexagonal centred, over time the packing
became dominated by body centred packing in the case 
where the interactions were longer reaching \cite{Tan2013}. 
They also confirmed that precursor formation was driven not
by local density variations, but instead was driven by the 
local bond order of the colloidal precursors \cite{Tan2013}, 
providing strong evidence that the colloidal pre-nucleation 
clusters undergo rearrangement prior to nucleation. Further
developments in $NaCl$ simulations have shown that by varying
the density fluctuations the precursor structure could either
be amorphous, the classical rock-salt structure, or a new 
structure similar to wurtzite. While this result is difficult
to confirm it provides a important conclusion, that simulations
can hint at new nucleation pathways that have not been confirmed
experimentally. This means that computer simulations can not
be merely used to confirm theories, but can be used to examine
and search for novel results.

Overall computer simulations make up a core area of research 
in the study of nucleation. The technique is however limited,
often struggling to examine nucleation over longer time scales
or larger number of particles. While sampling methods and 
enhanced simulation techniques has alleviated this somewhat 
they are still far from being able to describe the behaviour 
of larger molecules undergoing nucleation.

\subsection{Transmission Electron Microscope (TEM)}
One of the more well known techniques used for direct observations 
of crystallisation dynamics is using transmission electron 
microscopy (TEM). The basic working principle involves creating a 
focused beam of electrons that are directed onto a target sample.
Due to the wave-particle duality, the electrons can be treated 
as a unified beam of light, whose wavelength is dependent on the 
kinetic energy of the electrons. The wavelength of an individual 
electron is given by \cite{Williams2009}:
\begin{align}
	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
\end{align}

\noindent
When incident on a sample the wave is scattered by the sample 
and can then be detected by an imaging screen that measures 
the difference in intensity between the scattered and 
transmitted beams. The advantage of TEM is that one can bypass
the diffraction limit imposed by visible light, the maximum 
resolution that can be achieved by a microscope is given by
\cite{Champness2020}:
\begin{align}
	d = \frac{\lambda}{2NA}
\end{align}

\noindent
where NA is the numerical aperture of the objective lens, and
$\lambda$ is the wavelength of light used to illuminate a sample.
For a visible light LED with an average wavelength of $550~nm$, 
the upper limit for an objective lens (NA of 1.2) is around 
$230~nm$ whereas a similar lens focusing an electron beam 
with an average kinetic energy of $100~keV$ will have a maximum 
resolution of just over $4~pm$ \cite{Champness2020}. There is
an upper limit to the maximum resolution possible as electron 
lens are far inferior to what modern optics can achieve 
\cite{Champness2020} so there is a trade off between quality and 
resolution.

TEM technology has proven invaluable to the study of nucleation
at an atomic level. The samples are often only a few hundred 
atoms thick in order to optimise the imaging results. Much of 
the supporting evidence for non-classical nucleation has been 
demonstrated via TEM \cite{Ye2023}. For example when Cao 
\textit{et al} observed the formation of amorphous atomic 
cluster using $\gamma-$Fe,Au, and Re; all three of which 
underwent continuous transformation before eventually 
reaching an final ordered state \cite{Cao2020}. The amorphous 
cluster did not immediately transition to an ordered structure, 
instead fluctuating somewhat randomly until eventually settling 
into a crystalline structure \cite{Cao2020}. This was confirmed 
both in solid growth and liquid growth conditions \cite{Cao2020, 
	Loh2016}, suggesting that the non-classical nucleation path 
dominates in all cases of metal crystal formation \cite{Ye2023}. 
Confirmation is necessary for different solute choices, as in 
the case of organic solids it is believed that non-classical 
nucleation only dominates in low supersaturations whereas 
classical nucleation is dominate at higher supersaturations.

TEM analysis has not been extended to the study of organic 
crystals, proteins, or simple salts at an atomic level 
however. The main reason being that organic materials can be 
damaged by the electron beam and therefore need to be frozen 
prior to being imaged \cite{Champness2020}. This was used to 
study the polymorphic time dependence of glycine crystals, by 
varying the time prior to freezing Broadhurst \textit{et al} 
showed that $\beta$ glycine dominates initially but $\alpha$-
glycine is more stable after a longer period, $\gamma$-glycine 
was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
the time scales with which polymorphs form, it doesn't provide 
much information on the kinetics of crystal growth as these 
were snap shots and not direct observations of the nucleation 
process.

Overall TEM analysis is invaluable for the study of atomic 
crystal growth, the secondary scattering from electron 
beams also allows for chemical and structural information 
to be gleaned from the target. This has given strong evidence
for the non-classical nucleation theory, particularly in 
metal crystal formation. Due to the high energy of the 
electrons, organic materials have not received the same
focus outside of cryoTEM analysis. Non-invasive methods
are therefore necessary to gain an understanding of the
crystal growth and formation of organic crystals, 
particularly when both classical and multi-step nucleation 
can occur.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector 
\cite{Gieseler2020}. The laser used for this project was a $1064\ 
nm$ near infrared laser - provided by CNI Lasers – that was focused 
by a Nikon 100x oil immersion lens. The choice of an oil immersion 
lens is important as the optical oil used prevents a loss of focus 
when used on a glass cover slip. Now, experimental work has shown 
that the trapping efficiency increases with beam diameter up until 
it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
$D_{obj}$ is the diameter of the objective aperture. To expand the 
beam front we utilise a Galilean beam expansion arrangement 
(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
for high power laser applications. In our initial experiments the 
beam expansion provides a $4\times$ magnification. Whereas in later 
experiments we utilised a galvano-mirror the beam expansion is 
$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
- the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use of a 
Keplerian beam expansion arrangement which reduces the transmitted 
laser power due to localised heating of the air. Afterwards the 
laser is passed through a dichroic mirror that separates incoming 
infrared and visible light, this is to prevent the laser from 
damaging the CCD camera used for imaging the trapping plane. The 
laser is then focused to a diffraction limited spot by the objective. 
Utilizing a high numerical aperture objective enhances the gradient 
force at the focal point; the trade-off being that the for higher NA 
objectives the trapping depth is reduced due to spherical aberrations. 
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses 
the scattered laser light and also provide an aperture for an imaging 
LED to illuminate the focal plane. Samples are loaded onto a piezo 
driven table to that is inserted between the trapping and condensing 
objectives; the piezo drivers allow for sub-micron control of the 
beam focus position to a degree as small as a 10 nm. To detect and 
monitor the position of a trapped particle a quadrant photo diode 
(QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the work. 
		The focal lengths are $f_1= -20\ mm $ , $f_2=60\ mm$, $f_3=150\ 
		mm$, \& $f_4=200\ mm$ [M = mirrors, DM = dichroic mirrors, f = 
		focal lenses]. Diagram not drawn to scale.}
	\label{fig:setup}
\end{figure}

In order to rotate birefringent particles a quarter-wave plate was
installed via a rotating optic after $f_4$ (the exact placement in the 
is not specific). When the quarter wave plate's is aligned with the 
electric field of the laser we have linearly polarised light, and when
rotated by $45\deg$ the laser is now circularly polarised. Finding 
the point where the wave plate aligns with the laser was achieved by
measuring the power output and finding its maximum. When a birefringent
particle is trapped and the quarter wave plate is adjusted the particle 
will begin to rotate - assuming the fluid drag torque is less that the 
optical torque.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Galvano-Mirror}
Later on in this chapter we discuss the impact of a moving beam front
on the growth behaviour of a nucleating system. In order to do so we
need a means of rapidly displacing the beam focus, for our purposes
we chose a galvano-mirror. Galvano-mirror's are a method of beam 
steering, using a set of motorised mirror mounts the laser can be 
moved rapidly in any number of regular patterns. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth]{signal_generator.jpg}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth, height=0.7\linewidth]
		{galvano_mirror_controllers.jpg}
		\caption{}
	\end{subfigure}

	\caption{Signal generator galvano mirror controller, channel 1 controls the $x$-axis 
		mirror, while channel 2 controls the $y$-axis mirror. Both channels can be manipulated
		independently.}
	\label{fig:signal_generator}
\end{figure}

The signal generator used was supplied by 'MCS Test Equipment Ltd', 
allowing for dual channel signal control. This allowed us to precisely control the amplitude, phase, and frequency of both mirrors making 
alignment much easier.

While in theory the frequency can be increased until the mechanical 
limit of the mirrors is reached ($\approx 1000\ Hz$), the practical 
upper limit is determined by the trap strength. For a silica sphere 
this is on the order of $300\ Hz$ when suspended in water. Likewise 
while in theory the signal amplitude can be increased until the 
voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
of the focal lenses limits the maximum amplitude to $0.5 V$ as any 
greater will move the laser beyond the lens. For basic trapping 
calibration the galvano-mirrors were set to DC output, providing a 
fixed spot which operates like a typical optical trap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the galvano-mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

A 4f correlator is an example of Fourier optics in practice, 
understanding that a focused lens takes a Fourier transform 
of the light profile. Consider a laser with a circular Gaussian 
profile, if you were to place a detector there you would pick 
up the intensity as a function of its position within the beam. 
If however you focused the light into a single point (using a 
+ve focal lens) you are actually seeing a measurement of the 
phase of your laser with position, in which you would see a 
diffraction limited spot, indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be 
used to filter out unwanted scattering from an image (or in an 
inverse case differentiate between different images), the 
placement of each lens is shown in Fig.~\ref{fig:setup}.

For our applications a 4f correlator is utilised to ensure that the 
motion of the galvano-mirrors does not move the focal point of the 
laser, allowing for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after the galvano-mirror we have our two lenses 
- $f_3$ and $f_4$ - the former being installed $150\ mm$ from the 
second mirror of the galvano, and the latter being installed $200\ mm$ 
from the back focal plane of the trapping objective.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro-spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated 
using optical tweezers \cite{Parkin2009, Saito2022}. Over 
the course of the project two different micro spheres where 
investigated, Vaterite and liquid crystal droplets. Both can 
be readily synthesised in the lab and are will rotate at a 
variety of sizes (see \ref{sec:opt_torque}).

Vaterite is a polymorph of calcium carbonate that is rarely 
seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
All three polymorphs are inherently birefringent meaning 
that they can be rotated using circularly polarised light. 
However unlike its other polymorphs of calcite and aragonite, 
when synthesised vaterite will typically form small spherical 
particles making them ideal for optical trapping and rotation. 
Synthesis of Vaterite micro spheres requires fine control of 
the crystal growth process in order to maintain polymorphic 
stability. Though for the purposes of optical rotation the 
exact polymorph is not as important as the morphology as all 
3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts 
of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
aside for later. First a small vial was filled with $1.5mL$ of 
$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
$NaCO_3$ respectively, forming a seed solution. Next, a larger 
vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of 
Agepon was added to halt the reaction, the solution was filtered 
and washed 3 times with distilled water before being suspended 
in water.

When trapped in circularly polarised light, the anisotropic 
crystal lattice allows spin angular momentum to be transferred 
to the Vaterite particle, resulting in a rotation about the 
beam axis. Because the particle scatters light anisotropically 
the signal detected by the QPD shows periodic fluctuations 
\cite{Yogesha2012}. In addition due to the change in polarisation 
there is also a periodic variation that appears at twice the 
rotational frequency \cite{Monteiro2018}. Therefore, the resulting 
power spectrum is not a Lorentzian but now also displays peaks 
that appear at integer multiples of the particles rotational 
frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Vaterite sphere suspended in water and trapped by
		 circular polarised trap. (b) Collected power spectrum from 
		 rotating Vaterite, peaks in the power spectrum appear at integer 
		 multiples of the rotational frequency ($f_{rot} \approx 49.8$~Hz)}
	\label{fig:vaterite}
\end{figure}

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical 
assumption being that the particle perfectly rotates in the $x-y$ plane. 
It has long been suspected that birefringent microspheres experience torques 
outside of the $x-y$ plane \cite{Volpe2023} making it very difficult to 
characterise the behaviour of rotating birefringent microspheres without a 
proper understanding of the full optical torque being applied to it.

\subsection{Liquid Crystal Rotors}
Liquid crystals are an intriguing example of materials with mixed phase
properties. Unlike typical solutes such as Glycine, a liquid crystal
can still maintain some degree of order between its individual molecules
while in the liquid state. This is due to the fact that liquid crystals
are constructed of ordered molecules that demonstrate a long range
ordering. There are three main types of liquid crystal transition methods:
Thermotropic crystals will transition to their liquid crystal phase when 
sufficiently heated. Lyotropic materials can undergo this transition due 
to changes in temperature and concentration. And lastly, Metallotropic 
materials - which are composed of both organic and inorganic molecules - 
change phase according to the ratio of organic to inorganic molecules present.
Liquid crystal rotors are rather simple in their production, 
4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
and a small amount was added to a vial of distilled water. The solution 
was then heated in a water bath to $25^\circ$ in order to transition the
solid crystal into its liquid crystal state. The solution can then be 
loaded onto a sample cover slip and the individual droplets visualised.
The molecules of 7CB will align with a strong electric field, and due to
the spherical droplet geometry the droplets are inherently birefringent. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\linewidth]{LC_sample.png}
	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
\end{figure}

The liquid crystal droplets had a much faster rotation rate than comparable 
Vaterite spheres, due to their higher degree of birefringence and the fact
that the droplets are far closer to perfect spheres making angular momentum
transfer more efficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.

When it comes to optical rotation, a rotating particle will experience a
fluid drag torque that is proportional to its rotation rate, the maximum 
rotation rate is when the fluid drag is equal to the optical torque 
of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
to measure local temperature changes \cite{Millen2014}, or for understanding 
how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
one can use a beam steering arrangement to probe the drag force of the 
fluid, by understanding the trap strength (calibrating using a low 
frequency signal) one can measure the drag force experienced by the local 
fluid \cite{RobertsonAnderson2018}.

For any discussion of fluid flow its important that we know the fluid 
regime which is given by the Reynolds number. For a sphere submersed in a 
moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
density and viscosity respectively. In our case we do not have a fluid
moving around a sphere but a sphere moving through the fluid at some 
velocity $U$, assuming a no-slip boundary condition we can model the 
fluid velocity profile based on the velocity of the particle. Assuming
that for now the fluid properties are unaffected by the rotating particle
we see that for an individual sphere the Reynolds number can only change 
due to increasing/decreasing the fluid velocity. The fluid velocity is 
directly proportional to the particles rotation rate for low Reynolds
numbers \cite{Bruce2020}. Given the small size of the particles used 
($1-10\mu\ m$ in diameter), and the relatively low rotation rates that 
are possible with an optical tweezer, for the case of our analysis the
Reynolds number will not never be large enough to consider factors such
as turbulence. There are two possible avenues for generating shear flow 
with a trapped particle; rotation of birefringent particles, and fluid 
flow induced by particle motion.

Rotating birefringent particles are the more common method for generating 
and measuring fluid flow in a solution. To see if we can even achieve the 
theoretical maximum shear rate, Vaterite spheres were synthesised (see 
Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm 
laser at set to 450 mW. The rotation frequency was determined using the 
QPD, and the particle sizes were computed by image analysis. With the 
particle size and rotation frequency, the tangential rotation speed is 
calculated via \cite{Saito2022}:
\begin{align}
	\label{eq:birefringent_speed}
	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
\end{align}

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field). This approach was used previously with liquid crystal 
rotors \cite{Saito2022}:
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= 
	\frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}

One would expect that we set some reference distance where the fluid flow 
would be considered zero. However, at a microscopic level we are more 
concerned by the fluid shear for an individual nucleation event. We are
essentially computing what the shear rate would be for a small volume 
of fluid between two moving parallel plates. This is similar logic to Mura 
and Zaccone's work where they considered the shear on a subcritical nucleus
in a flow field \cite{Mura2016}. They go further in depth to discuss the 
impact of material transport, overall coming to the conclusion that 
increasing shear rate leads to an increase the nucleation rate up till some
maximum. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around a rotating particle in bulk fluid}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples 
were synthesised according to \cite{Parkin2009, Bishop2004} (see 
sec.~\ref{sec:vaterite}), and then suspended in distilled water. 
A sample of $200\ \mu L$ was pipetted and a single microsphere was 
captured via a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck 
together, fortunately individual sphere's were still present. 
Individual microspheres were trapped and their rotation rate was 
determined by looking at the peak frequency component of the 
collected power spectrum. Due to their anisotropic scattering the 
QPD signal will oscillate periodically. After taking the Fourier
transform of the QPD signal the power spectrum has distinctive
peaks at integer multiples of the rotational frequency. The largest 
peak is often twice the rotational frequency which is due to the 
polarization modulation of the scattered field, the first peak 
is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.

\begin{figure}[h!]
	\begin{subfigure}{0.49\linewidth}
		\includegraphics[width=\linewidth]{time_domain_data.png}
	\end{subfigure}
	\begin{subfigure}{0.49\linewidth}
		\includegraphics[width=\linewidth]{frequency_domain_data.png}
	\end{subfigure}
	\caption{Example of QPD signal being converted from the time 
		domain to the frequency domain. (left) Raw signal collected
		from the QPD over the 1st second of its trajectory showing 
		a periodic spacing with a period of $1/f_{rot}$. (right) 
		Power spectrum of the QPD signal, the y-axis is a linear 
		scale to demonstrate the clear peaks in the power spectra. 
		The rotational frequency is the first instance of a peak 
		forming, and after that the peaks appear in integer 
		multiples of $f_{rot}$}
	\label{fig:tdat_fdat}
\end{figure}

The accuracy of the rotational frequency is dependent on the sample
rate of the QPD and the sample duration \cite{BergSoerensen2004}. 
With higher sampling rates resulting in a wider frequency window 
(see Sec.\ref{sec:psd}) and longer sampling durations resulting in 
more distinct rotational peaks. The default QPD sampling rate is 
$2^{17}\ Hz$ which is more than adequate for detecting even the 
slightest low frequency signals. And the default sampling duration 
was $3\ s$ which for most cases gave us clear distinct peaks for 
even relatively low rotational frequencies. After trapping a 
vaterite microsphere the power spectrum was recorded and the 
rotation frequency was acquired. This was repeated several times
for each sphere trapped, to confirm that the rotational frequency
was constant and that the sphere had reached its maximum rotation 
rate.

After computing the rotational frequency the radial fluid speed 
could be estimated using Eq.~\eqref{eq:birefringent_speed}. From 
there we can estimate the shear flow experienced by a small volume 
of fluid within the flow field (see Eq.\eqref{eq:birefringent_shear}). 
Here we only consider the fluid flow as it propagates outward from 
the axis of rotation. While the cover slip will have some effect 
fluid flow it is difficult to quantify the effects of a hard 
boundary on a rotating object when the boundary is perpendicular to 
the axis of rotation. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.7\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.7\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
		\caption{}
	\end{subfigure}
	\caption{(Top) Fluid flow radiating out from the surface of a
		rotating Vaterite sphere. (Bottom) Shear rates computed 
		using Eq.\ref{eq:birefringent_shear}, optimal shear rate is 
		of $3000 s^{-1}$ is indicated by the dotted line. Vaterite 
		radii and rotation frequencies are shown, the laser power 
		was kept constant at 450 mW. Reported rotation rates, and 
		their corresponding fluid flow and shear rates, for Vaterite 
		are also plotted alongside lab results. Results from 
		\cite{Parkin2009, Arita2016} are included as well.}
	\label{fig:vaterite_shear}
\end{figure}

From Fig.\ref{fig:vaterite_shear} there is not a strong relationship 
between particle size and rotation rate, this is contrary to much of 
the theoretical predictions that predict an exponential decay with 
particle size. This can be in part due to the fact that synthesising 
perfectly spherical spheres that have uniform birefringence across the 
whole population is difficult. Despite our best efforts at controlling 
the growth rate the smallest particle ever synthesised was around $3\ 
\mu m$ in diameter. The Vaterite spheres would often stick together 
while suspended in water after a short period of time. The fastest 
reported rotation rate within a fluid was by \cite{Arita2016} 
that achieved a rotation rate of $5\ kHz$, this is plotted in Fig~\ref{fig:vaterite_shear}. In addition we added the results from 
\cite{Parkin2009} as a more realistic example.

The optimal shear rate predicted by \cite{Debuysschere2023} is 
plotted on Fig.~\ref{fig:vaterite_shear}(b), the outer dotted 
lines represent the point when the nucleation rate is less than 
90\% of the maximum. We can visualise the fluid around the 
rotating sphere by dividing it up into radial sections. Assuming 
the sphere is rotating sufficiently quickly there are three 
sections of concern, as demonstrated by fig.~\ref{fig:shear_diagram}. 
The fluid close to the sphere experiences a shear rate that 
is far greater than the optimal shear rate, in which case 
nucleation is suppressed \cite{Mura2016}. Secondly, there is 
a region of fluid that experiences a shear rate close to the 
optimal value ($\pm 5\%$) where the nucleation rate is enhanced 
significantly compared to the bulk fluid. Lastly, beyond the 
optimal region the shear rate drops off exponentially in which 
case the nucleation rate is barely different compared to the 
bulk fluid.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.875\linewidth]{shear_rate_diagram.pdf}
	\caption{Diagram demonstrating how fluid shear rate varies 
	radially from a rotating sphere (rotation rate = $f_{rot}$). 
	(a) the shear rate exceeds the optimal shear rate 
	($\dot{\gamma}_{opt}$) and suppresses the nucleation rate. 
	(b) shear rate is close to the optimal amount, enhancing the 
	nucleation rate. (c) shear rate is far below the optimal 
	amount and the nucleation rate is comparable to the bulk fluid.}
	\label{fig:shear_diagram}
\end{figure}

\section{Rotating spheres in Supersaturated solution}
If rotation rates in bulk solution are insufficient then a 
rotating sphere close to an artificial barrier may be able 
to improve the shear rate of the surrounding fluid. Of course 
placing a solid barrier in a supersaturated fluid may well 
encourage nucleation somewhere on the surface outside of our 
control. Instead we chose to use the droplet edge of the 
supersaturated solution, while not a hard barrier per say, 
the molecular mobility close to the droplet edge is reduced 
due to surface tension. Furthermore, it has been shown through 
multiple results that nucleation is enhanced at the air-solution 
interface \cite{Liao2022,Yuyama2010, Sugiyama2009}.

Estimating the distance from the trap focus to the cover slip
is difficult to measure directly, often requiring one fits a 
complex Lorentzian curve to the recorded power spectra in order 
to measure the fluid drag from being in close proximity to the 
cover slip \cite{BergSoerensen2004}. We estimate that the 
vaterite must be at least $15-20\mu\ m$ from the cover slip for 
us to reliably trap them. The contact angle between the cover 
slip and solution was previously measured in \cite{Flannigan2023},
while being loosely dependent on the supersaturation the 
measured contact angle for supersaturated solutions was between 
$30^{\circ}$ and $35^{\circ}$. This is shown in fig.\ref{fig:3.9}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating 
		birefringent particles in a supersaturated solution. 
		Left: side view of the trapping set up showing the 
		location of the trap focus at the edge of the droplet 
		of a supersaturated solution. The approximate contact 
		angle (calculated by \cite{Flannigan2023}) is used to 
		provide a scale of how close a trapped particle could 
		be located. Right: top down view of the glycine droplet 
		with a trapped birefringent particle shown close to 
		edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field 
		around itself (see Eq.~\ref{eq:birefringent_speed}).}
	\label{fig:3.9}
\end{figure}

Supersaturated solutions of glycine and water were prepared 
and stored in an incubator at $40^\circ C$ prior to use. 
When ready to be studied $15\ \mu L$ of vaterite suspension 
was pipetted into the solution and $20\ \mu L$ of the combined 
solution was pipetted onto the cover slip. A Vaterite sphere 
was located, trapped, and moved to the droplet edge where the 
solution meets the cover slip. This removed the need for 
hydrophilic coatings to achieve a flat fluid layer (as used by 
\cite{Liao2022, Rungsimanon2010}).

Particle sizing was achieved by converting the particle diameter 
from pixels to physical units within an accuracy of $\pm0.05\mu 
m$. The QPD was used to measure the frequency of fluctuations 
in the QPD signal. After measuring the rotational frequency the 
sphere was left to rotate for a period of ten minutes after which, 
if no nucleation event was observed the particle was released. 
The overall results are catalogued in table~\ref{tab:Nucleation}.
\begin{table}[h!]
	\centering
	\caption{Results from rotating Vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\label{tab:Nucleation}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu m$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 10.4 & $\times$ \\
		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
		\cline{2-4} & 5.67 & 9.63 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\cline{2-4} & 6.32 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 1.52 & $0.00$ & $\times$ \\
		\cline{2-4} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

It should be noted that rotational frequencies equal to 
$0.00\ Hz$ are not due to a rounding error. Simply looking
at the live video shows no rotation of the vaterite particles,
and even looking at the QPD signals shows no discernable 
fluctuations that would arise due to rotation. Trying to trap 
a particle close to the edge proved more challenging than 
expected. Unlike in previous reports where the beam is 
focused at the upper edge of the droplet \cite{Liao2022, 
Yuyama2010, Sugiyama2022}, we attempted trapping in close 
to the contact point of the droplet and the cover slip.

It is suspected that trapping is much harder at the 
interface due to increased surface tension and unpredictable 
scattering forces. The closest we could trap a microsphere to 
the droplet edge was in the range of $5-10\mu\ m$, at that 
distance the fluid flow is so low that even the presence of a 
hard boundary would be insufficient for shearing the fluid. 
Furthermore, as is evident in Table~\ref{tab:Nucleation}, 
the rotation rate drops off significantly with increased 
supersaturation, due to higher fluid viscosities. While in 
theory a sufficiently focused laser could rotate any 
microsphere to a fast enough to reach the shear rate 
predicted by \cite{Debuysschere2023} the localised intensity 
would be so large that even using $D_2O$ would see a significant 
increase in temperature.

It is not impossible that fluid shearing could be used in the 
future to localise nucleation; but from these results, using 
individual micro-rotors is not an appropriate method due to
two key factors. Firstly, the area of influence is far too 
small to see any noticeable increase in the nucleation rate 
this is demonstrated most clearly in fig.~\ref{fig:vaterite_shear}. 
Due to the exponential decay in the expected shear rate it is 
very difficult trying to shear a large volume of fluid. Even 
if the optimal shear rate at a micro-level was an order of 
magnitude less than what was predicted by \cite{Debuysschere2023}, 
only a small volume of fluid would even experience that shear 
rate. In order to properly understand the microscopic effects 
of fluid shear a method of localised shearing over a larger 
fluid volume is necessary.

Secondly, increased fluid viscosity significantly reduces 
the limits the maximum rotation rate possible. As indicated 
by \ref{tab:Nucleation}, the rotation rate achievable by a
vaterite sphere falls off significantly with increasing 
supersaturation. Experimental estimations of relative viscosities
showed a 12\% increase when looking at undersaturated glycine
solutions ($S\approx0.34$) \cite{Patyar2020}. As far as we
are aware there are no reported fluid viscosities for higher
supersaturations. Extrapolating the results from \cite{Patyar2020}
suggests that a saturated solution would have a 35\% increase 
in fluid viscosity (compared to pure water). This also does 
not account for any other factors such as the fact that close
to a boundary a rotating particle will experience a reaction
torque from the stationary surface \cite{Bruce2020}.

If multiple micro-rotors could be trapped in close proximity 
to one another they could create a large region of fluid where 
nucleation is more likely than the bulk fluid. Because optical
torque is not contingent on the fluid properties the only 
limiting factor would be the total angular momentum transferred 
to each particle. Micro-rotors have been created that allow for 
precise control of suspended micro-particles \cite{Butaite2019} 
and could potentially be used to generate sufficient shearing. 
However these could not be used in this project as we lacked the 
necessary hardware to form multiple gradient traps.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation did 
not result in any localised nucleation events even while 
in the proximity of the droplet edge. An alternative approach
was suggested, using a galvano mirror to move a particle 
quickly through a supersaturated solution. Preliminary 
calculations suggested a particle moving rapidly through 
a fluid could produce significantly greater shear rates 
(see \emph{Appendix X.X} for breakdown) than a rotating 
particle. While a particle is within the optical trap the 
gradient forces are significantly reduced meaning simply 
trapping a particle is insufficient for inducing nucleation \cite{Flannigan2023}. This is why we did not see any 
nucleation events while rotating the vaterite, there is 
not a large gradient force to draw in material.

During initial testing using silica beads we found that 
even in undersaturated solutions the optical trap would 
produce a crystal nucleus when the trap was empty. This 
has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but what is more interesting is how the beam's motion 
influenced the growth of the nucleus.

\subsubsection{Choice of Crystal growth rate units}
The growth rate of a crystal can be described using different
units depending on the situation. For example, seed crystals 
often use growth units in the form of \textit{length per unit
time}. This can either be used to describe the overall length
of the longest edge of a crystal, or to describe how individual
faces grow separately. This works well for crystals with a 
clear morphology, in regards to laser induced nucleation the 
crystal growth is heavily influenced by the local fluid 
conditions meaning tracking the length of an individual face 
is challenging. Instead we utilise image analysis software 
such as imageJ to measure the area of the crystal after each 
frame to get a rough estimate of the crystal growth rate in 
units of \textit{length$^2$ per unit time}. With the addition 
of the contact angle information from \cite{Flannigan2023} we 
can approximate the crystal height based on how close it is to 
the droplet edge.

\subsection{Stationary beam}
\label{sec:stationary}
Its well known that a supersaturated solution can nucleate
if irradiated by a focused beam with sufficient power 
\cite{Rungsimanon2010}. First we look at how a laser induced
nucleation event unfolds using a stationary focused beam. 
The laser was focused at the edge of a $20\mu L$ droplet 
consisting of water and dissolved glycine ($S=1.03$). The 
solution was monitored for 10 minutes, if a nucleation event 
was observed the event was recorded using a high speed CCD. 
If no nucleation event was observed after 10 minutes the 
solution was disposed of and a new sample was prepared. After 
ten repeats only 30\% of the samples observed a nucleation 
event at the focus. All of which displayed similar growth 
behaviour to one another.

Consider Fig.~\ref{fig:stationary_beam}, the frames taken from 
a nucleation event, the beam is a stationary being $\approx3.5 
\mu m$ from the droplet edge. After a period of roughly 5 minutes 
a nucleus forms at the trap focus, growing quickly (growth rate 
was approximated using imageJ to be on the order of $700\ \mu 
m^2/min$) from the focal point of the trap until after roughly 
$6$ seconds the crystal escapes. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to 
be held in place and is in fact still growing as the solution is supersaturated. Based on the contact angle measurements from \cite{Flannigan2023}, the droplet height could not have been 
any more than $2-5\mu m$ at the trap focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth, height=1.3\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
	the trap.}
	\label{fig:stationary_beam}
\end{figure}

Comparing to previous literature using optical tweezers shows 
that the growth rate is only loosely connected to the solutions 
supersaturation. Crystal growth rates for a solution of glycine 
and water ($S\approx1.00$) was found to vary from as much as 
$3900 \mu\ m^2/min$ to as low as $675\mu m^2/min$ 
\cite{Flannigan2023}. The reason for such high crystal growth 
rates is due primarily to the tweezer focus bringing in material 
due to gradient forces. A study of glycine crystals in water 
found that for a similarly supersaturated solution the growth 
rate along the {011} face was around $0.1\ \mu m/min$ where as 
the {010} face was found to have a growth rate of around $0.01\ 
\mu m/min$. While its difficult to extrapolate an exact area 
growth rate from these results we can approximate that if the 
we looked directly perpendicular to the crystal growth the area 
would increase at $\approx0.02\ \mu^2/min$ (see X.X for a 
breakdown of this approximation).

The key take away to remember is that the beam has no influence 
over the crystal shape, instead it grows outward from the trap 
focus. Furthermore, due to the fact that the solution is 
supersaturated the crystal growth cannot be contained to the 
trap focus. Instead the crystal escapes as its size exceeds the 
trap focus.

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the 
necessary shear rate for crystal nucleation we wanted to 
see if a trapped silica bead could be trapped and moved in 
an aqueous solution. $20\ \mu L$ of glycine and water ($S=
1.03$) was added to $10\ \mu L$ of a dilute water-silica 
mixture making the solution unsaturated ($S\approx0.7$). 
However, due to the beam's motion we instead encountered 
unexpected growth behaviour.

Shown below in Fig.~\ref{fig:eliptical_beam_1} where 
we have the laser focus moving in a small elliptical 
pattern. While nothing is seen directly entering the 
focus a nucleus forms close to the droplet edge, 
unlike in fig.~\ref{fig:stationary_beam} the crystal 
does not grow out from the focal point evenly. Due to 
the galvano mirror, the crystal is simultaneously 
being moved by and growing around the focal point of 
the trap. Because of this the crystal nucleus lacks 
a clear morphology at first. Until roughly $20\ s$ 
the crystal reaches a almost prismatic structure, 
with further irradiation increasing the size.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth 
		of a nucleus using a moving beam. Beam power is kept 
		at 700 mW and the supersaturation was estimated at 
		S=0.86. Initially the crystal shape is amorphous (a) 
		but eventually reaches a more regular shape (b). This crystal is still influenced by the optical trap as 
		even when not directly irradiated by the laser the 
		crystal rotates between (c) and (d). When the laser 
		is focused on a corner the crystal growth is localised 
		to that region, resulting in an elongated section 
		forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

Interestingly the galvano-mirror allows the trap to 
impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
though the crystal is not directly in the trap focus 
it rotates in the $x-y$ plane and gets trapped again 
at a corner. The rotation could not be due to fluid 
flow close to the surface of the crystal as the 
dipole moment of individual water molecules is too 
small to be influenced by an optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
crystal growth becomes localised to the corner. The 
area growth rate between figures \ref{fig:eliptical_beam_1}
(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures 
\ref{fig:eliptical_beam_1}(e) and (f) the growth rate 
at that particular edge was estimated at $42.10\ \mu m^2/min$.

Nucleation in undersaturated conditions has been 
reported previously in $D_2O$ \cite{Rungsimanon2010} 
and $H_2O$ \cite{Flannigan2023}, though not involving 
a moving beam. This modification allows for the crystal 
growth to be localised to a specific region of the bulk 
crystal whereas with a stationary beam there is no 
control over the crystal morphology. In fact this allows 
for a much finer control over the exact shape of the 
crystal nucleus, in some cases allowing for growth out
of the viewing plane as shown in fig.~\ref{fig:out_of_plane}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{out_of_plane.pdf}
	\caption{Interesting crystal growth on the surface of an
	existing crystal. Initially the crystal surface seems 
	flat (a) but eventually a different crystal structure 
	grows out off of the surface (b).}
	\label{fig:out_of_plane}
\end{figure}

Where a large outgrowth forms on the surface of a nucleus,
over a period of $30\ s$ the outgrowth develops into a 
rhombic crystal structure. The unique crystal growth 
demonstrated is likely a factor of the moving beam focus,
but this raises a question over how the laser can localise
the growth only around the focus and why the crystal does 
not grow or shrink when not directly within the trap. There
should be some material that is being drawn into the trap
as it scans across the camera frame.

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
In a repeat experiment a solution similar to sec.~
\ref{sec:stationary} was made up, but without any 
silica micro-spheres. Once again the beam was focused 
close to the droplet edge, this time the galvano 
mirror was scanning a circular path (as shown in fig.~\ref{fig:cluster_trapping}). After a few 
minutes of irradiation droplets were seen entering 
the camera frame. Because no silica had been added, 
and that the solutions were filtered, these droplets 
had to be from the glycine solution. Trapping 
individual droplets did not result in immediate 
nucleation even after several minutes being trapped. 
Trying to bring two droplets together resulted in 
nucleation between the two droplets, compared to 
\ref{sec:stationary} \& \ref{sec:moving} the growth 
is much slower, taking nearly 40 seconds before the 
crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping 
		of a glycine droplet. Solution is undersaturated glycine 
		and water (S = 0.86), with the laser power is set at $750\ 
		mW$. (a) shows a trapped droplet (outlined in black) being 
		brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth 
		is rather slow with the crystal having no clear defined 
		morphology through (c) and (d). Between frames (e) and (f) 
		the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates 
they must have a higher refractive index than the 
surrounding solution. Previous reports have shown that 
a focused trap will draw in solute material from the 
bulk \cite{Tsuboi2009, Gowayed2021}. This suggests that 
the droplets observed are concentrated spheres of glycine, 
which also explains how nucleation was initiated when those
two spheres were brought into contact with one another. The 
droplets must be providing material for the crystal growth.
It has been shown that the concentration of glycine solution 
is correlated with the refractive indices of the liquid 
\cite{Gowayed2021, Orttung1963}. This suggests that we could
measure the concentration of individual droplets based on the
trapping strength of each droplet. 
\newpage

\subsection{Influence of a moving beam front on seed crystals}
\label{sec:seed_crystals}
One potential application of this phenomena would be in using 
a moving the moving beam front as a method for shaping the 
final crystal morphology. Therefore, we wanted to test if a 
moving beam front could have any influence on the shape of a
seed crystal submerged in a bulk solution. Glycine seed crystals 
were grown via evaporative crystallisation over night. The 
resultant crystals were as wide as $0.5 mm$ in some cases. 
Individual crystals were collected and suspended in a water 
+ glycine solution ($S = 1.001$) and the laser was scanned 
along a narrow linear path. It should be noted that the seed 
crystal was fully submerged in the solution, meaning that there
is no interface between the air and solution close to the seed.
Five separate repeats were conducted, all demonstrating similar
behaviour to one another.

As shown by fig.~\ref{fig:seed_crystals} the laser does not 
in fact promote crystal growth but instead forces the crystal 
to dissolve into the bulk solution. The dissolution rate is not 
as substantial as the growth rate seen in previous sections. 
It takes over 20 minutes for the for the crystal surface to 
dissolve more than a few microns. Furthermore we do not report 
any sightings of droplets close to the surface of the crystal 
nor do we see anything enter the optical trap.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{seed_crystals.pdf}
	\caption{Frames from a seed crystal being irradiated by a 
		scanning laser. The laser path is shown below in red. 
		Initially the crystal edge is nearly perpendicular to 
		the laser path. Over the course of nearly twenty minutes 
		the crystal around the laser has dissolved leaving a 
		concave indent.}
	\label{fig:seed_crystals}
\end{figure}

This has some implications: Firstly, the lack of supposed 
glycine droplets suggests that the presence of an interface 
is crucial in order to promote the formation of these droplets. 
This is consistent with previous reports on laser nucleation, 
where a air-solution or solution-glass interface is necessary 
for the trap focus to have any noticeable effect on the solute \cite{Gowayed2021, Liao2022, Rungsimanon2010}. Secondly, the 
fact that the seed crystal dissolved in a saturated solution 
suggests that this could be due to a localised heating effect. 
A commonly used heating model for optical tweezers is the 
Peterman model \cite{Peterman2003} which accounts for the lateral 
distance between the optical trap and a local heat sink (such 
as the glass cover slip or microscope objective). Even with 
the heat sink the predicted temperature rise in pure water would 
be around $5-10 K$.  The fact that the seed crystal dissolves 
so slowly could simply be due to the fact that the local heating 
is lessened by the moving beam front.

\section{Summary of Moving Beam Phenomena}
To summarise, the introduction of a moving beam helps to 
accelerate the local growth of a newly formed crystal. It 
seems that this phenomena will only occur when close to 
interface between the solution and air. We note the presence
of droplets that can sometimes be seen entering the trap, 
though this is not always necessary for a nucleation event 
to occur. A plausible description of the phenomena is described
thusly.

Initial nucleation is similar to typical optical trapping 
induced nucleation, with the air solution interface limiting 
the molecular mobility of the solute molecules \cite{Liao2022, 
Sugiyama2009, Gowayed2021}. The moving beam front can influence 
the motion of the nucleus initially, but eventually the drag 
force means the crystal is not moved by the optical trap. 
Localised crystal growth occurs when the trap is close to or 
partially over the interface of the crystal (see fig.~
\ref{fig:local_nucleation}(a)). This suggests that the 
laser itself is bringing in new solute material that can then
adhere to the crystal surface. Interestingly even when the 
crystal growth is localised to a small section of the crystal
we do not observe the crystal dissolving. This is consistent 
with other observed laser induced nucleation results, where 
as long as the laser is active the crystal remains stable 
within the solution \cite{Flannigan2023, Rungsimanon2010,
Liao2022, Sugiyama2009}. It does raise further questions 
however, mainly in an undersaturated solution is a moving 
beam front able to increase the size of the formed nucleus
compared to a stationary beam? Additionally, it raises the 
question about what the role the 'glycine droplets' play in 
the nucleation process.

As shown in \ref{sec:clusters}, the optical trap can manipulate 
these droplets similar to microspheres. When in close proximity 
to the trap these droplets are brought towards the crystal 
surface (see fig.~\ref{fig:local_nucleation}(b)). We have 
already suggested that these droplets must contain glycine as 
they are to large to be a silica microsphere, and the solutions 
were filtered prior to being studied. The droplets provide 
material that grows the crystal around that region (see fig.~\ref{fig:local_nucleation}(c)). Eventually the local 
solution is either depleted of solute material or the crystal 
front has grown to fully encompass the trap, preventing further 
growth (see fig.~\ref{fig:local_nucleation}(d)). 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{galvano_diagram.pdf}
	\caption{Diagram outlining how a moving beam assists in the 
		growth of a crystal nucleus. (a) a crystal nucleus is 
		partially trapped by a moving beam with solute droplets 
		close to its surface. (b) droplets close to the laser
		focus can be drawn in by gradient forces and moved 
		towards the crystal surface. (c) the laser brings in 
		material that is then deposited on the surface of the 
		crystal. (d) eventually the crystal area either fully 
		surrounds the laser focus or the solution surrounding 
		the laser is depleted of solute material.}
	\label{fig:local_nucleation}
\end{figure}

What remains unclear is why the solution interface is necessary
for localised crystal growth. We know that the glycine droplets 
will form when the laser is close to the glass-solution interface
\cite{Gowayed2021, Yuyama2010, Yuyama2012}. This is consistent 
with our results as being able to image the droplet edge requires 
that the laser is brought close to the cover slip. However it is 
not clear how exactly these droplets contribute to localised 
crystal growth as in sec.~\ref{sec:moving} no droplets are seen 
entering the trap. It seems likely that the presence of droplets 
is necessary as we see that the seed crystals dissolve while 
there are no droplets present.

There are still several factors that need to be investigated. 
Firstly, there is the question of what conditions result in the 
production of concentrated droplets, it is not clear if the 
presence of a laser is required or if these droplets naturally 
occurring. Prior literature would suggest that the laser is 
required \cite{Liao2022, Tsuboi2009}, but this would not explain 
why in many cases the droplets are found far outside the influence 
of the optical trap. It has been shown that phase separation 
occurs when a trap is focused close to the cover slip 
\cite{Gowayed2021, Yuyama2010, Yuyama2012}, but there has been 
no cases of nucleation occurring at the cover slip. Furthermore
these experiments all occurred in supersaturated condition 
($S\ge 1.50$) with a much wider focus ($NA \le 1.00$) whereas 
our results occurred in undersaturated conditions with a 
tighter focus. We did not not any occurrence of a phase 
separation as reported by \cite{Gowayed2021, Yuyama2010, Yuyama2012}
and yet we see the appearance of droplets that we surmise 
are partially constructed of glycine.

Secondly, there is the question of how these droplets supply 
material to the bulk crystal. In some instances it is clear 
that the droplets are being drawn into the trap, however, in 
other instances while there are no droplets close to the 
vicinity of the optical trap the crystal continues to grow. 
If these droplets are a necessary precursor to induce crystal 
nucleation then understanding how they provide materials to 
the bulk crystal may help with our understanding of the 
kinetics of multi-step nucleation.

\section{Conclusion}
In conclusion, the experimental results from this chapter 
provide no affirmative evidence of shear induced crystal 
nucleation. This is in part due to the limited rotational 
speed that could be achieved using vaterite microspheres. 
Simply adding a micro-rotor to a supersaturated solution 
will not yield a significant fluid flow to induce crystal 
nucleation. Furthermore, the region in which the local 
shear rate is significant enough is so small that any 
nucleation events that could occur there may not grow large 
enough to be stable. This does not however, suggest that 
shear induced nucleation is impossible at a micro scale. 
In fact there have been several developments lately that
allow for the precise rotation of multiple rotors to 
control the precise motion of silica beads \cite{Butaite2019}. 
This would allow for direct control of the shear rate 
within a given volume of fluid, allowing one to study the 
impact of fluid flow on a nucleus at a micro-scale.

On the other hand, the results utilising the galvano mirror 
have some interesting implications for future research into 
crystal nucleation using optical tweezers. We noted that a 
moving beam front is capable of localising the crystal growth
around the beam focus in undersaturated conditions. Typically
in laser induced nucleation the final crystal morphology is 
influenced primarily by the local fluid conditions 
\cite{Flannigan2023, Liao2022}. If this phenomena could be 
fully understood we could have a means of directly manipulating 
the morphology of any nucleating crystal. We noted that in 
undersaturated conditions the presence of droplets that could 
be trapped and later nucleated when brought into contact with 
one another. There remains some uncertainty over the role these 
droplets play in a nucleation event. Further research is also 
necessary to understand how the fluid and laser parameters 
(i.e. supersaturation and numerical aperture) impact the 
formation of these droplets. The fact that these droplets could 
be trapped suggests they have a higher refractive index to the 
solution, given that the solutions were filtered prior to 
irradiation dismisses that they could be dust particles. Prior 
literature suggests that optical trapping of a supersaturated 
solution can lead to the production of nano droplets 
\cite{Gowayed2021, Tsuboi2009}.

### 2025-07-04 Latex_files/Thesis/chapters/chapter 5/main.tex
fleshing out QPD chapter and edits to new chp 3&6

@@
 
-\section{Overview}
-In this chapter, we begin with a brief discussion of the working 
-principle behind a position detection device. We then provide a 
-framework for replicating the scattering signal recorded by a 
-quadrant photo diode. We show that extending the framework to 
-capture information from an arbitrary shaped particle is simple 
-in principle but difficult to interpret, we demonstrate as such 
-by comparing the expected signal from a trapped sphere, to that 
-of a trapped dimer.
-
 \section{Introduction}
 
-Beyond merely simulating the forces experienced by a spherical 
-aggregate we also developed a simulative quadrant photo diode 
-to replicate the results from a typical calibration test. This 
-builds upon the work from \cite{Rohrbach2002} which applied 
-Lorenz-Mie theory to replicate the response signal of a QPD 
-being used in back focal-plane interferometry. Rather than be 
-constrained to individual spheres we can now consider the 
-expected response from any type of spherical aggregate. The
-goal of such simulation software would be to simulate the 
-expected experimental results from a complex particle whose 
-exact shape and size may be difficult to determine.
+As demonstrated in Chapter 3, the motion and behaviour of a dimer 
+is significantly different compared to that of a traditional sphere.
+While this is simple enough to demonstrate in a simulation, where the 
+position and orientation can be computed using the finite differences
+method, from an experimental perspective these details can go unnoticed.
+In order to accurately capture the dynamics of a trapped sphere, a 
+position detection system is typically used. This can capture the 
+translational motion of a trapped particle but relating the signals
+produced to the rotational motion is not as straight forward.
 
+The relationship between particle translations and light scattering 
+was first demonstrated by Rohrbach, where they showed a linear change 
+in the scattered light when the particle was displaced \cite{Rohrbach2002}. They further showed that the optical force exerted by the trapping 
+beam could be determined from the change in scattered light \cite{Rohrbach2002}. Rohrbach's model would later be used to develop 
+modern position detection systems, where by calibrating the scattering 
+signal using one of the many calibration techniques one could achieve a highly accurate estimation of the translational motion in the $X$-$Y$ 
+plane.The most common of which being the Power Spectral Density method, 
+one of the greatest advantages of the PSD method is that periodic 
+rotational motion is also captured in the power spectrum. As an 
+example consider an elongated cylindrical particle that is wobbling
+around the $Z$-axis at some periodic frequency 'f'. A power spectrum 
+of that trajectory will show a peak at that frequency. If however the 
+wobble is periodic but instead stochastic, then it is impossible to 
+separate out the rotational motion from the translational motion. This
+poses some challenges to experimental modelling of non-spherical 
+particles, if we cannot glean any information on the orientation
+of the trapped particle we cannot accurately distinguish between
+sphere's and dimers
+
+As a solution we developed a model similar to that of Rohrbach's but 
+now utilising a \textit{mstm} to provide the light scattering for a 
+dimer. We begin with a brief discussion of the working principle 
+behind a position detection device. We then provide a framework for replicating the scattering signal recorded by a quadrant photo diode. 
+We show that extending the framework to capture information from an 
+arbitrary shaped particle is simple in principle but difficult to 
+interpret, we demonstrate as such by comparing the expected signal 
+from a trapped sphere, to that of a trapped dimer.
+
+\subsection{Position Detection methods}
+There are 3 common methods of position detection: video-analysis, lateral-effect position sensing, and photodiodes. The former is 
+relies purely on cameras to track particle displacements; whereas 
+the latter two are examples of focal-plane interferometry. 
+Interferometry utilises the interference between the scattered and un-scattered light to measure displacement. Here we 
+provide a brief outline of their applications, and limitations.
+
+Video analysis is ideally suited for multiple traps or situations where 
+precision is not the top priority. In order for video analysis to 
+match the accuracy of back-focal plane interferometry methods requires 
+the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
+achieve while maintaining a decent resolution \cite{Gibson2008}. In 
+comparison off the shelf back-focal plane detectors can achieve temporal 
+resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
+
+A quadrant photo diode (QPD) is a frequently used position detection 
+system for optical tweezers due to their high sampling rate, high 
+degree of precision, and ease of set up. The QPD is constructed of 
+four photo diodes assembled in a quadrant formation, when a particle 
+is trapped the interference pattern produced is focused onto the QPD, 
+with the maximum intensity mapping to the particle's centre of mass. 
+By summing the voltages of the horizontal and vertical quadrants together 
+the particle's centre of mass is tracked in the $x-y$ plane. Axial 
+displacement can be estimated by observing the change in the total 
+voltage of the QPD. The outputted signal gives an indication of the 
+particle's relative displacement from the beam focus, but in order to 
+convert the signal to distance units the trap needs to be calibrated 
+(assuming a linear response curve).
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
+	\caption{Comparison between QPD and Lateral effect photodiodes.
+		The four quadrants of a QPD (left) experience different photocurrents
+		based on the total intensity of light incident on each section 
+		(labelled A, B, C, D). 
+		Whereas a Lateral effect sensor (right) uses the resistive properties
+		of the photodiode surface to vary the create different photocurrents 
+		passing through the anodes A, B, C, and D.}
+\end{figure}
+
+A lateral-effect sensor works using a single cell analogous to the focal plane of the trapping beam.
+The four corners of the sensor act as anodes connected to a base plate cathode, as the beam moves across the surface of the detector each anode will experience a different photocurrent depending
+on how close the centre of the interference pattern is to each anode. 
+The advantage of a lateral effect detector is that the linear regime 
+is much larger than a QPD making it much better for monitoring the 
+position of a trapped particle. However, Lateral-effect sensors are 
+often limited in their spacial resolution due to high signal-to-noise 
+ratios, requiring a high intensity of light on the sensor in order to 
+get a clean signal. As a result, most optical force measurements are 
+conducted using a QPD as opposed to a lateral-effect sensor, as often 
+the displacement is small enough that the signal-displacement curve 
+can be considered linear.
+
+The above informs how one can - to their best ability - measure small 
+rotational movements in a trajectory. Relying on video analysis is 
+unreliable as the frame rate required to accurately measure the 
+orientation results in a loss of resolution. Likewise, simply measuring
+the total intensity provides no means of separating out rotational and 
+translational motion.
+Similar to Chapter 4, we need to utilise contrasting signals to estimate the change in orientation, and see if at all it is possible to distinguish between translational and rotational motion.
+
+\section{Modelling the QPD response signal of a non-spherical particle}
 In order to simulate a typical experimental set up with a QPD 
@@
 the outer limit by which we evaluate the electric field, this is 
-analogous to our condenser lens removing noise from other light 
+analogous to a condenser lens removing noise from other light 
 sources by only accepting light at a specific acceptance angle 
@@
 The incident beam is simple enough to define given our set up 
-parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
-beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
-\textit{Ott} uses a point matching approach to approximate the beam shape 
-coefficients of the incident field by fitting it to the far field estimate. 
-From the QPD's perspective it is receiving light both from the incident 
-and scattered beam simultaneously, as such both fields must be expressed
-using outgoing vector spherical harmonics.
+parameters, for the sake of simplicity we assume that our beam is a 
+Gaussian beam. \textit{Ott} uses a point matching approach to 
+approximate the beam shape coefficients of the incident field by 
+fitting it to the far field estimate. From the QPD's perspective 
+it is receiving light both from the incident and scattered beam simultaneously, as such both fields must be expressed using outgoing 
+vector spherical harmonics.
 \begin{align}
@@
 long enough time period. By converting from signal units to length
-units (see \eqref{eq:conversion_factor}) the trap shape can be 
+units (see Eq.\eqref{eq:conversion_factor}) the trap shape can be 
 discerned from the QPD signals, assuming the fluid properties are
@@
 
-To confirm that our method is producing accurate results, we ran a comparison
-between our simulative QPD and the results from \cite{Rohrbach2002}. Where a 
-$300\ nm$ diameter sphere is scanned across the path of a focused Gaussian beam
-($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is
-suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
-set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
-produced the following response curve:
+To confirm that our method is producing accurate results, we ran a 
+comparison between our simulative QPD and the results from \cite{Rohrbach2002}. Where a $300\ nm$ diameter sphere is scanned 
+across the path of a focused Gaussian beam ($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
+set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis produced the following response curve:
 \begin{figure}[h!]
@@
 	\caption{Comparison between QPD response signal versus work conducted by 
-		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal 
-		produced by QPD using \textit{ott} and points represent the signal 
-		response collected from \cite{Rohrbach2002}.}
+		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded.
+Solid lines represent the signal produced by QPD using \textit{ott} and points represent the signal response collected from \cite{Rohrbach2002}.}
 	\label{fig:Rohrbach}
@@
 the error between Rohrbach's model and our own is inconsequential 
-for calibration purposes. With this any trajectory can be collected 
-from the QPD by displacing and rotating the beam accordingly, 
-later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of 
-accuracy that can be achieved using back focal plane interferometry 
-to characterise non-spherical targets.
-
-\subsection{Position detection}
-\label{sec:position_detection}
-In order to accurately capture the dynamics of a trapped particle, a 
-position detection system is required. There are 3 possible methods 
-of position detection: video-analysis, lateral-effect position sensing, 
-and photodiodes.
-
-Video analysis is ideally suited for multiple traps or situations where 
-precision is not the top priority. Whereas lateral-effect and photodiode 
-position detection are two examples of back-focal plane interferometry, 
-where the interference pattern produced by the target particle is 
-extrapolated to determine its position. In order for video analysis to 
-match the accuracy of back-focal plane interferometry methods requires 
-the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
-achieve while maintaining a decent resolution \cite{Gibson2008}. In 
-comparison off the shelf back-focal plane detectors can achieve temporal 
-resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
-
-A lateral-effect sensor has a similar output but works using a the 
-entire sensor as a single cell analogous to the focal plane of the 
-trapping beam. The four corners of the sensor act as anodes connected 
-to a base plate cathode, as the beam moves across the surface of the 
-detector each anode will experience a different photocurrent depending 
-on how close the centre of the interference pattern is to each anode. 
-The advantage of a lateral effect detector is that the linear regime 
-is much larger than a QPD making it much better for monitoring the 
-position of a trapped particle. However, Lateral-effect sensors are 
-often limited in their spacial resolution due to high signal-to-noise 
-ratios, requiring a high intensity of light on the sensor in order to 
-get a clean signal. As a result, most optical force measurements are 
-conducted using a QPD as opposed to a lateral-effect sensor, as often 
-the displacement is small enough that the signal-displacement curve 
-can be considered linear.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
-	\caption{Comparison between QPD and Lateral effect photodiodes.
-		The four quadrants of a QPD (left) experience different photocurrents
-		based on the total intensity of light incident on each section 
-		(labelled A, B, C, D). 
-		Whereas a Lateral effect sensor (right) uses the resistive properties
-		of the photodiode surface to vary the create different photocurrents 
-		passing through the anodes A, B, C, and D.}
-\end{figure}
-
-A quadrant photo diode (QPD) is a frequently used position detection 
-system for optical tweezers due to their high sampling rate, high 
-degree of precision, and ease of set up. The QPD is constructed of 
-four photo diodes assembled in a quadrant formation, when a particle 
-is trapped the interference pattern produced is focused onto the QPD, 
-with the maximum intensity mapping to the particle's centre of mass. 
-By summing the voltages of the horizontal and vertical quadrants together 
-the particle's centre of mass is tracked in the $x-y$ plane. Axial 
-displacement can be estimated by observing the change in the total 
-voltage of the QPD. The outputted signal gives an indication of the 
-particle's relative displacement from the beam focus, but in order to 
-convert the signal to distance units the trap needs to be calibrated 
-(assuming a linear response curve).
+for calibration purposes.
 
 \section{Characterisation of asymmetric dimer dynamics via PSD analysis}
-As discussed in \ref{sec:simulated_QPD}, one of the methods 
+As discussed in Sec.~\ref{sec:simulated_QPD}, one of the methods 
 developed to work in conjunction with \cite{Vigilante2020} 

### 2025-07-04 Latex_files/Thesis/chapters/chapter 6/main.tex
fleshing out QPD chapter and edits to new chp 3&6

@@
 surface tension of the surrounding fluid. This is reflected in 
-fig.~\ref{fig:free_energy} where we plot the free energy of the 
+Fig.~\ref{fig:free_energy} where we plot the free energy of the 
 system against nucleus size. This reveals a critical size above 
 which the gain in free energy exceeds the interfacial tension. 
-Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
+Furthermore Fig.~\ref{fig:free_energy} shows how increasing the 
 supersaturation of the system reduces said barrier. 
@@
 results regardless of the system parameters (i.e. varying 
-supersaturation, temperature, and solute choice). There have 
-been several different methods deployed to try and observe 
-crystallisation at a microscopic scale.
-
-\subsection{Computer Simulations}
-While not a direct observation of nucleation events, computer
-simulations have proved an invaluable tool for studying and 
-testing predictions about the pre-nucleated solution. By 
-comparing the simulative nucleation rate to the expected 
-rate given by current models, researchers can test newly 
-developed theories and glean information about the microscopic
-parameters of a system.
-
-The main challenge facing computer simulations is the issue
-of limited scope. Molecular dynamic simulations are constrained
-by their choice of time and length scale, for example simulations
-in a canonical ensemble face issues where the chemical potential
-is depleted to the point that the driving force for crystallisation
-is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
-addressed but requires an extensive increase in computational 
-resources to maintain the chemical potential equilibria. Like wise
-in most cases the choice of time scale is a crucial factor, shorter
-time scales provide a more accurate evolution of the solution; 
-however, in some cases the time scale magnitude for crystallisation
-is often far greater than the time scale used in the simulation. This
-has led to more advanced sampling techniques that allow for longer 
-time scales which has provided insights into the pre-nucleation 
-solution stage and formation of pre-nucleated clusters \cite{Finney2023}.
-
-The study of the pre-nucleated solution and the evolution of 
-precursors has been aided greatly by the use of simulative 
-studies. Early on it was shown in colloidal crystallisation 
-that the structure of the precursor was time dependent; 
-colloidal simulations showed that regardless of the molecular
-interactions between colloids the atomic packing pattern was 
-predominately hexagonal centred, over time the packing
-became dominated by body centred packing in the case 
-where the interactions were longer reaching \cite{Tan2013}. 
-They also confirmed that precursor formation was driven not
-by local density variations, but instead was driven by the 
-local bond order of the colloidal precursors \cite{Tan2013}, 
-providing strong evidence that the colloidal pre-nucleation 
-clusters undergo rearrangement prior to nucleation. Further
-developments in $NaCl$ simulations have shown that by varying
-the density fluctuations the precursor structure could either
-be amorphous, the classical rock-salt structure, or a new 
-structure similar to wurtzite. While this result is difficult
-to confirm it provides a important conclusion, that simulations
-can hint at new nucleation pathways that have not been confirmed
-experimentally. This means that computer simulations can not
-be merely used to confirm theories, but can be used to examine
-and search for novel results.
-
-Overall computer simulations make up a core area of research 
-in the study of nucleation. The technique is however limited,
-often struggling to examine nucleation over longer time scales
-or larger number of particles. While sampling methods and 
-enhanced simulation techniques has alleviated this somewhat 
-they are still far from being able to describe the behaviour 
-of larger molecules undergoing nucleation.
-
-\subsection{Transmission Electron Microscope (TEM)}
-One of the more well known techniques used for direct observations 
-of crystallisation dynamics is using transmission electron 
-microscopy (TEM). The basic working principle involves creating a 
-focused beam of electrons that are directed onto a target sample.
-Due to the wave-particle duality, the electrons can be treated 
-as a unified beam of light, whose wavelength is dependent on the 
-kinetic energy of the electrons. The wavelength of an individual 
-electron is given by \cite{Williams2009}:
-\begin{align}
-	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
-\end{align}
-
-\noindent
-When incident on a sample the wave is scattered by the sample 
-and can then be detected by an imaging screen that measures 
-the difference in intensity between the scattered and 
-transmitted beams. The advantage of TEM is that one can bypass
-the diffraction limit imposed by visible light, the maximum 
-resolution that can be achieved by a microscope is given by
-\cite{Champness2020}:
-\begin{align}
-	d = \frac{\lambda}{2NA}
-\end{align}
-
-\noindent
-where NA is the numerical aperture of the objective lens, and
-$\lambda$ is the wavelength of light used to illuminate a sample.
-For a visible light LED with an average wavelength of $550~nm$, 
-the upper limit for an objective lens (NA of 1.2) is around 
-$230~nm$ whereas a similar lens focusing an electron beam 
-with an average kinetic energy of $100~keV$ will have a maximum 
-resolution of just over $4~pm$ \cite{Champness2020}. There is
-an upper limit to the maximum resolution possible as electron 
-lens are far inferior to what modern optics can achieve 
-\cite{Champness2020} so there is a trade off between quality and 
-resolution.
-
-TEM technology has proven invaluable to the study of nucleation
-at an atomic level. The samples are often only a few hundred 
-atoms thick in order to optimise the imaging results. Much of 
-the supporting evidence for non-classical nucleation has been 
-demonstrated via TEM \cite{Ye2023}. For example when Cao 
-\textit{et al} observed the formation of amorphous atomic 
-cluster using $\gamma-$Fe,Au, and Re; all three of which 
-underwent continuous transformation before eventually 
-reaching an final ordered state \cite{Cao2020}. The amorphous 
-cluster did not immediately transition to an ordered structure, 
-instead fluctuating somewhat randomly until eventually settling 
-into a crystalline structure \cite{Cao2020}. This was confirmed 
-both in solid growth and liquid growth conditions \cite{Cao2020, 
-	Loh2016}, suggesting that the non-classical nucleation path 
-dominates in all cases of metal crystal formation \cite{Ye2023}. 
-Confirmation is necessary for different solute choices, as in 
-the case of organic solids it is believed that non-classical 
-nucleation only dominates in low supersaturations whereas 
-classical nucleation is dominate at higher supersaturations.
-
-TEM analysis has not been extended to the study of organic 
-crystals, proteins, or simple salts at an atomic level 
-however. The main reason being that organic materials can be 
-damaged by the electron beam and therefore need to be frozen 
-prior to being imaged \cite{Champness2020}. This was used to 
-study the polymorphic time dependence of glycine crystals, by 
-varying the time prior to freezing Broadhurst \textit{et al} 
-showed that $\beta$ glycine dominates initially but $\alpha$-
-glycine is more stable after a longer period, $\gamma$-glycine 
-was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
-the time scales with which polymorphs form, it doesn't provide 
-much information on the kinetics of crystal growth as these 
-were snap shots and not direct observations of the nucleation 
-process.
-
-Overall TEM analysis is invaluable for the study of atomic 
-crystal growth, the secondary scattering from electron 
-beams also allows for chemical and structural information 
-to be gleaned from the target. This has given strong evidence
-for the non-classical nucleation theory, particularly in 
-metal crystal formation. Due to the high energy of the 
-electrons, organic materials have not received the same
-focus outside of cryoTEM analysis. Non-invasive methods
-are therefore necessary to gain an understanding of the
-crystal growth and formation of organic crystals, 
-particularly when both classical and multi-step nucleation 
-can occur.
-
+supersaturation, temperature, and solute choice).
+
+This was one of the initial goals of the PhD, to localise nucleation
+using optical tweezing. While tweezing by itself can induce nucleation
+the exact mechanism that causes it is still unclear. Optical interactions
+with sub-wavelength particles are not possible so only a two-step 
+nucleation event can be occur via optical tweezing. In order to accurately 
+characterise the nucleation pathway taken at a given supersaturation, 
+a method is needed that is not dependent on the pathway taken. This is 
+what lead us to consider localised shearing, while still favoured by 
+two-step nucleation it is still possible for the classical nucleation 
+pathway to occur, given that the free energy barrier is low enough to 
+be traversed via localised shearing.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:4f_correlation}
-As shown in \ref{fig:setup}, after the galvano-mirror there are a pair
+As shown in Fig.~\ref{fig:setup}, after the galvano-mirror there are a pair
 of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
@@
 is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
-An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.
+An example of this is behaviour is shown in Fig.~\ref{fig:tdat_fdat}.
 
@@
 the sphere is rotating sufficiently quickly there are three 
-sections of concern, as demonstrated by fig.~\ref{fig:shear_diagram}. 
+sections of concern, as demonstrated by Fig.~\ref{fig:shear_diagram}. 
 The fluid close to the sphere experiences a shear rate that 
@@
 cover slip \cite{BergSoerensen2004}. We estimate that the 
-vaterite must be at least $15-20\mu\ m$ from the cover slip for 
+vaterite must be at least $15-20\ \mu m$ from the cover slip for 
 us to reliably trap them. The contact angle between the cover 
@@
 measured contact angle for supersaturated solutions was between 
-$30^{\circ}$ and $35^{\circ}$. This is shown in fig.\ref{fig:3.9}.
+$30^{\circ}$ and $35^{\circ}$. This is shown in Fig.\ref{fig:3.9}.
 \begin{figure}[h!]
@@
 interface due to increased surface tension and unpredictable 
-scattering forces. The closest we could trap a microsphere to 
-the droplet edge was in the range of $5-10\mu\ m$, at that 
-distance the fluid flow is so low that even the presence of a 
-hard boundary would be insufficient for shearing the fluid. 
-Furthermore, as is evident in Table~\ref{tab:Nucleation}, 
-the rotation rate drops off significantly with increased 
-supersaturation, due to higher fluid viscosities. While in 
-theory a sufficiently focused laser could rotate any 
-microsphere to a fast enough to reach the shear rate 
-predicted by \cite{Debuysschere2023} the localised intensity 
-would be so large that even using $D_2O$ would see a significant 
-increase in temperature.
+scattering forces. The drag coefficient is estimated to increase
+by 10\% when approaching a flat boundary \cite{Yao2009} so approaching 
+an angled surface would see an even greater increase.  The 
+closest we could trap a microsphere to the droplet edge was 
+in the range of $5-10\ \mu m$, at that distance the fluid flow 
+is so low that even the presence of a hard boundary would be 
+insufficient for shearing the fluid. Furthermore, as is evident 
+in Table~\ref{tab:Nucleation}, the rotation rate drops off 
+significantly with increased supersaturation, due to higher 
+fluid viscosities. Solution viscosity has been shown to increase 
+with a slight addition of glycine \cite{Patyar2020}, though as 
+best as we can find there have been no studies of solution 
+viscosity when the concentration reaches saturation levels. 
+One would suspect that when $S\geq1$ the chemical potential 
+driving the solute molecules to nucleate would also somewhat 
+resist the velocity field leading to an additional increase in 
+viscosity.
+
+While in theory a sufficiently focused laser could rotate any 
+microsphere to a fast enough to reach the shear rate predicted 
+by \cite{Debuysschere2023} the localised intensity would be so 
+large that even using $D_2O$ would see a significant increase 
+in temperature.
 
@@
 small to see any noticeable increase in the nucleation rate 
-this is demonstrated most clearly in fig.~\ref{fig:vaterite_shear}. 
+this is demonstrated most clearly in Fig.~\ref{fig:vaterite_shear}. 
 Due to the exponential decay in the expected shear rate it is 
@@
 a fluid could produce significantly greater shear rates 
-(see \emph{Appendix X.X} for breakdown) than a rotating 
-particle. While a particle is within the optical trap the 
-gradient forces are significantly reduced meaning simply 
-trapping a particle is insufficient for inducing nucleation \cite{Flannigan2023}. This is why we did not see any 
-nucleation events while rotating the vaterite, there is 
-not a large gradient force to draw in material.
+(see Appendix\ref{sec.shear_rate-calcs} for breakdown) 
+than a rotating particle. While a particle is within the 
+optical trap the gradient forces are significantly reduced 
+meaning simply trapping a particle is insufficient for 
+inducing nucleation \cite{Flannigan2023}. This is why we did 
+not see any nucleation events while rotating the vaterite, 
+there is not a large gradient force to draw in material.
 
@@
 
-\subsubsection{Choice of Crystal growth rate units}
+\subsection{Choice of Crystal growth rate units}
 The growth rate of a crystal can be described using different
@@
 focus a nucleus forms close to the droplet edge, 
-unlike in fig.~\ref{fig:stationary_beam} the crystal 
+unlike in Fig.~\ref{fig:stationary_beam} the crystal 
 does not grow out from the focal point evenly. Due to 
@@
 Interestingly the galvano-mirror allows the trap to 
-impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
+impart a slight torque on the crystal, as shown in Fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
 though the crystal is not directly in the trap focus 
@@
 dipole moment of individual water molecules is too 
-small to be influenced by an optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
+small to be influenced by an optical trap. In Figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
 crystal growth becomes localised to the corner. The 
-area growth rate between figures \ref{fig:eliptical_beam_1}
+area growth rate between Figures \ref{fig:eliptical_beam_1}
 (a) and (d) was approximated using imageJ at $45.03\ 
-\mu m^2 /min$, where as between figures 
-\ref{fig:eliptical_beam_1}(e) and (f) the growth rate 
-at that particular edge was estimated at $42.10\ \mu m^2/min$.
+\mu m^2 /min$, where as between Figures~.\ref{fig:eliptical_beam_1}
+(e) and (f) the growth rate at that particular edge was estimated 
+at $42.10\ \mu m^2/min$.
 
@@
 crystal nucleus, in some cases allowing for growth out
-of the viewing plane as shown in fig.~\ref{fig:out_of_plane}.
+of the viewing plane as shown in Fig.~\ref{fig:out_of_plane}.
 \begin{figure}[h!]
@@
 close to the droplet edge, this time the galvano 
-mirror was scanning a circular path (as shown in fig.~\ref{fig:cluster_trapping}). After a few 
+mirror was scanning a circular path (as shown in Fig.~\ref{fig:cluster_trapping}). After a few 
 minutes of irradiation droplets were seen entering 
@@
 Trying to bring two droplets together resulted in 
-nucleation between the two droplets, compared to 
+nucleation between the two droplets, compared to sections
 \ref{sec:stationary} \& \ref{sec:moving} the growth 
@@
 
-As shown by fig.~\ref{fig:seed_crystals} the laser does not 
+As shown by Fig.~\ref{fig:seed_crystals} the laser does not 
 in fact promote crystal growth but instead forces the crystal 
@@
 Localised crystal growth occurs when the trap is close to or 
-partially over the interface of the crystal (see fig.~
+partially over the interface of the crystal (see Fig.~
 \ref{fig:local_nucleation}(a)). This suggests that the 
@@
 to the trap these droplets are brought towards the crystal 
-surface (see fig.~\ref{fig:local_nucleation}(b)). We have 
+surface (see Fig.~\ref{fig:local_nucleation}(b)). We have 
 already suggested that these droplets must contain glycine as 
@@
 were filtered prior to being studied. The droplets provide 
-material that grows the crystal around that region (see fig.~\ref{fig:local_nucleation}(c)). Eventually the local 
+material that grows the crystal around that region (see Fig.~\ref{fig:local_nucleation}(c)). Eventually the local 
 solution is either depleted of solute material or the crystal 
 front has grown to fully encompass the trap, preventing further 
-growth (see fig.~\ref{fig:local_nucleation}(d)). 
+growth (see Fig.~\ref{fig:local_nucleation}(d)). 
 \begin{figure}[h!]

### 2025-07-04 Latex_files/Thesis/main_win.tex
updates to chapter 4, among others
27/06/2025 chapters 1,3,4 and structural changes
fleshing out QPD chapter and edits to new chp 3&6

@@
 \usepackage[hidelinks]{hyperref}
-%\usepackage{cite}
+\usepackage{textalpha}
 \usepackage{enumitem}
@@
 
+\input{chapters/chapter 6/main}
+
 \chapter{Closing Remarks}
@@
 \section{Estimation of fluid shear rates via galvano mirror}
+\label{sec.shear_rate_calcs}
 An alternative approach to generating fluid shear is to use a 
@@
 \section{Approximation of Area growth rate}
+\label{sec.growth_rate_calcs}
 Data collected from \cite{Li1992} provides face growth rates 

### 2025-07-11 Latex_files/Thesis/chapters/chapter 5/main.tex
updates to QPD chapter

@@
 \section{Introduction}
-
 As demonstrated in Chapter 3, the motion and behaviour of a dimer 
@@
 In order to accurately capture the dynamics of a trapped sphere, a 
-position detection system is typically used. This can capture the 
-translational motion of a trapped particle but relating the signals
-produced to the rotational motion is not as straight forward.
+position detection system is typically used - the most common of which is 
+a Quadrant Photo Diode (QPD). A QPD sends a signal measuring 
+the voltage difference between 4 photo diode quadrants. For purely 
+translational motion the diodes can be split into neighbouring pairs 
+(i.e. left and right, top and bottom - see Fig~\ref{fig:QPD_diagram})
+to estimate the total displacement made by the target particle. 
+However the same method cannot be applied to stochastic rotational 
+motion.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Position Detection methods}
-There are 3 common methods of position detection: video-analysis, lateral-effect position sensing, and photodiodes. The former is 
-relies purely on cameras to track particle displacements; whereas 
-the latter two are examples of focal-plane interferometry. 
-Interferometry utilises the interference between the scattered and un-scattered light to measure displacement. Here we 
-provide a brief outline of their applications, and limitations.
-
-Video analysis is ideally suited for multiple traps or situations where 
-precision is not the top priority. In order for video analysis to 
-match the accuracy of back-focal plane interferometry methods requires 
-the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
-achieve while maintaining a decent resolution \cite{Gibson2008}. In 
-comparison off the shelf back-focal plane detectors can achieve temporal 
-resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.
-
-A quadrant photo diode (QPD) is a frequently used position detection 
-system for optical tweezers due to their high sampling rate, high 
-degree of precision, and ease of set up. The QPD is constructed of 
-four photo diodes assembled in a quadrant formation, when a particle 
-is trapped the interference pattern produced is focused onto the QPD, 
-with the maximum intensity mapping to the particle's centre of mass. 
-By summing the voltages of the horizontal and vertical quadrants together 
-the particle's centre of mass is tracked in the $x-y$ plane. Axial 
-displacement can be estimated by observing the change in the total 
-voltage of the QPD. The outputted signal gives an indication of the 
-particle's relative displacement from the beam focus, but in order to 
-convert the signal to distance units the trap needs to be calibrated 
-(assuming a linear response curve).
+Here we discuss how one can record a trajectory using a QPD and 
+use it to interpret the particle's real time motion. The QPD is 
+constructed of four photo diodes assembled in a quadrant formation, 
+when a particle is trapped the interference pattern produced is 
+focused onto the QPD, with the maximum intensity mapping to the 
+particle's centre of mass. By summing the voltages of the horizontal 
+and vertical quadrants together the particle's centre of mass is 
+tracked in the $x-y$ plane. Axial displacement can be estimated 
+by observing the change in the total voltage of the QPD. The 
+outputted signal gives an indication of the particle's relative 
+displacement from the beam focus, but in order to convert the 
+signal to distance units the trap needs to be calibrated (assuming 
+a linear response curve).
 
@@
 	\centering
-	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
-	\caption{Comparison between QPD and Lateral effect photodiodes.
-		The four quadrants of a QPD (left) experience different photocurrents
-		based on the total intensity of light incident on each section 
-		(labelled A, B, C, D). 
-		Whereas a Lateral effect sensor (right) uses the resistive properties
-		of the photodiode surface to vary the create different photocurrents 
-		passing through the anodes A, B, C, and D.}
+	\includegraphics[height=0.5\linewidth, angle=270]{QPD_diagram.png}
+	\caption{Diagram depicting a typical QPD interface. The four 
+		quadrants of a QPD experience different photocurrents based 
+		on the total intensity of light incident on each section 
+		(labelled A, B, C, D).}
+	\label{fig:QPD_diagram}
 \end{figure}
 
-A lateral-effect sensor works using a single cell analogous to the focal plane of the trapping beam.
-The four corners of the sensor act as anodes connected to a base plate cathode, as the beam moves across the surface of the detector each anode will experience a different photocurrent depending
-on how close the centre of the interference pattern is to each anode. 
-The advantage of a lateral effect detector is that the linear regime 
-is much larger than a QPD making it much better for monitoring the 
-position of a trapped particle. However, Lateral-effect sensors are 
-often limited in their spacial resolution due to high signal-to-noise 
-ratios, requiring a high intensity of light on the sensor in order to 
-get a clean signal. As a result, most optical force measurements are 
-conducted using a QPD as opposed to a lateral-effect sensor, as often 
-the displacement is small enough that the signal-displacement curve 
-can be considered linear.
+Now while the signal returned by the QPD is proportional to the particle's
+displacement it cannot be directly interpreted as such until converted
+from voltage units to displacement units. This can be done by taking the 
+ratio of the expected diffusion coefficient $D_{exp}$ and the diffusion 
+coefficient calculated by calibration $D_{cal}$.
+\begin{align}
+	\label{eq:conversion_factor}
+	D_{exp} = \frac{k_BT}{6\pi\eta a} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{ext}}{D_{cal}}}
+\end{align}
 
-The above informs how one can - to their best ability - measure small 
-rotational movements in a trajectory. Relying on video analysis is 
-unreliable as the frame rate required to accurately measure the 
-orientation results in a loss of resolution. Likewise, simply measuring
-the total intensity provides no means of separating out rotational and 
-translational motion.
-Similar to Chapter 4, we need to utilise contrasting signals to estimate the change in orientation, and see if at all it is possible to distinguish between translational and rotational motion.
+The former depends on the exact size and shape of the target particle, 
+using image analysis one can get a fairly accurate estimation of the particles radius (a).
+The latter can be determined from fitting the QPD signal to a given model, for computing the diffusion coefficient
+computing the auto correlation function, the mean square displacement, 
+or taking the power spectrum are used. The former two are ideal for 
+simple position detection systems, where the QPD reports the signal at 
+each time step, however they often require longer recording times in 
+order to reach an ideal accuracy. More sophisticated QPD systems can calculate the power spectra of an incoming signal and immediately fit 
+it to a power spectrum. Once the conversion factor is known any further 
+QPD signals can be converted immediately into displacement units to monitor
+the position of the target particle, the spacial resolution being  
+limited by the accuracy of the calibration method. A full tutorial in 
+maximising the accuracy of the power spectral method is provided by 
+Berg and Sorensen \cite{BergSoerensen2004}
 
+For our consideration we will be relying on the power spectrum method 
+to characterise the motion of the trapped particle. Firstly because
+it is sometimes integrated into QPD systems. And secondly, because 
+on of the main advantages of the power spectrum method is its shorter 
+calibration times, and the ability to filter out large sources of signal 
+noise.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Modelling the QPD response signal of a non-spherical particle}
@@
 - even when collecting signals from isotropic scatterers. To prevent 
-this we need to rotate the total field via the inverse rotation matrix
-of the dimer.
+this we rotate the total field via the inverse rotation matrix of the 
+dimer.
 
 To confirm that our method is producing accurate results, we ran a
-comparison between our simulative QPD and the results from \cite{Rohrbach2002}. Where a $300\ nm$ diameter sphere is scanned 
+comparison test between our simulative QPD and the results from \cite{Rohrbach2002}. Where a $300\ nm$ diameter sphere is scanned 
 across the path of a focused Gaussian beam ($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Characterisation of asymmetric dimer dynamics via PSD analysis}
-As discussed in Sec.~\ref{sec:simulated_QPD}, one of the methods 
-developed to work in conjunction with \cite{Vigilante2020} 
-is a simulated quadrant photo diode for as a position 
-detection system. While it is possible to extract all of 
-the relevant dynamics from a simulation, from an experimental
-standpoint characterising those dynamics is dependent on what
-experimental techniques are used. Translational motion can be
-characterised via a position detection system but angular 
-motion is far more difficult to detect, let along characterise.
+While it is possible to extract all of the relevant dynamics from a simulation, from an experimental standpoint characterising those 
+dynamics is dependent on what experimental techniques are used. 
+Translational motion can be characterised via a position detection 
+system but angular motion is far more difficult to detect, let along characterise.
 

### 2025-08-14 Latex_files/Thesis/chapters/chapter 2/main.tex
14/08/25 changes, chp 5 done (ref needed)

@@
 \label{chapter:theory}
-\textit{\textbf{Add overview here}} \\
 This chapter covers the theory behind light and particle interactions, 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Harmonic Traps}
-\label{sec:harmonic_traps}
-When a sphere is located at the centre of a laser the gradient 
-force will try to recentre the sphere if it is displaced. If 
-the restoring force is proportional to the sphere's displacement 
-this is referred to as a harmonic trap. The ratio of the restoring 
-force to the displacement is called the trap stiffness (denoted 
-by $\kappa$), and is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap} demonstrates an example case of the
-force-displacement curve for a simple sphere in an optical trap. 
-Laser power and numerical aperture are omitted as the exact shape 
-is not of importance, only the linear region about the origin.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
-	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement
-		curve while moving along the x-axis, where $x=0$ is the 
-		centre of the beam. The harmonic regime is highlight by the 
-		linear fit given in black.}
-	\label{fig:harmonic_trap}
-\end{figure}
-
-The strength of the optical trap is directly proportional to 
-both the laser power and the contrast in refractive indices 
-between the medium and the particle. Fully describing the 
-optical forces is dependent on the ratio between the particle 
-size ($a$) and the trapping wavelength ($\lambda$). The 
-scattering theory that describes the optical force is dependent 
-on this ratio.  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and refractive index. The $T$-matrix method can be easily 
-applied to any solution of Maxwell's equations. In order to compute 
-the optical force and torque applied to a given object one simply 
-needs to know the momentum carried by the incident and scattered 
-fields. So long as we know the expansion coefficients of the 
-incident beam we can find the scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to compute the force.
-
+applied to any solution of Maxwell's equations. The $T$-matrix
+method allows us to relate directly the momentum carried by 
+the incoming wave \ref{eq:incident} to the momentum retained
+by the scattered wave \ref{eq:scattered}, where the difference 
+between these two quantities is the momentum transferred to
+the scattering object. The time averaged change change in 
+this momentum transfer provides an approximation of the force 
+imparted to the target scatterer.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{TO BE MOVED}
 And the time averaged linear momentum carried by the scattered 
@@
 the Rayleigh approximation.
+
+\textbf{END OF TO BE MOVED}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+This is in truth merely an extension of the Lorenz-Mie theory 
+as the T-matrix of a single sphere can be found from the electric 
+field solutions. The method developed by Waterman allows for any 
+arbitrary shaped particle to produce similar solutions. One of 
+the biggest advantages to the method is the fact that it can 
+leverage symmetries in particle geometries to speed up calculations.
+
 For this project we utilised an extension of the EBCM known as 
@@
 the electromagnetic field as a collection of individual 'rays' that 
-propagate and are refracted according to Snell's Law. Based on the 
-change in direction momentum is transferred to the target particle; 
-with rays closest to the centre of the beam having greater intensity 
-than those rays at the very edge of the beam. Consider a particle 
-struck by two rays in a Gaussian beam, one coming close to the centre, 
-and the other ray coming from the edge. As each ray is refracted by 
-the target sphere a force is imparted onto it, the total force 
-imparted is given by:
+propagate and are refracted according to Snell's Law. The Ray optics
+model is generally only used in cases where the particle size is far
+greater than the wavelength of light  Based on the change in direction 
+momentum is transferred to the target particle; with rays closest to 
+the centre of the beam having greater intensity than those rays at the 
+very edge of the beam. Consider a particle struck by two rays in a 
+Gaussian beam, one coming close to the centre, and the other ray 
+coming from the edge. As each ray is refracted by the target sphere a 
+force is imparted onto it, the total force imparted is given by:
 \begin{align}
@@
 Gaussian intensity distribution $P_i$ will fall off as you move from
-the centre of the beam. The ray optics model is ideal when dealing 
-with larger particles whose diameter far exceed the wavelength of light
-being scattered. While it can be useful in predicting the forces 
-experienced by said particle's it does not fully capture the behaviour
-of light when considering interference between different rays. And 
-as such no information is learned about the scattered field.
+the centre of the beam. While the ray optics model can be useful in 
+predicting the forces experienced by said particle's it does not fully 
+capture the behaviour of light when considering interference between 
+different rays. And as such no information is learned about the scattered 
+field.
 
@@
 The Rayleigh approximation is for describing a particle when 
-the particle radius is $\ll \lambda$. The underlying theory 
-is that a dielectric sphere can be treated as a dipole while 
-in the presence of the electromagnetic field. In which case 
-the scattering force is given simply by the scattering of 
-the induced dipole, and the gradient force is due to the 
-Lorentz force \cite{Gordon1973}. The gradient forces in the 
-principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
-rectangular force field:
+the particle radius is $\ll \lambda$, as the particle shape 
+decreases we find the effective scattered area approaches 
+zero as the incident light diffracts around the target particle. 
+As such we can approximate a dielectric sphere as a dipole while 
+in the presence of the electromagnetic field. In which case the 
+scattering force is given simply by the scattering of the induced 
+dipole, and the gradient force is due to the Lorentz force 
+\cite{Gordon1973}. The gradient forces in the principle Cartesian 
+axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS 
+units as a restorative rectangular force field:
 \begin{align}
@@
 There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
-\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- \eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
+\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- 
+\eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
 the state of the scattered field. This is because if the particle 
@@
 transfer calculation and simply compute the Lorenz force for a 
-dipole in an electric field. While not important for optical force calculation higher complexity scattering problems can be simplified 
-by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), in which case the scattered field is now significant. As the target particle gets larger this approximation 
-fails to accurately describe the trapping force \cite{Li2021}. For 
-larger particles the optical force and scattering is better described 
-via the Lorenz-Mie theory.
+dipole in an electric field. While not important for optical force 
+calculation higher complexity scattering problems can be simplified 
+by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
+in which case the scattered field is now significant. As the 
+target particle gets larger this approximation fails to accurately 
+describe the trapping force \cite{Li2021}. For larger particles 
+the optical force and scattering is better described via the 
+Lorenz-Mie theory.
 
@@
 experimental setting the observed motion is not as 
-simple. As such it is often far simpler to 
-characterise the optical trap by recording the particle's
-trajectory (see \ref{sec:position_detection} for 
-specifics regarding the equipment), and then fitting it 
-to our presumed description of the beam. The trajectory 
-is easily understood if we understand the particle's 
+simple. As such it is often far simpler to characterise 
+the optical trap by recording the particle's trajectory 
+(see \ref{sec:position_detection} for specifics 
+regarding the equipment), and then fitting it to our 
+presumed description of the beam. The trajectory is 
+easily understood if we understand the particle's 
 diffusive behaviour.
 
-\subsection{Langevin Equation}
-Describing any microscopic motion requires an understanding 
-of a molecules diffusive behaviour, for the case of optical 
-tweezers the most complete model of diffusion is the Langevin 
-equation. Models such as the Fickian, and Einstein derivations 
-are sufficient for macroscopic behaviours but the Langevin 
-model better describes the microscopic characteristics of 
-any diffusive behaviour.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Diffusion Models}
+For the case of optical tweezers the most complete model of 
+diffusion is the Langevin equation. Models such as the Fickian, 
+and Einstein derivations are sufficient for macroscopic 
+behaviours but the Langevin model better describes the 
+microscopic characteristics of any diffusive behaviour.
 
@@
 able to describe its motion over any time scale, unlike the 
-Einstein model. The standard form of the Langevin model for 
-a particle undergoing Brownian motion is given as:
-\begin{align}
-	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
+Einstein model. Unlike the Einstein model that views collisions
+as discrete but finite events, the Langevin model instead 
+considers the collisions as one random The standard form of the 
+Langevin model for a particle undergoing Brownian motion is 
+given as:
+\begin{align}
+	m\frac{d\vec{v}}{dt} = - \gamma_0 \vec{v} + F(t) + W(t)
 \end{align}
@@
 
-Where $w_i$ is a random real number between -1 and 1, we can say 
-that it is normally distributed around 0 for simulation purposes. 
+Where $w_i$ is a Gaussian random real number between -1 and 1, we 
+can say that it is normally distributed around 0 for simulation purposes. 
 We can rearrange this for $x_i$ to approximate the Brownian motion 
@@
 
-Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
-or \textit{ott} \cite{Nieminen2007} we can compute the beam 
-shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
-type. Using the results from \cite{Farsund1996} we can then 
-compute both the optical force and torque using the beam 
-coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
-which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
-\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
-total force in along the z-direction and the total torque about 
-the z-axis are provided:
-\begin{equation}
-	\label{eq:optical_force}
-	\begin{split}
-		\bold{F_z}
-		&=
-		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
-		\\ 
-		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-		\\
-		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
-	\end{split}
-\end{equation}
-\begin{equation}
-	\begin{split}
-		\bold{T_z}
-		&=
-		-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
-		\\ 
-		&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
-		\\
-		& \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
-	\end{split}
-\end{equation}
-
-where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
-coefficients of the incident and scattered fields respectively. 
-We can get the $x$ and $y$ force and torque components in a 
-similar form by applying a simple rotation transformation. With 
-the optical forces and torques computed all that remains is to 
-compute the Brownian forces and torques which are constrained 
-by the relation.
-\begin{align}
-	\langle q_iq_j\rangle =2D_{ij}\Delta t
-\end{align}
-
-Vigilante et al compiled together a python package that combines 
-both \textit{ott} and \textit{mstm} to simulate the behaviour of 
-spherical aggregates within a predefined optical trap 
-\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
-modules that can be used to model the scattering from different 
-shaped particles; however, for aggregates of spheres the resultant 
-force is nonsensical (see Chapter 4 for further details). As such 
-we rely on the \textit{mstm} for simulating dimers.  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Harmonic Traps}
+\label{sec:harmonic_traps}
+When a sphere is located at the centre of a laser the gradient 
+force will try to recentre the sphere if it is displaced. If 
+the restoring force is proportional to the sphere's displacement 
+this is referred to as a harmonic trap. The ratio of the restoring 
+force to the displacement is called the trap stiffness (denoted 
+by $\kappa$), and is an important characteristic of the trap. 
+Fig.~\ref{fig:harmonic_trap} demonstrates an example case of the
+force-displacement curve for a simple sphere in an optical trap. 
+Laser power and numerical aperture are omitted as the exact shape 
+is not of importance, only the linear region about the origin.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
+	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement
+		curve while moving along the x-axis, where $x=0$ is the 
+		centre of the beam. The harmonic regime is highlight by the 
+		linear fit given in black.}
+	\label{fig:harmonic_trap}
+\end{figure}
+
+The strength of the optical trap is directly proportional to 
+both the laser power and the contrast in refractive indices 
+between the medium and the particle. Fully describing the 
+optical forces is dependent on the ratio between the particle 
+size ($a$) and the trapping wavelength ($\lambda$). The 
+scattering theory that describes the optical force is dependent 
+on this ratio.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 period of time we can construct and experimental estimation
-of $\rho(x)$. By inserting \eqref{eq:prob_dist} into \eqref{eq:potential_well} we can fit the potential well in 
+of $\rho(x)$. By inserting \eqref{eq:prob_dist} into 
+\eqref{eq:potential_well} we can fit the potential well in 
 order to determine the trap strength $\kappa_x$. This has 
@@
 
-If we use \eqref{eq:anylitical_lang} for an optical tweezer we can 
-expand out the squared term to get an analytical expression for the 
-MSD as a function of time lags:
+If we sub in \eqref{eq:anylitical_lang} for $x(t)$ we can expand 
+out the squared term to get an analytical expression for the MSD 
+as a function of time lags:
 \begin{align}
 	\label{eq:MSD}
-	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
+	MSD(\tau) &= \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> \\
+	\rightarrow MSD(\tau) &=\frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
 \end{align}

### 2025-08-14 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to QPD chapter
14/08/25 changes, chp 5 done (ref needed)

@@
 \section{Simulation Details}
+\subsection{Set up}
 As a paradigmatic example, consider a dimer suspended in 
-water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
-a Gaussian beam (more specifically a Laguerre-Gaussian 
-beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
-is focused by a objective with numerical aperture of 1.2.
-The polarisation is defined using the Jones vector which 
-describes the magnitude of the x and y components of the 
-electric field (so the vector $[1.0, 0.0i]$ describes an 
-electric field that only oscillates in the x direction). 
+water ($n_p = 1.59$) located at the focus of a Gaussian 
+beam. The dimer is composed of two spheres with radii 
+$a_I$ and $a_{II}$, the size ratio between the two is 
+given as $a_{II}/a_{I}$ - unless stated otherwise we 
+kept $a_I$ fixed at $1\ \mu m$.
+
+The beam is focused by a objective with numerical aperture 
+of 1.2. The polarisation is defined using the Jones vector 
+which describes the magnitude of the x and y components of 
+the electric field (so the vector $[1.0, 0.0i]$ describes 
+an electric field that only oscillates in the x direction). 
 Unless stated otherwise the Jones vector is kept at $[1.0, 
-0.0i]$.
-
-As a paradigmatic example, we study a simple spherical 
-dimer composed of polystyrene, while there is the 
-possibility of considering absorption we choose to keep 
-the refractive index fixed at 1.59. The ratio of the two 
-sphere radii is given by $a_{I}/a_{II}$ where $a_{I}$ is 
-kept at $1\ \mu m$ unless specified otherwise. For 
-simulations we assume that the dimer is suspended in water 
-at room temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).
-
-In order to compute the magnitude of the Brownian forces we
-need to know the dimer's diffusion tensor \textbf{D}. Unlike
-for spheres where the translational and rotational diffusion 
-coefficients are equal along all three Cartesian axis', a
-dimer's translational and rotational diffusion is dependent
-on whether we consider the its long or short axis. We use
-the results from \cite{Nir1973} to compute the dimensionless 
-diffusion tensor and then scale it according to the radius of
-the largest sphere ($a_I$). The dimer's centre of diffusion
-is located a distance of $a_I\zeta$ from the point of contact
-between the two spheres, the coefficient $\zeta$ is based on
-the size ratio of the two spheres. We use the centre of 
-diffusion as the target origin that is provided to \textit{mstm}
-
-Where $U_3$ represents the orientation vector that passes
-through the centres of both sphere's (see \ref{sec:sim} 
-for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
-$U_3$ but have no physical representation for a dimer. We 
-can represent $U_3$ by it's spherical coordinates ($\theta$ 
-and $\phi$), where $\theta$ is the angle between the 
-orientation vector and the direction of propagation and 
-$\phi$ as the angle between the orientation vector and 
-the x-axis. We define the 'standard' orientation as 
-being aligned with the direction of beam propagation 
-direction ($\theta=0^\circ$) - and therefore the 'inverted' 
-orientation is defined when the dimer is orientated against 
-the direction of beam orientation ($\theta=180^\circ$) - see
-fig~\ref{fig:paradigmatic} for a visual of the laboratory and
-particle frame.
-
+0.0i]$. The conditions of the water are given as - $T = 298K,\ 
+\eta = 1.0x10^{-3}Pa\ s,\ n_m = 1.33$.
+
+\subsection{Coordinate system}
 Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
@@
 particle's centre of mass, this is also used to define it's 
-displacement from the origin of the laboratory frame. The diffusion 
-tensor of the dimer is more complex than a sphere's as we must 
-consider the rotational effects experienced by the dimer. We use 
-the analytical solution of Nir \& Acrivos \cite{Nir1973}, where 
-they provided the solution coefficients depending on the dimer's 
-size ratio ($a_I/a_{II}$). Rather than recalculate the coefficients 
-for every possible configuration of dimer we use a spline fitting 
-function to approximate the diffusion tensor.
-
-We utilise a python library that utilises \textit{mstm} and
-\textit{ott} to simulate the dynamical behaviour of spherical
-aggregates. By specifying the optical properties of the target 
-\textit{mstm} generates the targets T-matrix which can then be 
-used by \textit{ott} to compute the optical forces generated by 
-a focused beam. Using the Finite Differences approach we can solve
-the Langevin Equation to take discrete time steps and update the 
-dimer's position and orientation. A time step of $10 \mu s$ is 
-used in line with previous works \cite{Vigilante2020}.
+displacement from the origin of the laboratory frame.
+
+Where $U_3$ represents the orientation vector that passes
+through the centres of both sphere's (see \ref{sec:sim} 
+for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
+$U_3$ but have no physical representation for a dimer. We 
+can represent $U_3$ by it's spherical coordinates ($\theta$ 
+and $\phi$), where $\theta$ is the angle between the 
+orientation vector and the direction of propagation and 
+$\phi$ as the angle between the orientation vector and 
+the x-axis. We define the 'standard' orientation as 
+being aligned with the direction of beam propagation 
+direction ($\theta=0^\circ$) - and therefore the 'inverted' 
+orientation is defined when the dimer is orientated against 
+the direction of beam orientation ($\theta=180^\circ$) - see
+fig~\ref{fig:paradigmatic} for a visual of the laboratory and
+particle frame.
+
+\subsection{Diffusion Matrix}
+In order to compute the magnitude of the Brownian forces we need to 
+know the dimer's diffusion matrix \textbf{D}. Unlike for spheres 
+where the translational and rotational diffusion coefficients are 
+equal along all three Cartesian axis', a dimer's translational and 
+rotational diffusion is dependent on whether we consider the its 
+long or short axis. We use the results from \cite{Nir1973} to 
+compute the dimensionless diffusion matrix and then scale it 
+according to the radius of the largest sphere ($a_I$). The dimer's 
+centre of diffusion is located a distance of $a_I\zeta$ from the 
+point of contact between the two spheres, the coefficient $\zeta$ 
+is based on the size ratio of the two spheres. We use the centre of 
+diffusion as the target origin that is provided to \textit{mstm}
+
+\subsection{Equations of Motion}
+\label{sec:eq_of_motion}
+The main equation of motion for an optically trapped particle is the 
+Langevin equation. For a dimer the equation remains the same but is 
+altered slightly to account for coupling between rotational and 
+translational motion. 
+\begin{align}
+	\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + 
+	\sqrt{2\vec{D}_x}\eta(t)
+\end{align}
+
+Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
+motion is generalised to a 3 dimensional case. Here we omit the inertial term 
+as the restoration time of the optical trap is far greater than the characteristic time of the particle ($\kappa/\gamma \gg m/\gamma$).
+Now, the dimer is undergoing random rotational motion in addition to its Brownian
+translational motion the first term on the right hand side is no longer 
+purely a function of the dimer's position but also of its orientation. 
+The rotational form of the Langevin equation for a dipole within an external 
+potential is given as \cite{Mazo2008}:
+\begin{align}
+	\frac{{d}\vec{u}(t)}{{dt}}
+	=
+	\frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
+	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+
+Where $\vec{u}(t)$ is the unit vector aligned along the centres of 
+the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the 
+rotational friction coefficient which is given as $\gamma_R = 8\pi
+\eta r^3$ for a sphere. The term $E(t)$ corresponds to the time 
+dependent electric field vector, however we can assume that over a
+time step as large as ours the time variation in the electric field 
+is so insignificant that it can be treated as a constant. If we assume 
+that over the course of a trajectory the dimer has a 'equilibrium' orientation in which the optical torque is minimised we can say that 
+the first right hand term can be replaced with $\vec{\kappa}_u(\vec{r}
+(t))$ to represent a harmonic potential due to the torque. $\lambda(t)$ 
+is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian torque are normally distributed and are 
+also uncorrelated so that:
+\begin{align}
+	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
+\end{align}
+
+For an asymmetric scatterer whose radius is comparable to that of the 
+electric field's wavelength we now have a system of simultaneous equations:
+\begin{align}
+	\label{eq:full_langevin}
+	\frac{{d}\vec{r}(t)}{{dt}}
+	&=
+	\frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2k_BT\vec{D}}\eta(t)
+	\\
+	\frac{{d}\vec{u}(t)}{{dt}}
+	&=
+	\frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
+	+ \sqrt{2k_BT\vec{D}_R}\lambda(t)\times \vec{u}(t)
+\end{align}
+
+Fortunately, we do not need to solve these directly as the latter two
+random variables can be easily approximated if the thermal energy of
+the system is known, and the rate of change can be assumed as linear if
+we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
+\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
+to compute the optical force and torque applied to the dimer.
+
+\subsection{Computational methods}
+Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
+or \textit{ott} \cite{Nieminen2007} we can compute the beam 
+shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
+type. Using the results from \cite{Farsund1996} we can then 
+compute both the optical force and torque using the beam 
+coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
+which are found by calculating the dimer's $T$-matrix via \textit{mstm} 
+\cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
+total force in along the z-direction and the total torque about 
+the z-axis are provided:
+\begin{equation}
+	\label{eq:optical_force}
+	\begin{split}
+		\bold{F_z}
+		&=
+		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
+		\\ 
+		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+\begin{equation}
+	\begin{split}
+		\bold{T_z}
+		&=
+		-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
+		\\ 
+		&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
+		\\
+		& \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
+	\end{split}
+\end{equation}
+
+where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam 
+coefficients of the incident and scattered fields respectively. 
+We can get the $x$ and $y$ force and torque components in a 
+similar form by applying a simple rotation transformation. With 
+the optical forces and torques computed all that remains is to 
+compute the Brownian forces and torques which are constrained 
+by the relation.
+\begin{align}
+	\langle q_iq_j\rangle =2D_{ij}\Delta t
+\end{align}
+
+Vigilante et al compiled together a python package that combines 
+both \textit{ott} and \textit{mstm} to simulate the behaviour of 
+spherical aggregates within a predefined optical trap 
+\cite{Vigilante2020}. Now \textit{ott} does have inbuilt $T$-matrix 
+modules that can be used to model the scattering from different 
+shaped particles; however, for aggregates of spheres the resultant 
+force is nonsensical (see Chapter 4 for further details). As such 
+we rely on the \textit{mstm} for simulating dimers.
 
@@
 	\end{subfigure}
-	\caption{Plots of force vs displacement of the 
-		centre of mass of the dimer (µm) for the 
-		case of a dimer of size ratio 2. (a) is the 
-		case where the dimer is in its' 'standard' 
-		orientation, where the dimer is trapped at 
-		$z = 1.71\ \mu m$. (b) is the case where the 
-		dimer is in its' 'inverted' orientation, the 
-		dimer is trapped at two positions: $z = 1.20\ 
-		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
-		renders to visualise the dimer orientation 
-		are shown below each plot. The black lines 
-		on each force-curve is a linear fit with the 
-		slope being reported as the trap stiffness in 
-		the legend.}
+	\caption{Plots of force vs displacement of the centre of mass of the 
+		dimer (µm) for the case of a dimer of size ratio 2. (a) is the 
+		case where the dimer is in its' 'standard' orientation, where 
+		the dimer is trapped at $z = 1.71\ \mu m$. (b) is the case where 
+		the dimer is in its' 'inverted' orientation, the dimer is trapped 
+		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. On the 
+		left are renders to visualise the dimer orientation are shown 
+		below each plot. The black lines on each force-curve is a linear 
+		fit with the slope being reported as the trap stiffness in the 
+		legend.}
 	\label{fig:paradigmatic}
@@
 	%\captionsetup{hangindent=2.35cm}
-	\caption{Map of $z-\theta$ phase space using a dimer 
-		of size ratio 2 with a laser power of 500 mW 
-		($\theta=0^\circ$ is the 'standard' orientation 
-		and $\theta=180^\circ$ is the 'inverted' 
-		orientation). The stable configurations are 
-		indicated by the larger circles and the 
-		starting conditions are colour coded to match 
-		the stable point they end up in. Right hand 
-		render shows a dimer in its off-axis configuration.}
+	\caption{Map of $z-\theta$ phase space using a dimer of size ratio 
+		2 with a laser power of 500 mW ($\theta=0^\circ$ is the 
+		'standard' orientation and $\theta=180^\circ$ is the 'inverted' 
+		orientation). The stable configurations are indicated by the 
+		larger circles and the starting conditions are colour coded to 
+		match the stable point they end up in. Right hand render shows a 
+		dimer in its off-axis configuration.}
 	\label{fig:off_axis}

### 2025-08-14 Latex_files/Thesis/chapters/chapter 4/main.tex
14/08/25 changes, chp 5 done (ref needed)

@@
 \section{Interpretation of scattering data into orientation estimates}
+\label{sec:ref_orientations_explained}
 We now discuss the methodology for computing the instantaneous 

### 2025-08-14 Latex_files/Thesis/chapters/chapter 5/main.tex
14/08/25 changes, chp 5 done (ref needed)

@@
 \section{Characterisation of asymmetric dimer dynamics via PSD analysis}
+\label{sec:PSD_analysis}
 While it is possible to extract all of the relevant dynamics from a simulation, from an experimental standpoint characterising those 
@@
 
-The simulative QPD is composed of 4 photodiodes that measure the 
-intensity of light incident on their surfaces. Given that the 
-condensing lens will have a maximum acceptance angle given by 
-its numerical aperture ($\theta_{max} = \sin^{-1}\left(NA_c/n_
-{med}\right)$) we define the electric field incident on the QPD 
-surface as the elements of the scattered and incident field that 
-propagate within that angular range of $\pm \theta_{max}$. Using 
-\textit{ott} we can evaluate the total electric field to an 
-arbitrary spacial resolution across the surface of the QPD. The 
-reported signal is found by taking the difference between pairs 
-of quadrant signals (see \ref{sec:simulated_QPD} for further 
-detail). This does not provide a one-to-one result however as 
-hardware errors (i.e. internal resistance, external light sources, 
-and vibrations) would distort the signal. It can be used however to 
-show how if the reported power spectra is an accurate representation 
-of the true particle dynamics or if the additional degrees of freedom 
-have an adverse effect on the scattering. See chapter 2 for a breakdown
-of power spectral analysis.
+The tweezer setup we are simulating incorporates a dual objective 
+configuration, using one objective lens to focus the beam (NA=1.25)
+where as the condenser lens acts as a means of truncating the electric
+field ($NA_c=0.25$). The electric field falling onto the QPD surface is computed by considering the far field of the combined incident and 
+scattered beams. We can ignore any contributions from the electric
+field that falls outwith the maximum acceptance angle of our condenser
+lens ($\theta_{max}\approx\pm10.8\deg$). This is not perfectly analogous
+of a true QPD as external factors (i.e. internal resistance, external
+light sources, and vibrations) would also contribute to the detected QPD 
+signal. It does provide us a means of showing what elements of a particle's
+trajectory are represented in the power spectrum and which elements are 
+obscured.
 
@@
 	\captionof{table}{QPD fitting for single sphere}
+	\label{tab:sphere}
 	\begin{tabular}{ |c|c|c|c|c|c|c| } 
@@
 
-This coupling of the translational and orientational motion 
-has been highlighted previously \cite{Vigilante2020}, but 
-there effects have not been demonstrated in the context of
-an experimental situation. We can see from the difference in
-trajectory and QPD fittings that while our description of the
-trap shape is very similar (both methods give similar ellipticity
-values) but the magnitude of the trap strength are significantly
-different. This is partially explained by the fact that the
-QPD is not actually measuring the position of the particle
-but instead the intensity distribution of the total field 
-incident on the QPD surface. In which case if the dimer is 
-rotated slightly the QPD signal will change even if the dimers
-centre of diffusion remains stationary.
+There is still a concern in the fact that both produce similar
+power spectra, enough so that it would be difficult to determine
+if a trapped particle was actually a dimer or a sphere. This is
+because the rotational and translational motion, coupling of the 
+has been highlighted previously \cite{Vigilante2020}, but there 
+effects have not been demonstrated in the context of an experimental situation. We can see from the difference in trajectory and QPD 
+fittings that while our description of the trap shape is very 
+similar (both methods give similar ellipticity values) but the 
+magnitude of the trap strength are significantly different. This 
+is partially explained by the fact that the QPD is not actually 
+measuring the position of the particle but instead the intensity 
+distribution of the total field incident on the QPD surface. In 
+which case if the dimer is rotated slightly the QPD signal will 
+change even if the dimers centre of diffusion remains stationary. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{QPD for angular displacement detection}
+\section{QPD for angular displacement detection}
+\label{sec:angular_displacement}
 The next logical step is to consider whether or not a QPD can at 
@@
 
-However if we compare this to fig.~\ref{fig:rohrbach} we can see 
+However if we compare this to fig.~\ref{fig:Rohrbach} we can see 
 that the the signal change is dwarfed by the translational motion. 
@@
 QPD signal is not a simple task. This is significant because it 
-means that with regards to the power spectrum fittings in table 
-\ref{tab:dimer} there is no way of ensuring that the measured 
-trap strengths are accurate. The variance $\sigma(f_c)$ is only 
-the variance based upon our fitting parameters and not on the 
-actual dynamics of the dimer. In the case where the particle's
-shape is not known prior its not possible to determine whether 
-any inaccuracies in the power spectra are due to the fit or due
-to the particle's motion.
+means that with regards to the power spectrum fittings in Table 
+\ref{tab:dimer} there is no means of discerning between that of 
+a trapped sphere and that of a trapped dimer. Furthermore the 
+dynamics reported by the power spectrum fitting do not accurately 
+reflect the actual dynamics of the dimer, instead only providing 
+a orientation-averaged trap strength. But unless we have some means
+of at least estimating the degree of rotational freedom, calculating
+the actual trapping strength is impossible.
 
@@
 which takes as an input the voltage from the 4 quadrants and 
-tries to fit that to the dimer's trajectory. Fitting the QPD
-signal to the dimer's position shows promising results, shown in
-fig~\ref{fig:QPD_rotation} is a prediction of a symmetric dimer's 
-position based on the QPD signal. With the actual displacement 
-on the x-axis and the predicted result y-axis, and the dashed 
-line represents the y=x line - indicating an ideal prediction.
+tries to fit that to the dimer's trajectory. We utilised the 
+scikit-learn module available in Python to transform the 4D vector
+of the QPD signal to a 2D vector specifying the expected displacement
+of the target particle. Unlike in Chapter 4 where we utilised 
+machine learning to get a probabilistic estimation of the 
+particle's orientation we utilise vector regression as we 
+want to replicate QPD precision as closely as possible.
 
-Fig~\ref{fig:QPD_rotation} shows that its relatively trivial 
-to predict a particle's position based off the QPD signal, even 
-in axial direction which has not been done previously. It should 
-be noted that these displacements are relative to the trap focus 
-and so accurate tracking of the beam movements needs to be taken. 
-This result is partially due to the fact that displacement and 
+In order to fit the QPD signal to dimer displacements we need
+to train our model. For each size ratio considered a 3 second
+trajectory was ran (using the same conditions in 
+Sec.~\ref{sec:PSD_analysis}) and a normal distribution of the 
+dimer's position was taken and used to construct a training 
+data set of QPD signals. After the training results reached an
+error rate of less than 1\% the same trajectory was used to
+construct a test set that would allow our model to fit the QPD 
+signal to the positional data from the trajectory. Fitting the 
+QPD signal to the dimer's position showed promising results, 
+shown in fig~\ref{fig:QPD_rotation} is a prediction of a 
+symmetric dimer's position based on the QPD signal. With the 
+actual displacement on the x-axis and the predicted result y-axis, 
+and the dashed line represents the y=x line - indicating an 
+ideal prediction. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.9\linewidth]{predicted_displacements.png}
+	\caption{Scikit-Learn's best predictions of the displacement 
+		of a symmetric dimer placed in an optical trap, predictions 
+		are based on the QPD signal collected at the 4 quadrants. 
+		(Left) predictions of the $X$ \& $Y$ displacements. 
+		(Right) predictions of the $Z$ displacement, predictions 
+		have been scaled so they represent the displacement from 
+		the beam focus. The dashed black line represents when the 
+		predicted displacement = the true displacement.}
+	\label{fig:predicted_displacements}
+\end{figure}
+
+Fig~\ref{fig:predicted_displacements} shows that its relatively 
+trivial to predict a particle's position based off the QPD signal
+in the $X$-$Y$ plane but harder along the axial direction. If we 
+consider Fig~\ref{fig:Rohrbach} we note that the change in signal
+strength is only truly linear close to the beam focus, as shown in
+Chapter 3, dimers do not trap close to the beam focus but further 
+away. This would suggest that the model cannot accurately predict 
+the displacement of a dimer in the axial direction as there is a 
+non-linear relationship between the change in signal and the change 
+in axial position. It should be noted that even in the worst case 
+the difference between the predicted and true displacements is 
+around $~20 nm$.
+
+These displacements are relative to the trap focus and 
+so accurate tracking of the beam movements needs to be taken if 
+you want to consider the displacement relative to some larger object. 
+These results are likely due to the fact that displacement and 
 signal units are linearly related by \eqref{eq:conversion_factor}. Interestingly when we run the same protocol for different sized 
 dimer's there seems to be a cut off point when the fit begins to 
-fail. Plotting the $R^2$ for each fit vs its size ratio shows 
-that beyond a size ratio of 5 the machine learning begins to fall 
-off significantly.
+fail. We can compute the coefficient of determination ($R^2$) for 
+our model's prediction, the $R^2$ for each fit vs its size ratio 
+shows that beyond a size ratio of 5 the machine learning begins 
+to fall off significantly.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{score_results.png}
+	\caption{The coefficient of determination for different sized 
+	dimers using the scikit-learn vector regression model. Each line
+	plot shows the $R^2$ value for either the $X$, $Y$, or $Z$ 
+	displacements. When the size ratio is greater than 5 the $R^2$ 
+	value falls off dramatically, from around 0.95 down to as low as 
+	0.01.}
+\end{figure}
 
@@
 should be able to easily detect the dimer's translational motion.
-We can therefore only conclude that the dimer's rotational motion
-is contributing to the scattering to the scattered field, even if
-it is miniscule the contribution is enough to throw off a 
-position detection system. Detection of rotational motion was 
-attempted using machine learning, the model was trained on the 
-same trajectory mentioned in \ref{sec:off-axis}, where the dimer 
-is trapped in an off-axis orientation, from a top down perspective
-it should be easy enough to note that the dimer is not vertically 
-trapped even if the exact angle is difficult to determine.
+We can therefore only conclude that the contribution that the 
+second sphere makes to the scattered beam is not proportional to
+its size. What is interesting is that the axial prediction does 
+not experience as a large a drop in accuracy, suggesting that the
+total field incident on the QPD is not changing drastically, but
+rather that the second sphere is altering the distribution such
+that the $X-Y$ displacements are being incorrectly predicted. 
+This could simply be a limitation of the \textit{mstm} software being 
+unable to accurately represent the T-matrix of the second sphere, 
+however the documentation does not specify any limits on the relative 
+size differences so it is unclear if this is a known issue. The 
+problem is that it is difficult to verify the error between the 
+\textit{mstm} predicted scattering and the true scattering without 
+replicating the exact system experimentally. For now we can say that predictions for size ratios greater than 5 should be for the most 
+part disregarded.
+
+Detection of rotational motion was attempted using the same method, 
+the model was trained on the same trajectory mentioned in 
+\ref{sec:off-axis}, where the dimer is trapped in an off-axis 
+orientation. We chose this trajectory instead of a strictly vertical
+orientation in order to confirm that the model can distinguish between 
+the polar angle and the azimuthal angle. The results of which are 
+are shown in Fig.\ref{fig:predicted_rotations}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{angular_predictions.png}
+	\caption{Scikit-learn's prediction of the sphericical angles of
+	a dimer trapped in an off-axis configuration (see Sec\ref{sec:off-axis}).
+	(Left) predictions of the azimuthal angle, (Right) predictions 
+	of the Polar angle. Dashed black lines show the ideal prerformance 
+	where the predicted angle = true angle.}
+	\label{fig:predicted_rotations}
+\end{figure}
+
+From here we can see that the vector regression model performs rather
+poorly in predicting the rotations made by the dimer, even in a situation
+where visibly the dimer is an off-axis orientation. While the model can 
+roughly predict the median polar angle (~$102\deg$) it has no means of
+predicting the rotations made by the dimer. This result is reveals two important factors: Firstly, the change in the QPD signal for rotational
+motion is a full order of magnitude smaller than the change in signal 
+for translational motion meaning the model has no means of distinguishing
+between large and small rotations. Secondly, is the fact that unlike 
+translational motion, rotational motion is dependent on the prior 
+orientations in the trajectory. This was already discussed in Chapter 4
+where we improved our machine learning model by accounting for the 
+the priors $p(\hat{\bf n}_\alpha)$, where we weighted our estimation 
+by the distance from our prior estimate. Vector regression is powerful
+for predicting uncorrelated events but has no means for accounting for
+cases where one event is dependent on the next. At the same time, 
+utilising machine learning in the case of a QPD would not be able to 
+provide a high enough orientational resolution without first sampling 
+from a trajectory and creating a unique set of reference orientations
+(see Sec.~ \ref{sec:ref_orientations_explained} for definitions).
 
@@
 while restricting translational motion.
+
+\section{Conclusion}
+We have developed a program that allows one to compute the QPD 
+response signal produced by an arbitrary shaped particle. By 
+utilising \textit{mstm} we can construct any combination of 
+spherical particles and measure the total field incident on the
+focal plane of the condenser lens. This has immediate applications 
+in replicating power spectral data from complex particle structures 
+that experience periodic rotational motion. Furthermore, if all 
+contributions to the QPD response signal are known it would allow 
+researchers to compare the reported dynamics given by an experimental
+QPD to that of our simulated response signal.
+
+The QPD module shows promising results for dimers who have a size 
+ratio less than 5. Utilising vector regression models allowed for 
+accurate predictions of the displacements made by dimers while 
+confined by an optical trap. The use of vector regression is necessary
+to return a continuos range of positions instead of a probabilistic
+approximation of the particle's position vector.
+
+As shown in \ref{sec:angular_displacement} random rotational motion 
+is difficult to distinguish from random translational motion. Even
+when utilising a vector regression model the angular displacement
+predicted does not match the trajectory data. This means that at least
+for now utilising focal plane interferometry cannot provide an 
+accurate estimation of angular displacements. While static light 
+scattering methods - as seen in Chapter 4 - do present a possible 
+alternative, the limited angular resolution due to a high mixing 
+of signal and orientation spaces also make the method difficult to 
+implement. Measurements of angular displacement are crucial for 
+measuring torques exerted on trapped particles, current torque 
+measurements assume that only the the polarisation contributes to 
+the applied torque, measuring non-polarisation related torques still 
+eludes us.
+
+Overall, tracking the angular displacements of non-spherical 
+objects for the purposes of torque measurements will require 
+further study. Fully understanding the torques applied to an 
+arbitrary shaped particle would pave the way for high-precision 
+particle-manipulation techniques and a greater understanding 
+of the interactions between focused beams and aggregate particles.
